### 2008-04-15 docs/Godunov_notes/paper.tex
Working copy of notes for Steps 4b and 8b in Paper III.

\documentclass[11pt]{article}

\tolerance=600

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{color}

% Margins
\usepackage[lmargin=1.0in,rmargin=1.0in,tmargin=0.75in,bmargin=0.75in]{geometry}

\newcommand{\sfrac}[2]{\mathchoice
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
  {#1\!/#2}}

\def\half  {\frac{1}{2}}
\def\dt    {\Delta t}

\def\edge  {\rm EDGE}
\def\mac   {\rm MAC}
\def\trans {\rm TRANS}

\def\eb    {{\bf e}}
\def\ib    {{\bf i}}
\def\ub    {{\bf u}}
\def\Ub    {{\bf U}}

\def\Ubt   {\widetilde{\bf U}}
\def\wt    {\tilde{w}}

\title{Notes on the Godunov Step in {\tt MAESTRO}}

\begin{document}
\maketitle

These are working notes for the Godunov step in {\tt MAESTRO}.

\section{Computing $\ub^{\trans}$}
In {\tt advance\_premac.f90}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\ub^{\trans}$.  These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0}.  We only compute the normal component of velocity at each face.\\

The evolution equation for the perturbational velocity is:
\begin{equation}
\frac{\partial\Ubt}{\partial t} = -\Ubt\cdot\nabla\Ubt - w_0\frac{\partial\Ubt}{\partial r} - \wt\frac{\partial w_0}{\partial r}\eb_r - \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.
\end{equation}
We are going to use a 1D Taylor series extrapolation in space and time to compute face-centered normal velocities.  By 1D, we mean that we omit any spatial derivatives that are not in the direction of the extrapolation.  We first compute ``left'' and ``right'' states, and then use a Riemann solver to pick the final state.\\

For the x-faces, the extrapolation is as follows:
\begin{eqnarray}
u_L &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
u_R &=& u_{\ib}^n - \left(\half + \frac{\dt}{h}u_{\ib}^n\right)\Delta_x u_{\ib}^n,
\end{eqnarray}
where $\ib = (i,r)$ in 2D and $\ib = (i,j,r)$ in 3D.  For 3D problems only, for the y-faces, the extrapolation follows analogously:
\begin{eqnarray}
v_L &=& u_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_y}^n\right)\Delta_y u_{\ib-\eb_y}^n, \\
v_R &=& u_{\ib}^n - \left(\half + \frac{\dt}{h}u_{\ib}^n\right)\Delta_y u_{\ib}^n.
\end{eqnarray}
We pick the final edge states using:
\begin{equation}
u^{\trans}_{\ib-\half\eb_x} =
\begin{cases}
0, & (u_L \le 0 ~ {\rm AND} ~ u_R \ge 0) ~ {\rm OR} ~ |u_L + u_R| < \epsilon, \\
u_L, & u_L + u_R > 0, \\
u_R, & u_L + u_R < 0, \\
\end{cases}
\end{equation}
\begin{equation}
v^{\trans}_{\ib-\half\eb_y} =
\begin{cases}
0, & (v_L \le 0 ~ {\rm AND} ~ v_R \ge 0) ~ {\rm OR} ~ |v_L + v_R| < \epsilon, \\
v_L, & v_L + v_R > 0, \\
v_R, & v_L + v_R < 0. \\
\end{cases}
\end{equation}

For the radial faces, the extrapolation is more complicated due to the radial spatial derivatives in the velocity equation:
\begin{eqnarray}
\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{h}(\wt_{\ib-\eb_r}^n+w_{0,r-1})\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{r-1}, \\
\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{h}(\wt_{\ib}^n+w_{0,r})\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_r.
\end{eqnarray}
Note that $w_0$ is edge-centered, so we define:
\begin{equation}
w_{0,r} = \frac{w_{0,r+\half}+w_{0,r-\half}}{2}.
\end{equation}
The spatial derivatives of $w_0$ are two point differences:
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)_r = \frac{w_{0,r+\half}-w_{0,r-\half}}{h}.
\end{equation}
We pick the final edge states using:
\begin{equation}
\wt^{\trans}_{\ib-\half\eb_r} =
\begin{cases}
0, & (w_L \le 0 ~ {\rm AND} ~ w_R \ge 0) ~ {\rm OR} ~ |w_L + w_R| < \epsilon, \\
\wt_L, & w_L + w_R > 0, \\
\wt_R, & w_L + w_R < 0, \\
\end{cases}
\end{equation}
where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  As mentioned before, we now call {\tt addw0} to convert $\wt^{\trans}$ to $w^{\trans}.$
\section{Computing $\ub^{\mac}$}
In {\tt advance\_premac.f90}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac}$.
\section{Computing $s^{\edge}$}
\section{Computing $\ub^{\edge}$}
\end{document}

### 2008-04-15 docs/Godunov_notes/paper.tex
spherical notes done for utrans

@@
 \def\Ubt   {\widetilde{\bf U}}
+\def\ut    {\tilde{u}}
+\def\vt    {\tilde{v}}
 \def\wt    {\tilde{w}}
@@
 \begin{equation}
-\frac{\partial\Ubt}{\partial t} = -\Ubt\cdot\nabla\Ubt - w_0\frac{\partial\Ubt}{\partial r} - \wt\frac{\partial w_0}{\partial r}\eb_r - \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.
+\frac{\partial\Ubt}{\partial t} = -\Ubt\cdot\nabla\Ubt - w_0\frac{\partial\Ubt}{\partial r} - (\Ubt\cdot\eb_r)\frac{\partial w_0}{\partial r}\eb_r - \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.
 \end{equation}
-We are going to use a 1D Taylor series extrapolation in space and time to compute face-centered normal velocities.
-By 1D, we mean that we omit any spatial derivatives that are not in the direction of the extrapolation.
+We are going to use a 1D Taylor series extrapolation in space and time to compute the normal components of $\Ubt$ at face-centers.
+(Note that for Cartesian problems, $\Ubt = (u,v,\wt)$ and for spherical problems, $\Ubt = (\ut,\vt,\wt)$.)  By 1D, we mean that we omit any spatial derivatives that are not in the direction of the extrapolation.
 We first compute ``left'' and ``right'' states, and then use a Riemann solver to pick the final state.\\
 
-For the x-faces, the extrapolation is as follows:
+For Cartesian problems, for the x-faces, the extrapolation is as follows:
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $\ib = (i,r)$ in 2D and $\ib = (i,j,r)$ in 3D.  For 3D problems only, for the y-faces, the extrapolation follows analogously:
+where $\ib = (i,r)$ in 2D and $\ib = (i,j,r)$ in 3D.  For 3D Cartesian problems, for the y-faces, the extrapolation follows analogously:
 \begin{eqnarray}
@@
 
-For the radial faces, the extrapolation is more complicated due to the radial spatial derivatives in the velocity equation:
+For Cartesian problems, for the radial faces, we must include the radial derivative terms proportional to $w_0$ and $(\Ubt\cdot\eb_r) = \wt$:
 \begin{eqnarray}
-\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{h}(\wt_{\ib-\eb_r}^n+w_{0,r-1})\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{r-1}, \\
-\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{h}(\wt_{\ib}^n+w_{0,r})\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_r.
+\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{h}(\wt_{\ib-\eb_r}^n+w_{0,r-1})\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{r-1},\label{Radial utrans_L} \\
+\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{h}(\wt_{\ib}^n+w_{0,r})\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_r.\label{Radial utrans_R}
 \end{eqnarray}
@@
 \end{equation}
-where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  As mentioned before, we now call {\tt addw0} to convert $\wt^{\trans}$ to $w^{\trans}.$
+where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  As mentioned before, we now call {\tt addw0} to convert $\wt^{\trans}$ to $w^{\trans}.$\\
+
+For spherical problems, we first construct $\Ub$ by mapping $w_0$ onto a Cartesian array and adding it to $\Ubt$.
+Then we use an extrapolation analogous to equations equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}).
+For example, for x-faces:
+
+\begin{eqnarray}
+\ut_L &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_x}\right)\Delta_x\ut_{\ib-\eb_x}^n + \frac{\dt}{2}(\Ubt\cdot\eb_r)_{\ib-\eb_x}\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_x}n_x, \\
+\ut_R &=& \ut_{\ib}^n + \left(\half - \frac{\dt}{h}u_{\ib}\right)\Delta_x\ut_{\ib}^n + \frac{\dt}{2}(\Ubt\cdot\eb_r)_{\ib}\left(\frac{\partial w_0}{\partial r}\right)_{\ib}n_x,
+\end{eqnarray}
+where $n_i = (\eb_r\cdot\eb_i)$.  For the y and z-faces, simply replace each instance of ``x'' with ``y'' or ``z''.
 \section{Computing $\ub^{\mac}$}

### 2008-04-15 docs/Godunov_notes/paper.tex
cleanup utrans section

@@
 \begin{eqnarray}
-\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{h}(\wt_{\ib-\eb_r}^n+w_{0,r-1})\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{r-1},\label{Radial utrans_L} \\
-\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{h}(\wt_{\ib}^n+w_{0,r})\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_r.\label{Radial utrans_R}
+\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
+\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib},\label{Radial utrans_R}
 \end{eqnarray}
-Note that $w_0$ is edge-centered, so we define:
+where the total radial velocity is defined as:
 \begin{equation}
-w_{0,r} = \frac{w_{0,r+\half}+w_{0,r-\half}}{2}.
+w_{\ib}^n = \wt_{\ib}^n + \frac{w_{0,r+\half}+w_{0,r-\half}}{2}
 \end{equation}
@@
 \begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)_r = \frac{w_{0,r+\half}-w_{0,r-\half}}{h}.
+\left(\frac{\partial w_0}{\partial r}\right)_{\ib} = \frac{w_{0,r+\half}-w_{0,r-\half}}{h}.
 \end{equation}
@@
 For spherical problems, we first construct $\Ub$ by mapping $w_0$ onto a Cartesian array and adding it to $\Ubt$.
+We also map $\partial w_0/\partial r$ onto a Cartesian array.
 Then we use an extrapolation analogous to equations equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}).
@@
 \begin{eqnarray}
-\ut_L &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_x}\right)\Delta_x\ut_{\ib-\eb_x}^n + \frac{\dt}{2}(\Ubt\cdot\eb_r)_{\ib-\eb_x}\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_x}n_x, \\
-\ut_R &=& \ut_{\ib}^n + \left(\half - \frac{\dt}{h}u_{\ib}\right)\Delta_x\ut_{\ib}^n + \frac{\dt}{2}(\Ubt\cdot\eb_r)_{\ib}\left(\frac{\partial w_0}{\partial r}\right)_{\ib}n_x,
+\ut_L &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_x}^n\right)\Delta_x\ut_{\ib-\eb_x}^n + \frac{\dt}{2}(\Ubt\cdot\eb_r)_{\ib-\eb_x}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_x}(\eb_r\cdot\eb_x)_{\ib-\eb_x}, \\
+\ut_R &=& \ut_{\ib}^n + \left(\half - \frac{\dt}{h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n + \frac{\dt}{2}(\Ubt\cdot\eb_r)_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}(\eb_r\cdot\eb_x)_{\ib}.
 \end{eqnarray}
-where $n_i = (\eb_r\cdot\eb_i)$.  For the y and z-faces, simply replace each instance of ``x'' with ``y'' or ``z''.
+For the y and z-faces, simply replace each instance of ``x'' with ``y'' or ``z''.  Finally, we call {\tt addw0} to convert $\Ubt^{\trans}$ to $\Ub^{\trans}$.
 \section{Computing $\ub^{\mac}$}

### 2008-04-15 docs/Godunov_notes/paper.tex
more spherical cleanup
typo fix
more details about the differences between spherical and Cartesian

@@
 \end{equation}
-
 For Cartesian problems, for the radial faces, we must include the radial derivative terms proportional to $w_0$ and $(\Ubt\cdot\eb_r) = \wt$:
@@
 \begin{equation}
-w_{\ib}^n = \wt_{\ib}^n + \frac{w_{0,r+\half}+w_{0,r-\half}}{2}
+w_{\ib}^n = \wt_{\ib}^n + \frac{w_{0,r+\half}+w_{0,r-\half}}{2}\label{Total Radial Velocity}
 \end{equation}
@@
 \begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)_{\ib} = \frac{w_{0,r+\half}-w_{0,r-\half}}{h}.
+\left(\frac{\partial w_0}{\partial r}\right)_{\ib} = \frac{w_{0,r+\half}-w_{0,r-\half}}{h}.\label{Radial Derivative of w_0}
 \end{equation}
@@
 For spherical problems, we first construct $\Ub$ by mapping $w_0$ onto a Cartesian array and adding it to $\Ubt$.
+This is a spherical generalization of equation (\ref{Total Radial Velocity}).
 We also map $\partial w_0/\partial r$ onto a Cartesian array.
+This is a spherical generalization of equation (\ref{Radial Derivative of w_0}).
 Then we use an extrapolation analogous to equations equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}).
 For example, for x-faces:
-
 \begin{eqnarray}
-\ut_L &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_x}^n\right)\Delta_x\ut_{\ib-\eb_x}^n + \frac{\dt}{2}(\Ubt\cdot\eb_r)_{\ib-\eb_x}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_x}(\eb_r\cdot\eb_x)_{\ib-\eb_x}, \\
-\ut_R &=& \ut_{\ib}^n + \left(\half - \frac{\dt}{h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n + \frac{\dt}{2}(\Ubt\cdot\eb_r)_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}(\eb_r\cdot\eb_x)_{\ib}.
+\ut_L &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_x}^n\right)\Delta_x\ut_{\ib-\eb_x}^n + \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib-\eb_x}, \\
+\ut_R &=& \ut_{\ib}^n + \left(\half - \frac{\dt}{h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n + \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib}.
 \end{eqnarray}
-For the y and z-faces, simply replace each instance of ``x'' with ``y'' or ``z''.  Finally, we call {\tt addw0} to convert $\Ubt^{\trans}$ to $\Ub^{\trans}$.
+Finally, we call {\tt addw0} to convert $\Ubt^{\trans}$ to $\Ub^{\trans}$.
 \section{Computing $\ub^{\mac}$}

### 2008-04-15 docs/Godunov_notes/paper.tex
cleaned up some of the header stuff
add the forces
Major work on the utrans section; also first cut at the umac section

@@
 \usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{epsfig}
-\usepackage{color}
 
@@
 \usepackage[lmargin=1.0in,rmargin=1.0in,tmargin=0.75in,bmargin=0.75in]{geometry}
-
-\newcommand{\sfrac}[2]{\mathchoice
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
-  {#1\!/#2}}
 
@@
 \def\ib    {{\bf i}}
-\def\ub    {{\bf u}}
 \def\Ub    {{\bf U}}
+\def\er    {{\bf e}_r}
 
@@
 
-\section{Computing $\ub^{\trans}$}
-In {\tt advance\_premac.f90}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\ub^{\trans}$.
-These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0}.
-We only compute the normal component of velocity at each face.\\
+\section{Computing $\Ub^{\trans}$}
+In {\tt advance\_premac.f90}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.
+These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$.
+We will only compute the normal component of velocity at each face.\\
 
@@
 \begin{equation}
-\frac{\partial\Ubt}{\partial t} = -\Ubt\cdot\nabla\Ubt - w_0\frac{\partial\Ubt}{\partial r} - (\Ubt\cdot\eb_r)\frac{\partial w_0}{\partial r}\eb_r - \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.
+\frac{\partial\Ubt}{\partial t} = -\Ubt\cdot\nabla\Ubt - w_0\frac{\partial\Ubt}{\partial r} - (\Ubt\cdot\eb_r)\frac{\partial w_0}{\partial r}\eb_r - \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Perturbational Velocity Equation}
 \end{equation}
-We are going to use a 1D Taylor series extrapolation in space and time to compute the normal components of $\Ubt$ at face-centers.
-(Note that for Cartesian problems, $\Ubt = (u,v,\wt)$ and for spherical problems, $\Ubt = (\ut,\vt,\wt)$.)  By 1D, we mean that we omit any spatial derivatives that are not in the direction of the extrapolation.
-We first compute ``left'' and ``right'' states, and then use a Riemann solver to pick the final state.\\
+We are going to use a 1D Taylor series extrapolation in space and time to compute time-centered normal components of $\Ubt^{\trans}$ at face-centers.
+By 1D, we mean that we omit any spatial derivatives that are not in the direction of the extrapolation.
+We first compute ``left/right'' states, and then use a Riemann solver to pick the final state.\\
+\subsection{2D Cartesian Case}
+For the rest of this paper, for 2D Cartesian problems, $\Ub = (u,w)$ and $\Ubt = (u,\wt)$.  We will use the shorthand $\ib = (i,r)$.\\
 
-For Cartesian problems, for the x-faces, the extrapolation is as follows:
+For the x-faces, the extrapolation is as follows:
 \begin{eqnarray}
 u_L &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
-u_R &=& u_{\ib}^n - \left(\half + \frac{\dt}{h}u_{\ib}^n\right)\Delta_x u_{\ib}^n,
+u_R &=& u_{\ib}^n - \left(\half + \frac{\dt}{h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
 \end{eqnarray}
-where $\ib = (i,r)$ in 2D and $\ib = (i,j,r)$ in 3D.  For 3D Cartesian problems, for the y-faces, the extrapolation follows analogously:
-\begin{eqnarray}
-v_L &=& u_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_y}^n\right)\Delta_y u_{\ib-\eb_y}^n, \\
-v_R &=& u_{\ib}^n - \left(\half + \frac{\dt}{h}u_{\ib}^n\right)\Delta_y u_{\ib}^n.
-\end{eqnarray}
-We pick the final edge states using:
+We pick the final edge states using a Riemann solver:
 \begin{equation}
@@
 \end{equation}
-\begin{equation}
-v^{\trans}_{\ib-\half\eb_y} =
-\begin{cases}
-0, & (v_L \le 0 ~ {\rm AND} ~ v_R \ge 0) ~ {\rm OR} ~ |v_L + v_R| < \epsilon, \\
-v_L, & v_L + v_R > 0, \\
-v_R, & v_L + v_R < 0. \\
-\end{cases}
-\end{equation}
-For Cartesian problems, for the radial faces, we must include the radial derivative terms proportional to $w_0$ and $(\Ubt\cdot\eb_r) = \wt$:
+For the radial faces, we must include the radial derivative terms proportional to $w_0$ and $(\Ubt\cdot\eb_r) = \wt$.
+The term proportional to $w_0$ is included by using $w$ rather than $\wt$ as the advection speed.
+The term proportional to $\wt$ is an additional term:
 \begin{eqnarray}
 \wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
-\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib},\label{Radial utrans_R}
+\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.\label{Radial utrans_R}
 \end{eqnarray}
-where the total radial velocity is defined as:
+The total radial velocity is defined as:
 \begin{equation}
-w_{\ib}^n = \wt_{\ib}^n + \frac{w_{0,r+\half}+w_{0,r-\half}}{2}\label{Total Radial Velocity}
+w_{\ib}^n = \wt_{\ib}^n + \frac{w_{0,r+\half}+w_{0,r-\half}}{2}.\label{Total Radial Velocity}
 \end{equation}
-The spatial derivatives of $w_0$ are two point differences:
+The spatial derivatives of $w_0$ are given by the two-point centered difference:
 \begin{equation}
@@
 \end{equation}
-where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  As mentioned before, we now call {\tt addw0} to convert $\wt^{\trans}$ to $w^{\trans}.$\\
+where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  As mentioned before, we now call {\tt addw0} to convert $\wt^{\trans}$ to $w^{\trans}.$
+\subsection{3D Cartesian Case}
+For the rest of this paper, for 3D Cartesian problems, $\Ub = (u,v,w)$ and $\Ubt = (u,v,\wt)$.  We will use the shorthand $\ib = (i,j,r)$.\\
+
+The only difference between the 2D and 3D Cartesian cases is that we must also compute $v^{\trans}$.  The procedure is analogous to computing $u^{\trans}$:
+\begin{eqnarray}
+v_L &=& v_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{h}v_{\ib-\eb_y}^n\right)\Delta_y v_{\ib-\eb_y}^n, \\
+v_R &=& v_{\ib}^n - \left(\half + \frac{\dt}{h}v_{\ib}^n\right)\Delta_y v_{\ib}^n,
+\end{eqnarray}
+\begin{equation}
+v^{\trans}_{\ib-\half\eb_y} =
+\begin{cases}
+0, & (v_L \le 0 ~ {\rm AND} ~ v_R \ge 0) ~ {\rm OR} ~ |v_L + v_R| < \epsilon, \\
+v_L, & v_L + v_R > 0, \\
+v_R, & v_L + v_R < 0. \\
+\end{cases}
+\end{equation}
+
+\subsection{3D Spherical Case}
+For the rest of this paper, for 3D spherical problems, $\Ub = (u,v,w)$ and $\Ubt = (\ut,\vt,\wt)$.  We will use the shorthand $\ib = (i,j,k)$.\\
 
@@
 \end{eqnarray}
+Then, we use the same Riemann solver as the Cartesian case.
 Finally, we call {\tt addw0} to convert $\Ubt^{\trans}$ to $\Ub^{\trans}$.
-\section{Computing $\ub^{\mac}$}
+To compute $\vt^{\trans}$ or $\wt^{\trans}$, simply replace all instaces of $\ut$ with $\vt$ (or $\wt$) and all instaces of $x$ with $y$ (or $z$).
+\section{Computing $\Ub^{\mac}$}
 In {\tt advance\_premac.f90}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac}$.
+\subsection{2D Cartesian Case}
+Here are the steps to create $u^{\mac}$ for 2D Cartesian problems.  First, we need to compute a time-centered $u$ at transverse faces:
+\begin{eqnarray}
+u_L &=& u_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y u_{\ib-\eb_y}^n, \\
+u_R &=& u_{\ib} - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y u_{\ib}^n.
+\end{eqnarray}
+Then we solve a Riemann problem:
+\begin{equation}
+u_{\ib-\half\eb_y} =
+\begin{cases}
+\half(u_L + u_R), & \left|v^{\trans}_{\ib-\half\eb_y}\right| < \epsilon \\
+u_L, & v^{\trans}_{\ib-\half\eb_y} > 0, \\
+u_R, & v^{\trans}_{\ib-\half\eb_y} < 0. \\
+\end{cases}
+\end{equation}
+Then, we extrapolate the normal velocity to normal faces:
+\begin{eqnarray}
+u_L^{\edge} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{\ib-\eb_x} \nonumber \\
+&& - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\trans}+v_{\ib-\eb_x-\half\eb_y}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_y}+u_{\ib-\eb_x-\half\eb_y}\right), \\
+u_R^{\edge} &=& u_{\ib}^n + \left(\half - \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\trans}+v_{\ib-\half\eb_y}^{\trans}\right)\left(u_{\ib+\half\eb_y}+u_{\ib-\half\eb_y}\right). \\
+\end{eqnarray}
+Finally, we solve a Riemann problem:
+\begin{equation}
+u_{\ib-\half\eb_x}^{\mac} =
+\begin{cases}
+0, & \left(u_L^{\edge} \le 0 ~ {\rm AND} ~ u_R^{\edge} \ge 0\right) ~ {\rm OR} ~ \left|u_L^{\edge} + u_R^{\edge}\right| < \epsilon, \\
+u_L^{\edge}, & u_L^{\edge} + u_R^{\edge} > 0, \\
+u_R^{\edge}, & u_L^{\edge} + u_R^{\edge} < 0. 
+\end{cases}
+\end{equation}
+
+To compute $\vt^{\mac}$, we use an analogous procedure for computing $u^{\mac}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $\partial w_0/\partial r$.
+First, we need to compute a time-centered $\vt$ at transverse faces:
+\begin{eqnarray}
+\vt_L &=& \vt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \vt_{\ib-\eb_x}^n, \\
+\vt_R &=& \vt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \vt_{\ib}^n.
+\end{eqnarray}
+Then we solve a Riemann problem:
+\begin{equation}
+\vt_{\ib-\half\eb_x} =
+\begin{cases}
+\half(\vt_L + \vt_R), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
+\vt_L, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
+\vt_R, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
+\end{cases}
+\end{equation}
+Then, we extrapolate the normal velocity to normal faces, incorporating the extra term proportional to $\partial w_0/\partial r$\footnote{The underbraced terms could be replaced by the average of the neighboring $v^{\trans}$}:
+\begin{eqnarray}
+\vt_L^{\edge} &=& \vt_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}\right)\Delta_y \vt_{\ib-\eb_y}^n + \frac{\dt}{2}f_{\ib-\eb_y} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_y}^{\trans}+u_{\ib-\half\eb_x-\eb_y}^{\trans}\right)\left(\vt_{\ib+\half\eb_x-\eb_y}+\vt_{\ib-\half\eb_x-\eb_y}\right) \nonumber \\
+&& - \frac{\dt}{2h}\underbrace{\vt_{\ib-\eb_y}}_{?}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}), \\
+\vt_R^{\edge} &=& \vt_{\ib}^n + \left(\half - \frac{\dt}{2h}v_{\ib}\right)\Delta_y \vt_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\vt_{\ib+\half\eb_x}+\vt_{\ib-\half\eb_x}\right) \nonumber \\
+&& - \frac{\dt}{2h}\underbrace{\vt_{\ib}}_{?}(w_{0,j+\half} - w_{0,j-\half}).
+\end{eqnarray}
+Finally, we solve a Riemann problem:
+\begin{equation}
+\vt_{\ib-\half\eb_y}^{\mac} =
+\begin{cases}
+0, & \left(v_L^{\edge} \le 0 ~ {\rm AND} ~ v_R^{\edge} \ge 0\right) ~ {\rm OR} ~ \left|v_L^{\edge} + v_R^{\edge}\right| < \epsilon, \\
+\vt_L^{\edge}, & v_L^{\edge} + v_R^{\edge} > 0, \\
+\vt_R^{\edge}, & v_L^{\edge} + v_R^{\edge} < 0. 
+\end{cases}
+\end{equation}
+
 \section{Computing $s^{\edge}$}
-\section{Computing $\ub^{\edge}$}
+
+\section{Computing $\Ub^{\edge}$}
+
+\section{Forces}
+
+\cleardoublepage
+\begin{table*}
+\begin{center}
+\caption{Forcing term into make edge scal \newline}
+\begin{tabular}{|c|c|c|}
+\multicolumn{1}{c}{PRED X}   & \multicolumn{1}{c}{Force} \\
+\hline
+  &   \\
+T $(X)$ & $0$ \\
+  &   \\
+T $(\rho^\prime)$ & $-\rho^\prime \; \nabla \cdot (\ut+w_0 \er) - \nabla \cdot (\ut \rho_0) $  \\
+  &   \\
+\hline
+\end{tabular}
+\end{center}
+\begin{center}
+\begin{tabular}{|c|c|}
+\multicolumn{1}{c}{H PRED TYPE}   & \multicolumn{1}{c}{Force} \\
+\hline
+   & \\
+ 1 $((\rho h)^\prime)$ &  $(\tilde{w} \frac{\partial p_0}{\partial r}) - (\rho h)^\prime \; \nabla \cdot (\ut+w_0 \er) - \nabla \cdot (\ut (\rho h)_0) $  \\
+   & \\
+ 2 $(h)$ & $\frac{1}{\rho} (\psi + \tilde{w} \frac{\partial p_0}{\partial r})$  \\
+   & \\
+ 3 $(T)$ & $\frac{1}{\rho c_p} \left[ (1 - \rho h_p) (\psi + \tilde{w} \frac{\partial p_0}{\partial r}) \right]$ \\
+   & \\
+ 4 $(T)$ & $\frac{1}{\rho c_p} \left[ (1 - \rho h_p) (\psi + \tilde{w} \frac{\partial p_0}{\partial r}) \right]$ \\
+   & \\
+\hline
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
+\begin{table*}
+\begin{center}
+\caption{Quantity that goes into and out of make edge scal \newline}
+\begin{tabular}{|c|c|c|c|}
+\multicolumn{2}{c}{PREDICT} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
+\multicolumn{1}{c}{X} & \multicolumn{1}{c}{h pred type} & \multicolumn{1}{c}{Species } & \multicolumn{1}{c}{h/T } \\
+\hline
+T & 1 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
+T & 2 & $X$ and $\rho^\prime$ & $h$  \\
+T & 3 & $X$ and $\rho^\prime$ & $T$  \\
+T & 4 & $X$ and $\rho^\prime$ & $T$  \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
+\begin{table*}
+\begin{center}
+\caption{When predicting temp ... \newline}
+\begin{tabular}{|c|c|c|c|c|}
+\multicolumn{1}{c}{PRED X} & \multicolumn{1}{c}{type} & \multicolumn{1}{c}{Inputs } & 
+\multicolumn{1}{c}{EOS inputs} & \multicolumn{1}{c}{output} \\
+\hline
+T & 3 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ &  $(\rho h)^\prime = \rho h - (\rho h)_0$ \\
+  &   &                   & $   X = $ given & \\
+  &   &                   &                 & \\
+T & 4 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ & $h$ \\
+  &   &                   & $   X = $ given & \\
+  &   &                   &                 & \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
+\begin{table*}
+\begin{center}
+\caption{Quantity that goes into mkflux on edges \newline}
+\begin{tabular}{|c|c|c|c|c|}
+\multicolumn{2}{c}{PREDICT} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
+\multicolumn{1}{c}{X} & \multicolumn{1}{c}{h pred type} & \multicolumn{1}{c}{Species } & \multicolumn{1}{c}{h/T } & \multicolumn{1}{c}{$\rho^\prime$} \\
+\hline
+T & 1 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ & \\
+T & 2 & $X$ and $\rho^\prime$ & $h$ & \\
+T & 3 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ & \\
+T & 4 & $X$ and $\rho^\prime$ & $h$ & \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
+\begin{table*}
+\begin{center}
+\caption{Quantity that is created in mkflux and passed into update scal\newline}
+\begin{tabular}{|c|c|c|c|c|c|c|c|}
+\multicolumn{2}{c}{PREDICT}   & 
+\multicolumn{1}{c}{} & \multicolumn{1}{c}{}  
+& \multicolumn{2}{c}{Add db?} & \multicolumn{2}{c}{Forcing term }\\
+\multicolumn{1}{c}{X}   & \multicolumn{1}{c}{type} &
+\multicolumn{1}{c}{Species } & \multicolumn{1}{c}{$(\rho h)$} 
+& \multicolumn{1}{c}{X} & \multicolumn{1}{c}{h} & \multicolumn{1}{c}{$(\rho X)$} & \multicolumn{1}{c}{$(\rho h)$} \\
+\hline
+T & 1 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0)(\rho h)^\prime + \ut (\rho h)_0$ & N & Y & 0 & $ (\tilde{w} \frac{\partial p_0}{\partial r}) $ \\
+T & 2 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0) (\rho_0+\rho^\prime) h$ & N & N & 0 & $ (\psi + \tilde{w} \frac{\partial p_0}{\partial r}) $ \\
+T & 3 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0)(\rho h)^\prime + \ut (\rho h)_0$ & N & Y & 0 & $ (\tilde{w} \frac{\partial p_0}{\partial r}) $ \\
+T & 4 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0) (\rho_0+\rho^\prime) h$ & N & N & 0 & $ (\psi + \tilde{w} \frac{\partial p_0}{\partial r}) $ \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
 \end{document}

### 2008-04-15 docs/Godunov_notes/paper.tex
fixed a few typos

@@
 \end{eqnarray}
-Then, we use the same Riemann solver as the Cartesian case.
+Then, we use the same Riemann solver as the Cartesian case to obtain $\ut^{\trans}$.
+To compute $\vt^{\trans}$ or $\wt^{\trans}$, simply replace all instaces of $\ut$ with $\vt$ (or $\wt$) and all instaces of $x$ with $y$ (or $z$).
 Finally, we call {\tt addw0} to convert $\Ubt^{\trans}$ to $\Ub^{\trans}$.
-To compute $\vt^{\trans}$ or $\wt^{\trans}$, simply replace all instaces of $\ut$ with $\vt$ (or $\wt$) and all instaces of $x$ with $y$ (or $z$).
 \section{Computing $\Ub^{\mac}$}
@@
 \end{equation}
-Then, we extrapolate the normal velocity to normal faces, incorporating the extra term proportional to $\partial w_0/\partial r$\footnote{The underbraced terms could be replaced by the average of the neighboring $v^{\trans}$}:
+Then, we extrapolate the normal velocity to normal faces, incorporating the extra term proportional to $\partial w_0/\partial r$\footnote{The underbraced terms could be replaced by the average of the neighboring $v^{\trans}$.
+We still don't know what the best option is.}:
 \begin{eqnarray}

### 2008-04-15 docs/Godunov_notes/paper.tex
fixed some equation errors.  Changed notation to not use u^\edge, but use u^1D instead... it's more intuitive
more clarification on the extra terms needed in radial direction extrapolation

@@
 \subsection{2D Cartesian Case}
-Here are the steps to create $u^{\mac}$ for 2D Cartesian problems.  First, we need to compute a time-centered $u$ at transverse faces:
+Here are the steps to create $u^{\mac}$ for 2D Cartesian problems.  First, we need to compute a time-centered $u$ at transverse faces using a 1D Taylor series extrapolation:
 \begin{eqnarray}
-u_L &=& u_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y u_{\ib-\eb_y}^n, \\
-u_R &=& u_{\ib} - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y u_{\ib}^n.
+u_L^{1D} &=& u_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y u_{\ib-\eb_y}^n, \\
+u_R^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y u_{\ib}^n.
 \end{eqnarray}
@@
 \begin{equation}
-u_{\ib-\half\eb_y} =
+u_{\ib-\half\eb_y}^{1D} =
 \begin{cases}
-\half(u_L + u_R), & \left|v^{\trans}_{\ib-\half\eb_y}\right| < \epsilon \\
-u_L, & v^{\trans}_{\ib-\half\eb_y} > 0, \\
-u_R, & v^{\trans}_{\ib-\half\eb_y} < 0. \\
+\half(u_L^{1D} + u_R^{1D}), & \left|v^{\trans}_{\ib-\half\eb_y}\right| < \epsilon \\
+u_L^{1D}, & v^{\trans}_{\ib-\half\eb_y} > 0, \\
+u_R^{1D}, & v^{\trans}_{\ib-\half\eb_y} < 0. \\
 \end{cases}
@@
 \begin{eqnarray}
-u_L^{\edge} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{\ib-\eb_x} \nonumber \\
-&& - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\trans}+v_{\ib-\eb_x-\half\eb_y}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_y}+u_{\ib-\eb_x-\half\eb_y}\right), \\
-u_R^{\edge} &=& u_{\ib}^n + \left(\half - \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\trans}+v_{\ib-\half\eb_y}^{\trans}\right)\left(u_{\ib+\half\eb_y}+u_{\ib-\half\eb_y}\right). \\
+u_L &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{\ib-\eb_x} \nonumber \\
+&& - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\trans}+v_{\ib-\eb_x-\half\eb_y}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
+u_R &=& u_{\ib}^n + \left(\half - \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\trans}+v_{\ib-\half\eb_y}^{\trans}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right).
 \end{eqnarray}
@@
 \begin{cases}
-0, & \left(u_L^{\edge} \le 0 ~ {\rm AND} ~ u_R^{\edge} \ge 0\right) ~ {\rm OR} ~ \left|u_L^{\edge} + u_R^{\edge}\right| < \epsilon, \\
-u_L^{\edge}, & u_L^{\edge} + u_R^{\edge} > 0, \\
-u_R^{\edge}, & u_L^{\edge} + u_R^{\edge} < 0. 
+0, & \left(u_L \le 0 ~ {\rm AND} ~ u_R \ge 0\right) ~ {\rm OR} ~ \left|u_L + u_R\right| < \epsilon, \\
+u_L, & u_L + u_R > 0, \\
+u_R, & u_L + u_R < 0. 
 \end{cases}
@@
 
-To compute $\vt^{\mac}$, we use an analogous procedure for computing $u^{\mac}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $\partial w_0/\partial r$.
-First, we need to compute a time-centered $\vt$ at transverse faces:
+To compute $\vt^{\mac}$, we use an analogous procedure for computing $u^{\mac}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ and $\partial w_0/\partial r$.
+But first, we need to compute a time-centered $\vt$ at transverse faces using a 1D Taylor seriex extrapolation:
 \begin{eqnarray}
-\vt_L &=& \vt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \vt_{\ib-\eb_x}^n, \\
-\vt_R &=& \vt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \vt_{\ib}^n.
+\vt_L^{1D} &=& \vt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \vt_{\ib-\eb_x}^n, \\
+\vt_R^{1D} &=& \vt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \vt_{\ib}^n.
 \end{eqnarray}
@@
 \begin{equation}
-\vt_{\ib-\half\eb_x} =
+\vt_{\ib-\half\eb_x}^{1D} =
 \begin{cases}
-\half(\vt_L + \vt_R), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
-\vt_L, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
-\vt_R, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
+\half(\vt_L^{1D} + \vt_R^{1D}), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
+\vt_L^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
+\vt_R^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
 \end{cases}
 \end{equation}
-Then, we extrapolate the normal velocity to normal faces, incorporating the extra term proportional to $\partial w_0/\partial r$\footnote{The underbraced terms could be replaced by the average of the neighboring $v^{\trans}$.
+Then, we extrapolate the normal velocity to normal faces, incorporating the extra term proportional to $w_0$ by using $v$ instead of $\vt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term\footnote{The underbraced terms could be replaced by the average of the neighboring $v^{\trans}$.
 We still don't know what the best option is.}:
 \begin{eqnarray}
-\vt_L^{\edge} &=& \vt_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}\right)\Delta_y \vt_{\ib-\eb_y}^n + \frac{\dt}{2}f_{\ib-\eb_y} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_y}^{\trans}+u_{\ib-\half\eb_x-\eb_y}^{\trans}\right)\left(\vt_{\ib+\half\eb_x-\eb_y}+\vt_{\ib-\half\eb_x-\eb_y}\right) \nonumber \\
+\vt_L &=& \vt_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}\right)\Delta_y \vt_{\ib-\eb_y}^n + \frac{\dt}{2}f_{\ib-\eb_y} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_y}^{\trans}+u_{\ib-\half\eb_x-\eb_y}^{\trans}\right)\left(\vt_{\ib+\half\eb_x-\eb_y}^{1D} - \vt_{\ib-\half\eb_x-\eb_y}^{1D}\right) \nonumber \\
 && - \frac{\dt}{2h}\underbrace{\vt_{\ib-\eb_y}}_{?}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}), \\
-\vt_R^{\edge} &=& \vt_{\ib}^n + \left(\half - \frac{\dt}{2h}v_{\ib}\right)\Delta_y \vt_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\vt_{\ib+\half\eb_x}+\vt_{\ib-\half\eb_x}\right) \nonumber \\
+\vt_R &=& \vt_{\ib}^n + \left(\half - \frac{\dt}{2h}v_{\ib}\right)\Delta_y \vt_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\vt_{\ib+\half\eb_x}^{1D} - \vt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
 && - \frac{\dt}{2h}\underbrace{\vt_{\ib}}_{?}(w_{0,j+\half} - w_{0,j-\half}).
@@
 \begin{cases}
-0, & \left(v_L^{\edge} \le 0 ~ {\rm AND} ~ v_R^{\edge} \ge 0\right) ~ {\rm OR} ~ \left|v_L^{\edge} + v_R^{\edge}\right| < \epsilon, \\
-\vt_L^{\edge}, & v_L^{\edge} + v_R^{\edge} > 0, \\
-\vt_R^{\edge}, & v_L^{\edge} + v_R^{\edge} < 0. 
+0, & \left(v_L \le 0 ~ {\rm AND} ~ v_R \ge 0\right) ~ {\rm OR} ~ \left|v_L + v_R\right| < \epsilon, \\
+\vt_L, & v_L + v_R > 0, \\
+\vt_R, & v_L + v_R < 0. 
 \end{cases}

### 2008-04-15 docs/Godunov_notes/paper.tex
In 2D, we use U = (u,w), not U = (u,v).

@@
 In {\tt advance\_premac.f90}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.
-These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$.
+These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$..
 We will only compute the normal component of velocity at each face.\\
@@
 \end{equation}
-where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  As mentioned before, we now call {\tt addw0} to convert $\wt^{\trans}$ to $w^{\trans}.$
+where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  As mentioned before, we now call {\tt addw0} to compute $\wt^{\trans}$ from $w^{\trans}.$
 \subsection{3D Cartesian Case}
@@
 \end{equation}
-
 \subsection{3D Spherical Case}
@@
 To compute $\vt^{\trans}$ or $\wt^{\trans}$, simply replace all instaces of $\ut$ with $\vt$ (or $\wt$) and all instaces of $x$ with $y$ (or $z$).
-Finally, we call {\tt addw0} to convert $\Ubt^{\trans}$ to $\Ub^{\trans}$.
+Finally, we call {\tt addw0} to compute $\Ubt^{\trans}$ fromm $\Ub^{\trans}$.
 \section{Computing $\Ub^{\mac}$}
@@
 \begin{eqnarray}
-u_L^{1D} &=& u_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y u_{\ib-\eb_y}^n, \\
-u_R^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y u_{\ib}^n.
+u_L^{1D} &=& u_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_y}^n\right)\Delta_y u_{\ib-\eb_y}^n, \\
+u_R^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_y u_{\ib}^n.
 \end{eqnarray}
@@
 \begin{cases}
-\half(u_L^{1D} + u_R^{1D}), & \left|v^{\trans}_{\ib-\half\eb_y}\right| < \epsilon \\
-u_L^{1D}, & v^{\trans}_{\ib-\half\eb_y} > 0, \\
-u_R^{1D}, & v^{\trans}_{\ib-\half\eb_y} < 0. \\
+\half(u_L^{1D} + u_R^{1D}), & \left|w^{\trans}_{\ib-\half\eb_y}\right| < \epsilon \\
+u_L^{1D}, & w^{\trans}_{\ib-\half\eb_y} > 0, \\
+u_R^{1D}, & w^{\trans}_{\ib-\half\eb_y} < 0. \\
 \end{cases}
@@
 u_L &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{\ib-\eb_x} \nonumber \\
-&& - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\trans}+v_{\ib-\eb_x-\half\eb_y}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
+&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_y}^{\trans}+w_{\ib-\eb_x-\half\eb_y}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
 u_R &=& u_{\ib}^n + \left(\half - \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\trans}+v_{\ib-\half\eb_y}^{\trans}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right).
+&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_y}^{\trans}+w_{\ib-\half\eb_y}^{\trans}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right).
 \end{eqnarray}
@@
 
-To compute $\vt^{\mac}$, we use an analogous procedure for computing $u^{\mac}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ and $\partial w_0/\partial r$.
-But first, we need to compute a time-centered $\vt$ at transverse faces using a 1D Taylor seriex extrapolation:
+To compute $\wt^{\mac}$, we use an analogous procedure for computing $u^{\mac}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ and $\partial w_0/\partial r$.
+But first, we need to compute a time-centered $\wt$ at transverse faces using a 1D Taylor seriex extrapolation:
 \begin{eqnarray}
-\vt_L^{1D} &=& \vt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \vt_{\ib-\eb_x}^n, \\
-\vt_R^{1D} &=& \vt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \vt_{\ib}^n.
+\wt_L^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
+\wt_R^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \wt_{\ib}^n.
 \end{eqnarray}
@@
 \begin{equation}
-\vt_{\ib-\half\eb_x}^{1D} =
+\wt_{\ib-\half\eb_x}^{1D} =
 \begin{cases}
-\half(\vt_L^{1D} + \vt_R^{1D}), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
-\vt_L^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
-\vt_R^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
+\half(\wt_L^{1D} + \wt_R^{1D}), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
+\wt_L^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
+\wt_R^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
 \end{cases}
 \end{equation}
-Then, we extrapolate the normal velocity to normal faces, incorporating the extra term proportional to $w_0$ by using $v$ instead of $\vt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term\footnote{The underbraced terms could be replaced by the average of the neighboring $v^{\trans}$.
+Then, we extrapolate the normal velocity to normal faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term\footnote{The underbraced terms could be replaced by the average of the neighboring $w^{\trans}$.
 We still don't know what the best option is.}:
 \begin{eqnarray}
-\vt_L &=& \vt_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}\right)\Delta_y \vt_{\ib-\eb_y}^n + \frac{\dt}{2}f_{\ib-\eb_y} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_y}^{\trans}+u_{\ib-\half\eb_x-\eb_y}^{\trans}\right)\left(\vt_{\ib+\half\eb_x-\eb_y}^{1D} - \vt_{\ib-\half\eb_x-\eb_y}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{2h}\underbrace{\vt_{\ib-\eb_y}}_{?}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}), \\
-\vt_R &=& \vt_{\ib}^n + \left(\half - \frac{\dt}{2h}v_{\ib}\right)\Delta_y \vt_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\vt_{\ib+\half\eb_x}^{1D} - \vt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{2h}\underbrace{\vt_{\ib}}_{?}(w_{0,j+\half} - w_{0,j-\half}).
+\wt_L &=& \wt_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_y}\right)\Delta_y \wt_{\ib-\eb_y}^n + \frac{\dt}{2}f_{\ib-\eb_y} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_y}^{\trans}+u_{\ib-\half\eb_x-\eb_y}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_y}^{1D} - \wt_{\ib-\half\eb_x-\eb_y}^{1D}\right) \nonumber \\
+&& - \frac{\dt}{2h}\underbrace{\wt_{\ib-\eb_y}}_{?}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}), \\
+\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{2h}w_{\ib}\right)\Delta_y \wt_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
+&& - \frac{\dt}{2h}\underbrace{\wt_{\ib}}_{?}(w_{0,j+\half} - w_{0,j-\half}).
 \end{eqnarray}
-Finally, we solve a Riemann problem:
+Next, we solve a Riemann problem:
 \begin{equation}
-\vt_{\ib-\half\eb_y}^{\mac} =
+\wt_{\ib-\half\eb_y}^{\mac} =
 \begin{cases}
-0, & \left(v_L \le 0 ~ {\rm AND} ~ v_R \ge 0\right) ~ {\rm OR} ~ \left|v_L + v_R\right| < \epsilon, \\
-\vt_L, & v_L + v_R > 0, \\
-\vt_R, & v_L + v_R < 0. 
+0, & \left(w_L \le 0 ~ {\rm AND} ~ w_R \ge 0\right) ~ {\rm OR} ~ \left|w_L + w_R\right| < \epsilon, \\
+\wt_L, & w_L + w_R > 0, \\
+\wt_R, & w_L + w_R < 0. 
 \end{cases}
 \end{equation}
-
+Finally, we call {\tt addw0} to compute $\Ub^{\mac}$ from $\Ubt^{\trans}$.
+\subsection{3D Cartesian Case}
+\subsection{3D Spherical Case}
 \section{Computing $s^{\edge}$}
-
+\subsection{2D Cartesian Case}
+\subsection{3D Cartesian Case}
+\subsection{3D Spherical Case}
 \section{Computing $\Ub^{\edge}$}
-
+\subsection{2D Cartesian Case}
+\subsection{3D Cartesian Case}
+\subsection{3D Spherical Case}
 \section{Forces}

### 2008-04-15 docs/Godunov_notes/paper.tex
In 2D Cartesian, we use the coordiantes (x,r), not (x,y)

@@
 In {\tt advance\_premac.f90}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.
-These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$..
+These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
 We will only compute the normal component of velocity at each face.\\
@@
 \end{equation}
-where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  As mentioned before, we now call {\tt addw0} to compute $\wt^{\trans}$ from $w^{\trans}.$
+where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  Finally, we call {\tt addw0} to compute $\wt^{\trans}$ from $w^{\trans}.$
 \subsection{3D Cartesian Case}
@@
 \begin{eqnarray}
-u_L^{1D} &=& u_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_y}^n\right)\Delta_y u_{\ib-\eb_y}^n, \\
-u_R^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_y u_{\ib}^n.
+u_L^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
+u_R^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
 \end{eqnarray}
@@
 \begin{equation}
-u_{\ib-\half\eb_y}^{1D} =
+u_{\ib-\half\eb_r}^{1D} =
 \begin{cases}
-\half(u_L^{1D} + u_R^{1D}), & \left|w^{\trans}_{\ib-\half\eb_y}\right| < \epsilon \\
-u_L^{1D}, & w^{\trans}_{\ib-\half\eb_y} > 0, \\
-u_R^{1D}, & w^{\trans}_{\ib-\half\eb_y} < 0. \\
+\half(u_L^{1D} + u_R^{1D}), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
+u_L^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
+u_R^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
 \end{cases}
@@
 u_L &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{\ib-\eb_x} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_y}^{\trans}+w_{\ib-\eb_x-\half\eb_y}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
+&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
 u_R &=& u_{\ib}^n + \left(\half - \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_y}^{\trans}+w_{\ib-\half\eb_y}^{\trans}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right).
+&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right).
 \end{eqnarray}
@@
 \begin{eqnarray}
-\wt_L &=& \wt_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_y}\right)\Delta_y \wt_{\ib-\eb_y}^n + \frac{\dt}{2}f_{\ib-\eb_y} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_y}^{\trans}+u_{\ib-\half\eb_x-\eb_y}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_y}^{1D} - \wt_{\ib-\half\eb_x-\eb_y}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{2h}\underbrace{\wt_{\ib-\eb_y}}_{?}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}), \\
-\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{2h}w_{\ib}\right)\Delta_y \wt_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
+\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{\ib-\eb_r} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
+&& - \frac{\dt}{2h}\underbrace{\wt_{\ib-\eb_r}}_{?}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}), \\
+\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
@@
 \begin{equation}
-\wt_{\ib-\half\eb_y}^{\mac} =
+\wt_{\ib-\half\eb_r}^{\mac} =
 \begin{cases}
@@
 \end{equation}
-Finally, we call {\tt addw0} to compute $\Ub^{\mac}$ from $\Ubt^{\trans}$.
+Finally, we call {\tt addw0} to compute $\Ub^{\mac}$ from $\Ubt^{\mac}$.
 \subsection{3D Cartesian Case}

### 2008-04-15 docs/Godunov_notes/paper.tex
added text about the velocity forcing
more cleanup in the umac section

@@
 \def\eb    {{\bf e}}
+\def\fb    {{\bf f}}
 \def\ib    {{\bf i}}
 \def\Ub    {{\bf U}}
-\def\er    {{\bf e}_r}
 
@@
 
-\section{Computing $\Ub^{\trans}$}
+\section{Computing $\Ubt^{\trans}$}
 In {\tt advance\_premac.f90}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.
@@
 \end{equation}
-where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  Finally, we call {\tt addw0} to compute $\wt^{\trans}$ from $w^{\trans}.$
+where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  Finally, we call {\tt addw0} to compute $w^{\trans}$ from $\wt^{\trans}.$
 \subsection{3D Cartesian Case}
@@
 
-For spherical problems, we first construct $\Ub$ by mapping $w_0$ onto a Cartesian array and adding it to $\Ubt$.
+We first construct $\Ub$ by mapping $w_0$ onto a Cartesian array and adding it to $\Ubt$.
 This is a spherical generalization of equation (\ref{Total Radial Velocity}).
@@
 To compute $\vt^{\trans}$ or $\wt^{\trans}$, simply replace all instaces of $\ut$ with $\vt$ (or $\wt$) and all instaces of $x$ with $y$ (or $z$).
-Finally, we call {\tt addw0} to compute $\Ubt^{\trans}$ fromm $\Ub^{\trans}$.
-\section{Computing $\Ub^{\mac}$}
-In {\tt advance\_premac.f90}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac}$.
+Finally, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
+\section{Computing $\Ubt^{\mac,*}$}
+In {\tt advance\_premac.f90}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac,*}$.  We will only compute the normal component of velocity at each face.\\
 \subsection{2D Cartesian Case}
-Here are the steps to create $u^{\mac}$ for 2D Cartesian problems.  First, we need to compute a time-centered $u$ at transverse faces using a 1D Taylor series extrapolation:
+To compute $u^{\mac,*}$, first we need to compute a time-centered $u$ at transverse faces using a 1D Taylor series extrapolation:
 \begin{eqnarray}
@@
 \begin{eqnarray}
-u_L &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{\ib-\eb_x} \nonumber \\
+u_L &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
 && - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
-u_R &=& u_{\ib}^n + \left(\half - \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right).
+u_R &=& u_{\ib}^n + \left(\half - \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right),
 \end{eqnarray}
+where the force term is given by:\footnote{The code currently does not include the $\pi_0$ term.}
+\begin{equation}
+\fb =
+\left(\begin{array}{c}
+f_u \\
+f_w \\
+\end{array}\right)
+= \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.
+\end{equation}
 Finally, we solve a Riemann problem:
 \begin{equation}
-u_{\ib-\half\eb_x}^{\mac} =
+u_{\ib-\half\eb_x}^{\mac,*} =
 \begin{cases}
@@
 
-To compute $\wt^{\mac}$, we use an analogous procedure for computing $u^{\mac}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ and $\partial w_0/\partial r$.
-But first, we need to compute a time-centered $\wt$ at transverse faces using a 1D Taylor seriex extrapolation:
+To compute $\wt^{\mac,*}$, we use an analogous procedure for computing $u^{\mac,*}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ and $\partial w_0/\partial r$.
+But first, we need to compute a time-centered $\wt$ at transverse faces using a 1D Taylor series extrapolation:
 \begin{eqnarray}
@@
 \end{equation}
-Then, we extrapolate the normal velocity to normal faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term\footnote{The underbraced terms could be replaced by the average of the neighboring $w^{\trans}$.
-We still don't know what the best option is.}:
+Then, we extrapolate the normal velocity to normal faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term:\footnote{The underbraced terms could be replaced by the average of the neighboring $w^{\trans}$.
+We still don't know what the best option is.}
 \begin{eqnarray}
-\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{\ib-\eb_r} \nonumber \\
+\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{w,\ib-\eb_r} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
 && - \frac{\dt}{2h}\underbrace{\wt_{\ib-\eb_r}}_{?}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}), \\
-\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{\ib} \nonumber \\
+\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
@@
 \begin{equation}
-\wt_{\ib-\half\eb_r}^{\mac} =
+\wt_{\ib-\half\eb_r}^{\mac,*} =
 \begin{cases}
@@
 \end{equation}
-Finally, we call {\tt addw0} to compute $\Ub^{\mac}$ from $\Ubt^{\mac}$.
+Finally, we call {\tt addw0} to compute $\Ub^{\mac,*}$ from $\Ubt^{\mac,*}$.
 \subsection{3D Cartesian Case}
 \subsection{3D Spherical Case}
-\section{Computing $s^{\edge}$}
+\section{Computing $\rho^{'\edge}, X_k^{\edge}, h^{\edge},$ and $T^{\edge}$}
 \subsection{2D Cartesian Case}
@@
 \subsection{3D Spherical Case}
-\section{Computing $\Ub^{\edge}$}
+\section{Computing $\Ubt^{\edge}$}
+In {\tt velocity\_advance.f90}, we call {\tt make\_edge\_state} to compute $\Ubt^{\edge}$.
+Unlike when we computed $\Ubt^{\trans}$ and $\Ubt^{\mac,*}$, we need all components of velocity at each face.
 \subsection{2D Cartesian Case}
+To 
 \subsection{3D Cartesian Case}
 \subsection{3D Spherical Case}
+\cleardoublepage
 \section{Forces}
-
-\cleardoublepage
 \begin{table*}
@@
   &   \\
-T $(\rho^\prime)$ & $-\rho^\prime \; \nabla \cdot (\ut+w_0 \er) - \nabla \cdot (\ut \rho_0) $  \\
+T $(\rho^\prime)$ & $-\rho^\prime \; \nabla \cdot (\ut+w_0 \eb_r) - \nabla \cdot (\ut \rho_0) $  \\
   &   \\
@@
    & \\
- 1 $((\rho h)^\prime)$ &  $(\tilde{w} \frac{\partial p_0}{\partial r}) - (\rho h)^\prime \; \nabla \cdot (\ut+w_0 \er) - \nabla \cdot (\ut (\rho h)_0) $  \\
+ 1 $((\rho h)^\prime)$ &  $(\tilde{w} \frac{\partial p_0}{\partial r}) - (\rho h)^\prime \; \nabla \cdot (\ut+w_0 \eb_r) - \nabla \cdot (\ut (\rho h)_0) $  \\
    & \\

### 2008-04-15 docs/Godunov_notes/paper.tex
added sections concerning the differences between notes and code, and section about algorithm questions

@@
 
+\section{Differences Between These Notes and the Code}
+\begin{enumerate}
+\item In both equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}), we are omitting the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ (by using $\wt$ rather than $w$ as the advection speed) and $(\Ubt\cdot\eb_r)$ (by omitting the term altogether).
+\item In equation (\ref{Force Term}), we do not include the term proportional to $\partial\pi_0/\partial r$.
+\end{enumerate}
+\section{Questions}
+\begin{enumerate}
+\item We need to re-examine when to include the term proportional to $(\Ubt\cdot\eb_r)$ in equation (\ref{Perturbational Velocity Equation}) when we do a 1D extrapolation of the radial velocity.
+For example, equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}) include it, but the equations for the 1D extrapolation in the radial direction used for computing $\Ubt^{\edge}$ do not.
+\item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), we could also compute the term proportional to $\partial w_0/\partial r$ by averaging $w^{\trans}$.
+\end{enumerate}
 \section{Computing $\Ubt^{\trans}$}
@@
 \end{equation}
-Then, we extrapolate the normal velocity to normal faces:
+Then, we extrapolate the $u$ to normal faces:
 \begin{eqnarray}
@@
 \end{eqnarray}
-where the force term is given by:\footnote{The code currently does not include the $\pi_0$ term.}
+where the force term is given by:
 \begin{equation}
@@
 \end{array}\right)
-= \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.
+= \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Force Term}
 \end{equation}
@@
 \end{equation}
-Then, we extrapolate the normal velocity to normal faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term:\footnote{The underbraced terms could be replaced by the average of the neighboring $w^{\trans}$.
-We still don't know what the best option is.}
+Then, we extrapolate $\wt$ to normal faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term:
 \begin{eqnarray}
@@
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{2h}\underbrace{\wt_{\ib-\eb_r}}_{?}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}), \\
+&& - \frac{\dt}{2h}\wt_{\ib-\eb_r}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}),\label{wt_L Extrapolation} \\
 \wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{2h}\underbrace{\wt_{\ib}}_{?}(w_{0,j+\half} - w_{0,j-\half}).
+&& - \frac{\dt}{2h}\wt_{\ib}(w_{0,j+\half} - w_{0,j-\half}).\label{wt_R Extrapolation}
 \end{eqnarray}

### 2008-04-16 docs/Godunov_notes/paper.tex
lots of work done on Uedge.

@@
 \item We need to re-examine when to include the term proportional to $(\Ubt\cdot\eb_r)$ in equation (\ref{Perturbational Velocity Equation}) when we do a 1D extrapolation of the radial velocity.
-For example, equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}) include it, but the equations for the 1D extrapolation in the radial direction used for computing $\Ubt^{\edge}$ do not.
+For example, equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}) include it, but equations (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}) do not.
 \item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), we could also compute the term proportional to $\partial w_0/\partial r$ by averaging $w^{\trans}$.
@@
 
-For the x-faces, the extrapolation is as follows:
+For the x-faces, the 1D extrapolation of $u$ is:
 \begin{eqnarray}
@@
 The term proportional to $w_0$ is included by using $w$ rather than $\wt$ as the advection speed.
-The term proportional to $\wt$ is an additional term:
+The term proportional to $\wt$ is an additional term.
+For the radial faces, the 1D extrapolation of $\wt$ is:
 \begin{eqnarray}
@@
 
-The only difference between the 2D and 3D Cartesian cases is that we must also compute $v^{\trans}$.  The procedure is analogous to computing $u^{\trans}$:
+The only difference between the 2D and 3D Cartesian cases is that we must also compute $v^{\trans}$.
+The procedure is analogous to computing $u^{\trans}$.
+For y-faces, the 1D extrapolation of $v$ is:
 \begin{eqnarray}
@@
 This is a spherical generalization of equation (\ref{Radial Derivative of w_0}).
-Then we use an extrapolation analogous to equations equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}).
+Then we use a 1D extrapolation analogous to equations equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}).
 For example, for x-faces:
@@
 \subsection{2D Cartesian Case}
-To compute $u^{\mac,*}$, first we need to compute a time-centered $u$ at transverse faces using a 1D Taylor series extrapolation:
+We begin by computing a time-centered 1D extrapolation of $u$ to transverse faces:
 \begin{eqnarray}
@@
 \end{eqnarray}
-Then we solve a Riemann problem:
+Then we upwind based on $w^{\trans}$:
 \begin{equation}
@@
 \end{equation}
-Then, we extrapolate the $u$ to normal faces:
+Then, we do a full-dimensional extrapolation of $u$ to normal faces:
 \begin{eqnarray}
@@
 
-To compute $\wt^{\mac,*}$, we use an analogous procedure for computing $u^{\mac,*}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ and $\partial w_0/\partial r$.
-But first, we need to compute a time-centered $\wt$ at transverse faces using a 1D Taylor series extrapolation:
+To compute $\wt^{\mac,*}$, we use an analogous procedure for computing $u^{\mac,*}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ and $\partial w_0/\partial r$ into the full-dimensional extrapolation.
+We compute a time-centered 1D extrapolation of $\wt$ to transverse faces:
 \begin{eqnarray}
@@
 \end{eqnarray}
-Then we solve a Riemann problem:
+Then we upwind based on $u^{\trans}$:
 \begin{equation}
@@
 \end{equation}
-Then, we extrapolate $\wt$ to normal faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term:
+Then, we do a full-dimensional extrapolation of $\wt$ to normal faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term:
 \begin{eqnarray}
@@
 \subsection{2D Cartesian Case}
-To 
+We begin by computing a time-centered 1D extrapolation of $u$ to all faces.  The extrapolation of $u$ to x-faces is:
+\begin{eqnarray}
+u_{L,\ib-\half\eb_x}^{1D} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
+u_{R,\ib-\half\eb_x}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_r u_{\ib}^n.
+\end{eqnarray}
+Then we upwind based on $u^{\trans}$:
+\begin{equation}
+u_{\ib-\half\eb_x}^{1D} =
+\begin{cases}
+\half(u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
+u_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
+u_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
+\end{cases}
+\end{equation}
+The extrapolation of $u$ to radial faces is:
+\begin{eqnarray}
+u_{L,\ib-\half\eb_r}^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
+u_{R,\ib-\half\eb_r}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
+\end{eqnarray}
+Then we upwind based on $w^{\trans}$:
+\begin{equation}
+u_{\ib-\half\eb_r}^{1D} =
+\begin{cases}
+\half(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
+u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
+u_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
+\end{cases}
+\end{equation}
+We repeat this procedure to compute a time-centered 1D extrapolation of $\wt$ at each face.  The extrapolation of $\wt$ to x-faces is:
+\begin{eqnarray}
+\wt_{L,\ib-\half\eb_x}^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
+\wt_{R,\ib-\half\eb_x}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_r \wt_{\ib}^n.
+\end{eqnarray}
+Then we upwind based on $u^{\trans}$:
+\begin{equation}
+\wt_{\ib-\half\eb_x}^{1D} =
+\begin{cases}
+\half(\wt_{L,\ib-\half\eb_x}^{1D} + \wt_{R,\ib-\half\eb_x}^{1D}), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
+\wt_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
+\wt_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
+\end{cases}
+\end{equation}
+The extrapolation of $\wt$ to radial faces is:
+\begin{eqnarray}
+\wt_{L,\ib-\half\eb_r}^{1D} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n, \label{1D wt_L extrap} \\
+\wt_{R,\ib-\half\eb_r}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r wt_{\ib}^n. \label{1D wt_R extrap}
+\end{eqnarray}
+Then we upwind based on $w^{\trans}$:
+\begin{equation}
+\wt_{\ib-\half\eb_r}^{1D} =
+\begin{cases}
+\half(\wt_{L,\ib-\half\eb_r}^{1D} + \wt_{R,\ib-\half\eb_r}^{1D}), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
+\wt_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
+\wt_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
+\end{cases}
+\end{equation}
+Now we do a full-dimensional extrapolation of $u$ to each face.  The extrapolation of $u$ to x-faces is:
+\begin{eqnarray}
+u_{L,\ib-\half\eb_x}^{\edge} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
+&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
+u_{R,\ib-\half\eb_x}^{\edge} &=& u_{\ib}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right).
+\end{eqnarray}
+Then, we upwind based on $u^{\mac}$.
+\begin{equation}
+u_{\ib-\half\eb_x}^{\edge} =
+\begin{cases}
+\half(u_{L,\ib-\half\eb_x}^{\edge} + u_{R,\ib-\half\eb_x}^{\edge}), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+u_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
+u_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
+\end{cases}
+\end{equation}
+The full-dimensional extrapolation of $u$ to radial faces is:
+\begin{eqnarray}
+u_{L,\ib-\half\eb_r}^{\edge} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib-\eb_r}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_x-\eb_r}^{1D} - u_{\ib-\half\eb_x-\eb_r}^{1D}\right),  \\
+u_{R,\ib-\half\eb_r}^{\edge} &=& u_{\ib}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(u_{\ib+\half\eb_x}^{1D} - u_{\ib-\half\eb_x}^{1D}\right). 
+\end{eqnarray}
+Then, we upwind based on $w^{\mac}$.
+\begin{equation}
+u_{\ib-\half\eb_r}^{\edge} =
+\begin{cases}
+\half(u_{L,\ib-\half\eb_r}^{\edge} + u_{R,\ib-\half\eb_r}^{\edge}), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
+u_{L,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} > 0, \\
+u_{R,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} < 0.
+\end{cases}
+\end{equation}
+Now we do a full-dimensional extrapolatino of $\wt$ to each face.
+This is different from the extrapolation of $u$ to faces because we must include the term proportional to $\partial w_0/\partial r$ in equation (\ref{Perturbational Velocity Equation}).
+The full-dimensional extrapolation of $\wt$ to x-faces is:
+\begin{eqnarray}
+\wt_{L,\ib-\half\eb_x}^{\edge} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib-\eb_x}^n + \frac{\dt}{2}f_{w,\ib-\eb_x} \nonumber \\
+&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(\wt_{\ib-\eb_x+\half\eb_r}^{1D} - \wt_{\ib-\eb_x-\half\eb_r}^{1D}\right) \nonumber \\
+&& - \frac{\dt}{4h}\left(\wt_{\ib-\eb_x+\half\eb_r}^{\trans}+\wt_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_x}-w_{0,\ib-\half\eb_x}\right), \\
+\wt_{R,\ib-\half\eb_x}^{\edge} &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right) \nonumber \\
+&& - \frac{\dt}{4h}\left(\wt_{\ib-\eb_x+\half\eb_r}^{\trans}+\wt_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_x}-w_{0,\ib-\half\eb_x}\right). \\
+\end{eqnarray}
+Then, we upwind based on $u^{\mac}$:
+\begin{equation}
+\wt_{\ib-\half\eb_x}^{\edge} =
+\begin{cases}
+\half(\wt_{L,\ib-\half\eb_x}^{\edge} + \wt_{R,\ib-\half\eb_x}^{\edge}), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+\wt_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
+\wt_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
+\end{cases}
+\end{equation}
+The full-dimensional extrapolation of $\wt$ to radial faces is:
+
 \subsection{3D Cartesian Case}

### 2008-04-16 docs/Godunov_notes/paper.tex
add some MarginPars for Andy

@@
 
-\usepackage{amsmath}
+\usepackage{amsmath,color}
 
@@
 \title{Notes on the Godunov Step in {\tt MAESTRO}}
+
+\setlength{\marginparwidth}{0.75in}
+\newcommand{\MarginPar}[1]{\marginpar{%
+\vskip-\baselineskip %raise the marginpar a bit
+\raggedright\tiny\sffamily
+\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
 
@@
 u_R &=& u_{\ib}^n - \left(\half + \frac{\dt}{h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
-\end{eqnarray}
+\end{eqnarray} \MarginPar{should be $\dt/2h$}
 We pick the final edge states using a Riemann solver:
@@
 \wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.\label{Radial utrans_R}
-\end{eqnarray}
+\end{eqnarray} \MarginPar{should be $\dt / 2h$}
 The total radial velocity is defined as:
@@
 v_R &=& v_{\ib}^n - \left(\half + \frac{\dt}{h}v_{\ib}^n\right)\Delta_y v_{\ib}^n,
-\end{eqnarray}
+\end{eqnarray} \MarginPar{should be $\dt / 2h$}
 \begin{equation}
@@
 \ut_R &=& \ut_{\ib}^n + \left(\half - \frac{\dt}{h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n + \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib}.
-\end{eqnarray}
+\end{eqnarray} \MarginPar{should be $\dt / 2h$}
 Then, we use the same Riemann solver as the Cartesian case to obtain $\ut^{\trans}$.
@@
 Finally, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
+
 \section{Computing $\Ubt^{\mac,*}$}
@@
 \hline
-\hline
 \end{tabular}

### 2008-04-16 docs/Godunov_notes/paper.tex
fixed Mike's MarinPars
fix typo (should be dt/2h not dt/2  h)
fixed many more equation typos and errors

--- 2008-04-16 docs/Godunov_notes/paper.tex
@@
 
+\setlength{\marginparwidth}{0.75in}
+\newcommand{\MarginPar}[1]{\marginpar{%
+\vskip-\baselineskip %raise the marginpar a bit
+\raggedright\tiny\sffamily
+\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
+
 \begin{document}
+
 \maketitle
@@
 These are working notes for the Godunov step in {\tt MAESTRO}.
-
 \section{Differences Between These Notes and the Code}
@@
 By 1D, we mean that we omit any spatial derivatives that are not in the direction of the extrapolation.
-We first compute ``left/right'' states, and then use a Riemann solver to pick the final state.\\
+We first compute ``left/right'' states, and then use a Riemann solver to pick the final state.
 \subsection{2D Cartesian Case}
-For the rest of this paper, for 2D Cartesian problems, $\Ub = (u,w)$ and $\Ubt = (u,\wt)$.  We will use the shorthand $\ib = (i,r)$.\\
+For the rest of this paper, for 2D Cartesian problems, $\Ub = (u,w)$ and $\Ubt = (u,\wt)$.  We will use the shorthand $\ib = (x,r)$.\\
 
@@
 \begin{eqnarray}
-u_L &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
-u_R &=& u_{\ib}^n - \left(\half + \frac{\dt}{h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
+u_L &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
+u_R &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
 \end{eqnarray}
@@
 \end{equation}
-For the radial faces, we must include the radial derivative terms proportional to $w_0$ and $(\Ubt\cdot\eb_r) = \wt$.
+For the r-faces, we choose include the radial derivative terms proportional to $w_0$ and $(\Ubt\cdot\eb_r) = \wt$.
 The term proportional to $w_0$ is included by using $w$ rather than $\wt$ as the advection speed.
 The term proportional to $\wt$ is an additional term.
-For the radial faces, the 1D extrapolation of $\wt$ is:
+For the r-faces, the 1D extrapolation of $\wt$ is:
 \begin{eqnarray}
-\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
-\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.\label{Radial utrans_R}
+\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
+\wt_R &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.\label{Radial utrans_R}
 \end{eqnarray}
@@
 \end{equation}
-where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  Finally, we call {\tt addw0} to compute $w^{\trans}$ from $\wt^{\trans}.$
+where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  Finally, we call {\tt addw0} to compute $w^{\trans}$ from $\wt^{\trans}$.
 \subsection{3D Cartesian Case}
-For the rest of this paper, for 3D Cartesian problems, $\Ub = (u,v,w)$ and $\Ubt = (u,v,\wt)$.  We will use the shorthand $\ib = (i,j,r)$.\\
+For the rest of this paper, for 3D Cartesian problems, $\Ub = (u,v,w)$ and $\Ubt = (u,v,\wt)$.  We will use the shorthand $\ib = (x,y,r)$.\\
 
@@
 \begin{eqnarray}
-v_L &=& v_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{h}v_{\ib-\eb_y}^n\right)\Delta_y v_{\ib-\eb_y}^n, \\
-v_R &=& v_{\ib}^n - \left(\half + \frac{\dt}{h}v_{\ib}^n\right)\Delta_y v_{\ib}^n,
+v_L &=& v_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y v_{\ib-\eb_y}^n, \\
+v_R &=& v_{\ib}^n - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y v_{\ib}^n,
 \end{eqnarray}
@@
 \subsection{3D Spherical Case}
-For the rest of this paper, for 3D spherical problems, $\Ub = (u,v,w)$ and $\Ubt = (\ut,\vt,\wt)$.  We will use the shorthand $\ib = (i,j,k)$.\\
+For the rest of this paper, for 3D spherical problems, $\Ub = (u,v,w)$ and $\Ubt = (\ut,\vt,\wt)$.  We will use the shorthand $\ib = (x,y,z)$.\\
 
@@
 \begin{eqnarray}
-\ut_L &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_x}^n\right)\Delta_x\ut_{\ib-\eb_x}^n + \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib-\eb_x}, \\
-\ut_R &=& \ut_{\ib}^n + \left(\half - \frac{\dt}{h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n + \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib}.
+\ut_L &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x\ut_{\ib-\eb_x}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib-\eb_x}, \\
+\ut_R &=& \ut_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib}.
 \end{eqnarray}
 Then, we use the same Riemann solver as the Cartesian case to obtain $\ut^{\trans}$.
-To compute $\vt^{\trans}$ or $\wt^{\trans}$, simply replace all instaces of $\ut$ with $\vt$ (or $\wt$) and all instaces of $x$ with $y$ (or $z$).
+To compute $\vt^{\trans}$ (or $\wt^{\trans}$), simply replace all instaces of $\ut$ with $\vt$ (or $\wt$) and all instaces of $x$ with $y$ (or $z$).
 Finally, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
 \section{Computing $\Ubt^{\mac,*}$}
-In {\tt advance\_premac.f90}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac,*}$.  We will only compute the normal component of velocity at each face.\\
+In {\tt advance\_premac.f90}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac,*}$.  We will only compute the normal component of velocity at each face.
 \subsection{2D Cartesian Case}
-We begin by computing a time-centered 1D extrapolation of $u$ to transverse faces:
+We begin by computing a time-centered 1D extrapolation of $u$ to r-faces:
 \begin{eqnarray}
@@
 \end{equation}
-Then, we do a full-dimensional extrapolation of $u$ to normal faces:
+Then, we do a full-dimensional extrapolation of $u$ to x-faces:
 \begin{eqnarray}
@@
 && - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
-u_R &=& u_{\ib}^n + \left(\half - \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
+u_R &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right),
@@
 To compute $\wt^{\mac,*}$, we use an analogous procedure for computing $u^{\mac,*}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ and $\partial w_0/\partial r$ into the full-dimensional extrapolation.
-We compute a time-centered 1D extrapolation of $\wt$ to transverse faces:
+We compute a time-centered 1D extrapolation of $\wt$ to x-faces:
 \begin{eqnarray}
@@
 \end{equation}
-Then, we do a full-dimensional extrapolation of $\wt$ to normal faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term:
+Then, we do a full-dimensional extrapolation of $\wt$ to r-faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term:
 \begin{eqnarray}
@@
 && - \frac{\dt}{2h}\wt_{\ib-\eb_r}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}),\label{wt_L Extrapolation} \\
-\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
+\wt_R &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
@@
 u_{L,\ib-\half\eb_x}^{1D} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
-u_{R,\ib-\half\eb_x}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_r u_{\ib}^n.
+u_{R,\ib-\half\eb_x}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
 \end{eqnarray}
@@
 \end{equation}
-The extrapolation of $u$ to radial faces is:
+The extrapolation of $u$ to r-faces is:
 \begin{eqnarray}
@@
 \wt_{L,\ib-\half\eb_x}^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
-\wt_{R,\ib-\half\eb_x}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_r \wt_{\ib}^n.
+\wt_{R,\ib-\half\eb_x}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \wt_{\ib}^n.
 \end{eqnarray}
@@
 \end{equation}
-The extrapolation of $\wt$ to radial faces is:
+The extrapolation of $\wt$ to r-faces is:
 \begin{eqnarray}
 \wt_{L,\ib-\half\eb_r}^{1D} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n, \label{1D wt_L extrap} \\
-\wt_{R,\ib-\half\eb_r}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r wt_{\ib}^n. \label{1D wt_R extrap}
+\wt_{R,\ib-\half\eb_r}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n. \label{1D wt_R extrap}
 \end{eqnarray}
@@
 && - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
-u_{R,\ib-\half\eb_x}^{\edge} &=& u_{\ib}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
+u_{R,\ib-\half\eb_x}^{\edge} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right).
@@
 \end{equation}
-The full-dimensional extrapolation of $u$ to radial faces is:
+The full-dimensional extrapolation of $u$ to r-faces is:
 \begin{eqnarray}
-u_{L,\ib-\half\eb_r}^{\edge} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib-\eb_r}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
+u_{L,\ib-\half\eb_r}^{\edge} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib-\eb_r}^n + \frac{\dt}{2}f_{u,\ib-\eb_r} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_x-\eb_r}^{1D} - u_{\ib-\half\eb_x-\eb_r}^{1D}\right),  \\
-u_{R,\ib-\half\eb_r}^{\edge} &=& u_{\ib}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
+u_{R,\ib-\half\eb_r}^{\edge} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(u_{\ib+\half\eb_x}^{1D} - u_{\ib-\half\eb_x}^{1D}\right). 
@@
 && - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(\wt_{\ib-\eb_x+\half\eb_r}^{1D} - \wt_{\ib-\eb_x-\half\eb_r}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{4h}\left(\wt_{\ib-\eb_x+\half\eb_r}^{\trans}+\wt_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_x}-w_{0,\ib-\half\eb_x}\right), \\
-\wt_{R,\ib-\half\eb_x}^{\edge} &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{4h}\left(\wt_{\ib-\eb_x+\half\eb_r}^{\trans}+\wt_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_x}-w_{0,\ib-\half\eb_x}\right). \\
+&& - \frac{\dt}{4h}\left(\wt_{\ib-\eb_x+\half\eb_r}^{\trans}+\wt_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(w_{0,\ib-\eb_x+\half\eb_r}-w_{0,\ib-\eb_x-\half\eb_r}\right), \\
+\wt_{R,\ib-\half\eb_x}^{\edge} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_r}^{1D} - \wt_{\ib-\half\eb_r}^{1D}\right) \nonumber \\
+&& - \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right).
 \end{eqnarray}
@@
 \end{equation}
-The full-dimensional extrapolation of $\wt$ to radial faces is:
+The full-dimensional extrapolation of $\wt$ to r-faces is:
 
@@
    & \\
-\hline
 \hline

### 2008-04-17 docs/Godunov_notes/paper.tex
more 2D Cartesian Uedge stuff
Finished the U^EDGE section.  Needs proofreading.
finished the 2D Cartesian s^edge section

@@
 \section{Computing $\Ubt^{\trans}$}
-In {\tt advance\_premac.f90}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.
+In {\tt advance\_premac}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.
 These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
@@
 \section{Computing $\Ubt^{\mac,*}$}
-In {\tt advance\_premac.f90}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac,*}$.  We will only compute the normal component of velocity at each face.
+In {\tt advance\_premac}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac,*}$.  We will only compute the normal component of velocity at each face.\\
+
+The force terms are:
+\begin{equation}
+\fb =
+\left(\begin{array}{c}
+f_u \\
+f_w \\
+\end{array}\right)
+= \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Force Term}
+\end{equation}
 \subsection{2D Cartesian Case}
@@
 u_R &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right),
+&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right).
 \end{eqnarray}
-where the force term is given by:
-\begin{equation}
-\fb =
-\left(\begin{array}{c}
-f_u \\
-f_w \\
-\end{array}\right)
-= \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Force Term}
-\end{equation}
 Finally, we solve a Riemann problem:
@@
 \section{Computing $\rho^{'\edge}, X_k^{\edge}, h^{\edge},$ and $T^{\edge}$}
+In {\tt scalar\_advance}, we call {\tt make\_edge\_scal} to compute $\rho^{'\edge}, X_k^{\edge}, h^{\edge},$ and $T^{\edge}$ at each edge.
+The procedure is the same for each scalar, so we shall simply denote the scalar as $s$.
+\\
+
+The force terms are:
+\begin{eqnarray}
+f_{\rho'} &=& -\rho'\nabla\cdot\Ub^{\mac} - \nabla\cdot(\rho_0\Ubt^{\mac}), \\
+f_{X_k} &=& 0, \\
+f_h &=& \frac{1}{\rho}\left(\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} \right), \\
+f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
+\end{eqnarray}
 \subsection{2D Cartesian Case}
+We begin by computing a time-centered 1D extrapolation of $s$ to all faces.  The extrapolation of $s$ to x-faces is:
+\begin{eqnarray}
+s_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib-\eb_x}^n, \\
+s_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib}^n.
+\end{eqnarray}
+Then we upwind based on $u^{\mac}$:
+\begin{equation}
+s_{\ib-\half\eb_x}^{1D} =
+\begin{cases}
+\half(s_{L,\ib-\half\eb_x}^{1D} + s_{R,\ib-\half\eb_x}^{1D}), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+s_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
+s_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
+\end{cases}
+\end{equation}
+The extrapolation of $u$ to r-faces is completely analogous to the x-faces:
+\begin{eqnarray}
+s_{L,\ib-\half\eb_r}^{1D} &=& s_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib-\eb_r}^n, \\
+s_{R,\ib-\half\eb_r}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib}^n.
+\end{eqnarray}
+Then we upwind based on $w^{\mac}$:
+\begin{equation}
+s_{\ib-\half\eb_r}^{1D} =
+\begin{cases}
+\half(s_{L,\ib-\half\eb_r}^{1D} + s_{R,\ib-\half\eb_r}^{1D}), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
+u_{L,\ib-\half\eb_r}^{1D}, & w^{\mac}_{\ib-\half\eb_r} > 0, \\
+u_{R,\ib-\half\eb_r}^{1D}, & w^{\mac}_{\ib-\half\eb_r} < 0. \\
+\end{cases}
+\end{equation}
+Now we do a full-dimensional extrapolation of $s$ to each face.  The extrapolation of $s$ to x-faces is:
+\begin{eqnarray}
+s_{L,\ib-\half\eb_x}^{\edge} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib-\eb_x}^n + \frac{\dt}{2}f_{s,\ib-\eb_x} \nonumber \\
+&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\mac}+w_{\ib-\eb_x-\half\eb_r}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_r}^{1D} - s_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
+s_{R,\ib-\half\eb_x}^{\edge} &=& s_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib}^n + \frac{\dt}{2}f_{s,\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\mac}+w_{\ib-\half\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_r}^{1D} - s_{\ib-\half\eb_r}^{1D}\right).
+\end{eqnarray}
+Then, we upwind based on $u^{\mac}$.
+\begin{equation}
+s_{\ib-\half\eb_x}^{\edge} =
+\begin{cases}
+\half(s_{L,\ib-\half\eb_x}^{\edge} + s_{R,\ib-\half\eb_x}^{\edge}), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+s_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
+s_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
+\end{cases}
+\end{equation}
+The full-dimensional extrapolation of $s$ to r-faces is completely analogous to the x-faces:
+\begin{eqnarray}
+s_{L,\ib-\half\eb_r}^{\edge} &=& s_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib-\eb_r}^n + \frac{\dt}{2}f_{s,\ib-\eb_r} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\mac}+u_{\ib-\half\eb_x-\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_x-\eb_r}^{1D} - s_{\ib-\half\eb_x-\eb_r}^{1D}\right),  \\
+s_{R,\ib-\half\eb_r}^{\edge} &=& s_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib}^n + \frac{\dt}{2}f_{s,\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\mac}+u_{\ib-\half\eb_x}^{\mac}\right)\left(s_{\ib+\half\eb_x}^{1D} - s_{\ib-\half\eb_x}^{1D}\right). 
+\end{eqnarray}
+Then, we upwind based on $w^{\mac}$.
+\begin{equation}
+s_{\ib-\half\eb_r}^{\edge} =
+\begin{cases}
+\half(s_{L,\ib-\half\eb_r}^{\edge} + s_{R,\ib-\half\eb_r}^{\edge}), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
+s_{L,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} > 0, \\
+s_{R,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} < 0.
+\end{cases}
+\end{equation}
 \subsection{3D Cartesian Case}
@@
 \section{Computing $\Ubt^{\edge}$}
-In {\tt velocity\_advance.f90}, we call {\tt make\_edge\_state} to compute $\Ubt^{\edge}$.
-Unlike when we computed $\Ubt^{\trans}$ and $\Ubt^{\mac,*}$, we need all components of velocity at each face.
+In {\tt velocity\_advance}, we call {\tt make\_edge\_state} to compute $\Ubt^{\edge}$.  Unlike when we computed $\Ubt^{\trans}$ and $\Ubt^{\mac,*}$, we need all components of velocity at each face.
 \subsection{2D Cartesian Case}
@@
 The full-dimensional extrapolation of $\wt$ to r-faces is:
-
+\begin{eqnarray}
+\wt_{L,\ib-\half\eb_r}^{\edge} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{w,\ib-\eb_r} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
+&& -
+\frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)\left(w_{0,\ib-\half\eb_r}-w_{0,\ib-\frac{3}{2}\eb_r}\right),\\
+\wt_{R,\ib-\half\eb_r}^{\edge} &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
+&& -
+\frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib+\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right).
+\end{eqnarray}
+Then, we upwind based on $w^{\mac}$:
+\begin{equation}
+\wt_{\ib-\half\eb_r}^{\edge} =
+\begin{cases}
+\half(\wt_{L,\ib-\half\eb_r}^{\edge} + \wt_{R,\ib-\half\eb_r}^{\edge}), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
+\wt_{L,\ib-\half\eb_r}^{\edge}, & w^{\mac}_{\ib-\half\eb_r} > 0, \\
+\wt_{R,\ib-\half\eb_r}^{\edge}, & w^{\mac}_{\ib-\half\eb_r} < 0.
+\end{cases}
+\end{equation}
 \subsection{3D Cartesian Case}

### 2008-04-17 docs/Godunov_notes/paper.tex
added table of contents, and sections on the VARDEN integrators

@@
 
-\title{Notes on the Godunov Step in {\tt MAESTRO}}
+\title{Notes on the Godunov Step in {\tt MAESTRO} and {\tt VARDEN}}
 
@@
 \maketitle
+\tableofcontents
+\cleardoublepage
 
@@
 \end{enumerate}
-\section{Computing $\Ubt^{\trans}$}
+\cleardoublepage
+\section{Computing $\Ubt^{\trans}$ in {\tt MAESTRO}}
 In {\tt advance\_premac}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.
@@
 Finally, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
-\section{Computing $\Ubt^{\mac,*}$}
+\cleardoublepage
+\section{Computing $\Ubt^{\mac,*}$ in {\tt MAESTRO}}
 In {\tt advance\_premac}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac,*}$.  We will only compute the normal component of velocity at each face.\\
@@
 \subsection{3D Spherical Case}
-\section{Computing $\rho^{'\edge}, X_k^{\edge}, h^{\edge},$ and $T^{\edge}$}
+\cleardoublepage
+\section{Computing $\rho^{'\edge}, X_k^{\edge}, h^{\edge},$ and $T^{\edge}$ in {\tt MAESTRO}}
 In {\tt scalar\_advance}, we call {\tt make\_edge\_scal} to compute $\rho^{'\edge}, X_k^{\edge}, h^{\edge},$ and $T^{\edge}$ at each edge.
@@
 \subsection{3D Spherical Case}
-\section{Computing $\Ubt^{\edge}$}
+\cleardoublepage
+\section{Computing $\Ubt^{\edge}$ in {\tt MAESTRO}}
 In {\tt velocity\_advance}, we call {\tt make\_edge\_state} to compute $\Ubt^{\edge}$.  Unlike when we computed $\Ubt^{\trans}$ and $\Ubt^{\mac,*}$, we need all components of velocity at each face.
@@
 \cleardoublepage
-\section{Forces}
+\section{Computing $\Ub^{\mac}$ in {\tt VARDEN}}
+\subsection{2D Cartesian Case}
+\subsection{3D Cartesian Case}
+\cleardoublepage
+\section{Computing $\Ub^{\edge}$ and $s^{\edge}$ in {\tt VARDEN}}
+\subsection{2D Cartesian Case}
+\subsection{3D Cartesian Case}
+\cleardoublepage
+\section{Notes on {\tt enthalpy\_pred\_type} in {\tt MAESTRO}}
 \begin{table*}

### 2008-04-17 docs/Godunov_notes/paper.tex
listing some differences between MAESTRO and VARDEN

@@
 
-These are working notes for the Godunov step in {\tt MAESTRO}.
+These are working notes for the Godunov step in {\tt MAESTRO} and {\tt VARDEN}.
 \section{Differences Between These Notes and the Code}
@@
 \item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), we could also compute the term proportional to $\partial w_0/\partial r$ by averaging $w^{\trans}$.
+\end{enumerate}
+\section{Differences Between {\tt MAESTRO} and {\tt VARDEN}}
+\begin{enumerate}
+\item {\tt VARDEN} does not use the concept of $w_0$, so there is no such thing as $\Ubt$ either.  {\tt VARDEN} only does Cartesian problems.
+\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$, but rather in {\tt advance\_premac}, the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
+\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$.
+\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, but not $\Ub^{\edge}$
 \end{enumerate}

### 2008-04-17 docs/Godunov_notes/paper.tex
more work on varden umac section

@@
 \subsection{2D Cartesian Case}
+We do a 1D Taylor series extrapolation to get both components of velocity at the x-face:
+\begin{eqnarray}
+\Ub_{L,\ib-\half\eb_x}^{1D} &=& \Ub_{\ib-\eb_x} + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right){\rm max}(0,\Delta_x\Ub_{\ib-\eb_x}), \\
+\Ub_{R,\ib-\half\eb_x}^{1D} &=& \Ub_{\ib} + \left(\half - \frac{\dt}{2h}u_{\ib}\right){\rm min}(0,\Delta_x\Ub_{\ib}).
+\end{eqnarray}
+The normal velocity Riemann problem is:
+\begin{equation}
+u_{\ib-\half\eb_x}^{1D} =
+\begin{cases}
+0, & (u_{L,\ib-\half\eb_x}^{1D} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{1D} \ge 0 ~ {\rm OR} ~ |u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}| < \epsilon, \\
+u_{L,\ib-\half\eb_x}^{1D}, & u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D} > 0, \\
+u_{R,\ib-\half\eb_x}^{1D}, & u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D} < 0.
+\end{cases}
+\end{equation}
+We obtain the transverse velocity by upwinding based on $u_{\ib-\half\eb_x}^{1D}$:
+\begin{equation}
+v_{\ib-\half\eb_x}^{1D} =
+\begin{cases}
+\half(v_{L,\ib-\half\eb_x}^{1D} + v_{R,\ib-\half\eb_x}^{1D}), & \left|u_{\ib-\half\eb_x}^{1D}\right| < \epsilon \\
+v_{L,\ib-\half\eb_x}^{1D}, & u_{\ib-\half\eb_x}^{1D} > 0, \\
+v_{R,\ib-\half\eb_x}^{1D}, & u_{\ib-\half\eb_x}^{1D} < 0.
+\end{cases}
+\end{equation}
+Do the exact analogous thing to get both components of velocity at the y-face, $\Ub_{\ib-\half\eb_y}^{1D}$. \\
+
+Now we do a full-dimensional extrapolation to get only the normal velocity at each face:
+
 \subsection{3D Cartesian Case}

### 2008-04-17 docs/Godunov_notes/paper.tex
finished s^edge for varden
finished varden s^edge section, more equation cleanup

--- 2008-04-17 docs/Godunov_notes/paper.tex
@@
 \item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, but not $\Ub^{\edge}$
+\item The transverse derivative terms in {\tt VARDEN} are done in a conservative formulation, rather than advective.
 \end{enumerate}
@@
 \begin{eqnarray}
-u_L &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
-u_R &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
+u_L^{\trans} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
+u_R^{\trans} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
 \end{eqnarray}
@@
 \begin{cases}
-0, & (u_L \le 0 ~ {\rm AND} ~ u_R \ge 0) ~ {\rm OR} ~ |u_L + u_R| < \epsilon, \\
-u_L, & u_L + u_R > 0, \\
-u_R, & u_L + u_R < 0, \\
+0, & \left(u_L^{\trans} \le 0 ~ {\rm AND} ~ u_R^{\trans} \ge 0\right) ~ {\rm OR} ~ \left|u_L^{\trans} + u_R^{\trans}\right| < \epsilon, \\
+u_L^{\trans}, & u_L^{\trans} + u_R^{\trans} > 0, \\
+u_R^{\trans}, & u_L^{\trans} + u_R^{\trans} < 0, \\
 \end{cases}
@@
 \begin{eqnarray}
-\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
-\wt_R &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.\label{Radial utrans_R}
+\wt_L^{\trans} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
+\wt_R^{\trans} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.\label{Radial utrans_R}
 \end{eqnarray}
@@
 \begin{cases}
-0, & (w_L \le 0 ~ {\rm AND} ~ w_R \ge 0) ~ {\rm OR} ~ |w_L + w_R| < \epsilon, \\
-\wt_L, & w_L + w_R > 0, \\
-\wt_R, & w_L + w_R < 0, \\
+0, & \left(w_L^{\trans} \le 0 ~ {\rm AND} ~ w_R^{\trans} \ge 0\right) ~ {\rm OR} ~ \left|w_L^{\trans} + w_R^{\trans}\right| < \epsilon, \\
+\wt_L^{\trans}, & w_L^{\trans} + w_R^{\trans} > 0, \\
+\wt_R^{\trans}, & w_L^{\trans} + w_R^{\trans} < 0, \\
 \end{cases}
 \end{equation}
-where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  Finally, we call {\tt addw0} to compute $w^{\trans}$ from $\wt^{\trans}$.
+where $w_L^{\trans} = \wt_L^{\trans} + w_{0,r-\half}$ and $w_R^{\trans} = \wt_R^{\trans} + w_{0,r-\half}$.  Finally, we call {\tt addw0} to compute $w^{\trans}$ from $\wt^{\trans}$.
 \subsection{3D Cartesian Case}
@@
 \begin{eqnarray}
-v_L &=& v_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y v_{\ib-\eb_y}^n, \\
-v_R &=& v_{\ib}^n - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y v_{\ib}^n,
+v_L^{\trans} &=& v_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y v_{\ib-\eb_y}^n, \\
+v_R^{\trans} &=& v_{\ib}^n - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y v_{\ib}^n,
 \end{eqnarray}
@@
 \begin{cases}
-0, & (v_L \le 0 ~ {\rm AND} ~ v_R \ge 0) ~ {\rm OR} ~ |v_L + v_R| < \epsilon, \\
-v_L, & v_L + v_R > 0, \\
-v_R, & v_L + v_R < 0. \\
+0, & \left(v_L^{\trans} \le 0 ~ {\rm AND} ~ v_R^{\trans} \ge 0\right) ~ {\rm OR} ~ \left|v_L^{\trans} + v_R^{\trans}\right| < \epsilon, \\
+v_L^{\trans}, & v_L^{\trans} + v_R^{\trans} > 0, \\
+v_R^{\trans}, & v_L^{\trans} + v_R^{\trans} < 0. \\
 \end{cases}
@@
 \begin{eqnarray}
-\ut_L &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x\ut_{\ib-\eb_x}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib-\eb_x}, \\
-\ut_R &=& \ut_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib}.
+\ut_L^{\trans} &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x\ut_{\ib-\eb_x}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib-\eb_x}, \\
+\ut_R^{\trans} &=& \ut_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib}.
 \end{eqnarray}
@@
 \begin{cases}
-\half(u_L^{1D} + u_R^{1D}), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
+\half\left(u_L^{1D} + u_R^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
 u_L^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
@@
 \begin{eqnarray}
-u_L &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
+u_L^{\mac} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
 && - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
-u_R &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
+u_R^{\mac} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right).
@@
 \begin{cases}
-0, & \left(u_L \le 0 ~ {\rm AND} ~ u_R \ge 0\right) ~ {\rm OR} ~ \left|u_L + u_R\right| < \epsilon, \\
-u_L, & u_L + u_R > 0, \\
-u_R, & u_L + u_R < 0. 
+0, & \left(u_L^{\mac} \le 0 ~ {\rm AND} ~ u_R^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|u_L^{\mac} + u_R^{\mac}\right| < \epsilon, \\
+u_L^{\mac}, & u_L^{\mac} + u_R^{\mac} > 0, \\
+u_R^{\mac}, & u_L^{\mac} + u_R^{\mac} < 0. 
 \end{cases}
@@
 \begin{cases}
-\half(\wt_L^{1D} + \wt_R^{1D}), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
+\half\left(\wt_L^{1D} + \wt_R^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
 \wt_L^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
@@
 \begin{eqnarray}
-\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{w,\ib-\eb_r} \nonumber \\
+\wt_L^{\mac} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{w,\ib-\eb_r} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
 && - \frac{\dt}{2h}\wt_{\ib-\eb_r}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}),\label{wt_L Extrapolation} \\
-\wt_R &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
+\wt_R^{\mac} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
@@
 \begin{cases}
-0, & \left(w_L \le 0 ~ {\rm AND} ~ w_R \ge 0\right) ~ {\rm OR} ~ \left|w_L + w_R\right| < \epsilon, \\
-\wt_L, & w_L + w_R > 0, \\
-\wt_R, & w_L + w_R < 0. 
+0, & \left(w_L^{\mac} \le 0 ~ {\rm AND} ~ w_R^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|w_L^{\mac} + w_R^{\mac}\right| < \epsilon, \\
+\wt_L^{\mac}, & w_L^{\mac} + w_R^{\mac} > 0, \\
+\wt_R^{\mac}, & w_L^{\mac} + w_R^{\mac} < 0. 
 \end{cases}
@@
 \begin{cases}
-\half(s_{L,\ib-\half\eb_x}^{1D} + s_{R,\ib-\half\eb_x}^{1D}), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+\half\left(s_{L,\ib-\half\eb_x}^{1D} + s_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
 s_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
@@
 \begin{cases}
-\half(s_{L,\ib-\half\eb_r}^{1D} + s_{R,\ib-\half\eb_r}^{1D}), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
+\half\left(s_{L,\ib-\half\eb_r}^{1D} + s_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
 u_{L,\ib-\half\eb_r}^{1D}, & w^{\mac}_{\ib-\half\eb_r} > 0, \\
@@
 \begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib-\eb_x}^n + \frac{\dt}{2}f_{s,\ib-\eb_x} \nonumber \\
+s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib-\eb_x} \nonumber \\
 && - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\mac}+w_{\ib-\eb_x-\half\eb_r}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_r}^{1D} - s_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
-s_{R,\ib-\half\eb_x}^{\edge} &=& s_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib}^n + \frac{\dt}{2}f_{s,\ib} \nonumber \\
+s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\mac}+w_{\ib-\half\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_r}^{1D} - s_{\ib-\half\eb_r}^{1D}\right).
@@
 \begin{cases}
-\half(s_{L,\ib-\half\eb_x}^{\edge} + s_{R,\ib-\half\eb_x}^{\edge}), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+\half\left(s_{L,\ib-\half\eb_x}^{\edge} + s_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
 s_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
@@
 \begin{cases}
-\half(s_{L,\ib-\half\eb_r}^{\edge} + s_{R,\ib-\half\eb_r}^{\edge}), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
+\half\left(s_{L,\ib-\half\eb_r}^{\edge} + s_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
 s_{L,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} > 0, \\
@@
 \begin{cases}
-\half(u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
+\half\left(u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
 u_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
@@
 \begin{cases}
-\half(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
+\half\left(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
 u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
@@
 \begin{cases}
-\half(\wt_{L,\ib-\half\eb_x}^{1D} + \wt_{R,\ib-\half\eb_x}^{1D}), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
+\half\left(\wt_{L,\ib-\half\eb_x}^{1D} + \wt_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
 \wt_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
@@
 \begin{cases}
-\half(\wt_{L,\ib-\half\eb_r}^{1D} + \wt_{R,\ib-\half\eb_r}^{1D}), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
+\half\left(\wt_{L,\ib-\half\eb_r}^{1D} + \wt_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
 \wt_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
@@
 \begin{cases}
-\half(u_{L,\ib-\half\eb_x}^{\edge} + u_{R,\ib-\half\eb_x}^{\edge}), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+\half\left(u_{L,\ib-\half\eb_x}^{\edge} + u_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
 u_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
@@
 \begin{cases}
-\half(u_{L,\ib-\half\eb_r}^{\edge} + u_{R,\ib-\half\eb_r}^{\edge}), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
+\half\left(u_{L,\ib-\half\eb_r}^{\edge} + u_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
 u_{L,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} > 0, \\
@@
 \begin{cases}
-\half(\wt_{L,\ib-\half\eb_x}^{\edge} + \wt_{R,\ib-\half\eb_x}^{\edge}), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+\half\left(\wt_{L,\ib-\half\eb_x}^{\edge} + \wt_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
 \wt_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
@@
 \begin{cases}
-\half(\wt_{L,\ib-\half\eb_r}^{\edge} + \wt_{R,\ib-\half\eb_r}^{\edge}), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
+\half\left(\wt_{L,\ib-\half\eb_r}^{\edge} + \wt_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
 \wt_{L,\ib-\half\eb_r}^{\edge}, & w^{\mac}_{\ib-\half\eb_r} > 0, \\
@@
 \subsection{2D Cartesian Case}
+We do a 1D Taylor series extrapolation to get both components of velocity at the x-face:
+\begin{eqnarray}
+\Ub_{L,\ib-\half\eb_x}^{1D} &=& \Ub_{\ib-\eb_x} + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right){\rm max}(0,\Delta_x\Ub_{\ib-\eb_x}), \\
+\Ub_{R,\ib-\half\eb_x}^{1D} &=& \Ub_{\ib} + \left(\half - \frac{\dt}{2h}u_{\ib}\right){\rm min}(0,\Delta_x\Ub_{\ib}).
+\end{eqnarray}
+The normal velocity Riemann problem is:
+\begin{equation}
+u_{\ib-\half\eb_x}^{1D} =
+\begin{cases}
+0, & \left(u_{L,\ib-\half\eb_x}^{1D} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{1D} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}\right| < \epsilon, \\
+u_{L,\ib-\half\eb_x}^{1D}, & u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D} > 0, \\
+u_{R,\ib-\half\eb_x}^{1D}, & u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D} < 0.
+\end{cases}
+\end{equation}
+We obtain the transverse velocity by upwinding based on $u_{\ib-\half\eb_x}^{1D}$:
+\begin{equation}
+v_{\ib-\half\eb_x}^{1D} =
+\begin{cases}
+\half\left(v_{L,\ib-\half\eb_x}^{1D} + v_{R,\ib-\half\eb_x}^{1D}\right), & \left|u_{\ib-\half\eb_x}^{1D}\right| < \epsilon \\
+v_{L,\ib-\half\eb_x}^{1D}, & u_{\ib-\half\eb_x}^{1D} > 0, \\
+v_{R,\ib-\half\eb_x}^{1D}, & u_{\ib-\half\eb_x}^{1D} < 0.
+\end{cases}
+\end{equation}
+Do the exact analogous thing to get both components of velocity at the y-face, $\Ub_{\ib-\half\eb_y}^{1D}$. \\
+
+Now we do a full-dimensional extrapolation to get only the normal velocity at each face:
+\begin{eqnarray}
+u_L^{\mac} &=& \Ub_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
+&-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
+u_R^{\mac} &=& \Ub_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib} \nonumber \\
+&-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right).
+\end{eqnarray}
+Then, solve a Riemann problem:
+\begin{equation}
+u_{\ib-\half\eb_x}^{\mac} =
+\begin{cases}
+0, & \left(u_L^{\mac} \le 0 ~ {\rm AND} ~ u_R^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|u_L^{\mac} + u_R^{\mac}\right| < \epsilon, \\
+u_L^{\mac}, & u_L^{\mac} + u_R^{\mac} > 0, \\
+u_R^{\mac}, & u_L^{\mac} + u_R^{\mac} < 0.
+\end{cases}
+\end{equation}
+Use the exact analogous procedure to obtain $v^{\mac}_{\ib-\half\eb_y}$.
 \subsection{3D Cartesian Case}
 \cleardoublepage
-\section{Computing $\Ub^{\edge}$ and $s^{\edge}$ in {\tt VARDEN}}
+\section{Computing $\Ub^{\edge}$ and $\rho^{\edge}$ in {\tt VARDEN}}
+To compute $\Ub^{\edge}$, {\tt VARDEN} uses the exact same algorithm as the $s^{\edge}$ case in {\tt MAESTRO}.
+To the algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.
+In particular:
+\begin{eqnarray}
+\rho_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib-\eb_x}^n, \\
+\rho_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib}^n.
+\end{eqnarray}
+Then we upwind based on $u^{\mac}$:
+\begin{equation}
+\rho_{\ib-\half\eb_x}^{1D} =
+\begin{cases}
+\half\left(\rho_{L,\ib-\half\eb_x}^{1D} + \rho_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+\rho_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
+\rho_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
+\end{cases}
+\end{equation}
+Now we do a full-dimensional extrapolation of $\rho$ to each face.  The extrapolation of $\rho$ to x-faces is:
+\begin{eqnarray}
+\rho_{L,\ib-\half\eb_x}^{\edge} &=& \rho_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib-\eb_x} - \rho_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} \nonumber \\
+&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right), \nonumber \\
+&& \\
+\rho_{R,\ib-\half\eb_x}^{\edge} &=& \rho_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib} - \rho_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} \nonumber \\
+&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right).
+\end{eqnarray}
+Then, we upwind based on $u^{\mac}$.
+\begin{equation}
+\rho_{\ib-\half\eb_x}^{\edge} =
+\begin{cases}
+\half\left(\rho_{L,\ib-\half\eb_x}^{\edge} + \rho_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+\rho_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
+\rho_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
+\end{cases}
+\end{equation}
 \subsection{2D Cartesian Case}

### 2008-04-17 docs/Godunov_notes/paper.tex
cleaning up the notation so it is crystal clear what we are doing
Merging notes on enthalpy_pred_type into main document to remove some of the tables at the end.

@@
 \def\dt    {\Delta t}
+\def\kth   {k_{\rm th}}
 
@@
 \begin{eqnarray}
-u_L^{\trans} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
-u_R^{\trans} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
+u_{L,\ib-\half\eb_x}^{\trans} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
+u_{R,\ib-\half\eb_x}^{\trans} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
 \end{eqnarray}
@@
 \begin{cases}
-0, & \left(u_L^{\trans} \le 0 ~ {\rm AND} ~ u_R^{\trans} \ge 0\right) ~ {\rm OR} ~ \left|u_L^{\trans} + u_R^{\trans}\right| < \epsilon, \\
-u_L^{\trans}, & u_L^{\trans} + u_R^{\trans} > 0, \\
-u_R^{\trans}, & u_L^{\trans} + u_R^{\trans} < 0, \\
+0, & \left(u_{L,\ib-\half\eb_x}^{\trans} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\trans} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans}\right| < \epsilon, \\
+u_{L,\ib-\half\eb_x}^{\trans}, & u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans} > 0, \\
+u_{R,\ib-\half\eb_x}^{\trans}, & u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans} < 0, \\
 \end{cases}
@@
 \begin{eqnarray}
-\wt_L^{\trans} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
-\wt_R^{\trans} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.\label{Radial utrans_R}
+\wt_{L,\ib-\half\eb_r}^{\trans} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
+\wt_{R,\ib-\half\eb_r}^{\trans} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.\label{Radial utrans_R}
 \end{eqnarray}
@@
 0, & \left(w_L^{\trans} \le 0 ~ {\rm AND} ~ w_R^{\trans} \ge 0\right) ~ {\rm OR} ~ \left|w_L^{\trans} + w_R^{\trans}\right| < \epsilon, \\
-\wt_L^{\trans}, & w_L^{\trans} + w_R^{\trans} > 0, \\
-\wt_R^{\trans}, & w_L^{\trans} + w_R^{\trans} < 0, \\
+\wt_{L,\ib-\half\eb_r}^{\trans}, & w_L^{\trans} + w_R^{\trans} > 0, \\
+\wt_{R,\ib-\half\eb_r}^{\trans}, & w_L^{\trans} + w_R^{\trans} < 0, \\
 \end{cases}
 \end{equation}
-where $w_L^{\trans} = \wt_L^{\trans} + w_{0,r-\half}$ and $w_R^{\trans} = \wt_R^{\trans} + w_{0,r-\half}$.  Finally, we call {\tt addw0} to compute $w^{\trans}$ from $\wt^{\trans}$.
+where $w_L^{\trans} = \wt_{L,\ib-\half\eb_r}^{\trans} + w_{0,r-\half}$ and $w_R^{\trans} = \wt_{R,\ib-\half\eb_r}^{\trans} + w_{0,r-\half}$.
+Finally, we call {\tt addw0} to compute $w^{\trans}$ from $\wt^{\trans}$.
 \subsection{3D Cartesian Case}
@@
 \begin{eqnarray}
-v_L^{\trans} &=& v_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y v_{\ib-\eb_y}^n, \\
-v_R^{\trans} &=& v_{\ib}^n - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y v_{\ib}^n,
+v_{L,\ib-\half\eb_y}^{\trans} &=& v_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y v_{\ib-\eb_y}^n, \\
+v_{R,\ib-\half\eb_y}^{\trans} &=& v_{\ib}^n - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y v_{\ib}^n,
 \end{eqnarray}
@@
 \begin{cases}
-0, & \left(v_L^{\trans} \le 0 ~ {\rm AND} ~ v_R^{\trans} \ge 0\right) ~ {\rm OR} ~ \left|v_L^{\trans} + v_R^{\trans}\right| < \epsilon, \\
-v_L^{\trans}, & v_L^{\trans} + v_R^{\trans} > 0, \\
-v_R^{\trans}, & v_L^{\trans} + v_R^{\trans} < 0. \\
+0, & \left(v_{L,\ib-\half\eb_y}^{\trans} \le 0 ~ {\rm AND} ~ v_{R,\ib-\half\eb_y}^{\trans} \ge 0\right) ~ {\rm OR} ~ \left|v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans}\right| < \epsilon, \\
+v_{L,\ib-\half\eb_y}^{\trans}, & v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans} > 0, \\
+v_{R,\ib-\half\eb_y}^{\trans}, & v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans} < 0. \\
 \end{cases}
@@
 \begin{eqnarray}
-u_L^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
-u_R^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
+u_{L,\ib-\half\eb_r}^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
+u_{R,\ib-\half\eb_r}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
 \end{eqnarray}
@@
 \begin{cases}
-\half\left(u_L^{1D} + u_R^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_L^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
-u_R^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
+\half\left(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
+u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
+u_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
 \end{cases}
@@
 \begin{eqnarray}
-u_L^{\mac} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
+u_{L,\ib-\half\eb_x}^{\mac} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
 && - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
-u_R^{\mac} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
+u_{R,\ib-\half\eb_x}^{\mac} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right).
@@
 \begin{cases}
-0, & \left(u_L^{\mac} \le 0 ~ {\rm AND} ~ u_R^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|u_L^{\mac} + u_R^{\mac}\right| < \epsilon, \\
-u_L^{\mac}, & u_L^{\mac} + u_R^{\mac} > 0, \\
-u_R^{\mac}, & u_L^{\mac} + u_R^{\mac} < 0. 
+0, & \left(u_{L,\ib-\half\eb_x}^{\mac} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac}\right| < \epsilon, \\
+u_{L,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} > 0, \\
+u_{R,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} < 0. 
 \end{cases}
@@
 \begin{eqnarray}
-\wt_L^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
-\wt_R^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \wt_{\ib}^n.
+\wt_{L,\ib-\half\eb_x}^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
+\wt_{R,\ib-\half\eb_x}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \wt_{\ib}^n.
 \end{eqnarray}
@@
 \begin{cases}
-\half\left(\wt_L^{1D} + \wt_R^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
-\wt_L^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
-\wt_R^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
+\half\left(\wt_{L,\ib-\half\eb_x}^{1D} + \wt_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
+\wt_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
+\wt_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
 \end{cases}
@@
 \begin{eqnarray}
-\wt_L^{\mac} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{w,\ib-\eb_r} \nonumber \\
+\wt_{L,\ib-\half\eb_r}^{\mac} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{w,\ib-\eb_r} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
 && - \frac{\dt}{2h}\wt_{\ib-\eb_r}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}),\label{wt_L Extrapolation} \\
-\wt_R^{\mac} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
+\wt_{R,\ib-\half\eb_r}^{\mac} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
@@
 0, & \left(w_L^{\mac} \le 0 ~ {\rm AND} ~ w_R^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|w_L^{\mac} + w_R^{\mac}\right| < \epsilon, \\
-\wt_L^{\mac}, & w_L^{\mac} + w_R^{\mac} > 0, \\
-\wt_R^{\mac}, & w_L^{\mac} + w_R^{\mac} < 0. 
+\wt_{L,\ib-\half\eb_r}^{\mac}, & w_L^{\mac} + w_R^{\mac} > 0, \\
+\wt_{R,\ib-\half\eb_r}^{\mac}, & w_L^{\mac} + w_R^{\mac} < 0. 
 \end{cases}
@@
 Finally, we call {\tt addw0} to compute $\Ub^{\mac,*}$ from $\Ubt^{\mac,*}$.
-\subsection{3D Cartesian Case}
-\subsection{3D Spherical Case}
+%\subsection{3D Cartesian Case}
+%\subsection{3D Spherical Case}
 \cleardoublepage
@@
 The procedure is the same for each scalar, so we shall simply denote the scalar as $s$.
-\\
-
+We always need to compute $\rho'$ and $X_k$ to edges.
+\begin{itemize}
+\item For {\tt enthalpy\_pred\_type} = 1, we predict $(\rho h)'$ to edges.
+\item For {\tt enthalpy\_pred\_type} = 2, we predict $h$ to edges.
+\item For {\tt enthalpy\_pred\_type} = 3 and 4, we predict $T$ to edges.
+\end{itemize}
 The force terms are:
@@
 f_{X_k} &=& 0, \\
-f_h &=& \frac{1}{\rho}\left(\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} \right), \\
-f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
+f_{(\rho h)'} &=& (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} -(\rho h)'\nabla\cdot\Ub^{\mac} - \nabla\cdot\left[(\rho h)_0\Ubt^{\mac}\right] + \nabla\cdot\kth\nabla T \\
+f_h &=& \frac{1}{\rho}\left(\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T \right), \\
+f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext} + \nabla\cdot\kth\nabla T \right\}.
 \end{eqnarray}
@@
 \end{equation}
-\subsection{3D Cartesian Case}
-\subsection{3D Spherical Case}
+%\subsection{3D Cartesian Case}
+%\subsection{3D Spherical Case}
 \cleardoublepage
@@
 \begin{eqnarray}
-u_L^{\mac} &=& \Ub_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
+u_{L,\ib-\half\eb_x}^{\mac} &=& \Ub_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
 &-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
-u_R^{\mac} &=& \Ub_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib} \nonumber \\
+u_{R,\ib-\half\eb_x}^{\mac} &=& \Ub_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib} \nonumber \\
 &-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right).
@@
 \begin{cases}
-0, & \left(u_L^{\mac} \le 0 ~ {\rm AND} ~ u_R^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|u_L^{\mac} + u_R^{\mac}\right| < \epsilon, \\
-u_L^{\mac}, & u_L^{\mac} + u_R^{\mac} > 0, \\
-u_R^{\mac}, & u_L^{\mac} + u_R^{\mac} < 0.
+0, & \left(u_{L,\ib-\half\eb_x}^{\mac} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac}\right| < \epsilon, \\
+u_{L,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} > 0, \\
+u_{R,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} < 0.
 \end{cases}
@@
 Use the exact analogous procedure to obtain $v^{\mac}_{\ib-\half\eb_y}$.
-\subsection{3D Cartesian Case}
+%\subsection{3D Cartesian Case}
 \cleardoublepage
@@
 To the algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.
-In particular:
+\subsection{2D Cartesian Case}
+The 1D extrapolation is:
 \begin{eqnarray}
@@
 \end{equation}
-\subsection{2D Cartesian Case}
-\subsection{3D Cartesian Case}
+%\subsection{3D Cartesian Case}
 \cleardoublepage
 \section{Notes on {\tt enthalpy\_pred\_type} in {\tt MAESTRO}}
-\begin{table*}
-\begin{center}
-\caption{Forcing term into make edge scal \newline}
-\begin{tabular}{|c|c|c|}
-\multicolumn{1}{c}{PRED X}   & \multicolumn{1}{c}{Force} \\
-\hline
-  &   \\
-T $(X)$ & $0$ \\
-  &   \\
-T $(\rho^\prime)$ & $-\rho^\prime \; \nabla \cdot (\ut+w_0 \eb_r) - \nabla \cdot (\ut \rho_0) $  \\
-  &   \\
-\hline
-\end{tabular}
-\end{center}
-\begin{center}
-\begin{tabular}{|c|c|}
-\multicolumn{1}{c}{H PRED TYPE}   & \multicolumn{1}{c}{Force} \\
-\hline
-   & \\
- 1 $((\rho h)^\prime)$ &  $(\tilde{w} \frac{\partial p_0}{\partial r}) - (\rho h)^\prime \; \nabla \cdot (\ut+w_0 \eb_r) - \nabla \cdot (\ut (\rho h)_0) $  \\
-   & \\
- 2 $(h)$ & $\frac{1}{\rho} (\psi + \tilde{w} \frac{\partial p_0}{\partial r})$  \\
-   & \\
- 3 $(T)$ & $\frac{1}{\rho c_p} \left[ (1 - \rho h_p) (\psi + \tilde{w} \frac{\partial p_0}{\partial r}) \right]$ \\
-   & \\
- 4 $(T)$ & $\frac{1}{\rho c_p} \left[ (1 - \rho h_p) (\psi + \tilde{w} \frac{\partial p_0}{\partial r}) \right]$ \\
-   & \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-
-\begin{table*}
-\begin{center}
-\caption{Quantity that goes into and out of make edge scal \newline}
-\begin{tabular}{|c|c|c|c|}
-\multicolumn{2}{c}{PREDICT} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-\multicolumn{1}{c}{X} & \multicolumn{1}{c}{h pred type} & \multicolumn{1}{c}{Species } & \multicolumn{1}{c}{h/T } \\
-\hline
-T & 1 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
-T & 2 & $X$ and $\rho^\prime$ & $h$  \\
-T & 3 & $X$ and $\rho^\prime$ & $T$  \\
-T & 4 & $X$ and $\rho^\prime$ & $T$  \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
 

### 2008-04-17 docs/Godunov_notes/paper.tex
Merged the rest of the enthalpy_pred_type tables into the document

@@
 \end{eqnarray}
+If {\tt enthalpy\_pred\_type} = 3 or 4, we convert $T^{\edge}$ to $h^{\edge}$ using $h^{\edge} = h^{\edge}(\rho^{\edge},T^{\edge},X_k^{\edge})$, where $\rho^{\edge} = \rho^{'\edge} + \rho_0$.
+Furthermore, if {\tt enthalpy\_pred\_type} = 3, we compute $(\rho h)^{'\edge} = (\rho h)^{\edge} - (\rho h)_0$.
 \subsection{2D Cartesian Case}
@@
 %\subsection{3D Cartesian Case}
-\cleardoublepage
-\section{Notes on {\tt enthalpy\_pred\_type} in {\tt MAESTRO}}
-
-\begin{table*}
-\begin{center}
-\caption{When predicting temp ... \newline}
-\begin{tabular}{|c|c|c|c|c|}
-\multicolumn{1}{c}{PRED X} & \multicolumn{1}{c}{type} & \multicolumn{1}{c}{Inputs } & 
-\multicolumn{1}{c}{EOS inputs} & \multicolumn{1}{c}{output} \\
-\hline
-T & 3 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ &  $(\rho h)^\prime = \rho h - (\rho h)_0$ \\
-  &   &                   & $   X = $ given & \\
-  &   &                   &                 & \\
-T & 4 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ & $h$ \\
-  &   &                   & $   X = $ given & \\
-  &   &                   &                 & \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-
-\begin{table*}
-\begin{center}
-\caption{Quantity that goes into mkflux on edges \newline}
-\begin{tabular}{|c|c|c|c|c|}
-\multicolumn{2}{c}{PREDICT} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
-\multicolumn{1}{c}{X} & \multicolumn{1}{c}{h pred type} & \multicolumn{1}{c}{Species } & \multicolumn{1}{c}{h/T } & \multicolumn{1}{c}{$\rho^\prime$} \\
-\hline
-T & 1 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ & \\
-T & 2 & $X$ and $\rho^\prime$ & $h$ & \\
-T & 3 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ & \\
-T & 4 & $X$ and $\rho^\prime$ & $h$ & \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-
-\begin{table*}
-\begin{center}
-\caption{Quantity that is created in mkflux and passed into update scal\newline}
-\begin{tabular}{|c|c|c|c|c|c|c|c|}
-\multicolumn{2}{c}{PREDICT}   & 
-\multicolumn{1}{c}{} & \multicolumn{1}{c}{}  
-& \multicolumn{2}{c}{Add db?} & \multicolumn{2}{c}{Forcing term }\\
-\multicolumn{1}{c}{X}   & \multicolumn{1}{c}{type} &
-\multicolumn{1}{c}{Species } & \multicolumn{1}{c}{$(\rho h)$} 
-& \multicolumn{1}{c}{X} & \multicolumn{1}{c}{h} & \multicolumn{1}{c}{$(\rho X)$} & \multicolumn{1}{c}{$(\rho h)$} \\
-\hline
-T & 1 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0)(\rho h)^\prime + \ut (\rho h)_0$ & N & Y & 0 & $ (\tilde{w} \frac{\partial p_0}{\partial r}) $ \\
-T & 2 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0) (\rho_0+\rho^\prime) h$ & N & N & 0 & $ (\psi + \tilde{w} \frac{\partial p_0}{\partial r}) $ \\
-T & 3 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0)(\rho h)^\prime + \ut (\rho h)_0$ & N & Y & 0 & $ (\tilde{w} \frac{\partial p_0}{\partial r}) $ \\
-T & 4 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0) (\rho_0+\rho^\prime) h$ & N & N & 0 & $ (\psi + \tilde{w} \frac{\partial p_0}{\partial r}) $ \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-
 \end{document}

### 2008-04-17 docs/Godunov_notes/paper.tex
modified to account for the fact that we take the average of the wtilde terms for the wtilde dw_0/dr terms instead of the cell-centered wtilde

@@
 \begin{enumerate}
-\item In both equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}), we are omitting the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ (by using $\wt$ rather than $w$ as the advection speed) and $(\Ubt\cdot\eb_r)$ (by omitting the term altogether).
-\item In equation (\ref{Force Term}), we do not include the term proportional to $\partial\pi_0/\partial r$.
+\item Concerning both equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}), in the code we are omitting the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ (by using $\wt$ rather than $w$ as the advection speed) and $(\Ubt\cdot\eb_r)$ (by omitting the term altogether).
+\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$.
 \end{enumerate}
@@
 \begin{enumerate}
-\item We need to re-examine when to include the term proportional to $(\Ubt\cdot\eb_r)$ in equation (\ref{Perturbational Velocity Equation}) when we do a 1D extrapolation of the radial velocity.
+\item We need to re-examine when to include the term proportional to $(\Ubt\cdot\eb_r)$ in equation (\ref{Perturbational Velocity Equation}) when we do a 1D extrapolation of the radial velocity to r-faces.
 For example, equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}) include it, but equations (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}) do not.
-\item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), we could also compute the term proportional to $\partial w_0/\partial r$ by averaging $w^{\trans}$.
+\item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$.
 \end{enumerate}
@@
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{2h}\wt_{\ib-\eb_r}(w_{0,j-\half} - w_{0,j-\frac{3}{2}}),\label{wt_L Extrapolation} \\
+&& - \frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)(w_{0,\ib-\half\eb_r} - w_{0,\ib-\frac{3}{2}\eb_r}),\label{wt_L Extrapolation} \\
 \wt_{R,\ib-\half\eb_r}^{\mac} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{2h}\wt_{\ib}(w_{0,j+\half} - w_{0,j-\half}).\label{wt_R Extrapolation}
+&& - \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)(w_{0,\ib+\half\eb_r} - w_{0,\ib-\half\eb_r}).\label{wt_R Extrapolation}
 \end{eqnarray}
@@
 && - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(\wt_{\ib-\eb_x+\half\eb_r}^{1D} - \wt_{\ib-\eb_x-\half\eb_r}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{4h}\left(\wt_{\ib-\eb_x+\half\eb_r}^{\trans}+\wt_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(w_{0,\ib-\eb_x+\half\eb_r}-w_{0,\ib-\eb_x-\half\eb_r}\right), \\
+&& - \frac{\dt}{4h}\left(\wt_{\ib-\eb_x+\half\eb_r}^{\trans}+\wt_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(w_{0,\ib-\eb_x+\half\eb_r}-w_{0,\ib-\eb_x-\half\eb_r}\right), \label{wt_L x-edge} \\
 \wt_{R,\ib-\half\eb_x}^{\edge} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_r}^{1D} - \wt_{\ib-\half\eb_r}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right).
+&& - \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right). \label{wt_R x-edge}
 \end{eqnarray}
@@
 && -
-\frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)\left(w_{0,\ib-\half\eb_r}-w_{0,\ib-\frac{3}{2}\eb_r}\right),\\
+\frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)\left(w_{0,\ib-\half\eb_r}-w_{0,\ib-\frac{3}{2}\eb_r}\right), \label{wt_L r-edge} \\
 \wt_{R,\ib-\half\eb_r}^{\edge} &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
@@
 && -
-\frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib+\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right).
+\frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right). \label{wt_R r-edge}
 \end{eqnarray}

### 2008-04-17 docs/Godunov_notes/paper.tex
minor cleanup

@@
 For example, equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}) include it, but equations (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}) do not.
-\item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$.
+\item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$ rather than the average of the neighboring $\wt^{\trans}$.
 \end{enumerate}
@@
 f_h &=& \frac{1}{\rho}\left(\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T \right), \\
-f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext} + \nabla\cdot\kth\nabla T \right\}.
+f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext} + \nabla\cdot\kth\nabla T \right\}.\nonumber \\
 \end{eqnarray}

### 2008-04-17 docs/Godunov_notes/paper.tex
A few minor edits John suggested, and some of my own.

@@
 \item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, but not $\Ub^{\edge}$
-\item The transverse derivative terms in {\tt VARDEN} are done in a conservative formulation, rather than advective.
+\item The transverse derivative terms for $\rho$ in {\tt VARDEN} are done in a conservative formulation, rather than advective.
 \end{enumerate}
@@
 \begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
+u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
 && - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
-u_{R,\ib-\half\eb_x}^{\mac} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
+u_{R,\ib-\half\eb_x}^{\mac,*} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right).
@@
 \begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\mac} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} > 0, \\
-u_{R,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} < 0. 
+0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
+u_{L,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} > 0, \\
+u_{R,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} < 0. 
 \end{cases}
@@
 \begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{\mac} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{w,\ib-\eb_r} \nonumber \\
+\wt_{L,\ib-\half\eb_r}^{\mac,*} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{w,\ib-\eb_r} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
 && - \frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)(w_{0,\ib-\half\eb_r} - w_{0,\ib-\frac{3}{2}\eb_r}),\label{wt_L Extrapolation} \\
-\wt_{R,\ib-\half\eb_r}^{\mac} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
+\wt_{R,\ib-\half\eb_r}^{\mac,*} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
 && - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
@@
 \begin{cases}
-0, & \left(w_L^{\mac} \le 0 ~ {\rm AND} ~ w_R^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|w_L^{\mac} + w_R^{\mac}\right| < \epsilon, \\
-\wt_{L,\ib-\half\eb_r}^{\mac}, & w_L^{\mac} + w_R^{\mac} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{\mac}, & w_L^{\mac} + w_R^{\mac} < 0. 
+0, & \left(w_L^{\mac,*} \le 0 ~ {\rm AND} ~ w_R^{\mac,*} \ge 0\right) ~ {\rm OR} ~ \left|w_L^{\mac,*} + w_R^{\mac,*}\right| < \epsilon, \\
+\wt_{L,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} > 0, \\
+\wt_{R,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} < 0. 
 \end{cases}
 \end{equation}
-Finally, we call {\tt addw0} to compute $\Ub^{\mac,*}$ from $\Ubt^{\mac,*}$.
+We then project $\Ub^{\mac,*}$ to obtain $\Ub^{\mac}$ and we call {\tt addw0} to compute $\Ub^{\mac}$ from $\Ubt^{\mac}$.
 %\subsection{3D Cartesian Case}
@@
 \end{eqnarray}
-If {\tt enthalpy\_pred\_type} = 3 or 4, we convert $T^{\edge}$ to $h^{\edge}$ using $h^{\edge} = h^{\edge}(\rho^{\edge},T^{\edge},X_k^{\edge})$, where $\rho^{\edge} = \rho^{'\edge} + \rho_0$.
+If {\tt enthalpy\_pred\_type} = 3 or 4, we convert $T^{\edge}\rightarrow h^{\edge}$ using $h^{\edge} = h^{\edge}(\rho^{\edge},T^{\edge},X_k^{\edge})$, where $\rho^{\edge} = \rho^{'\edge} + \rho_0$.
 Furthermore, if {\tt enthalpy\_pred\_type} = 3, we compute $(\rho h)^{'\edge} = (\rho h)^{\edge} - (\rho h)_0$.
@@
 \cleardoublepage
-\section{Computing $\Ub^{\mac}$ in {\tt VARDEN}}
+\section{Computing $\Ub^{\mac,*}$ in {\tt VARDEN}}
 \subsection{2D Cartesian Case}
@@
 \begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac} &=& \Ub_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
+u_{L,\ib-\half\eb_x}^{\mac,*} &=& \Ub_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
 &-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
-u_{R,\ib-\half\eb_x}^{\mac} &=& \Ub_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib} \nonumber \\
+u_{R,\ib-\half\eb_x}^{\mac,*} &=& \Ub_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib} \nonumber \\
 &-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right).
@@
 \begin{equation}
-u_{\ib-\half\eb_x}^{\mac} =
+u_{\ib-\half\eb_x}^{\mac,*} =
 \begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\mac} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} > 0, \\
-u_{R,\ib-\half\eb_x}^{\mac}, & u_{L,\ib-\half\eb_x}^{\mac} + u_{R,\ib-\half\eb_x}^{\mac} < 0.
+0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
+u_{L,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} > 0, \\
+u_{R,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} < 0.
 \end{cases}
 \end{equation}
-Use the exact analogous procedure to obtain $v^{\mac}_{\ib-\half\eb_y}$.
+Use the exact analogous procedure to obtain $v^{\mac,*}_{\ib-\half\eb_y}$.
 %\subsection{3D Cartesian Case}
@@
 \begin{eqnarray}
-\rho_{L,\ib-\half\eb_x}^{\edge} &=& \rho_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib-\eb_x} - \rho_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} \nonumber \\
+\rho_{L,\ib-\half\eb_x}^{\edge} &=& \rho_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib-\eb_x} - (\rho\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} \nonumber \\
 &&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right), \nonumber \\
 && \\
-\rho_{R,\ib-\half\eb_x}^{\edge} &=& \rho_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib} - \rho_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} \nonumber \\
+\rho_{R,\ib-\half\eb_x}^{\edge} &=& \rho_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib} - (\rho\nabla\cdot\Ub^{\mac})_{\ib} \nonumber \\
 &&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right).

### 2008-04-17 docs/Godunov_notes/paper.tex
wrote the 3D corner coupling version of u^mac for varden

@@
 \end{equation}
-We then project $\Ub^{\mac,*}$ to obtain $\Ub^{\mac}$ and we call {\tt addw0} to compute $\Ub^{\mac}$ from $\Ubt^{\mac}$.
+We then project $\Ubt^{\mac,*}$ to obtain $\Ubt^{\mac}$ and we call {\tt addw0} to compute $\Ub^{\mac}$ from $\Ubt^{\mac}$.
 %\subsection{3D Cartesian Case}
@@
 \begin{eqnarray}
-\Ub_{L,\ib-\half\eb_x}^{1D} &=& \Ub_{\ib-\eb_x} + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right){\rm max}(0,\Delta_x\Ub_{\ib-\eb_x}), \\
+\Ub_{L,\ib-\half\eb_x}^{1D} &=& \Ub_{\ib-\eb_x} + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right){\rm max}(0,\Delta_x\Ub_{\ib-\eb_x}), \label{varden U_L^1D} \\
 \Ub_{R,\ib-\half\eb_x}^{1D} &=& \Ub_{\ib} + \left(\half - \frac{\dt}{2h}u_{\ib}\right){\rm min}(0,\Delta_x\Ub_{\ib}).
@@
 v_{R,\ib-\half\eb_x}^{1D}, & u_{\ib-\half\eb_x}^{1D} < 0.
-\end{cases}
+\end{cases}\label{Transverse Velocity Riemann Problem}
 \end{equation}
@@
 u_{R,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} < 0.
+\end{cases}\label{umac Riemann Problem}
+\end{equation}
+Use the exact analogous procedure to obtain $v^{\mac,*}_{\ib-\half\eb_y}$.
+\subsection{3D Cartesian Case}
+This is more complicated than the 2D case because we include corner coupling.
+We compute $\Ub_{\ib-\half\eb_x}, \Ub_{\ib-\half\eb_y}$, and $\Ub_{\ib-\half\eb_z}$ in an analogous manner as equations (\ref{varden U_L^1D})-(\ref{Transverse Velocity Riemann Problem}).
+Then, we compute an intermediate state, $u_{\ib-\half\eb_y}^{y|z}$, which is described as ``state $u_{\ib-\half\eb_y}^{1D}$ that has been updated to account for the transverse fluxes in the z direction'', using:
+\begin{eqnarray}
+u_{L,\ib-\half\eb_y}^{y|z} &=& u_{L,\ib-\half\eb_y}^{1D} - \frac{\dt}{6h}\left(w_{\ib-\eb_y+\half\eb_z}^{1D}+w_{\ib-\eb_y-\half\eb_z}^{1D}\right)\left(u_{\ib-\eb_y+\half\eb_z}^{1D}-u_{\ib-\eb_y-\half\eb_z}^{1D}\right), \\
+u_{R,\ib-\half\eb_y}^{y|z} &=& u_{R,\ib-\half\eb_y}^{1D} - \frac{\dt}{6h}\left(w_{\ib+\half\eb_z}^{1D}+w_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{1D}-u_{\ib-\half\eb_z}^{1D}\right).
+\end{eqnarray}
+Then, upwind based on $v_{\ib-\half\eb_y}^{1D}$:
+\begin{equation}
+u_{\ib-\half\eb_y}^{y|z} =
+\begin{cases}
+\half\left(u_{L,\ib-\half\eb_y}^{y|z} + u_{R,\ib-\half\eb_y}^{y|z}\right), & \left|v_{\ib-\half\eb_y}^{1D}\right| < \epsilon \\
+u_{L,\ib-\half\eb_y}^{y|z}, & v_{\ib-\half\eb_y}^{1D} > 0, \\
+u_{R,\ib-\half\eb_y}^{y|z}, & v_{\ib-\half\eb_y}^{1D} < 0.
 \end{cases}
 \end{equation}
-Use the exact analogous procedure to obtain $v^{\mac,*}_{\ib-\half\eb_y}$.
-%\subsection{3D Cartesian Case}
+We use an analogous procedure to compute five more intemediate states, $u_{\ib-\half\eb_z}^{z|y}, v_{\ib-\half\eb_x}^{x|z}, v_{\ib-\half\eb_z}^{z|x}, w_{\ib-\half\eb_x}^{x|y}$, and $w_{\ib-\half\eb_y}^{y|x}$.
+Then we do a full-dimensional extrapolation to get the MAC velocities at normal faces:
+\begin{eqnarray}
+u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
+&& - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{y|z}-u_{\ib-\eb_x-\half\eb_y}^{y|z}\right) \nonumber \\
+&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_z}^{1D}+v_{\ib-\eb_x-\half\eb_z}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_z}^{z|y}-u_{\ib-\eb_x-\half\eb_z}^{z|y}\right), \\
+u_{R,\ib-\half\eb_x}^{\mac,*} &=& u_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib} \nonumber \\
+&& - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{y|z}-u_{\ib-\half\eb_y}^{y|z}\right) \nonumber \\
+&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{1D}+v_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{z|y}-u_{\ib-\half\eb_z}^{z|y}\right).
+\end{eqnarray}
+Then, we use the Riemann solver given above for the 2D case (equation [\ref{umac Riemann Problem}]) to compute $u_{\ib-\half\eb_x}^{\mac,*}$.
+We use an analogous procedure to obtain $v_{\ib-\half\eb_y}^{\mac,*}$ and $w_{\ib-\half\eb_z}^{\mac,*}$.
 \cleardoublepage
@@
 \begin{eqnarray}
-\rho_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib-\eb_x}^n, \\
-\rho_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib}^n.
+\rho_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib-\eb_x}^n, \label{varden rho_L^1D}\\
+\rho_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib}^n. \label{varden rho_R^1D}
 \end{eqnarray}
@@
 \end{equation}
-Now we do a full-dimensional extrapolation of $\rho$ to each face.  The extrapolation of $\rho$ to x-faces is:
+The procedure to obtain $\rho_{\ib-\half\eb_y}^{1D}$ is completely analogous.  Now we do a full-dimensional extrapolation of $\rho$ to each face.  The extrapolation of $\rho$ to x-faces is:
 \begin{eqnarray}
@@
 \end{equation}
-%\subsection{3D Cartesian Case}
+The procedure to obtain $\rho_{\ib-\half\eb_y}^{\edge}$ is completely analogous.
+\subsection{3D Cartesian Case}
+This is more complicated than the 2D case because we include corner coupling.
+We compute $\rho_{\ib-\half\eb_x}^{1D}$, $\rho_{\ib-\half\eb_y}^{1D}$, and $\rho_{\ib-\half\eb_z}^{1D}$ in a completely analogous manner to equations (\ref{varden rho_L^1D}) and (\ref{varden rho_R^1D}).
 \end{document}

### 2008-04-18 docs/Godunov_notes/paper.tex
added notes on conservative vs non-conservative s^edge prediction in varden

@@
 \end{equation}
-The extrapolation of $u$ to r-faces is completely analogous to the x-faces:
+The extrapolation of $u$ to r-faces is analogous to the extrapolation to x-faces:
 \begin{eqnarray}
@@
 \end{equation}
-The full-dimensional extrapolation of $s$ to r-faces is completely analogous to the x-faces:
+The full-dimensional extrapolation of $s$ to r-faces is analogous to the extrapolation to x-faces:
 \begin{eqnarray}
@@
 \end{eqnarray}
-The normal velocity Riemann problem is:
+We obtain the normal velocity using the Riemann problem:
 \begin{equation}
@@
 \end{equation}
-Do the exact analogous thing to get both components of velocity at the y-face, $\Ub_{\ib-\half\eb_y}^{1D}$. \\
+We perform analogous operations to compute both components of velocity at the y-faces, $\Ub_{\ib-\half\eb_y}^{1D}$. \\
 
-Now we do a full-dimensional extrapolation to get only the normal velocity at each face:
+Now we do a full-dimensional extrapolation to get the MAC velocity at the x-faces (note that we only compute the normal components):
 \begin{eqnarray}
@@
 \end{eqnarray}
-Then, solve a Riemann problem:
+Then, we solve a Riemann problem:
 \begin{equation}
@@
 \end{equation}
-Use the exact analogous procedure to obtain $v^{\mac,*}_{\ib-\half\eb_y}$.
+We perform analogous operations to compute the normal velocity at the y-faces, $v^{\mac,*}_{\ib-\half\eb_y}$.
 \subsection{3D Cartesian Case}
 This is more complicated than the 2D case because we include corner coupling.
-We compute $\Ub_{\ib-\half\eb_x}, \Ub_{\ib-\half\eb_y}$, and $\Ub_{\ib-\half\eb_z}$ in an analogous manner as equations (\ref{varden U_L^1D})-(\ref{Transverse Velocity Riemann Problem}).
-Then, we compute an intermediate state, $u_{\ib-\half\eb_y}^{y|z}$, which is described as ``state $u_{\ib-\half\eb_y}^{1D}$ that has been updated to account for the transverse fluxes in the z direction'', using:
+We compute $\Ub_{\ib-\half\eb_x}^{1D}, \Ub_{\ib-\half\eb_y}^{1D}$, and $\Ub_{\ib-\half\eb_z}^{1D}$ in an analogous manner as equations (\ref{varden U_L^1D})-(\ref{Transverse Velocity Riemann Problem}).
+Then, we compute an intermediate state, $u_{\ib-\half\eb_y}^{y|z}$, which is described as ``state $u_{\ib-\half\eb_y}^{1D}$ that has been updated to account for the transverse derivatives in the z direction'', using:
 \begin{eqnarray}
@@
 && - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{y|z}-u_{\ib-\half\eb_y}^{y|z}\right) \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{1D}+v_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{z|y}-u_{\ib-\half\eb_z}^{z|y}\right).
+&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{1D}+w_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{z|y}-u_{\ib-\half\eb_z}^{z|y}\right).
 \end{eqnarray}
@@
 To compute $\Ub^{\edge}$, {\tt VARDEN} uses the exact same algorithm as the $s^{\edge}$ case in {\tt MAESTRO}.
-To the algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.
+The algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.
+Here, $s$ is used in place of either $\rho, u, v$, or $w$ (in 3D).
 \subsection{2D Cartesian Case}
@@
 \begin{eqnarray}
-\rho_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib-\eb_x}^n, \label{varden rho_L^1D}\\
-\rho_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \rho_{\ib}^n. \label{varden rho_R^1D}
+s_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib-\eb_x}^n, \label{varden s_L^1D}\\
+s_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib}^n. \label{varden s_R^1D}
 \end{eqnarray}
@@
 \begin{equation}
-\rho_{\ib-\half\eb_x}^{1D} =
+s_{\ib-\half\eb_x}^{1D} =
 \begin{cases}
-\half\left(\rho_{L,\ib-\half\eb_x}^{1D} + \rho_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-\rho_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-\rho_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
+\half\left(s_{L,\ib-\half\eb_x}^{1D} + s_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+s_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
+s_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
 \end{cases}
 \end{equation}
-The procedure to obtain $\rho_{\ib-\half\eb_y}^{1D}$ is completely analogous.  Now we do a full-dimensional extrapolation of $\rho$ to each face.  The extrapolation of $\rho$ to x-faces is:
+We use an analogous procedure to obtain $s_{\ib-\half\eb_y}^{1D}$.  Now we do a full-dimensional extrapolation of $s$ to each face.  The extrapolation of a ``conserved'' $s$ to x-faces is:
 \begin{eqnarray}
-\rho_{L,\ib-\half\eb_x}^{\edge} &=& \rho_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib-\eb_x} - (\rho\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} \nonumber \\
-&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right), \nonumber \\
+s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib-\eb_x} - (s\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} \nonumber \\
+&&- \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right), \nonumber \\
 && \\
-\rho_{R,\ib-\half\eb_x}^{\edge} &=& \rho_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{\ib} - (\rho\nabla\cdot\Ub^{\mac})_{\ib} \nonumber \\
-&&- \frac{\dt}{2h}\left[(\rho^{1D} v^{\mac})_{\ib+\half\eb_y} - (\rho^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2h}\rho_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right).
+s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib} - (s\nabla\cdot\Ub^{\mac})_{\ib} \nonumber \\
+&&- \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right).
+\end{eqnarray}
+The extrapolation of a ``non-conserved'' $s$ to x-faces is:
+\begin{eqnarray}
+s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib-\eb_x} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right), \nonumber \\
+&& \\
+s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right).
 \end{eqnarray}
@@
 \begin{equation}
-\rho_{\ib-\half\eb_x}^{\edge} =
+s_{\ib-\half\eb_x}^{\edge} =
 \begin{cases}
-\half\left(\rho_{L,\ib-\half\eb_x}^{\edge} + \rho_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-\rho_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-\rho_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
+\half\left(s_{L,\ib-\half\eb_x}^{\edge} + s_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+s_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
+s_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
 \end{cases}
 \end{equation}
-The procedure to obtain $\rho_{\ib-\half\eb_y}^{\edge}$ is completely analogous.
+We use an analogous procedure to compute $s_{\ib-\half\eb_y}^{\edge}$.
 \subsection{3D Cartesian Case}
 This is more complicated than the 2D case because we include corner coupling.
-We compute $\rho_{\ib-\half\eb_x}^{1D}$, $\rho_{\ib-\half\eb_y}^{1D}$, and $\rho_{\ib-\half\eb_z}^{1D}$ in a completely analogous manner to equations (\ref{varden rho_L^1D}) and (\ref{varden rho_R^1D}).
+We compute $s_{\ib-\half\eb_x}^{1D}$, $s_{\ib-\half\eb_y}^{1D}$, and $s_{\ib-\half\eb_z}^{1D}$ in an analogous manner to equations (\ref{varden s_L^1D}) and (\ref{varden s_R^1D}).
 \end{document}

### 2008-04-18 docs/Godunov_notes/paper.tex
typo in upwinding equation in U^MAC for varden
fixed a few equation typos, some formatting/cleanup
fixed a typo and re-aligned long equations to make them more readable

@@
 % Margins
-\usepackage[lmargin=1.0in,rmargin=1.0in,tmargin=0.75in,bmargin=0.75in]{geometry}
+\usepackage[lmargin=0.5in,rmargin=0.5in,tmargin=0.75in,bmargin=0.75in]{geometry}
 
@@
 \begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\trans} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\trans} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans}\right| < \epsilon, \\
+0, & \left(u_{L,\ib-\half\eb_x}^{\trans} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans}\right| < \epsilon, \\
 u_{L,\ib-\half\eb_x}^{\trans}, & u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans} > 0, \\
@@
 \begin{cases}
-0, & \left(w_L^{\trans} \le 0 ~ {\rm AND} ~ w_R^{\trans} \ge 0\right) ~ {\rm OR} ~ \left|w_L^{\trans} + w_R^{\trans}\right| < \epsilon, \\
+0, & \left(w_L^{\trans} \le 0 ~~ {\rm AND} ~~ w_R^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|w_L^{\trans} + w_R^{\trans}\right| < \epsilon, \\
 \wt_{L,\ib-\half\eb_r}^{\trans}, & w_L^{\trans} + w_R^{\trans} > 0, \\
@@
 \begin{cases}
-0, & \left(v_{L,\ib-\half\eb_y}^{\trans} \le 0 ~ {\rm AND} ~ v_{R,\ib-\half\eb_y}^{\trans} \ge 0\right) ~ {\rm OR} ~ \left|v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans}\right| < \epsilon, \\
+0, & \left(v_{L,\ib-\half\eb_y}^{\trans} \le 0 ~~ {\rm AND} ~~ v_{R,\ib-\half\eb_y}^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans}\right| < \epsilon, \\
 v_{L,\ib-\half\eb_y}^{\trans}, & v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans} > 0, \\
@@
 \begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
-u_{R,\ib-\half\eb_x}^{\mac,*} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right).
+u_{L,\ib-\half\eb_x}^{\mac,*} = u_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
+u_{R,\ib-\half\eb_x}^{\mac,*} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
 \end{eqnarray}
@@
 \begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
+0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
 u_{L,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} > 0, \\
@@
 \begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{\mac,*} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{w,\ib-\eb_r} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)(w_{0,\ib-\half\eb_r} - w_{0,\ib-\frac{3}{2}\eb_r}),\label{wt_L Extrapolation} \\
-\wt_{R,\ib-\half\eb_r}^{\mac,*} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)(w_{0,\ib+\half\eb_r} - w_{0,\ib-\half\eb_r}).\label{wt_R Extrapolation}
+\wt_{L,\ib-\half\eb_r}^{\mac,*} = \wt_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
+&-& \frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)(w_{0,\ib-\half\eb_r} - w_{0,\ib-\frac{3}{2}\eb_r}) + \frac{\dt}{2}f_{w,\ib-\eb_r},\label{wt_L Extrapolation} \\
+\wt_{R,\ib-\half\eb_r}^{\mac,*} = \wt_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
+&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)(w_{0,\ib+\half\eb_r} - w_{0,\ib-\half\eb_r}) + \frac{\dt}{2}f_{w,\ib}.\label{wt_R Extrapolation}
 \end{eqnarray}
@@
 \begin{cases}
-0, & \left(w_L^{\mac,*} \le 0 ~ {\rm AND} ~ w_R^{\mac,*} \ge 0\right) ~ {\rm OR} ~ \left|w_L^{\mac,*} + w_R^{\mac,*}\right| < \epsilon, \\
+0, & \left(w_L^{\mac,*} \le 0 ~~ {\rm AND} ~~ w_R^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|w_L^{\mac,*} + w_R^{\mac,*}\right| < \epsilon, \\
 \wt_{L,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} > 0, \\
@@
 \begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib-\eb_x} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\mac}+w_{\ib-\eb_x-\half\eb_r}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_r}^{1D} - s_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
-s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\mac}+w_{\ib-\half\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_r}^{1D} - s_{\ib-\half\eb_r}^{1D}\right).
+s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\mac}+w_{\ib-\eb_x-\half\eb_r}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_r}^{1D} - s_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
+s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\mac}+w_{\ib-\half\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_r}^{1D} - s_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}.
 \end{eqnarray}
@@
 \begin{eqnarray}
-s_{L,\ib-\half\eb_r}^{\edge} &=& s_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib-\eb_r}^n + \frac{\dt}{2}f_{s,\ib-\eb_r} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\mac}+u_{\ib-\half\eb_x-\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_x-\eb_r}^{1D} - s_{\ib-\half\eb_x-\eb_r}^{1D}\right),  \\
-s_{R,\ib-\half\eb_r}^{\edge} &=& s_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib}^n + \frac{\dt}{2}f_{s,\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\mac}+u_{\ib-\half\eb_x}^{\mac}\right)\left(s_{\ib+\half\eb_x}^{1D} - s_{\ib-\half\eb_x}^{1D}\right). 
+s_{L,\ib-\half\eb_r}^{\edge} = s_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib-\eb_r}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\mac}+u_{\ib-\half\eb_x-\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_x-\eb_r}^{1D} - s_{\ib-\half\eb_x-\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_r},  \\
+s_{R,\ib-\half\eb_r}^{\edge} = s_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\mac}+u_{\ib-\half\eb_x}^{\mac}\right)\left(s_{\ib+\half\eb_x}^{1D} - s_{\ib-\half\eb_x}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}. 
 \end{eqnarray}
@@
 \begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\edge} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib-\eb_x}^n + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right), \\
-u_{R,\ib-\half\eb_x}^{\edge} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right).
+u_{L,\ib-\half\eb_x}^{\edge} = u_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib-\eb_x}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
+u_{R,\ib-\half\eb_x}^{\edge} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
 \end{eqnarray}
@@
 \begin{eqnarray}
-u_{L,\ib-\half\eb_r}^{\edge} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib-\eb_r}^n + \frac{\dt}{2}f_{u,\ib-\eb_r} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_x-\eb_r}^{1D} - u_{\ib-\half\eb_x-\eb_r}^{1D}\right),  \\
-u_{R,\ib-\half\eb_r}^{\edge} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib}^n + \frac{\dt}{2}f_{u,\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(u_{\ib+\half\eb_x}^{1D} - u_{\ib-\half\eb_x}^{1D}\right). 
+u_{L,\ib-\half\eb_r}^{\edge} = u_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib-\eb_r}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_x-\eb_r}^{1D} - u_{\ib-\half\eb_x-\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_r},  \\
+u_{R,\ib-\half\eb_r}^{\edge} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(u_{\ib+\half\eb_x}^{1D} - u_{\ib-\half\eb_x}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}. 
 \end{eqnarray}
@@
 \begin{eqnarray}
-\wt_{L,\ib-\half\eb_x}^{\edge} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib-\eb_x}^n + \frac{\dt}{2}f_{w,\ib-\eb_x} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(\wt_{\ib-\eb_x+\half\eb_r}^{1D} - \wt_{\ib-\eb_x-\half\eb_r}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{4h}\left(\wt_{\ib-\eb_x+\half\eb_r}^{\trans}+\wt_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(w_{0,\ib-\eb_x+\half\eb_r}-w_{0,\ib-\eb_x-\half\eb_r}\right), \label{wt_L x-edge} \\
-\wt_{R,\ib-\half\eb_x}^{\edge} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_r}^{1D} - \wt_{\ib-\half\eb_r}^{1D}\right) \nonumber \\
-&& - \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right). \label{wt_R x-edge}
+\wt_{L,\ib-\half\eb_x}^{\edge} = \wt_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib-\eb_x}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(\wt_{\ib-\eb_x+\half\eb_r}^{1D} - \wt_{\ib-\eb_x-\half\eb_r}^{1D}\right) \nonumber \\
+&-& \frac{\dt}{4h}\left(\wt_{\ib-\eb_x+\half\eb_r}^{\trans}+\wt_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(w_{0,\ib-\eb_x+\half\eb_r}-w_{0,\ib-\eb_x-\half\eb_r}\right) + \frac{\dt}{2}f_{w,\ib-\eb_x}, \label{wt_L x-edge} \\
+\wt_{R,\ib-\half\eb_x}^{\edge} = \wt_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_r}^{1D} - \wt_{\ib-\half\eb_r}^{1D}\right) \nonumber \\
+&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right) + \frac{\dt}{2}f_{w,\ib}. \label{wt_R x-edge}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{\edge} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}f_{w,\ib-\eb_r} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
-&& -
-\frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)\left(w_{0,\ib-\half\eb_r}-w_{0,\ib-\frac{3}{2}\eb_r}\right), \label{wt_L r-edge} \\
-\wt_{R,\ib-\half\eb_r}^{\edge} &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}f_{w,\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
-&& -
-\frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right). \label{wt_R r-edge}
+\wt_{L,\ib-\half\eb_r}^{\edge} = \wt_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib-\eb_r}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
+&-& \frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)\left(w_{0,\ib-\half\eb_r}-w_{0,\ib-\frac{3}{2}\eb_r}\right) + \frac{\dt}{2}f_{w,\ib-\eb_r}, \label{wt_L r-edge} \\
+\wt_{R,\ib-\half\eb_r}^{\edge} = \wt_{\ib}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib}^n \nonumber \\
+&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
+&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right) + \frac{\dt}{2}f_{w,\ib}. \label{wt_R r-edge}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\Ub_{L,\ib-\half\eb_x}^{1D} &=& \Ub_{\ib-\eb_x} + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right){\rm max}(0,\Delta_x\Ub_{\ib-\eb_x}), \label{varden U_L^1D} \\
-\Ub_{R,\ib-\half\eb_x}^{1D} &=& \Ub_{\ib} + \left(\half - \frac{\dt}{2h}u_{\ib}\right){\rm min}(0,\Delta_x\Ub_{\ib}).
+u_{L,\ib-\half\eb_x}^{1D} &=& u_{\ib-\eb_x} + \left[\half - \frac{\dt}{2h}{\rm max}(0,u_{\ib-\eb_x})\right]\Delta_xu_{\ib-\eb_x}, \label{varden U_L^1D} \\
+u_{R,\ib-\half\eb_x}^{1D} &=& u_{\ib} + \left[\half - \frac{\dt}{2h}{\rm min}(0,u_{\ib})\right]\Delta_xu_{\ib}.
+\end{eqnarray}
+\begin{eqnarray}
+v_{L,\ib-\half\eb_x}^{1D} &=& v_{\ib-\eb_x} + \left[\half - \frac{\dt}{2h}{\rm max}(0,v_{\ib-\eb_x})\right]\Delta_xv_{\ib-\eb_x}, \label{varden U_L^1D} \\
+v_{R,\ib-\half\eb_x}^{1D} &=& v_{\ib} + \left[\half - \frac{\dt}{2h}{\rm min}(0,v_{\ib})\right]\Delta_xv_{\ib}.
 \end{eqnarray}
@@
 \begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{1D} \le 0 ~ {\rm AND} ~ u_{R,\ib-\half\eb_x}^{1D} \ge 0\right) ~ {\rm OR} ~ \left|u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}\right| < \epsilon, \\
+0, & \left(u_{L,\ib-\half\eb_x}^{1D} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{1D} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}\right| < \epsilon, \\
 u_{L,\ib-\half\eb_x}^{1D}, & u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D} > 0, \\
@@
 \begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} &=& \Ub_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
-&-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
-u_{R,\ib-\half\eb_x}^{\mac,*} &=& \Ub_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib} \nonumber \\
-&-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right).
+u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \nonumber \\
+&& \\
+u_{R,\ib-\half\eb_x}^{\mac,*} &=& u_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
 \end{eqnarray}
@@
 We use an analogous procedure to compute five more intemediate states, $u_{\ib-\half\eb_z}^{z|y}, v_{\ib-\half\eb_x}^{x|z}, v_{\ib-\half\eb_z}^{z|x}, w_{\ib-\half\eb_x}^{x|y}$, and $w_{\ib-\half\eb_y}^{y|x}$.
-Then we do a full-dimensional extrapolation to get the MAC velocities at normal faces:
+Then, we do a full-dimensional extrapolation to get the MAC velocities at normal faces:
 \begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib-\eb_x} \nonumber \\
-&& - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{y|z}-u_{\ib-\eb_x-\half\eb_y}^{y|z}\right) \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_z}^{1D}+v_{\ib-\eb_x-\half\eb_z}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_z}^{z|y}-u_{\ib-\eb_x-\half\eb_z}^{z|y}\right), \\
-u_{R,\ib-\half\eb_x}^{\mac,*} &=& u_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{u,\ib} \nonumber \\
-&& - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{y|z}-u_{\ib-\half\eb_y}^{y|z}\right) \nonumber \\
-&& - \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{1D}+w_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{z|y}-u_{\ib-\half\eb_z}^{z|y}\right).
+u_{L,\ib-\half\eb_x}^{\mac,*} = u_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{y|z}-u_{\ib-\eb_x-\half\eb_y}^{y|z}\right) \nonumber \\
+&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_z}^{1D}+v_{\ib-\eb_x-\half\eb_z}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_z}^{z|y}-u_{\ib-\eb_x-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
+u_{R,\ib-\half\eb_x}^{\mac,*} = u_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{y|z}-u_{\ib-\half\eb_y}^{y|z}\right) \nonumber \\
+&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{1D}+w_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{z|y}-u_{\ib-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{u,\ib}.
 \end{eqnarray}
@@
 \begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib-\eb_x} - (s\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} \nonumber \\
-&&- \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right), \nonumber \\
-&& \\
-s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib} - (s\nabla\cdot\Ub^{\mac})_{\ib} \nonumber \\
-&&- \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right).
+s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2}s_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right) \nonumber \\
+&-& \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
+s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2}s_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right) \nonumber \\
+&-& \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2}f_{s,\ib}.
 \end{eqnarray}
@@
 \begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib-\eb_x} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right), \nonumber \\
+s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \nonumber \\
 && \\
-s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} + \frac{\dt}{2}f_{s,\ib} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right).
+s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}.
 \end{eqnarray}

### 2008-04-18 docs/Godunov_notes/paper.tex
a couple small equation typos

@@
 \end{equation}
-The extrapolation of $u$ to r-faces is analogous to the extrapolation to x-faces:
+The extrapolation of $s$ to r-faces is analogous to the extrapolation to x-faces:
 \begin{eqnarray}
@@
 u_{L,\ib-\half\eb_x}^{\mac,*} = u_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{y|z}-u_{\ib-\eb_x-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_z}^{1D}+v_{\ib-\eb_x-\half\eb_z}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_z}^{z|y}-u_{\ib-\eb_x-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
+&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_z}^{1D}+w_{\ib-\eb_x-\half\eb_z}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_z}^{z|y}-u_{\ib-\eb_x-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
 u_{R,\ib-\half\eb_x}^{\mac,*} = u_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{y|z}-u_{\ib-\half\eb_y}^{y|z}\right) \nonumber \\
@@
 This is more complicated than the 2D case because we include corner coupling.
-We compute $s_{\ib-\half\eb_x}^{1D}$, $s_{\ib-\half\eb_y}^{1D}$, and $s_{\ib-\half\eb_z}^{1D}$ in an analogous manner to equations (\ref{varden s_L^1D}) and (\ref{varden s_R^1D}).
+We first compute $s_{\ib-\half\eb_x}^{1D}$, $s_{\ib-\half\eb_y}^{1D}$, and $s_{\ib-\half\eb_z}^{1D}$ in an analogous manner to equations (\ref{varden s_L^1D}) and (\ref{varden s_R^1D}).
 \end{document}

### 2008-04-18 docs/Godunov_notes/paper.tex
Finished the 3D Cartesian case for s^edge in VARDEN

@@
 \end{equation}
-We use an analogous procedure to obtain $s_{\ib-\half\eb_y}^{1D}$.  Now we do a full-dimensional extrapolation of $s$ to each face.  The extrapolation of a ``conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2}s_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right) \nonumber \\
-&-& \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2}s_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right) \nonumber \\
-&-& \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\half\eb_y}\right] + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-The extrapolation of a ``non-conserved'' $s$ to x-faces is:
+We use an analogous procedure to obtain $s_{\ib-\half\eb_y}^{1D}$.  Now we do a full-dimensional extrapolation of $s$ to each face.  The extrapolation of a ``non-conserved'' $s$ to x-faces is:
 \begin{eqnarray}
@@
 s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}.
+\end{eqnarray}
+The extrapolation of a ``conserved'' $s$ to x-faces is:
+\begin{eqnarray}
+s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] \nonumber \\
+&-& \frac{\dt}{2}s_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
+s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\half\eb_y}\right] \nonumber \\
+&-& \frac{\dt}{2}s_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right) + \frac{\dt}{2}f_{s,\ib}.
 \end{eqnarray}
@@
 s_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
+\end{cases}\label{varden s^edge upwind}
 \end{equation}
@@
 We first compute $s_{\ib-\half\eb_x}^{1D}$, $s_{\ib-\half\eb_y}^{1D}$, and $s_{\ib-\half\eb_z}^{1D}$ in an analogous manner to equations (\ref{varden s_L^1D}) and (\ref{varden s_R^1D}).
+Then we compute six intermediate states, $s_{\ib-\half\eb_x}^{x|y}, s_{\ib-\half\eb_x}^{x|z}, s_{\ib-\half\eb_y}^{y|x}, s_{\ib-\half\eb_y}^{y|z}, s_{\ib-\half\eb_z}^{z|x}$, and $s_{\ib-\half\eb_z}^{z|y}$.
+For the ``non-conservative case'', we use, for example:
+\begin{eqnarray}
+s_{L,\ib-\half\eb_x}^{x|y} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{6h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} + v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
+s_{R,\ib-\half\eb_x}^{x|y} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{6h}\left(v_{\ib+\half\eb_y}^{\mac} + v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right).
+\end{eqnarray}
+For the ``conservative'' case, we use, for example:
+\begin{eqnarray}
+s_{L,\ib-\half\eb_x}^{x|y} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{3h}\left[(sv^{\mac})_{\ib-\eb_x+\half\eb_y}-(sv^{\mac})_{\ib-\eb_x-\half\eb_y}\right], \\
+s_{R,\ib-\half\eb_x}^{x|y} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{3h}\left[(sv^{\mac})_{\ib+\half\eb_y}-(sv^{\mac})_{\ib-\half\eb_y}\right].
+\end{eqnarray}
+Then, we upwind based on $u^{\mac}$:
+\begin{equation}
+s_{\ib-\half\eb_x}^{\edge} =
+\begin{cases}
+\half\left(s_{L,\ib-\half\eb_x}^{x|y} + s_{R,\ib-\half\eb_x}^{x|y}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
+s_{L,\ib-\half\eb_x}^{x|y}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
+s_{R,\ib-\half\eb_x}^{x|y}, & u^{\mac}_{\ib-\half\eb_x} < 0.
+\end{cases}
+\end{equation}
+We use an analogous procedure to compute the other five intermediate states.
+Now we do a full-dimensional extrapolation of $s$ to each face.
+The extrapolation of a ``non-conserved'' $s$ to x-faces is:
+\begin{eqnarray}
+s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{y|z}-s_{\ib-\eb_x-\half\eb_y}^{y|z}\right) \nonumber \\
+&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_z}^{\mac}+w_{\ib-\eb_x-\half\eb_z}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_z}^{z|y}-s_{\ib-\eb_x-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
+s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{y|z}-s_{\ib-\half\eb_y}^{y|z}\right) \nonumber \\
+&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{\mac}+w_{\ib-\half\eb_z}^{\mac}\right)\left(s_{\ib+\half\eb_z}^{z|y}-s_{\ib-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{s,\ib}.
+\end{eqnarray}
+The extrapolation of a ``conserved'' $s$ to x-faces is:
+\begin{eqnarray}
+s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{y|z}v^{\mac})_{\ib-\eb_x+\half\eb_y}-({s^{y|z}v^{\mac})_{\ib-\eb_x-\half\eb_y}}\right] \nonumber \\
+&-& \frac{\dt}{2h}\left[(s^{z|y}w^{\mac})_{\ib-\eb_x+\half\eb_z}-({s^{z|y}w^{\mac})_{\ib-\eb_x-\half\eb_z}}\right] \nonumber \\
+&-& \frac{\dt}{2}s_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}-v_{\ib-\eb_x-\half\eb_y}^{\mac}+w_{\ib-\eb_x+\half\eb_z}^{\mac}-w_{\ib-\eb_x-\half\eb_z}^{\mac}\right) \nonumber \\
+&+& \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
+s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{y|z}v^{\mac})_{\ib+\half\eb_y}-({s^{y|z}v^{\mac})_{\ib-\half\eb_y}}\right] \nonumber \\
+&-& \frac{\dt}{2h}\left[(s^{z|y}w^{\mac})_{\ib+\half\eb_z}-({s^{z|y}w^{\mac})_{\ib-\half\eb_z}}\right] \nonumber \\
+&-& \frac{\dt}{2}s_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac}-v_{\ib-\half\eb_y}^{\mac}+w_{\ib+\half\eb_z}^{\mac}-w_{\ib-\half\eb_z}^{\mac}\right) \nonumber \\
+&+& \frac{\dt}{2}f_{s,\ib}.
+\end{eqnarray}
+Then we upwind based on $u^{\mac}$, as in equation (\ref{varden s^edge upwind}).  We use an analogous procedure to compute both $s_{\ib-\half\eb_y}^{\edge}$ and $s_{\ib-\half\eb_z}$.
 \end{document}

### 2008-04-18 docs/Godunov_notes/paper.tex
more notes about VARDEN and MAESTRO differences

@@
 \item Concerning both equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}), in the code we are omitting the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ (by using $\wt$ rather than $w$ as the advection speed) and $(\Ubt\cdot\eb_r)$ (by omitting the term altogether).
-\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$.
+\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction.
+However, this term is present in the cell-update.
 \end{enumerate}
@@
 \begin{enumerate}
-\item {\tt VARDEN} does not use the concept of $w_0$, so there is no such thing as $\Ubt$ either.  {\tt VARDEN} only does Cartesian problems.
+\item {\tt VARDEN} does not use the concept of $w_0$, so $\Ubt$ does not exist.
+\item {\tt VARDEN} only does Cartesian problems.
 \item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$, but rather in {\tt advance\_premac}, the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
-\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$.
-\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, but not $\Ub^{\edge}$
-\item The transverse derivative terms for $\rho$ in {\tt VARDEN} are done in a conservative formulation, rather than advective.
+\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$, whereas {\tt MAESTRO} does not.
+\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, and a ``non-FPU'' algorithm for $\Ub^{\edge}$
+\item {\tt VARDEN} treats $\rho$ in a conservative manner, whereas {\tt MAESTRO} treats $\rho'$ in a non-conservative manner.
+Therefore, {\tt MAESTRO} accounts for this difference by modifying the forcing term in {\tt modify\_scal\_force}
+\item {\tt MAESTRO} does not corner couple for 3D problems.
 \end{enumerate}
@@
 \cleardoublepage
-\section{Computing $\rho^{'\edge}, X_k^{\edge}, h^{\edge},$ and $T^{\edge}$ in {\tt MAESTRO}}
-In {\tt scalar\_advance}, we call {\tt make\_edge\_scal} to compute $\rho^{'\edge}, X_k^{\edge}, h^{\edge},$ and $T^{\edge}$ at each edge.
+\section{Computing $\rho^{'\edge}, X_k^{\edge}$, and $(\rho h)^{'\edge}/h^{\edge}/T^{\edge}$ in {\tt MAESTRO}}
+In {\tt scalar\_advance}, we call {\tt make\_edge\_scal} to compute $\rho^{'\edge}, X_k^{\edge}$, and $(\rho h)^{'\edge}/h^{\edge}/T^{\edge}$ at each edge.
 The procedure is the same for each scalar, so we shall simply denote the scalar as $s$.
-We always need to compute $\rho'$ and $X_k$ to edges.
+We always need to compute $\rho'$ and $X_k$ to edges, and the choice of energy prediction is as follows:
 \begin{itemize}
@@
 f_h &=& \frac{1}{\rho}\left(\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T \right), \\
-f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext} + \nabla\cdot\kth\nabla T \right\}.\nonumber \\
+f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext} + \nabla\cdot\kth\nabla T \right\}.
 \end{eqnarray}
@@
 \begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \nonumber \\
-&& \\
+u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
 u_{R,\ib-\half\eb_x}^{\mac,*} &=& u_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
@@
 \begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \nonumber \\
-&& \\
+s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
 s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}.

### 2008-04-21 docs/Godunov_notes/paper.tex
updated to account for latest cvs checkins.  Document is now consistent with code

@@
 \begin{enumerate}
-\item Concerning both equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}), in the code we are omitting the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ (by using $\wt$ rather than $w$ as the advection speed) and $(\Ubt\cdot\eb_r)$ (by omitting the term altogether).
+\item Concerning equations (\ref{Radial utrans_L}), (\ref{Radial utrans_R}), (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}), in the code we are omitting the term in equation (\ref{Perturbational Velocity Equation}) proportional to $(\Ubt\cdot\eb_r)$.
 \item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction.
@@
 \begin{enumerate}
-\item We need to re-examine when to include the term proportional to $(\Ubt\cdot\eb_r)$ in equation (\ref{Perturbational Velocity Equation}) when we do a 1D extrapolation of the radial velocity to r-faces.
-For example, equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}) include it, but equations (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}) do not.
-\item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$ rather than the average of the neighboring $\wt^{\trans}$.
+\item Concerning equations (\ref{Radial utrans_L}), (\ref{Radial utrans_R}), (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}), do we really need to include the term proportional to $(\Ubt\cdot\eb_r)$?
+\item Concerning (\ref{Force Term}), do we really need to include the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction?
+If so, do we use the $\pi_0$ or $w_0$ formulation (refer to Paper III equation [38])?
+\item Concerning equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$ rather than the average of the neighboring $\wt^{\trans}$.
 \end{enumerate}
@@
 \begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{1D} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n, \label{1D wt_L extrap} \\
-\wt_{R,\ib-\half\eb_r}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n. \label{1D wt_R extrap}
+\wt_{L,\ib-\half\eb_r}^{1D} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r}, \label{1D wt_L extrap} \\
+\wt_{R,\ib-\half\eb_r}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.
+\label{1D wt_R extrap}
 \end{eqnarray}

### 2008-04-21 docs/Godunov_notes/paper.tex
More details in the "Questions" section.

@@
 \item Concerning (\ref{Force Term}), do we really need to include the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction?
-If so, do we use the $\pi_0$ or $w_0$ formulation (refer to Paper III equation [38])?
+\item Do we want to do characteristic tracing to compute $\Ub^{\trans}$, analgous to how {\tt VARDEN} uses characteristic tracing to compute $\Ub^{\mac}$?
 \item Concerning equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$ rather than the average of the neighboring $\wt^{\trans}$.
+\item When we compute $\Ub^{\mac,*}$, do we want to replace all instances of ``trans'' velocities with ``MAC'' velocities?
+Doing this is simply a matter of replacing the call to {\tt make\_edge\_state} with {\tt make\_edge\_scal}.
 \end{enumerate}
@@
 \item {\tt VARDEN} only does Cartesian problems.
-\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$, but rather in {\tt advance\_premac}, the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
+\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$ in the computation of $\Ub^{\mac}$, so therefore the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
 \item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$, whereas {\tt MAESTRO} does not.

### 2008-04-21 docs/Godunov_notes/paper.tex
removed an item from the questions list - I believe we are happy with averaging the w^trans or w^mac terms for the wtilde dw_0/dr rather than using the CC velocity.

@@
 \item Do we want to do characteristic tracing to compute $\Ub^{\trans}$, analgous to how {\tt VARDEN} uses characteristic tracing to compute $\Ub^{\mac}$?
-\item Concerning equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$ rather than the average of the neighboring $\wt^{\trans}$.
 \item When we compute $\Ub^{\mac,*}$, do we want to replace all instances of ``trans'' velocities with ``MAC'' velocities?

### 2008-05-29 docs/lo_density/paper.tex
new document describing the difference between anelastic_cutoff, base_cutoff_density, and burning_cutoff_density

--- 2008-04-21 docs/Godunov_notes/paper.tex
@@
 
-\title{Notes on the Godunov Step in {\tt MAESTRO} and {\tt VARDEN}}
+\title{Notes on the low density parameters in {\tt MAESTRO}}
 
@@
 
-These are working notes for the Godunov step in {\tt MAESTRO} and {\tt VARDEN}.
-\section{Differences Between These Notes and the Code}
-\begin{enumerate}
-\item Concerning equations (\ref{Radial utrans_L}), (\ref{Radial utrans_R}), (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}), in the code we are omitting the term in equation (\ref{Perturbational Velocity Equation}) proportional to $(\Ubt\cdot\eb_r)$.
-\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction.
-However, this term is present in the cell-update.
-\end{enumerate}
-\section{Questions}
-\begin{enumerate}
-\item Concerning equations (\ref{Radial utrans_L}), (\ref{Radial utrans_R}), (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}), do we really need to include the term proportional to $(\Ubt\cdot\eb_r)$?
-\item Concerning (\ref{Force Term}), do we really need to include the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction?
-\item Do we want to do characteristic tracing to compute $\Ub^{\trans}$, analgous to how {\tt VARDEN} uses characteristic tracing to compute $\Ub^{\mac}$?
-\item When we compute $\Ub^{\mac,*}$, do we want to replace all instances of ``trans'' velocities with ``MAC'' velocities?
-Doing this is simply a matter of replacing the call to {\tt make\_edge\_state} with {\tt make\_edge\_scal}.
-\end{enumerate}
-\section{Differences Between {\tt MAESTRO} and {\tt VARDEN}}
-\begin{enumerate}
-\item {\tt VARDEN} does not use the concept of $w_0$, so $\Ubt$ does not exist.
-\item {\tt VARDEN} only does Cartesian problems.
-\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$ in the computation of $\Ub^{\mac}$, so therefore the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
-\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$, whereas {\tt MAESTRO} does not.
-\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, and a ``non-FPU'' algorithm for $\Ub^{\edge}$
-\item {\tt VARDEN} treats $\rho$ in a conservative manner, whereas {\tt MAESTRO} treats $\rho'$ in a non-conservative manner.
-Therefore, {\tt MAESTRO} accounts for this difference by modifying the forcing term in {\tt modify\_scal\_force}
-\item {\tt MAESTRO} does not corner couple for 3D problems.
-\end{enumerate}
-\cleardoublepage
-\section{Computing $\Ubt^{\trans}$ in {\tt MAESTRO}}
-In {\tt advance\_premac}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.
-These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
-We will only compute the normal component of velocity at each face.\\
-
-The evolution equation for the perturbational velocity is:
-\begin{equation}
-\frac{\partial\Ubt}{\partial t} = -\Ubt\cdot\nabla\Ubt - w_0\frac{\partial\Ubt}{\partial r} - (\Ubt\cdot\eb_r)\frac{\partial w_0}{\partial r}\eb_r - \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Perturbational Velocity Equation}
-\end{equation}
-We are going to use a 1D Taylor series extrapolation in space and time to compute time-centered normal components of $\Ubt^{\trans}$ at face-centers.
-By 1D, we mean that we omit any spatial derivatives that are not in the direction of the extrapolation.
-We first compute ``left/right'' states, and then use a Riemann solver to pick the final state.
-\subsection{2D Cartesian Case}
-For the rest of this paper, for 2D Cartesian problems, $\Ub = (u,w)$ and $\Ubt = (u,\wt)$.  We will use the shorthand $\ib = (x,r)$.\\
-
-For the x-faces, the 1D extrapolation of $u$ is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\trans} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
-u_{R,\ib-\half\eb_x}^{\trans} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
-\end{eqnarray}
-We pick the final edge states using a Riemann solver:
-\begin{equation}
-u^{\trans}_{\ib-\half\eb_x} =
-\begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\trans} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{\trans}, & u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans} > 0, \\
-u_{R,\ib-\half\eb_x}^{\trans}, & u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans} < 0, \\
-\end{cases}
-\end{equation}
-For the r-faces, we choose include the radial derivative terms proportional to $w_0$ and $(\Ubt\cdot\eb_r) = \wt$.
-The term proportional to $w_0$ is included by using $w$ rather than $\wt$ as the advection speed.
-The term proportional to $\wt$ is an additional term.
-For the r-faces, the 1D extrapolation of $\wt$ is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{\trans} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
-\wt_{R,\ib-\half\eb_r}^{\trans} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.\label{Radial utrans_R}
-\end{eqnarray}
-The total radial velocity is defined as:
-\begin{equation}
-w_{\ib}^n = \wt_{\ib}^n + \frac{w_{0,r+\half}+w_{0,r-\half}}{2}.\label{Total Radial Velocity}
-\end{equation}
-The spatial derivatives of $w_0$ are given by the two-point centered difference:
-\begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)_{\ib} = \frac{w_{0,r+\half}-w_{0,r-\half}}{h}.\label{Radial Derivative of w_0}
-\end{equation}
-We pick the final edge states using:
-\begin{equation}
-\wt^{\trans}_{\ib-\half\eb_r} =
-\begin{cases}
-0, & \left(w_L^{\trans} \le 0 ~~ {\rm AND} ~~ w_R^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|w_L^{\trans} + w_R^{\trans}\right| < \epsilon, \\
-\wt_{L,\ib-\half\eb_r}^{\trans}, & w_L^{\trans} + w_R^{\trans} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{\trans}, & w_L^{\trans} + w_R^{\trans} < 0, \\
-\end{cases}
-\end{equation}
-where $w_L^{\trans} = \wt_{L,\ib-\half\eb_r}^{\trans} + w_{0,r-\half}$ and $w_R^{\trans} = \wt_{R,\ib-\half\eb_r}^{\trans} + w_{0,r-\half}$.
-Finally, we call {\tt addw0} to compute $w^{\trans}$ from $\wt^{\trans}$.
-\subsection{3D Cartesian Case}
-For the rest of this paper, for 3D Cartesian problems, $\Ub = (u,v,w)$ and $\Ubt = (u,v,\wt)$.  We will use the shorthand $\ib = (x,y,r)$.\\
-
-The only difference between the 2D and 3D Cartesian cases is that we must also compute $v^{\trans}$.
-The procedure is analogous to computing $u^{\trans}$.
-For y-faces, the 1D extrapolation of $v$ is:
-\begin{eqnarray}
-v_{L,\ib-\half\eb_y}^{\trans} &=& v_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y v_{\ib-\eb_y}^n, \\
-v_{R,\ib-\half\eb_y}^{\trans} &=& v_{\ib}^n - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y v_{\ib}^n,
-\end{eqnarray}
-\begin{equation}
-v^{\trans}_{\ib-\half\eb_y} =
-\begin{cases}
-0, & \left(v_{L,\ib-\half\eb_y}^{\trans} \le 0 ~~ {\rm AND} ~~ v_{R,\ib-\half\eb_y}^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans}\right| < \epsilon, \\
-v_{L,\ib-\half\eb_y}^{\trans}, & v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans} > 0, \\
-v_{R,\ib-\half\eb_y}^{\trans}, & v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans} < 0. \\
-\end{cases}
-\end{equation}
-\subsection{3D Spherical Case}
-For the rest of this paper, for 3D spherical problems, $\Ub = (u,v,w)$ and $\Ubt = (\ut,\vt,\wt)$.  We will use the shorthand $\ib = (x,y,z)$.\\
-
-We first construct $\Ub$ by mapping $w_0$ onto a Cartesian array and adding it to $\Ubt$.
-This is a spherical generalization of equation (\ref{Total Radial Velocity}).
-We also map $\partial w_0/\partial r$ onto a Cartesian array.
-This is a spherical generalization of equation (\ref{Radial Derivative of w_0}).
-Then we use a 1D extrapolation analogous to equations equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}).
-For example, for x-faces:
-\begin{eqnarray}
-\ut_L^{\trans} &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x\ut_{\ib-\eb_x}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib-\eb_x}, \\
-\ut_R^{\trans} &=& \ut_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib}.
-\end{eqnarray}
-Then, we use the same Riemann solver as the Cartesian case to obtain $\ut^{\trans}$.
-To compute $\vt^{\trans}$ (or $\wt^{\trans}$), simply replace all instaces of $\ut$ with $\vt$ (or $\wt$) and all instaces of $x$ with $y$ (or $z$).
-Finally, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
-\cleardoublepage
-\section{Computing $\Ubt^{\mac,*}$ in {\tt MAESTRO}}
-In {\tt advance\_premac}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac,*}$.  We will only compute the normal component of velocity at each face.\\
-
-The force terms are:
-\begin{equation}
-\fb =
-\left(\begin{array}{c}
-f_u \\
-f_w \\
-\end{array}\right)
-= \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Force Term}
-\end{equation}
-\subsection{2D Cartesian Case}
-We begin by computing a time-centered 1D extrapolation of $u$ to r-faces:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_r}^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
-u_{R,\ib-\half\eb_r}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $w^{\trans}$:
-\begin{equation}
-u_{\ib-\half\eb_r}^{1D} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
-u_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
-\end{cases}
-\end{equation}
-Then, we do a full-dimensional extrapolation of $u$ to x-faces:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} = u_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
-u_{R,\ib-\half\eb_x}^{\mac,*} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
-\end{eqnarray}
-Finally, we solve a Riemann problem:
-\begin{equation}
-u_{\ib-\half\eb_x}^{\mac,*} =
-\begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} > 0, \\
-u_{R,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} < 0. 
-\end{cases}
-\end{equation}
-
-To compute $\wt^{\mac,*}$, we use an analogous procedure for computing $u^{\mac,*}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ and $\partial w_0/\partial r$ into the full-dimensional extrapolation.
-We compute a time-centered 1D extrapolation of $\wt$ to x-faces:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_x}^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
-\wt_{R,\ib-\half\eb_x}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \wt_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $u^{\trans}$:
-\begin{equation}
-\wt_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_x}^{1D} + \wt_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
-\wt_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-Then, we do a full-dimensional extrapolation of $\wt$ to r-faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{\mac,*} = \wt_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)(w_{0,\ib-\half\eb_r} - w_{0,\ib-\frac{3}{2}\eb_r}) + \frac{\dt}{2}f_{w,\ib-\eb_r},\label{wt_L Extrapolation} \\
-\wt_{R,\ib-\half\eb_r}^{\mac,*} = \wt_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)(w_{0,\ib+\half\eb_r} - w_{0,\ib-\half\eb_r}) + \frac{\dt}{2}f_{w,\ib}.\label{wt_R Extrapolation}
-\end{eqnarray}
-Next, we solve a Riemann problem:
-\begin{equation}
-\wt_{\ib-\half\eb_r}^{\mac,*} =
-\begin{cases}
-0, & \left(w_L^{\mac,*} \le 0 ~~ {\rm AND} ~~ w_R^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|w_L^{\mac,*} + w_R^{\mac,*}\right| < \epsilon, \\
-\wt_{L,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} < 0. 
-\end{cases}
-\end{equation}
-We then project $\Ubt^{\mac,*}$ to obtain $\Ubt^{\mac}$ and we call {\tt addw0} to compute $\Ub^{\mac}$ from $\Ubt^{\mac}$.
-%\subsection{3D Cartesian Case}
-%\subsection{3D Spherical Case}
-\cleardoublepage
-\section{Computing $\rho^{'\edge}, X_k^{\edge}$, and $(\rho h)^{'\edge}/h^{\edge}/T^{\edge}$ in {\tt MAESTRO}}
-In {\tt scalar\_advance}, we call {\tt make\_edge\_scal} to compute $\rho^{'\edge}, X_k^{\edge}$, and $(\rho h)^{'\edge}/h^{\edge}/T^{\edge}$ at each edge.
-The procedure is the same for each scalar, so we shall simply denote the scalar as $s$.
-We always need to compute $\rho'$ and $X_k$ to edges, and the choice of energy prediction is as follows:
-\begin{itemize}
-\item For {\tt enthalpy\_pred\_type} = 1, we predict $(\rho h)'$ to edges.
-\item For {\tt enthalpy\_pred\_type} = 2, we predict $h$ to edges.
-\item For {\tt enthalpy\_pred\_type} = 3 and 4, we predict $T$ to edges.
-\end{itemize}
-The force terms are:
-\begin{eqnarray}
-f_{\rho'} &=& -\rho'\nabla\cdot\Ub^{\mac} - \nabla\cdot(\rho_0\Ubt^{\mac}), \\
-f_{X_k} &=& 0, \\
-f_{(\rho h)'} &=& (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} -(\rho h)'\nabla\cdot\Ub^{\mac} - \nabla\cdot\left[(\rho h)_0\Ubt^{\mac}\right] + \nabla\cdot\kth\nabla T \\
-f_h &=& \frac{1}{\rho}\left(\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T \right), \\
-f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext} + \nabla\cdot\kth\nabla T \right\}.
-\end{eqnarray}
-If {\tt enthalpy\_pred\_type} = 3 or 4, we convert $T^{\edge}\rightarrow h^{\edge}$ using $h^{\edge} = h^{\edge}(\rho^{\edge},T^{\edge},X_k^{\edge})$, where $\rho^{\edge} = \rho^{'\edge} + \rho_0$.
-Furthermore, if {\tt enthalpy\_pred\_type} = 3, we compute $(\rho h)^{'\edge} = (\rho h)^{\edge} - (\rho h)_0$.
-\subsection{2D Cartesian Case}
-We begin by computing a time-centered 1D extrapolation of $s$ to all faces.  The extrapolation of $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib-\eb_x}^n, \\
-s_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $u^{\mac}$:
-\begin{equation}
-s_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{1D} + s_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-The extrapolation of $s$ to r-faces is analogous to the extrapolation to x-faces:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_r}^{1D} &=& s_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib-\eb_r}^n, \\
-s_{R,\ib-\half\eb_r}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $w^{\mac}$:
-\begin{equation}
-s_{\ib-\half\eb_r}^{1D} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_r}^{1D} + s_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_{L,\ib-\half\eb_r}^{1D}, & w^{\mac}_{\ib-\half\eb_r} > 0, \\
-u_{R,\ib-\half\eb_r}^{1D}, & w^{\mac}_{\ib-\half\eb_r} < 0. \\
-\end{cases}
-\end{equation}
-Now we do a full-dimensional extrapolation of $s$ to each face.  The extrapolation of $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\mac}+w_{\ib-\eb_x-\half\eb_r}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_r}^{1D} - s_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\mac}+w_{\ib-\half\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_r}^{1D} - s_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$.
-\begin{equation}
-s_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{\edge} + s_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
-\end{equation}
-The full-dimensional extrapolation of $s$ to r-faces is analogous to the extrapolation to x-faces:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_r}^{\edge} = s_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib-\eb_r}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\mac}+u_{\ib-\half\eb_x-\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_x-\eb_r}^{1D} - s_{\ib-\half\eb_x-\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_r},  \\
-s_{R,\ib-\half\eb_r}^{\edge} = s_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\mac}+u_{\ib-\half\eb_x}^{\mac}\right)\left(s_{\ib+\half\eb_x}^{1D} - s_{\ib-\half\eb_x}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}. 
-\end{eqnarray}
-Then, we upwind based on $w^{\mac}$.
-\begin{equation}
-s_{\ib-\half\eb_r}^{\edge} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_r}^{\edge} + s_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
-s_{L,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} > 0, \\
-s_{R,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} < 0.
-\end{cases}
-\end{equation}
-%\subsection{3D Cartesian Case}
-%\subsection{3D Spherical Case}
-\cleardoublepage
-\section{Computing $\Ubt^{\edge}$ in {\tt MAESTRO}}
-In {\tt velocity\_advance}, we call {\tt make\_edge\_state} to compute $\Ubt^{\edge}$.  Unlike when we computed $\Ubt^{\trans}$ and $\Ubt^{\mac,*}$, we need all components of velocity at each face.
-\subsection{2D Cartesian Case}
-We begin by computing a time-centered 1D extrapolation of $u$ to all faces.  The extrapolation of $u$ to x-faces is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{1D} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
-u_{R,\ib-\half\eb_x}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $u^{\trans}$:
-\begin{equation}
-u_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
-u_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
-u_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-The extrapolation of $u$ to r-faces is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_r}^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
-u_{R,\ib-\half\eb_r}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $w^{\trans}$:
-\begin{equation}
-u_{\ib-\half\eb_r}^{1D} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
-u_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
-\end{cases}
-\end{equation}
-We repeat this procedure to compute a time-centered 1D extrapolation of $\wt$ at each face.  The extrapolation of $\wt$ to x-faces is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_x}^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
-\wt_{R,\ib-\half\eb_x}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \wt_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $u^{\trans}$:
-\begin{equation}
-\wt_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_x}^{1D} + \wt_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
-\wt_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-The extrapolation of $\wt$ to r-faces is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{1D} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r}, \label{1D wt_L extrap} \\
-\wt_{R,\ib-\half\eb_r}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.
-\label{1D wt_R extrap}
-\end{eqnarray}
-Then we upwind based on $w^{\trans}$:
-\begin{equation}
-\wt_{\ib-\half\eb_r}^{1D} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_r}^{1D} + \wt_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
-\end{cases}
-\end{equation}
-Now we do a full-dimensional extrapolation of $u$ to each face.  The extrapolation of $u$ to x-faces is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\edge} = u_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib-\eb_x}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
-u_{R,\ib-\half\eb_x}^{\edge} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$.
-\begin{equation}
-u_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_x}^{\edge} + u_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-u_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-u_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
-\end{equation}
-The full-dimensional extrapolation of $u$ to r-faces is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_r}^{\edge} = u_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib-\eb_r}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_x-\eb_r}^{1D} - u_{\ib-\half\eb_x-\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_r},  \\
-u_{R,\ib-\half\eb_r}^{\edge} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(u_{\ib+\half\eb_x}^{1D} - u_{\ib-\half\eb_x}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}. 
-\end{eqnarray}
-Then, we upwind based on $w^{\mac}$.
-\begin{equation}
-u_{\ib-\half\eb_r}^{\edge} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_r}^{\edge} + u_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_{L,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} > 0, \\
-u_{R,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} < 0.
-\end{cases}
-\end{equation}
-Now we do a full-dimensional extrapolatino of $\wt$ to each face.
-This is different from the extrapolation of $u$ to faces because we must include the term proportional to $\partial w_0/\partial r$ in equation (\ref{Perturbational Velocity Equation}).
-The full-dimensional extrapolation of $\wt$ to x-faces is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_x}^{\edge} = \wt_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib-\eb_x}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(\wt_{\ib-\eb_x+\half\eb_r}^{1D} - \wt_{\ib-\eb_x-\half\eb_r}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib-\eb_x+\half\eb_r}^{\trans}+\wt_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(w_{0,\ib-\eb_x+\half\eb_r}-w_{0,\ib-\eb_x-\half\eb_r}\right) + \frac{\dt}{2}f_{w,\ib-\eb_x}, \label{wt_L x-edge} \\
-\wt_{R,\ib-\half\eb_x}^{\edge} = \wt_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_r}^{1D} - \wt_{\ib-\half\eb_r}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right) + \frac{\dt}{2}f_{w,\ib}. \label{wt_R x-edge}
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$:
-\begin{equation}
-\wt_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_x}^{\edge} + \wt_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-\wt_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
-\end{equation}
-The full-dimensional extrapolation of $\wt$ to r-faces is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{\edge} = \wt_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib-\eb_r}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)\left(w_{0,\ib-\half\eb_r}-w_{0,\ib-\frac{3}{2}\eb_r}\right) + \frac{\dt}{2}f_{w,\ib-\eb_r}, \label{wt_L r-edge} \\
-\wt_{R,\ib-\half\eb_r}^{\edge} = \wt_{\ib}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right) + \frac{\dt}{2}f_{w,\ib}. \label{wt_R r-edge}
-\end{eqnarray}
-Then, we upwind based on $w^{\mac}$:
-\begin{equation}
-\wt_{\ib-\half\eb_r}^{\edge} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_r}^{\edge} + \wt_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_r}^{\edge}, & w^{\mac}_{\ib-\half\eb_r} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{\edge}, & w^{\mac}_{\ib-\half\eb_r} < 0.
-\end{cases}
-\end{equation}
-%\subsection{3D Cartesian Case}
-%\subsection{3D Spherical Case}
-\cleardoublepage
-\section{Computing $\Ub^{\mac,*}$ in {\tt VARDEN}}
-\subsection{2D Cartesian Case}
-We do a 1D Taylor series extrapolation to get both components of velocity at the x-face:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{1D} &=& u_{\ib-\eb_x} + \left[\half - \frac{\dt}{2h}{\rm max}(0,u_{\ib-\eb_x})\right]\Delta_xu_{\ib-\eb_x}, \label{varden U_L^1D} \\
-u_{R,\ib-\half\eb_x}^{1D} &=& u_{\ib} + \left[\half - \frac{\dt}{2h}{\rm min}(0,u_{\ib})\right]\Delta_xu_{\ib}.
-\end{eqnarray}
-\begin{eqnarray}
-v_{L,\ib-\half\eb_x}^{1D} &=& v_{\ib-\eb_x} + \left[\half - \frac{\dt}{2h}{\rm max}(0,v_{\ib-\eb_x})\right]\Delta_xv_{\ib-\eb_x}, \label{varden U_L^1D} \\
-v_{R,\ib-\half\eb_x}^{1D} &=& v_{\ib} + \left[\half - \frac{\dt}{2h}{\rm min}(0,v_{\ib})\right]\Delta_xv_{\ib}.
-\end{eqnarray}
-We obtain the normal velocity using the Riemann problem:
-\begin{equation}
-u_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{1D} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{1D} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{1D}, & u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D} > 0, \\
-u_{R,\ib-\half\eb_x}^{1D}, & u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D} < 0.
-\end{cases}
-\end{equation}
-We obtain the transverse velocity by upwinding based on $u_{\ib-\half\eb_x}^{1D}$:
-\begin{equation}
-v_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(v_{L,\ib-\half\eb_x}^{1D} + v_{R,\ib-\half\eb_x}^{1D}\right), & \left|u_{\ib-\half\eb_x}^{1D}\right| < \epsilon \\
-v_{L,\ib-\half\eb_x}^{1D}, & u_{\ib-\half\eb_x}^{1D} > 0, \\
-v_{R,\ib-\half\eb_x}^{1D}, & u_{\ib-\half\eb_x}^{1D} < 0.
-\end{cases}\label{Transverse Velocity Riemann Problem}
-\end{equation}
-We perform analogous operations to compute both components of velocity at the y-faces, $\Ub_{\ib-\half\eb_y}^{1D}$. \\
-
-Now we do a full-dimensional extrapolation to get the MAC velocity at the x-faces (note that we only compute the normal components):
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
-u_{R,\ib-\half\eb_x}^{\mac,*} &=& u_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
-\end{eqnarray}
-Then, we solve a Riemann problem:
-\begin{equation}
-u_{\ib-\half\eb_x}^{\mac,*} =
-\begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} > 0, \\
-u_{R,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} < 0.
-\end{cases}\label{umac Riemann Problem}
-\end{equation}
-We perform analogous operations to compute the normal velocity at the y-faces, $v^{\mac,*}_{\ib-\half\eb_y}$.
-\subsection{3D Cartesian Case}
-This is more complicated than the 2D case because we include corner coupling.
-We compute $\Ub_{\ib-\half\eb_x}^{1D}, \Ub_{\ib-\half\eb_y}^{1D}$, and $\Ub_{\ib-\half\eb_z}^{1D}$ in an analogous manner as equations (\ref{varden U_L^1D})-(\ref{Transverse Velocity Riemann Problem}).
-Then, we compute an intermediate state, $u_{\ib-\half\eb_y}^{y|z}$, which is described as ``state $u_{\ib-\half\eb_y}^{1D}$ that has been updated to account for the transverse derivatives in the z direction'', using:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_y}^{y|z} &=& u_{L,\ib-\half\eb_y}^{1D} - \frac{\dt}{6h}\left(w_{\ib-\eb_y+\half\eb_z}^{1D}+w_{\ib-\eb_y-\half\eb_z}^{1D}\right)\left(u_{\ib-\eb_y+\half\eb_z}^{1D}-u_{\ib-\eb_y-\half\eb_z}^{1D}\right), \\
-u_{R,\ib-\half\eb_y}^{y|z} &=& u_{R,\ib-\half\eb_y}^{1D} - \frac{\dt}{6h}\left(w_{\ib+\half\eb_z}^{1D}+w_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{1D}-u_{\ib-\half\eb_z}^{1D}\right).
-\end{eqnarray}
-Then, upwind based on $v_{\ib-\half\eb_y}^{1D}$:
-\begin{equation}
-u_{\ib-\half\eb_y}^{y|z} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_y}^{y|z} + u_{R,\ib-\half\eb_y}^{y|z}\right), & \left|v_{\ib-\half\eb_y}^{1D}\right| < \epsilon \\
-u_{L,\ib-\half\eb_y}^{y|z}, & v_{\ib-\half\eb_y}^{1D} > 0, \\
-u_{R,\ib-\half\eb_y}^{y|z}, & v_{\ib-\half\eb_y}^{1D} < 0.
-\end{cases}
-\end{equation}
-We use an analogous procedure to compute five more intemediate states, $u_{\ib-\half\eb_z}^{z|y}, v_{\ib-\half\eb_x}^{x|z}, v_{\ib-\half\eb_z}^{z|x}, w_{\ib-\half\eb_x}^{x|y}$, and $w_{\ib-\half\eb_y}^{y|x}$.
-Then, we do a full-dimensional extrapolation to get the MAC velocities at normal faces:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} = u_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{y|z}-u_{\ib-\eb_x-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_z}^{1D}+w_{\ib-\eb_x-\half\eb_z}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_z}^{z|y}-u_{\ib-\eb_x-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
-u_{R,\ib-\half\eb_x}^{\mac,*} = u_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{y|z}-u_{\ib-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{1D}+w_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{z|y}-u_{\ib-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{u,\ib}.
-\end{eqnarray}
-Then, we use the Riemann solver given above for the 2D case (equation [\ref{umac Riemann Problem}]) to compute $u_{\ib-\half\eb_x}^{\mac,*}$.
-We use an analogous procedure to obtain $v_{\ib-\half\eb_y}^{\mac,*}$ and $w_{\ib-\half\eb_z}^{\mac,*}$.
-\cleardoublepage
-\section{Computing $\Ub^{\edge}$ and $\rho^{\edge}$ in {\tt VARDEN}}
-To compute $\Ub^{\edge}$, {\tt VARDEN} uses the exact same algorithm as the $s^{\edge}$ case in {\tt MAESTRO}.
-The algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.
-Here, $s$ is used in place of either $\rho, u, v$, or $w$ (in 3D).
-\subsection{2D Cartesian Case}
-The 1D extrapolation is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib-\eb_x}^n, \label{varden s_L^1D}\\
-s_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib}^n. \label{varden s_R^1D}
-\end{eqnarray}
-Then we upwind based on $u^{\mac}$:
-\begin{equation}
-s_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{1D} + s_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-We use an analogous procedure to obtain $s_{\ib-\half\eb_y}^{1D}$.  Now we do a full-dimensional extrapolation of $s$ to each face.  The extrapolation of a ``non-conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-The extrapolation of a ``conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] \nonumber \\
-&-& \frac{\dt}{2}s_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\half\eb_y}\right] \nonumber \\
-&-& \frac{\dt}{2}s_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right) + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$.
-\begin{equation}
-s_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{\edge} + s_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}\label{varden s^edge upwind}
-\end{equation}
-We use an analogous procedure to compute $s_{\ib-\half\eb_y}^{\edge}$.
-\subsection{3D Cartesian Case}
-This is more complicated than the 2D case because we include corner coupling.
-We first compute $s_{\ib-\half\eb_x}^{1D}$, $s_{\ib-\half\eb_y}^{1D}$, and $s_{\ib-\half\eb_z}^{1D}$ in an analogous manner to equations (\ref{varden s_L^1D}) and (\ref{varden s_R^1D}).
-Then we compute six intermediate states, $s_{\ib-\half\eb_x}^{x|y}, s_{\ib-\half\eb_x}^{x|z}, s_{\ib-\half\eb_y}^{y|x}, s_{\ib-\half\eb_y}^{y|z}, s_{\ib-\half\eb_z}^{z|x}$, and $s_{\ib-\half\eb_z}^{z|y}$.
-For the ``non-conservative case'', we use, for example:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{x|y} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{6h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} + v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
-s_{R,\ib-\half\eb_x}^{x|y} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{6h}\left(v_{\ib+\half\eb_y}^{\mac} + v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right).
-\end{eqnarray}
-For the ``conservative'' case, we use, for example:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{x|y} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{3h}\left[(sv^{\mac})_{\ib-\eb_x+\half\eb_y}-(sv^{\mac})_{\ib-\eb_x-\half\eb_y}\right], \\
-s_{R,\ib-\half\eb_x}^{x|y} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{3h}\left[(sv^{\mac})_{\ib+\half\eb_y}-(sv^{\mac})_{\ib-\half\eb_y}\right].
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$:
-\begin{equation}
-s_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{x|y} + s_{R,\ib-\half\eb_x}^{x|y}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{x|y}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{x|y}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
-\end{equation}
-We use an analogous procedure to compute the other five intermediate states.
-Now we do a full-dimensional extrapolation of $s$ to each face.
-The extrapolation of a ``non-conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{y|z}-s_{\ib-\eb_x-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_z}^{\mac}+w_{\ib-\eb_x-\half\eb_z}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_z}^{z|y}-s_{\ib-\eb_x-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{y|z}-s_{\ib-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{\mac}+w_{\ib-\half\eb_z}^{\mac}\right)\left(s_{\ib+\half\eb_z}^{z|y}-s_{\ib-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-The extrapolation of a ``conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{y|z}v^{\mac})_{\ib-\eb_x+\half\eb_y}-({s^{y|z}v^{\mac})_{\ib-\eb_x-\half\eb_y}}\right] \nonumber \\
-&-& \frac{\dt}{2h}\left[(s^{z|y}w^{\mac})_{\ib-\eb_x+\half\eb_z}-({s^{z|y}w^{\mac})_{\ib-\eb_x-\half\eb_z}}\right] \nonumber \\
-&-& \frac{\dt}{2}s_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}-v_{\ib-\eb_x-\half\eb_y}^{\mac}+w_{\ib-\eb_x+\half\eb_z}^{\mac}-w_{\ib-\eb_x-\half\eb_z}^{\mac}\right) \nonumber \\
-&+& \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{y|z}v^{\mac})_{\ib+\half\eb_y}-({s^{y|z}v^{\mac})_{\ib-\half\eb_y}}\right] \nonumber \\
-&-& \frac{\dt}{2h}\left[(s^{z|y}w^{\mac})_{\ib+\half\eb_z}-({s^{z|y}w^{\mac})_{\ib-\half\eb_z}}\right] \nonumber \\
-&-& \frac{\dt}{2}s_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac}-v_{\ib-\half\eb_y}^{\mac}+w_{\ib+\half\eb_z}^{\mac}-w_{\ib-\half\eb_z}^{\mac}\right) \nonumber \\
-&+& \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-Then we upwind based on $u^{\mac}$, as in equation (\ref{varden s^edge upwind}).  We use an analogous procedure to compute both $s_{\ib-\half\eb_y}^{\edge}$ and $s_{\ib-\half\eb_z}$.
+These are working notes for the low density parameters in {\tt MAESTRO}.
+\section{anelastic\_cutoff}
+\section{base\_cutoff\_density}
+\section{burning\_cutoff}
 \end{document}

### 2008-05-29 docs/lo_density/paper.tex
Updated doc.  I think I got all the cases in the code.

@@
 \tableofcontents
-\cleardoublepage
 
@@
 \section{anelastic\_cutoff}
+\begin{itemize}
+\item In {\tt base\_state}, we abort the program if {\tt anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
+\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance}, We compute the coordinate associated with {\tt anelastic\_cutoff}, $r_{\rm anel}$.
+This is done by setting $r_{\rm anel} = i$ as soon as $\rho_0(i) <$ {\tt anelastic\_cutoff}.
+\item In {\tt correct\_base}, we only update $\rho_0$ for $r\in[0,r_{\rm anel}-1]$.
+\item In {\tt make\_div\_coeff}, we change how we compute {\tt div\_coeff} at and above $r_{\rm anel}$.
+\item In {\tt make\_psi}, we only compute $\psi$ for $r\in[0,r_{\rm anel}-1]$.
+\item in {\tt make\_S}, we set {\tt delta\_gamma1\_term} and {\tt delta\_gamma1} to zero at and above $r_{\rm anel}$.
+\end{itemize}
 \section{base\_cutoff\_density}
+\begin{itemize}
+\item In {\tt base\_state}, we abort the program if {\tt anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
+\item In {\tt base\_state}, we compute a physical cutoff location, {\tt r\_cutoff\_loc}, which is defined as the physical location of the first cell-center at the coarsest level for which $\rho_0 <$ {\tt base\_cutoff\_density}.
+\item In {\tt base\_state}, when we are generating the initial background/base state, if we are above {\tt r\_cutoff\_loc}, just use the values for $\rho,T$, and $p$ at {\tt r\_cutoff\_loc}.
+\item In {\tt make\_hgrhs}, we only add the volume discrepancy if $\rho_0 >$ {\tt base\_cutoff\_density}.
+NOTE: This is only implemented for 2D Cartesian and needs to be implemented for 3D Cartesian and spherical.
+\item In {\tt make\_macrhs}, we only add the volume discrepancy if $\rho_0 >$ {\tt base\_cutoff\_density}.
+NOTE: This is only implemented for 2D Cartesian and needs to be implemented for 3D Cartesian and spherical.
+\item In {\tt mkscalforce} subroutine {\tt mkrhohforce}, once $\rho_0 \le$ {\tt base\_cutoff\_density}, we compute $\nabla p_0$ with a difference stencil instead of simply setting it to $\rho_0 g$.
+\item In {\tt rhoh\_vs\_t} function {\tt makeRhoHfromP}, the base state pressure at edges is computed differently based on where we are in relation to the {\tt base\_cutoff\_density}.
+If we are at a equal or higher density than {\tt base\_cutoff\_density}, we integrate $\rho g$ from the nearest cell-centered $p_0$ value.
+If we are at a lower density than {\tt base\_cutoff\_density}, we just use the cell-centered value.
+\item In {\tt update\_scal}, if the new $\rho$ is less than {\tt base\_cutoff\_density}, we call the EOS to compute $\rho h$.
+\end{itemize}
 \section{burning\_cutoff}
+\begin{itemize}
+\item In {\tt react\_state}, we only call the burner if $\rho >$ {\tt burning\_cutoff\_density}.
+\end{itemize}
 \end{document}

### 2008-05-29 Source/make_macrhs.f90
for the 3d sphr and cart cases, now the volume discrepancy term is only added if the density is less than the base_cutoff_density
updated document to reflect that I have fixed the projection right hand sides

@@ -72,8 +72,8 @@ contains
                                  p0_old(n,:),p0_new(n,:),pop(:,:,1,1),pnp(:,:,1,1), &
                                  pthermbar_old(n,:),pthermbar_new(n,:),dt)
           case (3)
-             call make_macrhs_3d(n,lo,hi,mp(:,:,:,1),sp(:,:,:,1),gp(:,:,:,1),Sbar(n,:), &
-                                 div_coeff(n,:),dx(n,:), &
+             call make_macrhs_3d(n,lo,hi,rho0(n,:),mp(:,:,:,1),sp(:,:,:,1),gp(:,:,:,1), &
+                                 Sbar(n,:),div_coeff(n,:),dx(n,:), &
                                  gamma1bar_old(n,:),gamma1bar_new(n,:), &
                                  p0_old(n,:),p0_new(n,:),pop(:,:,:,1),pnp(:,:,:,1), &
                                  pthermbar_old(n,:),pthermbar_new(n,:),dt)
@@ -117,29 +117,31 @@ contains
 
     if (dpdt_factor .gt. 0.0d0) then
        do j = lo(2),hi(2)
-          gamma1bar_p0_avg = 0.25d0*(gamma1bar_old(j)+gamma1bar_new(j))*(p0_old(j)+p0_new(j))
-          do i = lo(1),hi(1)
-             ptherm_diff = &
-                  0.5d0*(ptherm_old(i,j)+ptherm_new(i,j)-pthermbar_old(j)-pthermbar_new(j))
-             if (rho0(j) .gt. base_cutoff_density) then
+          if (rho0(j) .gt. base_cutoff_density) then
+             gamma1bar_p0_avg = 0.25d0 * &
+                  (gamma1bar_old(j)+gamma1bar_new(j))*(p0_old(j)+p0_new(j))
+             do i = lo(1),hi(1)
+                ptherm_diff = 0.5d0 * &
+                     (ptherm_old(i,j)+ptherm_new(i,j)-pthermbar_old(j)-pthermbar_new(j))
                 rhs(i,j) = rhs(i,j) + div_coeff(j) * &
                      (dpdt_factor / gamma1bar_p0_avg) * (ptherm_diff / dt)
-             end if
-          end do
+             end do
+          end if
        end do
     end if
 
   end subroutine make_macrhs_2d
 
-  subroutine make_macrhs_3d(n,lo,hi,rhs,Source,delta_gamma1_term,Sbar,div_coeff,dx, &
[…]

### 2008-05-29 docs/initial_models/paper.tex
start adding some notes on how to generate initial models for MAESTRO

--- 2008-05-29 docs/lo_density/paper.tex
@@
 
-\title{Notes on the low density parameters in {\tt MAESTRO}}
+\title{Notes on generating initial models for mapping into {\tt MAESTRO}}
 
@@
 \tableofcontents
-\cleardoublepage
 
-These are working notes for the low density parameters in {\tt MAESTRO}.
-\section{anelastic\_cutoff}
-\section{base\_cutoff\_density}
-\section{burning\_cutoff}
+\section{Creating the initial model}
+\begin{itemize}
+\item The initial model should be but into HSE according to $(p_{i+1} - p_i)/\Delta r = half (\rho_{i+1} - \rho_i) g_{i+\half}$
+
+\item The initial model should be created at the same resolution as the base state.  This means that
+when you change resolution of your simulation, use a new initial model.
+
+\item The initial model should be thermodynamically consistent with the EOS that {\tt MAESTRO}
+uses.
+
+\end{itemize}
+
+\section{Spherical notes}
+\begin{itemize}
+\item Spherical runs use $p$, $(\rho X)$, and $\rho$ from the initial model.  These are then
+mapped onto the Cartesian grid, and then $(rho h)$ is computed via the EOS.
+\end{itemize}
+
 \end{document}

### 2008-06-20 docs/lo_density/paper.tex
for the 3d sphr and cart cases, now the volume discrepancy term is only added if the density is less than the base_cutoff_density
updated document to reflect that I have fixed the projection right hand sides
the variable names now exactly match the variable names in the code

@@
 \item In {\tt base\_state}, we abort the program if {\tt anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
-\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance}, We compute the coordinate associated with {\tt anelastic\_cutoff}, $r_{\rm anel}$.
-This is done by setting $r_{\rm anel} = i$ as soon as $\rho_0(i) <$ {\tt anelastic\_cutoff}.
-\item In {\tt correct\_base}, we only update $\rho_0$ for $r\in[0,r_{\rm anel}-1]$.
-\item In {\tt make\_div\_coeff}, we change how we compute {\tt div\_coeff} at and above $r_{\rm anel}$.
-\item In {\tt make\_psi}, we only compute $\psi$ for $r\in[0,r_{\rm anel}-1]$.
-\item in {\tt make\_S}, we set {\tt delta\_gamma1\_term} and {\tt delta\_gamma1} to zero at and above $r_{\rm anel}$.
+\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance}, We compute the coordinate associated with {\tt anelastic\_cutoff}, {\tt anelastic\_cutoff\_coord}.
+This is done by setting ${\tt anelastic\_cutoff\_coord}= r$ as soon as $\rho_0(r) <$ {\tt anelastic\_cutoff}.
+\item In {\tt correct\_base}, we only update $\rho_0$ for $r\in[0,{\tt anelastic\_cutoff\_coord}-1]$.
+\item In {\tt make\_div\_coeff}, we change how we compute {\tt div\_coeff} at and above ${\tt anelastic\_cutoff\_coord}$.
+\item In {\tt make\_psi}, we only compute $\psi$ for $r\in[0,{\tt anelastic\_cutoff\_coord}-1]$.
+\item in {\tt make\_S}, we set {\tt delta\_gamma1\_term} and {\tt delta\_gamma1} to zero at and above ${\tt anelastic\_cutoff\_coord}$.
 \end{itemize}
@@
 \item In {\tt base\_state}, we abort the program if {\tt anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
-\item In {\tt base\_state}, we compute a physical cutoff location, {\tt r\_cutoff\_loc}, which is defined as the physical location of the first cell-center at the coarsest level for which $\rho_0 <$ {\tt base\_cutoff\_density}.
-\item In {\tt base\_state}, when we are generating the initial background/base state, if we are above {\tt r\_cutoff\_loc}, just use the values for $\rho,T$, and $p$ at {\tt r\_cutoff\_loc}.
+\item In {\tt base\_state}, we compute a physical cutoff location, {\tt base\_cutoff\_density\_loc}, which is defined as the physical location of the first cell-center at the coarsest level for which $\rho_0 <$ {\tt base\_cutoff\_density}.
+\item In {\tt base\_state}, when we are generating the initial background/base state, if we are above {\tt base\_cutoff\_density\_loc}, just use the values for $\rho,T$, and $p$ at {\tt base\_cutoff\_density\_loc}.
 \item In {\tt make\_hgrhs}, we only add the volume discrepancy if $\rho_0 >$ {\tt base\_cutoff\_density}.
-NOTE: This is only implemented for 2D Cartesian and needs to be implemented for 3D Cartesian and spherical.
 \item In {\tt make\_macrhs}, we only add the volume discrepancy if $\rho_0 >$ {\tt base\_cutoff\_density}.
-NOTE: This is only implemented for 2D Cartesian and needs to be implemented for 3D Cartesian and spherical.
 \item In {\tt mkscalforce} subroutine {\tt mkrhohforce}, once $\rho_0 \le$ {\tt base\_cutoff\_density}, we compute $\nabla p_0$ with a difference stencil instead of simply setting it to $\rho_0 g$.

### 2008-06-20 docs/lo_density/paper.tex
New global variables stored in geometry.f90: "base_cutoff_density_coord" and "burning_cutoff_density_coord".  We need coordinates so we can sync up how these affect the problem at all levels of refinement.  Just using a real number sometimes led to dynamics on fine grids being different from coarse grids right at the cutoff locations - in particular when the "higher" of two fine cells on top of a coarse cell was considered above the cutoff and the "lower" of the two fine cells was considered below the cutoff.  We compute the coordinaet locations at the same time we compute the anelastic_cutoff

@@
 \item In {\tt base\_state}, when we are generating the initial background/base state, if we are above {\tt base\_cutoff\_density\_loc}, just use the values for $\rho,T$, and $p$ at {\tt base\_cutoff\_density\_loc}.
-\item In {\tt make\_hgrhs}, we only add the volume discrepancy if $\rho_0 >$ {\tt base\_cutoff\_density}.
-\item In {\tt make\_macrhs}, we only add the volume discrepancy if $\rho_0 >$ {\tt base\_cutoff\_density}.
+\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance}, We compute the coordinate associated with {\tt base\_cutoff\_density}, {\tt base\_cutoff\_density\_coord}.
+This is done by setting ${\tt base\_cutoff\_density\_coord}= r$ as soon as $\rho_0(r) <$ {\tt base\_cutoff\_density}.
+\item In {\tt make\_hgrhs}, we only add the volume discrepancy if $r < {\tt base\_cutoff\_density\_coord}$.
+\item In {\tt make\_macrhs}, we only add the volume discrepancy if $r < {\tt base\_cutoff\_density\_coord}$.
 \item In {\tt mkscalforce} subroutine {\tt mkrhohforce}, once $\rho_0 \le$ {\tt base\_cutoff\_density}, we compute $\nabla p_0$ with a difference stencil instead of simply setting it to $\rho_0 g$.
@@
 \begin{itemize}
+\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance}, We compute the coordinate associated with {\tt burning\_cutoff\_density}, {\tt burning\_cutoff\_density\_coord}.
+This is done by setting ${\tt burning\_cutoff\_density\_coord}= r$ as soon as $\rho_0(r) <$ {\tt burning\_cutoff\_density}.
 \item In {\tt react\_state}, we only call the burner if $\rho >$ {\tt burning\_cutoff\_density}.

### 2008-06-20 Source/advance.f90

@@ -44,7 +44,7 @@ contains
     use add_react_to_thermal_module
     use variables, only: nscal, press_comp, temp_comp, rho_comp, foextrap_comp
     use geometry, only: spherical, nr_fine, r_start_coord, r_end_coord, &
-         anelastic_cutoff_coord
+         anelastic_cutoff_coord, base_cutoff_density_coord, burning_cutoff_density_coord
     use network, only: nspec
     use make_grav_module
     use make_eta_module
@@ -199,6 +199,32 @@ contains
     do n=2,nlevs
        anelastic_cutoff_coord(n) = 2*anelastic_cutoff_coord(n-1)
     end do
+
+    ! compute the coordinates of the base cutoff density
+    base_cutoff_density_coord(1) = r_end_coord(1)
+    do r=0,r_end_coord(1)
+       if (rho0_old(1,r) .lt. base_cutoff_density .and. &
+            base_cutoff_density_coord(1) .eq. r_end_coord(1)) then
+          base_cutoff_density_coord(1) = r
+          exit
+       end if
+    end do
+    do n=2,nlevs
+       base_cutoff_density_coord(n) = 2*base_cutoff_density_coord(n-1)
+    end do
+
+  ! compute the coordinates of the burning cutoff density
+  burning_cutoff_density_coord(1) = r_end_coord(1)
+  do r=0,r_end_coord(1)
+     if (rho0_old(1,r) .lt. burning_cutoff_density .and. &
+          burning_cutoff_density_coord(1) .eq. r_end_coord(1)) then
+        burning_cutoff_density_coord(1) = r
+        exit
+     end if
+  end do
+  do n=2,nlevs
+     burning_cutoff_density_coord(n) = 2*burning_cutoff_density_coord(n-1)
+  end do
 
     ! tempbar is only used as an initial guess for eos calls
     call average(mla,sold,tempbar,dx,temp_comp)

### 2008-06-20 Source/make_hgrhs.f90

@@ -363,7 +363,7 @@ contains
           hi =  upb(get_box(delta_p_term(n), i))
           select case (dm)
           case (2)
-             call create_correction_cc_2d(lo,hi,rho0(n,:),ccp(:,:,1,1),ptp(:,:,1,1), &
+             call create_correction_cc_2d(n,lo,hi,rho0(n,:),ccp(:,:,1,1),ptp(:,:,1,1), &
                                           div_coeff(n,:), &
                                           gamma1bar(n,:),p0(n,:),dt)
           case (3)
@@ -377,7 +377,7 @@ contains
                                                   r0p(:,:,:,1),dt)
 
              else
-                call create_correction_cc_3d_cart(lo,hi,rho0(n,:),ccp(:,:,:,1),ptp(:,:,:,1), &
+                call create_correction_cc_3d_cart(n,lo,hi,rho0(n,:),ccp(:,:,:,1),ptp(:,:,:,1), &
                                                   div_coeff(n,:),gamma1bar(n,:),p0(n,:), &
                                                   dt)
              end if
@@ -450,12 +450,13 @@ contains
     
   end subroutine correct_hgrhs
   
-  subroutine create_correction_cc_2d(lo,hi,rho0,correction_cc,delta_p_term,div_coeff,gamma1bar, &
-                                     p0,dt)
+  subroutine create_correction_cc_2d(n,lo,hi,rho0,correction_cc,delta_p_term,div_coeff, &
+                                     gamma1bar,p0,dt)
 
-    use probin_module, only: dpdt_factor, base_cutoff_density
+    use probin_module, only: dpdt_factor
+    use geometry, only: base_cutoff_density_coord
 
-    integer         , intent(in   ) :: lo(:), hi(:)
+    integer         , intent(in   ) :: n, lo(:), hi(:)
     real (kind=dp_t), intent(in   ) :: rho0(0:)
     real (kind=dp_t), intent(  out) :: correction_cc(lo(1)-1:,lo(2)-1:)
     real (kind=dp_t), intent(in   ) :: delta_p_term(lo(1):,lo(2):)
@@ -469,7 +470,7 @@ contains
     real(kind=dp_t) :: correction_factor
     
     do j = lo(2),hi(2)
-       if(rho0(j) .gt. base_cutoff_density) then
+       if(j .lt. base_cutoff_density_coord(n)) then
[…]

### 2008-06-20 Source/make_macrhs.f90

@@ -63,7 +63,7 @@ contains
           hi =  upb(get_box(Source(n), i))
           select case (dm)
           case (2)
-             call make_macrhs_2d(lo,hi,rho0(n,:),mp(:,:,1,1),sp(:,:,1,1),gp(:,:,1,1), &
+             call make_macrhs_2d(n,lo,hi,rho0(n,:),mp(:,:,1,1),sp(:,:,1,1),gp(:,:,1,1), &
                                  Sbar(n,:), &
                                  div_coeff(n,:),gamma1bar_old(n,:),gamma1bar_new(n,:), &
                                  p0_old(n,:),p0_new(n,:),pop(:,:,1,1),dt)
@@ -81,12 +81,13 @@ contains
 
   end subroutine make_macrhs
 
-  subroutine make_macrhs_2d(lo,hi,rho0,rhs,Source,delta_gamma1_term,Sbar,div_coeff, &
+  subroutine make_macrhs_2d(n,lo,hi,rho0,rhs,Source,delta_gamma1_term,Sbar,div_coeff, &
                             gamma1bar_old,gamma1bar_new,p0_old,p0_new,delta_p_term,dt)
 
-    use probin_module, only: dpdt_factor, base_cutoff_density
+    use probin_module, only: dpdt_factor
+    use geometry, only: base_cutoff_density_coord
 
-    integer         , intent(in   ) :: lo(:), hi(:)
+    integer         , intent(in   ) :: n, lo(:), hi(:)
     real (kind=dp_t), intent(in   ) :: rho0(0:) 
     real (kind=dp_t), intent(  out) :: rhs(lo(1):,lo(2):)  
     real (kind=dp_t), intent(in   ) :: Source(lo(1):,lo(2):)  
@@ -110,7 +111,7 @@ contains
 
     if (dpdt_factor .gt. 0.0d0) then
        do j = lo(2),hi(2)
-          if (rho0(j) .gt. base_cutoff_density) then
+          if (j .lt. base_cutoff_density_coord(n)) then
              gamma1bar_p0_avg = 0.25d0 * &
                   (gamma1bar_old(j)+gamma1bar_new(j))*(p0_old(j)+p0_new(j))
              do i = lo(1),hi(1)
@@ -127,7 +128,7 @@ contains
                             gamma1bar_old,gamma1bar_new,p0_old,p0_new, &
                             delta_p_term,dt)
 
-    use geometry, only: spherical
+    use geometry, only: spherical, base_cutoff_density_coord
[…]

### 2008-06-20 docs/lo_density/paper.tex
matches the latest code cleanups

--- 2008-06-20 docs/lo_density/paper.tex
@@
 \begin{itemize}
-\item In {\tt base\_state}, we abort the program if {\tt anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
-\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance}, We compute the coordinate associated with {\tt anelastic\_cutoff}, {\tt anelastic\_cutoff\_coord}.
+\item In {\tt init\_base\_state}, we abort the program if {\tt anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
+\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance\_timestep}, We compute the coordinate associated with {\tt anelastic\_cutoff}, {\tt anelastic\_cutoff\_coord}.
 This is done by setting ${\tt anelastic\_cutoff\_coord}= r$ as soon as $\rho_0(r) <$ {\tt anelastic\_cutoff}.
-\item In {\tt correct\_base}, we only update $\rho_0$ for $r\in[0,{\tt anelastic\_cutoff\_coord}-1]$.
-\item In {\tt make\_div\_coeff}, we change how we compute {\tt div\_coeff} at and above ${\tt anelastic\_cutoff\_coord}$.
-\item In {\tt make\_psi}, we only compute $\psi$ for $r\in[0,{\tt anelastic\_cutoff\_coord}-1]$.
-\item in {\tt make\_S}, we set {\tt delta\_gamma1\_term} and {\tt delta\_gamma1} to zero at and above ${\tt anelastic\_cutoff\_coord}$.
+\item In {\tt correct\_base}, we only update $\rho_0$ for $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+\item In {\tt make\_div\_coeff}, we change how we compute {\tt div\_coeff} for $r\in[{\tt anelastic\_cutoff\_coord},{\tt r\_end\_coord}]$.
+\item In {\tt make\_psi}, we only compute $\psi$ for $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+\item in {\tt make\_S}, we set {\tt delta\_gamma1\_term} and {\tt delta\_gamma1} to zero for $r\in[{\tt anelastic\_cutoff\_coord},{\tt r\_end\_coord}]$.
 \end{itemize}
@@
 \item In {\tt base\_state}, when we are generating the initial background/base state, if we are above {\tt base\_cutoff\_density\_loc}, just use the values for $\rho,T$, and $p$ at {\tt base\_cutoff\_density\_loc}.
-\item In {\tt make\_hgrhs}, we only add the volume discrepancy if $\rho_0 >$ {\tt base\_cutoff\_density}.
-\item In {\tt make\_macrhs}, we only add the volume discrepancy if $\rho_0 >$ {\tt base\_cutoff\_density}.
-\item In {\tt mkscalforce} subroutine {\tt mkrhohforce}, once $\rho_0 \le$ {\tt base\_cutoff\_density}, we compute $\nabla p_0$ with a difference stencil instead of simply setting it to $\rho_0 g$.
-\item In {\tt rhoh\_vs\_t} function {\tt makeRhoHfromP}, the base state pressure at edges is computed differently based on where we are in relation to the {\tt base\_cutoff\_density}.
-If we are at a equal or higher density than {\tt base\_cutoff\_density}, we integrate $\rho g$ from the nearest cell-centered $p_0$ value.
-If we are at a lower density than {\tt base\_cutoff\_density}, we just use the cell-centered value.
+\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance\_timestep}, We compute the coordinate associated with {\tt base\_cutoff\_density}, {\tt base\_cutoff\_density\_coord}.
+This is done by setting ${\tt base\_cutoff\_density\_coord}= r$ as soon as $\rho_0(r) <$ {\tt base\_cutoff\_density}.
+\item In {\tt make\_hgrhs}, we only add the volume discrepancy for $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+\item In {\tt make\_macrhs}, we only add the volume discrepancy for $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+\item In {\tt mkscalforce} subroutine {\tt mkrhohforce}, for $r\in[{\tt base\_cutoff\_density\_coord},{\tt r\_end\_coord}]$ we compute $\nabla p_0$ with a difference stencil instead of simply setting it to $\rho_0 g$.
+\item In {\tt makeRhoHfromP}, the base state pressure at edges is computed differently based on where we are in relation to the {\tt base\_cutoff\_density}.
+For $r\in[{\tt r\_start\_coord},{\tt base\_cutoff\_density\_coord}-1]$ we integrate $\rho g$ from the nearest cell-centered $p_0$ value.
+Otherwise, we just use the cell-centered value above the edge.
 \item In {\tt update\_scal}, if the new $\rho$ is less than {\tt base\_cutoff\_density}, we call the EOS to compute $\rho h$.
@@
 \begin{itemize}
+\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance\_timestep}, We compute the coordinate associated with {\tt burning\_cutoff\_density}, {\tt burning\_cutoff\_density\_coord}.
+This is done by setting ${\tt burning\_cutoff\_density\_coord}= r$ as soon as $\rho_0(r) <$ {\tt burning\_cutoff\_density}.
 \item In {\tt react\_state}, we only call the burner if $\rho >$ {\tt burning\_cutoff\_density}.
+%\item In {\tt react\_state}, we only call the burner for $r\in[{\tt r\_start\_coord},{\tt burning\_cutoff\_density\_coord}-1]$.
 \end{itemize}

### 2008-07-28 docs/pert_notes/paper.tex
new doc sorting out how rho, rho', (rho X), X, (rho h), (rho h)', h, h', T, and etarho relate to each other depending on predict_rho and enthalpy_pred_type

--- 2008-04-21 docs/Godunov_notes/paper.tex
@@
 
-\def\half  {\frac{1}{2}}
-\def\dt    {\Delta t}
-\def\kth   {k_{\rm th}}
-
-\def\edge  {\rm EDGE}
-\def\mac   {\rm MAC}
-\def\trans {\rm TRANS}
-
-\def\eb    {{\bf e}}
-\def\fb    {{\bf f}}
-\def\ib    {{\bf i}}
-\def\Ub    {{\bf U}}
-
-\def\Ubt   {\widetilde{\bf U}}
-\def\ut    {\tilde{u}}
-\def\vt    {\tilde{v}}
-\def\wt    {\tilde{w}}
-
-\title{Notes on the Godunov Step in {\tt MAESTRO} and {\tt VARDEN}}
+\title{Notes on Perturbational Quantities}
 
@@
 
-These are working notes for the Godunov step in {\tt MAESTRO} and {\tt VARDEN}.
-\section{Differences Between These Notes and the Code}
-\begin{enumerate}
-\item Concerning equations (\ref{Radial utrans_L}), (\ref{Radial utrans_R}), (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}), in the code we are omitting the term in equation (\ref{Perturbational Velocity Equation}) proportional to $(\Ubt\cdot\eb_r)$.
-\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction.
-However, this term is present in the cell-update.
-\end{enumerate}
-\section{Questions}
-\begin{enumerate}
-\item Concerning equations (\ref{Radial utrans_L}), (\ref{Radial utrans_R}), (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}), do we really need to include the term proportional to $(\Ubt\cdot\eb_r)$?
-\item Concerning (\ref{Force Term}), do we really need to include the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction?
-\item Do we want to do characteristic tracing to compute $\Ub^{\trans}$, analgous to how {\tt VARDEN} uses characteristic tracing to compute $\Ub^{\mac}$?
-\item When we compute $\Ub^{\mac,*}$, do we want to replace all instances of ``trans'' velocities with ``MAC'' velocities?
-Doing this is simply a matter of replacing the call to {\tt make\_edge\_state} with {\tt make\_edge\_scal}.
-\end{enumerate}
-\section{Differences Between {\tt MAESTRO} and {\tt VARDEN}}
-\begin{enumerate}
-\item {\tt VARDEN} does not use the concept of $w_0$, so $\Ubt$ does not exist.
-\item {\tt VARDEN} only does Cartesian problems.
-\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$ in the computation of $\Ub^{\mac}$, so therefore the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
-\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$, whereas {\tt MAESTRO} does not.
-\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, and a ``non-FPU'' algorithm for $\Ub^{\edge}$
-\item {\tt VARDEN} treats $\rho$ in a conservative manner, whereas {\tt MAESTRO} treats $\rho'$ in a non-conservative manner.
-Therefore, {\tt MAESTRO} accounts for this difference by modifying the forcing term in {\tt modify\_scal\_force}
-\item {\tt MAESTRO} does not corner couple for 3D problems.
-\end{enumerate}
-\cleardoublepage
-\section{Computing $\Ubt^{\trans}$ in {\tt MAESTRO}}
-In {\tt advance\_premac}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.
-These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
-We will only compute the normal component of velocity at each face.\\
-
-The evolution equation for the perturbational velocity is:
-\begin{equation}
-\frac{\partial\Ubt}{\partial t} = -\Ubt\cdot\nabla\Ubt - w_0\frac{\partial\Ubt}{\partial r} - (\Ubt\cdot\eb_r)\frac{\partial w_0}{\partial r}\eb_r - \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Perturbational Velocity Equation}
-\end{equation}
-We are going to use a 1D Taylor series extrapolation in space and time to compute time-centered normal components of $\Ubt^{\trans}$ at face-centers.
-By 1D, we mean that we omit any spatial derivatives that are not in the direction of the extrapolation.
-We first compute ``left/right'' states, and then use a Riemann solver to pick the final state.
-\subsection{2D Cartesian Case}
-For the rest of this paper, for 2D Cartesian problems, $\Ub = (u,w)$ and $\Ubt = (u,\wt)$.  We will use the shorthand $\ib = (x,r)$.\\
-
-For the x-faces, the 1D extrapolation of $u$ is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\trans} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
-u_{R,\ib-\half\eb_x}^{\trans} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
-\end{eqnarray}
-We pick the final edge states using a Riemann solver:
-\begin{equation}
-u^{\trans}_{\ib-\half\eb_x} =
-\begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\trans} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{\trans}, & u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans} > 0, \\
-u_{R,\ib-\half\eb_x}^{\trans}, & u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans} < 0, \\
-\end{cases}
-\end{equation}
-For the r-faces, we choose include the radial derivative terms proportional to $w_0$ and $(\Ubt\cdot\eb_r) = \wt$.
-The term proportional to $w_0$ is included by using $w$ rather than $\wt$ as the advection speed.
-The term proportional to $\wt$ is an additional term.
-For the r-faces, the 1D extrapolation of $\wt$ is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{\trans} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
-\wt_{R,\ib-\half\eb_r}^{\trans} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.\label{Radial utrans_R}
-\end{eqnarray}
-The total radial velocity is defined as:
-\begin{equation}
-w_{\ib}^n = \wt_{\ib}^n + \frac{w_{0,r+\half}+w_{0,r-\half}}{2}.\label{Total Radial Velocity}
-\end{equation}
-The spatial derivatives of $w_0$ are given by the two-point centered difference:
-\begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)_{\ib} = \frac{w_{0,r+\half}-w_{0,r-\half}}{h}.\label{Radial Derivative of w_0}
-\end{equation}
-We pick the final edge states using:
-\begin{equation}
-\wt^{\trans}_{\ib-\half\eb_r} =
-\begin{cases}
-0, & \left(w_L^{\trans} \le 0 ~~ {\rm AND} ~~ w_R^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|w_L^{\trans} + w_R^{\trans}\right| < \epsilon, \\
-\wt_{L,\ib-\half\eb_r}^{\trans}, & w_L^{\trans} + w_R^{\trans} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{\trans}, & w_L^{\trans} + w_R^{\trans} < 0, \\
-\end{cases}
-\end{equation}
-where $w_L^{\trans} = \wt_{L,\ib-\half\eb_r}^{\trans} + w_{0,r-\half}$ and $w_R^{\trans} = \wt_{R,\ib-\half\eb_r}^{\trans} + w_{0,r-\half}$.
-Finally, we call {\tt addw0} to compute $w^{\trans}$ from $\wt^{\trans}$.
-\subsection{3D Cartesian Case}
-For the rest of this paper, for 3D Cartesian problems, $\Ub = (u,v,w)$ and $\Ubt = (u,v,\wt)$.  We will use the shorthand $\ib = (x,y,r)$.\\
-
-The only difference between the 2D and 3D Cartesian cases is that we must also compute $v^{\trans}$.
-The procedure is analogous to computing $u^{\trans}$.
-For y-faces, the 1D extrapolation of $v$ is:
-\begin{eqnarray}
-v_{L,\ib-\half\eb_y}^{\trans} &=& v_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y v_{\ib-\eb_y}^n, \\
-v_{R,\ib-\half\eb_y}^{\trans} &=& v_{\ib}^n - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y v_{\ib}^n,
-\end{eqnarray}
-\begin{equation}
-v^{\trans}_{\ib-\half\eb_y} =
-\begin{cases}
-0, & \left(v_{L,\ib-\half\eb_y}^{\trans} \le 0 ~~ {\rm AND} ~~ v_{R,\ib-\half\eb_y}^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans}\right| < \epsilon, \\
-v_{L,\ib-\half\eb_y}^{\trans}, & v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans} > 0, \\
-v_{R,\ib-\half\eb_y}^{\trans}, & v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans} < 0. \\
-\end{cases}
-\end{equation}
-\subsection{3D Spherical Case}
-For the rest of this paper, for 3D spherical problems, $\Ub = (u,v,w)$ and $\Ubt = (\ut,\vt,\wt)$.  We will use the shorthand $\ib = (x,y,z)$.\\
-
-We first construct $\Ub$ by mapping $w_0$ onto a Cartesian array and adding it to $\Ubt$.
-This is a spherical generalization of equation (\ref{Total Radial Velocity}).
-We also map $\partial w_0/\partial r$ onto a Cartesian array.
-This is a spherical generalization of equation (\ref{Radial Derivative of w_0}).
-Then we use a 1D extrapolation analogous to equations equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}).
-For example, for x-faces:
-\begin{eqnarray}
-\ut_L^{\trans} &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x\ut_{\ib-\eb_x}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib-\eb_x}, \\
-\ut_R^{\trans} &=& \ut_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib}.
-\end{eqnarray}
-Then, we use the same Riemann solver as the Cartesian case to obtain $\ut^{\trans}$.
-To compute $\vt^{\trans}$ (or $\wt^{\trans}$), simply replace all instaces of $\ut$ with $\vt$ (or $\wt$) and all instaces of $x$ with $y$ (or $z$).
-Finally, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
-\cleardoublepage
-\section{Computing $\Ubt^{\mac,*}$ in {\tt MAESTRO}}
-In {\tt advance\_premac}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac,*}$.  We will only compute the normal component of velocity at each face.\\
-
-The force terms are:
-\begin{equation}
-\fb =
-\left(\begin{array}{c}
-f_u \\
-f_w \\
-\end{array}\right)
-= \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Force Term}
-\end{equation}
-\subsection{2D Cartesian Case}
-We begin by computing a time-centered 1D extrapolation of $u$ to r-faces:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_r}^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
-u_{R,\ib-\half\eb_r}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $w^{\trans}$:
-\begin{equation}
-u_{\ib-\half\eb_r}^{1D} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
-u_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
-\end{cases}
-\end{equation}
-Then, we do a full-dimensional extrapolation of $u$ to x-faces:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} = u_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
-u_{R,\ib-\half\eb_x}^{\mac,*} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
-\end{eqnarray}
-Finally, we solve a Riemann problem:
-\begin{equation}
-u_{\ib-\half\eb_x}^{\mac,*} =
-\begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} > 0, \\
-u_{R,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} < 0. 
-\end{cases}
-\end{equation}
-
-To compute $\wt^{\mac,*}$, we use an analogous procedure for computing $u^{\mac,*}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ and $\partial w_0/\partial r$ into the full-dimensional extrapolation.
-We compute a time-centered 1D extrapolation of $\wt$ to x-faces:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_x}^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
-\wt_{R,\ib-\half\eb_x}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \wt_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $u^{\trans}$:
-\begin{equation}
-\wt_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_x}^{1D} + \wt_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
-\wt_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-Then, we do a full-dimensional extrapolation of $\wt$ to r-faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{\mac,*} = \wt_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)(w_{0,\ib-\half\eb_r} - w_{0,\ib-\frac{3}{2}\eb_r}) + \frac{\dt}{2}f_{w,\ib-\eb_r},\label{wt_L Extrapolation} \\
-\wt_{R,\ib-\half\eb_r}^{\mac,*} = \wt_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)(w_{0,\ib+\half\eb_r} - w_{0,\ib-\half\eb_r}) + \frac{\dt}{2}f_{w,\ib}.\label{wt_R Extrapolation}
-\end{eqnarray}
-Next, we solve a Riemann problem:
-\begin{equation}
-\wt_{\ib-\half\eb_r}^{\mac,*} =
-\begin{cases}
-0, & \left(w_L^{\mac,*} \le 0 ~~ {\rm AND} ~~ w_R^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|w_L^{\mac,*} + w_R^{\mac,*}\right| < \epsilon, \\
-\wt_{L,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} < 0. 
-\end{cases}
-\end{equation}
-We then project $\Ubt^{\mac,*}$ to obtain $\Ubt^{\mac}$ and we call {\tt addw0} to compute $\Ub^{\mac}$ from $\Ubt^{\mac}$.
-%\subsection{3D Cartesian Case}
-%\subsection{3D Spherical Case}
-\cleardoublepage
-\section{Computing $\rho^{'\edge}, X_k^{\edge}$, and $(\rho h)^{'\edge}/h^{\edge}/T^{\edge}$ in {\tt MAESTRO}}
-In {\tt scalar\_advance}, we call {\tt make\_edge\_scal} to compute $\rho^{'\edge}, X_k^{\edge}$, and $(\rho h)^{'\edge}/h^{\edge}/T^{\edge}$ at each edge.
-The procedure is the same for each scalar, so we shall simply denote the scalar as $s$.
-We always need to compute $\rho'$ and $X_k$ to edges, and the choice of energy prediction is as follows:
-\begin{itemize}
-\item For {\tt enthalpy\_pred\_type} = 1, we predict $(\rho h)'$ to edges.
-\item For {\tt enthalpy\_pred\_type} = 2, we predict $h$ to edges.
-\item For {\tt enthalpy\_pred\_type} = 3 and 4, we predict $T$ to edges.
-\end{itemize}
-The force terms are:
-\begin{eqnarray}
-f_{\rho'} &=& -\rho'\nabla\cdot\Ub^{\mac} - \nabla\cdot(\rho_0\Ubt^{\mac}), \\
-f_{X_k} &=& 0, \\
-f_{(\rho h)'} &=& (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} -(\rho h)'\nabla\cdot\Ub^{\mac} - \nabla\cdot\left[(\rho h)_0\Ubt^{\mac}\right] + \nabla\cdot\kth\nabla T \\
-f_h &=& \frac{1}{\rho}\left(\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T \right), \\
-f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext} + \nabla\cdot\kth\nabla T \right\}.
-\end{eqnarray}
-If {\tt enthalpy\_pred\_type} = 3 or 4, we convert $T^{\edge}\rightarrow h^{\edge}$ using $h^{\edge} = h^{\edge}(\rho^{\edge},T^{\edge},X_k^{\edge})$, where $\rho^{\edge} = \rho^{'\edge} + \rho_0$.
-Furthermore, if {\tt enthalpy\_pred\_type} = 3, we compute $(\rho h)^{'\edge} = (\rho h)^{\edge} - (\rho h)_0$.
-\subsection{2D Cartesian Case}
-We begin by computing a time-centered 1D extrapolation of $s$ to all faces.  The extrapolation of $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib-\eb_x}^n, \\
-s_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $u^{\mac}$:
-\begin{equation}
-s_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{1D} + s_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-The extrapolation of $s$ to r-faces is analogous to the extrapolation to x-faces:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_r}^{1D} &=& s_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib-\eb_r}^n, \\
-s_{R,\ib-\half\eb_r}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $w^{\mac}$:
-\begin{equation}
-s_{\ib-\half\eb_r}^{1D} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_r}^{1D} + s_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_{L,\ib-\half\eb_r}^{1D}, & w^{\mac}_{\ib-\half\eb_r} > 0, \\
-u_{R,\ib-\half\eb_r}^{1D}, & w^{\mac}_{\ib-\half\eb_r} < 0. \\
-\end{cases}
-\end{equation}
-Now we do a full-dimensional extrapolation of $s$ to each face.  The extrapolation of $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\mac}+w_{\ib-\eb_x-\half\eb_r}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_r}^{1D} - s_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\mac}+w_{\ib-\half\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_r}^{1D} - s_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$.
-\begin{equation}
-s_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{\edge} + s_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
-\end{equation}
-The full-dimensional extrapolation of $s$ to r-faces is analogous to the extrapolation to x-faces:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_r}^{\edge} = s_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib-\eb_r}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\mac}+u_{\ib-\half\eb_x-\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_x-\eb_r}^{1D} - s_{\ib-\half\eb_x-\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_r},  \\
-s_{R,\ib-\half\eb_r}^{\edge} = s_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\mac}+u_{\ib-\half\eb_x}^{\mac}\right)\left(s_{\ib+\half\eb_x}^{1D} - s_{\ib-\half\eb_x}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}. 
-\end{eqnarray}
-Then, we upwind based on $w^{\mac}$.
-\begin{equation}
-s_{\ib-\half\eb_r}^{\edge} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_r}^{\edge} + s_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
-s_{L,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} > 0, \\
-s_{R,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} < 0.
-\end{cases}
-\end{equation}
-%\subsection{3D Cartesian Case}
-%\subsection{3D Spherical Case}
-\cleardoublepage
-\section{Computing $\Ubt^{\edge}$ in {\tt MAESTRO}}
-In {\tt velocity\_advance}, we call {\tt make\_edge\_state} to compute $\Ubt^{\edge}$.  Unlike when we computed $\Ubt^{\trans}$ and $\Ubt^{\mac,*}$, we need all components of velocity at each face.
-\subsection{2D Cartesian Case}
-We begin by computing a time-centered 1D extrapolation of $u$ to all faces.  The extrapolation of $u$ to x-faces is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{1D} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
-u_{R,\ib-\half\eb_x}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $u^{\trans}$:
-\begin{equation}
-u_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
-u_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
-u_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-The extrapolation of $u$ to r-faces is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_r}^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
-u_{R,\ib-\half\eb_r}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $w^{\trans}$:
-\begin{equation}
-u_{\ib-\half\eb_r}^{1D} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
-u_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
-\end{cases}
-\end{equation}
-We repeat this procedure to compute a time-centered 1D extrapolation of $\wt$ at each face.  The extrapolation of $\wt$ to x-faces is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_x}^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
-\wt_{R,\ib-\half\eb_x}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \wt_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $u^{\trans}$:
-\begin{equation}
-\wt_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_x}^{1D} + \wt_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
-\wt_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-The extrapolation of $\wt$ to r-faces is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{1D} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r}, \label{1D wt_L extrap} \\
-\wt_{R,\ib-\half\eb_r}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.
-\label{1D wt_R extrap}
-\end{eqnarray}
-Then we upwind based on $w^{\trans}$:
-\begin{equation}
-\wt_{\ib-\half\eb_r}^{1D} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_r}^{1D} + \wt_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
-\end{cases}
-\end{equation}
-Now we do a full-dimensional extrapolation of $u$ to each face.  The extrapolation of $u$ to x-faces is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\edge} = u_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib-\eb_x}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
-u_{R,\ib-\half\eb_x}^{\edge} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$.
-\begin{equation}
-u_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_x}^{\edge} + u_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-u_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-u_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
-\end{equation}
-The full-dimensional extrapolation of $u$ to r-faces is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_r}^{\edge} = u_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib-\eb_r}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_x-\eb_r}^{1D} - u_{\ib-\half\eb_x-\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_r},  \\
-u_{R,\ib-\half\eb_r}^{\edge} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(u_{\ib+\half\eb_x}^{1D} - u_{\ib-\half\eb_x}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}. 
-\end{eqnarray}
-Then, we upwind based on $w^{\mac}$.
-\begin{equation}
-u_{\ib-\half\eb_r}^{\edge} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_r}^{\edge} + u_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_{L,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} > 0, \\
-u_{R,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} < 0.
-\end{cases}
-\end{equation}
-Now we do a full-dimensional extrapolatino of $\wt$ to each face.
-This is different from the extrapolation of $u$ to faces because we must include the term proportional to $\partial w_0/\partial r$ in equation (\ref{Perturbational Velocity Equation}).
-The full-dimensional extrapolation of $\wt$ to x-faces is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_x}^{\edge} = \wt_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib-\eb_x}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(\wt_{\ib-\eb_x+\half\eb_r}^{1D} - \wt_{\ib-\eb_x-\half\eb_r}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib-\eb_x+\half\eb_r}^{\trans}+\wt_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(w_{0,\ib-\eb_x+\half\eb_r}-w_{0,\ib-\eb_x-\half\eb_r}\right) + \frac{\dt}{2}f_{w,\ib-\eb_x}, \label{wt_L x-edge} \\
-\wt_{R,\ib-\half\eb_x}^{\edge} = \wt_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_r}^{1D} - \wt_{\ib-\half\eb_r}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right) + \frac{\dt}{2}f_{w,\ib}. \label{wt_R x-edge}
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$:
-\begin{equation}
-\wt_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_x}^{\edge} + \wt_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-\wt_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
-\end{equation}
-The full-dimensional extrapolation of $\wt$ to r-faces is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{\edge} = \wt_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib-\eb_r}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)\left(w_{0,\ib-\half\eb_r}-w_{0,\ib-\frac{3}{2}\eb_r}\right) + \frac{\dt}{2}f_{w,\ib-\eb_r}, \label{wt_L r-edge} \\
-\wt_{R,\ib-\half\eb_r}^{\edge} = \wt_{\ib}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right) + \frac{\dt}{2}f_{w,\ib}. \label{wt_R r-edge}
-\end{eqnarray}
-Then, we upwind based on $w^{\mac}$:
-\begin{equation}
-\wt_{\ib-\half\eb_r}^{\edge} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_r}^{\edge} + \wt_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_r}^{\edge}, & w^{\mac}_{\ib-\half\eb_r} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{\edge}, & w^{\mac}_{\ib-\half\eb_r} < 0.
-\end{cases}
-\end{equation}
-%\subsection{3D Cartesian Case}
-%\subsection{3D Spherical Case}
-\cleardoublepage
-\section{Computing $\Ub^{\mac,*}$ in {\tt VARDEN}}
-\subsection{2D Cartesian Case}
-We do a 1D Taylor series extrapolation to get both components of velocity at the x-face:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{1D} &=& u_{\ib-\eb_x} + \left[\half - \frac{\dt}{2h}{\rm max}(0,u_{\ib-\eb_x})\right]\Delta_xu_{\ib-\eb_x}, \label{varden U_L^1D} \\
-u_{R,\ib-\half\eb_x}^{1D} &=& u_{\ib} + \left[\half - \frac{\dt}{2h}{\rm min}(0,u_{\ib})\right]\Delta_xu_{\ib}.
-\end{eqnarray}
-\begin{eqnarray}
-v_{L,\ib-\half\eb_x}^{1D} &=& v_{\ib-\eb_x} + \left[\half - \frac{\dt}{2h}{\rm max}(0,v_{\ib-\eb_x})\right]\Delta_xv_{\ib-\eb_x}, \label{varden U_L^1D} \\
-v_{R,\ib-\half\eb_x}^{1D} &=& v_{\ib} + \left[\half - \frac{\dt}{2h}{\rm min}(0,v_{\ib})\right]\Delta_xv_{\ib}.
-\end{eqnarray}
-We obtain the normal velocity using the Riemann problem:
-\begin{equation}
-u_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{1D} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{1D} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{1D}, & u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D} > 0, \\
-u_{R,\ib-\half\eb_x}^{1D}, & u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D} < 0.
-\end{cases}
-\end{equation}
-We obtain the transverse velocity by upwinding based on $u_{\ib-\half\eb_x}^{1D}$:
-\begin{equation}
-v_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(v_{L,\ib-\half\eb_x}^{1D} + v_{R,\ib-\half\eb_x}^{1D}\right), & \left|u_{\ib-\half\eb_x}^{1D}\right| < \epsilon \\
-v_{L,\ib-\half\eb_x}^{1D}, & u_{\ib-\half\eb_x}^{1D} > 0, \\
-v_{R,\ib-\half\eb_x}^{1D}, & u_{\ib-\half\eb_x}^{1D} < 0.
-\end{cases}\label{Transverse Velocity Riemann Problem}
-\end{equation}
-We perform analogous operations to compute both components of velocity at the y-faces, $\Ub_{\ib-\half\eb_y}^{1D}$. \\
-
-Now we do a full-dimensional extrapolation to get the MAC velocity at the x-faces (note that we only compute the normal components):
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
-u_{R,\ib-\half\eb_x}^{\mac,*} &=& u_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
-\end{eqnarray}
-Then, we solve a Riemann problem:
-\begin{equation}
-u_{\ib-\half\eb_x}^{\mac,*} =
-\begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} > 0, \\
-u_{R,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} < 0.
-\end{cases}\label{umac Riemann Problem}
-\end{equation}
-We perform analogous operations to compute the normal velocity at the y-faces, $v^{\mac,*}_{\ib-\half\eb_y}$.
-\subsection{3D Cartesian Case}
-This is more complicated than the 2D case because we include corner coupling.
-We compute $\Ub_{\ib-\half\eb_x}^{1D}, \Ub_{\ib-\half\eb_y}^{1D}$, and $\Ub_{\ib-\half\eb_z}^{1D}$ in an analogous manner as equations (\ref{varden U_L^1D})-(\ref{Transverse Velocity Riemann Problem}).
-Then, we compute an intermediate state, $u_{\ib-\half\eb_y}^{y|z}$, which is described as ``state $u_{\ib-\half\eb_y}^{1D}$ that has been updated to account for the transverse derivatives in the z direction'', using:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_y}^{y|z} &=& u_{L,\ib-\half\eb_y}^{1D} - \frac{\dt}{6h}\left(w_{\ib-\eb_y+\half\eb_z}^{1D}+w_{\ib-\eb_y-\half\eb_z}^{1D}\right)\left(u_{\ib-\eb_y+\half\eb_z}^{1D}-u_{\ib-\eb_y-\half\eb_z}^{1D}\right), \\
-u_{R,\ib-\half\eb_y}^{y|z} &=& u_{R,\ib-\half\eb_y}^{1D} - \frac{\dt}{6h}\left(w_{\ib+\half\eb_z}^{1D}+w_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{1D}-u_{\ib-\half\eb_z}^{1D}\right).
-\end{eqnarray}
-Then, upwind based on $v_{\ib-\half\eb_y}^{1D}$:
-\begin{equation}
-u_{\ib-\half\eb_y}^{y|z} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_y}^{y|z} + u_{R,\ib-\half\eb_y}^{y|z}\right), & \left|v_{\ib-\half\eb_y}^{1D}\right| < \epsilon \\
-u_{L,\ib-\half\eb_y}^{y|z}, & v_{\ib-\half\eb_y}^{1D} > 0, \\
-u_{R,\ib-\half\eb_y}^{y|z}, & v_{\ib-\half\eb_y}^{1D} < 0.
-\end{cases}
-\end{equation}
-We use an analogous procedure to compute five more intemediate states, $u_{\ib-\half\eb_z}^{z|y}, v_{\ib-\half\eb_x}^{x|z}, v_{\ib-\half\eb_z}^{z|x}, w_{\ib-\half\eb_x}^{x|y}$, and $w_{\ib-\half\eb_y}^{y|x}$.
-Then, we do a full-dimensional extrapolation to get the MAC velocities at normal faces:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} = u_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{y|z}-u_{\ib-\eb_x-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_z}^{1D}+w_{\ib-\eb_x-\half\eb_z}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_z}^{z|y}-u_{\ib-\eb_x-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
-u_{R,\ib-\half\eb_x}^{\mac,*} = u_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{y|z}-u_{\ib-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{1D}+w_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{z|y}-u_{\ib-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{u,\ib}.
-\end{eqnarray}
-Then, we use the Riemann solver given above for the 2D case (equation [\ref{umac Riemann Problem}]) to compute $u_{\ib-\half\eb_x}^{\mac,*}$.
-We use an analogous procedure to obtain $v_{\ib-\half\eb_y}^{\mac,*}$ and $w_{\ib-\half\eb_z}^{\mac,*}$.
-\cleardoublepage
-\section{Computing $\Ub^{\edge}$ and $\rho^{\edge}$ in {\tt VARDEN}}
-To compute $\Ub^{\edge}$, {\tt VARDEN} uses the exact same algorithm as the $s^{\edge}$ case in {\tt MAESTRO}.
-The algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.
-Here, $s$ is used in place of either $\rho, u, v$, or $w$ (in 3D).
-\subsection{2D Cartesian Case}
-The 1D extrapolation is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib-\eb_x}^n, \label{varden s_L^1D}\\
-s_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib}^n. \label{varden s_R^1D}
-\end{eqnarray}
-Then we upwind based on $u^{\mac}$:
-\begin{equation}
-s_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{1D} + s_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-We use an analogous procedure to obtain $s_{\ib-\half\eb_y}^{1D}$.  Now we do a full-dimensional extrapolation of $s$ to each face.  The extrapolation of a ``non-conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-The extrapolation of a ``conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] \nonumber \\
-&-& \frac{\dt}{2}s_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\half\eb_y}\right] \nonumber \\
-&-& \frac{\dt}{2}s_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right) + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$.
-\begin{equation}
-s_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{\edge} + s_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}\label{varden s^edge upwind}
-\end{equation}
-We use an analogous procedure to compute $s_{\ib-\half\eb_y}^{\edge}$.
-\subsection{3D Cartesian Case}
-This is more complicated than the 2D case because we include corner coupling.
-We first compute $s_{\ib-\half\eb_x}^{1D}$, $s_{\ib-\half\eb_y}^{1D}$, and $s_{\ib-\half\eb_z}^{1D}$ in an analogous manner to equations (\ref{varden s_L^1D}) and (\ref{varden s_R^1D}).
-Then we compute six intermediate states, $s_{\ib-\half\eb_x}^{x|y}, s_{\ib-\half\eb_x}^{x|z}, s_{\ib-\half\eb_y}^{y|x}, s_{\ib-\half\eb_y}^{y|z}, s_{\ib-\half\eb_z}^{z|x}$, and $s_{\ib-\half\eb_z}^{z|y}$.
-For the ``non-conservative case'', we use, for example:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{x|y} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{6h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} + v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
-s_{R,\ib-\half\eb_x}^{x|y} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{6h}\left(v_{\ib+\half\eb_y}^{\mac} + v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right).
-\end{eqnarray}
-For the ``conservative'' case, we use, for example:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{x|y} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{3h}\left[(sv^{\mac})_{\ib-\eb_x+\half\eb_y}-(sv^{\mac})_{\ib-\eb_x-\half\eb_y}\right], \\
-s_{R,\ib-\half\eb_x}^{x|y} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{3h}\left[(sv^{\mac})_{\ib+\half\eb_y}-(sv^{\mac})_{\ib-\half\eb_y}\right].
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$:
-\begin{equation}
-s_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{x|y} + s_{R,\ib-\half\eb_x}^{x|y}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{x|y}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{x|y}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
-\end{equation}
-We use an analogous procedure to compute the other five intermediate states.
-Now we do a full-dimensional extrapolation of $s$ to each face.
-The extrapolation of a ``non-conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{y|z}-s_{\ib-\eb_x-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_z}^{\mac}+w_{\ib-\eb_x-\half\eb_z}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_z}^{z|y}-s_{\ib-\eb_x-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{y|z}-s_{\ib-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{\mac}+w_{\ib-\half\eb_z}^{\mac}\right)\left(s_{\ib+\half\eb_z}^{z|y}-s_{\ib-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-The extrapolation of a ``conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{y|z}v^{\mac})_{\ib-\eb_x+\half\eb_y}-({s^{y|z}v^{\mac})_{\ib-\eb_x-\half\eb_y}}\right] \nonumber \\
-&-& \frac{\dt}{2h}\left[(s^{z|y}w^{\mac})_{\ib-\eb_x+\half\eb_z}-({s^{z|y}w^{\mac})_{\ib-\eb_x-\half\eb_z}}\right] \nonumber \\
-&-& \frac{\dt}{2}s_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}-v_{\ib-\eb_x-\half\eb_y}^{\mac}+w_{\ib-\eb_x+\half\eb_z}^{\mac}-w_{\ib-\eb_x-\half\eb_z}^{\mac}\right) \nonumber \\
-&+& \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{y|z}v^{\mac})_{\ib+\half\eb_y}-({s^{y|z}v^{\mac})_{\ib-\half\eb_y}}\right] \nonumber \\
-&-& \frac{\dt}{2h}\left[(s^{z|y}w^{\mac})_{\ib+\half\eb_z}-({s^{z|y}w^{\mac})_{\ib-\half\eb_z}}\right] \nonumber \\
-&-& \frac{\dt}{2}s_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac}-v_{\ib-\half\eb_y}^{\mac}+w_{\ib+\half\eb_z}^{\mac}-w_{\ib-\half\eb_z}^{\mac}\right) \nonumber \\
-&+& \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-Then we upwind based on $u^{\mac}$, as in equation (\ref{varden s^edge upwind}).  We use an analogous procedure to compute both $s_{\ib-\half\eb_y}^{\edge}$ and $s_{\ib-\half\eb_z}$.
 \end{document}

### 2008-07-29 docs/pert_notes/paper.tex
some density notes

@@
 % Margins
-\usepackage[lmargin=0.5in,rmargin=0.5in,tmargin=0.75in,bmargin=0.75in]{geometry}
+\usepackage[lmargin=0.5in,rmargin=1.5in,tmargin=1.0in,bmargin=1.0in]{geometry}
+
+\def\half  {\frac{1}{2}}
+
+\def\eb    {{\bf e}}
+\def\Ub    {{\bf U}}
+\def\Ubt   {\widetilde{\bf U}}
 
@@
 
-\setlength{\marginparwidth}{0.75in}
+\setlength{\marginparwidth}{1.0in}
 \newcommand{\MarginPar}[1]{\marginpar{%
@@
 
+\section{Density Evolution}
+Full density evolution equation:
+\begin{equation}
+\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub). \label{rho equation}
+\end{equation}
+\subsection{Predict $\rho'$ (false - default)}
+Base state density evolution equation:
+\begin{eqnarray}
+\frac{\partial\rho_0}{\partial t} &=& -\nabla\cdot(\rho_0 w_0 \eb_r) - \nabla\cdot(\eta_\rho \eb_r) \nonumber \\
+&=& -w_0\frac{\partial\rho_0}{\partial r} \underbrace{- \rho_0\frac{\partial w_0}{\partial r}} - \nabla\cdot(\eta_\rho \eb_r). \label{rho0 equation}
+\end{eqnarray}
+Note that $\Ub = \Ubt + w_o\eb_r$.  Subtract (\ref{rho0 equation}) from (\ref{rho equation}) to give perturbational enthalpy equation:
+\begin{eqnarray}
+\frac{\partial\rho'}{\partial t} &=& -\nabla\cdot(\rho'\Ub) - \nabla\cdot(\rho_0\Ubt) + \nabla\cdot(\eta_\rho \eb_r) \nonumber \\
+&=& -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub - \nabla\cdot(\rho_0\Ubt)} + \nabla\cdot(\eta_\rho \eb_r).\label{rho' equation}
+\end{eqnarray}
+\subsubsection{Predicing $\rho'$ at edges}
+We predict $\rho'$ to edges using make\_edge\_scal in scalar\_advance.
+The underbraced term in (\ref{rho' equation}) is computed in modify\_scal\_force.
+The $\eta_\rho$ term is not included, but this is correct because the procedure to compute $\rho$ at edges described in Section \ref{Computing rho at edges} makes use of an edge-centered, time-centered $\rho_0^{n+\half}$ that does not include an $\eta_\rho$ contribution either.
+\subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
+We first compute $\rho_0^{(2)}$ using advect\_base (but not correct\_base).
+We compute $\rho$ on edges by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at edges.
+For the non-radial faces, we can directly add $\rho_0^{n+\half}$ since it is a cell-centered quantity.
+For the radial faces, we interpolate to obtain $\rho_0^{n+\half}$ at radial faces.
+\subsection{Predict $\rho$ (true)}
+\subsection{$\rho_0^{n+\half,{\rm pred}}$}
+To compute $\eta_\rho$, we need $\rho_0$ at edge-centered, time-centered locations.
+However, we do NOT use the same edge-centered, time-centered $\rho_0^{n+\half}$ that was referred to in Section \ref{Computing rho at edges}.
+Instead, we use a $\rho_0$ computed with a Godunov method in make\_edge\_state\_1d.
+We call make\_edge\_state\_1d using the underbraced term in (\ref{rho0 equation}) as the forcing.
+The $\eta_\rho$ term is not included.\MarginPar{I think the $\eta_\rho$ term needs to be included here.}  This gives us the so-called $\rho_0^{n+\half,{\rm pred}}$.
+\section{Energy Evolution}
+\subsection{Predict $(\rho h)'$ (option 1 - default)}
+\subsection{Predict $h$ (option 2)}
+\subsection{Predict $T$ into $(\rho h)'$ (option 3)}
+\subsection{Predict $T$ into $h$ (option 4)}
 \end{document}

### 2008-07-29 docs/pert_notes/paper.tex
fixed typo
clarified flag options

@@
 \end{equation}
-\subsection{Predict $\rho'$ (false - default)}
+\subsection{Predicting $\rho'$ (predict\_rho = false - default)}
 Base state density evolution equation:
@@
 \end{eqnarray}
-Note that $\Ub = \Ubt + w_o\eb_r$.  Subtract (\ref{rho0 equation}) from (\ref{rho equation}) to give perturbational enthalpy equation:
+Note that $\Ub = \Ubt + w_o\eb_r$.  Subtract (\ref{rho0 equation}) from (\ref{rho equation}) to give perturbational density equation:
 \begin{eqnarray}
@@
 For the radial faces, we interpolate to obtain $\rho_0^{n+\half}$ at radial faces.
-\subsection{Predict $\rho$ (true)}
+\subsection{Predicting $\rho$ (predict\_rho = true)}
 \subsection{$\rho_0^{n+\half,{\rm pred}}$}
@@
 \section{Energy Evolution}
-\subsection{Predict $(\rho h)'$ (option 1 - default)}
-\subsection{Predict $h$ (option 2)}
-\subsection{Predict $T$ into $(\rho h)'$ (option 3)}
-\subsection{Predict $T$ into $h$ (option 4)}
+\subsection{Predicting $(\rho h)'$ (enthalpy\_pred\_type = 1 - default)}
+\subsection{Predicting $h$ (enthalpy\_pred\_type = 2)}
+\subsection{Predicting $T$ into $(\rho h)'$ (enthalpy\_pred\_type = 3)}
+\subsection{Predicting $T$ into $h$ (enthalpy\_pred\_type = 4)}
 \end{document}

### 2008-07-29 docs/pert_notes/paper.tex
Completed rho section.

@@
 \end{equation}
-\subsection{Predicting $\rho'$ (predict\_rho = false - default)}
+\subsection{Predicting $\rho'$ (predict\_rho = F - default)}
 Base state density evolution equation:
@@
 \end{eqnarray}
-\subsubsection{Predicing $\rho'$ at edges}
+\subsubsection{Predicting $\rho'$ at edges}
 We predict $\rho'$ to edges using make\_edge\_scal in scalar\_advance.
@@
 For the radial faces, we interpolate to obtain $\rho_0^{n+\half}$ at radial faces.
-\subsection{Predicting $\rho$ (predict\_rho = true)}
+\subsection{Predicting $\rho$ (predict\_rho = T)}
+\subsubsection{Predicting $\rho$ at edges}
+We predict $\rho$ to edges using make\_edge\_scal in scalar\_advance.
+The forcing term is set to zero, and therefor we set is\_conservative = T to alter the discretization in make\_edge\_scal.\MarginPar{I think there is a bug in the code since is\_conservative is always set to T, even if we are predicting $\rho'$.}
 \subsection{$\rho_0^{n+\half,{\rm pred}}$}
@@
 We call make\_edge\_state\_1d using the underbraced term in (\ref{rho0 equation}) as the forcing.
-The $\eta_\rho$ term is not included.\MarginPar{I think the $\eta_\rho$ term needs to be included here.}  This gives us the so-called $\rho_0^{n+\half,{\rm pred}}$.
+The $\eta_\rho$ term is not included.\MarginPar{I think there is a bug in the code and the $\eta_\rho$ term needs to be included here.}  This gives us the so-called $\rho_0^{n+\half,{\rm pred}}$.
 \section{Energy Evolution}

### 2008-07-29 docs/pert_notes/paper.tex
wrote rhohprime section

@@
 \def\Ubt   {\widetilde{\bf U}}
+\def\wt    {\widetilde{w}}
 
@@
 Full density evolution equation:
-\begin{equation}
-\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub). \label{rho equation}
-\end{equation}
+\begin{eqnarray}
+\frac{\partial\rho}{\partial t} &=& -\nabla\cdot(\rho\Ub) \nonumber \\
+&=& -\Ub\cdot\nabla\rho -\rho\nabla\cdot\Ub. \label{rho equation}
+\end{eqnarray}
 \subsection{Predicting $\rho'$ (predict\_rho = F - default)}
@@
 The forcing term is set to zero, and therefor we set is\_conservative = T to alter the discretization in make\_edge\_scal.\MarginPar{I think there is a bug in the code since is\_conservative is always set to T, even if we are predicting $\rho'$.}
-\subsection{$\rho_0^{n+\half,{\rm pred}}$}
+\subsection{Predicting $\rho_0$}
 To compute $\eta_\rho$, we need $\rho_0$ at edge-centered, time-centered locations.
@@
 \section{Energy Evolution}
+Full enthalpy equation:
+\begin{eqnarray}
+\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt} + \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext} \nonumber \\
+&=& -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt} + \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}.\label{rhoh equation}
+\end{eqnarray}
+Full temperature equation:
+\begin{equation}
+\frac{\partial T}{\partial t} = -\Ub\cdot\nabla T + \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
+\end{equation}
 \subsection{Predicting $(\rho h)'$ (enthalpy\_pred\_type = 1 - default)}
+These are working notes for what I'd like to do for the energy evolution.
+\subsubsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
+We can arbitrarily define a base state enthalpy evolution equation:
+\begin{eqnarray}
+\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] + \frac{D_0p_0}{Dt} + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} + \overline{\rho H_{\rm ext}} \nonumber \\
+&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} + \psi + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} + \overline{\rho H_{\rm ext}}.\label{rhoh0 equation}
+\end{eqnarray}
+We will predict both $(\rho h)_0$ and $(\rho h)'$ to edges using a Godunov method.
+Then we will add the two answers together to get $\rho h$ at edges.
+This is different from the density, in which we averaged an ``old'' and ``new'' base state density to get $\rho_0$ at edges.
+By computing $\rho h$ at edges in this way, we avoid having to compute the complete evolution of $(\rho h)_0$, and therefore avoid having to write a correct\_base and diffuse\_base function for $\rho h$.\\
+
+To simplify matters, the we have already called react\_state, so for now we will ignore the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms in the remaining derivations.
+Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and ignoring the aforementioned terms gives the perturbational enthalpy equation:
+\begin{eqnarray}
+\frac{\partial(\rho h)'}{\partial t} &=& -\nabla\cdot[(\rho h)'\Ub] - \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r} + (\nabla\cdot\kappa\nabla T - \overline{\nabla\cdot\kappa\nabla T}) \nonumber \\
+&=& -\Ub\cdot\nabla(\rho h)' \underbrace{- (\rho h)'\nabla\cdot\Ub - \nabla\cdot[(\rho h)_0\Ubt]} + \wt\frac{\partial p_0}{\partial r} + (\nabla\cdot\kappa\nabla T - \overline{\nabla\cdot\kappa\nabla T})\label{rhohprime equation}
+\end{eqnarray}
+We use make\_edge\_scal to get $(\rho h)'$ at edges.
+The function modify\_scal\_force computes the underbraced term in (\ref{rhohprime equation}).
+The function mkrhohforce computes the $\wt$ term.\MarginPar{I need to write a function to handle the thermal diffusion terms.}
+\subsubsection{Predicting $(\rho h)_0$ at edges}\label{Predicting rhoh0 at edges}
+After the call to react\_state, we set $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$.
+We use the Godunov algorithm in make\_edge\_state\_1d to predict $(\rho h)_0$ at edges, omitting the terms added in react\_state, so we are essentially evolving an abbreviated version of (\ref{rhoh0 equation}):
+\begin{eqnarray}
+\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] + \psi + \overline{\nabla\cdot\kappa\nabla T} \nonumber \\
+&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} + \psi + \overline{\nabla\cdot\kappa\nabla T}.\label{rhoh0 equation}
+\end{eqnarray}
+In advect\_base, the forcing term is set to the term proportional to $(\rho h)_0$.\MarginPar{We are currently omitting the $\psi$ and thermal conduction terms.  I think this is a bug.}
+\subsubsection{Computing $\rho h$ at edges}
+Simply add the results from Sections \ref{Predicting rhohprime at edges} and \ref{Predicting rhoh0 at edges}, using spatial interpolation to get $(\rho h)_0$ at non-radial faces.
 \subsection{Predicting $h$ (enthalpy\_pred\_type = 2)}

### 2008-07-29 docs/pert_notes/paper.tex
removed a marginpar that we addressed in the code

@@
 \subsubsection{Predicting $\rho$ at edges}
-We predict $\rho$ to edges using make\_edge\_scal in scalar\_advance.
-The forcing term is set to zero, and therefor we set is\_conservative = T to alter the discretization in make\_edge\_scal.\MarginPar{I think there is a bug in the code since is\_conservative is always set to T, even if we are predicting $\rho'$.}
+We predict $\rho$ to edges using make\_edge\_scal in scalar\_advance.  The forcing term is set to zero, and therefor we set is\_conservative = T to alter the discretization in make\_edge\_scal.
 \subsection{Predicting $\rho_0$}

### 2008-07-29 docs/pert_notes/paper.tex
made doc vi friendly

@@
 \begin{eqnarray}
-\frac{\partial\rho_0}{\partial t} &=& -\nabla\cdot(\rho_0 w_0 \eb_r) - \nabla\cdot(\eta_\rho \eb_r) \nonumber \\
-&=& -w_0\frac{\partial\rho_0}{\partial r} \underbrace{- \rho_0\frac{\partial w_0}{\partial r}} - \nabla\cdot(\eta_\rho \eb_r). \label{rho0 equation}
+\frac{\partial\rho_0}{\partial t} &=& -\nabla\cdot(\rho_0 w_0 \eb_r) - 
+\nabla\cdot(\eta_\rho \eb_r) \nonumber \\
+&=& -w_0\frac{\partial\rho_0}{\partial r} \underbrace{- 
+\rho_0\frac{\partial w_0}{\partial r}} 
+- \nabla\cdot(\eta_\rho \eb_r). \label{rho0 equation}
 \end{eqnarray}
-Note that $\Ub = \Ubt + w_o\eb_r$.  Subtract (\ref{rho0 equation}) from (\ref{rho equation}) to give perturbational density equation:
+Note that $\Ub = \Ubt + w_o\eb_r$.  Subtract (\ref{rho0 equation}) from 
+(\ref{rho equation}) to give perturbational density equation:
 \begin{eqnarray}
-\frac{\partial\rho'}{\partial t} &=& -\nabla\cdot(\rho'\Ub) - \nabla\cdot(\rho_0\Ubt) + \nabla\cdot(\eta_\rho \eb_r) \nonumber \\
-&=& -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub - \nabla\cdot(\rho_0\Ubt)} + \nabla\cdot(\eta_\rho \eb_r).\label{rho' equation}
+\frac{\partial\rho'}{\partial t} &=& -\nabla\cdot(\rho'\Ub) - \nabla\cdot(\rho_0\Ubt) 
++ \nabla\cdot(\eta_\rho \eb_r) \nonumber \\
+&=& -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub - \nabla\cdot(\rho_0\Ubt)} 
++ \nabla\cdot(\eta_\rho \eb_r).\label{rho' equation}
 \end{eqnarray}
 \subsubsection{Predicting $\rho'$ at edges}
-We predict $\rho'$ to edges using make\_edge\_scal in scalar\_advance.
-The underbraced term in (\ref{rho' equation}) is computed in modify\_scal\_force.
-The $\eta_\rho$ term is not included, but this is correct because the procedure to compute $\rho$ at edges described in Section \ref{Computing rho at edges} makes use of an edge-centered, time-centered $\rho_0^{n+\half}$ that does not include an $\eta_\rho$ contribution either.
+We predict $\rho'$ to edges using make\_edge\_scal in scalar\_advance.  The underbraced 
+term in (\ref{rho' equation}) is computed in modify\_scal\_force.  The $\eta_\rho$ term 
+is not included, but this is correct because the procedure to compute $\rho$ at edges 
+described in Section \ref{Computing rho at edges} makes use of an edge-centered, 
+time-centered $\rho_0^{n+\half}$ that does not include an $\eta_\rho$ contribution either.
 \subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
-We first compute $\rho_0^{(2)}$ using advect\_base (but not correct\_base).
-We compute $\rho$ on edges by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at edges.
-For the non-radial faces, we can directly add $\rho_0^{n+\half}$ since it is a cell-centered quantity.
-For the radial faces, we interpolate to obtain $\rho_0^{n+\half}$ at radial faces.
+We first compute $\rho_0^{(2)}$ using advect\_base (but not correct\_base).  We compute 
+$\rho$ on edges by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at 
+edges. For the non-radial faces, we can directly add $\rho_0^{n+\half}$ since it is a 
+cell-centered quantity.  For the radial faces, we interpolate to obtain $\rho_0^{n+\half}$ 
+at radial faces.
 \subsection{Predicting $\rho$ (predict\_rho = T)}
 \subsubsection{Predicting $\rho$ at edges}
-We predict $\rho$ to edges using make\_edge\_scal in scalar\_advance.  The forcing term is set to zero, and therefor we set is\_conservative = T to alter the discretization in make\_edge\_scal.
+We predict $\rho$ to edges using make\_edge\_scal in scalar\_advance.  The forcing term 
+is set to zero, and therefore we set is\_conservative = T to alter the discretization 
+in make\_edge\_scal.
 \subsection{Predicting $\rho_0$}
 To compute $\eta_\rho$, we need $\rho_0$ at edge-centered, time-centered locations.  
-However, we do NOT use the same edge-centered, time-centered $\rho_0^{n+\half}$ that was referred to in Section \ref{Computing rho at edges}.
-Instead, we use a $\rho_0$ computed with a Godunov method in make\_edge\_state\_1d.
-We call make\_edge\_state\_1d using the underbraced term in (\ref{rho0 equation}) as the forcing.
-The $\eta_\rho$ term is not included.\MarginPar{I think there is a bug in the code and the $\eta_\rho$ term needs to be included here.}  This gives us the so-called $\rho_0^{n+\half,{\rm pred}}$.
+However, we do NOT use the same edge-centered, time-centered $\rho_0^{n+\half}$ that was 
+referred to in Section \ref{Computing rho at edges}.  Instead, we use a $\rho_0$ computed 
+with a Godunov method in make\_edge\_state\_1d.  We call make\_edge\_state\_1d using the 
+underbraced term in (\ref{rho0 equation}) as the forcing.  The $\eta_\rho$ term is not 
+included.\MarginPar{I think there is a bug in the code and the $\eta_\rho$ term needs to 
+be included here.}  This gives us the so-called $\rho_0^{n+\half,{\rm pred}}$.
 \section{Energy Evolution}
@@
 \begin{eqnarray}
-\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt} + \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext} \nonumber \\
-&=& -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt} + \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}.\label{rhoh equation}
+\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt} 
++ \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext} \nonumber \\
+&=& -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub -\nabla\cdot(\rho h \Ub) 
++ \frac{Dp_0}{Dt} + \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} 
++ \rho H_{\rm ext}.\label{rhoh equation}
 \end{eqnarray}
@@
 \begin{equation}
-\frac{\partial T}{\partial t} = -\Ub\cdot\nabla T + \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
+\frac{\partial T}{\partial t} = -\Ub\cdot\nabla T
++ \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi
++ (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 
++ \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
 \end{equation}
@@
 \begin{eqnarray}
-\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] + \frac{D_0p_0}{Dt} + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} + \overline{\rho H_{\rm ext}} \nonumber \\
-&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} + \psi + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} + \overline{\rho H_{\rm ext}}.\label{rhoh0 equation}
+\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] 
++ \frac{D_0p_0}{Dt} + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} 
++ \overline{\rho H_{\rm ext}} \nonumber \\
+&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} 
++ \psi + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} 
++ \overline{\rho H_{\rm ext}}.\label{rhoh0 equation}
 \end{eqnarray}
-We will predict both $(\rho h)_0$ and $(\rho h)'$ to edges using a Godunov method.
-Then we will add the two answers together to get $\rho h$ at edges.
-This is different from the density, in which we averaged an ``old'' and ``new'' base state density to get $\rho_0$ at edges.
-By computing $\rho h$ at edges in this way, we avoid having to compute the complete evolution of $(\rho h)_0$, and therefore avoid having to write a correct\_base and diffuse\_base function for $\rho h$.\\
+We will predict both $(\rho h)_0$ and $(\rho h)'$ to edges using a Godunov method.  Then 
+we will add the two answers together to get $\rho h$ at edges.  This is different from 
+the density, in which we averaged an ``old'' and ``new'' base state density to get 
+$\rho_0$ at edges.  By computing $\rho h$ at edges in this way, we avoid having to 
+compute the complete evolution of $(\rho h)_0$, and therefore avoid having to write a 
+correct\_base and diffuse\_base function for $\rho h$.\\
 
-To simplify matters, the we have already called react\_state, so for now we will ignore the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms in the remaining derivations.
-Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and ignoring the aforementioned terms gives the perturbational enthalpy equation:
+To simplify matters, the we have already called react\_state, so for now we will ignore 
+the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms in the remaining derivations.  
+Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and ignoring the 
+aforementioned terms gives the perturbational enthalpy equation:
 \begin{eqnarray}
-\frac{\partial(\rho h)'}{\partial t} &=& -\nabla\cdot[(\rho h)'\Ub] - \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r} + (\nabla\cdot\kappa\nabla T - \overline{\nabla\cdot\kappa\nabla T}) \nonumber \\
-&=& -\Ub\cdot\nabla(\rho h)' \underbrace{- (\rho h)'\nabla\cdot\Ub - \nabla\cdot[(\rho h)_0\Ubt]} + \wt\frac{\partial p_0}{\partial r} + (\nabla\cdot\kappa\nabla T - \overline{\nabla\cdot\kappa\nabla T})\label{rhohprime equation}
+\frac{\partial(\rho h)'}{\partial t} &=& -\nabla\cdot[(\rho h)'\Ub] 
+- \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r} 
++ (\nabla\cdot\kappa\nabla T - \overline{\nabla\cdot\kappa\nabla T}) \nonumber \\
+&=& -\Ub\cdot\nabla(\rho h)' \underbrace{- (\rho h)'\nabla\cdot\Ub 
+- \nabla\cdot[(\rho h)_0\Ubt]} + \wt\frac{\partial p_0}{\partial r} 
++ (\nabla\cdot\kappa\nabla T 
+- \overline{\nabla\cdot\kappa\nabla T})\label{rhohprime equation}
 \end{eqnarray}
-We use make\_edge\_scal to get $(\rho h)'$ at edges.
-The function modify\_scal\_force computes the underbraced term in (\ref{rhohprime equation}).
-The function mkrhohforce computes the $\wt$ term.\MarginPar{I need to write a function to handle the thermal diffusion terms.}
+We use make\_edge\_scal to get $(\rho h)'$ at edges.  The function modify\_scal\_force 
+computes the underbraced term in (\ref{rhohprime equation}).  The function mkrhohforce 
+computes the $\wt$ term.\MarginPar{I need to write a function to handle the thermal 
+diffusion terms.}
 \subsubsection{Predicting $(\rho h)_0$ at edges}\label{Predicting rhoh0 at edges}
 After the call to react\_state, we set $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$.  
-We use the Godunov algorithm in make\_edge\_state\_1d to predict $(\rho h)_0$ at edges, omitting the terms added in react\_state, so we are essentially evolving an abbreviated version of (\ref{rhoh0 equation}):
+We use the Godunov algorithm in make\_edge\_state\_1d to predict $(\rho h)_0$ at edges, 
+omitting the terms added in react\_state, so we are essentially evolving an abbreviated 
+version of (\ref{rhoh0 equation}):
 \begin{eqnarray}
-\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] + \psi + \overline{\nabla\cdot\kappa\nabla T} \nonumber \\
-&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} + \psi + \overline{\nabla\cdot\kappa\nabla T}.\label{rhoh0 equation}
+\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] + \psi 
++ \overline{\nabla\cdot\kappa\nabla T} \nonumber \\
+&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} 
++ \psi + \overline{\nabla\cdot\kappa\nabla T}.\label{rhoh0 equation}
 \end{eqnarray}
-In advect\_base, the forcing term is set to the term proportional to $(\rho h)_0$.\MarginPar{We are currently omitting the $\psi$ and thermal conduction terms.  I think this is a bug.}
+In advect\_base, the forcing term is set to the term proportional to 
+$(\rho h)_0$.\MarginPar{We are currently omitting the $\psi$ and thermal conduction 
+terms.  I think this is a bug.}
 \subsubsection{Computing $\rho h$ at edges}
-Simply add the results from Sections \ref{Predicting rhohprime at edges} and \ref{Predicting rhoh0 at edges}, using spatial interpolation to get $(\rho h)_0$ at non-radial faces.
+Simply add the results from Sections \ref{Predicting rhohprime at edges} and 
+\ref{Predicting rhoh0 at edges}, using spatial interpolation to get $(\rho h)_0$ 
+at non-radial faces.
 \subsection{Predicting $h$ (enthalpy\_pred\_type = 2)}

### 2008-07-29 docs/pert_notes/paper.tex
more updates of rho section

@@
 \section{Density Evolution}
-Full density evolution equation:
-\begin{eqnarray}
-\frac{\partial\rho}{\partial t} &=& -\nabla\cdot(\rho\Ub) \nonumber \\
-&=& -\Ub\cdot\nabla\rho -\rho\nabla\cdot\Ub. \label{rho equation}
-\end{eqnarray}
+The full density evolution equation is
+\begin{equation}
+\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub) = -\Ub\cdot\nabla\rho 
+- \rho\nabla\cdot\Ub. \label{rho equation}
+\end{equation}
 \subsection{Predicting $\rho'$ (predict\_rho = F - default)}
-Base state density evolution equation:
-\begin{eqnarray}
-\frac{\partial\rho_0}{\partial t} &=& -\nabla\cdot(\rho_0 w_0 \eb_r) - 
-\nabla\cdot(\eta_\rho \eb_r) \nonumber \\
-&=& -w_0\frac{\partial\rho_0}{\partial r} \underbrace{- 
-\rho_0\frac{\partial w_0}{\partial r}} 
-- \nabla\cdot(\eta_\rho \eb_r). \label{rho0 equation}
-\end{eqnarray}
-Note that $\Ub = \Ubt + w_o\eb_r$.  Subtract (\ref{rho0 equation}) from 
-(\ref{rho equation}) to give perturbational density equation:
-\begin{eqnarray}
-\frac{\partial\rho'}{\partial t} &=& -\nabla\cdot(\rho'\Ub) - \nabla\cdot(\rho_0\Ubt) 
-+ \nabla\cdot(\eta_\rho \eb_r) \nonumber \\
-&=& -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub - \nabla\cdot(\rho_0\Ubt)} 
-+ \nabla\cdot(\eta_\rho \eb_r).\label{rho' equation}
-\end{eqnarray}
+The base state density evolution equation is
+\begin{equation}
+\frac{\partial\rho_0}{\partial t} = -\nabla\cdot(\rho_0 w_0 \eb_r) = 
+-w_0\frac{\partial\rho_0}{\partial r} - \rho_0\frac{\partial w_0}{\partial r}.
+\label{rho0 equation}
+\end{equation}
+Subtract (\ref{rho0 equation}) from (\ref{rho equation}) and rearrange, noting that 
+$\Ub = \Ubt + w_o\eb_r$, to obtain the perturbational density equation,
+\begin{equation}
+\frac{\partial\rho'}{\partial t} = -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub 
+- \nabla\cdot(\rho_0\Ubt)}_{"\rho' ~ \text{force}"}. 
+\label{rhoprime equation}
+\end{equation}
 \subsubsection{Predicting $\rho'$ at edges}
 We predict $\rho'$ to edges using make\_edge\_scal in scalar\_advance.  The underbraced 
-term in (\ref{rho' equation}) is computed in modify\_scal\_force.  The $\eta_\rho$ term 
-is not included, but this is correct because the procedure to compute $\rho$ at edges 
-described in Section \ref{Computing rho at edges} makes use of an edge-centered, 
-time-centered $\rho_0^{n+\half}$ that does not include an $\eta_\rho$ contribution either.
+forcing term in (\ref{rhoprime equation}) is computed in modify\_scal\_force.
 \subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
-We first compute $\rho_0^{(2)}$ using advect\_base (but not correct\_base).  We compute 
-$\rho$ on edges by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at 
-edges. For the non-radial faces, we can directly add $\rho_0^{n+\half}$ since it is a 
-cell-centered quantity.  For the radial faces, we interpolate to obtain $\rho_0^{n+\half}$ 
-at radial faces.
+We first compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base.  The $(2)$ in the 
+superscript indicates that we have not called correct\_base yet.  We compute $\rho$ on edges 
+by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at edges.  We use 
+$\rho^{(2)}$ rather than $\rho^{n+1}$ to be consistent with the fact that we omitted the 
+$\eta_\rho$ contribution to $\rho'$ at edges.  For the non-radial faces, we can directly add 
+$\rho_0^{n+\half}$ to $\rho'$ since it is a cell-centered quantity.  For the radial faces, we 
+interpolate to obtain $\rho_0^{n+\half}$ at edges.
 \subsection{Predicting $\rho$ (predict\_rho = T)}
@@
 We predict $\rho$ to edges using make\_edge\_scal in scalar\_advance.  The forcing term 
-is set to zero, and therefore we set is\_conservative = T to alter the discretization 
-in make\_edge\_scal.
+is set to zero, but then we set is\_conservative = T to use the conservative 
+discretization in make\_edge\_scal.
 \subsection{Predicting $\rho_0$}
@@
 with a Godunov method in make\_edge\_state\_1d.  We call make\_edge\_state\_1d using the 
-underbraced term in (\ref{rho0 equation}) as the forcing.  The $\eta_\rho$ term is not 
+term proportional to $\rho_0$ in  (\ref{rho0 equation}) as the forcing.  The $\eta_\rho$ 
+term is not 
 included.\MarginPar{I think there is a bug in the code and the $\eta_\rho$ term needs to 
@@
 + \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext} \nonumber \\
-&=& -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub -\nabla\cdot(\rho h \Ub) 
-+ \frac{Dp_0}{Dt} + \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} 
-+ \rho H_{\rm ext}.\label{rhoh equation}
+&=& \underbrace{-\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub}_{-\nabla\cdot(\rho h\Ub)} 
++ \underbrace{\psi + \Ubt\cdot\nabla p_0}_{\frac{Dp_0}{Dt}} + \nabla\cdot\kappa\nabla T 
++ \rho H_{\rm nuc} + \rho H_{\rm ext}\label{rhoh equation}
 \end{eqnarray}

### 2008-07-29 docs/pert_notes/paper.tex
more work on rho and rhoh sections

@@
 We first compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base.  The $(2)$ in the 
-superscript indicates that we have not called correct\_base yet.  We compute $\rho$ on edges 
-by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at edges.  We use 
-$\rho^{(2)}$ rather than $\rho^{n+1}$ to be consistent with the fact that we omitted the 
-$\eta_\rho$ contribution to $\rho'$ at edges.  For the non-radial faces, we can directly add 
-$\rho_0^{n+\half}$ to $\rho'$ since it is a cell-centered quantity.  For the radial faces, we 
-interpolate to obtain $\rho_0^{n+\half}$ at edges.
+superscript indicates that we have not called correct\_base yet (correct\_base takes 
+$\rho_0^{(2)}$ as an input and computes $\rho^{n+1}$ as the output).  We compute $\rho$ on 
+edges by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at edges.  We 
+use $\rho^{(2)}$ rather than $\rho^{n+1}$ to be consistent with the fact that we did not 
+include an $\eta_\rho$ contribution to compute $\rho'$ at edges.  For the non-radial edges, 
+we can directly add $\rho_0^{n+\half}$ to $\rho'$ since $\rho_0^{n+\half}$ is a cell-centered 
+quantity.  For the radial edges, we interpolate to obtain $\rho_0^{n+\half}$ at radial edges 
+before adding it to $\rho'$.
 \subsection{Predicting $\rho$ (predict\_rho = T)}
@@
 discretization in make\_edge\_scal.
-\subsection{Predicting $\rho_0$}
-To compute $\eta_\rho$, we need $\rho_0$ at edge-centered, time-centered locations.  
-However, we do NOT use the same edge-centered, time-centered $\rho_0^{n+\half}$ that was 
-referred to in Section \ref{Computing rho at edges}.  Instead, we use a $\rho_0$ computed 
-with a Godunov method in make\_edge\_state\_1d.  We call make\_edge\_state\_1d using the 
-term proportional to $\rho_0$ in  (\ref{rho0 equation}) as the forcing.  The $\eta_\rho$ 
-term is not 
-included.\MarginPar{I think there is a bug in the code and the $\eta_\rho$ term needs to 
-be included here.}  This gives us the so-called $\rho_0^{n+\half,{\rm pred}}$.
+\subsection{Predicting $\rho_0$ (predict\_rho = T or F)}
+To advect $\rho_0$ in advect\_base and to compute $\eta_\rho$ that will be used in 
+correct\_base, we need $\rho_0$ at edge-centered, time-centered locations.  However, we do 
+not use the same edge-centered, time-centered $\rho_0^{n+\half}$ referred to in Section 
+\ref{Computing rho at edges}.  Instead, we use a $\rho_0$ computed with a Godunov method in 
+make\_edge\_state\_1d.  We call make\_edge\_state\_1d using the term proportional to 
+$\rho_0$ in (\ref{rho0 equation}) as the forcing.  This gives us $\rho_0^{n+\half,{\rm pred}}$.
+\MarginPar{You could argue that including a time-lagged $\eta_\rho$ source term would
+increase the order of accuracy, but Ann has convinced me that this term isn't necessary.}
 \section{Energy Evolution}
-Full enthalpy equation:
+The full enthalpy equation is
 \begin{eqnarray}
@@
 &=& \underbrace{-\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub}_{-\nabla\cdot(\rho h\Ub)} 
-+ \underbrace{\psi + \Ubt\cdot\nabla p_0}_{\frac{Dp_0}{Dt}} + \nabla\cdot\kappa\nabla T 
-+ \rho H_{\rm nuc} + \rho H_{\rm ext}\label{rhoh equation}
++ \underbrace{\psi + \wt\frac{\partial p_0}{\partial r}}_{\frac{Dp_0}{Dt}} 
++ \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}. \label{rhoh equation}
 \end{eqnarray}
@@
 + \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi
-+ (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 
++ \wt\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 
 + \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
@@
 \ref{Predicting rhoh0 at edges}, using spatial interpolation to get $(\rho h)_0$ 
-at non-radial faces.
+at non-radial edges.
 \subsection{Predicting $h$ (enthalpy\_pred\_type = 2)}

### 2008-07-30 docs/pert_notes/paper.tex
Here's my (current) best guess about how to handle predicting (rho h)prime to edges.  It mirrors the treatment for rhoprime.  But there is a difference, since we compute (rho h)_0 by averaging (rho h)^{(1)}, we do not need to deal with any etarhoh stuff to correct rhoh.

@@
 \subsubsection{Predicting $\rho'$ at edges}
-We predict $\rho'$ to edges using make\_edge\_scal in scalar\_advance.  The underbraced 
-forcing term in (\ref{rhoprime equation}) is computed in modify\_scal\_force.
+We define $\rho'^n = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to edges using 
+make\_edge\_scal in scalar\_advance and the underbraced term in (\ref{rhoprime equation}) 
+as the forcing.
 \subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
-We first compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base.  The $(2)$ in the 
+We first compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base, which evolves 
+(\ref{rho0 equation}) through $\Delta t$ in time.  The $(2)$ in the 
 superscript indicates that we have not called correct\_base yet (correct\_base takes 
@@
 + \underbrace{\psi + \wt\frac{\partial p_0}{\partial r}}_{\frac{Dp_0}{Dt}} 
-+ \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}. \label{rhoh equation}
++ \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}.
 \end{eqnarray}
-Full temperature equation:
+The full temperature equation is:
 \begin{equation}
@@
 \subsection{Predicting $(\rho h)'$ (enthalpy\_pred\_type = 1 - default)}
-These are working notes for what I'd like to do for the energy evolution.
-\subsubsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
+These are working notes for how we are going to handle the energy evolution.  Due to the 
+operator splitting, the call to react\_state has already been made.  Hence, the goal is to 
+compute an edge state energy starting from $(\rho h)^{(1)}$ using an energy evolution 
+equation that does not include terms already accounted for in react\_state
+\begin{equation}
+\frac{\partial(\rho h)}{\partial t} = -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub 
++ \psi + \wt\frac{\partial p_0}{\partial r} + \nabla\cdot\kappa\nabla T. \label{rhoh equation}
+\end{equation}
 We can arbitrarily define a base state enthalpy evolution equation:
@@
 \frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] 
-+ \frac{D_0p_0}{Dt} + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} 
-+ \overline{\rho H_{\rm ext}} \nonumber \\
++ \frac{D_0p_0}{Dt} \nonumber \\
 &=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} 
-+ \psi + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} 
-+ \overline{\rho H_{\rm ext}}.\label{rhoh0 equation}
++ \psi.\label{rhoh0 equation}
 \end{eqnarray}
-We will predict both $(\rho h)_0$ and $(\rho h)'$ to edges using a Godunov method.  Then 
-we will add the two answers together to get $\rho h$ at edges.  This is different from 
-the density, in which we averaged an ``old'' and ``new'' base state density to get 
-$\rho_0$ at edges.  By computing $\rho h$ at edges in this way, we avoid having to 
-compute the complete evolution of $(\rho h)_0$, and therefore avoid having to write a 
-correct\_base and diffuse\_base function for $\rho h$.\\
-
-To simplify matters, the we have already called react\_state, so for now we will ignore 
-the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms in the remaining derivations.  
-Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and ignoring the 
-aforementioned terms gives the perturbational enthalpy equation:
+Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) gives the perturbational 
+enthalpy equation:
 \begin{eqnarray}
@@
 - \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r} 
-+ (\nabla\cdot\kappa\nabla T - \overline{\nabla\cdot\kappa\nabla T}) \nonumber \\
++ \nabla\cdot\kappa\nabla T\nonumber \\
 &=& -\Ub\cdot\nabla(\rho h)' \underbrace{- (\rho h)'\nabla\cdot\Ub 
-- \nabla\cdot[(\rho h)_0\Ubt]} + \wt\frac{\partial p_0}{\partial r} 
-+ (\nabla\cdot\kappa\nabla T 
-- \overline{\nabla\cdot\kappa\nabla T})\label{rhohprime equation}
+- \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r}
++ \nabla\cdot\kappa\nabla T}_{"(\rho h)' ~ \text{force}"} \label{rhohprime equation}
 \end{eqnarray}
-We use make\_edge\_scal to get $(\rho h)'$ at edges.  The function modify\_scal\_force 
-computes the underbraced term in (\ref{rhohprime equation}).  The function mkrhohforce 
-computes the $\wt$ term.\MarginPar{I need to write a function to handle the thermal 
-diffusion terms.}
-\subsubsection{Predicting $(\rho h)_0$ at edges}\label{Predicting rhoh0 at edges}
-After the call to react\_state, we set $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$.  
-We use the Godunov algorithm in make\_edge\_state\_1d to predict $(\rho h)_0$ at edges, 
-omitting the terms added in react\_state, so we are essentially evolving an abbreviated 
-version of (\ref{rhoh0 equation}):
-\begin{eqnarray}
-\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] + \psi 
-+ \overline{\nabla\cdot\kappa\nabla T} \nonumber \\
-&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} 
-+ \psi + \overline{\nabla\cdot\kappa\nabla T}.\label{rhoh0 equation}
-\end{eqnarray}
-In advect\_base, the forcing term is set to the term proportional to 
-$(\rho h)_0$.\MarginPar{We are currently omitting the $\psi$ and thermal conduction 
-terms.  I think this is a bug.}
+\subsubsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
+We define $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  and 
+$(\rho h)'^{(1)} = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'^{(1)}$ to 
+edges using make\_edge\_scal in scalar\_advance and the underbraced term in 
+(\ref{rhohprime equation}) as the 
+forcing.
 \subsubsection{Computing $\rho h$ at edges}
-Simply add the results from Sections \ref{Predicting rhohprime at edges} and 
-\ref{Predicting rhoh0 at edges}, using spatial interpolation to get $(\rho h)_0$ 
-at non-radial edges.
+We first compute $(\rho h)_0^{(2)}$ from $(\rho h)_0^{(1)}$ using advect\_base, which evolves 
+(\ref{rhoh equation}) through $\Delta t$ in time.  Then we follow an analogous procedure 
+used in Section \ref{Computing rho at edges} to obtain $\rho h$ at edges.
 \subsection{Predicting $h$ (enthalpy\_pred\_type = 2)}

### 2008-07-30 docs/pert_notes/paper.tex
more details

@@
 \subsubsection{Predicting $\rho'$ at edges}
-We define $\rho'^n = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to edges using 
+We define $\rho' = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to edges using 
 make\_edge\_scal in scalar\_advance and the underbraced term in (\ref{rhoprime equation}) 
 as the forcing.
-\subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
-We first compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base, which evolves 
-(\ref{rho0 equation}) through $\Delta t$ in time.  The $(2)$ in the 
-superscript indicates that we have not called correct\_base yet (correct\_base takes 
-$\rho_0^{(2)}$ as an input and computes $\rho^{n+1}$ as the output).  We compute $\rho$ on 
-edges by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at edges.  We 
-use $\rho^{(2)}$ rather than $\rho^{n+1}$ to be consistent with the fact that we did not 
-include an $\eta_\rho$ contribution to compute $\rho'$ at edges.  For the non-radial edges, 
-we can directly add $\rho_0^{n+\half}$ to $\rho'$ since $\rho_0^{n+\half}$ is a cell-centered 
-quantity.  For the radial edges, we interpolate to obtain $\rho_0^{n+\half}$ at radial edges 
-before adding it to $\rho'$.
-\subsection{Predicting $\rho$ (predict\_rho = T)}
-\subsubsection{Predicting $\rho$ at edges}
-We predict $\rho$ to edges using make\_edge\_scal in scalar\_advance.  The forcing term 
-is set to zero, but then we set is\_conservative = T to use the conservative 
-discretization in make\_edge\_scal.
-\subsection{Predicting $\rho_0$ (predict\_rho = T or F)}
-To advect $\rho_0$ in advect\_base and to compute $\eta_\rho$ that will be used in 
-correct\_base, we need $\rho_0$ at edge-centered, time-centered locations.  However, we do 
-not use the same edge-centered, time-centered $\rho_0^{n+\half}$ referred to in Section 
-\ref{Computing rho at edges}.  Instead, we use a $\rho_0$ computed with a Godunov method in 
-make\_edge\_state\_1d.  We call make\_edge\_state\_1d using the term proportional to 
-$\rho_0$ in (\ref{rho0 equation}) as the forcing.  This gives us $\rho_0^{n+\half,{\rm pred}}$.
+\subsubsection{Predicting $\rho_0$ at edges}
+There are two ways to predict $\rho_0$ at edges.\\ \\
+We can call make\_edge\_state\_1d using the term proportional to $\rho_0$ in 
+(\ref{rho0 equation}) as the forcing.  This gives us $\rho_0^{n+\half,{\rm pred}}$.
 \MarginPar{You could argue that including a time-lagged $\eta_\rho$ source term would 
 increase the order of accuracy, but Ann has convinced me that this term isn't necessary.}  
+This term is used to advect $\rho_0$ in advect\_base and to compute $\eta_\rho$ that will be 
+used in correct\_base.\\ \\
+We can also define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n + \rho_0^{(2)})/2$.  We first 
+compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base, which evolves (\ref{rho0 equation}) 
+through $\Delta t$ in time.  The $(2)$ in the superscript indicates that we have not called 
+correct\_base yet (correct\_base takes $\rho_0^{(2)}$ as an input and computes $\rho^{n+1}$ 
+as the output).  This term is used to construct $\rho$ at edges from $\rho'$ at edges, and 
+this $\rho$ at edges is used to compute fluxes for $\rho X_k$.  We use $\rho^{(2)}$ rather 
+than $\rho^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be consistent with the fact 
+that we do not include an $\eta_\rho$ contribution to compute $\rho'$ at edges.
+\subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
+For the non-radial edges, we can directly add $\rho_0^{n+\half,{\rm avg}}$ to $\rho'$ since 
+$\rho_0^{n+\half,{\rm avg}}$ is a cell-centered quantity.  For the radial edges, we 
+interpolate to obtain $\rho_0^{n+\half,{\rm avg}}$ at radial edges before adding it to $\rho'$.
+\subsubsection{Advancing $\rho X_k$}
+The evolution equation for $\rho X_k$, ignoring the reaction terms that were already 
+accounted for in react\_state, and the associated discretization is
+\begin{equation}
+\frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
+-\nabla\cdot\left\{\left[\left(\rho_0^{n+\half,{\rm avg}}+\rho'\right)X_k\right](\Ubt+w_0\eb_r)\right\}.
+\end{equation}
 \section{Energy Evolution}
@@
 - \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r}
-+ \nabla\cdot\kappa\nabla T}_{"(\rho h)' ~ \text{force}"} \label{rhohprime equation}
++ \nabla\cdot\kappa\nabla T}_{"(\rho h)' ~ \text{force}"}. \label{rhohprime equation}
 \end{eqnarray}
@@
 We define $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  and 
-$(\rho h)'^{(1)} = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'^{(1)}$ to 
+$(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'$ to 
 edges using make\_edge\_scal in scalar\_advance and the underbraced term in 
-(\ref{rhohprime equation}) as the 
-forcing.
+(\ref{rhohprime equation}) as the forcing.\\ \\
+{\bf What the code is currently doing:} Instead of defining 
+$(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries through the complete 
+evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ react\_base 
+$\rightarrow (\rho h)_0^{(1)} \rightarrow$ advect\_base 
+$\rightarrow (\rho h)_0^{(2)} \rightarrow$ react\_base $\rightarrow (\rho h)_0^{n+1}$.  
+There is no correct\_base function for $(\rho h)_0$ written yet, nor is any adjustment made 
+to sync up $(\rho h)_0$ after the thermal diffusion setp.
+\subsubsection{Predicting $(\rho h)_0$ at edges}
 \subsubsection{Computing $\rho h$ at edges}
@@
 used in Section \ref{Computing rho at edges} to obtain $\rho h$ at edges.
-\subsection{Predicting $h$ (enthalpy\_pred\_type = 2)}
-\subsection{Predicting $T$ into $(\rho h)'$ (enthalpy\_pred\_type = 3)}
-\subsection{Predicting $T$ into $h$ (enthalpy\_pred\_type = 4)}
+\subsubsection{Advancing $\rho h$}
+We would like to update the non-forcing part of the enthalpy equation using
+\begin{equation}
+\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) = -\nabla\cdot\left\{\left[(\rho h)_0^{n+\half,{\rm avg}}+(\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\}.
+\end{equation}
+{\bf What the code is currently doing:}
+\begin{equation}
+\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot\left[(\rho h)'\left(\Ubt + w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\right].
+\end{equation}
 \end{document}

### 2008-07-30 docs/pert_notes/paper.tex
this doc is in good shape now

@@
 \frac{\partial\rho_0}{\partial t} = -\nabla\cdot(\rho_0 w_0 \eb_r) = 
--w_0\frac{\partial\rho_0}{\partial r} - \rho_0\frac{\partial w_0}{\partial r}.
+-w_0\frac{\partial\rho_0}{\partial r} 
+\underbrace{-\rho_0\frac{\partial w_0}{\partial r}}_{"\rho_0 ~ \text{force}"}.
 \label{rho0 equation}
 \end{equation}
-Subtract (\ref{rho0 equation}) from (\ref{rho equation}) and rearrange, noting that 
+Subtract (\ref{rho0 equation}) from (\ref{rho equation}) and rearrange terms, noting that 
 $\Ub = \Ubt + w_o\eb_r$, to obtain the perturbational density equation,
@@
 as the forcing.
-\subsubsection{Predicting $\rho_0$ at edges}
-There are two ways to predict $\rho_0$ at edges.\\ \\
-We can call make\_edge\_state\_1d using the term proportional to $\rho_0$ in 
+\subsubsection{Predicting $\rho_0$ at edges}\label{Predicting rho0 at edges}
+There are two ways to predict $\rho_0$ at edges.
+\begin{enumerate}
+\item We call make\_edge\_state\_1d using the underbraced term in 
 (\ref{rho0 equation}) as the forcing.  This gives us $\rho_0^{n+\half,{\rm pred}}$.
@@
 increase the order of accuracy, but Ann has convinced me that this term isn't necessary.}  
-This term is used to advect $\rho_0$ in advect\_base and to compute $\eta_\rho$ that will be 
-used in correct\_base.\\ \\
-We can also define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n + \rho_0^{(2)})/2$.  We first 
-compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base, which evolves (\ref{rho0 equation}) 
-through $\Delta t$ in time.  The $(2)$ in the superscript indicates that we have not called 
-correct\_base yet (correct\_base takes $\rho_0^{(2)}$ as an input and computes $\rho^{n+1}$ 
-as the output).  This term is used to construct $\rho$ at edges from $\rho'$ at edges, and 
-this $\rho$ at edges is used to compute fluxes for $\rho X_k$.  We use $\rho^{(2)}$ rather 
-than $\rho^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be consistent with the fact 
-that we do not include an $\eta_\rho$ contribution to compute $\rho'$ at edges.
+This term is used to advect $\rho_0$ in advect\_base and to compute $\eta_\rho$, which will 
+be used in correct\_base.
+\item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n + \rho_0^{(2)})/2$.  We 
+compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base, which advances equation 
+(\ref{rho0 equation}) through $\Delta t$ in time.  The $(2)$ in the superscript indicates 
+that we have not called correct\_base yet, which computes  $\rho^{n+1}$ from $\rho_0^{(2)}$. 
+We use $\rho^{(2)}$ rather than $\rho^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be 
+consistent with the fact that we do not include an $\eta_\rho$ contribution to compute 
+$\rho'$ at edges.  $\rho_0^{n+\half,{\rm avg}}$ is used to construct $\rho$ at edges from 
+$\rho'$ at edges, and this $\rho$ at edges is used to compute fluxes for $\rho X_k$.  
+\end{enumerate}
 \subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
-For the non-radial edges, we can directly add $\rho_0^{n+\half,{\rm avg}}$ to $\rho'$ since 
+For the non-radial edges, we directly add $\rho_0^{n+\half,{\rm avg}}$ to $\rho'$ since 
 $\rho_0^{n+\half,{\rm avg}}$ is a cell-centered quantity.  For the radial edges, we 
 interpolate to obtain $\rho_0^{n+\half,{\rm avg}}$ at radial edges before adding it to $\rho'$.
-\subsubsection{Advancing $\rho X_k$}
+\subsubsection{Advancing $\rho X_k$}\label{Advancing rhoX_k}
 The evolution equation for $\rho X_k$, ignoring the reaction terms that were already 
@@
 \frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
--\nabla\cdot\left\{\left[\left(\rho_0^{n+\half,{\rm avg}}+\rho'\right)X_k\right](\Ubt+w_0\eb_r)\right\}.
+-\nabla\cdot\left\{\left[\left(\rho_0^{n+\half,{\rm avg}} 
++ \rho'\right)X_k\right](\Ubt+w_0\eb_r)\right\}.
 \end{equation}
@@
 \end{eqnarray}
-The full temperature equation is:
+The full temperature equation is
 \begin{equation}
@@
 \subsection{Predicting $(\rho h)'$ (enthalpy\_pred\_type = 1 - default)}
-These are working notes for how we are going to handle the energy evolution.  Due to the 
-operator splitting, the call to react\_state has already been made.  Hence, the goal is to 
-compute an edge state energy starting from $(\rho h)^{(1)}$ using an energy evolution 
-equation that does not include terms already accounted for in react\_state
+Due to operator splitting, the call to react\_state has already been made.  Hence, the goal 
+is to compute an edge state enthalpy starting from $(\rho h)^{(1)}$ using an enthalpy
+equation that does not include the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms, where were 
+already accounted for in react\_state
 \begin{equation}
@@
 \end{equation}
-We can arbitrarily define a base state enthalpy evolution equation:
+We arbitrarily define a base state enthalpy evolution equation
+\MarginPar{Does $\psi$ need to be included in the force?}
 \begin{eqnarray}
@@
 + \frac{D_0p_0}{Dt} \nonumber \\
-&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} 
+&=& -w_0\frac{\partial(\rho h)_0}{\partial r} 
+- \underbrace{(\rho h)_0\frac{\partial w_0}{\partial r}}_{"(\rho h)_0 ~ \text{force}"}
 + \psi.\label{rhoh0 equation}
 \end{eqnarray}
-Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) gives the perturbational 
-enthalpy equation:
+Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and rearranging terms gives 
+the perturbational enthalpy equation
 \begin{eqnarray}
@@
 (\ref{rhohprime equation}) as the forcing.\\ \\
-{\bf What the code is currently doing:} Instead of defining 
-$(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries through the complete 
-evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ react\_base 
+{\bf What the code is currently doing:}\MarginPar{Do we need to change this?}
+ Instead of defining $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries 
+through the complete evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ react\_base 
 $\rightarrow (\rho h)_0^{(1)} \rightarrow$ advect\_base 
@@
 There is no correct\_base function for $(\rho h)_0$ written yet, nor is any adjustment made 
-to sync up $(\rho h)_0$ after the thermal diffusion setp.
+to sync up $(\rho h)_0$ after the thermal diffusion step.
 \subsubsection{Predicting $(\rho h)_0$ at edges}
+We use an analogous procedure described in Section \ref{Predicting rho0 at edges} to 
+obtain $(\rho h)_0^{n+\half,\rm{pred}}$ and $(\rho h)_0^{n+\half,\rm{avg}}$.  Note that 
+$(\rho h)_0^{n+\half,{\rm avg}} = [(\rho h)_0^{(1)} + (\rho h)_0^{(2)}]/2$.
 \subsubsection{Computing $\rho h$ at edges}
-We first compute $(\rho h)_0^{(2)}$ from $(\rho h)_0^{(1)}$ using advect\_base, which evolves 
-(\ref{rhoh equation}) through $\Delta t$ in time.  Then we follow an analogous procedure 
-used in Section \ref{Computing rho at edges} to obtain $\rho h$ at edges.
+We use an analogous procedure described in Section \ref{Computing rho at edges} to compute
+$\rho h$ at edges.
 \subsubsection{Advancing $\rho h$}
-We would like to update the non-forcing part of the enthalpy equation using
+We would like to update the non-forcing part of the enthalpy equation analogously to
+Section \ref{Advancing rhoX_k}
 \begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) = -\nabla\cdot\left\{\left[(\rho h)_0^{n+\half,{\rm avg}}+(\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\}.
+\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) = 
+-\nabla\cdot\left\{\left[(\rho h)_0^{n+\half,{\rm avg}} 
++ (\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\}.
 \end{equation}
-{\bf What the code is currently doing:}
+{\bf What the code is currently doing:}\MarginPar{Do we need to change this?}
 \begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot\left[(\rho h)'\left(\Ubt + w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\right].
+\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot\left[(\rho h)'\left(\Ubt 
++ w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] 
+- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\right].
 \end{equation}

### 2008-07-30 docs/pert_notes/paper.tex
caught a few typos

@@
 \MarginPar{You could argue that including a time-lagged $\eta_\rho$ source term would 
-increase the order of accuracy, but Ann has convinced me that this term isn't necessary.}  
+decrease the error, but Ann has convinced me that this term isn't necessary.}  
 This term is used to advect $\rho_0$ in advect\_base and to compute $\eta_\rho$, which will 
@@
 \item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n + \rho_0^{(2)})/2$.  We 
-compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base, which advances equation 
+compute $\rho_0^{(2)}$ from $\rho_0^n$ using advect\_base, which advances equation 
 (\ref{rho0 equation}) through $\Delta t$ in time.  The $(2)$ in the superscript indicates 
-that we have not called correct\_base yet, which computes  $\rho^{n+1}$ from $\rho_0^{(2)}$. 
-We use $\rho^{(2)}$ rather than $\rho^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be 
+that we have not called correct\_base yet, which computes  $\rho_0^{n+1}$ from $\rho_0^{(2)}$. 
+We use $\rho_0^{(2)}$ rather than $\rho_0^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be 
 consistent with the fact that we do not include an $\eta_\rho$ contribution to compute 

### 2008-08-05 docs/pert_notes/paper.tex
typo
messing with the marginpars

@@
 (\ref{rho0 equation}) as the forcing.  This gives us $\rho_0^{n+\half,{\rm pred}}$.
-\MarginPar{You could argue that including a time-lagged $\eta_\rho$ source term would 
-decrease the error, but Ann has convinced me that this term isn't necessary.}  
-This term is used to advect $\rho_0$ in advect\_base and to compute $\eta_\rho$, which will 
-be used in correct\_base.
+This term is used to advect $\rho_0$ in advect\_base and therefore must also be used 
+to compute $\eta_\rho$, which will be used in correct\_base.
 \item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n + \rho_0^{(2)})/2$.  We 
@@
 We arbitrarily define a base state enthalpy evolution equation
-\MarginPar{Does $\psi$ need to be included in the force?}
+\MarginPar{Does $\psi$ need to be included in the force?
+I have written it in the base state equation for now, even though neither the base state or perturbational edge state prediction in the code uses $\psi$.
+How about evolving density first,  then computing a time-centered $\psi$ to be used as the energy forcing?}
 \begin{eqnarray}
@@
 \subsubsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
-We define $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  and 
-$(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'$ to 
-edges using make\_edge\_scal in scalar\_advance and the underbraced term in 
-(\ref{rhohprime equation}) as the forcing.\\ \\
 {\bf What the code is currently doing:}\MarginPar{Do we need to change this?}
@@
 There is no correct\_base function for $(\rho h)_0$ written yet, nor is any adjustment made 
-to sync up $(\rho h)_0$ after the thermal diffusion step.
+to sync up $(\rho h)_0$ after the thermal diffusion step.\\ \\
+{\bf What I'd like to do:}  Define $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  and 
+$(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'$ to 
+edges using make\_edge\_scal in scalar\_advance and the underbraced term in 
+(\ref{rhohprime equation}) as the forcing.
 \subsubsection{Predicting $(\rho h)_0$ at edges}
@@
 \subsubsection{Advancing $\rho h$}
-We would like to update the non-forcing part of the enthalpy equation analogously to
-Section \ref{Advancing rhoX_k}
+{\bf What the code is currently doing:}\MarginPar{Do we need to change this?}
+\begin{equation}
+\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot\left[(\rho h)'\left(\Ubt 
++ w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] 
+- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\eb_r\right].
+\end{equation}
+{\bf What we would do to be consistent with density:} We would like to update the 
+non-forcing part of the enthalpy equation analogously to Section \ref{Advancing rhoX_k}
 \begin{equation}
@@
 \end{equation}
-{\bf What the code is currently doing:}\MarginPar{Do we need to change this?}
-\begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot\left[(\rho h)'\left(\Ubt 
-+ w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] 
-- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\right].
-\end{equation}
+
 \end{document}

### 2008-08-05 docs/Godunov_notes/paper.tex
Added note about how the input flag "use_new_godunov=T" enables the four options I proposed in the Questions section.
typo fixes

@@
 \item Concerning (\ref{Force Term}), do we really need to include the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction?
-\item Do we want to do characteristic tracing to compute $\Ub^{\trans}$, analgous to how {\tt VARDEN} uses characteristic tracing to compute $\Ub^{\mac}$?
-\item When we compute $\Ub^{\mac,*}$, do we want to replace all instances of ``trans'' velocities with ``MAC'' velocities?
+\item Do we want to do characteristic tracing to compute $\Ubt^{\trans}$, analgous to how {\tt VARDEN} uses characteristic tracing to compute $\Ub^{\mac}$?
+\item When we compute $\Ubt^{\edge}$, do we want to replace all instances of ``trans'' velocities with ``MAC'' velocities?
 Doing this is simply a matter of replacing the call to {\tt make\_edge\_state} with {\tt make\_edge\_scal}.
+\item NOTE: The {\tt probin} parameter ``use\_new\_godunov'' enables all four of these features.
 \end{enumerate}

### 2008-08-06 docs/pert_notes/paper.tex
removed marginpars

@@
 \subsubsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
-{\bf What the code is currently doing:}\MarginPar{Do we need to change this?}
+{\bf What the code did before 8/6/08:}
  Instead of defining $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries 
@@
 to sync up $(\rho h)_0$ after the thermal diffusion step.\\ \\
-{\bf What I'd like to do:}  Define $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  and 
-$(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'$ to 
+{\bf What the code currently does:}  Define $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  
+and $(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'$ to 
 edges using make\_edge\_scal in scalar\_advance and the underbraced term in 
@@
 \subsubsection{Advancing $\rho h$}
-{\bf What the code is currently doing:}\MarginPar{Do we need to change this?}
+{\bf What the did before 8/6/08:}
 \begin{equation}
@@
 \end{equation}
-{\bf What we would do to be consistent with density:} We would like to update the 
+{\bf What the code does now:} We would like to update the 
 non-forcing part of the enthalpy equation analogously to Section \ref{Advancing rhoX_k}

### 2008-08-14 docs/pert_notes/paper.tex
correct the final enthalpy update part

@@
 \end{equation}
-\subsection{Predicting $\rho'$ (predict\_rho = F - default)}
+\subsection{Predicting $\rho'$ ({\tt predict\_rho = F} - default)}
 The base state density evolution equation is
@@
 We define $\rho' = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to edges using 
-make\_edge\_scal in scalar\_advance and the underbraced term in (\ref{rhoprime equation}) 
+{\tt make\_edge\_scal} in {\tt scalar\_advance} and the underbraced term in (\ref{rhoprime equation}) 
 as the forcing.
@@
 \begin{enumerate}
-\item We call make\_edge\_state\_1d using the underbraced term in 
+\item We call {\tt make\_edge\_state\_1d} using the underbraced term in 
 (\ref{rho0 equation}) as the forcing.  This gives us $\rho_0^{n+\half,{\rm pred}}$.
-This term is used to advect $\rho_0$ in advect\_base and therefore must also be used 
-to compute $\eta_\rho$, which will be used in correct\_base.
+This term is used to advect $\rho_0$ in {\tt advect\_base} and therefore must also be used 
+to compute $\eta_\rho$, which will be used in {\tt correct\_base}.
 \item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n + \rho_0^{(2)})/2$.  We 
-compute $\rho_0^{(2)}$ from $\rho_0^n$ using advect\_base, which advances equation 
+compute $\rho_0^{(2)}$ from $\rho_0^n$ using {\tt advect\_base}, which advances equation 
 (\ref{rho0 equation}) through $\Delta t$ in time.  The $(2)$ in the superscript indicates 
-that we have not called correct\_base yet, which computes  $\rho_0^{n+1}$ from $\rho_0^{(2)}$. 
+that we have not called {\tt correct\_base} yet, which computes  $\rho_0^{n+1}$ from $\rho_0^{(2)}$. 
 We use $\rho_0^{(2)}$ rather than $\rho_0^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be 
@@
 The evolution equation for $\rho X_k$, ignoring the reaction terms that were already 
-accounted for in react\_state, and the associated discretization is
+accounted for in {\tt react\_state}, and the associated discretization is
 \begin{equation}
@@
 \end{equation}
-\subsection{Predicting $(\rho h)'$ (enthalpy\_pred\_type = 1 - default)}
-Due to operator splitting, the call to react\_state has already been made.  Hence, the goal 
+\subsection{Predicting $(\rho h)'$ ({\tt enthalpy\_pred\_type = 1} - default)}
+Due to operator splitting, the call to {\tt react\_state} has already been made.  Hence, the goal 
 is to compute an edge state enthalpy starting from $(\rho h)^{(1)}$ using an enthalpy
 equation that does not include the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms, where were 
-already accounted for in react\_state
+already accounted for in {\tt react\_state}
 \begin{equation}
@@
  Instead of defining $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries 
-through the complete evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ react\_base 
-$\rightarrow (\rho h)_0^{(1)} \rightarrow$ advect\_base 
-$\rightarrow (\rho h)_0^{(2)} \rightarrow$ react\_base $\rightarrow (\rho h)_0^{n+1}$.  
-There is no correct\_base function for $(\rho h)_0$ written yet, nor is any adjustment made 
+through the complete evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ {\tt react\_base}
+$\rightarrow (\rho h)_0^{(1)} \rightarrow$ {\tt advect\_base}
+$\rightarrow (\rho h)_0^{(2)} \rightarrow$ {\tt react\_base} $\rightarrow (\rho h)_0^{n+1}$.  
+There is no {\tt correct\_base} function for $(\rho h)_0$ written yet, nor is any adjustment made 
 to sync up $(\rho h)_0$ after the thermal diffusion step.\\ \\
@@
 and $(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'$ to 
-edges using make\_edge\_scal in scalar\_advance and the underbraced term in 
+edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} and the underbraced term in 
 (\ref{rhohprime equation}) as the forcing.
@@
 + w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] 
-- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\eb_r\right].
+\underbrace{- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\eb_r\right] + \psi}_{\text{this term used to come in through} ~ (\rho h)_0^{n+1} - (\rho h)_0^n} + \wt\frac{\partial p_0}{\partial r} \enskip .
 \end{equation}
@@
 -\nabla\cdot\left\{\left[(\rho h)_0^{n+\half,{\rm avg}} 
-+ (\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\}.
++ (\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\} + \wt\frac{\partial p_0}{\partial r} + \psi  \enskip .
 \end{equation}

### 2008-08-14 docs/initial_models/paper.tex
minor fixes

@@
 \begin{itemize}
-\item The initial model should be but into HSE according to $(p_{i+1} - p_i)/\Delta r = half (\rho_{i+1} - \rho_i) g_{i+\half}$
+\item The initial model should be but into HSE according to $(p_{i+1} - p_i)/\Delta r = \half (\rho_{i+1} - \rho_i) g_{i+\half}$
 
@@
 \item Spherical runs use $p$, $(\rho X)$, and $\rho$ from the initial model.  These are then
-mapped onto the Cartesian grid, and then $(rho h)$ is computed via the EOS.
+mapped onto the Cartesian grid, and then $(\rho h)$ is computed via the EOS.
 \end{itemize}

### 2008-08-14 docs/initial_models/paper.tex
updates

@@
 
+\clearpage
+
 \section{Creating the initial model}
 \begin{itemize}
-\item The initial model should be but into HSE according to $(p_{i+1} - p_i)/\Delta r = \half (\rho_{i+1} - \rho_i) g_{i+\half}$
+\item The initial model should be but into HSE according to 
+$(p_{i+1} - p_i)/\Delta r = \half (\rho_{i+1} - \rho_i) g_{i+\half} \enskip .$
 
-\item The initial model should be created at the same resolution as the base state.  This means that
-when you change resolution of your simulation, use a new initial model.
+\item The initial model should be created at the same resolution as
+the base state.  This means that when you change resolution of your
+simulation, use a new initial model.
 
-\item The initial model should be thermodynamically consistent with the EOS that {\tt MAESTRO}
-uses.
+\item The initial model should be thermodynamically consistent with
+the EOS that {\tt MAESTRO} uses.
 
@@
 \begin{itemize}
-\item Spherical runs use $p$, $(\rho X)$, and $\rho$ from the initial model.  These are then
-mapped onto the Cartesian grid, and then $(\rho h)$ is computed via the EOS.
+\item Spherical runs use $p$, $(\rho X)$, and $\rho$ from the initial
+model.  These are then mapped onto the Cartesian grid, and then
+$(\rho h)$ is computed via the EOS.
+
+\item Spherical runs automatically set $\Delta x = \Delta r$.
+\end{itemize}
+
+\section{Running notes}
+\begin{itemize}
+\item When the code is run, it prints out the HSE error of the base
+state model---this should be checked to insure that we a
 \end{itemize}

### 2008-08-14 docs/pert_notes/paper.tex
merge in the flow_chart.tex tables to help make sense of all of this mess
make the tables reflect what is actually in the code

@@
 \def\Ubt   {\widetilde{\bf U}}
+\def\ut    {\widetilde{\bf U}}
 \def\wt    {\widetilde{w}}
+\def\er    {{\bf e}_r}
+\def\kth   {k_\mathrm{th}}
 
@@
 \frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt} 
-+ \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext} \nonumber \\
++ \nabla\cdot \kth \nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext} \nonumber \\
 &=& \underbrace{-\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub}_{-\nabla\cdot(\rho h\Ub)} 
-+ \underbrace{\psi + \wt\frac{\partial p_0}{\partial r}}_{\frac{Dp_0}{Dt}} 
-+ \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}.
++ \underbrace{\psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r}}_{\frac{Dp_0}{Dt}} 
++ \nabla\cdot\kth\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}.
 \end{eqnarray}
@@
 + \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi
-+ \wt\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 
++ (\ut \cdot \er )\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 
 + \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
@@
 \frac{\partial(\rho h)}{\partial t} = -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub 
-+ \psi + \wt\frac{\partial p_0}{\partial r} + \nabla\cdot\kappa\nabla T. \label{rhoh equation}
++ \psi + (\ut \dot \er) \frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T. \label{rhoh equation}
 \end{equation}
@@
 \frac{\partial(\rho h)'}{\partial t} &=& -\nabla\cdot[(\rho h)'\Ub] 
-- \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r} 
-+ \nabla\cdot\kappa\nabla T\nonumber \\
+- \nabla\cdot[(\rho h)_0\Ubt] + (\ut \cdot \er)\frac{\partial p_0}{\partial r} 
++ \nabla\cdot\kth\nabla T\nonumber \\
 &=& -\Ub\cdot\nabla(\rho h)' \underbrace{- (\rho h)'\nabla\cdot\Ub 
-- \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r}
-+ \nabla\cdot\kappa\nabla T}_{"(\rho h)' ~ \text{force}"}. \label{rhohprime equation}
+- \nabla\cdot[(\rho h)_0\Ubt] + (\ut \cdot \er)\frac{\partial p_0}{\partial r}
++ \nabla\cdot\kth\nabla T}_{"(\rho h)' ~ \text{force}"}. \label{rhohprime equation}
 \end{eqnarray}
@@
 edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} and the underbraced term in 
-(\ref{rhohprime equation}) as the forcing.
+(\ref{rhohprime equation}) as the forcing. \MarginPar{{\tt full\_rhoh0\_evolution} comes into play here}
 \subsubsection{Predicting $(\rho h)_0$ at edges}
@@
 + w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] 
-\underbrace{- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\eb_r\right] + \psi}_{\text{this term used to come in through} ~ (\rho h)_0^{n+1} - (\rho h)_0^n} + \wt\frac{\partial p_0}{\partial r} \enskip .
+\underbrace{- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\eb_r\right] + \psi}_{\text{this term used to come in through} ~ (\rho h)_0^{n+1} - (\rho h)_0^n} + (\ut \cdot \er)\frac{\partial p_0}{\partial r} \enskip .
 \end{equation}
@@
 -\nabla\cdot\left\{\left[(\rho h)_0^{n+\half,{\rm avg}} 
-+ (\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\} + \wt\frac{\partial p_0}{\partial r} + \psi  \enskip .
++ (\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\} + (\ut \cdot \er)\frac{\partial p_0}{\partial r} + \psi  \enskip .
 \end{equation}
 
+\subsection{Predicting $h$ ({\tt enthalpy\_pred\_type = 2})}
+
+\subsection{Predicting $T$ then $(\rho h)'$ ({\tt enthalpy\_pred\_type = 3})}
+
+\subsection{Predicting $T$ then $h$ ({\tt enthalpy\_pred\_type = 4})}
+
+\subsection{Summary of Forcing}
+
+\begin{table*}[h]
+\begin{center}
+\caption{Forcing term into {\tt make\_edge\_scal} \newline}
+\begin{tabular}{c|c}
+\hline
+\hline
+{H PRED TYPE} &   {Force} \\
+\hline \\[-3mm]
+1  $((\rho h)^\prime)$ &  $(\ut \cdot \er) \frac{\partial p_0}{\partial r} - 
+ (\rho h)^\prime \; \nabla \cdot (\ut+w_0 \er) - 
+ \nabla \cdot (\ut (\rho h)_0) + \nabla \cdot \kth \nabla T$ \\[2 mm]
+2  $(h)$ & $\frac{1}{\rho} \left [\psi + (\ut \cdot \er)
+  \frac{\partial p_0}{\partial r} + \nabla \cdot \kth \nabla T \right ]$ \\[2 mm]
+3  $(T)$ & $\frac{1}{\rho c_p} \left \{ (1 - \rho h_p) 
+   \left [\psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r} \right ] \nabla \cdot \kth \nabla T \right \}$ \\[2 mm]
+4  $(T)$ & $\frac{1}{\rho c_p} \left\{ (1 - \rho h_p) \left [\psi + (\ut \cdot \er)
+\frac{\partial p_0}{\partial r}\right ] \nabla \cdot \kth \nabla T \right\}$ \\[2 mm]
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
+\begin{table*}[h]
+\begin{center}
+\caption{Quantity that goes into and out of {\tt make\_edge\_scal} \newline}
+\begin{tabular}{c|c}
+\hline
+\hline
+{h pred type} & {h/T } \\
+\hline \\[-3mm]
+1 & $(\rho h)^\prime$ \\
+2 & $h$  \\
+3 & $T$  \\
+4 & $T$  \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
+\begin{table*}
+\begin{center}
+\caption{When predicting temp ... (conversion done by {\tt makeRhoHfromT}) \newline}
+\begin{tabular}{c|c|c|c}
+\hline
+\hline
+{type} & {Inputs } & {EOS inputs} & {output} \\
+\hline
+3 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ &  $(\rho h)^\prime = \rho h - (\rho h)_0$ \\
+   &                   & $   X = $ given & \\
+   &                   &                 & \\
+ 4 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ & $h$ \\
+   &                   & $   X = $ given & \\
+   &                   &                 & \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
+\begin{table*}
+\begin{center}
+\caption{Quantity that goes into {\tt mkflux} on edges \newline}
+\begin{tabular}{c|c|c}
+\hline
+\hline
+{h pred type} & {Species } & {h/T } \\
+\hline
+1 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
+2 & $X$ and $\rho^\prime$ & $h$ \\
+3 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
+4 & $X$ and $\rho^\prime$ & $h$ \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
+\begin{table*}
+\begin{center}
+\caption{Quantity that is created in mkflux and passed into update scal\newline\
+}
+\begin{tabular}{c|c|c|c|c}
+       &            &              &\multicolumn{2}{c}{Forcing Term} \\
+{type} & {Species } & {$(\rho h)$} & {$(\rho X)$} & {$(\rho h)$} \\
+\hline 
+1 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
+2 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
+3 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r}$ \\[2mm]
+4 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
 \end{document}

### 2008-08-14 docs/flowchart/paper.tex
the beginnings

\documentclass[11pt]{article}

\tolerance=600

\usepackage{amsmath,color}

% Margins
\usepackage[lmargin=0.5in,rmargin=1.5in,tmargin=1.0in,bmargin=1.0in]{geometry}

\newcommand{\rhozero}{\rho_0}
\newcommand{\pizero}{\pi_0}
\newcommand{\pizeroone}{\pi_0^{(1)}}
\newcommand{\pizerotwo}{\pi_0^{(2)}}
\newcommand{\gammabar}{\overline{\Gamma}_1}

\newcommand{\nablab}{\mathbf{\nabla}}
\newcommand{\cdotb}{\mathbf{\cdot}}

\newcommand{\sfrac}[2]{\mathchoice
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
  {#1\!/#2}}

\newcommand{\half}{\sfrac{1}{2}}
\newcommand{\myhalf}{\sfrac{1}{2}}
\newcommand{\nph}{{n + \myhalf}}
\newcommand{\nmh}{{n - \myhalf}}

\newcommand{\Hext}{{H_{\rm ext}}}
\newcommand{\Hnuc}{{H_{\rm nuc}}}

\newcommand{\Sbar}{\overline{S}}

\newcommand{\inp}{\mathrm{in}}
\newcommand{\initp}{\mathrm{init}}
\newcommand{\outp}{\mathrm{out}}

\newcommand{\uadvone}{\widetilde{\mathbf{U}}^{\mathrm{ADV},(1)}}
\newcommand{\uadvtwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV},(2)}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\V}{\mathbf{V}}
\newcommand{\uadvstar}{\widetilde{\mathbf{U}}^{\mathrm{ADV,\star}}}
\newcommand{\uadvstarone}{\widetilde{\mathbf{U}}^{\mathrm{ADV,(1),\star}}}
\newcommand{\uadvstartwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV,(2),\star}}}
\newcommand{\wadv}{w^{\mathrm{ADV}}}
\newcommand{\dt}{\Delta t}
\newcommand{\dr}{\Delta r}
\newcommand{\dx}{\Delta \x}
\newcommand{\dy}{\Delta r}
\newcommand{\etal}{\it et al.}

\newcommand{\etarho}{\eta_{\rho}}
\newcommand{\etaXk}{\eta_{\rho X_k}}
\newcommand{\etarhoh}{\eta_{\rho h}}

\newcommand{\ubold}{\mathbf{U}}
\newcommand{\ut}{\widetilde{\ubold}}
\newcommand{\wt}{\tilde{w}}

\newcommand{\hp}{\widetilde{h}}

\newcommand{\omegadot}{\dot{\omega}}
\newcommand{\er}{\mathbf{e}_r}
\newcommand{\nx}{\mathbf{n}_x}

\newcommand{\rhobar}{\overline{\rho}}

\newcommand{\x}{\mathbf{x}}
\newcommand{\rp}{r^\prime}
\newcommand{\rhop}{\rho^\prime}

\newcommand{\gcc}{\mathrm{g~cm^{-3} }}

\newcommand{\ext}{\mathrm{ext}}
\newcommand{\eff}{\mathrm{eff}}

\title{Flowchart}

\setlength{\marginparwidth}{1.0in}
\newcommand{\MarginPar}[1]{\marginpar{%
\vskip-\baselineskip %raise the marginpar a bit
\raggedright\tiny\sffamily
\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\begin{document}

\maketitle
\tableofcontents
\cleardoublepage

\section{Current Algorithm}
\subsection{Notation}

We make use of the following shorthand notations in outlining the algorithm.
\begin{enumerate}

\item For any quantity, $\phi,$ we define $\overline{\phi} = ${\bf Avg}$(\phi),$ 
the average over $\Omega_H,$ as
\begin{equation}
\overline{\phi}(r) = \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
\end{equation}

\item {\bf React State}$(\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp})\rightarrow$
$(\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp})$ \newline
is the process by which we evolve the species and enthalpy from $X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following system of equations 
over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip.
\end{eqnarray}
using the stiff ordinary differential equation
 integration methods provided by the {\tt VODE} package (cite VODE).
The absolute error tolerances are set to $10^{-12}$ for the species,
 and a relative tolerance of $10^{-5}$ is used for the temperature.  
The integration yields the new values of the mass fractions, $X_k^{\outp}$.  
Equation (ref{eq:tempreactC}) is derived from equation (ref{eq:tempenthalpy}) by assuming that the pressure is constant during the burn state.  
In evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  
In theory, this means evaluating the equation of state for each right-hand side evaluation that {\tt VODE} requires.  
In practice, we freeze $c_p$ and $\xi_k$ at the start of the integration time step and compute them using 
$\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs to the equation of state.  
We note that while temperature is evolved when solving these equations, we do not keep the final temperature, 
nor do we use it to compute the final change in enthalpy.  
Therefore, $T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged during 
the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

After the new mass fractions have been computed, the reaction rates are defined as:
\begin{equation}
(\rho\omegadot_k)^{\outp} = \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\Delta t /2} \enskip,
\end{equation}
and the nuclear energy generation rate is defined as
\begin{equation}
(\rho\Hnuc) = -\sum_k(\rho\omegadot_k)^{\outp} \; q_k \enskip.
\end{equation}
The enthalpy update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is updated by
\begin{equation}
(\rho h)^{\outp} = (\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc) + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
\end{equation}

\item {\bf Advect Base}$(\rho_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp} ) \rightarrow 
(\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ 
\newline
is the process by which we update the base state through $\dt$ in time given the
radial velocity $w_0^{\inp}.$ Here we discuss
the algorithm for plane-parallel geometries.  
The base state arrays are all one-dimensional 
in the radial coordinate; we think of the base state quantities as defined
at cell centers, with $w_0$ and $\beta_0$ defined halfway in-between.\\ \\
a)  For the density update, we discretize equation (ref{eq:rho0upd\_new}), neglecting the $\etarho$ term,
to compute the new base state density,
\begin{eqnarray}
\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
         \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ] \nonumber  \enskip,
\end{eqnarray}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the procedure described in
paper II, Appendix A, and are saved for use later in the overall algorithm.\\ \\
b)  For the pressure update, We discretize equation (ref{eq:p0upd\_new}), 
to compute the new base state pressure,
\begin{eqnarray}
p_{0,j}^\outp &=& p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+ w_{0,j-\myhalf}^\inp \right)
\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\inp \enskip,
\end{eqnarray}
where the interface states are again found via the procedure described in
paper II, Appendix A.
\\

\item {\bf Correct Base}$(\rho_0^{\inp} , \etarho^{\inp}) \rightarrow (\rho_0^{\outp})$
is the process by which we adjust the base state density given $\etarho$ using
\begin{eqnarray}
\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left( \eta_{\rho,j+\myhalf}^{\inp} - \eta_{\rho,j-\myhalf}^{\inp}  \right)  \enskip . 
\end{eqnarray} 
\end{enumerate}

\subsection{Time Advancement Algorithm}

We now describe the full time advancement algorithm, making frequent
use of the shorthand developed above.  Here, we assume that the
problem is already properly initialized.  We describe the details of
the initialization in \S~ref{Sec:Initialization}.

In paper II we discretized the density and enthalpy evolution equations in 
perturbational form, arguing that this would be less
susceptible to grid effects.  In the present algorithm, the analog would
be to use a perturbational form of the species equations.
Numerical testing has shown that the form of the algorithm presented
here is in fact more robust than the perturbational form 
when the base state mixing terms ($\etarho$) are included.  
In the algorithm described below, density is still 
extrapolated to edges in perturbational form, but 
we treat $X_k$ and $h$ in non-perturbational form.

%--------------------------------------------------------------------------
% STEP 1
%--------------------------------------------------------------------------
\noindent {\bf Step 1.} {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item At the beginning of each time step, we need an estimate for the time-centered
source term in the velocity divergence constraint (see eq. [ref{eq:constraint}]).
If this is the first step of the calculation ($n=0$), we set
\[
S^{\half,\star} = \frac{S^0 + S^1}{2} \enskip ,
\]
where $S^1$ is found through the iterative process that initializes the calculation.
Otherwise, following the method used in our small-scale low Mach number algorithm
(Cite Bell04), we extrapolate to the half-time using the source
terms at the previous and current time levels
\begin{equation}
S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
\end{equation}

\item Define
\[
\overline{S}^{\nph,\star} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip.
\]

\item Construct $w_0^{\nph,\star}$ by integrating equation (ref{eq:divw0}) using 
the lagged $\psi^{\nmh},$
\begin{equation}
\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S}^{\nph,\star} 
- \frac{1}{\gammabar^{n} p_0^{n}} \psi^{\nmh} \enskip . \nonumber
\end{equation}
%for plane-parallel geometries,  or the analogous equation (see eq. [ref{eq:B7}])
%for spherical configurations.  
For spherical geometries, this equation would be modified.
%In the case of spherical geometries use the current $p_0^n,$
%${\Gamma_1}^n_0 = \Gamma_1(p_0^n, \rho_0^n, X_k^n),$ and $g$ as computed from $\rho_0^n.$

\item Using equation (ref{eq:pizero}), define the scaled pressure gradient
\begin{equation}
-\left ( \frac{1}{\rho_0^n} \frac{\partial \pizeroone}{\partial r} \right ) = 
\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} 
+  w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star} \enskip ,
\end{equation} 
where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are
\begin{equation}
w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\end{equation}
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\end{equation}
If $n=0$, we use $\dt^{-1} = \dt^0$.

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 2
%--------------------------------------------------------------------------
\noindent {\bf Step 2.} {\em Construct the provisional edge-based advective velocity}, $\uadvone$:

The procedure to construct $\uadvone$ is described in detail in Appendix ref{app:advective\_vel}.
We note that  $\uadvone$ satisfies the discrete versions of equations
(ref{eq:udoterzero}) and (ref{eq:tildeconstraint}), 
in particular,
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
\enskip ,
\end{equation}

%--------------------------------------------------------------------------
% STEP 3
%--------------------------------------------------------------------------
\noindent {\bf Step 3.} {\em React the full state through the first time interval of $\dt / 2.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item {\bf React State}$(\rho^n, (\rho h)^n, X_k^n, T^n, (\rho^n \Hext))
                   \rightarrow$ $( \rho^{(1)}, (\rho h)^{(1)}, X_k^{(1)}, T^{(1)},
                                  (\rho \omegadot_k)^{(1)} ).$

\item
Define
\begin{eqnarray}
\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{(1)}, p_0^{n}, X_k^{(1)}) \right) 
                                                                \enskip , \nonumber \\
 {\beta   }_0^{(1)}    &=& \beta   \left(\rho_0^{n}, p_0^{n}, \gammabar^{(1)}\right) \enskip .
\end{eqnarray}

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 4
%--------------------------------------------------------------------------
\noindent {\bf Step 4.} {\em Advect the base state, then the full state, through a time interval of $\dt.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item {\bf Advect Base}$(\rhozero^{n}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph,\star}, \psi^{\nmh}) \rightarrow $
$(\rho_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}, p_0^{n+1,\star}).$

\item Compute the edge states, $(\rho X_k)^{(1),\nph,\star}$ and $(\rho h)^{(1),\nph,\star},$
      for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
      predict $\rho'$, $T$, and $X_k$ to the edges, using
      a second-order Taylor expansion in space and time, as described in
      paper II, Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  
      We explicitly include the reaction terms in the temperature prediction, 
      since we 
      did not update the temperature in {\bf React State}. 
      We use the equation of state and base state density, $\rho_0$,
      to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
      We define $\rho^{(1),\nph,\star} = \sum_k (\rho X_k)^{(1),\nph,\star}.$

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ and
$(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ without explicitly including the 
reaction terms,
\begin{eqnarray}
(\rho X_k)^{(2),\star} &=& (\rho X_k)^{(1)} 
 - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
  (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip , \\ \nonumber  \\
(\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
  (\rho h)^{(1),\nph,\star} \right) \right] \nonumber \\
&& + \dt \; \left(\uadvone \cdotb \er\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} 
+ \dt \; \psi^{\nmh} \enskip , \\
\rho^{(2),\star} &=& \sum_k (\rho X_k)^{(2),\star} \enskip , \\ 
X_k^{(2),\star} &=& (\rho X_k)^{(2),\star} / \rho^{(2),\star} \enskip , \\
T^{(2),\star} &=& T\left(\rho^{(2),\star}, (\rho h)^{(2),\star},  X_k^{(2),\star}\right) {\hbox{ \rm using the equation of state.}} 
\end{eqnarray}

%The construction of 
%$ \left[ \nabla \cdotb \left(  (\uadvone+w_0^{\nph,\star} \er)  (\rho X_k)^{(1)} \right) \right]^{\nph}$ \\
%and $ \left[ \nabla \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  (\rho h)^{(1)}\right)  \right]^{\nph}$
%also follows that described in paper II, Appendix A with $\V = \uadvone+w_0^{\nph,\star} \er.$ 
%We note that the time-centering of the $p_0$ terms in equations (ref{eq:step4_rhoh})
%matches as closely as possible the time-centering of the $\beta_0$ in 
%equation (ref{eq:uadvone_satisfies}).

\item {\bf Correct Base}$(\rho_0^{(2),\star}, \etarho^{\nmh}) \rightarrow (\rho_0^{n+1,\star})$.

\item Define an edge-centered $\etarho^{\nph,\star}$ and cell-centered $\psi_j^{\nph,\star}$:
\begin{eqnarray}
 \etarho^{\nph,\star}   &=&  {\rm {\bf Avg}} \left( (\uadvone \cdotb \er + w_0^{\nph,\star}) 
\rho^{(1),\nph,\star} \right) - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} 
 \nonumber \enskip , \\
 \psi_j^{\nph,\star}  &=&  \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph,\star} + \eta_{\rho,j+\myhalf}^{\nph,\star}\right) g \enskip .
\end{eqnarray}
For spherical geometries, $\psi_j^{\nph,\star}$ would have a different form. 
%be computed via equation (ref{eq:spherical_psi_num}).

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 5
%--------------------------------------------------------------------------
\noindent {\bf Step 5.} {\em React the full state through a second time interval of $\dt / 2.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item {\bf React State}$( \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, 
                             T^{(2),\star}, (\rho^{(2),\star} \Hext)) $\\
$\rightarrow ( \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, 
              (\rho \omegadot_k)^{(2),\star} ).$

\item Define
\begin{eqnarray}
 \gammabar^{n+1,\star}    &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{n+1,\star}, p_0^{n+1,\star}, 
                                                      X_k^{n+1,\star}) \right) \enskip , \nonumber \\
 {\beta   }_0^{n+1,\star}    &=& \beta   \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}\right) \enskip .
\end{eqnarray}

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 6
%--------------------------------------------------------------------------
\noindent {\bf Step 6.} {\em Define a new average expansion rate at time $t^\nph.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}
\item Define:
\begin{equation}
  S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + 
  \sigma \Hnuc^{(2),\star} +
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star}  
  + \sigma \Hext \enskip .
\end{equation} 
where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} / \rho^{(2),\star}$
and the remaining quantities are defined using $X_k^{n+1,\star},$ $\rho^{n+1,\star},$ 
and $T^{n+1,\star}$ from step 5.
Then define
\[
 S^\nph = \frac{S^n + S^{n+1,\star}}{2} \enskip. 
\]

\item Define:
\[
\overline{S}^{\nph} = {\mathrm{\bf Avg}} (S^{\nph}) \enskip.
\]

\item Construct $w_0^{\nph}$ by integrating equation (ref{eq:divw0}),
\begin{equation}
\frac{\partial w_0^{\nph}}{\partial r} =  \overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}} \psi^{\nph,\star} \enskip , \nonumber
\end{equation}
for plane-parallel geometries, where 
$\gammabar^{\nph,\star} = (\gammabar^{n} + \gammabar^{n+1,\star} ) / 2 $ and
$ p_0^{\nph,\star} = (p_0^{n} + p_0^{n+1,\star} ) / 2.$
%In spherical configurations we use the analogous equation (see eq. [ref{eq:B7}]). 
%In the case of spherical geometries use $p_0^{n+1,\star},$
%${\Gamma_1}^{n+1,\star}_0 = \Gamma_1(p_0^{n+1,\star}, \rho_0^{n+1,\star}, X_k^{n+1,\star}),$ and 
%$g$ as computed from $\rho_0^{n+1,\star}.$

\item Using equation (ref{eq:pizero}), define
\begin{equation}
-\left ( \frac{1}{\rho_0^n} \frac{\partial \pizerotwo}{\partial r} \right ) = 
\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
+ w_0^n \left(\frac{\partial w_0}{\partial r}\right)^n.
\end{equation}
where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined analogously to 
equations (ref{eq:w0nstar}) and (ref{eq:dw0drnstar}).

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 7
%--------------------------------------------------------------------------
\noindent {\bf Step 7.} {\em Construct the final edge-based advective velocity}, $\uadvtwo$:

The procedure to construct $\uadvtwo$ is described in detail in Appendix ref{app:advective\_vel}
and is analogous to the procedure used in step 2, but with updated values
for $w_0$ and $\pizero.$ We note that  $\uadvtwo$ satisfies the discrete versions of 
equations (ref{eq:udoterzero}) and (ref{eq:tildeconstraint}),
in particular
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) = \beta_0^{\nph,\star} 
\left(S^{\nph} - \overline{S}^{\nph}\right) \enskip .
\end{equation}
where $ \beta_0^{\nph,\star} = ( \beta_0^n +  \beta_0^{n+1,\star}) / 2.$

%--------------------------------------------------------------------------
% STEP 8
%--------------------------------------------------------------------------
\noindent {\bf Step 8.} {\em Advect the base state, then the full state, through a time interval of $\dt.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item {\bf Advect Base}$(\rho_0^{n}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph}, \psi^{\nph,\star} ) \rightarrow $
$(\rho_0^{(2)}, \rho_0^{\nph,{\rm pred}}, p_0^{n+1}).$

\item Compute the edge states, $(\rho X_k)^{(1),\nph}$ and $(\rho h)^{(1),\nph},$
      for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
      predict $\rho'$, $T$, and $X_k$ to the edges, using
      a second-order Taylor expansion in space and time, as described in
      paper II, Appendix A, using $\V = \uadvtwo+w_0^{\nph} \er$. 
      Again, we explicitly include the reaction terms in the temperature prediction, 
      since we did not update  
      the temperature in {\bf React State}.
      We use the equation of state and base state density, $\rho_0$,
      to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
      We define $\rho^{(1),\nph} = \sum_k (\rho X_k)^{(1),\nph}.$

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ and
$(\rho h)^{(1)} \rightarrow (\rho h)^{(2)},$ 
\begin{eqnarray}
(\rho X_k)^{(2)} &=& (\rho X_k)^{(1)} 
- \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
(\rho X_k)^{(1),\nph} \right) \right]  \enskip , \\ \nonumber \\
(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
(\rho h)^{(1),\nph} \right) \right] \nonumber \\
&& + \frac{\dt}{2} \; \left(\uadvtwo \cdotb \er\right)
\left [ \left(\frac{\partial p_0}{\partial r} \right)^{n}
      + \left(\frac{\partial p_0}{\partial r} \right)^{n+1}  \right ] 
+ \dt \; \psi^{\nph,\star} \enskip , \\ 
\rho^{(2)} &=& \sum_k (\rho X_k)^{(2)} \enskip , \\
X_k^{(2)} &=& (\rho X_k)^{(2)} / \rho^{(2)} \enskip , \\
T^{(2)} &=& T\left(\rho^{(2)}, (\rho h)^{(2)},  X_k^{(2)}\right) {\hbox{ \rm using the
equation of state.}} 
\end{eqnarray}

%The construction of 
%$ \left[ \nabla \cdotb \left(  (\uadvtwo+w_0^{\nph} \er)  (\rho X_k)^{(1)} \right) \right]^{\nph}$ \\
%and $ \left[ \nabla \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  (\rho h)^{(1)}\right)  \right]^{\nph}$
%is similar to that described in paper II, Appendix A with $\V = \uadvtwo+w_0^{\nph} \er.$ 
%For species advection we set $s = X_k,$ then multiply the edge states of $X_k$ by predicted
%edge states of $\rho.$  For the enthalpy update we predict temperature to the edges, convert
%$T$ to $h$ using the EOS, then multiply by $\rho.$
%We note that the time-centering of the $p_0$ terms in Equations (ref{eq:step8_rhoh})
%matches as closely as possible the time-centering of the $\beta_0$ in 
%Equation (ref{eq:uadvtwo_satisfies}).

\item {\bf Correct Base}$(\rho_0^{(2)}, \etarho^{\nph, \star}) \rightarrow (\rho_0^{n+1})$

\item Define an edge-centered $ \etarho^{\nph}$ and a cell-centered $\psi_j^{\nph}$,
\begin{eqnarray}
 \etarho^{\nph}   &=&  {\rm {\bf Avg}} \left( (\uadvtwo \cdotb \er + w_0^{\nph}) 
\rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} 
 \nonumber \enskip , \\
 \psi_j^{\nph}  &=&  \frac{1}{2} \left(\eta_{rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
\end{eqnarray}
%For spherical geometries, $\psi_j^{\nph}$ would be computed via equation (ref{eq:spherical_psi_num}).
\end{enumerate}

%--------------------------------------------------------------------------
% STEP 9
%--------------------------------------------------------------------------
\noindent {\bf Step 9.} {\em React the full state through a second time interval of $\dt / 2.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item {\bf React State}$(\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho^{(2)} \Hext) )
                   \rightarrow (\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, 
                               (\rho \omegadot_k)^{(2)} ).$

\item Define
\begin{eqnarray}
 \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left(\Gamma_1(\rho^{n+1}, p_0^{n+1}, {X_k}^{n+1}) \right) \enskip , \nonumber \\ 
 {\beta   }_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip .
\end{eqnarray}

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 10
%--------------------------------------------------------------------------
\noindent {\bf Step 10.} {\em Compute $S^{n+1}$ for the final projection.}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}
\item Define:
\begin{equation}
  S^{n+1} =  -\sigma  \sum_k  \xi_k (\omegadot_k)^{(2)}  + \sigma \Hnuc^{(2)} +
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2)}  
   + \sigma \Hext \enskip ,
\end{equation}
where $(\omegadot_k)^{(2)} = (\rho \omegadot_k)^{(2)} / \rho^{(2)}$
and the remaining quantities are defined using $X_k^{n+1},$ $\rho^{n+1},$
and $T^{n+1}$ from step 9.

\item Define
\[
\overline{S}^{n+1} = {\mathrm{\bf Avg}} (S^{n+1}) \enskip.
\]

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 11
%--------------------------------------------------------------------------
\noindent {\bf Step 11.} {\em Update the velocity}.

The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
We update the velocity field $\ut^n$ to $\ut^{n+1,\star}$ by discretizing
equation (ref{eq:utildeupd}), 
\begin{eqnarray}
\ut^{n+1,\star} &=& \ut^n - \dt \;
 \left[\left((\uadvtwo+ w_0^{\nph} \er) \cdotb \nablab\right) \ut \right]^\nph
      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
\end{eqnarray}
where $\rho^\nph = ( \rho^n + \rho^{n+1} )/2$
and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
The construction of $\left[((\uadvtwo+ w_0^{\nph} \er) \cdotb \nablab) \ut \right]^\nph$
is described in paper II, Appendix A, with $\V = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
component of $\ut^n$ individually.

Finally, we impose the divergence constraint from equation (ref{eq:tildeconstraint}),
\[
\nablab \cdotb \left(\beta_0^{\nph} \ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)\enskip,
\]
by defining ${\bf V} = \ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ and solving
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
\end{equation}
for nodal values of $\phi$, where
$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
$L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve the linear system of equation (ref{eq:nodal\_solve}) using multigrid V-cycles with Gauss-Seidel relaxation.

We determine the new time velocity field from
\begin{equation}
\ut^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
\end{equation}
and the new time-centered perturbational pressure from:
\[
  \pi^\nph = \frac{1}{\dt} \phi \enskip .
\]

%--------------------------------------------------------------------------
% STEP 12
%--------------------------------------------------------------------------
\noindent {\bf Step 12.} {\em Compute a new $\dt.$}

Compute $\dt$ for the next time step with the procedure described in \S~ref{sec:timesteps},
using $w_0$ as computed in step 6 and
$\ut^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.

\noindent This completes the time advancement of the algorithm.

\subsection{Initialization}

We start each calculation with user-specified initial values for
$\rho$, $X_k$ and $T,$ as well as an initial background state.  In
order for the low Mach number assumption to hold, the initial data
must be thermodynamically consistent with the initial background
state.  In addition, the initial velocity field must satisfy an
initial approximation to the divergence constraint.  We use an iterative
procedure to compute both an initial right-hand-side for the
constraint equation and an initial velocity field that satisfies
the constraint.  The user specifies the number of iterations,
$N_{\rm iters}^{S},$ in this first step of the initialization procedure.

The initial perturbational pressure also needs to be determined for
use in steps 2, 7 and 11. 
This is done through a second iterative procedure which follows the
time advancement algorithm as described in steps 1-11 in \S~ref{Sec:Alg}.  
The user specifies the number of iterations, 
$N_{\rm iters}^{\pi},$ in this second step of the initialization procedure.
The details for both iterations are given below.

%--------------------------------------------------------------------------
% STEP 0
%--------------------------------------------------------------------------
\noindent {\bf Step 0} {\em Initialization}

Start with initial data $X_k^{\initp}, \rho^{\initp},$ $T^{\initp},$ an 
initial base state, and an initial guess for the velocity, $\ubold^{\initp}.$
Use the equation of state to determine $(\rho h)^{\initp}$.  Set
$w_0^1 = 0$ as an initial approximation.  Compute $\beta_0^{\initp}$ as a function of 
the initial data.  Then, project $\ubold^{\initp}$ using $\beta_0^{\initp}$ and 
$S = \rho\Hext$, giving $\ubold^{0,1}$.  The next part of the initialization process 
proceeds as follows.

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}
%\renewcommand{\theenumii}{{\bf \roman{enumii}}}
\renewcommand{\labelenumii}{\roman{enumii}.}

\item {\bf Do} {$\nu = 1,...,N_{\rm iters}^{S}$.}
  \begin{enumerate}

  \item Estimate $\Delta t^\nu$ using $\ubold^{0,\nu}$ and $w_0^\nu.$

  \item {\bf React State}$( \rho^{\initp},(\rho h)^{\initp}, X_k^{\initp}, T^{\initp}, (\rho^{\initp} \Hext) )$\\
      $\rightarrow (\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{0,\nu} ).$

  \item Compute $S^{0,\nu}$ from equation (ref{eq:defineS}) 
        using $(\rho \omegadot_k)^{0,\nu}$ and the initial data.

  \item Compute $\overline{S}^{0,\nu} = {\mathrm{\bf Avg}} (S^{0,\nu}).$

  \item Compute $w_0^{\nu+1}$ as in step 1c using $\overline{S}^{0,\nu}$ and $\psi^{\nmh} = 0$.

  \item Project $\ubold^{0,\nu}$ using $\beta_0^{\initp}$ and 
        $(S^{0,\nu} - \overline{S}^{0,\nu})$ as the source term.  
        This yields $\ubold^{0,\nu+1}.$

  \end{enumerate}

  {\bf End do.}

  Define $S^0 = S^{0,N_{\rm iters}^S}$, $w_0^{-\myhalf} = w_0^{N_{\rm iters}^S+1}$, 
$\dt^0 = \Delta t^{N_{\rm iters}^S},$ and $\ubold^0 = \ubold^{0,N_{\rm iters}^S+1}.$

\end{enumerate}

Next, we need to construct an approximation to the time-centered perturbational pressure,
$\pi,$  and an approximation to the divergence constraint at the end of the
first time step.  As initial approximations, set $S^{1,0} = S^0$,
$\etarho^{-\half} = 0$, $\psi^{-\half} = 0$, and $\pi^{-\half} = 0.$
\begin{enumerate}
  \renewcommand{\theenumi}{{\bf \alph{enumi}}}
  \renewcommand{\labelenumii}{\roman{enumii}.}
  \addtocounter{enumi}{1}

\item {\bf Do} {$\nu = 0,...,N_{\rm iters}^{\pi}-1$.}

  \begin{enumerate}
  \item Perform steps 1--11 as described above, using 
    $S^{\half,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
    The only other difference in the time advancement is that in step 11
    we define ${\bf V} = (\ut^{1,\star} - \ut^0)$ and solve
    \begin{equation}  L_\beta^\rho \phi =
      D \left ( \beta_0^{\half} {\bf V} \right) - \beta_0^{\half} \left( (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right) \enskip . 
    \end{equation}
    (The motivation for this form of the projection in the initial pressure iterations
    is discussed in (cite almgren:bell:crutchfield).)
      We discard the new velocity resulting from this, but keep the new  
      value for $\pi^{\half} = \pi^{-\half} + (1 / \dt) \; \phi.$  
      These steps also yield new scalar data at time $\dt,$ which
      we discard,  and new values for $\etarho^{\half}$ (step 8), $\psi^{\half}$ (step 8), 
      $S^{1,\nu+1}$ (step 10), and $\pi^{\half}$ (step 11), which we keep.
    \item Set $\pi^{-\half} = \pi^{\half}$, $\etarho^{-\half} = \etarho^{\half}$,
      and $\psi^{-\half} = \psi^{\half}$. 
    \end{enumerate}

    {\bf End do.}

    Finally, we define $S^1 = S^{1,N_{\rm iters}^\pi}.$

  \end{enumerate}

\cleardoublepage

\section{Paper III Algorithm}
\subsection{Notation}

We make use of the following shorthand notations in outlining the algorithm.
\begin{enumerate}

\item For any quantity, $\phi,$ we define $\overline{\phi} = ${\bf Avg}$(\phi),$ 
the average over $\Omega_H,$ as
\begin{equation}
\overline{\phi}(r) = \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
\end{equation}

\item {\bf React State}$(\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp})\rightarrow$
$(\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp})$ \newline
is the process by which we evolve the species and enthalpy from $X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following system of equations 
over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip.
\end{eqnarray}
using the stiff ordinary differential equation
 integration methods provided by the {\tt VODE} package (cite VODE).
The absolute error tolerances are set to $10^{-12}$ for the species,
 and a relative tolerance of $10^{-5}$ is used for the temperature.  
The integration yields the new values of the mass fractions, $X_k^{\outp}$.  
Equation (ref{eq:tempreactC}) is derived from equation (ref{eq:tempenthalpy}) by assuming that the pressure is constant during the burn state.  
In evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  
In theory, this means evaluating the equation of state for each right-hand side evaluation that {\tt VODE} requires.  
In practice, we freeze $c_p$ and $\xi_k$ at the start of the integration time step and compute them using 
$\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs to the equation of state.  
We note that while temperature is evolved when solving these equations, we do not keep the final temperature, 
nor do we use it to compute the final change in enthalpy.  
Therefore, $T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged during 
the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

After the new mass fractions have been computed, the reaction rates are defined as:
\begin{equation}
(\rho\omegadot_k)^{\outp} = \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\Delta t /2} \enskip,
\end{equation}
and the nuclear energy generation rate is defined as
\begin{equation}
(\rho\Hnuc) = -\sum_k(\rho\omegadot_k)^{\outp} \; q_k \enskip.
\end{equation}
The enthalpy update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is updated by
\begin{equation}
(\rho h)^{\outp} = (\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc) + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
\end{equation}

\item {\bf Advect Base}$(\rho_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp} ) \rightarrow 
(\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ 
\newline
is the process by which we update the base state through $\dt$ in time given the
radial velocity $w_0^{\inp}.$ Here we discuss
the algorithm for plane-parallel geometries.  
The base state arrays are all one-dimensional 
in the radial coordinate; we think of the base state quantities as defined
at cell centers, with $w_0$ and $\beta_0$ defined halfway in-between.\\ \\
a)  For the density update, we discretize equation (ref{eq:rho0upd\_new}), neglecting the $\etarho$ term,
to compute the new base state density,
\begin{eqnarray}
\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
         \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ] \nonumber  \enskip,
\end{eqnarray}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the procedure described in
paper II, Appendix A, and are saved for use later in the overall algorithm.\\ \\
b)  For the pressure update, We discretize equation (ref{eq:p0upd\_new}), 
to compute the new base state pressure,
\begin{eqnarray}
p_{0,j}^\outp &=& p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+ w_{0,j-\myhalf}^\inp \right)
\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\inp \enskip,
\end{eqnarray}
where the interface states are again found via the procedure described in
paper II, Appendix A.
\\

\item {\bf Correct Base}$(\rho_0^{\inp} , \etarho^{\inp}) \rightarrow (\rho_0^{\outp})$
is the process by which we adjust the base state density given $\etarho$ using
\begin{eqnarray}
\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left( \eta_{\rho,j+\myhalf}^{\inp} - \eta_{\rho,j-\myhalf}^{\inp}  \right)  \enskip . 
\end{eqnarray} 
\end{enumerate}

\subsection{Time Advancement Algorithm}

We now describe the full time advancement algorithm, making frequent
use of the shorthand developed above.  Here, we assume that the
problem is already properly initialized.  We describe the details of
the initialization in \S~ref{Sec:Initialization}.

In paper II we discretized the density and enthalpy evolution equations in 
perturbational form, arguing that this would be less
susceptible to grid effects.  In the present algorithm, the analog would
be to use a perturbational form of the species equations.
Numerical testing has shown that the form of the algorithm presented
here is in fact more robust than the perturbational form 
when the base state mixing terms ($\etarho$) are included.  
In the algorithm described below, density is still 
extrapolated to edges in perturbational form, but 
we treat $X_k$ and $h$ in non-perturbational form.

%--------------------------------------------------------------------------
% STEP 1
%--------------------------------------------------------------------------
\noindent {\bf Step 1.} {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item At the beginning of each time step, we need an estimate for the time-centered
source term in the velocity divergence constraint (see eq. [ref{eq:constraint}]).
If this is the first step of the calculation ($n=0$), we set
\[
S^{\half,\star} = \frac{S^0 + S^1}{2} \enskip ,
\]
where $S^1$ is found through the iterative process that initializes the calculation.
Otherwise, following the method used in our small-scale low Mach number algorithm
(Cite Bell04), we extrapolate to the half-time using the source
terms at the previous and current time levels
\begin{equation}
S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
\end{equation}

\item Define
\[
\overline{S}^{\nph,\star} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip.
\]

\item Construct $w_0^{\nph,\star}$ by integrating equation (ref{eq:divw0}) using 
the lagged $\psi^{\nmh},$
\begin{equation}
\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S}^{\nph,\star} 
- \frac{1}{\gammabar^{n} p_0^{n}} \psi^{\nmh} \enskip . \nonumber
\end{equation}
%for plane-parallel geometries,  or the analogous equation (see eq. [ref{eq:B7}])
%for spherical configurations.  
For spherical geometries, this equation would be modified.
%In the case of spherical geometries use the current $p_0^n,$
%${\Gamma_1}^n_0 = \Gamma_1(p_0^n, \rho_0^n, X_k^n),$ and $g$ as computed from $\rho_0^n.$

\item Using equation (ref{eq:pizero}), define the scaled pressure gradient
\begin{equation}
-\left ( \frac{1}{\rho_0^n} \frac{\partial \pizeroone}{\partial r} \right ) = 
\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} 
+  w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star} \enskip ,
\end{equation} 
where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are
\begin{equation}
w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\end{equation}
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\end{equation}
If $n=0$, we use $\dt^{-1} = \dt^0$.

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 2
%--------------------------------------------------------------------------
\noindent {\bf Step 2.} {\em Construct the provisional edge-based advective velocity}, $\uadvone$:

The procedure to construct $\uadvone$ is described in detail in Appendix ref{app:advective\_vel}.
We note that  $\uadvone$ satisfies the discrete versions of equations
(ref{eq:udoterzero}) and (ref{eq:tildeconstraint}), 
in particular,
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
\enskip ,
\end{equation}

%--------------------------------------------------------------------------
% STEP 3
%--------------------------------------------------------------------------
\noindent {\bf Step 3.} {\em React the full state through the first time interval of $\dt / 2.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item {\bf React State}$(\rho^n, (\rho h)^n, X_k^n, T^n, (\rho^n \Hext))
                   \rightarrow$ $( \rho^{(1)}, (\rho h)^{(1)}, X_k^{(1)}, T^{(1)},
                                  (\rho \omegadot_k)^{(1)} ).$

\item
Define
\begin{eqnarray}
\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{(1)}, p_0^{n}, X_k^{(1)}) \right) 
                                                                \enskip , \nonumber \\
 {\beta   }_0^{(1)}    &=& \beta   \left(\rho_0^{n}, p_0^{n}, \gammabar^{(1)}\right) \enskip .
\end{eqnarray}

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 4
%--------------------------------------------------------------------------
\noindent {\bf Step 4.} {\em Advect the base state, then the full state, through a time interval of $\dt.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item {\bf Advect Base}$(\rhozero^{n}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph,\star}, \psi^{\nmh}) \rightarrow $
$(\rho_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}, p_0^{n+1,\star}).$

\item Compute the edge states, $(\rho X_k)^{(1),\nph,\star}$ and $(\rho h)^{(1),\nph,\star},$
      for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
      predict $\rho'$, $T$, and $X_k$ to the edges, using
      a second-order Taylor expansion in space and time, as described in
      paper II, Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  
      We explicitly include the reaction terms in the temperature prediction, 
      since we 
      did not update the temperature in {\bf React State}. 
      We use the equation of state and base state density, $\rho_0$,
      to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
      We define $\rho^{(1),\nph,\star} = \sum_k (\rho X_k)^{(1),\nph,\star}.$

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ and
$(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ without explicitly including the 
reaction terms,
\begin{eqnarray}
(\rho X_k)^{(2),\star} &=& (\rho X_k)^{(1)} 
 - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
  (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip , \\ \nonumber  \\
(\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
  (\rho h)^{(1),\nph,\star} \right) \right] \nonumber \\
&& + \dt \; \left(\uadvone \cdotb \er\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} 
+ \dt \; \psi^{\nmh} \enskip , \\
\rho^{(2),\star} &=& \sum_k (\rho X_k)^{(2),\star} \enskip , \\ 
X_k^{(2),\star} &=& (\rho X_k)^{(2),\star} / \rho^{(2),\star} \enskip , \\
T^{(2),\star} &=& T\left(\rho^{(2),\star}, (\rho h)^{(2),\star},  X_k^{(2),\star}\right) {\hbox{ \rm using the equation of state.}} 
\end{eqnarray}

%The construction of 
%$ \left[ \nabla \cdotb \left(  (\uadvone+w_0^{\nph,\star} \er)  (\rho X_k)^{(1)} \right) \right]^{\nph}$ \\
%and $ \left[ \nabla \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  (\rho h)^{(1)}\right)  \right]^{\nph}$
%also follows that described in paper II, Appendix A with $\V = \uadvone+w_0^{\nph,\star} \er.$ 
%We note that the time-centering of the $p_0$ terms in equations (ref{eq:step4_rhoh})
%matches as closely as possible the time-centering of the $\beta_0$ in 
%equation (ref{eq:uadvone_satisfies}).

\item {\bf Correct Base}$(\rho_0^{(2),\star}, \etarho^{\nmh}) \rightarrow (\rho_0^{n+1,\star})$.

\item Define an edge-centered $\etarho^{\nph,\star}$ and cell-centered $\psi_j^{\nph,\star}$:
\begin{eqnarray}
 \etarho^{\nph,\star}   &=&  {\rm {\bf Avg}} \left( (\uadvone \cdotb \er + w_0^{\nph,\star}) 
\rho^{(1),\nph,\star} \right) - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} 
 \nonumber \enskip , \\
 \psi_j^{\nph,\star}  &=&  \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph,\star} + \eta_{\rho,j+\myhalf}^{\nph,\star}\right) g \enskip .
\end{eqnarray}
For spherical geometries, $\psi_j^{\nph,\star}$ would have a different form. 
%be computed via equation (ref{eq:spherical_psi_num}).

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 5
%--------------------------------------------------------------------------
\noindent {\bf Step 5.} {\em React the full state through a second time interval of $\dt / 2.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item {\bf React State}$( \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, 
                             T^{(2),\star}, (\rho^{(2),\star} \Hext)) $\\
$\rightarrow ( \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, 
              (\rho \omegadot_k)^{(2),\star} ).$

\item Define
\begin{eqnarray}
 \gammabar^{n+1,\star}    &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{n+1,\star}, p_0^{n+1,\star}, 
                                                      X_k^{n+1,\star}) \right) \enskip , \nonumber \\
 {\beta   }_0^{n+1,\star}    &=& \beta   \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}\right) \enskip .
\end{eqnarray}

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 6
%--------------------------------------------------------------------------
\noindent {\bf Step 6.} {\em Define a new average expansion rate at time $t^\nph.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}
\item Define:
\begin{equation}
  S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + 
  \sigma \Hnuc^{(2),\star} +
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star}  
  + \sigma \Hext \enskip .
\end{equation} 
where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} / \rho^{(2),\star}$
and the remaining quantities are defined using $X_k^{n+1,\star},$ $\rho^{n+1,\star},$ 
and $T^{n+1,\star}$ from step 5.
Then define
\[
 S^\nph = \frac{S^n + S^{n+1,\star}}{2} \enskip. 
\]

\item Define:
\[
\overline{S}^{\nph} = {\mathrm{\bf Avg}} (S^{\nph}) \enskip.
\]

\item Construct $w_0^{\nph}$ by integrating equation (ref{eq:divw0}),
\begin{equation}
\frac{\partial w_0^{\nph}}{\partial r} =  \overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}} \psi^{\nph,\star} \enskip , \nonumber
\end{equation}
for plane-parallel geometries, where 
$\gammabar^{\nph,\star} = (\gammabar^{n} + \gammabar^{n+1,\star} ) / 2 $ and
$ p_0^{\nph,\star} = (p_0^{n} + p_0^{n+1,\star} ) / 2.$
%In spherical configurations we use the analogous equation (see eq. [ref{eq:B7}]). 
%In the case of spherical geometries use $p_0^{n+1,\star},$
%${\Gamma_1}^{n+1,\star}_0 = \Gamma_1(p_0^{n+1,\star}, \rho_0^{n+1,\star}, X_k^{n+1,\star}),$ and 
%$g$ as computed from $\rho_0^{n+1,\star}.$

\item Using equation (ref{eq:pizero}), define
\begin{equation}
-\left ( \frac{1}{\rho_0^n} \frac{\partial \pizerotwo}{\partial r} \right ) = 
\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
+ w_0^n \left(\frac{\partial w_0}{\partial r}\right)^n.
\end{equation}
where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined analogously to 
equations (ref{eq:w0nstar}) and (ref{eq:dw0drnstar}).

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 7
%--------------------------------------------------------------------------
\noindent {\bf Step 7.} {\em Construct the final edge-based advective velocity}, $\uadvtwo$:

The procedure to construct $\uadvtwo$ is described in detail in Appendix ref{app:advective\_vel}
and is analogous to the procedure used in step 2, but with updated values
for $w_0$ and $\pizero.$ We note that  $\uadvtwo$ satisfies the discrete versions of 
equations (ref{eq:udoterzero}) and (ref{eq:tildeconstraint}),
in particular
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) = \beta_0^{\nph,\star} 
\left(S^{\nph} - \overline{S}^{\nph}\right) \enskip .
\end{equation}
where $ \beta_0^{\nph,\star} = ( \beta_0^n +  \beta_0^{n+1,\star}) / 2.$

%--------------------------------------------------------------------------
% STEP 8
%--------------------------------------------------------------------------
\noindent {\bf Step 8.} {\em Advect the base state, then the full state, through a time interval of $\dt.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item {\bf Advect Base}$(\rho_0^{n}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph}, \psi^{\nph,\star} ) \rightarrow $
$(\rho_0^{(2)}, \rho_0^{\nph,{\rm pred}}, p_0^{n+1}).$

\item Compute the edge states, $(\rho X_k)^{(1),\nph}$ and $(\rho h)^{(1),\nph},$
      for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
      predict $\rho'$, $T$, and $X_k$ to the edges, using
      a second-order Taylor expansion in space and time, as described in
      paper II, Appendix A, using $\V = \uadvtwo+w_0^{\nph} \er$. 
      Again, we explicitly include the reaction terms in the temperature prediction, 
      since we did not update  
      the temperature in {\bf React State}.
      We use the equation of state and base state density, $\rho_0$,
      to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
      We define $\rho^{(1),\nph} = \sum_k (\rho X_k)^{(1),\nph}.$

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ and
$(\rho h)^{(1)} \rightarrow (\rho h)^{(2)},$ 
\begin{eqnarray}
(\rho X_k)^{(2)} &=& (\rho X_k)^{(1)} 
- \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
(\rho X_k)^{(1),\nph} \right) \right]  \enskip , \\ \nonumber \\
(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
(\rho h)^{(1),\nph} \right) \right] \nonumber \\
&& + \frac{\dt}{2} \; \left(\uadvtwo \cdotb \er\right)
\left [ \left(\frac{\partial p_0}{\partial r} \right)^{n}
      + \left(\frac{\partial p_0}{\partial r} \right)^{n+1}  \right ] 
+ \dt \; \psi^{\nph,\star} \enskip , \\ 
\rho^{(2)} &=& \sum_k (\rho X_k)^{(2)} \enskip , \\
X_k^{(2)} &=& (\rho X_k)^{(2)} / \rho^{(2)} \enskip , \\
T^{(2)} &=& T\left(\rho^{(2)}, (\rho h)^{(2)},  X_k^{(2)}\right) {\hbox{ \rm using the
equation of state.}} 
\end{eqnarray}

%The construction of 
%$ \left[ \nabla \cdotb \left(  (\uadvtwo+w_0^{\nph} \er)  (\rho X_k)^{(1)} \right) \right]^{\nph}$ \\
%and $ \left[ \nabla \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  (\rho h)^{(1)}\right)  \right]^{\nph}$
%is similar to that described in paper II, Appendix A with $\V = \uadvtwo+w_0^{\nph} \er.$ 
%For species advection we set $s = X_k,$ then multiply the edge states of $X_k$ by predicted
%edge states of $\rho.$  For the enthalpy update we predict temperature to the edges, convert
%$T$ to $h$ using the EOS, then multiply by $\rho.$
%We note that the time-centering of the $p_0$ terms in Equations (ref{eq:step8_rhoh})
%matches as closely as possible the time-centering of the $\beta_0$ in 
%Equation (ref{eq:uadvtwo_satisfies}).

\item {\bf Correct Base}$(\rho_0^{(2)}, \etarho^{\nph, \star}) \rightarrow (\rho_0^{n+1})$

\item Define an edge-centered $ \etarho^{\nph}$ and a cell-centered $\psi_j^{\nph}$,
\begin{eqnarray}
 \etarho^{\nph}   &=&  {\rm {\bf Avg}} \left( (\uadvtwo \cdotb \er + w_0^{\nph}) 
\rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} 
 \nonumber \enskip , \\
 \psi_j^{\nph}  &=&  \frac{1}{2} \left(\eta_{rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
\end{eqnarray}
%For spherical geometries, $\psi_j^{\nph}$ would be computed via equation (ref{eq:spherical_psi_num}).
\end{enumerate}

%--------------------------------------------------------------------------
% STEP 9
%--------------------------------------------------------------------------
\noindent {\bf Step 9.} {\em React the full state through a second time interval of $\dt / 2.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}

\item {\bf React State}$(\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho^{(2)} \Hext) )
                   \rightarrow (\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, 
                               (\rho \omegadot_k)^{(2)} ).$

\item Define
\begin{eqnarray}
 \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left(\Gamma_1(\rho^{n+1}, p_0^{n+1}, {X_k}^{n+1}) \right) \enskip , \nonumber \\ 
 {\beta   }_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip .
\end{eqnarray}

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 10
%--------------------------------------------------------------------------
\noindent {\bf Step 10.} {\em Compute $S^{n+1}$ for the final projection.}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}
\item Define:
\begin{equation}
  S^{n+1} =  -\sigma  \sum_k  \xi_k (\omegadot_k)^{(2)}  + \sigma \Hnuc^{(2)} +
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2)}  
   + \sigma \Hext \enskip ,
\end{equation}
where $(\omegadot_k)^{(2)} = (\rho \omegadot_k)^{(2)} / \rho^{(2)}$
and the remaining quantities are defined using $X_k^{n+1},$ $\rho^{n+1},$
and $T^{n+1}$ from step 9.

\item Define
\[
\overline{S}^{n+1} = {\mathrm{\bf Avg}} (S^{n+1}) \enskip.
\]

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 11
%--------------------------------------------------------------------------
\noindent {\bf Step 11.} {\em Update the velocity}.

The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
We update the velocity field $\ut^n$ to $\ut^{n+1,\star}$ by discretizing
equation (ref{eq:utildeupd}), 
\begin{eqnarray}
\ut^{n+1,\star} &=& \ut^n - \dt \;
 \left[\left((\uadvtwo+ w_0^{\nph} \er) \cdotb \nablab\right) \ut \right]^\nph
      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
\end{eqnarray}
where $\rho^\nph = ( \rho^n + \rho^{n+1} )/2$
and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
The construction of $\left[((\uadvtwo+ w_0^{\nph} \er) \cdotb \nablab) \ut \right]^\nph$
is described in paper II, Appendix A, with $\V = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
component of $\ut^n$ individually.

Finally, we impose the divergence constraint from equation (ref{eq:tildeconstraint}),
\[
\nablab \cdotb \left(\beta_0^{\nph} \ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)\enskip,
\]
by defining ${\bf V} = \ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ and solving
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
\end{equation}
for nodal values of $\phi$, where
$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
$L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve the linear system of equation (ref{eq:nodal\_solve}) using multigrid V-cycles with Gauss-Seidel relaxation.

We determine the new time velocity field from
\begin{equation}
\ut^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
\end{equation}
and the new time-centered perturbational pressure from:
\[
  \pi^\nph = \frac{1}{\dt} \phi \enskip .
\]

%--------------------------------------------------------------------------
% STEP 12
%--------------------------------------------------------------------------
\noindent {\bf Step 12.} {\em Compute a new $\dt.$}

Compute $\dt$ for the next time step with the procedure described in \S~ref{sec:timesteps},
using $w_0$ as computed in step 6 and
$\ut^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.

\noindent This completes the time advancement of the algorithm.

\subsection{Initialization}

We start each calculation with user-specified initial values for
$\rho$, $X_k$ and $T,$ as well as an initial background state.  In
order for the low Mach number assumption to hold, the initial data
must be thermodynamically consistent with the initial background
state.  In addition, the initial velocity field must satisfy an
initial approximation to the divergence constraint.  We use an iterative
procedure to compute both an initial right-hand-side for the
constraint equation and an initial velocity field that satisfies
the constraint.  The user specifies the number of iterations,
$N_{\rm iters}^{S},$ in this first step of the initialization procedure.

The initial perturbational pressure also needs to be determined for
use in steps 2, 7 and 11. 
This is done through a second iterative procedure which follows the
time advancement algorithm as described in steps 1-11 in \S~ref{Sec:Alg}.  
The user specifies the number of iterations, 
$N_{\rm iters}^{\pi},$ in this second step of the initialization procedure.
The details for both iterations are given below.

%--------------------------------------------------------------------------
% STEP 0
%--------------------------------------------------------------------------
\noindent {\bf Step 0} {\em Initialization}

Start with initial data $X_k^{\initp}, \rho^{\initp},$ $T^{\initp},$ an 
initial base state, and an initial guess for the velocity, $\ubold^{\initp}.$
Use the equation of state to determine $(\rho h)^{\initp}$.  Set
$w_0^1 = 0$ as an initial approximation.  Compute $\beta_0^{\initp}$ as a function of 
the initial data.  Then, project $\ubold^{\initp}$ using $\beta_0^{\initp}$ and 
$S = \rho\Hext$, giving $\ubold^{0,1}$.  The next part of the initialization process 
proceeds as follows.

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}
%\renewcommand{\theenumii}{{\bf \roman{enumii}}}
\renewcommand{\labelenumii}{\roman{enumii}.}

\item {\bf Do} {$\nu = 1,...,N_{\rm iters}^{S}$.}
  \begin{enumerate}

  \item Estimate $\Delta t^\nu$ using $\ubold^{0,\nu}$ and $w_0^\nu.$

  \item {\bf React State}$( \rho^{\initp},(\rho h)^{\initp}, X_k^{\initp}, T^{\initp}, (\rho^{\initp} \Hext) )$\\
      $\rightarrow (\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{0,\nu} ).$

  \item Compute $S^{0,\nu}$ from equation (ref{eq:defineS}) 
        using $(\rho \omegadot_k)^{0,\nu}$ and the initial data.

  \item Compute $\overline{S}^{0,\nu} = {\mathrm{\bf Avg}} (S^{0,\nu}).$

  \item Compute $w_0^{\nu+1}$ as in step 1c using $\overline{S}^{0,\nu}$ and $\psi^{\nmh} = 0$.

  \item Project $\ubold^{0,\nu}$ using $\beta_0^{\initp}$ and 
        $(S^{0,\nu} - \overline{S}^{0,\nu})$ as the source term.  
        This yields $\ubold^{0,\nu+1}.$

  \end{enumerate}

  {\bf End do.}

  Define $S^0 = S^{0,N_{\rm iters}^S}$, $w_0^{-\myhalf} = w_0^{N_{\rm iters}^S+1}$, 
$\dt^0 = \Delta t^{N_{\rm iters}^S},$ and $\ubold^0 = \ubold^{0,N_{\rm iters}^S+1}.$

\end{enumerate}

Next, we need to construct an approximation to the time-centered perturbational pressure,
$\pi,$  and an approximation to the divergence constraint at the end of the
first time step.  As initial approximations, set $S^{1,0} = S^0$,
$\etarho^{-\half} = 0$, $\psi^{-\half} = 0$, and $\pi^{-\half} = 0.$
\begin{enumerate}
  \renewcommand{\theenumi}{{\bf \alph{enumi}}}
  \renewcommand{\labelenumii}{\roman{enumii}.}
  \addtocounter{enumi}{1}

\item {\bf Do} {$\nu = 0,...,N_{\rm iters}^{\pi}-1$.}

  \begin{enumerate}
  \item Perform steps 1--11 as described above, using 
    $S^{\half,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
    The only other difference in the time advancement is that in step 11
    we define ${\bf V} = (\ut^{1,\star} - \ut^0)$ and solve
    \begin{equation}  L_\beta^\rho \phi =
      D \left ( \beta_0^{\half} {\bf V} \right) - \beta_0^{\half} \left( (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right) \enskip . 
    \end{equation}
    (The motivation for this form of the projection in the initial pressure iterations
    is discussed in (cite almgren:bell:crutchfield).)
      We discard the new velocity resulting from this, but keep the new  
      value for $\pi^{\half} = \pi^{-\half} + (1 / \dt) \; \phi.$  
      These steps also yield new scalar data at time $\dt,$ which
      we discard,  and new values for $\etarho^{\half}$ (step 8), $\psi^{\half}$ (step 8), 
      $S^{1,\nu+1}$ (step 10), and $\pi^{\half}$ (step 11), which we keep.
    \item Set $\pi^{-\half} = \pi^{\half}$, $\etarho^{-\half} = \etarho^{\half}$,
      and $\psi^{-\half} = \psi^{\half}$. 
    \end{enumerate}

    {\bf End do.}

    Finally, we define $S^1 = S^{1,N_{\rm iters}^\pi}.$

  \end{enumerate}

\end{document}

### 2008-08-15 docs/flowchart/paper.tex
no reason to include paper III algorithm anymore since it's updated in lowMach3/paper.tex
Added the equations that the algorithm section references

@@
 \newcommand{\uadvtwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV},(2)}}
-\newcommand{\U}{\mathbf{U}}
 \newcommand{\V}{\mathbf{V}}
-\newcommand{\uadvstar}{\widetilde{\mathbf{U}}^{\mathrm{ADV,\star}}}
-\newcommand{\uadvstarone}{\widetilde{\mathbf{U}}^{\mathrm{ADV,(1),\star}}}
-\newcommand{\uadvstartwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV,(2),\star}}}
-\newcommand{\wadv}{w^{\mathrm{ADV}}}
 \newcommand{\dt}{\Delta t}
 \newcommand{\dr}{\Delta r}
-\newcommand{\dx}{\Delta \x}
-\newcommand{\dy}{\Delta r}
-\newcommand{\etal}{\it et al.}
 
 \newcommand{\etarho}{\eta_{\rho}}
-\newcommand{\etaXk}{\eta_{\rho X_k}}
-\newcommand{\etarhoh}{\eta_{\rho h}}
 
@@
 \newcommand{\ut}{\widetilde{\ubold}}
-\newcommand{\wt}{\tilde{w}}
-
-\newcommand{\hp}{\widetilde{h}}
 
@@
 \newcommand{\er}{\mathbf{e}_r}
-\newcommand{\nx}{\mathbf{n}_x}
-
-\newcommand{\rhobar}{\overline{\rho}}
-
-\newcommand{\x}{\mathbf{x}}
-\newcommand{\rp}{r^\prime}
-\newcommand{\rhop}{\rho^\prime}
-
-\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
-
-\newcommand{\ext}{\mathrm{ext}}
-\newcommand{\eff}{\mathrm{eff}}
-
-\title{Flowchart}
 
@@
 
+\title{Flowchart of Current Algorithm}
+
 \begin{document}
@@
 \cleardoublepage
+
+\section{Equations}
+\begin{eqnarray}
+\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \label{eq:VODE1C} \\
+\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip. \label{eq:tempreactC}
+\end{eqnarray}
+\begin{equation}
+\label{eq:tempenthalpy}
+\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
+ - \sum_k \rho \xi_k {\omegadot}_k 
+ + \rho \Hnuc + \rho \Hext \right] \enskip , 
+\end{equation}
+\begin{eqnarray}
+\frac{\partial \rhozero}{\partial t} &=& - \nablab \cdotb \left( \rhozero w_0 \er \right)
+- \nablab \cdotb \left( \etarho \er \right) \enskip ,
+\label{eq:rho0upd_new}
+\end{eqnarray}
+\begin{equation}
+\frac{\partial p_0}{\partial t} = w_0 \rho_0 g  + \etarho g
+                                 = - w_0 \frac{\partial p_0}{\partial r} + \etarho g \enskip .
+\label{eq:p0upd_new} 
+\end{equation}
+\begin{equation}
+\label{eq:constraint}
+\nablab \cdotb \ubold + \alpha \left( \frac{\partial p_0}{\partial t} + \ubold \cdotb \nablab p_0 \right) =
+ -\sigma  \sum_k \xi_k \omegadot_k  + 
+  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc + \sigma \Hext \equiv S \enskip ,
+\end{equation}
+\begin{eqnarray}
+\frac{\partial w_0}{\partial r} &=& \frac{1}{\beta_0} \nablab \cdotb (\beta_0 w_0 \er) 
+                                    - w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}  \nonumber \\
+                                &=&   \left( \Sbar - \frac{1}{\gammabar p_0}
+                                      \frac{\partial p_0}{\partial t} \right)
+                                     - w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}  \nonumber \\
+                                &=&   \Sbar - \frac{1}{\gammabar p_0}
+                                      \left( \frac{\partial p_0}{\partial t}
+                                       + w_0 \frac{\partial p_0}{\partial r}  \right) \nonumber \\
+                                &=&   \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip \label{eq:divw0} ,
+\end{eqnarray}
+\begin{equation}
+- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
+\end{equation}
+\begin{equation}
+\int_{\Omega_H} \ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
+\end{equation}
+\begin{equation}
+\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )
+ \enskip .  \label{eq:tildeconstraint}
+\end{equation}
+\begin{equation}
+w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
+\end{equation}
+\begin{equation}
+\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
++ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\label{eq:dw0drnstar}
+\end{equation}
+\begin{eqnarray}
+\frac{\partial w_0}{\partial t} &=& - w_0 \frac{\partial w_0}{\partial r} - 
+\frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \label{eq:w0upd} \enskip , \\
+\frac{\partial\ut}{\partial t} &=& - \ut \cdotb \nablab \ut - w_0 \frac{\partial \ut}{\partial r}
+                                 - \left(\ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
+                                 - \frac{1}{\rho} \nablab\pi
+                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
+                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip ,
+\end{eqnarray}
+\begin{equation}
+ L_\beta^\rho \phi =
+   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
+\label{eq:nodal_solve}
+\end{equation}
+\begin{equation}
+  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
+  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext \enskip .
+\label{eq:defineS} 
+\end{equation}
 
@@
 The integration yields the new values of the mass fractions, $X_k^{\outp}$.  
-Equation (ref{eq:tempreactC}) is derived from equation (ref{eq:tempenthalpy}) by assuming that the pressure is constant during the burn state.  
+Equation (\ref{eq:tempreactC}) is derived from equation (\ref{eq:tempenthalpy}) by assuming that the pressure is constant during the burn state.  
 In evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  
@@
 at cell centers, with $w_0$ and $\beta_0$ defined halfway in-between.\\ \\
-a)  For the density update, we discretize equation (ref{eq:rho0upd\_new}), neglecting the $\etarho$ term,
+a)  For the density update, we discretize equation (\ref{eq:rho0upd_new}), neglecting the $\etarho$ term,
 to compute the new base state density,
@@
 paper II, Appendix A, and are saved for use later in the overall algorithm.\\ \\
-b)  For the pressure update, We discretize equation (ref{eq:p0upd\_new}), 
+b)  For the pressure update, We discretize equation (\ref{eq:p0upd_new}), 
 to compute the new base state pressure,
@@
 \item At the beginning of each time step, we need an estimate for the time-centered
-source term in the velocity divergence constraint (see eq. [ref{eq:constraint}]).
+source term in the velocity divergence constraint (see eq. [\ref{eq:constraint}]).
 If this is the first step of the calculation ($n=0$), we set
@@
 
-\item Construct $w_0^{\nph,\star}$ by integrating equation (ref{eq:divw0}) using 
+\item Construct $w_0^{\nph,\star}$ by integrating equation (\ref{eq:divw0}) using 
 the lagged $\psi^{\nmh},$
@@
 \end{equation}
-%for plane-parallel geometries,  or the analogous equation (see eq. [ref{eq:B7}])
-%for spherical configurations.  
 For spherical geometries, this equation would be modified.
-%In the case of spherical geometries use the current $p_0^n,$
-%${\Gamma_1}^n_0 = \Gamma_1(p_0^n, \rho_0^n, X_k^n),$ and $g$ as computed from $\rho_0^n.$
 
-\item Using equation (ref{eq:pizero}), define the scaled pressure gradient
+\item Using equation (\ref{eq:pizero}), define the scaled pressure gradient
 \begin{equation}
@@
 
-The procedure to construct $\uadvone$ is described in detail in Appendix ref{app:advective\_vel}.
+The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
 We note that  $\uadvone$ satisfies the discrete versions of equations
-(ref{eq:udoterzero}) and (ref{eq:tildeconstraint}), 
+(\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}), 
 in particular,
@@
 
-%The construction of 
-%$ \left[ \nabla \cdotb \left(  (\uadvone+w_0^{\nph,\star} \er)  (\rho X_k)^{(1)} \right) \right]^{\nph}$ \\
-%and $ \left[ \nabla \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  (\rho h)^{(1)}\right)  \right]^{\nph}$
-%also follows that described in paper II, Appendix A with $\V = \uadvone+w_0^{\nph,\star} \er.$ 
-%We note that the time-centering of the $p_0$ terms in equations (ref{eq:step4_rhoh})
-%matches as closely as possible the time-centering of the $\beta_0$ in 
-%equation (ref{eq:uadvone_satisfies}).
-
 \item {\bf Correct Base}$(\rho_0^{(2),\star}, \etarho^{\nmh}) \rightarrow (\rho_0^{n+1,\star})$.
@@
 For spherical geometries, $\psi_j^{\nph,\star}$ would have a different form.
-%be computed via equation (ref{eq:spherical_psi_num}).
 
@@
 
-\item Construct $w_0^{\nph}$ by integrating equation (ref{eq:divw0}),
+\item Construct $w_0^{\nph}$ by integrating equation (\ref{eq:divw0}),
 \begin{equation}
@@
 $ p_0^{\nph,\star} = (p_0^{n} + p_0^{n+1,\star} ) / 2.$
-%In spherical configurations we use the analogous equation (see eq. [ref{eq:B7}]). 
-%In the case of spherical geometries use $p_0^{n+1,\star},$
-%${\Gamma_1}^{n+1,\star}_0 = \Gamma_1(p_0^{n+1,\star}, \rho_0^{n+1,\star}, X_k^{n+1,\star}),$ and 
-%$g$ as computed from $\rho_0^{n+1,\star}.$
 
-\item Using equation (ref{eq:pizero}), define
+\item Using equation (\ref{eq:pizero}), define
 \begin{equation}
@@
 where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined analogously to 
-equations (ref{eq:w0nstar}) and (ref{eq:dw0drnstar}).
+equations (\ref{eq:w0nstar}) and (\ref{eq:dw0drnstar}).
 
@@
 
-The procedure to construct $\uadvtwo$ is described in detail in Appendix ref{app:advective\_vel}
+The procedure to construct $\uadvtwo$ is described in detail in Appendix B of paper III
 and is analogous to the procedure used in step 2, but with updated values
 for $w_0$ and $\pizero.$ We note that  $\uadvtwo$ satisfies the discrete versions of 
-equations (ref{eq:udoterzero}) and (ref{eq:tildeconstraint}),
+equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}),
 in particular
@@
 
-%The construction of 
-%$ \left[ \nabla \cdotb \left(  (\uadvtwo+w_0^{\nph} \er)  (\rho X_k)^{(1)} \right) \right]^{\nph}$ \\
-%and $ \left[ \nabla \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  (\rho h)^{(1)}\right)  \right]^{\nph}$
-%is similar to that described in paper II, Appendix A with $\V = \uadvtwo+w_0^{\nph} \er.$ 
-%For species advection we set $s = X_k,$ then multiply the edge states of $X_k$ by predicted
-%edge states of $\rho.$  For the enthalpy update we predict temperature to the edges, convert
-%$T$ to $h$ using the EOS, then multiply by $\rho.$
-%We note that the time-centering of the $p_0$ terms in Equations (ref{eq:step8_rhoh})
-%matches as closely as possible the time-centering of the $\beta_0$ in 
-%Equation (ref{eq:uadvtwo_satisfies}).
-
 \item {\bf Correct Base}$(\rho_0^{(2)}, \etarho^{\nph, \star}) \rightarrow (\rho_0^{n+1})$
@@
 \end{eqnarray}
-%For spherical geometries, $\psi_j^{\nph}$ would be computed via equation (ref{eq:spherical_psi_num}).
 \end{enumerate}
@@
 We update the velocity field $\ut^n$ to $\ut^{n+1,\star}$ by discretizing
-equation (ref{eq:utildeupd}), 
+equation (\ref{eq:utildeupd}), 
 \begin{eqnarray}
@@
 
-Finally, we impose the divergence constraint from equation (ref{eq:tildeconstraint}),
+Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}),
 \[
@@
 of this particular form of the projection operand.)  
-We solve the linear system of equation (ref{eq:nodal\_solve}) using multigrid V-cycles with Gauss-Seidel relaxation.
+We solve the linear system of equation (\ref{eq:nodal_solve}) using multigrid V-cycles with Gauss-Seidel relaxation.
 
@@
 \renewcommand{\theenumi}{{\bf \alph{enumi}}}
-%\renewcommand{\theenumii}{{\bf \roman{enumii}}}
 \renewcommand{\labelenumii}{\roman{enumii}.}
@@
 
-  \item Compute $S^{0,\nu}$ from equation (ref{eq:defineS}) 
-        using $(\rho \omegadot_k)^{0,\nu}$ and the initial data.
-
-  \item Compute $\overline{S}^{0,\nu} = {\mathrm{\bf Avg}} (S^{0,\nu}).$
-
-  \item Compute $w_0^{\nu+1}$ as in step 1c using $\overline{S}^{0,\nu}$ and $\psi^{\nmh} = 0$.
-
-  \item Project $\ubold^{0,\nu}$ using $\beta_0^{\initp}$ and 
-        $(S^{0,\nu} - \overline{S}^{0,\nu})$ as the source term.  
-        This yields $\ubold^{0,\nu+1}.$
-
-  \end{enumerate}
-
-  {\bf End do.}
-
-  Define $S^0 = S^{0,N_{\rm iters}^S}$, $w_0^{-\myhalf} = w_0^{N_{\rm iters}^S+1}$, 
-$\dt^0 = \Delta t^{N_{\rm iters}^S},$ and $\ubold^0 = \ubold^{0,N_{\rm iters}^S+1}.$
-
-\end{enumerate}
-
-Next, we need to construct an approximation to the time-centered perturbational pressure,
-$\pi,$  and an approximation to the divergence constraint at the end of the
-first time step.  As initial approximations, set $S^{1,0} = S^0$,
-$\etarho^{-\half} = 0$, $\psi^{-\half} = 0$, and $\pi^{-\half} = 0.$
-\begin{enumerate}
-  \renewcommand{\theenumi}{{\bf \alph{enumi}}}
-  \renewcommand{\labelenumii}{\roman{enumii}.}
-  \addtocounter{enumi}{1}
-
-\item {\bf Do} {$\nu = 0,...,N_{\rm iters}^{\pi}-1$.}
-
-  \begin{enumerate}
-  \item Perform steps 1--11 as described above, using 
-    $S^{\half,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
-    The only other difference in the time advancement is that in step 11
-    we define ${\bf V} = (\ut^{1,\star} - \ut^0)$ and solve
-    \begin{equation}  L_\beta^\rho \phi =
-      D \left ( \beta_0^{\half} {\bf V} \right) - \beta_0^{\half} \left( (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right) \enskip . 
-    \end{equation}
-    (The motivation for this form of the projection in the initial pressure iterations
-    is discussed in (cite almgren:bell:crutchfield).)
-      We discard the new velocity resulting from this, but keep the new  
-      value for $\pi^{\half} = \pi^{-\half} + (1 / \dt) \; \phi.$  
-      These steps also yield new scalar data at time $\dt,$ which
-      we discard,  and new values for $\etarho^{\half}$ (step 8), $\psi^{\half}$ (step 8), 
-      $S^{1,\nu+1}$ (step 10), and $\pi^{\half}$ (step 11), which we keep.
-    \item Set $\pi^{-\half} = \pi^{\half}$, $\etarho^{-\half} = \etarho^{\half}$,
-      and $\psi^{-\half} = \psi^{\half}$. 
-    \end{enumerate}
-
-    {\bf End do.}
-
-    Finally, we define $S^1 = S^{1,N_{\rm iters}^\pi}.$
-
-  \end{enumerate}
-
-\cleardoublepage
-
-\section{Paper III Algorithm}
-\subsection{Notation}
-
-We make use of the following shorthand notations in outlining the algorithm.
-\begin{enumerate}
-
-\item For any quantity, $\phi,$ we define $\overline{\phi} = ${\bf Avg}$(\phi),$ 
-the average over $\Omega_H,$ as
-\begin{equation}
-\overline{\phi}(r) = \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
-\end{equation}
-
-\item {\bf React State}$(\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp})\rightarrow$
-$(\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp})$ \newline
-is the process by which we evolve the species and enthalpy from $X_k^{\inp} \rightarrow X_k^{\outp}$ 
-and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following system of equations 
-over a time interval of  $\Delta t/2$,
-\begin{eqnarray}
-\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
-\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
-\end{eqnarray}
-  In particular, to evolve the species, we solve the system:
-\begin{eqnarray}
-\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
-\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip.
-\end{eqnarray}
-using the stiff ordinary differential equation
- integration methods provided by the {\tt VODE} package (cite VODE).
-The absolute error tolerances are set to $10^{-12}$ for the species,
- and a relative tolerance of $10^{-5}$ is used for the temperature.  
-The integration yields the new values of the mass fractions, $X_k^{\outp}$.  
-Equation (ref{eq:tempreactC}) is derived from equation (ref{eq:tempenthalpy}) by assuming that the pressure is constant during the burn state.  
-In evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  
-In theory, this means evaluating the equation of state for each right-hand side evaluation that {\tt VODE} requires.  
-In practice, we freeze $c_p$ and $\xi_k$ at the start of the integration time step and compute them using 
-$\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs to the equation of state.  
-We note that while temperature is evolved when solving these equations, we do not keep the final temperature, 
-nor do we use it to compute the final change in enthalpy.  
-Therefore, $T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged during 
-the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
-
-After the new mass fractions have been computed, the reaction rates are defined as:
-\begin{equation}
-(\rho\omegadot_k)^{\outp} = \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\Delta t /2} \enskip,
-\end{equation}
-and the nuclear energy generation rate is defined as
-\begin{equation}
-(\rho\Hnuc) = -\sum_k(\rho\omegadot_k)^{\outp} \; q_k \enskip.
-\end{equation}
-The enthalpy update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is updated by
-\begin{equation}
-(\rho h)^{\outp} = (\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc) + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
-\end{equation}
-
-\item {\bf Advect Base}$(\rho_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp} ) \rightarrow 
-(\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ 
-\newline
-is the process by which we update the base state through $\dt$ in time given the
-radial velocity $w_0^{\inp}.$ Here we discuss
-the algorithm for plane-parallel geometries.  
-The base state arrays are all one-dimensional 
-in the radial coordinate; we think of the base state quantities as defined
-at cell centers, with $w_0$ and $\beta_0$ defined halfway in-between.\\ \\
-a)  For the density update, we discretize equation (ref{eq:rho0upd\_new}), neglecting the $\etarho$ term,
-to compute the new base state density,
-\begin{eqnarray}
-\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
-\left [ \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
-         \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ] \nonumber  \enskip,
-\end{eqnarray}
-where $j$ refers to the one-dimensional index in the radial direction.
-The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the procedure described in
-paper II, Appendix A, and are saved for use later in the overall algorithm.\\ \\
-b)  For the pressure update, We discretize equation (ref{eq:p0upd\_new}), 
-to compute the new base state pressure,
-\begin{eqnarray}
-p_{0,j}^\outp &=& p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
-+ w_{0,j-\myhalf}^\inp \right)
-\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\inp \enskip,
-\end{eqnarray}
-where the interface states are again found via the procedure described in
-paper II, Appendix A.
-\\
-
-\item {\bf Correct Base}$(\rho_0^{\inp} , \etarho^{\inp}) \rightarrow (\rho_0^{\outp})$
-is the process by which we adjust the base state density given $\etarho$ using
-\begin{eqnarray}
-\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
-\left( \eta_{\rho,j+\myhalf}^{\inp} - \eta_{\rho,j-\myhalf}^{\inp}  \right)  \enskip . 
-\end{eqnarray} 
-\end{enumerate}
-
-\subsection{Time Advancement Algorithm}
-
-We now describe the full time advancement algorithm, making frequent
-use of the shorthand developed above.  Here, we assume that the
-problem is already properly initialized.  We describe the details of
-the initialization in \S~ref{Sec:Initialization}.
-
-In paper II we discretized the density and enthalpy evolution equations in 
-perturbational form, arguing that this would be less
-susceptible to grid effects.  In the present algorithm, the analog would
-be to use a perturbational form of the species equations.
-Numerical testing has shown that the form of the algorithm presented
-here is in fact more robust than the perturbational form 
-when the base state mixing terms ($\etarho$) are included.  
-In the algorithm described below, density is still 
-extrapolated to edges in perturbational form, but 
-we treat $X_k$ and $h$ in non-perturbational form.
-
-%--------------------------------------------------------------------------
-% STEP 1
-%--------------------------------------------------------------------------
-\noindent {\bf Step 1.} {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item At the beginning of each time step, we need an estimate for the time-centered
-source term in the velocity divergence constraint (see eq. [ref{eq:constraint}]).
-If this is the first step of the calculation ($n=0$), we set
-\[
-S^{\half,\star} = \frac{S^0 + S^1}{2} \enskip ,
-\]
-where $S^1$ is found through the iterative process that initializes the calculation.
-Otherwise, following the method used in our small-scale low Mach number algorithm
-(Cite Bell04), we extrapolate to the half-time using the source
-terms at the previous and current time levels
-\begin{equation}
-S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
-\end{equation}
-
-\item Define
-\[
-\overline{S}^{\nph,\star} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip.
-\]
-
-\item Construct $w_0^{\nph,\star}$ by integrating equation (ref{eq:divw0}) using 
-the lagged $\psi^{\nmh},$
-\begin{equation}
-\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S}^{\nph,\star} 
-- \frac{1}{\gammabar^{n} p_0^{n}} \psi^{\nmh} \enskip . \nonumber
-\end{equation}
-%for plane-parallel geometries,  or the analogous equation (see eq. [ref{eq:B7}])
-%for spherical configurations.  
-For spherical geometries, this equation would be modified.
-%In the case of spherical geometries use the current $p_0^n,$
-%${\Gamma_1}^n_0 = \Gamma_1(p_0^n, \rho_0^n, X_k^n),$ and $g$ as computed from $\rho_0^n.$
-
-\item Using equation (ref{eq:pizero}), define the scaled pressure gradient
-\begin{equation}
--\left ( \frac{1}{\rho_0^n} \frac{\partial \pizeroone}{\partial r} \right ) = 
-\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} 
-+  w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star} \enskip ,
-\end{equation} 
-where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are
-\begin{equation}
-w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\end{equation}
-\begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
-+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\end{equation}
-If $n=0$, we use $\dt^{-1} = \dt^0$.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 2
-%--------------------------------------------------------------------------
-\noindent {\bf Step 2.} {\em Construct the provisional edge-based advective velocity}, $\uadvone$:
-
-The procedure to construct $\uadvone$ is described in detail in Appendix ref{app:advective\_vel}.
-We note that  $\uadvone$ satisfies the discrete versions of equations
-(ref{eq:udoterzero}) and (ref{eq:tildeconstraint}), 
-in particular,
-\begin{equation}
-\nablab \cdotb \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
-\enskip ,
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 3
-%--------------------------------------------------------------------------
-\noindent {\bf Step 3.} {\em React the full state through the first time interval of $\dt / 2.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf React State}$(\rho^n, (\rho h)^n, X_k^n, T^n, (\rho^n \Hext))
-                   \rightarrow$ $( \rho^{(1)}, (\rho h)^{(1)}, X_k^{(1)}, T^{(1)},
-                                  (\rho \omegadot_k)^{(1)} ).$
-
-\item
-Define
-\begin{eqnarray}
-\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{(1)}, p_0^{n}, X_k^{(1)}) \right) 
-                                                                \enskip , \nonumber \\
- {\beta   }_0^{(1)}    &=& \beta   \left(\rho_0^{n}, p_0^{n}, \gammabar^{(1)}\right) \enskip .
-\end{eqnarray}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 4
-%--------------------------------------------------------------------------
-\noindent {\bf Step 4.} {\em Advect the base state, then the full state, through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf Advect Base}$(\rhozero^{n}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph,\star}, \psi^{\nmh}) \rightarrow $
-$(\rho_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}, p_0^{n+1,\star}).$
-
-\item Compute the edge states, $(\rho X_k)^{(1),\nph,\star}$ and $(\rho h)^{(1),\nph,\star},$
-      for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
-      predict $\rho'$, $T$, and $X_k$ to the edges, using
-      a second-order Taylor expansion in space and time, as described in
-      paper II, Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  
-      We explicitly include the reaction terms in the temperature prediction, 
-      since we 
-      did not update the temperature in {\bf React State}. 
-      We use the equation of state and base state density, $\rho_0$,
-      to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
-      We define $\rho^{(1),\nph,\star} = \sum_k (\rho X_k)^{(1),\nph,\star}.$
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ and
-$(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ without explicitly including the 
-reaction terms,
-\begin{eqnarray}
-(\rho X_k)^{(2),\star} &=& (\rho X_k)^{(1)} 
- - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
-  (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip , \\ \nonumber  \\
-(\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
-  (\rho h)^{(1),\nph,\star} \right) \right] \nonumber \\
-&& + \dt \; \left(\uadvone \cdotb \er\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} 
-+ \dt \; \psi^{\nmh} \enskip , \\
-\rho^{(2),\star} &=& \sum_k (\rho X_k)^{(2),\star} \enskip , \\ 
-X_k^{(2),\star} &=& (\rho X_k)^{(2),\star} / \rho^{(2),\star} \enskip , \\
-T^{(2),\star} &=& T\left(\rho^{(2),\star}, (\rho h)^{(2),\star},  X_k^{(2),\star}\right) {\hbox{ \rm using the equation of state.}} 
-\end{eqnarray}
-
-%The construction of 
-%$ \left[ \nabla \cdotb \left(  (\uadvone+w_0^{\nph,\star} \er)  (\rho X_k)^{(1)} \right) \right]^{\nph}$ \\
-%and $ \left[ \nabla \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  (\rho h)^{(1)}\right)  \right]^{\nph}$
-%also follows that described in paper II, Appendix A with $\V = \uadvone+w_0^{\nph,\star} \er.$ 
-%We note that the time-centering of the $p_0$ terms in equations (ref{eq:step4_rhoh})
-%matches as closely as possible the time-centering of the $\beta_0$ in 
-%equation (ref{eq:uadvone_satisfies}).
-
-\item {\bf Correct Base}$(\rho_0^{(2),\star}, \etarho^{\nmh}) \rightarrow (\rho_0^{n+1,\star})$.
-
-\item Define an edge-centered $\etarho^{\nph,\star}$ and cell-centered $\psi_j^{\nph,\star}$:
-\begin{eqnarray}
- \etarho^{\nph,\star}   &=&  {\rm {\bf Avg}} \left( (\uadvone \cdotb \er + w_0^{\nph,\star}) 
-\rho^{(1),\nph,\star} \right) - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} 
- \nonumber \enskip , \\
- \psi_j^{\nph,\star}  &=&  \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph,\star} + \eta_{\rho,j+\myhalf}^{\nph,\star}\right) g \enskip .
-\end{eqnarray}
-For spherical geometries, $\psi_j^{\nph,\star}$ would have a different form. 
-%be computed via equation (ref{eq:spherical_psi_num}).
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 5
-%--------------------------------------------------------------------------
-\noindent {\bf Step 5.} {\em React the full state through a second time interval of $\dt / 2.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf React State}$( \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, 
-                             T^{(2),\star}, (\rho^{(2),\star} \Hext)) $\\
-$\rightarrow ( \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, 
-              (\rho \omegadot_k)^{(2),\star} ).$
-
-\item Define
-\begin{eqnarray}
- \gammabar^{n+1,\star}    &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{n+1,\star}, p_0^{n+1,\star}, 
-                                                      X_k^{n+1,\star}) \right) \enskip , \nonumber \\
- {\beta   }_0^{n+1,\star}    &=& \beta   \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}\right) \enskip .
-\end{eqnarray}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 6
-%--------------------------------------------------------------------------
-\noindent {\bf Step 6.} {\em Define a new average expansion rate at time $t^\nph.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-\item Define:
-\begin{equation}
-  S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + 
-  \sigma \Hnuc^{(2),\star} +
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star}  
-  + \sigma \Hext \enskip .
-\end{equation} 
-where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} / \rho^{(2),\star}$
-and the remaining quantities are defined using $X_k^{n+1,\star},$ $\rho^{n+1,\star},$ 
-and $T^{n+1,\star}$ from step 5.
-Then define
-\[
- S^\nph = \frac{S^n + S^{n+1,\star}}{2} \enskip. 
-\]
-
-\item Define:
-\[
-\overline{S}^{\nph} = {\mathrm{\bf Avg}} (S^{\nph}) \enskip.
-\]
-
-\item Construct $w_0^{\nph}$ by integrating equation (ref{eq:divw0}),
-\begin{equation}
-\frac{\partial w_0^{\nph}}{\partial r} =  \overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}} \psi^{\nph,\star} \enskip , \nonumber
-\end{equation}
-for plane-parallel geometries, where 
-$\gammabar^{\nph,\star} = (\gammabar^{n} + \gammabar^{n+1,\star} ) / 2 $ and
-$ p_0^{\nph,\star} = (p_0^{n} + p_0^{n+1,\star} ) / 2.$
-%In spherical configurations we use the analogous equation (see eq. [ref{eq:B7}]). 
-%In the case of spherical geometries use $p_0^{n+1,\star},$
-%${\Gamma_1}^{n+1,\star}_0 = \Gamma_1(p_0^{n+1,\star}, \rho_0^{n+1,\star}, X_k^{n+1,\star}),$ and 
-%$g$ as computed from $\rho_0^{n+1,\star}.$
-
-\item Using equation (ref{eq:pizero}), define
-\begin{equation}
--\left ( \frac{1}{\rho_0^n} \frac{\partial \pizerotwo}{\partial r} \right ) = 
-\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
-+ w_0^n \left(\frac{\partial w_0}{\partial r}\right)^n.
-\end{equation}
-where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined analogously to 
-equations (ref{eq:w0nstar}) and (ref{eq:dw0drnstar}).
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 7
-%--------------------------------------------------------------------------
-\noindent {\bf Step 7.} {\em Construct the final edge-based advective velocity}, $\uadvtwo$:
-
-The procedure to construct $\uadvtwo$ is described in detail in Appendix ref{app:advective\_vel}
-and is analogous to the procedure used in step 2, but with updated values
-for $w_0$ and $\pizero.$ We note that  $\uadvtwo$ satisfies the discrete versions of 
-equations (ref{eq:udoterzero}) and (ref{eq:tildeconstraint}),
-in particular
-\begin{equation}
-\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) = \beta_0^{\nph,\star} 
-\left(S^{\nph} - \overline{S}^{\nph}\right) \enskip .
-\end{equation}
-where $ \beta_0^{\nph,\star} = ( \beta_0^n +  \beta_0^{n+1,\star}) / 2.$
-
-%--------------------------------------------------------------------------
-% STEP 8
-%--------------------------------------------------------------------------
-\noindent {\bf Step 8.} {\em Advect the base state, then the full state, through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf Advect Base}$(\rho_0^{n}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph}, \psi^{\nph,\star} ) \rightarrow $
-$(\rho_0^{(2)}, \rho_0^{\nph,{\rm pred}}, p_0^{n+1}).$
-
-\item Compute the edge states, $(\rho X_k)^{(1),\nph}$ and $(\rho h)^{(1),\nph},$
-      for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
-      predict $\rho'$, $T$, and $X_k$ to the edges, using
-      a second-order Taylor expansion in space and time, as described in
-      paper II, Appendix A, using $\V = \uadvtwo+w_0^{\nph} \er$. 
-      Again, we explicitly include the reaction terms in the temperature prediction, 
-      since we did not update  
-      the temperature in {\bf React State}.
-      We use the equation of state and base state density, $\rho_0$,
-      to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
-      We define $\rho^{(1),\nph} = \sum_k (\rho X_k)^{(1),\nph}.$
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ and
-$(\rho h)^{(1)} \rightarrow (\rho h)^{(2)},$ 
-\begin{eqnarray}
-(\rho X_k)^{(2)} &=& (\rho X_k)^{(1)} 
-- \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
-(\rho X_k)^{(1),\nph} \right) \right]  \enskip , \\ \nonumber \\
-(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
-(\rho h)^{(1),\nph} \right) \right] \nonumber \\
-&& + \frac{\dt}{2} \; \left(\uadvtwo \cdotb \er\right)
-\left [ \left(\frac{\partial p_0}{\partial r} \right)^{n}
-      + \left(\frac{\partial p_0}{\partial r} \right)^{n+1}  \right ] 
-+ \dt \; \psi^{\nph,\star} \enskip , \\ 
-\rho^{(2)} &=& \sum_k (\rho X_k)^{(2)} \enskip , \\
-X_k^{(2)} &=& (\rho X_k)^{(2)} / \rho^{(2)} \enskip , \\
-T^{(2)} &=& T\left(\rho^{(2)}, (\rho h)^{(2)},  X_k^{(2)}\right) {\hbox{ \rm using the
-equation of state.}} 
-\end{eqnarray}
-
-%The construction of 
-%$ \left[ \nabla \cdotb \left(  (\uadvtwo+w_0^{\nph} \er)  (\rho X_k)^{(1)} \right) \right]^{\nph}$ \\
-%and $ \left[ \nabla \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  (\rho h)^{(1)}\right)  \right]^{\nph}$
-%is similar to that described in paper II, Appendix A with $\V = \uadvtwo+w_0^{\nph} \er.$ 
-%For species advection we set $s = X_k,$ then multiply the edge states of $X_k$ by predicted
-%edge states of $\rho.$  For the enthalpy update we predict temperature to the edges, convert
-%$T$ to $h$ using the EOS, then multiply by $\rho.$
-%We note that the time-centering of the $p_0$ terms in Equations (ref{eq:step8_rhoh})
-%matches as closely as possible the time-centering of the $\beta_0$ in 
-%Equation (ref{eq:uadvtwo_satisfies}).
-
-\item {\bf Correct Base}$(\rho_0^{(2)}, \etarho^{\nph, \star}) \rightarrow (\rho_0^{n+1})$
-
-\item Define an edge-centered $ \etarho^{\nph}$ and a cell-centered $\psi_j^{\nph}$,
-\begin{eqnarray}
- \etarho^{\nph}   &=&  {\rm {\bf Avg}} \left( (\uadvtwo \cdotb \er + w_0^{\nph}) 
-\rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} 
- \nonumber \enskip , \\
- \psi_j^{\nph}  &=&  \frac{1}{2} \left(\eta_{rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
-\end{eqnarray}
-%For spherical geometries, $\psi_j^{\nph}$ would be computed via equation (ref{eq:spherical_psi_num}).
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 9
-%--------------------------------------------------------------------------
-\noindent {\bf Step 9.} {\em React the full state through a second time interval of $\dt / 2.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf React State}$(\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho^{(2)} \Hext) )
-                   \rightarrow (\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, 
-                               (\rho \omegadot_k)^{(2)} ).$
-
-\item Define
-\begin{eqnarray}
- \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left(\Gamma_1(\rho^{n+1}, p_0^{n+1}, {X_k}^{n+1}) \right) \enskip , \nonumber \\ 
- {\beta   }_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip .
-\end{eqnarray}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 10
-%--------------------------------------------------------------------------
-\noindent {\bf Step 10.} {\em Compute $S^{n+1}$ for the final projection.}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-\item Define:
-\begin{equation}
-  S^{n+1} =  -\sigma  \sum_k  \xi_k (\omegadot_k)^{(2)}  + \sigma \Hnuc^{(2)} +
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2)}  
-   + \sigma \Hext \enskip ,
-\end{equation}
-where $(\omegadot_k)^{(2)} = (\rho \omegadot_k)^{(2)} / \rho^{(2)}$
-and the remaining quantities are defined using $X_k^{n+1},$ $\rho^{n+1},$
-and $T^{n+1}$ from step 9.
-
-\item Define
-\[
-\overline{S}^{n+1} = {\mathrm{\bf Avg}} (S^{n+1}) \enskip.
-\]
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 11
-%--------------------------------------------------------------------------
-\noindent {\bf Step 11.} {\em Update the velocity}.
-
-The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
-We update the velocity field $\ut^n$ to $\ut^{n+1,\star}$ by discretizing
-equation (ref{eq:utildeupd}), 
-\begin{eqnarray}
-\ut^{n+1,\star} &=& \ut^n - \dt \;
- \left[\left((\uadvtwo+ w_0^{\nph} \er) \cdotb \nablab\right) \ut \right]^\nph
-      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
-   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
-        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
-        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
-\end{eqnarray}
-where $\rho^\nph = ( \rho^n + \rho^{n+1} )/2$
-and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
-The construction of $\left[((\uadvtwo+ w_0^{\nph} \er) \cdotb \nablab) \ut \right]^\nph$
-is described in paper II, Appendix A, with $\V = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
-component of $\ut^n$ individually.
-
-Finally, we impose the divergence constraint from equation (ref{eq:tildeconstraint}),
-\[
-\nablab \cdotb \left(\beta_0^{\nph} \ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)\enskip,
-\]
-by defining ${\bf V} = \ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ and solving
-\begin{equation}
- L_\beta^\rho \phi =
-   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
-\end{equation}
-for nodal values of $\phi$, where
-$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
-$L_\beta^\rho$ is the standard bilinear
-finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
-In this step, $D$ is a discrete second-order operator that approximates the 
-divergence at nodes from cell-centered data and satisfies
-$\mathbf{G} = -D^T.$ 
-(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
-approximate projection; see (cite almgren:bell:crutchfield) for a discussion
-of this particular form of the projection operand.)  
-We solve the linear system of equation (ref{eq:nodal\_solve}) using multigrid V-cycles with Gauss-Seidel relaxation.
-
-We determine the new time velocity field from
-\begin{equation}
-\ut^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
-\end{equation}
-and the new time-centered perturbational pressure from:
-\[
-  \pi^\nph = \frac{1}{\dt} \phi \enskip .
-\]
-
-%--------------------------------------------------------------------------
-% STEP 12
-%--------------------------------------------------------------------------
-\noindent {\bf Step 12.} {\em Compute a new $\dt.$}
-
-Compute $\dt$ for the next time step with the procedure described in \S~ref{sec:timesteps},
-using $w_0$ as computed in step 6 and
-$\ut^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.
-
-\noindent This completes the time advancement of the algorithm.
-
-\subsection{Initialization}
-
-We start each calculation with user-specified initial values for
-$\rho$, $X_k$ and $T,$ as well as an initial background state.  In
-order for the low Mach number assumption to hold, the initial data
-must be thermodynamically consistent with the initial background
-state.  In addition, the initial velocity field must satisfy an
-initial approximation to the divergence constraint.  We use an iterative
-procedure to compute both an initial right-hand-side for the
-constraint equation and an initial velocity field that satisfies
-the constraint.  The user specifies the number of iterations,
-$N_{\rm iters}^{S},$ in this first step of the initialization procedure.
-
-The initial perturbational pressure also needs to be determined for
-use in steps 2, 7 and 11. 
-This is done through a second iterative procedure which follows the
-time advancement algorithm as described in steps 1-11 in \S~ref{Sec:Alg}.  
-The user specifies the number of iterations, 
-$N_{\rm iters}^{\pi},$ in this second step of the initialization procedure.
-The details for both iterations are given below.
-
-%--------------------------------------------------------------------------
-% STEP 0
-%--------------------------------------------------------------------------
-\noindent {\bf Step 0} {\em Initialization}
-
-Start with initial data $X_k^{\initp}, \rho^{\initp},$ $T^{\initp},$ an 
-initial base state, and an initial guess for the velocity, $\ubold^{\initp}.$
-Use the equation of state to determine $(\rho h)^{\initp}$.  Set
-$w_0^1 = 0$ as an initial approximation.  Compute $\beta_0^{\initp}$ as a function of 
-the initial data.  Then, project $\ubold^{\initp}$ using $\beta_0^{\initp}$ and 
-$S = \rho\Hext$, giving $\ubold^{0,1}$.  The next part of the initialization process 
-proceeds as follows.
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-%\renewcommand{\theenumii}{{\bf \roman{enumii}}}
-\renewcommand{\labelenumii}{\roman{enumii}.}
-
-\item {\bf Do} {$\nu = 1,...,N_{\rm iters}^{S}$.}
-  \begin{enumerate}
-
-  \item Estimate $\Delta t^\nu$ using $\ubold^{0,\nu}$ and $w_0^\nu.$
-
-  \item {\bf React State}$( \rho^{\initp},(\rho h)^{\initp}, X_k^{\initp}, T^{\initp}, (\rho^{\initp} \Hext) )$\\
-      $\rightarrow (\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{0,\nu} ).$
-
-  \item Compute $S^{0,\nu}$ from equation (ref{eq:defineS}) 
+  \item Compute $S^{0,\nu}$ from equation (\ref{eq:defineS}) 
         using $(\rho \omegadot_k)^{0,\nu}$ and the initial data.

### 2008-08-15 docs/flowchart/paper.tex
merged the thermal diffusion stuff in

@@
 extrapolated to edges in perturbational form, but 
-we treat $X_k$ and $h$ in non-perturbational form.
+we treat $X_k$ and $h$ in non-perturbational form.\\
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 2.} {\em Construct the provisional edge-based advective velocity}, $\uadvone$:
+\noindent {\bf Step 2.} {\em Construct the provisional edge-based advective velocity}, $\uadvone$.\\
 
@@
   (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip , \\ \nonumber  \\
-(\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
+(\rho h)^{(2'),\star} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
   (\rho h)^{(1),\nph,\star} \right) \right] \nonumber \\
@@
 + \dt \; \psi^{\nmh} \enskip , \\
-\rho^{(2),\star} &=& \sum_k (\rho X_k)^{(2),\star} \enskip , \\ 
-X_k^{(2),\star} &=& (\rho X_k)^{(2),\star} / \rho^{(2),\star} \enskip , \\
-T^{(2),\star} &=& T\left(\rho^{(2),\star}, (\rho h)^{(2),\star},  X_k^{(2),\star}\right) {\hbox{ \rm using the equation of state.}} 
+\rho^{(2'),\star} &=& \sum_k (\rho X_k)^{(2),\star} \enskip , \\ 
+X_k^{(2),\star} &=& (\rho X_k)^{(2),\star} / \rho^{(2),\star}
 \end{eqnarray}
@@
 For spherical geometries, $\psi_j^{\nph,\star}$ would have a different form.
+
+%--------------------------------------------------------------------------
+% STEP 4.1
+%--------------------------------------------------------------------------
+
+\item Diffuse the enthalpy through a time interval of $\dt$.
+Compute $\kappa^{(1)}, c_p^{(1)}$, and $\xi_k^{(1)}$ from $\rho^{(1)}, T^{(1)}$, and $X_k^{(1)}$ as inputs to the equation of state.
+Note that density is constant in this step, i.e., $\rho^{(2'),\star}=\rho^{(2),\star}$.
+The update is given by
+\begin{eqnarray}
+(\rho h)^{(2),\star} &=& (\rho h)^{(2'),\star} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(2),\star} + \frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
+&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
+&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(2),\star} + \frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
+\end{eqnarray}
+which is numerically implemented as a diffusion equation for $h^{(2),\star}$,
+\begin{eqnarray}
+\left(\rho^{(2),\star} - \frac{\dt}{2}\nabla\cdot\frac{\kappa^{(1)}}{c_p^{(1)}}\nabla\right)h^{(2),\star} &=& (\rho h)^{(2'),\star} + \frac{\dt}{2}\nabla\cdot\frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
+&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
+&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(2),\star} + \frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
+\end{eqnarray}
+\item Update the temperature
+\begin{equation}
+T^{(2),\star} = T\left(\rho^{(2),\star}, h^{(2),\star}, X_k^{(2),\star}\right) {\hbox{ \rm using the equation of state.}}
+\end{equation}
 
@@
 \end{equation}
-where $ \beta_0^{\nph,\star} = ( \beta_0^n +  \beta_0^{n+1,\star}) / 2.$
+where $ \beta_0^{\nph,\star} = ( \beta_0^n +  \beta_0^{n+1,\star}) / 2.$\\
 
@@
 (\rho X_k)^{(1),\nph} \right) \right]  \enskip , \\ \nonumber \\
-(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
+(\rho h)^{(2')} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
 (\rho h)^{(1),\nph} \right) \right] \nonumber \\
@@
 + \dt \; \psi^{\nph,\star} \enskip , \\ 
-\rho^{(2)} &=& \sum_k (\rho X_k)^{(2)} \enskip , \\
-X_k^{(2)} &=& (\rho X_k)^{(2)} / \rho^{(2)} \enskip , \\
-T^{(2)} &=& T\left(\rho^{(2)}, (\rho h)^{(2)},  X_k^{(2)}\right) {\hbox{ \rm using the
-equation of state.}} 
+\rho^{(2')} &=& \sum_k (\rho X_k)^{(2)} \enskip , \\
+X_k^{(2)} &=& (\rho X_k)^{(2)} / \rho^{(2)}
 \end{eqnarray}
@@
 \end{eqnarray}
+
+%--------------------------------------------------------------------------
+% STEP 8.1
+%--------------------------------------------------------------------------
+\item Diffuse the enthalpy through a time interval of $\dt$.
+Compute $\kappa^{(2),\star}, c_p^{(2),\star}$, and $\xi_k^{(2),\star}$, from $\rho^{(2),\star}, T^{(2),\star}$, and $X_k^{(2),\star}$ as inputs to the equation of state.
+Note that density is constant in the diffusion step, i.e., $\rho^{(2')}=\rho^{(2)}$.
+The update is given by
+\begin{eqnarray}
+(\rho h)^{(2)} &=& (\rho h)^{(2')} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla h^{(2)} + \frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
+&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
+&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
+\end{eqnarray}
+which is numerically implemented as a diffusion equation for $h^{(2)}$,
+\begin{eqnarray}
+\left(\rho^{(2)} - \frac{\dt}{2}\nabla\cdot\frac{\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla\right)h^{(2)} &=& (\rho h)^{(2')} + \frac{\dt}{2}\nabla\cdot\frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
+&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
+&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
+\end{eqnarray}
+\item Update the temperature
+\begin{equation}
+T^{(2)} = T\left(\rho^{(2)}, h^{(2)}, X_k^{(2)}\right) {\hbox{ \rm using the equation of state.}}
+\end{equation}
+
 \end{enumerate}
@@
 $N_{\rm iters}^{\pi},$ in this second step of the initialization procedure.
-The details for both iterations are given below.
+The details for both iterations are given below.\\
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 0} {\em Initialization}
+\noindent {\bf Step 0.} {\em Initialization}
 

### 2008-08-15 docs/flowchart/paper.tex
kappa -> k_th

@@
 \newcommand{\Hnuc}{{H_{\rm nuc}}}
+\newcommand{\kth}{k_{\rm th}}
 
@@
 \item Diffuse the enthalpy through a time interval of $\dt$.
-Compute $\kappa^{(1)}, c_p^{(1)}$, and $\xi_k^{(1)}$ from $\rho^{(1)}, T^{(1)}$, and $X_k^{(1)}$ as inputs to the equation of state.
+Compute $\kth^{(1)}, c_p^{(1)}$, and $\xi_k^{(1)}$ from $\rho^{(1)}, T^{(1)}$, and $X_k^{(1)}$ as inputs to the equation of state.
 Note that density is constant in this step, i.e., $\rho^{(2'),\star}=\rho^{(2),\star}$.
@@
 \begin{eqnarray}
-(\rho h)^{(2),\star} &=& (\rho h)^{(2'),\star} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(2),\star} + \frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(2),\star} + \frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
+(\rho h)^{(2),\star} &=& (\rho h)^{(2'),\star} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(2),\star} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
+&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
+&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(2),\star} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
 \end{eqnarray}
@@
 \begin{eqnarray}
-\left(\rho^{(2),\star} - \frac{\dt}{2}\nabla\cdot\frac{\kappa^{(1)}}{c_p^{(1)}}\nabla\right)h^{(2),\star} &=& (\rho h)^{(2'),\star} + \frac{\dt}{2}\nabla\cdot\frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
-&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(2),\star} + \frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
+\left(\rho^{(2),\star} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla\right)h^{(2),\star} &=& (\rho h)^{(2'),\star} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
+&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
+&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(2),\star} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
 \end{eqnarray}
@@
 \item Diffuse the enthalpy through a time interval of $\dt$.
-Compute $\kappa^{(2),\star}, c_p^{(2),\star}$, and $\xi_k^{(2),\star}$, from $\rho^{(2),\star}, T^{(2),\star}$, and $X_k^{(2),\star}$ as inputs to the equation of state.
+Compute $\kth^{(2),\star}, c_p^{(2),\star}$, and $\xi_k^{(2),\star}$, from $\rho^{(2),\star}, T^{(2),\star}$, and $X_k^{(2),\star}$ as inputs to the equation of state.
 Note that density is constant in the diffusion step, i.e., $\rho^{(2')}=\rho^{(2)}$.
@@
 \begin{eqnarray}
-(\rho h)^{(2)} &=& (\rho h)^{(2')} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla h^{(2)} + \frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
+(\rho h)^{(2)} &=& (\rho h)^{(2')} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla h^{(2)} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
+&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
+&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
 \end{eqnarray}
@@
 \begin{eqnarray}
-\left(\rho^{(2)} - \frac{\dt}{2}\nabla\cdot\frac{\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla\right)h^{(2)} &=& (\rho h)^{(2')} + \frac{\dt}{2}\nabla\cdot\frac{\kappa^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
-&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kappa^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kappa^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
+\left(\rho^{(2)} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla\right)h^{(2)} &=& (\rho h)^{(2')} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
+&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
+&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
 \end{eqnarray}

### 2008-08-19 docs/initial_models/paper.tex
added some notes to myself about what needs to be done

@@
 % Margins
-\usepackage[lmargin=0.5in,rmargin=0.5in,tmargin=0.75in,bmargin=0.75in]{geometry}
+\usepackage[lmargin=0.5in,rmargin=1.5in,tmargin=0.5in,bmargin=0.5in]{geometry}
 
@@
 
-\section{Creating the initial model}
+\section{Initialization Procedure}
+\subsection{Creating the Initial Model}
+We denote the initial model returned by the code in the {\tt MAESTRO/initial\_models} directory with the superscript ``init''.
+\begin{itemize}
+\item Include a table of which file creates initial models for which examples.
+\item Indicate which variables are returned by the initial model.
+\item Describe how the initial model is developed such that the pressure and density are in hydrostatic equilibrium.
+\end{itemize}
+\subsection{Initialization Procedure}
+Given data from an initial model file (indicated with a superscript ``init''), this section describes how the base state ($\rho_0$, $p_0$) and full state are computed.
+
+\clearpage
+
+\section{Old Notes}
+\subsection{Creating the initial model}
 \begin{itemize}
@@
 $(p_{i+1} - p_i)/\Delta r = \half (\rho_{i+1} - \rho_i) g_{i+\half} \enskip .$
-
 \item The initial model should be created at the same resolution as
@@
 simulation, use a new initial model.
-
 \item The initial model should be thermodynamically consistent with
 the EOS that {\tt MAESTRO} uses.
-
 \end{itemize}
-
 \section{Spherical notes}
@@
 $(\rho h)$ is computed via the EOS.
-
 \item Spherical runs automatically set $\Delta x = \Delta r$.
 \end{itemize}
-
 \section{Running notes}
@@
 \end{itemize}
-
 \end{document}

### 2008-08-19 docs/initial_models/paper.tex
more notes, yet still lots more I need to add

--- 2008-08-14 docs/initial_models/paper.tex
@@
 \def\kth   {k_{\rm th}}
+\def\init  {\rm init}
 
@@
 
-\section{Creating the initial model}
-\begin{itemize}
-\item The initial model should be but into HSE according to 
-$(p_{i+1} - p_i)/\Delta r = \half (\rho_{i+1} - \rho_i) g_{i+\half} \enskip .$
+\section{Creating the Initial Model}
+The initial model returns the pressure, density, temperature, and species, denoted $p^{\init}, \rho^{\init}, T^{\init},$ and $X_k^{\init}$.
+The initial model uses an iterative procedure \MarginPar{add more details about the iterative procedure} to return a state that is both thermodynamically consistent with the {\tt MAESTRO} EOS, and in hydrostatic equilibrium,
+\begin{equation}
+\frac{p_r^{\init} - p_{r-1}^{\init}}{\Delta r} = \frac{\rho_r^{\init} + \rho_{r-1}^{\init}}{2}g,\label{HSE Discretization}
+\end{equation}
+to a user-defined tolerance.  Here is a list of which model file is used for each test problem:\MarginPar{This needs to be filled in and checked for accuracy}
+\begin{eqnarray}
+{\tt test2} & \rightarrow & {\tt ???} \nonumber \\
+{\tt test\_convect} & \rightarrow & {\tt ???} \nonumber \\
+{\tt wdconvect} & \rightarrow & {\tt ???} \nonumber \\
+{\tt spherical\_heat} & \rightarrow & {\tt ???} \nonumber \\
+{\tt xrb} & \rightarrow & {\tt initial\_models/xrb/init\_1d.f90} \nonumber
+\end{eqnarray}
+The initial model needs to be created at the same resolution as the base state (for single-level problems) and at the same resolution of the finest resolution base state (for multi-level plane-parallel problems).
 
-\item The initial model should be created at the same resolution as
-the base state.  This means that when you change resolution of your
-simulation, use a new initial model.
+\section{Initialization Procedure}
+Given data from an initial model file ($p^{\init}, \rho^{\init}, T^{\init},$ and $X_k^{\init}$), this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.
+At the end of the initialization, we require that $\rho_0 = \overline\rho$, the base state is in hydrostatic equilibrium according to equation (\ref{HSE Discretization}), and the full state is in thermodynamic equilibrium with $p_0$.
+Here are the steps:
+\begin{enumerate}
+\item Call the EOS to obtain a 1D $p^{\init},h^{\init} = p^{\init},h^{\init}(\rho^{\init},T^{\init},X^{\init})$.
+Note that this does overwrite the $p^{\init}$ from the initial model file, but this should not change $p^{\init}$ since this pressure was already both thermodynamically consistent and in hydrostatic equilibrium.
+\end{enumerate}
 
-\item The initial model should be thermodynamically consistent with
-the EOS that {\tt MAESTRO} uses.
+\clearpage
 
-\end{itemize}
-
+\section{Old Notes}
 \section{Spherical notes}

### 2008-08-19 docs/initial_models/paper.tex
more updates

@@
 The initial model returns the pressure, density, temperature, and species, denoted $p^{\init}, \rho^{\init}, T^{\init},$ and $X_k^{\init}$.
-The initial model uses an iterative procedure \MarginPar{add more details about the iterative procedure} to return a state that is both thermodynamically consistent with the {\tt MAESTRO} EOS, and in hydrostatic equilibrium,
+The initial model uses an iterative procedure \MarginPar{add an overview of the iterative procedure} to return a state that is both thermodynamically consistent with the {\tt MAESTRO} EOS, and in hydrostatic equilibrium,
 \begin{equation}
@@
 The initial model needs to be created at the same resolution as the base state (for single-level problems) and at the same resolution of the finest resolution base state (for multi-level plane-parallel problems).
+For spherical, we force $\Delta r = \Delta x$.
 
 \section{Initialization Procedure}
-Given data from an initial model file ($p^{\init}, \rho^{\init}, T^{\init},$ and $X_k^{\init}$), this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.
+Given data from an initial model file ($p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$), this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.
 At the end of the initialization, we require that $\rho_0 = \overline\rho$, the base state is in hydrostatic equilibrium according to equation (\ref{HSE Discretization}), and the full state is in thermodynamic equilibrium with $p_0$.
@@
 \item Call the EOS to obtain a 1D $p^{\init},h^{\init} = p^{\init},h^{\init}(\rho^{\init},T^{\init},X^{\init})$.
-Note that this does overwrite the $p^{\init}$ from the initial model file, but this should not change $p^{\init}$ since this pressure was already both thermodynamically consistent and in hydrostatic equilibrium.
+Note that this overwrites the $p^{\init}$ from the initial model file, but this should not change $p^{\init}$ since this pressure was already both thermodynamically consistent and in hydrostatic equilibrium.
+\item For plane-parallel problems, fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.
+For spherical problems, fill $\rho^{\init}, p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state $\rho, p$, and $X$, and then call the EOS to obtain the full state $h,T = h,T(\rho,p,X)$.
+\item For spherical problems, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
+\item If {\tt perturb\_model} = F, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
+\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
+\item Set $\rho_0 = \overline\rho$.
+\item Compute $p_0$ using equation (\ref{HSE Discretization}).
+\item Compute the full state $T,h = T,h(\rho,p_0,X)$.
 \end{enumerate}
-
-\clearpage
-
-\section{Old Notes}
-\section{Spherical notes}
-\begin{itemize}
-\item Spherical runs use $p$, $(\rho X)$, and $\rho$ from the initial
-model.  These are then mapped onto the Cartesian grid, and then
-$(\rho h)$ is computed via the EOS.
-\item Spherical runs automatically set $\Delta x = \Delta r$.
-\end{itemize}
-\section{Running notes}
-\begin{itemize}
-\item When the code is run, it prints out the HSE error of the base
-state model---this should be checked to insure that we a
-\end{itemize}
+Now the full state is in thermodynamic equilibrium with $p_0$, and the base state is in hydrostatic equilibrium.
 \end{document}

### 2008-08-19 docs/initial_models/paper.tex
changed the initial model for the xrb problem to initial_models/xrb/sorted_xrb.raw
initial_models/xrb/init_1d.f90 is the routine which uses the initial model file for mapping

@@
 {\tt spherical\_heat} & \rightarrow & {\tt ???} \nonumber \\
-{\tt xrb} & \rightarrow & {\tt initial\_models/xrb/init\_1d.f90} \nonumber
+{\tt xrb} & \rightarrow & {\tt initial\_models/xrb/sorted\_xrb.raw} \nonumber
 \end{eqnarray}

### 2008-08-19 docs/initial_models/paper.tex
updated - this is what I'm going to implement in the code

@@
 
-\section{Creating the Initial Model}
-The initial model returns the pressure, density, temperature, and species, denoted $p^{\init}, \rho^{\init}, T^{\init},$ and $X_k^{\init}$.
-The initial model uses an iterative procedure \MarginPar{add an overview of the iterative procedure} to return a state that is both thermodynamically consistent with the {\tt MAESTRO} EOS, and in hydrostatic equilibrium,
-\begin{equation}
-\frac{p_r^{\init} - p_{r-1}^{\init}}{\Delta r} = \frac{\rho_r^{\init} + \rho_{r-1}^{\init}}{2}g,\label{HSE Discretization}
-\end{equation}
-to a user-defined tolerance.  Here is a list of which model file is used for each test problem:\MarginPar{This needs to be filled in and checked for accuracy}
+\section{Creating the Initial Data}
+We begin with data from a model file, which we shall refer to as the ``model'' data.  Here is a list of the model data for each test problem:
 \begin{eqnarray}
@@
 \end{eqnarray}
-The initial model needs to be created at the same resolution as the base state (for single-level problems) and at the same resolution of the finest resolution base state (for multi-level plane-parallel problems).
-For spherical, we force $\Delta r = \Delta x$.
+We use a fortran subroutine to modify the model data to create ``initial'' data, denoted with the subcript ``init''.
+The model data needs to be modified because there is no guarantee that it was generated with the same equation of state (EOS) that {\tt MAESTRO} uses, nor is there any guarantee that the pressure and density are in hydrostatic equilibrium (HSE) according to our choice of discretization given in equation (\ref{HSE Discretization}).
+Here is a list of the fortran subroutine for each test problem:
+\begin{eqnarray}
+{\tt test2} & \rightarrow & {\tt ???} \nonumber \\
+{\tt test\_convect} & \rightarrow & {\tt ???} \nonumber \\
+{\tt wdconvect} & \rightarrow & {\tt ???} \nonumber \\
+{\tt spherical\_heat} & \rightarrow & {\tt ???} \nonumber \\
+{\tt xrb} & \rightarrow & {\tt initial\_models/xrb/init\_1d.f90} \nonumber
+\end{eqnarray}
+The fortran subroutine returns $p^{\init}, \rho^{\init}, T^{\init},$ and $X_k^{\init}$.
+The fortran subroutine uses an iterative procedure to return initial data that is both thermodynamically consistent with the {\tt MAESTRO} EOS, and in HSE,
+\begin{equation}
+\frac{p_r^{\init} - p_{r-1}^{\init}}{\Delta r} = \frac{\rho_r^{\init} + \rho_{r-1}^{\init}}{2}g,\label{HSE Discretization}
+\end{equation}
+to a user-defined tolerance.
 
-\section{Initialization Procedure}
-Given data from an initial model file ($p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$), this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.
-At the end of the initialization, we require that $\rho_0 = \overline\rho$, the base state is in hydrostatic equilibrium according to equation (\ref{HSE Discretization}), and the full state is in thermodynamic equilibrium with $p_0$.
+The initial data needs be generated at the same resolution as the base state for single-level problems and at the same resolution of the finest resolution base state for multi-level plane-parallel problems.
+For spherical problems, we force $\Delta r = \Delta x$.
+\section{Creating the Base State and Full State}
+Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.
+At the end of the initialization, we require that $\rho_0 = \overline\rho$, the base state is in HSE according to equation (\ref{HSE Discretization}), and the full state is in thermodynamic equilibrium with $p_0$.
 Here are the steps:
 \begin{enumerate}
-\item Call the EOS to obtain a 1D $p^{\init},h^{\init} = p^{\init},h^{\init}(\rho^{\init},T^{\init},X^{\init})$.
-Note that this overwrites the $p^{\init}$ from the initial model file, but this should not change $p^{\init}$ since this pressure was already both thermodynamically consistent and in hydrostatic equilibrium.
+\item Call the EOS to obtain a 1D $h^{\init} = h^{\init}(\rho^{\init},T^{\init},X^{\init})$.
+Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\init} = p^{\init}(\rho^{\init},T^{\init},X^{\init})$ at the end of the previous fortran subroutine.
 \item For plane-parallel problems, fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.
@@
 \end{enumerate}
-Now the full state is in thermodynamic equilibrium with $p_0$, and the base state is in hydrostatic equilibrium.
+Now $\rho_0 = {\rm Avg}(\rho)$, the base state is in HSE, and the full state is in thermodynamic equilibrium with $p_0$.
 \end{document}

### 2008-08-19 docs/initial_models/paper.tex
minor edits

@@
 Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.
-At the end of the initialization, we require that $\rho_0 = \overline\rho$, the base state is in HSE according to equation (\ref{HSE Discretization}), and the full state is in thermodynamic equilibrium with $p_0$.
+We require that $\rho_0 = \overline\rho$, the base state is in HSE according to equation (\ref{HSE Discretization}), and the full state is in thermodynamic equilibrium with $p_0$.
 Here are the steps:
@@
 \end{enumerate}
-Now $\rho_0 = {\rm Avg}(\rho)$, the base state is in HSE, and the full state is in thermodynamic equilibrium with $p_0$.
+Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is in thermodynamic equilibrium with $p_0$.
 \end{document}

### 2008-08-19 docs/initial_models/paper.tex
this version I think is much more clear

@@
 \def\init  {\rm init}
+\def\model {\rm model}
+\def\raw   {\rm raw}
 
@@
 
-\section{Creating the Initial Data}
-We begin with data from a model file, which we shall refer to as the ``model'' data.  Here is a list of the model data for each test problem:
+\section{Creating the Model Data from Raw Data}
+We begin with raw data, $\rho^{\raw}, T^{\raw}, p^{\raw}$, and $X^{\raw}$.  Here is the raw data file for each test problem:
 \begin{eqnarray}
@@
 \end{eqnarray}
-We use a fortran subroutine to modify the model data to create ``initial'' data, denoted with the subcript ``init''.
-The model data needs to be modified because there is no guarantee that it was generated with the same equation of state (EOS) that {\tt MAESTRO} uses, nor is there any guarantee that the pressure and density are in hydrostatic equilibrium (HSE) according to our choice of discretization given in equation (\ref{HSE Discretization}).
-Here is a list of the fortran subroutine for each test problem:
+Using the raw data, we need to generate model data at the same resolution as the base state.
+Note that for spherical problems we set $\Delta r = \Delta x$.
+We use a fortran subroutine to interpolate the raw data, yielding the model data, $\rho^{\model}, T^{\model}, p^{\model}$, and $X^{\model}$.
+Then, we modify the model data to ensure that it is thermodynamically consistent with the {\tt MAESTRO} equation of state (EOS), and also satisfies our chosen hydrostatic equilibrium (HSE) discretization,
+\begin{equation}
+\frac{p_r - p_{r-1}}{\Delta r} = \frac{\rho_r + \rho_{r-1}}{2}g,\label{HSE Discretization}
+\end{equation}
+to a user defined tolerance using an iterative procedure.  Here are the fortran subroutines for each test problem:
 \begin{eqnarray}
@@
 \end{eqnarray}
-The fortran subroutine returns $p^{\init}, \rho^{\init}, T^{\init},$ and $X_k^{\init}$.
-The fortran subroutine uses an iterative procedure to return initial data that is both thermodynamically consistent with the {\tt MAESTRO} EOS, and in HSE,
-\begin{equation}
-\frac{p_r^{\init} - p_{r-1}^{\init}}{\Delta r} = \frac{\rho_r^{\init} + \rho_{r-1}^{\init}}{2}g,\label{HSE Discretization}
-\end{equation}
-to a user-defined tolerance.
-
-The initial data needs be generated at the same resolution as the base state for single-level problems and at the same resolution of the finest resolution base state for multi-level plane-parallel problems.
-For spherical problems, we force $\Delta r = \Delta x$.
+\section{Creating the Initial Data from the Model Data}
+In {\tt initdata.f90}, we set the initial data $\rho^{\init}, T^{\init}, p^{\init}$ and $X^{\init}$ equal to the model data.
+Then, we set $h^{\init} = h^{\init}(\rho^{\init},T^{\init},X^{\init})$.
+Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\model} = p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the previous fortran subroutine.
 \section{Creating the Base State and Full State}
@@
 Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is in thermodynamic equilibrium with $p_0$.
+\subsection{Multilevel Case}
+Again, we are given 
 \end{document}

### 2008-08-19 docs/initial_models/paper.tex
more details.  this is complicated - and I haven't even gotten to multilevel yet

@@
 We use a fortran subroutine to interpolate the raw data, yielding the model data, $\rho^{\model}, T^{\model}, p^{\model}$, and $X^{\model}$.
-Then, we modify the model data to ensure that it is thermodynamically consistent with the {\tt MAESTRO} equation of state (EOS), and also satisfies our chosen hydrostatic equilibrium (HSE) discretization,
+The fortran subroutine then uses an iterative procedure to modify the model data so that it is thermodynamically consistent with the {\tt MAESTRO} equation of state (EOS), and also satisfies our chosen hydrostatic equilibrium (HSE) discretization,
 \begin{equation}
@@
 \end{equation}
-to a user defined tolerance using an iterative procedure.  Here are the fortran subroutines for each test problem:
+to a user defined tolerance.  Here are the fortran subroutines for each test problem:
 \begin{eqnarray}
@@
 \end{eqnarray}
+The model data is not generated at run-time - it must be generated in advance of running any MAESTRO examples.  The inputs file points to the file containing the model data.
 \section{Creating the Initial Data from the Model Data}
-In {\tt initdata.f90}, we set the initial data $\rho^{\init}, T^{\init}, p^{\init}$ and $X^{\init}$ equal to the model data.
+In {\tt base\_state.f90}, using the subroutine {\tt init\_base\_state.f90} (which is actually a terrible name since we are setting the initial data, not the base state), we set the initial data $\rho^{\init}, T^{\init}, p^{\init}$ and $X^{\init}$ equal to the model data.
 Then, we set $h^{\init} = h^{\init}(\rho^{\init},T^{\init},X^{\init})$.

### 2008-08-19 docs/initial_models/paper.tex
OK I think this is *finally* right for single level problems

@@
 \end{eqnarray}
-Using the raw data, we need to generate model data at the same resolution as the base state.
+Using the raw data, we need to generate model data with the same resolution as the base state.
 Note that for spherical problems we set $\Delta r = \Delta x$.
@@
 \end{eqnarray}
-The model data is not generated at run-time - it must be generated in advance of running any MAESTRO examples.  The inputs file points to the file containing the model data.
+The model data is not generated at run-time - it must be generated in advance of running any MAESTRO examples.  The inputs file should point to the file containing the model data.
+\subsection{Multilevel Plane-Parallel}
+We need to generate model data with the same resolution as the finest base state resolution.
 \section{Creating the Initial Data from the Model Data}
@@
 \begin{enumerate}
-\item Call the EOS to obtain a 1D $h^{\init} = h^{\init}(\rho^{\init},T^{\init},X^{\init})$.
-Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\init} = p^{\init}(\rho^{\init},T^{\init},X^{\init})$ at the end of the previous fortran subroutine.
 \item For plane-parallel problems, fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.
 For spherical problems, fill $\rho^{\init}, p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state $\rho, p$, and $X$, and then call the EOS to obtain the full state $h,T = h,T(\rho,p,X)$.
+Note that the multifab version of $p$ is temporary, and can be deallocated after this step.
 \item For spherical problems, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
 \item If {\tt perturb\_model} = F, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
-\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
+\item Perturb $T$, then compute $\rho = \rho(p^{\init},T,X)$.
 \item Set $\rho_0 = \overline\rho$.
-\item Compute $p_0$ using equation (\ref{HSE Discretization}).
+\item Compute $p_0$ using equation (\ref{HSE Discretization}), using $p_{0,r=0}$ as the initial condition.
 \item Compute the full state $T,h = T,h(\rho,p_0,X)$.
@@
 Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is in thermodynamic equilibrium with $p_0$.
-\subsection{Multilevel Case}
-Again, we are given 
 \end{document}

### 2008-08-20 docs/initial_models/paper.tex
added some multilevel notes

@@
 Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\model} = p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the previous fortran subroutine.
+\subsection{Multilevel Plane-Parallel}
+For the finest level, we can set the initial data equal to the model data.
+For the non-finest levels, we use linear interpolation to compute the initial data.
+For cells that are covered by finer cells, we average $\rho^{\init}, (\rho h)^{\init},$ and $(\rho X)^{\init}$ from the fine cell to get the coarse cell value.
+Then, for every cell not at the finest level, we call the EOS to compute $p^{\init},T^{\init} = p^{\init},T^{\init}(\rho^{\init},h^{\init},X^{\init})$.\MarginPar{Now $p^{\init}$ and $\rho^{\init}$ are not in HSE.
+Is this a problem?}
 \section{Creating the Base State and Full State}
@@
 Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is in thermodynamic equilibrium with $p_0$.
+\subsection{Multilevel Plane-Parallel}
 \end{document}

### 2008-08-20 docs/eta_notes/paper.tex
there are so many etas floating around I need to keep them straight

--- 2008-08-19 docs/initial_models/paper.tex
@@
 
-\def\half  {\frac{1}{2}}
-\def\dt    {\Delta t}
-\def\kth   {k_{\rm th}}
-\def\init  {\rm init}
-\def\model {\rm model}
-\def\raw   {\rm raw}
+\def\eb        {\bf{e}}
 
-\def\edge  {\rm EDGE}
-\def\mac   {\rm MAC}
-\def\trans {\rm TRANS}
+\def\etarho    {\eta_\rho}
+\def\etarhocc  {\etarho^{\rm cc}}
+\def\divetarho {\nabla\cdot(\etarho\eb_r)}
 
-\def\eb    {{\bf e}}
-\def\fb    {{\bf f}}
-\def\ib    {{\bf i}}
-\def\Ub    {{\bf U}}
-
-\def\Ubt   {\widetilde{\bf U}}
-\def\ut    {\tilde{u}}
-\def\vt    {\tilde{v}}
-\def\wt    {\tilde{w}}
-
-\title{Notes on generating initial models for mapping into {\tt MAESTRO}}
+\title{Notes on $\etarho$}
 
@@
 \maketitle
-\tableofcontents
 
-\clearpage
+We carry around three different quantities: $\etarho$ (edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$ (edge-centered).
+These notes discuss when each of these is used, and how they are computed, in both plane-parallel and spherical.
 
-\section{Creating the Model Data from Raw Data}
-We begin with raw data, $\rho^{\raw}, T^{\raw}, p^{\raw}$, and $X^{\raw}$.  Here is the raw data file for each test problem:
-\begin{eqnarray}
-{\tt test2} & \rightarrow & {\tt ???} \nonumber \\
-{\tt test\_convect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt wdconvect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt spherical\_heat} & \rightarrow & {\tt ???} \nonumber \\
-{\tt xrb} & \rightarrow & {\tt initial\_models/xrb/sorted\_xrb.raw} \nonumber
-\end{eqnarray}
-Using the raw data, we need to generate model data with the same resolution as the base state.
-Note that for spherical problems we set $\Delta r = \Delta x$.
-We use a fortran subroutine to interpolate the raw data, yielding the model data, $\rho^{\model}, T^{\model}, p^{\model}$, and $X^{\model}$.
-The fortran subroutine then uses an iterative procedure to modify the model data so that it is thermodynamically consistent with the {\tt MAESTRO} equation of state (EOS), and also satisfies our chosen hydrostatic equilibrium (HSE) discretization,
-\begin{equation}
-\frac{p_r - p_{r-1}}{\Delta r} = \frac{\rho_r + \rho_{r-1}}{2}g,\label{HSE Discretization}
-\end{equation}
-to a user defined tolerance.  Here are the fortran subroutines for each test problem:
-\begin{eqnarray}
-{\tt test2} & \rightarrow & {\tt ???} \nonumber \\
-{\tt test\_convect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt wdconvect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt spherical\_heat} & \rightarrow & {\tt ???} \nonumber \\
-{\tt xrb} & \rightarrow & {\tt initial\_models/xrb/init\_1d.f90} \nonumber
-\end{eqnarray}
-The model data is not generated at run-time - it must be generated in advance of running any MAESTRO examples.  The inputs file should point to the file containing the model data.
-\subsection{Multilevel Plane-Parallel}
-We need to generate model data with the same resolution as the finest base state resolution.
-\section{Creating the Initial Data from the Model Data}
-In {\tt base\_state.f90}, using the subroutine {\tt init\_base\_state.f90} (which is actually a terrible name since we are setting the initial data, not the base state), we set the initial data $\rho^{\init}, T^{\init}, p^{\init}$ and $X^{\init}$ equal to the model data.
-Then, we set $h^{\init} = h^{\init}(\rho^{\init},T^{\init},X^{\init})$.
-Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\model} = p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the previous fortran subroutine.
-\section{Creating the Base State and Full State}
-Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.
-We require that $\rho_0 = \overline\rho$, the base state is in HSE according to equation (\ref{HSE Discretization}), and the full state is in thermodynamic equilibrium with $p_0$.
-Here are the steps:
-\begin{enumerate}
-\item For plane-parallel problems, fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.
-For spherical problems, fill $\rho^{\init}, p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state $\rho, p$, and $X$, and then call the EOS to obtain the full state $h,T = h,T(\rho,p,X)$.
-Note that the multifab version of $p$ is temporary, and can be deallocated after this step.
-\item For spherical problems, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
-\item If {\tt perturb\_model} = F, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
-\item Perturb $T$, then compute $\rho = \rho(p^{\init},T,X)$.
-\item Set $\rho_0 = \overline\rho$.
-\item Compute $p_0$ using equation (\ref{HSE Discretization}), using $p_{0,r=0}$ as the initial condition.
-\item Compute the full state $T,h = T,h(\rho,p_0,X)$.
-\end{enumerate}
-Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is in thermodynamic equilibrium with $p_0$.
 \end{document}

### 2008-08-20 docs/eta_notes/paper.tex
Cartesian make_eta done.

@@
 
-\def\eb        {\bf{e}}
+\def\half   {\frac{1}{2}}
+
+\def\eb         {{\bf e}}
+\def\ib         {{\bf i}}
+\def\Ub         {{\bf U}}
 
@@
 We carry around three different quantities: $\etarho$ (edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$ (edge-centered).
+These are each 1D in the radial direction.
 These notes discuss when each of these is used, and how they are computed, in both plane-parallel and spherical.
 
+\section{Computing $\etarho, \etarhocc$, and $\divetarho$}
+This is done in {\tt make\_eta.f90}.
+\subsection{Plane-Parallel}
+We first compute a radial edge-centered multifab, $\eta_{\rho}^{\rm flux}$, using
+\begin{equation}
+\eta_{\rho,\ib+\half\eb_r}^{\rm flux} = \Ub_{\ib+\half\eb_r}^{n+\half} \rho_{\ib+\half\eb_r}^{n+\half} - w_{0,r+\half}^{n+\half}\rho_{0,r+\half}^{n+\half, {\rm pred}}
+\end{equation}
+$\etarho$ is the edge-centered ``average'' value of $\eta_{\rho}^{\rm flux}$,
+\begin{equation}
+\eta_{\rho,r+\half} = \overline{\eta_{\rho,\ib+\half\eb_r}^{\rm flux}}
+\end{equation}
+$\etarhocc$ is a cell-centered average of $\etarho$,
+\begin{equation}
+\eta_{\rho,r}^{\rm cc} = \frac{\eta_{\rho,r+\half} + \eta_{\rho,r-\half}}{2}.
+\end{equation}
+$\divetarho$ is the cell-centered divergence of $\etarho$,
+\begin{equation}
+\left[\divetarho\right]_r = \frac{\eta_{\rho,r+\half} - \eta_{\rho,r-\half}}{\Delta r}.
+\end{equation}
+\subsection{Spherical}
+
 \end{document}

### 2008-08-20 Source/advance.f90
passing in etarho_cc instead of etarho into make_psi.  The only thing we were doing inside of make_psi was average etarho to centers, which is exactly what etarho_cc is.  no change in results

@@ -592,7 +592,7 @@ contains
 
        endif
 
-       call make_psi(nlevs,etarho,psi,w0,gamma1bar,p0_old,p0_new,Sbar)
+       call make_psi(nlevs,etarho_cc,psi,w0,gamma1bar,p0_old,p0_new,Sbar)
     end if
 
 
@@ -1027,7 +1027,7 @@ contains
 
           endif
 
-          call make_psi(nlevs,etarho,psi,w0,gamma1bar,p0_old,p0_new,Sbar)
+          call make_psi(nlevs,etarho_cc,psi,w0,gamma1bar,p0_old,p0_new,Sbar)
        end if
 
        do n=1,nlevs

### 2008-08-20 Source/make_psi.f90

@@ -9,13 +9,13 @@ module make_psi_module
 
 contains
 
-  subroutine make_psi(nlevs,etarho,psi,w0,gamma1bar,p0_old,p0_new,Sbar_in)
+  subroutine make_psi(nlevs,etarho_cc,psi,w0,gamma1bar,p0_old,p0_new,Sbar_in)
 
     use bl_prof_module
     use geometry, only: spherical
 
     integer        , intent(in   ) :: nlevs
-    real(kind=dp_t), intent(in   ) :: etarho(:,0:)
+    real(kind=dp_t), intent(in   ) :: etarho_cc(:,0:)
     real(kind=dp_t), intent(inout) :: psi(:,0:)
     real(kind=dp_t), intent(in   ) :: w0(:,0:)
     real(kind=dp_t), intent(in   ) :: gamma1bar(:,0:)
@@ -31,7 +31,7 @@ contains
     
     if (spherical .eq. 0) then 
        do n = 1,nlevs
-          call make_psi_planar(n,etarho(n,0:),psi(n,0:))
+          call make_psi_planar(n,etarho_cc(n,0:),psi(n,0:))
        end do
     else
        do n = 1,nlevs
@@ -44,27 +44,25 @@ contains
        
   end subroutine make_psi
 
-  subroutine make_psi_planar(n,etarho,psi)
+  subroutine make_psi_planar(n,etarho_cc,psi)
 
     use bl_constants_module
     use geometry, only: anelastic_cutoff_coord, r_start_coord, r_end_coord, numdisjointchunks
     use probin_module, only: grav_const
 
     integer        , intent(in   ) :: n
-    real(kind=dp_t), intent(in   ) :: etarho(0:)
+    real(kind=dp_t), intent(in   ) :: etarho_cc(0:)
     real(kind=dp_t), intent(inout) :: psi(0:)
     
     ! Local variables
     integer         :: r,i
-    real(kind=dp_t) :: etarho_avg
    
     psi = ZERO
 
     do i=1,numdisjointchunks(n)
        do r = r_start_coord(n,i), r_end_coord(n,i)
           if (r .lt. anelastic_cutoff_coord(n)) then
-             etarho_avg = HALF * (etarho(r)+etarho(r+1))
-             psi(r) = etarho_avg * abs(grav_const)
+             psi(r) = etarho_cc(r) * abs(grav_const)
           end if
        end do
     end do

### 2008-08-20 docs/eta_notes/paper.tex
added a flow chart of when eta quantities are created and used

@@
 \def\etarhocc   {\etarho^{\rm cc}}
+\def\etarhoflux {\etarho^{\rm flux}}
 \def\divetarho  {\nabla\cdot(\etarho\eb_r)}
@@
 \maketitle
+\tableofcontents
+\clearpage
 
-We carry around three different quantities: $\etarho$ (edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$ (edge-centered).
-These are each 1D in the radial direction.
+We carry around three different 1D radial quantities: $\etarho$ (edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$ (edge-centered).
 These notes discuss when each of these is used, and how they are computed, in both plane-parallel and spherical.
+
+\section{Flow Chart}
+\begin{enumerate}
+\item Enter {\tt advance} with $[\etarho, \etarhocc, \divetarho]^{n-\half}$.
+\item Call {\tt make\_w0}.  The spherical version uses uses $\etarho$ and $\etarhocc$.
+\item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
+\item Call {\tt correct\_base}, which uses $\divetarho$.
+\item Compute $[\etarho, \etarhocc, \divetarho]^{n+\half,*}$.
+\item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
+\item Call {\tt make\_w0}.  The spherical version uses uses $\etarho$ and $\etarhocc$.
+\item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
+\item Call {\tt correct\_base}, which uses $\divetarho$.
+\item Compute $[\etarho, \etarhocc, \divetarho]^{n+\half}$.
+\item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
+\end{enumerate}
 
@@
 \subsection{Spherical}
+\section{Using $\etarho$}
+\subsection{Plane-Parallel}
+NOT USED.
+\subsection{Spherical}
+In {\tt make\_w0}
+\section{Using $\etarhocc$}
+\subsection{Plane-Parallel}
+In {\tt make\_psi}
+\subsection{Spherical}
+In {\tt make\_w0}
+\section{Using $\divetarho$}
+\subsection{Plane-Parallel}
+In {\tt correct\_base}
+\subsection{Spherical}
+In {\tt correct\_base}
 

### 2008-08-20 docs/eta_notes/paper.tex
more notes on spherical
fixed typo
Finished the usage details for each of the eta quantitites.  I think this document is complete.

@@
 \def\Ub         {{\bf U}}
+
+\newcommand{\Ubt}{\widetilde{\Ub}}
 
@@
 \begin{equation}
-\eta_{\rho,\ib+\half\eb_r}^{\rm flux} = \Ub_{\ib+\half\eb_r}^{n+\half} \rho_{\ib+\half\eb_r}^{n+\half} - w_{0,r+\half}^{n+\half}\rho_{0,r+\half}^{n+\half, {\rm pred}}
+\eta_{\rho,\ib+\half\eb_r}^{\rm flux} = \left[\left(\Ubt_{\ib+\half\eb_r}^{n+\half}\cdot\eb_r\right) + w_{0,r+\half}^{n+\half}\right] \rho_{\ib+\half\eb_r}^{n+\half} - w_{0,r+\half}^{n+\half}\rho_{0,r+\half}^{n+\half, {\rm pred}}
 \end{equation}
@@
 \subsection{Spherical}
+First, construct $\eta_{\rho}^{\rm cart} = [\rho'(\Ubt\cdot\eb_r)]^{n+\half}$.  Then, $\etarhocc$ is the cell-centered average of $\eta_{\rho}^{\rm cart}$,
+\begin{equation}
+\etarhocc = \overline{\eta_{\rho}^{\rm cart}}.
+\end{equation}
+On interior faces, $\etarho$ is the average of $\etarhocc$,
+\begin{equation}
+\eta_{\rho,r+\half} = \frac{\eta_{\rho,r}^{\rm cc} + \eta_{\rho,r+1}^{\rm cc}}{2}.
+\end{equation}
+At the upper and lower boundaries, we use
+\begin{eqnarray}
+\eta_{\rho,0} &=& 0, \\
+\eta_{\rho,{\rm nr}} &=& \eta_{\rho,{\rm nr}-1}^{\rm cc}.
+\end{eqnarray}
+$\divetarho$ is the cell-centered average of $\rho'^{n+1}$,
+\begin{equation}
+\divetarho = \overline{\rho'^{n+1}}
+\end{equation}
 \section{Using $\etarho$}
@@
 \subsection{Spherical}
-In {\tt make\_w0}
+In {\tt make\_w0}, $\etarho$ is used in the construction of the RHS for the $\delta w_0$ equation.
 \section{Using $\etarhocc$}
 \subsection{Plane-Parallel}
-In {\tt make\_psi}
+In {\tt make\_psi}, $\psi = \etarho g$.
 \subsection{Spherical}
-In {\tt make\_w0}
+In {\tt make\_w0} $\etarhocc$ is used in the construction of the RHS for the $\delta w_0$ equation.
 \section{Using $\divetarho$}
 \subsection{Plane-Parallel}
-In {\tt correct\_base}
+In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$.
 \subsection{Spherical}
-In {\tt correct\_base}
+In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$.
 

### 2008-08-20 docs/eta_notes/paper.tex
fix the spherical sections

@@
 \item Enter {\tt advance} with $[\etarho, \etarhocc, \divetarho]^{n-\half}$.
-\item Call {\tt make\_w0}.  The spherical version uses uses $\etarho$ and $\etarhocc$.
+\item Call {\tt make\_w0}.  The spherical version uses uses $\etarho$ and $\divetarho$.
 \item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
@@
 \item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
-\item Call {\tt make\_w0}.  The spherical version uses uses $\etarho$ and $\etarhocc$.
+\item Call {\tt make\_w0}.  The spherical version uses uses $\etarho$ and $\divetarho$.
 \item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
@@
 This is done in {\tt make\_eta.f90}.
+
 \subsection{Plane-Parallel}
@@
 \end{equation}
+
 \subsection{Spherical}
@@
 \end{eqnarray}
-$\divetarho$ is the cell-centered average of $\rho'^{n+1}$,
+$\divetarho$ is constructed from the cell-centered average of $\rho'^{n+1}$ as
 \begin{equation}
-\divetarho = \overline{\rho'^{n+1}}
+\divetarho = -\overline{\rho'^{n+1}}/\Delta t
 \end{equation}
+
 \section{Using $\etarho$}
+
 \subsection{Plane-Parallel}
 NOT USED.
+
 \subsection{Spherical}
 In {\tt make\_w0}, $\etarho$ is used in the construction of the RHS for the $\delta w_0$ equation.
+
 \section{Using $\etarhocc$}
+
 \subsection{Plane-Parallel}
 In {\tt make\_psi}, $\psi = \etarho g$.
+
 \subsection{Spherical}
-In {\tt make\_w0} $\etarhocc$ is used in the construction of the RHS for the $\delta w_0$ equation.
+NOT USED.
+
 \section{Using $\divetarho$}
+
 \subsection{Plane-Parallel}
 In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$.
+
 \subsection{Spherical}
-In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$.
-
+In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$. \\
+In {\tt make\_w0}, we average $\divetarho$ onto edges for the $\delta w_0$ equation.
 \end{document}

### 2008-08-22 docs/initial_models/paper.tex
derived a coarse-fine interface HSE stencil
added graphic of multilevel init

@@
 
-\usepackage{amsmath,color}
+\usepackage{amsmath,color,epsfig,epsf}
 
@@
 
+\newcommand{\sfrac}[2]{\mathchoice
+  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
+   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
+  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
+   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
+  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
+   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
+  {#1\!/#2}}
+
 \def\half  {\frac{1}{2}}
+\newcommand{\myhalf}{\sfrac{1}{2}}
+
 \def\dt    {\Delta t}
@@
 \subsection{Multilevel Plane-Parallel}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[tpb]
+\begin{center}
+\epsffile{./multilevel_init.eps}
+\caption{Multilevel Initialization}
+\label{Fig:Multilevel Initialization}
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Integrating the HSE discretization upward, when we encounter a change from coarse to fine, here is an equation to integrate up to the coarse-fine interface
+\begin{equation}
+\frac{p_{r-\myhalf}^l - p_{\sfrac{r}{2}-1}^{l-1}}{\Delta r^{l-1}/2} = \frac{\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}}{2}g ~~~ \rightarrow ~~~ p_{r-\myhalf}^l = p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^{l-1} g}{4}\left(\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}\right).
+\end{equation}
+Here is an equation to integrate from the coarse-fine interface up to the cell center
+\begin{equation}
+\frac{p_r^l - p_{r-\myhalf}^l}{\Delta r^l/2} = \frac{\rho_r^l + \rho_{r-\myhalf}^l}{2}g ~~~ \rightarrow ~~~ p_r^l = p_{r-\myhalf}^l + \frac{\Delta r^l g}{4}\left(\rho_r^l + \rho_{r-\myhalf}^l\right).
+\end{equation}
+Combining equations gives
+\begin{equation}
+p_r^l = p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^{l-1} g}{4}\left(\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}\right) + \frac{\Delta r^l g}{4}\left(\rho_r^l + \rho_{r-\myhalf}^l\right).
+\end{equation}
+We can simplify using
+\begin{equation}
+\Delta r^{l-1} = 2\Delta r^l,
+\end{equation}
+\begin{equation}
+\rho_{r-\myhalf}^l = \frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}.
+\end{equation}
+Simplifying
+\begin{eqnarray}
+p_r^l &=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^l g}{2}\left(\frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_{\sfrac{r}{2}-1}^{l-1}\right) + \frac{\Delta r^l g}{4}\left(\rho_r^l + \frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}\right) \nonumber \\
+&=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{3\Delta r^l g}{4}\left(\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_r^l\right).
+\end{eqnarray}
+When we encounter a change from fine to coarse, analogously
+\begin{equation}
+p_{r}^l = p_{2(r-1)+1}^{l+1} + \frac{3\Delta r^{l+1}g}{4}\left(\rho_{2(r-1)+1}^{l+1} + \rho_{r}^l\right).
+\end{equation}
+
 \end{document}

### 2008-08-22 docs/initial_models/paper.tex
Added multilevel plane-parallel notes for creating the base state and full state

@@
 Is this a problem?}
-\section{Creating the Base State and Full State}
+\section{Creating the Base State and Full State}\label{Sec:Creating the Base State and Full State}
 Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.
@@
 \item Set $\rho_0 = \overline\rho$.
-\item Compute $p_0$ using equation (\ref{HSE Discretization}), using $p_{0,r=0}$ as the initial condition.
+\item Compute $p_0$ using equation (\ref{HSE Discretization}), using $p_{0,r=0}$ as the initial condition.\MarginPar{We may change the location of the initial condition.}
 \item Compute the full state $T,h = T,h(\rho,p_0,X)$.
@@
 \subsection{Multilevel Plane-Parallel}
+Refer to figure \ref{Fig:Multilevel Initialization}.
+The dots represent the initial data, which was generated at the finest resolution on the domain, and is in HSE as well as thermodynamic equilibrium.
+The circles represent the fine level data points in the 1D array.
+The squares represent the coarse level data points in the 1D array.
+We modify step 1 in \S~\ref{Sec:Creating the Base State and Full State} as follows:
+\begin{enumerate}
+\item Step 1
+\begin{enumerate}
+\item Copy the initial data from the dots directly into the circles.  For the squares, compute only $\rho$ and $X$ by averaging $\rho$ and $(\rho X)$ from the two corresponding dots.
+\item Integrate to get $p$ in the squares.
+We need a fancy stencil at coarse-fine interfaces that I haven't finished deriving.
+Away from coarse-fine interfaces, we can use our standard HSE discretization.
+\item Call EOS to get $T$ and $h$ in the squares, i.e., $T,h = Th,h(\rho,p,X)$.
+\item We now copy this 1D profile onto the full multifab.
+\end{enumerate}
+\end{enumerate}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\subsubsection{Working Notes on Coarse-Fine HSE Discretization - Not Final}
 Integrating the HSE discretization upward, when we encounter a change from coarse to fine, here is an equation to integrate up to the coarse-fine interface

### 2008-08-22 docs/initial_models/paper.tex
instead of averaging model points to get a coarse value, I changed the text to say linearly interpolate between the two closest points.  This is more accurate for problems with 3+ levels
rewrote the fine -> coarse HSE pressure stencil to look more like how the code will be
finished notes on coarse-fine HSE discretizations

@@
 \begin{enumerate}
-\item Copy the initial data from the dots directly into the circles.  For the squares, compute only $\rho$ and $X$ by averaging $\rho$ and $(\rho X)$ from the two corresponding dots.
+\item Copy the initial data from the dots directly into the circles.
+For the squares, compute only $\rho$ and $X$ by using linear interpolation from the two nearest corresponding dots.\MarginPar{Maybe we should use quadratic interpolation?}
 \item Integrate to get $p$ in the squares.
-We need a fancy stencil at coarse-fine interfaces that I haven't finished deriving.
+We need the fancy stencils at coarse-fine interfaces in equations (\ref{Coarse-Fine Stencil}) and (\ref{Fine-Coarse Stencil}).
 Away from coarse-fine interfaces, we can use our standard HSE discretization.
-\item Call EOS to get $T$ and $h$ in the squares, i.e., $T,h = Th,h(\rho,p,X)$.
+\item Call EOS to get $T$ and $h$ in the squares, i.e., $T,h = T,h(\rho,p,X)$.
 \item We now copy this 1D profile onto the full multifab.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Working Notes on Coarse-Fine HSE Discretization - Not Final}
+\subsubsection{Coarse-Fine HSE Discretization}
 Integrating the HSE discretization upward, when we encounter a change from coarse to fine, here is an equation to integrate up to the coarse-fine interface
@@
 p_r^l &=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^l g}{2}\left(\frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_{\sfrac{r}{2}-1}^{l-1}\right) + \frac{\Delta r^l g}{4}\left(\rho_r^l + \frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}\right) \nonumber \\
-&=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{3\Delta r^l g}{4}\left(\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_r^l\right).
+&=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{3\Delta r^l g}{4}\left(\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_r^l\right).\label{Coarse-Fine Stencil}
 \end{eqnarray}
@@
 \begin{equation}
-p_{r}^l = p_{2(r-1)+1}^{l+1} + \frac{3\Delta r^{l+1}g}{4}\left(\rho_{2(r-1)+1}^{l+1} + \rho_{r}^l\right).
+p_{(r+1)/2}^{l-1} = p_{r}^l + \frac{3\Delta r^l g}{4}\left(\rho_{r}^l+\rho_{(r+1)/2}^{l-1}\right).\label{Fine-Coarse Stencil}
 \end{equation}
-
 \end{document}

### 2008-08-25 docs/initial_models/paper.tex
reflects what the code is currently doing

@@
 \item Set $\rho_0 = \overline\rho$.
-\item Compute $p_0$ using equation (\ref{HSE Discretization}), using $p_{0,r=0}$ as the initial condition.\MarginPar{We may change the location of the initial condition.}
+\item Compute $p_0$ using equation (\ref{HSE Discretization}), using the pressure at the top of the domain as the initial condition.
+Note that we still integrate upwards from the bottom, but then offset the final pressure so the pressure at the top of the domain remains unchanged.
 \item Compute the full state $T,h = T,h(\rho,p_0,X)$.
@@
 \begin{enumerate}
-\item Copy the initial data from the dots directly into the circles.
-For the squares, compute only $\rho$ and $X$ by using linear interpolation from the two nearest corresponding dots.\MarginPar{Maybe we should use quadratic interpolation?}
+\item Copy the initial data from the dots directly into the circles.  For the squares, compute only $\rho$ and $X$ by using linear interpolation from the two nearest corresponding dots.
 \item Integrate to get $p$ in the squares.
@@
 \subsubsection{Coarse-Fine HSE Discretization}
-Integrating the HSE discretization upward, when we encounter a change from coarse to fine, here is an equation to integrate up to the coarse-fine interface
+Integrating the HSE discretization upward, when we encounter a change from coarse to fine, here is an equation to integrate up to the coarse-fine interface ($l$ is the ``finer'' level, $l-1$ is the ``coarser'' level)
 \begin{equation}

### 2008-08-27 docs/initial_models/paper.tex
some more work... my multilevel notes aren't quite right in section 3 yet.  some sublte tweaking may have to be done in the algorithm as well

@@
 
-\section{Creating the Model Data from Raw Data}
+\section{Creating the Model Data from Raw Data}\label{Sec:Creating the Model Data from Raw Data}
 We begin with raw data, $\rho^{\raw}, T^{\raw}, p^{\raw}$, and $X^{\raw}$.  Here is the raw data file for each test problem:
@@
 \subsection{Multilevel Plane-Parallel}
-We need to generate model data with the same resolution as the finest base state resolution.
+We do the same thing as the single-level case except that we generate model data with the same resolution as the finest base state resolution.
 \section{Creating the Initial Data from the Model Data}
@@
 Then, we set $h^{\init} = h^{\init}(\rho^{\init},T^{\init},X^{\init})$.
-Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\model} = p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the previous fortran subroutine.
+Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\model} = p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the fortran subroutine in \S~\ref{Sec:Creating the Model Data from Raw Data}.
 \subsection{Multilevel Plane-Parallel}
-For the finest level, we can set the initial data equal to the model data.
-For the non-finest levels, we use linear interpolation to compute the initial data.
-For cells that are covered by finer cells, we average $\rho^{\init}, (\rho h)^{\init},$ and $(\rho X)^{\init}$ from the fine cell to get the coarse cell value.
-Then, for every cell not at the finest level, we call the EOS to compute $p^{\init},T^{\init} = p^{\init},T^{\init}(\rho^{\init},h^{\init},X^{\init})$.\MarginPar{Now $p^{\init}$ and $\rho^{\init}$ are not in HSE.
-Is this a problem?}
+We do the same thing as the single-level case except that we generage initial data at the non-finest levels by using linear interpolation from the two nearest model points.
+Then, after we compute $h^{\init} = h^{\init}(\rho^{\init},T^{\init},X^{\init})$, for each cell that is covered by finer cells, we average $\rho^{\init}, (\rho h)^{\init},$ and $(\rho X)^{\init}$ from the fine cell to get the coarse cell value.
+Then, for every cell not at the finest level, we call the EOS to compute $p^{\init},T^{\init} = p^{\init},T^{\init}(\rho^{\init},h^{\init},X^{\init})$.
+Note that the initial data is not in HSE.
+Later, when we compute the base state, we will make sure that the base state is in HSE.
 \section{Creating the Base State and Full State}\label{Sec:Creating the Base State and Full State}
@@
 For spherical problems, fill $\rho^{\init}, p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state $\rho, p$, and $X$, and then call the EOS to obtain the full state $h,T = h,T(\rho,p,X)$.
-Note that the multifab version of $p$ is temporary, and can be deallocated after this step.
+Note that this particular multifab version of $p$ is temporary, and can be deallocated after this step.
 \item For spherical problems, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
@@
 \subsection{Multilevel Plane-Parallel}
+
 Refer to figure \ref{Fig:Multilevel Initialization}.

### 2008-08-27 docs/initial_models/paper.tex
now reflects what's in cvs.  I'm much happier with this algorithm, it's very clean compared to what was there before.  All the test problems seem to perform well.

--- 2008-08-25 docs/initial_models/paper.tex
@@
 Then, we set $h^{\init} = h^{\init}(\rho^{\init},T^{\init},X^{\init})$.
-Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\model} = p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the previous fortran subroutine.
+Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\model} = p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the fortran subroutine in \S~\ref{Sec:Creating the Model Data from Raw Data}.
 \subsection{Multilevel Plane-Parallel}
-For the finest level, we can set the initial data equal to the model data.
-For the non-finest levels, we use linear interpolation to compute the initial data.
-For cells that are covered by finer cells, we average $\rho^{\init}, (\rho h)^{\init},$ and $(\rho X)^{\init}$ from the fine cell to get the coarse cell value.
-Then, for every cell not at the finest level, we call the EOS to compute $p^{\init},T^{\init} = p^{\init},T^{\init}(\rho^{\init},h^{\init},X^{\init})$.\MarginPar{Now $p^{\init}$ and $\rho^{\init}$ are not in HSE.
-Is this a problem?}
+We do the same thing as the single-level case except that we generate initial data at the non-finest levels by using linear interpolation from the two nearest model points (see Figure \ref{Fig:Multilevel Initialization}).
+Note that the initial data is not in HSE nor is it thermodynamically consistent on non-finest cells.
+We will enforce these on the base state and full state later.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[tpb]
+\begin{center}
+\epsffile{./multilevel_init.eps}
+\caption{Multilevel Initialization.
+The data from the initial model is represented by the dots on the right.
+The initial data at the finest level is represented by the circles.
+The initial data at non-finest levels is represented by the squares.
+We copy data from the dots directly into the circles.
+We use linear interpolation using the two nearest points to copy data from the dots to the squares.}
+\label{Fig:Multilevel Initialization}
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Creating the Base State and Full State}\label{Sec:Creating the Base State and Full State}
 Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.
-We require that $\rho_0 = \overline\rho$, the base state is in HSE according to equation (\ref{HSE Discretization}), and the full state is in thermodynamic equilibrium with $p_0$.
+We require that $\rho_0 = \overline\rho$, the base state is in HSE according to equation (\ref{HSE Discretization}), and the full state is thermodynamically consistent with $p_0$.
 Here are the steps:
@@
 For spherical problems, fill $\rho^{\init}, p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state $\rho, p$, and $X$, and then call the EOS to obtain the full state $h,T = h,T(\rho,p,X)$.
-Note that the multifab version of $p$ is temporary, and can be deallocated after this step.
+Note that this particular multifab version of $p$ is temporary, and can be deallocated after this step.
 \item For spherical problems, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
 \item If {\tt perturb\_model} = F, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
-\item Perturb $T$, then compute $\rho = \rho(p^{\init},T,X)$.
+\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
 \item Set $\rho_0 = \overline\rho$.
@@
 \item Compute the full state $T,h = T,h(\rho,p_0,X)$.
+\item Set $(\rho h)_0 = \overline{(\rho h)}$.
 \end{enumerate}
-Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is in thermodynamic equilibrium with $p_0$.
+Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is thermodynamically consistent with $p_0$.
 \subsection{Multilevel Plane-Parallel}
-Refer to figure \ref{Fig:Multilevel Initialization}.
-The dots represent the initial data, which was generated at the finest resolution on the domain, and is in HSE as well as thermodynamic equilibrium.
-The circles represent the fine level data points in the 1D array.
-The squares represent the coarse level data points in the 1D array.
-We modify step 1 in \S~\ref{Sec:Creating the Base State and Full State} as follows:
 \begin{enumerate}
-\item Step 1
-\begin{enumerate}
-\item Copy the initial data from the dots directly into the circles.  For the squares, compute only $\rho$ and $X$ by using linear interpolation from the two nearest corresponding dots.
-\item Integrate to get $p$ in the squares.
-We need the fancy stencils at coarse-fine interfaces in equations (\ref{Coarse-Fine Stencil}) and (\ref{Fine-Coarse Stencil}).
-Away from coarse-fine interfaces, we can use our standard HSE discretization.
-\item Call EOS to get $T$ and $h$ in the squares, i.e., $T,h = T,h(\rho,p,X)$.
-\item We now copy this 1D profile onto the full multifab.
+\item Fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.
+\item If {\tt perturb\_model} = F, skip the next two steps and instead set $\rho_0 = \rho^{\init}$.
+\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
+\item Set $\rho_0 = \overline\rho$.
+\item Compute $p_0$ using equation (\ref{HSE Discretization}), using the pressure at the top of the domain as the initial condition.
+Note that we still integrate upwards from the bottom, but then offset the final pressure so the pressure at the top of the domain remains unchanged.
+Use the coarse-fine interface discretizations given in \S~\ref{Sec:Coarse-Fine HSE Discretization}.
+\item Compute the full state $T,h = T,h(\rho,p_0,X)$.
+\item Set $(\rho h)_0 = \overline{(\rho h)}$.
 \end{enumerate}
-\end{enumerate}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[tpb]
-\begin{center}
-\epsffile{./multilevel_init.eps}
-\caption{Multilevel Initialization}
-\label{Fig:Multilevel Initialization}
-\end{center}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Coarse-Fine HSE Discretization}
+Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is thermodynamically consistent with $p_0$.
+\subsubsection{Coarse-Fine HSE Discretization}\label{Sec:Coarse-Fine HSE Discretization}
 Integrating the HSE discretization upward, when we encounter a change from coarse to fine, here is an equation to integrate up to the coarse-fine interface ($l$ is the ``finer'' level, $l-1$ is the ``coarser'' level)

### 2008-08-28 docs/thermo_notes/paper.tex
some thermo notes

\documentclass[11pt]{article}

\tolerance=600

\usepackage{amsmath,color,epsfig,epsf}

% Margins
\usepackage[lmargin=0.5in,rmargin=1.5in,tmargin=0.5in,bmargin=0.5in]{geometry}

\newcommand{\sfrac}[2]{\mathchoice
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
  {#1\!/#2}}

\def\half  {\frac{1}{2}}
\newcommand{\myhalf}{\sfrac{1}{2}}

\def\ptl   {\partial}

\title{Notes on thermodnyamic relations in {\tt MAESTRO}}

\setlength{\marginparwidth}{1.0in}
\newcommand{\MarginPar}[1]{\marginpar{%
\vskip-\baselineskip %raise the marginpar a bit
\raggedright\tiny\sffamily
\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\begin{document}

  \maketitle
  \tableofcontents

  \clearpage

  \section{Derivatives With Respect to Composition}
  \label{Sec:Derivatives With Respect to Composition}
  In the following we assume that the molar mass of species $i$ is given by its
  atomic mass number, $A_i = N_i + Z_i$ where $N_i$ is the number of neutrons
  and $Z_i$ is the number of protons in the isotope.  \MarginPar{Note that this
    is common practice and that our EOS also makes this approximation.}  This 
  is a slight approximation that ignores the mass difference between protons 
  and neutrons as well as some minor binding energy terms.

  The number density [cm$^-{3}$] of isotope $i$ is can be formed from the mass 
  density and the molar mass of that isotope as follows:
  \begin{equation}\label{eq:number density}
    n_i = \frac{\rho_i N_\text{A}}{A_i},
  \end{equation}
  where $N_\text{A}$ is Avogadro's number [\# / mole].  The molar abundance, 
  $Y_i$, is a measure of the number of moles of species $i$ per gram in the 
  system:
  \begin{equation}\label{eq:molar abundance}
    Y_i = \frac{n_i}{\rho N_\text{A}} = \frac{\rho_i}{\rho}\frac{1}{A_i} 
    \equiv \frac{X_i}{A_i}
  \end{equation}
  where we have defined the mass fraction, $X_i = \frac{\rho_i}{\rho}$.  Note
  \begin{equation}\label{eq:mass fraction sums to 1}
    \sum_i X_i = 1.
  \end{equation}

  We write the average molar mass and average proton number as:
  \begin{align}\label{eq:abar}
    \bar{A} &= \frac{\sum_i A_i n_i}{\sum_i n_i} = \left(\sum_i X_i\right)
    \left(\sum_i \frac{X_i}{A_i}\right)^{-1}\\
    \label{eq:zbar}
    \bar{Z} &= \frac{\sum_i Z_i n_i}{\sum_i n_i} = \left(\sum_i Z_i
    \frac{X_i}{A_i}\right)\left(\sum_i \frac{X_i}{A_i}\right)^{-1}.
  \end{align}
  Our algorithm requires terms involving the derivative thermodynamic variables
  ($p$ or $e$, e.g.) with respect to composition.  Our EOS does not return such
  derivatives but instead returns derivatives of these variables with respect 
  to $\bar{A}$ and $\bar{Z}$.  Using the chain rule, we have
  \begin{equation}\label{eq:p_xk}
    \frac{\ptl p}{\ptl X_k} = p_{X_k} = 
    \frac{\ptl p}{\ptl \bar{A}}\frac{\ptl \bar{A}}{\ptl X_k} + 
    \frac{\ptl p}{\ptl \bar{Z}}\frac{\ptl \bar{Z}}{\ptl X_k}.
  \end{equation}
  From \eqref{eq:abar} and \eqref{eq:zbar} we have
  \begin{align}\label{eq:abar_X_k}
    \frac{\ptl \bar{A}}{\ptl X_k} &= \left(\sum_i \frac{X_i}{A_i}\right)^{-1}
    - \frac{\bar{A}^2}{A_k} = -\frac{\bar{A}}{A_k}\left(\bar{A} - A_k\right)\\
    \label{eq:zbar_X_k}
    \frac{\ptl \bar{Z}}{\ptl X_k} &= 
    \left(\frac{Z_k}{A_k}\right)\left(\sum_i \frac{X_i}{A_i}\right)^{-1}
    - \frac{\bar{Z}}{A_k}\left(\sum_i \frac{X_i}{A_i}\right)^{-1} = 
    -\frac{\bar{A}}{A_k}\left(\bar{Z} - Z_k\right),
  \end{align}
  where after differentiation we have used \eqref{eq:mass fraction sums to 1}
  to write
  \[
  \left(\sum_i \frac{X_i}{A_i}\right)^{-1} = \bar{A}.
  \]
  We therefore have
  \begin{equation}\label{eq:p_Xk_full}
    p_{X_k} = -\frac{\bar{A}}{A_k}\left(\bar{A} - A_k\right)
    \frac{\ptl p}{\ptl\bar{A}} - \frac{\bar{A}}{A_k}\left(\bar{Z} - Z_k\right)
    \frac{\ptl p}{\ptl\bar{Z}}.
  \end{equation}
  Before it was brought to our attention by Frank Timmes, we were missing the
  second term in \eqref{eq:abar_X_k}.  The only place where such terms 
  appear in our algorithm is in a sum over all species, such as:
  \begin{equation}\label{eq:sum over species}
    \sum_i p_{X_i}\dot{\omega}_i = 
    -\bar{A}^2\frac{\ptl p}{\ptl\bar{A}}\sum_i \frac{\dot{\omega}_i}{A_i}
    +\bar{A}\frac{\ptl p}{\ptl\bar{A}}\sum_i \dot{\omega}_i
    -\bar{A}\bar{Z}\frac{\ptl p}{\ptl\bar{Z}}\sum_i \frac{\dot{\omega}_i}{A_i}
    +\bar{A}\frac{\ptl p}{\ptl \bar{Z}}\sum_i\frac{Z_i}{A_i}\dot{\omega}_i.
  \end{equation}
  \MarginPar{Note that if we are dealing with symmetric nuclei 
  (as we have been up through paper III) where $Z_k = N_k = \half A_k$ for all
  $k$, then the last term in \eqref{eq:sum over species} is identically zero
  as well.}
  The second term in \eqref{eq:sum over species} is identically zero because
  \[
  \sum_k \dot{\omega}_k \equiv 0.
  \]
  This second term arises from what was added to \eqref{eq:abar_X_k} by 
  Frank's correction.  Therefore, although important for individual derivatives
  with respect to composition, this correction term has no effect on our 
  solution.

\end{document}

### 2008-09-09 docs/flowchart/paper.tex
put a little order into the equations list
starting work on making this accurately reflect the current code

@@
 \section{Equations}
+
+\subsection{Reaction Network}
 \begin{eqnarray}
@@
 \end{eqnarray}
+
+\subsection{Energy}
+
 \begin{equation}
@@
 \end{equation}
-\begin{eqnarray}
-\frac{\partial \rhozero}{\partial t} &=& - \nablab \cdotb \left( \rhozero w_0 \er \right)
-- \nablab \cdotb \left( \etarho \er \right) \enskip ,
-\label{eq:rho0upd_new}
-\end{eqnarray}
+
+\subsection{Velocity}
+
 \begin{equation}
-\frac{\partial p_0}{\partial t} = w_0 \rho_0 g  + \etarho g
-                                 = - w_0 \frac{\partial p_0}{\partial r} + \etarho g \enskip .
-\label{eq:p0upd_new} 
+U = \ut + w_0 \er
 \end{equation}
-\begin{equation}
-\label{eq:constraint}
-\nablab \cdotb \ubold + \alpha \left( \frac{\partial p_0}{\partial t} + \ubold \cdotb \nablab p_0 \right) =
- -\sigma  \sum_k \xi_k \omegadot_k  + 
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc + \sigma \Hext \equiv S \enskip ,
-\end{equation}
-\begin{eqnarray}
-\frac{\partial w_0}{\partial r} &=& \frac{1}{\beta_0} \nablab \cdotb (\beta_0 w_0 \er) 
-                                    - w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}  \nonumber \\
-                                &=&   \left( \Sbar - \frac{1}{\gammabar p_0}
-                                      \frac{\partial p_0}{\partial t} \right)
-                                     - w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}  \nonumber \\
-                                &=&   \Sbar - \frac{1}{\gammabar p_0}
-                                      \left( \frac{\partial p_0}{\partial t}
-                                       + w_0 \frac{\partial p_0}{\partial r}  \right) \nonumber \\
-                                &=&   \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip \label{eq:divw0} ,
-\end{eqnarray}
-\begin{equation}
-- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
-\end{equation}
+
 \begin{equation}
@@
 \end{equation}
-\begin{equation}
-\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )
- \enskip .  \label{eq:tildeconstraint}
-\end{equation}
-\begin{equation}
-w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
-\end{equation}
-\begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
-+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\label{eq:dw0drnstar}
-\end{equation}
+
 \begin{eqnarray}
@@
 \end{eqnarray}
+
+\begin{equation}
+\label{eq:constraint}
+\nablab \cdotb \ubold + \alpha \left( \frac{\partial p_0}{\partial t} + \ubold \cdotb \nablab p_0 \right) =
+ -\sigma  \sum_k \xi_k \omegadot_k  + 
+  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc + \sigma \Hext \equiv S \enskip ,
+\end{equation}
+
+\begin{equation}
+  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
+  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext \enskip .
+\label{eq:defineS} 
+\end{equation}
+
+\begin{equation}
+\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )
+ \enskip .  \label{eq:tildeconstraint}
+\end{equation}
+
+\begin{equation}
+- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
+\end{equation}
+
+\subsection{Base State}
+
+\begin{eqnarray}
+\frac{\partial \rhozero}{\partial t} &=& - \nablab \cdotb \left( \rhozero w_0 \er \right)
+- \nablab \cdotb \left( \etarho \er \right) \enskip ,
+\label{eq:rho0upd_new}
+\end{eqnarray}
+
+\subsubsection{Plane-Parallel}
+\begin{equation}
+\frac{\partial p_0}{\partial t} = w_0 \rho_0 g  + \etarho g
+                                 = - w_0 \frac{\partial p_0}{\partial r} + \etarho g \enskip .
+\label{eq:p0upd_new} 
+\end{equation}
+
+\begin{equation}
+\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip \label{eq:divw0} ,
+\end{equation}
+
+\subsection{Misc.}
+
+\begin{equation}
+w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
+\end{equation}
+\begin{equation}
+\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
++ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\label{eq:dw0drnstar}
+\end{equation}
+
 \begin{equation}
@@
 \label{eq:nodal_solve}
-\end{equation}
-\begin{equation}
-  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext \enskip .
-\label{eq:defineS} 
 \end{equation}
@@
 
-\item {\bf React State}$(\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp})\rightarrow$
-$(\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp})$ \newline
+\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow$
+$[\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}]$ \newline
 is the process by which we evolve the species and enthalpy from $X_k^{\inp} \rightarrow X_k^{\outp}$ 
@@
 where the interface states are again found via the procedure described in
-paper II, Appendix A.
+paper II, Appendix A.\\ \\
+c)  For the enthalpy update, \MarginPar{We need enthalpy equations now that we are predicting $(\rho h)'$.}
 \\
@@
 problem is already properly initialized.  We describe the details of
-the initialization in \S~ref{Sec:Initialization}.
+the initialization in \S~\ref{Sec:Initialization}.
 
-In paper II we discretized the density and enthalpy evolution equations in 
-perturbational form, arguing that this would be less
-susceptible to grid effects.  In the present algorithm, the analog would
-be to use a perturbational form of the species equations.
-Numerical testing has shown that the form of the algorithm presented
-here is in fact more robust than the perturbational form 
-when the base state mixing terms ($\etarho$) are included.  
-In the algorithm described below, density is still 
-extrapolated to edges in perturbational form, but 
-we treat $X_k$ and $h$ in non-perturbational form.\\
+In paper III we discretized the enthalpy evolution equation in full, non-perturbation form.  
+We have discovered that discretizing the enthalpy evolution in non-perturbational form 
+leads to better numerical properties.  This is more like paper II.\\
 
@@
       for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
-      predict $\rho'$, $T$, and $X_k$ to the edges, using
+      predict $\rho'$, $(\rho h)'$, and $X_k$ to the edges, using
       a second-order Taylor expansion in space and time, as described in
       paper II, Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  
-      We explicitly include the reaction terms in the temperature prediction, 
-      since we 
-      did not update the temperature in {\bf React State}. 
-      We use the equation of state and base state density, $\rho_0$,
+      We do not include the reaction terms in the enthalpy prediction, 
+      since we accounted for them already in {\bf React State}. 
+      We use the base state density, $\rho_0$, and base state enthalpy, $(\rho h)_0$,
       to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
@@
       for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
-      predict $\rho'$, $T$, and $X_k$ to the edges, using
+      predict $\rho'$, $(\rho h)'$, and $X_k$ to the edges, using
       a second-order Taylor expansion in space and time, as described in
       paper II, Appendix A, using $\V = \uadvtwo+w_0^{\nph} \er$. 
-      Again, we explicitly include the reaction terms in the temperature prediction, 
-      since we did not update  
-      the temperature in {\bf React State}.
-      We use the equation of state and base state density, $\rho_0$,
+      Again, we do not include the reaction terms in the enthalpy prediction 
+      since we accounted for them already in {\bf React State}.
+      We use the base state density, $\rho_0$, and base state enthalpy, $(\rho h)_0$,
       to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
@@
 
-\subsection{Initialization}
+\subsection{Initialization}\label{Sec:Initialization}
 

### 2008-09-09 Source/advance.f90
cleaned up some comments

@@ -247,19 +247,13 @@ contains
                                the_bc_tower%bc_tower_array)
     end if
 
-    if (dm .eq. 3) then
-       do n=1,nlevs
-          do comp=1,dm
-             call multifab_build(w0mac(n,comp),mla%la(n),1,1,nodal=edge_nodal_flag(comp,:))
-             call setval(w0mac(n,comp), ZERO, all=.true.)
-          end do
-       end do
-    end if
-
     do n=1,nlevs
        call multifab_build(ptherm_old(n), mla%la(n), 1, 0)
     end do
     
+    ! compute p0_minus_pthermbar = p0_old - pthermbar (for RHS of projection)
+    ! and delta_p_term = ptherm_old - pthermbar_cart (for making w0)
+
     ! ptherm_old now holds the thermodynamic p computed from sold(rho,h,X)
     call makePfromRhoH(sold,ptherm_old,tempbar,mla,the_bc_tower%bc_tower_array,dx)
 
@@ -294,7 +288,10 @@ contains
     if (dm .eq. 3) then
        do n=1,nlevs
           call multifab_build(w0_force_cart_vec(n), mla%la(n), dm, 1)
-          call setval(w0_force_cart_vec(n),ZERO,all=.true.)
+          call multifab_build(w0mac(n,comp),        mla%la(n),  1, 1, &
+                              nodal=edge_nodal_flag(comp,:))
+          call setval(w0_force_cart_vec(n), ZERO, all=.true.)
+          call setval(w0mac(n,comp),        ZERO, all=.true.)
        end do
     end if
 
@@ -764,6 +761,9 @@ contains
        do n=1,nlevs
           call destroy(ptherm_new(n))
        end do
+
+       ! compute delta_p_term = ptherm_nph - pthermbar_cart (for RHS of projection)
+       ! and p0_minus_pthermbar = p0_nph - pthermbar (for making w0)
 
        ! compute pthermbar = Avg(ptherm_nph)
        call average(mla,ptherm_nph,pthermbar,dx,1)
@@ -1240,6 +1240,8 @@ contains
           do n=1,nlevs
              call multifab_build(ptherm_new(n), mla%la(n), 1, 0)
           enddo
+
+          ! compute delta_p_term = ptherm_new - pthermbar_cart (for RHS of projection)
 
           ! ptherm_new now holds the thermodynamic p computed from snew(rho h X)
           call makePfromRhoH(snew,ptherm_new,tempbar,mla, &

### 2008-09-09 docs/flowchart/paper.tex
added marginpars about missing dpdt_factor description

@@
 the lagged $\psi^{\nmh},$
+\MarginPar{Need dpdt\_factor stuff.}
 \begin{equation}
@@
 We note that  $\uadvone$ satisfies the discrete versions of equations
-(\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}), 
-in particular,
+(\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}), in particular,
+\MarginPar{Need dpdt\_factor stuff.}
 \begin{equation}
@@
 
-\item {\bf React State}$(\rho^n, (\rho h)^n, X_k^n, T^n, (\rho^n \Hext))
-                   \rightarrow$ $( \rho^{(1)}, (\rho h)^{(1)}, X_k^{(1)}, T^{(1)},
-                                  (\rho \omegadot_k)^{(1)} ).$
+\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho^n \Hext)]
+                   \rightarrow$ $[ \rho^{(1)}, (\rho h)^{(1)}, X_k^{(1)}, T^{(1)},
+                                  (\rho \omegadot_k)^{(1)} ].$
 
@@
 \item Construct $w_0^{\nph}$ by integrating equation (\ref{eq:divw0}),
+\MarginPar{Need dpdt\_factor stuff.}
 \begin{equation}
@@
 for $w_0$ and $\pizero.$ We note that  $\uadvtwo$ satisfies the discrete versions of 
-equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}),
-in particular
+equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}), in particular,
+\MarginPar{Need dpdt\_factor stuff.}
 \begin{equation}
@@
 Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}),
+\MarginPar{Need dpdt\_factor stuff.}
 \[

### 2008-09-19 docs/pert_notes/paper.tex
texism
starting on updates to the doc

@@
 \end{eqnarray}
-\subsubsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
-{\bf What the code did before 8/6/08:} 
+\subsubsection{Advancing $(\rho h)_0$}
+{\bf Using full\_rhoh0\_evolution = F:}  Define 
+$(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  
+and $(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then, $(\rho h)_0^{(1)} \rightarrow$ 
+{\tt advect\_base} $\rightarrow (\rho h)_0^{(2)}$.\\ \\
+{\bf Using full\_rhoh0\_evolution = T:}
  Instead of defining $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries 
@@
 $\rightarrow (\rho h)_0^{(2)} \rightarrow$ {\tt react\_base} $\rightarrow (\rho h)_0^{n+1}$.  
-There is no {\tt correct\_base} function for $(\rho h)_0$ written yet, nor is any adjustment made 
-to sync up $(\rho h)_0$ after the thermal diffusion step.\\ \\
-{\bf What the code currently does:}  Define $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  
-and $(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'$ to 
+There is no {\tt correct\_base} function for $(\rho h)_0$ written yet, nor is any 
+adjustment made to sync up $(\rho h)_0$ after the thermal diffusion step.
+\subsubsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
+We predict $(\rho h)'$ to 
 edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} and the underbraced term in 
-(\ref{rhohprime equation}) as the forcing. \MarginPar{{\tt full\_rhoh0\_evolution} comes into play here}
+(\ref{rhohprime equation}) as the forcing, except that we omit the $\psi$ in the forcing for plane-parallel, and we omit the $\partial w_0/\partial r$ part of $\psi$ in spherical.
+Note that we still use full $\psi$ to update the cell.
 \subsubsection{Predicting $(\rho h)_0$ at edges}
@@
 \begin{center}
-\caption{Quantity that is created in mkflux and passed into update scal\newline\
+\caption{Quantity that is created in {\tt mkflux} and passed into {\tt update\_scal}\newline\
 }

### 2008-09-26 docs/eta_notes/paper.tex
fixed some typos, added some clarification

@@
 
-We carry around three different 1D radial quantities: $\etarho$ (edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$ (edge-centered).
+We carry around three different 1D radial quantities: $\etarho$ (edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$ (cell-centered).
 These notes discuss when each of these is used, and how they are computed, in both plane-parallel and spherical.
@@
 \subsection{Spherical}
-First, construct $\eta_{\rho}^{\rm cart} = [\rho'(\Ubt\cdot\eb_r)]^{n+\half}$.  Then, $\etarhocc$ is the cell-centered average of $\eta_{\rho}^{\rm cart}$,
+First, construct $\eta_{\rho}^{\rm cart} = [\rho'(\Ubt\cdot\eb_r)]^{n+\half}$ using:
+\begin{equation}
+\left[\frac{\rho^n+\rho^{n+1}}{2}-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)^{\rm cart}\right] \sum_d\left(\frac{\Ubt_{\ib+\half\eb_d}^{n+\half}\cdot\eb_d+\Ubt_{\ib-\half\eb_d}^{n+\half}\cdot\eb_d}{2}\right)n_d.
+\end{equation}
+Then, $\etarhocc$ is the cell-centered average of $\eta_{\rho}^{\rm cart}$,
 \begin{equation}
@@
 \begin{equation}
-\eta_{\rho,r+\half} = \frac{\eta_{\rho,r}^{\rm cc} + \eta_{\rho,r+1}^{\rm cc}}{2}.
+\eta_{\rho,r-\half} = \frac{\eta_{\rho,r-1}^{\rm cc} + \eta_{\rho,r}^{\rm cc}}{2}.
 \end{equation}
@@
 \begin{eqnarray}
-\eta_{\rho,0} &=& 0, \\
-\eta_{\rho,{\rm nr}} &=& \eta_{\rho,{\rm nr}-1}^{\rm cc}.
+\eta_{\rho,-\half} &=& 0, \\
+\eta_{\rho,{\rm nr}-\half} &=& \eta_{\rho,{\rm nr}-1}^{\rm cc}.
 \end{eqnarray}
-$\divetarho$ is constructed from the cell-centered average of $\rho'^{n+1}$ as
+$\divetarho$ is constructed from the cell-centered average of $\rho'^{n+1} = \rho^{n+1} - \rho_0^{n+1}$ as
 \begin{equation}

### 2008-10-01 docs/eta_notes/paper.tex
div_etarho is not actually needed in spherical make_w0.  updated the eta_notes doc to reflect this

@@
 \def\etarho     {\eta_\rho}
+\def\etarhoec   {\etarho^{\rm ec}}
 \def\etarhocc   {\etarho^{\rm cc}}
@@
 
-We carry around three different 1D radial quantities: $\etarho$ (edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$ (cell-centered).
+We carry around three different 1D radial quantities: $\etarhoec$ (edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$ (cell-centered).
 These notes discuss when each of these is used, and how they are computed, in both plane-parallel and spherical.
@@
 \begin{enumerate}
-\item Enter {\tt advance} with $[\etarho, \etarhocc, \divetarho]^{n-\half}$.
-\item Call {\tt make\_w0}.  The spherical version uses uses $\etarho$ and $\divetarho$.
+\item Enter {\tt advance} with $[\etarhoec, \etarhocc, \divetarho]^{n-\half}$.
+\item Call {\tt make\_w0}.  The spherical version uses uses $\etarhoec$ and $\etarhocc$.
 \item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
 \item Call {\tt correct\_base}, which uses $\divetarho$.
-\item Compute $[\etarho, \etarhocc, \divetarho]^{n+\half,*}$.
+\item Compute $[\etarhoec, \etarhocc, \divetarho]^{n+\half,*}$.
 \item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
-\item Call {\tt make\_w0}.  The spherical version uses uses $\etarho$ and $\divetarho$.
+\item Call {\tt make\_w0}.  The spherical version uses uses $\etarhoec$ and $\etarhocc$.
 \item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
 \item Call {\tt correct\_base}, which uses $\divetarho$.
-\item Compute $[\etarho, \etarhocc, \divetarho]^{n+\half}$.
+\item Compute $[\etarhoec, \etarhocc, \divetarho]^{n+\half}$.
 \item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
@@
 
-\section{Computing $\etarho, \etarhocc$, and $\divetarho$}
+\section{Computing $\etarhoec, \etarhocc$, and $\divetarho$}
 This is done in {\tt make\_eta.f90}.
@@
 \end{equation}
-$\etarho$ is the edge-centered ``average'' value of $\eta_{\rho}^{\rm flux}$,
+$\etarhoec$ is the edge-centered ``average'' value of $\eta_{\rho}^{\rm flux}$,
 \begin{equation}
-\eta_{\rho,r+\half} = \overline{\eta_{\rho,\ib+\half\eb_r}^{\rm flux}}
+\eta_{\rho,r+\half}^{\rm ec} = \overline{\eta_{\rho,\ib+\half\eb_r}^{\rm flux}}
 \end{equation}
-$\etarhocc$ is a cell-centered average of $\etarho$,
+$\etarhocc$ is a cell-centered average of $\etarhoec$,
 \begin{equation}
-\eta_{\rho,r}^{\rm cc} = \frac{\eta_{\rho,r+\half} + \eta_{\rho,r-\half}}{2}.
+\eta_{\rho,r}^{\rm cc} = \frac{\eta_{\rho,r+\half}^{\rm ec} + \eta_{\rho,r-\half}^{\rm ec}}{2}.
 \end{equation}
-$\divetarho$ is the cell-centered divergence of $\etarho$,
+$\divetarho$ is the cell-centered divergence of $\etarhoec$,
 \begin{equation}
-\left[\divetarho\right]_r = \frac{\eta_{\rho,r+\half} - \eta_{\rho,r-\half}}{\Delta r}.
+\left[\divetarho\right]_r = \frac{\eta_{\rho,r+\half}^{\rm ec} - \eta_{\rho,r-\half}^{\rm ec}}{\Delta r}.
 \end{equation}
@@
 \end{equation}
-On interior faces, $\etarho$ is the average of $\etarhocc$,
+On interior faces, $\etarhoec$ is the average of $\etarhocc$,
 \begin{equation}
-\eta_{\rho,r-\half} = \frac{\eta_{\rho,r-1}^{\rm cc} + \eta_{\rho,r}^{\rm cc}}{2}.
+\eta_{\rho,r-\half}^{\rm ec} = \frac{\eta_{\rho,r-1}^{\rm cc} + \eta_{\rho,r}^{\rm cc}}{2}.
 \end{equation}
@@
 \begin{eqnarray}
-\eta_{\rho,-\half} &=& 0, \\
-\eta_{\rho,{\rm nr}-\half} &=& \eta_{\rho,{\rm nr}-1}^{\rm cc}.
+\eta_{\rho,-\half}^{\rm ec} &=& 0, \\
+\eta_{\rho,{\rm nr}-\half}^{\rm ec} &=& \eta_{\rho,{\rm nr}-1}^{\rm cc}.
 \end{eqnarray}
@@
 \end{equation}
-
-\section{Using $\etarho$}
-
+\section{Using $\etarhoec$}
 \subsection{Plane-Parallel}
 NOT USED.
-
 \subsection{Spherical}
-In {\tt make\_w0}, $\etarho$ is used in the construction of the RHS for the $\delta w_0$ equation.
-
+In {\tt make\_w0}, $\etarhoec$ is used in the construction of the RHS for the $\delta w_0$ equation.
 \section{Using $\etarhocc$}
-
 \subsection{Plane-Parallel}
-In {\tt make\_psi}, $\psi = \etarho g$.
-
+In {\tt make\_psi}, $\psi = \etarhocc g$.
 \subsection{Spherical}
-NOT USED.
-
+In {\tt make\_w0}, $\etarhocc$ is used in the construction of the RHS for the $\delta w_0$ equation.
 \section{Using $\divetarho$}
-
 \subsection{Plane-Parallel}
 In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$.
-
 \subsection{Spherical}
 In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$. \\
-In {\tt make\_w0}, we average $\divetarho$ onto edges for the $\delta w_0$ equation.
 \end{document}

### 2008-10-01 Source/advance.f90

@@ -289,7 +289,7 @@ contains
 
        call make_w0(w0,w0_old,w0_force,Sbar,rho0_old,rho0_old,p0_old,p0_old, &
                     gamma1bar,gamma1bar,p0_minus_pthermbar, &
-                    psi,etarho_ec,etarho_cc,div_etarho,dt,dtold)
+                    psi,etarho_ec,etarho_cc,dt,dtold)
 
        if (spherical .eq. 1) then
           call put_w0_on_edges(mla,w0,w0mac,dx,div_coeff_old,the_bc_tower)
@@ -808,7 +808,7 @@ contains
 
           call make_w0(w0,w0_old,w0_force,Sbar,rho0_old,rho0_new,p0_old,p0_new, &
                        gamma1bar_old,gamma1bar,p0_minus_pthermbar, &
-                       psi,etarho_ec,etarho_cc,div_etarho,dt,dtold)
+                       psi,etarho_ec,etarho_cc,dt,dtold)
 
           if (spherical .eq. 1) then
              call put_w0_on_edges(mla,w0,w0mac,dx,div_coeff_nph,the_bc_tower)

### 2008-10-01 Source/make_w0.f90

@@ -18,8 +18,7 @@ contains
 
   subroutine make_w0(w0,w0_old,w0_force,Sbar_in,rho0_old,rho0_new,p0_old,p0_new, &
                      gamma1bar_old,gamma1bar_new,p0_minus_pthermbar, &
-                     psi,etarho_ec,etarho_cc,div_etarho, &
-                     dt,dtold)
+                     psi,etarho_ec,etarho_cc,dt,dtold)
 
     use parallel
     use bl_prof_module
@@ -33,7 +32,6 @@ contains
     real(kind=dp_t), intent(in   ) :: psi(:,0:)
     real(kind=dp_t), intent(in   ) :: etarho_ec(:,0:)
     real(kind=dp_t), intent(in   ) :: etarho_cc(:,0:)
-    real(kind=dp_t), intent(in   ) :: div_etarho(:,0:)
     real(kind=dp_t), intent(inout) :: w0_force(:,0:)
     real(kind=dp_t), intent(in   ) :: rho0_old(:,0:), rho0_new(:,0:)
     real(kind=dp_t), intent(in   ) :: p0_old(:,0:), p0_new(:,0:)
@@ -64,8 +62,7 @@ contains
           call make_w0_spherical(n,w0(n,:),w0_old(n,0:),Sbar_in(n,0:), &
                                  rho0_old(n,:),rho0_new(n,:),p0_old(n,0:),p0_new(n,0:), &
                                  gamma1bar_old(n,0:),gamma1bar_new(n,0:), &
-                                 p0_minus_pthermbar(n,0:), &
-                                 etarho_ec(n,0:),etarho_cc(n,0:),div_etarho(n,0:), &
+                                 p0_minus_pthermbar(n,0:),etarho_ec(n,0:),etarho_cc(n,0:), &
                                  w0_force(n,0:),dt,dtold)
        end do
 
@@ -213,7 +210,7 @@ contains
 
   subroutine make_w0_spherical(n,w0,w0_old,Sbar_in,rho0,rho0_new,p0,p0_new, &
                                gamma1bar,gamma1bar_new,p0_minus_pthermbar, &
-                               etarho_ec,etarho_cc,div_etarho,w0_force,dt,dtold)
+                               etarho_ec,etarho_cc,w0_force,dt,dtold)
 
[…]

### 2008-10-11 docs/initial_models/paper.tex
update the locations of the various initial models and the generation routines.

@@
 \begin{eqnarray}
-{\tt test2} & \rightarrow & {\tt ???} \nonumber \\
-{\tt test\_convect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt wdconvect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt spherical\_heat} & \rightarrow & {\tt ???} \nonumber \\
+{\tt test2} & \rightarrow & \mathrm{comes~from~the~ {\tt lmn\_bubble\_burn}~setup~in~ {\tt FLASH}} \nonumber \\
+{\tt test\_convect} & \rightarrow & \mathrm{comes~from~the~ {\tt lmn\_convect}~setup~in~ {\tt FLASH}} \nonumber \\
+{\tt wdconvect} & \rightarrow & {\tt initial\_models/kepler/model\_6.e8.raw} \nonumber \\
+{\tt spherical\_heat} & \rightarrow & \mathrm{none-it~ is~ generated~ analytically} \nonumber \\
 {\tt xrb} & \rightarrow & {\tt initial\_models/xrb/sorted\_xrb.raw} \nonumber
@@
 \begin{eqnarray}
-{\tt test2} & \rightarrow & {\tt ???} \nonumber \\
-{\tt test\_convect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt wdconvect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt spherical\_heat} & \rightarrow & {\tt ???} \nonumber \\
+{\tt test2} & \rightarrow & \mathrm{comes~from~the~ {\tt lmn\_bubble\_burn}~setup~in~ {\tt FLASH}} \nonumber \\
+{\tt test\_convect} & \rightarrow & \mathrm{comes~from~the~ {\tt lmn\_convect}~setup~in~ {\tt FLASH}} \nonumber \\
+{\tt wdconvect} & \rightarrow & {\tt initial\_models/kepler/init\_1d.f90} \nonumber \\
+{\tt spherical\_heat} & \rightarrow & {\tt initial\_models/spherical/init\_1d.f90} \nonumber \\
 {\tt xrb} & \rightarrow & {\tt initial\_models/xrb/init\_1d.f90} \nonumber

### 2008-10-11 docs/initial_models/paper.tex
update how we define dr

--- 2008-08-27 docs/initial_models/paper.tex
@@
 \section{Creating the Model Data from Raw Data}\label{Sec:Creating the Model Data from Raw Data}
-We begin with raw data, $\rho^{\raw}, T^{\raw}, p^{\raw}$, and $X^{\raw}$.  Here is the raw data file for each test problem:
+
+We begin with raw data, $\rho^{\raw}, T^{\raw}, p^{\raw}$, and
+$X^{\raw}$.  Here is the raw data file for each test problem:
 \begin{eqnarray}
-{\tt test2} & \rightarrow & {\tt ???} \nonumber \\
-{\tt test\_convect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt wdconvect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt spherical\_heat} & \rightarrow & {\tt ???} \nonumber \\
+{\tt test2} & \rightarrow & \mathrm{comes~from~the~ {\tt lmn\_bubble\_burn}~setup~in~ {\tt FLASH}} \nonumber \\
+{\tt test\_convect} & \rightarrow & \mathrm{comes~from~the~ {\tt lmn\_convect}~setup~in~ {\tt FLASH}} \nonumber \\
+{\tt wdconvect} & \rightarrow & {\tt initial\_models/kepler/model\_6.e8.raw} \nonumber \\
+{\tt spherical\_heat} & \rightarrow & \mathrm{none-it~ is~ generated~ analytically} \nonumber \\
 {\tt xrb} & \rightarrow & {\tt initial\_models/xrb/sorted\_xrb.raw} \nonumber
 \end{eqnarray}
-Using the raw data, we need to generate model data with the same resolution as the base state.
-Note that for spherical problems we set $\Delta r = \Delta x$.
-We use a fortran subroutine to interpolate the raw data, yielding the model data, $\rho^{\model}, T^{\model}, p^{\model}$, and $X^{\model}$.
-The fortran subroutine then uses an iterative procedure to modify the model data so that it is thermodynamically consistent with the {\tt MAESTRO} equation of state (EOS), and also satisfies our chosen hydrostatic equilibrium (HSE) discretization,
+Using the raw data, we need to generate model data with the same
+resolution as the base state.  Note that for spherical problems we set
+$\Delta r = \Delta x/\mathtt{drdxfac}$.  We use a fortran subroutine to interpolate the
+raw data, yielding the model data, $\rho^{\model}, T^{\model},
+p^{\model}$, and $X^{\model}$.  The fortran subroutine then uses an
+iterative procedure to modify the model data so that it is
+thermodynamically consistent with the {\tt MAESTRO} equation of state
+(EOS), and also satisfies our chosen hydrostatic equilibrium (HSE)
+discretization,
 \begin{equation}
@@
 \begin{eqnarray}
-{\tt test2} & \rightarrow & {\tt ???} \nonumber \\
-{\tt test\_convect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt wdconvect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt spherical\_heat} & \rightarrow & {\tt ???} \nonumber \\
+{\tt test2} & \rightarrow & \mathrm{comes~from~the~ {\tt lmn\_bubble\_burn}~setup~in~ {\tt FLASH}} \nonumber \\
+{\tt test\_convect} & \rightarrow & \mathrm{comes~from~the~ {\tt lmn\_convect}~setup~in~ {\tt FLASH}} \nonumber \\
+{\tt wdconvect} & \rightarrow & {\tt initial\_models/kepler/init\_1d.f90} \nonumber \\
+{\tt spherical\_heat} & \rightarrow & {\tt initial\_models/spherical/init\_1d.f90} \nonumber \\
 {\tt xrb} & \rightarrow & {\tt initial\_models/xrb/init\_1d.f90} \nonumber
 \end{eqnarray}
-The model data is not generated at run-time - it must be generated in advance of running any MAESTRO examples.  The inputs file should point to the file containing the model data.
+The model data is not generated at run-time - it must be generated in
+advance of running any MAESTRO examples.  The inputs file should point
+to the file containing the model data.
+
 \subsection{Multilevel Plane-Parallel}
 We do the same thing as the single-level case except that we generate model data with the same resolution as the finest base state resolution.
+
 \section{Creating the Initial Data from the Model Data}
-In {\tt base\_state.f90}, using the subroutine {\tt init\_base\_state.f90} (which is actually a terrible name since we are setting the initial data, not the base state), we set the initial data $\rho^{\init}, T^{\init}, p^{\init}$ and $X^{\init}$ equal to the model data.
-Then, we set $h^{\init} = h^{\init}(\rho^{\init},T^{\init},X^{\init})$.
-Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\model} = p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the fortran subroutine in \S~\ref{Sec:Creating the Model Data from Raw Data}.
+In {\tt base\_state.f90}, using the subroutine {\tt
+  init\_base\_state.f90} (which is actually a terrible name since we
+are setting the initial data, not the base state), we set the initial
+data $\rho^{\init}, T^{\init}, p^{\init}$ and $X^{\init}$ equal to the
+model data.  Then, we set $h^{\init} =
+h^{\init}(\rho^{\init},T^{\init},X^{\init})$.  Note that in the code,
+$p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not
+change since we called $p^{\model} =
+p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the
+fortran subroutine in \S~\ref{Sec:Creating the Model Data from Raw
+  Data}.
+
 \subsection{Multilevel Plane-Parallel}
-We do the same thing as the single-level case except that we generate initial data at the non-finest levels by using linear interpolation from the two nearest model points (see Figure \ref{Fig:Multilevel Initialization}).
-Note that the initial data is not in HSE nor is it thermodynamically consistent on non-finest cells.
-We will enforce these on the base state and full state later.
+We do the same thing as the single-level case except that we generate
+initial data at the non-finest levels by using linear interpolation
+from the two nearest model points (see Figure \ref{Fig:Multilevel
+  Initialization}).  Note that the initial data is not in HSE nor is
+it thermodynamically consistent on non-finest cells.  We will enforce
+these on the base state and full state later.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \epsffile{./multilevel_init.eps}
-\caption{Multilevel Initialization.
-The data from the initial model is represented by the dots on the right.
-The initial data at the finest level is represented by the circles.
-The initial data at non-finest levels is represented by the squares.
-We copy data from the dots directly into the circles.
-We use linear interpolation using the two nearest points to copy data from the dots to the squares.}
+\caption{Multilevel Initialization.  The data from the initial model
+  is represented by the dots on the right.  The initial data at the
+  finest level is represented by the circles.  The initial data at
+  non-finest levels is represented by the squares.  We copy data from
+  the dots directly into the circles.  We use linear interpolation
+  using the two nearest points to copy data from the dots to the
+  squares.}
 \label{Fig:Multilevel Initialization}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Creating the Base State and Full State}\label{Sec:Creating the Base State and Full State}
-Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.
-We require that $\rho_0 = \overline\rho$, the base state is in HSE according to equation (\ref{HSE Discretization}), and the full state is thermodynamically consistent with $p_0$.
-Here are the steps:
+
+Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this
+section describes how the base state ($\rho_0$ and $p_0$) and full
+state ($\rho, h, X$, and $T$) are computed.  We require that $\rho_0 =
+\overline\rho$, the base state is in HSE according to equation
+(\ref{HSE Discretization}), and the full state is thermodynamically
+consistent with $p_0$.  Here are the steps:
 \begin{enumerate}
-\item For plane-parallel problems, fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.
-For spherical problems, fill $\rho^{\init}, p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state $\rho, p$, and $X$, and then call the EOS to obtain the full state $h,T = h,T(\rho,p,X)$.
-Note that this particular multifab version of $p$ is temporary, and can be deallocated after this step.
-\item For spherical problems, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
-\item If {\tt perturb\_model} = F, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
+\item For plane-parallel problems, fill $\rho^{\init}, h^{\init},
+  X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state
+  $\rho, h, X$, and $T$.  For spherical problems, fill $\rho^{\init},
+  p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state
+  $\rho, p$, and $X$, and then call the EOS to obtain the full state
+  $h,T = h,T(\rho,p,X)$.  Note that this particular multifab version
+  of $p$ is temporary, and can be deallocated after this step.
+\item For spherical problems, skip the remaining steps and instead set
+  $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
+\item If {\tt perturb\_model} = F, skip the remaining steps and
+  instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
 \item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
 \item Set $\rho_0 = \overline\rho$.
-\item Compute $p_0$ using equation (\ref{HSE Discretization}), using the pressure at the top of the domain as the initial condition.
-Note that we still integrate upwards from the bottom, but then offset the final pressure so the pressure at the top of the domain remains unchanged.
+\item Compute $p_0$ using equation (\ref{HSE Discretization}), using
+  the pressure at the top of the domain as the initial condition.
+  Note that we still integrate upwards from the bottom, but then
+  offset the final pressure so the pressure at the top of the domain
+  remains unchanged.
 \item Compute the full state $T,h = T,h(\rho,p_0,X)$.
@@
 \end{enumerate}
-Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is thermodynamically consistent with $p_0$.
+Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full
+state is thermodynamically consistent with $p_0$.
+
 \subsection{Multilevel Plane-Parallel}
 \begin{enumerate}
-\item Fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.
-\item If {\tt perturb\_model} = F, skip the next two steps and instead set $\rho_0 = \rho^{\init}$.
+\item Fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto
+  a multifab to obtain the full state $\rho, h, X$, and $T$.
+\item If {\tt perturb\_model} = F, skip the next two steps and instead
+  set $\rho_0 = \rho^{\init}$.
 \item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
 \item Set $\rho_0 = \overline\rho$.
-\item Compute $p_0$ using equation (\ref{HSE Discretization}), using the pressure at the top of the domain as the initial condition.
-Note that we still integrate upwards from the bottom, but then offset the final pressure so the pressure at the top of the domain remains unchanged.
-Use the coarse-fine interface discretizations given in \S~\ref{Sec:Coarse-Fine HSE Discretization}.
+\item Compute $p_0$ using equation (\ref{HSE Discretization}), using
+  the pressure at the top of the domain as the initial condition.
+  Note that we still integrate upwards from the bottom, but then
+  offset the final pressure so the pressure at the top of the domain
+  remains unchanged.  Use the coarse-fine interface discretizations
+  given in \S~\ref{Sec:Coarse-Fine HSE Discretization}.
 \item Compute the full state $T,h = T,h(\rho,p_0,X)$.
@@
 \end{enumerate}
-Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is thermodynamically consistent with $p_0$.
+Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full
+state is thermodynamically consistent with $p_0$.
+
 \subsubsection{Coarse-Fine HSE Discretization}\label{Sec:Coarse-Fine HSE Discretization}
-Integrating the HSE discretization upward, when we encounter a change from coarse to fine, here is an equation to integrate up to the coarse-fine interface ($l$ is the ``finer'' level, $l-1$ is the ``coarser'' level)
+Integrating the HSE discretization upward, when we encounter a change
+from coarse to fine, here is an equation to integrate up to the
+coarse-fine interface ($l$ is the ``finer'' level, $l-1$ is the
+``coarser'' level)
 \begin{equation}
@@
 \end{equation}
-Here is an equation to integrate from the coarse-fine interface up to the cell center
+Here is an equation to integrate from the coarse-fine interface up to
+the cell center
 \begin{equation}
@@
 \end{equation}
+
 \end{document}

### 2008-10-12 docs/flowchart/paper.tex
add MarginPar to note that we split scalar_advance
remove the random equations section.  Now instead inline them in the text and cite the equation number from paper III

@@
 
-\section{Equations}
-
-\subsection{Reaction Network}
-\begin{eqnarray}
-\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \label{eq:VODE1C} \\
-\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip. \label{eq:tempreactC}
-\end{eqnarray}
-
-\subsection{Energy}
-
-\begin{equation}
-\label{eq:tempenthalpy}
-\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
- - \sum_k \rho \xi_k {\omegadot}_k 
- + \rho \Hnuc + \rho \Hext \right] \enskip , 
-\end{equation}
-
-\subsection{Velocity}
-
-\begin{equation}
-U = \ut + w_0 \er
-\end{equation}
-
-\begin{equation}
-\int_{\Omega_H} \ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
-\end{equation}
-
-\begin{eqnarray}
-\frac{\partial w_0}{\partial t} &=& - w_0 \frac{\partial w_0}{\partial r} - 
-\frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \label{eq:w0upd} \enskip , \\
-\frac{\partial\ut}{\partial t} &=& - \ut \cdotb \nablab \ut - w_0 \frac{\partial \ut}{\partial r}
-                                 - \left(\ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
-                                 - \frac{1}{\rho} \nablab\pi
-                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
-                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip ,
-\end{eqnarray}
-
-\begin{equation}
-\label{eq:constraint}
-\nablab \cdotb \ubold + \alpha \left( \frac{\partial p_0}{\partial t} + \ubold \cdotb \nablab p_0 \right) =
- -\sigma  \sum_k \xi_k \omegadot_k  + 
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc + \sigma \Hext \equiv S \enskip ,
-\end{equation}
-
-\begin{equation}
-  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext \enskip .
-\label{eq:defineS} 
-\end{equation}
-
-\begin{equation}
-\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )
- \enskip .  \label{eq:tildeconstraint}
-\end{equation}
-
-\begin{equation}
-- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
-\end{equation}
-
-\subsection{Base State}
-
-\begin{eqnarray}
-\frac{\partial \rhozero}{\partial t} &=& - \nablab \cdotb \left( \rhozero w_0 \er \right)
-- \nablab \cdotb \left( \etarho \er \right) \enskip ,
-\label{eq:rho0upd_new}
-\end{eqnarray}
-
-\subsubsection{Plane-Parallel}
-\begin{equation}
-\frac{\partial p_0}{\partial t} = w_0 \rho_0 g  + \etarho g
-                                 = - w_0 \frac{\partial p_0}{\partial r} + \etarho g \enskip .
-\label{eq:p0upd_new} 
-\end{equation}
-
-\begin{equation}
-\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip \label{eq:divw0} ,
-\end{equation}
-
-\subsection{Misc.}
-
-\begin{equation}
-w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
-\end{equation}
-\begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
-+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\label{eq:dw0drnstar}
-\end{equation}
-
-\begin{equation}
- L_\beta^\rho \phi =
-   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
-\label{eq:nodal_solve}
-\end{equation}
-
 \section{Current Algorithm}
@@
 \frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
-\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip.
+\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip. \label{eq:reacttemp}
 \end{eqnarray}
-using the stiff ordinary differential equation
- integration methods provided by the {\tt VODE} package (cite VODE).
-The absolute error tolerances are set to $10^{-12}$ for the species,
- and a relative tolerance of $10^{-5}$ is used for the temperature.  
-The integration yields the new values of the mass fractions, $X_k^{\outp}$.  
-Equation (\ref{eq:tempreactC}) is derived from equation (\ref{eq:tempenthalpy}) by assuming that the pressure is constant during the burn state.  
-In evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  
-In theory, this means evaluating the equation of state for each right-hand side evaluation that {\tt VODE} requires.  
-In practice, we freeze $c_p$ and $\xi_k$ at the start of the integration time step and compute them using 
-$\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs to the equation of state.  
-We note that while temperature is evolved when solving these equations, we do not keep the final temperature, 
-nor do we use it to compute the final change in enthalpy.  
-Therefore, $T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged during 
-the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
+using the stiff ordinary differential equation integration methods
+provided by the {\tt VODE} package (cite VODE).  The absolute error
+tolerances are set to $10^{-12}$ for the species, and a relative
+tolerance of $10^{-5}$ is used for the temperature.  The integration
+yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
+(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
+\begin{equation}
+\label{eq:tempenthalpy}
+\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
+ - \sum_k \rho \xi_k {\omegadot}_k 
+ + \rho \Hnuc + \rho \Hext \right] \enskip , 
+\end{equation}
+(paper III; eq.~[8])
+by assuming that the pressure is constant during the burn state.  In
+evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
+theory, this means evaluating the equation of state for each
+right-hand side evaluation that {\tt VODE} requires.  In practice, we
+freeze $c_p$ and $\xi_k$ at the start of the integration time step and
+compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
+to the equation of state.  We note that while temperature is evolved
+when solving these equations, we do not keep the final temperature,
+nor do we use it to compute the final change in enthalpy.  Therefore,
+$T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged
+during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
 
@@
 \newline
-is the process by which we update the base state through $\dt$ in time given the
-radial velocity $w_0^{\inp}.$ Here we discuss
-the algorithm for plane-parallel geometries.  
-The base state arrays are all one-dimensional 
-in the radial coordinate; we think of the base state quantities as defined
-at cell centers, with $w_0$ and $\beta_0$ defined halfway in-between.\\ \\
-a)  For the density update, we discretize equation (\ref{eq:rho0upd_new}), neglecting the $\etarho$ term,
-to compute the new base state density,
+is the process by which we update the base state through $\dt$ in time \MarginPar{we need a separate section for spherical}
+given the radial velocity $w_0^{\inp}.$ Here we discuss the algorithm
+for plane-parallel geometries.  The base state arrays are all
+one-dimensional in the radial coordinate; we think of the base state
+quantities as defined at cell centers, with $w_0$ and $\beta_0$
+defined halfway in-between.\\ \\ a) For the density update, we
+discretize equation the base state density evolution equation
+\begin{eqnarray}
+\frac{\partial \rhozero}{\partial t} &=& - \nablab \cdotb \left( \rhozero w_0 \er \right)
+- \nablab \cdotb \left( \etarho \er \right) \enskip ,
+\label{eq:rho0upd_new}
+\end{eqnarray}
+(paper III; eq.~[29]), neglecting the $\etarho$ term, to compute the
+new base state density,
 \begin{eqnarray}
@@
 where $j$ refers to the one-dimensional index in the radial direction.
-The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the procedure described in
-paper II, Appendix A, and are saved for use later in the overall algorithm.\\ \\
-b)  For the pressure update, We discretize equation (\ref{eq:p0upd_new}), 
-to compute the new base state pressure,
+The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the
+procedure described in paper II, Appendix A, and are saved for use
+later in the overall algorithm.\\ \\ b) For the pressure update, we
+discretize the base state pressure evolution equation,
+\begin{equation}
+\frac{\partial p_0}{\partial t} = w_0 \rho_0 g  + \etarho g
+                                 = - w_0 \frac{\partial p_0}{\partial r} + \etarho g \enskip .
+\label{eq:p0upd_new}
+\end{equation}
+(paper III; eq.~[31]) to compute the new base state pressure,
 \begin{eqnarray}
@@
 \item At the beginning of each time step, we need an estimate for the time-centered
-source term in the velocity divergence constraint (see eq. [\ref{eq:constraint}]).
+source term in the velocity divergence constraint,
+\begin{equation}
+  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
+  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext \enskip ,
+\label{eq:defineS} 
+\end{equation}
+(paper III; eq.~[19]).
+
 If this is the first step of the calculation ($n=0$), we set
@@
 
-\item Construct $w_0^{\nph,\star}$ by integrating equation (\ref{eq:divw0}) using 
-the lagged $\psi^{\nmh},$
+\item Construct $w_0^{\nph,\star}$ by integrating
+\begin{equation}
+\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip \label{eq:divw0} ,
+\end{equation}
+using the lagged $\psi^{\nmh},$ \MarginPar{note: this is true only for plane-parallel}
 \MarginPar{Need dpdt\_factor stuff.}
@@
 
-\item Using equation (\ref{eq:pizero}), define the scaled pressure gradient
+\item Using 
+\begin{equation}
+- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
+\end{equation}
+define the scaled pressure gradient
 \begin{equation}
@@
 The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
-We note that  $\uadvone$ satisfies the discrete versions of equations
-(\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}), in particular,
+We note that  $\uadvone$ satisfies the discrete versions of 
+\begin{equation}
+\int_{\Omega_H} \ut \cdotb \er \; dA = 0 \label{eq:udoterzero} 
+\end{equation}
+(paper III; eq.~[21])
+and
+\begin{equation}
+\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )
+\label{eq:tildeconstraint}
+\end{equation}
+(paper III; eq.~[39]).
+In particular,
 \MarginPar{Need dpdt\_factor stuff.}
@@
       to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
-      We define $\rho^{(1),\nph,\star} = \sum_k (\rho X_k)^{(1),\nph,\star}.$
+      We define $\rho^{(1),\nph,\star} = \sum_k (\rho X_k)^{(1),\nph,\star}.$  \MarginPar{we actually split these now, and do $\rho$ first, compute $\psi$, and then do $(\rho h)$}
 
@@
 \item Construct $w_0^{\nph}$ by integrating equation (\ref{eq:divw0}),
+\MarginPar{this is plane-parallel only}
 \MarginPar{Need dpdt\_factor stuff.}
@@
 \end{equation}
-where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined analogously to 
-equations (\ref{eq:w0nstar}) and (\ref{eq:dw0drnstar}).
+where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined through
+\begin{equation}
+w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
+\end{equation}
+(paper III; eq.~[53]) and
+\begin{equation}
+\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
++ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\label{eq:dw0drnstar}
+\end{equation}
+(paper III; eq.~[54]).
 
@@
 We update the velocity field $\ut^n$ to $\ut^{n+1,\star}$ by discretizing
-equation (\ref{eq:utildeupd}), 
+\begin{equation}
+\frac{\partial\ut}{\partial t} = - \ut \cdotb \nablab \ut - w_0 \frac{\partial \ut}{\partial r}
+                                 - \left(\ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
+                                 - \frac{1}{\rho} \nablab\pi
+                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
+                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip\
+ ,
+\end{equation}
+(paper III; eq.~[37]) as 
 \begin{eqnarray}
@@
 of this particular form of the projection operand.)  
-We solve the linear system of equation (\ref{eq:nodal_solve}) using multigrid V-cycles with Gauss-Seidel relaxation.
+We solve the linear system of equation
+\begin{equation}
+ L_\beta^\rho \phi =
+   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
+\label{eq:nodal_solve}
+\end{equation}
+(paper III; eq.~[77]) using multigrid V-cycles with Gauss-Seidel relaxation.
 

### 2008-10-12 docs/flowchart/paper.tex
add more MarginPars pointing out where updating is needed

@@
 \section{Current Algorithm}
+
+Here we outline the algorithm currently implemented in the code.  Our
+starting point for this description are the series of papers describing
+the development of the algorithm:
+\begin{itemize}
+\item {\em Low Mach Number Modeling of Type Ia
+  Supernovae. I. Hydrodynamics}, A.~S.~Almgren, J.~B.~Bell, 
+  C.~A.~Rendleman, \& M.~Zingale 2006, ApJ, 637, 922 (henceforth
+  paper~I)
+\item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
+  Evolution} A.~S.~Almgren, J.~B.~Bell, C.~A.~Rendleman, \& M.~Zingale
+  2006, ApJ, 649, 927 (henceforth paper~II)
+\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions}
+A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale
+  2008, ApJ, 684, 449 (henceforth paper~III)
+\end{itemize}
+
 \subsection{Notation}
@@
 \end{equation}
+here $\Omega_H$ is a layer at constant radius $r$. \MarginPar{we should discuss the implementation of average for spherical here}
 
 \item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow$
-$[\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}]$ \newline
+$[\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \newline
 is the process by which we evolve the species and enthalpy from $X_k^{\inp} \rightarrow X_k^{\outp}$ 
@@
 \end{equation}
-The enthalpy update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is updated by
+for the case where only strong-mediated reactions are involved.  As
+$\Hnuc$ is an output of the reaction network, more general networks
+(i.e.\ involving weak interactions) can easily be added.  The enthalpy
+update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is
+updated by
 \begin{equation}
@@
 \newline
-is the process by which we update the base state through $\dt$ in time \MarginPar{we need a separate section for spherical}
-given the radial velocity $w_0^{\inp}.$ Here we discuss the algorithm
-for plane-parallel geometries.  The base state arrays are all
+is the process by which we update the base state through $\dt$ in
+time \MarginPar{we need a separate section for spherical} given the
+radial velocity $w_0^{\inp}.$ The base state arrays are all
 one-dimensional in the radial coordinate; we think of the base state
 quantities as defined at cell centers, with $w_0$ and $\beta_0$
-defined halfway in-between.\\ \\ a) For the density update, we
-discretize equation the base state density evolution equation
+defined halfway in-between.
+
+\begin{enumerate}
+\renewcommand{\theenumii}{{\bf \alph{enumii}}}
+
+\item plane-parallel
+
+\begin{enumerate}
+\item For the density update, we discretize equation the base state
+  density evolution equation
 \begin{eqnarray}
@@
 procedure described in paper II, Appendix A, and are saved for use
-later in the overall algorithm.\\ \\ b) For the pressure update, we
-discretize the base state pressure evolution equation,
+later in the overall algorithm.
+
+\item For the pressure update, we discretize the base state pressure
+  evolution equation,
 \begin{equation}
@@
 \end{eqnarray}
-where the interface states are again found via the procedure described in
-paper II, Appendix A.\\ \\
-c)  For the enthalpy update, \MarginPar{We need enthalpy equations now that we are predicting $(\rho h)'$.}
-\\
+where the interface states are again found via the procedure described
+in paper II, Appendix A.
+
+\item For the enthalpy update, \MarginPar{We need enthalpy equations
+  now that we are predicting $(\rho h)'$.}  
+\end{enumerate}
+
+\item spherical
+
+\MarginPar{we need to write the spherical description}
+
+\end{enumerate}
 
@@
 \end{enumerate}
+\MarginPar{need to add spherical}
 
@@
 \item Construct $w_0^{\nph,\star}$ by integrating
+
+\begin{enumerate}
+\renewcommand{\labelenumii}{\roman{enumii}.}
+\item plane-parallel
 \begin{equation}
@@
 \end{equation}
-For spherical geometries, this equation would be modified.
+
+\item spherical
+
+\MarginPar{need to add spherical}
+\end{enumerate}
 
@@
                    \rightarrow$ $[ \rho^{(1)}, (\rho h)^{(1)}, X_k^{(1)}, T^{(1)},
-                                  (\rho \omegadot_k)^{(1)} ].$
+                                  (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)} ].$
 
@@
 \item Define an edge-centered $\etarho^{\nph,\star}$ and cell-centered $\psi_j^{\nph,\star}$:
+
+\begin{enumerate}
+\renewcommand{\labelenumii}{\roman{enumii}.}
+
+\item plane-parallel
+
 \begin{eqnarray}
@@
 \end{eqnarray}
-For spherical geometries, $\psi_j^{\nph,\star}$ would have a different form.
+
+\item spherical
+
+\MarginPar{need to add spherical here}
+\end{enumerate}
 
@@
 $\rightarrow ( \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, 
-              (\rho \omegadot_k)^{(2),\star} ).$
+              (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ).$
 
@@
 \item Construct $w_0^{\nph}$ by integrating equation (\ref{eq:divw0}),
-\MarginPar{this is plane-parallel only}
+
+\begin{enumerate}
+\renewcommand{\labelenumii}{\roman{enumii}.}
+
+\item plane-parallel
+
 \MarginPar{Need dpdt\_factor stuff.}
@@
 $ p_0^{\nph,\star} = (p_0^{n} + p_0^{n+1,\star} ) / 2.$
+
+\item spherical
+
+\MarginPar{need to discuss spherical}
+
+\end{enumerate}
 
@@
       to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
-      We define $\rho^{(1),\nph} = \sum_k (\rho X_k)^{(1),\nph}.$
+      We define $\rho^{(1),\nph} = \sum_k (\rho X_k)^{(1),\nph}.$ \MarginPar{we split this into two now}
 
@@
 \item Define an edge-centered $ \etarho^{\nph}$ and a cell-centered $\psi_j^{\nph}$,
+
+\begin{enumerate}
+\renewcommand{\labelenumii}{\roman{enumii}.}
+
+\item plane-parallel
+
 \begin{eqnarray}
@@
 \end{eqnarray}
+
+\item spherical
+
+\end{enumerate}
 
@@
                    \rightarrow (\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, 
-                               (\rho \omegadot_k)^{(2)} ).$
+                               (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ).$
 

### 2008-10-12 docs/flowchart/paper.tex
switch over to a latex description for the individual steps

@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 1.} {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}
+
+\begin{description}
+
+\item[Step 1.] {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 2.} {\em Construct the provisional edge-based advective velocity}, $\uadvone$.\\
+\item[Step 2.] {\em Construct the provisional edge-based advective velocity}, $\uadvone$.\\
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 3.} {\em React the full state through the first time interval of $\dt / 2.$}
+\item[Step 3.] {\em React the full state through the first time interval of $\dt / 2.$}
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 4.} {\em Advect the base state, then the full state, through a time interval of $\dt.$}
+\item[Step 4.] {\em Advect the base state, then the full state, through a time interval of $\dt.$}
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 5.} {\em React the full state through a second time interval of $\dt / 2.$}
+\item[Step 5.] {\em React the full state through a second time interval of $\dt / 2.$}
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 6.} {\em Define a new average expansion rate at time $t^\nph.$}
+\item[Step 6.] {\em Define a new average expansion rate at time $t^\nph.$}
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 7.} {\em Construct the final edge-based advective velocity}, $\uadvtwo$:
+\item[Step 7.] {\em Construct the final edge-based advective velocity}, $\uadvtwo$:
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 8.} {\em Advect the base state, then the full state, through a time interval of $\dt.$}
+\item[Step 8.] {\em Advect the base state, then the full state, through a time interval of $\dt.$}
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 9.} {\em React the full state through a second time interval of $\dt / 2.$}
+\item[Step 9.] {\em React the full state through a second time interval of $\dt / 2.$}
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 10.} {\em Compute $S^{n+1}$ for the final projection.}
+\item[Step 10.] {\em Compute $S^{n+1}$ for the final projection.}
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 11.} {\em Update the velocity}.
+\item[Step 11.] {\em Update the velocity}.
 
@@
 %--------------------------------------------------------------------------
-\noindent {\bf Step 12.} {\em Compute a new $\dt.$}
+\item[Step 12.] {\em Compute a new $\dt.$}
 
@@
 $\ut^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.
+
+\end{description}
 

### 2008-10-12 docs/flowchart/paper.tex
more tex clean-ups

@@
 
-\begin{enumerate}
-\renewcommand{\theenumii}{{\bf \alph{enumii}}}
+\begin{description}
 
-\item plane-parallel
+\item[plane-parallel:] ~
 
 \begin{enumerate}
+\renewcommand{\labelenumii}{{\bf \alph{enumii}}.}
+
 \item For the density update, we discretize equation the base state
@@
 
-\item spherical
+\item[spherical:]
 
@@
 
-\end{enumerate}
+\end{description}
 
 \item {\bf Correct Base}$(\rho_0^{\inp} , \etarho^{\inp}) \rightarrow (\rho_0^{\outp})$
-is the process by which we adjust the base state density given $\etarho$ using
+is the process by which we adjust the base state density to account for
+large-scale mixing, through $\etarho$
+
+\begin{description}
+
+\item[plane-parallel:] We use
 \begin{eqnarray}
@@
 \end{eqnarray}
+
+\item[spherical:] ~
+
+\MarginPar{need to add spherical}
+
+\end{description}
+
 \end{enumerate}
-\MarginPar{need to add spherical}
 
@@
 
-\item Construct $w_0^{\nph,\star}$ by integrating
+\item Construct $w_0^{\nph,\star}$
 
-\begin{enumerate}
-\renewcommand{\labelenumii}{\roman{enumii}.}
-\item plane-parallel
+\begin{description}
+
+\item[plane-parallel:] We integrate
 \begin{equation}
@@
 
-\item spherical
+\item[spherical:] ~
 
 \MarginPar{need to add spherical}
-\end{enumerate}
+
+\end{description}
 
@@
 \item[Step 2.] {\em Construct the provisional edge-based advective velocity}, $\uadvone$.\\
+
+The local velocity field is described by
 
@@
 \end{eqnarray}
-
+\MarginPar{why $\rho^{(2'),\star}$ and not $\rho^{(2),\star}$?}
 \item {\bf Correct Base}$(\rho_0^{(2),\star}, \etarho^{\nmh}) \rightarrow (\rho_0^{n+1,\star})$.
@@
 
-\begin{enumerate}
-\renewcommand{\labelenumii}{\roman{enumii}.}
-
-\item plane-parallel
+\begin{description}
+\item[plane-parallel:]~
 
@@
 
-\item spherical
+\item[spherical:]~
 
 \MarginPar{need to add spherical here}
-\end{enumerate}
+\end{description}
 
@@
 
-\item {\bf React State}$( \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, 
-                             T^{(2),\star}, (\rho^{(2),\star} \Hext)) $\\
-$\rightarrow ( \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, 
-              (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ).$
+\item {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, 
+                             T^{(2),\star}, (\rho^{(2),\star} \Hext)] $\\
+$\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, 
+              (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
 
@@
 
-\item Construct $w_0^{\nph}$ by integrating equation (\ref{eq:divw0}),
+\item Construct $w_0^{\nph}$
 
-\begin{enumerate}
-\renewcommand{\labelenumii}{\roman{enumii}.}
+\begin{description}
 
-\item plane-parallel
+\item[plane-parallel:] We integrate equation (\ref{eq:divw0}),
 
@@
 
-\item spherical
+\item[spherical:]~
 
@@
 
-\end{enumerate}
+\end{description}
 
@@
 
-\begin{enumerate}
-\renewcommand{\labelenumii}{\roman{enumii}.}
+\begin{description}
 
-\item plane-parallel
+\item[plane-parallel:] ~
 
@@
 
-\item spherical
+\item[spherical:]~
 
-\end{enumerate}
+\end{description}
 
@@
 
-\item {\bf React State}$(\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho^{(2)} \Hext) )
-                   \rightarrow (\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, 
-                               (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ).$
+\item {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho^{(2)} \Hext) ]
+                   \rightarrow [\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, 
+                               (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$
 
@@
 
-  \item {\bf React State}$( \rho^{\initp},(\rho h)^{\initp}, X_k^{\initp}, T^{\initp}, (\rho^{\initp} \Hext) )$\\
-      $\rightarrow (\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{0,\nu} ).$
+  \item {\bf React State}$[ \rho^{\initp},(\rho h)^{\initp}, X_k^{\initp}, T^{\initp}, (\rho^{\initp} \Hext) ]$\\
+      $\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{0,\nu} ].$
 

### 2008-10-12 docs/flowchart/paper.tex
split the density and enthalpy updates

@@
 
-\item Compute the edge states, $(\rho X_k)^{(1),\nph,\star}$ and $(\rho h)^{(1),\nph,\star},$
-      for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
-      predict $\rho'$, $(\rho h)'$, and $X_k$ to the edges, using
-      a second-order Taylor expansion in space and time, as described in
-      paper II, Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  
-      We do not include the reaction terms in the enthalpy prediction, 
-      since we accounted for them already in {\bf React State}. 
-      We use the base state density, $\rho_0$, and base state enthalpy, $(\rho h)_0$,
-      to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
-      We define $\rho^{(1),\nph,\star} = \sum_k (\rho X_k)^{(1),\nph,\star}.$  \MarginPar{we actually split these now, and do $\rho$ first, compute $\psi$, and then do $(\rho h)$}
+\item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
+  for the conservative update of $(\rho X_k)$.  Here we predict
+  $\rho'$ and $X_k$ to the edges, using a second-order Taylor
+  expansion in space and time, as described in paper II, Appendix A,
+  using $\V = \uadvone+w_0^{\nph,\star} \er$.  We use the base state
+  density, $\rho_0$, to convert these to edge states for $(\rho X_k)$.
+  We define $\rho^{(1),\nph,\star} = \sum_k (\rho
+  X_k)^{(1),\nph,\star}.$
 
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ and
-$(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ without explicitly including the 
-reaction terms,
+\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$
+  without explicitly including the reaction terms,
 \begin{eqnarray}
@@
   (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip , \\ \nonumber  \\
-(\rho h)^{(2'),\star} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
-  (\rho h)^{(1),\nph,\star} \right) \right] \nonumber \\
-&& + \dt \; \left(\uadvone \cdotb \er\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} 
-+ \dt \; \psi^{\nmh} \enskip , \\
 \rho^{(2'),\star} &=& \sum_k (\rho X_k)^{(2),\star} \enskip , \\ 
@@
 \MarginPar{why $\rho^{(2'),\star}$ and not $\rho^{(2),\star}$?}
-\item {\bf Correct Base}$(\rho_0^{(2),\star}, \etarho^{\nmh}) \rightarrow (\rho_0^{n+1,\star})$.
 
@@
 \end{description}
+
+\item Compute the enthalpy edge state, $(\rho h)^{(1),\nph,\star},$
+  for the conservative update of $(\rho h).$ Here we predict $(\rho
+  h)'$ to the edges, using a second-order Taylor expansion in space
+  and time, as described in paper II, Appendix A, using $\V =
+  \uadvone+w_0^{\nph,\star} \er$.  We do not include the reaction
+  terms in the enthalpy prediction, since we accounted for them
+  already in {\bf React State}.  We use the base state enthalpy,
+  $(\rho h)_0$, to convert this to an edge state for $(\rho h)$.
+
+\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ without
+  explicitly including the reaction terms,
+\begin{eqnarray}
+(\rho h)^{(2'),\star} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
+  (\rho h)^{(1),\nph,\star} \right) \right] \nonumber \\
+&& + \dt \; \left(\uadvone \cdotb \er\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} 
++ \dt \; \psi^{\nmh} \enskip , 
+\end{eqnarray}
+
+\item {\bf Correct Base}$(\rho_0^{(2),\star}, \etarho^{\nmh}) \rightarrow (\rho_0^{n+1,\star})$.
 
@@
 
-\item Compute the edge states, $(\rho X_k)^{(1),\nph}$ and $(\rho h)^{(1),\nph},$
-      for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
-      predict $\rho'$, $(\rho h)'$, and $X_k$ to the edges, using
-      a second-order Taylor expansion in space and time, as described in
-      paper II, Appendix A, using $\V = \uadvtwo+w_0^{\nph} \er$. 
-      Again, we do not include the reaction terms in the enthalpy prediction 
-      since we accounted for them already in {\bf React State}.
-      We use the base state density, $\rho_0$, and base state enthalpy, $(\rho h)_0$,
-      to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
-      We define $\rho^{(1),\nph} = \sum_k (\rho X_k)^{(1),\nph}.$ \MarginPar{we split this into two now}
+\item Compute the species edge states, $(\rho X_k)^{(1),\nph}$, for
+  the conservative update of $(\rho X_k)$. Here we
+  predict $\rho'$ and $X_k$ to the edges, using a second-order Taylor
+  expansion in space and time, as described in paper II, Appendix A,
+  using $\V = \uadvtwo+w_0^{\nph} \er$.  We use the base state
+  density, $\rho_0$, to convert these to edge states for $(\rho X_k)$
+  and $(\rho h)$.  We define $\rho^{(1),\nph} = \sum_k (\rho
+  X_k)^{(1),\nph}$.
 
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ and
-$(\rho h)^{(1)} \rightarrow (\rho h)^{(2)},$ 
+\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ 
+without explicitly including the reaction terms,
 \begin{eqnarray}
@@
 (\rho X_k)^{(1),\nph} \right) \right]  \enskip , \\ \nonumber \\
-(\rho h)^{(2')} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
-(\rho h)^{(1),\nph} \right) \right] \nonumber \\
-&& + \frac{\dt}{2} \; \left(\uadvtwo \cdotb \er\right)
-\left [ \left(\frac{\partial p_0}{\partial r} \right)^{n}
-      + \left(\frac{\partial p_0}{\partial r} \right)^{n+1}  \right ] 
-+ \dt \; \psi^{\nph,\star} \enskip , \\ 
 \rho^{(2')} &=& \sum_k (\rho X_k)^{(2)} \enskip , \\
@@
 \end{eqnarray}
-
-\item {\bf Correct Base}$(\rho_0^{(2)}, \etarho^{\nph, \star}) \rightarrow (\rho_0^{n+1})$
 
@@
 \end{description}
+
+\item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, for the
+  conservative update of $(\rho h)$. Here we predict $(\rho h)'$ to
+  the edges, using a second-order Taylor expansion in space and time,
+  as described in paper II, Appendix A, using $\V =
+  \uadvtwo+w_0^{\nph} \er$.  Again, we do not include the reaction
+  terms in the enthalpy prediction since we accounted for them already
+  in {\bf React State}. We use the base state enthalpy, $(\rho h)_0$,
+  to convert these to edge states for $(\rho h)$.
+
+\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$ 
+without explicitly including the reaction terms,
+\begin{eqnarray}
+(\rho h)^{(2')} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
+(\rho h)^{(1),\nph} \right) \right] \nonumber \\
+&& + \frac{\dt}{2} \; \left(\uadvtwo \cdotb \er\right)
+\left [ \left(\frac{\partial p_0}{\partial r} \right)^{n}
+      + \left(\frac{\partial p_0}{\partial r} \right)^{n+1}  \right ] 
++ \dt \; \psi^{\nph,\star} \enskip ,
+\end{eqnarray}
+
+\item {\bf Correct Base}$(\rho_0^{(2)}, \etarho^{\nph, \star}) \rightarrow (\rho_0^{n+1})$
 

### 2008-10-12 docs/flowchart/paper.tex
update the scalar update

@@
 
+\item Update the species.  Here we do a difference approximation
+  to 
+  \begin{equation}
+  \frac{\partial (\rho X_k)}{\partial t} + \nablab \cdotb (\ubold \rho X_k) =
+         \rho {\omegadot}_k \enskip , \label{eq:species}
+  \end{equation}
+  neglecting the reaction terms.  The update consists of two steps:
+
+  \begin{enumerate}
+  \renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
+
   \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
@@
 
+  \end{enumerate}
+
 \item Define an edge-centered $\etarho^{\nph,\star}$ and cell-centered $\psi_j^{\nph,\star}$:
@@
 
+\item Update the enthalpy, using a difference approximation to
+
+  \begin{equation}
+  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\ubold \rho h)
++ \frac{Dp_0}{Dt} + \rho \Hnuc + \rho \Hext \enskip , \label{eq:rhohupd} 
+  \end{equation}
+  neglecting the reaction terms. \MarginPar{need to add diffusion.}
+
+  \begin{enumerate}
+  \renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
+
   \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph,\star},$
@@
   \begin{eqnarray}
-(\rho h)^{(2'),\star} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
-  (\rho h)^{(1),\nph,\star} \right) \right] \nonumber \\
-&& + \dt \; \left(\uadvone \cdotb \er\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} 
-+ \dt \; \psi^{\nmh} \enskip , 
+  (\rho h)^{(2'),\star} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab
+      \cdotb \left( (\uadvone+w_0^{\nph,\star} \er) (\rho
+      h)^{(1),\nph,\star} \right) \right] \nonumber \\ && + \dt \;
+    \left(\uadvone \cdotb \er\right) \left(\frac{\partial
+      p_0}{\partial r} \right)^{n} + \dt \; \psi^{\nmh} \enskip ,
   \end{eqnarray}
+  \end{enumerate}
 
@@
 
+\item Update the speces.  Here we do a difference approximation to
+  eq.~\ref{eq:species}, neglecting the reaction terms.  The
+  update consists of two steps:
+
+  \begin{enumerate}
+  \renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
+
   \item Compute the species edge states, $(\rho X_k)^{(1),\nph}$, for
-  the conservative update of $(\rho X_k)$. Here we
-  predict $\rho'$ and $X_k$ to the edges, using a second-order Taylor
-  expansion in space and time, as described in paper II, Appendix A,
-  using $\V = \uadvtwo+w_0^{\nph} \er$.  We use the base state
-  density, $\rho_0$, to convert these to edge states for $(\rho X_k)$
-  and $(\rho h)$.  We define $\rho^{(1),\nph} = \sum_k (\rho
+    the conservative update of $(\rho X_k)$. Here we predict $\rho'$
+    and $X_k$ to the edges, using a second-order Taylor expansion in
+    space and time, as described in paper II, Appendix A, using $\V =
+    \uadvtwo+w_0^{\nph} \er$.  We use the base state density,
+    $\rho_0$, to convert these to edge states for $(\rho X_k)$ and
+    $(\rho h)$.  We define $\rho^{(1),\nph} = \sum_k (\rho
     X_k)^{(1),\nph}$.
@@
 \end{eqnarray}
+
+   \end{enumerate}
 
@@
 \end{eqnarray}
+\MarginPar{this needs to be updated---we are not lagging things anymore.}
 
@@
 
-\item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, for the
-  conservative update of $(\rho h)$. Here we predict $(\rho h)'$ to
-  the edges, using a second-order Taylor expansion in space and time,
-  as described in paper II, Appendix A, using $\V =
+\item Update the enthalpy, using a difference approximation to
+  eq.~\ref{eq:rhohupd}, neglecting the reaction terms.
+
+  \begin{enumerate}
+  \renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
+
+  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, for
+    the conservative update of $(\rho h)$. Here we predict $(\rho h)'$
+    to the edges, using a second-order Taylor expansion in space and
+    time, as described in paper II, Appendix A, using $\V =
     \uadvtwo+w_0^{\nph} \er$.  Again, we do not include the reaction
-  terms in the enthalpy prediction since we accounted for them already
-  in {\bf React State}. We use the base state enthalpy, $(\rho h)_0$,
-  to convert these to edge states for $(\rho h)$.
+    terms in the enthalpy prediction since we accounted for them
+    already in {\bf React State}. We use the base state enthalpy,
+    $(\rho h)_0$, to convert these to edge states for $(\rho h)$.
 
-\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$ 
-without explicitly including the reaction terms,
+   \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$ without
+     explicitly including the reaction terms,
 \begin{eqnarray}
@@
 \end{eqnarray}
+   \end{enumerate}
+
+\MarginPar{we need to update the centering of $\psi$ in the enthalpy difference equation.}
 

### 2008-10-13 docs/flowchart/paper.tex
more MarginPars

@@
 (\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ 
+\MarginPar{shouldn't $h_0$ be in the interface?}
 \newline
@@
   \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
-  for the conservative update of $(\rho X_k)$.  Here we predict
-  $\rho'$ and $X_k$ to the edges, using a second-order Taylor
-  expansion in space and time, as described in paper II, Appendix A,
-  using $\V = \uadvone+w_0^{\nph,\star} \er$.  We use the base state
-  density, $\rho_0$, to convert these to edge states for $(\rho X_k)$.
-  We define $\rho^{(1),\nph,\star} = \sum_k (\rho
+  for the conservative update of $(\rho X_k)$. 
+  Here we predict $\rho'$ and $X_k$ to the edges, using a second-order
+  Taylor expansion in space and time, as described in paper II,
+  Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  We use the
+  base state density, $\rho_0$, to convert these to edge states for
+  $(\rho X_k)$.  We define $\rho^{(1),\nph,\star} = \sum_k (\rho
   X_k)^{(1),\nph,\star}.$
@@
     $(\rho h)_0$, to convert this to an edge state for $(\rho h)$.
+    \MarginPar{actually, there are lots of different prediction
+    methods that need to be listed here!}
 
   \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ without
-  explicitly including the reaction terms,
+  explicitly including the reaction terms, \MarginPar{shouldn't this be $(\rho h)^{(2'),\star}$?}
   \begin{eqnarray}
@@
   \end{eqnarray}
+  \MarginPar{we should also show the update in the case that $(\rho h)'$ is the flux}
   \end{enumerate}

### 2008-10-13 docs/flowchart/paper.tex
addressed some of Mike's marginpars, cleaned up all the time-lagging/centering of etarho and psi

@@
   (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip , \\ \nonumber  \\
-\rho^{(2'),\star} &=& \sum_k (\rho X_k)^{(2),\star} \enskip , \\ 
+\rho^{(2),\star} &=& \sum_k (\rho X_k)^{(2),\star} \enskip , \\ 
 X_k^{(2),\star} &=& (\rho X_k)^{(2),\star} / \rho^{(2),\star}
 \end{eqnarray}
-\MarginPar{why $\rho^{(2'),\star}$ and not $\rho^{(2),\star}$?}
 
@@
 
-  \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ without
-  explicitly including the reaction terms, \MarginPar{shouldn't this be $(\rho h)^{(2'),\star}$?}
+  \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2'),\star}$ without
+  explicitly including the reaction terms,
   \begin{eqnarray}
@@
     \left(\uadvone \cdotb \er\right) \left(\frac{\partial
-      p_0}{\partial r} \right)^{n} + \dt \; \psi^{\nmh} \enskip ,
+      p_0}{\partial r} \right)^{n} + \dt \; \psi^{\nph,*} \enskip ,
   \end{eqnarray}
@@
 
-\item {\bf Correct Base}$(\rho_0^{(2),\star}, \etarho^{\nmh}) \rightarrow (\rho_0^{n+1,\star})$.
+\item {\bf Correct Base}$(\rho_0^{(2),\star}, \etarho^{\nph,*}) \rightarrow (\rho_0^{n+1,\star})$.
 
@@
 Compute $\kth^{(1)}, c_p^{(1)}$, and $\xi_k^{(1)}$ from $\rho^{(1)}, T^{(1)}$, and $X_k^{(1)}$ as inputs to the equation of state.
-Note that density is constant in this step, i.e., $\rho^{(2'),\star}=\rho^{(2),\star}$.
-The update is given by
+The update is given by:
 \begin{eqnarray}
@@
 (\rho X_k)^{(1),\nph} \right) \right]  \enskip , \\ \nonumber \\
-\rho^{(2')} &=& \sum_k (\rho X_k)^{(2)} \enskip , \\
+\rho^{(2)} &=& \sum_k (\rho X_k)^{(2)} \enskip , \\
 X_k^{(2)} &=& (\rho X_k)^{(2)} / \rho^{(2)}
@@
  \nonumber \enskip , \\
- \psi_j^{\nph}  &=&  \frac{1}{2} \left(\eta_{rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
+ \psi_j^{\nph}  &=&  \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
 \end{eqnarray}
-\MarginPar{this needs to be updated---we are not lagging things anymore.}
 
@@
       + \left(\frac{\partial p_0}{\partial r} \right)^{n+1}  \right ] 
-+ \dt \; \psi^{\nph,\star} \enskip ,
++ \dt \; \psi^{\nph} \enskip ,
 \end{eqnarray}
+
 \end{enumerate}
 
-\MarginPar{we need to update the centering of $\psi$ in the enthalpy difference equation.}
-
-\item {\bf Correct Base}$(\rho_0^{(2)}, \etarho^{\nph, \star}) \rightarrow (\rho_0^{n+1})$
+\item {\bf Correct Base}$(\rho_0^{(2)}, \etarho^{\nph}) \rightarrow (\rho_0^{n+1})$
 
@@
 Compute $\kth^{(2),\star}, c_p^{(2),\star}$, and $\xi_k^{(2),\star}$, from $\rho^{(2),\star}, T^{(2),\star}$, and $X_k^{(2),\star}$ as inputs to the equation of state.
-Note that density is constant in the diffusion step, i.e., $\rho^{(2')}=\rho^{(2)}$.
-The update is given by
+The update is given by:
 \begin{eqnarray}
@@
 
-\item {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho^{(2)} \Hext) ]
-                   \rightarrow [\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, 
+\item {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho^{(2)} \Hext) ]$\\
+                   $\rightarrow [\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, 
                                (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$

### 2008-10-13 docs/flowchart/paper.tex
added dpdt factor part to w0 equations

@@
 \end{equation}
-using the lagged $\psi^{\nmh},$ \MarginPar{note: this is true only for plane-parallel}
-\MarginPar{Need dpdt\_factor stuff.}
+using the lagged $\psi^{\nmh}$ and the volume discrepancy correction with $f=0.3$:
 \begin{equation}
 \frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S}^{\nph,\star} 
-- \frac{1}{\gammabar^{n} p_0^{n}} \psi^{\nmh} \enskip . \nonumber
+- \frac{1}{\gammabar^{n} p_0^{n}}
+\left[ \psi^{\nmh} - f \left(\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right)\right]
+\enskip . \nonumber
 \end{equation}
@@
 
-\item[plane-parallel:] We integrate equation (\ref{eq:divw0}),
-
-\MarginPar{Need dpdt\_factor stuff.}
+\item[plane-parallel:] We integrate equation (\ref{eq:divw0}) with the 
+volume discrepancy correction with $f=0.3$:
 \begin{equation}
-\frac{\partial w_0^{\nph}}{\partial r} =  \overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}} \psi^{\nph,\star} \enskip , \nonumber
+\frac{\partial w_0^{\nph}}{\partial r} = 
+\overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
+\left[ \psi^{\nph,\star} - f \left(\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right)\right]
+\enskip , \nonumber
 \end{equation}
-for plane-parallel geometries, where 
-$\gammabar^{\nph,\star} = (\gammabar^{n} + \gammabar^{n+1,\star} ) / 2 $ and
-$ p_0^{\nph,\star} = (p_0^{n} + p_0^{n+1,\star} ) / 2.$
+\begin{equation}
+\gammabar^{\nph,\star} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}
+\end{equation}
+\begin{equation}
+p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}
+\end{equation}
+\begin{equation}
+\overline{p(\rho,h,X_k)^{\nph,\star}} = 
+\frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}
+\end{equation}
 

### 2008-10-13 docs/flowchart/paper.tex
added dpdt_factor equations everywhere and removed marginpars

@@
 If this is the first step of the calculation ($n=0$), we set
-\[
+\begin{equation}
 S^{\half,\star} = \frac{S^0 + S^1}{2} \enskip ,
-\]
+\end{equation}
 where $S^1$ is found through the iterative process that initializes the calculation.
@@
 \item Define
-\[
-\overline{S}^{\nph,\star} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip.
-\]
+\begin{equation}
+\overline{S}^{\nph,\star} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip .
+\end{equation}
 
@@
 \end{equation}
-using the lagged $\psi^{\nmh}$ and the volume discrepancy correction with $f=0.3$:
+using the lagged $\psi^{\nmh}$ and the volume discrepancy correction with $f=0.3$,
 \begin{equation}
@@
 \left[ \psi^{\nmh} - f \left(\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right)\right]
-\enskip . \nonumber
+\enskip .
 \end{equation}
@@
 \begin{equation}
-\int_{\Omega_H} \ut \cdotb \er \; dA = 0 \label{eq:udoterzero} 
+\int_{\Omega_H} \ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
 \end{equation}
@@
 \begin{equation}
-\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )
+\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )\enskip ,
 \label{eq:tildeconstraint}
 \end{equation}
-(paper III; eq.~[39]).
-In particular,
-\MarginPar{Need dpdt\_factor stuff.}
+(paper III; eq.~[39]).  The discretization with the volume discrepancy correction is
 \begin{equation}
-\nablab \cdotb \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
-\enskip ,
+\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
+\beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
++ \frac{f}{\gammabar^n p_0^n}
+\left(\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right) \enskip .
 \end{equation}
@@
 \begin{eqnarray}
-\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{(1)}, p_0^{n}, X_k^{(1)}) \right) 
-                                                                \enskip , \nonumber \\
+\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{(1)}, p_0^{n}, X_k^{(1)}) \right) \enskip , \\
  {\beta   }_0^{(1)}    &=& \beta   \left(\rho_0^{n}, p_0^{n}, \gammabar^{(1)}\right) \enskip .
@@
  - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
-  (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip , \\ \nonumber  \\
+  (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip , \nonumber \\ \\
 \rho^{(2),\star} &=& \sum_k (\rho X_k)^{(2),\star} \enskip , \\ 
@@
 \rho^{(1),\nph,\star} \right) - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} 
- \nonumber \enskip , \\
+ \enskip , \nonumber \\ \\
  \psi_j^{\nph,\star}  &=&  \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph,\star} + \eta_{\rho,j+\myhalf}^{\nph,\star}\right) g \enskip .
@@
   \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\ubold \rho h)
-+ \frac{Dp_0}{Dt} + \rho \Hnuc + \rho \Hext \enskip , \label{eq:rhohupd} 
++ \frac{Dp_0}{Dt} + \nabla\cdot\kth\nabla T + \rho \Hnuc + \rho \Hext 
+\enskip , \label{eq:rhohupd} 
   \end{equation}
-  neglecting the reaction terms. \MarginPar{need to add diffusion.}
+  neglecting the reaction terms.
 
@@
  \gammabar^{n+1,\star}    &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{n+1,\star}, p_0^{n+1,\star}, 
-                                                      X_k^{n+1,\star}) \right) \enskip , \nonumber \\
+                                                      X_k^{n+1,\star}) \right) \enskip , \\
  {\beta   }_0^{n+1,\star}    &=& \beta   \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}\right) \enskip .
@@
 Then define
-\[
+\begin{equation}
  S^\nph = \frac{S^n + S^{n+1,\star}}{2} \enskip. 
-\]
+\end{equation}
 
 \item Define:
-\[
+\begin{equation}
 \overline{S}^{\nph} = {\mathrm{\bf Avg}} (S^{\nph}) \enskip.
-\]
+\end{equation}
 
@@
 \item[plane-parallel:] We integrate equation (\ref{eq:divw0}) with the 
-volume discrepancy correction with $f=0.3$:
+volume discrepancy correction,
 \begin{equation}
@@
 \left[ \psi^{\nph,\star} - f \left(\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right)\right]
-\enskip , \nonumber
+\enskip ,
 \end{equation}
@@
 for $w_0$ and $\pizero.$  We note that $\uadvtwo$ satisfies the discrete versions of 
-equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}), in particular,
-\MarginPar{Need dpdt\_factor stuff.}
+equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}).   The discretization
+with the volume discrepancy correction is:
+\begin{eqnarray}
+\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) &=& 
+\beta_0^{\nph,\star} \left(S^{\nph} - \overline{S}^{\nph}\right) \nonumber \\
+&& + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
+\left(\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right)
+\enskip , \nonumber \\
+\end{eqnarray}
 \begin{equation}
-\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) = \beta_0^{\nph,\star} 
-\left(S^{\nph} - \overline{S}^{\nph}\right) \enskip .
+\beta_0^{\nph,\star} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2} \enskip .
 \end{equation}
-where $ \beta_0^{\nph,\star} = ( \beta_0^n +  \beta_0^{n+1,\star}) / 2.$\\
 
@@
  \etarho^{\nph}   &=&  {\rm {\bf Avg}} \left( (\uadvtwo \cdotb \er + w_0^{\nph}) 
-\rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} 
- \nonumber \enskip , \\
+\rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip , \\
  \psi_j^{\nph}  &=&  \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
@@
 \begin{eqnarray}
- \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left(\Gamma_1(\rho^{n+1}, p_0^{n+1}, {X_k}^{n+1}) \right) \enskip , \nonumber \\ 
+ \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left(\Gamma_1(\rho^{n+1}, p_0^{n+1}, {X_k}^{n+1}) \right) \enskip , \\ 
  {\beta   }_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip .
@@
 \item Define
-\[
+\begin{equation}
 \overline{S}^{n+1} = {\mathrm{\bf Avg}} (S^{n+1}) \enskip.
-\]
+\end{equation}
 
@@
 
-Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}),
-\MarginPar{Need dpdt\_factor stuff.}
-\[
-\nablab \cdotb \left(\beta_0^{\nph} \ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)\enskip,
-\]
-by defining ${\bf V} = \ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ and solving
+Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}) 
+including the volume discrepancy correction,
+\begin{equation}
+\nablab \cdotb \left(\beta_0^{\nph} \ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)
++ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
+\left(\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right)
+\enskip,
+\end{equation}
+by defining ${\bf V} = \ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
+and solving
 \begin{equation}
  L_\beta^\rho \phi =
-   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
+   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)
+- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
+\left(\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right) \enskip,
 \end{equation}
@@
 finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
+\MarginPar{Why are $\beta$ and $\rho$ time centered here instead using the new time value?  Does this have something to do with how $\phi$ relates to a lagged pressure?}
 In this step, $D$ is a discrete second-order operator that approximates the 
@@
 and the new time-centered perturbational pressure from:
-\[
+\begin{equation}
   \pi^\nph = \frac{1}{\dt} \phi \enskip .
-\]
+\end{equation}
 

### 2008-10-13 docs/flowchart/paper.tex
updated correct base to account for the fact that we pass in div etarho instead of etarho.  Also stated explicitly for plane-parallel what etarho is

@@
 
-\item {\bf Correct Base}$(\rho_0^{\inp} , \etarho^{\inp}) \rightarrow (\rho_0^{\outp})$
+\item {\bf Correct Base}
+$[\rho_0^{\inp}, (\nabla\cdot\etarho)^{\inp}] \rightarrow (\rho_0^{\outp})$
 is the process by which we adjust the base state density to account for
-large-scale mixing, through $\etarho$
-
-\begin{description}
-
-\item[plane-parallel:] We use
-\begin{eqnarray}
-\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
-\left( \eta_{\rho,j+\myhalf}^{\inp} - \eta_{\rho,j-\myhalf}^{\inp}  \right)  \enskip . 
-\end{eqnarray}
-
-\item[spherical:] ~
-
-\MarginPar{need to add spherical}
-
-\end{description}
+large-scale mixing, through $\etarho$.  For both plane-parallel and spherical
+we use
+\begin{equation}
+\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \dt (\nabla\cdot\etarho)^{\inp}_j \enskip . 
+\end{equation}
 
@@
 
-\item Define an edge-centered $\etarho^{\nph,\star}$ and cell-centered $\psi_j^{\nph,\star}$:
+\item Define an edge-centered $\etarho^{\nph,\star}$, cell-centered 
+$(\nabla\cdot\etarho)^{\nph,\star}$, and cell-centered $\psi_j^{\nph,\star}$:
 
@@
  \enskip , \nonumber \\ \\
+(\nabla\cdot\etarho)_j^{\nph,\star} &=& 
+\frac{\eta_{\rho,j+\myhalf}^{\nph,\star}-\eta_{\rho,j-\myhalf}^{\nph,\star}}{\Delta r} \enskip , \\
  \psi_j^{\nph,\star}  &=&  \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph,\star} + \eta_{\rho,j+\myhalf}^{\nph,\star}\right) g \enskip .
@@
 
-\item {\bf Correct Base}$(\rho_0^{(2),\star}, \etarho^{\nph,*}) \rightarrow (\rho_0^{n+1,\star})$.
+\item {\bf Correct Base}
+$[\rho_0^{(2),\star}, (\nabla\cdot\etarho)^{\nph,*}] \rightarrow (\rho_0^{n+1,\star})$.
 
@@
 
-\item Define an edge-centered $ \etarho^{\nph}$ and a cell-centered $\psi_j^{\nph}$,
+\item Define an edge-centered $\etarho^{\nph}$, cell-centered 
+$(\nabla\cdot\etarho)^{\nph}$, and a cell-centered $\psi_j^{\nph}$,
 
@@
  \etarho^{\nph}   &=&  {\rm {\bf Avg}} \left( (\uadvtwo \cdotb \er + w_0^{\nph}) 
-\rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip , \\
+\rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip , \nonumber \\ \\
+(\nabla\cdot\etarho)_j^{\nph} &=& 
+\frac{\eta_{\rho,j+\myhalf}^{\nph} - \eta_{\rho,j-\myhalf}^{\nph}}{\Delta r} \enskip , \\
  \psi_j^{\nph}  &=&  \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
@@
 
-\item {\bf Correct Base}$(\rho_0^{(2)}, \etarho^{\nph}) \rightarrow (\rho_0^{n+1})$
+\item {\bf Correct Base}
+$[\rho_0^{(2)}, (\nabla\cdot\etarho)^{\nph}] \rightarrow (\rho_0^{n+1})$
 

### 2008-10-13 docs/flowchart/paper.tex
some minor typos
in some places where we need separate spherical details, I'm going to refer to "paper IV" and use that as our working notes so we don't have to keep two documents updated.  Once we are more "final" in paper IV we can add it in here.

@@
   2008, ApJ, 684, 449 (henceforth paper~III)
+\item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection}
+A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale,
+  in perparation (henceforth paper~IV)
 \end{itemize}
@@
 \item {\bf Correct Base}
-$[\rho_0^{\inp}, (\nabla\cdot\etarho)^{\inp}] \rightarrow (\rho_0^{\outp})$
+$[\rho_0^{\inp}, (\nabla\cdot\etarho\er)^{\inp}] \rightarrow (\rho_0^{\outp})$
 is the process by which we adjust the base state density to account for
@@
 \begin{equation}
-\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \dt (\nabla\cdot\etarho)^{\inp}_j \enskip . 
+\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \dt (\nabla\cdot\etarho\er)^{\inp}_j \enskip . 
 \end{equation}
@@
 
-\MarginPar{need to add spherical}
+See Paper IV for working notes until we finalize this.
 
@@
 \item Define an edge-centered $\etarho^{\nph,\star}$, cell-centered 
-$(\nabla\cdot\etarho)^{\nph,\star}$, and cell-centered $\psi_j^{\nph,\star}$:
+$(\nabla\cdot\etarho\er)^{\nph,\star}$, and cell-centered $\psi_j^{\nph,\star}$:
 
@@
  \enskip , \nonumber \\ \\
-(\nabla\cdot\etarho)_j^{\nph,\star} &=& 
+(\nabla\cdot\etarho\er)_j^{\nph,\star} &=& 
 \frac{\eta_{\rho,j+\myhalf}^{\nph,\star}-\eta_{\rho,j-\myhalf}^{\nph,\star}}{\Delta r} \enskip , \\
@@
 \item {\bf Correct Base}
-$[\rho_0^{(2),\star}, (\nabla\cdot\etarho)^{\nph,*}] \rightarrow (\rho_0^{n+1,\star})$.
+$[\rho_0^{(2),\star}, (\nabla\cdot\etarho\er)^{\nph,*}] \rightarrow (\rho_0^{n+1,\star})$.
 
@@
 Compute $\kth^{(1)}, c_p^{(1)}$, and $\xi_k^{(1)}$ from $\rho^{(1)}, T^{(1)}$, and $X_k^{(1)}$ as inputs to the equation of state.
-The update is given by:
+The update is given by
 \begin{eqnarray}
@@
 \renewcommand{\theenumi}{{\bf \alph{enumi}}}
-\item Define:
+\item Define
 \begin{equation}
@@
 
-\item Define:
+\item Define
 \begin{equation}
@@
 %--------------------------------------------------------------------------
-\item[Step 7.] {\em Construct the final edge-based advective velocity}, $\uadvtwo$:
+\item[Step 7.] {\em Construct the final edge-based advective velocity}, $\uadvtwo$.
 
@@
 equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}).   The discretization
-with the volume discrepancy correction is:
+with the volume discrepancy correction is
 \begin{eqnarray}
@@
 \item Define an edge-centered $\etarho^{\nph}$, cell-centered 
-$(\nabla\cdot\etarho)^{\nph}$, and a cell-centered $\psi_j^{\nph}$,
+$(\nabla\cdot\etarho\er)^{\nph}$, and a cell-centered $\psi_j^{\nph}$,
 
@@
 \rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip , \nonumber \\ \\
-(\nabla\cdot\etarho)_j^{\nph} &=& 
+(\nabla\cdot\etarho\er)_j^{\nph} &=& 
 \frac{\eta_{\rho,j+\myhalf}^{\nph} - \eta_{\rho,j-\myhalf}^{\nph}}{\Delta r} \enskip , \\
@@
 \item {\bf Correct Base}
-$[\rho_0^{(2)}, (\nabla\cdot\etarho)^{\nph}] \rightarrow (\rho_0^{n+1})$
+$[\rho_0^{(2)}, (\nabla\cdot\etarho\er)^{\nph}] \rightarrow (\rho_0^{n+1})$
 
@@
 Compute $\kth^{(2),\star}, c_p^{(2),\star}$, and $\xi_k^{(2),\star}$, from $\rho^{(2),\star}, T^{(2),\star}$, and $X_k^{(2),\star}$ as inputs to the equation of state.
-The update is given by:
+The update is given by
 \begin{eqnarray}
@@
 \end{equation}
-and the new time-centered perturbational pressure from:
+and the new time-centered perturbational pressure from
 \begin{equation}

### 2008-10-13 docs/flowchart/paper.tex
the eta notes already exist in MAESTO/docs/eta_notes so I'm going to reference that document for the spherical part for now, they can be merged later.

@@
 \begin{equation}
-\overline{\phi}(r) = \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
+\overline{\phi}(r) 
+= \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
 \end{equation}
-here $\Omega_H$ is a layer at constant radius $r$. \MarginPar{we should discuss the implementation of average for spherical here}
+here $\Omega_H$ is a layer at constant radius $r$.
 
-\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow$
-$[\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \newline
-is the process by which we evolve the species and enthalpy from $X_k^{\inp} \rightarrow X_k^{\outp}$ 
-and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following system of equations 
-over a time interval of  $\Delta t/2$,
+For spherical problems, see Paper IV for details.
+
+\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \newline
+is the process by which we evolve the species and enthalpy from 
+$X_k^{\inp} \rightarrow X_k^{\outp}$ 
+and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
+system of equations over a time interval of  $\Delta t/2$,
 \begin{eqnarray}
@@
 \frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
-\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip. \label{eq:reacttemp}
+\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
+\enskip. \label{eq:reacttemp}
 \end{eqnarray}
@@
 \begin{equation}
-(\rho h)^{\outp} = (\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc) + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
+(\rho h)^{\outp} = 
+(\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc) + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
 \end{equation}
 
-\item {\bf Advect Base}$(\rho_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp} ) \rightarrow 
-(\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ 
+\item {\bf Advect Base}$(\rho_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp} ) \rightarrow (\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ 
 \MarginPar{shouldn't $h_0$ be in the interface?}
@@
 \left [ \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
-         \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ] \nonumber  \enskip,
+         \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ]
+         \nonumber \enskip,
 \end{eqnarray}
@@
 + w_{0,j-\myhalf}^\inp \right)
-\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\inp \enskip,
+\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\inp 
+\enskip,
 \end{eqnarray}
@@
 \begin{equation}
-\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip \label{eq:divw0} ,
+\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \etarho g 
+\enskip \label{eq:divw0} ,
 \end{equation}
@@
 - \frac{1}{\gammabar^{n} p_0^{n}}
-\left[ \psi^{\nmh} - f \left(\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right)\right]
+\left[ \psi^{\nmh} - 
+f \left(\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right)\right]
 \enskip .
@@
 
-\item[spherical:] ~
-
-See Paper IV for working notes until we finalize this.
+\item[spherical:] See Paper IV for working notes until we finalize this.
 
@@
 \begin{equation}
-- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
+- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} 
+= \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} 
+\enskip, \label{eq:pizero}
 \end{equation}
@@
 \begin{equation}
-w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\end{equation}
+w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}} 
+\enskip,\end{equation}
 \begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
-+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\end{equation}
+\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = 
+\frac{1}{\dt^n+\dt^{n-1} } 
+\left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
++ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] 
+\enskip.\end{equation}
 If $n=0$, we use $\dt^{-1} = \dt^0$.
@@
 \begin{eqnarray}
-\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{(1)}, p_0^{n}, X_k^{(1)}) \right) \enskip , \\
+\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{(1)}, p_0^{n}, X_k^{(1)}) 
+\right) \enskip , \\
  {\beta   }_0^{(1)}    &=& \beta   \left(\rho_0^{n}, p_0^{n}, \gammabar^{(1)}\right) \enskip .
@@
 %--------------------------------------------------------------------------
-\item[Step 4.] {\em Advect the base state, then the full state, through a time interval of $\dt.$}
+\item[Step 4.] {\em Advect the base state, then the full state, through a time interval 
+of $\dt.$}
 
@@
 
-\item[spherical:]~
+\item[spherical:] See {\tt MAESTO/docs/eta\_notes} for working notes until we finlize 
+the details.
 
-\MarginPar{need to add spherical here}
 \end{description}
@@
 
-\item[spherical:]~
-
-\MarginPar{need to discuss spherical}
+\item[spherical:] See Paper IV for working notes until we finalize this.
 

### 2008-10-13 docs/flowchart/paper.tex
fixed the density advance section so it more accurately describes what the code does

@@
   for the conservative update of $(\rho X_k)$. 
-  Here we predict $\rho'$ and $X_k$ to the edges, using a second-order
+  Here we predict $\rho'$ and $X_k$ to the edges to obtain 
+  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
   Taylor expansion in space and time, as described in paper II,
-  Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  We use the
-  base state density, $\rho_0$, to convert these to edge states for
-  $(\rho X_k)$.  We define $\rho^{(1),\nph,\star} = \sum_k (\rho
-  X_k)^{(1),\nph,\star}.$
+  Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.
 
@@
 \begin{eqnarray}
+\rho^{(1),\nph,\star} &=& 
+\left[\rho^{'(1),\nph,\star} + \left(\frac{\rho_0^n + \rho_0^{(2),\star}}{2}\right)\right] 
+\enskip , \\
 (\rho X_k)^{(2),\star} &=& (\rho X_k)^{(1)} 
@@
 
-\begin{eqnarray}
- \etarho^{\nph,\star}   &=&  {\rm {\bf Avg}} \left( (\uadvone \cdotb \er + w_0^{\nph,\star}) 
-\rho^{(1),\nph,\star} \right) - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} 
- \enskip , \nonumber \\ \\
-(\nabla\cdot\etarho\er)_j^{\nph,\star} &=& 
-\frac{\eta_{\rho,j+\myhalf}^{\nph,\star}-\eta_{\rho,j-\myhalf}^{\nph,\star}}{\Delta r} \enskip , \\
- \psi_j^{\nph,\star}  &=&  \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph,\star} + \eta_{\rho,j+\myhalf}^{\nph,\star}\right) g \enskip .
-\end{eqnarray}
+\begin{equation}
+ \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ (\uadvone \cdotb \er + w_0^{\nph,\star}) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
+\end{equation}
+\begin{equation}
+(\nabla\cdot\etarho\er)_j^{\nph,\star} = 
+\frac{\eta_{\rho,j+\myhalf}^{\nph,\star}-\eta_{\rho,j-\myhalf}^{\nph,\star}}{\Delta r} \enskip ,
+\end{equation}
+\begin{equation}
+ \psi_j^{\nph,\star} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph,\star} + \eta_{\rho,j+\myhalf}^{\nph,\star}\right) g \enskip .
+\end{equation}
 
@@
     the conservative update of $(\rho X_k)$.  Here we predict $\rho'$ 
-    and $X_k$ to the edges, using a second-order Taylor expansion in
+    and $X_k$ to the edges to obtain $\rho^{'(1),\nph}$ and 
+    $X_k^{(1),\nph}$ using a second-order Taylor expansion in
     space and time, as described in paper II, Appendix A, using $\V =
-    \uadvtwo+w_0^{\nph} \er$.  We use the base state density,
-    $\rho_0$, to convert these to edge states for $(\rho X_k)$ and
-    $(\rho h)$.  We define $\rho^{(1),\nph} = \sum_k (\rho
-    X_k)^{(1),\nph}$.
+    \uadvtwo+w_0^{\nph} \er$.
 
@@
 \begin{eqnarray}
+\rho^{(1),\nph} &=& 
+\left[\rho^{'(1),\nph} + \left(\frac{\rho_0^n + \rho_0^{(2)}}{2}\right)\right] 
+\enskip , \\
 (\rho X_k)^{(2)} &=& (\rho X_k)^{(1)} 
@@
 
-\begin{eqnarray}
- \etarho^{\nph}   &=&  {\rm {\bf Avg}} \left( (\uadvtwo \cdotb \er + w_0^{\nph}) 
-\rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip , \nonumber \\ \\
-(\nabla\cdot\etarho\er)_j^{\nph} &=& 
-\frac{\eta_{\rho,j+\myhalf}^{\nph} - \eta_{\rho,j-\myhalf}^{\nph}}{\Delta r} \enskip , \\
- \psi_j^{\nph}  &=&  \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
-\end{eqnarray}
+\begin{equation}
+ \etarho^{\nph} = {\rm {\bf Avg}} \sum_k \left [\uadvtwo \cdotb \er + w_0^{\nph}) (\rho X_k)^{(1),\nph} \right] - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip ,
+\end{equation}
+\begin{equation}
+(\nabla\cdot\etarho\er)_j^{\nph} = \frac{\eta_{\rho,j+\myhalf}^{\nph} - \eta_{\rho,j-\myhalf}^{\nph}}{\Delta r} \enskip , \\
+\end{equation}
+\begin{equation}
+ \psi_j^{\nph} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
+\end{equation}
 

### 2008-10-13 docs/flowchart/paper.tex
improved the enthalpy update notes

@@
 \end{equation}
+
+\MarginPar{Need React Base?}
 
@@
   Taylor expansion in space and time, as described in paper II,
-  Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.
+  Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  We use the base
+  state density, $\rho_0$, to convert this to an edge state for $\rho$,
+
+\begin{equation}
+\rho^{(1),\nph,\star} = 
+\rho^{'(1),\nph,\star} + \left(\frac{\rho_0^n + \rho_0^{(2),\star}}{2}\right) \enskip ,
+\end{equation}
 
@@
   without explicitly including the reaction terms,
-\begin{eqnarray}
-\rho^{(1),\nph,\star} &=& 
-\left[\rho^{'(1),\nph,\star} + \left(\frac{\rho_0^n + \rho_0^{(2),\star}}{2}\right)\right] 
-\enskip , \\
-(\rho X_k)^{(2),\star} &=& (\rho X_k)^{(1)} 
+\begin{equation}
+(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)} 
  - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
-  (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip , \nonumber \\ \\
-\rho^{(2),\star} &=& \sum_k (\rho X_k)^{(2),\star} \enskip , \\ 
-X_k^{(2),\star} &=& (\rho X_k)^{(2),\star} / \rho^{(2),\star}
-\end{eqnarray}
+  (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip ,
+\end{equation}
+\begin{equation}
+\rho^{(2),\star} = \sum_k (\rho X_k)^{(2),\star} \enskip ,
+\end{equation}
+\begin{equation}
+X_k^{(2),\star} = (\rho X_k)^{(2),\star} / \rho^{(2),\star}
+\end{equation}
 
@@
     for the conservative update of $(\rho h).$  Here we predict $(\rho
-    h)'$ to the edges, using a second-order Taylor expansion in space
+    h)'$ to the edges to obtain $(\rho h)^{'(1),\nph,\star}$, 
+    using a second-order Taylor expansion in space
     and time, as described in paper II, Appendix A, using $\V =
@@
     already in {\bf React State}.  We use the base state enthalpy,
-    $(\rho h)_0$, to convert this to an edge state for $(\rho h)$.
-    \MarginPar{actually, there are lots of different prediction
-    methods that need to be listed here!}
+    $(\rho h)_0$, to convert this to an edge state for $(\rho h)$,
+    \MarginPar{in these notes we currently assume pred type 1, $(\rho h)'$}
+\begin{equation}
+(\rho h)^{(1),\nph,\star} = 
+(\rho h)^{'(1),\nph,\star} + \left(\frac{(\rho h)_0^n + (\rho h)_0^{(2),\star}}{2}\right)
+\enskip .
+\end{equation}
 
@@
     space and time, as described in paper II, Appendix A, using $\V =
-    \uadvtwo+w_0^{\nph} \er$.
+    \uadvtwo+w_0^{\nph} \er$.  We use the base
+    state density, $\rho_0$, to convert this to an edge state for $\rho$,
+
+\begin{equation}
+\rho^{(1),\nph} = \rho^{'(1),\nph} + \left(\frac{\rho_0^n + \rho_0^{(2)}}{2}\right)\enskip ,
+\end{equation}
 
@@
       without explicitly including the reaction terms,
-\begin{eqnarray}
-\rho^{(1),\nph} &=& 
-\left[\rho^{'(1),\nph} + \left(\frac{\rho_0^n + \rho_0^{(2)}}{2}\right)\right] 
-\enskip , \\
-(\rho X_k)^{(2)} &=& (\rho X_k)^{(1)} 
+\begin{equation}
+(\rho X_k)^{(2)} = (\rho X_k)^{(1)} 
 - \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
-(\rho X_k)^{(1),\nph} \right) \right]  \enskip , \\ \nonumber \\
-\rho^{(2)} &=& \sum_k (\rho X_k)^{(2)} \enskip , \\
-X_k^{(2)} &=& (\rho X_k)^{(2)} / \rho^{(2)}
-\end{eqnarray}
+(\rho X_k)^{(1),\nph} \right) \right]  \enskip ,
+\end{equation}
+\begin{equation}
+\rho^{(2)} = \sum_k (\rho X_k)^{(2)} \enskip ,
+\end{equation}
+\begin{equation}
+X_k^{(2)} = (\rho X_k)^{(2)} / \rho^{(2)}
+\end{equation}
 
@@
     the conservative update of $(\rho h)$. Here we predict $(\rho h)'$
-    to the edges, using a second-order Taylor expansion in space and
+    to the edges to obtain $(\rho h)^{'(1),\nph}$, 
+    using a second-order Taylor expansion in space and
     time, as described in paper II, Appendix A, using $\V =
@@
     already in {\bf React State}. We use the base state enthalpy,
-    $(\rho h)_0$, to convert these to edge states for $(\rho h)$.
+    $(\rho h)_0$, to convert these to edge states for $(\rho h)$,
+\begin{equation}
+(\rho h)^{(1),\nph} = 
+(\rho h)^{'(1),\nph} + \left(\frac{(\rho h)_0^n + (\rho h)_0^{(2)}}{2}\right) \enskip .
+\end{equation}
 

### 2008-10-13 docs/flowchart/paper.tex
addressed more marginpars

@@
 
-\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \newline
+\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
 is the process by which we evolve the species and enthalpy from 
@@
 
-\MarginPar{Need React Base?}
-
-\item {\bf Advect Base}$(\rho_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp} ) \rightarrow (\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ 
-\MarginPar{shouldn't $h_0$ be in the interface?}
-\newline
+\item {\bf React Base}$[(\rho h)_0^{\inp}] \rightarrow [(\rho h)_0^{\outp}]$
+
+\item {\bf Advect Base}$[\rho_0^{\inp}, (\rho h)_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp}] \rightarrow (\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ \\
 is the process by which we update the base state through $\dt$ in
@@
 \item For the density update, we discretize equation the base state
-  density evolution equation
+  density evolution equation (paper III; eq.~[29]) without the $\etarho$ term,
 \begin{eqnarray}
 \frac{\partial \rhozero}{\partial t} &=& - \nablab \cdotb \left( \rhozero w_0 \er \right)
-- \nablab \cdotb \left( \etarho \er \right) \enskip ,
+\enskip ,
 \label{eq:rho0upd_new}
 \end{eqnarray}
-(paper III; eq.~[29]), neglecting the $\etarho$ term, to compute the
-new base state density,
+to compute the new base state density,
 \begin{eqnarray}
@@
 \item For the pressure update, we discretize the base state pressure
-  evolution equation,
-\begin{equation}
-\frac{\partial p_0}{\partial t} = w_0 \rho_0 g  + \etarho g
-                                 = - w_0 \frac{\partial p_0}{\partial r} + \etarho g \enskip .
+  evolution equation (paper III; eq.~[31]),
+\begin{equation}
+\frac{\partial p_0}{\partial t} = 
+-w_0 \frac{\partial p_0}{\partial r} + \psi \enskip ,
 \label{eq:p0upd_new}
 \end{equation}
-(paper III; eq.~[31]) to compute the new base state pressure,
-\begin{eqnarray}
-p_{0,j}^\outp &=& p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+to compute the new base state pressure,
+\begin{equation}
+p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
 + w_{0,j-\myhalf}^\inp \right)
@@
 \enskip,
-\end{eqnarray}
+\end{equation}
 where the interface states are again found via the procedure described
-in paper II, Appendix A.
-
-\item For the enthalpy update, \MarginPar{We need enthalpy equations
-  now that we are predicting $(\rho h)'$.}  
+in paper II, Appendix A.  Note that the $\psi$ term is not included in the forcing for
+the interface state calculation, but is used in the update.
+
+\item For the enthalpy update, we discretize
+\begin{equation}
+\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\er\right] + \psi \enskip ,
+\end{equation}
+to compute the new base state enthalpy,
+\begin{equation}
+(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
+\left [ \left( {(\rho h)_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
+         \left( {(\rho h)_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ] 
++ \dt\psi_j^{\inp} \enskip ,
+\end{equation}
+where the interface states are again found via the procedure described
+in paper II, Appendix A.  Note that the $\psi$ term is not included in 
+the forcing for the interface state calculation, but is used in the update.
+
 \end{enumerate}
@@
 \item {\bf Correct Base}
-$[\rho_0^{\inp}, (\nabla\cdot\etarho\er)^{\inp}] \rightarrow (\rho_0^{\outp})$
+$[\rho_0^{\inp}, (\nabla\cdot\etarho\er)^{\inp}] \rightarrow (\rho_0^{\outp})$\\
 is the process by which we adjust the base state density to account for
@@
 \begin{enumerate}
+\MarginPar{Need React Base}
 \renewcommand{\theenumi}{{\bf \alph{enumi}}}
@@
 
-\item {\bf Advect Base}$(\rhozero^{n}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph,\star}, \psi^{\nmh}) \rightarrow $
-$(\rho_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}, p_0^{n+1,\star}).$
+\item {\bf Advect Base}$(\rhozero^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph,\star}, \psi^{\nmh}) \rightarrow $
+$(\rho_0^{(2),\star}, (\rho h)_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}, p_0^{n+1,\star}).$
 
@@
 
-\item {\bf Advect Base}$(\rho_0^{n}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph}, \psi^{\nph,\star} ) \rightarrow $
-$(\rho_0^{(2)}, \rho_0^{\nph,{\rm pred}}, p_0^{n+1}).$
+\item {\bf Advect Base}$(\rho_0^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph}, \psi^{\nph,\star} ) \rightarrow $
+$(\rho_0^{(2)}, (\rho h)_0^{(2)}, \rho_0^{\nph,{\rm pred}}, p_0^{n+1}).$
 

### 2008-10-13 docs/flowchart/paper.tex
Made a noteworthy notational change:
\newcommand{\uadvone}{\widetilde{\mathbf{U}}^{\mathrm{ADV},(1)}}
\newcommand{\uadvtwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV},(2)}}

has now become

\newcommand{\uadvone}{\widetilde{\mathbf{U}}^{\mathrm{ADV},\star}}
\newcommand{\uadvtwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV}}}

I really think this change has to be made because

the ^{n}, ^{(1)}, ^{(2)}, ^{n+1} notation refers to different points in the splitting scheme.

The ^{\star} vs ^{no star} notation has to do with the predictor/corrector algorithm.

The old definitions of \uadveone and \uadvtwo was mixi

@@
 
-\newcommand{\uadvone}{\widetilde{\mathbf{U}}^{\mathrm{ADV},(1)}}
-\newcommand{\uadvtwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV},(2)}}
+\newcommand{\uadvone}{\widetilde{\mathbf{U}}^{\mathrm{ADV},\star}}
+\newcommand{\uadvtwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV}}}
 \newcommand{\V}{\mathbf{V}}
@@
 
-\item {\bf React Base}$[(\rho h)_0^{\inp}] \rightarrow [(\rho h)_0^{\outp}]$
+\item {\bf React Base}$[(\rho h)_0^{\inp},(\overline{\rho\Hnuc})^{\inp},(\overline{\rho\Hext})^{\inp}] \rightarrow [(\rho h)_0^{\outp}]$\\
+is the process by which we update the base state enthalpy through $\dt/2$ in time using
+\begin{equation}
+\frac{\partial(\rho h)_0}{\partial t} = (\overline{\rho\Hnuc})^{\inp} + (\overline{\rho\Hext})^{\inp}\enskip ,
+\end{equation}
+discretized as
+\begin{equation}
+(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} + \frac{\dt}{2}(\overline{\rho\Hnuc})_{j}^{\inp} + \frac{\dt}{2}(\overline{\rho\Hext})_{j}^{\inp} \enskip .
+\end{equation}
 
@@
 %--------------------------------------------------------------------------
-\item[Step 3.] {\em React the full state through the first time interval of $\dt / 2.$}
-
-\begin{enumerate}
-\MarginPar{Need React Base}
+\item[Step 3.] {\em React the full state and base state through the first time 
+interval of $\dt / 2.$}
+
+\begin{enumerate}
 \renewcommand{\theenumi}{{\bf \alph{enumi}}}
 
-\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho^n \Hext)]
+\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n]
                    \rightarrow$ $[ \rho^{(1)}, (\rho h)^{(1)}, X_k^{(1)}, T^{(1)},
                                   (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)} ].$
+
+\item {\bf React Base}$[(\rho h)_0^n,(\overline{\rho \Hnuc})^{(1)},(\overline{\rho\Hext})^n] \rightarrow [(\rho h)_0^{(1)}]$
 
@@
 \begin{enumerate}
-\MarginPar{Need React Base}
 \renewcommand{\theenumi}{{\bf \alph{enumi}}}
@@
 \item {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, 
-                             T^{(2),\star}, (\rho^{(2),\star} \Hext)] $\\
+                             T^{(2),\star}, (\rho\Hext)^{(2),\star}] $\\
 $\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, 
@@
 
-\item Update the speces.  Here we do a difference approximation to
+\item Update the species.  Here we do a difference approximation to
   eq.~\ref{eq:species}, neglecting the reaction terms.  The
@@
 %--------------------------------------------------------------------------
-\item[Step 9.] {\em React the full state through a second time interval of $\dt / 2.$}
-
-\begin{enumerate}
-\MarginPar{Need React Base}
+\item[Step 9.] {\em React the full state and base state through a second time 
+interval of $\dt / 2.$}
+
+\begin{enumerate}
 \renewcommand{\theenumi}{{\bf \alph{enumi}}}
@@
                                (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$
+
+\item {\bf React Base}$[(\rho h)_0^{(2)},(\overline{\rho \Hnuc})^{(2)},(\overline{\rho\Hext})^{(2)}] \rightarrow [(\rho h)_0^{n+1}]$
 

### 2008-10-13 Source/advance.f90

@@ -647,16 +647,20 @@ contains
     do n=1,nlevs
        call destroy(s2(n))
     end do
-
-    if (evolve_base_state .and. full_rhoh0_evolution) then
-       if (parallel_IOProcessor() .and. verbose .ge. 1) then
-          write(6,*) '            : react  base >>> '
+    
+    ! if we are doing the full algorithm, there is no need to call react_base since
+    ! rhoh0_new isn't actually used
+    if (do_half_alg) then
+       if (evolve_base_state .and. full_rhoh0_evolution) then
+          if (parallel_IOProcessor() .and. verbose .ge. 1) then
+             write(6,*) '            : react  base >>> '
+          end if
+          call average(mla,rho_Hnuc2,rho_Hnucbar,dx,1)
+          call average(mla,rho_Hext,rho_Hextbar,dx,1)
+          call react_base(rhoh0_2,rho_Hnucbar,rho_Hextbar,halfdt,rhoh0_new)
+       else
+          rhoh0_new = rhoh0_2
        end if
-       call average(mla,rho_Hnuc2,rho_Hnucbar,dx,1)
-       call average(mla,rho_Hext,rho_Hextbar,dx,1)
-       call react_base(rhoh0_2,rho_Hnucbar,rho_Hextbar,halfdt,rhoh0_new)
-    else
-       rhoh0_new = rhoh0_2
     end if
 
     if (evolve_base_state) then

### 2008-10-23 docs/flowchart/paper.tex
equation cleanup only

@@
 
-\item {\bf Advect Base}$[\rho_0^{\inp}, (\rho h)_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp}] \rightarrow (\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ \\
+\item {\bf Advect Base}$[\rho_0^{\inp}, (\rho h)_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp}] \rightarrow [\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp}]$ \\
 is the process by which we update the base state through $\dt$ in
@@
 (\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
-\left [ \left( {(\rho h)_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
-         \left( {(\rho h)_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ] 
+\left\{ \left[ {(\rho h)_0}^{\outp,\nph} {w_0}^{\inp}\right]_{j+\myhalf} -  
+         \left[ {(\rho h)_0}^{\outp,\nph} {w_0}^{\inp}\right]_{j-\myhalf} \right\}
 + \dt\psi_j^{\inp} \enskip ,
@@
 \item {\bf Correct Base}
-$[\rho_0^{\inp}, (\nabla\cdot\etarho\er)^{\inp}] \rightarrow (\rho_0^{\outp})$\\
+$[\rho_0^{\inp}, (\nabla\cdot\etarho\er)^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
 is the process by which we adjust the base state density to account for
@@
 - \frac{1}{\gammabar^{n} p_0^{n}}
-\left[ \psi^{\nmh} - 
-f \left(\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right)\right]
+\left\{ \psi^{\nmh} - 
+f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right]\right\}
 \enskip .
@@
 + \frac{f}{\gammabar^n p_0^n}
-\left(\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right) \enskip .
+\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right] \enskip .
 \end{equation}
@@
 \begin{eqnarray}
-\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{(1)}, p_0^{n}, X_k^{(1)}) 
-\right) \enskip , \\
+\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left[ \Gamma_1(\rho^{(1)}, p_0^{n}, X_k^{(1)}) \right] \enskip , \\
  {\beta   }_0^{(1)}    &=& \beta   \left(\rho_0^{n}, p_0^{n}, \gammabar^{(1)}\right) \enskip .
@@
 
-\item {\bf Advect Base}$(\rhozero^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph,\star}, \psi^{\nmh}) \rightarrow $
-$(\rho_0^{(2),\star}, (\rho h)_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}, p_0^{n+1,\star}).$
+\item {\bf Advect Base}$[\rhozero^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph,\star}, \psi^{\nmh}] \rightarrow $
+$[\rho_0^{(2),\star}, (\rho h)_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}, p_0^{n+1,\star}].$
 
@@
 \rho^{(1),\nph,\star} = 
-\rho^{'(1),\nph,\star} + \left(\frac{\rho_0^n + \rho_0^{(2),\star}}{2}\right) \enskip ,
+\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2),\star}}{2} \enskip ,
 \end{equation}
@@
 (\rho X_k)^{(2),\star} = (\rho X_k)^{(1)} 
- - \dt \; \left[ \nablab \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  
-  (\rho X_k)^{(1),\nph,\star} \right) \right] \enskip ,
+ - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right)  
+  (\rho X_k)^{(1),\nph,\star} \right] \right\} \enskip ,
 \end{equation}
@@
 \begin{equation}
- \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ (\uadvone \cdotb \er + w_0^{\nph,\star}) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
+ \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdotb \er + w_0^{\nph,\star}\right) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
 \end{equation}
@@
 (\rho h)^{(1),\nph,\star} = 
-(\rho h)^{'(1),\nph,\star} + \left(\frac{(\rho h)_0^n + (\rho h)_0^{(2),\star}}{2}\right)
+(\rho h)^{'(1),\nph,\star} + \frac{(\rho h)_0^n + (\rho h)_0^{(2),\star}}{2}
 \enskip .
@@
   \begin{eqnarray}
-  (\rho h)^{(2'),\star} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab
-      \cdotb \left( (\uadvone+w_0^{\nph,\star} \er) (\rho
-      h)^{(1),\nph,\star} \right) \right] \nonumber \\ && + \dt \;
+  (\rho h)^{(2'),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
+      \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right) (\rho
+      h)^{(1),\nph,\star} \right] \right\} \nonumber \\ && + \dt \;
     \left(\uadvone \cdotb \er\right) \left(\frac{\partial
@@
 \item {\bf Correct Base}
-$[\rho_0^{(2),\star}, (\nabla\cdot\etarho\er)^{\nph,*}] \rightarrow (\rho_0^{n+1,\star})$.
+$[\rho_0^{(2),\star}, (\nabla\cdot\etarho\er)^{\nph,*}] \rightarrow [\rho_0^{n+1,\star}]$.
 
@@
 \begin{eqnarray}
- \gammabar^{n+1,\star}    &=& {\rm{\bf Avg}} \left( \Gamma_1(\rho^{n+1,\star}, p_0^{n+1,\star}, 
-                                                      X_k^{n+1,\star}) \right) \enskip , \\
+ \gammabar^{n+1,\star}    &=& {\rm{\bf Avg}} \left[ \Gamma_1(\rho^{n+1,\star}, p_0^{n+1,\star}, 
+                                                      X_k^{n+1,\star}) \right] \enskip , \\
  {\beta   }_0^{n+1,\star}    &=& \beta   \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}\right) \enskip .
@@
 \overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
-\left[ \psi^{\nph,\star} - f \left(\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right)\right]
+\left\{ \psi^{\nph,\star} - f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\}
 \enskip ,
@@
 && + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
-\left(\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right)
+\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]
 \enskip , \nonumber \\
@@
 \item {\bf Advect Base}$(\rho_0^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph}, \psi^{\nph,\star} ) \rightarrow $
-$(\rho_0^{(2)}, (\rho h)_0^{(2)}, \rho_0^{\nph,{\rm pred}}, p_0^{n+1}).$
+$[\rho_0^{(2)}, (\rho h)_0^{(2)}, \rho_0^{\nph,{\rm pred}}, p_0^{n+1}].$
 
@@
 \begin{equation}
-\rho^{(1),\nph} = \rho^{'(1),\nph} + \left(\frac{\rho_0^n + \rho_0^{(2)}}{2}\right)\enskip ,
+\rho^{(1),\nph} = \rho^{'(1),\nph} + \frac{\rho_0^n + \rho_0^{(2)}}{2}\enskip ,
 \end{equation}
@@
 (\rho X_k)^{(2)} = (\rho X_k)^{(1)} 
-- \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
-(\rho X_k)^{(1),\nph} \right) \right]  \enskip ,
+- \dt \; \left\{ \nablab \cdotb \left[\left(\uadvtwo+w_0^{\nph} \er\right)  
+(\rho X_k)^{(1),\nph} \right] \right\}  \enskip ,
 \end{equation}
@@
 \begin{equation}
- \etarho^{\nph} = {\rm {\bf Avg}} \sum_k \left [\uadvtwo \cdotb \er + w_0^{\nph}) (\rho X_k)^{(1),\nph} \right] - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip ,
+ \etarho^{\nph} = {\rm {\bf Avg}} \sum_k \left [\left(\uadvtwo \cdotb \er + w_0^{\nph}\right) (\rho X_k)^{(1),\nph} \right] - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip ,
 \end{equation}
@@
 (\rho h)^{(1),\nph} = 
-(\rho h)^{'(1),\nph} + \left(\frac{(\rho h)_0^n + (\rho h)_0^{(2)}}{2}\right) \enskip .
+(\rho h)^{'(1),\nph} + \frac{(\rho h)_0^n + (\rho h)_0^{(2)}}{2} \enskip .
 \end{equation}
@@
 \begin{eqnarray}
-(\rho h)^{(2')} &=& (\rho h)^{(1)} - \dt \; \left[ \nablab \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  
-(\rho h)^{(1),\nph} \right) \right] \nonumber \\
+(\rho h)^{(2')} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
+(\rho h)^{(1),\nph} \right] \right\} \nonumber \\
 && + \frac{\dt}{2} \; \left(\uadvtwo \cdotb \er\right)
@@
 \item {\bf Correct Base}
-$[\rho_0^{(2)}, (\nabla\cdot\etarho\er)^{\nph}] \rightarrow (\rho_0^{n+1})$
+$[\rho_0^{(2)}, (\nabla\cdot\etarho\er)^{\nph}] \rightarrow [\rho_0^{n+1}]$
 
@@
 \begin{eqnarray}
- \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left(\Gamma_1(\rho^{n+1}, p_0^{n+1}, {X_k}^{n+1}) \right) \enskip , \\ 
+ \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left[\Gamma_1(\rho^{n+1}, p_0^{n+1}, {X_k}^{n+1}) \right] \enskip , \\ 
  {\beta   }_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip .
@@
 \ut^{n+1,\star} &=& \ut^n - \dt \;
- \left[\left((\uadvtwo+ w_0^{\nph} \er) \cdotb \nablab\right) \ut \right]^\nph
+ \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \ut \right]^\nph
       - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
@@
 and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
-The construction of $\left[((\uadvtwo+ w_0^{\nph} \er) \cdotb \nablab) \ut \right]^\nph$
+The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \ut ]^\nph$
 is described in paper II, Appendix A, with $\V = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
@@
 + \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left(\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right)
+\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]
 \enskip,
@@
 - \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left(\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right) \enskip,
+\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right] \enskip,
 \end{equation}
@@
 
-  \item {\bf React State}$[ \rho^{\initp},(\rho h)^{\initp}, X_k^{\initp}, T^{\initp}, (\rho^{\initp} \Hext) ]$\\
-      $\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{0,\nu} ].$
+  \item {\bf React State}$[ \rho^{\initp},(\rho h)^{\initp}, X_k^{\initp}, T^{\initp}, (\rho^{\initp} \Hext) ] \rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{0,\nu} ].$
 
@@
     \begin{equation}  L_\beta^\rho \phi =
-      D \left ( \beta_0^{\half} {\bf V} \right) - \beta_0^{\half} \left( (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right) \enskip . 
+      D \left ( \beta_0^{\half} {\bf V} \right) - \beta_0^{\half} \left[ (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right] \enskip . 
     \end{equation}

### 2008-10-23 docs/flowchart/paper.tex
add a notation table

@@
 
-\usepackage{amsmath,color}
+\usepackage{amsmath,color,longtable}
 
@@
 
+%-----------------------------------------------------------------------------
+% Notation
+%-----------------------------------------------------------------------------
+
 \subsection{Notation}
+
+%%%%%%%%%%%%%%%%
+% symbol table
+%%%%%%%%%%%%%%%%
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{4.0in}|l|}
+\caption[definition of symbols.]{definition of symbols.} \label{table:sym} \\
+
+\hline \multicolumn{1}{|c|}{\textbf{symbol}} & 
+       \multicolumn{1}{ c|}{\textbf{meaning}} & 
+       \multicolumn{1}{ c|}{\textbf{units}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+\hline \multicolumn{1}{|c|}{\textbf{symbol}} & 
+       \multicolumn{1}{ c|}{\textbf{meaning}} & 
+       \multicolumn{1}{ c|}{\textbf{units}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{Continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+$c_p$   & specific heat at constant pressure
+          ($c_p \equiv \left . \partial h / \partial T \right |_{p,X_k}$)
+        & erg~g$^{-1}$~K$^{-1}$ \\
+\hline
+$f$     & volume discrepancy factor ($0 \le f \le 1$) & - \\
+\hline
+$g$     & gravitational acceleration                 & cm~s$^{-2}$ \\
+\hline
+$h$     & specific enthalpy                          & erg~g$^{-1}$ \\
+\hline
+$\Hext$ & external heating energy generation rate    & erg~g$^{-1}$~s$^{-1}$ \\
+\hline
+$\Hnuc$ & nuclear energy generation rate             & erg~g$^{-1}$~s$^{-1}$ \\
+\hline
+$h_p$   & $h_p \equiv \left . \partial h / \partial p \right |_{T,X_k}$ & cm$^{3}$~g$^{-1}$ \\
+\hline
+$\kth$  & thermal conductivity                       & erg~cm$^{-1}$~s$^{-1}$~K$^{-1}$ \\
+\hline
+$p_0$   & base state pressure                        & erg~cm$^{-3}$ \\
+\hline
+$p_T$   & $p_T \equiv \left . \partial p / \partial T \right |_{\rho,X_k}$ & erg~cm$^{-3}$~K$^{-1}$ \\
+\hline
+$p_{X_k}$ & $p_{X_k} \equiv \left . \partial p / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~cm$^{-3}$ \\
+\hline
+$p_\rho$ & $p_\rho \equiv \left . \partial p / \partial \rho \right |_{T,X_k}$ & erg~g$^{-1}$ \\
+\hline
+$q_k$   & specific nuclear binding energy            & erg~g  \\
+\hline
+$r$     & radial coordinate (direction of gravity)   & cm \\
+\hline
+$s$     & specific entropy                           & erg~g$^{-1}$~K$^{-1}$ \\
+\hline
+$S$     & source term to the divergence constraint   & s$^{-1}$ \\
+\hline
+t       & time                                       & s \\
+\hline
+$T$     & temperature                                & K \\
+\hline
+$\ubold$     & total velocity ($\ubold = \ut + w_0 \er$) & cm~s$^{-1}$ \\
+\hline
+$\ut$   & local velocity                             & cm~s$^{-1}$ \\
+\hline
+$w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
+\hline
+$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & - \\
+\hline
+$\beta_0$ & density-like quantity that is the coefficient to velocity
+            in our constraint equation  & g~cm$^{-3}$ \\
+\hline
+$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & - \\
+\hline
+$\etarho$ & $\etarho \equiv \overline{ (\rho' \ut \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\
+\hline
+$\xi_k$ & $\xi_k \equiv \left . \partial h / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~g$^{-1}$ \\
+\hline 
+$\pi$   & dynamic pressure & erg~cm$^{-3}$ \\
+\hline
+$\pizero$ & base state dynamic pressure & erg~cm$^{-3}$ \\
+\hline
+$\rho$  & mass density  & g~cm$^{-3}$ \\
+\hline
+$\rho_0$  & base state mass density  & g~cm$^{-3}$ \\
+\hline
+$\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
+\hline
+$\sigma$ & $\sigma \equiv p_T/(\rho c_p p_\rho)$ & erg$^{-1}$~g \\
+\hline
+$\psi$  & $\psi \equiv D_0 p_0/Dt$ & erg~cm$^{-3}$~s$^{-1}$ \\
+\hline
+$\omegadot_k$ & creation rate for species $k$ ($\omegadot_k \equiv DX_k/Dt$) & s$^{-1}$ \\
+\end{longtable}
+\end{center}
+%
+\renewcommand{\arraystretch}{1.0}
 
@@
 
+%-----------------------------------------------------------------------------
+% Time Advancement Algorithm
+%-----------------------------------------------------------------------------
+
 \subsection{Time Advancement Algorithm}
@@
 
+%-----------------------------------------------------------------------------
+% Initialization
+%-----------------------------------------------------------------------------
+
 \subsection{Initialization}\label{Sec:Initialization}

### 2008-10-24 docs/flowchart/paper.tex
more equation updates

@@
 
+\newcommand{\uadv}{\widetilde{\mathbf{U}}^{\mathrm{ADV}}}
 \newcommand{\uadvone}{\widetilde{\mathbf{U}}^{\mathrm{ADV},\star}}
@@
 \caption[definition of symbols.]{definition of symbols.} \label{table:sym} \\
-
+%
 \hline \multicolumn{1}{|c|}{\textbf{symbol}} & 
@@
 \multicolumn{3}{c}%
-{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
+{{\tablename\ \thetable{}---continued}} \\
 \hline \multicolumn{1}{|c|}{\textbf{symbol}} & 
@@
 
-\multicolumn{3}{|r|}{{Continued on next page}} \\ \hline
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
 \endfoot
@@
 \hline
-$f$     & volume discrepancy factor ($0 \le f \le 1$) & - \\
+$f$     & volume discrepancy factor ($0 \le f \le 1$) & -- \\
 \hline
@@
 \hline
-$q_k$   & specific nuclear binding energy            & erg~g  \\
+$q_k$   & specific nuclear binding energy            & erg~g$^{-1}$  \\
 \hline
@@
 \hline
-t       & time                                       & s \\
+$t$     & time                                       & s \\
 \hline
@@
 \hline
+$\uadv$ & advective velocity (edge-centered)         & cm~s$^{-1}$ \\
+\hline
 $w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
 \hline
-$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & - \\
-\hline
-$\beta_0$ & density-like quantity that is the coefficient to velocity
-            in our constraint equation  & g~cm$^{-3}$ \\
-\hline
-$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & - \\
+$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & -- \\
+\hline
+$\beta_0$ & coefficient to velocity
+            in velocity constraint equation  & g~cm$^{-3}$ \\
+\hline
+$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & -- \\
 \hline
@@
 $\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
+\hline
+$(\rho h)_0$ & base state enthalpy density & erg~cm$^{-3}$  \\
+\hline
+$(\rho h)'$ & perturbational enthalpy density 
+              $ \left [(\rho h)' = \rho h - (\rho h)_0 \right ]$ & erg~cm$^{-3}$  \\
 \hline
@@
 
-We make use of the following shorthand notations in outlining the algorithm.
+We make use of the following shorthand notations in outlining the algorithm:
 \begin{enumerate}
@@
 
-\item Update the enthalpy, using a difference approximation to
-
+\item Update the enthalpy.  The full enthalpy equation is 
   \begin{equation}
@@
   \end{equation}
-  neglecting the reaction terms.
+Here we consider only the advection, neglecting the reaction terms and thermal diffusion terms,
+and solve a discretized version of
+  \begin{equation}
+  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\ubold \rho h)
++ \frac{Dp_0}{Dt} 
+\enskip , \label{eq:rhohupdadv} 
+  \end{equation}
 
@@
 
-\item Diffuse the enthalpy through a time interval of $\dt$.
-Compute $\kth^{(1)}, c_p^{(1)}$, and $\xi_k^{(1)}$ from $\rho^{(1)}, T^{(1)}$, and $X_k^{(1)}$ as inputs to the equation of state.
-The update is given by
+\item Diffuse the enthalpy through a time interval of $\dt$.  Here we begin
+with the enthalpy equation (eq.~[eq:rhohup]), but consider only the 
+diffusion terms,
+\begin{equation}
+  \frac{\partial (\rho h)}{\partial t}  = 
+ \nabla\cdot\kth\nabla T 
+\enskip . \label{eq:rhohupdthermal} 
+  \end{equation}
+We can recast this as an enthalpy-diffusion equation by applying the
+chain-rule to $h(p_0,T,X_k)$,
+\begin{equation}
+\nabla h = h_p \nabla p_0 + c_p \nabla T + \sum_k \xi_k \nabla X_k \enskip ,
+\end{equation}
+giving
+\begin{equation}
+  \frac{\partial (\rho h)}{\partial t}  = 
+ \nabla\cdot \frac{\kth}{c_p}\nabla h -  
+ \sum_k \nabla\cdot \frac{\xi_k \kth}{c_p}\nabla X_k -
+ \nabla\cdot \frac{h_p \kth}{c_p}\nabla p_0 
+\enskip . \label{eq:rhohupdthermal2} 
+  \end{equation}
+
+Compute $\kth^{(1)}, c_p^{(1)}$, and $\xi_k^{(1)}$ from $\rho^{(1)}, T^{(1)}$, and $X_k^{(1)}$ as inputs to the equation of state.  The update is given by
 \begin{eqnarray}

### 2008-10-24 docs/eta_notes/eta.tex
separate out the content so we can produce a master document

--- 2008-10-01 docs/eta_notes/paper.tex
@@
-\documentclass[11pt]{article}
+\section{Computing $\eta$}
 
-\tolerance=600
+We carry around three different 1D radial quantities: $\etarhoec$
+(edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$
+(cell-centered).  These notes discuss when each of these is used, and
+how they are computed, in both plane-parallel and spherical.
 
-\usepackage{amsmath,color}
-
-% Margins
-\usepackage[lmargin=0.5in,rmargin=1.5in,tmargin=0.5in,bmargin=0.5in]{geometry}
-
-\def\half   {\frac{1}{2}}
-
-\def\eb         {{\bf e}}
-\def\ib         {{\bf i}}
-\def\Ub         {{\bf U}}
-
-\newcommand{\Ubt}{\widetilde{\Ub}}
-
-\def\etarho     {\eta_\rho}
-\def\etarhoec   {\etarho^{\rm ec}}
-\def\etarhocc   {\etarho^{\rm cc}}
-\def\etarhoflux {\etarho^{\rm flux}}
-\def\divetarho  {\nabla\cdot(\etarho\eb_r)}
-
-\title{Notes on $\etarho$}
-
-\setlength{\marginparwidth}{1.0in}
-\newcommand{\MarginPar}[1]{\marginpar{%
-\vskip-\baselineskip %raise the marginpar a bit
-\raggedright\tiny\sffamily
-\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
-
-\begin{document}
-
-\maketitle
-\tableofcontents
-\clearpage
-
-We carry around three different 1D radial quantities: $\etarhoec$ (edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$ (cell-centered).
-These notes discuss when each of these is used, and how they are computed, in both plane-parallel and spherical.
-
-\section{Flow Chart}
+%-----------------------------------------------------------------------------
+% flow chart
+%-----------------------------------------------------------------------------
+\subsection{$\eta$ Flow Chart}
 \begin{enumerate}
@@
 
-\section{Computing $\etarhoec, \etarhocc$, and $\divetarho$}
+%-----------------------------------------------------------------------------
+% computing
+%-----------------------------------------------------------------------------
+\subsection{Computing $\etarhoec, \etarhocc$, and $\divetarho$}
 This is done in {\tt make\_eta.f90}.
 
-\subsection{Plane-Parallel}
+\subsubsection{Plane-Parallel}
 We first compute a radial edge-centered multifab, $\eta_{\rho}^{\rm flux}$, using
@@
 
-\subsection{Spherical}
-First, construct $\eta_{\rho}^{\rm cart} = [\rho'(\Ubt\cdot\eb_r)]^{n+\half}$ using:
+\subsubsection{Spherical}
+First, construct $\eta_{\rho}^{\rm cart} =
+[\rho'(\Ubt\cdot\eb_r)]^{n+\half}$ using:
 \begin{equation}
@@
 \end{eqnarray}
-$\divetarho$ is constructed from the cell-centered average of $\rho'^{n+1} = \rho^{n+1} - \rho_0^{n+1}$ as
+$\divetarho$ is constructed from the cell-centered average of
+$\rho'^{n+1} = \rho^{n+1} - \rho_0^{n+1}$ as
 \begin{equation}
@@
 \end{equation}
-\section{Using $\etarhoec$}
-\subsection{Plane-Parallel}
+
+%-----------------------------------------------------------------------------
+% etarho_ec
+%-----------------------------------------------------------------------------
+\subsection{Using $\etarhoec$}
+\subsubsection{Plane-Parallel}
 NOT USED.
-\subsection{Spherical}
-In {\tt make\_w0}, $\etarhoec$ is used in the construction of the RHS for the $\delta w_0$ equation.
-\section{Using $\etarhocc$}
-\subsection{Plane-Parallel}
+
+\subsubsection{Spherical}
+In {\tt make\_w0}, $\etarhoec$ is used in the construction of the RHS
+for the $\delta w_0$ equation.
+
+%-----------------------------------------------------------------------------
+% etarho_cc
+%-----------------------------------------------------------------------------
+\subsection{Using $\etarhocc$}
+\subsubsection{Plane-Parallel}
 In {\tt make\_psi}, $\psi = \etarhocc g$.
-\subsection{Spherical}
-In {\tt make\_w0}, $\etarhocc$ is used in the construction of the RHS for the $\delta w_0$ equation.
-\section{Using $\divetarho$}
-\subsection{Plane-Parallel}
+
+\subsubsection{Spherical}
+In {\tt make\_w0}, $\etarhocc$ is used in the construction of the RHS
+for the $\delta w_0$ equation.
+
+%-----------------------------------------------------------------------------
+% div etarho
+%-----------------------------------------------------------------------------
+\subsection{Using $\divetarho$}
+\subsubsection{Plane-Parallel}
 In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$.
-\subsection{Spherical}
+
+\subsubsection{Spherical}
 In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$. \\
-\end{document}

### 2008-10-24 docs/eta_notes/paper.tex
separate out the content so we can produce a master document

--- 2008-08-20 docs/eta_notes/paper.tex
@@
 
-We carry around three different 1D radial quantities: $\etarho$ (edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$ (edge-centered).
-These notes discuss when each of these is used, and how they are computed, in both plane-parallel and spherical.
+\input eta
 
-\section{Flow Chart}
-\begin{enumerate}
-\item Enter {\tt advance} with $[\etarho, \etarhocc, \divetarho]^{n-\half}$.
-\item Call {\tt make\_w0}.  The spherical version uses uses $\etarho$ and $\divetarho$.
-\item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
-\item Call {\tt correct\_base}, which uses $\divetarho$.
-\item Compute $[\etarho, \etarhocc, \divetarho]^{n+\half,*}$.
-\item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
-\item Call {\tt make\_w0}.  The spherical version uses uses $\etarho$ and $\divetarho$.
-\item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
-\item Call {\tt correct\_base}, which uses $\divetarho$.
-\item Compute $[\etarho, \etarhocc, \divetarho]^{n+\half}$.
-\item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
-\end{enumerate}
-
-\section{Computing $\etarho, \etarhocc$, and $\divetarho$}
-This is done in {\tt make\_eta.f90}.
-
-\subsection{Plane-Parallel}
-We first compute a radial edge-centered multifab, $\eta_{\rho}^{\rm flux}$, using
-\begin{equation}
-\eta_{\rho,\ib+\half\eb_r}^{\rm flux} = \left[\left(\Ubt_{\ib+\half\eb_r}^{n+\half}\cdot\eb_r\right) + w_{0,r+\half}^{n+\half}\right] \rho_{\ib+\half\eb_r}^{n+\half} - w_{0,r+\half}^{n+\half}\rho_{0,r+\half}^{n+\half, {\rm pred}}
-\end{equation}
-$\etarho$ is the edge-centered ``average'' value of $\eta_{\rho}^{\rm flux}$,
-\begin{equation}
-\eta_{\rho,r+\half} = \overline{\eta_{\rho,\ib+\half\eb_r}^{\rm flux}}
-\end{equation}
-$\etarhocc$ is a cell-centered average of $\etarho$,
-\begin{equation}
-\eta_{\rho,r}^{\rm cc} = \frac{\eta_{\rho,r+\half} + \eta_{\rho,r-\half}}{2}.
-\end{equation}
-$\divetarho$ is the cell-centered divergence of $\etarho$,
-\begin{equation}
-\left[\divetarho\right]_r = \frac{\eta_{\rho,r+\half} - \eta_{\rho,r-\half}}{\Delta r}.
-\end{equation}
-
-\subsection{Spherical}
-First, construct $\eta_{\rho}^{\rm cart} = [\rho'(\Ubt\cdot\eb_r)]^{n+\half}$.  Then, $\etarhocc$ is the cell-centered average of $\eta_{\rho}^{\rm cart}$,
-\begin{equation}
-\etarhocc = \overline{\eta_{\rho}^{\rm cart}}.
-\end{equation}
-On interior faces, $\etarho$ is the average of $\etarhocc$,
-\begin{equation}
-\eta_{\rho,r+\half} = \frac{\eta_{\rho,r}^{\rm cc} + \eta_{\rho,r+1}^{\rm cc}}{2}.
-\end{equation}
-At the upper and lower boundaries, we use
-\begin{eqnarray}
-\eta_{\rho,0} &=& 0, \\
-\eta_{\rho,{\rm nr}} &=& \eta_{\rho,{\rm nr}-1}^{\rm cc}.
-\end{eqnarray}
-$\divetarho$ is constructed from the cell-centered average of $\rho'^{n+1}$ as
-\begin{equation}
-\divetarho = -\overline{\rho'^{n+1}}/\Delta t
-\end{equation}
-
-\section{Using $\etarho$}
-
-\subsection{Plane-Parallel}
-NOT USED.
-
-\subsection{Spherical}
-In {\tt make\_w0}, $\etarho$ is used in the construction of the RHS for the $\delta w_0$ equation.
-
-\section{Using $\etarhocc$}
-
-\subsection{Plane-Parallel}
-In {\tt make\_psi}, $\psi = \etarho g$.
-
-\subsection{Spherical}
-NOT USED.
-
-\section{Using $\divetarho$}
-
-\subsection{Plane-Parallel}
-In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$.
-
-\subsection{Spherical}
-In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$. \\
-In {\tt make\_w0}, we average $\divetarho$ onto edges for the $\delta w_0$ equation.
 \end{document}

### 2008-10-24 docs/flowchart/paper.tex
separate out the content so we can produce a master document

--- 2008-10-23 docs/flowchart/paper.tex
@@
 
-\section{Current Algorithm}
-
-Here we outline the algorithm currently implemented in the code.  Our
-starting point for this description are the series of papers describing
-the development of the algorithm:
-\begin{itemize}
-\item {\em Low Mach Number Modeling of Type Ia
-  Supernovae. I. Hydrodynamics}, A.~S.~Almgren, J.~B.~Bell, 
-  C.~A.~Rendleman, \& M.~Zingale 2006, ApJ, 637, 922 (henceforth
-  paper~I)
-\item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
-  Evolution} A.~S.~Almgren, J.~B.~Bell, C.~A.~Rendleman, \& M.~Zingale
-  2006, ApJ, 649, 927 (henceforth paper~II)
-\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions}
-A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale
-  2008, ApJ, 684, 449 (henceforth paper~III)
-\item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection}
-A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale,
-  in perparation (henceforth paper~IV)
-\end{itemize}
-
-\subsection{Notation}
-
-We make use of the following shorthand notations in outlining the algorithm.
-\begin{enumerate}
-
-\item For any quantity, $\phi,$ we define $\overline{\phi} = ${\bf Avg}$(\phi),$ 
-the average over $\Omega_H,$ as
-\begin{equation}
-\overline{\phi}(r) 
-= \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
-\end{equation}
-here $\Omega_H$ is a layer at constant radius $r$.
-
-For spherical problems, see Paper IV for details.
-
-\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
-is the process by which we evolve the species and enthalpy from 
-$X_k^{\inp} \rightarrow X_k^{\outp}$ 
-and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
-system of equations over a time interval of  $\Delta t/2$,
-\begin{eqnarray}
-\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
-\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
-\end{eqnarray}
-  In particular, to evolve the species, we solve the system:
-\begin{eqnarray}
-\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
-\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
-\enskip. \label{eq:reacttemp}
-\end{eqnarray}
-using the stiff ordinary differential equation integration methods
-provided by the {\tt VODE} package (cite VODE).  The absolute error
-tolerances are set to $10^{-12}$ for the species, and a relative
-tolerance of $10^{-5}$ is used for the temperature.  The integration
-yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
-(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
-\begin{equation}
-\label{eq:tempenthalpy}
-\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
- - \sum_k \rho \xi_k {\omegadot}_k 
- + \rho \Hnuc + \rho \Hext \right] \enskip , 
-\end{equation}
-(paper III; eq.~[8])
-by assuming that the pressure is constant during the burn state.  In
-evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
-theory, this means evaluating the equation of state for each
-right-hand side evaluation that {\tt VODE} requires.  In practice, we
-freeze $c_p$ and $\xi_k$ at the start of the integration time step and
-compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
-to the equation of state.  We note that while temperature is evolved
-when solving these equations, we do not keep the final temperature,
-nor do we use it to compute the final change in enthalpy.  Therefore,
-$T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged
-during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
-
-After the new mass fractions have been computed, the reaction rates are defined as:
-\begin{equation}
-(\rho\omegadot_k)^{\outp} = \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\Delta t /2} \enskip,
-\end{equation}
-and the nuclear energy generation rate is defined as
-\begin{equation}
-(\rho\Hnuc) = -\sum_k(\rho\omegadot_k)^{\outp} \; q_k \enskip.
-\end{equation}
-for the case where only strong-mediated reactions are involved.  As
-$\Hnuc$ is an output of the reaction network, more general networks
-(i.e.\ involving weak interactions) can easily be added.  The enthalpy
-update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is
-updated by
-\begin{equation}
-(\rho h)^{\outp} = 
-(\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc) + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
-\end{equation}
-
-\item {\bf React Base}$[(\rho h)_0^{\inp},(\overline{\rho\Hnuc})^{\inp},(\overline{\rho\Hext})^{\inp}] \rightarrow [(\rho h)_0^{\outp}]$\\
-is the process by which we update the base state enthalpy through $\dt/2$ in time using
-\begin{equation}
-\frac{\partial(\rho h)_0}{\partial t} = (\overline{\rho\Hnuc})^{\inp} + (\overline{\rho\Hext})^{\inp}\enskip ,
-\end{equation}
-discretized as
-\begin{equation}
-(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} + \frac{\dt}{2}(\overline{\rho\Hnuc})_{j}^{\inp} + \frac{\dt}{2}(\overline{\rho\Hext})_{j}^{\inp} \enskip .
-\end{equation}
-
-\item {\bf Advect Base}$[\rho_0^{\inp}, (\rho h)_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp}] \rightarrow [\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp}]$ \\
-is the process by which we update the base state through $\dt$ in
-time \MarginPar{we need a separate section for spherical} given the
-radial velocity $w_0^{\inp}.$ The base state arrays are all
-one-dimensional in the radial coordinate; we think of the base state
-quantities as defined at cell centers, with $w_0$ and $\beta_0$
-defined halfway in-between.
-
-\begin{description}
-
-\item[plane-parallel:] ~
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \alph{enumii}}.}
-
-\item For the density update, we discretize equation the base state
-  density evolution equation (paper III; eq.~[29]) without the $\etarho$ term,
-\begin{eqnarray}
-\frac{\partial \rhozero}{\partial t} &=& - \nablab \cdotb \left( \rhozero w_0 \er \right)
-\enskip ,
-\label{eq:rho0upd_new}
-\end{eqnarray}
-to compute the new base state density,
-\begin{eqnarray}
-\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
-\left [ \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
-         \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ]
-         \nonumber \enskip,
-\end{eqnarray}
-where $j$ refers to the one-dimensional index in the radial direction.
-The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the
-procedure described in paper II, Appendix A, and are saved for use
-later in the overall algorithm.
-
-\item For the pressure update, we discretize the base state pressure
-  evolution equation (paper III; eq.~[31]),
-\begin{equation}
-\frac{\partial p_0}{\partial t} = 
--w_0 \frac{\partial p_0}{\partial r} + \psi \enskip ,
-\label{eq:p0upd_new}
-\end{equation}
-to compute the new base state pressure,
-\begin{equation}
-p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
-+ w_{0,j-\myhalf}^\inp \right)
-\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\inp 
-\enskip,
-\end{equation}
-where the interface states are again found via the procedure described
-in paper II, Appendix A.  Note that the $\psi$ term is not included in the forcing for
-the interface state calculation, but is used in the update.
-
-\item For the enthalpy update, we discretize
-\begin{equation}
-\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\er\right] + \psi \enskip ,
-\end{equation}
-to compute the new base state enthalpy,
-\begin{equation}
-(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
-\left\{ \left[ {(\rho h)_0}^{\outp,\nph} {w_0}^{\inp}\right]_{j+\myhalf} -  
-         \left[ {(\rho h)_0}^{\outp,\nph} {w_0}^{\inp}\right]_{j-\myhalf} \right\}
-+ \dt\psi_j^{\inp} \enskip ,
-\end{equation}
-where the interface states are again found via the procedure described
-in paper II, Appendix A.  Note that the $\psi$ term is not included in 
-the forcing for the interface state calculation, but is used in the update.
-
-\end{enumerate}
-
-\item[spherical:]
-
-\MarginPar{we need to write the spherical description}
-
-\end{description}
-
-\item {\bf Correct Base}
-$[\rho_0^{\inp}, (\nabla\cdot\etarho\er)^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
-is the process by which we adjust the base state density to account for
-large-scale mixing, through $\etarho$.  For both plane-parallel and spherical
-we use
-\begin{equation}
-\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \dt (\nabla\cdot\etarho\er)^{\inp}_j \enskip . 
-\end{equation}
-
-\end{enumerate}
-
-\subsection{Time Advancement Algorithm}
-
-We now describe the full time advancement algorithm, making frequent
-use of the shorthand developed above.  Here, we assume that the
-problem is already properly initialized.  We describe the details of
-the initialization in \S~\ref{Sec:Initialization}.
-
-In paper III we discretized the enthalpy evolution equation in full, non-perturbation form.  
-We have discovered that discretizing the enthalpy evolution in non-perturbational form 
-leads to better numerical properties.  This is more like paper II.\\
-
-%--------------------------------------------------------------------------
-% STEP 1
-%--------------------------------------------------------------------------
-
-\begin{description}
-
-\item[Step 1.] {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item At the beginning of each time step, we need an estimate for the time-centered
-source term in the velocity divergence constraint,
-\begin{equation}
-  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext \enskip ,
-\label{eq:defineS} 
-\end{equation}
-(paper III; eq.~[19]).
-
-If this is the first step of the calculation ($n=0$), we set
-\begin{equation}
-S^{\half,\star} = \frac{S^0 + S^1}{2} \enskip ,
-\end{equation}
-where $S^1$ is found through the iterative process that initializes the calculation.
-Otherwise, following the method used in our small-scale low Mach number algorithm
-(Cite Bell04), we extrapolate to the half-time using the source
-terms at the previous and current time levels
-\begin{equation}
-S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
-\end{equation}
-
-\item Define
-\begin{equation}
-\overline{S}^{\nph,\star} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip .
-\end{equation}
-
-\item Construct $w_0^{\nph,\star}$
-
-\begin{description}
-
-\item[plane-parallel:] We integrate
-\begin{equation}
-\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \etarho g 
-\enskip \label{eq:divw0} ,
-\end{equation}
-using the lagged $\psi^{\nmh}$ and the volume discrepancy correction with $f=0.3$,
-\begin{equation}
-\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S}^{\nph,\star} 
-- \frac{1}{\gammabar^{n} p_0^{n}}
-\left\{ \psi^{\nmh} - 
-f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right]\right\}
-\enskip .
-\end{equation}
-
-\item[spherical:] See Paper IV for working notes until we finalize this.
-
-\end{description}
-
-\item Using 
-\begin{equation}
-- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} 
-= \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} 
-\enskip, \label{eq:pizero}
-\end{equation}
-define the scaled pressure gradient
-\begin{equation}
--\left ( \frac{1}{\rho_0^n} \frac{\partial \pizeroone}{\partial r} \right ) = 
-\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} 
-+  w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star} \enskip ,
-\end{equation} 
-where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are
-\begin{equation}
-w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}} 
-\enskip,\end{equation}
-\begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = 
-\frac{1}{\dt^n+\dt^{n-1} } 
-\left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
-+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] 
-\enskip.\end{equation}
-If $n=0$, we use $\dt^{-1} = \dt^0$.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 2
-%--------------------------------------------------------------------------
-\item[Step 2.] {\em Construct the provisional edge-based advective velocity}, $\uadvone$.\\
-
-The local velocity field is described by
-
-The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
-We note that  $\uadvone$ satisfies the discrete versions of 
-\begin{equation}
-\int_{\Omega_H} \ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
-\end{equation}
-(paper III; eq.~[21])
-and
-\begin{equation}
-\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )\enskip ,
-\label{eq:tildeconstraint}
-\end{equation}
-(paper III; eq.~[39]).  The discretization with the volume discrepancy correction is
-\begin{equation}
-\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
-\beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
-+ \frac{f}{\gammabar^n p_0^n}
-\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right] \enskip .
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 3
-%--------------------------------------------------------------------------
-\item[Step 3.] {\em React the full state and base state through the first time 
-interval of $\dt / 2.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n]
-                   \rightarrow$ $[ \rho^{(1)}, (\rho h)^{(1)}, X_k^{(1)}, T^{(1)},
-                                  (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)} ].$
-
-\item {\bf React Base}$[(\rho h)_0^n,(\overline{\rho \Hnuc})^{(1)},(\overline{\rho\Hext})^n] \rightarrow [(\rho h)_0^{(1)}]$
-
-\item
-Define
-\begin{eqnarray}
-\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left[ \Gamma_1(\rho^{(1)}, p_0^{n}, X_k^{(1)}) \right] \enskip , \\
- {\beta   }_0^{(1)}    &=& \beta   \left(\rho_0^{n}, p_0^{n}, \gammabar^{(1)}\right) \enskip .
-\end{eqnarray}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 4
-%--------------------------------------------------------------------------
-\item[Step 4.] {\em Advect the base state, then the full state, through a time interval 
-of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf Advect Base}$[\rhozero^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph,\star}, \psi^{\nmh}] \rightarrow $
-$[\rho_0^{(2),\star}, (\rho h)_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}, p_0^{n+1,\star}].$
-
-\item Update the species.  Here we do a difference approximation
-  to 
-  \begin{equation}
-  \frac{\partial (\rho X_k)}{\partial t} + \nablab \cdotb (\ubold \rho X_k) =
-         \rho {\omegadot}_k \enskip , \label{eq:species}
-  \end{equation}
-  neglecting the reaction terms.  The update consists of two steps:
-
-  \begin{enumerate}
-  \renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
-  for the conservative update of $(\rho X_k)$. 
-  Here we predict $\rho'$ and $X_k$ to the edges to obtain 
-  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
-  Taylor expansion in space and time, as described in paper II,
-  Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  We use the base
-  state density, $\rho_0$, to convert this to an edge state for $\rho$,
-
-\begin{equation}
-\rho^{(1),\nph,\star} = 
-\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2),\star}}{2} \enskip ,
-\end{equation}
-
-  \item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$
-  without explicitly including the reaction terms,
-\begin{equation}
-(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)} 
- - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right)  
-  (\rho X_k)^{(1),\nph,\star} \right] \right\} \enskip ,
-\end{equation}
-\begin{equation}
-\rho^{(2),\star} = \sum_k (\rho X_k)^{(2),\star} \enskip ,
-\end{equation}
-\begin{equation}
-X_k^{(2),\star} = (\rho X_k)^{(2),\star} / \rho^{(2),\star}
-\end{equation}
-
-\end{enumerate}
-
-\item Define an edge-centered $\etarho^{\nph,\star}$, cell-centered 
-$(\nabla\cdot\etarho\er)^{\nph,\star}$, and cell-centered $\psi_j^{\nph,\star}$:
-
-\begin{description}
-\item[plane-parallel:]~
-
-\begin{equation}
- \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdotb \er + w_0^{\nph,\star}\right) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
-\end{equation}
-\begin{equation}
-(\nabla\cdot\etarho\er)_j^{\nph,\star} = 
-\frac{\eta_{\rho,j+\myhalf}^{\nph,\star}-\eta_{\rho,j-\myhalf}^{\nph,\star}}{\Delta r} \enskip ,
-\end{equation}
-\begin{equation}
- \psi_j^{\nph,\star} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph,\star} + \eta_{\rho,j+\myhalf}^{\nph,\star}\right) g \enskip .
-\end{equation}
-
-\item[spherical:] See {\tt MAESTO/docs/eta\_notes} for working notes until we finlize 
-the details.
-
-\end{description}
-
-\item Update the enthalpy, using a difference approximation to
-
-  \begin{equation}
-  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\ubold \rho h)
-+ \frac{Dp_0}{Dt} + \nabla\cdot\kth\nabla T + \rho \Hnuc + \rho \Hext 
-\enskip , \label{eq:rhohupd} 
-  \end{equation}
-  neglecting the reaction terms.
-
-  \begin{enumerate}
-  \renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph,\star},$
-    for the conservative update of $(\rho h).$  Here we predict $(\rho
-    h)'$ to the edges to obtain $(\rho h)^{'(1),\nph,\star}$, 
-    using a second-order Taylor expansion in space
-    and time, as described in paper II, Appendix A, using $\V =
-    \uadvone+w_0^{\nph,\star} \er$.  We do not include the reaction
-    terms in the enthalpy prediction, since we accounted for them
-    already in {\bf React State}.  We use the base state enthalpy,
-    $(\rho h)_0$, to convert this to an edge state for $(\rho h)$,
-    \MarginPar{in these notes we currently assume pred type 1, $(\rho h)'$}
-\begin{equation}
-(\rho h)^{(1),\nph,\star} = 
-(\rho h)^{'(1),\nph,\star} + \frac{(\rho h)_0^n + (\rho h)_0^{(2),\star}}{2}
-\enskip .
-\end{equation}
-
-  \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2'),\star}$ without
-  explicitly including the reaction terms,
-  \begin{eqnarray}
-  (\rho h)^{(2'),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
-      \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right) (\rho
-      h)^{(1),\nph,\star} \right] \right\} \nonumber \\ && + \dt \;
-    \left(\uadvone \cdotb \er\right) \left(\frac{\partial
-      p_0}{\partial r} \right)^{n} + \dt \; \psi^{\nph,*} \enskip ,
-  \end{eqnarray}
-  \MarginPar{we should also show the update in the case that $(\rho h)'$ is the flux}
-  \end{enumerate}
-
-\item {\bf Correct Base}
-$[\rho_0^{(2),\star}, (\nabla\cdot\etarho\er)^{\nph,*}] \rightarrow [\rho_0^{n+1,\star}]$.
-
-%--------------------------------------------------------------------------
-% STEP 4.1
-%--------------------------------------------------------------------------
-
-\item Diffuse the enthalpy through a time interval of $\dt$.
-Compute $\kth^{(1)}, c_p^{(1)}$, and $\xi_k^{(1)}$ from $\rho^{(1)}, T^{(1)}$, and $X_k^{(1)}$ as inputs to the equation of state.
-The update is given by
-\begin{eqnarray}
-(\rho h)^{(2),\star} &=& (\rho h)^{(2'),\star} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(2),\star} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(2),\star} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
-\end{eqnarray}
-which is numerically implemented as a diffusion equation for $h^{(2),\star}$,
-\begin{eqnarray}
-\left(\rho^{(2),\star} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla\right)h^{(2),\star} &=& (\rho h)^{(2'),\star} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
-&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(2),\star} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
-\end{eqnarray}
-\item Update the temperature
-\begin{equation}
-T^{(2),\star} = T\left(\rho^{(2),\star}, h^{(2),\star}, X_k^{(2),\star}\right) {\hbox{ \rm using the equation of state.}}
-\end{equation}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 5
-%--------------------------------------------------------------------------
-\item[Step 5.] {\em React the full state through a second time interval of $\dt / 2.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, 
-                             T^{(2),\star}, (\rho\Hext)^{(2),\star}] $\\
-$\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, 
-              (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
-
-\item Define
-\begin{eqnarray}
- \gammabar^{n+1,\star}    &=& {\rm{\bf Avg}} \left[ \Gamma_1(\rho^{n+1,\star}, p_0^{n+1,\star}, 
-                                                      X_k^{n+1,\star}) \right] \enskip , \\
- {\beta   }_0^{n+1,\star}    &=& \beta   \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}\right) \enskip .
-\end{eqnarray}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 6
-%--------------------------------------------------------------------------
-\item[Step 6.] {\em Define a new average expansion rate at time $t^\nph.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-\item Define
-\begin{equation}
-  S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + 
-  \sigma \Hnuc^{(2),\star} +
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star}  
-  + \sigma \Hext \enskip .
-\end{equation} 
-where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} / \rho^{(2),\star}$
-and the remaining quantities are defined using $X_k^{n+1,\star},$ $\rho^{n+1,\star},$ 
-and $T^{n+1,\star}$ from step 5.
-Then define
-\begin{equation}
- S^\nph = \frac{S^n + S^{n+1,\star}}{2} \enskip. 
-\end{equation}
-
-\item Define
-\begin{equation}
-\overline{S}^{\nph} = {\mathrm{\bf Avg}} (S^{\nph}) \enskip.
-\end{equation}
-
-\item Construct $w_0^{\nph}$
-
-\begin{description}
-
-\item[plane-parallel:] We integrate equation (\ref{eq:divw0}) with the 
-volume discrepancy correction,
-\begin{equation}
-\frac{\partial w_0^{\nph}}{\partial r} = 
-\overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
-\left\{ \psi^{\nph,\star} - f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\}
-\enskip ,
-\end{equation}
-\begin{equation}
-\gammabar^{\nph,\star} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}
-\end{equation}
-\begin{equation}
-p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}
-\end{equation}
-\begin{equation}
-\overline{p(\rho,h,X_k)^{\nph,\star}} = 
-\frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}
-\end{equation}
-
-\item[spherical:] See Paper IV for working notes until we finalize this.
-
-\end{description}
-
-\item Using equation (\ref{eq:pizero}), define
-\begin{equation}
--\left ( \frac{1}{\rho_0^n} \frac{\partial \pizerotwo}{\partial r} \right ) = 
-\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
-+ w_0^n \left(\frac{\partial w_0}{\partial r}\right)^n.
-\end{equation}
-where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined through
-\begin{equation}
-w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
-\end{equation}
-(paper III; eq.~[53]) and
-\begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
-+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\label{eq:dw0drnstar}
-\end{equation}
-(paper III; eq.~[54]).
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 7
-%--------------------------------------------------------------------------
-\item[Step 7.] {\em Construct the final edge-based advective velocity}, $\uadvtwo$.
-
-The procedure to construct $\uadvtwo$ is described in detail in Appendix B of paper III
-and is analogous to the procedure used in step 2, but with updated values
-for $w_0$ and $\pizero.$  We note that $\uadvtwo$ satisfies the discrete versions of 
-equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}).   The discretization
-with the volume discrepancy correction is
-\begin{eqnarray}
-\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) &=& 
-\beta_0^{\nph,\star} \left(S^{\nph} - \overline{S}^{\nph}\right) \nonumber \\
-&& + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
-\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]
-\enskip , \nonumber \\
-\end{eqnarray}
-\begin{equation}
-\beta_0^{\nph,\star} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2} \enskip .
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 8
-%--------------------------------------------------------------------------
-\item[Step 8.] {\em Advect the base state, then the full state, through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf Advect Base}$(\rho_0^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph}, \psi^{\nph,\star} ) \rightarrow $
-$[\rho_0^{(2)}, (\rho h)_0^{(2)}, \rho_0^{\nph,{\rm pred}}, p_0^{n+1}].$
-
-\item Update the species.  Here we do a difference approximation to
-  eq.~\ref{eq:species}, neglecting the reaction terms.  The
-  update consists of two steps:
-
-  \begin{enumerate}
-  \renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-  \item Compute the species edge states, $(\rho X_k)^{(1),\nph}$, for
-    the conservative update of $(\rho X_k)$.  Here we predict $\rho'$ 
-    and $X_k$ to the edges to obtain $\rho^{'(1),\nph}$ and 
-    $X_k^{(1),\nph}$ using a second-order Taylor expansion in
-    space and time, as described in paper II, Appendix A, using $\V =
-    \uadvtwo+w_0^{\nph} \er$.  We use the base
-    state density, $\rho_0$, to convert this to an edge state for $\rho$,
-
-\begin{equation}
-\rho^{(1),\nph} = \rho^{'(1),\nph} + \frac{\rho_0^n + \rho_0^{(2)}}{2}\enskip ,
-\end{equation}
-
-    \item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$
-      without explicitly including the reaction terms,
-\begin{equation}
-(\rho X_k)^{(2)} = (\rho X_k)^{(1)} 
-- \dt \; \left\{ \nablab \cdotb \left[\left(\uadvtwo+w_0^{\nph} \er\right)  
-(\rho X_k)^{(1),\nph} \right] \right\}  \enskip ,
-\end{equation}
-\begin{equation}
-\rho^{(2)} = \sum_k (\rho X_k)^{(2)} \enskip ,
-\end{equation}
-\begin{equation}
-X_k^{(2)} = (\rho X_k)^{(2)} / \rho^{(2)}
-\end{equation}
-
-   \end{enumerate}
-
-\item Define an edge-centered $\etarho^{\nph}$, cell-centered 
-$(\nabla\cdot\etarho\er)^{\nph}$, and a cell-centered $\psi_j^{\nph}$,
-
-\begin{description}
-
-\item[plane-parallel:] ~
-
-\begin{equation}
- \etarho^{\nph} = {\rm {\bf Avg}} \sum_k \left [\left(\uadvtwo \cdotb \er + w_0^{\nph}\right) (\rho X_k)^{(1),\nph} \right] - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip ,
-\end{equation}
-\begin{equation}
-(\nabla\cdot\etarho\er)_j^{\nph} = \frac{\eta_{\rho,j+\myhalf}^{\nph} - \eta_{\rho,j-\myhalf}^{\nph}}{\Delta r} \enskip , \\
-\end{equation}
-\begin{equation}
- \psi_j^{\nph} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
-\end{equation}
-
-\item[spherical:]~
-
-\end{description}
-
-\item Update the enthalpy, using a difference approximation to
-  eq.~\ref{eq:rhohupd}, neglecting the reaction terms.
-
-  \begin{enumerate}
-  \renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, for
-    the conservative update of $(\rho h)$. Here we predict $(\rho h)'$
-    to the edges to obtain $(\rho h)^{'(1),\nph}$, 
-    using a second-order Taylor expansion in space and
-    time, as described in paper II, Appendix A, using $\V =
-    \uadvtwo+w_0^{\nph} \er$.  Again, we do not include the reaction
-    terms in the enthalpy prediction since we accounted for them
-    already in {\bf React State}. We use the base state enthalpy,
-    $(\rho h)_0$, to convert these to edge states for $(\rho h)$,
-\begin{equation}
-(\rho h)^{(1),\nph} = 
-(\rho h)^{'(1),\nph} + \frac{(\rho h)_0^n + (\rho h)_0^{(2)}}{2} \enskip .
-\end{equation}
-
-   \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$ without
-     explicitly including the reaction terms,
-\begin{eqnarray}
-(\rho h)^{(2')} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
-(\rho h)^{(1),\nph} \right] \right\} \nonumber \\
-&& + \frac{\dt}{2} \; \left(\uadvtwo \cdotb \er\right)
-\left [ \left(\frac{\partial p_0}{\partial r} \right)^{n}
-      + \left(\frac{\partial p_0}{\partial r} \right)^{n+1}  \right ] 
-+ \dt \; \psi^{\nph} \enskip ,
-\end{eqnarray}
-
-\end{enumerate}
-
-\item {\bf Correct Base}
-$[\rho_0^{(2)}, (\nabla\cdot\etarho\er)^{\nph}] \rightarrow [\rho_0^{n+1}]$
-
-%--------------------------------------------------------------------------
-% STEP 8.1
-%--------------------------------------------------------------------------
-\item Diffuse the enthalpy through a time interval of $\dt$.
-Compute $\kth^{(2),\star}, c_p^{(2),\star}$, and $\xi_k^{(2),\star}$, from $\rho^{(2),\star}, T^{(2),\star}$, and $X_k^{(2),\star}$ as inputs to the equation of state.
-The update is given by
-\begin{eqnarray}
-(\rho h)^{(2)} &=& (\rho h)^{(2')} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla h^{(2)} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
-\end{eqnarray}
-which is numerically implemented as a diffusion equation for $h^{(2)}$,
-\begin{eqnarray}
-\left(\rho^{(2)} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla\right)h^{(2)} &=& (\rho h)^{(2')} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
-&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
-\end{eqnarray}
-\item Update the temperature
-\begin{equation}
-T^{(2)} = T\left(\rho^{(2)}, h^{(2)}, X_k^{(2)}\right) {\hbox{ \rm using the equation of state.}}
-\end{equation}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 9
-%--------------------------------------------------------------------------
-\item[Step 9.] {\em React the full state and base state through a second time 
-interval of $\dt / 2.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho^{(2)} \Hext) ]$\\
-                   $\rightarrow [\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, 
-                               (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$
-
-\item {\bf React Base}$[(\rho h)_0^{(2)},(\overline{\rho \Hnuc})^{(2)},(\overline{\rho\Hext})^{(2)}] \rightarrow [(\rho h)_0^{n+1}]$
-
-\item Define
-\begin{eqnarray}
- \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left[\Gamma_1(\rho^{n+1}, p_0^{n+1}, {X_k}^{n+1}) \right] \enskip , \\ 
- {\beta   }_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip .
-\end{eqnarray}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 10
-%--------------------------------------------------------------------------
-\item[Step 10.] {\em Compute $S^{n+1}$ for the final projection.}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-\item Define
-\begin{equation}
-  S^{n+1} =  -\sigma  \sum_k  \xi_k (\omegadot_k)^{(2)}  + \sigma \Hnuc^{(2)} +
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2)}  
-   + \sigma \Hext \enskip ,
-\end{equation}
-where $(\omegadot_k)^{(2)} = (\rho \omegadot_k)^{(2)} / \rho^{(2)}$
-and the remaining quantities are defined using $X_k^{n+1},$ $\rho^{n+1},$
-and $T^{n+1}$ from step 9.
-
-\item Define
-\begin{equation}
-\overline{S}^{n+1} = {\mathrm{\bf Avg}} (S^{n+1}) \enskip.
-\end{equation}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 11
-%--------------------------------------------------------------------------
-\item[Step 11.] {\em Update the velocity}.
-
-The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
-We update the velocity field $\ut^n$ to $\ut^{n+1,\star}$ by discretizing
-\begin{equation}
-\frac{\partial\ut}{\partial t} = - \ut \cdotb \nablab \ut - w_0 \frac{\partial \ut}{\partial r}
-                                 - \left(\ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
-                                 - \frac{1}{\rho} \nablab\pi
-                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
-                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip\
- ,
-\end{equation}
-(paper III; eq.~[37]) as 
-\begin{eqnarray}
-\ut^{n+1,\star} &=& \ut^n - \dt \;
- \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \ut \right]^\nph
-      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
-   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
-        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
-        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
-\end{eqnarray}
-where $\rho^\nph = ( \rho^n + \rho^{n+1} )/2$
-and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
-The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \ut ]^\nph$
-is described in paper II, Appendix A, with $\V = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
-component of $\ut^n$ individually.
-
-Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}) 
-including the volume discrepancy correction,
-\begin{equation}
-\nablab \cdotb \left(\beta_0^{\nph} \ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)
-+ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]
-\enskip,
-\end{equation}
-by defining ${\bf V} = \ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
-and solving
-\begin{equation}
- L_\beta^\rho \phi =
-   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)
-- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right] \enskip,
-\end{equation}
-for nodal values of $\phi$, where
-$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
-$L_\beta^\rho$ is the standard bilinear
-finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
-\MarginPar{Why are $\beta$ and $\rho$ time centered here instead using the new time value?  Does this have something to do with how $\phi$ relates to a lagged pressure?}
-In this step, $D$ is a discrete second-order operator that approximates the 
-divergence at nodes from cell-centered data and satisfies
-$\mathbf{G} = -D^T.$ 
-(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
-approximate projection; see (cite almgren:bell:crutchfield) for a discussion
-of this particular form of the projection operand.)  
-We solve the linear system of equation
-\begin{equation}
- L_\beta^\rho \phi =
-   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
-\label{eq:nodal_solve}
-\end{equation}
-(paper III; eq.~[77]) using multigrid V-cycles with Gauss-Seidel relaxation.
-
-We determine the new time velocity field from
-\begin{equation}
-\ut^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
-\end{equation}
-and the new time-centered perturbational pressure from
-\begin{equation}
-  \pi^\nph = \frac{1}{\dt} \phi \enskip .
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 12
-%--------------------------------------------------------------------------
-\item[Step 12.] {\em Compute a new $\dt.$}
-
-Compute $\dt$ for the next time step with the procedure described in \S~ref{sec:timesteps},
-using $w_0$ as computed in step 6 and
-$\ut^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.
-
-\end{description}
-
-\noindent This completes the time advancement of the algorithm.
-
-\subsection{Initialization}\label{Sec:Initialization}
-
-We start each calculation with user-specified initial values for
-$\rho$, $X_k$ and $T,$ as well as an initial background state.  In
-order for the low Mach number assumption to hold, the initial data
-must be thermodynamically consistent with the initial background
-state.  In addition, the initial velocity field must satisfy an
-initial approximation to the divergence constraint.  We use an iterative
-procedure to compute both an initial right-hand-side for the
-constraint equation and an initial velocity field that satisfies
-the constraint.  The user specifies the number of iterations,
-$N_{\rm iters}^{S},$ in this first step of the initialization procedure.
-
-The initial perturbational pressure also needs to be determined for
-use in steps 2, 7 and 11. 
-This is done through a second iterative procedure which follows the
-time advancement algorithm as described in steps 1-11 in \S~ref{Sec:Alg}.  
-The user specifies the number of iterations, 
-$N_{\rm iters}^{\pi},$ in this second step of the initialization procedure.
-The details for both iterations are given below.\\
-
-%--------------------------------------------------------------------------
-% STEP 0
-%--------------------------------------------------------------------------
-\noindent {\bf Step 0.} {\em Initialization}
-
-Start with initial data $X_k^{\initp}, \rho^{\initp},$ $T^{\initp},$ an 
-initial base state, and an initial guess for the velocity, $\ubold^{\initp}.$
-Use the equation of state to determine $(\rho h)^{\initp}$.  Set
-$w_0^1 = 0$ as an initial approximation.  Compute $\beta_0^{\initp}$ as a function of 
-the initial data.  Then, project $\ubold^{\initp}$ using $\beta_0^{\initp}$ and 
-$S = \rho\Hext$, giving $\ubold^{0,1}$.  The next part of the initialization process 
-proceeds as follows.
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-\renewcommand{\labelenumii}{\roman{enumii}.}
-
-\item {\bf Do} {$\nu = 1,...,N_{\rm iters}^{S}$.}
-  \begin{enumerate}
-
-  \item Estimate $\Delta t^\nu$ using $\ubold^{0,\nu}$ and $w_0^\nu.$
-
-  \item {\bf React State}$[ \rho^{\initp},(\rho h)^{\initp}, X_k^{\initp}, T^{\initp}, (\rho^{\initp} \Hext) ] \rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{0,\nu} ].$
-
-  \item Compute $S^{0,\nu}$ from equation (\ref{eq:defineS}) 
-        using $(\rho \omegadot_k)^{0,\nu}$ and the initial data.
-
-  \item Compute $\overline{S}^{0,\nu} = {\mathrm{\bf Avg}} (S^{0,\nu}).$
-
-  \item Compute $w_0^{\nu+1}$ as in step 1c using $\overline{S}^{0,\nu}$ and $\psi^{\nmh} = 0$.
-
-  \item Project $\ubold^{0,\nu}$ using $\beta_0^{\initp}$ and 
-        $(S^{0,\nu} - \overline{S}^{0,\nu})$ as the source term.  
-        This yields $\ubold^{0,\nu+1}.$
-
-  \end{enumerate}
-
-  {\bf End do.}
-
-  Define $S^0 = S^{0,N_{\rm iters}^S}$, $w_0^{-\myhalf} = w_0^{N_{\rm iters}^S+1}$, 
-$\dt^0 = \Delta t^{N_{\rm iters}^S},$ and $\ubold^0 = \ubold^{0,N_{\rm iters}^S+1}.$
-
-\end{enumerate}
-
-Next, we need to construct an approximation to the time-centered perturbational pressure,
-$\pi,$  and an approximation to the divergence constraint at the end of the
-first time step.  As initial approximations, set $S^{1,0} = S^0$,
-$\etarho^{-\half} = 0$, $\psi^{-\half} = 0$, and $\pi^{-\half} = 0.$
-\begin{enumerate}
-  \renewcommand{\theenumi}{{\bf \alph{enumi}}}
-  \renewcommand{\labelenumii}{\roman{enumii}.}
-  \addtocounter{enumi}{1}
-
-\item {\bf Do} {$\nu = 0,...,N_{\rm iters}^{\pi}-1$.}
-
-  \begin{enumerate}
-  \item Perform steps 1--11 as described above, using 
-    $S^{\half,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
-    The only other difference in the time advancement is that in step 11
-    we define ${\bf V} = (\ut^{1,\star} - \ut^0)$ and solve
-    \begin{equation}  L_\beta^\rho \phi =
-      D \left ( \beta_0^{\half} {\bf V} \right) - \beta_0^{\half} \left[ (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right] \enskip . 
-    \end{equation}
-    (The motivation for this form of the projection in the initial pressure iterations
-    is discussed in (cite almgren:bell:crutchfield).)
-      We discard the new velocity resulting from this, but keep the new  
-      value for $\pi^{\half} = \pi^{-\half} + (1 / \dt) \; \phi.$  
-      These steps also yield new scalar data at time $\dt,$ which
-      we discard,  and new values for $\etarho^{\half}$ (step 8), $\psi^{\half}$ (step 8), 
-      $S^{1,\nu+1}$ (step 10), and $\pi^{\half}$ (step 11), which we keep.
-    \item Set $\pi^{-\half} = \pi^{\half}$, $\etarho^{-\half} = \etarho^{\half}$,
-      and $\psi^{-\half} = \psi^{\half}$. 
-    \end{enumerate}
-
-    {\bf End do.}
-
-    Finally, we define $S^1 = S^{1,N_{\rm iters}^\pi}.$
-
-  \end{enumerate}
+\input flowchart
 

### 2008-10-24 docs/Godunov_notes/paper.tex
fixed the temperature forcing term
separate out the content so we can include this in a master document

--- 2008-04-21 docs/Godunov_notes/paper.tex
@@
 
-These are working notes for the Godunov step in {\tt MAESTRO} and {\tt VARDEN}.
-\section{Differences Between These Notes and the Code}
-\begin{enumerate}
-\item Concerning equations (\ref{Radial utrans_L}), (\ref{Radial utrans_R}), (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}), in the code we are omitting the term in equation (\ref{Perturbational Velocity Equation}) proportional to $(\Ubt\cdot\eb_r)$.
-\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction.
-However, this term is present in the cell-update.
-\end{enumerate}
-\section{Questions}
-\begin{enumerate}
-\item Concerning equations (\ref{Radial utrans_L}), (\ref{Radial utrans_R}), (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}), do we really need to include the term proportional to $(\Ubt\cdot\eb_r)$?
-\item Concerning (\ref{Force Term}), do we really need to include the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction?
-\item Do we want to do characteristic tracing to compute $\Ub^{\trans}$, analgous to how {\tt VARDEN} uses characteristic tracing to compute $\Ub^{\mac}$?
-\item When we compute $\Ub^{\mac,*}$, do we want to replace all instances of ``trans'' velocities with ``MAC'' velocities?
-Doing this is simply a matter of replacing the call to {\tt make\_edge\_state} with {\tt make\_edge\_scal}.
-\end{enumerate}
-\section{Differences Between {\tt MAESTRO} and {\tt VARDEN}}
-\begin{enumerate}
-\item {\tt VARDEN} does not use the concept of $w_0$, so $\Ubt$ does not exist.
-\item {\tt VARDEN} only does Cartesian problems.
-\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$ in the computation of $\Ub^{\mac}$, so therefore the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
-\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$, whereas {\tt MAESTRO} does not.
-\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, and a ``non-FPU'' algorithm for $\Ub^{\edge}$
-\item {\tt VARDEN} treats $\rho$ in a conservative manner, whereas {\tt MAESTRO} treats $\rho'$ in a non-conservative manner.
-Therefore, {\tt MAESTRO} accounts for this difference by modifying the forcing term in {\tt modify\_scal\_force}
-\item {\tt MAESTRO} does not corner couple for 3D problems.
-\end{enumerate}
-\cleardoublepage
-\section{Computing $\Ubt^{\trans}$ in {\tt MAESTRO}}
-In {\tt advance\_premac}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.
-These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
-We will only compute the normal component of velocity at each face.\\
+\input Godunov
 
-The evolution equation for the perturbational velocity is:
-\begin{equation}
-\frac{\partial\Ubt}{\partial t} = -\Ubt\cdot\nabla\Ubt - w_0\frac{\partial\Ubt}{\partial r} - (\Ubt\cdot\eb_r)\frac{\partial w_0}{\partial r}\eb_r - \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Perturbational Velocity Equation}
-\end{equation}
-We are going to use a 1D Taylor series extrapolation in space and time to compute time-centered normal components of $\Ubt^{\trans}$ at face-centers.
-By 1D, we mean that we omit any spatial derivatives that are not in the direction of the extrapolation.
-We first compute ``left/right'' states, and then use a Riemann solver to pick the final state.
-\subsection{2D Cartesian Case}
-For the rest of this paper, for 2D Cartesian problems, $\Ub = (u,w)$ and $\Ubt = (u,\wt)$.  We will use the shorthand $\ib = (x,r)$.\\
-
-For the x-faces, the 1D extrapolation of $u$ is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\trans} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
-u_{R,\ib-\half\eb_x}^{\trans} &=& u_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
-\end{eqnarray}
-We pick the final edge states using a Riemann solver:
-\begin{equation}
-u^{\trans}_{\ib-\half\eb_x} =
-\begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\trans} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{\trans}, & u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans} > 0, \\
-u_{R,\ib-\half\eb_x}^{\trans}, & u_{L,\ib-\half\eb_x}^{\trans} + u_{R,\ib-\half\eb_x}^{\trans} < 0, \\
-\end{cases}
-\end{equation}
-For the r-faces, we choose include the radial derivative terms proportional to $w_0$ and $(\Ubt\cdot\eb_r) = \wt$.
-The term proportional to $w_0$ is included by using $w$ rather than $\wt$ as the advection speed.
-The term proportional to $\wt$ is an additional term.
-For the r-faces, the 1D extrapolation of $\wt$ is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{\trans} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
-\wt_{R,\ib-\half\eb_r}^{\trans} &=& \wt_{\ib}^n - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.\label{Radial utrans_R}
-\end{eqnarray}
-The total radial velocity is defined as:
-\begin{equation}
-w_{\ib}^n = \wt_{\ib}^n + \frac{w_{0,r+\half}+w_{0,r-\half}}{2}.\label{Total Radial Velocity}
-\end{equation}
-The spatial derivatives of $w_0$ are given by the two-point centered difference:
-\begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)_{\ib} = \frac{w_{0,r+\half}-w_{0,r-\half}}{h}.\label{Radial Derivative of w_0}
-\end{equation}
-We pick the final edge states using:
-\begin{equation}
-\wt^{\trans}_{\ib-\half\eb_r} =
-\begin{cases}
-0, & \left(w_L^{\trans} \le 0 ~~ {\rm AND} ~~ w_R^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|w_L^{\trans} + w_R^{\trans}\right| < \epsilon, \\
-\wt_{L,\ib-\half\eb_r}^{\trans}, & w_L^{\trans} + w_R^{\trans} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{\trans}, & w_L^{\trans} + w_R^{\trans} < 0, \\
-\end{cases}
-\end{equation}
-where $w_L^{\trans} = \wt_{L,\ib-\half\eb_r}^{\trans} + w_{0,r-\half}$ and $w_R^{\trans} = \wt_{R,\ib-\half\eb_r}^{\trans} + w_{0,r-\half}$.
-Finally, we call {\tt addw0} to compute $w^{\trans}$ from $\wt^{\trans}$.
-\subsection{3D Cartesian Case}
-For the rest of this paper, for 3D Cartesian problems, $\Ub = (u,v,w)$ and $\Ubt = (u,v,\wt)$.  We will use the shorthand $\ib = (x,y,r)$.\\
-
-The only difference between the 2D and 3D Cartesian cases is that we must also compute $v^{\trans}$.
-The procedure is analogous to computing $u^{\trans}$.
-For y-faces, the 1D extrapolation of $v$ is:
-\begin{eqnarray}
-v_{L,\ib-\half\eb_y}^{\trans} &=& v_{\ib-\eb_y}^n + \left(\half - \frac{\dt}{2h}v_{\ib-\eb_y}^n\right)\Delta_y v_{\ib-\eb_y}^n, \\
-v_{R,\ib-\half\eb_y}^{\trans} &=& v_{\ib}^n - \left(\half + \frac{\dt}{2h}v_{\ib}^n\right)\Delta_y v_{\ib}^n,
-\end{eqnarray}
-\begin{equation}
-v^{\trans}_{\ib-\half\eb_y} =
-\begin{cases}
-0, & \left(v_{L,\ib-\half\eb_y}^{\trans} \le 0 ~~ {\rm AND} ~~ v_{R,\ib-\half\eb_y}^{\trans} \ge 0\right) ~~ {\rm OR} ~~ \left|v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans}\right| < \epsilon, \\
-v_{L,\ib-\half\eb_y}^{\trans}, & v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans} > 0, \\
-v_{R,\ib-\half\eb_y}^{\trans}, & v_{L,\ib-\half\eb_y}^{\trans} + v_{R,\ib-\half\eb_y}^{\trans} < 0. \\
-\end{cases}
-\end{equation}
-\subsection{3D Spherical Case}
-For the rest of this paper, for 3D spherical problems, $\Ub = (u,v,w)$ and $\Ubt = (\ut,\vt,\wt)$.  We will use the shorthand $\ib = (x,y,z)$.\\
-
-We first construct $\Ub$ by mapping $w_0$ onto a Cartesian array and adding it to $\Ubt$.
-This is a spherical generalization of equation (\ref{Total Radial Velocity}).
-We also map $\partial w_0/\partial r$ onto a Cartesian array.
-This is a spherical generalization of equation (\ref{Radial Derivative of w_0}).
-Then we use a 1D extrapolation analogous to equations equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}).
-For example, for x-faces:
-\begin{eqnarray}
-\ut_L^{\trans} &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x\ut_{\ib-\eb_x}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib-\eb_x}, \\
-\ut_R^{\trans} &=& \ut_{\ib}^n - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n - \frac{\dt}{2}\left[(\Ubt^n\cdot\eb_r)\left(\frac{\partial w_0}{\partial r}\right)(\eb_r\cdot\eb_x)\right]_{\ib}.
-\end{eqnarray}
-Then, we use the same Riemann solver as the Cartesian case to obtain $\ut^{\trans}$.
-To compute $\vt^{\trans}$ (or $\wt^{\trans}$), simply replace all instaces of $\ut$ with $\vt$ (or $\wt$) and all instaces of $x$ with $y$ (or $z$).
-Finally, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.
-\cleardoublepage
-\section{Computing $\Ubt^{\mac,*}$ in {\tt MAESTRO}}
-In {\tt advance\_premac}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac,*}$.  We will only compute the normal component of velocity at each face.\\
-
-The force terms are:
-\begin{equation}
-\fb =
-\left(\begin{array}{c}
-f_u \\
-f_w \\
-\end{array}\right)
-= \frac{1}{\rho}\nabla\pi + \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r - \frac{(\rho-\rho_0)}{\rho}g\eb_r.\label{Force Term}
-\end{equation}
-\subsection{2D Cartesian Case}
-We begin by computing a time-centered 1D extrapolation of $u$ to r-faces:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_r}^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
-u_{R,\ib-\half\eb_r}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $w^{\trans}$:
-\begin{equation}
-u_{\ib-\half\eb_r}^{1D} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
-u_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
-\end{cases}
-\end{equation}
-Then, we do a full-dimensional extrapolation of $u$ to x-faces:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} = u_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}\right)\Delta_x u_{\ib-\eb_x}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
-u_{R,\ib-\half\eb_x}^{\mac,*} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib}\right)\Delta_x u_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
-\end{eqnarray}
-Finally, we solve a Riemann problem:
-\begin{equation}
-u_{\ib-\half\eb_x}^{\mac,*} =
-\begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} > 0, \\
-u_{R,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} < 0. 
-\end{cases}
-\end{equation}
-
-To compute $\wt^{\mac,*}$, we use an analogous procedure for computing $u^{\mac,*}$, except that we need to incorporate the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ and $\partial w_0/\partial r$ into the full-dimensional extrapolation.
-We compute a time-centered 1D extrapolation of $\wt$ to x-faces:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_x}^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
-\wt_{R,\ib-\half\eb_x}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \wt_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $u^{\trans}$:
-\begin{equation}
-\wt_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_x}^{1D} + \wt_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
-\wt_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-Then, we do a full-dimensional extrapolation of $\wt$ to r-faces, incorporating the extra term proportional to $w_0$ by using $w$ instead of $\wt$ as the advection velocity and the extra term proportional to $\partial w_0/\partial r$ as a separate term:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{\mac,*} = \wt_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}\right)\Delta_r \wt_{\ib-\eb_r}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)(w_{0,\ib-\half\eb_r} - w_{0,\ib-\frac{3}{2}\eb_r}) + \frac{\dt}{2}f_{w,\ib-\eb_r},\label{wt_L Extrapolation} \\
-\wt_{R,\ib-\half\eb_r}^{\mac,*} = \wt_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib}\right)\Delta_r \wt_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)(w_{0,\ib+\half\eb_r} - w_{0,\ib-\half\eb_r}) + \frac{\dt}{2}f_{w,\ib}.\label{wt_R Extrapolation}
-\end{eqnarray}
-Next, we solve a Riemann problem:
-\begin{equation}
-\wt_{\ib-\half\eb_r}^{\mac,*} =
-\begin{cases}
-0, & \left(w_L^{\mac,*} \le 0 ~~ {\rm AND} ~~ w_R^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|w_L^{\mac,*} + w_R^{\mac,*}\right| < \epsilon, \\
-\wt_{L,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{\mac,*}, & w_L^{\mac,*} + w_R^{\mac,*} < 0. 
-\end{cases}
-\end{equation}
-We then project $\Ubt^{\mac,*}$ to obtain $\Ubt^{\mac}$ and we call {\tt addw0} to compute $\Ub^{\mac}$ from $\Ubt^{\mac}$.
-%\subsection{3D Cartesian Case}
-%\subsection{3D Spherical Case}
-\cleardoublepage
-\section{Computing $\rho^{'\edge}, X_k^{\edge}$, and $(\rho h)^{'\edge}/h^{\edge}/T^{\edge}$ in {\tt MAESTRO}}
-In {\tt scalar\_advance}, we call {\tt make\_edge\_scal} to compute $\rho^{'\edge}, X_k^{\edge}$, and $(\rho h)^{'\edge}/h^{\edge}/T^{\edge}$ at each edge.
-The procedure is the same for each scalar, so we shall simply denote the scalar as $s$.
-We always need to compute $\rho'$ and $X_k$ to edges, and the choice of energy prediction is as follows:
-\begin{itemize}
-\item For {\tt enthalpy\_pred\_type} = 1, we predict $(\rho h)'$ to edges.
-\item For {\tt enthalpy\_pred\_type} = 2, we predict $h$ to edges.
-\item For {\tt enthalpy\_pred\_type} = 3 and 4, we predict $T$ to edges.
-\end{itemize}
-The force terms are:
-\begin{eqnarray}
-f_{\rho'} &=& -\rho'\nabla\cdot\Ub^{\mac} - \nabla\cdot(\rho_0\Ubt^{\mac}), \\
-f_{X_k} &=& 0, \\
-f_{(\rho h)'} &=& (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} -(\rho h)'\nabla\cdot\Ub^{\mac} - \nabla\cdot\left[(\rho h)_0\Ubt^{\mac}\right] + \nabla\cdot\kth\nabla T \\
-f_h &=& \frac{1}{\rho}\left(\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T \right), \\
-f_T &=& \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext} + \nabla\cdot\kth\nabla T \right\}.
-\end{eqnarray}
-If {\tt enthalpy\_pred\_type} = 3 or 4, we convert $T^{\edge}\rightarrow h^{\edge}$ using $h^{\edge} = h^{\edge}(\rho^{\edge},T^{\edge},X_k^{\edge})$, where $\rho^{\edge} = \rho^{'\edge} + \rho_0$.
-Furthermore, if {\tt enthalpy\_pred\_type} = 3, we compute $(\rho h)^{'\edge} = (\rho h)^{\edge} - (\rho h)_0$.
-\subsection{2D Cartesian Case}
-We begin by computing a time-centered 1D extrapolation of $s$ to all faces.  The extrapolation of $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib-\eb_x}^n, \\
-s_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $u^{\mac}$:
-\begin{equation}
-s_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{1D} + s_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-The extrapolation of $s$ to r-faces is analogous to the extrapolation to x-faces:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_r}^{1D} &=& s_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib-\eb_r}^n, \\
-s_{R,\ib-\half\eb_r}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $w^{\mac}$:
-\begin{equation}
-s_{\ib-\half\eb_r}^{1D} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_r}^{1D} + s_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_{L,\ib-\half\eb_r}^{1D}, & w^{\mac}_{\ib-\half\eb_r} > 0, \\
-u_{R,\ib-\half\eb_r}^{1D}, & w^{\mac}_{\ib-\half\eb_r} < 0. \\
-\end{cases}
-\end{equation}
-Now we do a full-dimensional extrapolation of $s$ to each face.  The extrapolation of $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\mac}+w_{\ib-\eb_x-\half\eb_r}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_r}^{1D} - s_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\mac}+w_{\ib-\half\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_r}^{1D} - s_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$.
-\begin{equation}
-s_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{\edge} + s_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
-\end{equation}
-The full-dimensional extrapolation of $s$ to r-faces is analogous to the extrapolation to x-faces:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_r}^{\edge} = s_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib-\eb_r}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\mac}+u_{\ib-\half\eb_x-\eb_r}^{\mac}\right)\left(s_{\ib+\half\eb_x-\eb_r}^{1D} - s_{\ib-\half\eb_x-\eb_r}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_r},  \\
-s_{R,\ib-\half\eb_r}^{\edge} = s_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r s_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\mac}+u_{\ib-\half\eb_x}^{\mac}\right)\left(s_{\ib+\half\eb_x}^{1D} - s_{\ib-\half\eb_x}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}. 
-\end{eqnarray}
-Then, we upwind based on $w^{\mac}$.
-\begin{equation}
-s_{\ib-\half\eb_r}^{\edge} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_r}^{\edge} + s_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
-s_{L,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} > 0, \\
-s_{R,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} < 0.
-\end{cases}
-\end{equation}
-%\subsection{3D Cartesian Case}
-%\subsection{3D Spherical Case}
-\cleardoublepage
-\section{Computing $\Ubt^{\edge}$ in {\tt MAESTRO}}
-In {\tt velocity\_advance}, we call {\tt make\_edge\_state} to compute $\Ubt^{\edge}$.  Unlike when we computed $\Ubt^{\trans}$ and $\Ubt^{\mac,*}$, we need all components of velocity at each face.
-\subsection{2D Cartesian Case}
-We begin by computing a time-centered 1D extrapolation of $u$ to all faces.  The extrapolation of $u$ to x-faces is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{1D} &=& u_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x u_{\ib-\eb_x}^n, \\
-u_{R,\ib-\half\eb_x}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x u_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $u^{\trans}$:
-\begin{equation}
-u_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
-u_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
-u_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-The extrapolation of $u$ to r-faces is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_r}^{1D} &=& u_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r u_{\ib-\eb_r}^n, \\
-u_{R,\ib-\half\eb_r}^{1D} &=& u_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r u_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $w^{\trans}$:
-\begin{equation}
-u_{\ib-\half\eb_r}^{1D} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_r}^{1D} + u_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
-u_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
-\end{cases}
-\end{equation}
-We repeat this procedure to compute a time-centered 1D extrapolation of $\wt$ at each face.  The extrapolation of $\wt$ to x-faces is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_x}^{1D} &=& \wt_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\eb_x}^n\right)\Delta_x \wt_{\ib-\eb_x}^n, \\
-\wt_{R,\ib-\half\eb_x}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib}^n\right)\Delta_x \wt_{\ib}^n.
-\end{eqnarray}
-Then we upwind based on $u^{\trans}$:
-\begin{equation}
-\wt_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_x}^{1D} + \wt_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\trans}_{\ib-\half\eb_x}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} > 0, \\
-\wt_{R,\ib-\half\eb_x}^{1D}, & u^{\trans}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-The extrapolation of $\wt$ to r-faces is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{1D} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r}, \label{1D wt_L extrap} \\
-\wt_{R,\ib-\half\eb_r}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}.
-\label{1D wt_R extrap}
-\end{eqnarray}
-Then we upwind based on $w^{\trans}$:
-\begin{equation}
-\wt_{\ib-\half\eb_r}^{1D} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_r}^{1D} + \wt_{R,\ib-\half\eb_r}^{1D}\right), & \left|w^{\trans}_{\ib-\half\eb_r}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{1D}, & w^{\trans}_{\ib-\half\eb_r} < 0. \\
-\end{cases}
-\end{equation}
-Now we do a full-dimensional extrapolation of $u$ to each face.  The extrapolation of $u$ to x-faces is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\edge} = u_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib-\eb_x}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(u_{\ib-\eb_x+\half\eb_r}^{1D} - u_{\ib-\eb_x-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
-u_{R,\ib-\half\eb_x}^{\edge} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x u_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_r}^{1D} - u_{\ib-\half\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$.
-\begin{equation}
-u_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_x}^{\edge} + u_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-u_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-u_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
-\end{equation}
-The full-dimensional extrapolation of $u$ to r-faces is:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_r}^{\edge} = u_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib-\eb_r}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(u_{\ib+\half\eb_x-\eb_r}^{1D} - u_{\ib-\half\eb_x-\eb_r}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_r},  \\
-u_{R,\ib-\half\eb_r}^{\edge} = u_{\ib}^n &-& \left(\half + \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r u_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(u_{\ib+\half\eb_x}^{1D} - u_{\ib-\half\eb_x}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}. 
-\end{eqnarray}
-Then, we upwind based on $w^{\mac}$.
-\begin{equation}
-u_{\ib-\half\eb_r}^{\edge} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_r}^{\edge} + u_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
-u_{L,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} > 0, \\
-u_{R,\ib-\half\eb_r}^{\edge}, & u^{\mac}_{\ib-\half\eb_r} < 0.
-\end{cases}
-\end{equation}
-Now we do a full-dimensional extrapolatino of $\wt$ to each face.
-This is different from the extrapolation of $u$ to faces because we must include the term proportional to $\partial w_0/\partial r$ in equation (\ref{Perturbational Velocity Equation}).
-The full-dimensional extrapolation of $\wt$ to x-faces is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_x}^{\edge} = \wt_{\ib-\eb_x}^n &+& \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib-\eb_x}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_r}^{\trans}+w_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(\wt_{\ib-\eb_x+\half\eb_r}^{1D} - \wt_{\ib-\eb_x-\half\eb_r}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib-\eb_x+\half\eb_r}^{\trans}+\wt_{\ib-\eb_x-\half\eb_r}^{\trans}\right)\left(w_{0,\ib-\eb_x+\half\eb_r}-w_{0,\ib-\eb_x-\half\eb_r}\right) + \frac{\dt}{2}f_{w,\ib-\eb_x}, \label{wt_L x-edge} \\
-\wt_{R,\ib-\half\eb_x}^{\edge} = \wt_{\ib}^n &-& \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x \wt_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_r}^{\trans}+w_{\ib-\half\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_r}^{1D} - \wt_{\ib-\half\eb_r}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right) + \frac{\dt}{2}f_{w,\ib}. \label{wt_R x-edge}
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$:
-\begin{equation}
-\wt_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_x}^{\edge} + \wt_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-\wt_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
-\end{equation}
-The full-dimensional extrapolation of $\wt$ to r-faces is:
-\begin{eqnarray}
-\wt_{L,\ib-\half\eb_r}^{\edge} = \wt_{\ib-\eb_r}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib-\eb_r}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x-\eb_r}^{\trans}+u_{\ib-\half\eb_x-\eb_r}^{\trans}\right)\left(\wt_{\ib+\half\eb_x-\eb_r}^{1D} - \wt_{\ib-\half\eb_x-\eb_r}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib-\half\eb_r}^{\trans}+\wt_{\ib-\frac{3}{2}\eb_r}^{\trans}\right)\left(w_{0,\ib-\half\eb_r}-w_{0,\ib-\frac{3}{2}\eb_r}\right) + \frac{\dt}{2}f_{w,\ib-\eb_r}, \label{wt_L r-edge} \\
-\wt_{R,\ib-\half\eb_r}^{\edge} = \wt_{\ib}^n &+& \left(\half - \frac{\dt}{2h}w_{\ib-\half\eb_r}^{\mac}\right)\Delta_r \wt_{\ib}^n \nonumber \\
-&-& \frac{\dt}{4h}\left(u_{\ib+\half\eb_x}^{\trans}+u_{\ib-\half\eb_x}^{\trans}\right)\left(\wt_{\ib+\half\eb_x}^{1D} - \wt_{\ib-\half\eb_x}^{1D}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(\wt_{\ib+\half\eb_r}^{\trans}+\wt_{\ib-\half\eb_r}^{\trans}\right)\left(w_{0,\ib+\half\eb_r}-w_{0,\ib-\half\eb_r}\right) + \frac{\dt}{2}f_{w,\ib}. \label{wt_R r-edge}
-\end{eqnarray}
-Then, we upwind based on $w^{\mac}$:
-\begin{equation}
-\wt_{\ib-\half\eb_r}^{\edge} =
-\begin{cases}
-\half\left(\wt_{L,\ib-\half\eb_r}^{\edge} + \wt_{R,\ib-\half\eb_r}^{\edge}\right), & \left|w^{\mac}_{\ib-\half\eb_r}\right| < \epsilon \\
-\wt_{L,\ib-\half\eb_r}^{\edge}, & w^{\mac}_{\ib-\half\eb_r} > 0, \\
-\wt_{R,\ib-\half\eb_r}^{\edge}, & w^{\mac}_{\ib-\half\eb_r} < 0.
-\end{cases}
-\end{equation}
-%\subsection{3D Cartesian Case}
-%\subsection{3D Spherical Case}
-\cleardoublepage
-\section{Computing $\Ub^{\mac,*}$ in {\tt VARDEN}}
-\subsection{2D Cartesian Case}
-We do a 1D Taylor series extrapolation to get both components of velocity at the x-face:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{1D} &=& u_{\ib-\eb_x} + \left[\half - \frac{\dt}{2h}{\rm max}(0,u_{\ib-\eb_x})\right]\Delta_xu_{\ib-\eb_x}, \label{varden U_L^1D} \\
-u_{R,\ib-\half\eb_x}^{1D} &=& u_{\ib} + \left[\half - \frac{\dt}{2h}{\rm min}(0,u_{\ib})\right]\Delta_xu_{\ib}.
-\end{eqnarray}
-\begin{eqnarray}
-v_{L,\ib-\half\eb_x}^{1D} &=& v_{\ib-\eb_x} + \left[\half - \frac{\dt}{2h}{\rm max}(0,v_{\ib-\eb_x})\right]\Delta_xv_{\ib-\eb_x}, \label{varden U_L^1D} \\
-v_{R,\ib-\half\eb_x}^{1D} &=& v_{\ib} + \left[\half - \frac{\dt}{2h}{\rm min}(0,v_{\ib})\right]\Delta_xv_{\ib}.
-\end{eqnarray}
-We obtain the normal velocity using the Riemann problem:
-\begin{equation}
-u_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{1D} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{1D} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{1D}, & u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D} > 0, \\
-u_{R,\ib-\half\eb_x}^{1D}, & u_{L,\ib-\half\eb_x}^{1D} + u_{R,\ib-\half\eb_x}^{1D} < 0.
-\end{cases}
-\end{equation}
-We obtain the transverse velocity by upwinding based on $u_{\ib-\half\eb_x}^{1D}$:
-\begin{equation}
-v_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(v_{L,\ib-\half\eb_x}^{1D} + v_{R,\ib-\half\eb_x}^{1D}\right), & \left|u_{\ib-\half\eb_x}^{1D}\right| < \epsilon \\
-v_{L,\ib-\half\eb_x}^{1D}, & u_{\ib-\half\eb_x}^{1D} > 0, \\
-v_{R,\ib-\half\eb_x}^{1D}, & u_{\ib-\half\eb_x}^{1D} < 0.
-\end{cases}\label{Transverse Velocity Riemann Problem}
-\end{equation}
-We perform analogous operations to compute both components of velocity at the y-faces, $\Ub_{\ib-\half\eb_y}^{1D}$. \\
-
-Now we do a full-dimensional extrapolation to get the MAC velocity at the x-faces (note that we only compute the normal components):
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} &=& u_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{1D} - u_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
-u_{R,\ib-\half\eb_x}^{\mac,*} &=& u_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{1D} - u_{\ib-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{u,\ib}.
-\end{eqnarray}
-Then, we solve a Riemann problem:
-\begin{equation}
-u_{\ib-\half\eb_x}^{\mac,*} =
-\begin{cases}
-0, & \left(u_{L,\ib-\half\eb_x}^{\mac,*} \le 0 ~~ {\rm AND} ~~ u_{R,\ib-\half\eb_x}^{\mac,*} \ge 0\right) ~~ {\rm OR} ~~ \left|u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*}\right| < \epsilon, \\
-u_{L,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} > 0, \\
-u_{R,\ib-\half\eb_x}^{\mac,*}, & u_{L,\ib-\half\eb_x}^{\mac,*} + u_{R,\ib-\half\eb_x}^{\mac,*} < 0.
-\end{cases}\label{umac Riemann Problem}
-\end{equation}
-We perform analogous operations to compute the normal velocity at the y-faces, $v^{\mac,*}_{\ib-\half\eb_y}$.
-\subsection{3D Cartesian Case}
-This is more complicated than the 2D case because we include corner coupling.
-We compute $\Ub_{\ib-\half\eb_x}^{1D}, \Ub_{\ib-\half\eb_y}^{1D}$, and $\Ub_{\ib-\half\eb_z}^{1D}$ in an analogous manner as equations (\ref{varden U_L^1D})-(\ref{Transverse Velocity Riemann Problem}).
-Then, we compute an intermediate state, $u_{\ib-\half\eb_y}^{y|z}$, which is described as ``state $u_{\ib-\half\eb_y}^{1D}$ that has been updated to account for the transverse derivatives in the z direction'', using:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_y}^{y|z} &=& u_{L,\ib-\half\eb_y}^{1D} - \frac{\dt}{6h}\left(w_{\ib-\eb_y+\half\eb_z}^{1D}+w_{\ib-\eb_y-\half\eb_z}^{1D}\right)\left(u_{\ib-\eb_y+\half\eb_z}^{1D}-u_{\ib-\eb_y-\half\eb_z}^{1D}\right), \\
-u_{R,\ib-\half\eb_y}^{y|z} &=& u_{R,\ib-\half\eb_y}^{1D} - \frac{\dt}{6h}\left(w_{\ib+\half\eb_z}^{1D}+w_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{1D}-u_{\ib-\half\eb_z}^{1D}\right).
-\end{eqnarray}
-Then, upwind based on $v_{\ib-\half\eb_y}^{1D}$:
-\begin{equation}
-u_{\ib-\half\eb_y}^{y|z} =
-\begin{cases}
-\half\left(u_{L,\ib-\half\eb_y}^{y|z} + u_{R,\ib-\half\eb_y}^{y|z}\right), & \left|v_{\ib-\half\eb_y}^{1D}\right| < \epsilon \\
-u_{L,\ib-\half\eb_y}^{y|z}, & v_{\ib-\half\eb_y}^{1D} > 0, \\
-u_{R,\ib-\half\eb_y}^{y|z}, & v_{\ib-\half\eb_y}^{1D} < 0.
-\end{cases}
-\end{equation}
-We use an analogous procedure to compute five more intemediate states, $u_{\ib-\half\eb_z}^{z|y}, v_{\ib-\half\eb_x}^{x|z}, v_{\ib-\half\eb_z}^{z|x}, w_{\ib-\half\eb_x}^{x|y}$, and $w_{\ib-\half\eb_y}^{y|x}$.
-Then, we do a full-dimensional extrapolation to get the MAC velocities at normal faces:
-\begin{eqnarray}
-u_{L,\ib-\half\eb_x}^{\mac,*} = u_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{1D}+v_{\ib-\eb_x-\half\eb_y}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_y}^{y|z}-u_{\ib-\eb_x-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_z}^{1D}+w_{\ib-\eb_x-\half\eb_z}^{1D}\right)\left(u_{\ib-\eb_x+\half\eb_z}^{z|y}-u_{\ib-\eb_x-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{u,\ib-\eb_x}, \\
-u_{R,\ib-\half\eb_x}^{\mac,*} = u_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{1D}+v_{\ib-\half\eb_y}^{1D}\right)\left(u_{\ib+\half\eb_y}^{y|z}-u_{\ib-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{1D}+w_{\ib-\half\eb_z}^{1D}\right)\left(u_{\ib+\half\eb_z}^{z|y}-u_{\ib-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{u,\ib}.
-\end{eqnarray}
-Then, we use the Riemann solver given above for the 2D case (equation [\ref{umac Riemann Problem}]) to compute $u_{\ib-\half\eb_x}^{\mac,*}$.
-We use an analogous procedure to obtain $v_{\ib-\half\eb_y}^{\mac,*}$ and $w_{\ib-\half\eb_z}^{\mac,*}$.
-\cleardoublepage
-\section{Computing $\Ub^{\edge}$ and $\rho^{\edge}$ in {\tt VARDEN}}
-To compute $\Ub^{\edge}$, {\tt VARDEN} uses the exact same algorithm as the $s^{\edge}$ case in {\tt MAESTRO}.
-The algorithm for $\rho^{\edge}$ in {\tt VARDEN} is slightly different than the $s^{\edge}$ case in {\tt MAESTRO} since it uses a ``conservative'' formulation.
-Here, $s$ is used in place of either $\rho, u, v$, or $w$ (in 3D).
-\subsection{2D Cartesian Case}
-The 1D extrapolation is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{1D} &=& s_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib-\eb_x}^n, \label{varden s_L^1D}\\
-s_{R,\ib-\half\eb_x}^{1D} &=& s_{\ib} - \left(\half + \frac{\dt}{2h}u_{\ib-\half\eb_x}^{\mac}\right)\Delta_x s_{\ib}^n. \label{varden s_R^1D}
-\end{eqnarray}
-Then we upwind based on $u^{\mac}$:
-\begin{equation}
-s_{\ib-\half\eb_x}^{1D} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{1D} + s_{R,\ib-\half\eb_x}^{1D}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{1D}, & u^{\mac}_{\ib-\half\eb_x} < 0. \\
-\end{cases}
-\end{equation}
-We use an analogous procedure to obtain $s_{\ib-\half\eb_y}^{1D}$.  Now we do a full-dimensional extrapolation of $s$ to each face.  The extrapolation of a ``non-conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right) + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-The extrapolation of a ``conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib-\eb_x+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\eb_x-\half\eb_y}\right] \nonumber \\
-&-& \frac{\dt}{2}s_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} - v_{\ib-\eb_x-\half\eb_y}^{\mac}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{1D} v^{\mac})_{\ib+\half\eb_y} - (s^{1D} v^{\mac})_{\ib-\half\eb_y}\right] \nonumber \\
-&-& \frac{\dt}{2}s_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac} - v_{\ib-\half\eb_y}^{\mac}\right) + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$.
-\begin{equation}
-s_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{\edge} + s_{R,\ib-\half\eb_x}^{\edge}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{\edge}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}\label{varden s^edge upwind}
-\end{equation}
-We use an analogous procedure to compute $s_{\ib-\half\eb_y}^{\edge}$.
-\subsection{3D Cartesian Case}
-This is more complicated than the 2D case because we include corner coupling.
-We first compute $s_{\ib-\half\eb_x}^{1D}$, $s_{\ib-\half\eb_y}^{1D}$, and $s_{\ib-\half\eb_z}^{1D}$ in an analogous manner to equations (\ref{varden s_L^1D}) and (\ref{varden s_R^1D}).
-Then we compute six intermediate states, $s_{\ib-\half\eb_x}^{x|y}, s_{\ib-\half\eb_x}^{x|z}, s_{\ib-\half\eb_y}^{y|x}, s_{\ib-\half\eb_y}^{y|z}, s_{\ib-\half\eb_z}^{z|x}$, and $s_{\ib-\half\eb_z}^{z|y}$.
-For the ``non-conservative case'', we use, for example:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{x|y} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{6h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac} + v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{1D} - s_{\ib-\eb_x-\half\eb_y}^{1D}\right), \\
-s_{R,\ib-\half\eb_x}^{x|y} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{6h}\left(v_{\ib+\half\eb_y}^{\mac} + v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{1D} - s_{\ib-\half\eb_y}^{1D}\right).
-\end{eqnarray}
-For the ``conservative'' case, we use, for example:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{x|y} &=& s_{L,\ib-\half\eb_x}^{1D} - \frac{\dt}{3h}\left[(sv^{\mac})_{\ib-\eb_x+\half\eb_y}-(sv^{\mac})_{\ib-\eb_x-\half\eb_y}\right], \\
-s_{R,\ib-\half\eb_x}^{x|y} &=& s_{R,\ib-\half\eb_x}^{1D} - \frac{\dt}{3h}\left[(sv^{\mac})_{\ib+\half\eb_y}-(sv^{\mac})_{\ib-\half\eb_y}\right].
-\end{eqnarray}
-Then, we upwind based on $u^{\mac}$:
-\begin{equation}
-s_{\ib-\half\eb_x}^{\edge} =
-\begin{cases}
-\half\left(s_{L,\ib-\half\eb_x}^{x|y} + s_{R,\ib-\half\eb_x}^{x|y}\right), & \left|u^{\mac}_{\ib-\half\eb_x}\right| < \epsilon \\
-s_{L,\ib-\half\eb_x}^{x|y}, & u^{\mac}_{\ib-\half\eb_x} > 0, \\
-s_{R,\ib-\half\eb_x}^{x|y}, & u^{\mac}_{\ib-\half\eb_x} < 0.
-\end{cases}
-\end{equation}
-We use an analogous procedure to compute the other five intermediate states.
-Now we do a full-dimensional extrapolation of $s$ to each face.
-The extrapolation of a ``non-conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}+v_{\ib-\eb_x-\half\eb_y}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_y}^{y|z}-s_{\ib-\eb_x-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib-\eb_x+\half\eb_z}^{\mac}+w_{\ib-\eb_x-\half\eb_z}^{\mac}\right)\left(s_{\ib-\eb_x+\half\eb_z}^{z|y}-s_{\ib-\eb_x-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{4h}\left(v_{\ib+\half\eb_y}^{\mac}+v_{\ib-\half\eb_y}^{\mac}\right)\left(s_{\ib+\half\eb_y}^{y|z}-s_{\ib-\half\eb_y}^{y|z}\right) \nonumber \\
-&-& \frac{\dt}{4h}\left(w_{\ib+\half\eb_z}^{\mac}+w_{\ib-\half\eb_z}^{\mac}\right)\left(s_{\ib+\half\eb_z}^{z|y}-s_{\ib-\half\eb_z}^{z|y}\right) + \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-The extrapolation of a ``conserved'' $s$ to x-faces is:
-\begin{eqnarray}
-s_{L,\ib-\half\eb_x}^{\edge} = s_{L,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{y|z}v^{\mac})_{\ib-\eb_x+\half\eb_y}-({s^{y|z}v^{\mac})_{\ib-\eb_x-\half\eb_y}}\right] \nonumber \\
-&-& \frac{\dt}{2h}\left[(s^{z|y}w^{\mac})_{\ib-\eb_x+\half\eb_z}-({s^{z|y}w^{\mac})_{\ib-\eb_x-\half\eb_z}}\right] \nonumber \\
-&-& \frac{\dt}{2}s_{\ib-\eb_x}(\nabla\cdot\Ub^{\mac})_{\ib-\eb_x} + \frac{\dt}{2h}s_{\ib-\eb_x}\left(v_{\ib-\eb_x+\half\eb_y}^{\mac}-v_{\ib-\eb_x-\half\eb_y}^{\mac}+w_{\ib-\eb_x+\half\eb_z}^{\mac}-w_{\ib-\eb_x-\half\eb_z}^{\mac}\right) \nonumber \\
-&+& \frac{\dt}{2}f_{s,\ib-\eb_x}, \\
-s_{R,\ib-\half\eb_x}^{\edge} = s_{R,\ib-\half\eb_x}^{1D} &-& \frac{\dt}{2h}\left[(s^{y|z}v^{\mac})_{\ib+\half\eb_y}-({s^{y|z}v^{\mac})_{\ib-\half\eb_y}}\right] \nonumber \\
-&-& \frac{\dt}{2h}\left[(s^{z|y}w^{\mac})_{\ib+\half\eb_z}-({s^{z|y}w^{\mac})_{\ib-\half\eb_z}}\right] \nonumber \\
-&-& \frac{\dt}{2}s_{\ib}(\nabla\cdot\Ub^{\mac})_{\ib} + \frac{\dt}{2h}s_{\ib}\left(v_{\ib+\half\eb_y}^{\mac}-v_{\ib-\half\eb_y}^{\mac}+w_{\ib+\half\eb_z}^{\mac}-w_{\ib-\half\eb_z}^{\mac}\right) \nonumber \\
-&+& \frac{\dt}{2}f_{s,\ib}.
-\end{eqnarray}
-Then we upwind based on $u^{\mac}$, as in equation (\ref{varden s^edge upwind}).  We use an analogous procedure to compute both $s_{\ib-\half\eb_y}^{\edge}$ and $s_{\ib-\half\eb_z}$.
 \end{document}

### 2008-10-24 docs/flowchart/flowchart.tex
separate out the content so we can produce a master document
\ut -> \Ut to avoid multiple definitions

--- 2008-10-24 docs/flowchart/paper.tex
@@
-\documentclass[11pt]{article}
-
-\tolerance=600
-
-\usepackage{amsmath,color,longtable}
-
-% Margins
-\usepackage[lmargin=0.5in,rmargin=1.5in,tmargin=1.0in,bmargin=1.0in]{geometry}
-
-\newcommand{\rhozero}{\rho_0}
-\newcommand{\pizero}{\pi_0}
-\newcommand{\pizeroone}{\pi_0^{(1)}}
-\newcommand{\pizerotwo}{\pi_0^{(2)}}
-\newcommand{\gammabar}{\overline{\Gamma}_1}
-
-\newcommand{\nablab}{\mathbf{\nabla}}
-\newcommand{\cdotb}{\mathbf{\cdot}}
-
-\newcommand{\sfrac}[2]{\mathchoice
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
-  {#1\!/#2}}
-
-\newcommand{\half}{\sfrac{1}{2}}
-\newcommand{\myhalf}{\sfrac{1}{2}}
-\newcommand{\nph}{{n + \myhalf}}
-\newcommand{\nmh}{{n - \myhalf}}
-
-\newcommand{\Hext}{{H_{\rm ext}}}
-\newcommand{\Hnuc}{{H_{\rm nuc}}}
-\newcommand{\kth}{k_{\rm th}}
-
-\newcommand{\Sbar}{\overline{S}}
-
-\newcommand{\inp}{\mathrm{in}}
-\newcommand{\initp}{\mathrm{init}}
-\newcommand{\outp}{\mathrm{out}}
-
-\newcommand{\uadv}{\widetilde{\mathbf{U}}^{\mathrm{ADV}}}
-\newcommand{\uadvone}{\widetilde{\mathbf{U}}^{\mathrm{ADV},\star}}
-\newcommand{\uadvtwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV}}}
-\newcommand{\V}{\mathbf{V}}
-\newcommand{\dt}{\Delta t}
-\newcommand{\dr}{\Delta r}
-
-\newcommand{\etarho}{\eta_{\rho}}
-
-\newcommand{\ubold}{\mathbf{U}}
-\newcommand{\ut}{\widetilde{\ubold}}
-
-\newcommand{\omegadot}{\dot{\omega}}
-\newcommand{\er}{\mathbf{e}_r}
-
-\setlength{\marginparwidth}{1.0in}
-\newcommand{\MarginPar}[1]{\marginpar{%
-\vskip-\baselineskip %raise the marginpar a bit
-\raggedright\tiny\sffamily
-\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
-
-\title{Flowchart of Current Algorithm}
-
-\begin{document}
-
-\maketitle
-\tableofcontents
-\cleardoublepage
-
 \section{Current Algorithm}
@@
 \hline
-$\ubold$     & total velocity ($\ubold = \ut + w_0 \er$) & cm~s$^{-1}$ \\
-\hline
-$\ut$   & local velocity                             & cm~s$^{-1}$ \\
+$\ubold$     & total velocity ($\ubold = \Ut + w_0 \er$) & cm~s$^{-1}$ \\
+\hline
+$\Ut$   & local velocity                             & cm~s$^{-1}$ \\
 \hline
@@
 \hline
-$\etarho$ & $\etarho \equiv \overline{ (\rho' \ut \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\
+$\etarho$ & $\etarho \equiv \overline{ (\rho' \Ut \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\
 \hline
@@
 \begin{equation}
-\int_{\Omega_H} \ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
+\int_{\Omega_H} \Ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
 \end{equation}
@@
 \begin{equation}
-\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )\enskip ,
+\nablab \cdotb (\beta_0 \Ut )  = \beta_0 \left(S - \Sbar \right )\enskip ,
 \label{eq:tildeconstraint}
@@
 The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
-We update the velocity field $\ut^n$ to $\ut^{n+1,\star}$ by discretizing
-\begin{equation}
-\frac{\partial\ut}{\partial t} = - \ut \cdotb \nablab \ut - w_0 \frac{\partial \ut}{\partial r}
-                                 - \left(\ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
+We update the velocity field $\Ut^n$ to $\Ut^{n+1,\star}$ by discretizing
+\begin{equation}
+\frac{\partial\Ut}{\partial t} = - \Ut \cdotb \nablab \Ut - w_0 \frac{\partial \Ut}{\partial r}
+                                 - \left(\Ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
                                  - \frac{1}{\rho} \nablab\pi
@@
 \begin{eqnarray}
-\ut^{n+1,\star} &=& \ut^n - \dt \;
- \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \ut \right]^\nph
+\Ut^{n+1,\star} &=& \Ut^n - \dt \;
+ \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \Ut \right]^\nph
       - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
@@
 and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
-The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \ut ]^\nph$
+The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \Ut ]^\nph$
 is described in paper II, Appendix A, with $\V = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
-component of $\ut^n$ individually.
+component of $\Ut^n$ individually.
 
@@
 \begin{equation}
-\nablab \cdotb \left(\beta_0^{\nph} \ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)
+\nablab \cdotb \left(\beta_0^{\nph} \Ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)
 + \frac{f}{\gammabar^{n+1} p_0^{n+1}}
@@
 \end{equation}
-by defining ${\bf V} = \ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
+by defining ${\bf V} = \Ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
 and solving
@@
 \begin{equation}
-\ut^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
+\Ut^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
 \end{equation}
@@
 using $w_0$ as computed in step 6 and
-$\ut^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.
+$\Ut^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.
 
@@
     The only other difference in the time advancement is that in step 11
-    we define ${\bf V} = (\ut^{1,\star} - \ut^0)$ and solve
+    we define ${\bf V} = (\Ut^{1,\star} - \Ut^0)$ and solve
     \begin{equation}  L_\beta^\rho \phi =
@@
   \end{enumerate}
-
-\end{document}

### 2008-10-24 docs/initial_models/initial_models.tex
split out the content

--- 2008-10-11 docs/initial_models/paper.tex
@@
-\documentclass[11pt]{article}
+\section{Notes on generating initial models for mapping into {\tt MAESTRO}}
 
-\tolerance=600
-
-\usepackage{amsmath,color,epsfig,epsf}
-
-% Margins
-\usepackage[lmargin=0.5in,rmargin=1.5in,tmargin=0.5in,bmargin=0.5in]{geometry}
-
-\newcommand{\sfrac}[2]{\mathchoice
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
-  {#1\!/#2}}
-
-\def\half  {\frac{1}{2}}
-\newcommand{\myhalf}{\sfrac{1}{2}}
-
-\def\dt    {\Delta t}
-\def\kth   {k_{\rm th}}
-\def\init  {\rm init}
-\def\model {\rm model}
-\def\raw   {\rm raw}
-
-\def\edge  {\rm EDGE}
-\def\mac   {\rm MAC}
-\def\trans {\rm TRANS}
-
-\def\eb    {{\bf e}}
-\def\fb    {{\bf f}}
-\def\ib    {{\bf i}}
-\def\Ub    {{\bf U}}
-
-\def\Ubt   {\widetilde{\bf U}}
-\def\ut    {\tilde{u}}
-\def\vt    {\tilde{v}}
-\def\wt    {\tilde{w}}
-
-\title{Notes on generating initial models for mapping into {\tt MAESTRO}}
-
-\setlength{\marginparwidth}{1.0in}
-\newcommand{\MarginPar}[1]{\marginpar{%
-\vskip-\baselineskip %raise the marginpar a bit
-\raggedright\tiny\sffamily
-\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
-
-\begin{document}
-
-\maketitle
-\tableofcontents
-
-\clearpage
-
-\section{Creating the Model Data from Raw Data}\label{Sec:Creating the Model Data from Raw Data}
+\subsection{Creating the Model Data from Raw Data}\label{Sec:Creating the Model Data from Raw Data}
 
@@
 resolution as the base state.  Note that for spherical problems we set
-$\Delta r = \Delta x/\mathtt{drdxfac}$.  We use a fortran subroutine to interpolate the
-raw data, yielding the model data, $\rho^{\model}, T^{\model},
-p^{\model}$, and $X^{\model}$.  The fortran subroutine then uses an
-iterative procedure to modify the model data so that it is
-thermodynamically consistent with the {\tt MAESTRO} equation of state
-(EOS), and also satisfies our chosen hydrostatic equilibrium (HSE)
-discretization,
+$\Delta r = \Delta x/\mathtt{drdxfac}$.  We use a fortran subroutine
+to interpolate the raw data, yielding the model data, $\rho^{\model},
+T^{\model}, p^{\model}$, and $X^{\model}$.  The fortran subroutine
+then uses an iterative procedure to modify the model data so that it
+is thermodynamically consistent with the {\tt MAESTRO} equation of
+state (EOS), and also satisfies our chosen hydrostatic equilibrium
+(HSE) discretization,
 \begin{equation}
@@
 
-\subsection{Multilevel Plane-Parallel}
-We do the same thing as the single-level case except that we generate model data with the same resolution as the finest base state resolution.
+\subsubsection{Multilevel Plane-Parallel}
+We do the same thing as the single-level case except that we generate
+model data with the same resolution as the finest base state
+resolution.
 
-\section{Creating the Initial Data from the Model Data}
+\subsection{Creating the Initial Data from the Model Data}
 In {\tt base\_state.f90}, using the subroutine {\tt
@@
 
-\subsection{Multilevel Plane-Parallel}
+\subsubsection{Multilevel Plane-Parallel}
 We do the same thing as the single-level case except that we generate
@@
 
-\section{Creating the Base State and Full State}\label{Sec:Creating the Base State and Full State}
+\subsection{Creating the Base State and Full State}\label{Sec:Creating the Base State and Full State}
 
@@
 
-\subsection{Multilevel Plane-Parallel}
+\subsubsection{Multilevel Plane-Parallel}
 \begin{enumerate}

### 2008-10-24 docs/initial_models/paper.tex
split out the content

--- 2008-08-27 docs/initial_models/paper.tex
@@
 
-\section{Creating the Model Data from Raw Data}\label{Sec:Creating the Model Data from Raw Data}
-We begin with raw data, $\rho^{\raw}, T^{\raw}, p^{\raw}$, and $X^{\raw}$.  Here is the raw data file for each test problem:
-\begin{eqnarray}
-{\tt test2} & \rightarrow & {\tt ???} \nonumber \\
-{\tt test\_convect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt wdconvect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt spherical\_heat} & \rightarrow & {\tt ???} \nonumber \\
-{\tt xrb} & \rightarrow & {\tt initial\_models/xrb/sorted\_xrb.raw} \nonumber
-\end{eqnarray}
-Using the raw data, we need to generate model data with the same resolution as the base state.
-Note that for spherical problems we set $\Delta r = \Delta x$.
-We use a fortran subroutine to interpolate the raw data, yielding the model data, $\rho^{\model}, T^{\model}, p^{\model}$, and $X^{\model}$.
-The fortran subroutine then uses an iterative procedure to modify the model data so that it is thermodynamically consistent with the {\tt MAESTRO} equation of state (EOS), and also satisfies our chosen hydrostatic equilibrium (HSE) discretization,
-\begin{equation}
-\frac{p_r - p_{r-1}}{\Delta r} = \frac{\rho_r + \rho_{r-1}}{2}g,\label{HSE Discretization}
-\end{equation}
-to a user defined tolerance.  Here are the fortran subroutines for each test problem:
-\begin{eqnarray}
-{\tt test2} & \rightarrow & {\tt ???} \nonumber \\
-{\tt test\_convect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt wdconvect} & \rightarrow & {\tt ???} \nonumber \\
-{\tt spherical\_heat} & \rightarrow & {\tt ???} \nonumber \\
-{\tt xrb} & \rightarrow & {\tt initial\_models/xrb/init\_1d.f90} \nonumber
-\end{eqnarray}
-The model data is not generated at run-time - it must be generated in advance of running any MAESTRO examples.  The inputs file should point to the file containing the model data.
-\subsection{Multilevel Plane-Parallel}
-We do the same thing as the single-level case except that we generate model data with the same resolution as the finest base state resolution.
-\section{Creating the Initial Data from the Model Data}
-In {\tt base\_state.f90}, using the subroutine {\tt init\_base\_state.f90} (which is actually a terrible name since we are setting the initial data, not the base state), we set the initial data $\rho^{\init}, T^{\init}, p^{\init}$ and $X^{\init}$ equal to the model data.
-Then, we set $h^{\init} = h^{\init}(\rho^{\init},T^{\init},X^{\init})$.
-Note that in the code, $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not change since we called $p^{\model} = p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the fortran subroutine in \S~\ref{Sec:Creating the Model Data from Raw Data}.
-\subsection{Multilevel Plane-Parallel}
-We do the same thing as the single-level case except that we generate initial data at the non-finest levels by using linear interpolation from the two nearest model points (see Figure \ref{Fig:Multilevel Initialization}).
-Note that the initial data is not in HSE nor is it thermodynamically consistent on non-finest cells.
-We will enforce these on the base state and full state later.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[tpb]
-\begin{center}
-\epsffile{./multilevel_init.eps}
-\caption{Multilevel Initialization.
-The data from the initial model is represented by the dots on the right.
-The initial data at the finest level is represented by the circles.
-The initial data at non-finest levels is represented by the squares.
-We copy data from the dots directly into the circles.
-We use linear interpolation using the two nearest points to copy data from the dots to the squares.}
-\label{Fig:Multilevel Initialization}
-\end{center}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Creating the Base State and Full State}\label{Sec:Creating the Base State and Full State}
-Given $p^{\init}, \rho^{\init}, T^{\init},$ and $X^{\init}$, this section describes how the base state ($\rho_0$ and $p_0$) and full state ($\rho, h, X$, and $T$) are computed.
-We require that $\rho_0 = \overline\rho$, the base state is in HSE according to equation (\ref{HSE Discretization}), and the full state is thermodynamically consistent with $p_0$.
-Here are the steps:
-\begin{enumerate}
-\item For plane-parallel problems, fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.
-For spherical problems, fill $\rho^{\init}, p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state $\rho, p$, and $X$, and then call the EOS to obtain the full state $h,T = h,T(\rho,p,X)$.
-Note that this particular multifab version of $p$ is temporary, and can be deallocated after this step.
-\item For spherical problems, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
-\item If {\tt perturb\_model} = F, skip the remaining steps and instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
-\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
-\item Set $\rho_0 = \overline\rho$.
-\item Compute $p_0$ using equation (\ref{HSE Discretization}), using the pressure at the top of the domain as the initial condition.
-Note that we still integrate upwards from the bottom, but then offset the final pressure so the pressure at the top of the domain remains unchanged.
-\item Compute the full state $T,h = T,h(\rho,p_0,X)$.
-\item Set $(\rho h)_0 = \overline{(\rho h)}$.
-\end{enumerate}
-Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is thermodynamically consistent with $p_0$.
-\subsection{Multilevel Plane-Parallel}
-\begin{enumerate}
-\item Fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state $\rho, h, X$, and $T$.
-\item If {\tt perturb\_model} = F, skip the next two steps and instead set $\rho_0 = \rho^{\init}$.
-\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
-\item Set $\rho_0 = \overline\rho$.
-\item Compute $p_0$ using equation (\ref{HSE Discretization}), using the pressure at the top of the domain as the initial condition.
-Note that we still integrate upwards from the bottom, but then offset the final pressure so the pressure at the top of the domain remains unchanged.
-Use the coarse-fine interface discretizations given in \S~\ref{Sec:Coarse-Fine HSE Discretization}.
-\item Compute the full state $T,h = T,h(\rho,p_0,X)$.
-\item Set $(\rho h)_0 = \overline{(\rho h)}$.
-\end{enumerate}
-Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full state is thermodynamically consistent with $p_0$.
-\subsubsection{Coarse-Fine HSE Discretization}\label{Sec:Coarse-Fine HSE Discretization}
-Integrating the HSE discretization upward, when we encounter a change from coarse to fine, here is an equation to integrate up to the coarse-fine interface ($l$ is the ``finer'' level, $l-1$ is the ``coarser'' level)
-\begin{equation}
-\frac{p_{r-\myhalf}^l - p_{\sfrac{r}{2}-1}^{l-1}}{\Delta r^{l-1}/2} = \frac{\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}}{2}g ~~~ \rightarrow ~~~ p_{r-\myhalf}^l = p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^{l-1} g}{4}\left(\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}\right).
-\end{equation}
-Here is an equation to integrate from the coarse-fine interface up to the cell center
-\begin{equation}
-\frac{p_r^l - p_{r-\myhalf}^l}{\Delta r^l/2} = \frac{\rho_r^l + \rho_{r-\myhalf}^l}{2}g ~~~ \rightarrow ~~~ p_r^l = p_{r-\myhalf}^l + \frac{\Delta r^l g}{4}\left(\rho_r^l + \rho_{r-\myhalf}^l\right).
-\end{equation}
-Combining equations gives
-\begin{equation}
-p_r^l = p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^{l-1} g}{4}\left(\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}\right) + \frac{\Delta r^l g}{4}\left(\rho_r^l + \rho_{r-\myhalf}^l\right).
-\end{equation}
-We can simplify using
-\begin{equation}
-\Delta r^{l-1} = 2\Delta r^l,
-\end{equation}
-\begin{equation}
-\rho_{r-\myhalf}^l = \frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}.
-\end{equation}
-Simplifying
-\begin{eqnarray}
-p_r^l &=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^l g}{2}\left(\frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_{\sfrac{r}{2}-1}^{l-1}\right) + \frac{\Delta r^l g}{4}\left(\rho_r^l + \frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}\right) \nonumber \\
-&=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{3\Delta r^l g}{4}\left(\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_r^l\right).\label{Coarse-Fine Stencil}
-\end{eqnarray}
-When we encounter a change from fine to coarse, analogously
-\begin{equation}
-p_{(r+1)/2}^{l-1} = p_{r}^l + \frac{3\Delta r^l g}{4}\left(\rho_{r}^l+\rho_{(r+1)/2}^{l-1}\right).\label{Fine-Coarse Stencil}
-\end{equation}
+\input initial_models
 \end{document}

### 2008-10-24 docs/lo_density/lo_density.tex
split out the content

--- 2008-06-20 docs/lo_density/paper.tex
@@
-\documentclass[11pt]{article}
-
-\tolerance=600
-
-\usepackage{amsmath,color}
-
-% Margins
-\usepackage[lmargin=0.5in,rmargin=0.5in,tmargin=0.75in,bmargin=0.75in]{geometry}
-
-\def\half  {\frac{1}{2}}
-\def\dt    {\Delta t}
-\def\kth   {k_{\rm th}}
-
-\def\edge  {\rm EDGE}
-\def\mac   {\rm MAC}
-\def\trans {\rm TRANS}
-
-\def\eb    {{\bf e}}
-\def\fb    {{\bf f}}
-\def\ib    {{\bf i}}
-\def\Ub    {{\bf U}}
-
-\def\Ubt   {\widetilde{\bf U}}
-\def\ut    {\tilde{u}}
-\def\vt    {\tilde{v}}
-\def\wt    {\tilde{w}}
-
-\title{Notes on the low density parameters in {\tt MAESTRO}}
-
-\setlength{\marginparwidth}{0.75in}
-\newcommand{\MarginPar}[1]{\marginpar{%
-\vskip-\baselineskip %raise the marginpar a bit
-\raggedright\tiny\sffamily
-\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
-
-\begin{document}
-
-\maketitle
-\tableofcontents
+\section{Notes on the low density parameters in {\tt MAESTRO}}
 
 These are working notes for the low density parameters in {\tt MAESTRO}.
-\section{anelastic\_cutoff}
+
+\subsection{anelastic\_cutoff}
+
 \begin{itemize}
-\item In {\tt init\_base\_state}, we abort the program if {\tt anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
-\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance\_timestep}, We compute the coordinate associated with {\tt anelastic\_cutoff}, {\tt anelastic\_cutoff\_coord}.
-This is done by setting ${\tt anelastic\_cutoff\_coord}= r$ as soon as $\rho_0(r) <$ {\tt anelastic\_cutoff}.
-\item In {\tt correct\_base}, we only update $\rho_0$ for $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
-\item In {\tt make\_div\_coeff}, we change how we compute {\tt div\_coeff} for $r\in[{\tt anelastic\_cutoff\_coord},{\tt r\_end\_coord}]$.
-\item In {\tt make\_psi}, we only compute $\psi$ for $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
-\item in {\tt make\_S}, we set {\tt delta\_gamma1\_term} and {\tt delta\_gamma1} to zero for $r\in[{\tt anelastic\_cutoff\_coord},{\tt r\_end\_coord}]$.
+
+\item In {\tt init\_base\_state}, we abort the program if {\tt
+  anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
+
+\item After we initilize the base state in {\tt varden} and at the
+  beginning of {\tt advance\_timestep}, We compute the coordinate
+  associated with {\tt anelastic\_cutoff}, {\tt
+    anelastic\_cutoff\_coord}.  This is done by setting ${\tt
+    anelastic\_cutoff\_coord}= r$ as soon as $\rho_0(r) <$ {\tt
+    anelastic\_cutoff}.
+
+\item In {\tt correct\_base}, we only update $\rho_0$ for $r\in[{\tt
+    r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+
+\item In {\tt make\_div\_coeff}, we change how we compute {\tt
+  div\_coeff} for $r\in[{\tt anelastic\_cutoff\_coord},{\tt
+    r\_end\_coord}]$.
+
+\item In {\tt make\_psi}, we only compute $\psi$ for $r\in[{\tt
+    r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+
+\item in {\tt make\_S}, we set {\tt delta\_gamma1\_term} and {\tt
+  delta\_gamma1} to zero for $r\in[{\tt anelastic\_cutoff\_coord},{\tt
+    r\_end\_coord}]$.
 \end{itemize}
-\section{base\_cutoff\_density}
+
+\subsection{base\_cutoff\_density}
+
 \begin{itemize}
-\item In {\tt base\_state}, we abort the program if {\tt anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
-\item In {\tt base\_state}, we compute a physical cutoff location, {\tt base\_cutoff\_density\_loc}, which is defined as the physical location of the first cell-center at the coarsest level for which $\rho_0 <$ {\tt base\_cutoff\_density}.
-\item In {\tt base\_state}, when we are generating the initial background/base state, if we are above {\tt base\_cutoff\_density\_loc}, just use the values for $\rho,T$, and $p$ at {\tt base\_cutoff\_density\_loc}.
-\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance\_timestep}, We compute the coordinate associated with {\tt base\_cutoff\_density}, {\tt base\_cutoff\_density\_coord}.
-This is done by setting ${\tt base\_cutoff\_density\_coord}= r$ as soon as $\rho_0(r) <$ {\tt base\_cutoff\_density}.
-\item In {\tt make\_hgrhs}, we only add the volume discrepancy for $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
-\item In {\tt make\_macrhs}, we only add the volume discrepancy for $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
-\item In {\tt mkscalforce} subroutine {\tt mkrhohforce}, for $r\in[{\tt base\_cutoff\_density\_coord},{\tt r\_end\_coord}]$ we compute $\nabla p_0$ with a difference stencil instead of simply setting it to $\rho_0 g$.
-\item In {\tt makeRhoHfromP}, the base state pressure at edges is computed differently based on where we are in relation to the {\tt base\_cutoff\_density}.
-For $r\in[{\tt r\_start\_coord},{\tt base\_cutoff\_density\_coord}-1]$ we integrate $\rho g$ from the nearest cell-centered $p_0$ value.
-Otherwise, we just use the cell-centered value above the edge.
-\item In {\tt update\_scal}, if the new $\rho$ is less than {\tt base\_cutoff\_density}, we call the EOS to compute $\rho h$.
+
+\item In {\tt base\_state}, we abort the program if {\tt
+  anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
+
+\item In {\tt base\_state}, we compute a physical cutoff location,
+  {\tt base\_cutoff\_density\_loc}, which is defined as the physical
+  location of the first cell-center at the coarsest level for which
+  $\rho_0 <$ {\tt base\_cutoff\_density}.
+
+\item In {\tt base\_state}, when we are generating the initial
+  background/base state, if we are above {\tt
+    base\_cutoff\_density\_loc}, just use the values for $\rho,T$, and
+  $p$ at {\tt base\_cutoff\_density\_loc}.
+
+\item After we initilize the base state in {\tt varden} and at the
+  beginning of {\tt advance\_timestep}, We compute the coordinate
+  associated with {\tt base\_cutoff\_density}, {\tt
+    base\_cutoff\_density\_coord}.  This is done by setting ${\tt
+    base\_cutoff\_density\_coord}= r$ as soon as $\rho_0(r) <$ {\tt
+    base\_cutoff\_density}.
+
+\item In {\tt make\_hgrhs}, we only add the volume discrepancy for
+  $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+
+\item In {\tt make\_macrhs}, we only add the volume discrepancy for
+  $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+
+\item In {\tt mkscalforce} subroutine {\tt mkrhohforce}, for
+  $r\in[{\tt base\_cutoff\_density\_coord},{\tt r\_end\_coord}]$ we
+  compute $\nabla p_0$ with a difference stencil instead of simply
+  setting it to $\rho_0 g$.
+
+\item In {\tt makeRhoHfromP}, the base state pressure at edges is
+  computed differently based on where we are in relation to the {\tt
+    base\_cutoff\_density}.  For $r\in[{\tt r\_start\_coord},{\tt
+      base\_cutoff\_density\_coord}-1]$ we integrate $\rho g$ from the
+  nearest cell-centered $p_0$ value.  Otherwise, we just use the
+  cell-centered value above the edge.
+
+\item In {\tt update\_scal}, if the new $\rho$ is less than {\tt
+  base\_cutoff\_density}, we call the EOS to compute $\rho h$.
+
+\item In {\tt make\_grav}, if we are spherical, we do not add to the
+  enclosed mass, $M(r)$ if $\rho$ is less than {\tt base\_cutoff\_density}.
 \end{itemize}
-\section{burning\_cutoff}
+
+\subsection{burning\_cutoff}
+
 \begin{itemize}
-\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance\_timestep}, We compute the coordinate associated with {\tt burning\_cutoff\_density}, {\tt burning\_cutoff\_density\_coord}.
-This is done by setting ${\tt burning\_cutoff\_density\_coord}= r$ as soon as $\rho_0(r) <$ {\tt burning\_cutoff\_density}.
-\item In {\tt react\_state}, we only call the burner if $\rho >$ {\tt burning\_cutoff\_density}.
+\item After we initilize the base state in {\tt varden} and at the
+  beginning of {\tt advance\_timestep}, We compute the coordinate
+  associated with {\tt burning\_cutoff\_density}, {\tt
+    burning\_cutoff\_density\_coord}.  This is done by setting ${\tt
+    burning\_cutoff\_density\_coord}= r$ as soon as $\rho_0(r) <$ {\tt
+    burning\_cutoff\_density}.
+
+\item In {\tt react\_state}, we only call the burner if $\rho >$ {\tt
+  burning\_cutoff\_density}.
 %\item In {\tt react\_state}, we only call the burner for $r\in[{\tt r\_start\_coord},{\tt burning\_cutoff\_density\_coord}-1]$.
 \end{itemize}
-\end{document}

### 2008-10-24 docs/lo_density/paper.tex
update base_cutoff_density notes
split out the content

--- 2008-06-20 docs/lo_density/paper.tex
@@
 
-These are working notes for the low density parameters in {\tt MAESTRO}.
-\section{anelastic\_cutoff}
-\begin{itemize}
-\item In {\tt base\_state}, we abort the program if {\tt anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
-\item After we initilize the base state in {\tt varden} and at the beginning of {\tt advance}, We compute the coordinate associated with {\tt anelastic\_cutoff}, {\tt anelastic\_cutoff\_coord}.
-This is done by setting ${\tt anelastic\_cutoff\_coord}= r$ as soon as $\rho_0(r) <$ {\tt anelastic\_cutoff}.
-\item In {\tt correct\_base}, we only update $\rho_0$ for $r\in[0,{\tt anelastic\_cutoff\_coord}-1]$.
-\item In {\tt make\_div\_coeff}, we change how we compute {\tt div\_coeff} at and above ${\tt anelastic\_cutoff\_coord}$.
-\item In {\tt make\_psi}, we only compute $\psi$ for $r\in[0,{\tt anelastic\_cutoff\_coord}-1]$.
-\item in {\tt make\_S}, we set {\tt delta\_gamma1\_term} and {\tt delta\_gamma1} to zero at and above ${\tt anelastic\_cutoff\_coord}$.
-\end{itemize}
-\section{base\_cutoff\_density}
-\begin{itemize}
-\item In {\tt base\_state}, we abort the program if {\tt anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
-\item In {\tt base\_state}, we compute a physical cutoff location, {\tt base\_cutoff\_density\_loc}, which is defined as the physical location of the first cell-center at the coarsest level for which $\rho_0 <$ {\tt base\_cutoff\_density}.
-\item In {\tt base\_state}, when we are generating the initial background/base state, if we are above {\tt base\_cutoff\_density\_loc}, just use the values for $\rho,T$, and $p$ at {\tt base\_cutoff\_density\_loc}.
-\item In {\tt make\_hgrhs}, we only add the volume discrepancy if $\rho_0 >$ {\tt base\_cutoff\_density}.
-\item In {\tt make\_macrhs}, we only add the volume discrepancy if $\rho_0 >$ {\tt base\_cutoff\_density}.
-\item In {\tt mkscalforce} subroutine {\tt mkrhohforce}, once $\rho_0 \le$ {\tt base\_cutoff\_density}, we compute $\nabla p_0$ with a difference stencil instead of simply setting it to $\rho_0 g$.
-\item In {\tt rhoh\_vs\_t} function {\tt makeRhoHfromP}, the base state pressure at edges is computed differently based on where we are in relation to the {\tt base\_cutoff\_density}.
-If we are at a equal or higher density than {\tt base\_cutoff\_density}, we integrate $\rho g$ from the nearest cell-centered $p_0$ value.
-If we are at a lower density than {\tt base\_cutoff\_density}, we just use the cell-centered value.
-\item In {\tt update\_scal}, if the new $\rho$ is less than {\tt base\_cutoff\_density}, we call the EOS to compute $\rho h$.
-\end{itemize}
-\section{burning\_cutoff}
-\begin{itemize}
-\item In {\tt react\_state}, we only call the burner if $\rho >$ {\tt burning\_cutoff\_density}.
-\end{itemize}
+\input lo_density
+
 \end{document}

### 2008-10-24 docs/initial_models/initial_models.tex
remove extraneous \end{document}
define a macro for the figure path to allow this to be imbedded in other documents
switch over to book format for the master document -- this makes the sectioning more natural

--- 2008-10-11 docs/initial_models/paper.tex
@@
-\documentclass[11pt]{article}
-
-\tolerance=600
-
-\usepackage{amsmath,color,epsfig,epsf}
-
-% Margins
-\usepackage[lmargin=0.5in,rmargin=1.5in,tmargin=0.5in,bmargin=0.5in]{geometry}
-
-\newcommand{\sfrac}[2]{\mathchoice
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
-  {#1\!/#2}}
-
-\def\half  {\frac{1}{2}}
-\newcommand{\myhalf}{\sfrac{1}{2}}
-
-\def\dt    {\Delta t}
-\def\kth   {k_{\rm th}}
-\def\init  {\rm init}
-\def\model {\rm model}
-\def\raw   {\rm raw}
-
-\def\edge  {\rm EDGE}
-\def\mac   {\rm MAC}
-\def\trans {\rm TRANS}
-
-\def\eb    {{\bf e}}
-\def\fb    {{\bf f}}
-\def\ib    {{\bf i}}
-\def\Ub    {{\bf U}}
-
-\def\Ubt   {\widetilde{\bf U}}
-\def\ut    {\tilde{u}}
-\def\vt    {\tilde{v}}
-\def\wt    {\tilde{w}}
-
-\title{Notes on generating initial models for mapping into {\tt MAESTRO}}
-
-\setlength{\marginparwidth}{1.0in}
-\newcommand{\MarginPar}[1]{\marginpar{%
-\vskip-\baselineskip %raise the marginpar a bit
-\raggedright\tiny\sffamily
-\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
-
-\begin{document}
-
-\maketitle
-\tableofcontents
-
-\clearpage
-
 \section{Creating the Model Data from Raw Data}\label{Sec:Creating the Model Data from Raw Data}
@@
 \subsection{Multilevel Plane-Parallel}
-We do the same thing as the single-level case except that we generate model data with the same resolution as the finest base state resolution.
+We do the same thing as the single-level case except that we generate
+model data with the same resolution as the finest base state
+resolution.
 
@@
 \begin{center}
-\epsffile{./multilevel_init.eps}
+\epsffile{\figpath/multilevel_init.eps}
 \caption{Multilevel Initialization.  The data from the initial model
@@
 
-\subsubsection{Coarse-Fine HSE Discretization}\label{Sec:Coarse-Fine HSE Discretization}
+\subsection{Coarse-Fine HSE Discretization}\label{Sec:Coarse-Fine HSE Discretization}
 Integrating the HSE discretization upward, when we encounter a change
@@
 \end{equation}
-
-\end{document}

### 2008-10-24 docs/pert_notes/paper.tex
finish adding chapters

--- 2008-08-14 docs/pert_notes/paper.tex
@@
 
-\section{Density Evolution}
-The full density evolution equation is
-\begin{equation}
-\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub) = -\Ub\cdot\nabla\rho 
-- \rho\nabla\cdot\Ub. \label{rho equation}
-\end{equation}
-\subsection{Predicting $\rho'$ ({\tt predict\_rho = F} - default)}
-The base state density evolution equation is
-\begin{equation}
-\frac{\partial\rho_0}{\partial t} = -\nabla\cdot(\rho_0 w_0 \eb_r) = 
--w_0\frac{\partial\rho_0}{\partial r} 
-\underbrace{-\rho_0\frac{\partial w_0}{\partial r}}_{"\rho_0 ~ \text{force}"}.
-\label{rho0 equation}
-\end{equation}
-Subtract (\ref{rho0 equation}) from (\ref{rho equation}) and rearrange terms, noting that 
-$\Ub = \Ubt + w_o\eb_r$, to obtain the perturbational density equation,
-\begin{equation}
-\frac{\partial\rho'}{\partial t} = -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub 
-- \nabla\cdot(\rho_0\Ubt)}_{"\rho' ~ \text{force}"}. 
-\label{rhoprime equation}
-\end{equation}
-\subsubsection{Predicting $\rho'$ at edges}
-We define $\rho' = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to edges using 
-{\tt make\_edge\_scal} in {\tt scalar\_advance} and the underbraced term in (\ref{rhoprime equation}) 
-as the forcing.
-\subsubsection{Predicting $\rho_0$ at edges}\label{Predicting rho0 at edges}
-There are two ways to predict $\rho_0$ at edges.
-\begin{enumerate}
-\item We call {\tt make\_edge\_state\_1d} using the underbraced term in 
-(\ref{rho0 equation}) as the forcing.  This gives us $\rho_0^{n+\half,{\rm pred}}$.
-This term is used to advect $\rho_0$ in {\tt advect\_base} and therefore must also be used 
-to compute $\eta_\rho$, which will be used in {\tt correct\_base}.
-\item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n + \rho_0^{(2)})/2$.  We 
-compute $\rho_0^{(2)}$ from $\rho_0^n$ using {\tt advect\_base}, which advances equation 
-(\ref{rho0 equation}) through $\Delta t$ in time.  The $(2)$ in the superscript indicates 
-that we have not called {\tt correct\_base} yet, which computes  $\rho_0^{n+1}$ from $\rho_0^{(2)}$. 
-We use $\rho_0^{(2)}$ rather than $\rho_0^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be 
-consistent with the fact that we do not include an $\eta_\rho$ contribution to compute 
-$\rho'$ at edges.  $\rho_0^{n+\half,{\rm avg}}$ is used to construct $\rho$ at edges from 
-$\rho'$ at edges, and this $\rho$ at edges is used to compute fluxes for $\rho X_k$.  
-\end{enumerate}
-\subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
-For the non-radial edges, we directly add $\rho_0^{n+\half,{\rm avg}}$ to $\rho'$ since 
-$\rho_0^{n+\half,{\rm avg}}$ is a cell-centered quantity.  For the radial edges, we 
-interpolate to obtain $\rho_0^{n+\half,{\rm avg}}$ at radial edges before adding it to $\rho'$.
-\subsubsection{Advancing $\rho X_k$}\label{Advancing rhoX_k}
-The evolution equation for $\rho X_k$, ignoring the reaction terms that were already 
-accounted for in {\tt react\_state}, and the associated discretization is
-\begin{equation}
-\frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
--\nabla\cdot\left\{\left[\left(\rho_0^{n+\half,{\rm avg}} 
-+ \rho'\right)X_k\right](\Ubt+w_0\eb_r)\right\}.
-\end{equation}
-\section{Energy Evolution}
-The full enthalpy equation is
-\begin{eqnarray}
-\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt} 
-+ \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext} \nonumber \\
-&=& \underbrace{-\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub}_{-\nabla\cdot(\rho h\Ub)} 
-+ \underbrace{\psi + \wt\frac{\partial p_0}{\partial r}}_{\frac{Dp_0}{Dt}} 
-+ \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}.
-\end{eqnarray}
-The full temperature equation is
-\begin{equation}
-\frac{\partial T}{\partial t} = -\Ub\cdot\nabla T
-+ \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi
-+ \wt\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 
-+ \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
-\end{equation}
-\subsection{Predicting $(\rho h)'$ ({\tt enthalpy\_pred\_type = 1} - default)}
-Due to operator splitting, the call to {\tt react\_state} has already been made.  Hence, the goal 
-is to compute an edge state enthalpy starting from $(\rho h)^{(1)}$ using an enthalpy
-equation that does not include the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms, where were 
-already accounted for in {\tt react\_state}
-\begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub 
-+ \psi + \wt\frac{\partial p_0}{\partial r} + \nabla\cdot\kappa\nabla T. \label{rhoh equation}
-\end{equation}
-We arbitrarily define a base state enthalpy evolution equation
-\MarginPar{Does $\psi$ need to be included in the force?
-I have written it in the base state equation for now, even though neither the base state or perturbational edge state prediction in the code uses $\psi$.
-How about evolving density first,  then computing a time-centered $\psi$ to be used as the energy forcing?}
-\begin{eqnarray}
-\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] 
-+ \frac{D_0p_0}{Dt} \nonumber \\
-&=& -w_0\frac{\partial(\rho h)_0}{\partial r} 
-- \underbrace{(\rho h)_0\frac{\partial w_0}{\partial r}}_{"(\rho h)_0 ~ \text{force}"}
-+ \psi.\label{rhoh0 equation}
-\end{eqnarray}
-Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and rearranging terms gives 
-the perturbational enthalpy equation
-\begin{eqnarray}
-\frac{\partial(\rho h)'}{\partial t} &=& -\nabla\cdot[(\rho h)'\Ub] 
-- \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r} 
-+ \nabla\cdot\kappa\nabla T\nonumber \\
-&=& -\Ub\cdot\nabla(\rho h)' \underbrace{- (\rho h)'\nabla\cdot\Ub 
-- \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r}
-+ \nabla\cdot\kappa\nabla T}_{"(\rho h)' ~ \text{force}"}. \label{rhohprime equation}
-\end{eqnarray}
-\subsubsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
-{\bf What the code did before 8/6/08:}
- Instead of defining $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries 
-through the complete evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ {\tt react\_base}
-$\rightarrow (\rho h)_0^{(1)} \rightarrow$ {\tt advect\_base}
-$\rightarrow (\rho h)_0^{(2)} \rightarrow$ {\tt react\_base} $\rightarrow (\rho h)_0^{n+1}$.  
-There is no {\tt correct\_base} function for $(\rho h)_0$ written yet, nor is any adjustment made 
-to sync up $(\rho h)_0$ after the thermal diffusion step.\\ \\
-{\bf What the code currently does:}  Define $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  
-and $(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'$ to 
-edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} and the underbraced term in 
-(\ref{rhohprime equation}) as the forcing.
-\subsubsection{Predicting $(\rho h)_0$ at edges}
-We use an analogous procedure described in Section \ref{Predicting rho0 at edges} to 
-obtain $(\rho h)_0^{n+\half,\rm{pred}}$ and $(\rho h)_0^{n+\half,\rm{avg}}$.  Note that 
-$(\rho h)_0^{n+\half,{\rm avg}} = [(\rho h)_0^{(1)} + (\rho h)_0^{(2)}]/2$.
-\subsubsection{Computing $\rho h$ at edges}
-We use an analogous procedure described in Section \ref{Computing rho at edges} to compute
-$\rho h$ at edges.
-\subsubsection{Advancing $\rho h$}
-{\bf What the did before 8/6/08:}
-\begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot\left[(\rho h)'\left(\Ubt 
-+ w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] 
-\underbrace{- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\eb_r\right] + \psi}_{\text{this term used to come in through} ~ (\rho h)_0^{n+1} - (\rho h)_0^n} + \wt\frac{\partial p_0}{\partial r} \enskip .
-\end{equation}
-{\bf What the code does now:} We would like to update the 
-non-forcing part of the enthalpy equation analogously to Section \ref{Advancing rhoX_k}
-\begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) = 
--\nabla\cdot\left\{\left[(\rho h)_0^{n+\half,{\rm avg}} 
-+ (\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\} + \wt\frac{\partial p_0}{\partial r} + \psi  \enskip .
-\end{equation}
+\input pert
 

### 2008-10-24 docs/pert_notes/pert.tex
finish adding chapters

\section{Density Evolution}
The full density evolution equation is
\begin{equation}
\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub) = -\Ub\cdot\nabla\rho 
- \rho\nabla\cdot\Ub. \label{rho equation}
\end{equation}
\subsection{Predicting $\rho'$ ({\tt predict\_rho = F} - default)}
The base state density evolution equation is
\begin{equation}
\frac{\partial\rho_0}{\partial t} = -\nabla\cdot(\rho_0 w_0 \eb_r) = 
-w_0\frac{\partial\rho_0}{\partial r} 
\underbrace{-\rho_0\frac{\partial w_0}{\partial r}}_{"\rho_0 ~ \text{force}"}.
\label{rho0 equation}
\end{equation}
Subtract (\ref{rho0 equation}) from (\ref{rho equation}) and rearrange terms, noting that 
$\Ub = \Ubt + w_o\eb_r$, to obtain the perturbational density equation,
\begin{equation}
\frac{\partial\rho'}{\partial t} = -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub 
- \nabla\cdot(\rho_0\Ubt)}_{"\rho' ~ \text{force}"}. 
\label{rhoprime equation}
\end{equation}
\subsubsection{Predicting $\rho'$ at edges}
We define $\rho' = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to edges using 
{\tt make\_edge\_scal} in {\tt scalar\_advance} and the underbraced term in (\ref{rhoprime equation}) 
as the forcing.
\subsubsection{Predicting $\rho_0$ at edges}\label{Predicting rho0 at edges}
There are two ways to predict $\rho_0$ at edges.
\begin{enumerate}
\item We call {\tt make\_edge\_state\_1d} using the underbraced term in 
(\ref{rho0 equation}) as the forcing.  This gives us $\rho_0^{n+\half,{\rm pred}}$.
This term is used to advect $\rho_0$ in {\tt advect\_base} and therefore must also be used 
to compute $\eta_\rho$, which will be used in {\tt correct\_base}.
\item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n + \rho_0^{(2)})/2$.  We 
compute $\rho_0^{(2)}$ from $\rho_0^n$ using {\tt advect\_base}, which advances equation 
(\ref{rho0 equation}) through $\Delta t$ in time.  The $(2)$ in the superscript indicates 
that we have not called {\tt correct\_base} yet, which computes  $\rho_0^{n+1}$ from $\rho_0^{(2)}$. 
We use $\rho_0^{(2)}$ rather than $\rho_0^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be 
consistent with the fact that we do not include an $\eta_\rho$ contribution to compute 
$\rho'$ at edges.  $\rho_0^{n+\half,{\rm avg}}$ is used to construct $\rho$ at edges from 
$\rho'$ at edges, and this $\rho$ at edges is used to compute fluxes for $\rho X_k$.  
\end{enumerate}
\subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
For the non-radial edges, we directly add $\rho_0^{n+\half,{\rm avg}}$ to $\rho'$ since 
$\rho_0^{n+\half,{\rm avg}}$ is a cell-centered quantity.  For the radial edges, we 
interpolate to obtain $\rho_0^{n+\half,{\rm avg}}$ at radial edges before adding it to $\rho'$.
\subsubsection{Advancing $\rho X_k$}\label{Advancing rhoX_k}
The evolution equation for $\rho X_k$, ignoring the reaction terms that were already 
accounted for in {\tt react\_state}, and the associated discretization is
\begin{equation}
\frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
-\nabla\cdot\left\{\left[\left(\rho_0^{n+\half,{\rm avg}} 
+ \rho'\right)X_k\right](\Ubt+w_0\eb_r)\right\}.
\end{equation}
\section{Energy Evolution}
The full enthalpy equation is
\begin{eqnarray}
\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt} 
+ \nabla\cdot \kth \nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext} \nonumber \\
&=& \underbrace{-\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub}_{-\nabla\cdot(\rho h\Ub)} 
+ \underbrace{\psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r}}_{\frac{Dp_0}{Dt}} 
+ \nabla\cdot\kth\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}.
\end{eqnarray}
The full temperature equation is
\begin{equation}
\frac{\partial T}{\partial t} = -\Ub\cdot\nabla T
+ \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi
+ (\ut \cdot \er )\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 
+ \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
\end{equation}
\subsection{Predicting $(\rho h)'$ ({\tt enthalpy\_pred\_type = 1} - default)}
Due to operator splitting, the call to {\tt react\_state} has already been made.  Hence, the goal 
is to compute an edge state enthalpy starting from $(\rho h)^{(1)}$ using an enthalpy
equation that does not include the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms, where were 
already accounted for in {\tt react\_state}
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub 
+ \psi + (\ut \dot \er) \frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T. \label{rhoh equation}
\end{equation}
We arbitrarily define a base state enthalpy evolution equation
\MarginPar{Does $\psi$ need to be included in the force?  I have written it in the base state equation for now, even though neither the base state or perturbational edge state prediction in the code uses $\psi$.  How about evolving density first,  then computing a time-centered $\psi$ to be used as the energy forcing?}
\begin{eqnarray}
\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] 
+ \frac{D_0p_0}{Dt} \nonumber \\
&=& -w_0\frac{\partial(\rho h)_0}{\partial r} 
- \underbrace{(\rho h)_0\frac{\partial w_0}{\partial r}}_{"(\rho h)_0 ~ \text{force}"}
+ \psi.\label{rhoh0 equation}
\end{eqnarray}
Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and rearranging terms gives 
the perturbational enthalpy equation
\begin{eqnarray}
\frac{\partial(\rho h)'}{\partial t} &=& -\nabla\cdot[(\rho h)'\Ub] 
- \nabla\cdot[(\rho h)_0\Ubt] + (\ut \cdot \er)\frac{\partial p_0}{\partial r} 
+ \nabla\cdot\kth\nabla T\nonumber \\
&=& -\Ub\cdot\nabla(\rho h)' \underbrace{- (\rho h)'\nabla\cdot\Ub 
- \nabla\cdot[(\rho h)_0\Ubt] + (\ut \cdot \er)\frac{\partial p_0}{\partial r}
+ \nabla\cdot\kth\nabla T}_{"(\rho h)' ~ \text{force}"}. \label{rhohprime equation}
\end{eqnarray}
\subsubsection{Advancing $(\rho h)_0$}
{\bf Using full\_rhoh0\_evolution = F:}  Define 
$(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  
and $(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then, $(\rho h)_0^{(1)} \rightarrow$ 
{\tt advect\_base} $\rightarrow (\rho h)_0^{(2)}$.\\ \\
{\bf Using full\_rhoh0\_evolution = T:}
 Instead of defining $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries 
through the complete evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ {\tt react\_base}
$\rightarrow (\rho h)_0^{(1)} \rightarrow$ {\tt advect\_base}
$\rightarrow (\rho h)_0^{(2)} \rightarrow$ {\tt react\_base} $\rightarrow (\rho h)_0^{n+1}$.  
There is no {\tt correct\_base} function for $(\rho h)_0$ written yet, nor is any 
adjustment made to sync up $(\rho h)_0$ after the thermal diffusion step.
\subsubsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
We predict $(\rho h)'$ to 
edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} and the underbraced term in 
(\ref{rhohprime equation}) as the forcing, except that we omit the $\psi$ in the forcing for plane-parallel, and we omit the $\partial w_0/\partial r$ part of $\psi$ in spherical.  Note that we still use full $\psi$ to update the cell.
\subsubsection{Predicting $(\rho h)_0$ at edges}
We use an analogous procedure described in Section \ref{Predicting rho0 at edges} to 
obtain $(\rho h)_0^{n+\half,\rm{pred}}$ and $(\rho h)_0^{n+\half,\rm{avg}}$.  Note that 
$(\rho h)_0^{n+\half,{\rm avg}} = [(\rho h)_0^{(1)} + (\rho h)_0^{(2)}]/2$.
\subsubsection{Computing $\rho h$ at edges}
We use an analogous procedure described in Section \ref{Computing rho at edges} to compute
$\rho h$ at edges.
\subsubsection{Advancing $\rho h$}
{\bf What the did before 8/6/08:}
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot\left[(\rho h)'\left(\Ubt 
+ w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] 
\underbrace{- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\eb_r\right] + \psi}_{\text{this term used to come in through} ~ (\rho h)_0^{n+1} - (\rho h)_0^n} + (\ut \cdot \er)\frac{\partial p_0}{\partial r} \enskip .
\end{equation}
{\bf What the code does now:} We would like to update the 
non-forcing part of the enthalpy equation analogously to Section \ref{Advancing rhoX_k}
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) = 
-\nabla\cdot\left\{\left[(\rho h)_0^{n+\half,{\rm avg}} 
+ (\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\} + (\ut \cdot \er)\frac{\partial p_0}{\partial r} + \psi  \enskip .
\end{equation}

\subsection{Predicting $h$ ({\tt enthalpy\_pred\_type = 2})}

\subsection{Predicting $T$ then $(\rho h)'$ ({\tt enthalpy\_pred\_type = 3})}

\subsection{Predicting $T$ then $h$ ({\tt enthalpy\_pred\_type = 4})}

\subsection{Summary of Forcing}

\begin{table*}[h]
\begin{center}
\caption{Forcing term into {\tt make\_edge\_scal} \newline}
\begin{tabular}{c|c}
\hline
\hline
{H PRED TYPE} &   {Force} \\
\hline \\[-3mm]
1  $((\rho h)^\prime)$ &  $(\ut \cdot \er) \frac{\partial p_0}{\partial r} - 
 (\rho h)^\prime \; \nabla \cdot (\ut+w_0 \er) - 
 \nabla \cdot (\ut (\rho h)_0) + \nabla \cdot \kth \nabla T$ \\[2 mm]
2  $(h)$ & $\frac{1}{\rho} \left [\psi + (\ut \cdot \er)
  \frac{\partial p_0}{\partial r} + \nabla \cdot \kth \nabla T \right ]$ \\[2 mm]
3  $(T)$ & $\frac{1}{\rho c_p} \left \{ (1 - \rho h_p) 
   \left [\psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r} \right ] \nabla \cdot \kth \nabla T \right \}$ \\[2 mm]
4  $(T)$ & $\frac{1}{\rho c_p} \left\{ (1 - \rho h_p) \left [\psi + (\ut \cdot \er)
\frac{\partial p_0}{\partial r}\right ] \nabla \cdot \kth \nabla T \right\}$ \\[2 mm]
\hline
\end{tabular}
\end{center}
\end{table*}

\begin{table*}[h]
\begin{center}
\caption{Quantity that goes into and out of {\tt make\_edge\_scal} \newline}
\begin{tabular}{c|c}
\hline
\hline
{h pred type} & {h/T } \\
\hline \\[-3mm]
1 & $(\rho h)^\prime$ \\
2 & $h$  \\
3 & $T$  \\
4 & $T$  \\
\hline
\end{tabular}
\end{center}
\end{table*}

\begin{table*}
\begin{center}
\caption{When predicting temp ... (conversion done by {\tt makeRhoHfromT}) \newline}
\begin{tabular}{c|c|c|c}
\hline
\hline
{type} & {Inputs } & {EOS inputs} & {output} \\
\hline
3 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ &  $(\rho h)^\prime = \rho h - (\rho h)_0$ \\
   &                   & $   X = $ given & \\
   &                   &                 & \\
 4 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ & $h$ \\
   &                   & $   X = $ given & \\
   &                   &                 & \\
\hline
\end{tabular}
\end{center}
\end{table*}

\begin{table*}
\begin{center}
\caption{Quantity that goes into {\tt mkflux} on edges \newline}
\begin{tabular}{c|c|c}
\hline
\hline
{h pred type} & {Species } & {h/T } \\
\hline
1 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
2 & $X$ and $\rho^\prime$ & $h$ \\
3 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
4 & $X$ and $\rho^\prime$ & $h$ \\
\hline
\end{tabular}
\end{center}
\end{table*}

\begin{table*}
\begin{center}
\caption{Quantity that is created in {\tt mkflux} and passed into {\tt update\_scal}\newline\
}
\begin{tabular}{c|c|c|c|c}
       &            &              &\multicolumn{2}{c}{Forcing Term} \\
{type} & {Species } & {$(\rho h)$} & {$(\rho X)$} & {$(\rho h)$} \\
\hline 
1 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
2 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
3 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r}$ \\[2mm]
4 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
\hline
\end{tabular}
\end{center}
\end{table*}

### 2008-10-24 docs/flowchart/flowchart.tex
switch over to book format for the master document -- this makes the sectioning more natural
add MarginPars

@@
-\section{Current Algorithm}
-
 Here we outline the algorithm currently implemented in the code.  Our
@@
 
-\subsection{Notation}
+\section{Notation}
 
@@
 
-\subsection{Time Advancement Algorithm}
+\section{Time Advancement Algorithm}
 
@@
 interval of $\dt / 2.$}
+
+\MarginPar{don't we set $h_0 = \overline{h}$ somewhere around here?}
 
@@
 \item Update the enthalpy.  The full enthalpy equation is 
+\MarginPar{we should say something about {\tt use\_tfromp}, and {\tt do\_eos\_h\_above\_cutoff}}
   \begin{equation}
@@
 \item Diffuse the enthalpy through a time interval of $\dt$.  Here we begin
-with the enthalpy equation (eq.~[eq:rhohup]), but consider only the 
+with the enthalpy equation (eq.~[\ref{eq:rhohup}]), but consider only the 
 diffusion terms,
@@
 
-\subsection{Initialization}\label{Sec:Initialization}
+\section{Initialization}\label{Sec:Initialization}
 

### 2008-10-24 docs/pert_notes/pert.tex
add MarginPars

@@
+%-----------------------------------------------------------------------------
+% density
+%-----------------------------------------------------------------------------
 \section{Density Evolution}
@@
 \end{equation}
+
 \subsection{Predicting $\rho'$ ({\tt predict\_rho = F} - default)}
-The base state density evolution equation is
+The base state density evolution equation is \MarginPar{we don't have {\tt predict\_rho} anymore--is this still right?}
 \begin{equation}
@@
 \end{equation}
-Subtract (\ref{rho0 equation}) from (\ref{rho equation}) and rearrange terms, noting that 
-$\Ub = \Ubt + w_o\eb_r$, to obtain the perturbational density equation,
+Subtract (\ref{rho0 equation}) from (\ref{rho equation}) and rearrange
+terms, noting that $\Ub = \Ubt + w_o\eb_r$, to obtain the
+perturbational density equation,
 \begin{equation}
@@
 \end{equation}
+
 \subsubsection{Predicting $\rho'$ at edges}
-We define $\rho' = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to edges using 
-{\tt make\_edge\_scal} in {\tt scalar\_advance} and the underbraced term in (\ref{rhoprime equation}) 
-as the forcing.
+We define $\rho' = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to
+edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} and the
+underbraced term in (\ref{rhoprime equation}) as the forcing.
+
 \subsubsection{Predicting $\rho_0$ at edges}\label{Predicting rho0 at edges}
@@
 \begin{enumerate}
-\item We call {\tt make\_edge\_state\_1d} using the underbraced term in 
-(\ref{rho0 equation}) as the forcing.  This gives us $\rho_0^{n+\half,{\rm pred}}$.
-This term is used to advect $\rho_0$ in {\tt advect\_base} and therefore must also be used 
-to compute $\eta_\rho$, which will be used in {\tt correct\_base}.
-\item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n + \rho_0^{(2)})/2$.  We 
-compute $\rho_0^{(2)}$ from $\rho_0^n$ using {\tt advect\_base}, which advances equation 
-(\ref{rho0 equation}) through $\Delta t$ in time.  The $(2)$ in the superscript indicates 
-that we have not called {\tt correct\_base} yet, which computes  $\rho_0^{n+1}$ from $\rho_0^{(2)}$. 
-We use $\rho_0^{(2)}$ rather than $\rho_0^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be 
-consistent with the fact that we do not include an $\eta_\rho$ contribution to compute 
-$\rho'$ at edges.  $\rho_0^{n+\half,{\rm avg}}$ is used to construct $\rho$ at edges from 
-$\rho'$ at edges, and this $\rho$ at edges is used to compute fluxes for $\rho X_k$.  
+\item We call {\tt make\_edge\_state\_1d} using the underbraced term
+in (\ref{rho0 equation}) as the forcing.  This gives us
+$\rho_0^{n+\half,{\rm pred}}$.  This term is used to advect $\rho_0$
+in {\tt advect\_base} and therefore must also be used to compute
+$\eta_\rho$, which will be used in {\tt correct\_base}.
+\item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n +
+\rho_0^{(2)})/2$.  We compute $\rho_0^{(2)}$ from $\rho_0^n$ using
+{\tt advect\_base}, which advances equation (\ref{rho0 equation})
+through $\Delta t$ in time.  The $(2)$ in the superscript indicates
+that we have not called {\tt correct\_base} yet, which computes
+$\rho_0^{n+1}$ from $\rho_0^{(2)}$.  We use $\rho_0^{(2)}$ rather than
+$\rho_0^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be
+consistent with the fact that we do not include an $\eta_\rho$
+contribution to compute $\rho'$ at edges.  $\rho_0^{n+\half,{\rm
+avg}}$ is used to construct $\rho$ at edges from $\rho'$ at edges, and
+this $\rho$ at edges is used to compute fluxes for $\rho X_k$.
 \end{enumerate}
+
 \subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
-For the non-radial edges, we directly add $\rho_0^{n+\half,{\rm avg}}$ to $\rho'$ since 
-$\rho_0^{n+\half,{\rm avg}}$ is a cell-centered quantity.  For the radial edges, we 
-interpolate to obtain $\rho_0^{n+\half,{\rm avg}}$ at radial edges before adding it to $\rho'$.
+For the non-radial edges, we directly add $\rho_0^{n+\half,{\rm avg}}$
+to $\rho'$ since $\rho_0^{n+\half,{\rm avg}}$ is a cell-centered
+quantity.  For the radial edges, we interpolate to obtain
+$\rho_0^{n+\half,{\rm avg}}$ at radial edges before adding it to
+$\rho'$.
+
 \subsubsection{Advancing $\rho X_k$}\label{Advancing rhoX_k}
-The evolution equation for $\rho X_k$, ignoring the reaction terms that were already 
-accounted for in {\tt react\_state}, and the associated discretization is
+The evolution equation for $\rho X_k$, ignoring the reaction terms
+that were already accounted for in {\tt react\_state}, and the
+associated discretization is
 \begin{equation}
@@
 \end{equation}
+
+%-----------------------------------------------------------------------------
+% energy
+%-----------------------------------------------------------------------------
 \section{Energy Evolution}
@@
 \end{equation}
+
 \subsection{Predicting $(\rho h)'$ ({\tt enthalpy\_pred\_type = 1} - default)}
-Due to operator splitting, the call to {\tt react\_state} has already been made.  Hence, the goal 
-is to compute an edge state enthalpy starting from $(\rho h)^{(1)}$ using an enthalpy
-equation that does not include the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms, where were 
-already accounted for in {\tt react\_state}
+Due to operator splitting, the call to {\tt react\_state} has already
+been made.  Hence, the goal is to compute an edge state enthalpy
+starting from $(\rho h)^{(1)}$ using an enthalpy equation that does
+not include the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms, where
+were already accounted for in {\tt react\_state}
 \begin{equation}
@@
 We arbitrarily define a base state enthalpy evolution equation
-\MarginPar{Does $\psi$ need to be included in the force?
-I have written it in the base state equation for now, even though neither the base state or perturbational edge state prediction in the code uses $\psi$.
-How about evolving density first,  then computing a time-centered $\psi$ to be used as the energy forcing?}
+\MarginPar{Does $\psi$ need to be included in the force?  I have
+written it in the base state equation for now, even though neither the
+base state or perturbational edge state prediction in the code uses
+$\psi$.  How about evolving density first, then computing a
+time-centered $\psi$ to be used as the energy forcing?}
 \begin{eqnarray}
@@
 \end{eqnarray}
+
 \subsubsection{Advancing $(\rho h)_0$}
@@
 adjustment made to sync up $(\rho h)_0$ after the thermal diffusion step.
+
 \subsubsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
-We predict $(\rho h)'$ to 
-edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} and the underbraced term in 
-(\ref{rhohprime equation}) as the forcing, except that we omit the $\psi$ in the forcing for plane-parallel, and we omit the $\partial w_0/\partial r$ part of $\psi$ in spherical.
-Note that we still use full $\psi$ to update the cell.
+We predict $(\rho h)'$ to edges using {\tt make\_edge\_scal} in {\tt
+scalar\_advance} and the underbraced term in (\ref{rhohprime
+equation}) as the forcing, except that we omit the $\psi$ in the
+forcing for plane-parallel, and we omit the $\partial w_0/\partial r$
+part of $\psi$ in spherical.  Note that we still use full $\psi$ to
+update the cell.
+
 \subsubsection{Predicting $(\rho h)_0$ at edges}
-We use an analogous procedure described in Section \ref{Predicting rho0 at edges} to 
-obtain $(\rho h)_0^{n+\half,\rm{pred}}$ and $(\rho h)_0^{n+\half,\rm{avg}}$.  Note that 
-$(\rho h)_0^{n+\half,{\rm avg}} = [(\rho h)_0^{(1)} + (\rho h)_0^{(2)}]/2$.
+We use an analogous procedure described in Section \ref{Predicting
+rho0 at edges} to obtain $(\rho h)_0^{n+\half,\rm{pred}}$ and $(\rho
+h)_0^{n+\half,\rm{avg}}$.  Note that $(\rho h)_0^{n+\half,{\rm avg}} =
+[(\rho h)_0^{(1)} + (\rho h)_0^{(2)}]/2$.
+
 \subsubsection{Computing $\rho h$ at edges}
-We use an analogous procedure described in Section \ref{Computing rho at edges} to compute
-$\rho h$ at edges.
+We use an analogous procedure described in Section \ref{Computing rho
+at edges} to compute $\rho h$ at edges.
+
 \subsubsection{Advancing $\rho h$}
@@
 \end{equation}
-{\bf What the code does now:} We would like to update the 
-non-forcing part of the enthalpy equation analogously to Section \ref{Advancing rhoX_k}
+{\bf What the code does now:} We would like to update the non-forcing
+part of the enthalpy equation analogously to Section \ref{Advancing
+rhoX_k}
 \begin{equation}

### 2008-10-24 docs/pert_notes/pert.tex
addressed marginpar about predict_rho not being an option anymore

@@
 \end{equation}
-
-\subsection{Predicting $\rho'$ ({\tt predict\_rho = F} - default)}
-The base state density evolution equation is \MarginPar{we don't have {\tt predict\_rho} anymore--is this still right?}
+We predict both $\rho'$ and $\rho_0$ to edges and later reconstruct $\rho$.
+The base state density evolution equation is
 \begin{equation}
@@
 
-\subsubsection{Predicting $\rho'$ at edges}
+\subsection{Predicting $\rho'$ at edges}
 We define $\rho' = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to
@@
 
-\subsubsection{Predicting $\rho_0$ at edges}\label{Predicting rho0 at edges}
+\subsection{Predicting $\rho_0$ at edges}\label{Predicting rho0 at edges}
 There are two ways to predict $\rho_0$ at edges.
@@
 
-\subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
+\subsection{Computing $\rho$ at edges}\label{Computing rho at edges}
 For the non-radial edges, we directly add $\rho_0^{n+\half,{\rm avg}}$
@@
 
-\subsubsection{Advancing $\rho X_k$}\label{Advancing rhoX_k}
+\subsection{Advancing $\rho X_k$}\label{Advancing rhoX_k}
 The evolution equation for $\rho X_k$, ignoring the reaction terms

### 2008-10-24 docs/pert_notes/pert.tex
continue consolidation of macros - there are still some inconsistencies between documents

@@
 + w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] 
-\underbrace{- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\eb_r\right] + \psi}_{\text{this term used to come in through} ~ (\rho h)_0^{n+1} - (\rho h)_0^n} + (\ut \cdot \er)\frac{\partial p_0}{\partial r} \enskip .
+\underbrace{- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\eb_r\right] + \psi}_{\text{this term used to come in through} ~ (\rho h)_0^{n+1} - (\rho h)_0^n} + (\Ubt \cdot \er)\frac{\partial p_0}{\partial r} \enskip .
 \end{equation}
@@
 -\nabla\cdot\left\{\left[(\rho h)_0^{n+\half,{\rm avg}} 
-+ (\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\} + (\ut \cdot \er)\frac{\partial p_0}{\partial r} + \psi  \enskip .
++ (\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\} + (\Ubt \cdot \er)\frac{\partial p_0}{\partial r} + \psi  \enskip .
 \end{equation}
@@
 \hline \\[-3mm]
-1  $((\rho h)^\prime)$ &  $(\ut \cdot \er) \frac{\partial p_0}{\partial r} - 
- (\rho h)^\prime \; \nabla \cdot (\ut+w_0 \er) - 
- \nabla \cdot (\ut (\rho h)_0) + \nabla \cdot \kth \nabla T$ \\[2 mm]
-2  $(h)$ & $\frac{1}{\rho} \left [\psi + (\ut \cdot \er)
+1  $((\rho h)^\prime)$ &  $(\Ubt \cdot \er) \frac{\partial p_0}{\partial r} - 
+ (\rho h)^\prime \; \nabla \cdot (\Ubt+w_0 \er) - 
+ \nabla \cdot (\Ubt (\rho h)_0) + \nabla \cdot \kth \nabla T$ \\[2 mm]
+2  $(h)$ & $\frac{1}{\rho} \left [\psi + (\Ubt \cdot \er)
   \frac{\partial p_0}{\partial r} + \nabla \cdot \kth \nabla T \right ]$ \\[2 mm]
 3  $(T)$ & $\frac{1}{\rho c_p} \left \{ (1 - \rho h_p) 
-   \left [\psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r} \right ] \nabla \cdot \kth \nabla T \right \}$ \\[2 mm]
-4  $(T)$ & $\frac{1}{\rho c_p} \left\{ (1 - \rho h_p) \left [\psi + (\ut \cdot \er)
+   \left [\psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} \right ] \nabla \cdot \kth \nabla T \right \}$ \\[2 mm]
+4  $(T)$ & $\frac{1}{\rho c_p} \left\{ (1 - \rho h_p) \left [\psi + (\Ubt \cdot \er)
 \frac{\partial p_0}{\partial r}\right ] \nabla \cdot \kth \nabla T \right\}$ \\[2 mm]
@@
 \hline 
-1 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
-2 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
-3 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r}$ \\[2mm]
-4 & $(\ut+w_0)(\rho_0+\rho^\prime) X          $ & $(\ut+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
+1 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
+2 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
+3 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r}$ \\[2mm]
+4 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
 \hline

### 2008-10-24 docs/flowchart/flowchart.tex
more macro consolidation

@@
 \hline
-$\ubold$     & total velocity ($\ubold = \Ut + w_0 \er$) & cm~s$^{-1}$ \\
-\hline
-$\Ut$   & local velocity                             & cm~s$^{-1}$ \\
+$\Ub$     & total velocity ($\Ub = \Ubt + w_0 \er$) & cm~s$^{-1}$ \\
+\hline
+$\Ubt$   & local velocity                             & cm~s$^{-1}$ \\
 \hline
@@
 \hline
-$\etarho$ & $\etarho \equiv \overline{ (\rho' \Ut \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\
+$\etarho$ & $\etarho \equiv \overline{ (\rho' \Ubt \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\
 \hline
@@
 \begin{equation}
-\int_{\Omega_H} \Ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
+\int_{\Omega_H} \Ubt \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
 \end{equation}
@@
 \begin{equation}
-\nablab \cdotb (\beta_0 \Ut )  = \beta_0 \left(S - \Sbar \right )\enskip ,
+\nablab \cdotb (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
 \label{eq:tildeconstraint}
@@
   \begin{equation}
-  \frac{\partial (\rho X_k)}{\partial t} + \nablab \cdotb (\ubold \rho X_k) =
+  \frac{\partial (\rho X_k)}{\partial t} + \nablab \cdotb (\Ub \rho X_k) =
          \rho {\omegadot}_k \enskip , \label{eq:species}
@@
   Taylor expansion in space and time, as described in paper II,
-  Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  We use the base
+  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  We use the base
   state density, $\rho_0$, to convert this to an edge state for $\rho$,
@@
   \begin{equation}
-  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\ubold \rho h)
+  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
 + \frac{Dp_0}{Dt} + \nabla\cdot\kth\nabla T + \rho \Hnuc + \rho \Hext 
@@
   \begin{equation}
-  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\ubold \rho h)
+  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
 + \frac{Dp_0}{Dt} 
@@
     using a second-order Taylor expansion in space
-    and time, as described in paper II, Appendix A, using $\V =
+    and time, as described in paper II, Appendix A, using $\Vb =
     \uadvone+w_0^{\nph,\star} \er$.  We do not include the reaction
@@
     $X_k^{(1),\nph}$ using a second-order Taylor expansion in
-    space and time, as described in paper II, Appendix A, using $\V =
+    space and time, as described in paper II, Appendix A, using $\Vb =
     \uadvtwo+w_0^{\nph} \er$.  We use the base
@@
     using a second-order Taylor expansion in space and
-    time, as described in paper II, Appendix A, using $\V =
+    time, as described in paper II, Appendix A, using $\Vb =
     \uadvtwo+w_0^{\nph} \er$.  Again, we do not include the reaction
@@
 The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
-We update the velocity field $\Ut^n$ to $\Ut^{n+1,\star}$ by discretizing
-\begin{equation}
-\frac{\partial\Ut}{\partial t} = - \Ut \cdotb \nablab \Ut - w_0 \frac{\partial \Ut}{\partial r}
-                                 - \left(\Ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
+We update the velocity field $\Ubt^n$ to $\Ubt^{n+1,\star}$ by discretizing
+\begin{equation}
+\frac{\partial\Ubt}{\partial t} = - \Ubt \cdotb \nablab \Ubt - w_0 \frac{\partial \Ubt}{\partial r}
+                                 - \left(\Ubt \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
                                  - \frac{1}{\rho} \nablab\pi
@@
 \begin{eqnarray}
-\Ut^{n+1,\star} &=& \Ut^n - \dt \;
- \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \Ut \right]^\nph
+\Ubt^{n+1,\star} &=& \Ubt^n - \dt \;
+ \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \Ubt \right]^\nph
       - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
@@
 and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
-The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \Ut ]^\nph$
-is described in paper II, Appendix A, with $\V = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
-component of $\Ut^n$ individually.
+The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \Ubt ]^\nph$
+is described in paper II, Appendix A, with $\Vb = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
+component of $\Ubt^n$ individually.
 
@@
 \begin{equation}
-\nablab \cdotb \left(\beta_0^{\nph} \Ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)
+\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)
 + \frac{f}{\gammabar^{n+1} p_0^{n+1}}
@@
 \end{equation}
-by defining ${\bf V} = \Ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
+by defining ${\bf V} = \Ubt^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
 and solving
@@
 \begin{equation}
-\Ut^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
+\Ubt^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
 \end{equation}
@@
 using $w_0$ as computed in step 6 and
-$\Ut^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.
+$\Ubt^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.
 
@@
 Start with initial data $X_k^{\initp}, \rho^{\initp},$ $T^{\initp},$ an 
-initial base state, and an initial guess for the velocity, $\ubold^{\initp}.$
+initial base state, and an initial guess for the velocity, $\Ub^{\initp}.$
 Use the equation of state to determine $(\rho h)^{\initp}$.  Set
 $w_0^1 = 0$ as an initial approximation.  Compute $\beta_0^{\initp}$ as a function of 
-the initial data.  Then, project $\ubold^{\initp}$ using $\beta_0^{\initp}$ and 
-$S = \rho\Hext$, giving $\ubold^{0,1}$.  The next part of the initialization process 
+the initial data.  Then, project $\Ub^{\initp}$ using $\beta_0^{\initp}$ and 
+$S = \rho\Hext$, giving $\Ub^{0,1}$.  The next part of the initialization process 
 proceeds as follows.
@@
 
-  \item Estimate $\Delta t^\nu$ using $\ubold^{0,\nu}$ and $w_0^\nu.$
+  \item Estimate $\Delta t^\nu$ using $\Ub^{0,\nu}$ and $w_0^\nu.$
 
@@
 
-  \item Project $\ubold^{0,\nu}$ using $\beta_0^{\initp}$ and 
+  \item Project $\Ub^{0,\nu}$ using $\beta_0^{\initp}$ and 
         $(S^{0,\nu} - \overline{S}^{0,\nu})$ as the source term.  
-        This yields $\ubold^{0,\nu+1}.$
+        This yields $\Ub^{0,\nu+1}.$
 
@@
   Define $S^0 = S^{0,N_{\rm iters}^S}$, $w_0^{-\myhalf} = w_0^{N_{\rm iters}^S+1}$, 
-$\dt^0 = \Delta t^{N_{\rm iters}^S},$ and $\ubold^0 = \ubold^{0,N_{\rm iters}^S+1}.$
+$\dt^0 = \Delta t^{N_{\rm iters}^S},$ and $\Ub^0 = \Ub^{0,N_{\rm iters}^S+1}.$
 
@@
     The only other difference in the time advancement is that in step 11
-    we define ${\bf V} = (\Ut^{1,\star} - \Ut^0)$ and solve
+    we define ${\bf V} = (\Ubt^{1,\star} - \Ubt^0)$ and solve
     \begin{equation}  L_\beta^\rho \phi =

### 2008-10-24 docs/pert_notes/pert.tex
another macro fix

@@
 &=& \underbrace{-\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub}_{-\nabla\cdot(\rho h\Ub)} 
-+ \underbrace{\psi + (\ut \cdot \er) \frac{\partial p_0}{\partial r}}_{\frac{Dp_0}{Dt}} 
++ \underbrace{\psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r}}_{\frac{Dp_0}{Dt}} 
 + \nabla\cdot\kth\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}.
@@
 + \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi
-+ (\ut \cdot \er )\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 
++ (\Ubt \cdot \er )\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 
 + \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
@@
 \frac{\partial(\rho h)}{\partial t} = -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub 
-+ \psi + (\ut \dot \er) \frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T. \label{rhoh equation}
++ \psi + (\Ubt \dot \er) \frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T. \label{rhoh equation}
 \end{equation}
@@
 \frac{\partial(\rho h)'}{\partial t} &=& -\nabla\cdot[(\rho h)'\Ub] 
-- \nabla\cdot[(\rho h)_0\Ubt] + (\ut \cdot \er)\frac{\partial p_0}{\partial r} 
+- \nabla\cdot[(\rho h)_0\Ubt] + (\Ubt \cdot \er)\frac{\partial p_0}{\partial r} 
 + \nabla\cdot\kth\nabla T\nonumber \\
 &=& -\Ub\cdot\nabla(\rho h)' \underbrace{- (\rho h)'\nabla\cdot\Ub 
-- \nabla\cdot[(\rho h)_0\Ubt] + (\ut \cdot \er)\frac{\partial p_0}{\partial r}
+- \nabla\cdot[(\rho h)_0\Ubt] + (\Ubt \cdot \er)\frac{\partial p_0}{\partial r}
 + \nabla\cdot\kth\nabla T}_{"(\rho h)' ~ \text{force}"}. \label{rhohprime equation}

### 2008-10-24 docs/pert_notes/pert.tex
some pert_notes cleanup

@@
 \end{equation}
-We arbitrarily define a base state enthalpy evolution equation
+We define the base state enthalpy evolution equation as
 \MarginPar{Does $\psi$ need to be included in the force?  I have
@@
 \subsubsection{Advancing $(\rho h)_0$}
-{\bf Using full\_rhoh0\_evolution = F:}  Define 
-$(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  
+\begin{itemize}
+\item {\bf Using full\_rhoh0\_evolution = F:}\\
+ Define $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  
 and $(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then, $(\rho h)_0^{(1)} \rightarrow$ 
-{\tt advect\_base} $\rightarrow (\rho h)_0^{(2)}$.\\ \\
-{\bf Using full\_rhoh0\_evolution = T:}
+{\tt advect\_base} $\rightarrow (\rho h)_0^{(2)}$.
+\item {\bf Using full\_rhoh0\_evolution = T:}\\
  Instead of defining $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries 
-through the complete evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ {\tt react\_base}
-$\rightarrow (\rho h)_0^{(1)} \rightarrow$ {\tt advect\_base}
+through the complete evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ 
+{\tt react\_base} $\rightarrow (\rho h)_0^{(1)} \rightarrow$ {\tt advect\_base}
 $\rightarrow (\rho h)_0^{(2)} \rightarrow$ {\tt react\_base} $\rightarrow (\rho h)_0^{n+1}$.  
@@
 adjustment made to sync up $(\rho h)_0$ after the thermal diffusion step.
+\end{itemize}
 
@@
 \subsubsection{Advancing $\rho h$}
-{\bf What the did before 8/6/08:}
-\begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot\left[(\rho h)'\left(\Ubt 
-+ w_0\eb_r\right)\right] - \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm avg}}\Ubt\right] 
-\underbrace{- \nabla\cdot\left[(\rho h)_0^{n+\half,{\rm pred}}w_0\eb_r\right] + \psi}_{\text{this term used to come in through} ~ (\rho h)_0^{n+1} - (\rho h)_0^n} + (\Ubt \cdot \er)\frac{\partial p_0}{\partial r} \enskip .
-\end{equation}
-{\bf What the code does now:} We would like to update the non-forcing
-part of the enthalpy equation analogously to Section \ref{Advancing
-rhoX_k}
+We update the enthalpy analogously to the species update in 
+Section \ref{Advancing rhoX_k}.  The only difference is that we also 
+have a forcing term
 \begin{equation}

### 2008-10-24 docs/flowchart/flowchart.tex
add a bit about how the (rho h)_0 edge state is computed for spherical

@@
 \item {\em Low Mach Number Modeling of Type Ia
-  Supernovae. I. Hydrodynamics}, A.~S.~Almgren, J.~B.~Bell, 
+  Supernovae. I. Hydrodynamics,} A.~S.~Almgren, J.~B.~Bell, 
   C.~A.~Rendleman, \& M.~Zingale 2006, ApJ, 637, 922 (henceforth
@@
 \item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
-  Evolution} A.~S.~Almgren, J.~B.~Bell, C.~A.~Rendleman, \& M.~Zingale
+  Evolution,} A.~S.~Almgren, J.~B.~Bell, C.~A.~Rendleman, \& M.~Zingale
   2006, ApJ, 649, 927 (henceforth paper~II)
-\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions}
+\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
 A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale
   2008, ApJ, 684, 449 (henceforth paper~III)
-\item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection}
+\item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection,}
 A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale,
   in perparation (henceforth paper~IV)
+\end{itemize}
+We also carry over some ideas from our small-scale low Mach number algorithm
+for astrophysical flows:
+\begin{itemize}
+\item {\em Adaptive Low Mach Number Simulations of Nuclear Flames,}
+J.~B.~Bell, M.~S.~Day, C.~A.~Rendleman, S.~E.~Woosley, \& M.~Zingale
+2004, JCP, 195, 2, 677 (henceforth BDRWZ)
 \end{itemize}
@@
 
+\MarginPar{what about a $\eta_{\rho h}$}
 \end{enumerate}
@@
 
-In paper III we discretized the enthalpy evolution equation in full, non-perturbation form.  
-We have discovered that discretizing the enthalpy evolution in non-perturbational form 
-leads to better numerical properties.  This is more like paper II.\\
+There are several changes to the algorithm since it was last detailed
+in paper III.  The major changes are:
+\begin{itemize}
+\item In paper III we discretized the enthalpy evolution equation in
+full, non-perturbation form.  We have discovered that discretizing the
+enthalpy evolution in non-perturbational form leads to better
+numerical properties.  This is more like paper II.
+
+\item We split the update of the density and enthalpy in steps 4 and 8.
+Now, density is updated first, allowing us to compute new values of 
+$\psi$ that are then used in the enthalpy update.
+
+\item We discuss spherical geometries.
+
+\item We add thermal diffusion.
+\end{itemize}
+%
+The advance of the state through a single timestep appears as:
 
@@
 Otherwise, following the method used in our small-scale low Mach number algorithm
-(Cite Bell04), we extrapolate to the half-time using the source
+(BDRWZ), we extrapolate to the half-time using the source
 terms at the previous and current time levels
@@
 \end{equation}
+
+   We note that for spherical geometry, the edge state of $(\rho h)_0$
+   is computed by averging $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$
+   to the edge separately, and then computing 
+   ${(\rho h)_0}_{r+1/2} = {\rho_0}_{r+1/2} {h_0}_{r+1/2}$.
 

### 2008-10-24 docs/flowchart/flowchart.tex
update how the enthalpy update looks for spherical
fix an equation reference

@@
   \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
-+ \frac{Dp_0}{Dt} + \nabla\cdot\kth\nabla T + \rho \Hnuc + \rho \Hext 
++ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T + \rho \Hnuc + \rho \Hext 
 \enskip , \label{eq:rhohupd} 
@@
   \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
-+ \frac{Dp_0}{Dt} 
++ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} 
 \enskip , \label{eq:rhohupdadv} 
   \end{equation}
+For spherical flows, experience has shown that solving the algebraically-equivalent form
+\begin{equation}
+  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
++ \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt  
+\enskip , \label{eq:rhohupdadv2} 
+\end{equation}
+results in a better solution.
 
@@
   explicitly including the reaction terms,
+
+\begin{description}
+\item[plane-parallel:]
+
   \begin{eqnarray}
@@
   \MarginPar{we should also show the update in the case that $(\rho h)'$ is the flux}
+
+\item[spherical:]
+
+  \begin{eqnarray}
+  (\rho h)^{(2'),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
+      \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right) (\rho
+      h)^{(1),\nph,\star} \right] \right\} \nonumber \\ 
+    && + \dt \; \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) 
+       - p_0^{n} \nabla \cdot \uadvone \right \} \nonumber \\
+    && + \dt \; \psi^{\nph,*} \enskip ,
+  \end{eqnarray}
+
+\MarginPar{is it really $p_0^n$ here and not $p_0^{n+1/2}$?}
+\end{description}
+
   \end{enumerate}
@@
 \item Diffuse the enthalpy through a time interval of $\dt$.  Here we begin
-with the enthalpy equation (eq.~[\ref{eq:rhohup}]), but consider only the 
+with the enthalpy equation (eq.~[\ref{eq:rhohupd}]), but consider only the 
 diffusion terms,
@@
      explicitly including the reaction terms,
+
+\begin{description}
+\item[plane-parallel:]
+
 \begin{eqnarray}
@@
 \end{eqnarray}
+
+\item[spherical:]
+
+\begin{eqnarray}
+(\rho h)^{(2')} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
+(\rho h)^{(1),\nph} \right] \right\} \nonumber \\
+    && + \dt \; \left \{ \nabla \cdot \left (\uadvtwo p_0^{n+1/2} \right ) 
+       - p_0^{n+1/2} \nabla \cdot \uadvtwo \right \} \nonumber \\
+    && + \dt \; \psi^{\nph} \enskip ,
+\end{eqnarray}
+
+\end{description}
 

### 2008-10-24 docs/flowchart/flowchart.tex
fix a section reference

@@
 
-Compute $\dt$ for the next time step with the procedure described in \S~ref{sec:timesteps},
+Compute $\dt$ for the next time step with the procedure described in \S 3.4 of paper III
 using $w_0$ as computed in step 6 and

### 2008-10-24 docs/pert_notes/pert.tex
addressed the marginpar about the psi forcing in the (rho h)_0 equation.  now the code includes it, and the document reflects this.

@@
 \frac{\partial(\rho h)}{\partial t} = -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub 
-+ \psi + (\Ubt \dot \er) \frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T. \label{rhoh equation}
++ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T. \label{rhoh equation}
 \end{equation}
 We define the base state enthalpy evolution equation as
-\MarginPar{Does $\psi$ need to be included in the force?  I have
-written it in the base state equation for now, even though neither the
-base state or perturbational edge state prediction in the code uses
-$\psi$.  How about evolving density first, then computing a
-time-centered $\psi$ to be used as the energy forcing?}
 \begin{eqnarray}
@@
 &=& -w_0\frac{\partial(\rho h)_0}{\partial r} 
-- \underbrace{(\rho h)_0\frac{\partial w_0}{\partial r}}_{"(\rho h)_0 ~ \text{force}"}
-+ \psi.\label{rhoh0 equation}
+- \underbrace{(\rho h)_0\frac{\partial w_0}{\partial r}+ \psi}_{"(\rho h)_0 ~ \text{force}"}
+\enskip .\label{rhoh0 equation}
 \end{eqnarray}

### 2008-10-24 docs/flowchart/flowchart.tex
put in placeholders for Advect Base Density, Advect Base Pressure, and Advect Base Enthalpy

@@
 
-\item {\bf Correct Base}
-$[\rho_0^{\inp}, (\nabla\cdot\etarho\er)^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
+\item {\bf Correct Base}$[\rho_0^{\inp}, (\nabla\cdot\etarho\er)^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
 is the process by which we adjust the base state density to account for
@@
 
-\item {\bf Advect Base}$[\rhozero^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph,\star}, \psi^{\nmh}] \rightarrow $
-$[\rho_0^{(2),\star}, (\rho h)_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}, p_0^{n+1,\star}].$
+%\item {\bf Advect Base}$[\rhozero^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph,\star}, \psi^{\nmh}] \rightarrow [\rho_0^{(2),\star}, (\rho h)_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}, p_0^{n+1,\star}].$
+
+\item {\bf Advect Base Density}
 
@@
 
+\item (spherical only) {\bf Advect Base Pressure}
+
 \item Define an edge-centered $\etarho^{\nph,\star}$, cell-centered 
@@
 \end{description}
+
+\item (plane-parallel only only) {\bf Advect Base Pressure}
+
+\item {\bf Advect Base Enthalpy}
 
@@
 
-\item {\bf Correct Base}
-$[\rho_0^{(2),\star}, (\nabla\cdot\etarho\er)^{\nph,*}] \rightarrow [\rho_0^{n+1,\star}]$.
+\item {\bf Correct Base}$[\rho_0^{(2),\star}, (\nabla\cdot\etarho\er)^{\nph,*}] \rightarrow [\rho_0^{n+1,\star}]$.
 
@@
 
-\item {\bf Advect Base}$(\rho_0^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph}, \psi^{\nph,\star} ) \rightarrow $
-$[\rho_0^{(2)}, (\rho h)_0^{(2)}, \rho_0^{\nph,{\rm pred}}, p_0^{n+1}].$
+%\item {\bf Advect Base}$(\rho_0^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph}, \psi^{\nph,\star} ) \rightarrow [\rho_0^{(2)}, (\rho h)_0^{(2)}, \rho_0^{\nph,{\rm pred}}, p_0^{n+1}].$
+
+\item {\bf Advect Base Density}
 
@@
 
+\item (spherical only) {\bf Advect Base Pressure}
+
 \item Define an edge-centered $\etarho^{\nph}$, cell-centered 
@@
 \end{description}
+
+\item (plane-parallel only) {\bf Advect Base Pressure}
+
+\item {\bf Advect Base Enthalpy}
 
@@
 
-\item {\bf Correct Base}
-$[\rho_0^{(2)}, (\nabla\cdot\etarho\er)^{\nph}] \rightarrow [\rho_0^{n+1}]$
+\item {\bf Correct Base}$[\rho_0^{(2)}, (\nabla\cdot\etarho\er)^{\nph}] \rightarrow [\rho_0^{n+1}]$
 

### 2008-10-24 docs/flowchart/flowchart.tex
add plane-parallel description of Advect Base Density, Pressure, and Enthalpy

@@
 
-\item {\bf Advect Base}$[\rho_0^{\inp}, (\rho h)_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp}] \rightarrow [\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp}]$ \\
-is the process by which we update the base state through $\dt$ in
-time \MarginPar{we need a separate section for spherical} given the
-radial velocity $w_0^{\inp}.$ The base state arrays are all
-one-dimensional in the radial coordinate; we think of the base state
-quantities as defined at cell centers, with $w_0$ and $\beta_0$
-defined halfway in-between.
-
+\item {\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp]$
+$\rightarrow [\rho_0^\outp,\rho_0^{\outp,n+\myhalf}]$\\
+is the process by which we update the base state density through $\dt$ in time
+as well as compute time-centered interface states for use later in the algorithm.
 \begin{description}
 
-\item[plane-parallel:] ~
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \alph{enumii}}.}
-
-\item For the density update, we discretize equation the base state
-  density evolution equation (paper III; eq.~[29]) without the $\etarho$ term,
+\item[plane-parallel:] We discretize equation the base state density evolution 
+equation (paper III; eq.~[29]) without the $\etarho$ term,
 \begin{eqnarray}
@@
 The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the
-procedure described in paper II, Appendix A, and are saved for use
-later in the overall algorithm.
-
-\item For the pressure update, we discretize the base state pressure
+procedure described in paper II, Appendix A.
+
+\item[spherical:]
+
+\end{description}
+
+\item {\bf Advect Base Pressure}$[p_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [p_0^\outp]$\\
+is the process by which we update the base state pressure through $\dt$ in time.
+\begin{description}
+
+\item[plane-parallel:] We discretize the base state pressure
   evolution equation (paper III; eq.~[31]),
@@
 
-\item For the enthalpy update, we discretize
-\begin{equation}
-\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\er\right] + \psi \enskip ,
+\item[spherical:]
+
+\end{description}
+
+\item {\bf Advect Base Enthalpy}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [(\rho h)_0]^\outp$\\
+is the process by which we update the base state enthalpy through $\dt$ in time.
+\begin{description}
+
+\item[plane-parallel:] We discretize
+\begin{equation}
+\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\er\right] 
++ \psi \enskip ,
 \end{equation}
@@
 
-\end{enumerate}
-
 \item[spherical:]
-
-\MarginPar{we need to write the spherical description}
 

### 2008-10-24 docs/flowchart/flowchart.tex
equation cleanup
filled in the argument lists for Advect Base calls

@@
 \hline
-$\etarho$ & $\etarho \equiv \overline{ (\rho' \Ubt \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\
+$\etarho$ & $\etarho \equiv (\overline{ \rho' \Ubt \cdot \er})$ & g~cm$^{-2}$~s$^{-1}$ \\
 \hline
@@
 \left\{ \psi^{\nmh} - 
-f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right]\right\}
+f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)}^n}{\Delta t^n} \right]\right\}
 \enskip .
@@
 + \frac{f}{\gammabar^n p_0^n}
-\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right] \enskip .
+\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)}^n}{\Delta t^n}\right] \enskip .
 \end{equation}
@@
 
-%\item {\bf Advect Base}$[\rhozero^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph,\star}, \psi^{\nmh}] \rightarrow [\rho_0^{(2),\star}, (\rho h)_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}, p_0^{n+1,\star}].$
-
-\item {\bf Advect Base Density}
+\item {\bf Advect Base Density}$[\rhozero^{n},w_0^{\nph,\star}] \rightarrow$ 
+$[\rho_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}]$.
 
@@
 
-\item (spherical only) {\bf Advect Base Pressure}
+\item (spherical only) {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph,\star}, \psi^{\nmh}]$ 
+$\rightarrow [p_0^{n+1,\star}]$.  Note that $\psi$ is not used in the spherical case.
 
@@
 
-\item (plane-parallel only only) {\bf Advect Base Pressure}
-
-\item {\bf Advect Base Enthalpy}
+\item (plane-parallel only only) {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph,\star},$ 
+$\psi^{\nph,\star}] \rightarrow [p_0^{n+1,\star}]$.
+
+\item {\bf Advect Base Enthalpy}$[(\rho h)_0^{(1)}, w_0^{\nph,\star}, \psi^{\nph,\star}]$ 
+$\rightarrow [(\rho h)_0^{(2),\star}]$.
 
@@
 \overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
-\left\{ \psi^{\nph,\star} - f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\}
+\left\{ \psi^{\nph,\star} - f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]\right\}
 \enskip ,
@@
 \begin{equation}
-\overline{p(\rho,h,X_k)^{\nph,\star}} = 
-\frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}
+\overline{p(\rho,h,X_k)}^{\nph,\star} = 
+\frac{\overline{p(\rho,h,X_k)}^n + \overline{p(\rho,h,X_k)}^{n+1,\star}}{2}
 \end{equation}
@@
 && + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
-\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]
+\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]
 \enskip , \nonumber \\
@@
 
-%\item {\bf Advect Base}$(\rho_0^{n}, (\rho h)_0^{(1)}, p_0^{n}, \beta_0^{(1)}, w_0^{\nph}, \psi^{\nph,\star} ) \rightarrow [\rho_0^{(2)}, (\rho h)_0^{(2)}, \rho_0^{\nph,{\rm pred}}, p_0^{n+1}].$
-
-\item {\bf Advect Base Density}
+\item {\bf Advect Base Density}$[\rho_0^{n},w_0^{\nph}] \rightarrow$ 
+$[\rho_0^{(2)}, \rho_0^{\nph,{\rm pred}}]$.
 
@@
 
-\item (spherical only) {\bf Advect Base Pressure}
+\item (spherical only) {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph}, \psi^{\nph,\star}]$ 
+$\rightarrow [p_0^{n+1}]$.  Note that $\psi$ is not used in the spherical case.
 
@@
 
-\item (plane-parallel only) {\bf Advect Base Pressure}
-
-\item {\bf Advect Base Enthalpy}
+\item  (plane-parallel only) {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph}, \psi^{\nph}]$ 
+$\rightarrow [p_0^{n+1}]$.
+
+\item {\bf Advect Base Enthalpy}$[(\rho h)_0^{(1)}, w_0^{\nph}, \psi^{\nph}]$ 
+$\rightarrow [(\rho h)_0^{(2)}]$.
 
@@
 + \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]
+\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]
 \enskip,
@@
 - \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right] \enskip,
+\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right] \enskip,
 \end{equation}

### 2008-10-26 docs/flowchart/flowchart.tex
small updates
some more comments + spell check

@@
 A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale,
-  in perparation (henceforth paper~IV)
+  in preparation (henceforth paper~IV)
 \end{itemize}
@@
 \item We add thermal diffusion.
+
+\item We added the volume-discrepancy term to the velocity constraint
+equation.
 \end{itemize}
@@
 \end{eqnarray}
+Here $\beta_0$ is computed as described in paper~III, Appendix C.
 
@@
   state density, $\rho_0$, to convert this to an edge state for $\rho$,
-
+  \MarginPar{why don't we do something higher order here for plane-parallel?}
 \begin{equation}
@@
 
-\item[spherical:] See {\tt MAESTO/docs/eta\_notes} for working notes until we finlize 
+\item[spherical:] See {\tt MAESTRO/docs/eta\_notes} for working notes until we finalize 
 the details.
@@
    We note that for spherical geometry, the edge state of $(\rho h)_0$
-   is computed by averging $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$
+   is computed by averaging $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$
    to the edge separately, and then computing 
@@
 
+   We note that for spherical geometry, the edge state of $(\rho h)_0$
+   is computed by averaging $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$
+   to the edge separately, and then computing 
+   ${(\rho h)_0}_{r+1/2} = {\rho_0}_{r+1/2} {h_0}_{r+1/2}$.
    \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$ without

### 2008-10-26 docs/flowchart/flowchart.tex
add a section on outstanding questions about the algorithm.
fix some texisms

@@
 \rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2),\star}}{2} \enskip ,
+\label{eq:rhoXedgestate}
 \end{equation}
@@
    to the edge separately, and then computing 
-   ${(\rho h)_0}_{r+1/2} = {\rho_0}_{r+1/2} {h_0}_{r+1/2}$.
+   ${(\rho h)_0}_{r+\myhalf} = {\rho_0}_{r+\myhalf} {h_0}_{r+\myhalf}$.
 
@@
 
-\MarginPar{is it really $p_0^n$ here and not $p_0^{n+1/2}$?}
+\MarginPar{is it really $p_0^n$ here and not $p_0^{\nph}$?}
 \end{description}
@@
    to the edge separately, and then computing 
-   ${(\rho h)_0}_{r+1/2} = {\rho_0}_{r+1/2} {h_0}_{r+1/2}$.
+   ${(\rho h)_0}_{r+\myhalf} = {\rho_0}_{r+\myhalf} {h_0}_{r+\myhalf}$.
    \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$ without
@@
 (\rho h)^{(1),\nph} \right] \right\} \nonumber \\
-    && + \dt \; \left \{ \nabla \cdot \left (\uadvtwo p_0^{n+1/2} \right ) 
-       - p_0^{n+1/2} \nabla \cdot \uadvtwo \right \} \nonumber \\
+    && + \dt \; \left \{ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) 
+       - p_0^{\nph} \nabla \cdot \uadvtwo \right \} \nonumber \\
     && + \dt \; \psi^{\nph} \enskip ,
@@
   \end{enumerate}
+
+\section{Outstanding Questions}
+
+\begin{enumerate}
+\item Why do we only update the temperature at the end of steps 4 and
+      8.  This means that at the top of the algorithm, the temperature
+      going into {\bf React State} in step 2 does not include the
+      effects of the previous reactions.
+
+      To be thermodynamically consistent, I think that we need to
+      update the temperature at the very end of step 9.  That way we
+      end the time-level update with $T^{n+1}$ instead of $T^{(2)}$,
+      and we go into the next sequence, step 2 with a valid temperature.
+
+\item Why is there no equivalent of $\etarho$ for enthalpy---e.g.\ an
+      $\eta_{\rho h}$?
+
+\item Why for plane parallel, when constructing the full edge state
+      (eq.~\ref{eq:rhoXedgestate}) do we simply average to the interface
+      instead of doing the 4th order interpolation that is used for the
+      full state?
+
+\end{enumerate}

### 2008-10-26 docs/flowchart/flowchart.tex
added some MarginPars about notation, and moved an equation

@@
 \end{eqnarray}
+\MarginPar{why is the notation for the edge state ${{\rho}_0}^{\outp,\nph}$? shouldn't
+           it be ${{\rho}_0}^{\inp,\nph}$?}
 where $j$ refers to the one-dimensional index in the radial direction.
@@
 \end{equation}
+\MarginPar{shouldn't the edge states for $p_0$ have a $\inp$ or $\outp$ superscript?}
 where the interface states are again found via the procedure described
@@
 \end{equation}
+\MarginPar{why is the notation for the edge state ${{\rho h}_0}^{\outp,\nph}$? shouldn't
+           it be ${{\rho h}_0}^{\inp,\nph}$?}
 where the interface states are again found via the procedure described
@@
 The local velocity field is described by
+\begin{equation}
+\frac{\partial\Ubt}{\partial t} = - \Ubt \cdotb \nablab \Ubt - w_0 \frac{\partial \Ubt}{\partial r}
+                                 - \left(\Ubt \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
+                                 - \frac{1}{\rho} \nablab\pi
+                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
+                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip\
+ ,
+\end{equation}
+(paper III; eq.~[37]).
 
@@
 The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
-We update the velocity field $\Ubt^n$ to $\Ubt^{n+1,\star}$ by discretizing
-\begin{equation}
-\frac{\partial\Ubt}{\partial t} = - \Ubt \cdotb \nablab \Ubt - w_0 \frac{\partial \Ubt}{\partial r}
-                                 - \left(\Ubt \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
-                                 - \frac{1}{\rho} \nablab\pi
-                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
-                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip\
- ,
-\end{equation}
-(paper III; eq.~[37]) as 
+We update the velocity field $\Ubt^n$ to $\Ubt^{n+1,\star}$ by discretizing eq.~\ref{eq:utildeupd}
+as
 \begin{eqnarray}

### 2008-10-26 docs/flowchart/flowchart.tex
add use_tfromp

@@
 \end{eqnarray}
-\item Update the temperature
-\begin{equation}
-T^{(2),\star} = T\left(\rho^{(2),\star}, h^{(2),\star}, X_k^{(2),\star}\right) {\hbox{ \rm using the equation of state.}}
-\end{equation}
+\item Update the temperature using the equation of state.
+
+The default is to do
+\begin{equation}
+T^{(2),\star} = T\left(\rho^{(2),\star}, h^{(2),\star}, X_k^{(2),\star}\right) \enskip ,
+\end{equation}
+but if {\tt use\_tfromp = T}, then we do
+\begin{equation}
+T^{(2),\star} = T\left(\rho^{(2),\star}, p_0^{(2),\star}, X_k^{(2),\star}\right) \enskip .
+\end{equation}
+This has the effect of completely decoupling the enthalpy equation from the 
+evolution.
 
@@
 \end{eqnarray}
-\item Update the temperature
-\begin{equation}
-T^{(2)} = T\left(\rho^{(2)}, h^{(2)}, X_k^{(2)}\right) {\hbox{ \rm using the equation of state.}}
-\end{equation}
+\item Update the temperature using the equation of state
+
+The default is to do
+\begin{equation}
+T^{(2)} = T\left(\rho^{(2)}, h^{(2)}, X_k^{(2)}\right) \enskip ,
+\end{equation}
+but if {\tt use\_tfromp = T}, then we do
+\begin{equation}
+T^{(2)} = T\left(\rho^{(2)}, {p_0}^{(2)}, X_k^{(2)}\right) \enskip .
+\end{equation}
+This has the effect of completely decoupling the enthalpy equation from the 
+evolution.
 

### 2008-10-27 docs/flowchart/flowchart.tex
now reflects that psi is computed within advect_base_pres

@@
 
-\item {\bf Advect Base Pressure}$[p_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [p_0^\outp]$\\
-is the process by which we update the base state pressure through $\dt$ in time.
+\item {\bf Advect Base Pressure}$[p_0^\inp,w_0^\inp,\etarho^\inp] \rightarrow$ 
+$[p_0^\outp,\psi^\outp]$\\
+is the process by which we update the base state pressure through $\dt$ in time and
+compute $\psi$.
 \begin{description}
 
-\item[plane-parallel:] We discretize the base state pressure
-  evolution equation (paper III; eq.~[31]),
+\item[plane-parallel:] First we compute a cell-centered $\psi^\outp$ using
+\begin{equation}
+ \psi_j^\outp = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^\inp + \eta_{\rho,j+\myhalf}^\inp\right) g \enskip .
+\end{equation}
+Then, we discretize the base state pressure evolution equation (paper III; eq.~[31]),
 \begin{equation}
@@
 + w_{0,j-\myhalf}^\inp \right)
-\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\inp 
+\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\outp 
 \enskip,
@@
 where the interface states are again found via the procedure described
-in paper II, Appendix A.  Note that the $\psi$ term is not included in the forcing for
-the interface state calculation, but is used in the update.
+in paper II, Appendix A.
 
@@
 
-\item (spherical only) {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph,\star}, \psi^{\nmh}]$ 
-$\rightarrow [p_0^{n+1,\star}]$.  Note that $\psi$ is not used in the spherical case.
-
-\item Define an edge-centered $\etarho^{\nph,\star}$, cell-centered 
-$(\nabla\cdot\etarho\er)^{\nph,\star}$, and cell-centered $\psi_j^{\nph,\star}$:
+\item Define an edge-centered $\etarho^{\nph,\star}$ and  cell-centered 
+$(\nabla\cdot\etarho\er)^{\nph,\star}$:
 
@@
 \end{equation}
-\begin{equation}
- \psi_j^{\nph,\star} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph,\star} + \eta_{\rho,j+\myhalf}^{\nph,\star}\right) g \enskip .
-\end{equation}
 
@@
 
-\item (plane-parallel only only) {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph,\star},$ 
-$\psi^{\nph,\star}] \rightarrow [p_0^{n+1,\star}]$.
+\item {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]$ 
+$\rightarrow [p_0^{n+1,\star}, \psi^{\nph,\star}]$.
 
@@
 
-\item (spherical only) {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph}, \psi^{\nph,\star}]$ 
-$\rightarrow [p_0^{n+1}]$.  Note that $\psi$ is not used in the spherical case.
-
-\item Define an edge-centered $\etarho^{\nph}$, cell-centered 
-$(\nabla\cdot\etarho\er)^{\nph}$, and a cell-centered $\psi_j^{\nph}$,
+\item Define an edge-centered $\etarho^{\nph}$ and cell-centered 
+$(\nabla\cdot\etarho\er)^{\nph}$:
 
@@
 \end{equation}
-\begin{equation}
- \psi_j^{\nph} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
-\end{equation}
 
@@
 
-\item  (plane-parallel only) {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph}, \psi^{\nph}]$ 
-$\rightarrow [p_0^{n+1}]$.
+\item {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
+$\rightarrow [p_0^{n+1}, \psi^{n+\myhalf}]$.
 

### 2008-10-27 docs/flowchart/flowchart.tex
found typo

@@
 
-\item {\bf Advect Base Enthalpy}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [(\rho h)_0]^\outp$\\
+\item {\bf Advect Base Enthalpy}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [(\rho h)_0^\outp]$\\
 is the process by which we update the base state enthalpy through $\dt$ in time.

### 2008-10-27 docs/flowchart/flowchart.tex
add some marginpars about volume discrepancy
Mike caught a sign error on volume discrepancy
fixed equations with volume discrepancy - units weren't right and the equations were written wrong

@@
 - \frac{1}{\gammabar^{n} p_0^{n}}
-\left\{ \psi^{\nmh} - 
+\left\{ \psi^{\nmh} + 
 f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)}^n}{\Delta t^n} \right]\right\}
@@
 \nablab \cdotb \left(\beta_0^n \uadvone\right) = 
-\beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
+\beta_0^n \left\{ \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
 + \frac{f}{\gammabar^n p_0^n}
-\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)}^n}{\Delta t^n}\right] \enskip .
+\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)}^n}{\Delta t^n}\right]\right\} \enskip .
 \end{equation}
@@
 \overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
-\left\{ \psi^{\nph,\star} - f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]\right\}
+\left\{ \psi^{\nph,\star} + f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]\right\}
 \enskip ,
@@
 with the volume discrepancy correction is
-\begin{eqnarray}
-\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) &=& 
-\beta_0^{\nph,\star} \left(S^{\nph} - \overline{S}^{\nph}\right) \nonumber \\
-&& + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
-\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]
-\enskip , \nonumber \\
-\end{eqnarray}
+\begin{equation}
+\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) =
+\beta_0^{\nph,\star}\left\{\left(S^{\nph} - \overline{S}^{\nph}\right)
+ + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
+\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]\right\} \enskip ,
+\end{equation}
 \begin{equation}
@@
 \begin{equation}
-\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)
+\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S}^{n+1} \right)
 + \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]
+\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]\right\}
 \enskip,
@@
  L_\beta^\rho \phi =
-   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)
+   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph}\left\{ \left(S^{n+1}-\overline{S}^{n+1} \right)
 - \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right] \enskip,
+\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]\right\} \enskip,
 \end{equation}
@@
 of this particular form of the projection operand.)  
-We solve the linear system of equation
-\begin{equation}
- L_\beta^\rho \phi =
-   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
-\label{eq:nodal_solve}
-\end{equation}
-(paper III; eq.~[77]) using multigrid V-cycles with Gauss-Seidel relaxation.
+We solve this linear system of equations using multigrid V-cycles with 
+Gauss-Seidel relaxation.
 

### 2008-10-28 docs/flowchart/flowchart.tex
addressed Mike's MaringPars about the in/out notation of the edge states in the advect base descriptions.  Fixed one typo, where the (rho h)0 edge states had an "out" superscript - they should not have this.
-The reason why the rho0 edge states have an "out" is because we compute these states within advect_base_dens.

-The reason why the rhoh0 edge states have neither "in" nor "out" is because we compute them within the advect_base_pres and advect_base_enthalpy subroutines and do not pass them out since they aren't used outside of the subroutine (for now).

@@
 \end{eqnarray}
-\MarginPar{why is the notation for the edge state ${{\rho}_0}^{\outp,\nph}$? shouldn't
-           it be ${{\rho}_0}^{\inp,\nph}$?}
 where $j$ refers to the one-dimensional index in the radial direction.
@@
 \end{equation}
-\MarginPar{shouldn't the edge states for $p_0$ have a $\inp$ or $\outp$ superscript?}
 where the interface states are again found via the procedure described
@@
 (\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
-\left\{ \left[ {(\rho h)_0}^{\outp,\nph} {w_0}^{\inp}\right]_{j+\myhalf} -  
-         \left[ {(\rho h)_0}^{\outp,\nph} {w_0}^{\inp}\right]_{j-\myhalf} \right\}
+\left\{ \left[ {(\rho h)_0}^{\nph} {w_0}^{\inp}\right]_{j+\myhalf} -  
+         \left[ {(\rho h)_0}^{\nph} {w_0}^{\inp}\right]_{j-\myhalf} \right\}
 + \dt\psi_j^{\inp} \enskip ,
 \end{equation}
-\MarginPar{why is the notation for the edge state ${{\rho h}_0}^{\outp,\nph}$? shouldn't
-           it be ${{\rho h}_0}^{\inp,\nph}$?}
 where the interface states are again found via the procedure described

### 2008-10-30 docs/flowchart/flowchart.tex
added line about what full_rhoh0_evolution does
some minor cleanup

@@
 \begin{equation}
-S^{\half,\star} = \frac{S^0 + S^1}{2} \enskip ,
+S^{\myhalf,\star} = \frac{S^0 + S^1}{2} \enskip ,
 \end{equation}
@@
 
-\item {\bf React Base}$[(\rho h)_0^n,(\overline{\rho \Hnuc})^{(1)},(\overline{\rho\Hext})^n] \rightarrow [(\rho h)_0^{(1)}]$
-
+\item {\bf if} ~ {\tt full\_rhoh0\_evolution = T} ~ {\bf then}\\
+{\bf React Base}$[(\rho h)_0^n,(\overline{\rho \Hnuc})^{(1)},(\overline{\rho\Hext})^n] \rightarrow [(\rho h)_0^{(1)}]$ \\
+{\bf else}\\
+$(\rho h)_0^{(1)} = (\overline{\rho h})^{(1)}$\\
+{\bf end if}
 \item
@@
 first time step.  As initial approximations, set $S^{1,0} = S^0$,
-$\etarho^{-\half} = 0$, $\psi^{-\half} = 0$, and $\pi^{-\half} = 0.$
+$\etarho^{-\myhalf} = 0$, $\psi^{-\myhalf} = 0$, and $\pi^{-\myhalf} = 0.$
 \begin{enumerate}
@@
   \item Perform steps 1--11 as described above, using 
-    $S^{\half,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
+    $S^{\myhalf,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
     The only other difference in the time advancement is that in step 11
@@
     \begin{equation}  L_\beta^\rho \phi =
-      D \left ( \beta_0^{\half} {\bf V} \right) - \beta_0^{\half} \left[ (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right] \enskip . 
+      D \left ( \beta_0^{\myhalf} {\bf V} \right) - \beta_0^{\myhalf} \left[ (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right] \enskip . 
     \end{equation}
@@
       We discard the new velocity resulting from this, but keep the new  
-      value for $\pi^{\half} = \pi^{-\half} + (1 / \dt) \; \phi.$  
+      value for $\pi^{\myhalf} = \pi^{-\myhalf} + (1 / \dt) \; \phi.$  
       These steps also yield new scalar data at time $\dt,$ which
-      we discard,  and new values for $\etarho^{\half}$ (step 8), $\psi^{\half}$ (step 8), 
-      $S^{1,\nu+1}$ (step 10), and $\pi^{\half}$ (step 11), which we keep.
-    \item Set $\pi^{-\half} = \pi^{\half}$, $\etarho^{-\half} = \etarho^{\half}$,
-      and $\psi^{-\half} = \psi^{\half}$. 
+      we discard,  and new values for $\etarho^{\myhalf}$ (step 8), $\psi^{\myhalf}$ (step 8), 
+      $S^{1,\nu+1}$ (step 10), and $\pi^{\myhalf}$ (step 11), which we keep.
+    \item Set $\pi^{-\myhalf} = \pi^{\myhalf}$, $\etarho^{-\myhalf} = \etarho^{\myhalf}$,
+      and $\psi^{-\myhalf} = \psi^{\myhalf}$. 
     \end{enumerate}
@@
 \begin{enumerate}
+
 \item Why do we only update the temperature at the end of steps 4 and
@@
 
-\item Why for plane parallel, when constructing the full edge state
-      (eq.~\ref{eq:rhoXedgestate}) do we simply average to the interface
-      instead of doing the 4th order interpolation that is used for the
-      full state?
-
-\end{enumerate}
+\end{enumerate}

### 2008-11-03 docs/flowchart/flowchart.tex
addressed a marginpar, minor typo fix

@@
   Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  We use the base
-  state density, $\rho_0$, to convert this to an edge state for $\rho$,
-  \MarginPar{why don't we do something higher order here for plane-parallel?}
+  state density, $\rho_0$, to convert this to an edge state for $\rho$.
+  We use 4th-order spatial averaging to get $\rho_0$ on radial faces,
 \begin{equation}
@@
     already in {\bf React State}.  We use the base state enthalpy,
-    $(\rho h)_0$, to convert this to an edge state for $(\rho h)$,
+    $(\rho h)_0$, to convert this to an edge state for $(\rho h)$.
+    We use 4th-order spatial averaging to get $(\rho h)_0$ on radial faces,
     \MarginPar{in these notes we currently assume pred type 1, $(\rho h)'$}

### 2008-11-03 Source/mkutrans.f90
using characteristic tracing now treating dw0/dr term as a source term instead of an advection term (thus eliminating its presence here) updated document to reflect this

@@ -102,7 +102,6 @@ contains
     use bc_module
     use slope_module
     use geometry, only: nr
-    use probin_module, only: use_new_godunov
 
     integer,         intent(in   ) :: n,lo(:),hi(:),ng_u,ng_ut
     real(kind=dp_t), intent(in   ) ::    vel(lo(1)-ng_u :,lo(2)-ng_u :,:)
@@ -116,7 +115,7 @@ contains
     real(kind=dp_t) :: velx(lo(1)-1:hi(1)+1,lo(2)-1:hi(2)+1,1)
     real(kind=dp_t) :: vely(lo(1)-1:hi(1)+1,lo(2)-1:hi(2)+1,1)
     
-    real(kind=dp_t) hx, hy, dth, umax, dw0drhi, dw0drlo
+    real(kind=dp_t) hx, hy, dth, umax
     real(kind=dp_t) ulft,urgt,vbot,vtop,vlo,vhi,eps,abs_eps
 
     integer :: i,j,is,js,ie,je
@@ -157,13 +156,8 @@ contains
     do j = js,je
        do i = is,ie+1 
           
-          if (use_new_godunov) then
-             urgt = vel(i  ,j,1) - (HALF + dth*min(ZERO,vel(i  ,j,1))/hx) * velx(i  ,j,1)
-             ulft = vel(i-1,j,1) + (HALF - dth*max(ZERO,vel(i-1,j,1))/hx) * velx(i-1,j,1)
-          else
-             urgt = vel(i  ,j,1) - (HALF + dth*vel(i  ,j,1)/hx) * velx(i  ,j,1)
-             ulft = vel(i-1,j,1) + (HALF - dth*vel(i-1,j,1)/hx) * velx(i-1,j,1)
-          end if
+          urgt = vel(i  ,j,1) - (HALF + dth*min(ZERO,vel(i  ,j,1))/hx) * velx(i  ,j,1)
+          ulft = vel(i-1,j,1) + (HALF - dth*max(ZERO,vel(i-1,j,1))/hx) * velx(i-1,j,1)
           
           urgt = merge(vel(is-1,j,1),urgt,i.eq.is   .and. phys_bc(1,1) .eq. INLET)
           urgt = merge(vel(ie+1,j,1),urgt,i.eq.ie+1 .and. phys_bc(1,2) .eq. INLET)
@@ -193,29 +187,18 @@ contains
 
           if (j .eq. nr(n)) then
              vhi = vel(i,j,2) + w0(j)
[…]

### 2008-11-03 Source/velpred.f90
using characteristic tracing; updating doc to reflect ongoing code changes

@@ -168,7 +168,6 @@ contains
     use bc_module
     use slope_module
     use bl_constants_module
-    use probin_module, only: use_new_godunov
 
     integer        , intent(in   ) :: n,lo(:),hi(:),ng_u,ng_um,ng_ut,ng_f
     real(kind=dp_t), intent(in   ) ::      u(lo(1)-ng_u :,lo(2)-ng_u :,:)
@@ -245,13 +244,8 @@ contains
              vhi = u(i,j+1,2) + w0(j+1)
           end if
 
-          if (use_new_godunov) then
-             spbot = u(i,j  ,1) + (HALF - dth*max(ZERO,vlo)/hy) * slopey(i,j  ,1)
-             sptop = u(i,j+1,1) - (HALF + dth*min(ZERO,vhi)/hy) * slopey(i,j+1,1)
-          else
-             spbot = u(i,j  ,1) + (HALF - dth*vlo/hy) * slopey(i,j  ,1)
-             sptop = u(i,j+1,1) - (HALF + dth*vhi/hy) * slopey(i,j+1,1)
-          end if
+          spbot = u(i,j  ,1) + (HALF - dth*max(ZERO,vlo)/hy) * slopey(i,j  ,1)
+          sptop = u(i,j+1,1) - (HALF + dth*min(ZERO,vhi)/hy) * slopey(i,j+1,1)
 
           sptop = merge(u(i,je+1,1),sptop,j.eq.je .and. phys_bc(2,2) .eq. INLET)
           spbot = merge(u(i,je+1,1),spbot,j.eq.je .and. phys_bc(2,2) .eq. INLET)
@@ -273,13 +267,8 @@ contains
           end if
           vhi = u(i,j  ,2) + HALF * (w0(j  )+w0(j+1))
 
-          if (use_new_godunov) then
-             smtop = u(i,j  ,1) - (HALF + dth*min(ZERO,vhi)/hy) * slopey(i,j  ,1)
-             smbot = u(i,j-1,1) + (HALF - dth*max(ZERO,vlo)/hy) * slopey(i,j-1,1)
-          else
-             smtop = u(i,j  ,1) - (HALF + dth*vhi/hy) * slopey(i,j  ,1)
-             smbot = u(i,j-1,1) + (HALF - dth*vlo/hy) * slopey(i,j-1,1)
-          end if
[…]

### 2008-11-04 Source/make_edge_scal.f90
eliminated use_new_godunov flag - not needed anymore since the advection has been sufficiently upgraded (but I'm not quite done with some other changes that need to be made to make the code match the advection notes)

@@ -199,7 +199,6 @@ contains
     use bc_module
     use slope_module
     use bl_constants_module
-    use probin_module, only: use_new_godunov
 
     integer        , intent(in   ) :: n,lo(:),hi(:),ng_s,ng_se,ng_um,ng_f
     real(kind=dp_t), intent(in   ) ::      s(lo(1)-ng_s :,lo(2)-ng_s :,:)
@@ -227,7 +226,7 @@ contains
     real(kind=dp_t) :: hx,hy,dth,splus,sminus
     real(kind=dp_t) :: savg,st
     real(kind=dp_t) :: sptop,spbot,smtop,smbot,splft,sprgt,smlft,smrgt
-    real(kind=dp_t) :: abs_eps,eps,umax,dw0drhi,dw0drlo,vtilde
+    real(kind=dp_t) :: abs_eps,eps,umax
 
     integer :: i,j,is,js,ie,je
 
@@ -271,22 +270,6 @@ contains
           spbot = s(i,j  ,comp) + (HALF - dth*vmac(i,j+1)/hy) * slopey(i,j  ,1)
           sptop = s(i,j+1,comp) - (HALF + dth*vmac(i,j+1)/hy) * slopey(i,j+1,1)
 
-          if (use_new_godunov .and. is_vel .and. comp .eq. 2) then
-
-             if (j .ne. nr(n)-1) then
-                dw0drhi = (w0(j+2)-w0(j+1))/dx(2)
-             else
-                dw0drhi = (w0(j+1)-w0(j))/dx(2)
-             end if
-             dw0drlo = (w0(j+1)-w0(j))/dx(2)
-
-             vtilde = vmac(i,j+1) - w0(j+1)
-
-             spbot = spbot - dth*vtilde*dw0drlo
-             sptop = sptop - dth*vtilde*dw0drhi
-
-          end if
-
           sptop = merge(s(i,je+1,comp),sptop,j.eq.je .and. phys_bc(2,2) .eq. INLET)
           spbot = merge(s(i,je+1,comp),spbot,j.eq.je .and. phys_bc(2,2) .eq. INLET)
 
@@ -309,22 +292,6 @@ contains
 
           smtop = s(i,j  ,comp) - (HALF + dth*vmac(i,j)/hy) * slopey(i,j  ,1)
           smbot = s(i,j-1,comp) + (HALF - dth*vmac(i,j)/hy) * slopey(i,j-1,1)
-
[…]

### 2008-11-11 docs/flowchart/flowchart.tex
not about time-centering p0 updated
added thermal conduction term into S
some temperature notes clarifications

@@
   S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext \enskip ,
+  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext 
+  + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T \enskip ,
 \label{eq:defineS} 
@@
 
-\MarginPar{is it really $p_0^n$ here and not $p_0^{\nph}$?}
+\MarginPar{We could time-center $p_0$ now that we have split out {\tt advect\_base\_pres}}
 \end{description}
@@
   \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star}  
-  + \sigma \Hext \enskip .
+  + \sigma \Hext + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T^{n+1,\star} \enskip .
 \end{equation} 
 where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} / \rho^{(2),\star}$
-and the remaining quantities are defined using $X_k^{n+1,\star},$ $\rho^{n+1,\star},$ 
-and $T^{n+1,\star}$ from step 5.
+and the thermodynamic quantities are defined using $\rho^{n+1,\star}, X_k^{n+1,\star},$ 
+and $T^{n+1,\star}$ as inputs to the equation of state.
 Then define
@@
 with the volume discrepancy correction is
-\begin{equation}
-\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) =
+\begin{eqnarray}
+\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) &=&
 \beta_0^{\nph,\star}\left\{\left(S^{\nph} - \overline{S}^{\nph}\right)
- + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
-\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]\right\} \enskip ,
-\end{equation}
+ + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}\cdot \right. \nonumber \\
+&& \left.\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]\right\} \enskip ,
+\end{eqnarray}
 \begin{equation}
@@
   \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2)}  
-   + \sigma \Hext \enskip ,
+   + \sigma \Hext + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T^{n+1} \enskip ,
 \end{equation}
 where $(\omegadot_k)^{(2)} = (\rho \omegadot_k)^{(2)} / \rho^{(2)}$
-and the remaining quantities are defined using $X_k^{n+1},$ $\rho^{n+1},$
-and $T^{n+1}$ from step 9.
+and the thermodynamic quantities are defined using $\rho^{n+1}, X_k^{n+1},$
+and $T^{n+1}$ as inputs to the equation of state.
 

### 2008-11-12 docs/flowchart/flowchart.tex
typos and notation updates from Chris

@@
 \hline
-$\psi$  & $\psi \equiv D_0 p_0/Dt$ & erg~cm$^{-3}$~s$^{-1}$ \\
+$\psi$  & $\psi \equiv D_0 p_0/Dt = \partial p_0/\partial t + w_0\partial p_0/\partial r$ & erg~cm$^{-3}$~s$^{-1}$ \\
 \hline
@@
 
-\item[plane-parallel:] We discretize equation the base state density evolution 
+\item[plane-parallel:] We discretize the base state density evolution 
 equation (paper III; eq.~[29]) without the $\etarho$ term,
@@
 where the interface states are again found via the procedure described
-in paper II, Appendix A.  Note that the $\psi$ term is not included in 
-the forcing for the interface state calculation, but is used in the update.
+in paper II, Appendix A.
 
@@
 \item In paper III we discretized the enthalpy evolution equation in
-full, non-perturbation form.  We have discovered that discretizing the
-enthalpy evolution in non-perturbational form leads to better
+full, non-perturbation form.  Since then we have discovered that 
+discretizing the enthalpy evolution in perturbational form leads to better
 numerical properties.  This is more like paper II.
@@
 
-  \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2'),\star}$ without
+  \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2a),\star}$ without
   explicitly including the reaction terms,
@@
   \begin{eqnarray}
-  (\rho h)^{(2'),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
+  (\rho h)^{(2a),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
       \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right) (\rho
@@
 \begin{eqnarray}
-(\rho h)^{(2),\star} &=& (\rho h)^{(2'),\star} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(2),\star} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
+(\rho h)^{(2),\star} &=& (\rho h)^{(2a),\star} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(2),\star} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
 &&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
@@
 \begin{eqnarray}
-\left(\rho^{(2),\star} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla\right)h^{(2),\star} &=& (\rho h)^{(2'),\star} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
+\left(\rho^{(2),\star} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla\right)h^{(2),\star} &=& (\rho h)^{(2a),\star} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
 &&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
@@
    ${(\rho h)_0}_{r+\myhalf} = {\rho_0}_{r+\myhalf} {h_0}_{r+\myhalf}$.
-   \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$ without
+   \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2a)}$ without
      explicitly including the reaction terms,
@@
 \begin{eqnarray}
-(\rho h)^{(2')} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
+(\rho h)^{(2a)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
 (\rho h)^{(1),\nph} \right] \right\} \nonumber \\
@@
 \begin{eqnarray}
-(\rho h)^{(2)} &=& (\rho h)^{(2')} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla h^{(2)} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
+(\rho h)^{(2)} &=& (\rho h)^{(2a)} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla h^{(2)} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
 &&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
@@
 \begin{eqnarray}
-\left(\rho^{(2)} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla\right)h^{(2)} &=& (\rho h)^{(2')} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
+\left(\rho^{(2)} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla\right)h^{(2)} &=& (\rho h)^{(2a)} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
 &&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
@@
 - \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]\right\} \enskip,
+\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]\right\} \enskip, \label{Linear System for Multigrid}
 \end{equation}
@@
 of this particular form of the projection operand.)  
-We solve this linear system of equations using multigrid V-cycles with 
-Gauss-Seidel relaxation.
+We solve the linear system of equation (\ref{Linear System for Multigrid})
+using multigrid V-cycles with Gauss-Seidel relaxation.
 

### 2008-11-13 docs/flowchart/flowchart.tex
cleaned up notation a bit

@@
 \item {\bf Advect Base Density}$[\rhozero^{n},w_0^{\nph,\star}] \rightarrow$ 
-$[\rho_0^{(2),\star}, \rho_0^{\nph,\star,{\rm pred}}]$.
+$[\rho_0^{(2a),\star}, \rho_0^{\nph,\star,{\rm pred}}]$.
 
@@
 \rho^{(1),\nph,\star} = 
-\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2),\star}}{2} \enskip ,
+\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip ,
 \label{eq:rhoXedgestate}
@@
 
-  \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2a),\star}$ without
+  \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(1a),\star}$ without
   explicitly including the reaction terms,
@@
   \begin{eqnarray}
-  (\rho h)^{(2a),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
+  (\rho h)^{(1a),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
       \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right) (\rho
@@
 
-\item {\bf Correct Base}$[\rho_0^{(2),\star}, (\nabla\cdot\etarho\er)^{\nph,*}] \rightarrow [\rho_0^{n+1,\star}]$.
+\item {\bf Correct Base}$[\rho_0^{(2a),\star}, (\nabla\cdot\etarho\er)^{\nph,*}] \rightarrow [\rho_0^{n+1,\star}]$.
 
@@
 \begin{eqnarray}
-(\rho h)^{(2),\star} &=& (\rho h)^{(2a),\star} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(2),\star} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
+(\rho h)^{(2),\star} &=& (\rho h)^{(1a),\star} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(2),\star} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
 &&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
@@
 \begin{eqnarray}
-\left(\rho^{(2),\star} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla\right)h^{(2),\star} &=& (\rho h)^{(2a),\star} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
+\left(\rho^{(2),\star} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla\right)h^{(2),\star} &=& (\rho h)^{(1a),\star} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
 &&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
@@
 \item {\bf Advect Base Density}$[\rho_0^{n},w_0^{\nph}] \rightarrow$ 
-$[\rho_0^{(2)}, \rho_0^{\nph,{\rm pred}}]$.
+$[\rho_0^{(2a)}, \rho_0^{\nph,{\rm pred}}]$.
 
@@
 \begin{equation}
-\rho^{(1),\nph} = \rho^{'(1),\nph} + \frac{\rho_0^n + \rho_0^{(2)}}{2}\enskip ,
+\rho^{(1),\nph} = \rho^{'(1),\nph} + \frac{\rho_0^n + \rho_0^{(2a)}}{2}\enskip ,
 \end{equation}
@@
    ${(\rho h)_0}_{r+\myhalf} = {\rho_0}_{r+\myhalf} {h_0}_{r+\myhalf}$.
-   \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2a)}$ without
+   \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(1a)}$ without
      explicitly including the reaction terms,
@@
 \begin{eqnarray}
-(\rho h)^{(2a)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
+(\rho h)^{(1a)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
 (\rho h)^{(1),\nph} \right] \right\} \nonumber \\
@@
 
-\item {\bf Correct Base}$[\rho_0^{(2)}, (\nabla\cdot\etarho\er)^{\nph}] \rightarrow [\rho_0^{n+1}]$
+\item {\bf Correct Base}$[\rho_0^{(2a)}, (\nabla\cdot\etarho\er)^{\nph}] \rightarrow [\rho_0^{n+1}]$
 
@@
 \begin{eqnarray}
-(\rho h)^{(2)} &=& (\rho h)^{(2a)} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla h^{(2)} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
+(\rho h)^{(2)} &=& (\rho h)^{(1a)} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla h^{(2)} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
 &&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
@@
 \begin{eqnarray}
-\left(\rho^{(2)} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla\right)h^{(2)} &=& (\rho h)^{(2a)} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
+\left(\rho^{(2)} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla\right)h^{(2)} &=& (\rho h)^{(1a)} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
 &&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\

### 2008-11-14 docs/volume_discrepancy/volume_discrepancy.tex
add the notes on the volume discrepancy term

The volume discrepancy term is used in the constraint equation to
force the system back to the equation of state.  We write our velocity
constraint equation as
\begin{equation}
\nablab \cdotb (\beta_0 \Ub)  = \beta_0 \left(S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} - \frac{f}{\gammabar p_0} \frac{p_0 - p_\mathrm{EOS}}{\Delta t} \right ) \enskip .
\label{eq:fullconstraint}
\end{equation}
Here, $f$ is the volume discrepancy factor and ranges from 0 to 1, and
$p_\mathrm{EOS}$ is the thermodynamic pressure as returned by the EOS,
using the full state as input.
In practice we evaluate this as
\begin{equation}
p_\mathrm{EOS} = p(\rho,h,X_k)
\end{equation}

The idea behind this forcing term is that if $p_\mathrm{EOS} > p_0$ then
$\nablab \cdotb (\beta_0 \Ub) > 0$, and the cell will evacuate.  This
has the effect of returning us to $p_\mathrm{EOS} = p_0$.

In {\tt MAESTRO}, we decomponse the velocity into a base state component
and a local component.  The base state constraint equation is simply
the horizontal average of the full constraint.  Starting with 
$\Ub = \Ubt + w_0 \er$ in equation~\ref{eq:fullconstraint}, we have
\begin{equation}
\nablab \cdotb (\beta_0 w_0 \er) + \nablab \cdotb (\beta_0 \Ubt)  = \beta_0 \left(S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} - \frac{f}{\gammabar p_0} \frac{p_0 - p_\mathrm{EOS}}{\Delta t} \right ) \enskip .
\label{eq:fullconstraint2}
\end{equation}
Averaging this over a layer, we note that $\overline{\nablab \cdotb (\beta_0 \Ubt)} = 0$,
yielding
\begin{equation}
\nablab \cdotb (\beta_0 w_0 \er)  = \beta_0 \left(\overline{S} - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} - \frac{f}{\gammabar p_0} \frac{p_0 - \overline{p_\mathrm{EOS}}}{\Delta t} \right ) 
\label{eq:w0constraint_vd}
\end{equation}
and
\begin{equation}
\nablab \cdotb (\beta_0 \Ubt)  = \beta_0 \left(S - \overline{S} + \frac{f}{\gammabar p_0} \frac{p_\mathrm{EOS} - \overline{p_\mathrm{EOS}}}{\Delta t} \right ) \enskip .
\label{eq:Utconstraint_vd}
\end{equation}

### 2008-11-20 docs/radial_base/radial_base.tex
New set of notes about radial base state mapping

Here we talk about the radial base state.

\section{Mapping}

### 2008-11-20 docs/radial_base/radial_base.tex
starting point for these notes is the text from paper 4

@@
-Here we talk about the radial base state.
+Since the one-dimensional spherical base state is not aligned with the
+3-d Cartesian grid we use for the full star, computing quantities that
+involve both the base state and the full state becomes complicated.
+Therefore, when we go between the base state and the full state in
+various parts of the algorithm (such as the averaging operations), we
+need to map between the two geometries.
 
-\section{Mapping}
+As there is no alignment between the 3-d Cartesian grid and the 1-d
+spherical base state grid, we are free to choose their grid spacings
+independently.  Experimentation has shown that the best choice results
+from picking the 3-d Cartesian grid spacing, $\Delta x$ to be coarser
+than the 1-d spherical base state spacing, $\Delta r$.  A typical simulation
+may use $5 \Delta r = \Delta x$.
+We call the procedure to go from 1-d to 3-d {\tt fill\_3d} and the
+complementary procedure that takes us from 3-d to 1-d {\tt average}.
+\section{\tt fill\_3d}
+Figure XXX shows the Cartesian grid with the spherical
+base state overlayed (for simplicity, the figure is drawn in 2-d).  
+The {\tt fill\_3d} procedure computes the radius of the cell center
+from the center of the star,
+\begin{equation}
+r = \sqrt{(x_i - x_c)^2 + (y_j - y_c)^2 + (z_k - z_c)^2}
+\end{equation}
+where $(x_c, y_c, z_c)$ are the coordinates of the center of the star.
+We use this radius to find the corresponding radial bin as $n = \mathtt{int}(r
+/ \Delta r)$ (here, our convention is to use 0-based indexing for the
+base state).  We can then initialize a Cartesian cell quantity $q$ from its
+corresponding base state quantity, $q_0$ as $q_{i,j,k} = q_{0,n}$.
+\subsection{\tt put\_1d\_array\_on\_cart}
+\subsection{\tt put\_w0\_on\_edges}
+\section{\tt average}
+For the {\tt average} process, we again compute the radius of the
+Cartesian cell center, and compute the index, $n$, of the
+corresponding base state bin as above.  $q_{0,n}$ is then the average
+of all the $q_{i,j,k}$ whose Cartesian cell center maps into radial
+bin $n$.  However, if $\Delta r < \Delta x$, there may be some base
+state bins that are not filled.  In this case, we will fill them via
+interpolation. \MarginPar{more detail will be needed here.}

### 2008-11-20 docs/radial_base/radial_base.tex
more details, added Mike's fig

@@
 from picking the 3-d Cartesian grid spacing, $\Delta x$ to be coarser
-than the 1-d spherical base state spacing, $\Delta r$.  A typical simulation
-may use $5 \Delta r = \Delta x$.
+than the 1-d spherical base state spacing, $\Delta r$.
+A typical simulation may use $5 \Delta r = \Delta x$.
 We call the procedure to go from 1-d to 3-d {\tt fill\_3d} and the
@@
 \section{\tt fill\_3d}
-Figure XXX shows the Cartesian grid with the spherical
+There are three different {\tt fill\_3d} modes:
+\begin{enumerate}
+\item The 1D array is cell-centered, and we map to Cartesian cell-centers.
+\item The 1D array is edge-centered, and we map to Cartesian cell-centers.
+\item The 1D array is edge-centered, and we map to Cartesian edges.  
+\end{enumerate}
+\subsection{1D Cell-Centered to Cartesian Cell-Centered}
+This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
+Figure \ref{fig:mapping} shows the Cartesian grid with the spherical
 base state overlayed (for simplicity, the figure is drawn in 2-d).  
-The {\tt fill\_3d} procedure computes the radius of the cell center
-from the center of the star,
+We compute the radius of the cell center from the center of the star,
 \begin{equation}
@@
 corresponding base state quantity, $q_0$ as $q_{i,j,k} = q_{0,n}$.
-\subsection{\tt put\_1d\_array\_on\_cart}
-\subsection{\tt put\_w0\_on\_edges}
+\subsection{1D Edge-Centered to Cartesian Cell-Centered}
+This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
+\subsection{1D Edge-Centered to Cartesian Edges}
+This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
+This case is only used to create {\tt w0mac} from {\tt w0} in the 
+subroutine, {\tt put\_w0\_on\_edges}.
 \section{\tt average}
-For the {\tt average} process, we again compute the radius of the
-Cartesian cell center, and compute the index, $n$, of the
-corresponding base state bin as above.  $q_{0,n}$ is then the average
-of all the $q_{i,j,k}$ whose Cartesian cell center maps into radial
-bin $n$.  However, if $\Delta r < \Delta x$, there may be some base
-state bins that are not filled.  In this case, we will fill them via
-interpolation. \MarginPar{more detail will be needed here.}
+For the {\tt average} process, we first create a coarse 1D radial array 
+with $\Delta x = \Delta r$.  Then, we compute the radius of the Cartesian cell 
+center, and compute the index, $n$, of the corresponding coarse radial 
+array bin.  $q_{0,n}$ is then the average of all the $q_{i,j,k}$ whose 
+Cartesian cell center maps into coarse radial bin $n$.  Then, we 
+interpolate the coarse radial array onto the finer radial array.
+
+\clearpage
+
+\begin{figure}[tpb]
+\begin{center}
+\epsffile{./radial_base/mapping.eps}
+\caption{\label{fig:mapping} The Cartesian grid and spherical base
+state (shown here in 2-d for simplicity).  Here we represent the
+spherical base state as concentric shells (blue lines).  Since the
+base state is not aligned with the Cartesian grid we use on which
+we discretize the star, we need to map between the two configurations.
+The `$+$' symbols represent the Cartesian zone centers.  In our
+mapping, the zone marked with the red symbol maps into the gray-shaded
+radial bin. }
+\label{fig:mapping}
+\end{center}
+\end{figure}

### 2008-11-20 docs/radial_base/radial_base.tex
completed the verbal description of the fill and average operations for spherical.  I don't think a higher level of detail is really necessary unless inquiring minds want to know...

@@
 \begin{enumerate}
-\item The 1D array is cell-centered, and we map to Cartesian cell-centers.
+\item The 1D array is bin-centered, and we map to Cartesian cell-centers.
 \item The 1D array is edge-centered, and we map to Cartesian cell-centers.
@@
 base state overlayed (for simplicity, the figure is drawn in 2-d).  
-We compute the radius of the cell center from the center of the star,
+We compute the radius of the Cartesian cell center from the center of the star,
 \begin{equation}
@@
 base state).  We can then initialize a Cartesian cell quantity $q$ from its
-corresponding base state quantity, $q_0$ as $q_{i,j,k} = q_{0,n}$.
+corresponding base state quantity, $q_0$ as $q_{i,j,k} = q_{0,n}$.  In essence,
+we are doing a piecewise constant interpolation of the radial array onto the 
+Cartesian grid.  If the output is a vector quantity, rather than a scalar, we
+multiply the result by the unit vector in the normal direction.
 \subsection{1D Edge-Centered to Cartesian Cell-Centered}
 This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
+Again, for each Cartesian cell-center, we compute the corresponding radial
+bin.  We linearly interpolate from the two nearest radial edge values 
+to get the Cartesian cell-centered value.  In essence,
+we are doing a piecewise linear interpolation of the radial array onto the 
+Cartesian grid.  If the output is a vector quantity, rather than a scalar, we
+multiply the result by the unit vector in the normal direction.
 \subsection{1D Edge-Centered to Cartesian Edges}
@@
 This case is only used to create {\tt w0mac} from {\tt w0} in the 
-subroutine, {\tt put\_w0\_on\_edges}.
+subroutine, {\tt put\_w0\_on\_edges}.  There are several variations 
+of the interpolation, controlled by {\tt w0mac\_interp\_type}.  We 
+currently use the default, {\tt w0mac\_interp\_type} = 1.  We first 
+map $w_0$ onto Cartesian cell-centers, and multiply the result by 
+the unit vector in the normal direction.  Then, for each Cartesian 
+face, we obtain the normal velocity by averaging the corresponding 
+velocities from the Cartesian cell-centers adjacent to the face.
 \section{\tt average}
@@
 center, and compute the index, $n$, of the corresponding coarse radial 
-array bin.  $q_{0,n}$ is then the average of all the $q_{i,j,k}$ whose 
+bin.  $q_{0,n}$ is then the average of all the $q_{i,j,k}$ whose 
 Cartesian cell center maps into coarse radial bin $n$.  Then, we 

### 2008-11-25 docs/rotation/rotation.tex
Some initial remarks about rotation and how we handle this in spherical.
Plane-parallel still needs to be addressed

\section{Rotation}\label{Sec:Rotation}
To handle rotation in Maestro we move to the co-rotating reference frame.  Time
derivatives of a vector in the inertial frame are related to those in the 
co-rotating frame by
\begin{equation}\label{eq:derivative relations}
  \left(\frac{D}{Dt}\right)_\text{i} = 
  \left[\left(\frac{D}{Dt}\right)_\text{rot} + \Omegab\times\ \right],
\end{equation}
where i(rot) refers to the inertial(co-rotating) frame and $\Omegab$ is
the angular velocity vector.  Using (\ref{eq:derivative relations}) and 
assuming $\Omegab$ is constant, we have
\begin{equation}\label{eq:rotational velocity relation}
  \frac{Dv_\text{i}}{Dt} = \frac{Dv_\text{rot}}{Dt} + 
  2\Omegab \times \mathbf{v_\text{rot}} +
  \Omegab\times\left(\Omegab\times r_\text{rot}\right).
\end{equation}
Plugging this into the momentum equation and making use of the continuity 
equation we have a momentum equation in the co-rotating frame:
\begin{equation}\label{eq:momentum equation with rotation}
  \frac{\partial(\rho\mathbf{U})}{\partial t} + 
  \nabla\cdot\left(\mathbf{U}(\rho\mathbf{U}) + p \right) = 
  \underbrace{-2\rho\Omegab\times\mathbf{U}}_{\text{Coriolis}} -
  \underbrace{\rho\Omegab\times\left(\Omegab\times
    \rb\right)}_{\text{Centrifugal}} - 
  \rho \mathbf{g}.
\end{equation}
The Coriolis and Centrifugal terms are force terms that will be added to 
right hand side of the equations in {\tt mk\_vel\_force}.  Note that the 
Centrifugal term can be rewritten as a gradient of a potential and absorbed
into either the pressure or gravitational term to create an effective pressure
or effective gravitational potential.  Furthermore, because it can be written
as a gradient, this term would drop out completely in the projection if we were
doing incompressible flow.\MarginPar{Is this how John stated it?}
In what follows we will include the Centrifugal term explicitly.

\subsection{Using Spherical Geometry}\label{Sec:Using Spherical Geometry}
In spherical geometry implementing rotation is straightforward as we don't have
to worry about special boundary conditions.  We assume a geometry as shown in
Figure \ref{Fig:rotation in spherical} where the primed coordinate system is 
the simulation domain coordinate system, the unprimed system is the primed 
system translated by the vector $\rb_\text{c}$ and is added here for
clarity.  The  $\rb_\text{c}$ vector is given by {\tt center} in the 
{\tt geometry} module.  In spherical, the only runtime parameter of importance
is {\tt rotational\_frequency} in units of Hz.  This specifies the angular 
velocity vector which is assumed to be along the {\bf k} direction:
\[
\Omegab \equiv \Omega \text{\bf k} = 2\pi *
\left(\text{\tt rotational\_frequency}\right)\text{\bf k}.
\]
The direction of $\rb$ is given as the {\tt normal} vector which is 
passed into {\tt mk\_vel\_force}; in particular
\[
\cos\theta \equiv \frac{\rb\cdot\text{\bf k}}{r} = 
\text{\bf normal}\cdot\text{\bf k} = \text{\tt normal}(3).
\]
The magnitude of $\rb$ is calculated based on the current zone's location with
respect to the {\tt center}.
Using this notation we can write the Centrifugal term as
\begin{align*}
\Omegab\times\left(\Omegab\times\rb\right) &=
\left(\Omegab\cdot\rb\right)\Omegab - \left(\Omegab\cdot\Omegab\right)\rb\\
&= \Omega^2 r *\left[\text{\tt normal}(3)\right]\text{\bf k} -
\Omega^2 r *\text{\bf normal} = \left(
\begin{array}{c}
\Omega^2r*\left[\text{\tt normal}(1)\right]\\
\Omega^2r*\left[\text{\tt normal}(2)\right]\\
0 \end{array}\right).
\end{align*}
The Coriolis term is a straightforward cross-product:
\begin{align*}
\Omegab \times \Ub &= \left|
\begin{array}{ccc}
  \text{\bf{i}}&\text{\bf{j}}&\text{\bf{k}}\\
  0 & 0 & \Omega\\
  u & v & w
\end{array}\right|\\
&= \left(
\begin{array}{c}
-\Omega v\\ \Omega u \\ 0
\end{array}
\right).
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tpb]
\hspace{-0.25in}
\centering
\includegraphics[width=4.5in]{\figpath/rotation_spherical}
\begin{minipage}[h]{2.0in}
\vspace{-4in}
\caption{Geometry of rotation when {\tt spherical\_in} $=1$.  We assume the 
star to be rotating about the $z$ axis with rotational frequency $\Omega$.}
\end{minipage}
\label{Fig:rotation in spherical}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Using Plane-Parallel Geometry}\label{Sec:Using Plane-Parallel Geometry}

### 2009-01-20 docs/flowchart/flowchart.tex
note about updating temperature in react_state
addressed most of the remaining marginpars

@@
 during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
+\MarginPar{Currently we do update the temperature.}
 
@@
 
-\MarginPar{what about a $\eta_{\rho h}$}
+\MarginPar{what about an $\eta_{\rho h}$?}
 \end{enumerate}
@@
 
-\MarginPar{don't we set $h_0 = \overline{h}$ somewhere around here?}
-
 \begin{enumerate}
@@
 
-\item {\bf if} ~ {\tt full\_rhoh0\_evolution = T} ~ {\bf then}\\
-{\bf React Base}$[(\rho h)_0^n,(\overline{\rho \Hnuc})^{(1)},(\overline{\rho\Hext})^n] \rightarrow [(\rho h)_0^{(1)}]$ \\
-{\bf else}\\
-$(\rho h)_0^{(1)} = (\overline{\rho h})^{(1)}$\\
+\item {\bf if} ~ {\tt full\_rhoh0\_evolution = T} ~ {\bf then}
+
+~~~ {\bf React Base}$[(\rho h)_0^n,(\overline{\rho \Hnuc})^{(1)},(\overline{\rho\Hext})^n] \rightarrow [(\rho h)_0^{(1)}]$ \\
+{\bf else}
+
+~~~ $(\rho h)_0^{(1)} = (\overline{\rho h})^{(1)}$\\
 {\bf end if}
-\item
-Define
+\item Define
 \begin{eqnarray}
@@
 \item Update the enthalpy.  The full enthalpy equation is 
-\MarginPar{we should say something about {\tt use\_tfromp}, and {\tt do\_eos\_h\_above\_cutoff}}
   \begin{equation}
@@
     We use 4th-order spatial averaging to get $(\rho h)_0$ on radial faces,
-    \MarginPar{in these notes we currently assume pred type 1, $(\rho h)'$}
 \begin{equation}
@@
 \item[spherical:]
+\MarginPar{Should we time-center $p_0$ now that we have split out {\tt advect\_base\_pres}?}
 
@@
 
-\MarginPar{We could time-center $p_0$ now that we have split out {\tt advect\_base\_pres}}
 \end{description}
@@
 \end{enumerate}
+
+If {\tt do\_eos\_h\_above\_cutoff=T} (which is the default setting), then if
+$\rho^{(2),\star} < {\tt base\_cutoff\_density}$, then we recompute enthalpy using
+\begin{equation}
+(\rho h)^{(1a),\star} = \rho^{(2),\star}h\left(\rho^{(2),\star},p_0^{n+1,\star},X_k^{(2),\star}\right).
+\end{equation}
 
@@
 \end{eqnarray}
-\item Update the temperature using the equation of state.
-
-The default is to do
+\item Update the temperature using the equation of state.  For plane-parallel,
+we update using
 \begin{equation}
@@
 \end{equation}
-but if {\tt use\_tfromp = T}, then we do
+but in spherical, we set {\tt use\_tfromp = T}, and we use
 \begin{equation}
@@
 \end{equation}
-This has the effect of completely decoupling the enthalpy equation from the 
+The latter has the effect of completely decoupling the enthalpy equation from the 
 evolution.
@@
 
+If {\tt do\_eos\_h\_above\_cutoff=T} (which is the default setting), then if
+$\rho^{(2)} < {\tt base\_cutoff\_density}$, then we recompute enthalpy using
+\begin{equation}
+(\rho h)^{(1a)} = \rho^{(2)}h\left(\rho^{(2)},p_0^{n+\myhalf},X_k^{(2)}\right).
+\end{equation}
+
 \item {\bf Correct Base}$[\rho_0^{(2a)}, (\nabla\cdot\etarho\er)^{\nph}] \rightarrow [\rho_0^{n+1}]$
@@
 \end{eqnarray}
-\item Update the temperature using the equation of state
-
-The default is to do
+\item Update the temperature using the equation of state.  For plane-parallel we
+update using
 \begin{equation}

### 2009-01-20 docs/flowchart/flowchart.tex
added a mini flow chart of variable names at the end of this section

@@
 
-\section{Outstanding Questions}
-
-\begin{enumerate}
-
-\item Why do we only update the temperature at the end of steps 4 and
-      8.  This means that at the top of the algorithm, the temperature
-      going into {\bf React State} in step 2 does not include the
-      effects of the previous reactions.
-
-      To be thermodynamically consistent, I think that we need to
-      update the temperature at the very end of step 9.  That way we
-      end the time-level update with $T^{n+1}$ instead of $T^{(2)}$,
-      and we go into the next sequence, step 2 with a valid temperature.
-
-\item Why is there no equivalent of $\etarho$ for enthalpy---e.g.\ an
-      $\eta_{\rho h}$?
-
-\end{enumerate}
+In summary:
+\begin{equation}
+\left.\begin{array}{ccccccccc}
+& \text{React State} & & \text{Advect State} & & \text{Diffuse Enthalpy} & & \text{React State} & \\
+\rho^n & & & \longrightarrow & & & & & \rho^{n+1} \\
+(\rho X)^n & \longrightarrow & (\rho X)^{(1)} & \longrightarrow & & & (\rho X)^{(2)} & \longrightarrow & (\rho X)^{n+1} \\
+(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(2a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\
+T^n & \longrightarrow & T^{(1)} & & & \longrightarrow & T^{(2)} & \longrightarrow & T^{n+1} \\
+\end{array}\right.\nonumber
+\end{equation}
+\begin{equation}
+\left.\begin{array}{ccccccccc}
+& \text{React Base} & & \text{Advect Base} & & \text{Correct Base} & & \text{React Base} & \\
+\rho_0^n & & & \longrightarrow & \rho_0^{(2a)} & \longrightarrow & & & \rho_0^{n+1} \\
+(\rho h)_0^n & \longrightarrow & (\rho h)_0^{(1)} & \longrightarrow & & & (\rho h)_0^{(2)} & \longrightarrow & (\rho h)_0^{n+1} \\
+p_0^n & & & \longrightarrow & & & & & p_0^{n+1} \\
+\end{array}\right.
+\end{equation}
+Note that this chart indicates, for example, that $\rho^n = \rho^{(1)}$, $\rho^{(2a)} = \rho^{(2)} = \rho^{n+1}$, etc.

### 2009-01-20 Source/react_state.f90
If use_tfromp=T, we pass temperature through react_state. If use_tfromp=F, we update temperature in react_state using the eos.

@@ -152,7 +152,8 @@ contains
     use burner_module
     use variables, only: rho_comp, spec_comp, temp_comp, rhoh_comp, trac_comp, ntrac
     use network, only: nspec
-    use probin_module, ONLY: do_burning, burning_cutoff_density, enthalpy_pred_type
+    use probin_module, ONLY: do_burning, burning_cutoff_density, enthalpy_pred_type, &
+         use_tfromp
     use pred_parameters
     use bl_constants_module, only: zero
     use eos_module
@@ -207,24 +208,33 @@ contains
           s_out(i,j,rhoh_comp) = s_in(i,j,rhoh_comp) &
                + dt * rho_Hnuc(i,j) + dt * rho_Hext(i,j)
 
-          ! update the temperature with the eos
-          den_eos(1)  = s_out(i,j,rho_comp)
-          h_eos(1)    = s_out(i,j,rhoh_comp)/s_out(i,j,rho_comp)
-          xn_eos(1,:) = s_out(i,j,spec_comp:spec_comp+nspec-1)/s_out(i,j,rho_comp)
-          temp_eos(1) = T_in
+          if (use_tfromp) then
+
+             ! pass temperature through
+             s_out(i,j,temp_comp) = s_in(i,j,temp_comp)
+
+          else
+
+             ! update the temperature with the eos
+             den_eos(1)  = s_out(i,j,rho_comp)
+             h_eos(1)    = s_out(i,j,rhoh_comp)/s_out(i,j,rho_comp)
+             xn_eos(1,:) = s_out(i,j,spec_comp:spec_comp+nspec-1)/s_out(i,j,rho_comp)
+             temp_eos(1) = T_in
              
-          call eos(eos_input_rh, den_eos, temp_eos, &
-                   npts, nspec, &
-                   xn_eos, &
-                   p_eos, h_eos, e_eos, &
-                   cv_eos, cp_eos, xne_eos, eta_eos, pele_eos, &
-                   dpdt_eos, dpdr_eos, dedt_eos, dedr_eos, &
-                   dpdX_eos, dhdX_eos, &
-                   gam1_eos, cs_eos, s_eos, &
-                   dsdt_eos, dsdr_eos, &
-                   do_diag)
-          
-          s_out(i,j,temp_comp) = temp_eos(1)
+             call eos(eos_input_rh, den_eos, temp_eos, &
+                      npts, nspec, &
+                      xn_eos, &
+                      p_eos, h_eos, e_eos, &
[…]

### 2009-01-20 docs/flowchart/flowchart.tex
If use_tfromp=T, we pass temperature through react_state. If use_tfromp=F, we update temperature in react_state using the eos.
working on "Updates" Section

@@
 
+\section{Algorithm Updates}
+
+There are several changes to the algorithm since it was last detailed
+in paper III.  The major changes are:
+
+\begin{itemize}
+\item In paper III we discretized the enthalpy evolution equation in
+full, non-perturbation form.  Since then we have discovered that 
+discretizing the enthalpy evolution in perturbational form leads to better
+numerical properties.  This is more like paper II.
+
+\item We split the update of the density and enthalpy in steps 4 and 8.
+Now, density is updated first, allowing us to compute new values of 
+$\psi$ that are then used in the enthalpy update.
+
+\item We split {\bf Advect Base} into separate {\bf Advect Base Density},
+{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.  This is because we use
+the results from {\bf Advect Base Density} to compute a time-centered $\etarho$ which 
+is used as an input to {\bf Advect Base Pressure}.  Also, we use the results from
+{\bf Advect Base Pressure} to compute a time-centered $\psi$ which is used as an input
+to {\bf Advect Base Enthalpy}.
+
+\item We discuss spherical geometries.
+
+\item We add thermal diffusion.
+
+\item We added the volume-discrepancy term to the velocity constraint
+equation.
+\end{itemize}
+
 %-----------------------------------------------------------------------------
@@
 during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
-\MarginPar{Currently we do update the temperature.}
+\MarginPar{Currently we do update the temperature if {\tt use\_tfromp}=F .}
 
@@
 the initialization in \S~\ref{Sec:Initialization}.
-
-There are several changes to the algorithm since it was last detailed
-in paper III.  The major changes are:
-\begin{itemize}
-\item In paper III we discretized the enthalpy evolution equation in
-full, non-perturbation form.  Since then we have discovered that 
-discretizing the enthalpy evolution in perturbational form leads to better
-numerical properties.  This is more like paper II.
-
-\item We split the update of the density and enthalpy in steps 4 and 8.
-Now, density is updated first, allowing us to compute new values of 
-$\psi$ that are then used in the enthalpy update.
-
-\item We discuss spherical geometries.
-
-\item We add thermal diffusion.
-
-\item We added the volume-discrepancy term to the velocity constraint
-equation.
-\end{itemize}
-%
 The advance of the state through a single timestep appears as:
@@
     \begin{equation}  L_\beta^\rho \phi =
-      D \left ( \beta_0^{\myhalf} {\bf V} \right) - \beta_0^{\myhalf} \left[ (S^{1}-\overline{S}^{1}) - (S^{0}-\overline{S}^{0}) \right] \enskip . 
+      D \left ( \beta_0^{\myhalf} {\bf V} \right) - \beta_0^{\myhalf} \left[ \left(S^{1}-\overline{S}^{1}\right) - \left(S^{0}-\overline{S}^{0}\right) \right] \enskip . 
     \end{equation}
@@
 \rho^n & & & \longrightarrow & & & & & \rho^{n+1} \\
-(\rho X)^n & \longrightarrow & (\rho X)^{(1)} & \longrightarrow & & & (\rho X)^{(2)} & \longrightarrow & (\rho X)^{n+1} \\
+(\rho X_k)^n & \longrightarrow & (\rho X_k)^{(1)} & \longrightarrow & & & (\rho X_k)^{(2)} & \longrightarrow & (\rho X_k)^{n+1} \\
 (\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(2a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\

### 2009-01-20 docs/flowchart/flowchart.tex
removed another marginpar

@@
 during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
-\MarginPar{Currently we do update the temperature if {\tt use\_tfromp}=F .}
+\MarginPar{Currently we do update the temperature if {\tt use\_tfromp}=F}
 
@@
 \item[spherical:]
-\MarginPar{Should we time-center $p_0$ now that we have split out {\tt advect\_base\_pres}?}
 
@@
   \end{eqnarray}
+
+\MarginPar{Should we time-center $p_0$ now that we have split out {\tt advect\_base\_pres}?}
 
@@
 finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
-\MarginPar{Why are $\beta$ and $\rho$ time centered here instead using the new time value?  Does this have something to do with how $\phi$ relates to a lagged pressure?}
 In this step, $D$ is a discrete second-order operator that approximates the 

### 2009-01-21 docs/flowchart/flowchart.tex
merged the CHANGES file into the flowchart "Algorithm Updates" section

@@
 \begin{itemize}
+
+\item We have added thermal diffusion with {\tt use\_thermal\_diffusion = T}.
+
+\item We discuss spherical geometries, in paricular:
+\begin{itemize}
+\item We are working on {\bf Advect Base}, {\bf React Base}, and 
+{\bf Correct Base} details.
+\item We are working on $\etarho$ details.
+\item We compute the full state T on cell-centers after the advection-diffusion step 
+from $T=T(\rho,X_k,p_0)$ by setting {\tt use\_tfromp = T} rather than $T=T(\rho,h,X_k)$.
+\item We have changed the discretization of $\Ub\cdot\nabla p_0$ in the enthalpy
+update to instead be $\nabla\cdot(\Ub p_0) - p_0\nabla\cdot\Ub$.
+\end{itemize}
+
 \item In paper III we discretized the enthalpy evolution equation in
-full, non-perturbation form.  Since then we have discovered that 
-discretizing the enthalpy evolution in perturbational form leads to better
-numerical properties.  This is more like paper II.
+full, non-perturbation form, $\rho h$.  Since then we have discovered that 
+discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
+leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
+This is more like paper II.
 
@@
 
-\item We discuss spherical geometries.
-
-\item We add thermal diffusion.
-
-\item We added the volume-discrepancy term to the velocity constraint
-equation.
+\item We added the volume-discrepancy term to the velocity constraint equation.
+
+\item We have turned off the evolution of $h$ above the atmosphere and instead
+compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff} = T.
+
+\item We have created a {\tt burning\_cutoff\_density}, where the burning does
+not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
+
+\item We have added the flag, {\tt full\_rhoh0\_evolution}, when false, sets
+$(\rho h)_0 = (\overline{\rho h})^n$ before {\bf Advect Base}.  {\bf React Base}
+will not be called.  Note that we do not have a {\bf Correct Base} or 
+{\bf Diffuse Base} for enthalpy.
+
+\item We have given the advection scheme a face-lift, with some nifty options like
+corner coupling and various PPM limiters.
+
+\item We now update the density and enthalpy as full states with full-state fluxes, 
+not as perturbational quantities.
+
 \end{itemize}
@@
 \renewcommand{\arraystretch}{1.0}
-
+Here are some charts demonstrating what the superscript notation above scalars implies:
+\begin{equation}
+\left.\begin{array}{ccccccccc}
+& \text{React State} & & \text{Advect State} & & \text{Diffuse Enthalpy} & & \text{React State} & \\
+\rho^n & & & \longrightarrow & & & & & \rho^{n+1} \\
+(\rho X_k)^n & \longrightarrow & (\rho X_k)^{(1)} & \longrightarrow & & & (\rho X_k)^{(2)} & \longrightarrow & (\rho X_k)^{n+1} \\
+(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(2a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\
+T^n & \longrightarrow & T^{(1)} & & & \longrightarrow & T^{(2)} & \longrightarrow & T^{n+1} \\
+\end{array}\right.\nonumber
+\end{equation}
+\begin{equation}
+\left.\begin{array}{ccccccccc}
+& \text{React Base} & & \text{Advect Base} & & \text{Correct Base} & & \text{React Base} & \\
+\rho_0^n & & & \longrightarrow & \rho_0^{(2a)} & \longrightarrow & & & \rho_0^{n+1} \\
+(\rho h)_0^n & \longrightarrow & (\rho h)_0^{(1)} & \longrightarrow & & & (\rho h)_0^{(2)} & \longrightarrow & (\rho h)_0^{n+1} \\
+p_0^n & & & \longrightarrow & & & & & p_0^{n+1} \\
+\end{array}\right.\nonumber
+\end{equation}
+Note that this chart indicates, for example, that $\rho^n = \rho^{(1)}$, $\rho^{(2a)} = \rho^{(2)} = \rho^{n+1}$, etc.\\ \\
 We make use of the following shorthand notations in outlining the algorithm:
@@
   \end{enumerate}
-
-In summary:
-\begin{equation}
-\left.\begin{array}{ccccccccc}
-& \text{React State} & & \text{Advect State} & & \text{Diffuse Enthalpy} & & \text{React State} & \\
-\rho^n & & & \longrightarrow & & & & & \rho^{n+1} \\
-(\rho X_k)^n & \longrightarrow & (\rho X_k)^{(1)} & \longrightarrow & & & (\rho X_k)^{(2)} & \longrightarrow & (\rho X_k)^{n+1} \\
-(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(2a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\
-T^n & \longrightarrow & T^{(1)} & & & \longrightarrow & T^{(2)} & \longrightarrow & T^{n+1} \\
-\end{array}\right.\nonumber
-\end{equation}
-\begin{equation}
-\left.\begin{array}{ccccccccc}
-& \text{React Base} & & \text{Advect Base} & & \text{Correct Base} & & \text{React Base} & \\
-\rho_0^n & & & \longrightarrow & \rho_0^{(2a)} & \longrightarrow & & & \rho_0^{n+1} \\
-(\rho h)_0^n & \longrightarrow & (\rho h)_0^{(1)} & \longrightarrow & & & (\rho h)_0^{(2)} & \longrightarrow & (\rho h)_0^{n+1} \\
-p_0^n & & & \longrightarrow & & & & & p_0^{n+1} \\
-\end{array}\right.
-\end{equation}
-Note that this chart indicates, for example, that $\rho^n = \rho^{(1)}$, $\rho^{(2a)} = \rho^{(2)} = \rho^{n+1}$, etc.

### 2009-01-21 docs/flowchart/flowchart.tex
proper reference to spherical averaging

@@
 
-\item We have added thermal diffusion with {\tt use\_thermal\_diffusion = T}.
+\item We have added thermal diffusion with {\tt use\_thermal\_diffusion = TRUE}.
 
@@
 \begin{itemize}
+\item We are still working out the fill, average, and interpolate stencils.
 \item We are working on {\bf Advect Base}, {\bf React Base}, and 
@@
 \item We are working on $\etarho$ details.
-\item We compute the full state T on cell-centers after the advection-diffusion step 
-from $T=T(\rho,X_k,p_0)$ by setting {\tt use\_tfromp = T} rather than $T=T(\rho,h,X_k)$.
+\item We compute the full state temperature on cell-centers after the 
+advection-diffusion step using $T=T(\rho,X_k,p_0)$ by setting 
+{\tt use\_tfromp = TRUE} rather than using $T=T(\rho,h,X_k)$.
 \item We have changed the discretization of $\Ub\cdot\nabla p_0$ in the enthalpy
-update to instead be $\nabla\cdot(\Ub p_0) - p_0\nabla\cdot\Ub$.
+update to $\nabla\cdot(\Ub p_0) - p_0\nabla\cdot\Ub$.
 \end{itemize}
@@
 \item We split the update of the density and enthalpy in steps 4 and 8.
-Now, density is updated first, allowing us to compute new values of 
+Now, density is updated first, allowing us to compute time-centered values of 
 $\psi$ that are then used in the enthalpy update.
@@
 \item We split {\bf Advect Base} into separate {\bf Advect Base Density},
-{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.  This is because we use
-the results from {\bf Advect Base Density} to compute a time-centered $\etarho$ which 
-is used as an input to {\bf Advect Base Pressure}.  Also, we use the results from
-{\bf Advect Base Pressure} to compute a time-centered $\psi$ which is used as an input
-to {\bf Advect Base Enthalpy}.
-
-\item We added the volume-discrepancy term to the velocity constraint equation.
+{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.
+\begin{itemize}
+\item First, we call {\bf Advect Base Density}, which computes a time-centered
+$\etarho$.
+\item Second, we call {\bf Advect Base Pressure}, which takes $\etarho$ as an
+input and computes a time-centered $\psi$.
+\item Third, we call {\bf Advect Base Enthalpy}, which takes $\psi$ as an input.
+\end{itemize}
+
+\item We added the volume-discrepancy term to the velocity constraint equation,
+controlled by the input parameter, $f$.
 
 \item We have turned off the evolution of $h$ above the atmosphere and instead
-compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff} = T.
+compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
 
@@
 
-\item We have added the flag, {\tt full\_rhoh0\_evolution}, when false, sets
-$(\rho h)_0 = (\overline{\rho h})^n$ before {\bf Advect Base}.  {\bf React Base}
-will not be called.  Note that we do not have a {\bf Correct Base} or 
-{\bf Diffuse Base} for enthalpy.
+\item We have added the flag, {\tt full\_rhoh0\_evolution}, which defaults to 
+{\tt TRUE} (note that we do not have a {\bf Correct Base} or 
+{\bf Diffuse Base} for enthalpy).  When {\tt FALSE}, we do not call {\bf React Base},
+but instead set $(\rho h)_0 = (\overline{\rho h})^n$ before {\bf Advect Base}.
 
@@
 \end{equation}
-here $\Omega_H$ is a layer at constant radius $r$.
-
-For spherical problems, see Paper IV for details.
+here $\Omega_H$ is a layer at constant radius $r$.  In plane-parallel, this is 
+as straightforward as averaging cell values at a particular height.  For spherical 
+problems, see the description in Section \ref{Sec:Avg}.
 

### 2009-01-21 docs/flowchart/flowchart.tex
fixed some typos, added more detail about mapping and where it happens in the algorithm

@@
 & \text{React State} & & \text{Advect State} & & \text{Diffuse Enthalpy} & & \text{React State} & \\
-\rho^n & & & \longrightarrow & & & & & \rho^{n+1} \\
+\rho^n & = & \rho^{(1)} & \longrightarrow & & & \rho^{(2)} & = & \rho^{n+1} \\
 (\rho X_k)^n & \longrightarrow & (\rho X_k)^{(1)} & \longrightarrow & & & (\rho X_k)^{(2)} & \longrightarrow & (\rho X_k)^{n+1} \\
-(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(2a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\
+(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(1a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\
 T^n & \longrightarrow & T^{(1)} & & & \longrightarrow & T^{(2)} & \longrightarrow & T^{n+1} \\
@@
 \end{equation}
-Note that this chart indicates, for example, that $\rho^n = \rho^{(1)}$, $\rho^{(2a)} = \rho^{(2)} = \rho^{n+1}$, etc.\\ \\
+\\ \\
 We make use of the following shorthand notations in outlining the algorithm:
@@
 \begin{equation}
-(\rho\Hnuc) = -\sum_k(\rho\omegadot_k)^{\outp} \; q_k \enskip.
+(\rho\Hnuc)^{\outp} = -\sum_k(\rho\omegadot_k)^{\outp} \; q_k \enskip.
 \end{equation}
@@
 (\rho h)^{\outp} = 
-(\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc) + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
+(\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc)^{\outp} + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
 \end{equation}
@@
   Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  We use the base
-  state density, $\rho_0$, to convert this to an edge state for $\rho$.
-  We use 4th-order spatial averaging to get $\rho_0$ on radial faces,
+  state density, $\rho_0$, to convert this to an edge state for $\rho$,
 \begin{equation}
 \rho^{(1),\nph,\star} = 
-\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip ,
+\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip .
 \label{eq:rhoXedgestate}
 \end{equation}
+  In plane-parallel, we use 4th-order spatial averaging to compute $\rho_0$ on 
+  radial faces.  In spherical, we first map $\rho_0$ to Cartesian cell-centers 
+  using the mapping in Section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
+  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.
 
@@
     already in {\bf React State}.  We use the base state enthalpy,
-    $(\rho h)_0$, to convert this to an edge state for $(\rho h)$.
-    We use 4th-order spatial averaging to get $(\rho h)_0$ on radial faces,
+    $(\rho h)_0$, to convert this to an edge state for $(\rho h)$,
+
 \begin{equation}
@@
 \end{equation}
-
-   We note that for spherical geometry, the edge state of $(\rho h)_0$
-   is computed by averaging $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$
-   to the edge separately, and then computing 
-   ${(\rho h)_0}_{r+\myhalf} = {\rho_0}_{r+\myhalf} {h_0}_{r+\myhalf}$.
+  In plane-parallel, we use 4th-order spatial averaging to compute $(\rho h)_0$
+  on radial faces.  In spherical, we first map $h_0 \equiv (\rho h)_0/\rho_0$ 
+  and $\rho_0$ to Cartesian cell-centers using the mapping in Section 
+  \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, use 2nd-order spatial 
+  averaging to compute $h_0$ and $\rho_0$ on all faces, and then multiply these 
+  terms to get $(\rho h)_0$ on all faces.
 
@@
 \begin{equation}
-\rho^{(1),\nph} = \rho^{'(1),\nph} + \frac{\rho_0^n + \rho_0^{(2a)}}{2}\enskip ,
-\end{equation}
+\rho^{(1),\nph} = \rho^{'(1),\nph} + \frac{\rho_0^n + \rho_0^{(2a)}}{2}\enskip .
+\end{equation}
+  In plane-parallel, we use 4th-order spatial averaging to compute $\rho_0$ on 
+  radial faces.  In spherical, we first map $\rho_0$ to Cartesian cell-centers 
+  using the mapping in Section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
+  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.
 
@@
 \end{equation}
-
-   We note that for spherical geometry, the edge state of $(\rho h)_0$
-   is computed by averaging $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$
-   to the edge separately, and then computing 
-   ${(\rho h)_0}_{r+\myhalf} = {\rho_0}_{r+\myhalf} {h_0}_{r+\myhalf}$.
-   \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(1a)}$ without
-     explicitly including the reaction terms,
+  In plane-parallel, we use 4th-order spatial averaging to compute $(\rho h)_0$
+  on radial faces.  In spherical, we first map $h_0 \equiv (\rho h)_0/\rho_0$ 
+  and $\rho_0$ to Cartesian cell-centers using the mapping in Section 
+  \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, use 2nd-order spatial 
+  averaging to compute $h_0$ and $\rho_0$ on all faces, and then multiply these 
+  terms to get $(\rho h)_0$ on all faces.
 

### 2009-01-21 docs/radial_base/radial_base.tex
make the radial base figure work whether we build the main document or build just the radial base section alone
proper reference to spherical averaging
fixed some typos, added more detail about mapping and where it happens in the algorithm

@@
 \subsection{1D Cell-Centered to Cartesian Cell-Centered}
+\label{Sec:1D Cell-Centered to Cartesian Cell-Centered}
 This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
@@
 \subsection{1D Edge-Centered to Cartesian Cell-Centered}
+\label{Sec:1D Edge-Centered to Cartesian Cell-Centered}
 This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
@@
 \subsection{1D Edge-Centered to Cartesian Edges}
+\label{Sec:1D Edge-Centered to Cartesian Edges}
 This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
@@
 velocities from the Cartesian cell-centers adjacent to the face.
-\section{\tt average}
-For the {\tt average} process, we first create a coarse 1D radial array 
+\section{Avg}\label{Sec:Avg}
+For the {\bf Avg} process, we first create a coarse 1D radial array 
 with $\Delta x = \Delta r$.  Then, we compute the radius of the Cartesian cell 
@@
 \begin{center}
-\epsffile{./radial_base/mapping.eps}
+\epsffile{\radbasefigpath/mapping.eps}
 \caption{\label{fig:mapping} The Cartesian grid and spherical base

### 2009-01-22 docs/flowchart/flowchart.tex
some cosmetic changes to the equations just because I'm overly particular

@@
 {\bf Diffuse Base} for enthalpy).  When {\tt FALSE}, we do not call {\bf React Base},
-but instead set $(\rho h)_0 = (\overline{\rho h})^n$ before {\bf Advect Base}.
+but instead set $(\rho h)_0 = \overline{(\rho h)^n}$ before {\bf Advect Base}.
 
@@
 \hline
-$\etarho$ & $\etarho \equiv (\overline{ \rho' \Ubt \cdot \er})$ & g~cm$^{-2}$~s$^{-1}$ \\
+$\etarho$ & $\etarho \equiv \overline{(\rho' \Ubt \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\
 \hline
@@
 \begin{equation}
-\overline{\phi}(r) 
+\overline{\phi(r)}
 = \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
@@
 
-\item {\bf React Base}$[(\rho h)_0^{\inp},(\overline{\rho\Hnuc})^{\inp},(\overline{\rho\Hext})^{\inp}] \rightarrow [(\rho h)_0^{\outp}]$\\
+\item {\bf React Base}$[(\rho h)_0^{\inp},\overline{(\rho\Hnuc)^{\inp}},\overline{(\rho\Hext)^{\inp}}] \rightarrow [(\rho h)_0^{\outp}]$\\
 is the process by which we update the base state enthalpy through $\dt/2$ in time using
 \begin{equation}
-\frac{\partial(\rho h)_0}{\partial t} = (\overline{\rho\Hnuc})^{\inp} + (\overline{\rho\Hext})^{\inp}\enskip ,
+\frac{\partial(\rho h)_0}{\partial t} = \overline{(\rho\Hnuc)^{\inp}} + \overline{(\rho\Hext)^{\inp}}\enskip ,
 \end{equation}
@@
 \begin{equation}
-(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} + \frac{\dt}{2}(\overline{\rho\Hnuc})_{j}^{\inp} + \frac{\dt}{2}(\overline{\rho\Hext})_{j}^{\inp} \enskip .
+(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} + \frac{\dt}{2}\overline{(\rho\Hnuc)_{j}^{\inp}} + \frac{\dt}{2}\overline{(\rho\Hext)_{j}^{\inp}} \enskip .
 \end{equation}
@@
 
-\item Define
-\begin{equation}
-\overline{S}^{\nph,\star} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip .
+\item Compute
+\begin{equation}
+\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip .
 \end{equation}
@@
 \begin{equation}
-\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S}^{\nph,\star} 
+\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S^{\nph,\star}}
 - \frac{1}{\gammabar^{n} p_0^{n}}
 \left\{ \psi^{\nmh} + 
-f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)}^n}{\Delta t^n} \right]\right\}
+f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right]\right\}
 \enskip .
@@
 \nablab \cdotb \left(\beta_0^n \uadvone\right) = 
-\beta_0^n \left\{ \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
+\beta_0^n \left\{ \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right)
 + \frac{f}{\gammabar^n p_0^n}
-\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)}^n}{\Delta t^n}\right]\right\} \enskip .
+\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right]\right\} \enskip .
 \end{equation}
@@
 
-~~~ {\bf React Base}$[(\rho h)_0^n,(\overline{\rho \Hnuc})^{(1)},(\overline{\rho\Hext})^n] \rightarrow [(\rho h)_0^{(1)}]$ \\
+~~~ {\bf React Base}$[(\rho h)_0^n,\overline{(\rho \Hnuc)^{(1)}},\overline{(\rho\Hext)^n}] \rightarrow [(\rho h)_0^{(1)}]$ \\
 {\bf else}
 
-~~~ $(\rho h)_0^{(1)} = (\overline{\rho h})^{(1)}$\\
+~~~ $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$\\
 {\bf end if}
@@
 \begin{eqnarray}
- \gammabar^{n+1,\star}    &=& {\rm{\bf Avg}} \left[ \Gamma_1(\rho^{n+1,\star}, p_0^{n+1,\star}, 
-                                                      X_k^{n+1,\star}) \right] \enskip , \\
+ \gammabar^{n+1,\star}    &=& {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{n+1,\star}, p_0^{n+1,\star}, 
+                                                      X_k^{n+1,\star}\right) \right] \enskip , \\
  {\beta   }_0^{n+1,\star}    &=& \beta   \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}\right) \enskip .
@@
 
-\item Define
-\begin{equation}
-\overline{S}^{\nph} = {\mathrm{\bf Avg}} (S^{\nph}) \enskip.
+\item Compute
+\begin{equation}
+\overline{S^{\nph}} = {\mathrm{\bf Avg}} (S^{\nph}) \enskip.
 \end{equation}
@@
 \frac{\partial w_0^{\nph}}{\partial r} = 
-\overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
-\left\{ \psi^{\nph,\star} + f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]\right\}
+\overline{S^{\nph,\star}} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
+\left\{ \psi^{\nph,\star} + f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\}
 \enskip ,
@@
 \begin{equation}
-\overline{p(\rho,h,X_k)}^{\nph,\star} = 
-\frac{\overline{p(\rho,h,X_k)}^n + \overline{p(\rho,h,X_k)}^{n+1,\star}}{2}
+\overline{p(\rho,h,X_k)^{\nph,\star}} = 
+\frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}
 \end{equation}
@@
 \nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) &=&
-\beta_0^{\nph,\star}\left\{\left(S^{\nph} - \overline{S}^{\nph}\right)
+\beta_0^{\nph,\star}\left\{\left(S^{\nph} - \overline{S^{\nph}}\right)
  + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}\cdot \right. \nonumber \\
-&& \left.\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]\right\} \enskip ,
+&& \left.\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\} \enskip ,
 \end{eqnarray}
@@
 
-\item {\bf React Base}$[(\rho h)_0^{(2)},(\overline{\rho \Hnuc})^{(2)},(\overline{\rho\Hext})^{(2)}] \rightarrow [(\rho h)_0^{n+1}]$
+\item {\bf React Base}$[(\rho h)_0^{(2)},\overline{(\rho \Hnuc)^{(2)}},\overline{(\rho\Hext)^{(2)}}] \rightarrow [(\rho h)_0^{n+1}]$
 
@@
 \begin{eqnarray}
- \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left[\Gamma_1(\rho^{n+1}, p_0^{n+1}, {X_k}^{n+1}) \right] \enskip , \\ 
+ \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1}, {X_k}^{n+1}\right) \right] \enskip , \\ 
  {\beta   }_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip .
@@
 
-\item Define
-\begin{equation}
-\overline{S}^{n+1} = {\mathrm{\bf Avg}} (S^{n+1}) \enskip.
+\item Compute
+\begin{equation}
+\overline{S^{n+1}} = {\mathrm{\bf Avg}} (S^{n+1}) \enskip.
 \end{equation}
@@
 \begin{equation}
-\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S}^{n+1} \right)
+\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S^{n+1}} \right)
 + \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]\right\}
+\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\}
 \enskip,
@@
  L_\beta^\rho \phi =
-   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph}\left\{ \left(S^{n+1}-\overline{S}^{n+1} \right)
+   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph}\left\{ \left(S^{n+1}-\overline{S^{n+1}} \right)
 - \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]\right\} \enskip, \label{Linear System for Multigrid}
+\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\} \enskip, \label{Linear System for Multigrid}
 \end{equation}
@@
 
-  \item Compute $\overline{S}^{0,\nu} = {\mathrm{\bf Avg}} (S^{0,\nu}).$
-
-  \item Compute $w_0^{\nu+1}$ as in step 1c using $\overline{S}^{0,\nu}$ and $\psi^{\nmh} = 0$.
+  \item Compute $\overline{S^{0,\nu}} = {\mathrm{\bf Avg}} (S^{0,\nu}).$
+
+  \item Compute $w_0^{\nu+1}$ as in step 1c using $\overline{S^{0,\nu}}$ and $\psi^{\nmh} = 0$.
 
   \item Project $\Ub^{0,\nu}$ using $\beta_0^{\initp}$ and 
-        $(S^{0,\nu} - \overline{S}^{0,\nu})$ as the source term.  
+        $(S^{0,\nu} - \overline{S^{0,\nu}})$ as the source term.  
         This yields $\Ub^{0,\nu+1}.$
@@
     \begin{equation}  L_\beta^\rho \phi =
-      D \left ( \beta_0^{\myhalf} {\bf V} \right) - \beta_0^{\myhalf} \left[ \left(S^{1}-\overline{S}^{1}\right) - \left(S^{0}-\overline{S}^{0}\right) \right] \enskip . 
+      D \left ( \beta_0^{\myhalf} {\bf V} \right) - \beta_0^{\myhalf} \left[ \left(S^{1}-\overline{S^{1}}\right) - \left(S^{0}-\overline{S^{0}}\right) \right] \enskip . 
     \end{equation}

### 2009-01-29 docs/spherical_basestate/basestate.tex
add spherical notes

\section{\label{app:gravity} Modifications for a Spherical Self-Gravitating Star}

In papers II and III, we calculated the hydrostatic expansion of the base state
in plane-parallel geometry under the assumption that the weight of the
material above (or below) any given fluid parcel does not change
during hydrostatic expansion.  This assumption holds when the
gravitational acceleration is independent of location.  Here we discuss the
modifications to the algorithm in paper III required to treat a spherical 
self-gravitating star.

\subsection{Calculation of $w_0$}

%In the low Mach number model, we allow the density and
%temperature perturbations to be large compared to the background
%state, but we require the pressure perturbation to be small.
%Therefore, if the star expands, we need to incorporate that expansion
%into the background state.  Furthermore, we need to know the velocity of the
%base state expansion for the evolution of the perturbed
%quantities.

We begin with equation (23) in paper III written in spherical coordinates:
\[
\frac{1}{r^2}\frac{\partial}{\partial r} \left (r^2 \beta_0 w_0 \right ) =
\beta_0 \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )\enskip.
\]
We expand the spatial derivative and divide through by $\beta_0$:
\begin{equation}
\label{eq:sphconstraint}
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) +
w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r} =
 \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t}  \enskip .
\end{equation}
Recalling from paper I that;
\[
\frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} = \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}\enskip,
\]
and moving this term to the right hand side,
we can express equation (\ref{eq:sphconstraint}) as:
\begin{eqnarray}
%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) +
%w_0 \frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} &=&
% \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right ) \nonumber \\
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) &=&
  \Sbar - \frac{1}{\gammabar p_0} \left( 
\frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r} \right)\enskip. \label{eq:w01}
\end{eqnarray}
We first solve a piece of this equation for $\ow = w_0 - \dw$:
\begin{eqnarray}
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \ow \right ) &=& \Sbar \enskip.
\end{eqnarray}
Then we can write
\begin{eqnarray}
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \dw \right ) &=&
  - \frac{1}{\gammabar p_0} \left(
\frac{\partial p_0}{\partial t} + (\ow + \dw)
\frac{\partial p_0}{\partial r} \right)\enskip.\label{eq:w01a}
\end{eqnarray}

Multiplying equation (\ref{eq:w01a}) through by $\gammabar p_0 $,
taking another derivative with respect to $r$, and switching the order
of temporal and spatial derivatives, we get:
\begin{equation}
\label{eq:sphconstraint_2}
\frac{\partial}{\partial r} \left [ \frac{\gammabar p_0 }{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right ]
= -\frac{\partial}{\partial t} \frac{\partial p_0}{\partial r}
  -\frac{\partial}{\partial r} \left ((\ow+\dw) \frac{\partial p_0}{\partial r} \right )\enskip.
\end{equation}

To solve for $\dw$ we will need to substitute for the derivatives of $p_0.$
To do so we start with the equation of hydrostatic equilibrium in a spherical geometry:
\[
\frac{\partial p_0}{\partial r} = -\rho_0 g \, ; \quad
g = \frac{G m_\mathrm{encl}}{r^2}\enskip,
\]
where $m_\mathrm{encl}(r)$ is the mass enclosed at radius $r$ and $G$ is the
gravitational constant.  
Using this, we can then write equation (\ref{eq:sphconstraint_2}) as:
\begin{eqnarray}
\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=&
 \frac{\partial}{\partial t} \left (\rho_0 g \right )
+\frac{\partial}{\partial r} \left (w_0 \rho_0 g \right) \nonumber \\
&=&
  g \left[ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right]
+ \rho_0 \left( \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} \right) \enskip.\nonumber\\
&&
\label{eq:sphconstraint_3}
\end{eqnarray}
The mass enclosed inside any radius, $r$, is $m_\mathrm{encl}(r) = 4 \pi
\int_0^r \rhozero(s) s^2 ds$, or alternately, $\partial m_\mathrm{encl}/\partial r = 4\pi r^2 \rho_0$.  
The Lagrangian derivative of the enclosed mass is then:
\begin{eqnarray}
\frac{D_0 m_\mathrm{encl}}{D t} &=& \frac{\partial m_\mathrm{encl}}{\partial t} + w_0 \frac{\partial m_\mathrm{encl}}{\partial r} \nonumber \\
 &=& 4\pi \left( \frac{\partial}{\partial t} \int_0^r \rhozero(s) s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
&=& 4\pi \left( \int_0^r \frac{\partial \rho_0}{\partial t} s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
&=& 4\pi \left\{ -\int_0^r \left [ \frac{1}{s^2}\frac{\partial(s^2 \rho_0 w_0)}{\partial s} + \frac{1}{s^2}\frac{\partial (s^2 \etarho)}{\partial s} \right ]s^2 ds + w_0 r^2 \rho_0 \right\} \nonumber \\
&=&  \left . 4\pi \left(-s^2 \rho_0 w_0 \right |_0^r - \left . s^2 \etarho \right |_0^r + w_0 r^2 \rho_0 \right) \nonumber \\
&=& -4\pi r^2 \etarho \label{eq:dmdt}
\end{eqnarray}
where we used the spherical form of equation (29) in paper III, 
\begin{equation}
\label{eq:sph_continuity}
\frac{\partial \rhozero}{\partial t} + \frac{1}{r^2} \frac{\partial (r^2 \rho_0 w_0)}{\partial r} 
+ \frac{1}{r^2} \frac{\partial (r^2 \etarho )}{\partial r}  = 0 \enskip ,
\end{equation}
to eliminate $\partial \rho_0/\partial t$.  We note that in the
absence of any mixing, $\etarho=0$, and  $D_0 m_\mathrm{encl}/Dt = 0.$
%which is another way of saying that the mass inside a fluid element
%does not change during expansion.  
Equation (\ref{eq:dmdt})
allows us to write the Lagrangian change in the gravitational
acceleration as:
\begin{eqnarray}
\frac{D_0 g}{D t} = \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} &=& \frac{D_0}{Dt}\left(\frac{G m_\mathrm{encl}}{r^2}\right) \nonumber \\
&=& G m_\mathrm{encl}\frac{D_0}{D t}\left(\frac{1}{r^2}\right) + \frac{G}{r^2}\frac{D_0 m_\mathrm{encl}}{Dt} \nonumber \\
&=& -\frac{2 w_0 G m_\mathrm{encl}}{r^3} - 4 \pi G \etarho \nonumber \\
&=& -\frac{2 w_0 g}{r} - 4 \pi G \etarho \enskip .
\end{eqnarray}

Putting it all together, equation (\ref{eq:sphconstraint_3}) becomes:
\begin{equation}
 \frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] 
= 
  g \left [ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right ]
+ \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \label{eq:B5} \enskip.
\end{equation}
Finally, we can use equation (\ref{eq:sph_continuity})
to write equation (\ref{eq:B5}) as:
\begin{eqnarray}
\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] && \nonumber \\
&=&
 g \left [ -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 w_0 \rhozero)
           -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 \etarho)
           +              \frac{\partial}{\partial r} (w_0 \rhozero) \right ] + \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \nonumber \\
&=& 
- \frac{g}{r^2} \frac{\partial (r^2 \etarho)}{\partial r} - \frac{4 (\ow + \dw) \rho_0 g}{r} 
- 4 \pi G \rhozero \etarho \label{eq:B7} \enskip .
\end{eqnarray}

\noindent We discretize this elliptic equation in the radial dimension as:
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i} -
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i-1} \right\}
+ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{i-\half} \nonumber\\
&=& 
- \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
\left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
   - \left[ \frac{4 \ow \rho_0 g}{r} \right]_{i-\half}
   - \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip , \nonumber
\end{eqnarray}
where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
Then, using hydrostatic equilibrium, we expand this to 
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
 \left( \frac{\gammabar p_0}{r^2}\right)_{i  } \frac{[ (r^2 \dw)_{i+\half} - (r^2 \dw)_{i-\half} ]}{\Delta r}
-\left( \frac{\gammabar p_0}{r^2}\right)_{i-1} \frac{[ (r^2 \dw)_{i-\half} - (r^2 \dw)_{i-\thalf}]}{\Delta r} \right\} \nonumber \\
&&  - \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \dw)_{i-\half}  \nonumber \\
&=&   \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \ow)_{i-\half}
 - \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
  \left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
- \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip.
\end{eqnarray} \MarginPar{note in the RHS, we do a difference of cell-centered $\eta$'s, but $\eta$ is defined on edges -- what should we do?}

If we write this in matrix form, so that:
\begin{equation}
A_i (r^2 \dw)_{i-\thalf} + B_i (r^2 \dw)_{i-\half} + C_i (r^2 \dw)_{i+\half} = F_i\enskip,
\end{equation}
then:
\begin{eqnarray}
A_i &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{i-1}  \enskip, \nonumber  \\
B_i &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar p_0}{r^2}\right)_{i  }  
               +\left( \frac{\gammabar p_0}{r^2}\right)_{i-1} \right] 
              - \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} \nonumber \\
C_i &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{i}  \enskip, \nonumber  \\
F_i &=&  \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \ow)_{i-\half}
 - \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
\left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
- \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip \nonumber .
\end{eqnarray}
Here,
\begin{equation}
g_{j-\half} = \frac{G}{r_{j-\half}^2} \sum_{k=1}^{j} \frac{4}{3} \pi  (r_{k-\half}^3 - r_{k-\thalf}^3) \; {\rho_0}_{k-1} \enskip.
\end{equation}

We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
corresponds to $i=1,$ by setting:
\begin{eqnarray}
A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
B_1 &=& 1\enskip. \nonumber
\end{eqnarray}
We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $i=N,$
by setting:
\begin{eqnarray}
A_N &=& -1\enskip, \nonumber \\
B_N &=&  1\enskip, \nonumber \\
C_N = F_N &=& 0\enskip. \nonumber
\end{eqnarray}
\subsection{Calculation of $\psi$}

Rewriting Equation (\ref{eq:w01}), we can define
\begin{eqnarray}
\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
\end{eqnarray}
We discretize this locally as:
\begin{eqnarray}
\psi_j^{\nph,\star} &=& \gammabar^{(1)} \frac{\left( p_0^{n} + p_0^{n+1,\star} \right)_j}{2}
\left \{ \overline{S^{\nph,\star}_j} - 
 \frac{1}{r_j^2} \left [ (r^2 {w_0}^{\nph,\star})_{j+\myhalf} -
          (r^2 {w_0}^{\nph,\star})_{j-\myhalf} \right ] \right \} \enskip , \nonumber \\
&& \\
\psi_j^{\nph} &=& \gammabar^{n+1,\star} \frac{\left( p_0^{n} + p_0^{n+1} \right)_j}{2}
\left \{ \overline{S^{\nph}_j} - 
 \frac{1}{r_j^2} \left [ (r^2 {w_0}^{\nph})_{j+\myhalf} -
          (r^2 {w_0}^{\nph})_{j-\myhalf} \right ] \right \} \enskip .
\end{eqnarray}
\MarginPar{Why is $\gammabar$ not time centered in the $\psi_j^{\nph}$ equation?}

\subsection{{\bf Advect Base}}

\begin{enumerate}

\item {\bf Density Update:}

The base state density update now includes the area factors in the divergences:
\begin{eqnarray}
\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp}
 - \frac{1}{r_j^2} \frac{\dt}{\dr} 
 \left [ \left( {{r^2 \rho}_0}^{\inp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
          \left( {{r^2 \rho}_0}^{\inp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right]   \enskip. 
\end{eqnarray}
As before, the interface states are found via the procedure described in
paper II, Appendix A.

\item {\bf Pressure Update:}
\newline
We use a predictor-corrector formulation for the pressure update 
in order to improve the overall accuracy.  Beginning with equation~(\ref{eq:w01}),
we can write
\begin{eqnarray*}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
 \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) - \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
\end{eqnarray*}
We time-center the left hand side:
\[
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
   \frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} - p_0^{\inp}}{\dt}\enskip.
\]
We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
\begin{equation}
p_0^{\star} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
\end{equation}
with:
\begin{equation}
f = \frac{\dt}{2} \left [ \overline{S^{\inp}} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} \left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
\end{equation}
where the $(\partial p_0^\inp / \partial r)$ is computed by
interpolating $p_0^\inp$ from cell-centers to edges and then
differencing.

Now we begin the corrector step.  Using $p_0^\star$, we can compute a more accurate approximation
to equation~(\ref{eq:frhs}),
\begin{equation}
\tilde{f} = \frac{\dt}{2} \left [ \overline{S^{\inp}} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} \tilde{p_0}} w_0^{\inp} \left ( \frac{\partial \tilde{p_0}}{\partial r} \right ) \right ] \enskip ,
\end{equation}
where $\tilde{p} = (p_0^\inp + p_0^\star)/2$, and the difference is again computed
by first putting $\tilde{p}$ on edges.

This allows us to complete the corrector for the pressure update:
\begin{equation}
p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} \tilde{f}\right)}{\left(1 - \gammabar^{\inp} \tilde{f}\right)}\enskip .
\end{equation}
We note that since we do not time-center $\gammabar$, this is not
strictly second-order accurate.  However, as $\gammabar$ varies slowly
in time, we expect the effect to be small.  This is something that we 
defer until simulations indicate it is warranted. \MarginPar{new}

\end{enumerate}

\subsection{Correct Base}

The base state density correction now includes the area factors in the divergences,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp}  - \frac{1}{r_j^2} \frac{\dt}{\dr} \left [ \left(r^2 \etarho\right)_{j+\myhalf} - \left(r^2 \etarho\right)_{j-\myhalf} \right ] \enskip .
\end{equation}

\subsection{One-dimensional Results}

To test the spherical base state expansions, we inject heat at a
steady rate into a one-dimensional white dwarf model.  This is similar
to the first test in paper II, except now in spherical coordinates.
As in that test, the compressible method with which we compare the low Mach number method 
is the FLASH code's implementation of the
piecewise-parabolic method (PPM) in a one-dimensional spherical geometry.  
The initial conditions for the white dwarf are those described in
Section 4.1 of paper III for the central region.

%A simple initial model for a white dwarf was constructed by specifying
%a central density of $2.6\times 10^9~\gcc$, a central temperature of
%$7\times 10^8$~K, and a composition of 30\% carbon and 70\% oxygen,
%and integrating the equation of hydrostatic equilibrium outward (using
%spherically symmetric self-gravity) while constraining the entropy to
%be constant.  Once the temperature of the model falls to $10^7$~K, it
%is held constant---this happens only at the very outer region of the
%star.  Together with the equation of state, this completely determines
%the density, temperature, and pressure structure of the star.

In the expansion of a plane-parallel atmosphere, heating at a
height $r$ above the base does not affect the pressure or density 
below that height.  By contrast, in a spherical symmetric
self-gravitating star, heating at a radius $r$ will lead to a pressure
and density decrease at the center in addition to the expansion of the
outer layers (see Schwarzchild \& Harm, 1965, ApJ, 146, 855).

We apply a heating function of the form:
\begin{equation}
\Hext = H_0 \exp \left [-(r-r_0)^2 / W^2 \right ] \enskip ,
\end{equation}
with $r_0 = 4\times 10^7$~cm, $W = 10^7$~cm, and $H_0 = 1\times
10^{16}$~erg~g$^{-1}$~s$^{-1}$. This is the same functional form as used
in the first test of paper II, but with a lower amplitude.  Still, this
heating rate is far higher than what is expected during the convective
phase of Type Ia SNe.  The heating term is added to the enthalpy
equation in the low Mach number equations in the same fashion as
described in paper II.  In this test, we do not consider reactions.
Since this is a one-dimensional test all perturbational quantities,
as well as $\Ubt,$ are zero, so we are directly testing the computation of 
$w_0$ as and the base state update as described in
the {\bf Advect Base} procedure defined above.  
Both the PPM and low Mach calculation use 768 zones in a domain $5\times
10^8$~cm high.

Figure~\ref{fig:spherical768} shows the structure of the star after
heating for 10~s.  The gray line is the initial star before any
heating.  
%The solid black line is the PPM result and the dotted and
%dashed lines are the low Mach number model with advective CFL numbers
%of $0.5$ and $0.1$ respectively (the PPM calculation used a CFL number
%of $0.5$).  
We see that the compressible and low Mach number models
agree extremely well.  Both capture the decrease in the density and
pressure at the center of the star and the considerable expansion in
radius.  Only at the surface of the star do the temperatures differ slightly.
In all calculations, we set the minimum temperature to $5\times
10^6$~K.  The PPM simulation required 13488 steps and the low Mach
(CFL $=0.5$) calculation needed 203.  Over the course of the
simulation, the Mach number of the flow remained less than $0.35$, with the
maximum Mach number occurring at the surface of the star.  This Mach
number pushes the limits of validity of the low Mach number model;  a
smaller perturbation amplitude would result in a smaller Mach number.

Future improvements to the overall spherical base-state adjustment
algorithm will address the expansion in a simulation where the medium
outside the star is not brought down to arbitrarily low densities, but
instead a ``cutoff density'' is applied, as in the case of the
plane-parallel results presented in this paper. However, we expect
the changes to the overall method shown here to be small. \MarginPar{new}

{\color{red} Add a figure showing that we retain the correct solution 
even when we place higher density material outside the star.}

\clearpage

\begin{figure*}
\begin{center}
\epsffile{\sphericalfigpath/spherical_adjust_768.eps}
\end{center}
\caption{\label{fig:spherical768} Hydrostatic adjustment of a
spherically symmetric white dwarf with self-gravity.  The gray line
represents the initial model;  all other lines are after 10~s of heating.
The solid black line is the fully compressible solution, the dotted line is the
low Mach number solution with a CFL number of 0.5,  and the dashed line is
the low Mach number solution with a CFL number of 0.1.  All
simulations used 768 equally spaced zones.  We see excellent agreement
between the compressible and low Mach number models.  The only
differences appear at the top of the atmosphere, where the outer
boundary condition can influence the results.}
\end{figure*}

\end{document}

### 2009-01-29 docs/flowchart/flowchart.tex
added marginpar about not needed div etarho if we replace correct base with rho_0 = Avg(rho)
updated to reflect how we compute (rhoh)_0^{(1)} in the code
addressed some marginpars, put in section labels with proper referencing in the flowchart

@@
 
+\item We update temperature at the end of {\bf React State}.
+
 \end{itemize}
@@
 compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
-to the equation of state.  We note that while temperature is evolved
-when solving these equations, we do not keep the final temperature,
-nor do we use it to compute the final change in enthalpy.  Therefore,
-$T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged
+to the equation of state.  We update temperature using
+$T^{\outp} = T(\rho^\outp,h^\outp,X_k^\outp)$ if {\tt use\_tfromp = FALSE} or 
+$T^{\outp} = T(\rho^\outp,p_0,X_k^\outp)$  if {\tt use\_tfromp = TRUE}.  
+Note that the density remains unchanged
 during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
-\MarginPar{Currently we do update the temperature if {\tt use\_tfromp}=F}
 
@@
 
-\item[spherical:]
+\item[spherical:] See Section \ref{Sec: Advect Base Spherical} for working notes.
 
@@
 \item[spherical:] 
+\MarginPar{Need to write this up somewhere}
 
@@
 
+\MarginPar{Instead of calling Correct Base, if we set $\rho_0 = \overline\rho$, 
+we don't even need $\nabla\cdot\etarho\er$.}
 \item {\bf Correct Base}$[\rho_0^{\inp}, (\nabla\cdot\etarho\er)^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
@@
 
-\item[spherical:] See Paper IV for working notes until we finalize this.
+\item[spherical:] See Section \ref{Sec:Spherical w0 Calculation} for working notes.
 
@@
 
-~~~ $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$\\
+~~~ $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$ in plane-parallel
+
+~~~ $(\rho h)_0^{(1)} = (\rho h)_0^{n} - {\rm{\bf Avg}}\left[(\rho h)_0^{n} - (\rho h)^{(1)}\right] $ in spherical. \\
 {\bf end if}
@@
 \begin{eqnarray}
-\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left[ \Gamma_1(\rho^{(1)}, p_0^{n}, X_k^{(1)}) \right] \enskip , \\
+\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, X_k^{(1)}\right) \right] \enskip , \\
  {\beta   }_0^{(1)}    &=& \beta   \left(\rho_0^{n}, p_0^{n}, \gammabar^{(1)}\right) \enskip .
@@
 
-\item[spherical:] See {\tt MAESTRO/docs/eta\_notes} for working notes until we finalize 
-the details.
+\item[spherical:] See Section \ref{Sec:eta Spherical} for working notes.
 
@@
   \end{eqnarray}
-  \MarginPar{we should also show the update in the case that $(\rho h)'$ is the flux}
 
@@
 
-\MarginPar{Should we time-center $p_0$ now that we have split out {\tt advect\_base\_pres}?}
-
 \end{description}
@@
 
-\item[plane-parallel:] ~
+\item[plane-parallel:]
 
@@
 
-\item[spherical:]~
+\item[spherical:] See Section \ref{Sec:eta Spherical} for working notes.
 

### 2009-01-29 docs/flowchart/flowchart.tex
started a 'Future Considerations' section

@@
 %-----------------------------------------------------------------------------
+% Future Considerations
+%-----------------------------------------------------------------------------
+
+\section{Future Considerations}
+
+\begin{itemize}
+
+\item Replacing {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
+plane-parallel and $\rho_0 = \rho_0 - {\rm{\bf Avg}}(\rho_0 - \rho)$ in spherical.  
+This eliminates the need for $\nabla\cdot\etarho$.
+
+\item After {\bf React State}, setting $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
+plane-parallel and $(\rho h)_0 = (\rho h)_0 - {\rm{\bf Avg}}[(\rho h)_0 - (\rho h)]$ 
+in spherical by making {\tt full\_rhoh0\_evolution = FALSE} by default.  This means
+that {\bf Correct Base} no longer has to exist, and the intermediate states, 
+$(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
+{\bf Advect Base Enthalpy} advances $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ directly.
+
+\item Eliminating {\bf Advect Base Pressure} entirely, and instead enforcing HSE 
+directly after {\bf Correct Base}.  {\bf Correct Base} should occur directly after 
+creating $\etarho$, there is no need to delay the call as is currently being done.
+
+\end{itemize}
+
+%-----------------------------------------------------------------------------
 % Notation
@@
 
-\MarginPar{Instead of calling Correct Base, if we set $\rho_0 = \overline\rho$, 
-we don't even need $\nabla\cdot\etarho\er$.}
 \item {\bf Correct Base}$[\rho_0^{\inp}, (\nabla\cdot\etarho\er)^{\inp}] \rightarrow [\rho_0^{\outp}]$\\

### 2009-01-31 docs/FAQ/faq.tex
some useful documentation that doesn't seem to fit in anywhere else

\begin{enumerate}

\item {\em  How do we turn off all the initial projections to look at the
   initial velocity field as specified in initdata, instead of as
   modified by the velocity constraint?} 
%
\begin{verbatim}
    max_step  = 1
    init_iter = 0
    init_divu_iter = 0
    do_initial_projection = 0
\end{verbatim}

\item {\em How can we dump out a variable to a plotfile from any point in the
   code?} 
%
\begin{verbatim}
    use fabio_module

    call fabio_ml_multifab_write_d(uold,mla%mba%rr(:,1),"a_uold")
    call fabio_ml_multifab_write_d(umac(:,1),mla%mba%rr(:,1),"a_umacx")
\end{verbatim}

\item {\em Does it matter if we use a Fortran 90 compiler instead of a 
   Fortran 95 compiler?}

   On some machines it does---Fortran 95 specifies that any local
   allocated arrays are automatically deallocated when a subroutine
   ends.  Fortran 90 does not do this.  On IBM machines, for instance,
   using {\tt xlf90} instead of {\tt xlf95} for MAESTRO will produce
   memory leaks.

\end{enumerate}

### 2009-02-02 docs/flowchart/flowchart.tex
fixed some typos, addressed more marginpars
more future considerations

@@
 
+\item In {\bf Step 4} and {\bf Step 8}, move {\bf Correct Base} to immediately
+before {\bf Advect Base Pressure}.
+
+\item Should we use a predictor-corrector for updating the full-state density?
+In particular, after {\bf Correct Base}, do a full-state density advance and
+{\bf Correct Base} again?
+
 \item After {\bf React State}, setting $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
@@
 {\bf Advect Base Enthalpy} advances $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ directly.
+This also means we don't have to deal with $\eta_{\rho h}$ or a {\bf Diffuse Base}.
 
@@
 
-\item[spherical:] See Section \ref{Sec: Advect Base Spherical} for working notes.
+\item[spherical:] See Section \ref{Sec:Advect Base Spherical} for working notes.
 
@@
 
-\MarginPar{what about an $\eta_{\rho h}$?}
 \end{enumerate}

### 2009-02-03 docs/flowchart/flowchart.tex
clarified some of the future considerations

--- 2009-01-29 docs/flowchart/flowchart.tex
@@
 %-----------------------------------------------------------------------------
+% Future Considerations
+%-----------------------------------------------------------------------------
+
+\section{Future Considerations}
+
+\begin{itemize}
+
+\item Should we eeplace {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
+plane-parallel and $\rho_0 = \rho_0 - {\rm{\bf Avg}}(\rho_0 - \rho)$ in spherical?
+This eliminates the need for $\nabla\cdot\etarho$.
+
+\item In {\bf Step 4} and {\bf Step 8}, should we move {\bf Correct Base} so it 
+occurs immediately before {\bf Advect Base Pressure}?  Since the full
+enthalpy advance ``feels'' the effects of the time-centered $\etarho$, we should
+rely on a $\rho_0^{n+1}$ that also feels the effects.
+
+\item Should we use a predictor-corrector for updating the full-state density?
+Specifically, after calling {\bf Correct Base}, should we do a full-state density 
+advance and {\bf Correct Base} using the more accurate estimate of $\rho_0^{n+1}$?
+
+\item After {\bf React State}, should we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
+plane-parallel and $(\rho h)_0 = (\rho h)_0 - {\rm{\bf Avg}}[(\rho h)_0 - (\rho h)]$ 
+in spherical?  This can currently be done by making {\tt full\_rhoh0\_evolution = FALSE}.
+This means that {\bf React Base} no longer has to exist, and the intermediate states, 
+$(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
+{\bf Advect Base Enthalpy} will advance $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ 
+directly.  This also means we don't have to deal with $\eta_{\rho h}$ or a 
+{\bf Diffuse Base}.
+
+\item Should we eliminate {\bf Advect Base Pressure} entirely, and instead directly
+enforce HSE after {\bf Correct Base}?  {\bf Advect Base Pressure} was just a fancy
+way of enforcing HSE anyway, so why not just compute the new pressure directly?
+
+\end{itemize}
+
+%-----------------------------------------------------------------------------
 % Notation

### 2009-02-03 docs/flowchart/flowchart.tex
cleaning up the Initialization Step to 'look' more like the code

@@
 Use the equation of state to determine $(\rho h)^{\initp}$.  Set
-$w_0^1 = 0$ as an initial approximation.  Compute $\beta_0^{\initp}$ as a function of 
+$w_0^0 = 0$ as an initial approximation.  Compute $\beta_0^{\initp}$ as a function of 
 the initial data.  Then, project $\Ub^{\initp}$ using $\beta_0^{\initp}$ and 
-$S = \rho\Hext$, giving $\Ub^{0,1}$.  The next part of the initialization process 
+$S = \rho\Hext$, giving $\Ub^{0,0}$.  The next part of the initialization process 
 proceeds as follows.
@@
 
-  \item Estimate $\Delta t^\nu$ using $\Ub^{0,\nu}$ and $w_0^\nu.$
+  \item Estimate $\Delta t^\nu$ using $\Ub^{0,\nu-1}$ and $w_0^{\nu-1}.$
 
@@
 
-  \item Compute $w_0^{\nu+1}$ as in step 1c using $\overline{S^{0,\nu}}$ and $\psi^{\nmh} = 0$.
-
-  \item Project $\Ub^{0,\nu}$ using $\beta_0^{\initp}$ and 
+  \item Compute $w_0^{\nu}$ as in step 1c using $\overline{S^{0,\nu}}$ and $\psi=0$.
+
+  \item Project $\Ub^{0,\nu-1}$ using $\beta_0^{\initp}$ and 
         $(S^{0,\nu} - \overline{S^{0,\nu}})$ as the source term.  
-        This yields $\Ub^{0,\nu+1}.$
+        This yields $\Ub^{0,\nu}.$
 
@@
 
-  Define $S^0 = S^{0,N_{\rm iters}^S}$, $w_0^{-\myhalf} = w_0^{N_{\rm iters}^S+1}$, 
-$\dt^0 = \Delta t^{N_{\rm iters}^S},$ and $\Ub^0 = \Ub^{0,N_{\rm iters}^S+1}.$
+  Define $S^0 = S^{0,N_{\rm iters}^S}$, $w_0^{-\myhalf} = w_0^{N_{\rm iters}^S}$, 
+$\dt^0 = \Delta t^{N_{\rm iters}^S},$ and $\Ub^0 = \Ub^{0,N_{\rm iters}^S}.$
 
@@
 $\pi,$  and an approximation to the divergence constraint at the end of the
-first time step.  As initial approximations, set $S^{1,0} = S^0$,
-$\etarho^{-\myhalf} = 0$, $\psi^{-\myhalf} = 0$, and $\pi^{-\myhalf} = 0.$
+first time step.  As initial approximations, set $\etarho^{-\myhalf}=0$, $\psi^{-\myhalf}=0$,
+$S^{1,0}=S^0$, and $\pi^{-\myhalf}=0.$
 \begin{enumerate}
@@
 
-\item {\bf Do} {$\nu = 0,...,N_{\rm iters}^{\pi}-1$.}
+\item {\bf Do} {$\nu = 1,...,N_{\rm iters}^{\pi}$.}
 
@@
   \item Perform steps 1--11 as described above, using 
-    $S^{\myhalf,\star} = (S^0 + S^{1,\nu})/2$ in step 1 as described.
+    $S^{\myhalf,\star} = (S^0 + S^{1,\nu-1})/2$ in step 1 as described.
     The only other difference in the time advancement is that in step 11
@@
       we discard,  and new values for $\etarho^{\myhalf}$ (step 8), $\psi^{\myhalf}$ (step 8), 
-      $S^{1,\nu+1}$ (step 10), and $\pi^{\myhalf}$ (step 11), which we keep.
+      $S^{1,\nu}$ (step 10), and $\pi^{\myhalf}$ (step 11), which we keep.
     \item Set $\pi^{-\myhalf} = \pi^{\myhalf}$, $\etarho^{-\myhalf} = \etarho^{\myhalf}$,

### 2009-02-03 docs/flowchart/flowchart.tex
more Initialization cleanup

@@
 
-\section{Time Advancement Algorithm}
+\section{Time Advancement Algorithm}\label{Sec:Time Advancement Algorithm}
 
@@
 This is done through a second iterative procedure which follows the
-time advancement algorithm as described in steps 1-11 in \S~ref{Sec:Alg}.  
+time advancement algorithm as described in steps 1-11 in 
+\S~\ref{Sec:Time Advancement Algorithm}.  
 The user specifies the number of iterations, 
@@
 
-Start with initial data $X_k^{\initp}, \rho^{\initp},$ $T^{\initp},$ an 
-initial base state, and an initial guess for the velocity, $\Ub^{\initp}.$
-Use the equation of state to determine $(\rho h)^{\initp}$.  Set
-$w_0^0 = 0$ as an initial approximation.  Compute $\beta_0^{\initp}$ as a function of 
+First, we need to construct approximations to $S^0, w_0^{-\myhalf}, \Delta t^0$, 
+and $\Ub^0$.  Start with initial data $X_k^{\initp}, \rho^{\initp},$ $T^{\initp},$ an 
+initial base state, and an initial guess for the velocity, $\Ub^{\initp}$.
+Set $w_0^0 = 0$ as an initial approximation.  Use the equation of state to 
+determine $(\rho h)^{\initp}$.  Compute $\beta_0^{\initp}$ as a function of 
 the initial data.  Then, project $\Ub^{\initp}$ using $\beta_0^{\initp}$ and 
@@
 
-Next, we need to construct an approximation to the time-centered perturbational pressure,
-$\pi,$  and an approximation to the divergence constraint at the end of the
-first time step.  As initial approximations, set $\etarho^{-\myhalf}=0$, $\psi^{-\myhalf}=0$,
-$S^{1,0}=S^0$, and $\pi^{-\myhalf}=0.$
+Next, we need to construct approximations to $\etarho^{-\myhalf}, \psi^{-\myhalf}, S^1$,
+and $\pi^{-\myhalf}$.  As initial approximations, set 
+$\etarho^{-\myhalf}=0, \psi^{-\myhalf}=0, S^{1,0}=S^0$, and $\pi^{-\myhalf}=0.$
 \begin{enumerate}

### 2009-02-04 docs/flowchart/flowchart.tex
moved correct_base so it occurs before advect_base_pressure.  results will change slightly

@@
 
-\item Should we eeplace {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
+\item Should we replace {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
 plane-parallel and $\rho_0 = \rho_0 - {\rm{\bf Avg}}(\rho_0 - \rho)$ in spherical?
 This eliminates the need for $\nabla\cdot\etarho$.
-
-\item In {\bf Step 4} and {\bf Step 8}, should we move {\bf Correct Base} so it 
-occurs immediately before {\bf Advect Base Pressure}?  Since the full
-enthalpy advance ``feels'' the effects of the time-centered $\etarho$, we should
-rely on a $\rho_0^{n+1}$ that also feels the effects.
-
-\item Should we use a predictor-corrector for updating the full-state density?
-Specifically, after calling {\bf Correct Base}, should we do a full-state density 
-advance and {\bf Correct Base} using the more accurate estimate of $\rho_0^{n+1}$?
 
@@
 way of enforcing HSE anyway, so why not just compute the new pressure directly?
+
+\item Should we use a predictor-corrector for updating the full-state density?
+Specifically, after calling {\bf Correct Base}, should we do a full-state density 
+advance and {\bf Correct Base} using the more accurate estimate of $\rho_0^{n+1}$?
 
@@
 
+\item {\bf Correct Base}$[\rho_0^{(2a),\star}, (\nabla\cdot\etarho\er)^{\nph,*}] \rightarrow [\rho_0^{n+1,\star}]$.
+
 \item {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]$ 
@@
 
-\item {\bf Correct Base}$[\rho_0^{(2a),\star}, (\nabla\cdot\etarho\er)^{\nph,*}] \rightarrow [\rho_0^{n+1,\star}]$.
-
 %--------------------------------------------------------------------------
@@
 \end{description}
+
+\item {\bf Correct Base}$[\rho_0^{(2a)}, (\nabla\cdot\etarho\er)^{\nph}] \rightarrow [\rho_0^{n+1}]$
 
@@
 \end{equation}
-
-\item {\bf Correct Base}$[\rho_0^{(2a)}, (\nabla\cdot\etarho\er)^{\nph}] \rightarrow [\rho_0^{n+1}]$
 

### 2009-02-04 Source/advance.f90

@@ -510,6 +510,11 @@ contains
        endif
     end if
 
+    ! Correct the base state using the time-centered etarho and psi
+    if (use_etarho .and. evolve_base_state) then
+       call correct_base(rho0_new,div_etarho,dt)
+    end if
+
     if (evolve_base_state) then
        call advect_base_pres(w0,Sbar,p0_old,p0_new,gamma1bar,psi,etarho_cc,dx(:,dm),dt)
     else
@@ -540,11 +545,6 @@ contains
        call destroy(scal_force(n))
        call destroy(thermal(n))
     end do
-
-    ! Correct the base state using the time-centered etarho and psi
-    if (use_etarho .and. evolve_base_state) then
-       call correct_base(rho0_new,div_etarho,dt)
-    end if
 
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
 !! STEP 4a (Option I) -- Add thermal conduction (only enthalpy terms)
@@ -876,6 +876,11 @@ contains
        call destroy(etarhoflux(n))
     end do
 
+    ! Correct the base state using the time-centered etarho and psi
+    if (use_etarho .and. evolve_base_state) then
+       call correct_base(rho0_new,div_etarho,dt)
+    end if
+
     if (evolve_base_state) then
        call advect_base_pres(w0,Sbar,p0_old,p0_new,gamma1bar,psi,etarho_cc,dx(:,dm),dt)
     else
@@ -905,11 +910,6 @@ contains
        call destroy(scal_force(n))
        call destroy(thermal(n))
     end do
-
-    ! Correct the base state using the time-centered etarho and psi
-    if (use_etarho .and. evolve_base_state) then
-       call correct_base(rho0_new,div_etarho,dt)
-    end if
 
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
     !! STEP 8a (Option I) -- Add thermal conduction (only enthalpy terms)

### 2009-02-05 docs/flowchart/flowchart.tex
added note about moving correct base to Algorithm Updates Section
after react_state, set (rho h)_0 = Avg(rho h) in plane-parallel and (rho h)_0 = (rho h)_0 - Avg[(rho h)_0 - (rho h)] in spherical.
React Base no longer is called

rhoh_1 and rhoh_2 don't exist anymore.  Advect Base Enthalpy advances (rho h)^n to (rho h)^{n+1}

All of this is documented in the flowchart

@@
 \item We are still working out the fill, average, and interpolate stencils.
-\item We are working on {\bf Advect Base}, {\bf React Base}, and 
-{\bf Correct Base} details.
+\item We are working on {\bf Advect Base} and {\bf Correct Base} details.
 \item We are working on $\etarho$ details.
@@
 
-\item We have added the flag, {\tt full\_rhoh0\_evolution}, which defaults to 
-{\tt TRUE} (note that we do not have a {\bf Correct Base} or 
-{\bf Diffuse Base} for enthalpy).  When {\tt FALSE}, we do not call {\bf React Base},
-but instead set $(\rho h)_0 = \overline{(\rho h)^n}$ before {\bf Advect Base}.
-
 \item We have given the advection scheme a face-lift, with some nifty options like
@@
 
-\end{itemize}
-
-%-----------------------------------------------------------------------------
-% Future Considerations
-%-----------------------------------------------------------------------------
-
-\section{Future Considerations}
-
-\begin{itemize}
-
-\item Should we replace {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
-plane-parallel and $\rho_0 = \rho_0 - {\rm{\bf Avg}}(\rho_0 - \rho)$ in spherical?
-This eliminates the need for $\nabla\cdot\etarho$.
-
-\item After {\bf React State}, should we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
+\item In {\bf Step 4} and {\bf Step 8}, we moved {\bf Correct Base} so it
+occurs immediately before {\bf Advect Base Pressure}.  Since the full
+enthalpy advance ``feels'' the effects of the time-centered $\etarho$, we should
+rely on a $\rho_0^{n+1}$ that also feels the effects.
+
+\item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
 plane-parallel and $(\rho h)_0 = (\rho h)_0 - {\rm{\bf Avg}}[(\rho h)_0 - (\rho h)]$ 
-in spherical?  This can currently be done by making {\tt full\_rhoh0\_evolution = FALSE}.
-This means that {\bf React Base} no longer has to exist, and the intermediate states, 
+in spherical.  This means that {\bf React Base} no longer exists, and the intermediate states, 
 $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
@@
 {\bf Diffuse Base}.
+
+\end{itemize}
+
+%-----------------------------------------------------------------------------
+% Future Considerations
+%-----------------------------------------------------------------------------
+
+\section{Future Considerations}
+
+\begin{itemize}
+
+\item Should we replace {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
+plane-parallel and $\rho_0 = \rho_0 - {\rm{\bf Avg}}(\rho_0 - \rho)$ in spherical?
+This eliminates the need for $\nabla\cdot\etarho$.
 
@@
 \begin{equation}
-\left.\begin{array}{ccccccccc}
-& \text{React Base} & & \text{Advect Base} & & \text{Correct Base} & & \text{React Base} & \\
-\rho_0^n & & & \longrightarrow & \rho_0^{(2a)} & \longrightarrow & & & \rho_0^{n+1} \\
-(\rho h)_0^n & \longrightarrow & (\rho h)_0^{(1)} & \longrightarrow & & & (\rho h)_0^{(2)} & \longrightarrow & (\rho h)_0^{n+1} \\
-p_0^n & & & \longrightarrow & & & & & p_0^{n+1} \\
+\left.\begin{array}{ccccccc}
+& \text{Advect Base} & & \text{Correct Base} & \\
+\rho_0^n & \longrightarrow & \rho_0^{(2a)} & \longrightarrow & \rho_0^{n+1} \\
+(\rho h)_0^n & \longrightarrow & & & (\rho h)_0^{n+1} \\
+p_0^n & \longrightarrow & & & p_0^{n+1} \\
 \end{array}\right.\nonumber
@@
 
-\item {\bf React Base}$[(\rho h)_0^{\inp},\overline{(\rho\Hnuc)^{\inp}},\overline{(\rho\Hext)^{\inp}}] \rightarrow [(\rho h)_0^{\outp}]$\\
-is the process by which we update the base state enthalpy through $\dt/2$ in time using
-\begin{equation}
-\frac{\partial(\rho h)_0}{\partial t} = \overline{(\rho\Hnuc)^{\inp}} + \overline{(\rho\Hext)^{\inp}}\enskip ,
-\end{equation}
-discretized as
-\begin{equation}
-(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} + \frac{\dt}{2}\overline{(\rho\Hnuc)_{j}^{\inp}} + \frac{\dt}{2}\overline{(\rho\Hext)_{j}^{\inp}} \enskip .
-\end{equation}
-
 \item {\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp]$
@@
                                   (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)} ].$
-
-\item {\bf if} ~ {\tt full\_rhoh0\_evolution = T} ~ {\bf then}
-
-~~~ {\bf React Base}$[(\rho h)_0^n,\overline{(\rho \Hnuc)^{(1)}},\overline{(\rho\Hext)^n}] \rightarrow [(\rho h)_0^{(1)}]$ \\
-{\bf else}
-
-~~~ $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$ in plane-parallel
-
-~~~ $(\rho h)_0^{(1)} = (\rho h)_0^{n} - {\rm{\bf Avg}}\left[(\rho h)_0^{n} - (\rho h)^{(1)}\right] $ in spherical. \\
-{\bf end if}
+\item
+\begin{description}
+
+\item[plane-parallel:] Set $(\rho h)_0^{n} = \overline{(\rho h)^{(1)}}$.
+
+\item[spherical:] Set $(\rho h)_0^{n} = (\rho h)_0^{n} - {\rm{\bf Avg}}\left[(\rho h)_0^{n} - (\rho h)^{(1)}\right]$.
+
+\end{description}
+
 \item Define
@@
 
-\item {\bf Advect Base Enthalpy}$[(\rho h)_0^{(1)}, w_0^{\nph,\star}, \psi^{\nph,\star}]$ 
-$\rightarrow [(\rho h)_0^{(2),\star}]$.
+\item {\bf Advect Base Enthalpy}$[(\rho h)_0^{n}, w_0^{\nph,\star}, \psi^{\nph,\star}]$ 
+$\rightarrow [(\rho h)_0^{n+1,\star}]$.
 
@@
 (\rho h)^{(1),\nph,\star} = 
-(\rho h)^{'(1),\nph,\star} + \frac{(\rho h)_0^n + (\rho h)_0^{(2),\star}}{2}
+(\rho h)^{'(1),\nph,\star} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}
 \enskip .
@@
 
-\item {\bf Advect Base Enthalpy}$[(\rho h)_0^{(1)}, w_0^{\nph}, \psi^{\nph}]$ 
-$\rightarrow [(\rho h)_0^{(2)}]$.
+\item {\bf Advect Base Enthalpy}$[(\rho h)_0^{n}, w_0^{\nph}, \psi^{\nph}]$ 
+$\rightarrow [(\rho h)_0^{n+1}]$.
 
@@
 (\rho h)^{(1),\nph} = 
-(\rho h)^{'(1),\nph} + \frac{(\rho h)_0^n + (\rho h)_0^{(2)}}{2} \enskip .
+(\rho h)^{'(1),\nph} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1}}{2} \enskip .
 \end{equation}
@@
                                (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$
-
-\item {\bf React Base}$[(\rho h)_0^{(2)},\overline{(\rho \Hnuc)^{(2)}},\overline{(\rho\Hext)^{(2)}}] \rightarrow [(\rho h)_0^{n+1}]$
 

### 2009-02-05 Source/advance.f90

@@ -132,15 +132,11 @@ contains
     real(dp_t) ::             w0_force(nlevs_radial,0:nr_fine-1)
     real(dp_t) ::                 Sbar(nlevs_radial,0:nr_fine-1)
     real(dp_t) ::        div_coeff_nph(nlevs_radial,0:nr_fine-1)
-    real(dp_t) ::          rho_Hextbar(nlevs_radial,0:nr_fine-1)
-    real(dp_t) ::              rhoh0_1(nlevs_radial,0:nr_fine-1)
-    real(dp_t) ::              rhoh0_2(nlevs_radial,0:nr_fine-1)
     real(dp_t) ::        gamma1bar_old(nlevs_radial,0:nr_fine-1)
     real(dp_t) :: delta_gamma1_termbar(nlevs_radial,0:nr_fine-1)
     real(dp_t) ::               w0_old(nlevs_radial,0:nr_fine)
     real(dp_t) ::       div_coeff_edge(nlevs_radial,0:nr_fine)
     real(dp_t) ::  rho0_predicted_edge(nlevs_radial,0:nr_fine)
-    real(dp_t) ::          rho_Hnucbar(nlevs_radial,0:nr_fine-1)
     real(dp_t) ::          div_etarhoh(nlevs_radial,0:nr_fine-1)
 
     integer    :: r,n,comp,proj_type
@@ -388,45 +384,22 @@ contains
 
     do n=1,nlevs
        call destroy(rho_omegadot1(n))
-    end do
-    
-    if (full_rhoh0_evolution) then
-       if (evolve_base_state) then
-          if (parallel_IOProcessor() .and. verbose .ge. 1) then
-             write(6,*) '            : react  base >>> '
-          end if
-          call average(mla,rho_Hnuc1,rho_Hnucbar,dx,1)
-          call average(mla,rho_Hext, rho_Hextbar,dx,1)
-          call react_base(rhoh0_old,rho_Hnucbar,rho_Hextbar,halfdt,rhoh0_1)
-       else
-          rhoh0_1 = rhoh0_old
-       end if
-    else
-       if (evolve_base_state) then
-          if (spherical .eq. 0) then
-             call average(mla,s1,rhoh0_1,dx,rhoh_comp)
-          else
-             ! Simply setting rhoh0_1 = Avg(rhoh^{(1)}) doesn't work.
-             ! Instead, what we need to do first set
-             ! set rhoh0_1 = rhoh0_old
-             rhoh0_1 = rhoh0_old
[…]

### 2009-02-05 probin.f90

@@ -49,7 +49,7 @@ module probin_module
   integer, save            :: w0mac_interp_type
   integer, save            :: nOutFiles
   logical, save            :: lUsingNFiles
-  logical, save            :: use_tfromp, full_rhoh0_evolution, single_prec_plotfiles
+  logical, save            :: use_tfromp, single_prec_plotfiles
   logical, save            :: use_soundspeed_firstdt, use_divu_firstdt
   logical, save            :: smallscale_beta
   integer, save            :: use_ppm
@@ -143,7 +143,6 @@ module probin_module
   namelist /probin/ nOutFiles
   namelist /probin/ lUsingNFiles
   namelist /probin/ use_tfromp
-  namelist /probin/ full_rhoh0_evolution
   namelist /probin/ single_prec_plotfiles
   namelist /probin/ use_soundspeed_firstdt
   namelist /probin/ use_divu_firstdt
@@ -307,7 +306,6 @@ contains
     lUsingNFiles = .true.
 
     use_tfromp = .false.
-    full_rhoh0_evolution = .true.
 
     single_prec_plotfiles = .false.
 
@@ -788,11 +786,6 @@ contains
           farg = farg + 1
           call get_command_argument(farg, value = fname)
           read(fname, *) use_tfromp
-
-       case ('--full_rhoh0_evolution')
-          farg = farg + 1
-          call get_command_argument(farg, value = fname)
-          read(fname, *) full_rhoh0_evolution
 
        case ('--use_single_prec_plotfiles')
           farg = farg + 1

### 2009-02-05 docs/eta_notes/eta.tex
switch over to book format for the master document -- this makes the sectioning more natural
add MarginPars
fixed the etarho flow chart to reflect a time-centered eta
make the definitions of \half and \myhalf consistent throughout the notes
addressed some marginpars, put in section labels with proper referencing in the flowchart
put some notes about eta in spherical coordinate in here -- they were originally in paper IV

--- 2008-10-01 docs/eta_notes/paper.tex
@@
-\documentclass[11pt]{article}
+We carry around three different 1D radial quantities: $\etarhoec$
+(edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$
+(cell-centered).  These notes discuss when each of these is used, and
+how they are computed, in both plane-parallel and spherical.
 
-\tolerance=600
+\section{The Mixing Term, $\etarho$}
 
-\usepackage{amsmath,color}
+The base state evolves in response to heating and mixing in the star.
+The density evolution is governed by
+\begin{eqnarray}
+\frac{\partial \rho_0}{\partial t} &=& - 
+ \nablab \cdotb \left( \rho_0 w_0 \er \right)
+- \nablab \cdotb \left( \etarho \er \right) \enskip ,
+\label{eq:rho0upd_new}
+\end{eqnarray}
+with
+\begin{equation}
+\etarho(r) = \overline{\left(\rhop \ut \cdot \er \right)} = \frac{1}{A(\Omega_H)} 
+ \int_{\Omega_H}  (\rhop \ut \cdot \er ) \; dA \enskip , \label{eq:eta}
+\end{equation}
+designed to keep the average value of the full density, $\rho$, over a
+layer of constant radius in the star equal to $\rho_0$.  To complete
+the update of the base state, we need evolution equations for the
+pressure, $p_0$, and velocity, $w_0$.  For spherical geometry, the
+derivation of $w_0$ constraint equation is shown in
+Appendix~\ref{app:gravity}, resulting in the following system
+\begin{eqnarray}
+w_0 &=& \ow + \delta w_0 \\
+\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \ow \right ) &=& \Sbar \\
+\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=& - \frac{g}{r^2} \frac{\partial (r^2 \etarho)}{\partial r} - \frac{4 (\ow + \dw) \rho_0 g}{r}
+- 4 \pi G \rho_0 \etarho \label{eq:dw0constraint}
+\end{eqnarray}
 
-% Margins
-\usepackage[lmargin=0.5in,rmargin=1.5in,tmargin=0.5in,bmargin=0.5in]{geometry}
+In paper~III, we introduced a mixing term, $\etarho$, to the density
+evolution equation (eq.~[\ref{eq:rho0upd_new}]), with the objective of
+keeping the base state density equal to the average of the density
+over a layer.  For a spherical base state, it is best to define the 
+average in terms of spherical coordinates,
+\begin{equation}
+\overline{q} = \frac{1}{4\pi} \int_{\Omega_H} q(r,\theta,\phi) \; d\Omega
+\end{equation}
+where $\int_{\Omega_H} d\Omega = 4\pi$ represents the integral over
+the spherical $\theta$ and $\phi$ angles at constant radius.
 
-\def\half   {\frac{1}{2}}
+Recall that if initially $ \overline{\rhop} = 1 /
+{4\pi} \int_{\Omega_H} \rhop \; d\Omega = 0,$ 
+\MarginPar{this section is here temporarily, just to better understand
+what's going on here.  These notes came from an early version of
+paper~III} 
+there is no guarantee that $\overline{\rhop} = 0$ will hold at later
+time.  To see this we start with the equation for $\rho'$, 
+\begin{equation}
+\frac{\partial \rhop}{\partial t} = - \nablab \cdotb (\rho_0 \ut)
+-\nablab \cdotb (\rhop w_0 \er) - \nablab \cdotb (\rhop \ut )
+\label{eq:rhoprimeold} \enskip ,
+\end{equation}
+rewritten by moving one term to the left-hand-side:
+\begin{equation}
+\frac{\partial \rhop}{\partial t} + \nablab \cdotb (\rhop \ut) 
+= - \nablab \cdotb (\rho_0 \ut)
+  -\nablab \cdotb (\rhop w_0 \er) \enskip .
+\end{equation}
+We integrate this over a spherical shell of thickness $2h$ at radius $r_0$, 
+$\Omega_H \times (r_0-h, r_0+h)$, and normalize by the integration
+volume, which is $\sim 4\pi r_0^2 \, 2h$ for small $h$, to obtain:
+\begin{eqnarray}
+\frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H} 
+    \left(\frac{\partial \rhop}{\partial t} + \nabla \cdotb ( \rhop \ut ) \right) \; d\Omega 
+&=& - \frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} r^2 \; dr 
+\int_{\Omega_H} \nabla \cdotb \left( \rho_0 \ut + \rhop w_0 \er \right) \; d\Omega \nonumber  \\
+&=& \left .  - \frac{1}{4\pi r_0^2 \, 2h}  \int_{\Omega_H} 
+\left [ \rho_0  (\ut \cdot \er)  + \rhop w_0 \right] r^2
+  \; d\Omega  \; \right  |_{r_0-h}^{r_0+h} \nonumber  \\
+&=& 0 \nonumber \enskip ,
+\end{eqnarray} 
+where we have used the divergence theorem in spherical coordinates to transform
+the volume integrals into area integrals over $\Omega_H.$
+We see that both of the terms on the right hand side disappear. The first,
+$\int_{\Omega_H} \rho_0 (\ut \cdot \er) d\Omega = 0$ follows from the definition of $\ut,$
+and the second, $\int_{\Omega_H} \rhop w_0 d\Omega = w_0 \int_{\Omega_H} \rhop d\Omega = 0$
+follows from our initial assumption.  
+%\begin{eqnarray}
+%\frac{1}{4\pi r_0^2 \, 2h} \frac{\partial}{\partial t} 
+%  \int_{r_0-h}^{r_0+h} \int_{\Omega_H}  \rhop \; r^2 dr \; d\Omega &=&  
+%\frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} \int_{\Omega_H} \nabla \cdotb \left( \rho_0 \ut + \rhop w_0 \er \right)
+%\; r^2 dr \; d\Omega &=&
+%\\
+%&=& \left .  - \frac{1}{4\pi r_0^2 \, 2h} \int_{\Omega_H} 
+% \rhop (\ut \cdot \er)  \; r^2 d\Omega  \right  |_{r_0-h}^{r_0+h}
+%\enskip ,
+%\end{eqnarray} 
+Now expanding the remaining terms and taking the limit as $h \rightarrow 0,$ we can write 
+\begin{eqnarray} 
+0 &=& \lim_{h\rightarrow 0} \frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H} 
+    \left(\frac{\partial \rhop}{\partial t} + \nabla \cdotb ( \rhop \ut ) \right) \; d\Omega \nonumber \\
+&=& \frac{\partial}{\partial t} \left( \lim_{h\rightarrow 0} \frac{1}{4\pi r_0^2} \frac{1}{2h} 
+  \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H}  \rhop \;d\Omega \right) +
+\lim_{h\rightarrow 0} \left( \frac{1}{4\pi r_0^2 \, 2h}  
+  \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H}  \nabla \cdotb ( \rhop \ut ) \; d\Omega \right) \nonumber \\
+&=& \frac{\partial}{\partial t} \left(\frac{1}{4\pi} \int_{\Omega_H}  \rhop \; d\Omega \right) +
+\lim_{h\rightarrow 0} \left( \left .  \frac{1}{4\pi r_0^2 \, 2h}  
+\left [ \int_{\Omega_H} \rhop (\ut \cdot \er) \; d\Omega  \right]  r^2\; \right |_{r_0-h}^{r_0+h} \right) \nonumber  \\
+&=&  \frac{\partial}{\partial t} \overline{\rhop} + \lim_{h\rightarrow 0} \left( \left .  \frac{1}{r_0^2 \, 2h}
+\overline{[\rhop (\ut \cdot \er)]} \; r^2 \; \right |_{r_0-h}^{r_0+h} \right) \nonumber  \\
+&=&  \frac{\partial}{\partial t} \overline{\rhop} + 
+\lim_{h\rightarrow 0} \frac{1}{r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} \nabla \cdot \left[\overline{\left(\rhop\ut \cdot \er\right)} \er \right]\;  r^2 \; dr \nonumber \\
+&=&  \frac{\partial}{\partial t} \overline{\rhop} + \nabla \cdot \left[ \overline{\left(\rhop \ut \cdot \er\right)} \er \right]
+\end{eqnarray} 
+again using the divergence theorem, extracting the time derivative from the spatial integral,
+and switching the order of operations as appropriate.
 
-\def\eb         {{\bf e}}
-\def\ib         {{\bf i}}
-\def\Ub         {{\bf U}}
+In short, 
+\begin{equation}
+\frac{\partial}{\partial t} \overline{\rhop} = 
+- \nabla \cdot \left [ \overline{\left(\rhop \ut \cdot \er\right)} \, \er \right ] = 
+- \nabla \cdot ( \etarho \er )
+ \enskip .\label{eq:rhopbar} 
+\end{equation}
 
-\newcommand{\Ubt}{\widetilde{\Ub}}
+We need both $\etarho$ alone and its divergence for the
+various terms in the construction of $w_0$ and the correction to
+$\rho_0$.  The quantity $\etarho$ is edge-centered on our grid, and
+for Cartesian geometries, we constructed it by averaging the
+appropriate fluxes through the grid boundaries.  For a spherical base
+state, this does not work, since the spherical shells do not line up
+with the Cartesian grid boundaries.
 
-\def\etarho     {\eta_\rho}
-\def\etarhoec   {\etarho^{\rm ec}}
-\def\etarhocc   {\etarho^{\rm cc}}
-\def\etarhoflux {\etarho^{\rm flux}}
-\def\divetarho  {\nabla\cdot(\etarho\eb_r)}
+Therefore, we take a different approach.  We compute $\etarho$ by
+constructing the quantity $\rhop \ut \cdot \er$ in each cell, and then
+use our average routine to construct a 1-d, cell-centered
+$\eta_{\rho,r}$ (this is essentially numerically solving the integral
+in Eq.~[\ref{eq:eta}]).  The edge-centered values of $\etarho$,
+$\eta_{\rho,r+1/2}$ are then constructed by simple
+averaging: \MarginPar{should we be doing a volume-weighted average?}
+\begin{equation}
+\eta_{\rho,r+1/2} = \frac{\eta_{\rho,r} + \eta_{\rho,r+1}}{2} \enskip .
+\end{equation}
 
-\title{Notes on $\etarho$}
+Instead of differencing $\eta_{\rho,r+1/2}$ to construct the 
+divergence, we instead use equation~(\ref{eq:rhopbar}) directly, by writing:
+\begin{equation}
+\left [ \nabla \cdot (\etarho \er ) \right ]^{n+1/2}
+= - \frac{\overline{\rhop^{n+1}} - \overline{\rhop^n}}{\Delta t}
+= - \frac{\overline{\rhop^{n+1}}}{\Delta t} \enskip ,
+\end{equation}
+where we have made use of the fact that $\overline{\rhop^n} = 0$ by construction.
 
-\setlength{\marginparwidth}{1.0in}
-\newcommand{\MarginPar}[1]{\marginpar{%
-\vskip-\baselineskip %raise the marginpar a bit
-\raggedright\tiny\sffamily
-\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
-
-\begin{document}
-
-\maketitle
-\tableofcontents
-\clearpage
-
-We carry around three different 1D radial quantities: $\etarhoec$ (edge-centered), $\etarhocc$ (cell-centered), and $\divetarho$ (cell-centered).
-These notes discuss when each of these is used, and how they are computed, in both plane-parallel and spherical.
-
-\section{Flow Chart}
+%-----------------------------------------------------------------------------
+% flow chart
+%-----------------------------------------------------------------------------
+\section{$\eta$ Flow Chart}
 \begin{enumerate}
-\item Enter {\tt advance} with $[\etarhoec, \etarhocc, \divetarho]^{n-\half}$.
+\item Enter {\tt advance\_timestep} with $[\etarhoec, \etarhocc, \divetarho]^{n-\myhalf}$.
 \item Call {\tt make\_w0}.  The spherical version uses uses $\etarhoec$ and $\etarhocc$.
 \item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
+\item Call {\tt make\_etarho} to compute $[\etarhoec, \etarhocc, \divetarho]^{n+\myhalf,*}$
+\item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
 \item Call {\tt correct\_base}, which uses $\divetarho$.
-\item Compute $[\etarhoec, \etarhocc, \divetarho]^{n+\half,*}$.
-\item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
 \item Call {\tt make\_w0}.  The spherical version uses uses $\etarhoec$ and $\etarhocc$.
 \item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
+\item Call {\tt make\_etarho} to compute $[\etarhoec, \etarhocc, \divetarho]^{n+\myhalf}$
+\item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
 \item Call {\tt correct\_base}, which uses $\divetarho$.
-\item Compute $[\etarhoec, \etarhocc, \divetarho]^{n+\half}$.
-\item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
 \end{enumerate}
 
+%-----------------------------------------------------------------------------
+% computing
+%-----------------------------------------------------------------------------
 \section{Computing $\etarhoec, \etarhocc$, and $\divetarho$}
@@
 \begin{equation}
-\eta_{\rho,\ib+\half\eb_r}^{\rm flux} = \left[\left(\Ubt_{\ib+\half\eb_r}^{n+\half}\cdot\eb_r\right) + w_{0,r+\half}^{n+\half}\right] \rho_{\ib+\half\eb_r}^{n+\half} - w_{0,r+\half}^{n+\half}\rho_{0,r+\half}^{n+\half, {\rm pred}}
+\eta_{\rho,\ib+\myhalf\eb_r}^{\rm flux} = \left[\left(\Ubt_{\ib+\myhalf\eb_r}^{n+\myhalf}\cdot\eb_r\right) + w_{0,r+\myhalf}^{n+\myhalf}\right] \rho_{\ib+\myhalf\eb_r}^{n+\myhalf} - w_{0,r+\myhalf}^{n+\myhalf}\rho_{0,r+\myhalf}^{n+\myhalf, {\rm pred}}
 \end{equation}
@@
 \begin{equation}
-\eta_{\rho,r+\half}^{\rm ec} = \overline{\eta_{\rho,\ib+\half\eb_r}^{\rm flux}}
+\eta_{\rho,r+\myhalf}^{\rm ec} = \overline{\eta_{\rho,\ib+\myhalf\eb_r}^{\rm flux}}
 \end{equation}
@@
 \begin{equation}
-\eta_{\rho,r}^{\rm cc} = \frac{\eta_{\rho,r+\half}^{\rm ec} + \eta_{\rho,r-\half}^{\rm ec}}{2}.
+\eta_{\rho,r}^{\rm cc} = \frac{\eta_{\rho,r+\myhalf}^{\rm ec} + \eta_{\rho,r-\myhalf}^{\rm ec}}{2}.
 \end{equation}
@@
 \begin{equation}
-\left[\divetarho\right]_r = \frac{\eta_{\rho,r+\half}^{\rm ec} - \eta_{\rho,r-\half}^{\rm ec}}{\Delta r}.
+\left[\divetarho\right]_r = \frac{\eta_{\rho,r+\myhalf}^{\rm ec} - \eta_{\rho,r-\myhalf}^{\rm ec}}{\Delta r}.
 \end{equation}
 
-\subsection{Spherical}
-First, construct $\eta_{\rho}^{\rm cart} = [\rho'(\Ubt\cdot\eb_r)]^{n+\half}$ using:
+\subsection{Spherical}\label{Sec:eta Spherical}
+First, construct $\eta_{\rho}^{\rm cart} =
+[\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf}$ using:
 \begin{equation}
-\left[\frac{\rho^n+\rho^{n+1}}{2}-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)^{\rm cart}\right] \sum_d\left(\frac{\Ubt_{\ib+\half\eb_d}^{n+\half}\cdot\eb_d+\Ubt_{\ib-\half\eb_d}^{n+\half}\cdot\eb_d}{2}\right)n_d.
+\left[\frac{\rho^n+\rho^{n+1}}{2}-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)^{\rm cart}\right] \sum_d\left(\frac{\Ubt_{\ib+\myhalf\eb_d}^{n+\myhalf}\cdot\eb_d+\Ubt_{\ib-\myhalf\eb_d}^{n+\myhalf}\cdot\eb_d}{2}\right)n_d.
 \end{equation}
@@
 \begin{equation}
-\eta_{\rho,r-\half}^{\rm ec} = \frac{\eta_{\rho,r-1}^{\rm cc} + \eta_{\rho,r}^{\rm cc}}{2}.
+\eta_{\rho,r-\myhalf}^{\rm ec} = \frac{\eta_{\rho,r-1}^{\rm cc} + \eta_{\rho,r}^{\rm cc}}{2}.
 \end{equation}
@@
 \begin{eqnarray}
-\eta_{\rho,-\half}^{\rm ec} &=& 0, \\
-\eta_{\rho,{\rm nr}-\half}^{\rm ec} &=& \eta_{\rho,{\rm nr}-1}^{\rm cc}.
+\eta_{\rho,-\myhalf}^{\rm ec} &=& 0, \\
+\eta_{\rho,{\rm nr}-\myhalf}^{\rm ec} &=& \eta_{\rho,{\rm nr}-1}^{\rm cc}.
 \end{eqnarray}
-$\divetarho$ is constructed from the cell-centered average of $\rho'^{n+1} = \rho^{n+1} - \rho_0^{n+1}$ as
+$\divetarho$ is constructed from the cell-centered average of
+$\rho'^{n+1} = \rho^{n+1} - \rho_0^{n+1}$ as
 \begin{equation}
@@
 \end{equation}
+
+%-----------------------------------------------------------------------------
+% etarho_ec
+%-----------------------------------------------------------------------------
 \section{Using $\etarhoec$}
@@
 NOT USED.
+
 \subsection{Spherical}
-In {\tt make\_w0}, $\etarhoec$ is used in the construction of the RHS for the $\delta w_0$ equation.
+In {\tt make\_w0}, $\etarhoec$ is used in the construction of the RHS
+for the $\delta w_0$ equation.
+
+%-----------------------------------------------------------------------------
+% etarho_cc
+%-----------------------------------------------------------------------------
 \section{Using $\etarhocc$}
@@
 In {\tt make\_psi}, $\psi = \etarhocc g$.
+
 \subsection{Spherical}
-In {\tt make\_w0}, $\etarhocc$ is used in the construction of the RHS for the $\delta w_0$ equation.
+In {\tt make\_w0}, $\etarhocc$ is used in the construction of the RHS
+for the $\delta w_0$ equation.
+
+%-----------------------------------------------------------------------------
+% div etarho
+%-----------------------------------------------------------------------------
 \section{Using $\divetarho$}
@@
 In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$.
+
 \subsection{Spherical}
 In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$. \\
-\end{document}

### 2009-02-05 docs/eta_notes/eta.tex
fixed a macro

@@
 \begin{equation}
-\etarho(r) = \overline{\left(\rhop \ut \cdot \er \right)} = \frac{1}{A(\Omega_H)} 
- \int_{\Omega_H}  (\rhop \ut \cdot \er ) \; dA \enskip , \label{eq:eta}
+\etarho(r) = \overline{\left(\rhop \Ubt \cdot \er \right)} = \frac{1}{A(\Omega_H)} 
+ \int_{\Omega_H}  (\rhop \Ubt \cdot \er ) \; dA \enskip , \label{eq:eta}
 \end{equation}
@@
 \begin{equation}
-\frac{\partial \rhop}{\partial t} = - \nablab \cdotb (\rho_0 \ut)
--\nablab \cdotb (\rhop w_0 \er) - \nablab \cdotb (\rhop \ut )
+\frac{\partial \rhop}{\partial t} = - \nablab \cdotb (\rho_0 \Ubt)
+-\nablab \cdotb (\rhop w_0 \er) - \nablab \cdotb (\rhop \Ubt )
 \label{eq:rhoprimeold} \enskip ,
@@
 \begin{equation}
-\frac{\partial \rhop}{\partial t} + \nablab \cdotb (\rhop \ut) 
-= - \nablab \cdotb (\rho_0 \ut)
+\frac{\partial \rhop}{\partial t} + \nablab \cdotb (\rhop \Ubt) 
+= - \nablab \cdotb (\rho_0 \Ubt)
   -\nablab \cdotb (\rhop w_0 \er) \enskip .
@@
 \frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H} 
-    \left(\frac{\partial \rhop}{\partial t} + \nabla \cdotb ( \rhop \ut ) \right) \; d\Omega 
+    \left(\frac{\partial \rhop}{\partial t} + \nabla \cdotb ( \rhop \Ubt ) \right) \; d\Omega 
 &=& - \frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} r^2 \; dr 
-\int_{\Omega_H} \nabla \cdotb \left( \rho_0 \ut + \rhop w_0 \er \right) \; d\Omega \nonumber  \\
+\int_{\Omega_H} \nabla \cdotb \left( \rho_0 \Ubt + \rhop w_0 \er \right) \; d\Omega \nonumber  \\
 &=& \left .  - \frac{1}{4\pi r_0^2 \, 2h}  \int_{\Omega_H} 
-\left [ \rho_0  (\ut \cdot \er)  + \rhop w_0 \right] r^2
+\left [ \rho_0  (\Ubt \cdot \er)  + \rhop w_0 \right] r^2
   \; d\Omega  \; \right  |_{r_0-h}^{r_0+h} \nonumber  \\
@@
 We see that both of the terms on the right hand side disappear. The first,
-$\int_{\Omega_H} \rho_0 (\ut \cdot \er) d\Omega = 0$ follows from the definition of $\ut,$
+$\int_{\Omega_H} \rho_0 (\Ubt \cdot \er) d\Omega = 0$ follows from the definition of $\Ubt,$
 and the second, $\int_{\Omega_H} \rhop w_0 d\Omega = w_0 \int_{\Omega_H} \rhop d\Omega = 0$
@@
 %  \int_{r_0-h}^{r_0+h} \int_{\Omega_H}  \rhop \; r^2 dr \; d\Omega &=&  
-%\frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} \int_{\Omega_H} \nabla \cdotb \left( \rho_0 \ut + \rhop w_0 \er \right)
+%\frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} \int_{\Omega_H} \nabla \cdotb \left( \rho_0 \Ubt + \rhop w_0 \er \right)
 %\; r^2 dr \; d\Omega &=&
@@
 %&=& \left .  - \frac{1}{4\pi r_0^2 \, 2h} \int_{\Omega_H} 
-% \rhop (\ut \cdot \er)  \; r^2 d\Omega  \right  |_{r_0-h}^{r_0+h}
+% \rhop (\Ubt \cdot \er)  \; r^2 d\Omega  \right  |_{r_0-h}^{r_0+h}
 %\enskip ,
@@
 0 &=& \lim_{h\rightarrow 0} \frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H} 
-    \left(\frac{\partial \rhop}{\partial t} + \nabla \cdotb ( \rhop \ut ) \right) \; d\Omega \nonumber \\
+    \left(\frac{\partial \rhop}{\partial t} + \nabla \cdotb ( \rhop \Ubt ) \right) \; d\Omega \nonumber \\
 &=& \frac{\partial}{\partial t} \left( \lim_{h\rightarrow 0} \frac{1}{4\pi r_0^2} \frac{1}{2h} 
@@
 \lim_{h\rightarrow 0} \left( \frac{1}{4\pi r_0^2 \, 2h}  
-  \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H}  \nabla \cdotb ( \rhop \ut ) \; d\Omega \right) \nonumber \\
+  \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H}  \nabla \cdotb ( \rhop \Ubt ) \; d\Omega \right) \nonumber \\
 &=& \frac{\partial}{\partial t} \left(\frac{1}{4\pi} \int_{\Omega_H}  \rhop \; d\Omega \right) +
 \lim_{h\rightarrow 0} \left( \left .  \frac{1}{4\pi r_0^2 \, 2h}  
-\left [ \int_{\Omega_H} \rhop (\ut \cdot \er) \; d\Omega  \right]  r^2\; \right |_{r_0-h}^{r_0+h} \right) \nonumber  \\
+\left [ \int_{\Omega_H} \rhop (\Ubt \cdot \er) \; d\Omega  \right]  r^2\; \right |_{r_0-h}^{r_0+h} \right) \nonumber  \\
 &=&  \frac{\partial}{\partial t} \overline{\rhop} + \lim_{h\rightarrow 0} \left( \left .  \frac{1}{r_0^2 \, 2h}
-\overline{[\rhop (\ut \cdot \er)]} \; r^2 \; \right |_{r_0-h}^{r_0+h} \right) \nonumber  \\
+\overline{[\rhop (\Ubt \cdot \er)]} \; r^2 \; \right |_{r_0-h}^{r_0+h} \right) \nonumber  \\
 &=&  \frac{\partial}{\partial t} \overline{\rhop} + 
-\lim_{h\rightarrow 0} \frac{1}{r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} \nabla \cdot \left[\overline{\left(\rhop\ut \cdot \er\right)} \er \right]\;  r^2 \; dr \nonumber \\
-&=&  \frac{\partial}{\partial t} \overline{\rhop} + \nabla \cdot \left[ \overline{\left(\rhop \ut \cdot \er\right)} \er \right]
+\lim_{h\rightarrow 0} \frac{1}{r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} \nabla \cdot \left[\overline{\left(\rhop\Ubt \cdot \er\right)} \er \right]\;  r^2 \; dr \nonumber \\
+&=&  \frac{\partial}{\partial t} \overline{\rhop} + \nabla \cdot \left[ \overline{\left(\rhop \Ubt \cdot \er\right)} \er \right]
 \end{eqnarray} 
@@
 \frac{\partial}{\partial t} \overline{\rhop} = 
-- \nabla \cdot \left [ \overline{\left(\rhop \ut \cdot \er\right)} \, \er \right ] = 
+- \nabla \cdot \left [ \overline{\left(\rhop \Ubt \cdot \er\right)} \, \er \right ] = 
 - \nabla \cdot ( \etarho \er )
@@
 Therefore, we take a different approach.  We compute $\etarho$ by
-constructing the quantity $\rhop \ut \cdot \er$ in each cell, and then
+constructing the quantity $\rhop \Ubt \cdot \er$ in each cell, and then
 use our average routine to construct a 1-d, cell-centered

### 2009-02-05 Source/advance.f90
beginning the process of removing div_etarho from the code by first removing it from the checkpoint file.  It actually wasn't even needed - not sure why it was there in the first place.
You won't be able to restart with older checkpoint files since they have been modified

@@ -13,7 +13,7 @@ contains
                               rho0_new,rhoh0_new,p0_old,p0_new,tempbar,gamma1bar,w0, &
                               rho_omegadot2,rho_Hnuc2,div_coeff_old,div_coeff_new, &
                               grav_cell_old,dx,time,dt,dtold,the_bc_tower, &
-                              dSdt,Source_old,Source_new,etarho_ec,etarho_cc,div_etarho, &
+                              dSdt,Source_old,Source_new,etarho_ec,etarho_cc, &
                               psi,sponge,hgrhs)
 
     use bl_prof_module
@@ -87,7 +87,6 @@ contains
     type(multifab),  intent(inout) :: Source_new(:)
     real(dp_t)    ,  intent(inout) ::  etarho_ec(:,0:)
     real(dp_t)    ,  intent(inout) ::  etarho_cc(:,0:)
-    real(dp_t)    ,  intent(inout) :: div_etarho(:,0:)
     real(dp_t)    ,  intent(inout) ::        psi(:,0:)
     type(multifab),  intent(in   ) :: sponge(:)
     type(multifab),  intent(inout) ::  hgrhs(:)
@@ -136,6 +135,7 @@ contains
     real(dp_t) ::               w0_old(nlevs_radial,0:nr_fine)
     real(dp_t) ::       div_coeff_edge(nlevs_radial,0:nr_fine)
     real(dp_t) ::  rho0_predicted_edge(nlevs_radial,0:nr_fine)
+    real(dp_t) ::           div_etarho(nlevs_radial,0:nr_fine-1)
     real(dp_t) ::          div_etarhoh(nlevs_radial,0:nr_fine-1)
 
     integer    :: r,n,comp,proj_type

### 2009-02-05 Source/base_io.f90

@@ -11,8 +11,7 @@ module base_io_module
 contains
 
   subroutine write_base_state(state_name,w0_name,etarho_name,chk_name, &
-                              rho0,rhoh0,p0, &
-                              gamma1bar,w0,etarho_ec,etarho_cc,div_etarho, &
+                              rho0,rhoh0,p0,gamma1bar,w0,etarho_ec,etarho_cc, &
                               div_coeff,psi,problo)
     
     use parallel
@@ -30,7 +29,7 @@ contains
     real(kind=dp_t)  , intent(in) :: p0(:,0:),gamma1bar(:,0:)
     real(kind=dp_t)  , intent(in) :: div_coeff(:,0:), psi(:,0:)
     real(kind=dp_t)  , intent(in) :: w0(:,0:)
-    real(kind=dp_t)  , intent(in) :: etarho_ec(:,0:),etarho_cc(:,0:), div_etarho(:,0:)
+    real(kind=dp_t)  , intent(in) :: etarho_ec(:,0:),etarho_cc(:,0:)
 
     real(kind=dp_t) :: base_r, problo
     character(len=20) :: out_name
@@ -57,7 +56,7 @@ contains
                 base_r = problo + (dble(r)+HALF) * dr(n)
                 write(99,1000)  base_r, rho0(n,r), p0(n,r), gamma1bar(n,r), &
                      rhoh0(n,r), div_coeff(n,r), &
-                     psi(n,r), etarho_cc(n,r), div_etarho(n,r)
+                     psi(n,r), etarho_cc(n,r)
              end do
           end do
        end do
@@ -106,8 +105,7 @@ contains
 
   subroutine read_base_state(state_name,w0_name,etarho_name,chk_name, &
                              rho0,rhoh0,p0,gamma1bar,w0, &
-                             etarho_ec,etarho_cc,div_etarho, &
-                             div_coeff,psi)
+                             etarho_ec,etarho_cc,div_coeff,psi)
 
     use parallel
     use bl_prof_module
@@ -125,7 +123,7 @@ contains
     real(kind=dp_t)  , intent(inout) :: p0(:,0:),gamma1bar(:,0:)
     real(kind=dp_t)  , intent(inout) :: div_coeff(:,0:), psi(:,0:)
     real(kind=dp_t)  , intent(inout) :: w0(:,0:)
[…]

### 2009-02-05 Source/initialize.f90

@@ -25,7 +25,7 @@ contains
                                      dSdt,Source_old,rho_omegadot2,rho_Hnuc2,the_bc_tower, &
                                      div_coeff_old,div_coeff_new,gamma1bar,gamma1bar_hold, &
                                      s0_init,rho0_old,rhoh0_old,rho0_new,rhoh0_new,p0_init, &
-                                     p0_old,p0_new,w0,etarho_ec,etarho_cc,div_etarho,psi, &
+                                     p0_old,p0_new,w0,etarho_ec,etarho_cc,psi, &
                                      tempbar,grav_cell)
 
     use restart_module
@@ -46,8 +46,7 @@ contains
     real(dp_t)    , pointer       :: gamma1bar_hold(:,:),s0_init(:,:,:),rho0_old(:,:)
     real(dp_t)    , pointer       :: rhoh0_old(:,:),rho0_new(:,:),rhoh0_new(:,:),p0_init(:,:)
     real(dp_t)    , pointer       :: p0_old(:,:),p0_new(:,:),w0(:,:),etarho_ec(:,:)
-    real(dp_t)    , pointer       :: etarho_cc(:,:),div_etarho(:,:),psi(:,:),tempbar(:,:)
-    real(dp_t)    , pointer       :: grav_cell(:,:)
+    real(dp_t)    , pointer       :: etarho_cc(:,:),psi(:,:),tempbar(:,:),grav_cell(:,:)
 
     ! local
     type(ml_boxarray) :: mba
@@ -181,7 +180,7 @@ contains
     ! now that we have nr_fine we can allocate 1d arrays
     call initialize_1d_arrays(nlevs,div_coeff_old,div_coeff_new,gamma1bar,gamma1bar_hold, &
                               s0_init,rho0_old,rhoh0_old,rho0_new,rhoh0_new,p0_init, &
-                              p0_old,p0_new,w0,etarho_ec,etarho_cc,div_etarho,psi,tempbar, &
+                              p0_old,p0_new,w0,etarho_ec,etarho_cc,psi,tempbar, &
                               grav_cell)
 
     ! now that we have nr we can read in the base state
@@ -192,8 +191,7 @@ contains
     
     call read_base_state(base_state_name, base_w0_name, base_etarho_name, sd_name, &
                          rho0_old, rhoh0_old, p0_old, gamma1bar, w0, &
-                         etarho_ec, etarho_cc, div_etarho, &
[…]

### 2009-02-05 Source/varden.f90

@@ -98,14 +98,12 @@ subroutine varden()
   real(dp_t), pointer :: w0(:,:)
   real(dp_t), pointer :: etarho_ec(:,:)
   real(dp_t), pointer :: etarho_cc(:,:)
-  real(dp_t), pointer :: div_etarho(:,:)
   real(dp_t), pointer :: psi(:,:)
   real(dp_t), pointer :: tempbar(:,:)
   real(dp_t), pointer :: grav_cell(:,:)
 
   real(dp_t), allocatable :: psi_temp(:,:)
   real(dp_t), allocatable :: etarho_cc_temp(:,:)
-  real(dp_t), allocatable :: div_etarho_temp(:,:)
   real(dp_t), allocatable :: etarho_ec_temp(:,:)
   real(dp_t), allocatable :: w0_temp(:,:)
 
@@ -133,8 +131,7 @@ subroutine varden()
                                   dSdt,Source_old,rho_omegadot2,rho_Hnuc2,the_bc_tower, &
                                   div_coeff_old,div_coeff_new,gamma1bar,gamma1bar_hold, &
                                   s0_init,rho0_old,rhoh0_old,rho0_new,rhoh0_new,p0_init, &
-                                  p0_old,p0_new,w0,etarho_ec,etarho_cc,div_etarho,psi, &
-                                  tempbar,grav_cell)
+                                  p0_old,p0_new,w0,etarho_ec,etarho_cc,psi,tempbar,grav_cell)
      
 
   else if (test_set /= '') then
@@ -144,8 +141,7 @@ subroutine varden()
                                       div_coeff_old,div_coeff_new,gamma1bar, &
                                       gamma1bar_hold,s0_init,rho0_old,rhoh0_old, &
                                       rho0_new,rhoh0_new,p0_init,p0_old,p0_new,w0, &
-                                      etarho_ec,etarho_cc,div_etarho,psi,tempbar, &
-                                      grav_cell)
+                                      etarho_ec,etarho_cc,psi,tempbar,grav_cell)
 
   else
 
@@ -154,8 +150,7 @@ subroutine varden()
                                          div_coeff_old,div_coeff_new,gamma1bar, &
                                          gamma1bar_hold,s0_init,rho0_old,rhoh0_old, &
                                          rho0_new,rhoh0_new,p0_init,p0_old,p0_new,w0, &
-                                         etarho_ec,etarho_cc,div_etarho,psi,tempbar, &
[…]

### 2009-02-05 TEST_PROBLEMS/test_smallscale/base_io.f90

@@ -11,7 +11,7 @@ module base_io_module
 contains
 
   subroutine write_base_state(state_name,w0_name,etarho_name,chk_name,rho0,rhoh0,p0, &
-                              gamma1bar,w0,etarho_ec,etarho_cc,div_etarho,div_coeff,psi, &
+                              gamma1bar,w0,etarho_ec,etarho_cc,div_coeff,psi, &
                               problo)
     
     use parallel
@@ -29,7 +29,6 @@ contains
     real(kind=dp_t)  , intent(in) :: p0(:,0:),gamma1bar(:,0:)
     real(kind=dp_t)  , intent(in) :: div_coeff(:,0:), psi(:,0:)
     real(kind=dp_t)  , intent(in) :: w0(:,0:),etarho_ec(:,0:),etarho_cc(:,0:)
-    real(kind=dp_t)  , intent(in) :: div_etarho(:,0:)
 
     real(kind=dp_t) :: base_r, problo
     character(len=20) :: out_name
@@ -55,8 +54,7 @@ contains
              do r=r_start_coord(n,i),r_end_coord(n,i)
                 base_r = problo + (dble(r)+HALF) * dr(n)
                 write(99,1000)  base_r, rho0(n,r), p0(n,r), gamma1bar(n,r), &
-                     rhoh0(n,r), div_coeff(n,r), psi(n,r), etarho_cc(n,r), &
-                     div_etarho(n,r)
+                     rhoh0(n,r), div_coeff(n,r), psi(n,r), etarho_cc(n,r)
              end do
           end do
        end do
@@ -104,7 +102,7 @@ contains
 
 
   subroutine read_base_state(state_name,w0_name,etarho_name,chk_name, &
-                             rho0,rhoh0,p0,gamma1bar,w0,etarho_ec,etarho_cc,div_etarho, &
+                             rho0,rhoh0,p0,gamma1bar,w0,etarho_ec,etarho_cc, &
                              div_coeff,psi)
 
     use parallel
@@ -125,7 +123,6 @@ contains
     real(kind=dp_t)  , intent(inout) :: p0(:,0:),gamma1bar(:,0:)
     real(kind=dp_t)  , intent(inout) :: div_coeff(:,0:), psi(:,0:)
     real(kind=dp_t)  , intent(inout) :: w0(:,0:),etarho_ec(:,0:),etarho_cc(:,0:)
-    real(kind=dp_t)  , intent(inout) :: div_etarho(:,0:)
 
[…]

### 2009-02-05 docs/flowchart/flowchart.tex
removed div_etarho from code.  notes are still a work in progress to describe the change

@@
 \item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
-plane-parallel and $(\rho h)_0 = (\rho h)_0 - {\rm{\bf Avg}}[(\rho h)_0 - (\rho h)]$ 
+plane-parallel and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
 in spherical.  This means that {\bf React Base} no longer exists, and the intermediate states, 
@@
 
+\item We replaced {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
+plane-parallel and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical,
+eliminating the need for $\nabla\cdot\etarho$.
+
 \end{itemize}
@@
 \begin{itemize}
-
-\item Should we replace {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
-plane-parallel and $\rho_0 = \rho_0 - {\rm{\bf Avg}}(\rho_0 - \rho)$ in spherical?
-This eliminates the need for $\nabla\cdot\etarho$.
-
 \item Should we eliminate {\bf Advect Base Pressure} entirely, and instead directly
@@
 
-\item {\bf Correct Base}$[\rho_0^{\inp}, (\nabla\cdot\etarho\er)^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
+\item {\bf Correct Base}$[\rho_0^{\inp}, \rho^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
 is the process by which we adjust the base state density to account for
-large-scale mixing, through $\etarho$.  For both plane-parallel and spherical
-we use
-\begin{equation}
-\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \dt (\nabla\cdot\etarho\er)^{\inp}_j \enskip . 
-\end{equation}
+large-scale mixing.  It can also be used to adjust the base state enthalpy
+to account for large-scale mixing, reactions, and thermal diffusion.
+
+\begin{description}
+\item[plane-parallel:] Set $\rho_0^{\outp} = {\bf Avg}(\rho^{\inp})$.
+\item[spherical:] Set $\rho_0^{\outp} = \rho_0^{\inp} + {\bf Avg}(\rho^{\inp} - \rho_0^{\inp})$.
+\end{description}
 
@@
 
-\item[spherical:] Set $(\rho h)_0^{n} = (\rho h)_0^{n} - {\rm{\bf Avg}}\left[(\rho h)_0^{n} - (\rho h)^{(1)}\right]$.
+\item[spherical:] Set $(\rho h)_0^{n} = (\rho h)_0^{n} + {\rm{\bf Avg}}\left[(\rho h)^{(1)} - (\rho h)_0^{n}\right]$.
 

### 2009-02-05 Source/advance.f90

@@ -135,7 +135,6 @@ contains
     real(dp_t) ::               w0_old(nlevs_radial,0:nr_fine)
     real(dp_t) ::       div_coeff_edge(nlevs_radial,0:nr_fine)
     real(dp_t) ::  rho0_predicted_edge(nlevs_radial,0:nr_fine)
-    real(dp_t) ::           div_etarho(nlevs_radial,0:nr_fine-1)
     real(dp_t) ::          rhoprimebar(nlevs_radial,0:nr_fine-1)
     real(dp_t) ::         rhohprimebar(nlevs_radial,0:nr_fine-1)
 
@@ -395,7 +394,7 @@ contains
           ! set rhoh0_old = rhoh0_old - Avg(rhoh0_old - rhoh^{(1)})
           call make_sprimebar_spherical(s1,rhoh_comp,rhoh0_old,dx,rhohprimebar,mla, &
                                         the_bc_tower%bc_tower_array)
-          call correct_base(rhoh0_old,rhohprimebar,1.d0)
+          call correct_base(rhoh0_old,rhohprimebar)
        end if
     end if
 
@@ -467,11 +466,10 @@ contains
        if (use_etarho) then
 
           if (spherical .eq. 0) then
-             call make_etarho_planar(etarho_ec,etarho_cc,div_etarho,etarhoflux,mla)
+             call make_etarho_planar(etarho_ec,etarho_cc,etarhoflux,mla)
           else
              call make_etarho_spherical(s1,s2,umac,rho0_old,rho0_new,dx,dt,normal, &
-                                        etarho_ec,etarho_cc,div_etarho, &
-                                        mla,the_bc_tower%bc_tower_array)
+                                        etarho_ec,etarho_cc,mla,the_bc_tower%bc_tower_array)
           endif
 
        endif
@@ -485,7 +483,7 @@ contains
           ! set rho0_new = rho0_new - Avg(rho0_new - rho^{(2)})
           call make_sprimebar_spherical(s2,rho_comp,rho0_new,dx,rhoprimebar,mla, &
                                         the_bc_tower%bc_tower_array)
-          call correct_base(rho0_new,rhoprimebar,1.d0)
+          call correct_base(rho0_new,rhoprimebar)
        end if
     end if
 
@@ -831,11 +829,10 @@ contains
        if (use_etarho) then
 
           if (spherical .eq. 0) then
-             call make_etarho_planar(etarho_ec,etarho_cc,div_etarho,etarhoflux,mla)
[…]

### 2009-02-05 Source/make_eta.f90

@@ -1,26 +1,19 @@
 ! Compute eta_rho = Avg { rho' Utilde.e_r }  (see paper III, Eq. 30)
-! and also compute div_etarho = div {eta_rho}
 !
 ! We keep make three quantities here: 
 !    etarho     is edge-centered
 !    etarho_cc  is cell-centered
-!    div_etarho is cell-centered
 !
 ! For plane-parallel geometries, we compute etarho by averaging up 
 ! interface fluxes (etarho_flux) created in mkflux.  We compute etarho_cc
-! and div_etarho from etarho.
+! from etarho.
 !
 ! For spherical geometries, 
-!   1) We construct a multifab containing ! { rho' Utilde.e_r } and 
+!      We construct a multifab containing ! { rho' Utilde.e_r } and 
 !      use the average routine to put it in cell-centers 
 !      on the base state to get etarho_cc.  We compute etarho from these 
 !      cell-centered quantites by averaging to the center.  
-!   2) We compute div_etarho from the paper III, eq. 28,
-!
-!        d rho' / dt = - d eta_rho / dr
-!
-!      so div_etarho = - <rho'> / dt
-!
+
 
 module make_eta_module
 
@@ -39,7 +32,7 @@ contains
   ! plane-parallel geometry routines
   !---------------------------------------------------------------------------
 
-  subroutine make_etarho_planar(etarho_ec,etarho_cc,div_etarho,etarhoflux,mla)
+  subroutine make_etarho_planar(etarho_ec,etarho_cc,etarhoflux,mla)
 
     use bl_constants_module
     use geometry, only: spherical, nr_fine, r_start_coord, r_end_coord, numdisjointchunks, &
@@ -48,7 +41,6 @@ contains
 
     real(kind=dp_t)   , intent(  out) :: etarho_ec(:,0:)
     real(kind=dp_t)   , intent(  out) :: etarho_cc(:,0:)
-    real(kind=dp_t)   , intent(  out) :: div_etarho(:,0:)
     type(multifab)    , intent(in   ) :: etarhoflux(:)
     type(ml_layout)   , intent(in   ) :: mla
 
@@ -127,16 +119,6 @@ contains
        do i=1,numdisjointchunks(n)
           do r=r_start_coord(n,i),r_end_coord(n,i)
[…]

### 2009-02-05 correct_base.f90

@@ -10,15 +10,15 @@ module correct_base_module
 
 contains
 
-  subroutine correct_base(rho0_new,div_etarho,dt)
+  subroutine correct_base(rho0_new,sprimebar)
 
     use bl_prof_module
     use geometry, only: spherical, nlevs_radial
     use restrict_base_module
+    use bl_error_module
 
     real(kind=dp_t), intent(inout) :: rho0_new(:,0:)
-    real(kind=dp_t), intent(in   ) :: div_etarho(:,0:)
-    real(kind=dp_t), intent(in   ) :: dt
+    real(kind=dp_t), intent(in   ) :: sprimebar(:,0:)
     
     ! local
     integer :: n
@@ -29,9 +29,9 @@ contains
     
     do n=1,nlevs_radial
        if (spherical .eq. 1) then
-          call correct_base_state_spherical(n,rho0_new(n,0:),div_etarho(n,0:),dt)
+          call correct_base_state_spherical(n,rho0_new(n,0:),sprimebar(n,0:))
        else
-          call correct_base_state_planar(n,rho0_new(n,0:),div_etarho(n,0:),dt)
+          call bl_error("correct_base not defined for plane-parallel")
        end if
     enddo
 
@@ -42,40 +42,13 @@ contains
        
   end subroutine correct_base
 
-  subroutine correct_base_state_planar(n,rho0_new,div_etarho,dt)
-
-    use geometry, only: anelastic_cutoff_coord, r_start_coord, r_end_coord, numdisjointchunks
-
-    integer        , intent(in   ) :: n
-    real(kind=dp_t), intent(inout) :: rho0_new(0:)
-    real(kind=dp_t), intent(in   ) :: div_etarho(0:)
-    real(kind=dp_t), intent(in   ) :: dt
-    
-    ! Local variables
-    integer :: r,i
-   
-!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-! UPDATE RHO0
-!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-
-    do i=1,numdisjointchunks(n)
-       do r=r_start_coord(n,i),r_end_coord(n,i)
-          if (r .lt. anelastic_cutoff_coord(n)) then
-             rho0_new(r) = rho0_new(r) - dt*div_etarho(r)
-          end if
-       end do
-    end do
-    
[…]

### 2009-02-05 docs/eta_notes/eta.tex
beginning the process of removing div_etarho from the code by first removing it from the checkpoint file.  It actually wasn't even needed - not sure why it was there in the first place.
You won't be able to restart with older checkpoint files since they have been modified
div_etarho not needed in this flowchart

@@
 \begin{enumerate}
-\item Enter {\tt advance\_timestep} with $[\etarhoec, \etarhocc, \divetarho]^{n-\myhalf}$.
+\item Enter {\tt advance\_timestep} with $[\etarhoec, \etarhocc]^{n-\myhalf}$.
 \item Call {\tt make\_w0}.  The spherical version uses uses $\etarhoec$ and $\etarhocc$.
 \item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
-\item Call {\tt make\_etarho} to compute $[\etarhoec, \etarhocc, \divetarho]^{n+\myhalf,*}$
+\item Call {\tt make\_etarho} to compute $[\etarhoec, \etarhocc]$
 \item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
-\item Call {\tt correct\_base}, which uses $\divetarho$.
 \item Call {\tt make\_w0}.  The spherical version uses uses $\etarhoec$ and $\etarhocc$.
 \item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
-\item Call {\tt make\_etarho} to compute $[\etarhoec, \etarhocc, \divetarho]^{n+\myhalf}$
+\item Call {\tt make\_etarho} to compute $[\etarhoec, \etarhocc]^{n+\myhalf}$
 \item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
-\item Call {\tt correct\_base}, which uses $\divetarho$.
 \end{enumerate}
@@
 %-----------------------------------------------------------------------------
-\section{Computing $\etarhoec, \etarhocc$, and $\divetarho$}
+\section{Computing $\etarhoec$ and $\etarhocc$}
 This is done in {\tt make\_eta.f90}.
@@
 \eta_{\rho,r}^{\rm cc} = \frac{\eta_{\rho,r+\myhalf}^{\rm ec} + \eta_{\rho,r-\myhalf}^{\rm ec}}{2}.
-\end{equation}
-$\divetarho$ is the cell-centered divergence of $\etarhoec$,
-\begin{equation}
-\left[\divetarho\right]_r = \frac{\eta_{\rho,r+\myhalf}^{\rm ec} - \eta_{\rho,r-\myhalf}^{\rm ec}}{\Delta r}.
 \end{equation}
@@
 \end{eqnarray}
-$\divetarho$ is constructed from the cell-centered average of
-$\rho'^{n+1} = \rho^{n+1} - \rho_0^{n+1}$ as
-\begin{equation}
-\divetarho = -\overline{\rho'^{n+1}}/\Delta t
-\end{equation}
 
@@
 for the $\delta w_0$ equation.
-
-%-----------------------------------------------------------------------------
-% div etarho
-%-----------------------------------------------------------------------------
-\section{Using $\divetarho$}
-\subsection{Plane-Parallel}
-In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$.
-
-\subsection{Spherical}
-In {\tt correct\_base}, $\rho_0^{n+1} = \rho_0^{n+1} - \Delta t \divetarho$. \\

### 2009-02-05 docs/flowchart/flowchart.tex
add where we compute gravity in the main algorithm
more div_etarho updates
more div_etarho cleanup

@@
 plane-parallel and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical,
-eliminating the need for $\nabla\cdot\etarho$.
+eliminating the need for $\nabla\cdot\etarho$.  {\bf Correct Base} is now used
+analogously for setting $(\rho h)_0$ after {\bf React State}.
 
@@
                                   (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)} ].$
-\item
-\begin{description}
-
-\item[plane-parallel:] Set $(\rho h)_0^{n} = \overline{(\rho h)^{(1)}}$.
-
-\item[spherical:] Set $(\rho h)_0^{n} = (\rho h)_0^{n} + {\rm{\bf Avg}}\left[(\rho h)^{(1)} - (\rho h)_0^{n}\right]$.
-
-\end{description}
+
+\item {\bf Correct Base}$[(\rho h)_0^{n},(\rho h)^{(1)}] \rightarrow [(\rho h)_0^{n}]$.
 
@@
 
-\item Define an edge-centered $\etarho^{\nph,\star}$ and  cell-centered 
-$(\nabla\cdot\etarho\er)^{\nph,\star}$:
+\item Define an edge-centered $\etarho^{\nph,\star}$:
 
 \begin{description}
-\item[plane-parallel:]~
+\item[plane-parallel:]
 
@@
 \end{equation}
-\begin{equation}
-(\nabla\cdot\etarho\er)_j^{\nph,\star} = 
-\frac{\eta_{\rho,j+\myhalf}^{\nph,\star}-\eta_{\rho,j-\myhalf}^{\nph,\star}}{\Delta r} \enskip ,
-\end{equation}
 
@@
 
-\item {\bf Correct Base}$[\rho_0^{(2a),\star}, (\nabla\cdot\etarho\er)^{\nph,*}] \rightarrow [\rho_0^{n+1,\star}]$.
+\item {\bf Correct Base}$[\rho_0^{(2a),\star},\rho^{(2),\star}] \rightarrow [\rho_0^{n+1,\star}]$.
+
+\item Update gravity, $g^{n+1,\star} = g(\rho_0^{n+1,\star})$
 
@@
 
-\item Define an edge-centered $\etarho^{\nph}$ and cell-centered 
-$(\nabla\cdot\etarho\er)^{\nph}$:
+\item Define an edge-centered $\etarho^{\nph}$:
 
@@
 \end{equation}
-\begin{equation}
-(\nabla\cdot\etarho\er)_j^{\nph} = \frac{\eta_{\rho,j+\myhalf}^{\nph} - \eta_{\rho,j-\myhalf}^{\nph}}{\Delta r} \enskip , \\
-\end{equation}
 
@@
 
-\item {\bf Correct Base}$[\rho_0^{(2a)}, (\nabla\cdot\etarho\er)^{\nph}] \rightarrow [\rho_0^{n+1}]$
+\item {\bf Correct Base}$[\rho_0^{(2a)},\rho^{(2)}] \rightarrow [\rho_0^{n+1}]$.
+
+\item Update gravity, $g^{n+1} = g(\rho_0^{n+1})$; $g^{\nph} = g(\rho_0^\nph)$
 

### 2009-02-05 docs/flowchart/flowchart.tex
removed Advect Base Pressure and put in calls to enforce_HSE and make_psi

@@
 
+\item We eliminated {\bf Advect Base Pressure} entirely, and instead directly
+enforce HSE after {\bf Correct Base}.  {\bf Advect Base Pressure} was just a fancy
+way of enforcing HSE anyway, so we just compute the new pressure directly.
+
 \end{itemize}
@@
 \begin{itemize}
-\item Should we eliminate {\bf Advect Base Pressure} entirely, and instead directly
-enforce HSE after {\bf Correct Base}?  {\bf Advect Base Pressure} was just a fancy
-way of enforcing HSE anyway, so why not just compute the new pressure directly?
 
@@
 
-\item {\bf Advect Base Pressure}$[p_0^\inp,w_0^\inp,\etarho^\inp] \rightarrow$ 
-$[p_0^\outp,\psi^\outp]$\\
-is the process by which we update the base state pressure through $\dt$ in time and
-compute $\psi$.
-\begin{description}
-
-\item[plane-parallel:] First we compute a cell-centered $\psi^\outp$ using
-\begin{equation}
- \psi_j^\outp = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^\inp + \eta_{\rho,j+\myhalf}^\inp\right) g \enskip .
-\end{equation}
-Then, we discretize the base state pressure evolution equation (paper III; eq.~[31]),
-\begin{equation}
-\frac{\partial p_0}{\partial t} = 
--w_0 \frac{\partial p_0}{\partial r} + \psi \enskip ,
-\label{eq:p0upd_new}
-\end{equation}
-to compute the new base state pressure,
-\begin{equation}
-p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
-+ w_{0,j-\myhalf}^\inp \right)
-\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\outp 
-\enskip,
-\end{equation}
-where the interface states are again found via the procedure described
-in paper II, Appendix A.
-
-\item[spherical:] See Section \ref{Sec:Advect Base Spherical} for working notes.
-
-\end{description}
-
 \item {\bf Advect Base Enthalpy}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [(\rho h)_0^\outp]$\\
@@
 
-\item {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]$ 
-$\rightarrow [p_0^{n+1,\star}, \psi^{\nph,\star}]$.
+\item Update $p_0$ to enforce hydrostatic equilbrium, $\nabla p_0^{n+1,\star} = -\rho_0^{n+1,\star} g$
+
+\item Compute $\psi^{\nph,\star}$
 
@@
 
-\item {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
-$\rightarrow [p_0^{n+1}, \psi^{n+\myhalf}]$.
+\item Update $p_0$ to enforce hydrostatic equilbrium, $\nabla p_0^{n+1} = -\rho_0^{n+1} g$
+
+\item Compute $\psi^{\nph}$
 

### 2009-02-06 docs/lo_density/lo_density.tex
switch over to book format for the master document -- this makes the sectioning more natural
minor updates to anelastic cutoff section
updated the base cutoff density section

@@
-\section{Notes on the low density parameters in {\tt MAESTRO}}
-
 These are working notes for the low density parameters in {\tt MAESTRO}.
 
-\subsection{anelastic\_cutoff}
+\section{anelastic\_cutoff}
 
@@
 
-\item In {\tt init\_base\_state}, we abort the program if {\tt
-  anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
-
 \item After we initilize the base state in {\tt varden} and at the
   beginning of {\tt advance\_timestep}, We compute the coordinate
-  associated with {\tt anelastic\_cutoff}, {\tt
-    anelastic\_cutoff\_coord}.  This is done by setting ${\tt
-    anelastic\_cutoff\_coord}= r$ as soon as $\rho_0(r) <$ {\tt
-    anelastic\_cutoff}.
+  associated with {\tt anelastic\_cutoff}, {\tt anelastic\_cutoff\_coord}.  
+  This is done by setting ${\tt anelastic\_cutoff\_coord}= r$ as soon as 
+  $\rho_0(r) <$ {\tt anelastic\_cutoff}.
 
-\item In {\tt correct\_base}, we only update $\rho_0$ for $r\in[{\tt
-    r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+\item In {\tt correct\_base}, we only update $\rho_0$ for 
+  $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
 
-\item In {\tt make\_div\_coeff}, we change how we compute {\tt
-  div\_coeff} for $r\in[{\tt anelastic\_cutoff\_coord},{\tt
-    r\_end\_coord}]$.
+\item In {\tt init\_base\_state}, we abort the program if 
+  {\tt anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
 
-\item In {\tt make\_psi}, we only compute $\psi$ for $r\in[{\tt
-    r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+\item In {\tt make\_div\_coeff}, we change how we compute {\tt div\_coeff} 
+  for $r\in[{\tt anelastic\_cutoff\_coord},{\tt r\_end\_coord}]$.
 
-\item in {\tt make\_S}, we set {\tt delta\_gamma1\_term} and {\tt
-  delta\_gamma1} to zero for $r\in[{\tt anelastic\_cutoff\_coord},{\tt
-    r\_end\_coord}]$.
+\item In {\tt make\_psi}, we only compute $\psi$ for 
+  $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+
+\item in {\tt make\_S}, we set {\tt delta\_gamma1\_term} and {\tt delta\_gamma1} 
+  to zero for $r\in[{\tt anelastic\_cutoff\_coord},{\tt r\_end\_coord}]$.
+
+\item In {\tt sponge}, {\tt anelastic\_cutoff} is used in a problem
+  dependent way.
+
+\item In {\tt probin}, {\tt anelastic\_cutoff} is set to $3\times 10^6$
+  by default.
+
 \end{itemize}
 
-\subsection{base\_cutoff\_density}
+\section{base\_cutoff\_density}
 
 \begin{itemize}
+
+\item In {\tt probin}, {\tt base\_cutoff\_density} is set to $3\times 10^6$
+  by default.
 
@@
 \item In {\tt make\_hgrhs}, we only add the volume discrepancy for
-  $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+  $r\in[{\tt r\_start\_coord},{\tt base\_cutoff\_density\_coord}-1]$.
 
 \item In {\tt make\_macrhs}, we only add the volume discrepancy for
-  $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+  $r\in[{\tt r\_start\_coord},{\tt base\_cutoff\_density\_coord}-1]$.
+
+\item In {\tt make\_w0}, we only add the volume discrepancy for
+  $r\in[{\tt r\_start\_coord},{\tt base\_cutoff\_density\_coord}-1]$.
 
@@
   enclosed mass, $M(r)$ if $\rho$ is less than {\tt base\_cutoff\_density}.
+
+\item In {\tt enforce\_HSE}, we set $p_0(r) = p_0(r-1)$ for 
+  $r\in[{\tt base\_cutoff\_density\_coord}-1]$
+
 \end{itemize}
 
-\subsection{burning\_cutoff}
+\section{burning\_cutoff}
 

### 2009-02-10 docs/lo_density/lo_density.tex
fixed a minor error
for spherical, correct base everywhere, not just inside the anelastic cutoff

@@
 
-\item In {\tt correct\_base}, we only update $\rho_0$ for 
-  $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
-
 \item In {\tt init\_base\_state}, we abort the program if 
@@
 
-\item In {\tt make\_psi}, we only compute $\psi$ for 
+\item In {\tt make\_psi} for plane-parallel, we only compute $\psi$ for 
   $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.

### 2009-02-10 correct_base.f90

@@ -40,7 +40,7 @@ contains
 
   subroutine correct_base_state_spherical(s0_new,sprimebar)
 
-    use geometry, only: anelastic_cutoff_coord, r_cc_loc, r_edge_loc, dr
+    use geometry, only: r_cc_loc, r_edge_loc, dr, nr
 
     real(kind=dp_t), intent(inout) :: s0_new(0:)
     real(kind=dp_t), intent(in   ) :: sprimebar(0:)
@@ -48,7 +48,7 @@ contains
     ! Local variables
     integer :: r
 
-    do r=0,anelastic_cutoff_coord(1)-1
+    do r=0,nr(1)-1
        s0_new(r) = s0_new(r) + sprimebar(r)
     end do
     

### 2009-02-17 docs/lo_density/lo_density.tex
updated info about cutoff coordinates

@@
 These are working notes for the low density parameters in {\tt MAESTRO}.
+
+\section{Computing the Cutoff Values}
+
+At several points in the algorithm, we compute {\tt anelastic\_cutoff\_coord}, 
+{\tt base\_cutoff\_density\_coord}, and {\tt burning\_cutoff\_density\_coord} by 
+setting the corresponding coordinate equal to $r$ as soon as $\rho_0(r)$ is less than 
+that particular cutoff value.  We set the coordinates at the following points in the
+code:
+
+\begin{itemize}
+
+\item After we initialize the base state in {\tt varden}.
+
+\item At the beginning of {\tt advance\_timestep}.
+
+\item After each call to {\tt advect\_base\_dens}.
+
+\item After each call to {\tt correct\_base} in spherical 
+  (or call to {\tt average} in plane-parallel.
+
+\end{itemize}
 
@@
 \begin{itemize}
-
-\item After we initilize the base state in {\tt varden} and at the
-  beginning of {\tt advance\_timestep}, We compute the coordinate
-  associated with {\tt anelastic\_cutoff}, {\tt anelastic\_cutoff\_coord}.  
-  This is done by setting ${\tt anelastic\_cutoff\_coord}= r$ as soon as 
-  $\rho_0(r) <$ {\tt anelastic\_cutoff}.
 
@@
 
-\item After we initilize the base state in {\tt varden} and at the
-  beginning of {\tt advance\_timestep}, We compute the coordinate
-  associated with {\tt base\_cutoff\_density}, {\tt
-    base\_cutoff\_density\_coord}.  This is done by setting ${\tt
-    base\_cutoff\_density\_coord}= r$ as soon as $\rho_0(r) <$ {\tt
-    base\_cutoff\_density}.
-
 \item In {\tt make\_hgrhs}, we only add the volume discrepancy for
@@
 \item In {\tt enforce\_HSE}, we set $p_0(r) = p_0(r-1)$ for 
-  $r\in[{\tt base\_cutoff\_density\_coord}-1]$
+  $r\in[{\tt base\_cutoff\_density\_coord}-1,{\tt r\_end\_coord}]$
 
@@
 \begin{itemize}
-\item After we initilize the base state in {\tt varden} and at the
-  beginning of {\tt advance\_timestep}, We compute the coordinate
-  associated with {\tt burning\_cutoff\_density}, {\tt
-    burning\_cutoff\_density\_coord}.  This is done by setting ${\tt
-    burning\_cutoff\_density\_coord}= r$ as soon as $\rho_0(r) <$ {\tt
-    burning\_cutoff\_density}.
 
-\item In {\tt react\_state}, we only call the burner if $\rho >$ {\tt
-  burning\_cutoff\_density}.
-%\item In {\tt react\_state}, we only call the burner for $r\in[{\tt r\_start\_coord},{\tt burning\_cutoff\_density\_coord}-1]$.
+\item In {\tt react\_state}, we only call the burner if 
+  $\rho >$ {\tt burning\_cutoff\_density}.
+
 \end{itemize}

### 2009-02-17 docs/lo_density/lo_density.tex
more updates - this now fully reflects what the code is doing

@@
 
+\item In {\tt probin}, {\tt anelastic\_cutoff} and {\tt base\_cutoff\_density} are
+  set to $3\times 10^6$ by default.
+
+\item In {\tt probin}, {\tt burning\_cutoff\_density} is set to 
+  {\tt base\_cutoff\_density}.  There is no option to set 
+  {\tt burning\_cutoff\_density} using the inputs file.
+
 \item After we initialize the base state in {\tt varden}.
@@
 \item After each call to {\tt correct\_base} in spherical 
-  (or call to {\tt average} in plane-parallel.
+  (or call to {\tt average} in plane-parallel).
 
@@
 
-\item In {\tt probin}, {\tt anelastic\_cutoff} is set to $3\times 10^6$
-  by default.
-
 \end{itemize}
@@
 
-\item In {\tt probin}, {\tt base\_cutoff\_density} is set to $3\times 10^6$
-  by default.
-
-\item In {\tt base\_state}, we abort the program if {\tt
+\item In {\tt init\_base\_state}, we abort the program if {\tt
   anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
@@
   location of the first cell-center at the coarsest level for which
-  $\rho_0 <$ {\tt base\_cutoff\_density}.
+  $\rho_0 <$ {\tt base\_cutoff\_density}.  This is a trick used for making
+  the data consistent for multiple level problems.
 
 \item In {\tt base\_state}, when we are generating the initial
-  background/base state, if we are above {\tt
-    base\_cutoff\_density\_loc}, just use the values for $\rho,T$, and
-  $p$ at {\tt base\_cutoff\_density\_loc}.
+  background/base state, if we are above {\tt base\_cutoff\_density\_loc}, 
+  just use the values for $\rho,T$, and $p$ at {\tt base\_cutoff\_density\_loc}.
 
-\item In {\tt make\_hgrhs}, we only add the volume discrepancy for
+\item In {\tt make\_hgrhs}, {\tt make\_macrhs}, and {\tt make\_w0}, 
+  we only add the volume discrepancy for
   $r\in[{\tt r\_start\_coord},{\tt base\_cutoff\_density\_coord}-1]$.
 
-\item In {\tt make\_macrhs}, we only add the volume discrepancy for
-  $r\in[{\tt r\_start\_coord},{\tt base\_cutoff\_density\_coord}-1]$.
-
-\item In {\tt make\_w0}, we only add the volume discrepancy for
-  $r\in[{\tt r\_start\_coord},{\tt base\_cutoff\_density\_coord}-1]$.
-
-\item In {\tt mkscalforce} subroutine {\tt mkrhohforce}, for
+\item In {\tt mkrhohforce} for plane-parallel, for
   $r\in[{\tt base\_cutoff\_density\_coord},{\tt r\_end\_coord}]$ we
@@
 
-\item In {\tt makeRhoHfromP}, the base state pressure at edges is
-  computed differently based on where we are in relation to the {\tt
-    base\_cutoff\_density}.  For $r\in[{\tt r\_start\_coord},{\tt
-      base\_cutoff\_density\_coord}-1]$ we integrate $\rho g$ from the
-  nearest cell-centered $p_0$ value.  Otherwise, we just use the
-  cell-centered value above the edge.
-
-\item In {\tt update\_scal}, if the new $\rho$ is less than {\tt
-  base\_cutoff\_density}, we call the EOS to compute $\rho h$.
+\item In {\tt update\_scal}, if the new $\rho$ is less than 
+  {\tt base\_cutoff\_density} and {\tt do\_eos\_h\_above\_cutoff},  
+  we call the EOS to compute $h$.
 

### 2009-02-26 docs/flowchart/flowchart.tex
split the "changes" section into two parts:  Changes between paper 3 and paper 4, and Changes since paper 4
put advect base pres back in

--- 2009-02-05 docs/flowchart/flowchart.tex
@@
 
-\section{Algorithm Updates}
+%-----------------------------------------------------------------------------
+% Changes Between Paper 3 and Paper 4
+%-----------------------------------------------------------------------------
+
+\section{Changes Between Paper 3 and Paper 4}
 
@@
 
+\end{itemize}
+
+%-----------------------------------------------------------------------------
+% Changes Since Paper 4
+%-----------------------------------------------------------------------------
+
+\section{Changes Since Paper 4}
+
+\begin{itemize}
+
 \item In {\bf Step 4} and {\bf Step 8}, we moved {\bf Correct Base} so it
@@
 \begin{itemize}
+
 \item Should we eliminate {\bf Advect Base Pressure} entirely, and instead directly
 enforce HSE after {\bf Correct Base}?  {\bf Advect Base Pressure} was just a fancy
-way of enforcing HSE anyway, so why not just compute the new pressure directly?
+way of enforcing HSE anyway, so we just compute the new pressure directly.  The
+problems we have seen with enforcing HSE directly have to do with the roundoff
+error that accumulates, causing the temperature computed from the EOS to drift.
 

### 2009-02-26 docs/radial_base/radial_base.tex
spruced up this average and fill descriptions a bit - brought them a bit more in line with the wording in paper 4

--- 2008-11-20 docs/radial_base/radial_base.tex
@@
 \section{\tt fill\_3d}
-There are three different {\tt fill\_3d} modes:
+There are three different {\tt fill\_3d} modes, each performed with the
+subroutine, {\tt put\_1d\_array\_on\_cart}:
 \begin{enumerate}
-\item The 1D array is cell-centered, and we map to Cartesian cell-centers.
+\item The 1D array is bin-centered, and we map to Cartesian cell-centers.
 \item The 1D array is edge-centered, and we map to Cartesian cell-centers.
@@
 \subsection{1D Cell-Centered to Cartesian Cell-Centered}
-This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
+\label{Sec:1D Cell-Centered to Cartesian Cell-Centered}
 Figure \ref{fig:mapping} shows the Cartesian grid with the spherical
-base state overlayed (for simplicity, the figure is drawn in 2-d).  
-We compute the radius of the cell center from the center of the star,
+base state overlayed (for simplicity, the figure is drawn in 2-d 
+using $2 \Delta r = \Delta x$).  
+We compute the radius of the Cartesian cell center from the center of the star,
 \begin{equation}
@@
 base state).  We can then initialize a Cartesian cell quantity $q$ from its
-corresponding base state quantity, $q_0$ as $q_{i,j,k} = q_{0,n}$.
+corresponding base state quantity, $q_0$ as $q_{i,j,k} = q_{0,n}$.  In essence,
+we are doing a piecewise constant interpolation of the radial array onto the 
+Cartesian grid.  If the output is a vector quantity, rather than a scalar, we
+multiply the result by the unit vector in the normal direction.
 \subsection{1D Edge-Centered to Cartesian Cell-Centered}
-This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
+\label{Sec:1D Edge-Centered to Cartesian Cell-Centered}
+Again, for each Cartesian cell-center, we compute the corresponding radial
+bin.  We linearly interpolate from the two nearest radial edge values 
+to get the Cartesian cell-centered value.  In essence,
+we are doing a piecewise linear interpolation of the radial array onto the 
+Cartesian grid.  If the output is a vector quantity, rather than a scalar, we
+multiply the result by the unit vector in the normal direction.
 \subsection{1D Edge-Centered to Cartesian Edges}
-This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
+\label{Sec:1D Edge-Centered to Cartesian Edges}
 This case is only used to create {\tt w0mac} from {\tt w0} in the 
-subroutine, {\tt put\_w0\_on\_edges}.
-\section{\tt average}
-For the {\tt average} process, we first create a coarse 1D radial array 
-with $\Delta x = \Delta r$.  Then, we compute the radius of the Cartesian cell 
-center, and compute the index, $n$, of the corresponding coarse radial 
-array bin.  $q_{0,n}$ is then the average of all the $q_{i,j,k}$ whose 
-Cartesian cell center maps into coarse radial bin $n$.  Then, we 
-interpolate the coarse radial array onto the finer radial array.
+subroutine, {\tt put\_w0\_on\_edges}.  There are several variations 
+of the interpolation, controlled by {\tt w0mac\_interp\_type}.  We 
+currently use the default, {\tt w0mac\_interp\_type} = 1.  We first 
+map $w_0$ onto Cartesian cell-centers, and multiply the result by 
+the unit vector in the normal direction.  Then, for each Cartesian 
+face, we obtain the normal velocity by averaging the corresponding 
+velocities from the Cartesian cell-centers adjacent to the face.
+\subsection{1D Cell-Centered to Cartesian Edges}
+We don't have a subroutine that does this directly, but at various points
+in the algorithm we need base state quantities on edges.  For example, we 
+need $\rho_0$ at edges to compute the full state $\rho$ used to compute 
+fluxes.  More to come...
+\MarginPar{This section needs to be written.}
+\section{{\tt average}}\label{Sec:Avg}
+For the {\tt average} process, we first define a coarse 1D radial array 
+with $\Delta r_{\rm c} = \Delta x$.  Then, we compute the radius of each
+Cartesian cell center, and compute the index of the corresponding coarse radial 
+bin, $n_{\rm c} = \mathtt{int}(r / \Delta r_{\rm c})$.  
+We define $q_{0,n_{\rm c}}$ as the average of all the $q_{i,j,k}$ whose 
+Cartesian cell center maps into coarse radial bin $n_{\rm c}$.
+Next, we construct edge-centered states on the coarse radial bin using 
+$q_{0,n_{\rm c}+\myhalf} = (7/12)(q_{0,n_{\rm c}}+q_{0,n_{\rm c}+1}) 
+- (1/12)(q_{0,n_{\rm c}-1}+q_{0,n_{\rm c}+2})$ along with limiters to prevent
+new extrema.  Then, for each coarse radial bin
+we construct a quadratic profile using $q_{0,n_{\rm c}+\myhalf}, q_{0,n_{\rm c}}$ 
+and $q_{0,n_{\rm c}-\myhalf}$.  We use these quadratic profiles along with limiters
+to prevent new extrema to fill the fine radial bins.
 
@@
 \begin{center}
-\epsffile{./radial_base/mapping.eps}
+\epsffile{\radbasefigpath/mapping.eps}
 \caption{\label{fig:mapping} The Cartesian grid and spherical base
-state (shown here in 2-d for simplicity).  Here we represent the
+state (shown here in 2-d for simplicity using $2 \Delta r = \Delta x$).
+Here we represent the
 spherical base state as concentric shells (blue lines).  Since the

### 2009-02-26 docs/spherical_basestate/basestate.tex
addressed some marginpars, put in section labels with proper referencing in the flowchart
updated correct base description

@@
 
-\subsection{Calculation of $w_0$}
+\subsection{Calculation of $w_0$}\label{Sec:Spherical w0 Calculation}
 
@@
 
-\subsection{{\bf Advect Base}}
+\subsection{{\bf Advect Base}}\label{Sec:Advect Base Spherical}
 
@@
 
-\subsection{Correct Base}
+\subsection{\bf{Correct Base}}
 
-The base state density correction now includes the area factors in the divergences,
+The is the process by which we adjust the base state density to account for
+large-scale mixing.  It can also be used to adjust the base state enthalpy
+to account for large-scale mixing, reactions, and thermal diffusion.
+We have simplified the algorithm greatly by simply setting $\nabla\cdot\etarho$
+equal to the averaged difference between the base state and full state.
+
 \begin{equation}
-\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp}  - \frac{1}{r_j^2} \frac{\dt}{\dr} \left [ \left(r^2 \etarho\right)_{j+\myhalf} - \left(r^2 \etarho\right)_{j-\myhalf} \right ] \enskip .
+\rho_0^{\outp} = \rho_0^{\inp} + \underbrace{{\rm{\bf Avg}}(\rho^{\inp} - \rho_0^{\inp})}_{-\Delta t\nabla\cdot\etarho}.
 \end{equation}

### 2009-02-26 docs/pert_notes/pert.tex
updating in progress

@@
 The full density evolution equation is
-\begin{equation}
-\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub) = -\Ub\cdot\nabla\rho 
-- \rho\nabla\cdot\Ub. \label{rho equation}
-\end{equation}
-We predict both $\rho'$ and $\rho_0$ to edges and later reconstruct $\rho$.
-The base state density evolution equation is
+\begin{eqnarray}
+\frac{\partial\rho}{\partial t} &=& -\nabla\cdot(\rho\Ub) \nonumber \\
+&=& -\Ub\cdot\nabla\rho - \rho\nabla\cdot\Ub. \label{rho equation}
+\end{eqnarray}
+We predict both $\rho'$ and $\rho_0$ to edges separately and later use them to 
+reconstruct $\rho$ at edges.  The base state density evolution equation is
 \begin{equation}
@@
 \frac{\partial\rho'}{\partial t} = -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub 
-- \nabla\cdot(\rho_0\Ubt)}_{"\rho' ~ \text{force}"}. 
+- \nabla\cdot(\rho_0\Ubt)}_{f_{\rho'}},
 \label{rhoprime equation}
 \end{equation}
+where the underbraced term is the "$\rho'$ force'', which is computed 
+in {\tt modify\_scal\_force}.
 
 \subsection{Predicting $\rho'$ at edges}
-We define $\rho' = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to
+We define $\rho' = \rho^{(1)} - \rho_0^n$.  Then we predict $\rho'$ to
 edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} and the
@@
 \begin{enumerate}
+
 \item We call {\tt make\_edge\_state\_1d} using the underbraced term
@@
 $\rho_0^{n+\half,{\rm pred}}$.  This term is used to advect $\rho_0$
-in {\tt advect\_base} and therefore must also be used to compute
-$\eta_\rho$, which will be used in {\tt correct\_base}.
+in {\bf Advect Base Density}.  In plane-parallel geometries, we also use
+$\rho_0^{n+\half,{\rm pred}}$ to compute $\etarho$, which will be used 
+to compute $\psi$.
+
 \item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n +
 \rho_0^{(2)})/2$.  We compute $\rho_0^{(2)}$ from $\rho_0^n$ using
-{\tt advect\_base}, which advances equation (\ref{rho0 equation})
+{\bf Advect Base Density}, which advances equation (\ref{rho0 equation})
 through $\Delta t$ in time.  The $(2)$ in the superscript indicates
-that we have not called {\tt correct\_base} yet, which computes
+that we have not called {\bf Correct Base} yet, which computes
 $\rho_0^{n+1}$ from $\rho_0^{(2)}$.  We use $\rho_0^{(2)}$ rather than
@@
 \end{eqnarray}
+Note that the pressure term in the $(\rho h)'$ force can be represented as:
+\begin{equation}
+(\Ubt \cdot \er)\frac{\partial p_0}{\partial r} \equiv \Ubt\cdot\nabla p_0 = 
+\nabla\cdot(\Ubt p_0) - p_0\nabla\cdot\Ubt.
+\end{equation}
+For spherical problems, we use the RHS representation of this term, as we have
+seen it gives better results.
 
 \subsubsection{Advancing $(\rho h)_0$}
-\begin{itemize}
-\item {\bf Using full\_rhoh0\_evolution = F:}\\
- Define $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  
-and $(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then, $(\rho h)_0^{(1)} \rightarrow$ 
-{\tt advect\_base} $\rightarrow (\rho h)_0^{(2)}$.
-\item {\bf Using full\_rhoh0\_evolution = T:}\\
- Instead of defining $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$, the code carries 
-through the complete evolution of $(\rho h)_0$ using $(\rho h)_0^n \rightarrow$ 
-{\tt react\_base} $\rightarrow (\rho h)_0^{(1)} \rightarrow$ {\tt advect\_base}
-$\rightarrow (\rho h)_0^{(2)} \rightarrow$ {\tt react\_base} $\rightarrow (\rho h)_0^{n+1}$.  
-There is no {\tt correct\_base} function for $(\rho h)_0$ written yet, nor is any 
-adjustment made to sync up $(\rho h)_0$ after the thermal diffusion step.
-\end{itemize}
+Define $(\rho h)_0^{n} = \overline{(\rho h)^{(1)}}$  
+and $(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{n}$.  Then, $(\rho h)_0^{n} \rightarrow$ 
+{\bf Advect Base Enthalpy} $\rightarrow (\rho h)_0^{n+1}$.
 
@@
 \end{equation}
-
-\subsection{Predicting $h$ ({\tt enthalpy\_pred\_type = 2})}
-
-\subsection{Predicting $T$ then $(\rho h)'$ ({\tt enthalpy\_pred\_type = 3})}
-
-\subsection{Predicting $T$ then $h$ ({\tt enthalpy\_pred\_type = 4})}
 

### 2009-02-26 docs/pert_notes/pert.tex
more updates

@@
 \frac{\partial\rho'}{\partial t} = -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub 
-- \nabla\cdot(\rho_0\Ubt)}_{f_{\rho'}},
+- \nabla\cdot(\rho_0\Ubt)}_{\rho' ~ \text{force}},
 \label{rhoprime equation}
 \end{equation}
-where the underbraced term is the "$\rho'$ force'', which is computed 
-in {\tt modify\_scal\_force}.
+where the underbraced "$\rho'$ force'' is computed in {\tt modify\_scal\_force}.
 
@@
 in (\ref{rho0 equation}) as the forcing.  This gives us
-$\rho_0^{n+\half,{\rm pred}}$.  This term is used to advect $\rho_0$
+$\rho_0^{n+\myhalf,{\rm pred}}$.  This term is used to advect $\rho_0$
 in {\bf Advect Base Density}.  In plane-parallel geometries, we also use
-$\rho_0^{n+\half,{\rm pred}}$ to compute $\etarho$, which will be used 
+$\rho_0^{n+\myhalf,{\rm pred}}$ to compute $\etarho$, which will be used 
 to compute $\psi$.
 
-\item We define $\rho_0^{n+\half,{\rm avg}} = (\rho_0^n +
+\item We define $\rho_0^{n+\myhalf,{\rm avg}} = (\rho_0^n +
 \rho_0^{(2)})/2$.  We compute $\rho_0^{(2)}$ from $\rho_0^n$ using
@@
 $\rho_0^{n+1}$ from $\rho_0^{(2)}$.  We use $\rho_0^{(2)}$ rather than
-$\rho_0^{n+1}$ to construct $\rho_0^{n+\half,{\rm avg}}$ to be
-consistent with the fact that we do not include an $\eta_\rho$
-contribution to compute $\rho'$ at edges.  $\rho_0^{n+\half,{\rm
-avg}}$ is used to construct $\rho$ at edges from $\rho'$ at edges, and
+$\rho_0^{n+1}$ to construct $\rho_0^{n+\myhalf,{\rm avg}}$ since $\rho_0^{n+1}$
+is not available yet.  $\rho_0^{n+\myhalf,{\rm avg}}$ is used to construct 
+$\rho$ at edges from $\rho'$ at edges, and
 this $\rho$ at edges is used to compute fluxes for $\rho X_k$.
@@
 \subsection{Computing $\rho$ at edges}\label{Computing rho at edges}
-For the non-radial edges, we directly add $\rho_0^{n+\half,{\rm avg}}$
-to $\rho'$ since $\rho_0^{n+\half,{\rm avg}}$ is a cell-centered
+For the non-radial edges, we directly add $\rho_0^{n+\myhalf,{\rm avg}}$
+to $\rho'$ since $\rho_0^{n+\myhalf,{\rm avg}}$ is a cell-centered
 quantity.  For the radial edges, we interpolate to obtain
-$\rho_0^{n+\half,{\rm avg}}$ at radial edges before adding it to
+$\rho_0^{n+\myhalf,{\rm avg}}$ at radial edges before adding it to
 $\rho'$.
@@
 \frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
--\nabla\cdot\left\{\left[\left(\rho_0^{n+\half,{\rm avg}} 
+-\nabla\cdot\left\{\left[\left(\rho_0^{n+\myhalf,{\rm avg}} 
 + \rho'\right)X_k\right](\Ubt+w_0\eb_r)\right\}.
@@
 \section{Energy Evolution}
+We use {\tt enthalpy\_pred\_type} = 1, in which we predict $(\rho h)'$ at edges.
 The full enthalpy equation is
@@
 \end{eqnarray}
-The full temperature equation is
-\begin{equation}
-\frac{\partial T}{\partial t} = -\Ub\cdot\nabla T
-+ \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi
-+ (\Ubt \cdot \er )\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 
-+ \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
-\end{equation}
-
-\subsection{Predicting $(\rho h)'$ ({\tt enthalpy\_pred\_type = 1} - default)}
+%The full temperature equation is
+%\begin{equation}
+%\frac{\partial T}{\partial t} = -\Ub\cdot\nabla T
+%+ \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi
+%+ (\Ubt \cdot \er )\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 
+%+ \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
+%\end{equation}
 Due to operator splitting, the call to {\tt react\_state} has already
@@
 not include the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms, where
-were already accounted for in {\tt react\_state}
+were already accounted for in {\tt react\_state}, so our equation becomes
 \begin{equation}
@@
 - \nabla\cdot[(\rho h)_0\Ubt] + (\Ubt \cdot \er)\frac{\partial p_0}{\partial r}
-+ \nabla\cdot\kth\nabla T}_{"(\rho h)' ~ \text{force}"}. \label{rhohprime equation}
++ \nabla\cdot\kth\nabla T}_{"(\rho h)' ~ \text{force}"}, \label{rhohprime equation}
 \end{eqnarray}
-Note that the pressure term in the $(\rho h)'$ force can be represented as:
+where the first two terms in $(\rho h)'$ force are computed in 
+{\tt modify\_scal\_force}, and the last two terms are accounted for in
+{\tt mkrhohforce}.  For spherical problems, we have found that a different 
+representation of the pressure term in the $(\rho h)'$ force gives better
+results, namely:
 \begin{equation}
@@
 \end{equation}
-For spherical problems, we use the RHS representation of this term, as we have
-seen it gives better results.
 
-\subsubsection{Advancing $(\rho h)_0$}
-Define $(\rho h)_0^{n} = \overline{(\rho h)^{(1)}}$  
-and $(\rho h)' = (\rho h)^{(1)}-(\rho h)_0^{n}$.  Then, $(\rho h)_0^{n} \rightarrow$ 
-{\bf Advect Base Enthalpy} $\rightarrow (\rho h)_0^{n+1}$.
+\subsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
+We define $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$.  Then we predict 
+$(\rho h)'$ to edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} 
+and the underbraced term in (\ref{rhohprime equation}) as the forcing.
 
-\subsubsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
-We predict $(\rho h)'$ to edges using {\tt make\_edge\_scal} in {\tt
-scalar\_advance} and the underbraced term in (\ref{rhohprime
-equation}) as the forcing, except that we omit the $\psi$ in the
-forcing for plane-parallel, and we omit the $\partial w_0/\partial r$
-part of $\psi$ in spherical.  Note that we still use full $\psi$ to
-update the cell.
+\subsection{Predicting $(\rho h)_0$ at edges}
+We use an analogous procedure described in Section \ref{Predicting
+rho0 at edges} for computing $\rho_0^{n+\myhalf,\rm{avg}}$ to obtain 
+$(\rho h)_0^{n+\myhalf,\rm{avg}}$, i.e., 
+$(\rho h)_0^{n+\myhalf,{\rm avg}} = [(\rho h)_0^{n} + (\rho h)_0^{n+1}]/2$.
 
-\subsubsection{Predicting $(\rho h)_0$ at edges}
-We use an analogous procedure described in Section \ref{Predicting
-rho0 at edges} to obtain $(\rho h)_0^{n+\half,\rm{pred}}$ and $(\rho
-h)_0^{n+\half,\rm{avg}}$.  Note that $(\rho h)_0^{n+\half,{\rm avg}} =
-[(\rho h)_0^{(1)} + (\rho h)_0^{(2)}]/2$.
+\subsection{Computing $\rho h$ at edges}
+We use an analogous procedure described in Section \ref{Computing rho
+at edges} for computing $\rho$ at edges to compute $\rho h$ at edges.
+However, instead of computing $(\rho h)_0$ to edges directly, we compute $\rho$
+and $h$ separately, and multiply to get $(\rho h)_0$.
 
-\subsubsection{Computing $\rho h$ at edges}
-We use an analogous procedure described in Section \ref{Computing rho
-at edges} to compute $\rho h$ at edges.
-
-\subsubsection{Advancing $\rho h$}
+\subsection{Advancing $\rho h$}
 We update the enthalpy analogously to the species update in 
-Section \ref{Advancing rhoX_k}.  The only difference is that we also 
-have a forcing term
+Section \ref{Advancing rhoX_k}.  The forcing term does not include reaction
+source terms accounted for in {\bf React State}.
 \begin{equation}
 \frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) = 
--\nabla\cdot\left\{\left[(\rho h)_0^{n+\half,{\rm avg}} 
+-\nabla\cdot\left\{\left[(\rho h)_0^{n+\myhalf,{\rm avg}} 
 + (\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\} + (\Ubt \cdot \er)\frac{\partial p_0}{\partial r} + \psi  \enskip .

### 2009-04-01 docs/eos_notes/eos_notes.tex
new section on eos calls and temperature usage

\section{EOS Calls}
\section{Temperature Usage}

### 2009-04-01 docs/eos_notes/eos_notes.tex
put my temperature notes in the document

@@
 \section{EOS Calls}
+\subsection{{\tt advance\_timestep}}
+\begin{description}
+\item[Step 1.] {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}\\
+
+\item[Step 2.] {\em Construct the provisional edge-based advective velocity, $\uadvone.$}\\
+
+\item[Step 3.] {\em React the full state and base state through the first time interval 
+of $\dt/2.$}\\
+
+\item[Step 4.] {\em Advect the base state, then the full state, through a time interval 
+of $\dt.$}\\
+
+\item[Step 5.] {\em React the full state through a second time interval of $\dt/2.$}\\
+
+\item[Step 6.] {\em Define a new average expansion rate at time $t^\nph.$}\\
+
+\item[Step 7.] {\em Construct the final edge-based advective velocity, $\uadvtwo.$}\\
+
+\item[Step 8.] {\em Advect the base state, then the full state, through a time interval 
+of $\dt.$}\\
+
+\item[Step 9.] {\em React the full state and base state through a second time interval 
+of $\dt/2.$}\\
+
+\item[Step 10.] {\em Compute $S^{n+1}$ for the final projection.}\\
+
+\item[Step 11.] {\em Update the velocity.}\\
+
+\item[Step 12.] {\em Compute a new $\dt.$}\\
+
+\end{description}
+\subsection{Initialization}
+\subsection{{\tt make\_plotfile}}
 \section{Temperature Usage}
+\subsection{{\tt advance\_timestep}}
+\begin{description}
+\item[Step 1.] {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}
+\item[Step 2.] {\em Construct the provisional edge-based advective velocity, $\uadvone.$}
+\item[Step 3.] {\em React the full state and base state through the first time interval 
+of $\dt/2.$}
+\begin{itemize}
+\item{\tt react\_state} uses $T^n$ in burner and then computes $T^1$.
+\end{itemize}
+\item[Step 4.] {\em Advect the base state, then the full state, through a time interval 
+of $\dt.$}\\
+If {\tt use\_thermal\_diffusion} = T:
+\begin{itemize}
+\item {\tt make\_explicit\_thermal} computes coefficients for 
+$\nabla\cdot\kth\nabla h + \cdots$ using $(\rho,T,X)^1$.
+\item {\tt thermal\_conduct\_full\_alg} computes coefficients with $(\rho,T,X)^1$, then 
+computes $T^{2,*}$ from $(\rho,h,X)^{2,*}$ or $(\rho,p,X)^{2,*}$ if {\tt use\_tfromp} = T.
+\end{itemize}
+If {\tt use\_thermal\_diffusion} = F:
+\begin{itemize}
+\item {\tt enthalpy\_advance} computes $T^{2,*}$ from $(\rho,h,X)^{2,*}$ or $(\rho,p,X)^{2,*}$
+if {\tt use\_tfromp} = T.
+\end{itemize}
+\item[Step 5.] {\em React the full state through a second time interval of $\dt/2.$}
+\begin{itemize}
+\item {\tt react\_state} uses $T^{2,*}$ in burner and then computes $T^{n+1,*}$.
+\end{itemize}
+\item[Step 6.] {\em Define a new average expansion rate at time $t^\nph.$}
+\begin{itemize}
+\item {make\_explicit\_thermal} computes coefficients for 
+$\nabla\cdot\kth\nabla h + \cdots$ using $(\rho,T,X)^{n+1,*}$.
+\item {make\_S} thermodynamic variables are functions of $(\rho,T,X)^{n+1,*}$.
+\end{itemize}
+\item[Step 7.] {\em Construct the final edge-based advective velocity, $\uadvtwo.$}
+\item[Step 8.] {\em Advect the base state, then the full state, through a time interval 
+of $\dt.$}\\
+If {\tt use\_thermal\_diffusion} = T:
+\begin{itemize}
+\item {\tt make\_explicit\_thermal} computes coefficients for 
+$\nabla\cdot\kth\nabla h + \cdots$ using $(\rho,T,X)^1$.
+\item {\tt thermal\_conduct\_full\_alg} computes coefficients with $(\rho,T,X)^1$ and
+$(\rho,T,X)^{2,*}$, then computes $T^{2}$ from $(\rho,h,X)^{2}$ or $(\rho,p,X)^{2}$ if 
+{\tt use\_tfromp} = T.
+\end{itemize}
+If {\tt use\_thermal\_diffusion} = F:
+\begin{itemize}
+\item {\tt enthalpy\_advance} computes $T^2$ from $(\rho,h,X)^2$ or $(\rho,p,X)^2$
+if {\tt use\_tfromp} = T.
+\end{itemize}
+\item[Step 9.] {\em React the full state and base state through a second time interval 
+of $\dt/2.$}
+\begin{itemize}
+\item {\tt react\_state} uses $T^2$ in burner and computes $T^{n+1}$.
+\end{itemize}
+\item[Step 10.] {\em Compute $S^{n+1}$ for the final projection.}
+\begin{itemize}
+\item {\tt make\_explicit\_thermal} computes coefficients for 
+$\nabla\cdot\kth\nabla h + \cdots$ using $(\rho,T,X)^{n+1}$.
+\item {\tt make\_S} thermodynamic variables are functions of $(\rho,T,X)^{n+1}$.
+\end{itemize}
+\item[Step 11.] {\em Update the velocity.}
+\item[Step 12.] {\em Compute a new $\dt.$}
+\end{description}

### 2009-04-01 docs/eos_notes/eos_notes.tex
progress on the eos section

@@
 \section{EOS Calls}
+\subsection{Initialization}
 \subsection{{\tt advance\_timestep}}
 \begin{description}
-\item[Step 1.] {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}\\
-
-\item[Step 2.] {\em Construct the provisional edge-based advective velocity, $\uadvone.$}\\
-
+\item[Step 1.] {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}
+\begin{itemize}
+\item In {\tt makePfromRhoH}, we compute a thermodynamic $p^n$ for the volume discrepancy 
+term using $(\rho,h,X)^n$.
+\end{itemize}
+\item[Step 2.] {\em Construct the provisional edge-based advective velocity, $\uadvone.$}
 \item[Step 3.] {\em React the full state and base state through the first time interval 
-of $\dt/2.$}\\
-
+of $\dt/2.$}
+\begin{itemize}
+\item In {\tt react\_state} $\rightarrow$ {\tt burner}, we compute $c_p$ and $\xi_k$ 
+for inputs to {\tt VODE} using $(\rho,T,X)^n$.
+\item In {\tt make\_gamma}, we compute $\Gamma_1$ using $(\rho,X)^{(1)}$ and $p_0^n$.
+\end{itemize}
 \item[Step 4.] {\em Advect the base state, then the full state, through a time interval 
-of $\dt.$}\\
-
+of $\dt.$}
+\begin{itemize}
+\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
+\item In {\tt enthalpy\_advance} $\rightarrow$ {\tt update\_scal}, we compute $h$ above
+the {\tt base\_cutoff\_density} using $(\rho,X)^{(2),*}$ and $p_0^{n+1,*}$.
+\item In {\tt thermal\_conduct}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
+\item In {\tt thermal\_conduct}, we compute $T^{(2),*}$ using $(\rho,h,X)^{(2),*}$
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{(2),*}$ and $p_0^{n+1,*}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
 \item[Step 5.] {\em React the full state through a second time interval of $\dt/2.$}\\
@@
 \end{description}
-\subsection{Initialization}
 \subsection{{\tt make\_plotfile}}
@@
 \begin{itemize}
-\item{\tt react\_state} uses $T^n$ in burner and then computes $T^1$.
+\item{\tt react\_state} uses $T^n$ in burner and then computes $T^{(1)}$.
 \end{itemize}
@@
 \item {\tt thermal\_conduct\_full\_alg} computes coefficients with $(\rho,T,X)^1$, then 
-computes $T^{2,*}$ from $(\rho,h,X)^{2,*}$ or $(\rho,p,X)^{2,*}$ if {\tt use\_tfromp} = T.
+computes $T^{(2),*}$ from $(\rho,h,X)^{(2),*}$ or $(\rho,p,X)^{(2),*}$ if {\tt use\_tfromp} = T.
 \end{itemize}
@@
 \begin{itemize}
-\item {\tt enthalpy\_advance} computes $T^{2,*}$ from $(\rho,h,X)^{2,*}$ or $(\rho,p,X)^{2,*}$
+\item {\tt enthalpy\_advance} computes $T^{(2),*}$ from $(\rho,h,X)^{(2),*}$ or $(\rho,p,X)^{(2),*}$
 if {\tt use\_tfromp} = T.
@@
 \begin{itemize}
-\item {\tt react\_state} uses $T^{2,*}$ in burner and then computes $T^{n+1,*}$.
+\item {\tt react\_state} uses $T^{(2),*}$ in burner and then computes $T^{n+1,*}$.
 \end{itemize}
@@
 \item {\tt thermal\_conduct\_full\_alg} computes coefficients with $(\rho,T,X)^1$ and
-$(\rho,T,X)^{2,*}$, then computes $T^{2}$ from $(\rho,h,X)^{2}$ or $(\rho,p,X)^{2}$ if 
+$(\rho,T,X)^{(2),*}$, then computes $T^{2}$ from $(\rho,h,X)^{2}$ or $(\rho,p,X)^{2}$ if 
 {\tt use\_tfromp} = T.
@@
 \begin{itemize}
-\item {\tt enthalpy\_advance} computes $T^2$ from $(\rho,h,X)^2$ or $(\rho,p,X)^2$
+\item {\tt enthalpy\_advance} computes $T^{(2)}$ from $(\rho,h,X)^2$ or $(\rho,p,X)^2$
 if {\tt use\_tfromp} = T.
@@
 \begin{itemize}
-\item {\tt react\_state} uses $T^2$ in burner and computes $T^{n+1}$.
+\item {\tt react\_state} uses $T^{(2)}$ in burner and computes $T^{n+1}$.
 \end{itemize}

### 2009-04-01 docs/eos_notes/eos_notes.tex
finished the list of eos calls in advance.f90

--- 2009-04-01 docs/eos_notes/eos_notes.tex
@@
 \section{EOS Calls}
+\subsection{Initialization}
+\subsection{{\tt advance\_timestep}}
+\begin{description}
+\item[Step 1.] {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}
+\begin{itemize}
+\item In {\tt makePfromRhoH}, we compute a thermodynamic $p^n$ for the volume discrepancy 
+term using $(\rho,h,X)^n$.
+\end{itemize}
+\item[Step 2.] {\em Construct the provisional edge-based advective velocity, $\uadvone.$}
+\item[Step 3.] {\em React the full state and base state through the first time interval 
+of $\dt/2.$}
+\begin{itemize}
+\item In {\tt react\_state} $\rightarrow$ {\tt burner}, we compute $c_p$ and $\xi_k$ 
+for inputs to {\tt VODE} using $(\rho,T,X)^n$.
+\item In {\tt react\_state}, we compute $T^{(1)}$ using $(\rho,h,X)^{(1)}$ 
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{(1)}$ and $p_0^n$ (if {\tt use\_tfromp}=T).
+\item In {\tt make\_gamma}, we compute $\Gamma_1$ using $(\rho,X)^{(1)}$ and $p_0^n$.
+\end{itemize}
+\item[Step 4.] {\em Advect the base state, then the full state, through a time interval 
+of $\dt.$}\\
+If {\tt use\_thermal\_diffusion} = T:
+\begin{itemize}
+\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
+\item In {\tt enthalpy\_advance} $\rightarrow$ {\tt update\_scal}, we compute $h$ above
+the {\tt base\_cutoff\_density} using $(\rho,X)^{(2),*}$ and $p_0^{n+1,*}$.
+\item In {\tt thermal\_conduct}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
+\item In {\tt thermal\_conduct}, we compute $T^{(2),*}$ using $(\rho,h,X)^{(2),*}$
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{(2),*}$ and $p_0^{n+1,*}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+If {\tt use\_thermal\_diffusion} = F:
+\begin{itemize}
+\item In {\tt enthalpy\_advance} $\rightarrow$ {\tt update\_scal}, we compute $h$ above
+the {\tt base\_cutoff\_density} using $(\rho,X)^{(2),*}$ and $p_0^{n+1,*}$.
+\item In {\tt enthalpy\_advance}, we compute $T^{(2),*}$ using $(\rho,h,X)^{(2),*}$
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{(2),*}$ and $p_0^{n+1,*}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+\item[Step 5.] {\em React the full state through a second time interval of $\dt/2.$}
+\begin{itemize}
+\item In {\tt react\_state} $\rightarrow$ {\tt burner}, we compute $c_p$ and $\xi_k$ 
+for inputs to {\tt VODE} using $(\rho,T,X)^{(2),*}$.
+\item In {\tt react\_state}, we compute $T^{n+1,*}$ using $(\rho,h,X)^{n+1,*}$ 
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{n+1,*}$ and $p_0^{n+1,*}$ (if {\tt use\_tfromp}=T).
+\item In {\tt make\_gamma}, we compute $\Gamma_1$ using $(\rho,X)^{n+1,*}$ and $p_0^{n+1,*}$.
+\end{itemize}
+\item[Step 6.] {\em Define a new average expansion rate at time $t^\nph.$}
+\begin{itemize}
+\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{n+1,*}$.
+\item In {\tt make\_S}, we compute thermodynamic variables using $(\rho,T,X)^{n+1,*}$.
+\item In {\tt makePfromRhoH}, we compute a thermodynamic $p^{n+1,*}$ for the volume 
+discrepancy term using $(\rho,h,X)^{n+1,*}$.
+\end{itemize}
+\item[Step 7.] {\em Construct the final edge-based advective velocity, $\uadvtwo.$}
+\item[Step 8.] {\em Advect the base state, then the full state, through a time interval 
+of $\dt.$}\\
+If {\tt use\_thermal\_diffusion} = T:
+\begin{itemize}
+\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
+\item In {\tt enthalpy\_advance} $\rightarrow$ {\tt update\_scal}, we compute $h$ above
+the {\tt base\_cutoff\_density} using $(\rho,X)^{(2)}$ and $p_0^{n+1}$.
+\item In {\tt thermal\_conduct}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$ and $(\rho,T,X)^{(2),*}$.
+\item In {\tt thermal\_conduct}, we compute $T^{(2)}$ using $(\rho,h,X)^{(2)}$
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{(2)}$ and $p_0^{n+1}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+If {\tt use\_thermal\_diffusion} = F:
+\begin{itemize}
+\item In {\tt enthalpy\_advance} $\rightarrow$ {\tt update\_scal}, we compute $h$ above
+the {\tt base\_cutoff\_density} using $(\rho,X)^{(2)}$ and $p_0^{n+1}$.
+\item In {\tt enthalpy\_advance}, we compute $T^{(2)}$ using $(\rho,h,X)^{(2)}$
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{(2)}$ and $p_0^{n+1}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+\item[Step 9.] {\em React the full state and base state through a second time interval 
+of $\dt/2.$}
+\begin{itemize}
+\item In {\tt react\_state} $\rightarrow$ {\tt burner}, we compute $c_p$ and $\xi_k$ 
+for inputs to {\tt VODE} using $(\rho,T,X)^{(2)}$.
+\item In {\tt react\_state}, we compute $T^{n+1}$ using $(\rho,h,X)^{n+1}$ 
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{n+1}$ and $p_0^{n+1}$ (if {\tt use\_tfromp}=T).
+\item In {\tt make\_gamma}, we compute $\Gamma_1$ using $(\rho,X)^{n+1}$ and $p_0^{n+1}$.
+\end{itemize}
+\item[Step 10.] {\em Compute $S^{n+1}$ for the final projection.}
+\begin{itemize}
+\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{n+1}$.
+\item In {\tt make\_S}, we compute thermodynamic variables using $(\rho,T,X)^{n+1}$.
+\end{itemize}
+\item[Step 11.] {\em Update the velocity.}
+\item[Step 12.] {\em Compute a new $\dt.$}
+\end{description}
+\subsection{{\tt make\_plotfile}}
 \section{Temperature Usage}
+\subsection{{\tt advance\_timestep}}
+\begin{description}
+\item[Step 1.] {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}
+\item[Step 2.] {\em Construct the provisional edge-based advective velocity, $\uadvone.$}
+\item[Step 3.] {\em React the full state and base state through the first time interval 
+of $\dt/2.$}
+\begin{itemize}
+\item In {\tt react\_state} $\rightarrow$ {\tt burner}, we compute $c_p$ and $\xi_k$ 
+for inputs to {\tt VODE} using $(\rho,T,X)^n$.
+\item In {\tt react\_state}, we compute $T^{(1)}$ using $(\rho,h,X)^{(1)}$ 
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{(1)}$ and $p_0^n$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+\item[Step 4.] {\em Advect the base state, then the full state, through a time interval 
+of $\dt.$}\\
+If {\tt use\_thermal\_diffusion} = T:
+\begin{itemize}
+\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
+\item In {\tt thermal\_conduct}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
+\item In {\tt thermal\_conduct}, we compute $T^{(2),*}$ using $(\rho,h,X)^{(2),*}$
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{(2),*}$ and $p_0^{n+1,*}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+If {\tt use\_thermal\_diffusion} = F:
+\begin{itemize}
+\item In {\tt enthalpy\_advance}, we compute $T^{(2),*}$ using $(\rho,h,X)^{(2),*}$
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{(2),*}$ and $p_0^{n+1,*}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+\item[Step 5.] {\em React the full state through a second time interval of $\dt/2.$}
+\begin{itemize}
+\item In {\tt react\_state} $\rightarrow$ {\tt burner}, we compute $c_p$ and $\xi_k$ 
+for inputs to {\tt VODE} using $(\rho,T,X)^{(2),*}$.
+\item In {\tt react\_state}, we compute $T^{n+1,*}$ using $(\rho,h,X)^{n+1,*}$ 
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{n+1,*}$ and $p_0^{n+1,*}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+\item[Step 6.] {\em Define a new average expansion rate at time $t^\nph.$}
+\begin{itemize}
+\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{n+1,*}$.
+\item In {\tt make\_S}, we compute thermodynamic variables using $(\rho,T,X)^{n+1,*}$.
+\end{itemize}
+\item[Step 7.] {\em Construct the final edge-based advective velocity, $\uadvtwo.$}
+\item[Step 8.] {\em Advect the base state, then the full state, through a time interval 
+of $\dt.$}\\
+If {\tt use\_thermal\_diffusion} = T:
+\begin{itemize}
+\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
+\item In {\tt thermal\_conduct}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$ and $(\rho,T,X)^{(2),*}$.
+\item In {\tt thermal\_conduct}, we compute $T^{(2)}$ using $(\rho,h,X)^{(2)}$
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{(2)}$ and $p_0^{n+1}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+If {\tt use\_thermal\_diffusion} = F:
+\begin{itemize}
+\item In {\tt enthalpy\_advance}, we compute $T^{(2)}$ using $(\rho,h,X)^{(2)}$
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{(2)}$ and $p_0^{n+1}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+\item[Step 9.] {\em React the full state and base state through a second time interval 
+of $\dt/2.$}
+\begin{itemize}
+\item In {\tt react\_state} $\rightarrow$ {\tt burner}, we compute $c_p$ and $\xi_k$ 
+for inputs to {\tt VODE} using $(\rho,T,X)^{(2)}$.
+\item In {\tt react\_state}, we compute $T^{n+1}$ using $(\rho,h,X)^{n+1}$ 
+(if {\tt use\_tfromp}=F) or $(\rho,X)^{n+1}$ and $p_0^{n+1}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+\item[Step 10.] {\em Compute $S^{n+1}$ for the final projection.}
+\begin{itemize}
+\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{n+1}$.
+\item In {\tt make\_S}, we compute thermodynamic variables using $(\rho,T,X)^{n+1}$.
+\end{itemize}
+\item[Step 11.] {\em Update the velocity.}
+\item[Step 12.] {\em Compute a new $\dt.$}
+\end{description}

### 2009-04-02 docs/eos_notes/eos_notes.tex
missed one case added flag logic

@@
 \begin{description}
-\item[Step 1.] {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}
+\item[Step 1.] {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}\\ \\
+if {\tt dpdt\_factor} $>$ 0 then
 \begin{itemize}
@@
 \end{itemize}
+end if
 \item[Step 2.] {\em Construct the provisional edge-based advective velocity, $\uadvone.$}
@@
 (if {\tt use\_tfromp}=F) or $(\rho,X)^{(1)}$ and $p_0^n$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+if {\tt evolve\_base\_state} then
+\begin{itemize}
 \item In {\tt make\_gamma}, we compute $\Gamma_1$ using $(\rho,X)^{(1)}$ and $p_0^n$.
 \end{itemize}
+end if
 \item[Step 4.] {\em Advect the base state, then the full state, through a time interval 
-of $\dt.$}\\
-If {\tt use\_thermal\_diffusion} = T:
+of $\dt.$}\\ \\
+if {\tt use\_thermal\_diffusion} = T then
 \begin{itemize}
@@
 \end{itemize}
-If {\tt use\_thermal\_diffusion} = F:
+else
 \begin{itemize}
@@
 \end{itemize}
+end if
 \item[Step 5.] {\em React the full state through a second time interval of $\dt/2.$}
@@
 (if {\tt use\_tfromp}=F) or $(\rho,X)^{n+1,*}$ and $p_0^{n+1,*}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+if {\tt evolve\_base\_state} then
+\begin{itemize}
 \item In {\tt make\_gamma}, we compute $\Gamma_1$ using $(\rho,X)^{n+1,*}$ and $p_0^{n+1,*}$.
 \end{itemize}
-\item[Step 6.] {\em Define a new average expansion rate at time $t^\nph.$}
+end if
+\item[Step 6.] {\em Define a new average expansion rate at time $t^\nph.$}\\ \\
+if {\tt use\_thermal\_diffusion} then
 \begin{itemize}
@@
 $\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{n+1,*}$.
+\end{itemize}
+end if
+\begin{itemize}
 \item In {\tt make\_S}, we compute thermodynamic variables using $(\rho,T,X)^{n+1,*}$.
+\end{itemize}
+if {\tt dpdt\_factor} $>$ 0 then
+\begin{itemize}
 \item In {\tt makePfromRhoH}, we compute a thermodynamic $p^{n+1,*}$ for the volume 
@@
 \end{itemize}
+end if
 \item[Step 7.] {\em Construct the final edge-based advective velocity, $\uadvtwo.$}
 \item[Step 8.] {\em Advect the base state, then the full state, through a time interval 
-of $\dt.$}\\
-If {\tt use\_thermal\_diffusion} = T:
+of $\dt.$}\\ \\
+if {\tt use\_thermal\_diffusion} = T then
 \begin{itemize}
@@
 \end{itemize}
+end if
 \item[Step 9.] {\em React the full state and base state through a second time interval 
@@
 (if {\tt use\_tfromp}=F) or $(\rho,X)^{n+1}$ and $p_0^{n+1}$ (if {\tt use\_tfromp}=T).
+\end{itemize}
+if {\tt evolve\_base\_state} then
+\begin{itemize}
 \item In {\tt make\_gamma}, we compute $\Gamma_1$ using $(\rho,X)^{n+1}$ and $p_0^{n+1}$.
 \end{itemize}
-\item[Step 10.] {\em Compute $S^{n+1}$ for the final projection.}
+end if
+\item[Step 10.] {\em Compute $S^{n+1}$ for the final projection.}\\ \\
+if {\tt make\_explicit\_thermal} then
 \begin{itemize}
@@
 $\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{n+1}$.
+\end{itemize}
+end if
+\begin{itemize}
 \item In {\tt make\_S}, we compute thermodynamic variables using $(\rho,T,X)^{n+1}$.
 \end{itemize}
-\item[Step 11.] {\em Update the velocity.}
+\item[Step 11.] {\em Update the velocity.}\\ \\
+if {\tt dpdt\_factor} $>$ 0 then
+\begin{itemize}
+\item In {\tt makePfromRhoH}, we compute a thermodynamic $p^{n+1}$ for the volume 
+discrepancy term using $(\rho,h,X)^{n+1}$.
+\end{itemize}
+end if
 \item[Step 12.] {\em Compute a new $\dt.$}
@@
 \item[Step 4.] {\em Advect the base state, then the full state, through a time interval 
-of $\dt.$}\\
-If {\tt use\_thermal\_diffusion} = T:
+of $\dt.$}\\ \\
+if {\tt use\_thermal\_diffusion} = T then
 \begin{itemize}
@@
 \end{itemize}
-\item[Step 6.] {\em Define a new average expansion rate at time $t^\nph.$}
+\item[Step 6.] {\em Define a new average expansion rate at time $t^\nph.$}\\ \\
+if {\tt use\_thermal\_diffusion} then
 \begin{itemize}
@@
 $\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{n+1,*}$.
+\end{itemize}
+end if
+\begin{itemize}
 \item In {\tt make\_S}, we compute thermodynamic variables using $(\rho,T,X)^{n+1,*}$.
@@
 \item[Step 8.] {\em Advect the base state, then the full state, through a time interval 
-of $\dt.$}\\
-If {\tt use\_thermal\_diffusion} = T:
+of $\dt.$}\\ \\
+if {\tt use\_thermal\_diffusion} = T then
 \begin{itemize}
@@
 \end{itemize}
-\item[Step 10.] {\em Compute $S^{n+1}$ for the final projection.}
+\item[Step 10.] {\em Compute $S^{n+1}$ for the final projection.}\\ \\
+if {\tt make\_explicit\_thermal} then
 \begin{itemize}
@@
 $\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{n+1}$.
+\end{itemize}
+end if
+\begin{itemize}
 \item In {\tt make\_S}, we compute thermodynamic variables using $(\rho,T,X)^{n+1}$.

### 2009-04-02 Source/advance.f90
removed the call to make_explicit_thermal in step 8, and instead store the results from the call to make_explicit_thermal in step 4 since they were computing exactly the same thing.  Cuts down on another eos call with no changes to results or peak memory usage

@@ -98,7 +98,8 @@ contains
     type(multifab) ::             macphi(mla%nlevel)
     type(multifab) ::          hgrhs_old(mla%nlevel)
     type(multifab) ::         Source_nph(mla%nlevel)
-    type(multifab) ::            thermal(mla%nlevel)
+    type(multifab) ::           thermal1(mla%nlevel)
+    type(multifab) ::           thermal2(mla%nlevel)
     type(multifab) ::             s2star(mla%nlevel)
     type(multifab) ::                 s1(mla%nlevel)
     type(multifab) ::                 s2(mla%nlevel)
@@ -411,15 +412,15 @@ contains
     end if
 
     do n=1,nlevs
-       call multifab_build(thermal(n), mla%la(n), 1, 1)
+       call multifab_build(thermal1(n), mla%la(n), 1, 1)
     end do
     
     ! thermal is the forcing for rhoh or temperature
     if(use_thermal_diffusion) then
-       call make_explicit_thermal(mla,dx,thermal,s1,p0_old,the_bc_tower)
+       call make_explicit_thermal(mla,dx,thermal1,s1,p0_old,the_bc_tower)
     else
        do n=1,nlevs
-          call setval(thermal(n),ZERO,all=.true.)
+          call setval(thermal1(n),ZERO,all=.true.)
        end do
     end if
 
@@ -506,7 +507,7 @@ contains
        write(6,*) '            : enthalpy_advance >>> '
     end if
 
-    call enthalpy_advance(mla,1,uold,s1,s2,sedge,sflux,scal_force,thermal,umac,w0,w0mac, &
+    call enthalpy_advance(mla,1,uold,s1,s2,sedge,sflux,scal_force,thermal1,umac,w0,w0mac, &
                           normal,rho0_old,rhoh0_old,rho0_new,rhoh0_new,p0_old,p0_new, &
                           tempbar,psi,dx,dt,the_bc_tower%bc_tower_array)
 
@@ -517,7 +518,6 @@ contains
              call destroy(umac(n,comp))
        end do
        call destroy(scal_force(n))
-       call destroy(thermal(n))
     end do
 
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
@@ -585,14 +585,14 @@ contains
     end if
 
     do n=1,nlevs
[…]

### 2009-04-02 docs/eos_notes/eos_notes.tex
removed the call to make_explicit_thermal in step 8, and instead store the results from the call to make_explicit_thermal in step 4 since they were computing exactly the same thing.  Cuts down on another eos call with no changes to results or peak memory usage
consolidated calls to make_thermal_coeffs so we don't repeat the same calculations twice.  cuts about 7% off runtime

@@
 \begin{itemize}
-\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
+\item In {\tt advance} before {\tt make\_explicit\_thermal}, we compute the coefficients for 
 $\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
@@
 the {\tt base\_cutoff\_density} using $(\rho,X)^{(2),*}$ and $p_0^{n+1,*}$.
-\item In {\tt thermal\_conduct}, we compute the coefficients for 
-$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
 \item In {\tt thermal\_conduct}, we compute $T^{(2),*}$ using $(\rho,h,X)^{(2),*}$
@@
 \begin{itemize}
-\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
+\item In {\tt advance} before {\tt make\_explicit\_thermal}, we compute the coefficients for 
 $\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{n+1,*}$.
@@
 \begin{itemize}
-\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
-$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
 \item In {\tt enthalpy\_advance} $\rightarrow$ {\tt update\_scal}, we compute $h$ above
 the {\tt base\_cutoff\_density} using $(\rho,X)^{(2)}$ and $p_0^{n+1}$.
-\item In {\tt thermal\_conduct}, we compute the coefficients for 
-$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$ and $(\rho,T,X)^{(2),*}$.
+\item In {\tt advance} before {\tt thermal\_conduct}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(2),*}$.
 \item In {\tt thermal\_conduct}, we compute $T^{(2)}$ using $(\rho,h,X)^{(2)}$
@@
 \begin{itemize}
-\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
+\item In {\tt advance} before {\tt make\_explicit\_thermal}, we compute the coefficients for 
 $\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{n+1}$.
@@
 \begin{itemize}
-\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
-$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
-\item In {\tt thermal\_conduct}, we compute the coefficients for 
+\item In {\tt advance} before {\tt make\_explicit\_thermal}, we compute the coefficients for 
 $\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
@@
 \begin{itemize}
-\item In {\tt make\_explicit\_thermal}, we compute the coefficients for 
-$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$.
-\item In {\tt thermal\_conduct}, we compute the coefficients for 
-$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(1)}$ and $(\rho,T,X)^{(2),*}$.
+\item In {\tt advance} before {\tt thermal\_conduct}, we compute the coefficients for 
+$\nabla\cdot(\kth/c_p)\nabla h + \cdots$ using $(\rho,T,X)^{(2),*}$.
 \item In {\tt thermal\_conduct}, we compute $T^{(2)}$ using $(\rho,h,X)^{(2)}$

### 2009-04-02 Source/advance.f90

@@ -116,6 +116,14 @@ contains
     type(multifab) ::         ptherm_new(mla%nlevel)
     type(multifab) ::     pthermbar_cart(mla%nlevel)
     type(multifab) ::       delta_p_term(mla%nlevel)
+    type(multifab) ::            Tcoeff1(mla%nlevel)
+    type(multifab) ::            hcoeff1(mla%nlevel)
+    type(multifab) ::           Xkcoeff1(mla%nlevel)
+    type(multifab) ::            pcoeff1(mla%nlevel)
+    type(multifab) ::            Tcoeff2(mla%nlevel)
+    type(multifab) ::            hcoeff2(mla%nlevel)
+    type(multifab) ::           Xkcoeff2(mla%nlevel)
+    type(multifab) ::            pcoeff2(mla%nlevel)
     type(multifab) ::         scal_force(mla%nlevel)
     type(multifab) ::              w0mac(mla%nlevel,dm)
     type(multifab) ::               umac(mla%nlevel,dm)
@@ -417,7 +425,23 @@ contains
     
     ! thermal is the forcing for rhoh or temperature
     if(use_thermal_diffusion) then
-       call make_explicit_thermal(mla,dx,thermal1,s1,p0_old,the_bc_tower)
+
+       do n=1,nlevs
+          call multifab_build(Tcoeff1(n),  mla%la(n), 1,     1)
+          call multifab_build(hcoeff1(n),  mla%la(n), 1,     1)
+          call multifab_build(Xkcoeff1(n), mla%la(n), nspec, 1)
+          call multifab_build(pcoeff1(n),  mla%la(n), 1,     1)
+       end do
+
+       call make_thermal_coeffs(s1,Tcoeff1,hcoeff1,Xkcoeff1,pcoeff1)
+
+       call make_explicit_thermal(mla,dx,thermal1,s1,Tcoeff1,hcoeff1,Xkcoeff1,pcoeff1, &
+                                  p0_old,the_bc_tower)
+
+       do n=1,nlevs
+          call destroy(Tcoeff1(n))
+       end do
+
     else
        do n=1,nlevs
           call setval(thermal1(n),ZERO,all=.true.)
@@ -529,7 +553,8 @@ contains
           write(6,*) '<<< STEP  4a: thermal conduct >>>'
        end if
 
-       call thermal_conduct(mla,dx,dt,s1,s1,s2,p0_old,p0_new,the_bc_tower)
[…]

### 2009-04-02 Source/divu_iter.f90

@@ -68,6 +68,10 @@ contains
     type(multifab) :: rho_Hext(nlevs)
     type(multifab) :: div_coeff_3d(nlevs)
     type(multifab) :: w0mac(nlevs,dm)
+    type(multifab) :: Tcoeff(nlevs)
+    type(multifab) :: hcoeff(nlevs)
+    type(multifab) :: Xkcoeff(nlevs)
+    type(multifab) :: pcoeff(nlevs)
 
     real(dp_t) ::            etarho_ec(nlevs_radial,0:nr_fine)
     real(dp_t) ::                 Sbar(nlevs_radial,0:nr_fine-1)
@@ -115,7 +119,26 @@ contains
     end do
 
     if(use_thermal_diffusion) then
-       call make_explicit_thermal(mla,dx,thermal,sold,p0_old,the_bc_tower)
+
+       do n=1,nlevs
+          call multifab_build(Tcoeff(n),  mla%la(n), 1,     1)
+          call multifab_build(hcoeff(n),  mla%la(n), 1,     1)
+          call multifab_build(Xkcoeff(n), mla%la(n), nspec, 1)
+          call multifab_build(pcoeff(n),  mla%la(n), 1,     1)
+       end do
+
+       call make_thermal_coeffs(sold,Tcoeff,hcoeff,Xkcoeff,pcoeff)
+
+       call make_explicit_thermal(mla,dx,thermal,sold,Tcoeff,hcoeff,Xkcoeff,pcoeff, &
+                                  p0_old,the_bc_tower)
+
+       do n=1,nlevs
+          call destroy(Tcoeff(n))
+          call destroy(hcoeff(n))
+          call destroy(Xkcoeff(n))
+          call destroy(pcoeff(n))
+       end do
+
     else
        do n=1,nlevs
           call setval(thermal(n), ZERO, all=.true.)

### 2009-04-02 Source/initial_proj.f90

@@ -52,6 +52,10 @@ contains
     type(multifab) :: rho_Hnuc1(nlevs)
     type(multifab) :: rho_Hext(nlevs)
     type(multifab) :: div_coeff_3d(nlevs)
+    type(multifab) :: Tcoeff(nlevs)
+    type(multifab) :: hcoeff(nlevs)
+    type(multifab) :: Xkcoeff(nlevs)
+    type(multifab) :: pcoeff(nlevs)
 
     real(dp_t) ::                  psi(nlevs_radial,0:nr_fine-1)
     real(dp_t) ::                 Sbar(nlevs_radial,0:nr_fine-1)
@@ -70,7 +74,26 @@ contains
     end do
     
     if(use_thermal_diffusion) then
-       call make_explicit_thermal(mla,dx,thermal,sold,p0,the_bc_tower)
+
+       do n=1,nlevs
+          call multifab_build(Tcoeff(n),  mla%la(n), 1,     1)
+          call multifab_build(hcoeff(n),  mla%la(n), 1,     1)
+          call multifab_build(Xkcoeff(n), mla%la(n), nspec, 1)
+          call multifab_build(pcoeff(n),  mla%la(n), 1,     1)
+       end do
+
+       call make_thermal_coeffs(sold,Tcoeff,hcoeff,Xkcoeff,pcoeff)
+
+       call make_explicit_thermal(mla,dx,thermal,sold,Tcoeff,hcoeff,Xkcoeff,pcoeff, &
+                                  p0,the_bc_tower)
+
+       do n=1,nlevs
+          call destroy(Tcoeff(n))
+          call destroy(hcoeff(n))
+          call destroy(Xkcoeff(n))
+          call destroy(pcoeff(n))
+       end do
+
     else
        do n=1,nlevs
           call setval(thermal(n), ZERO, all=.true.)

### 2009-04-02 Source/make_explicit_thermal.f90

@@ -19,7 +19,8 @@ contains
 ! Otherwise, compute thermal with grad h + grad X_k + grad p_0 formulation
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
 
-  subroutine make_explicit_thermal(mla,dx,thermal,s,p0,the_bc_tower)
+  subroutine make_explicit_thermal(mla,dx,thermal,s,Tcoeff,hcoeff,Xkcoeff,pcoeff, &
+                                   p0,the_bc_tower)
 
     use bc_module
     use bl_prof_module
@@ -39,19 +40,18 @@ contains
     real(dp_t)     , intent(in   ) :: dx(:,:)
     type(multifab) , intent(inout) :: thermal(:)
     type(multifab) , intent(in   ) :: s(:)
+    type(multifab) , intent(in   ) :: Tcoeff(:)
+    type(multifab) , intent(in   ) :: hcoeff(:)
+    type(multifab) , intent(in   ) :: Xkcoeff(:)
+    type(multifab) , intent(in   ) :: pcoeff(:)
     real(kind=dp_t), intent(in   ) :: p0(:,0:)
     type(bc_tower) , intent(in   ) :: the_bc_tower
 
     ! Local
-    type(multifab) :: phi(mla%nlevel),alpha(mla%nlevel),beta(mla%nlevel),Xkcoeff(mla%nlevel)
-    type(multifab) :: Tcoeff(mla%nlevel),hcoeff(mla%nlevel),pcoeff(mla%nlevel)
+    type(multifab) :: phi(mla%nlevel),alpha(mla%nlevel),beta(mla%nlevel)
     type(multifab) :: resid(mla%nlevel)
 
     integer                     :: i,comp,n,stencil_order
-    integer                     :: lo(dm),hi(dm)
-    real(kind=dp_t), pointer    :: betap(:,:,:,:)
-    real(kind=dp_t), pointer    :: Tcoeffp(:,:,:,:),hcoeffp(:,:,:,:)
-    real(kind=dp_t), pointer    :: Xkcoeffp(:,:,:,:),pcoeffp(:,:,:,:)
 
     type(bl_prof_timer), save :: bpt
 
@@ -60,22 +60,10 @@ contains
     stencil_order = 2
 
     do n=1,nlevs
[…]

### 2009-04-02 Source/thermal_conduct.f90

@@ -20,7 +20,8 @@ contains
   ! Crank-Nicholson solve for enthalpy, taking into account only the
   ! enthalpy-diffusion terms in the temperature conduction term.
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-  subroutine thermal_conduct(mla,dx,dt,s1,s_for_new_coeff,s2,p0_old,p0_new,the_bc_tower)
+  subroutine thermal_conduct(mla,dx,dt,s1,hcoeff1,Xkcoeff1,pcoeff1, &
+                             hcoeff2,Xkcoeff2,pcoeff2,s2,p0_old,p0_new,the_bc_tower)
 
     use variables, only: foextrap_comp, rho_comp, spec_comp, rhoh_comp
     use macproject_module
@@ -37,7 +38,12 @@ contains
     type(ml_layout), intent(inout) :: mla
     real(dp_t)     , intent(in   ) :: dx(:,:),dt
     type(multifab) , intent(in   ) :: s1(:)
-    type(multifab) , intent(in   ) :: s_for_new_coeff(:)
+    type(multifab) , intent(in   ) :: hcoeff1(:)
+    type(multifab) , intent(in   ) :: Xkcoeff1(:)
+    type(multifab) , intent(in   ) :: pcoeff1(:)
+    type(multifab) , intent(in   ) :: hcoeff2(:)
+    type(multifab) , intent(in   ) :: Xkcoeff2(:)
+    type(multifab) , intent(in   ) :: pcoeff2(:)
     type(multifab) , intent(inout) :: s2(:)
     real(kind=dp_t), intent(in   ) :: p0_old(:,0:),p0_new(:,0:)
     type(bc_tower) , intent(in   ) :: the_bc_tower
@@ -46,17 +52,10 @@ contains
     type(multifab) :: rhsalpha(mla%nlevel),lhsalpha(mla%nlevel)
     type(multifab) :: rhsbeta(mla%nlevel),lhsbeta(mla%nlevel)
     type(multifab) :: phi(mla%nlevel),Lphi(mla%nlevel),rhs(mla%nlevel)
-    type(multifab) :: hcoeff1(mla%nlevel),hcoeff2(mla%nlevel)
[…]

### 2009-04-10 docs/runtime_parameters/runtime_parameters.tex
an automatically generated skeleton of a runtime parameters table

\section{Runtime Parameters}

%%%%%%%%%%%%%%%%
% symbol table
%%%%%%%%%%%%%%%%

\renewcommand{\arraystretch}{1.5}
%
% note: list generated via:
% grep namelist probin.f90 | awk '{print ``\\verb= ``, $3, `` = &   &   \\\\"}' | sort

\begin{center}
\begin{longtable}{|l|p{3.0in}|l|}
\caption[runtime parameters]{runtime parameters.} \label{table:runtime} \\
%
\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
       \multicolumn{1}{ c|}{\textbf{description}} & 
       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\tablename\ \thetable{}---continued}} \\
\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
       \multicolumn{1}{ c|}{\textbf{description}} & 
       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
\endhead

\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
\endfoot

\hline 
\endlastfoot

\verb=  anelastic_cutoff  = &   &   \\
\verb=  base_cutoff_density  = &   &   \\
\verb=  bcx_hi  = &   &   \\
\verb=  bcx_lo  = &   &   \\
\verb=  bcy_hi  = &   &   \\
\verb=  bcy_lo  = &   &   \\
\verb=  bcz_hi  = &   &   \\
\verb=  bcz_lo  = &   &   \\
\verb=  beta_type  = &   &   \\
\verb=  cflfac  = &   &   \\
\verb=  cg_verbose  = &   &   \\
\verb=  check_base_name  = &   &   \\
\verb=  chk_int  = &   &   \\
\verb=  co_latitude  = &   &   \\
\verb=  dm_in  = &   &   \\
\verb=  do_alltoallv  = &   &   \\
\verb=  do_burning  = &   &   \\
\verb=  do_eos_h_above_cutoff  = &   &   \\
\verb=  do_initial_projection  = &   &   \\
\verb=  do_sponge  = &   &   \\
\verb=  dpdt_factor  = &   &   \\
\verb=  drdxfac  = &   &   \\
\verb=  enthalpy_pred_type  = &   &   \\
\verb=  evolve_base_state  = &   &   \\
\verb=  fixed_dt  = &   &   \\
\verb=  grav_const  = &   &   \\
\verb=  hg_bottom_solver  = &   &   \\
\verb=  hg_dense_stencil  = &   &   \\
\verb=  init_divu_iter  = &   &   \\
\verb=  init_iter  = &   &   \\
\verb=  init_shrink  = &   &   \\
\verb=  interp_type_radial_bin_to_cart  = &   &   \\
\verb=  job_name  = &   &   \\
\verb=  lUsingNFiles  = &   &   \\
\verb=  max_dt_growth  = &   &   \\
\verb=  max_grid_size  = &   &   \\
\verb=  max_levs  = &   &   \\
\verb=  max_step  = &   &   \\
\verb=  mg_bottom_solver  = &   &   \\
\verb=  mg_verbose  = &   &   \\
\verb=  min_eff  = &   &   \\
\verb=  min_width  = &   &   \\
\verb=  model_file  = &   &   \\
\verb=  n_cellx  = &   &   \\
\verb=  n_celly  = &   &   \\
\verb=  n_cellz  = &   &   \\
\verb=  nOutFiles  = &   &   \\
\verb=  perturb_model  = &   &   \\
\verb=  plot_base  = &   &   \\
\verb=  plot_base_name  = &   &   \\
\verb=  plot_deltat  = &   &   \\
\verb=  plot_int  = &   &   \\
\verb=  plot_spec  = &   &   \\
\verb=  plot_trac  = &   &   \\
\verb=  pmask_x  = &   &   \\
\verb=  pmask_xyz  = &   &   \\
\verb=  pmask_y  = &   &   \\
\verb=  pmask_z  = &   &   \\
\verb=  prob_hi_x  = &   &   \\
\verb=  prob_hi_y  = &   &   \\
\verb=  prob_hi_z  = &   &   \\
\verb=  prob_lo_x  = &   &   \\
\verb=  prob_lo_y  = &   &   \\
\verb=  prob_lo_z  = &   &   \\
\verb=  radius  = &   &   \\
\verb=  ref_ratio  = &   &   \\
\verb=  regrid_int  = &   &   \\
\verb=  restart  = &   &   \\
\verb=  rotational_frequency  = &   &   \\
\verb=  single_prec_plotfiles  = &   &   \\
\verb=  slope_order  = &   &   \\
\verb=  small_dens  = &   &   \\
\verb=  small_temp  = &   &   \\
\verb=  spherical_in  = &   &   \\
\verb=  sponge_kappa  = &   &   \\
\verb=  stop_time  = &   &   \\
\verb=  temp_diffusion_formulation  = &   &   \\
\verb=  test_set  = &   &   \\
\verb=  the_knapsack_verbosity  = &   &   \\
\verb=  thermal_diffusion_type  = &   &   \\
\verb=  the_sfc_threshold  = &   &   \\
\verb=  use_delta_gamma1_term  = &   &   \\
\verb=  use_divu_firstdt  = &   &   \\
\verb=  use_eos_coulomb  = &   &   \\
\verb=  use_etarho  = &   &   \\
\verb=  use_ppm  = &   &   \\
\verb=  use_soundspeed_firstdt  = &   &   \\
\verb=  use_tfromp  = &   &   \\
\verb=  use_thermal_diffusion  = &   &   \\
\verb=  velpert_amplitude  = &   &   \\
\verb=  velpert_radius  = &   &   \\
\verb=  velpert_scale  = &   &   \\
\verb=  velpert_steep  = &   &   \\
\verb=  verbose  = &   &   \\
\verb=  w0mac_interp_type  = &   &   \\
\verb=  xrb_pert_factor  = &   &   \\
\verb=  xrb_pert_size  = &   &   \\
\verb=  xrb_pert_type  = &   &   \\
\verb=  xrb_use_bottom_sponge  = &   &   \\

\end{longtable}
\end{center}
%

### 2009-04-10 docs/runtime_parameters/rp.sh
a simple script that attempts to make the table and list the defaults

@@ -1,0 +1,11 @@
+#!/bin/ksh
+
+parameters=$(grep namelist probin.f90 | awk '{print $3}' | sort)
+
+for i in ${parameters}
+do
+    value=$(grep $i probin.f90 | grep = | grep -v "len=" | cut -d'=' -f2)
+
+    echo "\\verb= " $i "=" " &   & " ${value}
+done
+

### 2009-04-10 docs/runtime_parameters/rp.sh
rp.sh now includes the latex header info to generate the table fixed a typo in previous version updated the tex file to reflect the output of rp.sh

@@ -1,11 +1,55 @@
 #!/bin/ksh
 
+# dump the header
+cat <<EOF
+
+\section{Runtime Parameters}
+
+%%%%%%%%%%%%%%%%
+% symbol table
+%%%%%%%%%%%%%%%%
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{3.0in}|l|}
+\caption[runtime parameters]{runtime parameters.} \label{table:runtime} \\\\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\\\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\\\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\\\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\\\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+EOF
+
+# make the table entries
 parameters=$(grep namelist probin.f90 | awk '{print $3}' | sort)
 
 for i in ${parameters}
 do
     value=$(grep -m 1 "$i =" probin.f90 | cut -d'=' -f2 | cut -d"!" -f1)
 
-    echo "\\verb= " $i "=" " &   & " `echo ${value} | sed 's/\_/\\\_/g'` "//"
+    echo "\\verb= " $i "=" " &   & " `echo ${value} | sed 's/\_/\\\_/g'` "\\\\"
 done
 
+# dump the end
+cat <<EOF
+
+\end{longtable}
+\end{center}
+%
+
+EOF

### 2009-04-10 docs/runtime_parameters/runtime_parameters.tex
add the defaults
rp.sh now includes the latex header info to generate the table fixed a typo in previous version updated the tex file to reflect the output of rp.sh
start filling in some descriptions

@@
 
+{\small
+
 \renewcommand{\arraystretch}{1.5}
 %
-% note: list generated via:
-% grep namelist probin.f90 | awk '{print ``\\verb= ``, $3, `` = &   &   \\\\"}' | sort
-
 \begin{center}
-\begin{longtable}{|l|p{3.0in}|l|}
+\begin{longtable}{|l|p{3.25in}|l|}
 \caption[runtime parameters]{runtime parameters.} \label{table:runtime} \\
@@
 
-\verb=  anelastic_cutoff  = &   &   \\
-\verb=  base_cutoff_density  = &   &   \\
-\verb=  bcx_hi  = &   &   \\
-\verb=  bcx_lo  = &   &   \\
-\verb=  bcy_hi  = &   &   \\
-\verb=  bcy_lo  = &   &   \\
-\verb=  bcz_hi  = &   &   \\
-\verb=  bcz_lo  = &   &   \\
-\verb=  beta_type  = &   &   \\
-\verb=  cflfac  = &   &   \\
-\verb=  cg_verbose  = &   &   \\
-\verb=  check_base_name  = &   &   \\
-\verb=  chk_int  = &   &   \\
-\verb=  co_latitude  = &   &   \\
-\verb=  dm_in  = &   &   \\
-\verb=  do_alltoallv  = &   &   \\
-\verb=  do_burning  = &   &   \\
-\verb=  do_eos_h_above_cutoff  = &   &   \\
+\verb=  anelastic_cutoff =  &   &  3.e6 \\
+\verb=  base_cutoff_density =  &   &  3.e6 \\
+\verb=  bcx_hi =  & $+x$ boundary condition (valid values are listed in {\tt  boxlib/bc.f90})  &  SLIP\_WALL \\
+\verb=  bcx_lo =  & $-x$ boundary condition  &  SLIP\_WALL \\
+\verb=  bcy_hi =  & $+y$ boundary condition  &  SLIP\_WALL \\
+\verb=  bcy_lo =  & $-y$ boundary condition  &  SLIP\_WALL \\
+\verb=  bcz_hi =  & $+z$ boundary condition  &  SLIP\_WALL \\
+\verb=  bcz_lo =  & $-z$ boundary condition  &  SLIP\_WALL \\
+\verb=  beta_type =  & what type of coefficient to use inside the velocity divergence constraint.  
+                       {\tt beta\_type} = 1 uses $\beta_0$; {\tt beta\_type} = 2 uses $\rho_0$ (anelastic);
+                       {\tt beta\_type} = 3 uses 1 (small-scale combustion).  &  1 \\
+\verb=  cflfac =  & CFL factor to use in the computation of the advection timestep constraint  &  0.5 \\
+\verb=  cg_verbose =  &   &  0 \\
+\verb=  check_base_name =  & prefix to use in checkpoint file names  &  "chk" \\
+\verb=  chk_int =  & number of timesteps between writing a checkpoint file  &  \\
+\verb=  co_latitude =  &   &  ZERO \\
+\verb=  dm_in =  & dimensionality (valid values are 2 or 3)  &  2 \\
+\verb=  do_alltoallv =  &   &  .false. \\
+\verb=  do_burning =  &   &  .true. \\
+\verb=  do_eos_h_above_cutoff =  &   &  .true. \\
 \verb=  do_initial_projection =  &   &  \\
-\verb=  do_sponge  = &   &   \\
-\verb=  dpdt_factor  = &   &   \\
-\verb=  drdxfac  = &   &   \\
-\verb=  enthalpy_pred_type  = &   &   \\
-\verb=  evolve_base_state  = &   &   \\
-\verb=  fixed_dt  = &   &   \\
-\verb=  grav_const  = &   &   \\
-\verb=  hg_bottom_solver  = &   &   \\
-\verb=  hg_dense_stencil  = &   &   \\
-\verb=  init_divu_iter  = &   &   \\
-\verb=  init_iter  = &   &   \\
-\verb=  init_shrink  = &   &   \\
-\verb=  interp_type_radial_bin_to_cart  = &   &   \\
-\verb=  job_name  = &   &   \\
-\verb=  lUsingNFiles  = &   &   \\
-\verb=  max_dt_growth  = &   &   \\
-\verb=  max_grid_size  = &   &   \\
-\verb=  max_levs  = &   &   \\
+\verb=  do_sponge =  &   &  .false. \\
+\verb=  dpdt_factor =  &   &  0.0d0 \\
+\verb=  drdxfac =  &   &  1 \\
+\verb=  enthalpy_pred_type =  &   &  predict\_rhohprime \\
+\verb=  evolve_base_state =  &   &  .true. \\
+\verb=  fixed_dt =  &   &  -1.0d0 \\
+\verb=  grav_const =  &   &  -1.5d10 \\
+\verb=  hg_bottom_solver =  &   &  -1 \\
+\verb=  hg_dense_stencil =  &   &  .true. \\
+\verb=  init_divu_iter =  &   &  4 \\
+\verb=  init_iter =  &   &  4 \\
+\verb=  init_shrink =  &   &  1.0 \\
+\verb=  interp_type_radial_bin_to_cart =  &   &  1 \\
+\verb=  job_name =  &   &  "" \\
+\verb=  lUsingNFiles =  &   &  .true. \\
+\verb=  max_dt_growth =  &   &  1.1D0 \\
+\verb=  max_grid_size =  &   &  64 \\
+\verb=  max_levs =  &   &  1 \\
 \verb=  max_step =  &   &  \\
-\verb=  mg_bottom_solver  = &   &   \\
-\verb=  mg_verbose  = &   &   \\
-\verb=  min_eff  = &   &   \\
-\verb=  min_width  = &   &   \\
+\verb=  mg_bottom_solver =  &   &  -1 \\
+\verb=  mg_verbose =  &   &  0 \\
+\verb=  min_eff =  &   &  0.7d0 \\
+\verb=  min_width =  &   &  16 \\
 \verb=  model_file =  &   &  \\
@@
 \verb=  nOutFiles =  &   &  \\
-\verb=  perturb_model  = &   &   \\
+\verb=  perturb_model =  &   &  .false. \\
 \verb=  plot_base =  &   &  \\
-\verb=  plot_base_name  = &   &   \\
-\verb=  plot_deltat  = &   &   \\
+\verb=  plot_base_name =  &   &  "plt" \\
+\verb=  plot_deltat =  &   &  0.0 \\
 \verb=  plot_int =  &   &  \\
@@
 \verb=  plot_trac =  &   &  \\
-\verb=  pmask_x  = &   &   \\
-\verb=  pmask_xyz  = &   &   \\
-\verb=  pmask_y  = &   &   \\
-\verb=  pmask_z  = &   &   \\
+\verb=  pmask_x =  &   &  .false. \\
+\verb=  pmask_y =  &   &  .false. \\
+\verb=  pmask_z =  &   &  .false. \\
+\verb=  ppm_type =  &   &  0 \\
 \verb=  prob_hi_x =  &   &  \\
@@
 \verb=  prob_hi_z =  &   &  \\
-\verb=  prob_lo_x  = &   &   \\
-\verb=  prob_lo_y  = &   &   \\
-\verb=  prob_lo_z  = &   &   \\
-\verb=  radius  = &   &   \\
-\verb=  ref_ratio  = &   &   \\
-\verb=  regrid_int  = &   &   \\
+\verb=  prob_lo_x =  &   &  ZERO \\
+\verb=  prob_lo_y =  &   &  ZERO \\
+\verb=  prob_lo_z =  &   &  ZERO \\
+\verb=  radius =  &   &  0.75d8 \\
+\verb=  ref_ratio =  &   &  2 \\
+\verb=  regrid_int =  &   &  -1 \\
 \verb=  restart =  &   &  \\
-\verb=  rotational_frequency  = &   &   \\
-\verb=  single_prec_plotfiles  = &   &   \\
-\verb=  slope_order  = &   &   \\
-\verb=  small_dens  = &   &   \\
-\verb=  small_temp  = &   &   \\
-\verb=  spherical_in  = &   &   \\
-\verb=  sponge_kappa  = &   &   \\
-\verb=  stop_time  = &   &   \\
-\verb=  temp_diffusion_formulation  = &   &   \\
-\verb=  test_set  = &   &   \\
-\verb=  the_knapsack_verbosity  = &   &   \\
-\verb=  thermal_diffusion_type  = &   &   \\
+\verb=  rotational_frequency =  &   &  ZERO \\
+\verb=  single_prec_plotfiles =  &   &  .false. \\
+\verb=  slope_order =  &   &  4 \\
+\verb=  small_dens =  &   &  1.d-5 \\
+\verb=  small_temp =  &   &  5.d6 \\
+\verb=  spherical_in =  &   &  0 \\
+\verb=  sponge_kappa =  &   &  10.d0 \\
+\verb=  stop_time =  &   &  -1.d0 \\
+\verb=  temp_diffusion_formulation =  &   &  .false. \\
+\verb=  test_set =  &   &  '' \\
+\verb=  the_copy_cache_max =  &   &  \\
+\verb=  the_knapsack_verbosity =  &   &  .false. \\
+\verb=  the_layout_verbosity =  &   &  0 \\
+\verb=  thermal_diffusion_type =  &   &  1 \\
 \verb=  the_sfc_threshold =  &   &  \\
-\verb=  use_delta_gamma1_term  = &   &   \\
-\verb=  use_divu_firstdt  = &   &   \\
-\verb=  use_eos_coulomb  = &   &   \\
-\verb=  use_etarho  = &   &   \\
-\verb=  use_ppm  = &   &   \\
-\verb=  use_soundspeed_firstdt  = &   &   \\
-\verb=  use_tfromp  = &   &   \\
-\verb=  use_thermal_diffusion  = &   &   \\
-\verb=  velpert_amplitude  = &   &   \\
-\verb=  velpert_radius  = &   &   \\
-\verb=  velpert_scale  = &   &   \\
-\verb=  velpert_steep  = &   &   \\
-\verb=  verbose  = &   &   \\
-\verb=  w0mac_interp_type  = &   &   \\
-\verb=  xrb_pert_factor  = &   &   \\
-\verb=  xrb_pert_size  = &   &   \\
-\verb=  xrb_pert_type  = &   &   \\
-\verb=  xrb_use_bottom_sponge  = &   &   \\
+\verb=  use_delta_gamma1_term =  &   &  .false. \\
+\verb=  use_divu_firstdt =  &   &  .false. \\
+\verb=  use_eos_coulomb =  &   &  .false. \\
+\verb=  use_etarho =  &   &  .true. \\
+\verb=  use_soundspeed_firstdt =  &   &  .false. \\
+\verb=  use_tfromp =  &   &  .false. \\
+\verb=  use_thermal_diffusion =  &   &  .false. \\
+\verb=  velpert_amplitude =  &   &  0.d0 \\
+\verb=  velpert_radius =  &   &  0.75d8 \\
+\verb=  velpert_scale =  &   &  0.8d8 \\
+\verb=  velpert_steep =  &   &  1.d0 \\
+\verb=  verbose =  &   &  0 \\
+\verb=  w0mac_interp_type =  &   &  1 \\
+\verb=  xrb_pert_factor =  &   &  1.d-2 \\
+\verb=  xrb_pert_size =  &   &  5.0d1 \\
+\verb=  xrb_pert_type =  &   &  1 \\
+\verb=  xrb_use_bottom_sponge =  &   &  .true. \\
 
@@
 \end{center}
+
+} % ends \small
+
 %

### 2009-04-10 docs/runtime_parameters/runtime_parameters.tex
more descriptions

@@
 \verb=  do_sponge =  &   &  .false. \\
-\verb=  dpdt_factor =  &   &  0.0d0 \\
-\verb=  drdxfac =  &   &  1 \\
+\verb=  dpdt_factor =  & factor in front of the volume discrepancy term (0.0 = off)  &  0.0d0 \\
+\verb=  drdxfac =  & ratio of radial base state zones to Cartesian full state zones for spherical geometry  &  1 \\
 \verb=  enthalpy_pred_type =  &   &  predict\_rhohprime \\
-\verb=  evolve_base_state =  &   &  .true. \\
+\verb=  evolve_base_state =  & turn on (.true.) or off (.false.) basestate evolution  &  .true. \\
 \verb=  fixed_dt =  &   &  -1.0d0 \\
-\verb=  grav_const =  &   &  -1.5d10 \\
+\verb=  grav_const =  & the gravitational acceleration (cm~s$^{-2}$) for plane-parallel geometry  &  -1.5d10 \\
 \verb=  hg_bottom_solver =  &   &  -1 \\
@@
 \verb=  init_iter =  &   &  4 \\
-\verb=  init_shrink =  &   &  1.0 \\
+\verb=  init_shrink =  & the multiplicative factor ($\le 1$) to reduce the initial timestep as computed by the various timestep estimators   &  1.0 \\
 \verb=  interp_type_radial_bin_to_cart =  &   &  1 \\

### 2009-04-11 docs/runtime_parameters/runtime_parameters.tex
add in the parameters that the modified script found
more progress

@@
 
-\verb=  anelastic_cutoff =  &   &  3.e6 \\
-\verb=  base_cutoff_density =  &   &  3.e6 \\
+\verb=  anelastic_cutoff =  & Refer to Section \ref{Sec:Anelastic Cutoff}.  &  3.e6 \\
+\verb=  base_cutoff_density =  & Refer to Section \ref{Sec:Base Cutoff Density}  &  3.e6 \\
 \verb=  bcx_hi =  & $+x$ boundary condition (valid values are listed in {\tt  boxlib/bc.f90})  &  SLIP\_WALL \\
@@
 \verb=  cflfac =  & CFL factor to use in the computation of the advection timestep constraint  &  0.5 \\
-\verb=  cg_verbose =  &   &  0 \\
+\verb=  cg_verbose =  & Verbosity of bottom solver.  &  0 \\
 \verb=  check_base_name =  & prefix to use in checkpoint file names  &  "chk" \\
-\verb=  chk_int =  & number of timesteps between writing a checkpoint file  &  \\
-\verb=  co_latitude =  &   &  ZERO \\
+\verb=  chk_int =  & number of timesteps between writing a checkpoint file  & 0 \\
+\verb=  co_latitude =  & latitude, in radians, for problems with rotation where the domain is only a subset
+                             of a full star.  &  ZERO \\
 \verb=  dm_in =  & dimensionality (valid values are 2 or 3)  &  2 \\
 \verb=  do_alltoallv =  &   &  .false. \\
-\verb=  do_burning =  &   &  .true. \\
-\verb=  do_eos_h_above_cutoff =  &   &  .true. \\
-\verb=  do_initial_projection =  &   &  \\
-\verb=  do_sponge =  &   &  .false. \\
+\verb=  do_burning =  & turn on (.true.) or off (.false.) burning  &  .true. \\
+\verb=  do_eos_h_above_cutoff =  & After the advective enthalpy update, recompute the enthalpy if we are 
+                                         below the base cutoff density.   &  .true. \\
+\verb=  do_initial_projection =  & Do the initial projection.  & 1 \\
+\verb=  do_sponge =  & Use sponging.  &  .false. \\
 \verb=  dpdt_factor =  & factor in front of the volume discrepancy term (0.0 = off)  &  0.0d0 \\
 \verb=  drdxfac =  & ratio of radial base state zones to Cartesian full state zones for spherical geometry  &  1 \\
-\verb=  enthalpy_pred_type =  &   &  predict\_rhohprime \\
+\verb=  enthalpy_pred_type =  & predict\_rhohprime = 1; predict\_h = 2; predict\_T\_then\_rhohprime = 3;
+                                      predict\_T\_then\_h = 4; predict\_hprime = 5; predict\_Tprime\_then\_h = 6;
+                                    &  predict\_rhohprime \\
 \verb=  evolve_base_state =  & turn on (.true.) or off (.false.) basestate evolution  &  .true. \\
-\verb=  fixed_dt =  &   &  -1.0d0 \\
+\verb=  fixed_dt =  & Fix the time step.  If -1.0, then use the standard time step.  &  -1.0d0 \\
 \verb=  grav_const =  & the gravitational acceleration (cm~s$^{-2}$) for plane-parallel geometry  &  -1.5d10 \\
 \verb=  hg_bottom_solver =  &   &  -1 \\
-\verb=  hg_dense_stencil =  &   &  .true. \\
-\verb=  init_divu_iter =  &   &  4 \\
-\verb=  init_iter =  &   &  4 \\
-\verb=  init_shrink =  & the multiplicative factor ($\le 1$) to reduce the initial timestep as computed by the various timestep estimators   &  1.0 \\
-\verb=  interp_type_radial_bin_to_cart =  &   &  1 \\
+\verb=  hg_dense_stencil =  & In hgproject, 
+                                   In 2D, use a 9 point Laplacian (.true.) or 5-point Laplacian (.false.).
+                                   In 3D, use a 27 point Laplacian (.true.) or 7-point Laplacian (.false.). &  .true. \\
+\verb=  init_divu_iter =  & Number of initial divu iterations.  &  4 \\
+\verb=  init_iter =  & Number of initial pressure iterations.  &  4 \\
+\verb=  init_shrink =  & the multiplicative factor ($\le 1$) to reduce the initial timestep as computed by the 
+                             various timestep estimators   &  1.0 \\
+\verb=  interp_type_radial_bin_to_cart =  & The interpolation for filling a full multifab from a 1D 
+                                                     bin-centered array.  1 = piecewise constant; 
+                                                     2 = piecewise linear.  &  1 \\
 \verb=  job_name =  &   &  "" \\
-\verb=  lUsingNFiles =  &   &  .true. \\
-\verb=  max_dt_growth =  &   &  1.1D0 \\
-\verb=  max_grid_size =  &   &  64 \\
-\verb=  max_levs =  &   &  1 \\
-\verb=  max_step =  &   &  \\
+\verb=  lUsingNFiles =  & If .true., use nOutFiles processors to write checkpoint and plotfiles.  
+                              Fortran has the unfortunate feature of each processor only being able to write out 
+                              1-2GB each without crashing.  &  .true. \\
+\verb=  max_dt_growth =  & The maximum scale factor that the time step is allowed to grow by per time step.  
+                              &  1.1D0 \\
+\verb=  max_grid_size =  & The largest grid size that will be created using make\_new\_grids.  &  64 \\
+\verb=  max_levs =  & Total number of levels.  1 = single level.  &  1 \\
+\verb=  max_step =  & Maximum number of steps in the simulation.  & 1 \\
 \verb=  mg_bottom_solver =  &   &  -1 \\
-\verb=  mg_verbose =  &   &  0 \\
+\verb=  mg_verbose =  & Verbsoity of the multigrid solver, but not the bottom solver.  &  0 \\
 \verb=  min_eff =  &   &  0.7d0 \\
-\verb=  min_width =  &   &  16 \\
-\verb=  model_file =  &   &  \\
-\verb=  n_cellx =  &   &  \\
-\verb=  n_celly =  &   &  \\
-\verb=  n_cellz =  &   &  \\
-\verb=  nOutFiles =  &   &  \\
-\verb=  perturb_model =  &   &  .false. \\
-\verb=  plot_base =  &   &  \\
+\verb=  min_width =  & The minimum size on a side  for a grid created using make\_new\_grids.  &  16 \\
+\verb=  model_file =  &   & ``model.hse'' \\
+\verb=  n_cellx =  & Number of cells for the base level in the x-direction  &  \\
+\verb=  n_celly =  & Number of cells for the base level in the y-direction  &  \\
+\verb=  n_cellz =  & Number of cells for the base level in the z-direction  &  \\
+\verb=  nOutFiles =  & If lUsingNFiles = .true., use this many processors to write checkpoint and plotfiles.  
+                           Fortran has the unfortunate feature of each processor only being able to write out 
+                           1-2GB each without crashing.  & 64 \\
+\verb=  perturb_model =  & Turn on a perturbation in the initial data.  Problem specific.  &  .false. \\
+\verb=  plot_base =  &   & .false. \\
 \verb=  plot_base_name =  &   &  "plt" \\
 \verb=  plot_deltat =  &   &  0.0 \\
-\verb=  plot_int =  &   &  \\
-\verb=  plot_spec =  &   &  \\
-\verb=  plot_trac =  &   &  \\
+\verb=  plot_int =  &   & 0 \\
+\verb=  plot_spec =  &   & .true. \\
+\verb=  plot_trac =  &   & .false. \\
 \verb=  pmask_x =  &   &  .false. \\
@@
 \verb=  prob_lo_z =  &   &  ZERO \\
-\verb=  radius =  &   &  0.75d8 \\
+\verb=  radius =  &   &  1.0e6 \\
 \verb=  ref_ratio =  &   &  2 \\
 \verb=  regrid_int =  &   &  -1 \\
-\verb=  restart =  &   &  \\
+\verb=  restart =  &   & -1 \\
 \verb=  rotational_frequency =  &   &  ZERO \\
@@
 \verb=  test_set =  &   &  '' \\
-\verb=  the_copy_cache_max =  &   &  \\
+\verb=  the_copy_cache_max =  &   & 50 \\
 \verb=  the_knapsack_verbosity =  &   &  .false. \\
@@
 \verb=  thermal_diffusion_type =  &   &  1 \\
-\verb=  the_sfc_threshold =  &   &  \\
+\verb=  the_sfc_threshold =  &   & 5 \\
 \verb=  use_delta_gamma1_term =  &   &  .false. \\

### 2009-04-11 Source/varden.f90
made use_eos_coulomb=.true. by default in probin
made do_initial_projection a logical instead of int.

modified do_initial_projection in all inputs files

@@ -312,7 +312,7 @@ subroutine varden()
 
      call make_div_coeff(div_coeff_old,rho0_old,p0_old,gamma1bar,grav_cell)
 
-     if(do_initial_projection > 0) then
+     if(do_initial_projection) then
         call initial_proj(uold,sold,pres,gpres,Source_old,hgrhs, &
                           div_coeff_old,p0_old,gamma1bar,dx,the_bc_tower,mla)
      end if

### 2009-04-11 TEST_PROBLEMS/test_eta_stuff/varden.f90

@@ -108,8 +108,8 @@ subroutine varden()
 
   call probin_init()
 
-  if(parallel_IOProcessor() .and. do_initial_projection .eq. 0) then
-     print*,'Warning: do_initial_projection = 0'
+  if(parallel_IOProcessor() .and. do_initial_projection) then
+     print*,'Warning: do_initial_projection = F'
      print*,'Setting init_divu_iter = init_iter = max_step = 0'
      print*,'So the output will be the loaded in data'
 
@@ -515,7 +515,7 @@ subroutine varden()
      ! Do an initial projection with omegadot = 0 and rho_Hext = 0
      !----------------------------------------------------------------------
 
-     if(do_initial_projection > 0) then
+     if(do_initial_projection) then
         call initial_proj(nlevs,uold,sold,pres,gpres,Source_old,hgrhs, &
                           div_coeff_old,s0_old,p0_old,gamma1bar(:,:,1),dx,the_bc_tower,mla)
      end if

### 2009-04-11 probin.f90

@@ -18,15 +18,14 @@ module probin_module
   real(dp_t), save         :: prob_hi_x, prob_hi_y, prob_hi_z
   integer, save            :: max_step, plot_int
   real(dp_t), save         :: plot_deltat
-  integer, save            :: chk_int, init_iter, init_divu_iter
+  integer, save            :: chk_int, init_iter, init_divu_iter, restart
   real(dp_t), save         :: cflfac, init_shrink
   character(len=128), save :: test_set
-  integer, save            :: restart, do_initial_projection
   integer, save            :: bcx_lo, bcx_hi, bcy_lo, bcy_hi, bcz_lo, bcz_hi
   logical, save            :: pmask_x, pmask_y, pmask_z, pmask_xyz(MAX_SPACEDIM)
   integer, save            :: verbose, mg_verbose, cg_verbose
   integer, save            :: hg_bottom_solver, mg_bottom_solver
-  logical, save            :: do_sponge, hg_dense_stencil
+  logical, save            :: do_sponge, hg_dense_stencil, do_initial_projection
   real(dp_t), save         :: anelastic_cutoff, base_cutoff_density, dpdt_factor
   integer, save            :: spherical_in, dm_in
   logical, save            :: perturb_model, plot_spec, plot_trac, plot_base
@@ -241,7 +240,7 @@ contains
 
     init_shrink = 1.0
 
-    do_initial_projection  = 1
+    do_initial_projection  = .true.
 
     need_inputs = .true.
     test_set = ''
@@ -307,7 +306,7 @@ contains
     xrb_pert_type = 1
     xrb_use_bottom_sponge = .true.
 
-    use_eos_coulomb = .false.
+    use_eos_coulomb = .true.
 
     small_temp = 5.d6
     small_dens = 1.d-5

### 2009-04-11 docs/runtime_parameters/runtime_parameters.tex
made use_eos_coulomb=.true. by default in probin
made do_initial_projection a logical instead of int.

modified do_initial_projection in all inputs files
some formatting

@@
 \verb=  anelastic_cutoff =  & Refer to Section \ref{Sec:Anelastic Cutoff}.  &  3.e6 \\
-\verb=  base_cutoff_density =  & Refer to Section \ref{Sec:Base Cutoff Density}  &  3.e6 \\
-\verb=  bcx_hi =  & $+x$ boundary condition (valid values are listed in {\tt  boxlib/bc.f90})  &  SLIP\_WALL \\
+\verb=  base_cutoff_density =  & Refer to Section \ref{Sec:Base Cutoff Density}  
+                                     &  3.e6 \\
+\verb=  bcx_hi =  & $+x$ boundary condition (valid values are listed in 
+                       {\tt  boxlib/bc.f90})  &  SLIP\_WALL \\
 \verb=  bcx_lo =  & $-x$ boundary condition  &  SLIP\_WALL \\
@@
 \verb=  bcz_lo =  & $-z$ boundary condition  &  SLIP\_WALL \\
-\verb=  beta_type =  & what type of coefficient to use inside the velocity divergence constraint.  
-                           {\tt beta\_type} = 1 uses $\beta_0$; {\tt beta\_type} = 2 uses $\rho_0$ (anelastic);
+\verb=  beta_type =  & what type of coefficient to use inside the velocity divergence 
+                           constraint.  {\tt beta\_type} = 1 uses $\beta_0$; 
+                           {\tt beta\_type} = 2 uses $\rho_0$ (anelastic);
                            {\tt beta\_type} = 3 uses 1 (small-scale combustion).  &  1 \\
-\verb=  cflfac =  & CFL factor to use in the computation of the advection timestep constraint  &  0.5 \\
+\verb=  cflfac =  & CFL factor to use in the computation of the advection timestep 
+                       constraint  &  0.5 \\
 \verb=  cg_verbose =  & Verbosity of bottom solver.  &  0 \\
@@
 \verb=  chk_int =  & number of timesteps between writing a checkpoint file  & 0 \\
-\verb=  co_latitude =  & latitude, in radians, for problems with rotation where the domain is only a subset
-                             of a full star.  &  ZERO \\
+\verb=  co_latitude =  & latitude, in radians, for problems with rotation where the 
+                             domain is only a subset of a full star.  &  ZERO \\
 \verb=  dm_in =  & dimensionality (valid values are 2 or 3)  &  2 \\
@@
 \verb=  do_burning =  & turn on (.true.) or off (.false.) burning  &  .true. \\
-\verb=  do_eos_h_above_cutoff =  & After the advective enthalpy update, recompute the enthalpy if we are 
-                                         below the base cutoff density.   &  .true. \\
-\verb=  do_initial_projection =  & Do the initial projection.  & 1 \\
+\verb=  do_eos_h_above_cutoff =  & After the advective enthalpy update, recompute 
+                                          the enthalpy if we are below the base cutoff 
+                                          density.   &  .true. \\
+\verb=  do_initial_projection =  & Do the initial projection.  & .true. \\
 \verb=  do_sponge =  & Use sponging.  &  .false. \\
-\verb=  dpdt_factor =  & factor in front of the volume discrepancy term (0.0 = off)  &  0.0d0 \\
-\verb=  drdxfac =  & ratio of radial base state zones to Cartesian full state zones for spherical geometry  &  1 \\
-\verb=  enthalpy_pred_type =  & predict\_rhohprime = 1; predict\_h = 2; predict\_T\_then\_rhohprime = 3;
-                                      predict\_T\_then\_h = 4; predict\_hprime = 5; predict\_Tprime\_then\_h = 6;
-                                    &  predict\_rhohprime \\
-\verb=  evolve_base_state =  & turn on (.true.) or off (.false.) basestate evolution  &  .true. \\
-\verb=  fixed_dt =  & Fix the time step.  If -1.0, then use the standard time step.  &  -1.0d0 \\
-\verb=  grav_const =  & the gravitational acceleration (cm~s$^{-2}$) for plane-parallel geometry  &  -1.5d10 \\
+\verb=  dpdt_factor =  & factor in front of the volume discrepancy term (0.0 = off)  
+                           &  0.0d0 \\
+\verb=  drdxfac =  & ratio of radial base state zones to Cartesian full state zones for 
+                        spherical geometry  &  1 \\
+\verb=  enthalpy_pred_type =  & predict\_rhohprime = 1; predict\_h = 2; 
+                                      predict\_T\_then\_rhohprime = 3;
+                                      predict\_T\_then\_h = 4; predict\_hprime = 5; 
+                                      predict\_Tprime\_then\_h = 6;  &  predict\_rhohprime \\
+\verb=  evolve_base_state =  & turn on (.true.) or off (.false.) basestate evolution  
+                                   &  .true. \\
+\verb=  fixed_dt =  & Fix the time step.  If -1.0, then use the standard time step.  
+                        &  -1.0d0 \\
+\verb=  grav_const =  & the gravitational acceleration (cm~s$^{-2}$) for plane-parallel 
+                            geometry  &  -1.5d10 \\
 \verb=  hg_bottom_solver =  &   &  -1 \\
-\verb=  hg_dense_stencil =  & In hgproject, 
-                                   In 2D, use a 9 point Laplacian (.true.) or 5-point Laplacian (.false.).
-                                   In 3D, use a 27 point Laplacian (.true.) or 7-point Laplacian (.false.). &  .true. \\
+\verb=  hg_dense_stencil =  & In hgproject, in 2D, use a 9 point Laplacian (.true.) 
+                                    or 5-point Laplacian (.false.).  In 3D, use a 27 point 
+                                    Laplacian (.true.) or 7-point Laplacian (.false.). 
+                                  &  .true. \\
 \verb=  init_divu_iter =  & Number of initial divu iterations.  &  4 \\
 \verb=  init_iter =  & Number of initial pressure iterations.  &  4 \\
-\verb=  init_shrink =  & the multiplicative factor ($\le 1$) to reduce the initial timestep as computed by the 
-                             various timestep estimators   &  1.0 \\
-\verb=  interp_type_radial_bin_to_cart =  & The interpolation for filling a full multifab from a 1D 
-                                                     bin-centered array.  1 = piecewise constant; 
+\verb=  init_shrink =  & the multiplicative factor ($\le 1$) to reduce the initial 
+                             timestep as computed by the various timestep estimators   
+                            &  1.0 \\
+\verb=  interp_type_radial_bin_to_cart =  & The interpolation for filling a full 
+                                                     multifab from a 1D bin-centered array.  
+                                                     1 = piecewise constant; 
                                                      2 = piecewise linear.  &  1 \\
 \verb=  job_name =  &   &  "" \\
-\verb=  lUsingNFiles =  & If .true., use nOutFiles processors to write checkpoint and plotfiles.  
-                              Fortran has the unfortunate feature of each processor only being able to write out 
+\verb=  lUsingNFiles =  & If .true., use nOutFiles processors to write checkpoint and 
+                               plotfiles.  Fortran has the unfortunate feature of each 
+                               processor only being able to write out 
                                1-2GB each without crashing.  &  .true. \\
-\verb=  max_dt_growth =  & The maximum scale factor that the time step is allowed to grow by per time step.  
-                              &  1.1D0 \\
-\verb=  max_grid_size =  & The largest grid size that will be created using make\_new\_grids.  &  64 \\
+\verb=  max_dt_growth =  & The maximum scale factor that the time step is allowed to 
+                                grow by per time step.  &  1.1D0 \\
+\verb=  max_grid_size =  & The largest grid size that will be created using 
+                                make\_new\_grids.  &  64 \\
 \verb=  max_levs =  & Total number of levels.  1 = single level.  &  1 \\
@@
 \verb=  mg_bottom_solver =  &   &  -1 \\
-\verb=  mg_verbose =  & Verbsoity of the multigrid solver, but not the bottom solver.  &  0 \\
+\verb=  mg_verbose =  & Verbsoity of the multigrid solver, but not the bottom solver.  
+                          &  0 \\
 \verb=  min_eff =  &   &  0.7d0 \\
-\verb=  min_width =  & The minimum size on a side  for a grid created using make\_new\_grids.  &  16 \\
+\verb=  min_width =  & The minimum size on a side  for a grid created using 
+                           make\_new\_grids.  &  16 \\
 \verb=  model_file =  &   & ``model.hse'' \\
@@
 \verb=  n_cellz =  & Number of cells for the base level in the z-direction  &  \\
-\verb=  nOutFiles =  & If lUsingNFiles = .true., use this many processors to write checkpoint and plotfiles.  
-                           Fortran has the unfortunate feature of each processor only being able to write out 
-                           1-2GB each without crashing.  & 64 \\
-\verb=  perturb_model =  & Turn on a perturbation in the initial data.  Problem specific.  &  .false. \\
+\verb=  nOutFiles =  & If lUsingNFiles = .true., use this many processors to write 
+                           checkpoint and plotfiles.  Fortran has the unfortunate feature 
+                           of each processor only being able to write out 1-2GB each 
+                           without crashing.  & 64 \\
+\verb=  perturb_model =  & Turn on a perturbation in the initial data.  
+                                Problem specific.  &  .false. \\
 \verb=  plot_base =  &   & .false. \\

### 2009-04-13 docs/runtime_parameters/runtime_parameters.tex
more progress

@@
                                       predict\_T\_then\_h = 4; predict\_hprime = 5; 
-                                      predict\_Tprime\_then\_h = 6;  &  predict\_rhohprime \\
+                                      predict\_Tprime\_then\_h = 6.  &  predict\_rhohprime \\
 \verb=  evolve_base_state =  & turn on (.true.) or off (.false.) basestate evolution  
@@
 \verb=  stop_time =  &   &  -1.d0 \\
-\verb=  temp_diffusion_formulation =  &   &  .false. \\
-\verb=  test_set =  &   &  '' \\
-\verb=  the_copy_cache_max =  &   & 50 \\
-\verb=  the_knapsack_verbosity =  &   &  .false. \\
+\verb=  temp_diffusion_formulation =  & How to compute the explicit thermal diffusion
+                                                term.  .true. = in terms of $T$; .false. =
+                                                in terms of $\rho,p_0,X$.  &  .false. \\
+\verb=  test_set =  & Fixed grid file.  &  '' \\
+\verb=  the_copy_cache_max =  & Number of boxassoc layouts we keep in memory to avoid 
+                                      having to recompute the boxassoc, which is 
+                                      computationally expensive.  & 50 \\
+\verb=  the_knapsack_verbosity =  & Verbosity of the knapscak processor-to-grid
+                                           algorithm.  &  .false. \\
 \verb=  the_layout_verbosity =  &   &  0 \\
-\verb=  thermal_diffusion_type =  &   &  1 \\
-\verb=  the_sfc_threshold =  &   & 5 \\
+\verb=  thermal_diffusion_type =  & In the thermal diffusion solver, 1 = 
+                                           Crank-Nicholson; 2 = Backward Euler.  &  1 \\
+\verb=  the_sfc_threshold =  & When assigning processors for grids, this determines 
+                                     whether we use the sfc algorithm or knapsack 
+                                     algorithm.  If the total number of grids divided by 
+                                     the number of processors is greater than this number, 
+                                     use sfc.  & 5 \\
 \verb=  use_delta_gamma1_term =  &   &  .false. \\
-\verb=  use_divu_firstdt =  &   &  .false. \\
+\verb=  use_divu_firstdt =  & Use the divu constraint when computing the first
+                                   time step.  &  .false. \\
 \verb=  use_eos_coulomb =  &   &  .false. \\
 \verb=  use_etarho =  &   &  .true. \\
-\verb=  use_soundspeed_firstdt =  &   &  .false. \\
-\verb=  use_tfromp =  &   &  .false. \\
-\verb=  use_thermal_diffusion =  &   &  .false. \\
-\verb=  velpert_amplitude =  &   &  0.d0 \\
-\verb=  velpert_radius =  &   &  0.75d8 \\
-\verb=  velpert_scale =  &   &  0.8d8 \\
-\verb=  velpert_steep =  &   &  1.d0 \\
-\verb=  verbose =  &   &  0 \\
-\verb=  w0mac_interp_type =  &   &  1 \\
-\verb=  xrb_pert_factor =  &   &  1.d-2 \\
-\verb=  xrb_pert_size =  &   &  5.0d1 \\
-\verb=  xrb_pert_type =  &   &  1 \\
-\verb=  xrb_use_bottom_sponge =  &   &  .true. \\
+\verb=  use_soundspeed_firstdt =  & Use the soundspeed constraint when computing the
+                                           first time step.  &  .false. \\
+\verb=  use_tfromp =  & When updating temperature, use $T=T(\rho,p_0,X) $ rather than
+                            $T=T(\rho,h,X)$. &  .false. \\
+\verb=  use_thermal_diffusion =  & Use thermal diffusion.  &  .false. \\
+\verb=  velpert_amplitude =  & (spherical stars only) Amplitude of random Fourier
+                                     modes.  &  0.d0 \\
+\verb=  velpert_radius =  & (spherical stars only) Cutoff radius where we start 
+                                 applying a tanh function to decay the random Fourier
+                                 modes. &  0.75d8 \\
+\verb=  velpert_scale =  & (spherical stars only) Square length which defines the 
+                                random Fourier modes correspond to. &  0.8d8 \\
+\verb=  velpert_steep =  & (spherical stars only) Steepness of the cutoff of the random
+                                Fourier modes. &  1.d0 \\
+\verb=  verbose =  & General verbosity  &  0 \\
+\verb=  w0mac_interp_type =  & 0 = Compute w0 at edges using a projection; 1 = 
+                                     Interpolate w0 to cell centers, then average to edges; 
+                                     2 = Interpolate w0 to edges directly; 3 = Interpolate 
+                                     w0 to nodes, then average to edges.  &  1 \\
+\verb=  xrb_pert_factor =  & (xrb problem only) Increase in amplitude of 
+                                  perturbational quantity relative to ambient. &  1.d-2 \\
+\verb=  xrb_pert_size =  & (xrb problem only)  Width of Gaussian distribution &  5.0d1 \\
+\verb=  xrb_pert_type =  & (xrb problem only) 1 = perturb temperature; 2 = perturb 
+                                density. &  1 \\
+\verb=  xrb_use_bottom_sponge =  & (xrb problem only)  Put a sponge at the bottom of 
+                                          the domain.  &  .true. \\
 

### 2009-04-14 docs/runtime_parameters/runtime_parameters.tex
replaced the shell script with a python script that will read in from fParallel/MAESTRO/_parameters and dump out the LaTeX formatted table for use in documentation.
comments within the _parameters file should begin with a "#" symbol and
should be LaTeX friendly (e.g. "$\le$" instead of "<=").

there should also be a space between the header info in _parameters and the
first parameter:

# This file contains...
# .
# .
# .
# file in the problem directory.

# This is used to specify the model file to map into multidimensions
model_file                          character          "model.hse"

etc.

--- 2009-04-11 docs/runtime_parameters/runtime_parameters.tex
@@
 \verb=  beta_type  = &   what type of coefficient to use inside the velocity divergence constraint.
-                           {\tt beta\_type} = 1 uses $\beta_0$; {\tt beta\_type} = 2 uses $\rho_0$ (anelastic);
-                           {\tt beta\_type} = 3 uses 1 (small-scale combustion).  &  1 \\
+{\tt beta\_type} = 1 uses $\beta_0$; {\tt beta\_type} = 2 uses $\rho_0$ (anelastic); {\tt beta\_type} = 3 uses 1 (small-scale combustion).
+&  1 \\
 \verb=  cflfac  = &   CFL factor to use in the computation of the advection timestep constraint  &  0.5 \\
-\verb=  cg_verbose =  & Verbosity of bottom solver.  &  0 \\
+\verb=  cg_verbose  = &   Verbosity of bottom solver  &  0 \\
 \verb=  check_base_name  = &   prefix to use in checkpoint file names  &  "chk" \\
 \verb=  chk_int  = &   number of timesteps between writing a checkpoint file  &  0 \\
-\verb=  co_latitude =  & latitude, in radians, for problems with rotation where the domain is only a subset
-                             of a full star.  &  ZERO \\
+\verb=  co_latitude  = &   latitude, in radians, for problems with rotation where the domain is only a subset of a full star.  &  ZERO \\
 \verb=  dm_in  = &   dimensionality (valid values are 2 or 3)  &  2 \\
@@
 \verb=  do_burning  = &   turn on (.true.) or off (.false.) burning  &  .true. \\
-\verb=  do_eos_h_above_cutoff =  & After the advective enthalpy update, recompute the enthalpy if we are 
-                                         below the base cutoff density.   &  .true. \\
-\verb=  do_initial_projection =  & Do the initial projection.  & 1 \\
+\verb=  do_eos_h_above_cutoff  = &   After the advective enthalpy update, recompute the enthalpy if we are below the base cutoff density.  &  .true. \\
+\verb=  do_initial_projection  = &   Do the initial projection.  &  .true. \\
 \verb=  do_sponge  = &   Use sponging.  &  .false. \\
@@
 \verb=  drdxfac  = &   ratio of radial base state zones to Cartesian full state zones for spherical geometry  &  1 \\
-\verb=  enthalpy_pred_type =  & predict\_rhohprime = 1; predict\_h = 2; predict\_T\_then\_rhohprime = 3;
-                                      predict\_T\_then\_h = 4; predict\_hprime = 5; predict\_Tprime\_then\_h = 6;
+\verb=  enthalpy_pred_type  = &   predict\_rhohprime = 1; predict\_h = 2; predict\_T\_then\_rhohprime = 3; predict\_T\_then\_h = 4; predict\_hprime = 5; predict\_Tprime\_then\_h = 6.
 &  predict\_rhohprime \\
@@
 \verb=  grav_const  = &   the gravitational acceleration (cm~s$^{-2}$) for plane-parallel geometry  &  -1.5d10 \\
-\verb=  hg_bottom_solver =  &   &  -1 \\
-\verb=  hg_dense_stencil =  & In hgproject, 
-                                   In 2D, use a 9 point Laplacian (.true.) or 5-point Laplacian (.false.).
-                                   In 3D, use a 27 point Laplacian (.true.) or 7-point Laplacian (.false.). &  .true. \\
+\verb=  hg_bottom_solver  = &   valid values are $\le$ 0  &  -1 \\
+\verb=  hg_dense_stencil  = &   In hgproject, in 2D, use a 9 point Laplacian (.true.) or 5-point Laplacian (.false.).
+In 3D, use a 27 point Laplacian (.true.) or 7-point Laplacian (.false.).
+&  .true.
+\\
 \verb=  init_divu_iter  = &   Number of initial divu iterations.  &  4 \\
 \verb=  init_iter  = &   Number of initial pressure iterations.  &  4 \\
-\verb=  init_shrink =  & the multiplicative factor ($\le 1$) to reduce the initial timestep as computed by the 
-                             various timestep estimators   &  1.0 \\
-\verb=  interp_type_radial_bin_to_cart =  & The interpolation for filling a full multifab from a 1D 
-                                                     bin-centered array.  1 = piecewise constant; 
-                                                     2 = piecewise linear.  &  1 \\
+\verb=  init_shrink  = &   the multiplicative factor ($\le 1$) to reduce the initial timestep as computed by the various timestep estimators  &  1.0 \\
+\verb=  interp_type_radial_bin_to_cart  = &   The interpolation for filling a full multifab from a 1D bin-centered array. 1 = piecewise constant; 2 = piecewise linear.  &  1 \\
 \verb=  job_name  = &    &  "" \\
 \verb=  lUsingNFiles  = &   If .true., use nOutFiles processors to write checkpoint and plotfiles.
-                              Fortran has the unfortunate feature of each processor only being able to write out 
-                              1-2GB each without crashing.  &  .true. \\
-\verb=  max_dt_growth =  & The maximum scale factor that the time step is allowed to grow by per time step.  
-                              &  1.1D0 \\
+Fortran has the unfortunate feature of each processor only being able to write out 1-2GB each without crashing.
+&  .true.
+\\
+\verb=  max_dt_growth  = &   The maximum scale factor that the time step is allowed to grow by per time step.  &  1.1D0 \\
 \verb=  max_grid_size  = &    The largest grid size that will be created using make\_new\_grids.  &  64 \\
@@
 \verb=  max_step  = &   Maximum number of steps in the simulation.  &  1 \\
-\verb=  mg_bottom_solver =  &   &  -1 \\
+\verb=  mg_bottom_solver  = &   valid values are $\le$ 0  &  -1 \\
 \verb=  mg_verbose  = &   Verbsoity of the multigrid solver, but not the bottom solver.  &  0 \\
@@
 \verb=  min_width  = &   The minimum size on a side for a grid created using make\_new\_grids.  &  16 \\
-\verb=  model_file =  &   & ``model.hse'' \\
-\verb=  n_cellx =  & Number of cells for the base level in the x-direction  &  \\
-\verb=  n_celly =  & Number of cells for the base level in the y-direction  &  \\
-\verb=  n_cellz =  & Number of cells for the base level in the z-direction  &  \\
+\verb=  model_file  = &    &  "model.hse" \\
 \verb=  nOutFiles  = &   If lUsingNFiles = .true., use this many processors to write checkpoint and plotfiles.
-                           Fortran has the unfortunate feature of each processor only being able to write out 
-                           1-2GB each without crashing.  & 64 \\
+Fortran has the unfortunate feature of each processor only being able to write out 1-2GB each without crashing.
+&  64 \\
+\verb=  n_cellx  = &   Number of cells for the base level in the x-direction  &  -1 \\
+\verb=  n_celly  = &   Number of cells for the base level in the y-direction  &  -1 \\
+\verb=  n_cellz  = &   Number of cells for the base level in the z-direction  &  -1 \\
 \verb=  perturb_model  = &   Turn on a perturbation in the initial data.  Problem specific.  &  .false. \\
@@
 \verb=  pmask_z  = &    &  .false. \\
-\verb=  ppm_type =  &   &  0 \\
-\verb=  prob_hi_x =  &   &  \\
-\verb=  prob_hi_y =  &   &  \\
-\verb=  prob_hi_z =  &   &  \\
+\verb=  ppm_type  = &   0 = no ppm 1 = 1985 ppm 2 = 2009 ppm  &  0 \\
+\verb=  prob_hi_x  = &    &  1.0 \\
+\verb=  prob_hi_y  = &    &  1.0 \\
+\verb=  prob_hi_z  = &    &  1.0 \\
 \verb=  prob_lo_x  = &    &  ZERO \\
@@
 \verb=  prob_lo_z  = &    &  ZERO \\
-\verb=  radius =  &   &  1.0e6 \\
+\verb=  radius  = &    &  1.0e6\_dp\_t \\
 \verb=  ref_ratio  = &    &  2 \\
@@
 \verb=  stop_time  = &    &  -1.d0 \\
-\verb=  temp_diffusion_formulation =  &   &  .false. \\
-\verb=  test_set =  &   &  '' \\
-\verb=  the_copy_cache_max =  &   & 50 \\
-\verb=  the_knapsack_verbosity =  &   &  .false. \\
+\verb=  temp_diffusion_formulation  = &   How to compute the explicit thermal diffusion term. .true. = in terms of $T$; .false. = in terms of $\rho,p_0,X$.  &  .false. \\
+\verb=  test_set  = &   Fixed grid file.  &  '' \\
+\verb=  the_copy_cache_max  = &    Number of boxassoc layouts we keep in memory to avoid having to recompute the boxassoc, which is computationally expensive.  &  50 \\
+\verb=  the_knapsack_verbosity  = &   Verbosity of the knapscak processor-to-grid algorithm.  &  .false. \\
 \verb=  the_layout_verbosity  = &    &  0 \\
-\verb=  thermal_diffusion_type =  &   &  1 \\
-\verb=  the_sfc_threshold =  &   & 5 \\
+\verb=  the_sfc_threshold  = &   When assigning processors for grids, this determines whether we use the sfc algorithm or knapsack algorithm.
+If the total number of grids divided by the number of processors is greater than this number, use sfc.
+&  5 \\
+\verb=  thermal_diffusion_type  = &   In the thermal diffusion solver, 1 = Crank-Nicholson; 2 = Backward Euler.  &  1 \\
 \verb=  use_delta_gamma1_term  = &    &  .false. \\
-\verb=  use_divu_firstdt =  &   &  .false. \\
-\verb=  use_eos_coulomb =  &   &  .false. \\
+\verb=  use_divu_firstdt  = &   Use the divu constraint when computing the first time step.  &  .false. \\
+\verb=  use_eos_coulomb  = &    &  .true. \\
 \verb=  use_etarho  = &    &  .true. \\
-\verb=  use_soundspeed_firstdt =  &   &  .false. \\
-\verb=  use_tfromp =  &   &  .false. \\
-\verb=  use_thermal_diffusion =  &   &  .false. \\
-\verb=  velpert_amplitude =  &   &  0.d0 \\
-\verb=  velpert_radius =  &   &  0.75d8 \\
-\verb=  velpert_scale =  &   &  0.8d8 \\
-\verb=  velpert_steep =  &   &  1.d0 \\
-\verb=  verbose =  &   &  0 \\
-\verb=  w0mac_interp_type =  &   &  1 \\
-\verb=  xrb_pert_factor =  &   &  1.d-2 \\
-\verb=  xrb_pert_size =  &   &  5.0d1 \\
-\verb=  xrb_pert_type =  &   &  1 \\
-\verb=  xrb_use_bottom_sponge =  &   &  .true. \\
+\verb=  use_soundspeed_firstdt  = &   Use the soundspeed constraint when computing the first time step.  &  .false. \\
+\verb=  use_tfromp  = &   When updating temperature, use $T=T(\rho,p_0,X) $ rather than $T=T(\rho,h,X)$.  &  .false. \\
+\verb=  use_thermal_diffusion  = &   Use thermal diffusion.  &  .false. \\
+\verb=  verbose  = &   General verbosity  &  0 \\
+\verb=  w0mac_interp_type  = &   0 = Compute w0 at edges using a projection; 1 = Interpolate w0 to cell centers, then average to edges; 2 = Interpolate w0 to edges directly; 3 = Interpolate w0 to nodes, then average to edges.
+&  1 \\
 

### 2009-04-14 docs/runtime_parameters/rp.py

@@ -1,0 +1,135 @@
+#!/usr/bin/env python
+import sys
+
+# tex format stuff
+header=r"""
+\section{Runtime Parameters}
+
+%%%%%%%%%%%%%%%%
+% symbol table
+%%%%%%%%%%%%%%%%
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{3.25in}|l|}
+\caption[runtime parameters]{runtime parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+"""
+
+footer=r"""
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+%
+
+"""
+
+fParallel="../../.."
+paramFile=fParallel + "/MAESTRO/_parameters"
+
+
+# container class for the parameters
+class Parameter:
+    def __init__(self):
+        self.var=""
+        self.default=""
+        self.description=[]
+
+
+
+def make_tex_table():
+
+    # open the file
+    try: f = open(paramFile, "r")
+    except IOError:
+        print "ERROR: %s does not exist" % paramFile
+        sys.exit(2)
+
+    # local storage for the parameters
+    paramsDict={}
+    paramsList=[]
+    descr=r""
+
+    # read in the file
+    # skip all lines before the first empty line
+    foundFirstParam = False
+
+    for line in f:
+        if not foundFirstParam:
+            if line.isspace():
+                # this is the first empty line and we begin reading the file 
+                # from here on out
+                foundFirstParam = True
+                continue
+            # no blank line found yet, keep going
+            continue
+
+        # land here once we have found the first parameter
+        currentParam = Parameter()
+        
+        # skip blank lines
+        if line.isspace(): 
[…]

### 2009-04-14 docs/runtime_parameters/rp.sh

@@ -1,60 +1,0 @@
-#!/bin/ksh
-
-# dump the header
-cat <<EOF
-
-\section{Runtime Parameters}
-
-%%%%%%%%%%%%%%%%
-% symbol table
-%%%%%%%%%%%%%%%%
-
-{\small
-
-\renewcommand{\arraystretch}{1.5}
-%
-\begin{center}
-\begin{longtable}{|l|p{3.25in}|l|}
-\caption[runtime parameters]{runtime parameters.} \label{table:runtime} \\\\
-%
-\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
-       \multicolumn{1}{ c|}{\textbf{description}} & 
-       \multicolumn{1}{ c|}{\textbf{default value}} \\\\ \hline 
-\endfirsthead
-
-\multicolumn{3}{c}%
-{{\tablename\ \thetable{}---continued}} \\\\
-\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
-       \multicolumn{1}{ c|}{\textbf{description}} & 
-       \multicolumn{1}{ c|}{\textbf{default value}} \\\\ \hline 
-\endhead
-
-\multicolumn{3}{|r|}{{\em continued on next page}} \\\\ \hline
-\endfoot
-
-\hline 
-\endlastfoot
-
-EOF
-
-# make the table entries
-parameters=$(grep namelist probin.f90 | awk '{print $3}' | sort)
-
-for i in ${parameters}
-do
-    value=$(grep -E -m 1 "^[ ]*$i +=" probin.f90 | cut -d'=' -f2 | cut -d"!" -f1)
-
-    echo "\\verb= " $i "=" " &   & " `echo ${value} | sed 's/\_/\\\_/g'` "\\\\"
-done
-
-# dump the end
-cat <<EOF
-
-\end{longtable}
-\end{center}
-
-} % ends \small
-
-%
-
-EOF

### 2009-04-15 docs/runtime_parameters/runtime_parameters.tex
fixed the formatting of the tex file
now to make the description explicitly multiline, either add a "\newline" in
description, or use the shortcut "@@" which will be parsed by rp.py.
for example:

# 0 = Compute w0 at edges using a projection; @@
# 1 = Interpolate w0 to cell centers, then average to edges; @@
# 2 = Interpolate w0 to edges directly; @@
# 3 = Interpolate w0 to nodes, then average to edges.
w0mac_interp_type                   integer            1

or

# 0 = Compute w0 at edges using a projection; \newline
# 1 = Interpolate w0 to cell centers, then average to edges; \newline
# 2 =

@@
 \verb=  beta_type  = &   what type of coefficient to use inside the velocity divergence constraint.
-{\tt beta\_type} = 1 uses $\beta_0$; {\tt beta\_type} = 2 uses $\rho_0$ (anelastic); {\tt beta\_type} = 3 uses 1 (small-scale combustion).
+\newline {\tt beta\_type} = 1 uses $\beta_0$; \newline {\tt beta\_type} = 2 uses $\rho_0$ (anelastic); \newline {\tt beta\_type} = 3 uses 1 (small-scale combustion).
 &  1 \\
@@
 \verb=  check_base_name  = &   prefix to use in checkpoint file names  &  "chk" \\
-\verb=  chk_int  = &   number of timesteps between writing a checkpoint file  &  0 \\
+\verb=  chk_int  = &   Number of timesteps between writing a checkpoint file  &  0 \\
 \verb=  co_latitude  = &   latitude, in radians, for problems with rotation where the domain is only a subset of a full star.  &  ZERO \\
@@
 \verb=  drdxfac  = &   ratio of radial base state zones to Cartesian full state zones for spherical geometry  &  1 \\
-\verb=  enthalpy_pred_type  = &   predict\_rhohprime = 1; predict\_h = 2; predict\_T\_then\_rhohprime = 3; predict\_T\_then\_h = 4; predict\_hprime = 5; predict\_Tprime\_then\_h = 6.
+\verb=  enthalpy_pred_type  = &   predict\_rhohprime = 1; \newline predict\_h = 2; \newline predict\_T\_then\_rhohprime = 3; \newline predict\_T\_then\_h = 4; \newline predict\_hprime = 5; \newline predict\_Tprime\_then\_h = 6.
 &  predict\_rhohprime \\
@@
 \verb=  init_shrink  = &   the multiplicative factor ($\le 1$) to reduce the initial timestep as computed by the various timestep estimators  &  1.0 \\
-\verb=  interp_type_radial_bin_to_cart  = &   The interpolation for filling a full multifab from a 1D bin-centered array. 1 = piecewise constant; 2 = piecewise linear.  &  1 \\
+\verb=  interp_type_radial_bin_to_cart  = &   The interpolation for filling a full multifab from a 1D bin-centered array. \newline 1 = piecewise constant; \newline 2 = piecewise linear.  &  1 \\
 \verb=  job_name  = &    &  "" \\
@@
 \verb=  pmask_z  = &    &  .false. \\
-\verb=  ppm_type  = &   0 = no ppm 1 = 1985 ppm 2 = 2009 ppm  &  0 \\
+\verb=  ppm_type  = &   0 = no ppm \newline 1 = 1985 ppm \newline 2 = 2009 ppm  &  0 \\
 \verb=  prob_hi_x  = &    &  1.0 \\
@@
 \verb=  stop_time  = &    &  -1.d0 \\
-\verb=  temp_diffusion_formulation  = &   How to compute the explicit thermal diffusion term. .true. = in terms of $T$; .false. = in terms of $\rho,p_0,X$.  &  .false. \\
+\verb=  temp_diffusion_formulation  = &   How to compute the explicit thermal diffusion term.  \newline .true. = in terms of $T$; \newline .false. = in terms of $\rho,p_0,X$.  &  .false. \\
 \verb=  test_set  = &   Fixed grid file.  &  '' \\
@@
 \verb=  verbose  = &   General verbosity  &  0 \\
-\verb=  w0mac_interp_type  = &   0 = Compute w0 at edges using a projection; 1 = Interpolate w0 to cell centers, then average to edges; 2 = Interpolate w0 to edges directly; 3 = Interpolate w0 to nodes, then average to edges.
+\verb=  w0mac_interp_type  = &   0 = Compute w0 at edges using a projection; \newline 1 = Interpolate w0 to cell centers, then average to edges; \newline 2 = Interpolate w0 to edges directly;                \newline 3 = Interpolate w0 to nodes, then average to edges.
 &  1 \\

### 2009-04-15 docs/runtime_parameters/rp.py

@@ -97,7 +97,7 @@ def make_tex_table():
 
         elif line.startswith("#"):
             # handle descriptions here
-            descr+=line[1:].rstrip()
+            descr+=line[1:].rstrip().replace("@@",r"\newline")
             continue
 
         else:

### 2009-04-17 docs/runtime_parameters/runtime_parameters.tex
-fixed outdated comments in advance.f90 -more progress on runtime_parameters.tex -added new flag to _parameters: p0_update_type, default 1  # How we update the base state pressure and compute psi. @@  # 1 = call {\tt advect\_base\_pres}; @@  # 2 = call {\tt enforce\_HSE}.

@@
 \verb=  max_levs  = &   Total number of levels.  1 = single level.  &  1 \\
+\verb=  max_mg_bottom_nlevels  = &   if mg\_bottom\_solver == 4, then how many mg levels can the bottom solver mgt object have  &  -1 \\
 \verb=  max_step  = &   Maximum number of steps in the simulation.  &  1 \\
@@
 \verb=  min_width  = &   The minimum size on a side for a grid created using make\_new\_grids.  &  16 \\
-\verb=  model_file  = &    &  "model.hse" \\
+\verb=  model_file  = &   input model file  &  "model.hse" \\
 \verb=  nOutFiles  = &   If lUsingNFiles = .true., use this many processors to write checkpoint and plotfiles.
@@
 \verb=  n_cellz  = &   Number of cells for the base level in the z-direction  &  -1 \\
+\verb=  p0_update_type  = &   How we update the base state pressure and compute psi. \newline 1 = call {\tt advect\_base\_pres}; \newline 2 = call {\tt enforce\_HSE}.  &  1 \\
 \verb=  perturb_model  = &   Turn on a perturbation in the initial data.  Problem specific.  &  .false. \\
-\verb=  plot_base  = &    &  .false. \\
-\verb=  plot_base_name  = &    &  "plt" \\
-\verb=  plot_deltat  = &    &  0.0 \\
-\verb=  plot_int  = &    &  0 \\
-\verb=  plot_spec  = &    &  .true. \\
-\verb=  plot_trac  = &    &  .false. \\
-\verb=  pmask_x  = &    &  .false. \\
-\verb=  pmask_y  = &    &  .false. \\
-\verb=  pmask_z  = &    &  .false. \\
+\verb=  plot_base  = &   plot w0\_x, w0\_y, w0\_z, divw0, rho0, rhoh0, h0, and p0 in plotfile  &  .false. \\
+\verb=  plot_base_name  = &   prefix to use in plotfile file names  &  "plt" \\
+\verb=  plot_deltat  = &   rather than use a plot interva, plot a file after the solution has advanced past plot\_deltat in time  &  0.0 \\
+\verb=  plot_int  = &   plot interval  &  0 \\
+\verb=  plot_spec  = &   plot species and omegadot in plotfile  &  .true. \\
+\verb=  plot_trac  = &   plot tracers in plotfile  &  .false. \\
+\verb=  pmask_x  = &   indicates whether the problem domain periodic in the x-direction  &  .false. \\
+\verb=  pmask_y  = &   indicates whether the problem domain periodic in the y-direction  &  .false. \\
+\verb=  pmask_z  = &   indicates whether the problem domain periodic in the z-direction  &  .false. \\
 \verb=  ppm_type  = &   0 = no ppm \newline 1 = 1985 ppm \newline 2 = 2009 ppm  &  0 \\
-\verb=  prob_hi_x  = &    &  1.0 \\
-\verb=  prob_hi_y  = &    &  1.0 \\
-\verb=  prob_hi_z  = &    &  1.0 \\
-\verb=  prob_lo_x  = &    &  ZERO \\
-\verb=  prob_lo_y  = &    &  ZERO \\
-\verb=  prob_lo_z  = &    &  ZERO \\
+\verb=  prob_hi_x  = &   physical coordinates of hi-x corner of problem domain  &  1.0 \\
+\verb=  prob_hi_y  = &   physical coordinates of hi-y corner of problem domain  &  1.0 \\
+\verb=  prob_hi_z  = &   physical coordinates of hi-z corner of problem domain  &  1.0 \\
+\verb=  prob_lo_x  = &   physical coordinates of lo-x corner of problem domain  &  ZERO \\
+\verb=  prob_lo_y  = &   physical coordinates of lo-y corner of problem domain  &  ZERO \\
+\verb=  prob_lo_z  = &   physical coordinates of lo-z corner of problem domain  &  ZERO \\
 \verb=  radius  = &    &  1.0e6\_dp\_t \\
@@
 \verb=  regrid_int  = &    &  -1 \\
-\verb=  restart  = &    &  -1 \\
+\verb=  restart  = &   which file to restart from.  -1 means do not restart  &  -1 \\
 \verb=  rotational_frequency  = &    &  ZERO \\
@@
 \verb=  small_temp  = &    &  5.d6 \\
-\verb=  spherical_in  = &    &  0 \\
-\verb=  sponge_kappa  = &    &  10.d0 \\
-\verb=  stop_time  = &    &  -1.d0 \\
+\verb=  spherical_in  = &   set to 1 if you are doing a spherical problem  &  0 \\
+\verb=  sponge_kappa  = &   Parameter for sponge.  Problem dependent.  &  10.d0 \\
+\verb=  stop_time  = &   simulation stop time  &  -1.d0 \\
 \verb=  temp_diffusion_formulation  = &   How to compute the explicit thermal diffusion term.  \newline .true. = in terms of $T$; \newline .false. = in terms of $\rho,p_0,X$.  &  .false. \\

### 2009-04-17 Source/advance.f90

@@ -171,23 +171,15 @@ contains
        end if
     end if
 
-    ! Set this to zero so if evolve_base_state = F there is no effect in update_vel
-    w0_force = ZERO
-
-    ! Set Sbar to zero so if evolve_base_state = F then it doesn't affect rhs of projections
-    Sbar = ZERO
-
-    ! Set these to results from last
+    ! Initialize these to previous values
     w0_old = w0
     gamma1bar_old = gamma1bar
-
-    ! Set this to zero since we only compute this if dpdt_factor > 0
-    p0_minus_pthermbar = ZERO
 
     halfdt = half*dt
 
     call compute_cutoff_coords(rho0_old)
     
+    ! compute tempbar by "averaging"
     if (spherical .eq. 0) then
        call average(mla,sold,tempbar,dx,temp_comp)
     else
@@ -224,11 +216,10 @@ contains
        call setval(delta_p_term(n),ZERO,all=.true.)
     end do
 
+    ! compute p0_minus_pthermbar = p0_old - pthermbar (for making w0) and
+    ! compute delta_p_term = ptherm_old - pthermbar_cart (for RHS of projections)
     if (dpdt_factor .gt. ZERO ) then
     
-       ! compute p0_minus_pthermbar = p0_old - pthermbar (for making w0)
-       ! and delta_p_term = ptherm_old - pthermbar_cart (for RHS of projection)
-
        ! ptherm_old now holds the thermodynamic p computed from sold(rho,h,X)
        call makePfromRhoH(sold,sold,ptherm_old,mla,the_bc_tower%bc_tower_array)
 
@@ -242,7 +233,7 @@ contains
           call multifab_build(pthermbar_cart(n), mla%la(n), 1, 0)
        end do
        
-       ! compute pthermbar_cart = fill(pthermbar)
+       ! compute pthermbar_cart from pthermbar
        call put_1d_array_on_cart(pthermbar,pthermbar_cart,foextrap_comp, &
                                  .false.,.false.,dx,the_bc_tower%bc_tower_array,mla)
 
@@ -255,6 +246,11 @@ contains
        do n=1,nlevs
           call destroy(pthermbar_cart(n))
        end do
+
+    else
+
+       ! this should have no effect if dpdt_factor .le. 0
+       p0_minus_pthermbar = ZERO
 
     end if
 
@@ -285,6 +281,12 @@ contains
[…]

### 2009-04-17 docs/runtime_parameters/runtime_parameters.tex
changed "radius" in _parameters to "rotation_radius" so we don't confuse with the many local radius we declare throughout the code

@@
 \verb=  interp_type_radial_bin_to_cart  = &   The interpolation for filling a full multifab from a 1D bin-centered array. \newline 1 = piecewise constant; \newline 2 = piecewise linear.  &  1 \\
-\verb=  job_name  = &    &  "" \\
+\verb=  job_name  = &   job name printed in output  &  "" \\
 \verb=  lUsingNFiles  = &   If .true., use nOutFiles processors to write checkpoint and plotfiles.
@@
 \verb=  mg_verbose  = &   Verbsoity of the multigrid solver, but not the bottom solver.  &  0 \\
-\verb=  min_eff  = &    &  0.7d0 \\
+\verb=  min_eff  = &   parameter for cluster algorithm for making new grids in adaptive problems  &  0.7d0 \\
 \verb=  min_width  = &   The minimum size on a side for a grid created using make\_new\_grids.  &  16 \\
@@
 \verb=  prob_lo_z  = &   physical coordinates of lo-z corner of problem domain  &  ZERO \\
-\verb=  radius  = &    &  1.0e6\_dp\_t \\
-\verb=  ref_ratio  = &    &  2 \\
-\verb=  regrid_int  = &    &  -1 \\
+\verb=  ref_ratio  = &   Refinement ratio for multilevel problems  &  2 \\
+\verb=  regrid_int  = &   How often we regrid.  &  -1 \\
 \verb=  restart  = &   which file to restart from.  -1 means do not restart  &  -1 \\
-\verb=  rotational_frequency  = &    &  ZERO \\
+\verb=  rotation_radius  = &   radius used for computing centrifugal term in rotation problems  &  1.0e6\_dp\_t \\
+\verb=  rotational_frequency  = &   rotational frequency used for computing centrifugal term in rotation problems.  &  ZERO \\
 \verb=  single_prec_plotfiles  = &    &  .false. \\
-\verb=  slope_order  = &    &  4 \\
+\verb=  slope_order  = &   order of slopes in Godunov algorithm.  Options are 0, 2, or 4.  &  4 \\
 \verb=  small_dens  = &    &  1.d-5 \\
@@
 \verb=  thermal_diffusion_type  = &   In the thermal diffusion solver, 1 = Crank-Nicholson; 2 = Backward Euler.  &  1 \\
-\verb=  use_delta_gamma1_term  = &    &  .false. \\
+\verb=  use_delta_gamma1_term  = &   turns on second order correction to delta gamma1 term  &  .false. \\
 \verb=  use_divu_firstdt  = &   Use the divu constraint when computing the first time step.  &  .false. \\
 \verb=  use_eos_coulomb  = &    &  .true. \\
-\verb=  use_etarho  = &    &  .true. \\
+\verb=  use_etarho  = &   turn on the etarho term as described in flow chart  &  .true. \\
 \verb=  use_soundspeed_firstdt  = &   Use the soundspeed constraint when computing the first time step.  &  .false. \\

### 2009-04-17 Source/geometry.f90

@@ -334,7 +334,7 @@ contains
 
   subroutine init_rotation()
 
-    use probin_module, only: rotational_frequency, co_latitude, radius
+    use probin_module, only: rotational_frequency, co_latitude, rotation_radius
     use bl_constants_module, only: M_PI, ZERO
 
     real(dp_t) :: theta_in_rad
@@ -346,9 +346,9 @@ contains
 
     omega = 2 * M_PI * rotational_frequency
 
-    centrifugal_term(1) = - omega * omega * radius * sin_theta * cos_theta
+    centrifugal_term(1) = - omega * omega * rotation_radius * sin_theta * cos_theta
     centrifugal_term(2) = ZERO
-    centrifugal_term(3) = - omega * omega * radius * sin_theta * sin_theta
+    centrifugal_term(3) = - omega * omega * rotation_radius * sin_theta * sin_theta
 
   end subroutine init_rotation
 

### 2009-04-17 docs/initial_models/initial_models.tex
put the caption next to the figure so it takes less space
simplification to spherical initialization.  Before, we would fill rho, X, and p0 onto a multifab and then call the eos for h and T.  Since fill is just a piecewise constant operation, might as well use the base state h and T we already had.  Changes initial data by solver tolerance, but at least the code is doing the right thing now.

@@
 \begin{figure}[tpb]
-\begin{center}
-\epsffile{\figpath/multilevel_init.eps}
+\centering
+\includegraphics[width=0.4in]{\figpath/multilevel_init}\hspace{0.2in}
+\begin{minipage}[b]{5.0in}
 \caption{Multilevel Initialization.  The data from the initial model
@@
   using the two nearest points to copy data from the dots to the
-  squares.}
+  squares.\vspace{2em}}
+\end{minipage}
 \label{Fig:Multilevel Initialization}
-\end{center}
 \end{figure}
@@
 \begin{enumerate}
-\item For plane-parallel problems, fill $\rho^{\init}, h^{\init},
-  X^{\init}$, and $T^{\init}$ onto a multifab to obtain the full state
-  $\rho, h, X$, and $T$.  For spherical problems, fill $\rho^{\init},
-  p^{\init}$, and $X^{\init}$ onto a multifab to obtain the full state
-  $\rho, p$, and $X$, and then call the EOS to obtain the full state
-  $h,T = h,T(\rho,p,X)$.  Note that this particular multifab version
-  of $p$ is temporary, and can be deallocated after this step.
-\item For spherical problems, skip the remaining steps and instead set
-  $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
+\item Fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto a 
+  multifab to obtain the full state $\rho, h, X$, and $T$.
 \item If {\tt perturb\_model} = F, skip the remaining steps and
@@
 \item Compute $p_0$ using equation (\ref{HSE Discretization}), using
-  the pressure at the top of the domain as the initial condition.
-  Note that we still integrate upwards from the bottom, but then
+  the $p^{\init}$ at the top of the domain as the one required boundary 
+  condition.  Note that we still integrate upwards from the bottom, but then
   offset the final pressure so the pressure at the top of the domain

### 2009-04-17 SCIENCE/wdconvect/initdata.f90

@@ -142,7 +142,6 @@ contains
     real(kind=dp_t) :: x,y,z
     real(kind=dp_t) :: dens_pert, rhoh_pert, temp_pert
     real(kind=dp_t) :: rhoX_pert(nspec), trac_pert(ntrac)
-    real(kind=dp_t), allocatable :: p0_cart(:,:,:,:)
 
     if (perturb_model) then
        call bl_error('perturb_model not written for initscalardata_3d_sphr')
@@ -151,21 +150,15 @@ contains
     ! initial the domain with the base state
     s = ZERO
 
-    ! if we are spherical, we want to make sure that p0 is good, since that is
-    ! what is needed for HSE.  Therefore, we will put p0 onto a cart array and
-    ! then initialize h from rho, X, and p0.
-    allocate(p0_cart(lo(1):hi(1),lo(2):hi(2),lo(3):hi(3),1))
-
     ! initialize the scalars
     call put_1d_array_on_cart_3d_sphr(.false.,.false.,s0_init(:,rho_comp), &
                                       s(:,:,:,rho_comp:),lo,hi,dx,ng,0)
 
+    call put_1d_array_on_cart_3d_sphr(.false.,.false.,s0_init(:,rhoh_comp), &
+                                      s(:,:,:,rhoh_comp:),lo,hi,dx,ng,0)
+
     call put_1d_array_on_cart_3d_sphr(.false.,.false.,s0_init(:,temp_comp), &
                                       s(:,:,:,temp_comp:),lo,hi,dx,ng,0)
-
-    ! initialize p0_cart
-    call put_1d_array_on_cart_3d_sphr(.false.,.false.,p0_background(:), &
-                                      p0_cart(:,:,:,1:),lo,hi,dx,0,0)
 
     ! initialize species
     do comp = spec_comp, spec_comp+nspec-1
@@ -178,36 +171,6 @@ contains
        call put_1d_array_on_cart_3d_sphr(.false.,.false.,s0_init(:,comp), &
                                          s(:,:,:,comp:),lo,hi,dx,ng,0)
     end do
-
-    ! initialize (rho h) using the EOS
-    do k = lo(3), hi(3)
-       do j = lo(2), hi(2)
-          do i = lo(1), hi(1)
-
-             temp_eos(1) = s(i,j,k,temp_comp)
-             p_eos(1) = p0_cart(i,j,k,1)
-             den_eos(1) = s(i,j,k,rho_comp)
[…]

### 2009-04-17 TEST_PROBLEMS/spherical_heat/initdata.f90

@@ -46,13 +46,13 @@ contains
           
           select case (dm)
           case (2)
-             call bl_error('initscalardata not written for 2d')
+             call bl_error('initscalardata_2d not written')
           case (3)
              if (spherical .eq. 1) then
                 call initscalardata_3d_sphr(sop(:,:,:,:), lo, hi, ng, dx(n,:), &
                                             s0_init(1,:,:), p0_background(1,:))
              else
-                call bl_error('initscalardata not written for 3d plane parallel')
+                call bl_error('initscalardata_3d not written')
              end if
           end select
        end do
@@ -111,12 +111,12 @@ contains
        hi =  upb(get_box(s,i))
        select case (dm)
        case (2)
-          call bl_error('initscalardata_on_level not written for 2d')
+          call bl_error('initscalardata_2d not written')
        case (3)
           if (spherical .eq. 1) then
              call initscalardata_3d_sphr(sop(:,:,:,:),lo,hi,ng,dx,s0_init,p0_background)
           else
-             call bl_error('initscalardata_on_level not written for 3d plane parallel')
+             call bl_error('initscalardata_3d not written')
           end if
        end select
     end do
@@ -142,7 +142,6 @@ contains
     real(kind=dp_t) :: x,y,z
     real(kind=dp_t) :: dens_pert, rhoh_pert, temp_pert
     real(kind=dp_t) :: rhoX_pert(nspec), trac_pert(ntrac)
-    real(kind=dp_t), allocatable :: p0_cart(:,:,:,:)
 
     if (perturb_model) then
        call bl_error('perturb_model not written for initscalardata_3d_sphr')
@@ -151,21 +150,15 @@ contains
     ! initial the domain with the base state
     s = ZERO
 
-    ! if we are spherical, we want to make sure that p0 is good, since that is
-    ! what is needed for HSE.  Therefore, we will put p0 onto a cart array and
-    ! then initialize h from rho, X, and p0.
-    allocate(p0_cart(lo(1):hi(1),lo(2):hi(2),lo(3):hi(3),1))
-
     ! initialize the scalars
     call put_1d_array_on_cart_3d_sphr(.false.,.false.,s0_init(:,rho_comp), &
                                       s(:,:,:,rho_comp:),lo,hi,dx,ng,0)
 
[…]

### 2009-04-22 docs/eta_notes/eta.tex
more detail

@@
 \item Enter {\tt advance\_timestep} with $[\etarhoec, \etarhocc]^{n-\myhalf}$.
-\item Call {\tt make\_w0}.  The spherical version uses uses $\etarhoec$ and $\etarhocc$.
-\item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
-\item Call {\tt make\_etarho} to compute $[\etarhoec, \etarhocc]$
-\item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
-\item Call {\tt make\_w0}.  The spherical version uses uses $\etarhoec$ and $\etarhocc$.
-\item Call {\tt density\_advance}.  The plane-parallel version computes $\etarhoflux$.
-\item Call {\tt make\_etarho} to compute $[\etarhoec, \etarhocc]^{n+\myhalf}$
-\item Call {\tt make\_psi}.  The plane-parallel version uses $\etarhocc$.
+\item Call {\tt make\_w0}.  The spherical version uses uses $\etarho^{{\rm ec},n-\myhalf}$ and $\etarho^{{\rm cc},n-\myhalf}$.
+\item Call {\tt density\_advance}.  The plane-parallel version computes $\etarho^{{\rm flux},n+\myhalf,*}$.
+\item Call {\tt make\_etarho} to compute $[\etarhoec, \etarhocc]^{n+\myhalf,*}$.  The plane-parallel version uses $\etarho^{{\rm flux},n+\myhalf,*}$.
+\item Call {\tt make\_psi}.  The plane-parallel version uses $\etarho^{{\rm cc},n+\myhalf,*}$.
+\item Call {\tt make\_w0}.  The spherical version uses uses $\etarho^{{\rm ec},n+\myhalf,*}$ and $\etarho^{{\rm cc},n+\myhalf,*}$.
+\item Call {\tt density\_advance}.  The plane-parallel version computes $\etarho^{{\rm flux},n+\myhalf}$.
+\item Call {\tt make\_etarho} to compute $[\etarhoec, \etarhocc]^{n+\myhalf}$.  The plane-parallel version uses $\etarho^{{\rm flux},n+\myhalf}$.
+\item Call {\tt make\_psi}.  The plane-parallel version uses $\etarho^{{\rm cc},n+\myhalf}$.
 \end{enumerate}
@@
 \subsection{Plane-Parallel}
-We first compute a radial edge-centered multifab, $\eta_{\rho}^{\rm flux}$, using
+We first compute a radial edge-centered multifab, $\etarho^{\rm flux}$, using
 \begin{equation}

### 2009-04-22 docs/flowchart/flowchart.tex
minor cleanup
starting to merge spherical base state stuff into flowchart

@@
 \end{equation}
-\\ \\
+\section{Shorthand Functions}
 We make use of the following shorthand notations in outlining the algorithm:
@@
 equation (paper III; eq.~[29]) without the $\etarho$ term,
-\begin{eqnarray}
-\frac{\partial \rhozero}{\partial t} &=& - \nablab \cdotb \left( \rhozero w_0 \er \right)
+\begin{equation}
+\frac{\partial \rhozero}{\partial t} = - \nablab \cdotb \left( \rhozero w_0 \er \right)
 \enskip ,
 \label{eq:rho0upd_new}
-\end{eqnarray}
+\end{equation}
 to compute the new base state density,
-\begin{eqnarray}
-\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
-\left [ \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
-         \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ]
+\begin{equation}
+\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
+\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
+         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ]
          \nonumber \enskip,
-\end{eqnarray}
+\end{equation}
 where $j$ refers to the one-dimensional index in the radial direction.
-The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the
-procedure described in paper II, Appendix A.
-
-\item[spherical:] See Section \ref{Sec:Advect Base Spherical} for working notes.
+The interface states, $\rho_0^{\outp,\nph}$, are found via the
+procedure described in paper II, Appendix A.\\
+
+\item[spherical:] The base state density update now includes the area factors in the 
+divergences:
+\begin{equation}
+\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp}
+ - \frac{1}{r_j^2} \frac{\dt}{\dr} 
+ \left [ \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
+          \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right] \enskip. 
+\end{equation}
 
@@
 (\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
-\left\{ \left[ {(\rho h)_0}^{\nph} {w_0}^{\inp}\right]_{j+\myhalf} -  
-         \left[ {(\rho h)_0}^{\nph} {w_0}^{\inp}\right]_{j-\myhalf} \right\}
+\left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} -  
+         \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\}
 + \dt\psi_j^{\inp} \enskip ,
@@
 is the process by which we adjust the base state density to account for
-large-scale mixing.  It can also be used to adjust the base state enthalpy
+large-scale mixing.  It is also used to adjust the base state enthalpy
 to account for large-scale mixing, reactions, and thermal diffusion.
@@
 \begin{equation}
-T^{(2)} = T\left(\rho^{(2)}, {p_0}^{(2)}, X_k^{(2)}\right) \enskip .
+T^{(2)} = T\left(\rho^{(2)}, p_0^{(2)}, X_k^{(2)}\right) \enskip .
 \end{equation}
@@
 \begin{eqnarray}
- \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1}, {X_k}^{n+1}\right) \right] \enskip , \\ 
+ \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1}, X_k^{n+1}\right) \right] \enskip , \\ 
  {\beta   }_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip .

### 2009-04-22 docs/spherical_basestate/basestate.tex
shrunk the image so the caption is readable
removed a marginpar we've addressed
starting to merge spherical base state stuff into flowchart

@@
 - \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip.
-\end{eqnarray} \MarginPar{note in the RHS, we do a difference of cell-centered $\eta$'s, but $\eta$ is defined on edges -- what should we do?}
+\end{eqnarray}
 
@@
 \begin{enumerate}
-
-\item {\bf Density Update:}
-
-The base state density update now includes the area factors in the divergences:
-\begin{eqnarray}
-\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp}
- - \frac{1}{r_j^2} \frac{\dt}{\dr} 
- \left [ \left( {{r^2 \rho}_0}^{\inp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
-          \left( {{r^2 \rho}_0}^{\inp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right]   \enskip. 
-\end{eqnarray}
-As before, the interface states are found via the procedure described in
-paper II, Appendix A.
 
@@
 \begin{center}
+\epsfxsize=4.0in
 \epsffile{\sphericalfigpath/spherical_adjust_768.eps}

### 2009-04-22 docs/flowchart/flowchart.tex
more work on spherical base state

@@
 
-\item {\bf Advect Base Pressure}$[p_0^\inp,w_0^\inp,\etarho^\inp] \rightarrow$ 
-$[p_0^\outp,\psi^\outp]$\\
+\item {\bf Advect Base Pressure}
 is the process by which we update the base state pressure through $\dt$ in time and
-compute $\psi$.
+compute $\psi$.  The input arguments are different for plane-parallel and spherical.
 \begin{description}
-
-\item[plane-parallel:] First we compute a cell-centered $\psi^\outp$ using
+\item {\bf Advect Base Pressure Planar}$[p_0^\inp,w_0^\inp,\etarho^\inp] \rightarrow 
+[p_0^\outp,\psi^\outp]$:
+
+First we compute a cell-centered $\psi^\outp$ using
 \begin{equation}
@@
 where the interface states are again found via the procedure described
-in paper II, Appendix A.
-
-\item[spherical:] See Section \ref{Sec:Advect Base Spherical} for working notes.
-
+in paper II, Appendix A.\\
+
+\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},\Sbar^{\inp}] 
+\rightarrow [p_0^{\outp},\psi^{\outp}]$:
+
+First use a predictor-corrector formulation for the pressure update 
+in order to improve the overall accuracy.  Beginning with equation~(\ref{eq:w01}),
+we can write
+\begin{eqnarray*}
+\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
+ \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) - \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
+\end{eqnarray*}
+We time-center the left hand side:
+\[
+\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
+   \frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} - p_0^{\inp}}{\dt}\enskip.
+\]
+We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
+\begin{equation}
+p_0^{\star} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
+\end{equation}
+with:
+\begin{equation}
+f = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
+                                             \frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} \left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
+\end{equation}
+where the $(\partial p_0^\inp / \partial r)$ is computed by
+interpolating $p_0^\inp$ from cell-centers to edges and then
+differencing.  Now we begin the corrector step.  Using $p_0^\star$, we can compute a more 
+accurate approximation to equation~(\ref{eq:frhs}),
+\begin{equation}
+\tilde{f} = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
+                                             \frac{1}{\gammabar^{\inp} \tilde{p_0}} w_0^{\inp} \left ( \frac{\partial \tilde{p_0}}{\partial r} \right ) \right ] \enskip ,
+\end{equation}
+where $\tilde{p} = (p_0^\inp + p_0^\star)/2$, and the difference is again computed
+by first putting $\tilde{p}$ on edges.
+
+This allows us to complete the corrector for the pressure update:
+\begin{equation}
+p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} \tilde{f}\right)}{\left(1 - \gammabar^{\inp} \tilde{f}\right)}\enskip .
+\end{equation}
+We note that since we do not time-center $\gammabar$, this is not
+strictly second-order accurate.  However, as $\gammabar$ varies slowly
+in time, we expect the effect to be small.  This is something that we 
+defer until simulations indicate it is warranted.
+
+Next, we compute $\psi^{\outp}$.  Rewriting Equation (\ref{eq:w01}), we can define
+\begin{eqnarray}
+\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
+       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
+\end{eqnarray}
+We discretize this locally as:
+\begin{equation}
+\psi_j^{\outp} = 
+\left[\gammabar^{\inp} \frac{\left( p_0^{\inp} + p_0^{\outp} \right)}{2}\right]_j
+\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
+          (r^2 w_0^{\outp})_{j-\myhalf} \right ] \right \} \enskip .
+\end{equation}
 \end{description}
@@
 \item[spherical:]
-\MarginPar{Need to write this up somewhere}
 

### 2009-04-22 docs/spherical_basestate/basestate.tex
more work on spherical base state

@@
 \end{eqnarray}
-\subsection{Calculation of $\psi$}
-
-Rewriting Equation (\ref{eq:w01}), we can define
-\begin{eqnarray}
-\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
-       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
-\end{eqnarray}
-We discretize this locally as:
-\begin{eqnarray}
-\psi_j^{\nph,\star} &=& \gammabar^{(1)} \frac{\left( p_0^{n} + p_0^{n+1,\star} \right)_j}{2}
-\left \{ \overline{S^{\nph,\star}_j} - 
- \frac{1}{r_j^2} \left [ (r^2 {w_0}^{\nph,\star})_{j+\myhalf} -
-          (r^2 {w_0}^{\nph,\star})_{j-\myhalf} \right ] \right \} \enskip , \nonumber \\
-&& \\
-\psi_j^{\nph} &=& \gammabar^{n+1,\star} \frac{\left( p_0^{n} + p_0^{n+1} \right)_j}{2}
-\left \{ \overline{S^{\nph}_j} - 
- \frac{1}{r_j^2} \left [ (r^2 {w_0}^{\nph})_{j+\myhalf} -
-          (r^2 {w_0}^{\nph})_{j-\myhalf} \right ] \right \} \enskip .
-\end{eqnarray}
-\MarginPar{Why is $\gammabar$ not time centered in the $\psi_j^{\nph}$ equation?}
-
-\subsection{{\bf Advect Base}}\label{Sec:Advect Base Spherical}
-
-\begin{enumerate}
-
-\item {\bf Pressure Update:}
-\newline
-We use a predictor-corrector formulation for the pressure update 
-in order to improve the overall accuracy.  Beginning with equation~(\ref{eq:w01}),
-we can write
-\begin{eqnarray*}
-\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
- \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) - \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
-\end{eqnarray*}
-We time-center the left hand side:
-\[
-\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
-   \frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} - p_0^{\inp}}{\dt}\enskip.
-\]
-We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
-\begin{equation}
-p_0^{\star} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
-\end{equation}
-with:
-\begin{equation}
-f = \frac{\dt}{2} \left [ \overline{S^{\inp}} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
-                                             \frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} \left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
-\end{equation}
-where the $(\partial p_0^\inp / \partial r)$ is computed by
-interpolating $p_0^\inp$ from cell-centers to edges and then
-differencing.
-
-Now we begin the corrector step.  Using $p_0^\star$, we can compute a more accurate approximation
-to equation~(\ref{eq:frhs}),
-\begin{equation}
-\tilde{f} = \frac{\dt}{2} \left [ \overline{S^{\inp}} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
-                                             \frac{1}{\gammabar^{\inp} \tilde{p_0}} w_0^{\inp} \left ( \frac{\partial \tilde{p_0}}{\partial r} \right ) \right ] \enskip ,
-\end{equation}
-where $\tilde{p} = (p_0^\inp + p_0^\star)/2$, and the difference is again computed
-by first putting $\tilde{p}$ on edges.
-
-This allows us to complete the corrector for the pressure update:
-\begin{equation}
-p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} \tilde{f}\right)}{\left(1 - \gammabar^{\inp} \tilde{f}\right)}\enskip .
-\end{equation}
-We note that since we do not time-center $\gammabar$, this is not
-strictly second-order accurate.  However, as $\gammabar$ varies slowly
-in time, we expect the effect to be small.  This is something that we 
-defer until simulations indicate it is warranted. \MarginPar{new}
-
-\end{enumerate}
-
-\subsection{\bf{Correct Base}}
-
-The is the process by which we adjust the base state density to account for
-large-scale mixing.  It can also be used to adjust the base state enthalpy
-to account for large-scale mixing, reactions, and thermal diffusion.
-We have simplified the algorithm greatly by simply setting $\nabla\cdot\etarho$
-equal to the averaged difference between the base state and full state.
-
-\begin{equation}
-\rho_0^{\outp} = \rho_0^{\inp} + \underbrace{{\rm{\bf Avg}}(\rho^{\inp} - \rho_0^{\inp})}_{-\Delta t\nabla\cdot\etarho}.
-\end{equation}
 

### 2009-04-23 docs/flowchart/flowchart.tex
for spherical, we need to compute psi at time n and n+1/2
more flowchart progress
rewrote the "Changes since" section so it's accurate

@@
 
-%-----------------------------------------------------------------------------
-% Changes Between Paper 3 and Paper 4
-%-----------------------------------------------------------------------------
-
 \section{Changes Between Paper 3 and Paper 4}
-
-There are several changes to the algorithm since it was last detailed
-in paper III.  The major changes are:
-
-\begin{itemize}
-
-\item We have added thermal diffusion with {\tt use\_thermal\_diffusion = TRUE}.
-
-\item We discuss spherical geometries, in paricular:
-\begin{itemize}
-\item We are still working out the fill, average, and interpolate stencils.
-\item We are working on {\bf Advect Base} and {\bf Correct Base} details.
-\item We are working on $\etarho$ details.
-\item We compute the full state temperature on cell-centers after the 
-advection-diffusion step using $T=T(\rho,X_k,p_0)$ by setting 
-{\tt use\_tfromp = TRUE} rather than using $T=T(\rho,h,X_k)$.
+\begin{enumerate}
+\item {\tt fill\_3d} and {\tt average} algorithms.
+\item We update temperature after each {\bf React State} as well as the advection step.
+\item We have created a {\tt burning\_cutoff\_density}, where the burning does
+not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
+\item Use corner coupling in advection.
+\item We use {\tt use\_tfromp} to update temperature using $T=T(\rho,X_k,p_0)$ rather 
+than $T=T(\rho,h,X_k)$.
+\end{enumerate}
+Changes which modify enthalpy update.  These are separate because in this paper,
+enthalpy is only used as a diagnostic since we use {\tt use\_tfromp}.  If we don't
+mention these items in paper IV, they should me noted in paper 4.5.
+\begin{enumerate}
 \item We have changed the discretization of $\Ub\cdot\nabla p_0$ in the enthalpy
 update to $\nabla\cdot(\Ub p_0) - p_0\nabla\cdot\Ub$.
-\end{itemize}
-
 \item In paper III we discretized the enthalpy evolution equation in
@@
 This is more like paper II.
-
+\item We have turned off the evolution of $h$ above the atmosphere and instead
+compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
+\end{enumerate}
+\section{Changes Between Papers 4 and Papers 4.5}
+\begin{enumerate}
 \item We split the update of the density and enthalpy in steps 4 and 8.
@@
 $\psi$ that are then used in the enthalpy update.
-
 \item We split {\bf Advect Base} into separate {\bf Advect Base Density},
-{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.
-\begin{itemize}
-\item First, we call {\bf Advect Base Density}, which computes a time-centered
-$\etarho$.
-\item Second, we call {\bf Advect Base Pressure}, which takes $\etarho$ as an
-input and computes a time-centered $\psi$.
-\item Third, we call {\bf Advect Base Enthalpy}, which takes $\psi$ as an input.
-\end{itemize}
-
-\item We added the volume-discrepancy term to the velocity constraint equation,
-controlled by the input parameter, $f$.
-
-\item We have turned off the evolution of $h$ above the atmosphere and instead
-compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
-
-\item We have created a {\tt burning\_cutoff\_density}, where the burning does
-not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
-
-\item We have given the advection scheme a face-lift, with some nifty options like
-corner coupling and various PPM limiters.
-
-\item We now update the density and enthalpy as full states with full-state fluxes, 
-not as perturbational quantities.
-
-\item We update temperature at the end of {\bf React State}.
-
-\end{itemize}
-
-%-----------------------------------------------------------------------------
-% Changes Since Paper 4
-%-----------------------------------------------------------------------------
-
-\section{Changes Since Paper 4}
-
-\begin{itemize}
-
-\item In {\bf Step 4} and {\bf Step 8}, we moved {\bf Correct Base} so it
-occurs immediately before {\bf Advect Base Pressure}.  Since the full
-enthalpy advance ``feels'' the effects of the time-centered $\etarho$, we should
-rely on a $\rho_0^{n+1}$ that also feels the effects.
-
-\item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
-plane-parallel and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
-in spherical.  This means that {\bf React Base} no longer exists, and the intermediate states, 
-$(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
-{\bf Advect Base Enthalpy} will advance $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ 
-directly.  This also means we don't have to deal with $\eta_{\rho h}$ or a 
-{\bf Diffuse Base}.
-
+{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.  The calling sequence
+is in the main flow chart.
 \item We replaced {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
@@
 analogously for setting $(\rho h)_0$ after {\bf React State}.
-
-\end{itemize}
+\item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
+plane-parallel and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
+in spherical.  This means that {\bf React Base} no longer exists, and the intermediate 
+states, $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
+{\bf Advect Base Enthalpy} will advance $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ 
+directly.  This also means we don't have to deal with $\eta_{\rho h}$ or a 
+{\bf Diffuse Base}.
+\item In {\bf Step 4} and {\bf Step 8}, we moved {\bf Correct Base} so it
+occurs after the full state density update and  before {\bf Advect Base Pressure}.  
+Since the full enthalpy advance ``feels'' the effects of the time-centered $\etarho$, 
+we should rely on a $\rho_0^{n+1}$ that also feels the effects.
+\end{enumerate}
+Changes specific to spherical:
+\begin{enumerate}
+\item Discuss {\bf Advect Base} routines and {\bf Correct Base} routine.
+\item $\etarho$ and $\psi$ equations.
+\end{enumerate}
+\section{Changes Between Papers 4.5 and Paper 5}
+\begin{enumerate}
+\item Use PPM limiters in advection.
+\end{enumerate}
+\section{Changes Between Paper 5 and XRB Paper}
+\begin{enumerate}
+\item We have added thermal diffusion, controlled by {\tt use\_thermal\_diffusion},
+{\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
+\item We added the volume-discrepancy term to the velocity constraint equation,
+controlled by the input parameter, {\tt dpdt\_factor}.
+\end{enumerate}
 
@@
 \item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},\Sbar^{\inp}] 
-\rightarrow [p_0^{\outp},\psi^{\outp}]$:
+\rightarrow [p_0^{\outp},\psi^{n,\outp},\psi^{n+\myhalf,\outp}]$:
 
@@
 
-Next, we compute $\psi^{\outp}$.  Rewriting Equation (\ref{eq:w01}), we can define
+Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.  Rewriting Equation (\ref{eq:w01}), we can define
 \begin{eqnarray}
@@
 \begin{equation}
-\psi_j^{\outp} = 
+\psi_j^{\outp,n} = 
+\left(\gammabar^{\inp}p_0^{\inp}\right)_j
+\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
+          (r^2 w_0^{\outp})_{j-\myhalf} \right ] \right \} \enskip ,
+\end{equation}
+\begin{equation}
+\psi_j^{\outp,n+\myhalf} = 
 \left[\gammabar^{\inp} \frac{\left( p_0^{\inp} + p_0^{\outp} \right)}{2}\right]_j
@@
 
-\item {\bf Advect Base Enthalpy}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [(\rho h)_0^\outp]$\\
+\item {\bf Advect Base Enthalpy}
 is the process by which we update the base state enthalpy through $\dt$ in time.
+The input arguments are different for plane-parallel and spherical.
 \begin{description}
-
-\item[plane-parallel:] We discretize
+\item {\bf Advect Base Enthalpy Planar}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow 
+[(\rho h)_0^\outp]$
+
+We discretize the base state enthalpy equation,
 \begin{equation}
@@
 where the interface states are again found via the procedure described
-in paper II, Appendix A.
-
-\item[spherical:]
+in paper II, Appendix A.\\
+
+\item {\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^\inp,w_0^\inp,\psi^{\inp,n},
+\psi^{\inp,n+\myhalf}] \rightarrow [(\rho h)_0^\outp]$
+
+The base state enthalpy update now includes the area factors in the divergences:
+\begin{equation}
+(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp}
+ - \frac{1}{r_j^2} \frac{\dt}{\dr} 
+ \left \{ \left[ r^2 (\rho h)_0^{\outp} w_0^{\inp}\right]_{j+\myhalf} -  
+          \left[ r^2 (\rho h)_0^{\outp} w_0^{\inp}\right]_{j-\myhalf} \right\}
++\Delta t\psi^{\inp,n+\myhalf} \enskip. 
+\end{equation}
+The other difference is that the interface states are computed using 
+$\psi^{\inp,n}$ as the forcing.
 
@@
 to account for large-scale mixing, reactions, and thermal diffusion.
-
+Unlike paper III where we used $\nabla\cdot\etarho$ as an adjustment to $\rho_0$, 
+we instead simply set $\rho_0=\overline{\rho}$, which is analytically equivalent.
 \begin{description}
@@
 \end{eqnarray}
-Here $\beta_0$ is computed as described in paper~III, Appendix C.
+Here $\beta_0$ is computed as described in paper III, Appendix C.
 

### 2009-04-23 docs/flowchart/flowchart.tex
more progress on spherical base state

@@
 we should rely on a $\rho_0^{n+1}$ that also feels the effects.
+\item {\bf AMR!!!}
 \end{enumerate}
@@
 (\ref{eq:reacttemp}) is derived from the full temperature evolution equation
+from equation (8) in paper III,
 \begin{equation}
@@
 \end{equation}
-(paper III; eq.~[8])
 by assuming that the pressure is constant during the burn state.  In
@@
 
+\item {\bf Compute $w_0$}\\
+is the process by which we compute $w_0$.  The arguments are different for plane-parallel
+and spherical.
+\begin{description}
+\item {\bf Compute $w_0$ Planar}$[\Sbar^{\inp},\gammabar^{\inp},p_0^{\inp},\psi^{\inp}] 
+\rightarrow [w_0^{\outp}]$:
+
+We integrate
+\begin{equation}
+\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \psi
+\enskip \label{eq:divw0} ,
+\end{equation}
+with the following discretization:
+\begin{equation}
+\frac{\partial w_0^{\outp}}{\partial r} =  \Sbar^{\inp}
+- \frac{1}{\gammabar^{\inp} p_0^{\inp}}\psi^{\inp}
+\enskip .
+\end{equation}
+% VOLUME DISCREPANCY STUFF
+%\begin{equation}
+%\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S^{\nph,\star}}
+%- \frac{1}{\gammabar^{n} p_0^{n}}
+%\left\{ \psi^{\nmh} + 
+%f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right]\right\}
+%\enskip .
+%\end{equation}
+
+\item {\bf Compute $w_0$ Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},
+\etarho^{\inp}] \rightarrow[w_0^{\outp}]$:
+
+We begin with equation (23) in paper III written in spherical coordinates:
+\[
+\frac{1}{r^2}\frac{\partial}{\partial r} \left (r^2 \beta_0 w_0 \right ) =
+\beta_0 \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )\enskip.
+\]
+We expand the spatial derivative and divide through by $\beta_0$:
+\begin{equation}
+\label{eq:sphconstraint}
+\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) +
+w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r} =
+ \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t}  \enskip .
+\end{equation}
+Recalling from paper I that;
+\[
+\frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} = \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}\enskip,
+\]
+and moving this term to the right hand side,
+we can express equation (\ref{eq:sphconstraint}) as:
+\begin{eqnarray}
+%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) +
+%w_0 \frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} &=&
+% \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right ) \nonumber \\
+\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) &=&
+  \Sbar - \frac{1}{\gammabar p_0} \left( 
+\frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r} \right)\enskip. \label{eq:w01}
+\end{eqnarray}
+We first solve a piece of this equation for $\ow = w_0 - \dw$:
+\begin{eqnarray}
+\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \ow \right ) &=& \Sbar \enskip.
+\end{eqnarray}
+Then we can write
+\begin{eqnarray}
+\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \dw \right ) &=&
+  - \frac{1}{\gammabar p_0} \left(
+\frac{\partial p_0}{\partial t} + (\ow + \dw)
+\frac{\partial p_0}{\partial r} \right)\enskip.\label{eq:w01a}
+\end{eqnarray}
+
+Multiplying equation (\ref{eq:w01a}) through by $\gammabar p_0 $,
+taking another derivative with respect to $r$, and switching the order
+of temporal and spatial derivatives, we get:
+\begin{equation}
+\label{eq:sphconstraint_2}
+\frac{\partial}{\partial r} \left [ \frac{\gammabar p_0 }{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right ]
+= -\frac{\partial}{\partial t} \frac{\partial p_0}{\partial r}
+  -\frac{\partial}{\partial r} \left ((\ow+\dw) \frac{\partial p_0}{\partial r} \right )\enskip.
+\end{equation}
+
+To solve for $\dw$ we will need to substitute for the derivatives of $p_0.$
+To do so we start with the equation of hydrostatic equilibrium in a spherical geometry:
+\[
+\frac{\partial p_0}{\partial r} = -\rho_0 g \, ; \quad
+g = \frac{G m_\mathrm{encl}}{r^2}\enskip,
+\]
+where $m_\mathrm{encl}(r)$ is the mass enclosed at radius $r$ and $G$ is the
+gravitational constant.  
+Using this, we can then write equation (\ref{eq:sphconstraint_2}) as:
+\begin{eqnarray}
+\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=&
+ \frac{\partial}{\partial t} \left (\rho_0 g \right )
++\frac{\partial}{\partial r} \left (w_0 \rho_0 g \right) \nonumber \\
+&=&
+  g \left[ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right]
++ \rho_0 \left( \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} \right) \enskip.\nonumber\\
+&&
+\label{eq:sphconstraint_3}
+\end{eqnarray}
+The mass enclosed inside any radius, $r$, is $m_\mathrm{encl}(r) = 4 \pi
+\int_0^r \rhozero(s) s^2 ds$, or alternately, $\partial m_\mathrm{encl}/\partial r = 4\pi r^2 \rho_0$.  
+The Lagrangian derivative of the enclosed mass is then:
+\begin{eqnarray}
+\frac{D_0 m_\mathrm{encl}}{D t} &=& \frac{\partial m_\mathrm{encl}}{\partial t} + w_0 \frac{\partial m_\mathrm{encl}}{\partial r} \nonumber \\
+ &=& 4\pi \left( \frac{\partial}{\partial t} \int_0^r \rhozero(s) s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
+&=& 4\pi \left( \int_0^r \frac{\partial \rho_0}{\partial t} s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
+&=& 4\pi \left\{ -\int_0^r \left [ \frac{1}{s^2}\frac{\partial(s^2 \rho_0 w_0)}{\partial s} + \frac{1}{s^2}\frac{\partial (s^2 \etarho)}{\partial s} \right ]s^2 ds + w_0 r^2 \rho_0 \right\} \nonumber \\
+&=&  \left . 4\pi \left(-s^2 \rho_0 w_0 \right |_0^r - \left . s^2 \etarho \right |_0^r + w_0 r^2 \rho_0 \right) \nonumber \\
+&=& -4\pi r^2 \etarho \label{eq:dmdt}
+\end{eqnarray}
+where we used the spherical form of equation (29) in paper III, 
+\begin{equation}
+\label{eq:sph_continuity}
+\frac{\partial \rhozero}{\partial t} + \frac{1}{r^2} \frac{\partial (r^2 \rho_0 w_0)}{\partial r} 
++ \frac{1}{r^2} \frac{\partial (r^2 \etarho )}{\partial r}  = 0 \enskip ,
+\end{equation}
+to eliminate $\partial \rho_0/\partial t$.  We note that in the
+absence of any mixing, $\etarho=0$, and  $D_0 m_\mathrm{encl}/Dt = 0.$
+%which is another way of saying that the mass inside a fluid element
+%does not change during expansion.  
+Equation (\ref{eq:dmdt})
+allows us to write the Lagrangian change in the gravitational
+acceleration as:
+\begin{eqnarray}
+\frac{D_0 g}{D t} = \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} &=& \frac{D_0}{Dt}\left(\frac{G m_\mathrm{encl}}{r^2}\right) \nonumber \\
+&=& G m_\mathrm{encl}\frac{D_0}{D t}\left(\frac{1}{r^2}\right) + \frac{G}{r^2}\frac{D_0 m_\mathrm{encl}}{Dt} \nonumber \\
+&=& -\frac{2 w_0 G m_\mathrm{encl}}{r^3} - 4 \pi G \etarho \nonumber \\
+&=& -\frac{2 w_0 g}{r} - 4 \pi G \etarho \enskip .
+\end{eqnarray}
+
+Putting it all together, equation (\ref{eq:sphconstraint_3}) becomes:
+\begin{equation}
+ \frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] 
+= 
+  g \left [ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right ]
++ \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \label{eq:B5} \enskip.
+\end{equation}
+Finally, we can use equation (\ref{eq:sph_continuity})
+to write equation (\ref{eq:B5}) as:
+\begin{eqnarray}
+\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] && \nonumber \\
+&=&
+ g \left [ -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 w_0 \rhozero)
+           -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 \etarho)
+           +              \frac{\partial}{\partial r} (w_0 \rhozero) \right ] + \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \nonumber \\
+&=& 
+- \frac{g}{r^2} \frac{\partial (r^2 \etarho)}{\partial r} - \frac{4 (\ow + \dw) \rho_0 g}{r} 
+- 4 \pi G \rhozero \etarho \label{eq:B7} \enskip .
+\end{eqnarray}
+
+\noindent We discretize this elliptic equation in the radial dimension as:
+\begin{eqnarray}
+&& \frac{1}{\Delta r} \left\{
+\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i} -
+\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i-1} \right\}
++ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{i-\half} \nonumber\\
+&=& 
+- \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
+\left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
+   - \left[ \frac{4 \ow \rho_0 g}{r} \right]_{i-\half}
+   - \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip , \nonumber
+\end{eqnarray}
+where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
+Then, using hydrostatic equilibrium, we expand this to 
+\begin{eqnarray}
+&& \frac{1}{\Delta r} \left\{
+ \left( \frac{\gammabar p_0}{r^2}\right)_{i  } \frac{[ (r^2 \dw)_{i+\half} - (r^2 \dw)_{i-\half} ]}{\Delta r}
+-\left( \frac{\gammabar p_0}{r^2}\right)_{i-1} \frac{[ (r^2 \dw)_{i-\half} - (r^2 \dw)_{i-\thalf}]}{\Delta r} \right\} \nonumber \\
+&&  - \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \dw)_{i-\half}  \nonumber \\
+&=&   \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \ow)_{i-\half}
+ - \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
+  \left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
+- \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip.
+\end{eqnarray}
+
+If we write this in matrix form, so that:
+\begin{equation}
+A_i (r^2 \dw)_{i-\thalf} + B_i (r^2 \dw)_{i-\half} + C_i (r^2 \dw)_{i+\half} = F_i\enskip,
+\end{equation}
+then:
+\begin{eqnarray}
+A_i &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{i-1}  \enskip, \nonumber  \\
+B_i &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar p_0}{r^2}\right)_{i  }  
+               +\left( \frac{\gammabar p_0}{r^2}\right)_{i-1} \right] 
+              - \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} \nonumber \\
+C_i &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{i}  \enskip, \nonumber  \\
+F_i &=&  \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \ow)_{i-\half}
+ - \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
+\left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
+- \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip \nonumber .
+\end{eqnarray}
+Here,
+\begin{equation}
+g_{j-\half} = \frac{G}{r_{j-\half}^2} \sum_{k=1}^{j} \frac{4}{3} \pi  (r_{k-\half}^3 - r_{k-\thalf}^3) \; {\rho_0}_{k-1} \enskip.
+\end{equation}
+
+We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
+corresponds to $i=1,$ by setting:
+\begin{eqnarray}
+A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
+B_1 &=& 1\enskip. \nonumber
+\end{eqnarray}
+We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $i=N,$
+by setting:
+\begin{eqnarray}
+A_N &=& -1\enskip, \nonumber \\
+B_N &=&  1\enskip, \nonumber \\
+C_N = F_N &=& 0\enskip. \nonumber
+\end{eqnarray}
+\end{description}
+
 \item {\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp]$
@@
 \item[plane-parallel:] We discretize the base state density evolution 
-equation (paper III; eq.~[29]) without the $\etarho$ term,
+equation from paper III equation (29), without the $\etarho$ term,
 \begin{equation}
@@
 \end{equation}
-Then, we discretize the base state pressure evolution equation (paper III; eq.~[31]),
+Then, we discretize the base state pressure evolution equation from paper III equation (31),
 \begin{equation}
@@
 \item At the beginning of each time step, we need an estimate for the time-centered
-source term in the velocity divergence constraint,
+source term in the velocity divergence constraint, as given in paper III equation (19),
 \begin{equation}
@@
   \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext 
-  + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T \enskip ,
+  + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T \enskip .
 \label{eq:defineS} 
 \end{equation}
-(paper III; eq.~[19]).
 
@@
 
-\item Construct $w_0^{\nph,\star}$
-
-\begin{description}
-
-\item[plane-parallel:] We integrate
-\begin{equation}
-\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \etarho g 
-\enskip \label{eq:divw0} ,
-\end{equation}
-using the lagged $\psi^{\nmh}$ and the volume discrepancy correction with $f=0.3$,
-\begin{equation}
-\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S^{\nph,\star}}
-- \frac{1}{\gammabar^{n} p_0^{n}}
-\left\{ \psi^{\nmh} + 
-f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right]\right\}
-\enskip .
-\end{equation}
-
-\item[spherical:] See Section \ref{Sec:Spherical w0 Calculation} for working notes.
-
-\end{description}
+\item {\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,\psi^{n-\myhalf}]
+\rightarrow [w_0^{n+\myhalf,*}]$
 
@@
 
-The local velocity field is described by
+The local velocity field is described by paper III equation (37),
 \begin{equation}
@@
                                  - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip\
- ,
-\end{equation}
-(paper III; eq.~[37]).
+ .
+\end{equation}
 
 The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
-We note that  $\uadvone$ satisfies the discrete versions of 
+We note that  $\uadvone$ satisfies the discrete versions of paper III equation (21),
 \begin{equation}
@@
 \end{equation}
-(paper III; eq.~[21])
-and
-\begin{equation}
-\nablab \cdotb (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
+and paper III equation (39),
+\begin{equation}
+\nablab \cdotb (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip .
 \label{eq:tildeconstraint}
 \end{equation}
-(paper III; eq.~[39]).  The discretization with the volume discrepancy correction is
+The discretization with the volume discrepancy correction is
 \begin{equation}
@@
 \item Diffuse the enthalpy through a time interval of $\dt$.  Here we begin
-with the enthalpy equation (eq.~[\ref{eq:rhohupd}]), but consider only the 
+with the enthalpy equation (\ref{eq:rhohupd}), but consider only the 
 diffusion terms,
@@
 
-\item Construct $w_0^{\nph}$
-
-\begin{description}
-
-\item[plane-parallel:] We integrate equation (\ref{eq:divw0}) with the 
-volume discrepancy correction,
-\begin{equation}
-\frac{\partial w_0^{\nph}}{\partial r} = 
-\overline{S^{\nph,\star}} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
-\left\{ \psi^{\nph,\star} + f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\}
-\enskip ,
-\end{equation}
-\begin{equation}
-\gammabar^{\nph,\star} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}
-\end{equation}
-\begin{equation}
-p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}
-\end{equation}
-\begin{equation}
-\overline{p(\rho,h,X_k)^{\nph,\star}} = 
-\frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}
-\end{equation}
-
-\item[spherical:] See Section \ref{Sec:Spherical w0 Calculation} for working notes.
-
-\end{description}
+\item First, define:
+\begin{equation}
+\gammabar^{\nph,\star} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}\enskip ,
+\end{equation}
+\begin{equation}
+p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip .
+\end{equation}
+{\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},\gammabar^{\nph,\star},p_0^{\nph,\star},
+\psi^{\nph,\star}] \rightarrow [w_0^{\nph}]$.
+% VOLUME DISCREPANCY STUFF
+%\begin{equation}
+%\overline{p(\rho,h,X_k)^{\nph,\star}} = 
+%\frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}\enskip .
+%\end{equation}
+%\begin{equation}
+%\frac{\partial w_0^{\nph}}{\partial r} = 
+%\overline{S^{\nph}} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
+%\left\{ \psi^{\nph,\star} + f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\}
+%\enskip ,
 
@@
 \end{equation}
-where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined through
+where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined through paper III equation (53),
 \begin{equation}
@@
 \end{equation}
-(paper III; eq.~[53]) and
+and paper III equation (54),
 \begin{equation}
@@
 \end{equation}
-(paper III; eq.~[54]).
 
@@
 \item Update the species.  Here we do a difference approximation to
-  eq.~\ref{eq:species}, neglecting the reaction terms.  The
+  equation (\ref{eq:species}), neglecting the reaction terms.  The
   update consists of two steps:
@@
 \item Update the enthalpy, using a difference approximation to
-  eq.~\ref{eq:rhohupd}, neglecting the reaction terms.
+  equation (\ref{eq:rhohupd}), neglecting the reaction terms.
 
@@
 The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
-We update the velocity field $\Ubt^n$ to $\Ubt^{n+1,\star}$ by discretizing eq.~\ref{eq:utildeupd}
-as
+We update the velocity field $\Ubt^n$ to $\Ubt^{n+1,\star}$ by discretizing 
+equation (\ref{eq:utildeupd}) as
 \begin{eqnarray}

### 2009-04-23 docs/spherical_basestate/basestate.tex
more progress on spherical base state

@@
 self-gravitating star.
-
-\subsection{Calculation of $w_0$}\label{Sec:Spherical w0 Calculation}
-
-%In the low Mach number model, we allow the density and
-%temperature perturbations to be large compared to the background
-%state, but we require the pressure perturbation to be small.
-%Therefore, if the star expands, we need to incorporate that expansion
-%into the background state.  Furthermore, we need to know the velocity of the
-%base state expansion for the evolution of the perturbed
-%quantities.
-
-We begin with equation (23) in paper III written in spherical coordinates:
-\[
-\frac{1}{r^2}\frac{\partial}{\partial r} \left (r^2 \beta_0 w_0 \right ) =
-\beta_0 \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )\enskip.
-\]
-We expand the spatial derivative and divide through by $\beta_0$:
-\begin{equation}
-\label{eq:sphconstraint}
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) +
-w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r} =
- \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t}  \enskip .
-\end{equation}
-Recalling from paper I that;
-\[
-\frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} = \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}\enskip,
-\]
-and moving this term to the right hand side,
-we can express equation (\ref{eq:sphconstraint}) as:
-\begin{eqnarray}
-%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) +
-%w_0 \frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} &=&
-% \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right ) \nonumber \\
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) &=&
-  \Sbar - \frac{1}{\gammabar p_0} \left( 
-\frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r} \right)\enskip. \label{eq:w01}
-\end{eqnarray}
-We first solve a piece of this equation for $\ow = w_0 - \dw$:
-\begin{eqnarray}
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \ow \right ) &=& \Sbar \enskip.
-\end{eqnarray}
-Then we can write
-\begin{eqnarray}
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \dw \right ) &=&
-  - \frac{1}{\gammabar p_0} \left(
-\frac{\partial p_0}{\partial t} + (\ow + \dw)
-\frac{\partial p_0}{\partial r} \right)\enskip.\label{eq:w01a}
-\end{eqnarray}
-
-Multiplying equation (\ref{eq:w01a}) through by $\gammabar p_0 $,
-taking another derivative with respect to $r$, and switching the order
-of temporal and spatial derivatives, we get:
-\begin{equation}
-\label{eq:sphconstraint_2}
-\frac{\partial}{\partial r} \left [ \frac{\gammabar p_0 }{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right ]
-= -\frac{\partial}{\partial t} \frac{\partial p_0}{\partial r}
-  -\frac{\partial}{\partial r} \left ((\ow+\dw) \frac{\partial p_0}{\partial r} \right )\enskip.
-\end{equation}
-
-To solve for $\dw$ we will need to substitute for the derivatives of $p_0.$
-To do so we start with the equation of hydrostatic equilibrium in a spherical geometry:
-\[
-\frac{\partial p_0}{\partial r} = -\rho_0 g \, ; \quad
-g = \frac{G m_\mathrm{encl}}{r^2}\enskip,
-\]
-where $m_\mathrm{encl}(r)$ is the mass enclosed at radius $r$ and $G$ is the
-gravitational constant.  
-Using this, we can then write equation (\ref{eq:sphconstraint_2}) as:
-\begin{eqnarray}
-\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=&
- \frac{\partial}{\partial t} \left (\rho_0 g \right )
-+\frac{\partial}{\partial r} \left (w_0 \rho_0 g \right) \nonumber \\
-&=&
-  g \left[ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right]
-+ \rho_0 \left( \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} \right) \enskip.\nonumber\\
-&&
-\label{eq:sphconstraint_3}
-\end{eqnarray}
-The mass enclosed inside any radius, $r$, is $m_\mathrm{encl}(r) = 4 \pi
-\int_0^r \rhozero(s) s^2 ds$, or alternately, $\partial m_\mathrm{encl}/\partial r = 4\pi r^2 \rho_0$.  
-The Lagrangian derivative of the enclosed mass is then:
-\begin{eqnarray}
-\frac{D_0 m_\mathrm{encl}}{D t} &=& \frac{\partial m_\mathrm{encl}}{\partial t} + w_0 \frac{\partial m_\mathrm{encl}}{\partial r} \nonumber \\
- &=& 4\pi \left( \frac{\partial}{\partial t} \int_0^r \rhozero(s) s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
-&=& 4\pi \left( \int_0^r \frac{\partial \rho_0}{\partial t} s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
-&=& 4\pi \left\{ -\int_0^r \left [ \frac{1}{s^2}\frac{\partial(s^2 \rho_0 w_0)}{\partial s} + \frac{1}{s^2}\frac{\partial (s^2 \etarho)}{\partial s} \right ]s^2 ds + w_0 r^2 \rho_0 \right\} \nonumber \\
-&=&  \left . 4\pi \left(-s^2 \rho_0 w_0 \right |_0^r - \left . s^2 \etarho \right |_0^r + w_0 r^2 \rho_0 \right) \nonumber \\
-&=& -4\pi r^2 \etarho \label{eq:dmdt}
-\end{eqnarray}
-where we used the spherical form of equation (29) in paper III, 
-\begin{equation}
-\label{eq:sph_continuity}
-\frac{\partial \rhozero}{\partial t} + \frac{1}{r^2} \frac{\partial (r^2 \rho_0 w_0)}{\partial r} 
-+ \frac{1}{r^2} \frac{\partial (r^2 \etarho )}{\partial r}  = 0 \enskip ,
-\end{equation}
-to eliminate $\partial \rho_0/\partial t$.  We note that in the
-absence of any mixing, $\etarho=0$, and  $D_0 m_\mathrm{encl}/Dt = 0.$
-%which is another way of saying that the mass inside a fluid element
-%does not change during expansion.  
-Equation (\ref{eq:dmdt})
-allows us to write the Lagrangian change in the gravitational
-acceleration as:
-\begin{eqnarray}
-\frac{D_0 g}{D t} = \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} &=& \frac{D_0}{Dt}\left(\frac{G m_\mathrm{encl}}{r^2}\right) \nonumber \\
-&=& G m_\mathrm{encl}\frac{D_0}{D t}\left(\frac{1}{r^2}\right) + \frac{G}{r^2}\frac{D_0 m_\mathrm{encl}}{Dt} \nonumber \\
-&=& -\frac{2 w_0 G m_\mathrm{encl}}{r^3} - 4 \pi G \etarho \nonumber \\
-&=& -\frac{2 w_0 g}{r} - 4 \pi G \etarho \enskip .
-\end{eqnarray}
-
-Putting it all together, equation (\ref{eq:sphconstraint_3}) becomes:
-\begin{equation}
- \frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] 
-= 
-  g \left [ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right ]
-+ \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \label{eq:B5} \enskip.
-\end{equation}
-Finally, we can use equation (\ref{eq:sph_continuity})
-to write equation (\ref{eq:B5}) as:
-\begin{eqnarray}
-\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] && \nonumber \\
-&=&
- g \left [ -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 w_0 \rhozero)
-           -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 \etarho)
-           +              \frac{\partial}{\partial r} (w_0 \rhozero) \right ] + \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \nonumber \\
-&=& 
-- \frac{g}{r^2} \frac{\partial (r^2 \etarho)}{\partial r} - \frac{4 (\ow + \dw) \rho_0 g}{r} 
-- 4 \pi G \rhozero \etarho \label{eq:B7} \enskip .
-\end{eqnarray}
-
-\noindent We discretize this elliptic equation in the radial dimension as:
-\begin{eqnarray}
-&& \frac{1}{\Delta r} \left\{
-\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i} -
-\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i-1} \right\}
-+ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{i-\half} \nonumber\\
-&=& 
-- \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
-\left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
-   - \left[ \frac{4 \ow \rho_0 g}{r} \right]_{i-\half}
-   - \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip , \nonumber
-\end{eqnarray}
-where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
-Then, using hydrostatic equilibrium, we expand this to 
-\begin{eqnarray}
-&& \frac{1}{\Delta r} \left\{
- \left( \frac{\gammabar p_0}{r^2}\right)_{i  } \frac{[ (r^2 \dw)_{i+\half} - (r^2 \dw)_{i-\half} ]}{\Delta r}
--\left( \frac{\gammabar p_0}{r^2}\right)_{i-1} \frac{[ (r^2 \dw)_{i-\half} - (r^2 \dw)_{i-\thalf}]}{\Delta r} \right\} \nonumber \\
-&&  - \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \dw)_{i-\half}  \nonumber \\
-&=&   \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \ow)_{i-\half}
- - \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
-  \left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
-- \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip.
-\end{eqnarray}
-
-If we write this in matrix form, so that:
-\begin{equation}
-A_i (r^2 \dw)_{i-\thalf} + B_i (r^2 \dw)_{i-\half} + C_i (r^2 \dw)_{i+\half} = F_i\enskip,
-\end{equation}
-then:
-\begin{eqnarray}
-A_i &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{i-1}  \enskip, \nonumber  \\
-B_i &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar p_0}{r^2}\right)_{i  }  
-               +\left( \frac{\gammabar p_0}{r^2}\right)_{i-1} \right] 
-              - \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} \nonumber \\
-C_i &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{i}  \enskip, \nonumber  \\
-F_i &=&  \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \ow)_{i-\half}
- - \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
-\left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
-- \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip \nonumber .
-\end{eqnarray}
-Here,
-\begin{equation}
-g_{j-\half} = \frac{G}{r_{j-\half}^2} \sum_{k=1}^{j} \frac{4}{3} \pi  (r_{k-\half}^3 - r_{k-\thalf}^3) \; {\rho_0}_{k-1} \enskip.
-\end{equation}
-
-We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
-corresponds to $i=1,$ by setting:
-\begin{eqnarray}
-A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
-B_1 &=& 1\enskip. \nonumber
-\end{eqnarray}
-We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $i=N,$
-by setting:
-\begin{eqnarray}
-A_N &=& -1\enskip, \nonumber \\
-B_N &=&  1\enskip, \nonumber \\
-C_N = F_N &=& 0\enskip. \nonumber
-\end{eqnarray}
 

### 2009-04-23 docs/flowchart/flowchart.tex
made notation consisten with new w0 spherical stuff

@@
 
-\item {\bf Compute $w_0$}\\
-is the process by which we compute $w_0$.  The arguments are different for plane-parallel
-and spherical.
-\begin{description}
-\item {\bf Compute $w_0$ Planar}$[\Sbar^{\inp},\gammabar^{\inp},p_0^{\inp},\psi^{\inp}] 
-\rightarrow [w_0^{\outp}]$:
-
-We integrate
-\begin{equation}
-\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \psi
-\enskip \label{eq:divw0} ,
-\end{equation}
-with the following discretization:
-\begin{equation}
-\frac{\partial w_0^{\outp}}{\partial r} =  \Sbar^{\inp}
-- \frac{1}{\gammabar^{\inp} p_0^{\inp}}\psi^{\inp}
-\enskip .
-\end{equation}
-% VOLUME DISCREPANCY STUFF
-%\begin{equation}
-%\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S^{\nph,\star}}
-%- \frac{1}{\gammabar^{n} p_0^{n}}
-%\left\{ \psi^{\nmh} + 
-%f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right]\right\}
-%\enskip .
-%\end{equation}
-
-\item {\bf Compute $w_0$ Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},
-\etarho^{\inp}] \rightarrow[w_0^{\outp}]$:
-
-We begin with equation (23) in paper III written in spherical coordinates:
-\[
-\frac{1}{r^2}\frac{\partial}{\partial r} \left (r^2 \beta_0 w_0 \right ) =
-\beta_0 \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )\enskip.
-\]
-We expand the spatial derivative and divide through by $\beta_0$:
-\begin{equation}
-\label{eq:sphconstraint}
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) +
-w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r} =
- \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t}  \enskip .
-\end{equation}
-Recalling from paper I that;
-\[
-\frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} = \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}\enskip,
-\]
-and moving this term to the right hand side,
-we can express equation (\ref{eq:sphconstraint}) as:
-\begin{eqnarray}
-%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) +
-%w_0 \frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} &=&
-% \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right ) \nonumber \\
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) &=&
-  \Sbar - \frac{1}{\gammabar p_0} \left( 
-\frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r} \right)\enskip. \label{eq:w01}
-\end{eqnarray}
-We first solve a piece of this equation for $\ow = w_0 - \dw$:
-\begin{eqnarray}
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \ow \right ) &=& \Sbar \enskip.
-\end{eqnarray}
-Then we can write
-\begin{eqnarray}
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \dw \right ) &=&
-  - \frac{1}{\gammabar p_0} \left(
-\frac{\partial p_0}{\partial t} + (\ow + \dw)
-\frac{\partial p_0}{\partial r} \right)\enskip.\label{eq:w01a}
-\end{eqnarray}
-
-Multiplying equation (\ref{eq:w01a}) through by $\gammabar p_0 $,
-taking another derivative with respect to $r$, and switching the order
-of temporal and spatial derivatives, we get:
-\begin{equation}
-\label{eq:sphconstraint_2}
-\frac{\partial}{\partial r} \left [ \frac{\gammabar p_0 }{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right ]
-= -\frac{\partial}{\partial t} \frac{\partial p_0}{\partial r}
-  -\frac{\partial}{\partial r} \left ((\ow+\dw) \frac{\partial p_0}{\partial r} \right )\enskip.
-\end{equation}
-
-To solve for $\dw$ we will need to substitute for the derivatives of $p_0.$
-To do so we start with the equation of hydrostatic equilibrium in a spherical geometry:
-\[
-\frac{\partial p_0}{\partial r} = -\rho_0 g \, ; \quad
-g = \frac{G m_\mathrm{encl}}{r^2}\enskip,
-\]
-where $m_\mathrm{encl}(r)$ is the mass enclosed at radius $r$ and $G$ is the
-gravitational constant.  
-Using this, we can then write equation (\ref{eq:sphconstraint_2}) as:
-\begin{eqnarray}
-\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=&
- \frac{\partial}{\partial t} \left (\rho_0 g \right )
-+\frac{\partial}{\partial r} \left (w_0 \rho_0 g \right) \nonumber \\
-&=&
-  g \left[ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right]
-+ \rho_0 \left( \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} \right) \enskip.\nonumber\\
-&&
-\label{eq:sphconstraint_3}
-\end{eqnarray}
-The mass enclosed inside any radius, $r$, is $m_\mathrm{encl}(r) = 4 \pi
-\int_0^r \rhozero(s) s^2 ds$, or alternately, $\partial m_\mathrm{encl}/\partial r = 4\pi r^2 \rho_0$.  
-The Lagrangian derivative of the enclosed mass is then:
-\begin{eqnarray}
-\frac{D_0 m_\mathrm{encl}}{D t} &=& \frac{\partial m_\mathrm{encl}}{\partial t} + w_0 \frac{\partial m_\mathrm{encl}}{\partial r} \nonumber \\
- &=& 4\pi \left( \frac{\partial}{\partial t} \int_0^r \rhozero(s) s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
-&=& 4\pi \left( \int_0^r \frac{\partial \rho_0}{\partial t} s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
-&=& 4\pi \left\{ -\int_0^r \left [ \frac{1}{s^2}\frac{\partial(s^2 \rho_0 w_0)}{\partial s} + \frac{1}{s^2}\frac{\partial (s^2 \etarho)}{\partial s} \right ]s^2 ds + w_0 r^2 \rho_0 \right\} \nonumber \\
-&=&  \left . 4\pi \left(-s^2 \rho_0 w_0 \right |_0^r - \left . s^2 \etarho \right |_0^r + w_0 r^2 \rho_0 \right) \nonumber \\
-&=& -4\pi r^2 \etarho \label{eq:dmdt}
-\end{eqnarray}
-where we used the spherical form of equation (29) in paper III, 
-\begin{equation}
-\label{eq:sph_continuity}
-\frac{\partial \rhozero}{\partial t} + \frac{1}{r^2} \frac{\partial (r^2 \rho_0 w_0)}{\partial r} 
-+ \frac{1}{r^2} \frac{\partial (r^2 \etarho )}{\partial r}  = 0 \enskip ,
-\end{equation}
-to eliminate $\partial \rho_0/\partial t$.  We note that in the
-absence of any mixing, $\etarho=0$, and  $D_0 m_\mathrm{encl}/Dt = 0.$
-%which is another way of saying that the mass inside a fluid element
-%does not change during expansion.  
-Equation (\ref{eq:dmdt})
-allows us to write the Lagrangian change in the gravitational
-acceleration as:
-\begin{eqnarray}
-\frac{D_0 g}{D t} = \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} &=& \frac{D_0}{Dt}\left(\frac{G m_\mathrm{encl}}{r^2}\right) \nonumber \\
-&=& G m_\mathrm{encl}\frac{D_0}{D t}\left(\frac{1}{r^2}\right) + \frac{G}{r^2}\frac{D_0 m_\mathrm{encl}}{Dt} \nonumber \\
-&=& -\frac{2 w_0 G m_\mathrm{encl}}{r^3} - 4 \pi G \etarho \nonumber \\
-&=& -\frac{2 w_0 g}{r} - 4 \pi G \etarho \enskip .
-\end{eqnarray}
-
-Putting it all together, equation (\ref{eq:sphconstraint_3}) becomes:
-\begin{equation}
- \frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] 
-= 
-  g \left [ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right ]
-+ \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \label{eq:B5} \enskip.
-\end{equation}
-Finally, we can use equation (\ref{eq:sph_continuity})
-to write equation (\ref{eq:B5}) as:
-\begin{eqnarray}
-\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] && \nonumber \\
-&=&
- g \left [ -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 w_0 \rhozero)
-           -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 \etarho)
-           +              \frac{\partial}{\partial r} (w_0 \rhozero) \right ] + \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \nonumber \\
-&=& 
-- \frac{g}{r^2} \frac{\partial (r^2 \etarho)}{\partial r} - \frac{4 (\ow + \dw) \rho_0 g}{r} 
-- 4 \pi G \rhozero \etarho \label{eq:B7} \enskip .
-\end{eqnarray}
-
-\noindent We discretize this elliptic equation in the radial dimension as:
-\begin{eqnarray}
-&& \frac{1}{\Delta r} \left\{
-\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i} -
-\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i-1} \right\}
-+ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{i-\half} \nonumber\\
-&=& 
-- \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
-\left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
-   - \left[ \frac{4 \ow \rho_0 g}{r} \right]_{i-\half}
-   - \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip , \nonumber
-\end{eqnarray}
-where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
-Then, using hydrostatic equilibrium, we expand this to 
-\begin{eqnarray}
-&& \frac{1}{\Delta r} \left\{
- \left( \frac{\gammabar p_0}{r^2}\right)_{i  } \frac{[ (r^2 \dw)_{i+\half} - (r^2 \dw)_{i-\half} ]}{\Delta r}
--\left( \frac{\gammabar p_0}{r^2}\right)_{i-1} \frac{[ (r^2 \dw)_{i-\half} - (r^2 \dw)_{i-\thalf}]}{\Delta r} \right\} \nonumber \\
-&&  - \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \dw)_{i-\half}  \nonumber \\
-&=&   \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \ow)_{i-\half}
- - \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
-  \left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
-- \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip.
-\end{eqnarray}
-
-If we write this in matrix form, so that:
-\begin{equation}
-A_i (r^2 \dw)_{i-\thalf} + B_i (r^2 \dw)_{i-\half} + C_i (r^2 \dw)_{i+\half} = F_i\enskip,
-\end{equation}
-then:
-\begin{eqnarray}
-A_i &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{i-1}  \enskip, \nonumber  \\
-B_i &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar p_0}{r^2}\right)_{i  }  
-               +\left( \frac{\gammabar p_0}{r^2}\right)_{i-1} \right] 
-              - \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} \nonumber \\
-C_i &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{i}  \enskip, \nonumber  \\
-F_i &=&  \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \ow)_{i-\half}
- - \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
-\left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
-- \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip \nonumber .
-\end{eqnarray}
-Here,
-\begin{equation}
-g_{j-\half} = \frac{G}{r_{j-\half}^2} \sum_{k=1}^{j} \frac{4}{3} \pi  (r_{k-\half}^3 - r_{k-\thalf}^3) \; {\rho_0}_{k-1} \enskip.
-\end{equation}
-
-We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
-corresponds to $i=1,$ by setting:
-\begin{eqnarray}
-A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
-B_1 &=& 1\enskip. \nonumber
-\end{eqnarray}
-We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $i=N,$
-by setting:
-\begin{eqnarray}
-A_N &=& -1\enskip, \nonumber \\
-B_N &=&  1\enskip, \nonumber \\
-C_N = F_N &=& 0\enskip. \nonumber
-\end{eqnarray}
-\end{description}
-
 \item {\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp]$
@@
 \end{equation}
-
 \end{description}
-
 \item {\bf Advect Base Pressure}
@@
 
+\item {\bf Compute $w_0$}\\
+is the process by which we compute $w_0$.  The arguments are different for plane-parallel
+and spherical.
+\begin{description}
+\item {\bf Compute $w_0$ Planar}$[\Sbar^{\inp},\gammabar^{\inp},p_0^{\inp},\psi^{\inp}] 
+\rightarrow [w_0^{\outp}]$:
+
+We integrate
+\begin{equation}
+\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \psi
+\enskip \label{eq:divw0} ,
+\end{equation}
+with the following discretization:
+\begin{equation}
+\frac{\partial w_0^{\outp}}{\partial r} =  \Sbar^{\inp}
+- \frac{1}{\gammabar^{\inp} p_0^{\inp}}\psi^{\inp}
+\enskip .
+\end{equation}
+% VOLUME DISCREPANCY STUFF
+%\begin{equation}
+%\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S^{\nph,\star}}
+%- \frac{1}{\gammabar^{n} p_0^{n}}
+%\left\{ \psi^{\nmh} + 
+%f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right]\right\}
+%\enskip .
+%\end{equation}
+
+\item {\bf Compute $w_0$ Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},
+\etarho^{\inp}] \rightarrow[w_0^{\outp}]$:
+
+We begin with equation (23) in paper III written in spherical coordinates:
+\[
+\frac{1}{r^2}\frac{\partial}{\partial r} \left (r^2 \beta_0 w_0 \right ) =
+\beta_0 \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )\enskip.
+\]
+We expand the spatial derivative and divide through by $\beta_0$:
+\begin{equation}
+\label{eq:sphconstraint}
+\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) +
+w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r} =
+ \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t}  \enskip .
+\end{equation}
+Recalling from paper I that;
+\[
+\frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} = \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}\enskip,
+\]
+and moving this term to the right hand side,
+we can express equation (\ref{eq:sphconstraint}) as:
+\begin{eqnarray}
+\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) &=&
+  \Sbar - \frac{1}{\gammabar p_0} \left( 
+\frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r} \right)\enskip. \label{eq:w01}
+\end{eqnarray}
+We first solve a piece of this equation for $\ow = w_0 - \dw$:
+\begin{eqnarray}
+\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \ow \right ) &=& \Sbar \enskip.
+\end{eqnarray}
+Then we can write
+\begin{eqnarray}
+\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \dw \right ) &=&
+  - \frac{1}{\gammabar p_0} \left[
+\frac{\partial p_0}{\partial t} + (\ow + \dw)
+\frac{\partial p_0}{\partial r} \right]\enskip.\label{eq:w01a}
+\end{eqnarray}
+
+Multiplying equation (\ref{eq:w01a}) through by $\gammabar p_0 $,
+taking another derivative with respect to $r$, and switching the order
+of temporal and spatial derivatives, we get:
+\begin{equation}
+\label{eq:sphconstraint_2}
+\frac{\partial}{\partial r} \left [ \frac{\gammabar p_0 }{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right ]
+= -\frac{\partial}{\partial t} \frac{\partial p_0}{\partial r}
+  -\frac{\partial}{\partial r} \left [(\ow+\dw) \frac{\partial p_0}{\partial r} \right ]\enskip.
+\end{equation}
+
+To solve for $\dw$ we will need to substitute for the derivatives of $p_0.$
+To do so we start with the equation of hydrostatic equilibrium in a spherical geometry:
+\[
+\frac{\partial p_0}{\partial r} = -\rho_0 g \, ; \quad
+g = \frac{G m_\mathrm{encl}}{r^2}\enskip,
+\]
+where $m_\mathrm{encl}(r)$ is the mass enclosed at radius $r$ and $G$ is the
+gravitational constant.  
+Using this, we can then write equation (\ref{eq:sphconstraint_2}) as:
+\begin{eqnarray}
+\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=&
+ \frac{\partial}{\partial t} \left (\rho_0 g \right )
++\frac{\partial}{\partial r} \left (w_0 \rho_0 g \right) \nonumber \\
+&=&
+  g \left[ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right]
++ \rho_0 \left( \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} \right) \enskip.\nonumber\\
+&&
+\label{eq:sphconstraint_3}
+\end{eqnarray}
+The mass enclosed inside any radius, $r$, is $m_\mathrm{encl}(r) = 4 \pi
+\int_0^r \rhozero(s) s^2 ds$, or alternately, $\partial m_\mathrm{encl}/\partial r = 4\pi r^2 \rho_0$.  
+The Lagrangian derivative of the enclosed mass is then:
+\begin{eqnarray}
+\frac{D_0 m_\mathrm{encl}}{D t} &=& \frac{\partial m_\mathrm{encl}}{\partial t} + w_0 \frac{\partial m_\mathrm{encl}}{\partial r} \nonumber \\
+ &=& 4\pi \left( \frac{\partial}{\partial t} \int_0^r \rhozero(s) s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
+&=& 4\pi \left( \int_0^r \frac{\partial \rho_0}{\partial t} s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
+&=& 4\pi \left\{ -\int_0^r \left [ \frac{1}{s^2}\frac{\partial(s^2 \rho_0 w_0)}{\partial s} + \frac{1}{s^2}\frac{\partial (s^2 \etarho)}{\partial s} \right ]s^2 ds + w_0 r^2 \rho_0 \right\} \nonumber \\
+&=&  \left . 4\pi \left(-s^2 \rho_0 w_0 \right |_0^r - \left . s^2 \etarho \right |_0^r + w_0 r^2 \rho_0 \right) \nonumber \\
+&=& -4\pi r^2 \etarho \label{eq:dmdt}
+\end{eqnarray}
+where we used the spherical form of equation (29) in paper III, 
+\begin{equation}
+\label{eq:sph_continuity}
+\frac{\partial \rhozero}{\partial t} + \frac{1}{r^2} \frac{\partial (r^2 \rho_0 w_0)}{\partial r} 
++ \frac{1}{r^2} \frac{\partial (r^2 \etarho )}{\partial r}  = 0 \enskip ,
+\end{equation}
+to eliminate $\partial \rho_0/\partial t$.  We note that in the
+absence of any mixing, $\etarho=0$, and  $D_0 m_\mathrm{encl}/Dt = 0.$
+Equation (\ref{eq:dmdt})
+allows us to write the Lagrangian change in the gravitational
+acceleration as:
+\begin{eqnarray}
+\frac{D_0 g}{D t} = \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} &=& \frac{D_0}{Dt}\left(\frac{G m_\mathrm{encl}}{r^2}\right) \nonumber \\
+&=& G m_\mathrm{encl}\frac{D_0}{D t}\left(\frac{1}{r^2}\right) + \frac{G}{r^2}\frac{D_0 m_\mathrm{encl}}{Dt} \nonumber \\
+&=& -\frac{2 w_0 G m_\mathrm{encl}}{r^3} - 4 \pi G \etarho \nonumber \\
+&=& -\frac{2 w_0 g}{r} - 4 \pi G \etarho \enskip .
+\end{eqnarray}
+
+Putting it all together, equation (\ref{eq:sphconstraint_3}) becomes:
+\begin{equation}
+ \frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] 
+= 
+  g \left [ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right ]
++ \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \label{eq:B5} \enskip.
+\end{equation}
+Finally, we can use equation (\ref{eq:sph_continuity})
+to write equation (\ref{eq:B5}) as:
+\begin{eqnarray}
+\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=&
+ g \left [ -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 w_0 \rhozero)
+           -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 \etarho)
+           + \frac{\partial}{\partial r} (w_0 \rhozero) \right ] \nonumber \\
+&& + \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \nonumber \\
+&=& - \frac{g}{r^2} \frac{\partial (r^2 \etarho)}{\partial r} 
+- \frac{4 (\ow + \dw) \rho_0 g}{r} - 4 \pi G \rhozero \etarho \label{eq:B7} \enskip .
+\end{eqnarray}
+
+\noindent We discretize this elliptic equation in the radial dimension as:
+\begin{eqnarray}
+&& \frac{1}{\Delta r} \left\{
+\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{j} -
+\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{j-1} \right\}
++ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{j-\myhalf} \nonumber\\
+&=& 
+- \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
+\left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
+   - \left( \frac{4 \ow \rho_0 g}{r} \right)_{j-\myhalf}
+   - \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip , \nonumber
+\end{eqnarray}
+where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
+Then, using hydrostatic equilibrium, we expand this to 
+\begin{eqnarray}
+&& \frac{1}{\Delta r} \left\{
+ \left( \frac{\gammabar p_0}{r^2}\right)_{j  } \frac{\left[ (r^2 \dw)_{j+\myhalf} - (r^2 \dw)_{j-\myhalf} \right]}{\Delta r}
+-\left( \frac{\gammabar p_0}{r^2}\right)_{j-1} \frac{\left[ (r^2 \dw)_{j-\myhalf} - (r^2 \dw)_{j-\thalf}\right]}{\Delta r} \right\} \nonumber \\
+&&  - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,i} - p_{0,i-1}}{\dr} \right) (r^2 \dw)_{j-\myhalf}  \nonumber \\
+&=&   \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,i} - p_{0,i-1}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
+ - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
+  \left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
+- \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip. \nonumber \\
+\end{eqnarray}
+
+If we write this in matrix form, so that:
+\begin{equation}
+A_j (r^2 \dw)_{j-\thalf} + B_j (r^2 \dw)_{j-\myhalf} + C_j (r^2 \dw)_{j+\myhalf} = F_j\enskip,
+\end{equation}
+then:
+\begin{eqnarray}
+A_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{j-1}  \enskip, \nonumber  \\
+B_j &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar p_0}{r^2}\right)_{j  }  
+               +\left( \frac{\gammabar p_0}{r^2}\right)_{j-1} \right] 
+              - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,i} - p_{0,i-1}}{\dr} \right) \nonumber \\
+C_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{j}  \enskip, \nonumber  \\
+F_j &=&  \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,i} - p_{0,i-1}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
+ - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
+\left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
+- \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip \nonumber .
+\end{eqnarray}
+Here,
+\begin{equation}
+g_{j-\myhalf} = \frac{G}{r_{j-\myhalf}^2} \sum_{k=1}^{j} \frac{4}{3} \pi  (r_{k-\myhalf}^3 - r_{k-\thalf}^3) \; {\rho_0}_{k-1} \enskip.
+\end{equation}
+
+We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
+corresponds to $j=1,$ by setting:
+\begin{eqnarray}
+A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
+B_1 &=& 1\enskip. \nonumber
+\end{eqnarray}
+We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $j=N,$
+by setting:
+\begin{eqnarray}
+A_N &=& -1\enskip, \nonumber \\
+B_N &=&  1\enskip, \nonumber \\
+C_N = F_N &=& 0\enskip. \nonumber
+\end{eqnarray}
+\end{description}
+
 \end{enumerate}

### 2009-04-23 docs/radial_base/radial_base.tex
sync up with paper IV, add notes about future text needed

--- 2009-01-21 docs/radial_base/radial_base.tex
@@
 Since the one-dimensional spherical base state is not aligned with the
-3-d Cartesian grid we use for the full star, computing quantities that
+3-d Cartesian grid we use for full star spherical geometry, computing quantities that
 involve both the base state and the full state becomes complicated.
-Therefore, when we go between the base state and the full state in
-various parts of the algorithm (such as the averaging operations), we
-need to map between the two geometries.
+Therefore, various parts of the algorithm (such as the averaging operations)
+require a mapping between the base state and the full state.
 
@@
 independently.  Experimentation has shown that the best choice results
-from picking the 3-d Cartesian grid spacing, $\Delta x$ to be coarser
-than the 1-d spherical base state spacing, $\Delta r$.
-A typical simulation may use $5 \Delta r = \Delta x$.
-We call the procedure to go from 1-d to 3-d {\tt fill\_3d} and the
-complementary procedure that takes us from 3-d to 1-d {\tt average}.
+from picking the 3-d Cartesian grid spacing, $\Delta x,$ to be coarser
+than the 1-d spherical base state spacing, $\Delta r$.  (Here we
+assume $\Delta x = \Delta y = \Delta z.$)  For the paper IV simulations,
+we use $5 \Delta r = \Delta x$.  We refer to the procedure that maps data 
+from 1-d to 3-d as {\tt fill\_3d}, and the
+complementary procedure that maps from 3-d to 1-d as {\tt average}.
 \section{\tt fill\_3d}
-There are three different {\tt fill\_3d} modes:
+There are three different {\tt fill\_3d} modes, each performed with the
+subroutine, {\tt put\_1d\_array\_on\_cart}:
 \begin{enumerate}
-\item The 1D array is bin-centered, and we map to Cartesian cell-centers.
+\item The 1D array is cell-centered, and we map to Cartesian cell-centers.  
+This is the most commonly used mode of {\tt fill\_3d}, and in paper IV, 
+we refer to this mode as {\tt fill\_3d}.
 \item The 1D array is edge-centered, and we map to Cartesian cell-centers.
@@
 \label{Sec:1D Cell-Centered to Cartesian Cell-Centered}
-This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
-Figure \ref{fig:mapping} shows the Cartesian grid with the spherical
-base state overlayed (for simplicity, the figure is drawn in 2-d).  
-We compute the radius of the Cartesian cell center from the center of the star,
+
+Figure  \ref{fig:mapping} shows the Cartesian grid overlaid by the 
+spherical base state (for simplicity, the figure is drawn in 2-d 
+using $2 \Delta r = \Delta x$).
+The {\tt fill\_3d} procedure computes the distance of the center of
+cell indexed by $(i,j,k)$ from the center of the star,
 \begin{equation}
-r = \sqrt{(x_i - x_c)^2 + (y_j - y_c)^2 + (z_k - z_c)^2}
+r = \sqrt{(x_i - x_c)^2 + (y_j - y_c)^2 + (z_k - z_c)^2},
 \end{equation}
@@
 base state).  We can then initialize a Cartesian cell quantity $q$ from its
-corresponding base state quantity, $q_0$ as $q_{i,j,k} = q_{0,n}$.  In essence,
-we are doing a piecewise constant interpolation of the radial array onto the 
-Cartesian grid.  If the output is a vector quantity, rather than a scalar, we
-multiply the result by the unit vector in the normal direction.
+corresponding base state quantity, $q_0,$ as $q_{i,j,k} = q_{0,n}$.
+
+In essence, we are doing a piecewise constant interpolation of the radial array 
+onto the Cartesian grid.  If the output is a vector quantity, rather than a 
+scalar, we multiply the result by the unit vector in the normal direction.
 \subsection{1D Edge-Centered to Cartesian Cell-Centered}
 \label{Sec:1D Edge-Centered to Cartesian Cell-Centered}
-This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
 Again, for each Cartesian cell-center, we compute the corresponding radial
@@
 \label{Sec:1D Edge-Centered to Cartesian Edges}
-This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
 This case is only used to create {\tt w0mac} from {\tt w0} in the 
@@
 velocities from the Cartesian cell-centers adjacent to the face.
-\section{Avg}\label{Sec:Avg}
-For the {\bf Avg} process, we first create a coarse 1D radial array 
-with $\Delta x = \Delta r$.  Then, we compute the radius of the Cartesian cell 
-center, and compute the index, $n$, of the corresponding coarse radial 
-bin.  $q_{0,n}$ is then the average of all the $q_{i,j,k}$ whose 
-Cartesian cell center maps into coarse radial bin $n$.  Then, we 
-interpolate the coarse radial array onto the finer radial array.
+\section{1D Cell-Centered to Cartesian Edges}
+This is not part of the {\tt put\_1d\_array\_on\_cart} options, 
+but at various points in the algorithm we need base state quantities on 
+edges.  For example, we need $\rho_0$ at edges to compute the full 
+state $\rho$ used to compute fluxes.  More details to come.
+\MarginPar{This section needs to be written.}
+\section{{\tt average}}\label{Sec:Avg}
+For the {\tt average} process, we first define a coarse 1-d radial array 
+with $\Delta r_{\rm c} = \Delta x$.  Then, for each cell indexed by
+$(i,j,k)$ we again compute the radius, $r,$ as above, and 
+define the index of the corresponding coarse radial bin, 
+$n_{\rm c} = \mathtt{int}(r / \Delta r_{\rm c})$.  
+We define $q_{0,n_{\rm c}}$ as the average of all the $q_{i,j,k}$ whose 
+Cartesian cell centers map into coarse radial bin $n_{\rm c}$.
+Next, we construct edge-centered states on the coarse radial bin using 
+the fourth order approximation,
+$q_{0,n_{\rm c}+\myhalf} = (7/12)(q_{0,n_{\rm c}}+q_{0,n_{\rm c}+1}) 
+- (1/12)(q_{0,n_{\rm c}-1}+q_{0,n_{\rm c}+2})$.  Finally, for each coarse radial bin
+we construct a quadratic profile using $q_{0,n_{\rm c}-\myhalf}, q_{0,n_{\rm c}}$ 
+and $q_{0,n_{\rm c}+\myhalf}$.  This is based on the interpolating polynomial used by 
+the PPM scheme to find edge states (Colella \& Woodward 1984).  Specifically, for 
+$n_{\rm c}\Delta r_{\rm c} \le r \le (n_{\rm c}+1)\Delta r_{\rm c}$, the 
+interpolating polynomial is
+\MarginPar{Need to add description of limiters}
+\begin{equation}
+q_0(r) = q_{0,n_{\rm c}-\myhalf} + \xi(r)\left\{\Delta q_{n_{\rm c}} + q_{6,n_{\rm c}}[1 - \xi(r)]\right\},
+\end{equation}
+with
+\begin{equation}
+\xi(r) = \frac{r - n_{\rm c}\Delta r_{\rm c}}{\Delta r_{\rm c}},
+\end{equation}
+\begin{equation}
+\Delta q_{n_{\rm c}} = q_{0,n_{\rm c}+\myhalf} - q_{0,n_{\rm c}-\myhalf},
+\end{equation}
+and
+\begin{equation}
+q_{6,n_{\rm c}} = 6\left[q_{0,n_{\rm c}} - 
+\frac{1}{2}\left(q_{0,n_{\rm c}+\myhalf} + q_{0,n_{\rm c}-\myhalf}\right)\right].
+\end{equation}
 
@@
 \caption{\label{fig:mapping} The Cartesian grid and spherical base
-state (shown here in 2-d for simplicity).  Here we represent the
+state (shown here in 2-d for simplicity using $2 \Delta r = \Delta x$).
+Here we represent the
 spherical base state as concentric shells (blue lines).  Since the

### 2009-04-23 docs/radial_base/radial_base.tex
updated fill_3d and other related items

@@
 \section{\tt fill\_3d}
-There are three different {\tt fill\_3d} modes, each performed with the
+There are two different {\tt fill\_3d} modes, each performed with the
 subroutine, {\tt put\_1d\_array\_on\_cart}:
@@
 \item The 1D array is edge-centered, and we map to Cartesian cell-centers.
-\item The 1D array is edge-centered, and we map to Cartesian edges.  
 \end{enumerate}
@@
 multiply the result by the unit vector in the normal direction.
-\subsection{1D Edge-Centered to Cartesian Edges}
+We use this function for:
+\begin{enumerate}
+\item Putting $\nabla p_0$ from radial edges onto cell centers as a vector.
+\item Computing radial velocity for plotfile.
+\end{enumerate}
+\section{1D Edge-Centered to Cartesian Edges}
 \label{Sec:1D Edge-Centered to Cartesian Edges}
@@
 \section{1D Cell-Centered to Cartesian Edges}
-This is not part of the {\tt put\_1d\_array\_on\_cart} options, 
-but at various points in the algorithm we need base state quantities on 
-edges.  For example, we need $\rho_0$ at edges to compute the full 
-state $\rho$ used to compute fluxes.  More details to come.
-\MarginPar{This section needs to be written.}
+At various points in the algorithm we need to map cell-centered radial 
+quantities onto Cartesian edges.  For example, we need $\rho_0$ at edges 
+to compute the full state $\rho$ used to compute fluxes.  What we do
+is call {\tt fill\_3d}, and then average neighboring cell-centers onto
+edges.
 \section{{\tt average}}\label{Sec:Avg}

### 2009-04-23 docs/flowchart/flowchart.tex
done hooking in the w0 stuff for spherical

@@
 - \frac{1}{\gammabar^{\inp} p_0^{\inp}}\psi^{\inp}
-\enskip .
-\end{equation}
+\enskip ,
+\end{equation}
+and $w_0=0$ at $r=0$.
 % VOLUME DISCREPANCY STUFF
@@
 \begin{eqnarray}
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \ow \right ) &=& \Sbar \enskip.
+\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \ow \right ) &=& \Sbar^{\inp} \enskip.
 \end{eqnarray}
@@
 -\left( \frac{\gammabar p_0}{r^2}\right)_{j-1} \frac{\left[ (r^2 \dw)_{j-\myhalf} - (r^2 \dw)_{j-\thalf}\right]}{\Delta r} \right\} \nonumber \\
-&&  - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,i} - p_{0,i-1}}{\dr} \right) (r^2 \dw)_{j-\myhalf}  \nonumber \\
-&=&   \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,i} - p_{0,i-1}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
+&&  - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \dw)_{j-\myhalf}  \nonumber \\
+&=&   \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
  - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
   \left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
-- \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip. \nonumber \\
-\end{eqnarray}
+- \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip, \nonumber \\
+\end{eqnarray}
+where $\etarho$ at cell-centers is the average of the neighboring edge-values.
 
@@
 \begin{eqnarray}
-A_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{j-1}  \enskip, \nonumber  \\
-B_j &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar p_0}{r^2}\right)_{j  }  
-               +\left( \frac{\gammabar p_0}{r^2}\right)_{j-1} \right] 
-              - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,i} - p_{0,i-1}}{\dr} \right) \nonumber \\
-C_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar p_0}{r^2}\right)_{j}  \enskip, \nonumber  \\
-F_j &=&  \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,i} - p_{0,i-1}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
+A_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1}  \enskip, \nonumber  \\
+B_j &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j  }  
+               +\left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1} \right] 
+              - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) \nonumber \\
+C_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j}  \enskip, \nonumber  \\
+F_j &=&  \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
  - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
-\left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
-- \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip \nonumber .
+\left[ \left( r^2 \etarho^{\inp} \right)_{j} - \left( r^2 \etarho^{\inp}  \right)_{j-1} \right] 
+- \left(  4 \pi G \rhozero^{\inp} \etarho^{\inp}   \right)_{j-\myhalf} \enskip \nonumber .
 \end{eqnarray}
@@
 \begin{equation}
-g_{j-\myhalf} = \frac{G}{r_{j-\myhalf}^2} \sum_{k=1}^{j} \frac{4}{3} \pi  (r_{k-\myhalf}^3 - r_{k-\thalf}^3) \; {\rho_0}_{k-1} \enskip.
+g_{j-\myhalf} = \frac{G}{r_{j-\myhalf}^2} \sum_{k=1}^{j} \frac{4}{3} \pi  (r_{k-\myhalf}^3 - r_{k-\thalf}^3) \; \rho_{0,k-1}^{\inp} \enskip.
 \end{equation}
@@
 \end{eqnarray}
+Finally, $w_0^{\outp} = \ow + \delta w_0$.
 \end{description}
-
 \end{enumerate}
@@
 \end{equation}
-
 \item Compute
@@
 \end{equation}
-
-\item {\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,\psi^{n-\myhalf}]
-\rightarrow [w_0^{n+\myhalf,*}]$
-
-\item Using 
+\item Compute $w_0^{n+\myhalf,*}$:
+\begin{description}
+\item[plane-parallel:]{\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},
+\gammabar^n,p_0^n,\psi^{n-\myhalf}]\rightarrow [w_0^{n+\myhalf,*}]$.
+\item[spherical:]{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,
+\rhozero^n,p_0^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,*}]$.
+\end{description}
+
+\item Define the scaled pressure gradient using
 \begin{equation}
@@
 \end{equation}
-define the scaled pressure gradient
+with the following discretization:
 \begin{equation}
@@
 
-\item First, define:
-\begin{equation}
-\gammabar^{\nph,\star} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}\enskip ,
-\end{equation}
-\begin{equation}
+\item Define:
+\begin{equation}
+\gammabar^{\nph,\star} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}\enskip , ~~~
+\rhozero^{\nph,\star} = \frac{\rhozero^{n} + \rhozero^{n+1,\star}}{2}\enskip , ~~~
 p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip .
 \end{equation}
-{\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},\gammabar^{\nph,\star},p_0^{\nph,\star},
-\psi^{\nph,\star}] \rightarrow [w_0^{\nph}]$.
+\item Compute $w_0^{n+\myhalf}$:
+\begin{description}
+\item[plane-parallel:]{\bf Compute $w_0$ Planar}$[\overline{S^{\nph}},
+\gammabar^{n+\myhalf,*},p_0^{n+\myhalf,*},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf,*}]$.
+\item[spherical:]{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf,*},
+\rhozero^{n+\myhalf,*},p_0^{n+\myhalf,*},\etarho^{n+\myhalf,\star}] \rightarrow [w_0^{n+\myhalf}]$.
+\end{description}
 % VOLUME DISCREPANCY STUFF
@@
 
-\item Using equation (\ref{eq:pizero}), define
+\item Using equation (\ref{eq:pizero}), define the scaled pressure gradient using the
+following discretization:
 \begin{equation}

### 2009-04-23 docs/flowchart/flowchart.tex
hooked in the rest of the spherical base state special calls

@@
 where $\etarho$ at cell-centers is the average of the neighboring edge-values.  
-
 If we write this in matrix form, so that:
@@
 \begin{description}
-\item[plane-parallel:]{\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},
-\gammabar^n,p_0^n,\psi^{n-\myhalf}]\rightarrow [w_0^{n+\myhalf,*}]$.
-\item[spherical:]{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,
-\rhozero^n,p_0^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,*}]$.
+\item For plane-parallel,\\
+{\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,\psi^{n-\myhalf}] 
+\rightarrow [w_0^{n+\myhalf,*}]$.
+\item For spherical,\\
+{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,\rhozero^n,p_0^n,
+\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,*}]$.
 \end{description}
@@
 \begin{description}
-\item[plane-parallel:]
+\item For plane-parallel,
 
@@
 
-\item[spherical:] See Section \ref{Sec:eta Spherical} for working notes.
+\item For spherical, (See Section \ref{Sec:eta Spherical} for working notes).
 
@@
 
-\item Update gravity, $g^{n+1,\star} = g(\rho_0^{n+1,\star})$
-
-\item {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]$ 
-$\rightarrow [p_0^{n+1,\star}, \psi^{\nph,\star}]$.
-
-\item {\bf Advect Base Enthalpy}$[(\rho h)_0^{n}, w_0^{\nph,\star}, \psi^{\nph,\star}]$ 
-$\rightarrow [(\rho h)_0^{n+1,\star}]$.
+\item For spherical problems, update gravity, $g^{n+1,\star} = g(\rho_0^{n+1,\star})$
+
+\item Update $p_0$ and compute $\psi$.
+\begin{description}
+\item For plane-parallel,\\
+{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
+\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
+\item For spherical,\\
+{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,*},\gammabar^{(1)},
+\overline{S^{\nph,\star}}] \rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
+\end{description}
+
+\item Update $(\rho h)_0$.
+\begin{description}
+\item For plane-parallel,\\
+{\bf Advect Base Enthalpy Planar}$[(\rho h)_0^{n}, w_0^{\nph,\star}, \psi^{\nph,\star}]
+\rightarrow [(\rho h)_0^{n+1,\star}]$.
+\item For spherical,\\
+{\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^{n}, w_0^{\nph,\star}, 
+\psi^{n,\star},\psi^{\nph,\star}] \rightarrow [(\rho h)_0^{n+1,\star}]$.
+\end{description}
 
@@
   \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(1a),\star}$ without
-  explicitly including the reaction terms,
+  explicitly including the reaction terms.
 
 \begin{description}
-\item[plane-parallel:]
+\item For plane-parallel,
 
@@
 
-\item[spherical:]
+\item For spherical,
 
@@
 \begin{description}
-\item[plane-parallel:]{\bf Compute $w_0$ Planar}$[\overline{S^{\nph}},
+\item For plane-parallel,\\
+{\bf Compute $w_0$ Planar}$[\overline{S^{\nph}},
 \gammabar^{n+\myhalf,*},p_0^{n+\myhalf,*},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf,*}]$.
-\item[spherical:]{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf,*},
+\item For spherical,\\
+{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf,*},
 \rhozero^{n+\myhalf,*},p_0^{n+\myhalf,*},\etarho^{n+\myhalf,\star}] \rightarrow [w_0^{n+\myhalf}]$.
@@
 
-\item[plane-parallel:]
-
+\item For plane-parallel,
 \begin{equation}
@@
 
-\item Update gravity, $g^{n+1} = g(\rho_0^{n+1})$; $g^{\nph} = g(\rho_0^\nph)$
-
-\item {\bf Advect Base Pressure}$[p_0^{n}, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
-$\rightarrow [p_0^{n+1}, \psi^{n+\myhalf}]$.
-
-\item {\bf Advect Base Enthalpy}$[(\rho h)_0^{n}, w_0^{\nph}, \psi^{\nph}]$ 
-$\rightarrow [(\rho h)_0^{n+1}]$.
+\item For spherical problems, update gravity, 
+$g^{n+1} = g(\rho_0^{n+1})$; $g^{\nph} = g(\rho_0^\nph)$
+
+\item Update $p_0$ and compute $\psi$.
+\begin{description}
+\item For plane-parallel,\\
+{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
+$\rightarrow [p_0^{n+1},\psi^{\nph}]$.
+\item For spherical,\\
+{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{n+1,*},\overline{S^{\nph,\star}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
+\end{description}
+\MarginPar{Rework Advect Base Pressure Spherical to take time-centered $\gammabar$?  Also, look at how $\psi^n$ is computed.  The base time centering doesn't make sense.}
+
+\item Update $(\rho h)_0$.
+\begin{description}
+\item For plane-parallel,\\
+{\bf Advect Base Enthalpy Planar}$[(\rho h)_0^n, w_0^{\nph}, \psi^{\nph}]
+\rightarrow [(\rho h)_0^{n+1}]$.
+\item For spherical,\\
+{\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^n, w_0^{\nph}, \psi^n, \psi^{\nph}]
+\rightarrow [(\rho h)_0^{n+1}]$.
+\end{description}
 
@@
 \begin{description}
-\item[plane-parallel:]
-
+\item For plane-parallel,
 \begin{eqnarray}
@@
 
-\item[spherical:]
-
+\item For spherical,
 \begin{eqnarray}

### 2009-04-24 docs/flowchart/flowchart.tex
some notation fixes
fixed some errors about how we describe computation of beta

@@
 \item Define
-\begin{eqnarray}
-\gammabar^{(1)}        &=& {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, X_k^{(1)}\right) \right] \enskip , \\
- {\beta   }_0^{(1)}    &=& \beta   \left(\rho_0^{n}, p_0^{n}, \gammabar^{(1)}\right) \enskip .
-\end{eqnarray}
-Here $\beta_0$ is computed as described in paper III, Appendix C.
+\begin{equation}
+\gammabar^{(1)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, X_k^{(1)}\right) \right] \enskip .
+\end{equation}
 
@@
 &&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(2),\star} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(2),\star} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
+&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n+1,\star} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n}\right),
 \end{eqnarray}
@@
 \begin{equation}
-T^{(2),\star} = T\left(\rho^{(2),\star}, p_0^{(2),\star}, X_k^{(2),\star}\right) \enskip .
+T^{(2),\star} = T\left(\rho^{(2),\star}, p_0^{n+1,\star}, X_k^{(2),\star}\right) \enskip .
 \end{equation}
@@
                              T^{(2),\star}, (\rho\Hext)^{(2),\star}] $\\
-$\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, 
-              (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, 
+X_k^{n+1,\star}, T^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
 
@@
                                                       X_k^{n+1,\star}\right) \right] \enskip , \\
- {\beta   }_0^{n+1,\star}    &=& \beta   \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}\right) \enskip .
-\end{eqnarray}
+ \beta_0^{n+1,\star}    &=& \beta   \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}\right) \enskip ,
+\end{eqnarray}
+where $\beta_0$ is computed as described in paper III, Appendix C.
 
@@
 &&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
-&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
+&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{n+1} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{n}\right),
 \end{eqnarray}
@@
 \begin{equation}
-T^{(2)} = T\left(\rho^{(2)}, p_0^{(2)}, X_k^{(2)}\right) \enskip .
+T^{(2)} = T\left(\rho^{(2)}, p_0^{n+1}, X_k^{(2)}\right) \enskip .
 \end{equation}
@@
 
-\item {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho^{(2)} \Hext) ]$\\
-                   $\rightarrow [\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, 
-                               (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$
+\item {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho\Hext)^{(2)} ]$\\
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\rightarrow [\rho^{n+1}, 
+(\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$
 
@@
  \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1}, X_k^{n+1}\right) \right] \enskip , \\ 
- {\beta   }_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip .
-\end{eqnarray}
+ \beta_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip ,
+\end{eqnarray}
+where $\beta_0$ is computed as described in paper III, Appendix C.
 
@@
 Set $w_0^0 = 0$ as an initial approximation.  Use the equation of state to 
-determine $(\rho h)^{\initp}$.  Compute $\beta_0^{\initp}$ as a function of 
-the initial data.  Then, project $\Ub^{\initp}$ using $\beta_0^{\initp}$ and 
+determine $(\rho h)^{\initp}$.  Compute $\beta_0^0$ as a function of 
+the initial data.  Then, project $\Ub^{\initp}$ using $\beta_0^0$ and 
 $S = \rho\Hext$, giving $\Ub^{0,0}$.  The next part of the initialization process 
@@
 
-  \item Project $\Ub^{0,\nu-1}$ using $\beta_0^{\initp}$ and 
+  \item Project $\Ub^{0,\nu-1}$ using $\beta_0^0$ and 
         $(S^{0,\nu} - \overline{S^{0,\nu}})$ as the source term.  

### 2009-04-24 docs/flowchart/flowchart.tex
more detailed notes (in red) about proposed changes to Advect Base Pressure Spherical

@@
 
+\begin{color}{red}
+Andy's proposal: We need to pass in two different $\gammabar$ objects, $\gammabar^{\inp,n}$ and $\gammabar^{\inp,n+\myhalf}$.
+The pressure update will use $\gammabar^{\inp,n+\myhalf}$.
+To compute $\psi^{\outp,n}$, we use $\gammabar^{\inp,n}$.
+To compute $\psi^{\outp,n+\myhalf}$, we use $\gammabar^{\inp,n+\myhalf}$.
+In the first half of the algorithm, we pass in $\gammabar^n$ for both arguments.
+For the second half of the algorithm, we pass in $\gammabar^n$ and $(\gammabar^n+\gammabar^{n+1,*})/2$.
+I don't think we ever need $\gammabar^{(1)}$, as the only place it's currently used is as an input to the first call to {\bf Advect Base Pressure Spherical}.
+\end{color}
+
 First use a predictor-corrector formulation for the pressure update 
@@
 \end{description}
-\MarginPar{Rework Advect Base Pressure Spherical to take time-centered $\gammabar$?  Also, look at how $\psi^n$ is computed.  The base time centering doesn't make sense.}
 

### 2009-04-27 docs/flowchart/flowchart.tex
added eta computation in spherical

@@
 \begin{color}{red}
-Andy's proposal: We need to pass in two different $\gammabar$ objects, $\gammabar^{\inp,n}$ and $\gammabar^{\inp,n+\myhalf}$.
+Proposal: We need to pass in two different $\gammabar$ objects, $\gammabar^{\inp,n}$ and $\gammabar^{\inp,n+\myhalf}$.
 The pressure update will use $\gammabar^{\inp,n+\myhalf}$.
@@
 In the first half of the algorithm, we pass in $\gammabar^n$ for both arguments.
-For the second half of the algorithm, we pass in $\gammabar^n$ and $(\gammabar^n+\gammabar^{n+1,*})/2$.
+For the second half of the algorithm, we pass in $\gammabar^n$ and $(\gammabar^n+\gammabar^{n+1,\star})/2$.
 I don't think we ever need $\gammabar^{(1)}$, as the only place it's currently used is as an input to the first call to {\bf Advect Base Pressure Spherical}.
@@
 we can write
-\begin{eqnarray*}
+\begin{equation}
 \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
  \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) - \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
-\end{eqnarray*}
+\end{equation}
 We time-center the left hand side:
-\[
+\begin{equation}
 \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
    \frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} - p_0^{\inp}}{\dt}\enskip.
-\]
+\end{equation}
 We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
@@
 We begin with equation (23) in paper III written in spherical coordinates:
-\[
+\begin{equation}
 \frac{1}{r^2}\frac{\partial}{\partial r} \left (r^2 \beta_0 w_0 \right ) =
 \beta_0 \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )\enskip.
-\]
+\end{equation}
 We expand the spatial derivative and divide through by $\beta_0$:
@@
 Recalling from paper I that;
-\[
+\begin{equation}
 \frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} = \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}\enskip,
-\]
+\end{equation}
 and moving this term to the right hand side,
@@
 To do so we start with the equation of hydrostatic equilibrium in a spherical geometry:
-\[
+\begin{equation}
 \frac{\partial p_0}{\partial r} = -\rho_0 g \, ; \quad
 g = \frac{G m_\mathrm{encl}}{r^2}\enskip,
-\]
+\end{equation}
 where $m_\mathrm{encl}(r)$ is the mass enclosed at radius $r$ and $G$ is the
@@
 \end{equation}
-\item Compute $w_0^{n+\myhalf,*}$:
+\item Compute $w_0^{n+\myhalf,\star}$:
 \begin{description}
@@
 {\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,\psi^{n-\myhalf}] 
-\rightarrow [w_0^{n+\myhalf,*}]$.
+\rightarrow [w_0^{n+\myhalf,\star}]$.
 \item For spherical,\\
 {\bf Compute $w_0$ Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,\rhozero^n,p_0^n,
-\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,*}]$.
+\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
 \end{description}
@@
 \item For plane-parallel,
-
 \begin{equation}
@@
 \end{equation}
-
-\item For spherical, (See Section \ref{Sec:eta Spherical} for working notes).
-
+\item For spherical, first construct 
+$\eta_{\rho}^{{\rm cart},\nph,\star} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf,\star}$ using:
+\begin{equation}
+\eta_{\rho}^{{\rm cart},\nph,\star} = \left[\left(\frac{\rho^{(1)}+\rho^{(2),\star}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a),\star}}{2}\right)\right] \sum_d \left(\uadvone \cdot \eb_d\right)n_d.
+\end{equation}
+Then, $\etarho^{\nph,\star}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph,\star}$,
+\begin{equation}
+\etarho^{\nph,\star} = \overline{\eta_{\rho}^{{\rm cart},\nph,\star}}.
+\end{equation}
+This gives a radial bin-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
+radial edges, average the two neighboring radial bins.
 \end{description}
@@
 \item For spherical,\\
-{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,*},\gammabar^{(1)},
+{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
 \overline{S^{\nph,\star}}] \rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
@@
     \left(\uadvone \cdotb \er\right) \left(\frac{\partial
-      p_0}{\partial r} \right)^{n} + \dt \; \psi^{\nph,*} \enskip ,
+      p_0}{\partial r} \right)^{n} + \dt \; \psi^{\nph,\star} \enskip ,
   \end{eqnarray}
@@
        - p_0^{n} \nabla \cdot \uadvone \right \} \nonumber \\
-    && + \dt \; \psi^{\nph,*} \enskip ,
+    && + \dt \; \psi^{\nph,\star} \enskip ,
   \end{eqnarray}
@@
 {\bf Compute $w_0$ Planar}$[\overline{S^{\nph}},
-\gammabar^{n+\myhalf,*},p_0^{n+\myhalf,*},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf,*}]$.
+\gammabar^{n+\myhalf,\star},p_0^{n+\myhalf,\star},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf,\star}]$.
 \item For spherical,\\
-{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf,*},
-\rhozero^{n+\myhalf,*},p_0^{n+\myhalf,*},\etarho^{n+\myhalf,\star}] \rightarrow [w_0^{n+\myhalf}]$.
+{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf,\star},
+\rhozero^{n+\myhalf,\star},p_0^{n+\myhalf,\star},\etarho^{n+\myhalf,\star}] \rightarrow [w_0^{n+\myhalf}]$.
 \end{description}
@@
 \end{equation}
-
-\item For spherical, (See Section \ref{Sec:eta Spherical} for working notes).
-
+\item For spherical, first construct 
+$\eta_{\rho}^{{\rm cart},\nph} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf}$ using:
+\begin{equation}
+\eta_{\rho}^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2),\star}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a),\star}}{2}\right)\right] \sum_d \left(\uadvtwo \cdot \eb_d\right)n_d.
+\end{equation}
+Then, $\etarho^{\nph}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph}$,
+\begin{equation}
+\etarho^{\nph} = \overline{\eta_{\rho}^{{\rm cart},\nph}}.
+\end{equation}
+This gives a radial bin-centered $\etarho^{\nph}$.  To get $\etarho^{\nph}$ at 
+radial edges, average the two neighboring radial bins.
 \end{description}
@@
 \item For spherical,\\
-{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{n+1,*},\overline{S^{\nph,\star}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
+{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{n+1,\star},\overline{S^{\nph,\star}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
 \end{description}

### 2009-04-27 docs/flowchart/flowchart.tex
somewhat major changes to integrate the volume discrepancy correction term in a more seamless manner

@@
 {\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
-\item We added the volume-discrepancy term to the velocity constraint equation,
+\item We added the volume discrepancy term to the velocity constraint equation,
 controlled by the input parameter, {\tt dpdt\_factor}.
@@
 \end{equation}
-and $w_0=0$ at $r=0$.
-% VOLUME DISCREPANCY STUFF
-%\begin{equation}
-%\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S^{\nph,\star}}
-%- \frac{1}{\gammabar^{n} p_0^{n}}
-%\left\{ \psi^{\nmh} + 
-%f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n} \right]\right\}
-%\enskip .
-%\end{equation}
+and $w_0=0$ at $r=0$.\\
+
+The discretization with the volume discrepancy correction require an additional argument,
+$\overline{p^{\inp}}$, and the discretization becomes:
+\begin{equation}
+\frac{\partial w_0^{\nph,\outp}}{\partial r} = 
+\overline{S^{\nph,\inp}} - \frac{1}{\gammabar^{\inp} p_0^{\inp}}
+\left[ \psi^{\inp} + f \left(\frac{p_0^{\inp} - \overline{p^{\inp}}}{\Delta t^n}\right)\right]
+\enskip .
+\end{equation}
 
@@
 \rightarrow [w_0^{n+\myhalf,\star}]$.
+
+If we are using the volume discrepancy term, compute 
+$\overline{p^{\inp}} = \overline{p(\rho,h,X_k)^n}$.
 \item For spherical,\\
@@
 \begin{equation}
-\nablab \cdotb (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip .
+\nablab \cdotb (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
 \label{eq:tildeconstraint}
+\end{equation}
+with the discretization
+\begin{equation}
+\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
+\beta_0^n \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right) \enskip .
 \end{equation}
@@
 \begin{equation}
-\gammabar^{\nph,\star} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}\enskip , ~~~
-\rhozero^{\nph,\star} = \frac{\rhozero^{n} + \rhozero^{n+1,\star}}{2}\enskip , ~~~
-p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip .
+\gammabar^{\nph} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}\enskip , ~~~
+\rhozero^{\nph} = \frac{\rhozero^{n} + \rhozero^{n+1,\star}}{2}\enskip , ~~~
+p_0^{\nph} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip .
 \end{equation}
@@
 {\bf Compute $w_0$ Planar}$[\overline{S^{\nph}},
-\gammabar^{n+\myhalf,\star},p_0^{n+\myhalf,\star},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf,\star}]$.
+\gammabar^{n+\myhalf},p_0^{n+\myhalf},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf}]$.
+
+If we are using the volume discrepancy term, compute
+\begin{equation}
+\overline{p^{\inp}} = \frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}.
+\end{equation}
 \item For spherical,\\
-{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf,\star},
-\rhozero^{n+\myhalf,\star},p_0^{n+\myhalf,\star},\etarho^{n+\myhalf,\star}] \rightarrow [w_0^{n+\myhalf}]$.
+{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf},
+\rhozero^{n+\myhalf},p_0^{n+\myhalf},\etarho^{n+\myhalf}] \rightarrow [w_0^{n+\myhalf}]$.
 \end{description}
-% VOLUME DISCREPANCY STUFF
-%\begin{equation}
-%\overline{p(\rho,h,X_k)^{\nph,\star}} = 
-%\frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}\enskip .
-%\end{equation}
-%\begin{equation}
-%\frac{\partial w_0^{\nph}}{\partial r} = 
-%\overline{S^{\nph}} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
-%\left\{ \psi^{\nph,\star} + f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\}
-%\enskip ,
 
@@
 \begin{equation}
-w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
+w_0^n = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
 \end{equation}
@@
 \begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
-+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\label{eq:dw0drnstar}
+\left(\frac{\partial w_0}{\partial r}\right)^{n} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
++ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph} \right ] \enskip.\label{eq:dw0drnstar}
 \end{equation}
@@
 for $w_0$ and $\pizero.$  We note that $\uadvtwo$ satisfies the discrete versions of 
-equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}).   The discretization
-with the volume discrepancy correction is
-\begin{eqnarray}
-\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) &=&
-\beta_0^{\nph,\star}\left\{\left(S^{\nph} - \overline{S^{\nph}}\right)
- + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}\cdot \right. \nonumber \\
-&& \left.\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\} \enskip ,
-\end{eqnarray}
-\begin{equation}
-\beta_0^{\nph,\star} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2} \enskip .
+equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}), with the discretization
+\begin{equation}
+\nablab \cdotb \left(\beta_0^{\nph} \uadvtwo\right) =
+\beta_0^{\nph}\left(S^{\nph} - \overline{S^{\nph}}\right) \enskip ,
+\end{equation}
+\begin{equation}
+\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2} \enskip .
+\end{equation}
+The discretization with the volume discrepancy correction is
+\begin{equation}
+\nablab \cdotb \left(\beta_0^{\nph} \uadvtwo\right) =
+\beta_0^{\nph}\left\{\left(S^{\nph} - \overline{S^{\nph}}\right)
+ + \frac{f}{\gammabar^{\nph} p_0^{\nph}}\cdot 
+\left[\frac{p(\rho,h,X_k)^{\nph} - \overline{p(\rho,h,X_k)^{\nph}}}{\Delta t^n}\right]\right\} \enskip ,
 \end{equation}
@@
 \item For spherical,\\
-{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{n+1,\star},\overline{S^{\nph,\star}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
+{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{n+1,\star},\overline{S^{\nph}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
 \end{description}
@@
 \end{eqnarray}
-where $\rho^\nph = ( \rho^n + \rho^{n+1} )/2$
+where $\rho^{\nph}$ is defined as
+\begin{equation}
+\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}\enskip ,
+\end{equation}
 and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
@@
 
-Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}) 
-including the volume discrepancy correction,
+Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}):
+\begin{equation}
+\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right) 
+= \beta_0^{\nph} \left(S^{n+1} - \overline{S^{n+1}} \right) \enskip,
+\end{equation}
+where $\beta_0^{n+\myhalf}$ is now defined as
+\begin{equation}
+\beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1}}{2}\enskip .
+\end{equation}
+The discretization with the volume discrepancy term is:
 \begin{equation}
@@
 \left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\}
-\enskip,
-\end{equation}
-by defining ${\bf V} = \Ubt^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
-and solving
+\enskip .
+\end{equation}
+
+To solve this equation, we define
+${\bf V} = \Ubt^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
+and solve
+\begin{equation}
+ L_\beta^\rho \phi = D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} 
+\left(S^{n+1}-\overline{S^{n+1}} \right) \enskip, \label{Linear System for Multigrid}
+\end{equation}
+for nodal values of $\phi$, where $L_\beta^\rho$ is the standard bilinear
+finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
+The discretization with the volume discrepancy term is:
 \begin{equation}
@@
 - \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\} \enskip, \label{Linear System for Multigrid}
-\end{equation}
-for nodal values of $\phi$, where
-$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
-$L_\beta^\rho$ is the standard bilinear
-finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
+\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\} \enskip.
+\end{equation}
 In this step, $D$ is a discrete second-order operator that approximates the 

### 2009-04-27 docs/flowchart/flowchart.tex
fixed a typo reworked the text so the thermal diffusion section can be commented out without having to redefine variable superscripts

@@
 
-  \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(1a),\star}$ without
+  \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ without
   explicitly including the reaction terms.
@@
   \begin{eqnarray}
-  (\rho h)^{(1a),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
+  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
       \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right) (\rho
@@
 \begin{equation}
-(\rho h)^{(1a),\star} = \rho^{(2),\star}h\left(\rho^{(2),\star},p_0^{n+1,\star},X_k^{(2),\star}\right).
+(\rho h)^{(2),\star} = \rho^{(2),\star}h\left(\rho^{(2),\star},p_0^{n+1,\star},X_k^{(2),\star}\right).
 \end{equation}
@@
 
-\item Diffuse the enthalpy through a time interval of $\dt$.  Here we begin
+\item If we are using thermal diffusion, diffuse the enthalpy through a time interval of 
+$\dt$.  First, define $(\rho h)^{(1a),\star} = (\rho h)^{(2),\star}$.  We recompute
+$(\rho h)^{(2),\star}$ to account for thermal diffusion.  Here we begin
 with the enthalpy equation (\ref{eq:rhohupd}), but consider only the 
@@
 \begin{eqnarray}
-(\rho h)^{(1a)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
+(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
 (\rho h)^{(1),\nph} \right] \right\} \nonumber \\
@@
 \begin{equation}
-(\rho h)^{(1a)} = \rho^{(2)}h\left(\rho^{(2)},p_0^{n+\myhalf},X_k^{(2)}\right).
+(\rho h)^{(2)} = \rho^{(2)}h\left(\rho^{(2)},p_0^{n+1},X_k^{(2)}\right).
 \end{equation}
@@
 %--------------------------------------------------------------------------
-\item Diffuse the enthalpy through a time interval of $\dt$.
-Compute $\kth^{(2),\star}, c_p^{(2),\star}$, and $\xi_k^{(2),\star}$, from $\rho^{(2),\star}, T^{(2),\star}$, and $X_k^{(2),\star}$ as inputs to the equation of state.
-The update is given by
+\item If we are using thermal diffusion, diffuse the enthalpy through a time interval of 
+$\dt$.  First, define $(\rho h)^{(1a)} = (\rho h)^{(2)}$.  We recompute $(\rho h)^{(2)}$ to 
+account for thermal diffusion.  Compute $\kth^{(2),\star}, c_p^{(2),\star}$, and 
+$\xi_k^{(2),\star}$, from $\rho^{(2),\star}, T^{(2),\star}$, and $X_k^{(2),\star}$ as inputs to 
+the equation of state.  The update is given by
 \begin{eqnarray}

### 2009-04-27 docs/flowchart/flowchart.tex
added a Compute Gravity function

@@
 \end{description}
+
+\item {\bf Compute Gravity}$[\rhozero^{\inp}] \rightarrow [g^{\outp}]$\\
+is the process by which we compute gravity at both radial cell-centers and edges.
+\begin{description}
+\item For plane-parallel, we hold the gravity constant in space and time.
+\item For spherical, we define the enclosed mass at edges, and then cell-centers:
+\begin{equation}
+m_{j-\myhalf} = \sum_{k=1}^j \frac{4}{3}\pi\left(r_{k-\myhalf}^3 - r_{k-\sfrac{3}{2}}^3\right)\rho_{0,k-1}^{\inp}; ~~~~~ m_j = m_{j-\myhalf} + \frac{4}{3}\pi \left(r_k^3 - r_{k-\myhalf}^3\right)\rho_{0,k}^{\inp} \enskip .
+\end{equation}
+Then, we define the gravity at both edges and cell-centers:
+\begin{equation}
+g_{j-\myhalf} = \frac{Gm_{j-\myhalf}}{r_{j-\myhalf}^2}; ~~~~~ g_j = \frac{Gm_j}{r_j^2}\enskip .
+\end{equation}
+\end{description}
+
 \end{enumerate}
@@
 \end{equation}
-This gives a radial bin-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
+This gives a radial cell-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
 radial edges, average the two neighboring radial bins.
@@
 
-\item For spherical problems, update gravity, $g^{n+1,\star} = g(\rho_0^{n+1,\star})$
+\item {\bf Compute Gravity}$[\rho_0^{n+1,\star}] \rightarrow [g^{n+1,\star}]$.
 
@@
                                                       X_k^{n+1,\star}\right) \right] \enskip , \\
- \beta_0^{n+1,\star}    &=& \beta   \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}\right) \enskip ,
+ \beta_0^{n+1,\star}    &=& \beta   \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}, g^{n+1,\star}\right) \enskip ,
 \end{eqnarray}
@@
 \end{equation}
-This gives a radial bin-centered $\etarho^{\nph}$.  To get $\etarho^{\nph}$ at 
+This gives a radial cell-centered $\etarho^{\nph}$.  To get $\etarho^{\nph}$ at 
 radial edges, average the two neighboring radial bins.
@@
 
-\item For spherical problems, update gravity, 
-$g^{n+1} = g(\rho_0^{n+1})$; $g^{\nph} = g(\rho_0^\nph)$
+\item {\bf Compute Gravity}$[\rho_0^{n+1}] \rightarrow [g^{n+1}]$.
+
+Then, set $g^{n+\myhalf} = (g^n + g^{n+1})/2$.
 
@@
  \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1}, X_k^{n+1}\right) \right] \enskip , \\ 
- \beta_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip ,
+ \beta_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1}, \gammabar^{n+1}, g^{n+1}\right) \enskip ,
 \end{eqnarray}
@@
         + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
-        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
+        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g^{n+\myhalf} \er \right] \enskip ,
 \end{eqnarray}

### 2009-04-29 docs/flowchart/flowchart.tex
typo
changed some notation and added details about how to fix the gamma1bar problem in advect base spherical

@@
 \begin{color}{red}
-Proposal: We need to pass in two different $\gammabar$ objects, $\gammabar^{\inp,n}$ and $\gammabar^{\inp,n+\myhalf}$.
+%Proposal: We need to pass in two different $\gammabar$ objects, $\gammabar^{\inp,n}$ and $\gammabar^{\inp,n+\myhalf}$.
 The pressure update will use $\gammabar^{\inp,n+\myhalf}$.
@@
 interpolating $p_0^\inp$ from cell-centers to edges and then
-differencing.  Now we begin the corrector step.  Using $p_0^\star$, we can compute a more 
+differencing.  Now we begin the corrector step.
+\begin{color}{red} We need to compute 
+$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{(2)},p_0^\star,X_k^{(2)})]$ 
+and $\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$, and then use 
+$\gammabar^\dagger$ in the $f^\dagger$ equation. \end{color} 
+Using $p_0^\star$, we can compute a more 
 accurate approximation to equation~(\ref{eq:frhs}),
 \begin{equation}
-\tilde{f} = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
-                                             \frac{1}{\gammabar^{\inp} \tilde{p_0}} w_0^{\inp} \left ( \frac{\partial \tilde{p_0}}{\partial r} \right ) \right ] \enskip ,
-\end{equation}
-where $\tilde{p} = (p_0^\inp + p_0^\star)/2$, and the difference is again computed
-by first putting $\tilde{p}$ on edges.
+f^\dagger = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
+                                             \frac{1}{\gammabar^{\inp} p_0^\dagger} w_0^{\inp} \left ( \frac{\partial p_0^\dagger}{\partial r} \right ) \right ] \enskip ,
+\end{equation}
+where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$, and the difference is again computed
+by first putting $p_0^\dagger$ on edges.
 
@@
 in time, we expect the effect to be small.  This is something that we 
-defer until simulations indicate it is warranted.
+defer until simulations indicate it is warranted.\\
 
@@
 \left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
-          (r^2 w_0^{\outp})_{j-\myhalf} \right ] \right \} \enskip ,
+          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip ,
 \end{equation}
@@
 \left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
-          (r^2 w_0^{\outp})_{j-\myhalf} \right ] \right \} \enskip .
+          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip .
 \end{equation}
@@
 \begin{equation}
-m_{j-\myhalf} = \sum_{k=1}^j \frac{4}{3}\pi\left(r_{k-\myhalf}^3 - r_{k-\sfrac{3}{2}}^3\right)\rho_{0,k-1}^{\inp}; ~~~~~ m_j = m_{j-\myhalf} + \frac{4}{3}\pi \left(r_k^3 - r_{k-\myhalf}^3\right)\rho_{0,k}^{\inp} \enskip .
+m_{j-\myhalf} = \sum_{k=1}^j \frac{4}{3}\pi\left(r_{k-\myhalf}^3 - r_{k-\sfrac{3}{2}}^3\right)\rho_{0,k-1}^{\inp}; ~~~~~ m_j = m_{j-\myhalf} + \frac{4}{3}\pi \left(r_j^3 - r_{j-\myhalf}^3\right)\rho_{0,j}^{\inp} \enskip .
 \end{equation}
@@
 \overline{S^{\nph,\star}}] \rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
+
 \end{description}
@@
 $\rightarrow [p_0^{n+1},\psi^{\nph}]$.
-\item For spherical,\\
+\item For spherical,\MarginPar{$\gammabar^{(1)}$ instead of $\gammabar^{n+1,\star}$?}\\
 {\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{n+1,\star},\overline{S^{\nph}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.

### 2009-04-29 docs/flowchart/flowchart.tex
pass in gamma1bar^(1) instead of gamma1bar^{n+1,star} into the second call to advect base pressure

@@
 $\rightarrow [p_0^{n+1},\psi^{\nph}]$.
-\item For spherical,\MarginPar{$\gammabar^{(1)}$ instead of $\gammabar^{n+1,\star}$?}\\
-{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{n+1,\star},\overline{S^{\nph}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
+\item For spherical,\\
+{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{(1)},\overline{S^{\nph}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
 \end{description}

### 2009-04-29 Source/advance.f90

@@ -140,6 +140,7 @@ contains
     real(dp_t) ::                 Sbar(nlevs_radial,0:nr_fine-1)
     real(dp_t) ::        div_coeff_nph(nlevs_radial,0:nr_fine-1)
     real(dp_t) ::        gamma1bar_old(nlevs_radial,0:nr_fine-1)
+    real(dp_t) ::          gamma1bar_1(nlevs_radial,0:nr_fine-1)
     real(dp_t) :: delta_gamma1_termbar(nlevs_radial,0:nr_fine-1)
     real(dp_t) ::               w0_old(nlevs_radial,0:nr_fine)
     real(dp_t) ::       div_coeff_edge(nlevs_radial,0:nr_fine)
@@ -271,8 +272,8 @@ contains
 
        call average(mla,Source_nph,Sbar,dx,1)
 
-       call make_w0(w0,w0_old,w0_force,Sbar,rho0_old,rho0_old,p0_old,p0_old,gamma1bar, &
-                    gamma1bar,p0_minus_pthermbar,psi,etarho_ec,etarho_cc,dt,dtold)
+       call make_w0(w0,w0_old,w0_force,Sbar,rho0_old,rho0_old,p0_old,p0_old,gamma1bar_old, &
+                    gamma1bar_old,p0_minus_pthermbar,psi,etarho_ec,etarho_cc,dt,dtold)
 
        if (spherical .eq. 1) then
           call put_w0_on_edges(mla,w0,w0mac,dx,div_coeff_old,the_bc_tower)
@@ -321,7 +322,7 @@ contains
     end do
 
     call make_macrhs(macrhs,rho0_old,Source_nph,delta_gamma1_term,Sbar,div_coeff_old,dx, &
-                     gamma1bar,gamma1bar,p0_old,p0_old,delta_p_term,dt)
+                     gamma1bar_old,gamma1bar_old,p0_old,p0_old,delta_p_term,dt)
 
     do n=1,nlevs
        call destroy(delta_gamma1_term(n))
@@ -401,7 +402,7 @@ contains
        end do
        
        call make_gamma(mla,gamma1,s1,p0_old,dx)
-       call average(mla,gamma1,gamma1bar,dx,1)
+       call average(mla,gamma1,gamma1bar_1,dx,1)
 
        do n=1,nlevs
           call destroy(gamma1(n))
@@ -509,7 +510,7 @@ contains
 
        if (p0_update_type .eq. 1) then
 
[…]

### 2009-04-29 Source/make_w0.f90

@@ -216,7 +216,7 @@ contains
   end subroutine make_w0_planar
 
   subroutine make_w0_spherical(w0,w0_old,Sbar_in,rho0,rho0_new,p0,p0_new, &
-                               gamma1bar,gamma1bar_new,p0_minus_pthermbar, &
+                               gamma1bar_old,gamma1bar_new,p0_minus_pthermbar, &
                                etarho_ec,etarho_cc,w0_force,dt,dtold)
 
     use geometry, only: r_cc_loc, nr_fine, r_edge_loc, dr
@@ -232,7 +232,7 @@ contains
     real(kind=dp_t), intent(in   ) ::           rho0_new(0:)
     real(kind=dp_t), intent(in   ) ::                 p0(0:)
     real(kind=dp_t), intent(in   ) ::             p0_new(0:)
-    real(kind=dp_t), intent(in   ) ::          gamma1bar(0:)
+    real(kind=dp_t), intent(in   ) ::      gamma1bar_old(0:)
     real(kind=dp_t), intent(in   ) ::      gamma1bar_new(0:)
     real(kind=dp_t), intent(in   ) :: p0_minus_pthermbar(0:)
     real(kind=dp_t), intent(in   ) ::          etarho_ec(0:)
@@ -263,7 +263,7 @@ contains
     do r = 0, nr_fine-1
        p0_nph(r)        = HALF*(p0(r)        + p0_new(r))
        rho0_nph(1,r)    = HALF*(rho0(r)      + rho0_new(r))
-       gamma1bar_nph(r) = HALF*(gamma1bar(r) + gamma1bar_new(r))       
+       gamma1bar_nph(r) = HALF*(gamma1bar_old(r) + gamma1bar_new(r))       
     enddo
 
     ! NOTE:  we first solve for the w0 resulting only from Sbar -- then we will

### 2009-04-29 docs/flowchart/flowchart.tex
notes on how to time-center  gamma1bar in advect_base_pres_spherical.  not implemented in code yet

@@
 
-\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},\Sbar^{\inp}] 
-\rightarrow [p_0^{\outp},\psi^{n,\outp},\psi^{n+\myhalf,\outp}]$:
-
-\begin{color}{red}
-%Proposal: We need to pass in two different $\gammabar$ objects, $\gammabar^{\inp,n}$ and $\gammabar^{\inp,n+\myhalf}$.
-The pressure update will use $\gammabar^{\inp,n+\myhalf}$.
-To compute $\psi^{\outp,n}$, we use $\gammabar^{\inp,n}$.
-To compute $\psi^{\outp,n+\myhalf}$, we use $\gammabar^{\inp,n+\myhalf}$.
-In the first half of the algorithm, we pass in $\gammabar^n$ for both arguments.
-For the second half of the algorithm, we pass in $\gammabar^n$ and $(\gammabar^n+\gammabar^{n+1,\star})/2$.
-I don't think we ever need $\gammabar^{(1)}$, as the only place it's currently used is as an input to the first call to {\bf Advect Base Pressure Spherical}.
-\end{color}
+\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},
+\Sbar^{\inp}, \rho^\inp, X_k^\inp] \rightarrow [p_0^{\outp},\psi^{n,\outp},\psi^{n+\myhalf,\outp}]$:
 
@@
 interpolating $p_0^\inp$ from cell-centers to edges and then
-differencing.  Now we begin the corrector step.
-\begin{color}{red} We need to compute 
-$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{(2)},p_0^\star,X_k^{(2)})]$ 
-and $\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$, and then use 
-$\gammabar^\dagger$ in the $f^\dagger$ equation. \end{color} 
-Using $p_0^\star$, we can compute a more 
-accurate approximation to equation~(\ref{eq:frhs}),
+differencing.  Next, we compute a provisional updated $\gammabar$ using
+$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^\star,X_k^{\inp})]$.\\
+
+Now we begin the corrector step.  Using $p_0^\star$ and $\gammabar^\star$, 
+we can compute a more accurate approximation to equation~(\ref{eq:frhs}),
 \begin{equation}
 f^\dagger = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
-                                             \frac{1}{\gammabar^{\inp} p_0^\dagger} w_0^{\inp} \left ( \frac{\partial p_0^\dagger}{\partial r} \right ) \right ] \enskip ,
-\end{equation}
-where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$, and the difference is again computed
-by first putting $p_0^\dagger$ on edges.
+                                             \frac{1}{\gammabar^{\dagger} p_0^\dagger} w_0^{\inp} \left ( \frac{\partial p_0^\dagger}{\partial r} \right ) \right ] \enskip ,
+\end{equation}
+where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$, 
+$\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$, 
+and the difference is again computed by first putting $p_0^\dagger$ on edges.
 
@@
 {\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
-\overline{S^{\nph,\star}}] \rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
+\overline{S^{\nph,\star}}, \rho^{(2),\star}, X_k^{(2),\star}]$\\
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+\rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
 
@@
 \item For spherical,\\
-{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{(1)},\overline{S^{\nph}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
+{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{(1)},\overline{S^{\nph}},
+\rho^{(2)},X_k^{(2)}]$\\
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+ \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
 \end{description}

### 2009-04-29 docs/flowchart/flowchart.tex
finished documenting the change to advect base spherical

@@
 differencing.  Next, we compute a provisional updated $\gammabar$ using
-$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^\star,X_k^{\inp})]$.\\
+$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^\star,X_k^{\inp})]$,
+where $\rho^\inp$ and $X_k^\inp$ are time-advanced versions of the $\rho$ and
+$X_k$ used to compute $\gammabar^\inp$.\\
 
@@
 \end{equation}
-where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$, 
-$\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$, 
-and the difference is again computed by first putting $p_0^\dagger$ on edges.
-
+where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$ and
+$\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$.
 This allows us to complete the corrector for the pressure update:
 \begin{equation}
-p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} \tilde{f}\right)}{\left(1 - \gammabar^{\inp} \tilde{f}\right)}\enskip .
+p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\dagger} f^\dagger\right)}{\left(1 - \gammabar^{\dagger} f^\dagger\right)}\enskip .
 \end{equation}
@@
 
-Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.  Rewriting Equation (\ref{eq:w01}), we can define
+Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.  Rewriting Equation 
+(\ref{eq:w01}), we can define
 \begin{eqnarray}

### 2009-04-29 docs/flowchart/flowchart.tex
we now compute the time-centered psi in spherical with a time-centered gamma1bar rather than a base time gamma1bar

@@
 interpolating $p_0^\inp$ from cell-centers to edges and then
-differencing.  Next, we compute a provisional updated $\gammabar$ using
+differencing.  Next, we compute a provisional updated $\gammabar$ using the
+provisional updated pressure,
 $\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^\star,X_k^{\inp})]$,
 where $\rho^\inp$ and $X_k^\inp$ are time-advanced versions of the $\rho$ and
-$X_k$ used to compute $\gammabar^\inp$.\\
+$X_k$ which were used to compute $\gammabar^\inp$.\\
 
@@
 \end{equation}
-We note that since we do not time-center $\gammabar$, this is not
-strictly second-order accurate.  However, as $\gammabar$ varies slowly
-in time, we expect the effect to be small.  This is something that we 
-defer until simulations indicate it is warranted.\\
-
-Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.  Rewriting Equation 
-(\ref{eq:w01}), we can define
+Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.
+First, we compute an updated $\gammabar$ using the update pressure, $\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^{\outp},X_k^{\inp})]$.
+Then, rewriting equation (\ref{eq:w01}), we can define
 \begin{eqnarray}
@@
 \psi_j^{\outp,n} = 
-\left(\gammabar^{\inp}p_0^{\inp}\right)_j
+\left(\gammabar^\inp p_0^\inp\right)_j
 \left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
@@
 \psi_j^{\outp,n+\myhalf} = 
-\left[\gammabar^{\inp} \frac{\left( p_0^{\inp} + p_0^{\outp} \right)}{2}\right]_j
+\left(\frac{\gammabar^\inp+\gammabar^{\star}}{2}\right)_j
+\left(\frac{p_0^\inp+p_0^\outp}{2}\right)_j
 \left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
@@
 \item {\bf Advect Base Enthalpy Planar}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow 
-[(\rho h)_0^\outp]$
+[(\rho h)_0^\outp]$:
 
@@
 \item {\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^\inp,w_0^\inp,\psi^{\inp,n},
-\psi^{\inp,n+\myhalf}] \rightarrow [(\rho h)_0^\outp]$
+\psi^{\inp,n+\myhalf}] \rightarrow [(\rho h)_0^\outp]$:
 

### 2009-04-29 Source/advance.f90

@@ -527,6 +527,7 @@ contains
              call make_psi_planar(etarho_cc,psi)
           else
              call make_psi_spherical(psi_old,w0,gamma1bar_1,p0_old,p0_old,Sbar)
+             ! this isn't right yet.  gamma1bar needs to be time-centered.
              call make_psi_spherical(psi,w0,gamma1bar_1,p0_old,p0_new,Sbar)
           end if
           
@@ -932,6 +933,7 @@ contains
              call make_psi_planar(etarho_cc,psi)
           else
              call make_psi_spherical(psi_old,w0,gamma1bar_1,p0_old,p0_old,Sbar)
+             ! this isn't right yet.  gamma1bar needs to be time-centered.
              call make_psi_spherical(psi,w0,gamma1bar_1,p0_old,p0_new,Sbar)
           end if
           

### 2009-04-29 Source/advect_base.f90

@@ -138,7 +138,7 @@ contains
                               dt,dx,mla)
 
     use bl_prof_module
-    use geometry, only: spherical, nlevs
+    use geometry, only: spherical, nlevs, nr_fine
     use make_psi_module
     use restrict_base_module
     use multifab_module
@@ -160,6 +160,8 @@ contains
     ! local
     type(bl_prof_timer), save :: bpt
 
+    real(kind=dp_t) :: gamma1bar_avg(1,0:nr_fine-1)
+
     call build(bpt, "advect_base_pres")
 
     if (spherical .eq. 0) then
@@ -177,10 +179,11 @@ contains
     else
 
        ! advect p0
-       call advect_base_pres_spherical(w0,Sbar_in,p0_old,p0_new,gamma1bar,s,dt,dx,mla)
+       call advect_base_pres_spherical(w0,Sbar_in,p0_old,p0_new,gamma1bar,gamma1bar_avg, &
+                                       s,dt,dx,mla)
 
        ! make psi
-       call make_psi_spherical(psi,w0,gamma1bar,p0_old,p0_new,Sbar_in)
+       call make_psi_spherical(psi,w0,gamma1bar_avg,p0_old,p0_new,Sbar_in)
        call make_psi_spherical(psi_old,w0,gamma1bar,p0_old,p0_old,Sbar_in)
 
     end if
@@ -227,7 +230,8 @@ contains
 
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
   
-  subroutine advect_base_pres_spherical(w0,Sbar_in,p0_old,p0_new,gamma1bar,s,dt,dx,mla)
+  subroutine advect_base_pres_spherical(w0,Sbar_in,p0_old,p0_new,gamma1bar,gamma1bar_avg, &
+                                        s,dt,dx,mla)
 
     use bl_constants_module
     use make_edge_state_module
@@ -245,6 +249,7 @@ contains
     real(kind=dp_t), intent(in   ) ::    p0_old(:,0:)
     real(kind=dp_t), intent(  out) ::    p0_new(:,0:)
     real(kind=dp_t), intent(in   ) :: gamma1bar(:,0:)
+    real(kind=dp_t), intent(  out) :: gamma1bar_avg(:,0:)
     type(multifab) , intent(in   ) :: s(:)
[…]

### 2009-05-03 docs/flowchart/flowchart.tex
added more necessary details
updating the language so it's more journal friendly

@@
 First use a predictor-corrector formulation for the pressure update 
-in order to improve the overall accuracy.  Beginning with equation~(\ref{eq:w01}),
+in order to improve the overall accuracy.  Beginning with equation (\ref{eq:w01}),
 we can write
@@
 Now we begin the corrector step.  Using $p_0^\star$ and $\gammabar^\star$, 
-we can compute a more accurate approximation to equation~(\ref{eq:frhs}),
+we can compute a more accurate approximation to equation (\ref{eq:frhs}),
 \begin{equation}
@@
 problem is already properly initialized.  We describe the details of
-the initialization in \S~\ref{Sec:Initialization}.
+the initialization in \S \ref{Sec:Initialization}.
 The advance of the state through a single timestep appears as:
@@
 
-The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
-We note that  $\uadvone$ satisfies the discrete versions of paper III equation (21),
+We construct time-centered velocities at edges, $\uadvone$, 
+using a second-order unsplit Godunov procedure, as described
+in Appendix B of paper III.  We note that  $\uadvone$ satisfies 
+the discrete versions of paper III equation (21),
 \begin{equation}
@@
   \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
-  for the conservative update of $(\rho X_k)$. 
-  Here we predict $\rho'$ and $X_k$ to the edges to obtain 
+  for the conservative update of $(\rho X_k)^{(1)}$. 
+  Here we predict $\rho^{'(1)} = \rho^{(1)} - \rhozero^n$ and 
+  $X_k^{(1)} = (\rho X_k)^{(1)} / \rho^{(1)}$ to time-centered edges to obtain 
   $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
-  Taylor expansion in space and time, as described in paper II,
+  unsplit Godunov procedure, as described in paper II,
   Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  We use the base
@@
   \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph,\star},$
-    for the conservative update of $(\rho h).$  Here we predict $(\rho
-    h)'$ to the edges to obtain $(\rho h)^{'(1),\nph,\star}$, 
-    using a second-order Taylor expansion in space
-    and time, as described in paper II, Appendix A, using $\Vb =
+    for the conservative update of $(\rho h)^{(1)}.$  Here we predict 
+    $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0$ to time-centered edges to obtain 
+    $(\rho h)^{'(1),\nph,\star}$, 
+    using a second-order unsplit Godunov procedure, 
+    as described in paper II, Appendix A, using $\Vb =
     \uadvone+w_0^{\nph,\star} \er$.  We do not include the reaction
@@
   \item Compute the species edge states, $(\rho X_k)^{(1),\nph}$, for
-    the conservative update of $(\rho X_k)$.  Here we predict $\rho'$ 
-    and $X_k$ to the edges to obtain $\rho^{'(1),\nph}$ and 
-    $X_k^{(1),\nph}$ using a second-order Taylor expansion in
-    space and time, as described in paper II, Appendix A, using $\Vb =
-    \uadvtwo+w_0^{\nph} \er$.  We use the base
+    the conservative update of $(\rho X_k)^{(1)}$.  Here we predict 
+    $\rho^{'(1)} = \rho^{(1)} - \rhozero^n$
+    and $X_k^{(1)} = (\rho X_k)^{(1)} / \rho^{(1)}$ to time-centered edges to 
+    obtain $\rho^{'(1),\nph}$ and $X_k^{(1),\nph}$ using a second-order 
+    unsplit Godunov procedure, as described in paper II, Appendix A, 
+    using $\Vb = \uadvtwo+w_0^{\nph} \er$.  We use the base
     state density, $\rho_0$, to convert this to an edge state for $\rho$,
@@
   \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, for
-    the conservative update of $(\rho h)$. Here we predict $(\rho h)'$
-    to the edges to obtain $(\rho h)^{'(1),\nph}$, 
-    using a second-order Taylor expansion in space and
-    time, as described in paper II, Appendix A, using $\Vb =
+    the conservative update of $(\rho h)^{(1)}$. Here we predict 
+    $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0$
+    to time-centered edges to obtain $(\rho h)^{'(1),\nph}$, 
+    using a second-order unsplit Godunov procedure, 
+    as described in paper II, Appendix A, using $\Vb =
     \uadvtwo+w_0^{\nph} \er$.  Again, we do not include the reaction
@@
 The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \Ubt ]^\nph$
-is described in paper II, Appendix A, with $\Vb = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
-component of $\Ubt^n$ individually.
+is described in paper II, Appendix A, with $\Vb = \uadvtwo+w_0^{\nph} \er$ and $s$ 
+set to each component of $\Ubt^n$ individually.  Here, the $\star$ superscript 
+does not refer to the predictor-corrector aspect of the algorithm, but rather 
+than the updated velocity does not satisfy the divergence constraint.
 

### 2009-05-04 docs/flowchart/flowchart.tex
more updates, mostly language

@@
 error that accumulates, causing the temperature computed from the EOS to drift.
+However, using {\bf Advect Base Pressure}, the latest version of {\bf test\_basestate}
+does not work after some bug fixes.
 
@@
 
-\item[Step 1.] {\em Define the average expansion at time $t^\nph$ and the new $w_0.$}
+\item[Step 1.] {\em Define the provisional time-centered expansion, $S^{n+\myhalf,\star}$, 
+base state velocity, $w_0^{n+\myhalf,\star}$, and base state velocity forcing.}
 
@@
 %--------------------------------------------------------------------------
-\item[Step 2.] {\em Construct the provisional edge-based advective velocity}, $\uadvone$.\\
+\item[Step 2.] {\em Construct the provisional time-centered advective velocity on 
+edges, $\uadvone$.}
 
@@
 %--------------------------------------------------------------------------
-\item[Step 6.] {\em Define a new average expansion rate at time $t^\nph.$}
+\item[Step 6.] {\em Define the final time-centered expansion, $S^{n+\myhalf}$, 
+base state velocity, $w_0^{n+\myhalf}$, and base state velocity forcing.}
 
@@
 %--------------------------------------------------------------------------
-\item[Step 7.] {\em Construct the final edge-based advective velocity}, $\uadvtwo$.
+\item[Step 7.] {\em Construct the final time-centered advective velocity on 
+edges, $\uadvtwo$.}
 
@@
 time advancement algorithm as described in steps 1-11 in 
-\S~\ref{Sec:Time Advancement Algorithm}.  
+\S \ref{Sec:Time Advancement Algorithm}.  
 The user specifies the number of iterations, 

### 2009-05-04 docs/flowchart/flowchart.tex
lots of changes, primarily, switched the advect_base_pres stuff and put in the enforce_HSE method

@@
 
-\item For any quantity, $\phi,$ we define $\overline{\phi} = ${\bf Avg}$(\phi),$ 
-the average over $\Omega_H,$ as
+\item For any quantity, $\phi$, we define $\overline{\phi} = ${\bf Avg}$(\phi)$
+as the average over a layer at constant radius r, $\Omega_H$, as
 \begin{equation}
@@
 \end{equation}
-here $\Omega_H$ is a layer at constant radius $r$.  In plane-parallel, this is 
-as straightforward as averaging cell values at a particular height.  For spherical 
-problems, see the description in Section \ref{Sec:Avg}.
-
-\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
+In plane-parallel, this is as straightforward as averaging cell values at 
+a particular height.  For spherical problems, see the description in 
+paper IV, section 2.2 (or section \ref{Sec:Avg}).
+
+\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}, 
+p_0^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, 
+(\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
 is the process by which we evolve the species and enthalpy from 
@@
 using the stiff ordinary differential equation integration methods
-provided by the {\tt VODE} package (cite VODE).  The absolute error
+provided by the {\tt VODE} package (cite vode).  The absolute error
 tolerances are set to $10^{-12}$ for the species, and a relative
@@
 \begin{equation}
-(\rho\Hnuc)^{\outp} = -\sum_k(\rho\omegadot_k)^{\outp} \; q_k \enskip.
+(\rho\Hnuc)^{\outp} = -\sum_k q_k (\rho\omegadot_k)^{\outp} \enskip,
 \end{equation}
@@
 $\Hnuc$ is an output of the reaction network, more general networks
-(i.e.\ involving weak interactions) can easily be added.  The enthalpy
+(i.e., involving weak interactions) can easily be added.  The enthalpy
 update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is
@@
 is the process by which we update the base state density through $\dt$ in time
-as well as compute time-centered interface states for use later in the algorithm.
+as well as compute time-centered interface states, $\rho_0^{\outp,n+\myhalf}$, 
+for use later in the algorithm.
 \begin{description}
@@
 \item[plane-parallel:] We discretize the base state density evolution 
-equation from paper III equation (29), without the $\etarho$ term,
+equation from paper III equation (29) without the $\etarho$ term,
 \begin{equation}
@@
 where $j$ refers to the one-dimensional index in the radial direction.
-The interface states, $\rho_0^{\outp,\nph}$, are found via the
+The interface states, $\rho_0^{\outp,\nph}$, are computed using the
 procedure described in paper II, Appendix A.\\
@@
 \end{description}
-\item {\bf Advect Base Pressure}
-is the process by which we update the base state pressure through $\dt$ in time and
-compute $\psi$.  The input arguments are different for plane-parallel and spherical.
-\begin{description}
-\item {\bf Advect Base Pressure Planar}$[p_0^\inp,w_0^\inp,\etarho^\inp] \rightarrow 
-[p_0^\outp,\psi^\outp]$:
-
-First we compute a cell-centered $\psi^\outp$ using
-\begin{equation}
- \psi_j^\outp = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^\inp + \eta_{\rho,j+\myhalf}^\inp\right) g \enskip .
-\end{equation}
-Then, we discretize the base state pressure evolution equation from paper III equation (31),
-\begin{equation}
-\frac{\partial p_0}{\partial t} = 
--w_0 \frac{\partial p_0}{\partial r} + \psi \enskip ,
-\label{eq:p0upd_new}
-\end{equation}
-to compute the new base state pressure,
-\begin{equation}
-p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
-+ w_{0,j-\myhalf}^\inp \right)
-\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\outp 
-\enskip,
-\end{equation}
-where the interface states are again found via the procedure described
-in paper II, Appendix A.\\
-
-\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},
-\Sbar^{\inp}, \rho^\inp, X_k^\inp] \rightarrow [p_0^{\outp},\psi^{n,\outp},\psi^{n+\myhalf,\outp}]$:
-
-First use a predictor-corrector formulation for the pressure update 
-in order to improve the overall accuracy.  Beginning with equation (\ref{eq:w01}),
-we can write
-\begin{equation}
-\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
- \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) - \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
-\end{equation}
-We time-center the left hand side:
-\begin{equation}
-\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
-   \frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} - p_0^{\inp}}{\dt}\enskip.
-\end{equation}
-We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
-\begin{equation}
-p_0^{\star} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
-\end{equation}
-with:
-\begin{equation}
-f = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
-                                             \frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} \left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
-\end{equation}
-where the $(\partial p_0^\inp / \partial r)$ is computed by
-interpolating $p_0^\inp$ from cell-centers to edges and then
-differencing.  Next, we compute a provisional updated $\gammabar$ using the
-provisional updated pressure,
-$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^\star,X_k^{\inp})]$,
-where $\rho^\inp$ and $X_k^\inp$ are time-advanced versions of the $\rho$ and
-$X_k$ which were used to compute $\gammabar^\inp$.\\
-
-Now we begin the corrector step.  Using $p_0^\star$ and $\gammabar^\star$, 
-we can compute a more accurate approximation to equation (\ref{eq:frhs}),
-\begin{equation}
-f^\dagger = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
-                                             \frac{1}{\gammabar^{\dagger} p_0^\dagger} w_0^{\inp} \left ( \frac{\partial p_0^\dagger}{\partial r} \right ) \right ] \enskip ,
-\end{equation}
-where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$ and
-$\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$.
-This allows us to complete the corrector for the pressure update:
-\begin{equation}
-p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\dagger} f^\dagger\right)}{\left(1 - \gammabar^{\dagger} f^\dagger\right)}\enskip .
-\end{equation}
-Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.
-First, we compute an updated $\gammabar$ using the update pressure, $\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^{\outp},X_k^{\inp})]$.
-Then, rewriting equation (\ref{eq:w01}), we can define
-\begin{eqnarray}
-\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
-       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
-\end{eqnarray}
-We discretize this locally as:
-\begin{equation}
-\psi_j^{\outp,n} = 
-\left(\gammabar^\inp p_0^\inp\right)_j
-\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
-          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip ,
-\end{equation}
-\begin{equation}
-\psi_j^{\outp,n+\myhalf} = 
-\left(\frac{\gammabar^\inp+\gammabar^{\star}}{2}\right)_j
-\left(\frac{p_0^\inp+p_0^\outp}{2}\right)_j
-\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
-          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip .
-\end{equation}
-\end{description}
+%\item {\bf Advect Base Pressure}
+%is the process by which we update the base state pressure through $\dt$ in time and
+%compute $\psi$.  The input arguments are different for plane-parallel and spherical.
+%\begin{description}
+%\item {\bf Advect Base Pressure Planar}$[p_0^\inp,w_0^\inp,\etarho^\inp] \rightarrow 
+%[p_0^\outp,\psi^\outp]$:
+%
+%First we compute a cell-centered $\psi^\outp$ using
+%\begin{equation}
+%\psi_j^\outp = 
+%\frac{1}{2} \left(\eta_{\rho,j-\myhalf}^\inp + \eta_{\rho,j+\myhalf}^\inp\right) g \enskip .
+%\end{equation}
+%Then, we discretize the base state pressure evolution equation from paper III equation (31),
+%\begin{equation}
+%\frac{\partial p_0}{\partial t} = 
+%-w_0 \frac{\partial p_0}{\partial r} + \psi \enskip , \label{eq:p0upd_new}
+%\end{equation}
+%to compute the new base state pressure,
+%\begin{equation}
+%p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+%+ w_{0,j-\myhalf}^\inp \right)
+%\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) 
+%+ \dt \; \psi_j^\outp \enskip,
+%\end{equation}
+%where the interface states, $p_0^{n+\myhalf}$, are again found via the procedure 
+%described in paper II, Appendix A.\\
+%
+%\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},
+%\Sbar^{\inp}, \rho^\inp, X_k^\inp] \rightarrow [p_0^{\outp},\psi^{n,\outp}, 
+%\psi^{n+\myhalf,\outp}]$:
+%
+%First use a predictor-corrector formulation for the pressure update in order to improve 
+%the overall accuracy.  Beginning with equation (\ref{eq:w01}), we can write
+%\begin{equation}
+%\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
+% \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) 
+%- \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
+%\end{equation}
+%We time-center the left hand side:
+%\begin{equation}
+%\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
+%\frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} 
+%- p_0^{\inp}}{\dt}\enskip.
+%\end{equation}
+%We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
+%\begin{equation}
+%p_0^{\star} = p_0^{\inp} 
+%\frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
+%\end{equation}
+%with:
+%\begin{equation}
+%f = \frac{\dt}{2} \left [ \Sbar^{\inp} - 
+%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
+%\frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} 
+%\left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
+%\end{equation}
+%where the $(\partial p_0^\inp / \partial r)$ is computed by interpolating $p_0^\inp$ from 
+%cell-centers to edges and then differencing.  Next, we compute a provisional updated 
+%$\gammabar$ using the provisional updated pressure,
+%$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^\star,X_k^{\inp})]$,
+%where $\rho^\inp$ and $X_k^\inp$ are time-advanced versions of the $\rho$ and
+%$X_k$ which were used to compute $\gammabar^\inp$.\\
+%
+%Now we begin the corrector step.  Using $p_0^\star$ and $\gammabar^\star$, 
+%we can compute a more accurate approximation to equation (\ref{eq:frhs}),
+%\begin{equation}
+%f^\dagger = \frac{\dt}{2} \left [ \Sbar^{\inp} -
+%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
+%\frac{1}{\gammabar^{\dagger} p_0^\dagger} w_0^{\inp} 
+%\left ( \frac{\partial p_0^\dagger}{\partial r} \right ) \right ] \enskip ,
+%\end{equation}
+%where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$ and
+%$\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$.
+%This allows us to complete the corrector for the pressure update:
+%\begin{equation}
+%p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\dagger} f^\dagger\right)}{\left(1 - \gammabar^{\dagger} f^\dagger\right)}\enskip .
+%\end{equation}
+%Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.  First, we compute an 
+%updated $\gammabar$ using the update pressure, 
+%$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^{\outp},X_k^{\inp})]$.  Then, 
+%rewriting equation (\ref{eq:w01}), we can define
+%\begin{eqnarray}
+%\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
+%       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
+%\end{eqnarray}
+%We discretize this locally as:
+%\begin{equation}
+%\psi_j^{\outp,n} = 
+%\left(\gammabar^\inp p_0^\inp\right)_j
+%\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
+%          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip ,
+%\end{equation}
+%\begin{equation}
+%\psi_j^{\outp,n+\myhalf} = 
+%\left(\frac{\gammabar^\inp+\gammabar^{\star}}{2}\right)_j
+%\left(\frac{p_0^\inp+p_0^\outp}{2}\right)_j
+%\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
+%          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip .
+%\end{equation}
+%\end{description}
+\item {\bf Enforce HSE}$[p_0^{\inp},\rho_0^{\inp},g^{\inp}] \rightarrow [p_0^{\outp}]$
+has replaced {\bf Advect Base Pressure} from paper III as the process by which we 
+update the base state pressure.  Rather than discretize the evolution equation
+for $p_0$, we enforce hydrostatic equilibrium directly, which is simpler and
+analytically equivalent.  We update $p_0$ using:
+\begin{equation}
+p_{0,j}^{\outp} = p_{0,j-1}^{\outp} + 
+\frac{\Delta r}{2}\left(\rho_{0,j}^{inp}+\rho_{0,j-1}^{\inp}\right)g_{j-\myhalf}^{\inp}\enskip .
+\end{equation}
+Once the density is less than the cutoff density, we set $p_{0,j}^\outp = p_{0,j-1}^\outp$.
+We use the boundary condition, $p_{0,jmax}^\outp = p_{0,jmax}^\inp$.
 
@@
 \end{equation}
-where the interface states are again found via the procedure described
-in paper II, Appendix A.\\
+where the interface states $(\rho h)_0^{n+\myhalf}$, are again found via the procedure 
+described in paper II, Appendix A.\\
 
@@
 \end{equation}
-The other difference is that the interface states are computed using 
-$\psi^{\inp,n}$ as the forcing.
+To compute the interface states, we use $\psi^{\inp,n}$ as the forcing term, whereas in the
+enthalpy update, we use $\psi^{\inp,n+\myhalf}$ as the forcing.
 
@@
 is the process by which we adjust the base state density to account for
-large-scale mixing.  It is also used to adjust the base state enthalpy
-to account for large-scale mixing, reactions, and thermal diffusion.
+large-scale mixing.  It is also used to reset the base state enthalpy to be
+the average of the full state enthalpy since we do not need to compute the complete
+evolution of the base state enthalpy.
 Unlike paper III where we used $\nabla\cdot\etarho$ as an adjustment to $\rho_0$, 
@@
 \left[ \left( r^2 \etarho^{\inp} \right)_{j} - \left( r^2 \etarho^{\inp}  \right)_{j-1} \right] 
-- \left(  4 \pi G \rhozero^{\inp} \etarho^{\inp}   \right)_{j-\myhalf} \enskip \nonumber .
+\nonumber \\
+&& - \left(  4 \pi G \rhozero^{\inp} \etarho^{\inp}   \right)_{j-\myhalf} \enskip \nonumber .
 \end{eqnarray}
@@
 \begin{description}
-\item For plane-parallel, we hold the gravity constant in space and time.
+\item For plane-parallel, gravity constant in space and time.
 \item For spherical, we define the enclosed mass at edges, and then cell-centers:
@@
 \begin{equation}
-g_{j-\myhalf} = \frac{Gm_{j-\myhalf}}{r_{j-\myhalf}^2}; ~~~~~ g_j = \frac{Gm_j}{r_j^2}\enskip .
+g_{j-\myhalf}^\outp = \frac{Gm_{j-\myhalf}}{r_{j-\myhalf}^2}; ~~~~~ 
+g_j^\outp = \frac{Gm_j}{r_j^2}\enskip .
 \end{equation}
@@
 %--------------------------------------------------------------------------
-\item[Step 3.] {\em React the full state and base state through the first time 
-interval of $\dt / 2.$}
+\item[Step 3.] {\em React the full state through the first time interval of $\dt / 2.$}
 
@@
 
-\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n]
+\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n]
                    \rightarrow$ $[ \rho^{(1)}, (\rho h)^{(1)}, X_k^{(1)}, T^{(1)},
@@
   radial faces.  In spherical, we first map $\rho_0$ to Cartesian cell-centers 
-  using the mapping in Section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
+  using the mapping in section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
   and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.
@@
 
-\item Update $p_0$ and compute $\psi$.
+%{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
+%\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
+%\item For spherical,\\
+%{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
+%\overline{S^{\nph,\star}}, \rho^{(2),\star}, X_k^{(2),\star}]$\\
+%$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+%\rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
+\item {\bf Enforce HSE}$[p_0^n,\rho_0^{n+1,\star},g^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
+
+\item Compute $\psi$.
 \begin{description}
-\item For plane-parallel,\\
-{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
-\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
-\item For spherical,\\
-{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
-\overline{S^{\nph,\star}}, \rho^{(2),\star}, X_k^{(2),\star}]$\\
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-\rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
-
+\item For plane-parallel, 
+\begin{equation}
+\psi_j^{n+\myhalf,\star} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{n+\myhalf,\star} 
++ \eta_{\rho,j+\myhalf}^{n+\myhalf,\star}\right) g^{n+1,\star} \enskip .
+\end{equation}
+\item For spherical, first compute:
+\begin{equation}
+\gammabar^{(2)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2)}, p_0^{n+1,\star}, 
+X_k^{(2)}\right) \right] \enskip .
+\end{equation}
+Then, define a base time and time-centered $\psi$:
+\begin{equation}
+\psi_j^{n,\star} = 
+\left(\gammabar^{(1)} p_0^n\right)_j
+\left \{ \Sbar_j^{n+\myhalf,\star} - 
+\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
+          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip ,
+\end{equation}
+\begin{eqnarray}
+\psi_j^{n+\myhalf,\star} &=& 
+\left(\frac{\gammabar^{(1)}+\gammabar^{(2)}}{2}\right)_j
+\left(\frac{p_0^n+p_0^{n+1,\star}}{2}\right)_j \nonumber \\
+&& \left \{ \Sbar_j^{n+\myhalf,\star} - 
+\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
+          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip .
+\end{eqnarray}
 \end{description}
@@
   on radial faces.  In spherical, we first map $h_0 \equiv (\rho h)_0/\rho_0$ 
-  and $\rho_0$ to Cartesian cell-centers using the mapping in Section 
+  and $\rho_0$ to Cartesian cell-centers using the mapping in section 
   \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, use 2nd-order spatial 
@@
 \item {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, 
-                             T^{(2),\star}, (\rho\Hext)^{(2),\star}] $\\
+                             T^{(2),\star}, (\rho\Hext)^{(2),\star}, p_0^{n+1,\star}] $\\
 $~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, 
@@
 
-\item Update $p_0$ and compute $\psi$.
+%\item Update $p_0$ and compute $\psi$.
+%\begin{description}
+%\item For plane-parallel,\\
+%{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
+%$\rightarrow [p_0^{n+1},\psi^{\nph}]$.
+%\item For spherical,\\
+%{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{(1)},\overline{S^{\nph}},
+%\rho^{(2)},X_k^{(2)}]$\\
+%$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+% \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
+%\end{description}
+\item {\bf Enforce HSE}$[p_0^n,\rho_0^{n+1},g^{n+1}] \rightarrow [p_0^{n+1}]$.
+
+\item Compute $\psi$.
 \begin{description}
-\item For plane-parallel,\\
-{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
-$\rightarrow [p_0^{n+1},\psi^{\nph}]$.
-\item For spherical,\\
-{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{(1)},\overline{S^{\nph}},
-\rho^{(2)},X_k^{(2)}]$\\
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
- \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
+\item For plane-parallel, 
+\begin{equation}
+\psi_j^{n+\myhalf} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{n+\myhalf} 
++ \eta_{\rho,j+\myhalf}^{n+\myhalf}\right) g^{n+1} \enskip .
+\end{equation}
+\item For spherical, first compute:
+\begin{equation}
+\gammabar^{(2),\star} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2),\star}, p_0^{n+1}, 
+X_k^{(2),\star}\right) \right] \enskip .
+\end{equation}
+Then, define a base time and time-centered $\psi$:
+\begin{equation}
+\psi_j^{n} = 
+\left(\gammabar^{(1)} p_0^n\right)_j
+\left \{ \Sbar_j^{n+\myhalf} - 
+\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf})_{j+\myhalf} -
+          (r^2 w_0^{n+\myhalf})_{j-\myhalf} \right ] \right \} \enskip ,
+\end{equation}
+\begin{eqnarray}
+\psi_j^{n+\myhalf} &=& 
+\left(\frac{\gammabar^{(1)}+\gammabar^{(2),\star}}{2}\right)_j
+\left(\frac{p_0^n+p_0^{n+1}}{2}\right)_j \nonumber \\
+&& \left \{ \Sbar_j^{n+\myhalf} - 
+\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf})_{j+\myhalf} -
+          (r^2 w_0^{n+\myhalf})_{j-\myhalf} \right ] \right \} \enskip .
+\end{eqnarray}
 \end{description}
@@
 %--------------------------------------------------------------------------
-\item[Step 9.] {\em React the full state and base state through a second time 
-interval of $\dt / 2.$}
+\item[Step 9.] {\em React the full state through a second time interval of $\dt / 2.$}
 
@@
 
-\item {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho\Hext)^{(2)} ]$\\
+\item {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho\Hext)^{(2)},
+p_0^{n+1}]$\\
 $~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\rightarrow [\rho^{n+1}, 
@@
 
-  \item {\bf React State}$[ \rho^{\initp},(\rho h)^{\initp}, X_k^{\initp}, T^{\initp}, (\rho^{\initp} \Hext) ] \rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{0,\nu} ].$
+  \item {\bf React State}$[ \rho^{\initp},(\rho h)^{\initp}, X_k^{\initp}, T^{\initp}, 
+(\rho^{\initp} \Hext), p_0^{\initp}] \rightarrow [\rho^{\outp}, (\rho h)^{\outp}, 
+X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{0,\nu} ].$
 

### 2009-05-04 Source/advance.f90
setting (rho h)_0^n = avg((rho h)^{(1)}) in step 4 rather than step 3 - makes more sense in the flow of the alg.  no changes to results

@@ -385,18 +385,6 @@ contains
        call destroy(rho_Hext(n))
     end do
 
-    ! compute rhoh0_old by "averaging"
-    if (evolve_base_state) then
-       if (spherical .eq. 0) then
-          call average(mla,s1,rhoh0_old,dx,rhoh_comp)
-       else
-          ! set rhoh0_old = rhoh0_old - Avg(rhoh0_old - rhoh^{(1)})
-          call make_sprimebar_spherical(s1,rhoh_comp,rhoh0_old,dx,rhohprimebar,mla, &
-                                        the_bc_tower%bc_tower_array)
-          call correct_base(rhoh0_old,rhohprimebar)
-       end if
-    end if
-
     if (evolve_base_state) then
        do n=1,nlevs
           call multifab_build(gamma1(n), mla%la(n), 1, 0)
@@ -557,6 +545,17 @@ contains
     end if
 
     if (evolve_base_state) then
+
+       ! compute rhoh0_old by "averaging"
+       if (spherical .eq. 0) then
+          call average(mla,s1,rhoh0_old,dx,rhoh_comp)
+       else
+          ! set rhoh0_old = rhoh0_old - Avg(rhoh0_old - rhoh^{(1)})
+          call make_sprimebar_spherical(s1,rhoh_comp,rhoh0_old,dx,rhohprimebar,mla, &
+                                        the_bc_tower%bc_tower_array)
+          call correct_base(rhoh0_old,rhohprimebar)
+       end if
+
        call advect_base_enthalpy(w0,rho0_old,rhoh0_old,rhoh0_new, &
                                  rho0_predicted_edge,psi,psi_old,dt)
     else

### 2009-05-04 Source/advance.f90
another simplification - moved computation of gamma1bar^{(1)} from step 3 into step 4, and only do it if you are in spherical since it's not used in plane-parallel.  flow chart reads better.  no changes to results

@@ -385,19 +385,6 @@ contains
        call destroy(rho_Hext(n))
     end do
 
-    if (evolve_base_state) then
-       do n=1,nlevs
-          call multifab_build(gamma1(n), mla%la(n), 1, 0)
-       end do
-       
-       call make_gamma(mla,gamma1,s1,p0_old,dx)
-       call average(mla,gamma1,gamma1bar_1,dx,1)
-
-       do n=1,nlevs
-          call destroy(gamma1(n))
-       end do
-    end if
-
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
 !! STEP 4 -- advect the base state and full state through dt
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
@@ -499,6 +486,19 @@ contains
     call make_grav_cell(grav_cell_new,rho0_new)
 
     if (evolve_base_state) then
+
+       if (spherical .eq. 1) then
+          do n=1,nlevs
+             call multifab_build(gamma1(n), mla%la(n), 1, 0)
+          end do
+          
+          call make_gamma(mla,gamma1,s1,p0_old,dx)
+          call average(mla,gamma1,gamma1bar_1,dx,1)
+          
+          do n=1,nlevs
+             call destroy(gamma1(n))
+          end do
+       end if
 
        if (p0_update_type .eq. 1) then
 

### 2009-05-04 docs/flowchart/flowchart.tex
setting (rho h)_0^n = avg((rho h)^{(1)}) in step 4 rather than step 3 - makes more sense in the flow of the alg.  no changes to results
another simplification - moved computation of gamma1bar^{(1)} from step 3 into step 4, and only do it if you are in spherical since it's not used in plane-parallel.  flow chart reads better.  no changes to results
minor

@@
 \begin{equation}
-\frac{\partial\Ubt}{\partial t} = - \Ubt \cdotb \nablab \Ubt - w_0 \frac{\partial \Ubt}{\partial r}
+\frac{\partial\Ubt}{\partial t} = 
+- \left(\Ubt+w_0\right) \cdotb \nablab \Ubt
 - \left(\Ubt \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
@@
 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
-                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip\
- .
+- \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip .
 \end{equation}
@@
 We construct time-centered velocities at edges, $\uadvone$, 
-using a second-order unsplit Godunov procedure, as described
-in Appendix B of paper III.  We note that  $\uadvone$ satisfies 
-the discrete versions of paper III equation (21),
+using a second-order unsplit Godunov procedure followed by an elliptic
+solve, as described in Appendix B of paper III.  We note that 
+$\uadvone$ satisfies the discrete versions of paper III equation (21),
 \begin{equation}
@@
 
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n]
-                   \rightarrow$ $[ \rho^{(1)}, (\rho h)^{(1)}, X_k^{(1)}, T^{(1)},
-                                  (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)} ].$
-
-\item {\bf Correct Base}$[(\rho h)_0^{n},(\rho h)^{(1)}] \rightarrow [(\rho h)_0^{n}]$.
-
-\item Define
-\begin{equation}
-\gammabar^{(1)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, X_k^{(1)}\right) \right] \enskip .
-\end{equation}
-
-\end{enumerate}
+{\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n] \rightarrow
+[ \rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)} ]$.
 
@@
 \begin{equation}
-\gammabar^{(2)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2)}, p_0^{n+1,\star}, 
-X_k^{(2)}\right) \right] \enskip .
+\gammabar^{(1)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, 
+X_k^{(1)}\right) \right] \enskip ,
+\end{equation}
+\begin{equation}
+\gammabar^{(2),\star} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2),\star}, p_0^{n+1,\star}, 
+X_k^{(2),\star}\right) \right] \enskip .
 \end{equation}
@@
 \psi_j^{n+\myhalf,\star} &=& 
-\left(\frac{\gammabar^{(1)}+\gammabar^{(2)}}{2}\right)_j
+\left(\frac{\gammabar^{(1)}+\gammabar^{(2),\star}}{2}\right)_j
 \left(\frac{p_0^n+p_0^{n+1,\star}}{2}\right)_j \nonumber \\
@@
 \end{description}
+
+\item {\bf Correct Base}$[(\rho h)_0^{n},(\rho h)^{(1)}] \rightarrow [(\rho h)_0^{n}]$.
 
@@
 \begin{equation}
-\gammabar^{(2),\star} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2),\star}, p_0^{n+1}, 
-X_k^{(2),\star}\right) \right] \enskip .
+\gammabar^{(2)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2)}, p_0^{n+1}, 
+X_k^{(2)}\right) \right] \enskip .
 \end{equation}
@@
 \psi_j^{n+\myhalf} &=& 
-\left(\frac{\gammabar^{(1)}+\gammabar^{(2),\star}}{2}\right)_j
+\left(\frac{\gammabar^{(1)}+\gammabar^{(2)}}{2}\right)_j
 \left(\frac{p_0^n+p_0^{n+1}}{2}\right)_j \nonumber \\

### 2009-05-05 docs/flowchart/flowchart.tex
more cleanup

@@
 p_0^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, 
-(\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
+(\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$
 is the process by which we evolve the species and enthalpy from 
@@
 \item {\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp]$
-$\rightarrow [\rho_0^\outp,\rho_0^{\outp,n+\myhalf}]$\\
+$\rightarrow [\rho_0^\outp,\rho_0^{\outp,n+\myhalf}]$
 is the process by which we update the base state density through $\dt$ in time
@@
 p_{0,j}^{\outp} = p_{0,j-1}^{\outp} + 
-\frac{\Delta r}{2}\left(\rho_{0,j}^{inp}+\rho_{0,j-1}^{\inp}\right)g_{j-\myhalf}^{\inp}\enskip .
+\frac{\Delta r}{2}\left(\rho_{0,j}^{\inp}+\rho_{0,j-1}^{\inp}\right)g_{j-\myhalf}^{\inp}\enskip .
 \end{equation}
@@
  - \frac{1}{r_j^2} \frac{\dt}{\dr} 
- \left \{ \left[ r^2 (\rho h)_0^{\outp} w_0^{\inp}\right]_{j+\myhalf} -  
-          \left[ r^2 (\rho h)_0^{\outp} w_0^{\inp}\right]_{j-\myhalf} \right\}
+ \left \{ \left[ r^2 (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} -  
+          \left[ r^2 (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\}
 +\Delta t\psi^{\inp,n+\myhalf} \enskip. 
 \end{equation}
-To compute the interface states, we use $\psi^{\inp,n}$ as the forcing term, whereas in the
-enthalpy update, we use $\psi^{\inp,n+\myhalf}$ as the forcing.
+To compute the interface states, we use $\psi^{\inp,n}$ as the forcing term, 
+whereas in the enthalpy update, we use $\psi^{\inp,n+\myhalf}$ as the forcing.
 
@@
 
-\item {\bf Correct Base}$[\rho_0^{\inp}, \rho^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
+\item {\bf Correct Base}$[\rho_0^{\inp}, \rho^{\inp}] \rightarrow [\rho_0^{\outp}]$
 is the process by which we adjust the base state density to account for
@@
 
-\item {\bf Compute $w_0$}\\
-is the process by which we compute $w_0$.  The arguments are different for plane-parallel
-and spherical.
+\item {\bf Compute $w_0$} is the process by which we compute $w_0$.
+The arguments are different for plane-parallel and spherical.
 \begin{description}
@@
 \item[Step 1.] {\em Define the provisional time-centered expansion, $S^{n+\myhalf,\star}$, 
-base state velocity, $w_0^{n+\myhalf,\star}$, and base state velocity forcing.}
+provisional base state velocity, $w_0^{n+\myhalf,\star}$, and provisional base state velocity 
+forcing.}
 
@@
 Otherwise, following the method used in our small-scale low Mach number algorithm
-(BDRWZ), we extrapolate to the half-time using the source
+(cite BDRWZ), we extrapolate to the half-time using the source
 terms at the previous and current time levels
@@
 
-\item Define the scaled pressure gradient using
-\begin{equation}
-- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} 
+\item Define the base state velocity forcing using equation (38) from paper III,
+\begin{equation}
+-\frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} 
 = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} 
@@
 \begin{equation}
--\left ( \frac{1}{\rho_0^n} \frac{\partial \pizeroone}{\partial r} \right ) = 
-\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} 
-+  w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star} \enskip ,
-\end{equation} 
-where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are
+\left ( \frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} \right )^{n,\star} = 
+-\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} 
+- w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star} \enskip ,
+\end{equation} 
+where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are defined through 
+paper III equation (53),
 \begin{equation}
 w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}} 
-\enskip,\end{equation}
+\enskip,
+\end{equation}
 \begin{equation}
@@
 + \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] 
-\enskip.\end{equation}
+\enskip.
+\end{equation}
 If $n=0$, we use $\dt^{-1} = \dt^0$.
@@
 
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, 
-                             T^{(2),\star}, (\rho\Hext)^{(2),\star}, p_0^{n+1,\star}] $\\
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, 
+{\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, T^{(2),\star}, 
+(\rho\Hext)^{(2),\star}, p_0^{n+1,\star}] $\\
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, 
 X_k^{n+1,\star}, T^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
 
-\item Define
-\begin{eqnarray}
- \gammabar^{n+1,\star}    &=& {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{n+1,\star}, p_0^{n+1,\star}, 
-                                                      X_k^{n+1,\star}\right) \right] \enskip , \\
- \beta_0^{n+1,\star}    &=& \beta   \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}, g^{n+1,\star}\right) \enskip ,
-\end{eqnarray}
-where $\beta_0$ is computed as described in paper III, Appendix C.
-
-\end{enumerate}
-
 %--------------------------------------------------------------------------
@@
 %--------------------------------------------------------------------------
-\item[Step 6.] {\em Define the final time-centered expansion, $S^{n+\myhalf}$, 
-base state velocity, $w_0^{n+\myhalf}$, and base state velocity forcing.}
+\item[Step 6.] {\em Define the time-centered expansion, $S^{n+\myhalf}$, base state
+velocity, $w_0^{n+\myhalf}$, and base state velocity forcing.}
 
@@
   \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star}  
-  + \sigma \Hext + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T^{n+1,\star} \enskip .
+  + \sigma \Hext + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T^{n+1,\star} \enskip ,
 \end{equation} 
@@
 \rhozero^{\nph} = \frac{\rhozero^{n} + \rhozero^{n+1,\star}}{2}\enskip , ~~~
-p_0^{\nph} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip .
-\end{equation}
+p_0^{\nph} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip ,
+\end{equation}
+with
+\begin{equation}
+ \gammabar^{n+1,\star} = {\rm{\bf Avg}} 
+\left[ \Gamma_1\left(\rho^{n+1,\star}, p_0^{n+1,\star}, X_k^{n+1,\star}\right) \right] 
+\enskip .
+\end{equation}
+
 \item Compute $w_0^{n+\myhalf}$:
@@
 
-\item Using equation (\ref{eq:pizero}), define the scaled pressure gradient using the
-following discretization:
-\begin{equation}
--\left ( \frac{1}{\rho_0^n} \frac{\partial \pizerotwo}{\partial r} \right ) = 
-\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
-+ w_0^n \left(\frac{\partial w_0}{\partial r}\right)^n.
-\end{equation}
-where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined through paper III equation (53),
-\begin{equation}
-w_0^n = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
-\end{equation}
-and paper III equation (54),
+\item Define the base state velocity forcing using equation (\ref{eq:pizero}) 
+with the following discretization:
+\begin{equation}
+\left ( \frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} \right )^n = 
+-\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
+- w_0^n \left(\frac{\partial w_0}{\partial r}\right)^n,
+\end{equation}
+where $w_0^{n}$ and $(\partial w_0 / \partial r)^{n}$ are defined as
+\begin{equation}
+w_0^n = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph}}{\dt^n+\dt^{n-1}}\enskip,
+\end{equation}
 \begin{equation}
 \left(\frac{\partial w_0}{\partial r}\right)^{n} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
-+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph} \right ] \enskip.\label{eq:dw0drnstar}
-\end{equation}
++ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph} \right ] \enskip.
+\end{equation}
+If $n=0$, we use $\dt^{-1} = \dt^0$.
 
@@
 %--------------------------------------------------------------------------
-\item[Step 7.] {\em Construct the final time-centered advective velocity on 
-edges, $\uadvtwo$.}
+\item[Step 7.] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}
 
@@
 \begin{equation}
-\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2} \enskip .
-\end{equation}
+\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2}; ~~~~~
+ \beta_0^{n+1,\star} = \beta \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}, g^{n+1,\star}\right) \enskip ,
+\end{equation}
+where $\beta_0$ is computed as described in paper III, Appendix C.
+
 The discretization with the volume discrepancy correction is
@@
 
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho\Hext)^{(2)},
-p_0^{n+1}]$\\
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\rightarrow [\rho^{n+1}, 
+{\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho\Hext)^{(2)}, p_0^{n+1}]$\\
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\rightarrow [\rho^{n+1}, 
 (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$
 
-\item Define
-\begin{eqnarray}
- \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1}, X_k^{n+1}\right) \right] \enskip , \\ 
- \beta_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1}, \gammabar^{n+1}, g^{n+1}\right) \enskip ,
-\end{eqnarray}
-where $\beta_0$ is computed as described in paper III, Appendix C.
-
-\end{enumerate}
-
 %--------------------------------------------------------------------------
@@
 %--------------------------------------------------------------------------
-\item[Step 10.] {\em Compute $S^{n+1}$ for the final projection.}
+\item[Step 10.] {\em Define the new time expansion $S^{n+1}$, and $\gammabar^{n+1}$.}
 
@@
 \overline{S^{n+1}} = {\mathrm{\bf Avg}} (S^{n+1}) \enskip.
+\end{equation}
+
+\item Define
+\begin{equation}
+ \gammabar^{n+1} = {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1}, 
+X_k^{n+1}\right) \right] \enskip .
 \end{equation}
@@
 \begin{equation}
-\beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1}}{2}\enskip .
-\end{equation}
+\beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1}}{2}; ~~~~~ 
+\beta_0^{n+1} = \beta \left(\rho_0^{n+1}, p_0^{n+1}, \gammabar^{n+1}, g^{n+1}\right)\enskip .
+\end{equation}
+
 The discretization with the volume discrepancy term is:

### 2009-05-05 docs/flowchart/flowchart.tex
minor fix - source term for enthalpy is utilde dot grad p0 not u dot grad p0
updated to show that we like enforce_HSE rather than advect_base_pres, since test_basestate never actually worked with advect_base_pres after I fixed a bug.  You get negative pressure at the edge of the star after one call to advect_base_pres.  Full code still defaults to using advect_base_pres though.

@@
 \begin{enumerate}
-\item We have changed the discretization of $\Ub\cdot\nabla p_0$ in the enthalpy
-update to $\nabla\cdot(\Ub p_0) - p_0\nabla\cdot\Ub$.
+\item We have changed the discretization of $\Ubt\cdot\nabla p_0$ in the enthalpy
+update to $\nabla\cdot(\Ubt p_0) - p_0\nabla\cdot\Ubt$.
 \item In paper III we discretized the enthalpy evolution equation in
-full, non-perturbation form, $\rho h$.  Since then we have discovered that 
+terms of $T$.  Since then we have discovered that 
 discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
@@
 is in the main flow chart.
+\item We have eliminated {\bf Advect Base Pressure} entirely, and instead directly
+enforce HSE after {\bf Correct Base}.  {\bf Advect Base Pressure} was just a fancy
+way of enforcing HSE anyway, so we just compute the new pressure directly.  The
+problems we have seen with enforcing HSE directly may have to do with the roundoff
+error that accumulates, causing the temperature computed from the EOS to drift.
+However, using {\bf Advect Base Pressure}, the latest version of {\bf test\_basestate}
+generates negative pressures, so that is even worse.
 \item We replaced {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
@@
 \begin{itemize}
-
-\item Should we eliminate {\bf Advect Base Pressure} entirely, and instead directly
-enforce HSE after {\bf Correct Base}?  {\bf Advect Base Pressure} was just a fancy
-way of enforcing HSE anyway, so we just compute the new pressure directly.  The
-problems we have seen with enforcing HSE directly have to do with the roundoff
-error that accumulates, causing the temperature computed from the EOS to drift.
-However, using {\bf Advect Base Pressure}, the latest version of {\bf test\_basestate}
-does not work after some bug fixes.
 

### 2009-05-07 docs/flowchart/flowchart.tex
minor
don't need to add temporal location of g in planar since g is constant
fixed a notational inconsistency
more cleanup
a lot of wordiness in step 8 that doesn't need to be repeated since we say the exact same thing in step 4.  Step 8 now refers to step 4 in certain parts for explanations

@@
 \left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
-         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ]
-         \nonumber \enskip,
+         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ] \enskip,
 \end{equation}
@@
 
-\item {\bf Compute $w_0$} is the process by which we compute $w_0$.
+\item {\bf Compute} {\boldmath $w_0$} is the process by which we compute $w_0$.
 The arguments are different for plane-parallel and spherical.
 \begin{description}
-\item {\bf Compute $w_0$ Planar}$[\Sbar^{\inp},\gammabar^{\inp},p_0^{\inp},\psi^{\inp}] 
-\rightarrow [w_0^{\outp}]$:
+\item {\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\Sbar^{\inp},\gammabar^{\inp},p_0^{\inp},
+\psi^{\inp}] \rightarrow [w_0^{\outp}]$:
 
@@
 
-\item {\bf Compute $w_0$ Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},
+\item {\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},
 \etarho^{\inp}] \rightarrow[w_0^{\outp}]$:
@@
   g \left[ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right]
-+ \rho_0 \left( \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} \right) \enskip.\nonumber\\
-&&
++ \rho_0 \left( \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} \right) \enskip.
 \label{eq:sphconstraint_3}
@@
    - \left( \frac{4 \ow \rho_0 g}{r} \right)_{j-\myhalf}
-   - \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip , \nonumber
+   - \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip ,
 \end{eqnarray}
@@
 \begin{eqnarray}
-&& \frac{1}{\Delta r} \left\{
- \left( \frac{\gammabar p_0}{r^2}\right)_{j  } \frac{\left[ (r^2 \dw)_{j+\myhalf} - (r^2 \dw)_{j-\myhalf} \right]}{\Delta r}
--\left( \frac{\gammabar p_0}{r^2}\right)_{j-1} \frac{\left[ (r^2 \dw)_{j-\myhalf} - (r^2 \dw)_{j-\thalf}\right]}{\Delta r} \right\} \nonumber \\
-&&  - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \dw)_{j-\myhalf}  \nonumber \\
-&=&   \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
- - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
+& \frac{1}{\Delta r} \left\{
+ \left( \frac{\gammabar p_0}{r^2}\right)_{j  } \frac{\left[ (r^2 \dw)_{j+\myhalf} - (r^2 \dw)_{j-\myhalf} \right]}{\Delta r} -\left( \frac{\gammabar p_0}{r^2}\right)_{j-1} \frac{\left[ (r^2 \dw)_{j-\myhalf} - (r^2 \dw)_{j-\thalf}\right]}{\Delta r} \right\} & \nonumber \\
+&  - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \dw)_{j-\myhalf} = \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \ow)_{j-\myhalf} & \nonumber \\
+& - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
   \left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
-- \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip, \nonumber \\
+- \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip, &
 \end{eqnarray}
@@
 \begin{eqnarray}
-A_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1}  \enskip, \nonumber  \\
+A_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1}  \enskip, \\
 B_j &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j  }  
                +\left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1} \right] 
-              - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) \nonumber \\
-C_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j}  \enskip, \nonumber  \\
+              - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) \enskip , \\
+C_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j}  \enskip, \\
 F_j &=&  \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
@@
 \nonumber \\
-&& - \left(  4 \pi G \rhozero^{\inp} \etarho^{\inp}   \right)_{j-\myhalf} \enskip \nonumber .
+&& - \left(  4 \pi G \rhozero^{\inp} \etarho^{\inp}   \right)_{j-\myhalf} \enskip .
 \end{eqnarray}
@@
 A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
-B_1 &=& 1\enskip. \nonumber
+B_1 &=& 1\enskip.
 \end{eqnarray}
@@
 B_N &=&  1\enskip, \nonumber \\
-C_N = F_N &=& 0\enskip. \nonumber
+C_N = F_N &=& 0\enskip.
 \end{eqnarray}
@@
 \item For plane-parallel,\\
-{\bf Compute $w_0$ Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,\psi^{n-\myhalf}] 
-\rightarrow [w_0^{n+\myhalf,\star}]$.
+{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,
+\psi^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
 
@@
 \item For spherical,\\
-{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,\rhozero^n,p_0^n,
-\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
+{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,
+\rhozero^n,p_0^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
 \end{description}
@@
 
-\item Update the species.  Here we do a difference approximation
-  to 
-  \begin{equation}
-  \frac{\partial (\rho X_k)}{\partial t} + \nablab \cdotb (\Ub \rho X_k) =
-         \rho {\omegadot}_k \enskip , \label{eq:species}
-  \end{equation}
-  neglecting the reaction terms.  The update consists of two steps:
+\item \label{step:species update} 
+Update the species.  Here we consider only the advection terms, 
+neglecting the reaction terms, and solve a discretized version of
+\begin{equation}
+\frac{\partial (\rho X_k)}{\partial t} + \nablab \cdotb (\Ub \rho X_k) = 0 \enskip .
+\label{eq:species}
+\end{equation}
+The update consists of two steps:
 
@@
   unsplit Godunov procedure, as described in paper II,
-  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  We use the base
-  state density, $\rho_0$, to convert this to an edge state for $\rho$,
+  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  Convert the 
+  perturbational edge state to a full state using
 \begin{equation}
@@
 \end{equation}
-This gives a radial cell-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
-radial edges, average the two neighboring radial bins.
+This gives a cell-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
+radial edges, average the two neighboring radial cell-centers.
 \end{description}
@@
 \psi_j^{n+\myhalf,\star} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{n+\myhalf,\star} 
-+ \eta_{\rho,j+\myhalf}^{n+\myhalf,\star}\right) g^{n+1,\star} \enskip .
++ \eta_{\rho,j+\myhalf}^{n+\myhalf,\star}\right) g \enskip .
 \end{equation}
@@
 
-\item Update the enthalpy.  The full enthalpy equation is 
-  \begin{equation}
-  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
-+ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T + \rho \Hnuc + \rho \Hext 
-\enskip , \label{eq:rhohupd} 
-  \end{equation}
-Here we consider only the advection, neglecting the reaction terms and thermal diffusion terms,
-and solve a discretized version of
+\item Update the enthalpy.  Here we consider only the advective terms, 
+neglecting the reaction terms and thermal diffusion terms, and solve a discretized version of
   \begin{equation}
@@
   \end{equation}
-For spherical flows, experience has shown that solving the algebraically-equivalent form
+For spherical flows, experience has shown that solving the analytically equivalent form,
 \begin{equation}
@@
     terms in the enthalpy prediction, since we accounted for them
-    already in {\bf React State}.  We use the base state enthalpy,
-    $(\rho h)_0$, to convert this to an edge state for $(\rho h)$,
+    already in {\bf React State}.  Convert the perturbational edge
+    state to a full state using
 
@@
 $(\rho h)^{(2),\star}$ to account for thermal diffusion.  Here we begin
-with the enthalpy equation (\ref{eq:rhohupd}), but consider only the 
+with the enthalpy equation, but consider only the 
 diffusion terms,
@@
 \item For plane-parallel,\\
-{\bf Compute $w_0$ Planar}$[\overline{S^{\nph}},
+{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph}},
 \gammabar^{n+\myhalf},p_0^{n+\myhalf},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf}]$.
@@
 \item For spherical,\\
-{\bf Compute $w_0$ Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf},
+{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf},
 \rhozero^{n+\myhalf},p_0^{n+\myhalf},\etarho^{n+\myhalf}] \rightarrow [w_0^{n+\myhalf}]$.
@@
 
-\item Update the species.  Here we do a difference approximation to
-  equation (\ref{eq:species}), neglecting the reaction terms.  The
-  update consists of two steps:
+\item Update the species.  This step is identical to step 4b with the following changes:
 
@@
 
-  \item Compute the species edge states, $(\rho X_k)^{(1),\nph}$, for
-    the conservative update of $(\rho X_k)^{(1)}$.  Here we predict 
-    $\rho^{'(1)} = \rho^{(1)} - \rhozero^n$
-    and $X_k^{(1)} = (\rho X_k)^{(1)} / \rho^{(1)}$ to time-centered edges to 
-    obtain $\rho^{'(1),\nph}$ and $X_k^{(1),\nph}$ using a second-order 
-    unsplit Godunov procedure, as described in paper II, Appendix A, 
-    using $\Vb = \uadvtwo+w_0^{\nph} \er$.  We use the base
-    state density, $\rho_0$, to convert this to an edge state for $\rho$,
-
+\item Compute the species edge states, $(\rho X_k)^{(1),n+\myhalf}$, 
+using $\Vb = \uadvtwo+w_0^{\nph}\er$.  Convert the perturbational edge 
+state to a full state using 
 \begin{equation}
@@
 \end{equation}
-  In plane-parallel, we use 4th-order spatial averaging to compute $\rho_0$ on 
-  radial faces.  In spherical, we first map $\rho_0$ to Cartesian cell-centers 
-  using the mapping in section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
-  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.
-
-    \item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$
-      without explicitly including the reaction terms,
+
+\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ using
 \begin{equation}
@@
 \begin{equation}
-\eta_{\rho}^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2),\star}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a),\star}}{2}\right)\right] \sum_d \left(\uadvtwo \cdot \eb_d\right)n_d.
+\eta_{\rho}^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a)}}{2}\right)\right] \sum_d \left(\uadvtwo \cdot \eb_d\right)n_d.
 \end{equation}
@@
 \end{equation}
-This gives a radial cell-centered $\etarho^{\nph}$.  To get $\etarho^{\nph}$ at 
-radial edges, average the two neighboring radial bins.
+This gives a cell-centered $\etarho^{\nph}$.  To get $\etarho^{\nph}$ at 
+radial edges, average the two neighboring cell-centers.
 \end{description}
@@
 \psi_j^{n+\myhalf} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{n+\myhalf} 
-+ \eta_{\rho,j+\myhalf}^{n+\myhalf}\right) g^{n+1} \enskip .
++ \eta_{\rho,j+\myhalf}^{n+\myhalf}\right) g\enskip .
 \end{equation}
@@
 
-\item Update the enthalpy, using a difference approximation to
-  equation (\ref{eq:rhohupd}), neglecting the reaction terms.
+\item Update the enthalpy.  This step is identical to step 4j with the following changes:
 
@@
 
-  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, for
-    the conservative update of $(\rho h)^{(1)}$. Here we predict 
-    $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0$
-    to time-centered edges to obtain $(\rho h)^{'(1),\nph}$, 
-    using a second-order unsplit Godunov procedure, 
-    as described in paper II, Appendix A, using $\Vb =
-    \uadvtwo+w_0^{\nph} \er$.  Again, we do not include the reaction
-    terms in the enthalpy prediction since we accounted for them
-    already in {\bf React State}. We use the base state enthalpy,
-    $(\rho h)_0$, to convert these to edge states for $(\rho h)$,
+\item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, using
+$\Vb = \uadvtwo+w_0^{\nph} \er$.  Convert the perturbational edge 
+state to a full state using
 \begin{equation}
@@
 \end{equation}
-  In plane-parallel, we use 4th-order spatial averaging to compute $(\rho h)_0$
-  on radial faces.  In spherical, we first map $h_0 \equiv (\rho h)_0/\rho_0$ 
-  and $\rho_0$ to Cartesian cell-centers using the mapping in section 
-  \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, use 2nd-order spatial 
-  averaging to compute $h_0$ and $\rho_0$ on all faces, and then multiply these 
-  terms to get $(\rho h)_0$ on all faces.
+
+\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$
+without explicitly including the reaction terms,
 
@@
    &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
-        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
+        + \frac{1}{\rho_0} {\mathbf{G}} \pi_0^n
         - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g^{n+\myhalf} \er \right] \enskip ,

### 2009-05-18 docs/flowchart/flowchart.tex
moved the call to (rho h)_0 = avg(rho h)^{(1)} to step 3.  This way steps 4 and 8 exactly mirror each other.
making notation consistent with future paper 4.5

@@
   Evolution,} A.~S.~Almgren, J.~B.~Bell, C.~A.~Rendleman, \& M.~Zingale
-  2006, ApJ, 649, 927 (henceforth paper~II)
-\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
+  2006, ApJ, 649, 927 (henceforth paper~II)\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
 A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale
@@
 \item We replaced {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
-plane-parallel and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical,
+planar geometry and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical geometry,
 eliminating the need for $\nabla\cdot\etarho$.  {\bf Correct Base} is now used
@@
 \item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
-plane-parallel and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
-in spherical.  This means that {\bf React Base} no longer exists, and the intermediate 
-states, $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
+planar geometry and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
+in spherical geometry.  This means that {\bf React Base} no longer exists, and the 
+intermediate states, $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
 {\bf Advect Base Enthalpy} will advance $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ 
@@
 \end{enumerate}
-Changes specific to spherical:
+Changes specific to spherical geometry:
 \begin{enumerate}
@@
 \end{equation}
-In plane-parallel, this is as straightforward as averaging cell values at 
-a particular height.  For spherical problems, see the description in 
-paper IV, section 2.2 (or section \ref{Sec:Avg}).
+\begin{description}
+\item[planar:] We simply average cells at constant height.
+\item[spherical:] See the description in paper IV, section 2.2 
+(or section \ref{Sec:Avg} in these notes).
+\end{description}
 
@@
 
-\item[plane-parallel:] We discretize the base state density evolution 
+\item[planar:] We discretize the base state density evolution 
 equation from paper III equation (29) without the $\etarho$ term,
@@
 %is the process by which we update the base state pressure through $\dt$ in time and
-%compute $\psi$.  The input arguments are different for plane-parallel and spherical.
+%compute $\psi$.  The input arguments are different for planar and spherical geometries.
 %\begin{description}
@@
 is the process by which we update the base state enthalpy through $\dt$ in time.
-The input arguments are different for plane-parallel and spherical.
+The input arguments are different for planar and spherical geometries.
 \begin{description}
@@
 \begin{description}
-\item[plane-parallel:] Set $\rho_0^{\outp} = {\bf Avg}(\rho^{\inp})$.
+\item[planar:] Set $\rho_0^{\outp} = {\bf Avg}(\rho^{\inp})$.
 \item[spherical:] Set $\rho_0^{\outp} = \rho_0^{\inp} + {\bf Avg}(\rho^{\inp} - \rho_0^{\inp})$.
@@
 \item {\bf Compute} {\boldmath $w_0$} is the process by which we compute $w_0$.
-The arguments are different for plane-parallel and spherical.
+The arguments are different for planar and spherical geometries.
 \begin{description}
@@
 \begin{description}
-\item For plane-parallel, gravity constant in space and time.
-\item For spherical, we define the enclosed mass at edges, and then cell-centers:
+\item [planar:] Gravity constant in space and time.
+\item [spherical:] We define the enclosed mass at edges, and then cell-centers:
 \begin{equation}
@@
 \begin{description}
-\item For plane-parallel,\\
+\item For planar geometry,\\
 {\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,
@@
 $\overline{p^{\inp}} = \overline{p(\rho,h,X_k)^n}$.
-\item For spherical,\\
+\item For spherical geometry,\\
 {\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,
@@
 
-{\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n] \rightarrow
+\begin{enumerate}
+\renewcommand{\theenumi}{{\bf \alph{enumi}}}
+
+\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n] \rightarrow
 [ \rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)} ]$.
+
+\item {\bf Correct Base}$[(\rho h)_0^{n},(\rho h)^{(1)}] \rightarrow [(\rho h)_0^{n}]$.
+This sets $(\rho h)_0$ to be the lateral average of $(\rho h)^{(1)}$
+
+\end{enumerate}
 
@@
 \end{equation}
-  In plane-parallel, we use 4th-order spatial averaging to compute $\rho_0$ on 
-  radial faces.  In spherical, we first map $\rho_0$ to Cartesian cell-centers 
+  In planar geometry, we use 4th-order spatial averaging to compute $\rho_0$ on 
+  radial faces.  In spherical geometry, we first map $\rho_0$ to Cartesian cell-centers 
   using the mapping in section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
@@
 \begin{description}
-\item For plane-parallel,
+\item For planar geometry,
 \begin{equation}
@@
 \end{equation}
-\item For spherical, first construct 
+\item For spherical geometry, first construct 
 $\eta_{\rho}^{{\rm cart},\nph,\star} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf,\star}$ using:
@@
 %\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
-%\item For spherical,\\
+%\item For spherical geometry,\\
 %{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
@@
 \begin{description}
-\item For plane-parallel, 
+\item For planar geometry,
 \begin{equation}
@@
 \end{equation}
-\item For spherical, first compute:
+\item For spherical geometry, first compute:
 \begin{equation}
@@
 
-\item {\bf Correct Base}$[(\rho h)_0^{n},(\rho h)^{(1)}] \rightarrow [(\rho h)_0^{n}]$.
-
 \item Update $(\rho h)_0$.
 \begin{description}
-\item For plane-parallel,\\
+\item For planar geometry,\\
 {\bf Advect Base Enthalpy Planar}$[(\rho h)_0^{n}, w_0^{\nph,\star}, \psi^{\nph,\star}]
 \rightarrow [(\rho h)_0^{n+1,\star}]$.
-\item For spherical,\\
+\item For spherical geometry,\\
 {\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^{n}, w_0^{\nph,\star}, 
@@
   \end{equation}
-For spherical flows, experience has shown that solving the analytically equivalent form,
+For spherical geometry, experience has shown that solving the analytically equivalent form,
 \begin{equation}
@@
 \end{equation}
-  In plane-parallel, we use 4th-order spatial averaging to compute $(\rho h)_0$
-  on radial faces.  In spherical, we first map $h_0 \equiv (\rho h)_0/\rho_0$ 
+  For planar geometry, we use 4th-order spatial averaging to compute $(\rho h)_0$
+  on radial faces.  In spherical geometry, we first map $h_0 \equiv (\rho h)_0/\rho_0$ 
   and $\rho_0$ to Cartesian cell-centers using the mapping in section 
@@
 \begin{description}
-\item For plane-parallel,
+\item For planar geometry,
 
@@
 
-\item For spherical,
+\item For spherical geometry,
 
@@
 \end{eqnarray}
-\item Update the temperature using the equation of state.  For plane-parallel,
+\item Update the temperature using the equation of state.  For planar geometry,
 we update using
@@
 \end{equation}
-but in spherical, we set {\tt use\_tfromp = T}, and we use
+but in spherical geometry, we set {\tt use\_tfromp = T}, and we use
 \begin{equation}
@@
 \begin{description}
-\item For plane-parallel,\\
+\item For planar geometry,\\
 {\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph}},
@@
 \end{equation}
-\item For spherical,\\
+\item For spherical geometry,\\
 {\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf},
@@
 
-\item For plane-parallel,
+\item For planar geometry,
 \begin{equation}
@@
 \end{equation}
-\item For spherical, first construct 
+\item For spherical geometry, first construct 
 $\eta_{\rho}^{{\rm cart},\nph} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf}$ using:
@@
 %\begin{description}
-%\item For plane-parallel,\\
+%\item For planar geometry,\\
 %{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
 %$\rightarrow [p_0^{n+1},\psi^{\nph}]$.
-%\item For spherical,\\
+%\item For spherical geometry,\\
 %{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{(1)},\overline{S^{\nph}},
@@
 \begin{description}
-\item For plane-parallel, 
+\item For planar geometry, 
 \begin{equation}
@@
 \begin{description}
-\item For plane-parallel,\\
+\item For planar geometry,\\
 {\bf Advect Base Enthalpy Planar}$[(\rho h)_0^n, w_0^{\nph}, \psi^{\nph}]
 \rightarrow [(\rho h)_0^{n+1}]$.
-\item For spherical,\\
+\item For spherical geometry,\\
 {\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^n, w_0^{\nph}, \psi^n, \psi^{\nph}]
@@
 
-\item Update the enthalpy.  This step is identical to step 4j with the following changes:
+\item Update the enthalpy.  This step is identical to step 4i with the following changes:
 
@@
 \begin{description}
-\item For plane-parallel,
+\item For planar geometry,
 \begin{eqnarray}
@@
 
-\item For spherical,
+\item For spherical geometry,
 \begin{eqnarray}
@@
 \end{eqnarray}
-\item Update the temperature using the equation of state.  For plane-parallel we
+\item Update the temperature using the equation of state.  For planar geometry we
 update using

### 2009-05-18 docs/flowchart/flowchart.tex
minor fixes to description

@@
 \hline
-$\Ub$     & total velocity ($\Ub = \Ubt + w_0 \er$) & cm~s$^{-1}$ \\
+$\Ub$     & total velocity ($\Ub = \Ubt + w_0 \eb_r$) & cm~s$^{-1}$ \\
 \hline
@@
 \hline
-$\etarho$ & $\etarho \equiv \overline{(\rho' \Ubt \cdot \er)}$ & g~cm$^{-2}$~s$^{-1}$ \\
+$\etarho$ & $\etarho \equiv \overline{(\rho' \Ubt \cdot \eb_r)}$ & g~cm$^{-2}$~s$^{-1}$ \\
 \hline
@@
 \begin{description}
-\item[planar:] We simply average cells at constant height.
+\item[planar:] This is a straightforward average since the radial coordinate is 
+in alignment with the lateral coordinate.
 \item[spherical:] See the description in paper IV, section 2.2 
@@
 $T^{\outp} = T(\rho^\outp,h^\outp,X_k^\outp)$ if {\tt use\_tfromp = FALSE} or 
-$T^{\outp} = T(\rho^\outp,p_0,X_k^\outp)$  if {\tt use\_tfromp = TRUE}.  
+$T^{\outp} = T(\rho^\outp,p_0^\inp,X_k^\outp)$  if {\tt use\_tfromp = TRUE}.  
 Note that the density remains unchanged
@@
 \begin{equation}
-\frac{\partial \rhozero}{\partial t} = - \nablab \cdotb \left( \rhozero w_0 \er \right)
+\frac{\partial \rhozero}{\partial t} = - \nabla \cdot \left( \rhozero w_0 \eb_r \right)
 \enskip ,
@@
 \begin{equation}
-\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\er\right] 
+\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\eb_r\right] 
 + \psi \enskip ,
@@
 \frac{\partial\Ubt}{\partial t} = 
-- \left(\Ubt+w_0\right) \cdotb \nablab \Ubt
-- \left(\Ubt \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
-- \frac{1}{\rho} \nablab\pi
-+ \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
-- \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip .
+- \left(\Ubt+w_0\right) \cdot \nabla \Ubt
+- \left(\Ubt \cdot \eb_r\right) \frac{\partial w_0}{\partial r} \eb_r
+- \frac{1}{\rho} \nabla\pi
++ \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \eb_r
+- \frac{(\rho-\rhozero)}{\rho} \; g \; \eb_r  \label{eq:utildeupd}  \enskip .
 \end{equation}
@@
 \begin{equation}
-\int_{\Omega_H} \Ubt \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
+\int_{\Omega_H} \Ubt \cdot \eb_r \; dA = 0 \label{eq:udoterzero} \enskip ,
 \end{equation}
@@
 \begin{equation}
-\nablab \cdotb (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
+\nabla \cdot (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
 \label{eq:tildeconstraint}
@@
 \begin{equation}
-\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
+\nabla \cdot \left(\beta_0^n \uadvone\right) = 
 \beta_0^n \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right) \enskip .
@@
 \begin{equation}
-\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
+\nabla \cdot \left(\beta_0^n \uadvone\right) = 
 \beta_0^n \left\{ \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right)
@@
 %--------------------------------------------------------------------------
-\item[Step 4.] {\em Advect the base state, then the full state, through a time interval 
-of $\dt.$}
+\item[Step 4.] {\em Advect the base state and full state, through a time interval of $\dt.$}
 
@@
 \begin{equation}
-\frac{\partial (\rho X_k)}{\partial t} + \nablab \cdotb (\Ub \rho X_k) = 0 \enskip .
+\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\Ub \rho X_k) = 0 \enskip .
 \label{eq:species}
@@
   unsplit Godunov procedure, as described in paper II,
-  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \er$.  Convert the 
+  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \eb_r$.  Convert the 
   perturbational edge state to a full state using
@@
 (\rho X_k)^{(2),\star} = (\rho X_k)^{(1)} 
- - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right)  
+ - \dt \; \left\{ \nabla \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right)  
   (\rho X_k)^{(1),\nph,\star} \right] \right\} \enskip ,
@@
 \begin{equation}
- \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdotb \er + w_0^{\nph,\star}\right) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
+ \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdot \eb_r + w_0^{\nph,\star}\right) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
 \end{equation}
@@
   \begin{equation}
-  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
-+ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} 
+  \frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h)
++ \psi + (\Ubt \cdot \eb_r) \frac{\partial p_0}{\partial r} 
 \enskip , \label{eq:rhohupdadv} 
@@
 \begin{equation}
-  \frac{\partial (\rho h)}{\partial t}  = - \nablab \cdotb (\Ub \rho h)
+  \frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h)
 + \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt  
@@
     as described in paper II, Appendix A, using $\Vb =
-    \uadvone+w_0^{\nph,\star} \er$.  We do not include the reaction
+    \uadvone+w_0^{\nph,\star} \eb_r$.  We do not include the reaction
     terms in the enthalpy prediction, since we accounted for them
@@
   \begin{eqnarray}
-  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
-      \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right) (\rho
+  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nabla
+      \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right) (\rho
       h)^{(1),\nph,\star} \right] \right\} \nonumber \\ && + \dt \;
-    \left(\uadvone \cdotb \er\right) \left(\frac{\partial
+    \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial
       p_0}{\partial r} \right)^{n} + \dt \; \psi^{\nph,\star} \enskip ,
@@
   \begin{eqnarray}
-  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab
-      \cdotb \left[ \left(\uadvone+w_0^{\nph,\star} \er\right) (\rho
+  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nabla
+      \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right) (\rho
       h)^{(1),\nph,\star} \right] \right\} \nonumber \\ 
@@
 \begin{equation}
-\nablab \cdotb \left(\beta_0^{\nph} \uadvtwo\right) =
+\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
 \beta_0^{\nph}\left(S^{\nph} - \overline{S^{\nph}}\right) \enskip ,
@@
 \begin{equation}
-\nablab \cdotb \left(\beta_0^{\nph} \uadvtwo\right) =
+\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
 \beta_0^{\nph}\left\{\left(S^{\nph} - \overline{S^{\nph}}\right)
@@
 %--------------------------------------------------------------------------
-\item[Step 8.] {\em Advect the base state, then the full state, through a time interval of $\dt.$}
+\item[Step 8.] {\em Advect the base state and full state through a time interval of $\dt.$}
 
@@
 \item Compute the species edge states, $(\rho X_k)^{(1),n+\myhalf}$, 
-using $\Vb = \uadvtwo+w_0^{\nph}\er$.  Convert the perturbational edge 
+using $\Vb = \uadvtwo+w_0^{\nph}\eb_r$.  Convert the perturbational edge 
 state to a full state using 
@@
 (\rho X_k)^{(2)} = (\rho X_k)^{(1)} 
-- \dt \; \left\{ \nablab \cdotb \left[\left(\uadvtwo+w_0^{\nph} \er\right)  
+- \dt \; \left\{ \nabla \cdot \left[\left(\uadvtwo+w_0^{\nph} \eb_r\right)  
 (\rho X_k)^{(1),\nph} \right] \right\}  \enskip ,
@@
 \begin{equation}
- \etarho^{\nph} = {\rm {\bf Avg}} \sum_k \left [\left(\uadvtwo \cdotb \er + w_0^{\nph}\right) (\rho X_k)^{(1),\nph} \right] - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip ,
+ \etarho^{\nph} = {\rm {\bf Avg}} \sum_k \left [\left(\uadvtwo \cdot \eb_r + w_0^{\nph}\right) (\rho X_k)^{(1),\nph} \right] - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip ,
 \end{equation}
@@
 \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, using
-$\Vb = \uadvtwo+w_0^{\nph} \er$.  Convert the perturbational edge 
+$\Vb = \uadvtwo+w_0^{\nph} \eb_r$.  Convert the perturbational edge 
 state to a full state using
@@
 \begin{eqnarray}
-(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
+(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nabla \cdot \left[ \left(\uadvtwo+w_0^{\nph} \eb_r\right)  
 (\rho h)^{(1),\nph} \right] \right\} \nonumber \\
-&& + \frac{\dt}{2} \; \left(\uadvtwo \cdotb \er\right)
+&& + \frac{\dt}{2} \; \left(\uadvtwo \cdot \eb_r\right)
 \left [ \left(\frac{\partial p_0}{\partial r} \right)^{n}
@@
 \begin{eqnarray}
-(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nablab \cdotb \left[ \left(\uadvtwo+w_0^{\nph} \er\right)  
+(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nabla \cdot \left[ \left(\uadvtwo+w_0^{\nph} \eb_r\right)  
 (\rho h)^{(1),\nph} \right] \right\} \nonumber \\
@@
 \Ubt^{n+1,\star} &=& \Ubt^n - \dt \;
- \left[\left(\uadvtwo+ w_0^{\nph} \er\right) \cdotb \nablab \Ubt \right]^\nph
-      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
+ \left[\left(\uadvtwo+ w_0^{\nph} \eb_r\right) \cdot \nabla \Ubt \right]^\nph
+      - \dt \; \left(\uadvtwo \cdot \eb_r\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \eb_r \nonumber \\
    &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
         + \frac{1}{\rho_0} {\mathbf{G}} \pi_0^n
-        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g^{n+\myhalf} \er \right] \enskip ,
+        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g^{n+\myhalf} \eb_r \right] \enskip ,
 \end{eqnarray}
@@
 and $\mathbf{G}$ approximates a cell-centered gradient from nodal data.
-The construction of $[(\uadvtwo+ w_0^{\nph} \er \cdotb \nablab) \Ubt ]^\nph$
-is described in paper II, Appendix A, with $\Vb = \uadvtwo+w_0^{\nph} \er$ and $s$ 
+The construction of $[(\uadvtwo+ w_0^{\nph} \eb_r \cdot \nabla) \Ubt ]^\nph$
+is described in paper II, Appendix A, with $\Vb = \uadvtwo+w_0^{\nph} \eb_r$ and $s$ 
 set to each component of $\Ubt^n$ individually.  Here, the $\star$ superscript 
@@
 \begin{equation}
-\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right) 
+\nabla \cdot \left(\beta_0^{\nph} \Ubt^{n+1} \right) 
 = \beta_0^{\nph} \left(S^{n+1} - \overline{S^{n+1}} \right) \enskip,
@@
 \begin{equation}
-\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S^{n+1}} \right)
+\nabla \cdot \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S^{n+1}} \right)
 + \frac{f}{\gammabar^{n+1} p_0^{n+1}}
@@
 for nodal values of $\phi$, where $L_\beta^\rho$ is the standard bilinear
-finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
+finite element approximation to $\nabla \cdot ({\beta_0^\nph}/{\rho^\nph}) \nabla.$
 The discretization with the volume discrepancy term is:

### 2009-05-18 docs/flowchart/flowchart.tex
got rid of correct base shorthand and just put the average calls in

@@
 
-\item {\bf Correct Base}$[\rho_0^{\inp}, \rho^{\inp}] \rightarrow [\rho_0^{\outp}]$
-is the process by which we adjust the base state density to account for
-large-scale mixing.  It is also used to reset the base state enthalpy to be
-the average of the full state enthalpy since we do not need to compute the complete
-evolution of the base state enthalpy.
-Unlike paper III where we used $\nabla\cdot\etarho$ as an adjustment to $\rho_0$, 
-we instead simply set $\rho_0=\overline{\rho}$, which is analytically equivalent.
-\begin{description}
-\item[planar:] Set $\rho_0^{\outp} = {\bf Avg}(\rho^{\inp})$.
-\item[spherical:] Set $\rho_0^{\outp} = \rho_0^{\inp} + {\bf Avg}(\rho^{\inp} - \rho_0^{\inp})$.
-\end{description}
-
 \item {\bf Compute} {\boldmath $w_0$} is the process by which we compute $w_0$.
@@
 
-\item {\bf Correct Base}$[(\rho h)_0^{n},(\rho h)^{(1)}] \rightarrow [(\rho h)_0^{n}]$.
-This sets $(\rho h)_0$ to be the lateral average of $(\rho h)^{(1)}$
+\item Set $(\rho h)_0^n$ to be the lateral average of $(\rho h)^{(1)}$:
+\begin{description}
+\item For planar geometry, set $(\rho h)_0^n = $ {\bf Avg}$[(\rho h)^{(1)}]$
+\item For spherical geometry, set $(\rho h)_0^n = (\rho h)_0^n +$ {\bf Avg}$[(\rho h)^{(1)} - (\rho h)_0^n]$
+\end{description}
 
@@
 
-\item {\bf Correct Base}$[\rho_0^{(2a),\star},\rho^{(2),\star}] \rightarrow [\rho_0^{n+1,\star}]$.
+\item Set $\rho_0^{n+1,\star}$ to be the lateral average of $\rho^{(2),\star}$:
+\begin{description}
+\item For planar geometry, set $\rho_0^{n+1,\star} =$ {\bf Avg}$(\rho^{(2),\star})$
+\item For spherical geometry, set $\rho_0^{n+1,\star} = \rho_0^{(2a),\star} +$ {\bf Avg}$(\rho^{(2),\star} - \rho_0^{(2a),\star})$
+\end{description}
 
@@
 
-\item {\bf Correct Base}$[\rho_0^{(2a)},\rho^{(2)}] \rightarrow [\rho_0^{n+1}]$.
+\item Set $\rho_0^{n+1}$ to be the lateral average of $\rho^{(2)}$:
+\begin{description}
+\item For planar geometry, set $\rho_0^{n+1} =$ {\bf Avg}$(\rho^{(2)})$
+\item For spherical geometry, set $\rho_0^{n+1} = \rho_0^{(2a)} +$ {\bf Avg}$(\rho^{(2)} - \rho_0^{(2a)})$
+\end{description}
 

### 2009-06-05 docs/lo_density/lo_density.tex
add section reference for runtime_parameters
we compute cutoff using less than or equal to rather than less than

@@
 setting the corresponding coordinate equal to $r$ as soon as $\rho_0(r)$ is less than 
-that particular cutoff value.  We set the coordinates at the following points in the
-code:
+or equal to that particular cutoff value.  We set the coordinates at the following 
+points in the code:
 
@@
 
-\section{anelastic\_cutoff}
+\section{anelastic\_cutoff}\label{Sec:Anelastic Cutoff}
 
@@
 
-\section{base\_cutoff\_density}
+\section{base\_cutoff\_density}\label{Sec:Base Cutoff Density}
 

### 2009-06-08 Source/advect_base.f90
make the planar version of advect_base_dens consistent with the spherical in that we update density everywhere, not just below the cutoff.  this is still a work in progress

@@ -45,8 +45,7 @@ contains
 
     use bl_constants_module
     use make_edge_state_module
-    use geometry, only: nr_fine, r_start_coord, r_end_coord, numdisjointchunks, nlevs, dr, &
-         base_cutoff_density_coord
+    use geometry, only: nr_fine, r_start_coord, r_end_coord, numdisjointchunks, nlevs, dr
 
     real(kind=dp_t), intent(in   ) ::                  w0(:,0:)
     real(kind=dp_t), intent(in   ) ::            rho0_old(:,0:)
@@ -82,12 +81,8 @@ contains
     do n=1,nlevs
        do i=1,numdisjointchunks(n)
           do r=r_start_coord(n,i),r_end_coord(n,i)
-             if (r .le. base_cutoff_density_coord(n)) then
-                rho0_new(n,r) = rho0_old(n,r) &
-                     - dt / dr(n) * (edge(n,r+1) * w0(n,r+1) - edge(n,r) * w0(n,r)) 
-             else
-                rho0_new(n,r) = rho0_old(n,r)
-             end if
+             rho0_new(n,r) = rho0_old(n,r) &
+                  - dt / dr(n) * (edge(n,r+1) * w0(n,r+1) - edge(n,r) * w0(n,r)) 
           end do
        end do
     end do
@@ -100,7 +95,7 @@ contains
 
     use bl_constants_module
     use make_edge_state_module
-    use geometry, only: r_cc_loc, r_edge_loc, dr, nr_fine, base_cutoff_density_coord
+    use geometry, only: r_cc_loc, r_edge_loc, dr, nr_fine
     
     real(kind=dp_t), intent(in   ) ::                  w0(:,0:)
     real(kind=dp_t), intent(in   ) ::            rho0_old(:,0:)

### 2009-06-08 docs/lo_density/lo_density.tex
updated notes

@@
 setting the corresponding coordinate equal to $r$ as soon as $\rho_0(r)$ is less than 
-or equal to that particular cutoff value.  We set the coordinates at the following 
-points in the code:
+or equal to that particular cutoff value.  We compute the cutoff coordinate at the
+coarsest level, and then set the coordinate at the next finer level to be twice that
+of the coarser level.  We set the coordinates at the following points in the code:
 
@@
 
-\item In {\tt probin}, {\tt anelastic\_cutoff} and {\tt base\_cutoff\_density} are
-  set to $3\times 10^6$ by default.
-
-\item In {\tt probin}, {\tt burning\_cutoff\_density} is set to 
-  {\tt base\_cutoff\_density}.  There is no option to set 
-  {\tt burning\_cutoff\_density} using the inputs file.
-
-\item After we initialize the base state in {\tt varden}.
+\item After we call {\tt initialize} in {\tt varden}.
 
@@
 
-\item After each call to {\tt correct\_base} in spherical 
-  (or call to {\tt average} in plane-parallel).
+\item After we ``{\tt correct base}'' (by setting $\rho_0 = \overline{\rho}$ in
+  plane-parallel and $\rho_0 = \rho_0 + {\bf{Avg}}(\rho - \rho_0)$).
+
+\item At the beginning of the second-half of the algorithm (Step 6), we reset
+  the coordinates using $\rho_0^n$.
 
@@
 
-\item In {\tt init\_base\_state}, we abort the program if 
-  {\tt anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
+\item In {\tt probin}, {\tt anelastic\_cutoff} is set to $3\times 10^6$ by default.
 
-\item In {\tt make\_div\_coeff}, we change how we compute {\tt div\_coeff} 
-  for $r\in[{\tt anelastic\_cutoff\_coord},{\tt r\_end\_coord}]$.
+\item In {\tt init\_base\_state}, we give a warning of {\tt anelastic\_cutoff}
+  is smaller than the minimum model density.  Then, if {\tt anelastic\_cutoff} 
+  $<$ {\tt base\_cutoff\_density} we abort the program.
+
+\item In {\tt make\_div\_coeff}, for 
+  $r \ge {\tt anelastic\_cutoff\_coord}$, we set
+  ${\tt div\_coeff}(n,r) = {\tt div\_coeff}(n,r-1) * \rho_0(n,r)/\rho_0(n,r-1)$.
 
 \item In {\tt make\_psi} for plane-parallel, we only compute $\psi$ for 
-  $r\in[{\tt r\_start\_coord},{\tt anelastic\_cutoff\_coord}-1]$.
+  $r < {\tt anelastic\_cutoff\_coord}$.
+\MarginPar{I think we want to change this to include {\tt anelatic\_cutoff\_coord}.}
 
 \item in {\tt make\_S}, we set {\tt delta\_gamma1\_term} and {\tt delta\_gamma1} 
-  to zero for $r\in[{\tt anelastic\_cutoff\_coord},{\tt r\_end\_coord}]$.
+  to zero for $r \ge {\tt anelastic\_cutoff\_coord}$.
 
@@
 
-\item In {\tt init\_base\_state}, we abort the program if {\tt
-  anelastic\_cutoff} $<$ {\tt base\_cutoff\_density}.
+\item In {\tt probin}, {\tt base\_cutoff\_density} is set to $3\times 10^6$ by default.
+
+\item In {\tt init\_base\_state}, we give a warning of {\tt base\_cutoff\_density}
+  is smaller than the minimum model density.  Then, if {\tt anelastic\_cutoff} 
+  $<$ {\tt base\_cutoff\_density} we abort the program.
 
@@
   location of the first cell-center at the coarsest level for which
-  $\rho_0 <$ {\tt base\_cutoff\_density}.  This is a trick used for making
-  the data consistent for multiple level problems.
-
-\item In {\tt base\_state}, when we are generating the initial
-  background/base state, if we are above {\tt base\_cutoff\_density\_loc}, 
+  $\rho_0 \le {\tt base\_cutoff\_density}$.  This is a trick used for making
+  the data consistent for multiple level problems.  When we are generating the 
+  initial background/base state, if we are above {\tt base\_cutoff\_density\_loc}, 
   just use the values for $\rho,T$, and $p$ at {\tt base\_cutoff\_density\_loc}.
+  When we check whether we are in HSE, we use {\tt base\_cutoff\_density\_loc}.
 
 \item In {\tt make\_hgrhs}, {\tt make\_macrhs}, and {\tt make\_w0}, 
-  we only add the volume discrepancy for
-  $r\in[{\tt r\_start\_coord},{\tt base\_cutoff\_density\_coord}-1]$.
+  we only add the volume discrepancy for $r < {\tt base\_cutoff\_density\_coord}$
+  (in plane parallel) and if $\rho_0 > {\tt base\_cutoff\_density}$ (in spherical).
+  \MarginPar{Revisit this.}
 
 \item In {\tt mkrhohforce} for plane-parallel, for
-  $r\in[{\tt base\_cutoff\_density\_coord},{\tt r\_end\_coord}]$ we
+  $r \ge {\tt base\_cutoff\_density\_coord}$, we
   compute $\nabla p_0$ with a difference stencil instead of simply
@@
 
-\item In {\tt make\_grav}, if we are spherical, we do not add to the
-  enclosed mass, $M(r)$ if $\rho$ is less than {\tt base\_cutoff\_density}.
+\item In {\tt make\_grav} for spherical, we only add the enclosed mass if
+  $\rho_0 > {\tt base\_cutoff\_density}$.
 
 \item In {\tt enforce\_HSE}, we set $p_0(r) = p_0(r-1)$ for 
-  $r\in[{\tt base\_cutoff\_density\_coord}-1,{\tt r\_end\_coord}]$
+  $r \ge {\tt base\_cutoff\_density\_coord}+1$.
 
@@
 
+\item In {\tt probin}, {\tt burning\_cutoff\_density} is set to 
+  {\tt base\_cutoff\_density}.  There is no option to set 
+  {\tt burning\_cutoff\_density} using the inputs file.
+
 \item In {\tt react\_state}, we only call the burner if 

### 2009-06-08 docs/lo_density/lo_density.tex
added marginpars, cleaned up some notation

@@
 \item After we ``{\tt correct base}'' (by setting $\rho_0 = \overline{\rho}$ in
-  plane-parallel and $\rho_0 = \rho_0 + {\bf{Avg}}(\rho - \rho_0)$).
+  plane-parallel and $\rho_0 = \rho_0 + {\bf{Avg}}(\rho - \rho_0)$ in spherical).
 
 \item At the beginning of the second-half of the algorithm (Step 6), we reset
-  the coordinates using $\rho_0^n$.
+  the coordinates to the base-time values using $\rho_0^n$.
 
@@
 
-\item In {\tt init\_base\_state}, we give a warning of {\tt anelastic\_cutoff}
-  is smaller than the minimum model density.  Then, if {\tt anelastic\_cutoff} 
-  $<$ {\tt base\_cutoff\_density} we abort the program.
+\item In {\tt init\_base\_state}, we give a warning if {\tt anelastic\_cutoff}
+  is smaller than the minimum model density.  Then, if ${\tt anelastic\_cutoff} 
+  < {\tt base\_cutoff\_density}$, we abort the program.
 
@@
   ${\tt div\_coeff}(n,r) = {\tt div\_coeff}(n,r-1) * \rho_0(n,r)/\rho_0(n,r-1)$.
+  \MarginPar{I think we want $r > {\tt anelastic\_cutoff\_coord}$.}
 
@@
   $r < {\tt anelastic\_cutoff\_coord}$.
-\MarginPar{I think we want to change this to include {\tt anelatic\_cutoff\_coord}.}
+  \MarginPar{I think we want to change this to $r \le {\tt anelastic\_cutoff\_coord}$.}
 
@@
   to zero for $r \ge {\tt anelastic\_cutoff\_coord}$.
+  \MarginPar{I think we want $r > {\tt anelastic\_cutoff\_coord}$.}
 
@@
   setting it to $\rho_0 g$.
+  \MarginPar{I think we want $r > {\tt base\_cutoff\_density\_coord}$.}
 
@@
 \item In {\tt enforce\_HSE}, we set $p_0(r) = p_0(r-1)$ for 
-  $r \ge {\tt base\_cutoff\_density\_coord}+1$.
+  $r > {\tt base\_cutoff\_density\_coord}$.
 

### 2009-06-08 docs/lo_density/lo_density.tex
more marginpars

@@
 \item In {\tt init\_base\_state}, we give a warning of {\tt base\_cutoff\_density}
-  is smaller than the minimum model density.  Then, if {\tt anelastic\_cutoff} 
-  $<$ {\tt base\_cutoff\_density} we abort the program.
+  is smaller than the minimum model density.  Then, if ${\tt anelastic\_cutoff} 
+  < {\tt base\_cutoff\_density}$ we abort the program.
 
@@
   we only add the volume discrepancy for $r < {\tt base\_cutoff\_density\_coord}$
-  (in plane parallel) and if $\rho_0 > {\tt base\_cutoff\_density}$ (in spherical).
-  \MarginPar{Revisit this.}
+  (in plane parallel) and if $\rho_0^{\rm cart} > {\tt base\_cutoff\_density}$ 
+  (in spherical).
+  \MarginPar{Spherical should cue off the coordinate.}
 
@@
 
-\item In {\tt update\_scal}, if the new $\rho$ is less than 
-  {\tt base\_cutoff\_density} and {\tt do\_eos\_h\_above\_cutoff},  
-  we call the EOS to compute $h$.
+\item In {\tt update\_scal}, if $\rho < {\tt base\_cutoff\_density}$
+   and {\tt do\_eos\_h\_above\_cutoff}, we call the EOS to compute $h$.
+  \MarginPar{Perhaps $\rho \le {\tt base\_cutoff\_density}$}
 
@@
   $\rho_0 > {\tt base\_cutoff\_density}$.
+  \MarginPar{Cue off the coordinate?}
 

### 2009-06-08 docs/lo_density/lo_density.tex
added a cute figure about the cutoff density and coordinate

@@
 setting the corresponding coordinate equal to $r$ as soon as $\rho_0(r)$ is less than 
-or equal to that particular cutoff value.  We compute the cutoff coordinate at the
+or equal to that particular cutoff value (see Figure \ref{Fig:Cutoff}.
+We compute the cutoff coordinate at the
 coarsest level, and then set the coordinate at the next finer level to be twice that
 of the coarser level.  We set the coordinates at the following points in the code:
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[hpb]
+\centering
+\includegraphics[width=4.0in]{\lodensfigpath/cutoff}\hspace{0.2in}
+\begin{minipage}[b]{5.0in}
+\caption{Image of how the cutoff density and cutoff coordinates
+are related.\vspace{2em}}
+\end{minipage}
+\label{Fig:Cutoff}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2009-06-08 docs/lo_density/lo_density.tex
finished with my comments for how I think we can make the cutoff treatment consistent.  These are subtle details which shouldn't "fix" test_basestate.

@@
   (in spherical).
-  \MarginPar{Spherical should cue off the coordinate.}
+  \MarginPar{For plane-parallel, I think we should use 
+    $r \le {\tt base\_cutoff\_density\_coord}$.}
 
@@
    and {\tt do\_eos\_h\_above\_cutoff}, we call the EOS to compute $h$.
-  \MarginPar{Perhaps $\rho \le {\tt base\_cutoff\_density}$}
+  \MarginPar{I think we want $\rho \le {\tt base\_cutoff\_density}$}
 
@@
   $\rho_0 > {\tt base\_cutoff\_density}$.
-  \MarginPar{Cue off the coordinate?}
+  \MarginPar{I think this needs to be reworked to include mass up to the 
+    cell-centered location of {\tt base\_cutoff\_density\_coord} for both cell 
+    and edge gravity.}
 

### 2009-06-21 docs/flowchart/flowchart.tex
pass in gravity to make_w0_spherical
fix the lower boundary condition index for w0 in spherical
update paper IV info

@@
 \item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection,}
-A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale,
-  in preparation (henceforth paper~IV)
+M.~Zingale, A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& S.~E.~Woosley
+  2009, ApJ, submitted (henceforth paper~IV)
 \end{itemize}
@@
 
-\item {\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},
-\etarho^{\inp}] \rightarrow[w_0^{\outp}]$:
+\item {\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},g^\inp,\etarho^{\inp}] \rightarrow[w_0^{\outp}]$:
 
@@
 F_j &=&  \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
- - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
+ - \frac{g_{j-\myhalf}^\inp}{r_{j-\myhalf}^2 \Delta r} 
 \left[ \left( r^2 \etarho^{\inp} \right)_{j} - \left( r^2 \etarho^{\inp}  \right)_{j-1} \right] 
@@
 \end{eqnarray}
-Here,
-\begin{equation}
-g_{j-\myhalf} = \frac{G}{r_{j-\myhalf}^2} \sum_{k=1}^{j} \frac{4}{3} \pi  (r_{k-\myhalf}^3 - r_{k-\thalf}^3) \; \rho_{0,k-1}^{\inp} \enskip.
-\end{equation}
-
 We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
-corresponds to $j=1,$ by setting:
-\begin{eqnarray}
-A_1 = C_1 = F_1 &=& 0\enskip, \nonumber \\
-B_1 &=& 1\enskip.
+corresponds to $j=0,$ by setting:
+\begin{eqnarray}
+A_0 = C_0 = F_0 &=& 0\enskip, \nonumber \\
+B_0 &=& 1\enskip.
 \end{eqnarray}
@@
 {\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,
-\rhozero^n,p_0^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
+\rhozero^n,p_0^n,g^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
 \end{description}
@@
 \item Define:
-\begin{equation}
-\gammabar^{\nph} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}\enskip , ~~~
-\rhozero^{\nph} = \frac{\rhozero^{n} + \rhozero^{n+1,\star}}{2}\enskip , ~~~
-p_0^{\nph} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip ,
-\end{equation}
+\begin{eqnarray}
+\gammabar^{\nph} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}\enskip, \qquad
+\rhozero^{\nph} = \frac{\rhozero^{n} + \rhozero^{n+1,\star}}{2}\enskip, \nonumber \\
+p_0^{\nph} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip, \qquad
+g^\nph = \frac{g^n + g^{n+1,\star}}{2}\enskip ,
+\end{eqnarray}
 with
@@
 {\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph}},
-\gammabar^{n+\myhalf},p_0^{n+\myhalf},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf}]$.
+\gammabar^{n+\myhalf},p_0^{n+\myhalf},g^{n+\myhalf},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf}]$.
 

### 2009-07-01 docs/runtime_parameters/runtime_parameters.tex
updated with script

@@
 \verb=  init_shrink  = &   the multiplicative factor ($\le 1$) to reduce the initial timestep as computed by the various timestep estimators  &  1.0 \\
-\verb=  interp_type_radial_bin_to_cart  = &   The interpolation for filling a full multifab from a 1D bin-centered array. \newline 1 = piecewise constant; \newline 2 = piecewise linear.  &  1 \\
 \verb=  job_name  = &   job name printed in output  &  "" \\
@@
 \verb=  max_levs  = &   Total number of levels.  1 = single level.  &  1 \\
-\verb=  max_mg_bottom_nlevels  = &   if mg\_bottom\_solver == 4, then how many mg levels can the bottom solver mgt object have  &  -1 \\
+\verb=  max_mg_bottom_nlevels  = &   if mg\_bottom\_solver == 4, then how many mg levels can the bottom solver mgt object have  &  1000 \\
 \verb=  max_step  = &   Maximum number of steps in the simulation.  &  1 \\
@@
 \verb=  n_cellz  = &   Number of cells for the base level in the z-direction  &  -1 \\
-\verb=  p0_update_type  = &   How we update the base state pressure and compute psi. \newline 1 = call {\tt advect\_base\_pres}; \newline 2 = call {\tt enforce\_HSE}.  &  1 \\
+\verb=  p0_update_type  = &   How we update the base state pressure and compute psi. \newline 1 = call {\tt advect\_base\_pres}; \newline 2 = call {\tt enforce\_HSE}.  &  2 \\
 \verb=  perturb_model  = &   Turn on a perturbation in the initial data.  Problem specific.  &  .false. \\
@@
 \verb=  rotational_frequency  = &   rotational frequency used for computing centrifugal term in rotation problems.  &  ZERO \\
+\verb=  s0_interp_type  = &   The interpolation for filling a cell-centered multifab from a 1D bin-centered array. \newline 1 = piecewise constant; \newline 2 = piecewise linear.  &  1 \\
+\verb=  s0mac_interp_type  = &   The interpolation for filling an edge based multifab from a 1D bin-centered array.
+\newline 1 = Interpolate s0 to cell centers, then average to edges; \newline 2 = Interpolate s0 to edges directly;                \newline 3 = Interpolate s0 to nodes, then average to edges.
+&  1 \\
 \verb=  single_prec_plotfiles  = &    &  .false. \\
@@
 \verb=  verbose  = &   General verbosity  &  0 \\
-\verb=  w0mac_interp_type  = &   0 = Compute w0 at edges using a projection; \newline 1 = Interpolate w0 to cell centers, then average to edges; \newline 2 = Interpolate w0 to edges directly;                \newline 3 = Interpolate w0 to nodes, then average to edges.
+\verb=  w0mac_interp_type  = &   The interpolation for putting w0 on edges.
+We only compute the normal component.
+\newline 1 = Interpolate w0 to cell centers, then average to edges; \newline 2 = Interpolate w0 to edges directly;                \newline 3 = Interpolate w0 to nodes, then average to edges.
 &  1 \\

### 2009-07-02 Source/advance.f90
removed p0_update_type=1 code

@@ -488,48 +488,41 @@ contains
           end do
        end if
 
-       if (p0_update_type .eq. 1) then
+       ! set new p0 through HSE
+       p0_new = p0_old
+       call enforce_HSE(rho0_new,p0_new,grav_cell_new)
 
-          call advect_base_pres(w0,Sbar,p0_old,p0_new,gamma1bar_1,psi,psi_old,etarho_cc, &
-                                s2,dt,dx,mla)
+       ! make psi
+       if (spherical .eq. 0) then
+          call make_psi_planar(etarho_cc,psi)
+       else
+          ! compute p0_nph
+          p0_nph = HALF*(p0_old+p0_new)
 
-       else if (p0_update_type .eq. 2) then
+          ! compute gamma1bar_2 and store it in gamma1bar_nph
+          do n=1,nlevs
+             call multifab_build(gamma1(n), mla%la(n), 1, 0)
+          end do
 
-          ! set new p0 through HSE
-          p0_new = p0_old
-          call enforce_HSE(rho0_new,p0_new,grav_cell_new)
+          call make_gamma(mla,gamma1,s2,p0_new,dx)
+          call average(mla,gamma1,gamma1bar_nph,dx,1)
 
-          ! make psi
-          if (spherical .eq. 0) then
-             call make_psi_planar(etarho_cc,psi)
-          else
-             ! compute p0_nph
-             p0_nph = HALF*(p0_old+p0_new)
+          do n=1,nlevs
+             call destroy(gamma1(n))
+          end do
 
-             ! compute gamma1bar_2 and store it in gamma1bar_nph
-             do n=1,nlevs
-                call multifab_build(gamma1(n), mla%la(n), 1, 0)
-             end do
-             
-             call make_gamma(mla,gamma1,s2,p0_new,dx)
-             call average(mla,gamma1,gamma1bar_nph,dx,1)
-             
-             do n=1,nlevs
-                call destroy(gamma1(n))
-             end do
+          ! compute gamma1bar_nph
+          gamma1bar_nph = HALF*(gamma1bar_1+gamma1bar_nph)
 
-             ! compute gamma1bar_nph
-             gamma1bar_nph = HALF*(gamma1bar_1+gamma1bar_nph)
-
-             ! make base time and time-centered psi
-             call make_psi_spherical(psi_old,w0,gamma1bar_1  ,p0_old,Sbar)
[…]

### 2009-07-02 Source/advect_base.f90

@@ -6,7 +6,7 @@ module advect_base_module
 
   private
 
-  public :: advect_base_dens, advect_base_pres, advect_base_enthalpy
+  public :: advect_base_dens, advect_base_enthalpy
 
 contains
 
@@ -131,229 +131,6 @@ contains
     end do
     
   end subroutine advect_base_dens_spherical
-
-!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-
-  subroutine advect_base_pres(w0,Sbar_in,p0_old,p0_new,gamma1bar,psi,psi_old,etarho_cc,s, &
-                              dt,dx,mla)
-
-    use bl_prof_module
-    use geometry, only: spherical, nlevs, nr_fine
-    use make_psi_module
-    use restrict_base_module
-    use multifab_module
-    use ml_layout_module
-
-    real(kind=dp_t), intent(in   ) ::        w0(:,0:)
-    real(kind=dp_t), intent(in   ) ::   Sbar_in(:,0:)
-    real(kind=dp_t), intent(in   ) ::    p0_old(:,0:)
-    real(kind=dp_t), intent(  out) ::    p0_new(:,0:)
-    real(kind=dp_t), intent(in   ) :: gamma1bar(:,0:)
-    real(kind=dp_t), intent(inout) ::       psi(:,0:)
-    real(kind=dp_t), intent(inout) ::   psi_old(:,0:)
-    real(kind=dp_t), intent(in   ) :: etarho_cc(:,0:)
-    type(multifab) , intent(in   ) :: s(:)
-    real(kind=dp_t), intent(in   ) :: dt
-    real(kind=dp_t), intent(in   ) :: dx(:,:)
-    type(ml_layout), intent(in   ) :: mla
-    
-    ! local
-    type(bl_prof_timer), save :: bpt
-
-    real(kind=dp_t) :: gamma1bar_nph(1,0:nr_fine-1)
-    real(kind=dp_t) :: p0_nph(1,0:nr_fine-1)
-
-    call build(bpt, "advect_base_pres")
-
-    if (spherical .eq. 0) then
-
-       ! make psi
-       call make_psi_planar(etarho_cc,psi)
-       call restrict_base(psi,.true.)
[…]

### 2009-07-02 UNIT_TESTS/test_basestate/varden.f90

@@ -319,41 +319,33 @@ subroutine varden()
      ! update pressure
      !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
 
-     if (p0_update_type .eq. 1) then
+     ! update pressure
+     p0_new = p0_old
+     call enforce_HSE(s0_new(:,:,rho_comp),p0_new,grav_cell)
 
-        call bl_error("We do not support p0_update_type = 1 for this example")
+     ! compute gamma1bar_new
+     do r=0,nr_fine-1
 
-     else
+        ! (rho, p) --> gamma1bar
+        den_eos(1)  = s0_new(1,r,rho_comp)
+        p_eos(1)    = p0_new(1,r)
+        xn_eos(1,:) = s0_new(1,r,spec_comp:spec_comp-1+nspec)/s0_new(1,r,rho_comp)
 
-        ! update pressure
-        p0_new = p0_old
-        call enforce_HSE(s0_new(:,:,rho_comp),p0_new,grav_cell)
+        temp_eos(1) = s0_old(1,r,temp_comp)
 
-        ! compute gamma1bar_new
-        do r=0,nr_fine-1
-           
-           ! (rho, p) --> gamma1bar
-           den_eos(1)  = s0_new(1,r,rho_comp)
-           p_eos(1)    = p0_new(1,r)
-           xn_eos(1,:) = s0_new(1,r,spec_comp:spec_comp-1+nspec)/s0_new(1,r,rho_comp)
-           
-           temp_eos(1) = s0_old(1,r,temp_comp)
-           
-           call eos(eos_input_rp, den_eos, temp_eos, NP, nspec, &
-                xn_eos, &
-                p_eos, h_eos, e_eos, &
-                cv_eos, cp_eos, xne_eos, eta_eos, pele_eos, &
-                dpdt_eos, dpdr_eos, dedt_eos, dedr_eos, &
-                dpdX_eos, dhdX_eos, &
-                gam1_eos, cs_eos, s_eos, &
-                dsdt_eos, dsdr_eos, &
-                do_diag)
-           
-           gamma1bar_new(1,r) = gam1_eos(1)
-           
-        end do
+        call eos(eos_input_rp, den_eos, temp_eos, NP, nspec, &
+             xn_eos, &
[…]

### 2009-07-07 Source/make_explicit_thermal.f90
to be consistent with our naming convention, made temp_diffusion_formulation an integer instead of a logical

@@ -15,7 +15,7 @@ module make_explicit_thermal_module
 contains 
 
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-! Compute thermal = del dot kappa grad T (if temp_diffusion_formulation)
+! Compute thermal = del dot kappa grad T (if temp_diffusion_formulation .eq. 1)
 ! Otherwise, compute thermal with grad h + grad X_k + grad p_0 formulation
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
 
@@ -69,7 +69,7 @@ contains
     !
     ! (alpha MINUS del dot beta grad) phi = RHS
 
-    if(temp_diffusion_formulation) then
+    if(temp_diffusion_formulation .eq. 1) then
 
        do n=1,nlevs
           call multifab_build(phi(n), mla%la(n), 1,  1)
@@ -108,7 +108,7 @@ contains
           call destroy(resid(n))
        enddo
 
-    else ! the if(.not. temp_diffusion_formulation) case
+    else ! if (temp_diffusion_formulation .eq. 2) case
        
        do n=1,nlevs
           call multifab_build(phi(n),  mla%la(n), 1,  1)
@@ -211,7 +211,7 @@ contains
           call destroy(resid(n))
        enddo
 
-    endif ! end if(temp_diffusion_formulation) logic
+    endif ! end temp_diffusion_formulation logic
     
     if (nlevs .eq. 1) then
 

### 2009-07-08 docs/architecture/architecture.tex
start of some notes on how MAESTRO works code-wise.  This is for people unfamiliar with boxlib stuff (like new students).

\section{MAESTRO Architecture}

\subsection{MAESTRO Directory Structure}

\begin{figure}[h]
\centering
\includegraphics[width=5.0in]{\archfigpath/maestro_directory2}
\caption{The basic MAESTRO directory structure}
\end{figure}

\subsection{BoxLib Data Structures}

\subsubsection{\boxtype}

A \boxtype\ is simply a rectangular domain in space.  The datatype for
a \boxtype\ is
\begin{verbatim}
  type box
     integer :: dim  = 0
     integer :: lo(MAX_SPACEDIM) =  Huge(1)
     integer :: hi(MAX_SPACEDIM) = -Huge(1)
  end type box
\end{verbatim}

\noindent Here:
\begin{itemize}
\item {\tt dim} is the dimensionality of the data
\item {\tt lo} is the index of the lower bound of the box
\item {\tt hi} is the index of the upper bound of the box
\end{itemize}

Note that boxes do not hold the state data themselves.

add description of index-space

\subsubsection{\fab}

A \fab\ is a ``Fortran Array Box''.  It contains the state data in a
multidimensional array and several \boxtype-types to describe where in
index-space it lives.  The datatype of a \fab\ is
\begin{verbatim}
  type fab
     integer   :: dim = 0
     type(box) :: bx
     type(box) :: pbx
     type(box) :: ibx
     integer   :: nc = 1
     real(kind = dp_t), pointer, dimension(:,:,:,:) :: p => Null()
  end type fab
\end{verbatim}

\noindent Here:
\begin{itemize}
\item {\tt dim} is the dimensionality of the data
\item {\tt bx} is a \boxtype\ describing where the data in this
  \fab\ is defined.
\item {\tt pbx} is a \boxtype\ that extends {\tt bx} to include
  any ghostcells
\item {\tt ibx}
\item {\tt nc} is the number of components (how many variables at  
  each grid location).
\item {\tt p} is a pointer to the data.
\end{itemize}

Note that alternate datatypes exist for data that is not
double-precision.

Also note that all state data is stored in a four-dimensional array,
{\tt (nx,ny,nz,nc)} in size, regardless of the dimensionality of the
problem.  For 2D problems, {\tt nz=1}.

\subsubsection{\multifab}

A \multifab\ is a collection of \fab s at the same level of
refinement.

\subsubsection{\boxarray}

\subsubsection{\mlboxarray}

\subsubsection{\mllayout}

\subsubsection{\bctower}

\subsection{Example: Accessing State Data}

In MAESTRO, the state data is stored in a multifab array (the array
index refers to the AMR level).  A typical way to extract the state data
array looks like:

\begin{verbatim}
  subroutine example(s,dx,dt)

    use bl_types
    use multifab_module
    use geometry, only: dm, nlevs
    use variables, only: rho_comp
\end{verbatim}

\noindent Here, the {\tt bl\_types} and {\tt multifab\_types} modules
bring in the basic boxlib data types Specifically, here, {\tt
bl\_types} defines {\tt dp\_t} which is the {\tt kind} used for
declaring double precision data, and {\tt multifab\_module} defines
the \multifab\ data type.  The {\tt geometry} module is a MAESTRO
module that provides the dimensionality ({\tt dm}) and the number of 
levels ({\tt nlevs}).  The {\tt variables} module is a MAESTRO
module that provides integer keys for indexing the state arrays.  In
this case the integer {\tt rho\_comp} refers to the location in the
state array corresponding to density.

\begin{verbatim}
    type(multifab) , intent(inout) :: s(:)
    real(kind=dp_t), intent(in   ) :: dx(:,:),dt
\end{verbatim}

\noindent Next we declare the subroutine arguments.  Here, {\tt s(:)}
is our \multifab\ array that holds the state data.  The array index
in {\tt s} refers to the AMR level.

\begin{verbatim}
    ! Local variables
    real(kind=dp_t), pointer :: sp(:,:,:,:)
    integer :: i,n,ng_sp
    integer :: lo(dm),hi(dm)
\end{verbatim}

\noindent Amongst the local variables we define here are a pointer,
{\tt sp}, that will point to a single \fab\ from the
\multifab\ {\tt s}.

\begin{verbatim}
    ng_sp = s(1)%ng
\end{verbatim}

\noindent Here we get the number of ghostcells for this particular
\multifab.  This will be needed to access the data stored in the
\fab s.  Note that all levels in a \multifab\ will have the same
number of ghostcells, so we can use {\tt s(1)} here.

\begin{verbatim}
    do n=1,nlevs

       do i = 1, s(n)%nboxes
\end{verbatim}

\noindent To access the data, we loop over all the levels, and
all the boxes in the given level.  {\tt s(n)\%nboxes} is simply
the number of boxes in level {\tt n}.

\begin{verbatim}
          if ( multifab_remote(s(n), i) ) cycle
          sp => dataptr(s(n), i)
          lo =  lwb(get_box(s(n), i))
          hi =  upb(get_box(s(n), i))
\end{verbatim}

\noindent For parallel runs, there is no guarantee that the particular
box is on the current processor.  The function {\tt multifab\_remote}
returns {\tt .true.} if the box is off-processor, in which case, we 
simply skip to the next box.

The actual data array is accessed through the {\tt dataptr} function,
which takes a \multifab\ (in this case, {\tt s(n)}) and the index of
the \boxtype\ ({\tt i}) we want.  {\tt sp} is always four-dimensional:
3 spatial dimensions and 1 for the components.

Finally, the index bounds of the box (just the data, not the ghostcells) are 
stored in the {\tt dm}-dimensional arrays {\tt lo} and {\tt hi}

\begin{verbatim}
          select case (dm)
          case (2)
             call example_2d(sp(:,:,1,rho_comp),ng_sp,lo,hi,dx(n,:),dt)
          case (3)
             call example_3d(sp(:,:,:,rho_comp),ng_sp,lo,hi,dx(n,:),dt)
          end select
       end do

    end do

  end subroutine example
\end{verbatim}

\noindent Finally, in this example, we call either the function
{\tt example\_2d} for two-dimensional data or {\tt example\_3d}
for three-dimensional data.  Note that in the two-dimensional
case, we index the data as {\tt sp(:,:,1,rho\_comp)}.  Here a
`{\tt 1}' is used as the `z'-coordinate spatial index, since this
is a 2D problem, and the density component of the state is selected
(using the integer key {\tt rho\_comp}).

This routine will be supplimented with {\tt example\_2d} and {\tt
example\_3d}, which actually operate on the data.  The form of 
these functions is:

\begin{verbatim}
  subroutine example_2d(density,ng_sp,lo,hi,dx,dt)

    use bl_constants_module
    use probin_module, only: prob_lo

    integer        , intent(in) :: lo(:),hi(:), ng_sp
    real(kind=dp_t), intent(in) :: density(lo(1)-ng_sp:,lo(2)-ng_sp:)
    real(kind=dp_t), intent(in) :: dx(:),dt

    integer         :: i, j
    real(kind=dp_t) :: x, y

    do j = lo(2), hi(2)
       y = prob_lo(2) + (dble(j) + HALF)*dx(2)

       do i = lo(1), hi(1)
          y = prob_lo(2) + (dble(j) + HALF)*dx(2)

          ! operate on density(i,j)

       enddo
    end do

  end subroutine mk_sponge_2d
\end{verbatim}

\noindent In this function, the bounds of the {\tt density} array
take into account the {\tt ng\_sp} ghostcells.  The {\tt j} and {\tt i}
loops loop over all the valid zones.  Coordinate information is 
computed from {\tt dx} and {\tt prob\_lo} which is the physical
lower bound of the domain.  {\tt bl\_constants\_module} declares
useful double-precision constants, like {\tt HALF} (0.5).

The three-dimensional case is similar, with the {\tt density} array
declared as {\tt density(lo(1)-ng\_sp:,lo(2)-ng\_sp:,lo(3)-ng\_sp)},
and an additional loop over the `z' coordinate (from {\tt lo(3)} to
{\tt hi(3)}).

\subsection{Working with Multifabs}

\subsection{Filling Ghostcells}

\subsection{Boundary Conditions}

\section{Adding A Problem}

### 2009-07-08 docs/architecture/architecture.tex
minor fixes

@@
     do n=1,nlevs
-
        do i = 1, s(n)%nboxes
@@
        do i = lo(1), hi(1)
-          y = prob_lo(2) + (dble(j) + HALF)*dx(2)
+          x = prob_lo(1) + (dble(i) + HALF)*dx(1)
 
@@
 The three-dimensional case is similar, with the {\tt density} array
-declared as {\tt density(lo(1)-ng\_sp:,lo(2)-ng\_sp:,lo(3)-ng\_sp)},
+declared as 
+\begin{verbatim}
+  density(lo(1)-ng_sp:,lo(2)-ng_sp:,lo(3)-ng_sp)
+\end{verbatim}
 and an additional loop over the `z' coordinate (from {\tt lo(3)} to

### 2009-07-08 docs/architecture/architecture.tex
describe the MAESTRO directory structure

@@
 \end{figure}
+
+All the files needed to build MAESTRO are contained in the {\tt
+fParallel} directory structure.  Briefly, the sub-directories
+are:
+\begin{itemize}
+\item {\tt boxlib/}
+
+ boxlib is a library for describing meshes consisting of a union
+ of boxes.  The boxlib modules define the basic datatypes used
+ in MAESTRO.
+
+\item {\tt data\_processing/}
+
+ Simple Fortran-based analysis routines (e.g.\ extract a line from a
+ multidimensional dataset) that operate on boxlib plotfiles.
+
+\item {\tt doc/}
+
+ Documentation describing the basic algorithm.
+
+\item {\tt extern/}
+
+ External modules, like ODE integrators, equations of state, reaction
+ networks
+
+\item {\tt MAESTRO/}
+
+ The main MAESTRO algorithm files.  Here you will find the main driver,
+ the advection routines, etc.
+
+ Important directories under {\tt MAESTRO/} include:
+
+ \begin{itemize}
+
+ \item {\tt initial\_models}
+
+   Several sub-directories containing routines to generate
+   1D initial conditions in hydrostatic equilibrium for 
+   mapping onto the MAESTRO grid.
+
+ \item problem directories: {\tt test}, {\tt test2}, {\tt
+   test\_spherical}, $\ldots$
+
+   Each problem in MAESTRO gets it own subdirectory.  The {\tt
+   GNUmakefile} includes the instructions on how to build the
+   executable, including what modules in {\tt extern} are used.
+
+   Any file that you place in a sub-directory here takes 
+   precedence over a file of the same name in {\tt MAESTRO/}.
+   This allows problems to have custom versions of the main
+   MAESTRO routines (e.g.\ initial conditions via {\tt initdata.f90}.
+
+ \end{itemize}
+
+\item {\tt mg/}
+
+  The multigrid solver.
+
+\item {\tt mk/}
+
+  The generic Makefiles that store the compilation flags for
+  various platforms.
+
+\item {\tt scripts/}
+
+  Some simple scripts that are useful for building, running,
+  maintaining MAESTRO.
+
+\end{itemize}
 

### 2009-07-09 docs/architecture/architecture.tex
a figure that shows boxes on a single level and the integer indices
reference figure

@@
 
-A \boxtype\ is simply a rectangular domain in space.  The datatype for
-a \boxtype\ is
+A \boxtype\ is simply a rectangular domain in space.  Note that boxes
+do not hold the state data themselves.  The datatype for a
+\boxtype\ is
 \begin{verbatim}
@@
 
-Note that boxes do not hold the state data themselves.
+\begin{figure}[h]
+\centering
+\includegraphics[width=4.0in]{\archfigpath/index_grid2}
+\caption{\label{fig:boxes} Three boxes that comprise a single level.
+At this resolution, the domain is 20$\times$18 zones.}
+\end{figure}
 
-add description of index-space
+The computational domain is divided into boxes.  The collection of
+boxes with the same resolution comprise a level.
+Figure~\ref{fig:boxes} shows three boxes in the same level of
+refinement.  The position of the boxes is with respect to a global
+index space at that level.  For example, box 1 in the figure has {\tt
+  lo} = (3,7) and {\tt hi} = (9,12).
 

### 2009-07-09 docs/architecture/architecture.tex
more...

@@
 A \multifab\ is a collection of \fab s at the same level of
-refinement.
+refinement.  The data type of a \multifab\ is:
+\begin{verbatim}
+  type multifab
+     integer :: dim = 0
+     integer :: nboxes = 0
+     integer :: nc = 1
+     integer :: ng = 0
+     logical, pointer :: nodal(:) => Null()
+     type(layout) :: la
+     type(fab), pointer :: fbs(:) => Null()
+  end type multifab
+\end{verbatim}
 
@@
 
+A \boxarray\ is an array of boxes
+
 \subsubsection{\mlboxarray}
+
+A \mlboxarray\ is a collection of \boxarray s at different levels of
+refinement.
 

### 2009-07-09 docs/architecture/architecture.tex
more intro + fix sectioning

@@
-\section{MAESTRO Architecture}
-
-\subsection{MAESTRO Directory Structure}
+\section{MAESTRO Directory Structure}
 
@@
 
-\subsection{BoxLib Data Structures}
+\section{BoxLib Data Structures}
 
-\subsubsection{\boxtype}
+MAESTRO's gridding is handled by the boxlib library.  The underlying
+idea in boxlib is to allow for adaptive mesh refinement.  The domain
+is broken into boxes.  At the coarsest (base) level of refinement, the
+entire computational domain is covered, either by one box or broken
+across many boxes.  Higher levels of refinement have finer zones
+(typically $2\times$).  Only a portion of the domain may be covered by
+the higher levels of refinement.  \MarginPar{Nesting?}
+For parallel computations, the boxes are spread across processors, in
+a fashion designed to put roughly equal amounts of work on each
+processor (load balancing).
+
+To simplify the description of the underlying AMR grid, boxlib
+provides a number of Fortran types.  We briefly summarize the major
+data types below.
+
+\subsection{\boxtype}
 
@@
 
-\subsubsection{\fab}
+\subsection{\fab}
 
@@
 
-\subsubsection{\multifab}
+\subsection{\multifab}
 
@@
 
-\subsubsection{\boxarray}
+\subsection{\boxarray}
 
@@
 
-\subsubsection{\mlboxarray}
+\subsection{\mlboxarray}
 
@@
 
-\subsubsection{\mllayout}
+\subsection{\mllayout}
 
-\subsubsection{\bctower}
+\subsection{\bctower}
 
-\subsection{Example: Accessing State Data}
+\section{Example: Accessing State Data}
 
@@
 
-\subsection{Working with Multifabs}
+\section{Working with Multifabs}
 
-\subsection{Filling Ghostcells}
+\section{Filling Ghostcells}
 
-\subsection{Boundary Conditions}
+\section{Boundary Conditions}
 

### 2009-07-09 docs/architecture/architecture.tex
fixed type multifab_types --> multifab_module

@@
 
-\noindent Here, the {\tt bl\_types} and {\tt multifab\_types} modules
+\noindent Here, the {\tt bl\_types} and {\tt multifab\_module} modules
 bring in the basic boxlib data types Specifically, here, {\tt

### 2009-07-09 docs/architecture/architecture.tex
define the acronym AMR change mk_spong_2d --> example_2d for consistency
stuff for the far future

@@
 MAESTRO's gridding is handled by the boxlib library.  The underlying
-idea in boxlib is to allow for adaptive mesh refinement.  The domain
-is broken into boxes.  At the coarsest (base) level of refinement, the
-entire computational domain is covered, either by one box or broken
-across many boxes.  Higher levels of refinement have finer zones
-(typically $2\times$).  Only a portion of the domain may be covered by
-the higher levels of refinement.  \MarginPar{Nesting?}
+idea in boxlib is to allow for adaptive mesh refinement---different
+regions of the domain can have different spatial resolutions.  The
+domain is broken into boxes.  At the coarsest (base) level of
+refinement, the entire computational domain is covered, either by one
+box or broken across many boxes.  Higher levels of refinement have
+finer zones (typically $2\times$).  Only a portion of the domain may
+be covered by the higher levels of refinement.  \MarginPar{Nesting?}
 For parallel computations, the boxes are spread across processors, in
@@
 \caption{\label{fig:boxes} Three boxes that comprise a single level.
-At this resolution, the domain is 20$\times$18 zones.}
+At this resolution, the domain is 20$\times$18 zones.  Note that the
+indexing in boxlib starts with $0$.}
 \end{figure}
@@
           end select
-       end do
+       enddo    ! end loop over boxes
 
-    end do
+    enddo    ! end loop over levels
 
@@
           ! operate on density(i,j)
+          ! ...
 
        enddo
-    end do
+    enddo
 
-  end subroutine mk_sponge_2d
+  end subroutine example_2d
 \end{verbatim}
@@
 
+\section{Multigrid}
+
+\section{MAESTRO Modules}
+
+\subsection{{\tt fill\_3d\_module}}
+
+\subsection{{\tt geometry}}
+
+\subsection{{\tt variables}}
+
 \section{Adding A Problem}
+
+\subsection{Plane-Parallel Geometry}
+
+\subsection{Full Star / Spherical}
+
+\subsection{Small-scale Flame}

### 2009-07-09 docs/architecture/architecture.tex
add a figure showing different data centerings

@@
 
+\subsubsection{Common Operations on a \boxtype}
+
+For a \boxtype\ declared as:
+\begin{verbatim}
+  type(box) :: mybox
+\end{verbatim}
+
+\begin{itemize}
+
+\item {\tt lo = lwb(mybox)} returns an array, {\tt lo(dm)}, with
+     the box lower bounds
+
+\item {\tt hi = upb(mybox)} returns an array, {\tt hi(dm)}, with
+     the box lower bounds
+
+\end{itemize}
+
 \subsection{\fab}
@@
   each grid location).
-\item {\tt p} is a pointer to the data.
+\item {\tt p} is a pointer to the data, in this case double precision
+  ({\tt kind=dp\_t}).  Other \fab\ types exists for other Fortran
+  data types.
 \end{itemize}
 
-Note that alternate datatypes exist for data that is not
-double-precision.
-
-Also note that all state data is stored in a four-dimensional array,
+Note that all state data is stored in a four-dimensional array,
 {\tt (nx,ny,nz,nc)} in size, regardless of the dimensionality of the
@@
 
+In MAESTRO, we don't usually deal with \fab s alone, but rather
+through \multifab s, described next.
+
 \subsection{\multifab}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=6.5in]{\archfigpath/data_loc2}
+\caption{\label{fig:dataloc} Some of the different data-centerings: (a) cell-centered, (b) nodal in the $x$-direction, and (c) nodal in both the $x$- and $y$-directions.}
+\end{figure}
 
@@
 
+\subsection{{\tt probin\_module}}
+
 \subsection{{\tt variables}}
@@
 \subsection{Small-scale Flame}
+
+\section{Visualization / Analysis}

### 2009-07-09 docs/architecture/architecture.tex
more stubs

@@
 processor (load balancing).
+
+On a grid, the data can be stored at cell-centers, on a face/edge, or
+on the corners.  In boxlib, data that is on an edge is termed `nodal'
+in that direction (see figure~\ref{fig:dataloc}.  Data that is on the
+corners is nodal in all spatial directions.  In MAESTRO, the state
+data (density, enthalpy, velocity, $\ldots$) is generally
+cell-centered.  Fluxes are nodal in the direction they represent.
+A few quantities are nodal in all directions (e.g.\ $\phi$ used in
+the final velocity projection).
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=6.5in]{\archfigpath/data_loc2}
+\caption{\label{fig:dataloc} Some of the different data-centerings:
+  (a) cell-centered, (b) nodal in the $x$-direction, and (c) nodal in
+  both the $x$- and $y$-directions.  Note that for nodal data, the
+  integer index corresponds to the lower boundary in that direction.
+  In each of these centerings, the red point has the same indices:\ (1,2).
+  Not show is the case where data is nodal in the $y$-direction only.}
+\end{figure}
 
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=6.5in]{\archfigpath/data_loc2}
-\caption{\label{fig:dataloc} Some of the different data-centerings: (a) cell-centered, (b) nodal in the $x$-direction, and (c) nodal in both the $x$- and $y$-directions.}
-\end{figure}
-
 A \multifab\ is a collection of \fab s at the same level of
@@
 \subsection{\bctower}
+
+\section{MAESTRO Data Organization}
+
+The state of the star in MAESTRO is described by both a
+multidimensional state and the 1D base state.  The full
+multidimensional state is stored in \multifab s while the base state
+is simply stored in Fortran arrays.
+
+\subsection{`{\tt s}' and `{\tt u}' \multifab s}
+
+\subsection{\tt umac}
+
+\subsection{Base State Arrays}
+
+\subsection{Other Quantities}
 
@@
 
-\section{MAESTRO Modules}
+\section{MAESTRO Helper Modules}
 
@@
 \subsection{{\tt variables}}
+
+The {\tt variables} module provides
 

### 2009-07-10 docs/architecture/architecture.tex
more about data

@@
 MAESTRO's gridding is handled by the boxlib library.  The underlying
-idea in boxlib is to allow for adaptive mesh refinement---different
+idea in boxlib is to allow for adaptive mesh refinement (AMR)---different
 regions of the domain can have different spatial resolutions.  The
@@
 
-\subsection{`{\tt s}' and `{\tt u}' \multifab s}
+\subsection{`{\tt s}' and `{\tt u}' \multifab s (fluid state and velocity)}
 
-\subsection{\tt umac}
+The fluid state (density, enthalpy, species) are stored together in a
+multi-component \multifab, typically named {\tt sold}, {\tt s1}, {\tt
+s2}, or {\tt snew} (depending on which time-level it represents).
+
+\subsection{{\tt umac} (the MAC velocity)}
+
+In creating the advective fluxes, we need the velocity through the
+faces of the zone---the $x$-velocity on the $x$-edges, the
+$y$-velocity on the $y$-edges, etc (see figure~\ref{fig:mac}.  This
+type of velocity discretization is termed the MAC velocity (after the
+``marker-and-cell'' method for free boundaries in incompressible
+flows, Harlow and Welch 1965).
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=2.5in]{\archfigpath/mac2}
+\hspace{0.1in}
+\begin{minipage}[b]{3.8in}
+\caption{\label{fig:mac} The MAC grid for the velocity.  
+Here the $x$-velocity is on the $x$-edges (shown as the 
+blue points) and the $y$-velocity is on the $y$-edges
+(shown as the red points).
+}\ \\
+\end{minipage}
+\end{figure}
+
+At each level of refinement, this is done by making a \multifab\
+array where each of the {\tt dm} components is nodal in
+its respective direction:
+\begin{verbatim}
+  type(multifab) :: umac(dm)
+
+  do comp=1,dm
+     call multifab_build(umac(n,comp), mla%la(n),1,1, &
+                         nodal=edge_nodal_flag(comp,:))
+  enddo
+\end{verbatim}
 

### 2009-07-10 docs/architecture/architecture.tex
more arch stuff...

@@
 
-\subsection{`{\tt s}' and `{\tt u}' \multifab s (fluid state and velocity)}
+\subsection{`{\tt s}' \multifab s (fluid state)}
 
 The fluid state (density, enthalpy, species) are stored together in a
-multi-component \multifab, typically named {\tt sold}, {\tt s1}, {\tt
-s2}, or {\tt snew} (depending on which time-level it represents).
+cell-centered multi-component \multifab, typically named {\tt sold},
+{\tt s1}, {\tt s2}, or {\tt snew} (depending on which time-level it
+represents).  The enthalpy is stored as $(\rho h)$, and the species
+are stored as partial-densities $(\rho X_k)$.
+
+Individual state variables should be indexed using the {\tt variables}
+module, with {\tt rho\_comp} referring to the density, {\tt rhoh\_comp}
+referring to the enthalpy quantity $(\rho h)$, and {\tt spec\_comp}
+referring to the start of the species (partial densities).
+
+To find a particular species, a query can be made through the {\tt network}
+module, such as:
+\begin{verbatim}
+  ic12 = network_species_index("carbon-12")
+\end{verbatim}
+and then the \fab\ can be indexed using {\tt spec\_comp-1+ic12} to
+get ``carbon-12''.
+
+\subsection{`{\tt u}' \multifab s (fluid velocity)}
+
+The fluid velocity at time-levels $n$ and $n+1$ is stored in
+a cell-centered multi-component \multifab, typically named
+{\tt uold} or {\tt unew}.  Here the {\tt dm}
+components correspond to each coordinate direction.
 
@@
 
-In creating the advective fluxes, we need the velocity through the
-faces of the zone---the $x$-velocity on the $x$-edges, the
-$y$-velocity on the $y$-edges, etc (see figure~\ref{fig:mac}.  This
+In creating the advective fluxes, we need the time-centered velocity
+through the faces of the zone---the $x$-velocity on the $x$-edges, the
+$y$-velocity on the $y$-edges, etc.\ (see figure~\ref{fig:mac}).  This
 type of velocity discretization is termed the MAC velocity (after the
@@
 
-At each level of refinement, this is done by making a \multifab\
-array where each of the {\tt dm} components is nodal in
-its respective direction:
+The MAC velocities are allocated at each level of refinement,{\tt n},
+by making a \multifab\ array where each of the {\tt dm} components is
+nodal in its respective direction:
 \begin{verbatim}

### 2009-07-13 Source/make_edge_state.f90
starting to implement some of the ppm changes that withstand the flame test problem

@@ -295,11 +295,6 @@ contains
                     diff3 = s(n,r+1)-s(n,r)
                     diff4 = s(n,r)-s(n,r-1)
 
-                    if (abs(diff1) .le. rel_eps) diff1 = ZERO
-                    if (abs(diff2) .le. rel_eps) diff2 = ZERO
-                    if (abs(diff3) .le. rel_eps) diff3 = ZERO
-                    if (abs(diff4) .le. rel_eps) diff4 = ZERO
-
                     if (diff1*diff2 .le. ZERO .or. &
                          diff3*diff4 .le. ZERO ) then
                        s6 = SIX*s(n,r) - THREE*(sm(n,r)+sp(n,r))
@@ -321,17 +316,17 @@ contains
                     else
                        alphap = sp(n,r)-s(n,r)
                        alpham = sm(n,r)-s(n,r)
-                       if (abs(alphap) .ge. TWO*abs(alpham)) then
+                       if (abs(alphap) .gt. TWO*abs(alpham)) then
                           dI = -alphap**2 / (FOUR*(alphap+alpham))
                           ds = s(n,r+1)-s(n,r)
-                          sgn = sign(ONE,s(n,r+1)-s(n,r-1))
+                          sgn = sign(ONE,alpham)
                           if (sgn*dI .ge. sgn*ds) then
                              sp(n,r) = s(n,r) - (TWO*ds + TWO*sgn*sqrt(ds**2 - ds*alpham))
                           end if
-                       else if (abs(alpham) .ge. TWO*abs(alphap)) then
+                       else if (abs(alpham) .gt. TWO*abs(alphap)) then
                           dI = -alpham**2 / (FOUR*(alphap+alpham))
                           ds = s(n,r-1)-s(n,r)
-                          sgn = sign(ONE,s(n,r+1)-s(n,r-1))
+                          sgn = sign(ONE,alphap)
                           if (sgn*dI .ge. sgn*ds) then
                              sm(n,r) = s(n,r) - (TWO*ds + TWO*sgn*sqrt(ds**2 - ds*alphap))
                           end if

### 2009-07-13 Source/ppm.f90

@@ -153,11 +153,6 @@ contains
              diff3 = s(i+1,j)-s(i,j)
              diff4 = s(i,j)-s(i-1,j)
 
-             if (abs(diff1) .le. rel_eps) diff1 = ZERO
-             if (abs(diff2) .le. rel_eps) diff2 = ZERO
-             if (abs(diff3) .le. rel_eps) diff3 = ZERO
-             if (abs(diff4) .le. rel_eps) diff4 = ZERO
-
              if (diff1*diff2 .le. ZERO .or. &
                  diff3*diff4 .le. ZERO ) then
                 s6 = SIX*s(i,j) - THREE*(sm(i,j)+sp(i,j))
@@ -179,17 +174,17 @@ contains
              else
                 alphap = sp(i,j)-s(i,j)
                 alpham = sm(i,j)-s(i,j)
-                if (abs(alphap) .ge. TWO*abs(alpham)) then
+                if (abs(alphap) .gt. TWO*abs(alpham)) then
                    dI = -alphap**2 / (FOUR*(alphap+alpham))
                    ds = s(i+1,j)-s(i,j)
-                   sgn = sign(ONE,s(i+1,j)-s(i-1,j))
+                   sgn = sign(ONE,alpham)
                    if (sgn*dI .ge. sgn*ds) then
                       sp(i,j) = s(i,j) - (TWO*ds + TWO*sgn*sqrt(ds**2 - ds*alpham))
                    end if
-                else if (abs(alpham) .ge. TWO*abs(alphap)) then
+                else if (abs(alpham) .gt. TWO*abs(alphap)) then
                    dI = -alpham**2 / (FOUR*(alphap+alpham))
                    ds = s(i-1,j)-s(i,j)
-                   sgn = sign(ONE,s(i+1,j)-s(i-1,j))
+                   sgn = sign(ONE,alphap)
                    if (sgn*dI .ge. sgn*ds) then
                       sm(i,j) = s(i,j) - (TWO*ds + TWO*sgn*sqrt(ds**2 - ds*alphap))
                    end if
@@ -368,11 +363,6 @@ contains
              diff3 = s(i,j+1)-s(i,j)
              diff4 = s(i,j)-s(i,j-1)
 
-             if (abs(diff1) .le. rel_eps) diff1 = ZERO
-             if (abs(diff2) .le. rel_eps) diff2 = ZERO
-             if (abs(diff3) .le. rel_eps) diff3 = ZERO
[…]

### 2009-07-22 docs/FAQ/faq.tex
made use_eos_coulomb=.true. by default in probin
made do_initial_projection a logical instead of int.

modified do_initial_projection in all inputs files
add another handy tip

@@
     init_divu_iter = 0
-    do_initial_projection = 0
+    do_initial_projection = T
 \end{verbatim}
@@
 
+\item {\em How can I tell from a plotfile what runtime parameters were
+   used for its run? or when it was created?}
+
+   In each plotfile directory, there is a file called {\tt job\_info}
+   (e.g.\ {\tt plt00000/job\_info}) that lists the build directory and
+   date, as well as the value of every runtime parameter for the run.
+
 \end{enumerate}

### 2009-07-23 docs/FAQ/faq.tex
add the .dump_restart

@@
+\section{Compiling}
+
+\begin{enumerate}
+
+\item {\em Does it matter if we use a Fortran 90 compiler instead of a 
+   Fortran 95 compiler?}
+
+   On some machines it does---Fortran 95 specifies that any local
+   allocated arrays are automatically deallocated when a subroutine
+   ends.  Fortran 90 does not do this.  Most MAESTRO routines rely 
+   on this Fortran 95 feature.
+
+   On IBM machines, for instance, using {\tt xlf90} instead of {\tt
+     xlf95} for MAESTRO will produce memory leaks.
+
+\end{enumerate}
+
+\section{Running}
+
 \begin{enumerate}
@@
 
+\end{enumerate}
+
+\section{I/O}
+
+\begin{enumerate}
+
 \item {\em How can we dump out a variable to a plotfile from any point in the
@@
 
-\item {\em Does it matter if we use a Fortran 90 compiler instead of a 
-   Fortran 95 compiler?}
-
-   On some machines it does---Fortran 95 specifies that any local
-   allocated arrays are automatically deallocated when a subroutine
-   ends.  Fortran 90 does not do this.  On IBM machines, for instance,
-   using {\tt xlf90} instead of {\tt xlf95} for MAESTRO will produce
-   memory leaks.
-
 \item {\em How can I tell from a plotfile what runtime parameters were
@@
 
+\item {\em How can I force the code to output a plotfile at the next
+   step?}
+
+   In the output directory (where the code is running) do `{\tt touch
+     .dump\_plotfile}'.  This will create an empty file called {\tt
+     .dump\_plotfile}.  At the end of each step, if the code finds
+   that file, it will output a plotfile.  Simply delete the file to
+   restore the code to its normal plotfile behavior.
+
 \end{enumerate}

### 2009-07-30 docs/runtime_parameters/runtime_parameters.tex
removed p0_update_type=1 code
to be consistent with our naming convention, made temp_diffusion_formulation an integer instead of a logical
updated with rp.py

@@
 &  1 \\
+\verb=  burner_threshold_cutoff  = &   Mass fraction cutoff for burner\_threshold\_species  used in burner threshold  &  1.e-10 \\
+\verb=  burner_threshold_species  = &   Name of the species to be used in burner threshold  &  "" \\
 \verb=  cflfac  = &   CFL factor to use in the computation of the advection timestep constraint  &  0.5 \\
@@
 \verb=  do_initial_projection  = &   Do the initial projection.  &  .true. \\
+\verb=  do_smallscale  = &   force $\rho_0 = (\rho h)_0 = 0$, {\tt evolve\_base\_state = F} and {\tt beta\_type} = 3  &  .false. \\
 \verb=  do_sponge  = &   Use sponging.  &  .false. \\
@@
 \verb=  max_grid_size  = &   The largest grid size that will be created using make\_new\_grids.  &  64 \\
+\verb=  max_grid_size_base  = &   The largest grid size that will be created using make\_new\_grids for the coarsest level.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
 \verb=  max_levs  = &   Total number of levels.  1 = single level.  &  1 \\
@@
 \verb=  n_cellz  = &   Number of cells for the base level in the z-direction  &  -1 \\
-\verb=  p0_update_type  = &   How we update the base state pressure and compute psi. \newline 1 = call {\tt advect\_base\_pres}; \newline 2 = call {\tt enforce\_HSE}.  &  2 \\
 \verb=  perturb_model  = &   Turn on a perturbation in the initial data.  Problem specific.  &  .false. \\
@@
 \verb=  stop_time  = &   simulation stop time  &  -1.d0 \\
-\verb=  temp_diffusion_formulation  = &   How to compute the explicit thermal diffusion term.  \newline .true. = in terms of $T$; \newline .false. = in terms of $\rho,p_0,X$.  &  .false. \\
+\verb=  temp_diffusion_formulation  = &   How to compute the explicit thermal diffusion term.  \newline 1 = in terms of $T$; \newline 2 = in terms of $\rho,p_0,X$.  &  2 \\
 \verb=  test_set  = &   Fixed grid file.  &  '' \\

### 2009-08-03 docs/thermo_notes/paper.tex
split-up for inclusion into the master document

@@
 
-  \section{Derivatives With Respect to Composition}
-  \label{Sec:Derivatives With Respect to Composition}
-  In the following we assume that the molar mass of species $i$ is given by its
-  atomic mass number, $A_i = N_i + Z_i$ where $N_i$ is the number of neutrons
-  and $Z_i$ is the number of protons in the isotope.  \MarginPar{Note that this
-    is common practice and that our EOS also makes this approximation.}  This 
-  is a slight approximation that ignores the mass difference between protons 
-  and neutrons as well as some minor binding energy terms.
-
-  The number density [cm$^-{3}$] of isotope $i$ is can be formed from the mass 
-  density and the molar mass of that isotope as follows:
-  \begin{equation}\label{eq:number density}
-    n_i = \frac{\rho_i N_\text{A}}{A_i},
-  \end{equation}
-  where $N_\text{A}$ is Avogadro's number [\# / mole].  The molar abundance, 
-  $Y_i$, is a measure of the number of moles of species $i$ per gram in the 
-  system:
-  \begin{equation}\label{eq:molar abundance}
-    Y_i = \frac{n_i}{\rho N_\text{A}} = \frac{\rho_i}{\rho}\frac{1}{A_i} 
-    \equiv \frac{X_i}{A_i}
-  \end{equation}
-  where we have defined the mass fraction, $X_i = \frac{\rho_i}{\rho}$.  Note
-  \begin{equation}\label{eq:mass fraction sums to 1}
-    \sum_i X_i = 1.
-  \end{equation}
-
-  We write the average molar mass and average proton number as:
-  \begin{align}\label{eq:abar}
-    \bar{A} &= \frac{\sum_i A_i n_i}{\sum_i n_i} = \left(\sum_i X_i\right)
-    \left(\sum_i \frac{X_i}{A_i}\right)^{-1}\\
-    \label{eq:zbar}
-    \bar{Z} &= \frac{\sum_i Z_i n_i}{\sum_i n_i} = \left(\sum_i Z_i
-    \frac{X_i}{A_i}\right)\left(\sum_i \frac{X_i}{A_i}\right)^{-1}.
-  \end{align}
-  Our algorithm requires terms involving the derivative thermodynamic variables
-  ($p$ or $e$, e.g.) with respect to composition.  Our EOS does not return such
-  derivatives but instead returns derivatives of these variables with respect 
-  to $\bar{A}$ and $\bar{Z}$.  Using the chain rule, we have
-  \begin{equation}\label{eq:p_xk}
-    \frac{\ptl p}{\ptl X_k} = p_{X_k} = 
-    \frac{\ptl p}{\ptl \bar{A}}\frac{\ptl \bar{A}}{\ptl X_k} + 
-    \frac{\ptl p}{\ptl \bar{Z}}\frac{\ptl \bar{Z}}{\ptl X_k}.
-  \end{equation}
-  From \eqref{eq:abar} and \eqref{eq:zbar} we have
-  \begin{align}\label{eq:abar_X_k}
-    \frac{\ptl \bar{A}}{\ptl X_k} &= \left(\sum_i \frac{X_i}{A_i}\right)^{-1}
-    - \frac{\bar{A}^2}{A_k} = -\frac{\bar{A}}{A_k}\left(\bar{A} - A_k\right)\\
-    \label{eq:zbar_X_k}
-    \frac{\ptl \bar{Z}}{\ptl X_k} &= 
-    \left(\frac{Z_k}{A_k}\right)\left(\sum_i \frac{X_i}{A_i}\right)^{-1}
-    - \frac{\bar{Z}}{A_k}\left(\sum_i \frac{X_i}{A_i}\right)^{-1} = 
-    -\frac{\bar{A}}{A_k}\left(\bar{Z} - Z_k\right),
-  \end{align}
-  where after differentiation we have used \eqref{eq:mass fraction sums to 1}
-  to write
-  \[
-  \left(\sum_i \frac{X_i}{A_i}\right)^{-1} = \bar{A}.
-  \]
-  We therefore have
-  \begin{equation}\label{eq:p_Xk_full}
-    p_{X_k} = -\frac{\bar{A}}{A_k}\left(\bar{A} - A_k\right)
-    \frac{\ptl p}{\ptl\bar{A}} - \frac{\bar{A}}{A_k}\left(\bar{Z} - Z_k\right)
-    \frac{\ptl p}{\ptl\bar{Z}}.
-  \end{equation}
-  Before it was brought to our attention by Frank Timmes, we were missing the
-  second term in \eqref{eq:abar_X_k}.  The only place where such terms 
-  appear in our algorithm is in a sum over all species, such as:
-  \begin{equation}\label{eq:sum over species}
-    \sum_i p_{X_i}\dot{\omega}_i = 
-    -\bar{A}^2\frac{\ptl p}{\ptl\bar{A}}\sum_i \frac{\dot{\omega}_i}{A_i}
-    +\bar{A}\frac{\ptl p}{\ptl\bar{A}}\sum_i \dot{\omega}_i
-    -\bar{A}\bar{Z}\frac{\ptl p}{\ptl\bar{Z}}\sum_i \frac{\dot{\omega}_i}{A_i}
-    +\bar{A}\frac{\ptl p}{\ptl \bar{Z}}\sum_i\frac{Z_i}{A_i}\dot{\omega}_i.
-  \end{equation}
-  \MarginPar{Note that if we are dealing with symmetric nuclei 
-  (as we have been up through paper III) where $Z_k = N_k = \half A_k$ for all
-  $k$, then the last term in \eqref{eq:sum over species} is identically zero
-  as well.}
-  The second term in \eqref{eq:sum over species} is identically zero because
-  \[
-  \sum_k \dot{\omega}_k \equiv 0.
-  \]
-  This second term arises from what was added to \eqref{eq:abar_X_k} by 
-  Frank's correction.  Therefore, although important for individual derivatives
-  with respect to composition, this correction term has no effect on our 
-  solution.
+\input thermo_notes
 

### 2009-08-03 docs/thermo_notes/thermo_notes.tex
split-up for inclusion into the master document

  \section{Derivatives With Respect to Composition}
  \label{Sec:Derivatives With Respect to Composition}
  In the following we assume that the molar mass of species $i$ is given by its
  atomic mass number, $A_i = N_i + Z_i$ where $N_i$ is the number of neutrons
  and $Z_i$ is the number of protons in the isotope.  \MarginPar{Note that this
    is common practice and that our EOS also makes this approximation.}  This 
  is a slight approximation that ignores the mass difference between protons 
  and neutrons as well as some minor binding energy terms.

  The number density [cm$^-{3}$] of isotope $i$ is can be formed from the mass 
  density and the molar mass of that isotope as follows:
  \begin{equation}\label{eq:number density}
    n_i = \frac{\rho_i N_\text{A}}{A_i},
  \end{equation}
  where $N_\text{A}$ is Avogadro's number [\# / mole].  The molar abundance, 
  $Y_i$, is a measure of the number of moles of species $i$ per gram in the 
  system:
  \begin{equation}\label{eq:molar abundance}
    Y_i = \frac{n_i}{\rho N_\text{A}} = \frac{\rho_i}{\rho}\frac{1}{A_i} 
    \equiv \frac{X_i}{A_i}
  \end{equation}
  where we have defined the mass fraction, $X_i = \frac{\rho_i}{\rho}$.  Note
  \begin{equation}\label{eq:mass fraction sums to 1}
    \sum_i X_i = 1.
  \end{equation}

  We write the average molar mass and average proton number as:
  \begin{align}\label{eq:abar}
    \bar{A} &= \frac{\sum_i A_i n_i}{\sum_i n_i} = \left(\sum_i X_i\right)
    \left(\sum_i \frac{X_i}{A_i}\right)^{-1}\\
    \label{eq:zbar}
    \bar{Z} &= \frac{\sum_i Z_i n_i}{\sum_i n_i} = \left(\sum_i Z_i
    \frac{X_i}{A_i}\right)\left(\sum_i \frac{X_i}{A_i}\right)^{-1}.
  \end{align}
  Our algorithm requires terms involving the derivative thermodynamic variables
  ($p$ or $e$, e.g.) with respect to composition.  Our EOS does not return such
  derivatives but instead returns derivatives of these variables with respect 
  to $\bar{A}$ and $\bar{Z}$.  Using the chain rule, we have
  \begin{equation}\label{eq:p_xk}
    \frac{\ptl p}{\ptl X_k} = p_{X_k} = 
    \frac{\ptl p}{\ptl \bar{A}}\frac{\ptl \bar{A}}{\ptl X_k} + 
    \frac{\ptl p}{\ptl \bar{Z}}\frac{\ptl \bar{Z}}{\ptl X_k}.
  \end{equation}
  From \eqref{eq:abar} and \eqref{eq:zbar} we have
  \begin{align}\label{eq:abar_X_k}
    \frac{\ptl \bar{A}}{\ptl X_k} &= \left(\sum_i \frac{X_i}{A_i}\right)^{-1}
    - \frac{\bar{A}^2}{A_k} = -\frac{\bar{A}}{A_k}\left(\bar{A} - A_k\right)\\
    \label{eq:zbar_X_k}
    \frac{\ptl \bar{Z}}{\ptl X_k} &= 
    \left(\frac{Z_k}{A_k}\right)\left(\sum_i \frac{X_i}{A_i}\right)^{-1}
    - \frac{\bar{Z}}{A_k}\left(\sum_i \frac{X_i}{A_i}\right)^{-1} = 
    -\frac{\bar{A}}{A_k}\left(\bar{Z} - Z_k\right),
  \end{align}
  where after differentiation we have used \eqref{eq:mass fraction sums to 1}
  to write
  \[
  \left(\sum_i \frac{X_i}{A_i}\right)^{-1} = \bar{A}.
  \]
  We therefore have
  \begin{equation}\label{eq:p_Xk_full}
    p_{X_k} = -\frac{\bar{A}}{A_k}\left(\bar{A} - A_k\right)
    \frac{\ptl p}{\ptl\bar{A}} - \frac{\bar{A}}{A_k}\left(\bar{Z} - Z_k\right)
    \frac{\ptl p}{\ptl\bar{Z}}.
  \end{equation}
  Before it was brought to our attention by Frank Timmes, we were missing the
  second term in \eqref{eq:abar_X_k}.  The only place where such terms 
  appear in our algorithm is in a sum over all species, such as:
  \begin{equation}\label{eq:sum over species}
    \sum_i p_{X_i}\dot{\omega}_i = 
    -\bar{A}^2\frac{\ptl p}{\ptl\bar{A}}\sum_i \frac{\dot{\omega}_i}{A_i}
    +\bar{A}\frac{\ptl p}{\ptl\bar{A}}\sum_i \dot{\omega}_i
    -\bar{A}\bar{Z}\frac{\ptl p}{\ptl\bar{Z}}\sum_i \frac{\dot{\omega}_i}{A_i}
    +\bar{A}\frac{\ptl p}{\ptl \bar{Z}}\sum_i\frac{Z_i}{A_i}\dot{\omega}_i.
  \end{equation}
  \MarginPar{Note that if we are dealing with symmetric nuclei 
  (as we have been up through paper III) where $Z_k = N_k = \half A_k$ for all
  $k$, then the last term in \eqref{eq:sum over species} is identically zero
  as well.}
  The second term in \eqref{eq:sum over species} is identically zero because
  \[
  \sum_k \dot{\omega}_k \equiv 0.
  \]
  This second term arises from what was added to \eqref{eq:abar_X_k} by 
  Frank's correction.  Therefore, although important for individual derivatives
  with respect to composition, this correction term has no effect on our 
  solution.

### 2009-08-21 docs/analysis/analysis.tex
documentation of some of the analysis routines in data_processing

The {\tt fParallel/data\_processing/} directory contains a large
number of Fortran-based analysis routines for BoxLib datasets.  Many
of these can be used with both MAESTRO and the compressible
astrophysics code, Castro.

\section{General Analysis Routines}

The following routines are generally applicable for any BoxLib-based
plotfile.  Typing the executable names without any arguments will
provide usage examples.

\begin{itemize}

\item {\tt faverage2d.f90}, {\tt faverage3d.f90}

  Laterally average each of the variables in a plotfile (separate
  routines exist for 2-d and 3-d).  This is written with MAESTRO
  plane-paralle geometry plotfiles in mind, and the averaging is done
  over the coordinate direction(s) perpendicular to gravity.

\item {\tt fcompare.f90}

  Compare two plotfiles, zone-by-zone to machine precision, and report
  the L2-norm of the error (both absolute and relative) for each
  variable.  This assumes that the grids are identical. \\[-3mm]

  This is used by in the regression test suite in {\tt
  Parallel/util/regtests/}.

\item {\tt fextract.f90}

  Extract a 1-d line through a dataset (1-, 2-, or 3-d).  This works
  with both uniformly-gridded or AMR datasets.  For multi-dimensional
  datasets, the coordinate direction to extract along can be specified.
  The line is always taken through the center of the domain.  Either
  a single variable or all variables, along with the coordinate 
  information, are output to a file.

\item {\tt fextrema.f90}

  Report the min and max of each variable (or only a single variable)
  in one or more plotfiles.

\item {\tt fIDLdump.f90}

  Create a binary file containing a uniformly gridded, single variable
  from a plotfile.  This is can then be read into IDL using the routines
  in {\tt fParallel/scripts/idlbl/}.

\item {\tt fsnapshot2d.f90}, {\tt fsnapshot3d.f90}

  Create an image (PPM file) of a single variable in a plotfile.  For
  3-d, the slice plane through the center of the domain is specified.
  Separate routines exist for 2-d and 3-d datasets.

\item {\tt ftime.f90}

  For each plotfile, simply print the simulation time.

\end{itemize}

### 2009-08-24 docs/runtime_parameters/runtime_parameters.tex
made s0_interp_type=3 the default

@@
 
-\verb=  anelastic_cutoff  = &   Refer to Section \ref{Sec:Anelastic Cutoff}.  &  3.e6 \\
-\verb=  base_cutoff_density  = &   Refer to Section \ref{Sec:Base Cutoff Density}  &  3.e6 \\
+\verb=  anelastic_cutoff  = &   The density below which we modify the constraint to look like the anelastic constraint, instead of the low Mach constraint.
+This prevents velocities from getting out of hand at the edge of the star.
+Refer to Section \ref{Sec:Anelastic Cutoff}.
+&  3.d6 \\
+\verb=  base_cutoff_density  = &   The density below which we keep the initial model constant. Refer to Section \ref{Sec:Base Cutoff Density}  &  3.d6 \\
 \verb=  bcx_hi  = &   $+x$ boundary condition (valid values are listed in {\tt  boxlib/bc.f90})  &  SLIP\_WALL \\
@@
 &  1 \\
-\verb=  burner_threshold_cutoff  = &   Mass fraction cutoff for burner\_threshold\_species  used in burner threshold  &  1.e-10 \\
+\verb=  buoyancy_cutoff_factor  = &   The multiplicative factor (over base\_cutoff\_density) below which we do zero out the buoyancy term in the momentum equation.  &  5.0 \\
+\verb=  burner_threshold_cutoff  = &   Mass fraction cutoff for burner\_threshold\_species  used in burner threshold  &  1.d-10 \\
 \verb=  burner_threshold_species  = &   Name of the species to be used in burner threshold  &  "" \\
-\verb=  cflfac  = &   CFL factor to use in the computation of the advection timestep constraint  &  0.5 \\
+\verb=  cflfac  = &   CFL factor to use in the computation of the advection timestep constraint  &  0.5d0 \\
 \verb=  cg_verbose  = &   Verbosity of bottom solver  &  0 \\
@@
 \verb=  do_sponge  = &   Use sponging.  &  .false. \\
-\verb=  dpdt_factor  = &   factor in front of the volume discrepancy term (0.0 = off)  &  0.0d0 \\
+\verb=  dpdt_factor  = &   factor in front of the volume discrepancy term (0.0 = off)  &  0.d0 \\
 \verb=  drdxfac  = &   ratio of radial base state zones to Cartesian full state zones for spherical geometry  &  1 \\
@@
 \verb=  init_iter  = &   Number of initial pressure iterations.  &  4 \\
-\verb=  init_shrink  = &   the multiplicative factor ($\le 1$) to reduce the initial timestep as computed by the various timestep estimators  &  1.0 \\
+\verb=  init_shrink  = &   the multiplicative factor ($\le 1$) to reduce the initial timestep as computed by the various timestep estimators  &  1.d0 \\
 \verb=  job_name  = &   job name printed in output  &  "" \\
@@
 \\
-\verb=  max_dt_growth  = &   The maximum scale factor that the time step is allowed to grow by per time step.  &  1.1D0 \\
+\verb=  max_dt_growth  = &   The maximum scale factor that the time step is allowed to grow by per time step.  &  1.1d0 \\
 \verb=  max_grid_size  = &   The largest grid size that will be created using make\_new\_grids.  &  64 \\
@@
 \verb=  plot_base_name  = &   prefix to use in plotfile file names  &  "plt" \\
-\verb=  plot_deltat  = &   rather than use a plot interva, plot a file after the solution has advanced past plot\_deltat in time  &  0.0 \\
+\verb=  plot_deltat  = &   rather than use a plot interva, plot a file after the solution has advanced past plot\_deltat in time  &  0.d0 \\
 \verb=  plot_int  = &   plot interval  &  0 \\
@@
 \verb=  ppm_type  = &   0 = no ppm \newline 1 = 1985 ppm \newline 2 = 2009 ppm  &  0 \\
-\verb=  prob_hi_x  = &   physical coordinates of hi-x corner of problem domain  &  1.0 \\
-\verb=  prob_hi_y  = &   physical coordinates of hi-y corner of problem domain  &  1.0 \\
-\verb=  prob_hi_z  = &   physical coordinates of hi-z corner of problem domain  &  1.0 \\
+\verb=  prob_hi_x  = &   physical coordinates of hi-x corner of problem domain  &  1.d0 \\
+\verb=  prob_hi_y  = &   physical coordinates of hi-y corner of problem domain  &  1.d0 \\
+\verb=  prob_hi_z  = &   physical coordinates of hi-z corner of problem domain  &  1.d0 \\
 \verb=  prob_lo_x  = &   physical coordinates of lo-x corner of problem domain  &  ZERO \\
@@
 \verb=  restart  = &   which file to restart from.  -1 means do not restart  &  -1 \\
-\verb=  rotation_radius  = &   radius used for computing centrifugal term in rotation problems  &  1.0e6\_dp\_t \\
+\verb=  rotation_radius  = &   radius used for computing centrifugal term in rotation problems  &  1.0d6 \\
 \verb=  rotational_frequency  = &   rotational frequency used for computing centrifugal term in rotation problems.  &  ZERO \\
-\verb=  s0_interp_type  = &   The interpolation for filling a cell-centered multifab from a 1D bin-centered array. \newline 1 = piecewise constant; \newline 2 = piecewise linear.  &  1 \\
+\verb=  s0_interp_type  = &   The interpolation for filling a cell-centered multifab from a 1D bin-centered array.
+\newline 1 = piecewise constant; \newline 2 = piecewise linear; \newline 3 = quadratic  &  1 \\
 \verb=  s0mac_interp_type  = &   The interpolation for filling an edge based multifab from a 1D bin-centered array.
-\newline 1 = Interpolate s0 to cell centers, then average to edges; \newline 2 = Interpolate s0 to edges directly;                \newline 3 = Interpolate s0 to nodes, then average to edges.
+\newline 1 = Interpolate s0 to cell centers (with s0\_interp\_type), then average to edges; \newline 2 = Interpolate s0 to edges directly using linear interpolation; \newline 3 = Interpolate s0 to edges directly using quadratic interpolation.
 &  1 \\
@@
 \verb=  verbose  = &   General verbosity  &  0 \\
+\verb=  w0_interp_type  = &   The interpolation for filling a cell-centered multifab from a 1D edge-centered array.
+\newline 1 = piecewise constant; \newline 2 = piecewise linear; \newline 3 = quadratic  &  2 \\
 \verb=  w0mac_interp_type  = &   The interpolation for putting w0 on edges.
 We only compute the normal component.
-\newline 1 = Interpolate w0 to cell centers, then average to edges; \newline 2 = Interpolate w0 to edges directly;                \newline 3 = Interpolate w0 to nodes, then average to edges.
+\newline 1 = Interpolate w0 to cell centers (with w0\_interp\_type), then average to edges; \newline 2 = Interpolate w0 to edges directly using linear interpolation; \newline 3 = Interpolate w0 to edges directly using quadratic interpolation; \newline 4 = Interpolate w0 to nodes using linear interpolation, then average to edges.
 &  1 \\

### 2009-08-26 Source/sponge.f90
in the middle of consolidating all of the sponge.f90 into one file (except for XRB)

@@ -3,6 +3,7 @@
 module sponge_module
 
   use bl_types
+  use bl_constants_module
   use multifab_module
   use ml_layout_module
 
@@ -19,9 +20,20 @@ contains
 
   subroutine init_sponge(rho0,dx,prob_lo_r)
 
-    use geometry, only: dr, r_end_coord
+    ! The sponge has a HALF * ( 1 - cos( (r - r_sp)/L)) profile, where
+    ! the width, L, is r_tp - r_sp.
+    !
+    ! The center of the sponge, r_md, is set to the radius where r =
+    ! sponge_center_density
+    !
+    ! The start of the sponge, r_sp, (moving outward from the center)
+    ! is the radius where r = sponge_start_factor * sponge_center_density
+    ! 
+    ! The top of the sponge is then 2 * r_md - r_tp
+
+    use geometry, only: dr, r_end_coord, spherical
     use bl_constants_module
-    use probin_module, only: anelastic_cutoff, prob_hi, verbose
+    use probin_module, only: verbose, sponge_start_factor, sponge_center_density
 
     real(kind=dp_t), intent(in   ) :: rho0(0:),prob_lo_r
     real(kind=dp_t), intent(in   ) :: dx(:)
@@ -33,35 +45,38 @@ contains
     r_top = prob_lo_r + dble(r_end_coord(1,1)+1) * dr(1)
     r_sp = r_top
 
+    ! set r_sp
     do r=0,r_end_coord(1,1)
        rloc = prob_lo_r + (dble(r)+HALF) * dr(1)
-       if (rho0(r) < 10.d0*anelastic_cutoff) then
+       if (rho0(r) < sponge_start_factor*sponge_center_density) then
           r_sp = rloc
           exit
        endif
     enddo
 
+    ! set r_md
     r_md = r_top
     do r=0,r_end_coord(1,1)
        rloc = prob_lo_r + (dble(r)+HALF) * dr(1)
-       if (rho0(r) < anelastic_cutoff) then
+       if (rho0(r) < sponge_center_density) then
           r_md = rloc
           exit
        endif
     enddo
 
-    r_tp = 2.d0 * r_md - r_sp
+    ! set r_tp
+    r_tp = TWO * r_md - r_sp
 
-    r_tp_outer = HALF * max(prob_hi(1),prob_hi(2)) 
-    if (size(prob_hi,dim=1) .eq. 2) then
[…]

### 2009-08-26 TEST_PROBLEMS/spherical_heat/sponge.f90

@@ -1,188 +1,0 @@
-! The sponge acts to damp the velocities at the edge of the star.
-
-module sponge_module
-
-  use bl_types
-  use bl_constants_module
-  use multifab_module
-  use ml_layout_module
-
-  implicit none
-
-  real(dp_t), save :: r_sp, r_md, r_tp
-  real(dp_t), save :: r_sp_outer, r_tp_outer
-
-  private
-
-  public :: init_sponge, make_sponge
-
-contains
-
-  subroutine init_sponge(rho0,dx,prob_lo_r)
-
-    ! The sponge has a HALF * ( 1 - cos( (r - r_sp)/L)) profile, where
-    ! the width, L, is r_tp - r_sp.
-    !
-    ! The center of the sponge, r_md, is set to the radius where r =
-    ! sponge_center_density
-    !
-    ! The start of the sponge, r_sp, (moving outward from the center)
-    ! is the radius where r = sponge_start_factor * sponge_center_density
-    ! 
-    ! The top of the sponge is then 2 * r_md - r_tp
-
-    use geometry, only: dr, r_end_coord
-    use bl_constants_module
-    use probin_module, only: anelastic_cutoff, prob_hi, verbose, &
-         sponge_start_factor, sponge_center_density
-
-    real(kind=dp_t), intent(in   ) :: rho0(0:),prob_lo_r
-    real(kind=dp_t), intent(in   ) :: dx(:)
-
-    real (kind = dp_t) :: rloc
-    real (kind = dp_t) :: r_top
-    integer            :: r
-
-    r_top = prob_lo_r + dble(r_end_coord(1,1)+1) * dr(1)
-    r_sp = r_top
-
-    ! set r_sp
-    do r=0,r_end_coord(1,1)
-       rloc = prob_lo_r + (dble(r)+HALF) * dr(1)
-       if (rho0(r) < sponge_start_factor*sponge_center_density) then
-          r_sp = rloc
-          exit
-       endif
-    enddo
-
-    ! set r_md
-    r_md = r_top
-    do r=0,r_end_coord(1,1)
-       rloc = prob_lo_r + (dble(r)+HALF) * dr(1)
-       if (rho0(r) < sponge_center_density) then
-          r_md = rloc
-          exit
-       endif
-    enddo
-
-    ! set r_tp
-    r_tp = TWO * r_md - r_sp
-
-    ! outer sponge parameters
-    r_sp_outer = r_tp
[…]

### 2009-08-26 TEST_PROBLEMS/test2_spherical/sponge.f90

@@ -1,191 +1,0 @@
-! The sponge acts to damp the velocities at the edge of the star.
-
-module sponge_module
-
-  use bl_types
-  use bl_constants_module
-  use multifab_module
-  use ml_layout_module
-
-  implicit none
-
-  real(dp_t), save :: r_sp, r_md, r_tp
-  real(dp_t), save :: r_sp_outer, r_tp_outer
-
-  private
-
-  public :: init_sponge, make_sponge
-
-contains
-
-  subroutine init_sponge(rho0,dx,prob_lo_r)
-
-    ! The sponge has a HALF * ( 1 - cos( (r - r_sp)/L)) profile, where
-    ! the width, L, is r_tp - r_sp.
-    !
-    ! The center of the sponge, r_md, is set to the radius where rho =
-    ! sponge_center_density
-    !
-    ! The start of the sponge, r_sp, (moving outward from the center)
-    ! is the radius where rho = sponge_start_factor * sponge_center_density
-    ! 
-
-    use geometry, only: dr, r_end_coord
-    use bl_constants_module
-    use probin_module, only: anelastic_cutoff, prob_hi, verbose, &
-         sponge_start_factor, sponge_center_density
-
-    real(kind=dp_t), intent(in   ) :: rho0(0:),prob_lo_r
-    real(kind=dp_t), intent(in   ) :: dx(:)
-
-    real (kind = dp_t) :: rloc
-    real (kind = dp_t) :: r_top
-    integer            :: r
-
-    r_top = prob_lo_r + dble(r_end_coord(1,1)+1) * dr(1)
-    r_sp = r_top
-
-    ! set r_sp
-    do r=0,r_end_coord(1,1)
-       rloc = prob_lo_r + (dble(r)+HALF) * dr(1)
-       if (rho0(r) < sponge_start_factor*sponge_center_density) then
-          r_sp = rloc
-          exit
-       endif
-    enddo
-
-    ! set r_md
-    r_md = r_top
-    do r=0,r_end_coord(1,1)
-       rloc = prob_lo_r + (dble(r)+HALF) * dr(1)
-       if (rho0(r) < sponge_center_density) then
-          r_md = rloc
-          exit
-       endif
-    enddo
-
-    ! set r_tp
-    r_tp = TWO * r_md - r_sp
-
-    ! outer sponge parameters
-    r_tp_outer = HALF * max(prob_hi(1),prob_hi(2)) 
-    if (size(prob_hi,dim=1) .eq. 2) then
[…]

### 2009-08-26 TEST_PROBLEMS/test_convect/sponge.f90

@@ -1,176 +1,0 @@
-! a module for storing the geometric information so we don't have to pass it
-
-module sponge_module
-
-  use bl_types
-  use multifab_module
-  use ml_layout_module
-
-  implicit none
-
-  real(dp_t), save :: r_sp, r_md, r_tp
-  real(dp_t), save :: r_sp_outer, r_tp_outer
-
-  private
-
-  public :: init_sponge, make_sponge
-
-contains
-
-  subroutine init_sponge(rho0,dx,prob_lo_r)
-
-    use probin_module, only: prob_hi
-
-    real(kind=dp_t), intent(in   ) :: rho0(0:),prob_lo_r
-    real(kind=dp_t), intent(in   ) :: dx(:)
-
-    r_sp = 2.19140625d8
-    r_tp = 2.97265625d8
-
-    if ( parallel_IOProcessor() ) write(6,1000) r_sp, r_tp
-
-1000 format('inner sponge: r_sp      , r_tp      : ',e20.12,2x,e20.12)
-
-  end subroutine init_sponge
-
-  subroutine make_sponge(sponge,dx,dt,mla)
-
-    use bl_constants_module
-    use ml_restriction_module, only: ml_cc_restriction
-    use geometry, only: dm, nlevs
-
-    type(multifab) , intent(inout) :: sponge(:)
-    real(kind=dp_t), intent(in   ) :: dx(:,:),dt
-    type(ml_layout), intent(in   ) :: mla
-
-    ! Local variables
-    real(kind=dp_t), pointer :: sp(:,:,:,:)
-    integer :: i,n,ng_sp
-    integer :: lo(dm),hi(dm)
-
-    ng_sp = sponge(1)%ng
-
-    do n=1,nlevs
-
-       do i = 1, sponge(n)%nboxes
-          if ( multifab_remote(sponge(n), i) ) cycle
-          sp => dataptr(sponge(n), i)
-          lo =  lwb(get_box(sponge(n), i))
-          hi =  upb(get_box(sponge(n), i))
-          select case (dm)
-          case (2)
-             call mk_sponge_2d(sp(:,:,1,1),ng_sp,lo,hi,dx(n,:),dt)
-          case (3)
-             call mk_sponge_3d(sp(:,:,:,1),ng_sp,lo,hi,dx(n,:),dt)
-          end select
-       end do
-
-    end do
-
-    ! the loop over nlevs must count backwards to make sure the finer grids are done first
-    do n=nlevs,2,-1
[…]

### 2009-08-26 TEST_PROBLEMS/test_eta_stuff/sponge.f90

@@ -1,214 +1,0 @@
-! a module for storing the geometric information so we don't have to pass it
- 
-module sponge_module
- 
-  use bl_types
-  use multifab_module
-  use ml_layout_module
- 
-  implicit none
- 
-  real(dp_t), save :: r_sp, r_md, r_tp
-  real(dp_t), save :: r_sp_outer, r_tp_outer
-
-  private
-
-  public :: init_sponge, make_sponge
- 
-contains
-
-  subroutine init_sponge (nlevs,rho0,dx,prob_lo_r)
-
-    use geometry, only: nr, dr
-    use bl_constants_module
-    use probin_module, only: anelastic_cutoff, prob_hi
-
-    integer        , intent(in   ) :: nlevs
-    real(kind=dp_t), intent(in   ) :: rho0(0:),prob_lo_r
-    real(kind=dp_t), intent(in   ) :: dx(:)
-
-    real (kind = dp_t) :: r
-    real (kind = dp_t) :: r_top
-    integer            :: j
-
-    r_top = prob_lo_r + dble(nr(nlevs)) * dr(nlevs)
-    r_sp = r_top
-
-    do j = 0, nr(nlevs)-1
-       r = prob_lo_r + (dble(j)+HALF) * dr(nlevs)
-       if (rho0(j) < 10.d0*anelastic_cutoff) then
-          r_sp = r
-          exit
-       endif
-    enddo
-
-    r_md = r_top
-    do j = 0,nr(nlevs)-1
-       r = prob_lo_r + (dble(j)+HALF) * dr(nlevs)
-       if (rho0(j) < anelastic_cutoff) then
-          r_md = r
-          exit
-       endif
-    enddo
-
-    r_tp = 2.d0 * r_md - r_sp
-
-    r_tp = 1.9e8
-    r_sp = 1.75e8
-
-    r_tp_outer = HALF * max(prob_hi(1),prob_hi(2))
-    if (size(prob_hi,dim=1) .eq. 2) then
-      r_sp_outer = r_tp_outer - 4.d0 * dx(2)
-    else
-      r_tp_outer = max(r_tp_outer, HALF * prob_hi(3))
-      r_sp_outer = r_tp_outer - 4.d0 * dx(3)
-    end if
-
-    if ( parallel_IOProcessor() ) write(6,1000) r_sp, r_tp
-    if ( parallel_IOProcessor() ) write(6,1001) r_sp_outer, r_tp_outer
-
[…]

### 2009-08-26 TEST_PROBLEMS/test_longtime_convect/sponge.f90

@@ -1,213 +1,0 @@
-! a module for storing the geometric information so we don't have to pass it
-
-module sponge_module
-
-  use bl_types
-  use multifab_module
-  use ml_layout_module
-
-  implicit none
-
-  real(dp_t), save :: r_sp, r_md, r_tp
-  real(dp_t), save :: r_sp_outer, r_tp_outer
-
-  private
-
-  public :: init_sponge, make_sponge
-
-contains
-
-  subroutine init_sponge(rho0,dx,prob_lo_r)
-
-    use geometry, only: dr, r_end_coord
-    use bl_constants_module
-    use probin_module, only: anelastic_cutoff, prob_hi
-
-    real(kind=dp_t), intent(in   ) :: rho0(0:),prob_lo_r
-    real(kind=dp_t), intent(in   ) :: dx(:)
-
-    real (kind = dp_t) :: rloc
-    real (kind = dp_t) :: r_top, r_hw
-    integer            :: r
-
-    r_top = prob_lo_r + dble(r_end_coord(1)+1) * dr(1)
-
-    r_hw = r_top
-    do r = 0, r_end_coord(1)
-       rloc = prob_lo_r + (dble(r)+HALF) * dr(1)
-       if (rho0(r) < 1.d7) then
-          r_hw = rloc
-          exit
-       endif
-    enddo
-
-    r_md = r_top
-    do r = 0,r_end_coord(1)
-       rloc = prob_lo_r + (dble(r)+HALF) * dr(1)
-       if (rho0(r) < anelastic_cutoff) then
-          r_md = rloc
-          exit
-       endif
-    enddo
-
-    r_tp = r_md + 2*(r_md - r_hw)
-    r_sp = r_md - 2*(r_md - r_hw)
-
-    r_tp_outer = HALF * max(prob_hi(1),prob_hi(2)) 
-    if (size(prob_hi,dim=1) .eq. 2) then
-       r_sp_outer = r_tp_outer - 4.d0 * dx(2)
-    else
-       r_tp_outer = max(r_tp_outer, HALF * prob_hi(3)) 
-       r_sp_outer = r_tp_outer - 4.d0 * dx(3)
-    end if
-
-    if ( parallel_IOProcessor() ) write(6,1000) r_sp, r_tp
-    if ( parallel_IOProcessor() ) write(6,1001) r_sp_outer, r_tp_outer
-
[…]

### 2009-08-28 docs/runtime_parameters/runtime_parameters.tex
in the middle of consolidating all of the sponge.f90 into one file (except for XRB)
make ppm_type=1 the default for all problems except xrb

@@
 \verb=  pmask_z  = &   indicates whether the problem domain periodic in the z-direction  &  .false. \\
-\verb=  ppm_type  = &   0 = no ppm \newline 1 = 1985 ppm \newline 2 = 2009 ppm  &  0 \\
+\verb=  ppm_type  = &   0 = no ppm \newline 1 = 1985 ppm \newline 2 = 2009 ppm  &  1 \\
 \verb=  prob_hi_x  = &   physical coordinates of hi-x corner of problem domain  &  1.d0 \\
@@
 \verb=  s0_interp_type  = &   The interpolation for filling a cell-centered multifab from a 1D bin-centered array.
-\newline 1 = piecewise constant; \newline 2 = piecewise linear; \newline 3 = quadratic  &  1 \\
+\newline 1 = piecewise constant; \newline 2 = piecewise linear; \newline 3 = quadratic  &  3 \\
 \verb=  s0mac_interp_type  = &   The interpolation for filling an edge based multifab from a 1D bin-centered array.
@@
 \verb=  spherical_in  = &   set to 1 if you are doing a spherical problem  &  0 \\
+\verb=  sponge_center_density  = &   Center of the inner sponge.  &  3.d6 \\
 \verb=  sponge_kappa  = &   Parameter for sponge.  Problem dependent.  &  10.d0 \\
+\verb=  sponge_start_factor  = &   The sponge begins at sponge\_center\_density * sponge\_start\_factor.  &  3.333d0 \\
 \verb=  stop_time  = &   simulation stop time  &  -1.d0 \\

### 2009-09-16 Source/make_psi.f90
we trigger the point where we turn off psi using base_cutoff_density rather than anelastic_cutoff_density.  now all parameters relevant to multilevel paper, except for div_coeff, cut off base_cutoff_density

@@ -12,7 +12,7 @@ contains
   subroutine make_psi_planar(etarho_cc,psi)
 
     use bl_constants_module
-    use geometry, only: anelastic_cutoff_coord, r_start_coord, r_end_coord, &
+    use geometry, only: base_cutoff_density_coord, r_start_coord, r_end_coord, &
          numdisjointchunks, nlevs_radial
     use probin_module, only: grav_const
     use restrict_base_module
@@ -28,7 +28,7 @@ contains
     do n=1,nlevs_radial
        do i=1,numdisjointchunks(n)
           do r = r_start_coord(n,i), r_end_coord(n,i)
-             if (r .lt. anelastic_cutoff_coord(n)) then
+             if (r .lt. base_cutoff_density_coord(n)) then
                 psi(n,r) = etarho_cc(n,r) * abs(grav_const)
              end if
           end do

### 2009-09-16 docs/lo_density/lo_density.tex
updated figure
we trigger the point where we turn off psi using base_cutoff_density rather than anelastic_cutoff_density.  now all parameters relevant to multilevel paper, except for div_coeff, cut off base_cutoff_density
typo

@@
 \caption{Image of how the cutoff density and cutoff coordinates
-are related.\vspace{2em}}
+are related.  This is drawn assuming {\tt drdxfac}=5.\vspace{2em}}
 \end{minipage}
@@
   \MarginPar{I think we want $r > {\tt anelastic\_cutoff\_coord}$.}
-
-\item In {\tt make\_psi} for plane-parallel, we only compute $\psi$ for 
-  $r < {\tt anelastic\_cutoff\_coord}$.
-  \MarginPar{I think we want to change this to $r \le {\tt anelastic\_cutoff\_coord}$.}
 
@@
 
+\item In {\tt make\_psi} for plane-parallel, we only compute $\psi$ for 
+  $r < {\tt base\_cutoff\_density\_coord}$.
+  \MarginPar{I think we want to change this to $r \le {\tt base\_cutoff\_density\_coord}$.}
+
 \end{itemize}

### 2009-09-17 docs/lo_density/lo_density.tex
if do_eos_h_above_cutoff=T, recompute h for rho <= base_cutoff_density rather than rho < base_cutoff_density
in short, now every instance of using a cutoff density in the code is consistent between the <= and < inequalities.

@@
 
+\section{Density vs. Coordinate}
+
+Note that saying $r\ge$ {\tt anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)\le$ {\tt anelastic\_cutoff}.
+Also, saying $r<$ {\tt anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)>$ {\tt anelastic\_cutoff}.
+Ditto for {\tt base\_cutoff\_density} and {\tt base\_cutoff\_density\_coord}.
+
 \section{anelastic\_cutoff}\label{Sec:Anelastic Cutoff}
@@
   ${\tt div\_coeff}(n,r) = {\tt div\_coeff}(n,r-1) * \rho_0(n,r)/\rho_0(n,r-1)$.
-  \MarginPar{I think we want $r > {\tt anelastic\_cutoff\_coord}$.}
 
@@
   to zero for $r \ge {\tt anelastic\_cutoff\_coord}$.
-  \MarginPar{I think we want $r > {\tt anelastic\_cutoff\_coord}$.}
 
@@
   (in spherical).
-  \MarginPar{For plane-parallel, I think we should use 
-    $r \le {\tt base\_cutoff\_density\_coord}$.}
 
@@
   setting it to $\rho_0 g$.
-  \MarginPar{I think we want $r > {\tt base\_cutoff\_density\_coord}$.}
 
-\item In {\tt update\_scal}, if $\rho < {\tt base\_cutoff\_density}$
+\item In {\tt update\_scal}, if $\rho \le {\tt base\_cutoff\_density}$
    and {\tt do\_eos\_h\_above\_cutoff}, we call the EOS to compute $h$.
-  \MarginPar{I think we want $\rho \le {\tt base\_cutoff\_density}$}
 
@@
   $\rho_0 > {\tt base\_cutoff\_density}$.
-  \MarginPar{I think this needs to be reworked to include mass up to the 
-    cell-centered location of {\tt base\_cutoff\_density\_coord} for both cell 
-    and edge gravity.}
 
-\item In {\tt enforce\_HSE}, we set $p_0(r) = p_0(r-1)$ for 
-  $r > {\tt base\_cutoff\_density\_coord}$.
+\item In {\tt enforce\_HSE}, we set $p_0(r+1) = p_0(r)$ for 
+  $r \ge {\tt base\_cutoff\_density\_coord}$.
 
@@
   $r < {\tt base\_cutoff\_density\_coord}$.
-  \MarginPar{I think we want to change this to $r \le {\tt base\_cutoff\_density\_coord}$.}
 

### 2009-09-17 Source/update_scal.f90

@@ -197,7 +197,7 @@ contains
        do j = lo(2), hi(2)
           do i = lo(1), hi(1)
              
-             if (snew(i,j,rho_comp) < base_cutoff_density) then
+             if (snew(i,j,rho_comp) .le. base_cutoff_density) then
                 den_eos(1) = snew(i,j,rho_comp)
                 temp_eos(1) = sold(i,j,temp_comp)
                 p_eos(1) = p0(j)
@@ -325,7 +325,7 @@ contains
           do j = lo(2), hi(2)
              do i = lo(1), hi(1)
 
-                if (snew(i,j,k,rho_comp) < base_cutoff_density) then
+                if (snew(i,j,k,rho_comp) .le. base_cutoff_density) then
                    den_eos(1) = snew(i,j,k,rho_comp)
                    temp_eos(1) = sold(i,j,k,temp_comp)
                    p_eos(1) = p0(k)
@@ -460,7 +460,7 @@ contains
           do j = lo(2), hi(2)
              do i = lo(1), hi(1)
 
-                if (snew(i,j,k,rho_comp) < base_cutoff_density) then
+                if (snew(i,j,k,rho_comp) .le. base_cutoff_density) then
                    den_eos(1) = snew(i,j,k,rho_comp)
                    temp_eos(1) = sold(i,j,k,temp_comp)
                    p_eos(1) = p0_new_cart(i,j,k)

### 2009-09-25 docs/eta_notes/eta.tex
updated

@@
 
-Recall that if initially $ \overline{\rhop} = 1 /
-{4\pi} \int_{\Omega_H} \rhop \; d\Omega = 0,$ 
-\MarginPar{this section is here temporarily, just to better understand
-what's going on here.  These notes came from an early version of
-paper~III} 
-there is no guarantee that $\overline{\rhop} = 0$ will hold at later
-time.  To see this we start with the equation for $\rho'$, 
+In Paper III, we introduced a mixing term, $\etarho$, to the base state density
+evolution equation (see eq. [29] of Paper III), with the objective of
+keeping the base state density equal to the average of the density
+over a layer.  Here we derive $\etarho$ for spherical problems.
+For this derivation, it is best to define the average in terms of spherical coordinates,
 \begin{equation}
-\frac{\partial \rhop}{\partial t} = - \nablab \cdotb (\rho_0 \Ubt)
--\nablab \cdotb (\rhop w_0 \er) - \nablab \cdotb (\rhop \Ubt )
-\label{eq:rhoprimeold} \enskip ,
+\overline{\phi(r)} = \frac{1}{4\pi} \int_{\Omega_H} \phi(r,\theta,\phi) d\Omega
 \end{equation}
-rewritten by moving one term to the left-hand-side:
+where $\int_{\Omega_H} d\Omega = 4\pi$ represents the integral over
+the angles $\theta$ and $\phi$ at constant radius.
+
+Recall from Paper III that if initially $\overline{\rho'} = 0$,
+there is no guarantee that $\overline{\rho'} = 0$ will hold at later
+time.  To see this, recall equation (\ref{eq:Perturbational Density}),
+written in a slightly different form:
 \begin{equation}
-\frac{\partial \rhop}{\partial t} + \nablab \cdotb (\rhop \Ubt) 
-= - \nablab \cdotb (\rho_0 \Ubt)
-  -\nablab \cdotb (\rhop w_0 \er) \enskip .
+\frac{\partial\rho'}{\partial t} + \nabla\cdot\left(\rho'\Ub\right) = -\nabla\cdot\left(\rho_0\Ubt\right).
 \end{equation}
-We integrate this over a spherical shell of thickness $2h$ at radius $r_0$, 
+We integrate this over a spherical shell of thickness $2h$ at radius $r_0$, i.e.,
 $\Omega_H \times (r_0-h, r_0+h)$, and normalize by the integration
-volume, which is $\sim 4\pi r_0^2 \, 2h$ for small $h$, to obtain:
+volume, which is $\sim 4\pi r_0^2  2h$ for small $h$, to obtain:
 \begin{eqnarray}
-\frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H} 
-    \left(\frac{\partial \rhop}{\partial t} + \nabla \cdotb ( \rhop \Ubt ) \right) \; d\Omega 
-&=& - \frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} r^2 \; dr 
-\int_{\Omega_H} \nabla \cdotb \left( \rho_0 \Ubt + \rhop w_0 \er \right) \; d\Omega \nonumber  \\
-&=& \left .  - \frac{1}{4\pi r_0^2 \, 2h}  \int_{\Omega_H} 
-\left [ \rho_0  (\Ubt \cdot \er)  + \rhop w_0 \right] r^2
-  \; d\Omega  \; \right  |_{r_0-h}^{r_0+h} \nonumber  \\
-&=& 0 \nonumber \enskip ,
+\frac{1}{4\pi r_0^2 2h}\int_{r_0-h}^{r_0+h}r^2 dr\int_{\Omega_H}\left[\frac{\partial\rho'}{\partial t} + \nabla\cdot\left(\rho'\Ub\right)\right]d\Omega 
+&=& - \frac{1}{4\pi r_0^2 2h}\int_{r_0-h}^{r_0+h}r^2 dr\int_{\Omega_H}\nabla\cdot\left(\rho_0\Ubt\right)d\Omega \nonumber \\
+&=& \left. -\frac{1}{4\pi r_0^2 2h}\int_{\Omega_H}\left[\rho_0\left(\Ubt\cdot\eb_r\right)\right]r^2 d\Omega\right|_{r_0-h}^{r_0+h} \nonumber \\
+&=& 0,
 \end{eqnarray} 
 where we have used the divergence theorem in spherical coordinates to transform
-the volume integrals into area integrals over $\Omega_H.$
-We see that both of the terms on the right hand side disappear. The first,
-$\int_{\Omega_H} \rho_0 (\Ubt \cdot \er) d\Omega = 0$ follows from the definition of $\Ubt,$
-and the second, $\int_{\Omega_H} \rhop w_0 d\Omega = w_0 \int_{\Omega_H} \rhop d\Omega = 0$
-follows from our initial assumption.  
-%\begin{eqnarray}
-%\frac{1}{4\pi r_0^2 \, 2h} \frac{\partial}{\partial t} 
-%  \int_{r_0-h}^{r_0+h} \int_{\Omega_H}  \rhop \; r^2 dr \; d\Omega &=&  
-%\frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} \int_{\Omega_H} \nabla \cdotb \left( \rho_0 \Ubt + \rhop w_0 \er \right)
-%\; r^2 dr \; d\Omega &=&
-%\\
-%&=& \left .  - \frac{1}{4\pi r_0^2 \, 2h} \int_{\Omega_H} 
-% \rhop (\Ubt \cdot \er)  \; r^2 d\Omega  \right  |_{r_0-h}^{r_0+h}
-%\enskip ,
-%\end{eqnarray} 
-Now expanding the remaining terms and taking the limit as $h \rightarrow 0,$ we can write 
+the volume integral on the right hand side into an area integral over $\Omega_H$.  
+We see that the right hand side disappears since 
+$\int_{\Omega_H}\rho_0(\Ubt\cdot\eb_r)d\Omega=0$,  which follows from the definition 
+of $\Ubt$.  Now, expanding the remaining terms and taking the limit as 
+$h\rightarrow 0,$ we can write
 \begin{eqnarray} 
-0 &=& \lim_{h\rightarrow 0} \frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H} 
-    \left(\frac{\partial \rhop}{\partial t} + \nabla \cdotb ( \rhop \Ubt ) \right) \; d\Omega \nonumber \\
-&=& \frac{\partial}{\partial t} \left( \lim_{h\rightarrow 0} \frac{1}{4\pi r_0^2} \frac{1}{2h} 
-  \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H}  \rhop \;d\Omega \right) +
-\lim_{h\rightarrow 0} \left( \frac{1}{4\pi r_0^2 \, 2h}  
-  \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H}  \nabla \cdotb ( \rhop \Ubt ) \; d\Omega \right) \nonumber \\
-&=& \frac{\partial}{\partial t} \left(\frac{1}{4\pi} \int_{\Omega_H}  \rhop \; d\Omega \right) +
-\lim_{h\rightarrow 0} \left( \left .  \frac{1}{4\pi r_0^2 \, 2h}  
-\left [ \int_{\Omega_H} \rhop (\Ubt \cdot \er) \; d\Omega  \right]  r^2\; \right |_{r_0-h}^{r_0+h} \right) \nonumber  \\
-&=&  \frac{\partial}{\partial t} \overline{\rhop} + \lim_{h\rightarrow 0} \left( \left .  \frac{1}{r_0^2 \, 2h}
-\overline{[\rhop (\Ubt \cdot \er)]} \; r^2 \; \right |_{r_0-h}^{r_0+h} \right) \nonumber  \\
-&=&  \frac{\partial}{\partial t} \overline{\rhop} + 
-\lim_{h\rightarrow 0} \frac{1}{r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} \nabla \cdot \left[\overline{\left(\rhop\Ubt \cdot \er\right)} \er \right]\;  r^2 \; dr \nonumber \\
-&=&  \frac{\partial}{\partial t} \overline{\rhop} + \nabla \cdot \left[ \overline{\left(\rhop \Ubt \cdot \er\right)} \er \right]
+0 &=& \lim_{h\rightarrow 0} \frac{1}{4\pi r_0^2 2h} \int_{r_0-h}^{r_0+h} r^2  dr \int_{\Omega_H} \left[\frac{\partial \rho'}{\partial t} + \nabla\cdot(\rho'\Ub)\right] d\Omega \nonumber \\
+&=& \frac{\partial}{\partial t} \left( \lim_{h\rightarrow 0} \frac{1}{4\pi r_0^2} \frac{1}{2h} \int_{r_0-h}^{r_0+h} r^2  dr \int_{\Omega_H}  \rho' d\Omega \right) + \lim_{h\rightarrow 0} \left[\frac{1}{4\pi r_0^2  2h} \int_{r_0-h}^{r_0+h} r^2  dr \int_{\Omega_H}  \nabla \cdot ( \rho' \Ub )  d\Omega\right] \nonumber \\
+&=& \frac{\partial}{\partial t} \left(\frac{1}{4\pi} \int_{\Omega_H}  \rho'  d\Omega \right) + \lim_{h\rightarrow 0} \left\{\left.\frac{1}{r_0^2 2h}\left[\frac{1}{4\pi}\int_{\Omega_H} \rho' (\Ub \cdot \eb_r)  d\Omega  \right]  r^2 \right |_{r_0-h}^{r_0+h} \right\} \nonumber  \\
+&=&  \frac{\partial}{\partial t} \overline{\rho'} + \lim_{h\rightarrow 0} \left\{ \left .
+\frac{1}{r_0^2  2h} \left[ \overline{\rho' (\Ub \cdot \eb_r)} \right]  r^2  \right |_{r_0-h}^{r_0+h} \right\} \nonumber  \\
+&=&  \frac{\partial}{\partial t} \overline{\rho'} + \lim_{h\rightarrow 0} \frac{1}{r_0^2 2h} \int_{r_0-h}^{r_0+h} \nabla \cdot \left[\overline{\rho' (\Ub \cdot \eb_r)} \eb_r \right]  r^2  dr \nonumber \\
+&=&  \frac{\partial}{\partial t} \overline{\rho'} + \nabla \cdot \left[ \overline{\rho' (\Ub \cdot \eb_r)} \eb_r \right]
 \end{eqnarray} 
@@
 \begin{equation}
-\frac{\partial}{\partial t} \overline{\rhop} = 
-- \nabla \cdot \left [ \overline{\left(\rhop \Ubt \cdot \er\right)} \, \er \right ] = 
-- \nabla \cdot ( \etarho \er )
- \enskip .\label{eq:rhopbar} 
+\frac{\partial}{\partial t} \overline{\rho'} = - \nabla\cdot\left[\overline{\rho'\left(\Ub\cdot\eb_r\right)}\eb_r\right] = -\nabla\cdot\left(\etarho\eb_r\right),
 \end{equation} 
+and thus, $\etarho = \overline{\left(\rho'\Ub\cdot\eb_r\right)}$.
 

### 2009-10-19 docs/flowchart/flowchart.tex
minor updates

@@
   Evolution,} A.~S.~Almgren, J.~B.~Bell, C.~A.~Rendleman, \& M.~Zingale
-  2006, ApJ, 649, 927 (henceforth paper~II)\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
+  2006, ApJ, 649, 927 (henceforth paper~II)
+\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
   A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale
@@
   M.~Zingale, A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& S.~E.~Woosley
-  2009, ApJ, submitted (henceforth paper~IV)
+  2009, ApJ, 704, 196 (henceforth paper~IV)
+\item {\em MAESTRO: An Adaptive Low Mach Number Hydrodynamics Algorithm for Stellar
+  Flows,} A.~Nonaka, A.~S.~Almgren, J.~B.~Bell, M.~J.~Lijewski, C.~M.~Malone,
+  \& M.~Zingale 2009, {\it Submited to SIAM J. Sci. Comput.}
+  (henceforth ``the multilevel paper'')
 \end{itemize}
@@
 \hline
-$\etarho$ & $\etarho \equiv \overline{(\rho' \Ubt \cdot \eb_r)}$ & g~cm$^{-2}$~s$^{-1}$ \\
+$\etarho$ & $\etarho \equiv \overline{(\rho' \Ub \cdot \eb_r)}$ & g~cm$^{-2}$~s$^{-1}$ \\
 \hline
@@
 Otherwise, following the method used in our small-scale low Mach number algorithm
-(cite BDRWZ), we extrapolate to the half-time using the source
+(BDRWZ), we extrapolate to the half-time using the source
 terms at the previous and current time levels

### 2009-10-19 docs/flowchart/flowchart.tex
chopped out outdated material that appears in the multilevel paper.  this can be added later

@@
 \item {\em Low Mach Number Modeling of Type Ia
-  Supernovae. I. Hydrodynamics,} A.~S.~Almgren, J.~B.~Bell, 
-  C.~A.~Rendleman, \& M.~Zingale 2006, ApJ, 637, 922 (henceforth
-  paper~I)
+  Supernovae. I. Hydrodynamics,} A. S. Almgren, J. B. Bell, 
+  C. A. Rendleman, \& M. Zingale 2006, ApJ, 637, 922 (henceforth
+  paper I)
 \item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
-  Evolution,} A.~S.~Almgren, J.~B.~Bell, C.~A.~Rendleman, \& M.~Zingale
-  2006, ApJ, 649, 927 (henceforth paper~II)
+  Evolution,} A. S. Almgren, J. B. Bell, C. A. Rendleman, \& M. Zingale
+  2006, ApJ, 649, 927 (henceforth paper II)
 \item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
-  A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale
-  2008, ApJ, 684, 449 (henceforth paper~III)
+  A. S. Almgren, J. B. Bell, A. Nonaka, \& M. Zingale
+  2008, ApJ, 684, 449 (henceforth paper III)
 \item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection,}
-  M.~Zingale, A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& S.~E.~Woosley
-  2009, ApJ, 704, 196 (henceforth paper~IV)
+  M. Zingale, A. S. Almgren, J. B. Bell, A. Nonaka, \& S. E. Woosley
+  2009, ApJ, 704, 196 (henceforth paper IV)
 \item {\em MAESTRO: An Adaptive Low Mach Number Hydrodynamics Algorithm for Stellar
-  Flows,} A.~Nonaka, A.~S.~Almgren, J.~B.~Bell, M.~J.~Lijewski, C.~M.~Malone,
-  \& M.~Zingale 2009, {\it Submited to SIAM J. Sci. Comput.}
+  Flows,} A. Nonaka, A. S. Almgren, J. B. Bell, M. J. Lijewski, C. M. Malone,
+  \& M. Zingale 2009, submitted to SIAM J. Sci. Comput. 
   (henceforth ``the multilevel paper'')
@@
 \item {\em Adaptive Low Mach Number Simulations of Nuclear Flames,}
-J.~B.~Bell, M.~S.~Day, C.~A.~Rendleman, S.~E.~Woosley, \& M.~Zingale
+J. B. Bell, M. S. Day, C. A. Rendleman, S. E. Woosley, \& M. Zingale
 2004, JCP, 195, 2, 677 (henceforth BDRWZ)
@@
 \begin{enumerate}
-\item {\tt fill\_3d} and {\tt average} algorithms.
-\item We update temperature after each {\bf React State} as well as the advection step.
+\item We defined the mapping of data between a 1D radial array and the 3D Cartesian
+grid for spherical problems (which we improve upon in the multilevel paper).
+\item We update $T$ after the call to {\bf React State}.
 \item We have created a {\tt burning\_cutoff\_density}, where the burning does
@@
 than $T=T(\rho,h,X_k)$.
-\end{enumerate}
-Changes which modify enthalpy update.  These are separate because in this paper,
-enthalpy is only used as a diagnostic since we use {\tt use\_tfromp}.  If we don't
-mention these items in paper IV, they should me noted in paper 4.5.
-\begin{enumerate}
-\item We have changed the discretization of $\Ubt\cdot\nabla p_0$ in the enthalpy
-update to $\nabla\cdot(\Ubt p_0) - p_0\nabla\cdot\Ubt$.
+\item For spherical problems, we have changed the discretization of 
+$\Ubt\cdot\nabla p_0$ in the enthalpy update to 
+$\nabla\cdot(\Ubt p_0) - p_0\nabla\cdot\Ubt$.
 \item In paper III we discretized the enthalpy evolution equation in
@@
 \end{enumerate}
-\section{Changes Between Papers 4 and Papers 4.5}
-\begin{enumerate}
-\item We split the update of the density and enthalpy in steps 4 and 8.
-Now, density is updated first, allowing us to compute time-centered values of 
-$\psi$ that are then used in the enthalpy update.
-\item We split {\bf Advect Base} into separate {\bf Advect Base Density},
-{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.  The calling sequence
-is in the main flow chart.
-\item We have eliminated {\bf Advect Base Pressure} entirely, and instead directly
-enforce HSE after {\bf Correct Base}.  {\bf Advect Base Pressure} was just a fancy
-way of enforcing HSE anyway, so we just compute the new pressure directly.  The
-problems we have seen with enforcing HSE directly may have to do with the roundoff
-error that accumulates, causing the temperature computed from the EOS to drift.
-However, using {\bf Advect Base Pressure}, the latest version of {\bf test\_basestate}
-generates negative pressures, so that is even worse.
-\item We replaced {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
-planar geometry and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical geometry,
-eliminating the need for $\nabla\cdot\etarho$.  {\bf Correct Base} is now used
-analogously for setting $(\rho h)_0$ after {\bf React State}.
-\item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
-planar geometry and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
-in spherical geometry.  This means that {\bf React Base} no longer exists, and the 
-intermediate states, $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
-{\bf Advect Base Enthalpy} will advance $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ 
-directly.  This also means we don't have to deal with $\eta_{\rho h}$ or a 
-{\bf Diffuse Base}.
-\item In {\bf Step 4} and {\bf Step 8}, we moved {\bf Correct Base} so it
-occurs after the full state density update and  before {\bf Advect Base Pressure}.  
-Since the full enthalpy advance ``feels'' the effects of the time-centered $\etarho$, 
-we should rely on a $\rho_0^{n+1}$ that also feels the effects.
-\item {\bf AMR!!!}
-\end{enumerate}
-Changes specific to spherical geometry:
-\begin{enumerate}
-\item Discuss {\bf Advect Base} routines and {\bf Correct Base} routine.
-\item $\etarho$ and $\psi$ equations.
-\end{enumerate}
+\section{Changes Between Paper 4 and the Multilevel Paper}
+See the multilevel paper for the latest.
 \section{Changes Between Papers 4.5 and Paper 5}
 \begin{enumerate}
-\item Use PPM limiters in advection.
+\item Added rotation.
 \end{enumerate}
@@
 \end{center}
-%
+
 \renewcommand{\arraystretch}{1.0}
-Here are some charts demonstrating what the superscript notation above scalars implies:
-\begin{equation}
-\left.\begin{array}{ccccccccc}
-& \text{React State} & & \text{Advect State} & & \text{Diffuse Enthalpy} & & \text{React State} & \\
-\rho^n & = & \rho^{(1)} & \longrightarrow & & & \rho^{(2)} & = & \rho^{n+1} \\
-(\rho X_k)^n & \longrightarrow & (\rho X_k)^{(1)} & \longrightarrow & & & (\rho X_k)^{(2)} & \longrightarrow & (\rho X_k)^{n+1} \\
-(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(1a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\
-T^n & \longrightarrow & T^{(1)} & & & \longrightarrow & T^{(2)} & \longrightarrow & T^{n+1} \\
-\end{array}\right.\nonumber
-\end{equation}
-\begin{equation}
-\left.\begin{array}{ccccccc}
-& \text{Advect Base} & & \text{Correct Base} & \\
-\rho_0^n & \longrightarrow & \rho_0^{(2a)} & \longrightarrow & \rho_0^{n+1} \\
-(\rho h)_0^n & \longrightarrow & & & (\rho h)_0^{n+1} \\
-p_0^n & \longrightarrow & & & p_0^{n+1} \\
-\end{array}\right.\nonumber
-\end{equation}
+There are some great charts in the appendix of the multilevel paper demonstrating what 
+the superscript notation represents.
 \section{Shorthand Functions}
-We make use of the following shorthand notations in outlining the algorithm:
-\begin{enumerate}
-
-\item For any quantity, $\phi$, we define $\overline{\phi} = ${\bf Avg}$(\phi)$
-as the average over a layer at constant radius r, $\Omega_H$, as
-\begin{equation}
-\overline{\phi(r)}
-= \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
-\end{equation}
-\begin{description}
-\item[planar:] This is a straightforward average since the radial coordinate is 
-in alignment with the lateral coordinate.
-\item[spherical:] See the description in paper IV, section 2.2 
-(or section \ref{Sec:Avg} in these notes).
-\end{description}
-
-\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}, 
-p_0^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, 
-(\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$
-is the process by which we evolve the species and enthalpy from 
-$X_k^{\inp} \rightarrow X_k^{\outp}$ 
-and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
-system of equations over a time interval of  $\Delta t/2$,
-\begin{eqnarray}
-\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
-\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
-\end{eqnarray}
-  In particular, to evolve the species, we solve the system:
-\begin{eqnarray}
-\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
-\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
-\enskip. \label{eq:reacttemp}
-\end{eqnarray}
-using the stiff ordinary differential equation integration methods
-provided by the {\tt VODE} package (cite vode).  The absolute error
-tolerances are set to $10^{-12}$ for the species, and a relative
-tolerance of $10^{-5}$ is used for the temperature.  The integration
-yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
-(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
-from equation (8) in paper III,
-\begin{equation}
-\label{eq:tempenthalpy}
-\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
- - \sum_k \rho \xi_k {\omegadot}_k 
- + \rho \Hnuc + \rho \Hext \right] \enskip , 
-\end{equation}
-by assuming that the pressure is constant during the burn state.  In
-evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
-theory, this means evaluating the equation of state for each
-right-hand side evaluation that {\tt VODE} requires.  In practice, we
-freeze $c_p$ and $\xi_k$ at the start of the integration time step and
-compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
-to the equation of state.  We update temperature using
-$T^{\outp} = T(\rho^\outp,h^\outp,X_k^\outp)$ if {\tt use\_tfromp = FALSE} or 
-$T^{\outp} = T(\rho^\outp,p_0^\inp,X_k^\outp)$  if {\tt use\_tfromp = TRUE}.  
-Note that the density remains unchanged
-during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
-
-After the new mass fractions have been computed, the reaction rates are defined as:
-\begin{equation}
-(\rho\omegadot_k)^{\outp} = \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\Delta t /2} \enskip,
-\end{equation}
-and the nuclear energy generation rate is defined as
-\begin{equation}
-(\rho\Hnuc)^{\outp} = -\sum_k q_k (\rho\omegadot_k)^{\outp} \enskip,
-\end{equation}
-for the case where only strong-mediated reactions are involved.  As
-$\Hnuc$ is an output of the reaction network, more general networks
-(i.e., involving weak interactions) can easily be added.  The enthalpy
-update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is
-updated by
-\begin{equation}
-(\rho h)^{\outp} = 
-(\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc)^{\outp} + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
-\end{equation}
-
-\item {\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp]$
-$\rightarrow [\rho_0^\outp,\rho_0^{\outp,n+\myhalf}]$
-is the process by which we update the base state density through $\dt$ in time
-as well as compute time-centered interface states, $\rho_0^{\outp,n+\myhalf}$, 
-for use later in the algorithm.
-\begin{description}
-
-\item[planar:] We discretize the base state density evolution 
-equation from paper III equation (29) without the $\etarho$ term,
-\begin{equation}
-\frac{\partial \rhozero}{\partial t} = - \nabla \cdot \left( \rhozero w_0 \eb_r \right)
-\enskip ,
-\label{eq:rho0upd_new}
-\end{equation}
-to compute the new base state density,
-\begin{equation}
-\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
-\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
-         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ] \enskip,
-\end{equation}
-where $j$ refers to the one-dimensional index in the radial direction.
-The interface states, $\rho_0^{\outp,\nph}$, are computed using the
-procedure described in paper II, Appendix A.\\
-
-\item[spherical:] The base state density update now includes the area factors in the 
-divergences:
-\begin{equation}
-\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp}
- - \frac{1}{r_j^2} \frac{\dt}{\dr} 
- \left [ \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
-          \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right] \enskip. 
-\end{equation}
-\end{description}
-%\item {\bf Advect Base Pressure}
-%is the process by which we update the base state pressure through $\dt$ in time and
-%compute $\psi$.  The input arguments are different for planar and spherical geometries.
-%\begin{description}
-%\item {\bf Advect Base Pressure Planar}$[p_0^\inp,w_0^\inp,\etarho^\inp] \rightarrow 
-%[p_0^\outp,\psi^\outp]$:
-%
-%First we compute a cell-centered $\psi^\outp$ using
-%\begin{equation}
-%\psi_j^\outp = 
-%\frac{1}{2} \left(\eta_{\rho,j-\myhalf}^\inp + \eta_{\rho,j+\myhalf}^\inp\right) g \enskip .
-%\end{equation}
-%Then, we discretize the base state pressure evolution equation from paper III equation (31),
-%\begin{equation}
-%\frac{\partial p_0}{\partial t} = 
-%-w_0 \frac{\partial p_0}{\partial r} + \psi \enskip , \label{eq:p0upd_new}
-%\end{equation}
-%to compute the new base state pressure,
-%\begin{equation}
-%p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
-%+ w_{0,j-\myhalf}^\inp \right)
-%\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) 
-%+ \dt \; \psi_j^\outp \enskip,
-%\end{equation}
-%where the interface states, $p_0^{n+\myhalf}$, are again found via the procedure 
-%described in paper II, Appendix A.\\
-%
-%\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},
-%\Sbar^{\inp}, \rho^\inp, X_k^\inp] \rightarrow [p_0^{\outp},\psi^{n,\outp}, 
-%\psi^{n+\myhalf,\outp}]$:
-%
-%First use a predictor-corrector formulation for the pressure update in order to improve 
-%the overall accuracy.  Beginning with equation (\ref{eq:w01}), we can write
-%\begin{equation}
-%\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
-% \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) 
-%- \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
-%\end{equation}
-%We time-center the left hand side:
-%\begin{equation}
-%\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
-%\frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} 
-%- p_0^{\inp}}{\dt}\enskip.
-%\end{equation}
-%We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
-%\begin{equation}
-%p_0^{\star} = p_0^{\inp} 
-%\frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
-%\end{equation}
-%with:
-%\begin{equation}
-%f = \frac{\dt}{2} \left [ \Sbar^{\inp} - 
-%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
-%\frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} 
-%\left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
-%\end{equation}
-%where the $(\partial p_0^\inp / \partial r)$ is computed by interpolating $p_0^\inp$ from 
-%cell-centers to edges and then differencing.  Next, we compute a provisional updated 
-%$\gammabar$ using the provisional updated pressure,
-%$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^\star,X_k^{\inp})]$,
-%where $\rho^\inp$ and $X_k^\inp$ are time-advanced versions of the $\rho$ and
-%$X_k$ which were used to compute $\gammabar^\inp$.\\
-%
-%Now we begin the corrector step.  Using $p_0^\star$ and $\gammabar^\star$, 
-%we can compute a more accurate approximation to equation (\ref{eq:frhs}),
-%\begin{equation}
-%f^\dagger = \frac{\dt}{2} \left [ \Sbar^{\inp} -
-%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
-%\frac{1}{\gammabar^{\dagger} p_0^\dagger} w_0^{\inp} 
-%\left ( \frac{\partial p_0^\dagger}{\partial r} \right ) \right ] \enskip ,
-%\end{equation}
-%where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$ and
-%$\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$.
-%This allows us to complete the corrector for the pressure update:
-%\begin{equation}
-%p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\dagger} f^\dagger\right)}{\left(1 - \gammabar^{\dagger} f^\dagger\right)}\enskip .
-%\end{equation}
-%Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.  First, we compute an 
-%updated $\gammabar$ using the update pressure, 
-%$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^{\outp},X_k^{\inp})]$.  Then, 
-%rewriting equation (\ref{eq:w01}), we can define
-%\begin{eqnarray}
-%\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
-%       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
-%\end{eqnarray}
-%We discretize this locally as:
-%\begin{equation}
-%\psi_j^{\outp,n} = 
-%\left(\gammabar^\inp p_0^\inp\right)_j
-%\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
-%          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip ,
-%\end{equation}
-%\begin{equation}
-%\psi_j^{\outp,n+\myhalf} = 
-%\left(\frac{\gammabar^\inp+\gammabar^{\star}}{2}\right)_j
-%\left(\frac{p_0^\inp+p_0^\outp}{2}\right)_j
-%\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
-%          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip .
-%\end{equation}
-%\end{description}
-\item {\bf Enforce HSE}$[p_0^{\inp},\rho_0^{\inp},g^{\inp}] \rightarrow [p_0^{\outp}]$
-has replaced {\bf Advect Base Pressure} from paper III as the process by which we 
-update the base state pressure.  Rather than discretize the evolution equation
-for $p_0$, we enforce hydrostatic equilibrium directly, which is simpler and
-analytically equivalent.  We update $p_0$ using:
-\begin{equation}
-p_{0,j}^{\outp} = p_{0,j-1}^{\outp} + 
-\frac{\Delta r}{2}\left(\rho_{0,j}^{\inp}+\rho_{0,j-1}^{\inp}\right)g_{j-\myhalf}^{\inp}\enskip .
-\end{equation}
-Once the density is less than the cutoff density, we set $p_{0,j}^\outp = p_{0,j-1}^\outp$.
-We use the boundary condition, $p_{0,jmax}^\outp = p_{0,jmax}^\inp$.
-
-\item {\bf Advect Base Enthalpy}
-is the process by which we update the base state enthalpy through $\dt$ in time.
-The input arguments are different for planar and spherical geometries.
-\begin{description}
-\item {\bf Advect Base Enthalpy Planar}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow 
-[(\rho h)_0^\outp]$:
-
-We discretize the base state enthalpy equation,
-\begin{equation}
-\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\eb_r\right] 
-+ \psi \enskip ,
-\end{equation}
-to compute the new base state enthalpy,
-\begin{equation}
-(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
-\left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} -  
-         \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\}
-+ \dt\psi_j^{\inp} \enskip ,
-\end{equation}
-where the interface states $(\rho h)_0^{n+\myhalf}$, are again found via the procedure 
-described in paper II, Appendix A.\\
-
-\item {\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^\inp,w_0^\inp,\psi^{\inp,n},
-\psi^{\inp,n+\myhalf}] \rightarrow [(\rho h)_0^\outp]$:
-
-The base state enthalpy update now includes the area factors in the divergences:
-\begin{equation}
-(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp}
- - \frac{1}{r_j^2} \frac{\dt}{\dr} 
- \left \{ \left[ r^2 (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} -  
-          \left[ r^2 (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\}
-+\Delta t\psi^{\inp,n+\myhalf} \enskip. 
-\end{equation}
-To compute the interface states, we use $\psi^{\inp,n}$ as the forcing term, 
-whereas in the enthalpy update, we use $\psi^{\inp,n+\myhalf}$ as the forcing.
-
-\end{description}
-
-\item {\bf Compute} {\boldmath $w_0$} is the process by which we compute $w_0$.
-The arguments are different for planar and spherical geometries.
-\begin{description}
-\item {\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\Sbar^{\inp},\gammabar^{\inp},p_0^{\inp},
-\psi^{\inp}] \rightarrow [w_0^{\outp}]$:
-
-We integrate
-\begin{equation}
-\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \psi
-\enskip \label{eq:divw0} ,
-\end{equation}
-with the following discretization:
-\begin{equation}
-\frac{\partial w_0^{\outp}}{\partial r} =  \Sbar^{\inp}
-- \frac{1}{\gammabar^{\inp} p_0^{\inp}}\psi^{\inp}
-\enskip ,
-\end{equation}
-and $w_0=0$ at $r=0$.\\
-
-The discretization with the volume discrepancy correction require an additional argument,
-$\overline{p^{\inp}}$, and the discretization becomes:
-\begin{equation}
-\frac{\partial w_0^{\nph,\outp}}{\partial r} = 
-\overline{S^{\nph,\inp}} - \frac{1}{\gammabar^{\inp} p_0^{\inp}}
-\left[ \psi^{\inp} + f \left(\frac{p_0^{\inp} - \overline{p^{\inp}}}{\Delta t^n}\right)\right]
-\enskip .
-\end{equation}
-
-\item {\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},g^\inp,\etarho^{\inp}] \rightarrow[w_0^{\outp}]$:
-
-We begin with equation (23) in paper III written in spherical coordinates:
-\begin{equation}
-\frac{1}{r^2}\frac{\partial}{\partial r} \left (r^2 \beta_0 w_0 \right ) =
-\beta_0 \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )\enskip.
-\end{equation}
-We expand the spatial derivative and divide through by $\beta_0$:
-\begin{equation}
-\label{eq:sphconstraint}
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) +
-w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r} =
- \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t}  \enskip .
-\end{equation}
-Recalling from paper I that;
-\begin{equation}
-\frac{1}{\gammabar p_0}  \frac{\partial p_0}{\partial r} = \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}\enskip,
-\end{equation}
-and moving this term to the right hand side,
-we can express equation (\ref{eq:sphconstraint}) as:
-\begin{eqnarray}
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) &=&
-  \Sbar - \frac{1}{\gammabar p_0} \left( 
-\frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r} \right)\enskip. \label{eq:w01}
-\end{eqnarray}
-We first solve a piece of this equation for $\ow = w_0 - \dw$:
-\begin{eqnarray}
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \ow \right ) &=& \Sbar^{\inp} \enskip.
-\end{eqnarray}
-Then we can write
-\begin{eqnarray}
-\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \dw \right ) &=&
-  - \frac{1}{\gammabar p_0} \left[
-\frac{\partial p_0}{\partial t} + (\ow + \dw)
-\frac{\partial p_0}{\partial r} \right]\enskip.\label{eq:w01a}
-\end{eqnarray}
-
-Multiplying equation (\ref{eq:w01a}) through by $\gammabar p_0 $,
-taking another derivative with respect to $r$, and switching the order
-of temporal and spatial derivatives, we get:
-\begin{equation}
-\label{eq:sphconstraint_2}
-\frac{\partial}{\partial r} \left [ \frac{\gammabar p_0 }{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right ]
-= -\frac{\partial}{\partial t} \frac{\partial p_0}{\partial r}
-  -\frac{\partial}{\partial r} \left [(\ow+\dw) \frac{\partial p_0}{\partial r} \right ]\enskip.
-\end{equation}
-
-To solve for $\dw$ we will need to substitute for the derivatives of $p_0.$
-To do so we start with the equation of hydrostatic equilibrium in a spherical geometry:
-\begin{equation}
-\frac{\partial p_0}{\partial r} = -\rho_0 g \, ; \quad
-g = \frac{G m_\mathrm{encl}}{r^2}\enskip,
-\end{equation}
-where $m_\mathrm{encl}(r)$ is the mass enclosed at radius $r$ and $G$ is the
-gravitational constant.  
-Using this, we can then write equation (\ref{eq:sphconstraint_2}) as:
-\begin{eqnarray}
-\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=&
- \frac{\partial}{\partial t} \left (\rho_0 g \right )
-+\frac{\partial}{\partial r} \left (w_0 \rho_0 g \right) \nonumber \\
-&=&
-  g \left[ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right]
-+ \rho_0 \left( \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} \right) \enskip.
-\label{eq:sphconstraint_3}
-\end{eqnarray}
-The mass enclosed inside any radius, $r$, is $m_\mathrm{encl}(r) = 4 \pi
-\int_0^r \rhozero(s) s^2 ds$, or alternately, $\partial m_\mathrm{encl}/\partial r = 4\pi r^2 \rho_0$.  
-The Lagrangian derivative of the enclosed mass is then:
-\begin{eqnarray}
-\frac{D_0 m_\mathrm{encl}}{D t} &=& \frac{\partial m_\mathrm{encl}}{\partial t} + w_0 \frac{\partial m_\mathrm{encl}}{\partial r} \nonumber \\
- &=& 4\pi \left( \frac{\partial}{\partial t} \int_0^r \rhozero(s) s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
-&=& 4\pi \left( \int_0^r \frac{\partial \rho_0}{\partial t} s^2 ds + w_0 r^2 \rho_0 \right) \nonumber \\
-&=& 4\pi \left\{ -\int_0^r \left [ \frac{1}{s^2}\frac{\partial(s^2 \rho_0 w_0)}{\partial s} + \frac{1}{s^2}\frac{\partial (s^2 \etarho)}{\partial s} \right ]s^2 ds + w_0 r^2 \rho_0 \right\} \nonumber \\
-&=&  \left . 4\pi \left(-s^2 \rho_0 w_0 \right |_0^r - \left . s^2 \etarho \right |_0^r + w_0 r^2 \rho_0 \right) \nonumber \\
-&=& -4\pi r^2 \etarho \label{eq:dmdt}
-\end{eqnarray}
-where we used the spherical form of equation (29) in paper III, 
-\begin{equation}
-\label{eq:sph_continuity}
-\frac{\partial \rhozero}{\partial t} + \frac{1}{r^2} \frac{\partial (r^2 \rho_0 w_0)}{\partial r} 
-+ \frac{1}{r^2} \frac{\partial (r^2 \etarho )}{\partial r}  = 0 \enskip ,
-\end{equation}
-to eliminate $\partial \rho_0/\partial t$.  We note that in the
-absence of any mixing, $\etarho=0$, and  $D_0 m_\mathrm{encl}/Dt = 0.$
-Equation (\ref{eq:dmdt})
-allows us to write the Lagrangian change in the gravitational
-acceleration as:
-\begin{eqnarray}
-\frac{D_0 g}{D t} = \frac{\partial g}{\partial t} + w_0 \frac{\partial g}{\partial r} &=& \frac{D_0}{Dt}\left(\frac{G m_\mathrm{encl}}{r^2}\right) \nonumber \\
-&=& G m_\mathrm{encl}\frac{D_0}{D t}\left(\frac{1}{r^2}\right) + \frac{G}{r^2}\frac{D_0 m_\mathrm{encl}}{Dt} \nonumber \\
-&=& -\frac{2 w_0 G m_\mathrm{encl}}{r^3} - 4 \pi G \etarho \nonumber \\
-&=& -\frac{2 w_0 g}{r} - 4 \pi G \etarho \enskip .
-\end{eqnarray}
-
-Putting it all together, equation (\ref{eq:sphconstraint_3}) becomes:
-\begin{equation}
- \frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] 
-= 
-  g \left [ \frac{\partial \rho_0}{\partial t} + \frac{\partial}{\partial r} (w_0 \rho_0) \right ]
-+ \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \label{eq:B5} \enskip.
-\end{equation}
-Finally, we can use equation (\ref{eq:sph_continuity})
-to write equation (\ref{eq:B5}) as:
-\begin{eqnarray}
-\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=&
- g \left [ -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 w_0 \rhozero)
-           -\frac{1}{r^2} \frac{\partial}{\partial r} (r^2 \etarho)
-           + \frac{\partial}{\partial r} (w_0 \rhozero) \right ] \nonumber \\
-&& + \rho_0 \left(\frac{-2 w_0 g}{r} - 4 \pi G \etarho \right) \nonumber \\
-&=& - \frac{g}{r^2} \frac{\partial (r^2 \etarho)}{\partial r} 
-- \frac{4 (\ow + \dw) \rho_0 g}{r} - 4 \pi G \rhozero \etarho \label{eq:B7} \enskip .
-\end{eqnarray}
-
-\noindent We discretize this elliptic equation in the radial dimension as:
-\begin{eqnarray}
-&& \frac{1}{\Delta r} \left\{
-\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{j} -
-\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{j-1} \right\}
-+ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{j-\myhalf} \nonumber\\
-&=& 
-- \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
-\left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
-   - \left( \frac{4 \ow \rho_0 g}{r} \right)_{j-\myhalf}
-   - \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip ,
-\end{eqnarray}
-where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
-Then, using hydrostatic equilibrium, we expand this to 
-\begin{eqnarray}
-& \frac{1}{\Delta r} \left\{
- \left( \frac{\gammabar p_0}{r^2}\right)_{j  } \frac{\left[ (r^2 \dw)_{j+\myhalf} - (r^2 \dw)_{j-\myhalf} \right]}{\Delta r} -\left( \frac{\gammabar p_0}{r^2}\right)_{j-1} \frac{\left[ (r^2 \dw)_{j-\myhalf} - (r^2 \dw)_{j-\thalf}\right]}{\Delta r} \right\} & \nonumber \\
-&  - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \dw)_{j-\myhalf} = \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j} - p_{0,j-1}}{\dr} \right) (r^2 \ow)_{j-\myhalf} & \nonumber \\
-& - \frac{g_{j-\myhalf}}{r_{j-\myhalf}^2 \Delta r} 
-  \left[ \left( r^2 \etarho \right)_{j} - \left( r^2 \etarho  \right)_{j-1} \right] 
-- \left(  4 \pi G \rhozero \etarho   \right)_{j-\myhalf} \enskip, &
-\end{eqnarray}
-where $\etarho$ at cell-centers is the average of the neighboring edge-values.  
-If we write this in matrix form, so that:
-\begin{equation}
-A_j (r^2 \dw)_{j-\thalf} + B_j (r^2 \dw)_{j-\myhalf} + C_j (r^2 \dw)_{j+\myhalf} = F_j\enskip,
-\end{equation}
-then:
-\begin{eqnarray}
-A_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1}  \enskip, \\
-B_j &=& -\frac{1}{\Delta r^2} \left[ \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j  }  
-               +\left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j-1} \right] 
-              - \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) \enskip , \\
-C_j &=& \frac{1}{\Delta r^2} \left( \frac{\gammabar^{\inp} p_0^{\inp}}{r^2}\right)_{j}  \enskip, \\
-F_j &=&  \left( \frac{4}{r_{j-\myhalf} ^3} \frac{p_{0,j}^{\inp} - p_{0,j-1}^{\inp}}{\dr} \right) (r^2 \ow)_{j-\myhalf}
- - \frac{g_{j-\myhalf}^\inp}{r_{j-\myhalf}^2 \Delta r} 
-\left[ \left( r^2 \etarho^{\inp} \right)_{j} - \left( r^2 \etarho^{\inp}  \right)_{j-1} \right] 
-\nonumber \\
-&& - \left(  4 \pi G \rhozero^{\inp} \etarho^{\inp}   \right)_{j-\myhalf} \enskip .
-\end{eqnarray}
-We define the lower boundary condition, $ \dw = 0$ at $r=0,$ which
-corresponds to $j=0,$ by setting:
-\begin{eqnarray}
-A_0 = C_0 = F_0 &=& 0\enskip, \nonumber \\
-B_0 &=& 1\enskip.
-\end{eqnarray}
-We also specify $\partial (r^2 \dw) / \partial r = 0$ at the the upper boundary, which corresponds to $j=N,$
-by setting:
-\begin{eqnarray}
-A_N &=& -1\enskip, \nonumber \\
-B_N &=&  1\enskip, \nonumber \\
-C_N = F_N &=& 0\enskip.
-\end{eqnarray}
-Finally, $w_0^{\outp} = \ow + \delta w_0$.
-\end{description}
-
-\item {\bf Compute Gravity}$[\rhozero^{\inp}] \rightarrow [g^{\outp}]$\\
-is the process by which we compute gravity at both radial cell-centers and edges.
-\begin{description}
-\item [planar:] Gravity constant in space and time.
-\item [spherical:] We define the enclosed mass at edges, and then cell-centers:
-\begin{equation}
-m_{j-\myhalf} = \sum_{k=1}^j \frac{4}{3}\pi\left(r_{k-\myhalf}^3 - r_{k-\sfrac{3}{2}}^3\right)\rho_{0,k-1}^{\inp}; ~~~~~ m_j = m_{j-\myhalf} + \frac{4}{3}\pi \left(r_j^3 - r_{j-\myhalf}^3\right)\rho_{0,j}^{\inp} \enskip .
-\end{equation}
-Then, we define the gravity at both edges and cell-centers:
-\begin{equation}
-g_{j-\myhalf}^\outp = \frac{Gm_{j-\myhalf}}{r_{j-\myhalf}^2}; ~~~~~ 
-g_j^\outp = \frac{Gm_j}{r_j^2}\enskip .
-\end{equation}
-\end{description}
-
-\end{enumerate}
+See the multilevel paper for the latest.
 

### 2009-10-19 docs/radial_base/radial_base.tex
get the figure working with the PDF version
i'm a sick person
added more radial mapping data
more mapping data
fixed section label
some cubic mapping stencils
added cubic newton tableau
solved cubic stencil at center
this section was totally outdated, so I copied in the multilevel paper text in

--- 2009-02-26 docs/radial_base/radial_base.tex
@@
-Since the one-dimensional spherical base state is not aligned with the
-3-d Cartesian grid we use for the full star, computing quantities that
-involve both the base state and the full state becomes complicated.
-Therefore, when we go between the base state and the full state in
-various parts of the algorithm (such as the averaging operations), we
-need to map between the two geometries.
+We refer to the procedure for mapping a cell-centered Cartesian field to a
+cell-centered radial array as a lateral ``average''.  We refer to the procedures for mapping
+an edge- or cell-centered radial field to an edge- or cell-centered Cartesian grid
+as a ``fill''.
+\subsubsection{Average}\label{Sec:Average}
+For any Cartesian cell-centered field, $\phi$, we define 
+$\overline{\phi} = ${\bf Avg}$(\phi)$ as the average over a layer at 
+constant radius $r$, $\Omega_H$, as
+\begin{equation}
+\overline{\phi(r)} = \frac{1}{A(\Omega_H)}\int_{\Omega_H}\phi(r,\xb)d\Omega; \qquad
+A(\Omega_H) = \int_{\Omega_H}d\Omega.
+\end{equation}
+\begin{description}
+\item[planar:] 
+  This is a straightforward arithmetic average of cells at a particular
+  height since the radial cell centers are in alignment with the Cartesian 
+  grid cell centers.
+\item[spherical:] It can be shown that any Cartesian cell center is a
+  radius $\hat r_m = \Delta x\sqrt{\sfrac{3}{4} + 2m}$ from the center of
+  the star, where $m \ge 0$ is an integer.  
+  For example, the Cartesian cell with coordinates $(i,j,k) = (1,1,1)$ relative
+  to the center of the star lies at a distance of $\Delta x\sqrt{\sfrac{3}{4}+6}$
+  from the center of the star, corresponding to $m=3$.  The Cartesian cells
+  with coordinates $(i,j,k) = (2,0,0), (0,2,0)$, or $(0,0,2)$ relative to
+  the center of the star also lie at that same 
+  distance.  For the 384$^3$ resolution
+  examples in this paper, we have verified that a non-zero set of
+  Cartesian cell centers map into each radius $\hat r_m$ until $m$ is 
+  large enough to correspond to a radius larger
+  than half the width of the computational domain (i.e., the edge of the
+  domain, not the corner of the domain).  Figure 
+  \ref{Fig:test_average3} shows the number of Cartesian cells that map into each 
+  radius $\hat r_m$, which we refer to as the ``hit count'', for a 384$^3$ domain.  
+  We use this mapping to help construct the lateral average, using the following steps:
+\begin{enumerate}
+\item Create an itemized list, $\hat\phi_m$, where each element is
+  associated with a radius $\hat r_m = \Delta x\sqrt{\sfrac{3}{4} + 2m}$ from the 
+  center of the star.
+\item For each $\hat\phi_m$, compute the arithmetic average value of the Cartesian 
+  cells whose centers lie at the associated radius.  As an additional 
+  element in the itemized list, include the center of the star (corresponding
+  to a radius of $r=0$).  Compute this additional value of $\hat\phi$
+  at this location using quadratic interpolation with $\hat\phi_0$, $\hat\phi_1$,
+  and a homogeneous Neumann condition at $r=0$ as the stencil points.  Note
+  that for very large values of $m$, it is possible that no Cartesian cell centers
+  exist at a radius $\hat r_m$ (i.e., the hit count is zero).  
+  If so, we say that $\hat\phi_m$ has an 
+  undefined/invalid value, and we ignore such values for the rest of this procedure.
+\item To compute the lateral average, use quadratic interpolation using the 
+  value in the itemized list with the closest associated radius, $\hat\phi_k$,
+  and the nearest values above and below, $\hat\phi_{k_+}$ and $\hat\phi_{k_-}$,
+  using divided differences:
+  \begin{equation}
+    \overline{\phi(r)} = \hat\phi_{k_-} + \frac{\hat\phi_k-\hat\phi_{k_-}}{\hat r_k-\hat r_{k_-}}(r-\hat r_{k_-}) + \frac{\frac{\hat\phi_{k_+}-\hat\phi_k}{\hat r_{k_+}-\hat r_k}-\frac{\hat\phi_k-\hat\phi_{k_-}}{\hat r_k-\hat r_{k_-}}}{\hat r_{k_+}-\hat r_{k_-}}(r-\hat r_{k_-})(r-\hat r_k)\nonumber,
+  \end{equation}
+  where $\hat r_{k_-}$, $\hat r_k$, and $\hat r_{k_+}$ are the three radii
+  associated with $\hat\phi_{k_-}$, $\hat\phi_k$, and  $\hat\phi_{k_+}$.
+  Finally, constrain $\overline{\phi(r)}$ to lie within the range of
+  $\hat\phi_{k_-}$, $\hat\phi_k$, and $\hat\phi_{k_+}$ so as to not introduce
+  any new maxima or minima.
+\end{enumerate}
+  In \S \ref{Sec:Mapping Test}, we show the improvement of this averaging procedure 
+  over the Paper IV procedure.
+\end{description}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[h]
+\centering
+\includegraphics{\radbasefigpath/test_average3}
+\caption{\label{Fig:test_average3}
+The number of Cartesian cells whose centers lie
+at a radius $\hat r_m$ (i.e., the ``hit count'') for a 384$^3$ domain vs. the
+itemized list index, $m$.  
+Indices $m<18,432$ correspond to locations within half the width 
+of the computational domain.  A non-zero set of Cartesian cell centers maps 
+into the radius associated with every $m\le 37,912$, which corresponds to 
+approximately 0.72 times the width of the computational domain.  The inset plot
+is a zoom-in of the innermost 75 values of $m$.}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Fill}\label{Sec:Fill}
+There are four different mappings from a 1D radial array to the 3D Cartesian grid;
+below we describe the procedures for planar and spherical geometries separately.
+\begin{description}
+\item[planar:]
+\end{description}
+\begin{enumerate}
+\item To map a cell-centered radial array onto Cartesian cell centers,
+  we use direct-injection since the radial cell centers are in
+  alignment with the Cartesian cell centers. 
+\item To map an edge-centered radial array onto Cartesian 
+  cell centers, we average the two nearest radial edge-centered values.
+\item To map a cell-centered radial array onto Cartesian edges with normal in 
+  the radial direction, we use 4th order spatial interpolation:
+\begin{equation}
+\phi_{i,j,k+\myhalf} = \frac{7}{12}(\phi_{j}+\phi_{j+1}) - \frac{1}{12}(\phi_{j-1}+\phi_{j+2}).
+\end{equation}
+  We constrain $\phi_{i,j,k+\myhalf}$ to lie between the interpolated values, and 
+  lower the order of interpolation near domain boundaries.  For the 
+  Cartesian edges transverse to the base state direction, we use direct-injection 
+  since the radial cell centers are in alignment with these 
+  Cartesian edges.
+\item To map an edge-centered radial array onto Cartesian edges, 
+  we use direct-injection on Cartesian edges normal to the base state direction
+  since the radial edges are in alignment with these Cartesian edges.  
+  For the remaining Cartesian edges, we average the two nearest radial 
+  edge-centered values.
+\end{enumerate}
+\begin{description}
+\item[spherical:] 
+\end{description}
+\begin{enumerate}
+\item To map a cell-centered radial array onto Cartesian cell centers,
+      we use quadratic interpolation from the nearest 
+      three radial cell centers (see Figure \ref{Fig:fill}a).
+      This is a departure from Paper IV, in which we used 
+      piecewise constant interpolation.  
+\item To map an edge-centered radial array onto Cartesian cell centers, 
+      we use linear interpolation from the nearest two points
+      (see Figure \ref{Fig:fill}b).
+\item To map a cell-centered radial array onto Cartesian edges,
+      we first map the radial array onto Cartesian cell centers (see 1.), 
+      then average the two neighboring centers to obtain the Cartesian edge values
+      (see Figure \ref{Fig:fill}c).
+\item To map an edge-centered radial array onto Cartesian edges,
+      we first map the radial array onto Cartesian cell centers (see 2.), 
+      then average the two neighboring centers to obtain the Cartesian edge values
+      (see Figure \ref{Fig:fill}c).
+\end{enumerate}
 
-As there is no alignment between the 3-d Cartesian grid and the 1-d
-spherical base state grid, we are free to choose their grid spacings
-independently.  Experimentation has shown that the best choice results
-from picking the 3-d Cartesian grid spacing, $\Delta x$ to be coarser
-than the 1-d spherical base state spacing, $\Delta r$.
-A typical simulation may use $5 \Delta r = \Delta x$.
-We call the procedure to go from 1-d to 3-d {\tt fill\_3d} and the
-complementary procedure that takes us from 3-d to 1-d {\tt average}.
-\section{\tt fill\_3d}
-There are three different {\tt fill\_3d} modes, each performed with the
-subroutine, {\tt put\_1d\_array\_on\_cart}:
-\begin{enumerate}
-\item The 1D array is bin-centered, and we map to Cartesian cell-centers.
-\item The 1D array is edge-centered, and we map to Cartesian cell-centers.
-\item The 1D array is edge-centered, and we map to Cartesian edges.  
-\end{enumerate}
-\subsection{1D Cell-Centered to Cartesian Cell-Centered}
-\label{Sec:1D Cell-Centered to Cartesian Cell-Centered}
-Figure \ref{fig:mapping} shows the Cartesian grid with the spherical
-base state overlayed (for simplicity, the figure is drawn in 2-d 
-using $2 \Delta r = \Delta x$).  
-We compute the radius of the Cartesian cell center from the center of the star,
-\begin{equation}
-r = \sqrt{(x_i - x_c)^2 + (y_j - y_c)^2 + (z_k - z_c)^2}
-\end{equation}
-where $(x_c, y_c, z_c)$ are the coordinates of the center of the star.
-We use this radius to find the corresponding radial bin as $n = \mathtt{int}(r
-/ \Delta r)$ (here, our convention is to use 0-based indexing for the
-base state).  We can then initialize a Cartesian cell quantity $q$ from its
-corresponding base state quantity, $q_0$ as $q_{i,j,k} = q_{0,n}$.  In essence,
-we are doing a piecewise constant interpolation of the radial array onto the 
-Cartesian grid.  If the output is a vector quantity, rather than a scalar, we
-multiply the result by the unit vector in the normal direction.
-\subsection{1D Edge-Centered to Cartesian Cell-Centered}
-\label{Sec:1D Edge-Centered to Cartesian Cell-Centered}
-Again, for each Cartesian cell-center, we compute the corresponding radial
-bin.  We linearly interpolate from the two nearest radial edge values 
-to get the Cartesian cell-centered value.  In essence,
-we are doing a piecewise linear interpolation of the radial array onto the 
-Cartesian grid.  If the output is a vector quantity, rather than a scalar, we
-multiply the result by the unit vector in the normal direction.
-\subsection{1D Edge-Centered to Cartesian Edges}
-\label{Sec:1D Edge-Centered to Cartesian Edges}
-This case is only used to create {\tt w0mac} from {\tt w0} in the 
-subroutine, {\tt put\_w0\_on\_edges}.  There are several variations 
-of the interpolation, controlled by {\tt w0mac\_interp\_type}.  We 
-currently use the default, {\tt w0mac\_interp\_type} = 1.  We first 
-map $w_0$ onto Cartesian cell-centers, and multiply the result by 
-the unit vector in the normal direction.  Then, for each Cartesian 
-face, we obtain the normal velocity by averaging the corresponding 
-velocities from the Cartesian cell-centers adjacent to the face.
-\subsection{1D Cell-Centered to Cartesian Edges}
-We don't have a subroutine that does this directly, but at various points
-in the algorithm we need base state quantities on edges.  For example, we 
-need $\rho_0$ at edges to compute the full state $\rho$ used to compute 
-fluxes.  More to come...
-\MarginPar{This section needs to be written.}
-\section{{\tt average}}\label{Sec:Avg}
-For the {\tt average} process, we first define a coarse 1D radial array 
-with $\Delta r_{\rm c} = \Delta x$.  Then, we compute the radius of each
-Cartesian cell center, and compute the index of the corresponding coarse radial 
-bin, $n_{\rm c} = \mathtt{int}(r / \Delta r_{\rm c})$.  
-We define $q_{0,n_{\rm c}}$ as the average of all the $q_{i,j,k}$ whose 
-Cartesian cell center maps into coarse radial bin $n_{\rm c}$.
-Next, we construct edge-centered states on the coarse radial bin using 
-$q_{0,n_{\rm c}+\myhalf} = (7/12)(q_{0,n_{\rm c}}+q_{0,n_{\rm c}+1}) 
-- (1/12)(q_{0,n_{\rm c}-1}+q_{0,n_{\rm c}+2})$ along with limiters to prevent
-new extrema.  Then, for each coarse radial bin
-we construct a quadratic profile using $q_{0,n_{\rm c}+\myhalf}, q_{0,n_{\rm c}}$ 
-and $q_{0,n_{\rm c}-\myhalf}$.  We use these quadratic profiles along with limiters
-to prevent new extrema to fill the fine radial bins.
-
-\clearpage
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[tpb]
-\begin{center}
-\epsffile{\radbasefigpath/mapping.eps}
-\caption{\label{fig:mapping} The Cartesian grid and spherical base
-state (shown here in 2-d for simplicity using $2 \Delta r = \Delta x$).
-Here we represent the
-spherical base state as concentric shells (blue lines).  Since the
-base state is not aligned with the Cartesian grid we use on which
-we discretize the star, we need to map between the two configurations.
-The `$+$' symbols represent the Cartesian zone centers.  In our
-mapping, the zone marked with the red symbol maps into the gray-shaded
-radial bin. }
-\label{fig:mapping}
-\end{center}
+\centering
+\includegraphics[width=1.75in]{\radbasefigpath/fill1}\hspace{0.1in}
+\includegraphics[width=1.75in]{\radbasefigpath/fill2}\hspace{0.1in}
+\includegraphics[width=1.75in]{\radbasefigpath/fill3}
+\caption{\label{Fig:fill} Illustrations of the fill operation for
+  spherical geometry.  (a) To fill the Cartesian cell center (fill
+  type 1), represented by the square, from radial cell-centered data,
+  represented by the circles, we use quadratic interpolation from the
+  nearest three points.  (b) To fill the Cartesian cell center
+  (fill type 2), represented by the square, from radial edge-centered
+  data, represented by the circles, we use linear interpolation from
+  the nearest two points. (c) To fill a Cartesian edge (fill
+  types 3 and 4), represented by the squares, first fill the 
+  Cartesian cell centers, represented by the circles, then average
+  the two neighboring cell centers.}
 \end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2009-10-19 docs/flowchart/flowchart.tex
updates

@@
 the superscript notation represents.
-\section{Shorthand Functions}
-See the multilevel paper for the latest.
 
@@
 \section{Time Advancement Algorithm}\label{Sec:Time Advancement Algorithm}
-
-We now describe the full time advancement algorithm, making frequent
-use of the shorthand developed above.  Here, we assume that the
-problem is already properly initialized.  We describe the details of
-the initialization in \S \ref{Sec:Initialization}.
-The advance of the state through a single timestep appears as:
-
-%--------------------------------------------------------------------------
-% STEP 1
-%--------------------------------------------------------------------------
-
-\begin{description}
-
-\item[Step 1.] {\em Define the provisional time-centered expansion, $S^{n+\myhalf,\star}$, 
-provisional base state velocity, $w_0^{n+\myhalf,\star}$, and provisional base state velocity 
-forcing.}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item At the beginning of each time step, we need an estimate for the time-centered
-source term in the velocity divergence constraint, as given in paper III equation (19),
-\begin{equation}
-  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext 
-  + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T \enskip .
-\label{eq:defineS} 
-\end{equation}
-
-If this is the first step of the calculation ($n=0$), we set
-\begin{equation}
-S^{\myhalf,\star} = \frac{S^0 + S^1}{2} \enskip ,
-\end{equation}
-where $S^1$ is found through the iterative process that initializes the calculation.
-Otherwise, following the method used in our small-scale low Mach number algorithm
-(BDRWZ), we extrapolate to the half-time using the source
-terms at the previous and current time levels
-\begin{equation}
-S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
-\end{equation}
-\item Compute
-\begin{equation}
-\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star}\right) \enskip .
-\end{equation}
-\item Compute $w_0^{n+\myhalf,\star}$:
-\begin{description}
-\item For planar geometry,\\
-{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star}},\gammabar^n,p_0^n,
-\psi^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
-
-If we are using the volume discrepancy term, compute 
-$\overline{p^{\inp}} = \overline{p(\rho,h,X_k)^n}$.
-\item For spherical geometry,\\
-{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\gammabar^n,
-\rhozero^n,p_0^n,g^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{n+\myhalf,\star}]$.
-\end{description}
-
-\item Define the base state velocity forcing using equation (38) from paper III,
-\begin{equation}
--\frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} 
-= \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} 
-\enskip, \label{eq:pizero}
-\end{equation}
-with the following discretization:
-\begin{equation}
-\left ( \frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} \right )^{n,\star} = 
--\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} 
-- w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star} \enskip ,
-\end{equation} 
-where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are defined through 
-paper III equation (53),
-\begin{equation}
-w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}} 
-\enskip,
-\end{equation}
-\begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = 
-\frac{1}{\dt^n+\dt^{n-1} } 
-\left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
-+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] 
-\enskip.
-\end{equation}
-If $n=0$, we use $\dt^{-1} = \dt^0$.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 2
-%--------------------------------------------------------------------------
-\item[Step 2.] {\em Construct the provisional time-centered advective velocity on 
-edges, $\uadvone$.}
-
-The local velocity field is described by paper III equation (37),
-\begin{equation}
-\frac{\partial\Ubt}{\partial t} = 
-- \left(\Ubt+w_0\right) \cdot \nabla \Ubt
-- \left(\Ubt \cdot \eb_r\right) \frac{\partial w_0}{\partial r} \eb_r
-- \frac{1}{\rho} \nabla\pi
-+ \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \eb_r
-- \frac{(\rho-\rhozero)}{\rho} \; g \; \eb_r  \label{eq:utildeupd}  \enskip .
-\end{equation}
-
-We construct time-centered velocities at edges, $\uadvone$, 
-using a second-order unsplit Godunov procedure followed by an elliptic
-solve, as described in Appendix B of paper III.  We note that 
-$\uadvone$ satisfies the discrete versions of paper III equation (21),
-\begin{equation}
-\int_{\Omega_H} \Ubt \cdot \eb_r \; dA = 0 \label{eq:udoterzero} \enskip ,
-\end{equation}
-and paper III equation (39),
-\begin{equation}
-\nabla \cdot (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
-\label{eq:tildeconstraint}
-\end{equation}
-with the discretization
+Here are changes to the algorithm from the multilevel paper description.
+The initialization has also been included with more detail than was given in paper III.
+\subsection{Volume Discrepancy Changes}
+\begin{itemize}
+\item In {\bf Step 2B}, to compute $w_0$, we need to account for the volume discrepancy
+term by computing $p_{\rm EOS} = \overline{p(\rho,h,X_k)^n}$.
+\item In {\bf Step 3}, the MAC projection should account for the volume discrepancy term:
 \begin{equation}
 \nabla \cdot \left(\beta_0^n \uadvone\right) = 
-\beta_0^n \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right) \enskip .
+\beta_0^n \left\{ \left(S^{\nph,\star\star} - \overline{S^{\nph,\star\star}}\right)
++ \frac{f}{\gammabar^n p_0^n}
+\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right]\right\}.
 \end{equation}
-The discretization with the volume discrepancy correction is
+\item In {\bf Step 6B}, to compute $w_0$, we need to account for the volume discrepancy
+term by computing $p_{\rm EOS} = \left[\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}\right]/2$.
+\item In {\bf Step 7}, the MAC projection should account for the volume discrepancy term:
 \begin{equation}
-\nabla \cdot \left(\beta_0^n \uadvone\right) = 
-\beta_0^n \left\{ \left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right)
-+ \frac{f}{\gammabar^n p_0^n}
-\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right]\right\} \enskip .
+\nabla \cdot \left(\beta_0^{\nph,\star} \uadvtwo\right) = \beta_0^{\nph,\star}\left\{\left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right) + \frac{f}{\overline{\Gamma_1^{\nph,\star}} p_0^{\nph,\star}} \left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\},
 \end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 3
-%--------------------------------------------------------------------------
-\item[Step 3.] {\em React the full state through the first time interval of $\dt / 2.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n] \rightarrow
-[ \rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)} ]$.
-
-\item Set $(\rho h)_0^n$ to be the lateral average of $(\rho h)^{(1)}$:
-\begin{description}
-\item For planar geometry, set $(\rho h)_0^n = $ {\bf Avg}$[(\rho h)^{(1)}]$
-\item For spherical geometry, set $(\rho h)_0^n = (\rho h)_0^n +$ {\bf Avg}$[(\rho h)^{(1)} - (\rho h)_0^n]$
-\end{description}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 4
-%--------------------------------------------------------------------------
-\item[Step 4.] {\em Advect the base state and full state, through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf Advect Base Density}$[\rhozero^{n},w_0^{\nph,\star}] \rightarrow$ 
-$[\rho_0^{(2a),\star}, \rho_0^{\nph,\star,{\rm pred}}]$.
-
-\item \label{step:species update} 
-Update the species.  Here we consider only the advection terms, 
-neglecting the reaction terms, and solve a discretized version of
+where $p(\rho,h,X_k)^{\nph,\star} = \left[p(\rho,h,X_k)^n +p(\rho,h,X_k)^{n+1,\star}\right]/2$.
+\item In {\bf Step 11}, the approximate projection should account for the volume
+discrepancy term:
 \begin{equation}
-\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\Ub \rho X_k) = 0 \enskip .
-\label{eq:species}
+\nabla \cdot \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S^{n+1}} \right)
++ \frac{f}{\overline{\Gamma_1^{n+1}} p_0^{n+1}}
+\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\}.
 \end{equation}
-The update consists of two steps:
-
-  \begin{enumerate}
-  \renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-  \item Compute the species edge states, $(\rho X_k)^{(1),\nph,\star}$,
-  for the conservative update of $(\rho X_k)^{(1)}$. 
-  Here we predict $\rho^{'(1)} = \rho^{(1)} - \rhozero^n$ and 
-  $X_k^{(1)} = (\rho X_k)^{(1)} / \rho^{(1)}$ to time-centered edges to obtain 
-  $\rho^{'(1),\nph,\star}$ and $X_k^{(1),\nph,\star}$ using a second-order
-  unsplit Godunov procedure, as described in paper II,
-  Appendix A, using $\Vb = \uadvone+w_0^{\nph,\star} \eb_r$.  Convert the 
-  perturbational edge state to a full state using
-\begin{equation}
-\rho^{(1),\nph,\star} = 
-\rho^{'(1),\nph,\star} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2} \enskip .
-\label{eq:rhoXedgestate}
-\end{equation}
-  In planar geometry, we use 4th-order spatial averaging to compute $\rho_0$ on 
-  radial faces.  In spherical geometry, we first map $\rho_0$ to Cartesian cell-centers 
-  using the mapping in section \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, 
-  and then use 2nd-order spatial averaging to compute $\rho_0$ on all faces.
-
-  \item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$
-  without explicitly including the reaction terms,
-\begin{equation}
-(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)} 
- - \dt \; \left\{ \nabla \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right)  
-  (\rho X_k)^{(1),\nph,\star} \right] \right\} \enskip ,
-\end{equation}
-\begin{equation}
-\rho^{(2),\star} = \sum_k (\rho X_k)^{(2),\star} \enskip ,
-\end{equation}
-\begin{equation}
-X_k^{(2),\star} = (\rho X_k)^{(2),\star} / \rho^{(2),\star}
-\end{equation}
-
-\end{enumerate}
-
-\item Define an edge-centered $\etarho^{\nph,\star}$:
-
-\begin{description}
-\item For planar geometry,
-\begin{equation}
- \etarho^{\nph,\star} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdot \eb_r + w_0^{\nph,\star}\right) (\rho X_k)^{(1),\nph,\star} \right] - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} \enskip ,
-\end{equation}
-\item For spherical geometry, first construct 
-$\eta_{\rho}^{{\rm cart},\nph,\star} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf,\star}$ using:
-\begin{equation}
-\eta_{\rho}^{{\rm cart},\nph,\star} = \left[\left(\frac{\rho^{(1)}+\rho^{(2),\star}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a),\star}}{2}\right)\right] \sum_d \left(\uadvone \cdot \eb_d\right)n_d.
-\end{equation}
-Then, $\etarho^{\nph,\star}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph,\star}$,
-\begin{equation}
-\etarho^{\nph,\star} = \overline{\eta_{\rho}^{{\rm cart},\nph,\star}}.
-\end{equation}
-This gives a cell-centered $\etarho^{\nph,\star}$.  To get $\etarho^{\nph,\star}$ at 
-radial edges, average the two neighboring radial cell-centers.
-\end{description}
-
-\item Set $\rho_0^{n+1,\star}$ to be the lateral average of $\rho^{(2),\star}$:
-\begin{description}
-\item For planar geometry, set $\rho_0^{n+1,\star} =$ {\bf Avg}$(\rho^{(2),\star})$
-\item For spherical geometry, set $\rho_0^{n+1,\star} = \rho_0^{(2a),\star} +$ {\bf Avg}$(\rho^{(2),\star} - \rho_0^{(2a),\star})$
-\end{description}
-
-\item {\bf Compute Gravity}$[\rho_0^{n+1,\star}] \rightarrow [g^{n+1,\star}]$.
-
-%{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
-%\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
-%\item For spherical geometry,\\
-%{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
-%\overline{S^{\nph,\star}}, \rho^{(2),\star}, X_k^{(2),\star}]$\\
-%$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-%\rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
-\item {\bf Enforce HSE}$[p_0^n,\rho_0^{n+1,\star},g^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
-
-\item Compute $\psi$.
-\begin{description}
-\item For planar geometry,
-\begin{equation}
-\psi_j^{n+\myhalf,\star} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{n+\myhalf,\star} 
-+ \eta_{\rho,j+\myhalf}^{n+\myhalf,\star}\right) g \enskip .
-\end{equation}
-\item For spherical geometry, first compute:
-\begin{equation}
-\gammabar^{(1)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, 
-X_k^{(1)}\right) \right] \enskip ,
-\end{equation}
-\begin{equation}
-\gammabar^{(2),\star} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2),\star}, p_0^{n+1,\star}, 
-X_k^{(2),\star}\right) \right] \enskip .
-\end{equation}
-Then, define a base time and time-centered $\psi$:
-\begin{equation}
-\psi_j^{n,\star} = 
-\left(\gammabar^{(1)} p_0^n\right)_j
-\left \{ \Sbar_j^{n+\myhalf,\star} - 
-\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
-          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip ,
-\end{equation}
-\begin{eqnarray}
-\psi_j^{n+\myhalf,\star} &=& 
-\left(\frac{\gammabar^{(1)}+\gammabar^{(2),\star}}{2}\right)_j
-\left(\frac{p_0^n+p_0^{n+1,\star}}{2}\right)_j \nonumber \\
-&& \left \{ \Sbar_j^{n+\myhalf,\star} - 
-\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf,\star})_{j+\myhalf} -
-          (r^2 w_0^{n+\myhalf,\star})_{j-\myhalf} \right ] \right \} \enskip .
-\end{eqnarray}
-\end{description}
-
-\item Update $(\rho h)_0$.
-\begin{description}
-\item For planar geometry,\\
-{\bf Advect Base Enthalpy Planar}$[(\rho h)_0^{n}, w_0^{\nph,\star}, \psi^{\nph,\star}]
-\rightarrow [(\rho h)_0^{n+1,\star}]$.
-\item For spherical geometry,\\
-{\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^{n}, w_0^{\nph,\star}, 
-\psi^{n,\star},\psi^{\nph,\star}] \rightarrow [(\rho h)_0^{n+1,\star}]$.
-\end{description}
-
-\item Update the enthalpy.  Here we consider only the advective terms, 
-neglecting the reaction terms and thermal diffusion terms, and solve a discretized version of
-  \begin{equation}
-  \frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h)
-+ \psi + (\Ubt \cdot \eb_r) \frac{\partial p_0}{\partial r} 
-\enskip , \label{eq:rhohupdadv} 
-  \end{equation}
-For spherical geometry, experience has shown that solving the analytically equivalent form,
-\begin{equation}
-  \frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h)
-+ \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt  
-\enskip , \label{eq:rhohupdadv2} 
-\end{equation}
-results in a better solution.
-
-  \begin{enumerate}
-  \renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-  \item Compute the enthalpy edge state, $(\rho h)^{(1),\nph,\star},$
-    for the conservative update of $(\rho h)^{(1)}.$  Here we predict 
-    $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0$ to time-centered edges to obtain 
-    $(\rho h)^{'(1),\nph,\star}$, 
-    using a second-order unsplit Godunov procedure, 
-    as described in paper II, Appendix A, using $\Vb =
-    \uadvone+w_0^{\nph,\star} \eb_r$.  We do not include the reaction
-    terms in the enthalpy prediction, since we accounted for them
-    already in {\bf React State}.  Convert the perturbational edge
-    state to a full state using
-
-\begin{equation}
-(\rho h)^{(1),\nph,\star} = 
-(\rho h)^{'(1),\nph,\star} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}
-\enskip .
-\end{equation}
-  For planar geometry, we use 4th-order spatial averaging to compute $(\rho h)_0$
-  on radial faces.  In spherical geometry, we first map $h_0 \equiv (\rho h)_0/\rho_0$ 
-  and $\rho_0$ to Cartesian cell-centers using the mapping in section 
-  \ref{Sec:1D Cell-Centered to Cartesian Cell-Centered}, use 2nd-order spatial 
-  averaging to compute $h_0$ and $\rho_0$ on all faces, and then multiply these 
-  terms to get $(\rho h)_0$ on all faces.
-
-  \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ without
-  explicitly including the reaction terms.
-
-\begin{description}
-\item For planar geometry,
-
-  \begin{eqnarray}
-  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nabla
-      \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right) (\rho
-      h)^{(1),\nph,\star} \right] \right\} \nonumber \\ && + \dt \;
-    \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial
-      p_0}{\partial r} \right)^{n} + \dt \; \psi^{\nph,\star} \enskip ,
-  \end{eqnarray}
-
-\item For spherical geometry,
-
-  \begin{eqnarray}
-  (\rho h)^{(2),\star} &=& (\rho h)^{(1)} - \dt \; \left\{ \nabla
-      \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right) (\rho
-      h)^{(1),\nph,\star} \right] \right\} \nonumber \\ 
-    && + \dt \; \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) 
-       - p_0^{n} \nabla \cdot \uadvone \right \} \nonumber \\
-    && + \dt \; \psi^{\nph,\star} \enskip ,
-  \end{eqnarray}
-
-\end{description}
-
-\end{enumerate}
-
-If {\tt do\_eos\_h\_above\_cutoff=T} (which is the default setting), then if
-$\rho^{(2),\star} < {\tt base\_cutoff\_density}$, then we recompute enthalpy using
-\begin{equation}
-(\rho h)^{(2),\star} = \rho^{(2),\star}h\left(\rho^{(2),\star},p_0^{n+1,\star},X_k^{(2),\star}\right).
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 4.1
-%--------------------------------------------------------------------------
-
-\item If we are using thermal diffusion, diffuse the enthalpy through a time interval of 
-$\dt$.  First, define $(\rho h)^{(1a),\star} = (\rho h)^{(2),\star}$.  We recompute
-$(\rho h)^{(2),\star}$ to account for thermal diffusion.  Here we begin
+\end{itemize}
+\subsection{Thermal Diffusion Changes}
+Immediately after {\bf Step 4H}, diffuse the enthalpy through 
+a time interval of $\dt$.  First, define $(\rho h)^{(1a),\star} = (\rho h)^{(2),\star}$.  
+We recompute $(\rho h)^{(2),\star}$ to account for thermal diffusion.  Here we begin
 with the enthalpy equation, but consider only the 
@@
 \begin{equation}
-  \frac{\partial (\rho h)}{\partial t}  = 
- \nabla\cdot\kth\nabla T 
-\enskip . \label{eq:rhohupdthermal} 
+\frac{\partial (\rho h)}{\partial t} = \nabla\cdot\kth\nabla T.
   \end{equation}
@@
  \sum_k \nabla\cdot \frac{\xi_k \kth}{c_p}\nabla X_k -
- \nabla\cdot \frac{h_p \kth}{c_p}\nabla p_0 
-\enskip . \label{eq:rhohupdthermal2} 
+ \nabla\cdot \frac{h_p \kth}{c_p}\nabla p_0.
   \end{equation}
-
 Compute $\kth^{(1)}, c_p^{(1)}$, and $\xi_k^{(1)}$ from $\rho^{(1)}, T^{(1)}$, and $X_k^{(1)}$ as inputs to the equation of state.  The update is given by
@@
 \end{eqnarray}
-\item Update the temperature using the equation of state.  For planar geometry,
-we update using
-\begin{equation}
-T^{(2),\star} = T\left(\rho^{(2),\star}, h^{(2),\star}, X_k^{(2),\star}\right) \enskip ,
-\end{equation}
-but in spherical geometry, we set {\tt use\_tfromp = T}, and we use
-\begin{equation}
-T^{(2),\star} = T\left(\rho^{(2),\star}, p_0^{n+1,\star}, X_k^{(2),\star}\right) \enskip .
-\end{equation}
-The latter has the effect of completely decoupling the enthalpy equation from the 
-evolution.
 
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 5
-%--------------------------------------------------------------------------
-\item[Step 5.] {\em React the full state through a second time interval of $\dt / 2.$}
-
-{\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, T^{(2),\star}, 
-(\rho\Hext)^{(2),\star}, p_0^{n+1,\star}] $\\
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, 
-X_k^{n+1,\star}, T^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
-
-%--------------------------------------------------------------------------
-% STEP 6
-%--------------------------------------------------------------------------
-\item[Step 6.] {\em Define the time-centered expansion, $S^{n+\myhalf}$, base state
-velocity, $w_0^{n+\myhalf}$, and base state velocity forcing.}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-\item Define
-\begin{equation}
-  S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + 
-  \sigma \Hnuc^{(2),\star} +
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star}  
-  + \sigma \Hext + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T^{n+1,\star} \enskip ,
-\end{equation} 
-where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} / \rho^{(2),\star}$
-and the thermodynamic quantities are defined using $\rho^{n+1,\star}, X_k^{n+1,\star},$ 
-and $T^{n+1,\star}$ as inputs to the equation of state.
-Then define
-\begin{equation}
- S^\nph = \frac{S^n + S^{n+1,\star}}{2} \enskip. 
-\end{equation}
-
-\item Compute
-\begin{equation}
-\overline{S^{\nph}} = {\mathrm{\bf Avg}} (S^{\nph}) \enskip.
-\end{equation}
-
-\item Define:
-\begin{eqnarray}
-\gammabar^{\nph} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}\enskip, \qquad
-\rhozero^{\nph} = \frac{\rhozero^{n} + \rhozero^{n+1,\star}}{2}\enskip, \nonumber \\
-p_0^{\nph} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}\enskip, \qquad
-g^\nph = \frac{g^n + g^{n+1,\star}}{2}\enskip ,
-\end{eqnarray}
-with
-\begin{equation}
- \gammabar^{n+1,\star} = {\rm{\bf Avg}} 
-\left[ \Gamma_1\left(\rho^{n+1,\star}, p_0^{n+1,\star}, X_k^{n+1,\star}\right) \right] 
-\enskip .
-\end{equation}
-
-\item Compute $w_0^{n+\myhalf}$:
-\begin{description}
-\item For planar geometry,\\
-{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph}},
-\gammabar^{n+\myhalf},p_0^{n+\myhalf},g^{n+\myhalf},\psi^{n+\myhalf,\star}]\rightarrow [w_0^{n+\myhalf}]$.
-
-If we are using the volume discrepancy term, compute
-\begin{equation}
-\overline{p^{\inp}} = \frac{\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}}{2}.
-\end{equation}
-\item For spherical geometry,\\
-{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph}},\gammabar^{n+\myhalf},
-\rhozero^{n+\myhalf},p_0^{n+\myhalf},\etarho^{n+\myhalf}] \rightarrow [w_0^{n+\myhalf}]$.
-\end{description}
-
-\item Define the base state velocity forcing using equation (\ref{eq:pizero}) 
-with the following discretization:
-\begin{equation}
-\left ( \frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} \right )^n = 
--\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
-- w_0^n \left(\frac{\partial w_0}{\partial r}\right)^n,
-\end{equation}
-where $w_0^{n}$ and $(\partial w_0 / \partial r)^{n}$ are defined as
-\begin{equation}
-w_0^n = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph}}{\dt^n+\dt^{n-1}}\enskip,
-\end{equation}
-\begin{equation}
-\left(\frac{\partial w_0}{\partial r}\right)^{n} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
-+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph} \right ] \enskip.
-\end{equation}
-If $n=0$, we use $\dt^{-1} = \dt^0$.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 7
-%--------------------------------------------------------------------------
-\item[Step 7.] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}
-
-The procedure to construct $\uadvtwo$ is described in detail in Appendix B of paper III
-and is analogous to the procedure used in step 2, but with updated values
-for $w_0$ and $\pizero.$  We note that $\uadvtwo$ satisfies the discrete versions of 
-equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}), with the discretization
-\begin{equation}
-\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph}\left(S^{\nph} - \overline{S^{\nph}}\right) \enskip ,
-\end{equation}
-\begin{equation}
-\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2}; ~~~~~
- \beta_0^{n+1,\star} = \beta \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \gammabar^{n+1,\star}, g^{n+1,\star}\right) \enskip ,
-\end{equation}
-where $\beta_0$ is computed as described in paper III, Appendix C.
-
-The discretization with the volume discrepancy correction is
-\begin{equation}
-\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph}\left\{\left(S^{\nph} - \overline{S^{\nph}}\right)
- + \frac{f}{\gammabar^{\nph} p_0^{\nph}}\cdot 
-\left[\frac{p(\rho,h,X_k)^{\nph} - \overline{p(\rho,h,X_k)^{\nph}}}{\Delta t^n}\right]\right\} \enskip ,
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 8
-%--------------------------------------------------------------------------
-\item[Step 8.] {\em Advect the base state and full state through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-
-\item {\bf Advect Base Density}$[\rho_0^{n},w_0^{\nph}] \rightarrow$ 
-$[\rho_0^{(2a)}, \rho_0^{\nph,{\rm pred}}]$.
-
-\item Update the species.  This step is identical to step 4b with the following changes:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the species edge states, $(\rho X_k)^{(1),n+\myhalf}$, 
-using $\Vb = \uadvtwo+w_0^{\nph}\eb_r$.  Convert the perturbational edge 
-state to a full state using 
-\begin{equation}
-\rho^{(1),\nph} = \rho^{'(1),\nph} + \frac{\rho_0^n + \rho_0^{(2a)}}{2}\enskip .
-\end{equation}
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ using
-\begin{equation}
-(\rho X_k)^{(2)} = (\rho X_k)^{(1)} 
-- \dt \; \left\{ \nabla \cdot \left[\left(\uadvtwo+w_0^{\nph} \eb_r\right)  
-(\rho X_k)^{(1),\nph} \right] \right\}  \enskip ,
-\end{equation}
-\begin{equation}
-\rho^{(2)} = \sum_k (\rho X_k)^{(2)} \enskip ,
-\end{equation}
-\begin{equation}
-X_k^{(2)} = (\rho X_k)^{(2)} / \rho^{(2)}
-\end{equation}
-
-\end{enumerate}
-
-\item Define an edge-centered $\etarho^{\nph}$:
-
-\begin{description}
-
-\item For planar geometry,
-\begin{equation}
- \etarho^{\nph} = {\rm {\bf Avg}} \sum_k \left [\left(\uadvtwo \cdot \eb_r + w_0^{\nph}\right) (\rho X_k)^{(1),\nph} \right] - w_0^{\nph} \rho_0^{\nph,{\rm pred}} \enskip ,
-\end{equation}
-\item For spherical geometry, first construct 
-$\eta_{\rho}^{{\rm cart},\nph} = [\rho'(\Ubt\cdot\eb_r)]^{n+\myhalf}$ using:
-\begin{equation}
-\eta_{\rho}^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a)}}{2}\right)\right] \sum_d \left(\uadvtwo \cdot \eb_d\right)n_d.
-\end{equation}
-Then, $\etarho^{\nph}$ is the cell-centered average of $\eta_{\rho}^{{\rm cart},\nph}$,
-\begin{equation}
-\etarho^{\nph} = \overline{\eta_{\rho}^{{\rm cart},\nph}}.
-\end{equation}
-This gives a cell-centered $\etarho^{\nph}$.  To get $\etarho^{\nph}$ at 
-radial edges, average the two neighboring cell-centers.
-\end{description}
-
-\item Set $\rho_0^{n+1}$ to be the lateral average of $\rho^{(2)}$:
-\begin{description}
-\item For planar geometry, set $\rho_0^{n+1} =$ {\bf Avg}$(\rho^{(2)})$
-\item For spherical geometry, set $\rho_0^{n+1} = \rho_0^{(2a)} +$ {\bf Avg}$(\rho^{(2)} - \rho_0^{(2a)})$
-\end{description}
-
-\item {\bf Compute Gravity}$[\rho_0^{n+1}] \rightarrow [g^{n+1}]$.
-
-Then, set $g^{n+\myhalf} = (g^n + g^{n+1})/2$.
-
-%\item Update $p_0$ and compute $\psi$.
-%\begin{description}
-%\item For planar geometry,\\
-%{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
-%$\rightarrow [p_0^{n+1},\psi^{\nph}]$.
-%\item For spherical geometry,\\
-%{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{(1)},\overline{S^{\nph}},
-%\rho^{(2)},X_k^{(2)}]$\\
-%$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-% \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
-%\end{description}
-\item {\bf Enforce HSE}$[p_0^n,\rho_0^{n+1},g^{n+1}] \rightarrow [p_0^{n+1}]$.
-
-\item Compute $\psi$.
-\begin{description}
-\item For planar geometry, 
-\begin{equation}
-\psi_j^{n+\myhalf} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{n+\myhalf} 
-+ \eta_{\rho,j+\myhalf}^{n+\myhalf}\right) g\enskip .
-\end{equation}
-\item For spherical geometry, first compute:
-\begin{equation}
-\gammabar^{(2)} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2)}, p_0^{n+1}, 
-X_k^{(2)}\right) \right] \enskip .
-\end{equation}
-Then, define a base time and time-centered $\psi$:
-\begin{equation}
-\psi_j^{n} = 
-\left(\gammabar^{(1)} p_0^n\right)_j
-\left \{ \Sbar_j^{n+\myhalf} - 
-\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf})_{j+\myhalf} -
-          (r^2 w_0^{n+\myhalf})_{j-\myhalf} \right ] \right \} \enskip ,
-\end{equation}
-\begin{eqnarray}
-\psi_j^{n+\myhalf} &=& 
-\left(\frac{\gammabar^{(1)}+\gammabar^{(2)}}{2}\right)_j
-\left(\frac{p_0^n+p_0^{n+1}}{2}\right)_j \nonumber \\
-&& \left \{ \Sbar_j^{n+\myhalf} - 
-\frac{1}{r_j^2} \left [ (r^2 w_0^{n+\myhalf})_{j+\myhalf} -
-          (r^2 w_0^{n+\myhalf})_{j-\myhalf} \right ] \right \} \enskip .
-\end{eqnarray}
-\end{description}
-
-\item Update $(\rho h)_0$.
-\begin{description}
-\item For planar geometry,\\
-{\bf Advect Base Enthalpy Planar}$[(\rho h)_0^n, w_0^{\nph}, \psi^{\nph}]
-\rightarrow [(\rho h)_0^{n+1}]$.
-\item For spherical geometry,\\
-{\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^n, w_0^{\nph}, \psi^n, \psi^{\nph}]
-\rightarrow [(\rho h)_0^{n+1}]$.
-\end{description}
-
-\item Update the enthalpy.  This step is identical to step 4i with the following changes:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the enthalpy edge state, $(\rho h)^{(1),\nph}$, using
-$\Vb = \uadvtwo+w_0^{\nph} \eb_r$.  Convert the perturbational edge 
-state to a full state using
-\begin{equation}
-(\rho h)^{(1),\nph} = 
-(\rho h)^{'(1),\nph} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1}}{2} \enskip .
-\end{equation}
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$
-without explicitly including the reaction terms,
-
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
-(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nabla \cdot \left[ \left(\uadvtwo+w_0^{\nph} \eb_r\right)  
-(\rho h)^{(1),\nph} \right] \right\} \nonumber \\
-&& + \frac{\dt}{2} \; \left(\uadvtwo \cdot \eb_r\right)
-\left [ \left(\frac{\partial p_0}{\partial r} \right)^{n}
-      + \left(\frac{\partial p_0}{\partial r} \right)^{n+1}  \right ] 
-+ \dt \; \psi^{\nph} \enskip ,
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2)} &=& (\rho h)^{(1)} - \dt \; \left\{ \nabla \cdot \left[ \left(\uadvtwo+w_0^{\nph} \eb_r\right)  
-(\rho h)^{(1),\nph} \right] \right\} \nonumber \\
-    && + \dt \; \left \{ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) 
-       - p_0^{\nph} \nabla \cdot \uadvtwo \right \} \nonumber \\
-    && + \dt \; \psi^{\nph} \enskip ,
-\end{eqnarray}
-
-\end{description}
-
-\end{enumerate}
-
-If {\tt do\_eos\_h\_above\_cutoff=T} (which is the default setting), then if
-$\rho^{(2)} < {\tt base\_cutoff\_density}$, then we recompute enthalpy using
-\begin{equation}
-(\rho h)^{(2)} = \rho^{(2)}h\left(\rho^{(2)},p_0^{n+1},X_k^{(2)}\right).
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 8.1
-%--------------------------------------------------------------------------
-\item If we are using thermal diffusion, diffuse the enthalpy through a time interval of 
+Immediately after {\bf Step 8H}, diffuse the enthalpy through a time interval of 
 $\dt$.  First, define $(\rho h)^{(1a)} = (\rho h)^{(2)}$.  We recompute $(\rho h)^{(2)}$ to 
@@
 \end{eqnarray}
-which is numerically implemented as a diffusion equation for $h^{(2)}$,
+which is numerically implemented as a diffusion equation for $h^{(2)}$.
 \begin{eqnarray}
@@
 \end{eqnarray}
-\item Update the temperature using the equation of state.  For planar geometry we
-update using
-\begin{equation}
-T^{(2)} = T\left(\rho^{(2)}, h^{(2)}, X_k^{(2)}\right) \enskip ,
-\end{equation}
-but in spherical geometry, we set {\tt use\_tfromp = T}, and we use
-\begin{equation}
-T^{(2)} = T\left(\rho^{(2)}, p_0^{n+1}, X_k^{(2)}\right) \enskip .
-\end{equation}
-The latter has the effect of completely decoupling the enthalpy equation from the 
-evolution.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 9
-%--------------------------------------------------------------------------
-\item[Step 9.] {\em React the full state through a second time interval of $\dt / 2.$}
-
-{\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho\Hext)^{(2)}, p_0^{n+1}]$\\
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\rightarrow [\rho^{n+1}, 
-(\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$
-
-%--------------------------------------------------------------------------
-% STEP 10
-%--------------------------------------------------------------------------
-\item[Step 10.] {\em Define the new time expansion $S^{n+1}$, and $\gammabar^{n+1}$.}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \alph{enumi}}}
-\item Define
-\begin{equation}
-  S^{n+1} =  -\sigma  \sum_k  \xi_k (\omegadot_k)^{(2)}  + \sigma \Hnuc^{(2)} +
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2)}  
-   + \sigma \Hext + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T^{n+1} \enskip ,
-\end{equation}
-where $(\omegadot_k)^{(2)} = (\rho \omegadot_k)^{(2)} / \rho^{(2)}$
-and the thermodynamic quantities are defined using $\rho^{n+1}, X_k^{n+1},$
-and $T^{n+1}$ as inputs to the equation of state.
-
-\item Compute
-\begin{equation}
-\overline{S^{n+1}} = {\mathrm{\bf Avg}} (S^{n+1}) \enskip.
-\end{equation}
-
-\item Define
-\begin{equation}
- \gammabar^{n+1} = {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1}, 
-X_k^{n+1}\right) \right] \enskip .
-\end{equation}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 11
-%--------------------------------------------------------------------------
-\item[Step 11.] {\em Update the velocity}.
-
-The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
-We update the velocity field $\Ubt^n$ to $\Ubt^{n+1,\star}$ by discretizing 
-equation (\ref{eq:utildeupd}) as
-\begin{eqnarray}
-\Ubt^{n+1,\star} &=& \Ubt^n - \dt \;
- \left[\left(\uadvtwo+ w_0^{\nph} \eb_r\right) \cdot \nabla \Ubt \right]^\nph
-      - \dt \; \left(\uadvtwo \cdot \eb_r\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \eb_r \nonumber \\
-   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
-        + \frac{1}{\rho_0} {\mathbf{G}} \pi_0^n
-        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g^{n+\myhalf} \eb_r \right] \enskip ,
-\end{eqnarray}
-where $\rho^{\nph}$ is defined as
-\begin{equation}
-\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}\enskip ,
-\end{equation}
-and $\mathbf{G}$ approximates a cell-centered gradient from nodal data.
-The construction of $[(\uadvtwo+ w_0^{\nph} \eb_r \cdot \nabla) \Ubt ]^\nph$
-is described in paper II, Appendix A, with $\Vb = \uadvtwo+w_0^{\nph} \eb_r$ and $s$ 
-set to each component of $\Ubt^n$ individually.  Here, the $\star$ superscript 
-does not refer to the predictor-corrector aspect of the algorithm, but rather 
-than the updated velocity does not satisfy the divergence constraint.
-
-Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}):
-\begin{equation}
-\nabla \cdot \left(\beta_0^{\nph} \Ubt^{n+1} \right) 
-= \beta_0^{\nph} \left(S^{n+1} - \overline{S^{n+1}} \right) \enskip,
-\end{equation}
-where $\beta_0^{n+\myhalf}$ is now defined as
-\begin{equation}
-\beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1}}{2}; ~~~~~ 
-\beta_0^{n+1} = \beta \left(\rho_0^{n+1}, p_0^{n+1}, \gammabar^{n+1}, g^{n+1}\right)\enskip .
-\end{equation}
-
-The discretization with the volume discrepancy term is:
-\begin{equation}
-\nabla \cdot \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S^{n+1}} \right)
-+ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\}
-\enskip .
-\end{equation}
-
-To solve this equation, we define
-${\bf V} = \Ubt^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
-and solve
-\begin{equation}
- L_\beta^\rho \phi = D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} 
-\left(S^{n+1}-\overline{S^{n+1}} \right) \enskip, \label{Linear System for Multigrid}
-\end{equation}
-for nodal values of $\phi$, where $L_\beta^\rho$ is the standard bilinear
-finite element approximation to $\nabla \cdot ({\beta_0^\nph}/{\rho^\nph}) \nabla.$
-The discretization with the volume discrepancy term is:
-\begin{equation}
- L_\beta^\rho \phi =
-   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph}\left\{ \left(S^{n+1}-\overline{S^{n+1}} \right)
-- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
-\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)^{n+1}}}{\Delta t^n}\right]\right\} \enskip.
-\end{equation}
-In this step, $D$ is a discrete second-order operator that approximates the 
-divergence at nodes from cell-centered data and satisfies
-$\mathbf{G} = -D^T.$ 
-(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
-approximate projection; see (cite almgren:bell:crutchfield) for a discussion
-of this particular form of the projection operand.)  
-We solve the linear system of equation (\ref{Linear System for Multigrid})
-using multigrid V-cycles with Gauss-Seidel relaxation.
-
-We determine the new time velocity field from
-\begin{equation}
-\Ubt^{n+1} = {\bf V} - \frac{1}{\rho^\nph} \mathbf{G} \phi \enskip ,
-\end{equation}
-and the new time-centered perturbational pressure from
-\begin{equation}
-  \pi^\nph = \frac{1}{\dt} \phi \enskip .
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 12
-%--------------------------------------------------------------------------
-\item[Step 12.] {\em Compute a new $\dt.$}
-
-Compute $\dt$ for the next time step with the procedure described in \S 3.4 of paper III
-using $w_0$ as computed in step 6 and
-$\Ubt^{n+1}$ as computed in step 11.  We use this $\dt$ in the next time step.
-
-\end{description}
-
-\noindent This completes the time advancement of the algorithm.
 
@@
 The initial perturbational pressure also needs to be determined for
-use in steps 2, 7 and 11. 
+use in {\bf Steps 2}, {\bf 7}, and {\bf 11}. 
 This is done through a second iterative procedure which follows the
-time advancement algorithm as described in steps 1-11 in 
+time advancement algorithm as described in {\bf Steps 1-11} in 
 \S \ref{Sec:Time Advancement Algorithm}.  
@@
 
-  \item Compute $w_0^{\nu}$ as in step 1c using $\overline{S^{0,\nu}}$ and $\psi=0$.
+  \item Compute $w_0^{\nu}$ as in {\bf Step 2B} using $\overline{S^{0,\nu}}$ and $\psi=0$.
 
@@
   \begin{enumerate}
-  \item Perform steps 1--11 as described above, using 
-    $S^{\myhalf,\star} = (S^0 + S^{1,\nu-1})/2$ in step 1 as described.
-    The only other difference in the time advancement is that in step 11
+  \item Perform {\bf Steps 1-11} as described above, using 
+    $S^{\myhalf,\star\star} = (S^0 + S^{1,\nu-1})/2$ in {\bf Step 2} as described.
+    The only other difference in the time advancement is that in {\bf Step 11}
     we define ${\bf V} = (\Ubt^{1,\star} - \Ubt^0)$ and solve
-    \begin{equation}  L_\beta^\rho \phi =
-      D \left ( \beta_0^{\myhalf} {\bf V} \right) - \beta_0^{\myhalf} \left[ \left(S^{1}-\overline{S^{1}}\right) - \left(S^{0}-\overline{S^{0}}\right) \right] \enskip . 
+    \begin{equation}  L_\beta^\rho \phi = D \left ( \beta_0^{\myhalf} {\bf V} \right) - \beta_0^{\myhalf} \left[ \left(S^{1}-\overline{S^{1}}\right) - \left(S^{0}-\overline{S^{0}}\right) \right] \enskip .
     \end{equation}
@@
       These steps also yield new scalar data at time $\dt,$ which
-      we discard,  and new values for $\etarho^{\myhalf}$ (step 8), $\psi^{\myhalf}$ (step 8), 
-      $S^{1,\nu}$ (step 10), and $\pi^{\myhalf}$ (step 11), which we keep.
+      we discard, and new values for $\etarho^{\myhalf}$ ({\bf Step 8}), 
+      $\psi^{\myhalf}$ ({\bf step 8}), 
+      $S^{1,\nu}$ ({\bf step 10}), and $\pi^{\myhalf}$ ({\bf step 11}), which we keep.
     \item Set $\pi^{-\myhalf} = \pi^{\myhalf}$, $\etarho^{-\myhalf} = \etarho^{\myhalf}$,

### 2009-10-20 docs/flowchart/flowchart.tex
more updates

@@
 The initial perturbational pressure also needs to be determined for
-use in {\bf Steps 2}, {\bf 7}, and {\bf 11}. 
+use in {\bf Steps 3}, {\bf 7}, and {\bf 11}. 
 This is done through a second iterative procedure which follows the
@@
 \addtocounter{enumi}{1}
-
 \item {\bf Do} {$\nu = 1,...,N_{\rm iters}^{\pi}$.}
-
   \begin{enumerate}
@@
       These steps also yield new scalar data at time $\dt,$ which
-      we discard, and new values for $\etarho^{\myhalf}$ ({\bf Step 8}), 
-      $\psi^{\myhalf}$ ({\bf step 8}), 
-      $S^{1,\nu}$ ({\bf step 10}), and $\pi^{\myhalf}$ ({\bf step 11}), which we keep.
+      we discard, and new values for $\etarho^{\myhalf}$ ({\bf Step 8C}), 
+      $\psi^{\myhalf}$ ({\bf Step 8F}), 
+      $S^{1,\nu}$ ({\bf Step 10A}), and $\pi^{\myhalf}$ ({\bf Step 11}), which we keep.
     \item Set $\pi^{-\myhalf} = \pi^{\myhalf}$, $\etarho^{-\myhalf} = \etarho^{\myhalf}$,

### 2010-01-20 docs/architecture/architecture.tex
updates

@@
 on the corners.  In boxlib, data that is on an edge is termed `nodal'
-in that direction (see figure~\ref{fig:dataloc}.  Data that is on the
+in that direction (see Figure~\ref{fig:dataloc}).  Data that is on the
 corners is nodal in all spatial directions.  In MAESTRO, the state
@@
   In each of these centerings, the red point has the same indices:\ (1,2).
-  Not show is the case where data is nodal in the $y$-direction only.}
+  Not shown is the case where data is nodal in the $y$-direction only.}
 \end{figure}
@@
 \item {\tt dim} is the dimensionality of the data
-\item {\tt bx} is a \boxtype\ describing where the data in this
-  \fab\ is defined.
-\item {\tt pbx} is a \boxtype\ that extends {\tt bx} to include
-  any ghostcells
-\item {\tt ibx}
+\item {\tt bx} is a \boxtype\ describing the index space for which this \fab\ is defined
+\item {\tt ibx} is a \boxtype\ that describes the index range of the valid data
+  for the \fab.  This differes from {\tt bx} when the box is nodal in one more more 
+  directions.
+\item {\tt pbx} is a \boxtype\ that extends {\tt ibx} to include
+  any ghostcells, or equivalently, the physical box for the \fab\
 \item {\tt nc} is the number of components (how many variables at  
-  each grid location).
+  each grid location)
 \item {\tt p} is a pointer to the data, in this case double precision

### 2010-02-01 docs/managing_jobs/managingjobs.tex
add documentation on how to run the process scripts to manage jobs

\section{Automatic Restarting and Archiving of Data}

The submission script {\tt jaguar.run} and shell script
{\tt process.jaguar} in {\tt wdconvect/job\_scripts/}
are designed to allow you to run MAESTRO with minimal interaction,
while being assured that the data is archived to HPSS on the NCCS
machines.

To use the scripts, first create a directory in HPSS that has the same
name as the directory on lustre you are running in (just the directory
name, not the full path).  E.g.\ if you are running in a directory
call {\tt wdconvect\_run}, then do:
\begin{verbatim}
hsi
mkdir wdconvect_run
\end{verbatim}

The script {\tt process.jaguar} is called from {\tt jaguar.run} and
will run in the background and continually wait until checkpoint or
plotfiles are created (actually, it always leaves the most recent one
alone, since data may still be written to it, so it waits until there
are more than 1 in the directory).

Then the script will use {\tt htar} to archive the plotfiles and checkpoints
to HPSS.  If the {\tt htar} command was successful, then the plotfiles are
copied into a {\tt plotfile/} subdirectory.  This is actually important,
since you don't want to try archiving the data a second time and
overwriting the stored copy, especially if a purge took place.

Additionally, if you have the path to the {\tt ftime} executable set in the
script ({\tt ftime.f90} lives in {\tt fParallel/data\_processing/}), then the
script will create a file called {\tt ftime.out} that lists the name of the
plotfile and the corresponding simulation time.

Finally, right when the job is submitted, the script will tar up all
of the diagnostic files created by {\tt diag.f90} and archive them on HPSS.
The {\tt .tar} file is given a name that contains the date-string to allow
multiple archives to co-exist.

This really allows you to run the job and have all of the data stored
long term automatically.  This way you don't need to worry about
filesystem purges.  It seems to work well.

Also, the {\tt jaguar.run} submission script has code in it that will look
at the most recently generated checkpoint files, make sure that they
were written out correctly (it looks to see if there is a Header file,
since that is the last thing written), and automatically set the
{\tt --restart} flag on the run line to restart from the most recent
checkpoint file.

This allows you to job-chain a bunch of submission and have them wait
until the previous job finished and then automatically queue up:
\begin{verbatim}
qsub -W depend=afterany:<JOB-ID>  <QSUB SCRIPT>
\end{verbatim}

This way you can queue up a bunch of runs and literally leave things
alone and it will restart from the right place automatically and store
the data as it is generated.

When {\tt process.jaguar} is running, it creates a lockfile (called
{\tt process.pid}) that ensures that only one instance of the script
is running at any one time.  Sometimes if the machine crashes, the
{\tt process.pid} file will be left behind, in which case, the script
aborts.  Just delete that if you know the script is not running.

Analogous scripts exist for running on Atlas, with the same general
procedure.

### 2010-02-11 docs/architecture/architecture.tex
some additions from a while back that were never checked in

@@
 
+\subsection{{\tt eos\_module}}
+
 \subsection{{\tt fill\_3d\_module}}
@@
 \subsection{{\tt geometry}}
+
+\subsection{{\tt network}}
 
@@
 \section{Visualization / Analysis}
+
+\subsection{Amrvis}
+
+\subsection{{\tt data\_processing} scripts}
+
+Several useful analysis scripts can be found in {\tt
+fParallel/data\_processing}.  The makefile there needs
+to be edited to indicate which of the tools to build.
+
+\begin{itemize}
+
+\item {\tt fextract3d}
+
+  Extracts a 1-d slice from a plotfile, through the center of the
+  domain in any coordinate direction.  Either all of the variables or
+  a single variable is output.
+
+\item {\tt fextrema}
+
+  Outputs the min/max of each of the quantities in the plotfile.
+
+\item {\tt ftime}
+
+  Outputs the simulation time for one or more plotfiles.
+
+\end{itemize}
+
+\subsection{VisIt}

### 2010-03-04 docs/analysis/analysis.tex
restructuring
more analysis routine documentation

@@
+\section{The {\tt data\_processing} Routines}
+
 The {\tt fParallel/data\_processing/} directory contains a large
@@
 
-\section{General Analysis Routines}
+To compile any of the individual routines, edit the {\tt GNUmakefile}
+add uncomment the line beginning with `{\tt programs +=}' containing
+the routine you want to build.
+
+\subsection{General Analysis Routines}
 
@@
 
-\item {\tt faverage2d.f90}, {\tt faverage3d.f90}
+\item {\tt faverage.f90}
 
-  Laterally average each of the variables in a plotfile (separate
-  routines exist for 2-d and 3-d).  This is written with MAESTRO
-  plane-paralle geometry plotfiles in mind, and the averaging is done
+  Laterally average each of the variables in a plotfile (works for
+  both 2-d and 3-d).  This is written with MAESTRO
+  plane-parallel geometry plotfiles in mind, and the averaging is done
   over the coordinate direction(s) perpendicular to gravity.
+
+\item {\tt fboxinfo.f90}
+
+  Print out some basic information about the number of boxes on each
+  refinement level and (optionally) the bounds of each of the boxes.
 
@@
 
-\item {\tt fIDLdump.f90}
+\item {\tt fIDLdump.f90}, {\tt fIDLdump3d.f90}
 
@@
 
+\subsection{Data Processing Example}
+
+A well-commented example processing routine is provided in {\tt
+  data\_processing/tutorial} as {\tt fspeciesmass2d.f90}.  This
+routine simply computes the total mass of a particular species on the
+entire domain for a 2-d dataset.  It is written to understand a
+multilevel (AMR) dataset, and only considers the finest-available data
+at any physical location in the computational domain.
+
+{\tt fspeciesmass2d.f90} should provide a good starting point for
+writing a new analysis routine for BoxLib data.
+
 \end{itemize}

### 2010-03-05 docs/FAQ/faq.tex
fix options in the initial projection question
add Andy's profiling tip

@@
     init_divu_iter = 0
-    do_initial_projection = T
+    do_initial_projection = F
 \end{verbatim}
+
+\item {\em How can I obtain profiling infomation for my run?}
+
+  The code is already instrumented with timers.  Simply compile with
+  {\tt PROF=TRUE} in the {\tt GNUmakefile}, or equvalently do 
+  {\tt make PROF=t}
 

### 2010-03-10 docs/managing_jobs/managingjobs.tex
update job-chaining magic for atlas

@@
 \end{verbatim}
-
+where {\tt <JOB-ID>} is the id number of the job that must complete
+before the new submission runs and {\tt QSUB SCRIPT} is the submission
+script (e.g.\ {\tt jaguar.run}).
 This way you can queue up a bunch of runs and literally leave things
@@
 Analogous scripts exist for running on Atlas, with the same general
-procedure.
+procedure.  The command to chain a job on atlas is:
+\begin{verbatim}
+msub -l depend=<JOB-ID> <MSUB SCRIPT>
+\end{verbatim}
+where, again, {\tt <JOB-ID>} is the id number of the job that must
+complete before the current submission runs and {\tt <MSUB SCRIPT>}
+is the job submission script (e.g.\ {\tt atlas.run}).

### 2010-03-25 docs/architecture/architecture.tex
add a description on the GNUmakefile

@@
 
+Different problems in MAESTRO are contained in individual
+sub-directories under {\tt MAESTRO/}.  The minimum requirement to
+define a new problem would be a {\tt GNUmakefile} which describes how
+to build the application and an input file which lists the runtime
+parameters.  The problem-specific executable is built in the problem
+directory by typing {\tt make}.  Source files are found automatically
+by searching the directories listed in the {\tt GNUmakefile}.
+
+\subsection{The {\tt GNUmakefile}}
+
+A basic {\tt GNUmakefile} begins with:
+\begin{verbatim}
+  NDEBUG := t
+  MPI    :=
+  OMP    :=
+\end{verbatim}
+Here, {\tt NDEBUG} is true is we are building an optimized executable.  Otherwise,
+the debug version is built---this typically uses less optimization and adds various
+runtime checks through compiler flags.  {\tt MPI} and {\tt OMP} are set to true if
+we want to use either MPI or OpenMP for parallelization.
+
+The MAESTRO build system knows what options to use for various compiler families.
+The {\tt COMP} flag specifies which compiler to us.  Allowed values include
+{\tt Intel}, {\tt gfortran}, {\tt PGI}, and {\tt PathScale}.  The specific details
+of these choices are defined in the {\tt MAESTRO/mk/} directory.
+\begin{verbatim}
+  COMP := Intel
+  MKVERBOSE := t
+\end{verbatim}
+{\tt MKVERBOSE} set to true will echo the build commands to the terminal as the
+are executed.
+
+The next set of lines tell the build system where to look for source
+files.  A MAESTRO application is built from several pacakges (the
+multigrid solver, an EOS, a reaction network, etc.).  
+\begin{verbatim}
+  FPARALLEL := ../..
+
+  include $(FPARALLEL)/mk/GMakedefs.mak
+
+  Fmdirs := boxlib \
+            mg \
+            MAESTRO \
+            extern/EOS/helmeos \
+            extern/EOS/conducteos \
+            extern/constants \
+            extern/networks/ignition \
+            extern/VODE
+
+  Fmincludes := extern/EOS/helmeos
+\end{verbatim}
+The {\tt FPARALLEL} variable should point to the top-level directory
+of the source tree (i.e.\ {\tt fParallel/}).  Next we include the
+compiler definitions, {\tt mk/GMakedefs.mak}.  The {\tt Fmdirs}
+variable lists the individual directories in the {\tt fParallel/}
+source tree that are used to build the problem.  Generally, one does
+not need to include the problem directory itself, unless there are
+unique source files found there, described in a {\tt GPackage.mak}
+file.  The build system will attempt to build all of the files listed
+in the various {\tt GPackage.mak} files found in the {\tt Fmdirs}
+directories.  Furthermore, {\tt Fmdirs} will be will be added to the
+{\tt make} {\tt VPATH}, which is the list of directories to search for
+source files.  The problem directory will always be put first in the
+{\tt VPATH}, so any source files placed there override those with the
+same name found elsewhere in the source tree.  The {\tt Fmincludes} variable
+lists all those directories that contain include files that are inserted into
+the Fortran source at compile time via the {\tt include} statement.  Presently,
+the only instance of this is with the Helmholtz general equation of state
+found in {\tt extern/EOS/helmeos/}.
+
+Runtime parameters listed in the {\tt MAESTRO/\_parameters} file are
+parsed at compile time and the file {\tt probin.f90} is written and 
+compiled.  This is a Fortran module that holds the values of the runtime
+parameters and makes them available to any routine.  The variable
+{\tt PROBIN\_PARAMETERS} should list all of the files that contain
+runtime parameters for the current problem.  You must always
+include the main {\tt ../\_parameters} file (i.e.\ {\tt MAESTRO/\_parameters}).
+Sometimes a problem defines their own runtime parameters, in which
+case, its parameter file should also appear in {\tt PROBIN\_PARAMETERS}.
+\begin{verbatim}
+  PROBIN_PARAMETERS := ../_parameters
+\end{verbatim}
+
+The final lines in the {\tt GNUmakefile} include the rules to actually
+build the executable.
+\begin{verbatim}
+  include $(FPARALLEL)/MAESTRO/GMaestro.mak
+  include $(FPARALLEL)/mk/GMakerules.mak
+\end{verbatim}
+
+As mentioned above, any source files placed in the problem directory
+override a files with the same name found elsewhere in the source
+tree.  This allows you to create a problem-specific version of any
+routine.  Source files that are unique to this problem (i.e.\ there is
+no file with the same name elsewhere in the source tree) need to be
+listed in a file {\tt GPackage.mak} in the problem directory.
+
 \subsection{Plane-Parallel Geometry}

### 2010-03-25 docs/architecture/architecture.tex
reorganization
another reorg fix
add reference

@@
 \end{itemize}
+
+\section{Running A Problem}
+
+Different problems in MAESTRO are contained in individual
+sub-directories under {\tt MAESTRO/}.  The {\tt GNUmakefile}
+in each problem directory lists the components of {\tt MAESTRO}
+that are used to build the executable.  Full details of the
+{\tt GNUmakefile} can be found in \S~\ref{sec:adding_problems}.
+
+Some of the test problems available are:
+\begin{itemize}
+\item {\tt flame}
+\item {\tt planar\_heat}
+\item {\tt spherical\_heat}
+\item {\tt test2}
+\item {\tt test\_convect}
+\item {\tt test\_spherical}
+\end{itemize}
+
+\subsection{Unit Tests}
+
+In addition to the problems descibed above which use the full
+capabilities of MAESTRO, there are a number of unit tests that
+exercise only specific components of the MAESTRO solvers.  The
+tests have their own drivers (a custom {\tt varden.f90}) that
+initialize only the data needed for the specific test and call
+specific MAESTRO routines directly.
+
+\begin{itemize}
+\item {\tt test\_advect} \\[-3mm]
+
+  This test initializes a Gaussian density field (no other scalar
+  quantities are used) and a uniform velocity field in any one of the
+  coordinate directions.  The Gaussian profile is advected through
+  the period domain exactly once and the error in the density profile
+  (L2 norm) is computed.  The driver for this problem does this 
+  for every dimension twice (once with a positive velocity and once
+  with a negative velocity).  After all coordinate directions are 
+  tested, the norms are compared to ensure that the error does
+  not show any directional bias.
+
+\item {\tt test\_average} \\[-3mm]
+
+  This test initializes a 1-d radial base state quantity with a
+  Gaussian distribution, maps it into the 3-d domain (assuming a
+  spherical geometry) using the routines provided by the {\tt
+    fill\_3d\_module} module, and then calls {\tt average} to put it
+  back onto a 1-d radial array.  This way we test the accuracy of our
+  procedure to map between the 1-d radial and 3-d Cartesian states.
+  The output from this test was described in detail in
+  \cite{multilevel}.
+
+\item {\tt test\_basestate}
+\item {\tt test\_diffusion}
+\end{itemize}
+
+\section{Adding A New Problem}
+\label{sec:adding_problems}
+
+The minimum requirement to
+define a new problem would be a {\tt GNUmakefile} which describes how
+to build the application and an input file which lists the runtime
+parameters.  The problem-specific executable is built in the problem
+directory by typing {\tt make}.  Source files are found automatically
+by searching the directories listed in the {\tt GNUmakefile}.
+
+\subsection{The {\tt GNUmakefile}}
+
+A basic {\tt GNUmakefile} begins with:
+\begin{verbatim}
+  NDEBUG := t
+  MPI    :=
+  OMP    :=
+\end{verbatim}
+Here, {\tt NDEBUG} is true is we are building an optimized executable.  Otherwise,
+the debug version is built---this typically uses less optimization and adds various
+runtime checks through compiler flags.  {\tt MPI} and {\tt OMP} are set to true if
+we want to use either MPI or OpenMP for parallelization.
+
+The MAESTRO build system knows what options to use for various compiler families.
+The {\tt COMP} flag specifies which compiler to us.  Allowed values include
+{\tt Intel}, {\tt gfortran}, {\tt PGI}, and {\tt PathScale}.  The specific details
+of these choices are defined in the {\tt MAESTRO/mk/} directory.
+\begin{verbatim}
+  COMP := Intel
+  MKVERBOSE := t
+\end{verbatim}
+{\tt MKVERBOSE} set to true will echo the build commands to the terminal as the
+are executed.
+
+The next set of lines tell the build system where to look for source
+files.  A MAESTRO application is built from several pacakges (the
+multigrid solver, an EOS, a reaction network, etc.).  
+\begin{verbatim}
+  FPARALLEL := ../..
+
+  include $(FPARALLEL)/mk/GMakedefs.mak
+
+  Fmdirs := boxlib \
+            mg \
+            MAESTRO \
+            extern/EOS/helmeos \
+            extern/EOS/conducteos \
+            extern/constants \
+            extern/networks/ignition \
+            extern/VODE
+
+  Fmincludes := extern/EOS/helmeos
+\end{verbatim}
+The {\tt FPARALLEL} variable should point to the top-level directory
+of the source tree (i.e.\ {\tt fParallel/}).  Next we include the
+compiler definitions, {\tt mk/GMakedefs.mak}.  The {\tt Fmdirs}
+variable lists the individual directories in the {\tt fParallel/}
+source tree that are used to build the problem.  Generally, one does
+not need to include the problem directory itself, unless there are
+unique source files found there, described in a {\tt GPackage.mak}
+file.  The build system will attempt to build all of the files listed
+in the various {\tt GPackage.mak} files found in the {\tt Fmdirs}
+directories.  Furthermore, {\tt Fmdirs} will be will be added to the
+{\tt make} {\tt VPATH}, which is the list of directories to search for
+source files.  The problem directory will always be put first in the
+{\tt VPATH}, so any source files placed there override those with the
+same name found elsewhere in the source tree.  The {\tt Fmincludes} variable
+lists all those directories that contain include files that are inserted into
+the Fortran source at compile time via the {\tt include} statement.  Presently,
+the only instance of this is with the Helmholtz general equation of state
+found in {\tt extern/EOS/helmeos/}.
+
+Runtime parameters listed in the {\tt MAESTRO/\_parameters} file are
+parsed at compile time and the file {\tt probin.f90} is written and 
+compiled.  This is a Fortran module that holds the values of the runtime
+parameters and makes them available to any routine.  The variable
+{\tt PROBIN\_PARAMETERS} should list all of the files that contain
+runtime parameters for the current problem.  You must always
+include the main {\tt ../\_parameters} file (i.e.\ {\tt MAESTRO/\_parameters}).
+Sometimes a problem defines their own runtime parameters, in which
+case, its parameter file should also appear in {\tt PROBIN\_PARAMETERS}.
+\begin{verbatim}
+  PROBIN_PARAMETERS := ../_parameters
+\end{verbatim}
+
+The final lines in the {\tt GNUmakefile} include the rules to actually
+build the executable.
+\begin{verbatim}
+  include $(FPARALLEL)/MAESTRO/GMaestro.mak
+  include $(FPARALLEL)/mk/GMakerules.mak
+\end{verbatim}
+
+As mentioned above, any source files placed in the problem directory
+override a files with the same name found elsewhere in the source
+tree.  This allows you to create a problem-specific version of any
+routine.  Source files that are unique to this problem (i.e.\ there is
+no file with the same name elsewhere in the source tree) need to be
+listed in a file {\tt GPackage.mak} in the problem directory.
+
+\subsection{Plane-Parallel Geometry}
+
+\subsection{Full Star / Spherical}
+
+\subsection{Small-scale Flame}
+
+\section{Visualization / Analysis}
+
+\subsection{Amrvis}
+
+\subsection{{\tt data\_processing} scripts}
+
+Several useful analysis scripts can be found in {\tt
+fParallel/data\_processing}.  The makefile there needs
+to be edited to indicate which of the tools to build.
+
+\begin{itemize}
+
+\item {\tt fextract3d}
+
+  Extracts a 1-d slice from a plotfile, through the center of the
+  domain in any coordinate direction.  Either all of the variables or
+  a single variable is output.
+
+\item {\tt fextrema}
+
+  Outputs the min/max of each of the quantities in the plotfile.
+
+\item {\tt ftime}
+
+  Outputs the simulation time for one or more plotfiles.
+
+\end{itemize}
+
+\subsection{VisIt}
 
@@
 The {\tt variables} module provides
-
-\section{Adding A Problem}
-
-Different problems in MAESTRO are contained in individual
-sub-directories under {\tt MAESTRO/}.  The minimum requirement to
-define a new problem would be a {\tt GNUmakefile} which describes how
-to build the application and an input file which lists the runtime
-parameters.  The problem-specific executable is built in the problem
-directory by typing {\tt make}.  Source files are found automatically
-by searching the directories listed in the {\tt GNUmakefile}.
-
-\subsection{The {\tt GNUmakefile}}
-
-A basic {\tt GNUmakefile} begins with:
-\begin{verbatim}
-  NDEBUG := t
-  MPI    :=
-  OMP    :=
-\end{verbatim}
-Here, {\tt NDEBUG} is true is we are building an optimized executable.  Otherwise,
-the debug version is built---this typically uses less optimization and adds various
-runtime checks through compiler flags.  {\tt MPI} and {\tt OMP} are set to true if
-we want to use either MPI or OpenMP for parallelization.
-
-The MAESTRO build system knows what options to use for various compiler families.
-The {\tt COMP} flag specifies which compiler to us.  Allowed values include
-{\tt Intel}, {\tt gfortran}, {\tt PGI}, and {\tt PathScale}.  The specific details
-of these choices are defined in the {\tt MAESTRO/mk/} directory.
-\begin{verbatim}
-  COMP := Intel
-  MKVERBOSE := t
-\end{verbatim}
-{\tt MKVERBOSE} set to true will echo the build commands to the terminal as the
-are executed.
-
-The next set of lines tell the build system where to look for source
-files.  A MAESTRO application is built from several pacakges (the
-multigrid solver, an EOS, a reaction network, etc.).  
-\begin{verbatim}
-  FPARALLEL := ../..
-
-  include $(FPARALLEL)/mk/GMakedefs.mak
-
-  Fmdirs := boxlib \
-            mg \
-            MAESTRO \
-            extern/EOS/helmeos \
-            extern/EOS/conducteos \
-            extern/constants \
-            extern/networks/ignition \
-            extern/VODE
-
-  Fmincludes := extern/EOS/helmeos
-\end{verbatim}
-The {\tt FPARALLEL} variable should point to the top-level directory
-of the source tree (i.e.\ {\tt fParallel/}).  Next we include the
-compiler definitions, {\tt mk/GMakedefs.mak}.  The {\tt Fmdirs}
-variable lists the individual directories in the {\tt fParallel/}
-source tree that are used to build the problem.  Generally, one does
-not need to include the problem directory itself, unless there are
-unique source files found there, described in a {\tt GPackage.mak}
-file.  The build system will attempt to build all of the files listed
-in the various {\tt GPackage.mak} files found in the {\tt Fmdirs}
-directories.  Furthermore, {\tt Fmdirs} will be will be added to the
-{\tt make} {\tt VPATH}, which is the list of directories to search for
-source files.  The problem directory will always be put first in the
-{\tt VPATH}, so any source files placed there override those with the
-same name found elsewhere in the source tree.  The {\tt Fmincludes} variable
-lists all those directories that contain include files that are inserted into
-the Fortran source at compile time via the {\tt include} statement.  Presently,
-the only instance of this is with the Helmholtz general equation of state
-found in {\tt extern/EOS/helmeos/}.
-
-Runtime parameters listed in the {\tt MAESTRO/\_parameters} file are
-parsed at compile time and the file {\tt probin.f90} is written and 
-compiled.  This is a Fortran module that holds the values of the runtime
-parameters and makes them available to any routine.  The variable
-{\tt PROBIN\_PARAMETERS} should list all of the files that contain
-runtime parameters for the current problem.  You must always
-include the main {\tt ../\_parameters} file (i.e.\ {\tt MAESTRO/\_parameters}).
-Sometimes a problem defines their own runtime parameters, in which
-case, its parameter file should also appear in {\tt PROBIN\_PARAMETERS}.
-\begin{verbatim}
-  PROBIN_PARAMETERS := ../_parameters
-\end{verbatim}
-
-The final lines in the {\tt GNUmakefile} include the rules to actually
-build the executable.
-\begin{verbatim}
-  include $(FPARALLEL)/MAESTRO/GMaestro.mak
-  include $(FPARALLEL)/mk/GMakerules.mak
-\end{verbatim}
-
-As mentioned above, any source files placed in the problem directory
-override a files with the same name found elsewhere in the source
-tree.  This allows you to create a problem-specific version of any
-routine.  Source files that are unique to this problem (i.e.\ there is
-no file with the same name elsewhere in the source tree) need to be
-listed in a file {\tt GPackage.mak} in the problem directory.
-
-\subsection{Plane-Parallel Geometry}
-
-\subsection{Full Star / Spherical}
-
-\subsection{Small-scale Flame}
-
-\section{Visualization / Analysis}
-
-\subsection{Amrvis}
-
-\subsection{{\tt data\_processing} scripts}
-
-Several useful analysis scripts can be found in {\tt
-fParallel/data\_processing}.  The makefile there needs
-to be edited to indicate which of the tools to build.
-
-\begin{itemize}
-
-\item {\tt fextract3d}
-
-  Extracts a 1-d slice from a plotfile, through the center of the
-  domain in any coordinate direction.  Either all of the variables or
-  a single variable is output.
-
-\item {\tt fextrema}
-
-  Outputs the min/max of each of the quantities in the plotfile.
-
-\item {\tt ftime}
-
-  Outputs the simulation time for one or more plotfiles.
-
-\end{itemize}
-
-\subsection{VisIt}

### 2010-03-25 docs/architecture/architecture.tex
describe some test problems

@@
 
-\item {\tt test\_basestate}
+\item {\tt test\_basestate} \\[-3mm]
+
+  This test initializes the base state to contain a hydrostatic
+  model and then evolves the state with heating to watch the 
+  hydrostatic adjustment of the atmosphere.  In particular,
+  the base state velocity, $w_0$, is computed in response to 
+  the heating and this is used to advect the base state density
+  and compute the new pressure, $p_0$.  An early version of 
+  this routine was used for the plane-parallel expansion test
+  in \cite{lowMach2}.  This version of the test was also shown
+  for a spherical, self-gravitating star in \cite{multilevel}.
+
 \item {\tt test\_diffusion}
+
+  This test initializes a Gaussian temperature profile and calls
+  the thermal diffusion routines in MAESTRO to evolve the state 
+  considering only diffusion.  The driver estimates a timestep
+  based on the explicit thermal diffusion timescale and loops
+  over calls to the thermal diffusion solver.  A Gaussian remains
+  Gaussian when diffusing, so an explicit error can be computed
+  by comparing to the analytic solution.  This test is 
+  described in \cite{xrb}.
+
 \end{itemize}
@@
 \item {\tt ibx} is a \boxtype\ that describes the index range of the valid data
-  for the \fab.  This differes from {\tt bx} when the box is nodal in one more more 
+  for the \fab.  This differes from {\tt bx} when the box is nodal in one or more 
   directions.
@@
 
-The MAC velocities are allocated at each level of refinement,{\tt n},
+The MAC velocities are allocated at each level of refinement, {\tt n},
 by making a \multifab\ array where each of the {\tt dm} components is
@@
 \noindent Here, the {\tt bl\_types} and {\tt multifab\_module} modules
-bring in the basic boxlib data types Specifically, here, {\tt
+bring in the basic boxlib data types. Specifically, here, {\tt
 bl\_types} defines {\tt dp\_t} which is the {\tt kind} used for

### 2010-03-25 docs/architecture/architecture.tex
more on running

@@
 \begin{itemize}
-\item {\tt flame}
-\item {\tt planar\_heat}
-\item {\tt spherical\_heat}
-\item {\tt test2}
-\item {\tt test\_convect}
-\item {\tt test\_spherical}
+\item {\tt flame} \\[-3mm]
+
+The {\tt flame} problem models a planar laminar flame.  Initially cool
+fuel and hot ash are put in pressure equilibrium.  Thermal diffusion
+transfers heat to the fuel, driving a burning front.
+
+\item {\tt planar\_heat} \\[-3mm]
+
+\item {\tt spherical\_heat} \\[-3mm]
+
+\item {\tt test2} \\[-3mm]
+
+{\tt test2} places 3 hots spots in a plane-parallel atmosphere.
+Burning makes these buoyant bubbles which role up.  This problem was
+used in \cite{lowMach3} to compare with compressible solvers.
+
+\item {\tt test\_convect} \\[-3mm]
+
+\item {\tt test\_spherical} \\[-3mm]
 \end{itemize}
+
+To build the MAESTRO executable, simply type `{\tt make}' in the
+problem directory.  The executable will have a name like
+{\tt main.Linux.Intel.exe}, where the specific parts of 
+the name depend on the compilers and OS used.  Each problem
+will have one or more inputs files (for example, {\tt test2/inputs\_2d}
+which override the default values for one or more runtime parameters.
+The code is run simply as:
+\begin{verbatim}
+  main.Linux.Intel.exe inputs_2d
+\end{verbatim}
 

### 2010-03-26 docs/architecture/architecture.tex
some more reorgaization and more on the 'standard' test problems

@@
 
+\subsection{`Standard' Test Problems}
+
 Different problems in MAESTRO are contained in individual
@@
 
-\item {\tt planar\_heat} \\[-3mm]
+\item {\tt spherical\_heat} \\[-3mm]
 
-\item {\tt spherical\_heat} \\[-3mm]
+Spherical heat maps a spherical star (a Chandrasekhar-mass white dwarf)
+onto the Cartesian domain and uses an external heat source to cause
+the star to expand.  This is the 3-d analogue of the {\tt test\_basestate}
+unit test.  This problem was discussed in \cite{multilevel}.
 
@@
 
+{\tt test\_convect} drives convection through a plane-parallel
+atmosphere using an externally-specified heat source.  This problem
+was used to compare with compressible solvers in \cite{lowMach3}
+and to test the multilevel algorithm in \cite{multilevel}.
+
 \item {\tt test\_spherical} \\[-3mm]
+
+{\tt test\_spherical} sets up an isentropically stratified star
+and stirs it up with a random velocity field.  The low Mach number
+constraint is replaced with the anelastic constraint (through
+the {\tt beta\_type} runtime parameter).  Analytically, under
+these conditions, the density of the star should not change.
+This test problem was discussed in \cite{wdconvect}.
+
 \end{itemize}
@@
   main.Linux.Intel.exe inputs_2d
+\end{verbatim}
+As the code runs, it will output both plotfiles and checkpoint files.
+By default, the plotfiles will be named {\tt plt}{\em nnnnn}, where
+the number {\em nnnnn} is the timestep number when the file was
+outputted.  Similarly, the checkpoint files are named {\tt chk}{\em
+  nnnnn}.
+
+To restart a calculation from a checkpoint file, either edit the inputs
+file, setting
+\begin{verbatim}
+  restart = n
+\end{verbatim}
+where {\tt n} is the step number of the checkpoint file to restart
+from.  Equivalently, you can override the value in the inputs file
+by adding {\tt --restart n} to the commandline, as 
+\begin{verbatim}
+  main.Linux.Intel.exe inputs_2d --restart n
 \end{verbatim}
@@
 listed in a file {\tt GPackage.mak} in the problem directory.
+
+\subsection{Defining Runtime Parameters}
 
@@
 
+\section{MAESTRO Helper Modules}
+
+\subsection{{\tt eos\_module}}
+
+\subsection{{\tt fill\_3d\_module}}
+
+\subsection{{\tt geometry}}
+
+\subsection{{\tt network}}
+
+\subsection{{\tt probin\_module}}
+
+\subsection{{\tt variables}}
+
+The {\tt variables} module provides
+
 \section{Example: Accessing State Data}
@@
 \section{Multigrid}
-
-\section{MAESTRO Helper Modules}
-
-\subsection{{\tt eos\_module}}
-
-\subsection{{\tt fill\_3d\_module}}
-
-\subsection{{\tt geometry}}
-
-\subsection{{\tt network}}
-
-\subsection{{\tt probin\_module}}
-
-\subsection{{\tt variables}}
-
-The {\tt variables} module provides

### 2010-03-26 docs/architecture/architecture.tex
start documenting the helper modules

@@
 
-To find a particular species, a query can be made through the {\tt network}
-module, such as:
-\begin{verbatim}
-  ic12 = network_species_index("carbon-12")
-\end{verbatim}
-and then the \fab\ can be indexed using {\tt spec\_comp-1+ic12} to
-get ``carbon-12''.
-
 \subsection{`{\tt u}' \multifab s (fluid velocity)}
@@
 
+The {\tt network} module defines the number species advected by the
+code ({\tt nspec}), their ordering, and gives their basic properties
+(like atomic number, $Z$, and atomic mass, $A$).  All MAESTRO problems
+require a {\tt network} module, even if there are no reactions
+modeled.  Many different reaction modules (containing different sets
+of isotopes) exist in {\tt extern/networks}.  The particular network
+used by a problem is defined in the problem's {\tt GNUmakefile}.
+
 \subsection{{\tt probin\_module}}
+
+{\tt probin\_module} provides access to the runtime parameters.
 
@@
 
-The {\tt variables} module provides
+The {\tt variables} module provides integer keys to index the state
+multifabs and other arrays dealing with the scalar quantities.  The
+most commonly used keys are are listed in table~\ref{tab:variable_keys}.
+\begin{table}[h!]
+\centering
+\begin{tabular}{ll}
+\hline
+{\tt rho\_comp}  & density \\
+{\tt rhoh\_comp} & density * specific enthalpy, $(\rho h)$ \\
+{\tt spec\_comp} & starting index of the species \\
+{\tt temp\_comp} & temperature \\
+\hline
+\end{tabular}
+\caption{\label{tab:variable_keys} {\tt variables} module common keys}
+\end{table}
+
+To find a particular species, a query can be made through the {\tt
+  network} module, such as:
+\begin{verbatim}
+  ic12 = network_species_index("carbon-12")
+\end{verbatim}
+and then the \fab\ can be indexed using {\tt spec\_comp-1+ic12} to
+get ``carbon-12''.
+The {\tt variables} module also provides keys for the plotfile
+variables and boundary condition types.
 

### 2010-03-26 docs/architecture/architecture.tex
more documentation...

@@
 
-\subsection{Plane-Parallel Geometry}
+\subsection{Preparing the Initial Model}
 
-\subsection{Full Star / Spherical}
+MAESTRO models subsonic flows that are in hydrostatic equilibrium.
+The solution in MAESTRO is broken up into a 1-d base state and the 2-
+or 3-d full state.  The job of the 1-d base state in the algorithm is
+to represent the hydrostatic structure.  The full, Cartesian state
+carries the departures from hydrostatic equilibrium.  The underlying
+formulation of the low Mach number equations assumes that the base
+state is in hydrostatic equilibrium.  At the start of a simulation,
+the initial model is read in and taken as the base state.  Therefore,
+any initial model needs to already be in hydrostatic equilibrium.
 
-\subsection{Small-scale Flame}
+The routines in {\tt MAESTRO/initial\_model/} take a 1-d initial
+model and map it onto a uniform grid, at the desired resolution, using
+the equation of state in MAESTRO, and using MAESTRO's discretization of
+hydrostatic equilibrium.
 
@@
 
-Individual state variables should be indexed using the {\tt variables}
-module, with {\tt rho\_comp} referring to the density, {\tt rhoh\_comp}
-referring to the enthalpy quantity $(\rho h)$, and {\tt spec\_comp}
-referring to the start of the species (partial densities).
+Individual state variables should be indexed using the integer keys provided 
+by the {\tt variables} module (see \S \ref{sec:variables_module}).  For example,
+the integer {\tt rho\_comp} will always refer to the density component of the state.
 
@@
 ``marker-and-cell'' method for free boundaries in incompressible
-flows, Harlow and Welch 1965).
+flows \cite{harlowwelch:1965}).
 
@@
 
+The {\tt eos\_module} provides the interface to the equation of 
+state to connect the state variables thermodynamically.  It 
+gets the information about the fluid species from the {\tt network}
+module (for example, the atomic number, $Z$, and atomic weight, $A$,
+of the nuclei).
+
 \subsection{{\tt fill\_3d\_module}}
@@
 
+To find the location of a particular species (for instance, ``carbon-12'')
+in the allowed range of {\tt 1:nspec}, you do the following query:
+\begin{verbatim}
+  ic12 = network_species_index("carbon-12")
+\end{verbatim}
+If the resulting index is {\tt -1}, then the requested species was not
+found.
+
 \subsection{{\tt probin\_module}}
@@
 \subsection{{\tt variables}}
+
+\label{sec:variables_module}
 
@@
 {\tt rho\_comp}  & density \\
-{\tt rhoh\_comp} & density * specific enthalpy, $(\rho h)$ \\
-{\tt spec\_comp} & starting index of the species \\
+{\tt rhoh\_comp} & density $\times$ specific enthalpy, $(\rho h)$ \\
+{\tt spec\_comp} & first species partial density, $(\rho X_1)$ \\
 {\tt temp\_comp} & temperature \\
@@
 \end{tabular}
+
 \caption{\label{tab:variable_keys} {\tt variables} module common keys}
@@
 
+The species indices are contiguous in the state array, spanning {\tt spec\_comp:spec\_comp-1+nspec}.
 To find a particular species, a query can be made through the {\tt
@@
 variables and boundary condition types.
+
+\section{BoxLib Helper Modules}
+
+\subsection{{\tt bl\_types}}
+
+\subsection{{\tt bl\_constants}}
 

### 2010-03-26 docs/architecture/architecture.tex
talk about runtime parameters

@@
 these conditions, the density of the star should not change.
-This test problem was discussed in \cite{wdconvect}.
+This test problem was discussed in \cite{lowMach4}.
 
@@
 
+\label{sec:makefile}
+
 A basic {\tt GNUmakefile} begins with:
@@
 \subsection{Defining Runtime Parameters}
+
+The runtime parameters for the core MAESTRO algorithm are listed
+in {\tt MAESTRO/\_parameters}.  That file is parsed at compile-time
+by the {\tt MAESTRO/write\_probin.py} script
+(along with any other parameter files listed in the {\tt PROBIN\_PARAMETERS}
+variable in the {GNUmakefile}, see \S \ref{sec:makefile}).  The script
+outputs the {\tt probin.f90} source file.  Each line
+in the {\tt \_parameters} file has the form:
+
+{\em parameter} \hskip 10em  {\em data-type} \hskip 10em  {\em value}
+
+\noindent where {\em parameter} is the name of the runtime parameter,
+{\em data-type} is one of \{{\tt character}, {\tt real}, {\tt
+  integer}, {\tt logical}\}, and the {\em value} specifies the default
+value for the runtime parameter.  Comments are indicated by a `{\tt
+  \#}' character and a used to produce documentation about the
+available runtime parameters.
+
+At runtime, the default values for the parameters can be overridden
+either through the inputs file (by adding a line of the form: {\tt
+  parameter = value}) or through a command-line argument (taking the
+form: {\tt --parameter value}).  The {\tt probin\_module} makes the
+values of the runtime parameters available to the various functions
+in the code (see \S~\ref{sec:probin}).
 
@@
 
+Presently there are 2 equations of state that work with MAESTRO:
+\begin{itemize}
+\item {\tt extern/EOS/helmeos/} represents a general stellar equation 
+      of state, consisting of nuclei (as an ideal gas), radiation,
+      and electrons (with arbitrary degeneracy and degree of relativity).
+      This equation of state is that described in \cite{timmes_eos}.
+
+\item {\tt extern/EOS/gamma\_law\_general} assumes an ideal gas with a mixed 
+     composition and a constant ratio of specific heats, $\gamma$:
+      \begin{eqnarray}
+      e &=& \frac{p}{\rho (\gamma - 1) } \\
+      p &=& \frac{\rho k_B T}{\mu m_p} 
+      \end{eqnarray}
+     where $k_B$ is Boltzmann's constant and $m_p$ is the mass of the
+     proton.
+     The mean molecular weight, $\mu$, is computed either assuming 
+     electrically neutral atoms:
+     \begin{equation}
+     \mu = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}
+     \end{equation}
+     or complete ionization:
+     \begin{equation}
+     \mu = \left ( \sum_k \frac{X_k (Z_k + 1)}{A_k} \right )^{-1}
+     \end{equation}
+     The default is neutral, and presently one must change a parameter
+     in the source code to switch between these limits.
+\end{itemize}
+
 \subsection{{\tt fill\_3d\_module}}
@@
 \subsection{{\tt probin\_module}}
+
+\label{sec:probin}
 
@@
 \subsection{{\tt bl\_constants}}
+
+\subsection{{\tt parallel}}
 

### 2010-03-27 docs/architecture/architecture.tex
more reorganization to make it more into a User's Manual

@@
 
-\section{Running A Problem}
+\section{Getting Started}
+
+In this section we describe some of the standard problems that come
+with MAESTRO, how to run the code, and how to look at the output.
 
@@
 
-Spherical heat maps a spherical star (a Chandrasekhar-mass white dwarf)
-onto the Cartesian domain and uses an external heat source to cause
-the star to expand.  This is the 3-d analogue of the {\tt test\_basestate}
-unit test.  This problem was discussed in \cite{multilevel}.
+Spherical heat maps a spherical star (a Chandrasekhar-mass white
+dwarf) onto the Cartesian domain and uses an external heat source to
+cause the star to expand.  This is the 3-d analogue of the {\tt
+  test\_basestate} unit test.  This problem was discussed in
+\cite{multilevel}.
 
@@
 \end{itemize}
+
+\subsection{Compiling and Running}
 
@@
 \end{verbatim}
+
+\subsection{Common Runtime Parameters}
+
+\subsection{Working with the Output}
+
+Visualization and analysis are done with on the plotfiles.  The
+frequency of plotfile outputting can be changed using the runtime
+parameters {\tt plot\_int}, which specifies the number of steps
+between output, and {\tt plot\_deltat} which specifies the simulation
+time between output.  Setting {\tt plot\_int} to {\tt -1} disables
+step-based outputting; setting {\tt plot\_deltat} to {\tt 0.0}
+disables time-based outputting.
+
+\subsubsection{Amrvis}
+
+\subsubsection{{\tt data\_processing} scripts}
+
+Several useful analysis scripts can be found in {\tt
+fParallel/data\_processing}.  The makefile there needs
+to be edited to indicate which of the tools to build.
+
+\begin{itemize}
+
+\item {\tt fextract3d}
+
+  Extracts a 1-d slice from a plotfile, through the center of the
+  domain in any coordinate direction.  Either all of the variables or
+  a single variable is output.
+
+\item {\tt fextrema}
+
+  Outputs the min/max of each of the quantities in the plotfile.
+
+\item {\tt ftime}
+
+  Outputs the simulation time for one or more plotfiles.
+
+\end{itemize}
+
+\subsubsection{VisIt}
 
@@
 \end{verbatim}
-Here, {\tt NDEBUG} is true is we are building an optimized executable.  Otherwise,
-the debug version is built---this typically uses less optimization and adds various
-runtime checks through compiler flags.  {\tt MPI} and {\tt OMP} are set to true if
-we want to use either MPI or OpenMP for parallelization.
+Here, {\tt NDEBUG} is true is we are building an optimized executable.
+Otherwise, the debug version is built---this typically uses less
+optimization and adds various runtime checks through compiler flags.
+{\tt MPI} and {\tt OMP} are set to true if we want to use either MPI
+or OpenMP for parallelization.
 
-The MAESTRO build system knows what options to use for various compiler families.
-The {\tt COMP} flag specifies which compiler to us.  Allowed values include
-{\tt Intel}, {\tt gfortran}, {\tt PGI}, and {\tt PathScale}.  The specific details
-of these choices are defined in the {\tt MAESTRO/mk/} directory.
+The MAESTRO build system knows what options to use for various
+compiler families.  The {\tt COMP} flag specifies which compiler to
+us.  Allowed values include {\tt Intel}, {\tt gfortran}, {\tt PGI},
+and {\tt PathScale}.  The specific details of these choices are
+defined in the {\tt MAESTRO/mk/} directory.
 \begin{verbatim}
@@
 \end{verbatim}
-{\tt MKVERBOSE} set to true will echo the build commands to the terminal as the
-are executed.
+{\tt MKVERBOSE} set to true will echo the build commands to the
+terminal as the are executed.
 
@@
 {\tt VPATH}, so any source files placed there override those with the
-same name found elsewhere in the source tree.  The {\tt Fmincludes} variable
-lists all those directories that contain include files that are inserted into
-the Fortran source at compile time via the {\tt include} statement.  Presently,
-the only instance of this is with the Helmholtz general equation of state
-found in {\tt extern/EOS/helmeos/}.
+same name found elsewhere in the source tree.  The {\tt Fmincludes}
+variable lists all those directories that contain include files that
+are inserted into the Fortran source at compile time via the {\tt
+  include} statement.  Presently, the only instance of this is with
+the Helmholtz general equation of state found in {\tt
+  extern/EOS/helmeos/}.
 
@@
 parsed at compile time and the file {\tt probin.f90} is written and
-compiled.  This is a Fortran module that holds the values of the runtime
-parameters and makes them available to any routine.  The variable
-{\tt PROBIN\_PARAMETERS} should list all of the files that contain
-runtime parameters for the current problem.  You must always
-include the main {\tt ../\_parameters} file (i.e.\ {\tt MAESTRO/\_parameters}).
-Sometimes a problem defines their own runtime parameters, in which
-case, its parameter file should also appear in {\tt PROBIN\_PARAMETERS}.
+compiled.  This is a Fortran module that holds the values of the
+runtime parameters and makes them available to any routine.  The
+variable {\tt PROBIN\_PARAMETERS} should list all of the files that
+contain runtime parameters for the current problem.  You must always
+include the main {\tt ../\_parameters} file (i.e.\ {\tt
+  MAESTRO/\_parameters}).  Sometimes a problem defines their own
+runtime parameters, in which case, its parameter file should also
+appear in {\tt PROBIN\_PARAMETERS}.
 \begin{verbatim}
@@
 hydrostatic equilibrium.
-
-\section{Visualization / Analysis}
-
-\subsection{Amrvis}
-
-\subsection{{\tt data\_processing} scripts}
-
-Several useful analysis scripts can be found in {\tt
-fParallel/data\_processing}.  The makefile there needs
-to be edited to indicate which of the tools to build.
-
-\begin{itemize}
-
-\item {\tt fextract3d}
-
-  Extracts a 1-d slice from a plotfile, through the center of the
-  domain in any coordinate direction.  Either all of the variables or
-  a single variable is output.
-
-\item {\tt fextrema}
-
-  Outputs the min/max of each of the quantities in the plotfile.
-
-\item {\tt ftime}
-
-  Outputs the simulation time for one or more plotfiles.
-
-\end{itemize}
-
-\subsection{VisIt}
 
@@
      composition and a constant ratio of specific heats, $\gamma$:
-      \begin{eqnarray}
-      e &=& \frac{p}{\rho (\gamma - 1) } \\
-      p &=& \frac{\rho k_B T}{\mu m_p} 
-      \end{eqnarray}
+      \begin{equation}
+      p = \rho e (\gamma - 1) = \frac{\rho k_B T}{\mu m_p} 
+      \end{equation}
      where $k_B$ is Boltzmann's constant and $m_p$ is the mass of the

### 2010-03-28 docs/initial_models/initial_models.tex
test2 and test_convect are now generated by initial_models/test2/

@@
 \begin{eqnarray}
-{\tt test2} & \rightarrow & \mathrm{comes~from~the~ {\tt lmn\_bubble\_burn}~setup~in~ {\tt FLASH}} \nonumber \\
-{\tt test\_convect} & \rightarrow & \mathrm{comes~from~the~ {\tt lmn\_convect}~setup~in~ {\tt FLASH}} \nonumber \\
+{\tt test2} & \rightarrow & \mathrm{none-it~ is~ generated~ analytically} \nonumber \nonumber \\
+{\tt test\_convect} & \rightarrow & \mathrm{none-it~ is~ generated~ analytically} \nonumber \nonumber \\
 {\tt wdconvect} & \rightarrow & {\tt initial\_models/kepler/model\_6.e8.raw} \nonumber \\
@@
 \begin{eqnarray}
-{\tt test2} & \rightarrow & \mathrm{comes~from~the~ {\tt lmn\_bubble\_burn}~setup~in~ {\tt FLASH}} \nonumber \\
-{\tt test\_convect} & \rightarrow & \mathrm{comes~from~the~ {\tt lmn\_convect}~setup~in~ {\tt FLASH}} \nonumber \\
+{\tt test2} & \rightarrow & {\tt initial\_models/test2/init\_1d.f90} \nonumber \\
+{\tt test\_convect} & \rightarrow & {\tt initial\_models/test2/init\_1d.f90} \nonumber \\
 {\tt wdconvect} & \rightarrow & {\tt initial\_models/kepler/init\_1d.f90} \nonumber \\

### 2010-03-28 docs/flowchart/flowchart.tex
remove the table of symbols -- that'll be in the introduction

@@
 \end{itemize}
-
-%-----------------------------------------------------------------------------
-% Notation
-%-----------------------------------------------------------------------------
-
-\section{Notation}
-
-%%%%%%%%%%%%%%%%
-% symbol table
-%%%%%%%%%%%%%%%%
-
-\renewcommand{\arraystretch}{1.5}
-%
-\begin{center}
-\begin{longtable}{|l|p{4.0in}|l|}
-\caption[definition of symbols.]{definition of symbols.} \label{table:sym} \\
-%
-\hline \multicolumn{1}{|c|}{\textbf{symbol}} & 
-       \multicolumn{1}{ c|}{\textbf{meaning}} & 
-       \multicolumn{1}{ c|}{\textbf{units}} \\ \hline 
-\endfirsthead
-
-\multicolumn{3}{c}%
-{{\tablename\ \thetable{}---continued}} \\
-\hline \multicolumn{1}{|c|}{\textbf{symbol}} & 
-       \multicolumn{1}{ c|}{\textbf{meaning}} & 
-       \multicolumn{1}{ c|}{\textbf{units}} \\ \hline 
-\endhead
-
-\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
-\endfoot
-
-\hline 
-\endlastfoot
-
-$c_p$   & specific heat at constant pressure
-          ($c_p \equiv \left . \partial h / \partial T \right |_{p,X_k}$)
-        & erg~g$^{-1}$~K$^{-1}$ \\
-\hline
-$f$     & volume discrepancy factor ($0 \le f \le 1$) & -- \\
-\hline
-$g$     & gravitational acceleration                 & cm~s$^{-2}$ \\
-\hline
-$h$     & specific enthalpy                          & erg~g$^{-1}$ \\
-\hline
-$\Hext$ & external heating energy generation rate    & erg~g$^{-1}$~s$^{-1}$ \\
-\hline
-$\Hnuc$ & nuclear energy generation rate             & erg~g$^{-1}$~s$^{-1}$ \\
-\hline
-$h_p$   & $h_p \equiv \left . \partial h / \partial p \right |_{T,X_k}$ & cm$^{3}$~g$^{-1}$ \\
-\hline
-$\kth$  & thermal conductivity                       & erg~cm$^{-1}$~s$^{-1}$~K$^{-1}$ \\
-\hline
-$p_0$   & base state pressure                        & erg~cm$^{-3}$ \\
-\hline
-$p_T$   & $p_T \equiv \left . \partial p / \partial T \right |_{\rho,X_k}$ & erg~cm$^{-3}$~K$^{-1}$ \\
-\hline
-$p_{X_k}$ & $p_{X_k} \equiv \left . \partial p / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~cm$^{-3}$ \\
-\hline
-$p_\rho$ & $p_\rho \equiv \left . \partial p / \partial \rho \right |_{T,X_k}$ & erg~g$^{-1}$ \\
-\hline
-$q_k$   & specific nuclear binding energy            & erg~g$^{-1}$  \\
-\hline
-$r$     & radial coordinate (direction of gravity)   & cm \\
-\hline
-$s$     & specific entropy                           & erg~g$^{-1}$~K$^{-1}$ \\
-\hline
-$S$     & source term to the divergence constraint   & s$^{-1}$ \\
-\hline
-$t$     & time                                       & s \\
-\hline
-$T$     & temperature                                & K \\
-\hline
-$\Ub$     & total velocity ($\Ub = \Ubt + w_0 \eb_r$) & cm~s$^{-1}$ \\
-\hline
-$\Ubt$   & local velocity                             & cm~s$^{-1}$ \\
-\hline
-$\uadv$ & advective velocity (edge-centered)         & cm~s$^{-1}$ \\
-\hline
-$w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
-\hline
-$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & -- \\
-\hline
-$\beta_0$ & coefficient to velocity
-            in velocity constraint equation  & g~cm$^{-3}$ \\
-\hline
-$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & -- \\
-\hline
-$\etarho$ & $\etarho \equiv \overline{(\rho' \Ub \cdot \eb_r)}$ & g~cm$^{-2}$~s$^{-1}$ \\
-\hline
-$\xi_k$ & $\xi_k \equiv \left . \partial h / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~g$^{-1}$ \\
-\hline 
-$\pi$   & dynamic pressure & erg~cm$^{-3}$ \\
-\hline
-$\pizero$ & base state dynamic pressure & erg~cm$^{-3}$ \\
-\hline
-$\rho$  & mass density  & g~cm$^{-3}$ \\
-\hline
-$\rho_0$  & base state mass density  & g~cm$^{-3}$ \\
-\hline
-$\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
-\hline
-$(\rho h)_0$ & base state enthalpy density & erg~cm$^{-3}$  \\
-\hline
-$(\rho h)'$ & perturbational enthalpy density 
-              $ \left [(\rho h)' = \rho h - (\rho h)_0 \right ]$ & erg~cm$^{-3}$  \\
-\hline
-$\sigma$ & $\sigma \equiv p_T/(\rho c_p p_\rho)$ & erg$^{-1}$~g \\
-\hline
-$\psi$  & $\psi \equiv D_0 p_0/Dt = \partial p_0/\partial t + w_0\partial p_0/\partial r$ & erg~cm$^{-3}$~s$^{-1}$ \\
-\hline
-$\omegadot_k$ & creation rate for species $k$ ($\omegadot_k \equiv DX_k/Dt$) & s$^{-1}$ \\
-\end{longtable}
-\end{center}
-
-\renewcommand{\arraystretch}{1.0}
-There are some great charts in the appendix of the multilevel paper demonstrating what 
-the superscript notation represents.
 

### 2010-03-28 docs/introduction/introduction.tex
move the table of symbols to the new introduction

%-----------------------------------------------------------------------------
% Notation
%-----------------------------------------------------------------------------

\section{Notation}

%%%%%%%%%%%%%%%%
% symbol table
%%%%%%%%%%%%%%%%

\renewcommand{\arraystretch}{1.5}
%
\begin{center}
\begin{longtable}{|l|p{4.0in}|l|}
\caption[definition of symbols.]{definition of symbols.} \label{table:sym} \\
%
\hline \multicolumn{1}{|c|}{\textbf{symbol}} & 
       \multicolumn{1}{ c|}{\textbf{meaning}} & 
       \multicolumn{1}{ c|}{\textbf{units}} \\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\tablename\ \thetable{}---continued}} \\
\hline \multicolumn{1}{|c|}{\textbf{symbol}} & 
       \multicolumn{1}{ c|}{\textbf{meaning}} & 
       \multicolumn{1}{ c|}{\textbf{units}} \\ \hline 
\endhead

\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
\endfoot

\hline 
\endlastfoot

$c_p$   & specific heat at constant pressure
          ($c_p \equiv \left . \partial h / \partial T \right |_{p,X_k}$)
        & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$f$     & volume discrepancy factor ($0 \le f \le 1$) & -- \\
\hline
$g$     & gravitational acceleration                 & cm~s$^{-2}$ \\
\hline
$h$     & specific enthalpy                          & erg~g$^{-1}$ \\
\hline
$\Hext$ & external heating energy generation rate    & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$\Hnuc$ & nuclear energy generation rate             & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$h_p$   & $h_p \equiv \left . \partial h / \partial p \right |_{T,X_k}$ & cm$^{3}$~g$^{-1}$ \\
\hline
$\kth$  & thermal conductivity                       & erg~cm$^{-1}$~s$^{-1}$~K$^{-1}$ \\
\hline
$p_0$   & base state pressure                        & erg~cm$^{-3}$ \\
\hline
$p_T$   & $p_T \equiv \left . \partial p / \partial T \right |_{\rho,X_k}$ & erg~cm$^{-3}$~K$^{-1}$ \\
\hline
$p_{X_k}$ & $p_{X_k} \equiv \left . \partial p / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~cm$^{-3}$ \\
\hline
$p_\rho$ & $p_\rho \equiv \left . \partial p / \partial \rho \right |_{T,X_k}$ & erg~g$^{-1}$ \\
\hline
$q_k$   & specific nuclear binding energy            & erg~g$^{-1}$  \\
\hline
$r$     & radial coordinate (direction of gravity)   & cm \\
\hline
$s$     & specific entropy                           & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$S$     & source term to the divergence constraint   & s$^{-1}$ \\
\hline
$t$     & time                                       & s \\
\hline
$T$     & temperature                                & K \\
\hline
$\Ub$     & total velocity ($\Ub = \Ubt + w_0 \eb_r$) & cm~s$^{-1}$ \\
\hline
$\Ubt$   & local velocity                             & cm~s$^{-1}$ \\
\hline
$\uadv$ & advective velocity (edge-centered)         & cm~s$^{-1}$ \\
\hline
$w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
\hline
$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & -- \\
\hline
$\beta_0$ & coefficient to velocity
            in velocity constraint equation  & g~cm$^{-3}$ \\
\hline
$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & -- \\
\hline
$\etarho$ & $\etarho \equiv \overline{(\rho' \Ub \cdot \eb_r)}$ & g~cm$^{-2}$~s$^{-1}$ \\
\hline
$\xi_k$ & $\xi_k \equiv \left . \partial h / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~g$^{-1}$ \\
\hline 
$\pi$   & dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\pizero$ & base state dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\rho$  & mass density  & g~cm$^{-3}$ \\
\hline
$\rho_0$  & base state mass density  & g~cm$^{-3}$ \\
\hline
$\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
\hline
$(\rho h)_0$ & base state enthalpy density & erg~cm$^{-3}$  \\
\hline
$(\rho h)'$ & perturbational enthalpy density 
              $ \left [(\rho h)' = \rho h - (\rho h)_0 \right ]$ & erg~cm$^{-3}$  \\
\hline
$\sigma$ & $\sigma \equiv p_T/(\rho c_p p_\rho)$ & erg$^{-1}$~g \\
\hline
$\psi$  & $\psi \equiv D_0 p_0/Dt = \partial p_0/\partial t + w_0\partial p_0/\partial r$ & erg~cm$^{-3}$~s$^{-1}$ \\
\hline
$\omegadot_k$ & creation rate for species $k$ ($\omegadot_k \equiv DX_k/Dt$) & s$^{-1}$ \\
\end{longtable}
\end{center}

\renewcommand{\arraystretch}{1.0}
There are some great charts in the appendix of the multilevel paper demonstrating what 
the superscript notation represents.

### 2010-03-28 docs/flowchart/flowchart.tex
add the graphical flowcharts from the multilevel paper

@@
 %-----------------------------------------------------------------------------
+% graphical flowcharts
+%-----------------------------------------------------------------------------
+\begin{figure}[tb]
+\centering
+\includegraphics[scale=0.6]{flowchart}
+\caption{\label{Fig:flowchart} A flowchart of the algorithm.  The
+  thermodynamic state variables, base state variables, and local velocity are
+  indicated in each step.  Red text indicates that quantity was
+  updated during that step.  The predictor-corrector steps are
+  outlined by the dotted box.  The blue text indicates state
+  variables that are the same in {\bf Step 6} as they are in
+  {\bf Step 2}, i.e., they are unchanged by the predictor steps.}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                               
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                               
+\begin{figure}[tb]                                                              
+\centering
+\includegraphics[scale=0.6]{flowchart_4_8}
+\caption{\label{Fig:flowchart48}
+A flowchart for {\bf Steps 4} and {\bf 8}.  The thermodynamic state variables
+and base state variables are indicated in each step.  Red text indicates that q\
+uantity
+was updated during that step.  Note, for {\bf Step 4}, the updated
+quantities should also have a $\star$ superscript, e.g., {\bf Step 8I} defines
+$T^{(2)}$ while {\bf Step 4I} defines $T^{(2),\star}$ .}
+\end{figure}
+
+%-----------------------------------------------------------------------------
 % Initialization

### 2010-03-28 docs/architecture/architecture.tex
more additions

@@
 fuel and hot ash are put in pressure equilibrium.  Thermal diffusion
-transfers heat to the fuel, driving a burning front.
+transfers heat to the fuel, driving a burning front.  This problem
+does not use the traditional MAESTRO elliptic constraint, but
+rather enforces $\nabla \cdot U = S$ as discussed in \cite{SNe}.
 
@@
 
+\subsection{Defining the Grid and Boundary Conditions}
+
 \subsection{Common Runtime Parameters}
+
+\begin{itemize}
+
+\item {\tt model\_file }
+\item {\tt spherical\_in }
+\item {\tt ppm\_type }
+\item {\tt max\_levs }
+\item {\tt n\_cellx }, {\tt n\_celly }, {\tt n\_cellz }
+\item {\tt max\_grid\_size }
+\item {\tt prob\_lo\_x }, {\tt prob\_lo\_y }, {\tt prob\_lo\_z }
+\item {\tt prob\_hi\_x }, {\tt prob\_hi\_y }, {\tt prob\_hi\_z }
+\item {\tt anelastic\_cutoff }
+\item {\tt base\_cutoff\_density }
+\item {\tt do\_sponge }
+\item {\tt grav\_const }
+\item {\tt max\_step }
+\item {\tt init\_iter }
+\item {\tt stop\_time }
+\item {\tt plot\_base }
+\item {\tt plot\_deltat }
+\item {\tt chk\_int }
+\item {\tt cflfac }
+\item {\tt init\_shrink }
+\item {\tt bcx\_lo }, {\tt bcy\_lo }, {\tt bcz\_lo }, 
+      {\tt bcx\_hi }, {\tt bcy\_hi }, {\tt bcz\_hi }
+\item {\tt pmask\_x }, {\tt pmask\_y }, {\tt pmask\_z }
+\item {\tt verbose }
+\item {\tt mg\_verbose }
+\item {\tt cg\_verbose }
+\item {\tt do\_initial\_projection }
+\item {\tt init\_divu\_iter }
+\item {\tt perturb\_model }
+
+\end{itemize}
 
@@
 in the {\tt \_parameters} file has the form: 
-
+\vskip 3mm
 {\em parameter} \hskip 10em  {\em data-type} \hskip 10em  {\em value} 
-
+\vskip 3mm
 \noindent where {\em parameter} is the name of the runtime parameter,
@@
 
+The base state is defined by $\rho_0$, $p_0$, and $w_0$.  There is no
+base state composition.  Other arrays are defined as needed, such as
+$h_0$, the base state enthalpy.
+
 \subsection{Other Quantities}
@@
 \section{MAESTRO Helper Modules}
+
+A number of MAESTRO modules appear frequently throughout the source.
+Below, we describe some of the more common functionality of the most
+popular modules.
+
+\subsection{\tt average}
 
@@
 \end{itemize}
+
+eos module variables
+
+eos input types
 
@@
 \section{BoxLib Helper Modules}
+
+There are a large number of modules in {\tt boxlib/} that provide
+the core functionality for managing grids.  Here we describe
+the most popular such modules.
 
@@
 
+need to describe the routines for adding two multifabs, subtracting,
+etc. and norm, etc.
+
 \section{Filling Ghostcells}

### 2010-04-01 docs/architecture/architecture.tex
more reorganization
lots of additions

--- 2010-03-27 docs/architecture/architecture.tex
@@
 
- boxlib is a library for describing meshes consisting of a union
- of boxes.  The boxlib modules define the basic datatypes used
+ BoxLib is a library for describing meshes consisting of a union
+ of boxes.  The BoxLib modules define the basic datatypes used
  in MAESTRO.
@@
  Simple Fortran-based analysis routines (e.g.\ extract a line from a
- multidimensional dataset) that operate on boxlib plotfiles.
+ multidimensional dataset) that operate on BoxLib datasets.
 
@@
  networks
+
+ Important subdirectories include:
+ \begin{itemize}
+ \item {\tt constants/}
+
+ A simple module to define physical constants
+
+ \item {\tt EOS/}
+
+ A collection of various EOSes for use with the code.  The two
+ important ones for MAESTRO are {\tt helmeos/} and {\tt
+ gamma\_law\_general/}.
+
+ \item {\tt networks/}
+
+ Various reaction networks for MAESTRO problems.  In addition to
+ providing a routine to evolve the nuclear species due to reactions,
+ the networks also define the species that are advected by the code.
+
+ \item {\tt VODE/}
+
+ An integration package for ODEs.  At the moment, this is used 
+ for integrating various reaction networks.
+
+ \end{itemize}
 
@@
    1D initial conditions in hydrostatic equilibrium for 
-   mapping onto the MAESTRO grid.
+   mapping onto the MAESTRO grid.  See \S~\ref{sec:initial_models}
+   for details.
 
@@
    MAESTRO routines (e.g.\ initial conditions via {\tt initdata.f90}.
+   See \S~\ref{sec:makefile} for details on the build system.
 
@@
 In this section we describe some of the standard problems that come
-with MAESTRO, how to run the code, and how to look at the output.
+with MAESTRO, how to run the code, some basic runtime parameters, and
+how to look at the output.
 
@@
 fuel and hot ash are put in pressure equilibrium.  Thermal diffusion
-transfers heat to the fuel, driving a burning front.
+transfers heat to the fuel, driving a burning front.  This problem
+does not use the traditional MAESTRO elliptic constraint, but
+rather enforces $\nabla \cdot U = S$ (through the runtime parameter
+{\tt do\_smallscale = T}) as discussed in \cite{SNe}.
 
@@
 
-Spherical heat maps a spherical star (a Chandrasekhar-mass white
+{\tt spherical\_heat} maps a spherical star (a Chandrasekhar-mass white
 dwarf) onto the Cartesian domain and uses an external heat source to
 cause the star to expand.  This is the 3-d analogue of the {\tt
-  test\_basestate} unit test.  This problem was discussed in
-\cite{multilevel}.
+  test\_basestate} unit test (see \S~\ref{sec:unit_tests}).  This
+problem was discussed in \cite{multilevel}.
 
@@
 the name depend on the compilers and OS used.  Each problem
-will have one or more inputs files (for example, {\tt test2/inputs\_2d}
+will have one or more inputs files (for example, {\tt test2/inputs\_2d})
 which override the default values for one or more runtime parameters.
@@
 \end{verbatim}
-As the code runs, it will output both plotfiles and checkpoint files.
+As the code runs, it will output both plotfiles and checkpoints.
 By default, the plotfiles will be named {\tt plt}{\em nnnnn}, where
 the number {\em nnnnn} is the timestep number when the file was
-outputted.  Similarly, the checkpoint files are named {\tt chk}{\em
-  nnnnn}.
+outputted.  Similarly, the checkpoints are named {\tt chk}{\em
+  nnnnn}.  BoxLib plotfiles and checkpoints are actually directories,
+with the data stored in subdirectories grouped by refinement level.
+Details of the simulation (number of processors used, output date,
+output directory, runtime parameter values, ...) are stored in
+the {\tt job\_info} file in each plotfile directory.
 
-To restart a calculation from a checkpoint file, either edit the inputs
-file, setting
-\begin{verbatim}
-  restart = n
-\end{verbatim}
-where {\tt n} is the step number of the checkpoint file to restart
-from.  Equivalently, you can override the value in the inputs file
-by adding {\tt --restart n} to the commandline, as 
-\begin{verbatim}
-  main.Linux.Intel.exe inputs_2d --restart n
-\end{verbatim}
+A large number of runtime parameters affect the code's behavior.  Any
+runtime parameter can be set either in the inputs file, as {\tt parameter
+  = value} or on the command line by adding {\tt --parameter value} after
+listing the inputs file on the command line.  A full list of runtime
+parameters is detailed in \S~\ref{sec:runtime_parameters}.  Below we
+outline the most common runtime parameters.
 
-\subsection{Common Runtime Parameters}
+\subsection{Controlling Timestepping and Output}
+
+Parameters that set the maximum time for the simulation to run
+include:
+\begin{itemize}
+\item {\tt stop\_time} is the maximum simulation time, in seconds,
+      to evolve the system for.
+
+\item {\tt max\_step} is the maximum number of steps to take.
+\end{itemize}
+
+\noindent Parameters affecting the size of the timestep include:
+\begin{itemize}
+\item {\tt cflfac} is a multiplicative factor ({\tt $\le 1$}) 
+      applied to the advective CFL timestep
+
+\item {\tt init\_shrink} is the factor ({\tt $\le 1$}) by which to reduce 
+      the initial timestep from the estimated first timestep.
+\end{itemize}
+
+\noindent Parameters affecting output and restart include:
+\begin{itemize}
+
+\item {\tt restart} tells MAESTRO to restart from a checkpoint.  The
+      value of this parameter should be the file number to restart from.
+
+\item {\tt plot\_int} is the number of steps to take between
+  outputting a plotfile
+
+\item {\tt plot\_deltat} is the simulation time to evolve between
+  outputting a plotfile.  Note: to output only based on simulation
+  time, set {\tt plot\_int = -1}.
+
+\item {\tt check\_int} is the number of steps to take between
+  outputting a checkpoint.
+
+\end{itemize}
+
+\subsection{Defining the Grid and Boundary Conditions}
+
+Parameters that determine the spatial extent of the grid, 
+the types of boundaries, and the number of computational cells include:
+\begin{itemize}
+
+\item {\tt max\_levs } is the maximum number of grid levels in the AMR
+  hierarchy to us.  {\tt max\_levs = 1} indicates running with only a
+  single level spanning the whole domain.
+
+\item {\tt n\_cellx }, {\tt n\_celly }, {\tt n\_cellz } the size of
+  base level in terms of number of cells, in the $x$, $y$, and $z$
+  coordinate directions.
+
+\item {\tt max\_grid\_size } the maximum extend of a grid, in any
+  coordinate direction, as measured in terms of number of cells.
+
+\item {\tt prob\_lo\_x }, {\tt prob\_lo\_y }, {\tt prob\_lo\_z } is
+  the physical coordinate of the lower extent of the domain boundary
+  in the $x$, $y$, and $z$ coordinate directions.
+
+\item {\tt prob\_hi\_x }, {\tt prob\_hi\_y }, {\tt prob\_hi\_z } is
+  the physical coordinate of the upper extent of the domain boundary
+  in the $x$, $y$, and $z$ coordinate directions.
+
+\item {\tt bcx\_lo }, {\tt bcy\_lo }, {\tt bcz\_lo }, 
+      {\tt bcx\_hi }, {\tt bcy\_hi }, {\tt bcz\_hi } are the boundary
+   condition types at the lower (`{\tt lo}') and upper (`{\tt hi}')
+   domain boundaries in the $x$, $y$, and $z$ coordinate directions.
+   The different types are set via integer flags noted below:
+
+   \begin{center}
+   \begin{tabular}{ll}
+   \hline
+   BC type    & integer flag \\
+   \hline
+   periodic             & -1 \\
+   inlet (user-defined) & 11 \\
+   outlet               & 12 \\
+   symmetry             & 13 \\
+   slip wall            & 14 \\
+   no-slip wall         & 15 \\
+   \hline
+   \end{tabular}
+   \end{center}
+
+   Note: for periodic boundaries, one must also set the corresponding
+   {\tt pmask\_}\{{\em x,y,z}\} parameter.
+
+\item {\tt pmask\_x }, {\tt pmask\_y }, {\tt pmask\_z } is set to 
+   true ({\tt T}) to indicate that the boundary is periodic in the
+   $x$, $y$, or $z$ coordinate directions.
+
+\end{itemize}
+
+Note that grid cells must be square, i.e. $\Delta x = \Delta y = \Delta z$
+where $\Delta x$ on the base grid is computed as $({\tt prob\_hi\_x}
+- {\tt prob\_lo\_x})/{\tt n\_cellx}$.
+
+\subsection{Refinement Criteria}
 
@@
 
-Visualization and analysis are done with on the plotfiles.  The
-frequency of plotfile outputting can be changed using the runtime
-parameters {\tt plot\_int}, which specifies the number of steps
-between output, and {\tt plot\_deltat} which specifies the simulation
-time between output.  Setting {\tt plot\_int} to {\tt -1} disables
-step-based outputting; setting {\tt plot\_deltat} to {\tt 0.0}
-disables time-based outputting.
+Visualization and analysis are done with on the plotfiles.  A
+number of in-house and externally developed tools can work 
+with BoxLib-formatted plotfiles.
 
-\subsubsection{Amrvis}
+\subsubsection{\tt Amrvis}
+
+{\tt Amrvis} is a visualization tool developed at LBL for 2- and 3-d
+datasets which can plot slices through 3-d datasets as well as
+volume-renderings.  It is distributed separately from the MAESTRO
+distribution.
 
@@
 
-Several useful analysis scripts can be found in {\tt
-fParallel/data\_processing}.  The makefile there needs
-to be edited to indicate which of the tools to build.
-
-\begin{itemize}
-
-\item {\tt fextract3d}
-
-  Extracts a 1-d slice from a plotfile, through the center of the
-  domain in any coordinate direction.  Either all of the variables or
-  a single variable is output.
-
-\item {\tt fextrema}
-
-  Outputs the min/max of each of the quantities in the plotfile.
-
-\item {\tt ftime}
-
-  Outputs the simulation time for one or more plotfiles.
-
-\end{itemize}
+Several useful analysis scripts (written in Fortran 90) can be found
+in {\tt fParallel/data\_processing}.  The {\tt GNUmakefile} there
+needs to be edited to indicate which of the tools to build.  These
+routines are described in \S~\ref{sec:analysis}.
 
@@
 
+VisIt is a powerful, DOE-supported visualization tool for 2- and 3-d
+datasets.  It can do contouring, volume rendering, streamlines, ...\
+directly from BoxLib plotfiles (to open a plotfile in VisIt, point to
+the {\tt Header} file inside the plotfile directory.  Details on
+visit can be found at {\tt https://wci.llnl.gov/codes/visit/home.html}\,.
+
 \subsection{Unit Tests}
+
+\label{sec:unit_tests}
 
@@
 
+Different MAESTRO problems are defined in subdirectories under
+{\tt MAESTRO/}.  To add a problem, start by creating a new
+sub-directory---this is where you will compile your problem and
+store all the problem-specific files.
+
 The minimum requirement to
@@
 by searching the directories listed in the {\tt GNUmakefile}.
+Customised versions of any source files placed in the problem-directory
+override those with the same name found elsewhere.  Any unique
+source files (and not simply a custom version of a file found
+elsewhere) needs to be listed in a file call {\tt GPackage.mak}
+in the problem-directory.
 
@@
   MAESTRO/\_parameters}).  Sometimes a problem defines their own
-runtime parameters, in which case, its parameter file should also
-appear in {\tt PROBIN\_PARAMETERS}.
+runtime parameters (see \S~\ref{sec:def_runtime_param}), in which
+case, its parameter file should also appear in {\tt
+  PROBIN\_PARAMETERS}.
 \begin{verbatim}
@@
 
+\label{sec:def_runtime_param}
+
 The runtime parameters for the core MAESTRO algorithm are listed
@@
 
+Problem-specify runtime parameters should be defined in the
+problem-directory in a file called {\tt
+  \_parameters}\{{\em.problem-name}\}.  This file should be added
+to the {\tt PROBIN\_PARAMETERS} variable in the {\tt GNUmakefile}.
+
 \subsection{Preparing the Initial Model}
+
+\label{sec:initial_models}
 
@@
 the equation of state in MAESTRO, and using MAESTRO's discretization of
-hydrostatic equilibrium.
+hydrostatic equilibrium:
+\begin{equation}
+\frac{p_{i+1} - p_i}{\Delta r} = \frac{1}{2} (\rho_i + \rho_{i+1})
+g_{i+1/2}
+\end{equation}
 
@@
 
-MAESTRO's gridding is handled by the boxlib library.  The underlying
-idea in boxlib is to allow for adaptive mesh refinement (AMR)---different
+MAESTRO's gridding is handled by the BoxLib library.  The underlying
+idea in BoxLib is to allow for adaptive mesh refinement (AMR)---different
 regions of the domain can have different spatial resolutions.  The
@@
 On a grid, the data can be stored at cell-centers, on a face/edge, or
-on the corners.  In boxlib, data that is on an edge is termed `nodal'
+on the corners.  In BoxLib, data that is on an edge is termed `nodal'
 in that direction (see Figure~\ref{fig:dataloc}).  Data that is on the
@@
 
-To simplify the description of the underlying AMR grid, boxlib
+To simplify the description of the underlying AMR grid, BoxLib
 provides a number of Fortran types.  We briefly summarize the major
@@
 At this resolution, the domain is 20$\times$18 zones.  Note that the
-indexing in boxlib starts with $0$.}
+indexing in BoxLib starts with $0$.}
 \end{figure}
@@
 \item {\tt hi = upb(mybox)} returns an array, {\tt hi(dm)}, with
-     the box lower bounds
+     the box upper bounds
 
@@
 multidimensional state is stored in \multifab s while the base state
-is simply stored in Fortran arrays.
+is simply stored in Fortran arrays.  Here we describe the
+major MAESTRO data-structures.
 
@@
      proton.
-     The mean molecular weight, $\mu$, is computed either assuming 
+     The mean molecular weight, $\mu$, is computed assuming 
      electrically neutral atoms:
@@
      \end{equation}
-     or complete ionization:
-     \begin{equation}
-     \mu = \left ( \sum_k \frac{X_k (Z_k + 1)}{A_k} \right )^{-1}
-     \end{equation}
-     The default is neutral, and presently one must change a parameter
-     in the source code to switch between these limits.
+     An option in the source code itself exists for treating the
+     species as fully-ionized, but there is no runtime-parameter to
+     make this switch.
 \end{itemize}
+
+eos module variables
+
+eos input types
 
@@
 {\tt probin\_module} provides access to the runtime parameters.
+The runtime parameters appear simply as module variables.  To get the 
+value of a parameter, one simply needs to `{\tt use probin\_module}'.
+The preferred method is to add the `{\tt only}' clause to the
+{\tt use} statement and explicitly list only those parameters that
+are used in the routine.  Defining new runtime parameters is
+described in \S~\ref{sec:def_runtime_param}.
 
@@
 multifabs and other arrays dealing with the scalar quantities.  The
-most commonly used keys are are listed in table~\ref{tab:variable_keys}.
-\begin{table}[h!]
-\centering
+most commonly used keys are are:
+
+\begin{center}
 \begin{tabular}{ll}
@@
 \end{tabular}
-
-\caption{\label{tab:variable_keys} {\tt variables} module common keys}
-\end{table}
+\end{center}
 
@@
 \noindent Here, the {\tt bl\_types} and {\tt multifab\_module} modules
-bring in the basic boxlib data types. Specifically, here, {\tt
+bring in the basic BoxLib data types. Specifically, here, {\tt
 bl\_types} defines {\tt dp\_t} which is the {\tt kind} used for

### 2010-04-01 docs/preface/preface.tex
add a preface describing support for Maestro

MAESTRO is developed at the Center for Computational Sciences and
Engineering at Lawrence Berkeley National Laboratory and at Stony
Brook University.  Development was supported by the SciDAC Program of
the DOE Office of Mathematics, Information, and Computational Sciences
under the U.S. Department of Energy under contract
No.\ DE-AC02-05CH11231 to LBL and by a DOE/Office of Nuclear Physics
Outstanding Junior Investigator award, grant No.\ DE-FG02-06ER41448,
to Stony Brook.

The current version of this User's Guide can be found in the MAESTRO
source tree under {\tt fParallel/MAESTRO/docs/}.

### 2010-04-02 docs/architecture/architecture.tex
add a list of figures

@@
 \includegraphics[width=5.0in]{\archfigpath/maestro_directory2}
-\caption{The basic MAESTRO directory structure}
+\caption[MAESTRO directory structure]
+{The basic MAESTRO directory structure}
 \end{figure}
@@
 \includegraphics[width=6.5in]{\archfigpath/data_loc2}
-\caption{\label{fig:dataloc} Some of the different data-centerings:
+\caption[Data-centerings on the grid]
+  {\label{fig:dataloc} Some of the different data-centerings:
   (a) cell-centered, (b) nodal in the $x$-direction, and (c) nodal in
@@
 \includegraphics[width=4.0in]{\archfigpath/index_grid2}
-\caption{\label{fig:boxes} Three boxes that comprise a single level.
-At this resolution, the domain is 20$\times$18 zones.  Note that the
+\caption[Single-level grid structure]
+{\label{fig:boxes} Three boxes that comprise a single level.  At this
+  resolution, the domain is 20$\times$18 zones.  Note that the
   indexing in BoxLib starts with $0$.}
@@
 \begin{minipage}[b]{3.8in}
-\caption{\label{fig:mac} The MAC grid for the velocity.  
+\caption[The MAC grid]
+{\label{fig:mac} The MAC grid for the velocity.  
 Here the $x$-velocity is on the $x$-edges (shown as the 

### 2010-04-02 docs/flowchart/flowchart.tex
fix how figures are found
add a list of figures

@@
 \centering
-\includegraphics[scale=0.6]{flowchart}
-\caption{\label{Fig:flowchart} A flowchart of the algorithm.  The
+\includegraphics[scale=0.6]{\flowfigpath/flowchart}
+\caption[Graphical flowchart of MAESTRO]
+  {\label{Fig:flowchart} A flowchart of the algorithm.  The
   thermodynamic state variables, base state variables, and local velocity are
@@
 \centering
-\includegraphics[scale=0.6]{flowchart_4_8}
-\caption{\label{Fig:flowchart48}
-A flowchart for {\bf Steps 4} and {\bf 8}.  The thermodynamic state variables
-and base state variables are indicated in each step.  Red text indicates that q\
-uantity
-was updated during that step.  Note, for {\bf Step 4}, the updated
-quantities should also have a $\star$ superscript, e.g., {\bf Step 8I} defines
-$T^{(2)}$ while {\bf Step 4I} defines $T^{(2),\star}$ .}
+\includegraphics[scale=0.6]{\flowfigpath/flowchart_4_8}
+\caption[Graphical flowchart of the density and enthalpy update steps]
+{\label{Fig:flowchart48} A flowchart for {\bf Steps 4} and {\bf 8}.
+  The thermodynamic state variables and base state variables are
+  indicated in each step.  Red text indicates that q\ uantity was
+  updated during that step.  Note, for {\bf Step 4}, the updated
+  quantities should also have a $\star$ superscript, e.g., {\bf Step
+    8I} defines $T^{(2)}$ while {\bf Step 4I} defines $T^{(2),\star}$
+  .}
 \end{figure}

### 2010-04-02 docs/lo_density/lo_density.tex
update cutoff document to reflect new density floor
add a list of figures

@@
 \begin{minipage}[b]{5.0in}
-\caption{Image of how the cutoff density and cutoff coordinates
+\caption[Relation between cutoff density and cutoff coordinates]
+{Image of how the cutoff density and cutoff coordinates
 are related.  This is drawn assuming {\tt drdxfac}=5.\vspace{2em}}
@@
 
+\item In {\tt update\_scal}, if $\rho \le {\tt base\_cutoff\_density}/2$
+   we set it to ${\tt base\_cutoff\_density}/2$.
+
 \item In {\tt make\_grav} for spherical, we only add the enclosed mass if

### 2010-04-04 docs/flowchart/flowchart.tex
add the flowchart from the multilevel paper -- as the algorithm evolves, this version can be updated in step

@@
 \section{Time Advancement Algorithm}\label{Sec:Time Advancement Algorithm}
-Here are changes to the algorithm from the multilevel paper description.
+Here is the current description of the algorithm, based from the description
+in the multilevel paper.
 The initialization has also been included with more detail than was given in paper III.
+
+\begin{description}
+
+%--------------------------------------------------------------------------
+% STEP 0
+%--------------------------------------------------------------------------
+
+\item[Step 0.] {\em Initialization}
+
+This step remains unchanged from Paper III.
+The initialization step only occurs at the beginning of the simulation.  
+The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0, 
+\rho_0^0, p_0^0$, and $\overline{\Gamma_1^0}$ are specified from the problem-dependent 
+initial conditions.  The initial time step, $\dt^0$, is computed as in 
+Paper III.  Finally, initial
+values for $w_0^{-\myhalf}, \etarho^{-\myhalf}, \psi^{-\myhalf}, 
+\pi^{-\myhalf}, S^0$, and $S^1$ come from a preliminary pass through
+the algorithm.
+
+%--------------------------------------------------------------------------
+% STEP 1
+%--------------------------------------------------------------------------
+\item[Step 1.] {\em React the full state through the first time interval of $\dt / 2.$}
+
+Call {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n]$\\
+\phantom{ }\hfill $\rightarrow [\rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)}]$.
+
+%--------------------------------------------------------------------------
+% STEP 2
+%--------------------------------------------------------------------------
+
+\item[Step 2.] {\em Compute the provisional time-centered expansion,
+  $S^{\nph,\star\star}$, provisional base state velocity,
+  $w_0^{\nph,\star}$, and provisional base state velocity forcing.}
+
+\begin{enumerate}
+\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
+
+\item Compute $S^{\nph,\star\star}$.  We compute an estimate for the 
+  time-centered expansion term in the velocity 
+  divergence constraint (eq. [\ref{eq:utilde divergence}]).  For the  
+  first time step ($n=0$), we set
+\begin{equation}
+S^{n+\myhalf,\star\star} = \frac{S^0 + S^1}{2},
+\end{equation}
+  where $S^1$ is found during initialization.  For other time steps
+  $(n \ne 0)$, following \cite{Bell:2004}, we extrapolate
+  to the half-time using $S$ at the previous and current
+  time levels
+\begin{equation}
+S^{\nph,\star\star} = S^n + \frac{\dt^n}{2} \frac{S^n - S^{n-1}}{\dt^{n-1}}.
+\end{equation}
+  Next, compute
+\begin{equation}
+\overline{S^{\nph,\star\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star\star}\right).
+\end{equation}
+
+\item Compute $w_0^{\nph,\star}$.
+\begin{description}
+\item For planar geometry, call\\
+{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star\star}},\overline{\Gamma_1^n},p_0^n,\psi^{n-\myhalf}] \rightarrow [w_0^{\nph,\star}]$.
+\item For spherical geometry, call\\
+{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star\star}},\overline{\Gamma_1^n},\rho_0^n,p_0^n,\etarho^{n-\myhalf}] \rightarrow [w_0^{\nph,\star}]$.
+\end{description}
+
+\item Compute the provisional base state velocity forcing.
+  Rearrange equation (\ref{eq:w0 evolution}),
+\begin{equation}
+-\frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r}, \label{eq:pizero}
+\end{equation}
+  with the following discretization:
+\begin{equation}
+\left ( \frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} \right )^{n,\star} = -\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} - w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star},
+\end{equation} 
+  where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are defined as
+\begin{eqnarray}
+w_0^{n,\star} &=& \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}, \\
+\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} &=& \frac{1}{\dt^n+\dt^{n-1}}\left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh} + \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ].\nonumber \\
+\end{eqnarray}
+  If $n=0$, we use $\dt^{-1} = \dt^0$.
+\end{enumerate}
+
+%--------------------------------------------------------------------------
+% STEP 3
+%--------------------------------------------------------------------------
+\item[Step 3.] {\em Construct the provisional time-centered advective velocity on 
+edges, $\uadvone$.}
+
+Using equation (\ref{eq:utildeupd}), we compute time-centered edge 
+velocities, $\uadvonedag$, using 
+$\Ub = \Ubt^n + w_0^{\nph,\star}$.  The $\dagger$ superscript refers to the 
+fact that the predicted velocity field does not satisfy the divergence 
+constraint.  We then construct $\uadvone$ from $\uadvonedag$
+using a MAC projection, as described in detail in Appendix B of Paper III.  
+We note that $\uadvone$ satisfies the discrete version of
+$\overline{(\uadvone\cdot\eb_r)}=0$ as well as
+\begin{eqnarray}
+\nabla \cdot \left(\beta_0^n \uadvone\right) &=& \beta_0^n \left(S^{\nph,\star\star} - \overline{S^{\nph,\star\star}}\right),\\
+ \beta_0^n &=& \beta_0 \left(\rho_0^n, p_0^n, \overline{\Gamma_1^n}\right),
+\end{eqnarray}
+where $\beta_0$ is computed as described in Appendix C of Paper III.
+
+%--------------------------------------------------------------------------
+% STEP 4
+%--------------------------------------------------------------------------
+\item[Step 4.] {\em Advect the base state and full state through a time interval of $\dt.$}
+
+\begin{enumerate}
+\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
+
+\item Update $\rho_0$, saving the time-centered density at radial edges by calling
+
+{\bf Advect Base Density}$[\rho_0^{n},w_0^{\nph,\star}] \rightarrow [\rho_0^{(2a),\star}, \rho_0^{\nph,\star,\pred}]$.
+
+\item Update $(\rho X_k)$ using a discretized version of equation 
+(\ref{eq:species}) omitting the reaction terms, which were already 
+accounted for in {\bf React State}.  The update consists of two steps:
+
+\begin{enumerate}
+\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
+
+\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\star,\pred}$,
+  for the conservative update of $(\rho X_k)^{(1)}$.  We use equations 
+  (\ref{eq:Perturbational Density}) and (\ref{eq:Primitive Species}) to 
+  predict $\rho^{'(1)} = \rho^{(1)} - \rho_0^n$ and 
+  $X_k^{(1)} = (\rho  X_k)^{(1)} / \rho^{(1)}$ to time-centered edges using 
+  $\Ub = \uadvone+w_0^{\nph,\star}\eb_r$, yielding $\rho^{'\nph,\star,\pred}$ 
+  and $X_k^{\nph,\star,\pred}$.
+  We convert the perturbational density full state density using
+\begin{equation}
+\rho^{\nph,\star,\pred} = \rho^{'\nph,\star,\pred} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2},
+\end{equation}
+  where the base state density terms are mapped to Cartesian edges.
+  Then,\\
+  $(\rho X_k)^{\nph,\star,\pred} = (\rho^{\nph,\star,\pred}X_k^{\nph,\star,\pred})$.
+
+\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ using
+\begin{eqnarray}
+(\rho X_k)^{(2),\star} &=& (\rho X_k)^{(1)} \nonumber \\
+&& - \dt \left\{ \nabla \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right) (\rho X_k)^{\nph,\star,\pred} \right] \right\},\nonumber \\
+\end{eqnarray}
+\begin{equation}
+\rho^{(2),\star} = \sum_k (\rho X_k)^{(2),\star},
+\qquad
+X_k^{(2),\star} = (\rho X_k)^{(2),\star} / \rho^{(2),\star}.
+\end{equation}
+
+\end{enumerate}
+
+\item Define a radial edge-centered $\etarho^{\nph,\star}$.
+
+\begin{description}
+\item For planar geometry, since $\etarho = \overline{\rho'(\Ub\cdot\eb_r)} = \overline{\rho(\Ub\cdot\eb_r)}-\overline{\rho_0(\Ub\cdot\eb_r}) = \overline{\rho(\Ub\cdot\eb_r)} - \rho_0w_0$,
+\begin{eqnarray}
+ \etarho^{\nph,\star} &=&  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvone \cdot \eb_r + w_0^{\nph,\star}\right) (\rho X_k)^{\nph,\star,\pred} \right]\nonumber\\
+&& - w_0^{\nph,\star} \rho_0^{\nph,\star,\pred},
+\end{eqnarray}
+\item For spherical geometry, first construct 
+$\etarho^{{\rm cart},\nph,\star} = [\rho'(\Ub\cdot\eb_r)]^{\nph,\star}$ on Cartesian cell centers using:
+\begin{eqnarray}
+\etarho^{{\rm cart},\nph,\star} &=& \left[\left(\frac{\rho^{(1)}+\rho^{(2),\star}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a),\star}}{2}\right)\right] \nonumber \\
+&&\cdot \left( \uadvone \cdot \eb_r  + w_0^{\nph,\star}\right).
+\end{eqnarray}
+Then,
+\begin{equation}
+\etarho^{\nph,\star} = {\rm {\bf Avg}}\left(\etarho^{{\rm cart},\nph,\star}\right).
+\end{equation}
+This gives a radial cell-centered $\etarho^{\nph,\star}$.  To get
+$\etarho^{\nph,\star}$ at radial edges, average the two neighboring
+radial cell-centered values.
+\end{description}
+
+\item Correct $\rho_0$ by setting $\rho_0^{n+1,\star} =$ {\bf Avg}$(\rho^{(2),\star})$.
+
+\item Update $p_0$ by calling 
+{\bf Enforce HSE}$[p_0^n,\rho_0^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
+
+\item Compute $\psi^{\nph,\star}$.
+\begin{description}
+\item For planar geometry,
+\begin{equation}
+\psi_j^{\nph,\star} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph,\star} 
++ \eta_{\rho,j+\myhalf}^{\nph,\star}\right) g.
+\end{equation}
+\item For spherical geometry, first compute:
+\begin{eqnarray}
+\overline{\Gamma_1^{(1)}} &=& {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(1)}, p_0^{n}, X_k^{(1)}\right) \right]  , \\
+\overline{\Gamma_1^{(2),\star}} &=& {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2),\star}, p_0^{n+1,\star}, X_k^{(2),\star}\right) \right].
+\end{eqnarray}
+Then, define $\psi^{\nph,\star}$ using equation (\ref{eq:psi def})
+\begin{eqnarray}
+\psi_j^{\nph,\star} 
+&=& \left(\frac{\overline{\Gamma_1^{(1)}}+\overline{\Gamma_1^{(2),\star}}}{2}\right)_j
+\left(\frac{p_0^n+p_0^{n+1,\star}}{2}\right)_j \nonumber \\
+&& \left \{ \overline{S_j^{\nph,\star}} - \frac{1}{r_j^2} \left [ \left(r^2 w_0^{\nph,\star}\right)_{j+\myhalf} - \left(r^2 w_0^{\nph,\star}\right)_{j-\myhalf} \right ] \right \}.\nonumber \\
+\end{eqnarray}
+\end{description}
+
+\item Update $(\rho h)_0$.  First, compute $(\rho h)_0^n = $ {\bf Avg}$[(\rho h)^{(1)}]$.
+Then, call\\
+{\bf Advect Base Enthalpy}$[(\rho h)_0^{n}, w_0^{\nph,\star}, \psi^{\nph,\star}] \rightarrow [(\rho h)_0^{n+1,\star}]$.
+
+\item Update the enthalpy using a discretized version of equation
+(\ref{eq:enthalpy}), again omitting the reaction and heating terms
+since we already accounted for
+them in {\bf React State}.  This equation takes the form:
+\begin{equation}
+\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \psi + (\Ubt \cdot \eb_r) \frac{\partial p_0}{\partial r}.
+\end{equation}
+For spherical geometry, we solve the
+analytically equivalent form,
+\begin{equation}
+\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt,
+\end{equation}
+which experience has shown to minimize the drift from thermodynamic equilibrium.  The update consists of two steps:
+
+\begin{enumerate}
+\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
+
+\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\star,\pred},$
+  for the conservative update of $(\rho h)^{(1)}$.  We use equation 
+  (\ref{eq:Perturbational Enthalpy}) to predict
+  $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$ to time-centered edges, 
+  using $\Ub = \uadvone+w_0^{\nph,\star} \eb_r$,
+  yielding $(\rho h)^{'\nph,\star,\pred}$.  We convert the perturbational 
+  enthalpy to a full state enthalpy using
+\begin{equation}
+(\rho h)^{\nph,\star,\pred} = (\rho h)^{'\nph,\star,\pred} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}.
+\end{equation}
+  For planar geometry, we map $(\rho h)_0$ directly to Cartesian edges.
+  In spherical geometry, our experience has shown that a slightly different
+  approach leads to reduced discretization errors.  We first map 
+  $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$ to Cartesian edges separately, 
+  and then multiply these terms to get $(\rho h)_0$.
+
+\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$.
+\begin{description}
+\item For planar geometry,
+\begin{eqnarray}
+(\rho h)^{(2),\star} 
+&=& (\rho h)^{(1)} \nonumber \\
+&&- \dt \left\{ \nabla \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right) (\rho h)^{\nph,\star,\pred} \right] \right\} \nonumber \\ 
+&& + \dt \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} + \dt \psi^{\nph,\star},
+\end{eqnarray}
+
+\item For spherical geometry,
+\begin{eqnarray}
+(\rho h)^{(2),\star} 
+&=& (\rho h)^{(1)} \nonumber \\
+&&- \dt \left\{ \nabla \cdot \left[ \left(\uadvone+w_0^{\nph,\star} \eb_r\right) (\rho h)^{\nph,\star,\pred} \right] \right\} \nonumber \\ 
+&& + \dt \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) - p_0^{n} \nabla \cdot \uadvone \right \} \nonumber \\
+&&+ \dt \psi^{\nph,\star},
+\end{eqnarray}
+\end{description}
+
+\end{enumerate}
+
+Then, for each Cartesian cell where $\rho^{(2),\star} < \rho_\mathrm{cutoff}$,
+we recompute enthalpy using
+\begin{equation}
+(\rho h)^{(2),\star} = \rho^{(2),\star}h\left(\rho^{(2),\star},p_0^{n+1,\star},X_k^{(2),\star}\right).
+\end{equation}
+
+\item Update the temperature using the equation of state:
+$T^{(2),\star} = 
+  T(\rho^{(2),\star}, h^{(2),\star}, X_k^{(2),\star})$ (planar geometry) or
+$T^{(2),\star} = 
+  T(\rho^{(2),\star}, p_0^{n+1,\star}, X_k^{(2),\star})$ (spherical geometry).
+\end{enumerate}
+
+%--------------------------------------------------------------------------
+% STEP 5
+%--------------------------------------------------------------------------
+\item[Step 5.] {\em React the full state through a second time interval of $\dt / 2.$}
+
+Call {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, T^{(2),\star}, 
+(\rho\Hext)^{(2),\star}, p_0^{n+1,\star}]$\\
+\phantom{ }\hfill$\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
+
+%--------------------------------------------------------------------------
+% STEP 6
+%--------------------------------------------------------------------------
+\item[Step 6.] {\em Compute the time-centered expansion, $S^{\nph,\star}$, base state
+velocity, $w_0^{\nph}$, and base state velocity forcing.}
+
+\begin{enumerate}
+\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
+
+\item Compute $S^{\nph,\star}$.  First, compute $S^{n+1,\star}$ with
+\begin{equation}
+S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + \frac{1}{\rho^{n+1,\star} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star} + \sigma \Hnuc^{(2),\star} + \sigma \Hext^{(2),\star},
+\end{equation}
+  where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} /
+  \rho^{(2),\star}$ and the thermodynamic quantities are defined using
+  $\rho^{n+1,\star}, X_k^{n+1,\star},$ and $T^{n+1,\star}$ as inputs to
+  the equation of state.  Then, define
+\begin{equation}
+\overline{S^{\nph,\star}} = {\mathrm{\bf Avg}} (S^{\nph,\star}),
+\qquad
+ S^{\nph.\star} = \frac{S^n + S^{n+1,\star}}{2},
+\end{equation}
+\item Compute $w_0^{\nph}$.  First, define
+\begin{equation}
+\overline{\Gamma_1^{\nph,\star}} = \frac{\overline{\Gamma_1^n} + \overline{\Gamma_1^{n+1,\star}}}{2}, 
+\quad
+\rho_0^{\nph,\star} = \frac{\rho_0^{n} + \rho_0^{n+1,\star}}{2},
+\quad
+p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2},
+\end{equation}
+  with
+\begin{equation}
+ \overline{\Gamma_1^{n+1,\star}} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{n+1,\star}, p_0^{n+1,\star}, X_k^{n+1,\star}\right) \right].
+\end{equation}
+\begin{description}
+\item For planar geometry, call\\
+{\bf Compute} {\boldmath $w_0$} {\bf Planar}$[\overline{S^{\nph,\star}},\overline{\Gamma_1^{\nph,\star}},p_0^{\nph,\star},\psi^{\nph,\star}]\rightarrow [w_0^{\nph}]$.
+\item For spherical geometry, call\\
+{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\overline{\Gamma_1^{\nph,\star}},\rho_0^{\nph,\star},p_0^{\nph,\star},\etarho^{\nph,\star}]$\\
+\phantom{ }\hfill$\rightarrow [w_0^{\nph}]$.
+\end{description}
+
+\item Compute the base state velocity forcing.  Rearrange equation (\ref{eq:pizero}),
+\begin{equation}
+\left ( \frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} \right )^n = 
+-\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
+- w_0^n \left(\frac{\partial w_0}{\partial r}\right)^n,
+\end{equation}
+  where $w_0^{n}$ and $(\partial w_0 / \partial r)^{n}$ are defined as
+\begin{eqnarray}
+w_0^n &=& \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph}}{\dt^n+\dt^{n-1}}, \\
+\left(\frac{\partial w_0}{\partial r}\right)^{n} &=& \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh} + \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph} \right ].\nonumber \\
+\end{eqnarray}
+  If $n=0$, we use $\dt^{-1} = \dt^0$.
+
+\end{enumerate}
+
+%--------------------------------------------------------------------------
+% STEP 7
+%--------------------------------------------------------------------------
+\item[Step 7.] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}
+
+The procedure to construct $\uadvtwodag$ is identical to the procedure
+for computing $\uadvonedag$ in {\bf Step 3}, but uses 
+the updated values $w_0^{\nph}$ and $\pi_0^n$ rather than $w_0^{\nph,\star}$ 
+and $\pi_0^{n,\star}$.  We note that $\uadvtwo$ satisfies the discrete version of
+$\overline{(\uadvtwo\cdot\eb_r)}=0$ as well as
+\begin{equation}
+\nabla \cdot \left(\beta_0^{\nph,\star} \uadvtwo\right) =
+\beta_0^{\nph,\star}\left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right),
+\end{equation}
+\begin{equation}
+\beta_0^{\nph,\star} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2};
+\qquad
+ \beta_0^{n+1,\star} = \beta_0 \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \overline{\Gamma_1^{n+1,\star}}\right).
+\end{equation}
+
+%--------------------------------------------------------------------------
+% STEP 8
+%--------------------------------------------------------------------------
+\item[Step 8.] {\em Advect the base state and full state through a time interval of $\dt.$}
+
+\begin{enumerate}
+\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
+
+\item Update $\rho_0$, saving the time-centered density at radial edges by calling
+
+{\bf Advect Base Density}$[\rho_0^{n},w_0^{\nph}] \rightarrow [\rho_0^{(2a)}, \rho_0^{\nph,\pred}]$.
+
+\item Update $(\rho X_k)$.  This step is identical to {\bf Step 4B} except we use
+  the updated values $w_0^{\nph}, \uadvtwo$, and $\rho_0^{(2a)}$ rather than 
+  $w_0^{\nph,\star}, \uadvone$, and $\rho_0^{(2a),\star}$.  In particular:
+
+\begin{enumerate}
+\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
+
+\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
+  for the conservative update of $(\rho X_k)^{(1)}$.  We use equations 
+  (\ref{eq:Perturbational Density}) and (\ref{eq:Primitive Species}) to 
+  predict $\rho^{'(1)} = \rho^{(1)} - \rho_0^n$ and 
+  $X_k^{(1)} = (\rho  X_k)^{(1)} / \rho^{(1)}$ to time-centered edges
+  with $\Ub = \uadvtwo+w_0^{\nph} \eb_r$,
+  yielding $\rho^{'\nph,\pred}$ and $X_k^{\nph,\pred}$.  
+  We convert the perturbational density to a full state density using
+\begin{equation}
+\rho^{\nph,\pred} = \rho^{'\nph,\pred} + \frac{\rho_0^n + \rho_0^{(2a)}}{2}.
+\end{equation}
+  Then, $(\rho X_k)^{\nph,\pred} = (\rho^{\nph,\pred}X_k^{\nph,\pred})$.
+
+\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ using
+\begin{equation}
+(\rho X_k)^{(2)} = (\rho X_k)^{(1)} 
+- \dt \left\{ \nabla \cdot \left[\left(\uadvtwo+w_0^{\nph} \eb_r\right)  
+(\rho X_k)^{\nph,\pred} \right] \right\},
+\end{equation}
+\begin{equation}
+\rho^{(2)} = \sum_k (\rho X_k)^{(2)},
+\qquad
+X_k^{(2)} = (\rho X_k)^{(2)} / \rho^{(2)}.
+\end{equation}
+
+\end{enumerate}
+
+\item Define a radial edge-centered $\etarho^{\nph}$.
+\begin{description}
+\item For planar geometry,
+\begin{eqnarray}
+ \etarho^{\nph} &=& {\rm {\bf Avg}} \sum_k \left [\left(\uadvtwo \cdot \eb_r + w_0^{\nph}\right) (\rho X_k)^{\nph,\pred} \right] \nonumber \\
+&&- w_0^{\nph} \rho_0^{\nph,\pred},
+\end{eqnarray}
+\item For spherical geometry, first construct 
+$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian
+  cell centers using:
+\begin{equation}
+\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{(2a)}}{2}\right)\right] \left(\uadvtwo \cdot \eb_r + w_0^{\nph}\right).
+\end{equation}
+Then,
+\begin{equation}
+\etarho^{\nph} = {\rm {\bf Avg}}\left(\etarho^{{\rm cart},\nph}\right).
+\end{equation}
+This gives a radial cell-centered $\etarho^{\nph}$.  To get
+$\etarho^{\nph}$ at radial edges, average the two neighboring
+cell-centered values.
+\end{description}
+
+\item Correct $\rho_0$ by setting $\rho_0^{n+1} =$ {\bf Avg}$(\rho^{(2)})$.
+
+\item Update $p_0$ by calling
+{\bf Enforce HSE}$[p_0^n,\rho_0^{n+1}] \rightarrow [p_0^{n+1}]$.
+
+\item Compute $\psi^{\nph}$.
+\begin{description}
+\item For planar geometry, 
+\begin{equation}
+\psi_j^{\nph} = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph} 
++ \eta_{\rho,j+\myhalf}^{\nph}\right) g.
+\end{equation}
+\item For spherical geometry, first compute:
+\begin{equation}
+\overline{\Gamma_1^{(2)}} = {\rm{\bf Avg}} \left[ \Gamma_1\left(\rho^{(2)}, p_0^{n+1}, 
+X_k^{(2)}\right) \right].
+\end{equation}
+Then, define $\psi^{\nph}$ using equation (\ref{eq:psi def}):
+\begin{eqnarray}
+\psi_j^{\nph} 
+&=& \left(\frac{\overline{\Gamma_1^{(1)}}+\overline{\Gamma_1^{(2)}}}{2}\right)_j \left(\frac{p_0^n+p_0^{n+1}}{2}\right)_j \nonumber \\
+&& \left \{ \overline{S_j^{\nph}} - \frac{1}{r_j^2} \left [ \left(r^2 w_0^{\nph}\right)_{j+\myhalf} - \left(r^2 w_0^{\nph}\right)_{j-\myhalf} \right ] \right \}.
+\end{eqnarray}
+\end{description}
+
+\item Update $(\rho h)_0$ by calling 
+{\bf Advect Base Enthalpy}$[(\rho h)_0^n, w_0^{\nph}, \psi^{\nph}] \rightarrow [(\rho h)_0^{n+1}]$.
+
+\item Update the enthalpy.  This step is identical to {\bf Step 4H} except we use
+  the updated values $w_0^{\nph}, \uadvtwo, \rho_0^{n+1}, (\rho h)_0^{n+1}, p_0^{n+\myhalf}$, 
+  and $\psi^{n+\myhalf}$ rather than\\
+  $w_0^{\nph,\star}, \uadvone, \rho_0^{n+1,\star}, (\rho h)_0^{n+1,\star}, p_0^n$, 
+  and $\psi^{n+\myhalf,\star}$.  In particular:
+
+\begin{enumerate}
+\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
+
+\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
+  for the conservative update of $(\rho h)^{(1)}$.  We use equation 
+  (\ref{eq:Perturbational Enthalpy}) to predict
+  $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$ to time-centered edges
+  with $\Ub = \uadvtwo+w_0^{\nph} \eb_r$,
+  yielding $(\rho h)^{'\nph,\pred}$.  
+  We convert the perturbational enthalpy to a full state enthalpy using
+\begin{equation}
+(\rho h)^{\nph,\pred} = (\rho h)^{'\nph,\pred} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1}}{2}.
+\end{equation}
+
+\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$.
+\begin{description}
+\item For planar geometry,
+\begin{eqnarray}
+(\rho h)^{(2)} 
+&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \left(\uadvtwo+w_0^{\nph} \eb_r\right)  (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
+&& + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph + \dt \psi^{\nph},
+\end{eqnarray}
+
+\item For spherical geometry,
+\begin{eqnarray}
+(\rho h)^{(2)} 
+&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \left(\uadvtwo+w_0^{\nph} \eb_r\right)  (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
+&& + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right] + \dt \psi^{\nph},\nonumber \\
+\end{eqnarray}
+\end{description}
+where $p_0^\nph$ is defined as $p_0^\nph = (p_0^n+p_0^{n+1})/2$.
+
+\end{enumerate}
+
+Then, for each Cartesian cell where $\rho^{(2)} < \rho_\mathrm{cutoff}$, we recompute enthalpy using
+\begin{equation}
+(\rho h)^{(2)} = \rho^{(2)}h\left(\rho^{(2)},p_0^{n+1},X_k^{(2)}\right).
+\end{equation}
+
+\item Update the temperature using the equation of state:
+$T^{(2)} = 
+   T(\rho^{(2)}, h^{(2)}, X_k^{(2)})$ (planar geometry) or
+$T^{(2)} = 
+   T(\rho^{(2)}, p_0^{n+1}, X_k^{(2)})$ (spherical geometry).
+\end{enumerate}
+
+%--------------------------------------------------------------------------
+% STEP 9
+%--------------------------------------------------------------------------
+\item[Step 9.] {\em React the full state through a second time interval of $\dt / 2.$}
+
+Call {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho\Hext)^{(2)}, p_0^{n+1}]$\\
+\phantom{ }\hfill$\rightarrow [\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$
+
+%--------------------------------------------------------------------------
+% STEP 10
+%--------------------------------------------------------------------------
+\item[Step 10.] {\em Define the new time expansion, $S^{n+1}$, and $\overline{\Gamma_1^{n+1}}$.}
+
+\begin{enumerate}
+\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
+\item Define
+\begin{equation}
+  S^{n+1} =  -\sigma  \sum_k  \xi_k (\omegadot_k)^{(2)}  + \sigma \Hnuc^{(2)} +
+  \frac{1}{\rho^{n+1} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2)}  
+   + \sigma \Hext^{(2)},
+\end{equation}
+where $(\omegadot_k)^{(2)} = (\rho \omegadot_k)^{(2)} / \rho^{(2)}$
+and the thermodynamic quantities are defined using $\rho^{n+1}$,
+$X_k^{n+1}$, and $T^{n+1}$ as inputs to the equation of state.
+Then, compute
+\begin{equation}
+\overline{S^{n+1}} = {\mathrm{\bf Avg}} (S^{n+1}).
+\end{equation}
+
+\item Define
+\begin{equation}
+\overline{\Gamma_1^{n+1}} = {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1}, 
+X_k^{n+1}\right) \right].
+\end{equation}
+
+\end{enumerate}
+
+%--------------------------------------------------------------------------
+% STEP 11
+%--------------------------------------------------------------------------
+\item[Step 11.] {\em Update the velocity}.
+
+First, we compute the time-centered edge velocities, $\Ubt^{\nph,\pred}$.
+Then, we define
+\begin{equation}
+\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
+\end{equation}
+We update the velocity field $\Ubt^n$ to $\Ubt^{n+1,\dagger}$ by discretizing 
+equation (\ref{eq:utildeupd}) as
+\begin{eqnarray}
+\Ubt^{n+1,\dagger} 
+&=& \Ubt^n - \dt \left[\left(\uadvtwo+ w_0^{\nph} \eb_r\right) \cdot \nabla \Ubt^{\nph,\pred} \right] \nonumber \\
+&&- \dt \left(\uadvtwo \cdot \eb_r\right)  \left(\frac{\partial w_0}{\partial r} \right)^\nph \eb_r \nonumber \\
+&& + \dt \left[ - \frac{1}{\rho^\nph} \mathbf{G} \pi^\nmh + \left(\frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\right)^n \eb_r - \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],\nonumber \\
+\end{eqnarray}
+where $\mathbf{G}$ approximates a cell-centered gradient from nodal
+data.  Again, the $\dagger$ superscript refers 
+to the fact that the updated velocity does not satisfy the divergence 
+constraint.
+
+Finally, we use an approximate nodal projection to define $\Ubt^{n+1}$
+from $\Ubt^{n+1,\dagger},$  such that $\Ubt^{n+1}$ approximately
+satisfies 
+\begin{equation}
+\nabla \cdot \left(\beta_0^{\nph} \Ubt^{n+1} \right) 
+= \beta_0^{\nph} \left(S^{n+1} - \overline{S^{n+1}} \right),
+\end{equation}
+where $\beta_0^{\nph}$ is defined as
+\begin{equation}
+\beta_0^{\nph} = \frac{\beta_0^n + \beta_0^{n+1}}{2}; \qquad
+\beta_0^{n+1} = \beta \left(\rho_0^{n+1}, p_0^{n+1}, \overline{\Gamma_1^{n+1}}, g^{n+1}\right).
+\end{equation}
+As part of the projection we also define the new-time perturbational pressure,
+$\pi^\nph.$  This projection necessarily differs from the MAC projection used in 
+{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
+on edges and $\Ubt^{n+1}$ is defined at cell centers, requiring different divergence
+and gradient operators.  Details of the approximate projection are given in Paper III.
+
+%--------------------------------------------------------------------------
+% STEP 12
+%--------------------------------------------------------------------------
+\item[Step 12.] {\em Compute a new $\dt.$}
+
+Compute $\dt$ for the next time step with the procedure described in
+\S 3.4 of Paper III using $w_0$ as computed in {\bf Step 6} and $\Ubt^{n+1}$
+as computed in {\bf Step 11}.
+
+\end{description}
+
+\noindent This completes one step of the algorithm.
+
 \subsection{Volume Discrepancy Changes}

### 2010-04-05 docs/flowchart/flowchart.tex
add in some equations for context

@@
 \item Compute $S^{\nph,\star\star}$.  We compute an estimate for the
-  time-centered expansion term in the velocity 
-  divergence constraint (eq. [\ref{eq:utilde divergence}]).  For the  
-  first time step ($n=0$), we set
+  time-centered expansion term in the velocity divergence constraint,
+  as given in paper~III equation (19),
+\begin{equation}
+  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
+  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext 
+  + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T \enskip .
+\label{eq:defineS} 
+\end{equation}
+
+  For the first time step ($n=0$), we set
 \begin{equation}
@@
 
-\item Compute the provisional base state velocity forcing.
-  Rearrange equation (\ref{eq:w0 evolution}),
+\item Compute the provisional base state velocity forcing, using equation (38)
+  from paper~III,
 \begin{equation}
@@
 
-Using equation (\ref{eq:utildeupd}), we compute time-centered edge 
+The local velocity field is described by paper III equation (37),
+\begin{equation}
+\frac{\partial\Ubt}{\partial t} = 
+- \left(\Ubt+w_0\right) \cdot \nabla \Ubt
+- \left(\Ubt \cdot \eb_r\right) \frac{\partial w_0}{\partial r} \eb_r
+- \frac{1}{\rho} \nabla\pi
++ \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \eb_r
+- \frac{(\rho-\rhozero)}{\rho} \; g \; \eb_r  \label{eq:utildeupd}  \enskip .
+\end{equation}
+From this, we compute time-centered edge 
 velocities, $\uadvonedag$, using 
@@
 
-\item Update $(\rho X_k)$ using a discretized version of equation 
-(\ref{eq:species}) omitting the reaction terms, which were already 
+\item Update $(\rho X_k)$ using a discretized version of the species 
+continuity equation,
+\begin{equation}
+\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\Ub \rho X_k) = 0 \enskip .
+\label{eq:species}
+\end{equation}
+where we omit the reaction terms, which were already 
 accounted for in {\bf React State}.  The update consists of two steps:
@@
 \item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\star,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$.  We use equations 
-  (\ref{eq:Perturbational Density}) and (\ref{eq:Primitive Species}) to 
+  for the conservative update of $(\rho X_k)^{(1)}$.  We use the equations 
+\begin{eqnarray}
+\frac{\partial\rho'}{\partial t} &=& -\Ub\cdot\nabla\rho' - 
+     \rho'\nabla\cdot\Ub - \nabla\cdot\left(\rho_0\Ubt\right),
+     \label{eq:Perturbational Density}  \\
+\frac{\partial X_k}{\partial t} &=& -\Ub\cdot\nabla X_k + 
+     \omegadot_k. \label{eq:Primitive Species}
+\end{eqnarray}
+  to
   predict $\rho^{'(1)} = \rho^{(1)} - \rho_0^n$ and 
@@
 
-\item Update the enthalpy using a discretized version of equation
-(\ref{eq:enthalpy}), again omitting the reaction and heating terms
+\item Update the enthalpy using a discretized version of the enthalpy
+evolution equation, again omitting the reaction and heating terms
 since we already accounted for
@@
 \end{equation}
-which experience has shown to minimize the drift from thermodynamic equilibrium.  The update consists of two steps:
+which experience has shown to minimize the drift from thermodynamic
+equilibrium.  The update consists of two steps:
 
@@
 \item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\star,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$.  We use equation 
-  (\ref{eq:Perturbational Enthalpy}) to predict
+  for the conservative update of $(\rho h)^{(1)}$.  The perturbational
+  enthalpy equation, neglecting reactions,
+\begin{equation}
+\frac{\partial(\rho h)'}{\partial t} = -\Ub\cdot\nabla(\rho h)' - 
+   (\rho h)'\nabla\cdot\Ub - \nabla\cdot\left[(\rho h)_0\Ubt\right] + \Ubt\cdot\nabla p_0 
+    \label{eq:Perturbational Enthalpy}.
+\end{equation}
+  is used to predict
   $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$ to time-centered edges, 

### 2010-04-05 docs/flowchart/flowchart.tex
move the initial description of the MAESTRO papers into the intro

@@
-Here we outline the algorithm currently implemented in the code.  Our
-starting point for this description are the series of papers describing
-the development of the algorithm:
-\begin{itemize}
-\item {\em Low Mach Number Modeling of Type Ia
-  Supernovae. I. Hydrodynamics,} A. S. Almgren, J. B. Bell, 
-  C. A. Rendleman, \& M. Zingale 2006, ApJ, 637, 922 (henceforth
-  paper I)
-\item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
-  Evolution,} A. S. Almgren, J. B. Bell, C. A. Rendleman, \& M. Zingale
-  2006, ApJ, 649, 927 (henceforth paper II)
-\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
-  A. S. Almgren, J. B. Bell, A. Nonaka, \& M. Zingale
-  2008, ApJ, 684, 449 (henceforth paper III)
-\item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection,}
-  M. Zingale, A. S. Almgren, J. B. Bell, A. Nonaka, \& S. E. Woosley
-  2009, ApJ, 704, 196 (henceforth paper IV)
-\item {\em MAESTRO: An Adaptive Low Mach Number Hydrodynamics Algorithm for Stellar
-  Flows,} A. Nonaka, A. S. Almgren, J. B. Bell, M. J. Lijewski, C. M. Malone,
-  \& M. Zingale 2009, submitted to SIAM J. Sci. Comput. 
-  (henceforth ``the multilevel paper'')
-\end{itemize}
-We also carry over some ideas from our small-scale low Mach number algorithm
-for astrophysical flows:
-\begin{itemize}
-\item {\em Adaptive Low Mach Number Simulations of Nuclear Flames,}
-J. B. Bell, M. S. Day, C. A. Rendleman, S. E. Woosley, \& M. Zingale
-2004, JCP, 195, 2, 677 (henceforth BDRWZ)
-\end{itemize}
+Here we outline the algorithm currently implemented in the code.  The latest
+published reference for MAESTRO is the multilevel paper \cite{multilevel}.  In
+this description, we make frequent reference to paper~I \cite{lowMach}, paper~II \cite{lowMach2},
+paper~III \cite{lowMach3}, and paper~IV \cite{lowMach4}.  We also make reference the BDRZW paper \cite{SNe}
+which describes the small-scale astrophysical combustion algrothm.
 

### 2010-04-05 docs/introduction/introduction.tex
move the initial description of the MAESTRO papers into the intro

@@
+\section{History of MAESTRO}
+
+MAESTRO describes the evolution of low Mach number flows that are in
+hydrostatic equilibrium.  The idea for MAESTRO grew out of our
+small-scale astrophysical combustion algorithm detailed in
+\begin{itemize}
+\item {\em Adaptive Low Mach Number Simulations of Nuclear Flames,}
+J. B. Bell, M. S. Day, C. A. Rendleman, S. E. Woosley, \& M. Zingale
+2004, JCP, 195, 2, 677 (henceforth BDRWZ)
+\end{itemize}
+
+\noindent MAESTRO was developed initially for modeling the convecting
+phase in a white dwarf preceding the ignition of a Type Ia supernovae.
+The low Mach number equations were developed in a series of papers
+leading up to the first application to this problem:
+\begin{itemize}
+\item {\em Low Mach Number Modeling of Type Ia
+  Supernovae. I. Hydrodynamics,} A. S. Almgren, J. B. Bell, 
+  C. A. Rendleman, \& M. Zingale 2006, ApJ, 637, 922 (henceforth
+  paper I)
+\item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
+  Evolution,} A. S. Almgren, J. B. Bell, C. A. Rendleman, \& M. Zingale
+  2006, ApJ, 649, 927 (henceforth paper II)
+\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
+  A. S. Almgren, J. B. Bell, A. Nonaka, \& M. Zingale
+  2008, ApJ, 684, 449 (henceforth paper III)
+\item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection,}
+  M. Zingale, A. S. Almgren, J. B. Bell, A. Nonaka, \& S. E. Woosley
+  2009, ApJ, 704, 196 (henceforth paper IV)
+\end{itemize}
+
+\noindent The current version of the algorithm is described in our multilevel paper:
+\begin{itemize}
+\item {\em MAESTRO: An Adaptive Low Mach Number Hydrodynamics Algorithm for Stellar
+  Flows,} A. Nonaka, A. S. Almgren, J. B. Bell, M. J. Lijewski, C. M. Malone,
+  \& M. Zingale 2009, submitted to SIAM J. Sci. Comput. 
+  (henceforth ``the multilevel paper'')
+\end{itemize}
+
 %-----------------------------------------------------------------------------
@@
 \section{Notation}
+
+Throughout the papers describing MAESTRO, we've largely kept our
+notation consistent.  Table~\ref{table:sym} defines the
+frequently-used quantities and provides their units.
 

### 2010-04-05 docs/architecture/architecture.tex
no longer need to set pmask
more on the initial models

@@
    \end{center}
-
-   Note: for periodic boundaries, one must also set the corresponding
-   {\tt pmask\_}\{{\em x,y,z}\} parameter.
-
-\item {\tt pmask\_x }, {\tt pmask\_y }, {\tt pmask\_z } is set to 
-   true ({\tt T}) to indicate that the boundary is periodic in the
-   $x$, $y$, or $z$ coordinate directions.
 
@@
 
-The routines in {\tt MAESTRO/initial\_model/} take a 1-d initial
-model and map it onto a uniform grid, at the desired resolution, using
-the equation of state in MAESTRO, and using MAESTRO's discretization of
-hydrostatic equilibrium:
+The routines in {\tt MAESTRO/initial\_model/} prepare an initial model
+for MAESTRO.  There are two types of routines here, the first type
+take an existing 1-d initial model produced somewhere else (e.g.\ a
+1-d stellar evolution code), and and map it onto a uniform grid, at
+the desired resolution, using the equation of state in MAESTRO, and
+using MAESTRO's discretization of hydrostatic equilibrium.  The second
+type generate the initial model internally, by integrating the
+condition of hydrostatic equilibrium together with a simplifying
+assumption on the energy (e.g. isothermal or isentropic).  In
+both cases hydrostatic equilibrium is enforced as:
 \begin{equation}
@@
 \end{equation}
+
+The {\tt toy\_atm} example provides a simple approximation for a thin
+(plane-parallel) convectively-unstable accreted layer on the surface
+of a star.  This can be used as the starting point for a more complex
+model.  Further details on the initial model routines can be 
+found in \S~\ref{sec:initial_models_main}.
 

### 2010-04-06 docs/introduction/introduction.tex
added references to our in-progress papers

@@
   Flows,} A. Nonaka, A. S. Almgren, J. B. Bell, M. J. Lijewski, C. M. Malone,
-  \& M. Zingale 2009, submitted to SIAM J. Sci. Comput. 
+  \& M. Zingale 2010, submitted to ApJ
   (henceforth ``the multilevel paper'')
+\end{itemize}
+
+\noindent We have two papers in preparation:
+\begin{itemize}
+\item {\em The Convective Phase Preceding Type Ia Supernovae,}
+  M. Zingale, A. S. Almgren, J. B. Bell, C. M. Malone, A. Nonaka, \& S. E. Woosley 2010,
+  in preparation (henceforth paper V)
+\item {\em Multidimensional Modeling of Type I X-ray Bursts,}
+  C. M. Malone, A. Nonaka, A. S. Almgren, J. B. Bell, \& M. Zingale 2010,
+  in preparation (henceforth ``the XRB paper'')
 \end{itemize}

### 2010-04-06 docs/flowchart/flowchart.tex
added references to our in-progress papers
more touchups

@@
 \item Use corner coupling in advection.
-\item We use {\tt use\_tfromp} to update temperature using $T=T(\rho,X_k,p_0)$ rather 
-than $T=T(\rho,h,X_k)$.
+\item We have an option, controlled by {\tt use\_tfromp}, to update temperature 
+using $T=T(\rho,X_k,p_0)$ rather than $T=T(\rho,h,X_k)$.  The former completely 
+decouples enthalpy from our system.  For spherical problems, we use 
+{\tt use\_tfromp = TRUE}, for planar problems, we use {\tt use\_tfromp = FALSE}.
 \item For spherical problems, we have changed the discretization of 
@@
 See the multilevel paper for the latest.
-\section{Changes Between Papers 4.5 and Paper 5}
+\section{Changes Between the Multilevel Paper and Paper 5}
 \begin{enumerate}
@@
 \end{enumerate}
-\section{Changes Between Paper 5 and XRB Paper}
+\section{Changes Between Paper 5 and the XRB Paper}
 \begin{enumerate}
@@
 controlled by the input parameter, {\tt dpdt\_factor}.
+\item For certain problems, we need to set {\tt do\_eos\_h\_above\_cutoff = FALSE}
+to prevent large, unphysical velocities from appearing near the edge of the star.
 \end{enumerate}
@@
 \begin{itemize}
-
 \item Should we use a predictor-corrector for updating the full-state density?
@@
 advance and {\bf Correct Base} using the more accurate estimate of $\rho_0^{n+1}$?
-
+\item We are still exploring the effects of {\tt use\_tfromp = FALSE} for spherical
+problems.  We would eventually like to run in this mode, but $T=T(\rho,X_k,p_0)$ 
+and $T=T(\rho,h,X_k)$ drift away from each other more than we would like.  Our
+attempts at incorporating a {\tt dpdt\_factor} for spherical problems have not 
+been successful.
 \end{itemize}
@@
 
-This step remains unchanged from Paper III.
-The initialization step only occurs at the beginning of the simulation.  
+This step remains unchanged from Paper III.  See \S \ref{Sec:Initialization}
+for details.  The initialization step only occurs at the beginning of the simulation.  
 The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0, 
@@
   where $S^1$ is found during initialization.  For other time steps
-  $(n \ne 0)$, following \cite{Bell:2004}, we extrapolate
+  $(n \ne 0)$, following \cite{SNe}, we extrapolate
   to the half-time using $S$ at the previous and current

### 2010-04-06 docs/runtime_parameters/runtime_parameters.tex
add intro
changed "min_width" to "cluster_min_width".  The other use of min_width, corresponding to the minimum size a grid can be at the coarsest level of multigrid, remains unchanged.

@@
 \verb=  chk_int  = &   Number of timesteps between writing a checkpoint file  &  0 \\
+\verb=  cluster_min_width  = &   The minimum size on a side for a grid created using make\_new\_grids.  &  16 \\
 \verb=  co_latitude  = &   latitude, in radians, for problems with rotation where the domain is only a subset of a full star.  &  ZERO \\
+\verb=  diag_buf_size  = &   number of timesteps to buffer diagnostic output information before writing (note: not implemented for all problems)  &  10 \\
 \verb=  dm_in  = &   dimensionality (valid values are 2 or 3)  &  2 \\
@@
 \verb=  min_eff  = &   parameter for cluster algorithm for making new grids in adaptive problems  &  0.7d0 \\
-\verb=  min_width  = &   The minimum size on a side for a grid created using make\_new\_grids.  &  16 \\
 \verb=  model_file  = &   input model file  &  "model.hse" \\
@@
 \verb=  plot_trac  = &   plot tracers in plotfile  &  .false. \\
-\verb=  pmask_x  = &   indicates whether the problem domain periodic in the x-direction  &  .false. \\
-\verb=  pmask_y  = &   indicates whether the problem domain periodic in the y-direction  &  .false. \\
-\verb=  pmask_z  = &   indicates whether the problem domain periodic in the z-direction  &  .false. \\
 \verb=  ppm_type  = &   0 = no ppm \newline 1 = 1985 ppm \newline 2 = 2009 ppm  &  1 \\
@@
 \verb=  restart  = &   which file to restart from.  -1 means do not restart  &  -1 \\
+\verb=  restart_into_finer  = &   restart and add a level of refinement  &  .false. \\
 \verb=  rotation_radius  = &   radius used for computing centrifugal term in rotation problems  &  1.0d6 \\

### 2010-04-07 docs/architecture/architecture.tex
started writing up fancy bottom solver - will finish when I get a chance

@@
 \section{Multigrid}
+\subsection{\tt bottom\_solver\_type = 4}
+We have a special bottom solver type that greatly speeds up large problems.
+In fact, you should use this bottom solver whenever possible, because in 
+testing, it always speeds up the multigrid, even if the problem is relatively 
+small.  What this solver essentially does is take the grid structure and
+data from the coarsest level of multigrid, and copy it onto a new grid
+structure with fewer, larger grids, with the intent of further coarsening
+this ``new'' problem so the ``new'' bottom solver has fewer data points on
+a smaller number of grids.  If you want to enable this solver for the nodal 
+solver, set {\tt hg\_bottom\_solver = 4} in your inputs file.  If you want 
+to enable this solver for the cell-centered solver, set 
+{\tt mg\_bottom\_solver = 4} in your inputs file.
+
+To understand how this bottom solver work, the first thing you need to know
+is what the grid structure of the coarsest level of your multigrid V-cycle
+looks like.  Now, figure out what the size of a single box would be if you
+wanted it to fit all the data at the bottom level.  Now, figure out what
+the largest integer $n$ is so that you can evenly divide this single box
+by $2^n$ in every coordinate direction.  If $n < 2$, the program will abort
+since the grid structure is not suitable for this bottom solver.
+
+{\tt max\_mg\_bottom\_nlevs}

### 2010-04-07 docs/architecture/architecture.tex
finished a description of the fancy bottom solver, with examples

--- 2010-04-05 docs/architecture/architecture.tex
@@
 \section{Multigrid}
+\subsection{Special Bottom Solver}
+We have a special bottom solver type that greatly speeds up large problems
+and speeds up smaller problems to a lesser degree.
+In fact, you should use this bottom solver whenever possible, because I
+have yet to find a case that performs worse than our standard bottom solver.
+What this special solver essentially does is take the data from the coarsest 
+level of the original multigrid V-cycle and copy it onto a new grid structure 
+with the same number of total cells in each direction, but with a fewer number of larger 
+grids.  A new V-cycle begins from this point, 
+so we are essentially coarsening this ``new'' problem.
+Now, the coarsest level of the multigrid V-cycle in the ``new'' problem has 
+fewer cells and fewer grids as compared to the original coarsest level.
+
+To enable this solver, set {\tt hg\_bottom\_solver = 4} (for the nodal
+projections) and/or {\tt mg\_bottom\_solver = 4} (for the cell-centered
+projections) in your inputs file.
+
+To understand how this bottom solver works, the first thing you need to know
+is what the grid structure of the coarsest level of your multigrid V-cycle
+looks like (we need to put a description of this in the section above).
+Now, figure out the size of the box you would need if you
+wanted it to fit all the data on the coarsest level.  Now, figure out what
+the largest integer $n$ is so that you can evenly divide the length of this box
+by $2^n$ in every coordinate direction.  If $n < 2$, the program will abort
+since the grid structure is not suitable for this bottom solver.
+
+Now the code will set up a ``new'' problem, using the data at the coarsest level
+of the original problem as the initial data.  The grid structure for this new
+problem has the same number of cells as the coarsest level of the original problem,
+but the data is copied onto a grid structure where each grid has $2^n$ cells
+on each side.  The new V-cycle continues down to the new coarsest level, in
+which each grid has 2 cells on each side.  If you wish to impose a limit on
+the maximum value that $n$ can have, you can do so by setting 
+{\tt max\_mg\_bottom\_nlevs} equal to that value.
+
+{\bf Example 1:} A 3D problem with $768^3$ cells divided into $32^3$ grids, i.e., 
+there is a $12\times 12\times 12$ block of $32^3$ grids.  The coarsest level of the 
+multigrid V-cycle contains $12\times 12\times 12$ grids that have $2^3$ cells, so the 
+entire problem domain has $24^3$ cells.  We see that $n=3$, and create a new problem 
+domain with a $3\times 3\times 3$ block of $8^3$ grids.  The coarsest level of the 
+multigrid V-cycle for the ``new'' problem will be a $3\times 3\times 3$ block of 
+$2^3$ grids.
+
+{\bf Example 2:} A 2D problem with $96\times 384$ cells into $48^2$ grids, i.e., 
+there is a $2\times 8$ block of $48^2$ grids.  The coarsest level of the multigrid 
+V-cycle contains $2\times 8$ grids that have $3^2$ cells, so the entire problem 
+domain has $6\times 24$ cells.  We see that $n=1$, so the program aborts since this grid
+structure is not appropriate for the fancy bottom solver.

### 2010-04-07 docs/architecture/architecture.tex
typo
MG intro

@@
 \section{Multigrid}
+
+MAESTRO uses the multigrid solver to enforce the divergence constraint
+both on the half-time edge-centered advective velocities (the ``MAC
+projection'') and on the final cell-centered velocities (the ``HG
+projection'').  For the MAC projection, since the velocity data is
+edge-centered (the MAC grid), the projection is cell-centered.  For
+the HG projection, since the velocity data is cell-centered, the
+projection is node-centered.  \MarginPar{should draw a figure} The multigrid solver performs a number
+of V-cycles until the residual drops by 10-12 orders-of-magnitude.
+There are several options that affect how the multigrid solver
+behaves, which we describe below.
+
 \subsection{Special Bottom Solver}
@@
 
-{\bf Example 2:} A 2D problem with $96\times 384$ cells into $48^2$ grids, i.e., 
+{\bf Example 2:} A 2D problem with $96\times 384$ cells divided into $48^2$ grids, i.e., 
 there is a $2\times 8$ block of $48^2$ grids.  The coarsest level of the multigrid 

### 2010-04-08 docs/architecture/architecture.tex
change example 1 of the special bottom solver so the numbers match up
some updates to the initial model description

@@
 of a star.  This can be used as the starting point for a more complex
-model.  Further details on the initial model routines can be 
-found in \S~\ref{sec:initial_models_main}.
+model.
+
+MAESTRO initial models are read in by the {\tt init\_base\_state}
+routine.  This expects the initial model to contain a header giving
+the number of variables and their names, followed by rows of data
+giving the coordinate and data values at that coordinate.  The initial
+model should contain the same species data (as mass fractions) as
+defined in the {\tt network} module used by the MAESTRO problem.
+
+Full details on which initial model routine matches each problem and
+how the initial models are used to initialize the full state data can
+be found in \S~\ref{sec:initial_models_main}.
 
@@
 
-{\bf Example 1:} A 3D problem with $768^3$ cells divided into $32^3$ grids, i.e., 
+{\bf Example 1:} A 3D problem with $384^3$ cells divided into $32^3$ grids, i.e., 
 there is a $12\times 12\times 12$ block of $32^3$ grids.  The coarsest level of the 

### 2010-04-08 docs/initial_models/initial_models.tex
add a list of figures
more on the initial models
some updates to the initial model description

@@
+Here we briefly describe the various routines for generating initial models
+for the main MAESTRO problems and how the initial model is used to initialize
+both the base state and the full Cartesian state.
+
 \section{Creating the Model Data from Raw Data}\label{Sec:Creating the Model Data from Raw Data}
+\label{sec:initial_models_main}
 
@@
 In {\tt base\_state.f90}, using the subroutine {\tt
-  init\_base\_state.f90} (which is actually a terrible name since we
+  init\_base\_state} (which is actually a terrible name since we
 are setting the initial data, not the base state), we set the initial
@@
 model data.  Then, we set $h^{\init} =
-h^{\init}(\rho^{\init},T^{\init},X^{\init})$.  Note that in the code,
+h(\rho^{\init},T^{\init},X^{\init})$ through the EOS.  Note that in the code,
 $p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not
@@
 p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the
-fortran subroutine in \S~\ref{Sec:Creating the Model Data from Raw
+initial model generation routines in \S~\ref{Sec:Creating the Model Data from Raw
   Data}.
@@
 \begin{minipage}[b]{5.0in}
-\caption{Multilevel Initialization.  The data from the initial model
+\caption[Multi-level initial model initialization]
+ {Multilevel Initialization.  The data from the initial model
   is represented by the dots on the right.  The initial data at the
@@
 section describes how the base state ($\rho_0$ and $p_0$) and full
-state ($\rho, h, X$, and $T$) are computed.  We require that $\rho_0 =
-\overline\rho$, the base state is in HSE according to equation
-(\ref{HSE Discretization}), and the full state is thermodynamically
-consistent with $p_0$.  Here are the steps:
+state ($\rho, h, X$, and $T$) are computed.  The base state is, general, not
+simply a direct copy of the initial model data, since we require that
+$\rho_0 = \overline\rho$.  Additionally, we require thebase state to
+be HSE according to equation (\ref{HSE Discretization}), and that the full
+state is thermodynamically consistent with $p_0$.  Overall we do:
 \begin{enumerate}

### 2010-04-13 docs/introduction/introduction.tex
a non-mathematician's introduction to what a projection method is.  Hopefully I didn't say anything too wrong.

@@
 MAESTRO describes the evolution of low Mach number flows that are in
-hydrostatic equilibrium.  The idea for MAESTRO grew out of our
-small-scale astrophysical combustion algorithm detailed in
+hydrostatic equilibrium.  The idea for MAESTRO grew out of our success
+in applying low Mach number combustion methods to small-scale
+astrophysical flows.  Our original small-scale astrophysical
+combustion algorithm is detailed in
 \begin{itemize}
@@
 phase in a white dwarf preceding the ignition of a Type Ia supernovae.
-The low Mach number equations were developed in a series of papers
-leading up to the first application to this problem:
+As such, we needed to incorporate the compressibility effects due to
+large-scale stratification in the star.  The method closest in spirit
+to MAESTRO is the pseudo-incompressible method of
+Durran~\cite{durran}, developed for terrestrial atmospheric flows
+(assuming an ideal gas).  Part of the complexity of the equations in
+MAESTRO stems from the need to descibe a general equation of state.
+Additionally, since reactions can significantly alter the hydrostatic
+structure of a star, we incorporated extensions that capture the
+expansion of the background state~\cite{almgren:2000}.  The low Mach
+number equations for stellar flows were developed in a series of
+papers leading up to the first application to this problem:
 \begin{itemize}
@@
 
-\noindent The current version of the algorithm is described in our multilevel paper:
+\noindent The current version of the algorithm is described in our
+multilevel paper:
 \begin{itemize}
@@
 
-\noindent We have two papers in preparation:
+\noindent We have two papers in preparation that describe applications of the
+method:
 \begin{itemize}
@@
 \end{itemize}
+
+\section{Projection Methods 101}
+
+Most astrophysical codes hydrodynamics codes
+(e.g.\ CASTRO~\cite{castro} or FLASH~\cite{flash}) solve the
+compressible Euler equations, which can be written in the form:
+\begin{equation}
+U_t + \nabla \cdot F(U) = 0
+\end{equation}
+where U is the vector of conserved quantities, $U = (\rho, \rho u,
+\rho E)$, with $\rho$ the density, $u$ the velocity, and $E$ the total
+energy per unit mass.  This system of equations can be expressed 
+as a system of advection equations:
+\begin{equation}
+w_t + A(w) w_x = 0
+\end{equation}
+where $w$ are called the characteristic variables, and $A$ is the
+Jacobian, $A \equiv \partial F / \partial U$.  The eigenvalues of the
+matrix $A$ are the characteristic speeds---the speeds at which
+information propagates.  For the Euler equations, these are $u$ and $u
+\pm c$, where $c$ is the sound speed.  Solution methods for the
+compressible equations make use of this wave-nature to compute fluxes
+at the interfaces of grid cells to update the state in time.  An
+excellent introduction to these methods is provided by LeVeque's book
+\cite{leveque}.  The timestep for these methods is limited by the time
+it takes for maximum characteristic speed to traverse one grid cell.
+For very subsonic flows, this means that the timestep is dominated by
+the propagation of soundwaves, which may not be important to the
+overall dynamics of the flow.
+
+In contrast, low Mach number systems (including the equations of
+incompressible hydrodynamics) typically solve one or more
+advection-like equations (representing, e.g.\ conservation of mass and
+momentum) coupled with a divergence constraint on the velocity field.
+For example, the equations of constant-density incompressible flow
+are:
+\begin{eqnarray}
+U_t &=& -U \cdot \nabla U + \nabla p \label{eq:incompressible_u} \\
+\nabla \cdot U &=& 0
+\end{eqnarray}
+Here, $U$ represents the velocity vector%
+%
+\footnote{Here we see an unfortunately conflict
+of notation between the compressible hydro community and the
+incompressible community.  In papers on compressible hydrodynamics,
+$U$ will usually mean the vector of conserved quantities.  In 
+incompressible / low speed papers, $U$ will mean the velocity vector.}
+%
+and $p$ is the dynamical pressure.  The time-evolution equation for
+the velocity (eq.~[\ref{eq:incompressible_u}]) can be solved using
+techniques similar to those developed for compressible hydrodynamics,
+updating the old velocity, $U^n$, to the new time-level, $U^\star$.
+Here the `$^\star$' indicates that the updated velocity does not, in
+general, satisfy the divergence constraint.  A projection method will
+take this updated velocity and force it to obey the constraint
+equation.  The basic idea follows from the idea that any velocity
+field can be expressed as the sum of a divergence-free quantity and
+the gradient of a scalar.  For the velocity, we can write:
+\begin{equation}
+U^\star = U^d + \nabla \phi \label{eq:decomposition}
+\end{equation}
+where $U^d$ is the divergence free portion of the velocity vector,
+$U^\star$, and $\phi$ is a scalar.  Taking the divergence of
+eq~(\ref{eq:decomposition}), we have
+\begin{equation}
+\nabla^2 \phi = \nabla \cdot U^\star
+\end{equation}
+(where we used $\nabla \cdot U^d = 0$).
+With appropriate boundary conditions, this Poisson equation can be
+solved for $\phi$, and the final, divergence free velocity can 
+be computed as
+\begin{equation}
+U^{n+1} = U^\star - \nabla \phi
+\end{equation}
+
+Extensions to variable-density incompressible
+flows~\cite{bellMarcus:1992b} involve a slightly different
+decomposition of the velocity field and, as a result, a slightly
+different Poisson equation.  There is also a variety of different ways
+to express what is being projected~\cite{almgren:bell:crutchfield},
+and different discretizations of the divergence and gradient operators
+lead to slightly different mathematical properties of the methods
+(leading to ``approximate
+projections''~\cite{almgrenBellSzymczak:1996}).  Finally, for
+second-order methods, two projections are typically done per timestep.
+The first (the `MAC' projection~\cite{bellColellaHowell:1991})
+operates one the half-time, edge-centered advective velocities, making
+sure that they satisfy the divergence constraint.  These advective
+velocities are used to construct the fluxes through the interfaces to
+advance the solution to the new time.  The second/final projection
+operates on the cell-centered velocities at the new time, again
+enforcing the divergence constraint.  The MAESTRO algorithm performs
+both of these projections.
+
+The MAESTRO algorithm builds upon these ideas, using a different
+velocity constraint equation that captures the compressibility
+due to local sources and large-scale stratification.
 
@@
         & erg~g$^{-1}$~K$^{-1}$ \\
-\hline
+%\hline
 $f$     & volume discrepancy factor ($0 \le f \le 1$) & -- \\
-\hline
+%\hline
 $g$     & gravitational acceleration                 & cm~s$^{-2}$ \\
-\hline
+%\hline
 $h$     & specific enthalpy                          & erg~g$^{-1}$ \\
-\hline
+%\hline
 $\Hext$ & external heating energy generation rate    & erg~g$^{-1}$~s$^{-1}$ \\
-\hline
+%\hline
 $\Hnuc$ & nuclear energy generation rate             & erg~g$^{-1}$~s$^{-1}$ \\
-\hline
+%\hline
 $h_p$   & $h_p \equiv \left . \partial h / \partial p \right |_{T,X_k}$ & cm$^{3}$~g$^{-1}$ \\
-\hline
+%\hline
 $\kth$  & thermal conductivity                       & erg~cm$^{-1}$~s$^{-1}$~K$^{-1}$ \\
-\hline
+%\hline
 $p_0$   & base state pressure                        & erg~cm$^{-3}$ \\
-\hline
+%\hline
 $p_T$   & $p_T \equiv \left . \partial p / \partial T \right |_{\rho,X_k}$ & erg~cm$^{-3}$~K$^{-1}$ \\
-\hline
+%\hline
 $p_{X_k}$ & $p_{X_k} \equiv \left . \partial p / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~cm$^{-3}$ \\
-\hline
+%\hline
 $p_\rho$ & $p_\rho \equiv \left . \partial p / \partial \rho \right |_{T,X_k}$ & erg~g$^{-1}$ \\
-\hline
+%\hline
 $q_k$   & specific nuclear binding energy            & erg~g$^{-1}$  \\
-\hline
+%\hline
 $r$     & radial coordinate (direction of gravity)   & cm \\
-\hline
+%\hline
 $s$     & specific entropy                           & erg~g$^{-1}$~K$^{-1}$ \\
-\hline
+%\hline
 $S$     & source term to the divergence constraint   & s$^{-1}$ \\
-\hline
+%\hline
 $t$     & time                                       & s \\
-\hline
+%\hline
 $T$     & temperature                                & K \\
-\hline
+%\hline
 $\Ub$     & total velocity ($\Ub = \Ubt + w_0 \eb_r$) & cm~s$^{-1}$ \\
-\hline
+%\hline
 $\Ubt$   & local velocity                             & cm~s$^{-1}$ \\
-\hline
+%\hline
 $\uadv$ & advective velocity (edge-centered)         & cm~s$^{-1}$ \\
-\hline
+%\hline
 $w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
-\hline
+%\hline
 $X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & -- \\
-\hline
+%\hline
 $\beta_0$ & coefficient to velocity
             in velocity constraint equation  & g~cm$^{-3}$ \\
-\hline
+%\hline
 $\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & -- \\
-\hline
+%\hline
 $\etarho$ & $\etarho \equiv \overline{(\rho' \Ub \cdot \eb_r)}$ & g~cm$^{-2}$~s$^{-1}$ \\
-\hline
+%\hline
 $\xi_k$ & $\xi_k \equiv \left . \partial h / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~g$^{-1}$ \\
-\hline 
+%\hline 
 $\pi$   & dynamic pressure & erg~cm$^{-3}$ \\
-\hline
+%\hline
 $\pizero$ & base state dynamic pressure & erg~cm$^{-3}$ \\
-\hline
+%\hline
 $\rho$  & mass density  & g~cm$^{-3}$ \\
-\hline
+%\hline
 $\rho_0$  & base state mass density  & g~cm$^{-3}$ \\
-\hline
+%\hline
 $\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
-\hline
+%\hline
 $(\rho h)_0$ & base state enthalpy density & erg~cm$^{-3}$  \\
-\hline
+%\hline
 $(\rho h)'$ & perturbational enthalpy density 
               $ \left [(\rho h)' = \rho h - (\rho h)_0 \right ]$ & erg~cm$^{-3}$  \\
-\hline
+%\hline
 $\sigma$ & $\sigma \equiv p_T/(\rho c_p p_\rho)$ & erg$^{-1}$~g \\
-\hline
+%\hline
 $\psi$  & $\psi \equiv D_0 p_0/Dt = \partial p_0/\partial t + w_0\partial p_0/\partial r$ & erg~cm$^{-3}$~s$^{-1}$ \\
-\hline
+%\hline
 $\omegadot_k$ & creation rate for species $k$ ($\omegadot_k \equiv DX_k/Dt$) & s$^{-1}$ \\
@@
 \renewcommand{\arraystretch}{1.0}
-There are some great charts in the appendix of the multilevel paper demonstrating what 
-the superscript notation represents.

### 2010-04-13 docs/introduction/introduction.tex
Fix tiny typo -- otherwise looks good!

@@
 The first (the `MAC' projection~\cite{bellColellaHowell:1991})
-operates one the half-time, edge-centered advective velocities, making
+operates on the half-time, edge-centered advective velocities, making
 sure that they satisfy the divergence constraint.  These advective

### 2010-04-13 docs/introduction/introduction.tex
typo

@@
 
-Most astrophysical codes hydrodynamics codes
+Most astrophysical hydrodynamics codes
 (e.g.\ CASTRO~\cite{castro} or FLASH~\cite{flash}) solve the

### 2010-04-13 docs/introduction/introduction.tex
rewording

@@
 
-In contrast, low Mach number systems (including the equations of
-incompressible hydrodynamics) typically solve one or more
+In contrast, solving low Mach number systems (including the equations of
+incompressible hydrodynamics) typically involves solving one or more
 advection-like equations (representing, e.g.\ conservation of mass and

### 2010-04-22 docs/eta_notes/eta.tex
fixed some small errors

@@
 the spherical $\theta$ and $\phi$ angles at constant radius.
-
-In Paper III, we introduced a mixing term, $\etarho$, to the base state density
-evolution equation (see eq. [29] of Paper III), with the objective of
-keeping the base state density equal to the average of the density
-over a layer.  Here we derive $\etarho$ for spherical problems.
-For this derivation, it is best to define the average in terms of spherical coordinates,
-\begin{equation}
-\overline{\phi(r)} = \frac{1}{4\pi} \int_{\Omega_H} \phi(r,\theta,\phi) d\Omega
-\end{equation}
-where $\int_{\Omega_H} d\Omega = 4\pi$ represents the integral over
-the angles $\theta$ and $\phi$ at constant radius.
 

### 2010-04-22 docs/introduction/introduction.tex
generalize decomposition of velocity field to decomposition of any vector field
fixed some small errors

@@
 %
-\footnote{Here we see an unfortunately conflict
+\footnote{Here we see an unfortunate conflict
 of notation between the compressible hydro community and the
@@
 take this updated velocity and force it to obey the constraint
-equation.  The basic idea follows from the idea that any velocity
+equation.  The basic idea follows from the idea that any vector
 field can be expressed as the sum of a divergence-free quantity and

### 2010-04-22 docs/runtime_parameters/runtime_parameters.tex
fixed some small errors

@@
 \verb=  dpdt_factor  = &   factor in front of the volume discrepancy term (0.0 = off)  &  0.d0 \\
-\verb=  drdxfac  = &   ratio of radial base state zones to Cartesian full state zones for spherical geometry  &  1 \\
+\verb=  drdxfac  = &   ratio of radial base state cells to Cartesian full state cells for spherical geometry, (i.e. $\Delta r = \Delta x/ \mathtt{drdxfac}$)  &  1 \\
 \verb=  enthalpy_pred_type  = &   predict\_rhohprime = 1; \newline predict\_h = 2; \newline predict\_T\_then\_rhohprime = 3; \newline predict\_T\_then\_h = 4; \newline predict\_hprime = 5; \newline predict\_Tprime\_then\_h = 6.
@@
 \verb=  use_eos_coulomb  = &    &  .true. \\
-\verb=  use_etarho  = &   turn on the etarho term as described in flow chart  &  .true. \\
+\verb=  use_etarho  = &   turn on the $\etarho$ term as described in the flow chart  &  .true. \\
 \verb=  use_soundspeed_firstdt  = &   Use the soundspeed constraint when computing the first time step.  &  .false. \\

### 2010-04-29 docs/architecture/architecture.tex
add Ann's explaination of a layout

@@
 
-\subsection{\boxarray}
-
-A \boxarray\ is an array of boxes
-
-\subsection{\mlboxarray}
-
-A \mlboxarray\ is a collection of \boxarray s at different levels of
-refinement.
-
-\subsection{\mllayout}
+\subsection{\boxarray\ and \mlboxarray}
+
+A \boxarray\ is an array of boxes.  A \mlboxarray\ is a collection of
+\boxarray s at different levels of refinement.
+
+\subsection{\layout\ and \mllayout}
+
+A \layout\ is basically a \boxarray\ that knows information about other
+boxes, or box ``connectivity.''  It contains additional information
+that is used in filling ghost cells from other fine grids or from
+coarser grids.  This information is stored as long as the layout
+exists so that we don't have to recompute intersections every time we
+do some operation with two multifabs that have that layout, for
+example.
 

### 2010-05-05 docs/FAQ/faq.tex
add Ann's answer to Brendan's pi time-centering question

@@
   {\tt PROF=TRUE} in the {\tt GNUmakefile}, or equvalently do 
-  {\tt make PROF=t}
+  {\tt make PROF=t}.  A file containing a summary of the timings will
+  be output in the run directory.
+
+  An alternate way to get single-processor timings, when using 
+  the GCC compilers is to add {\tt -pg} to the compilation flags
+  for both {\tt gfortran} and {\tt gcc}.  Upon completion, a
+  file names {\tt gmon.out} will be produced.  This can be 
+  processed by {\tt gprof} running
+
+  {\tt~~~ gprof} {\em exec-name}
+
+  where {\em exec-name} is the name of the executable.
 
@@
 \end{enumerate}
+
+\section{Algorithm}
+
+\begin{enumerate}
+
+\item {\em In the final projection in the algorithm, we project
+  $U^{n+1}$, using a time-centered $\beta_0$, a time-centered $\rho_0$, but
+  an ``$n+1$''-centered $S$.  When then is the resulting $\phi$ (which then
+  defines $\pi$) is at ``$n+1/2$''?}
+
+  The short answer to this question is that you should think of this
+  as really projecting  $(U^{n+1} - U^n)$ and the right hand side as having
+  $(S^{n+1} - S^n)$.  This is because the pressure enters the dynamic equations as
+  $(U^{n+1} - U^n) = \ldots + \frac{1}{\rho^{n+1/2}} \nabla \pi^{n+1/2}$.
+  (We approximate $\pi^{n+1/2}$ by $\pi^{n-1/2}$ then do the projection to fix the
+  $\pi$ as well as the $U$.)
+
+  So everything is in fact time-centered.
+
+\end{enumerate}

### 2010-05-31 docs/initial_models/initial_models.tex
added a suggestion

@@
 
+\MarginPar{I think it would be a good idea to mention some where in here that MAESTRO outputs info that's useful to check --dr of base state and dr of the input file, which should be the same at the finest level; and Maximum HSE Error, which should be some small number.}
+
 \section{Creating the Model Data from Raw Data}\label{Sec:Creating the Model Data from Raw Data}
@@
 
-We begin with raw data, $\rho^{\raw}, T^{\raw}, p^{\raw}$, and
+MAESTRO needs to be given model data with the same resolution as the base state.
+(Note that for spherical problems we set 
+$\Delta r = \Delta x/\mathtt{drdxfac}$.)  
+We need to generate this from the raw data,
+ $\rho^{\raw}, T^{\raw}, p^{\raw}$, and
 $X^{\raw}$.  Here is the raw data file for each test problem:
@@
 \end{eqnarray}
-Using the raw data, we need to generate model data with the same
-resolution as the base state.  Note that for spherical problems we set
-$\Delta r = \Delta x/\mathtt{drdxfac}$.  We use a fortran subroutine
+
+We use a fortran subroutine
 to interpolate the raw data, yielding the model data, $\rho^{\model},
@@
 to the file containing the model data.
+
+During runtime, MAESTRO will outout information about the Check that ``dr of MAESTRO base state'' and ``dr of input file data''
 

### 2010-06-07 docs/architecture/architecture.tex
more vis

@@
 
-{\tt Amrvis} is a visualization tool developed at LBL for 2- and 3-d
-datasets which can plot slices through 3-d datasets as well as
-volume-renderings.  It is distributed separately from the MAESTRO
-distribution.
+{\tt Amrvis} is an easy-to-use visualization tool developed at LBL for
+2- and 3-d datasets which can plot slices through 3-d datasets as well
+as volume-renderings.  It can also very easily extract 1-d lines
+through the dataset along any coordinate direction.  It is distributed
+separately from the MAESTRO distribution.
 
@@
 directly from BoxLib plotfiles (to open a plotfile in VisIt, point to
-the {\tt Header} file inside the plotfile directory.  Details on
+the {\tt Header} file inside the plotfile directory).  Details on
 visit can be found at {\tt https://wci.llnl.gov/codes/visit/home.html}\,.

### 2010-06-08 docs/introduction/introduction.tex
characteristic -> primitive
update multilevel paper ref

@@
   Flows,} A. Nonaka, A. S. Almgren, J. B. Bell, M. J. Lijewski, C. M. Malone,
-  \& M. Zingale 2010, submitted to ApJ
-  (henceforth ``the multilevel paper'')
+  \& M. Zingale 2010, ApJS, 188, 358 (henceforth ``the multilevel paper'')
 \end{itemize}
@@
 \begin{equation}
-w_t + A(w) w_x = 0
+q_t + A(q) q_x = 0
 \end{equation}
-where $w$ are called the characteristic variables, and $A$ is the
+where $q$ are called the primitive variables, and $A$ is the
 Jacobian, $A \equiv \partial F / \partial U$.  The eigenvalues of the

### 2010-06-08 docs/architecture/architecture.tex
update docs to reflect new Makefile organization

@@
 
-The next set of lines tell the build system where to look for source
-files.  A MAESTRO application is built from several pacakges (the
-multigrid solver, an EOS, a reaction network, etc.).  
+The next two lines define where the top of the source tree is
+include the basic Makefile information (like compiler
+flags).
 \begin{verbatim}
@@
   include $(FPARALLEL)/mk/GMakedefs.mak
-
-  Fmdirs := boxlib \
-            mg \
-            MAESTRO \
-            extern/EOS/helmeos \
-            extern/EOS/conducteos \
-            extern/constants \
-            extern/networks/ignition \
-            extern/VODE
-
-  Fmincludes := extern/EOS/helmeos
-\end{verbatim}
-The {\tt FPARALLEL} variable should point to the top-level directory
-of the source tree (i.e.\ {\tt fParallel/}).  Next we include the
-compiler definitions, {\tt mk/GMakedefs.mak}.  The {\tt Fmdirs}
-variable lists the individual directories in the {\tt fParallel/}
-source tree that are used to build the problem.  Generally, one does
-not need to include the problem directory itself, unless there are
-unique source files found there, described in a {\tt GPackage.mak}
-file.  The build system will attempt to build all of the files listed
-in the various {\tt GPackage.mak} files found in the {\tt Fmdirs}
-directories.  Furthermore, {\tt Fmdirs} will be will be added to the
-{\tt make} {\tt VPATH}, which is the list of directories to search for
-source files.  The problem directory will always be put first in the
-{\tt VPATH}, so any source files placed there override those with the
-same name found elsewhere in the source tree.  The {\tt Fmincludes}
-variable lists all those directories that contain include files that
-are inserted into the Fortran source at compile time via the {\tt
-  include} statement.  Presently, the only instance of this is with
-the Helmholtz general equation of state found in {\tt
-  extern/EOS/helmeos/}.
+\end{verbatim}
+
+A MAESTRO application is built from several pacakges (the
+multigrid solver, an EOS, a reaction network, etc.).  The core
+MAESTRO packages are always included, so a problem only needs
+to define the EOS, reaction network, and conductivities to
+use, as well as any extra, problem specific files.  
+\begin{verbatim}
+EOS_DIR := extern/EOS/helmeos   
+CONDUCTIVITY_DIR := extern/EOS/conducteos 
+NETWORK_DIR := extern/networks/ignition_chamulak
+
+EXTRA_DIR := extern/random
+\end{verbatim}
+Generally, one does not need to include the problem directory itself
+in {\tt EXTRA\_DIR}, unless there are unique source files found there,
+described in a {\tt GPackage.mak} file.  These variables are
+interpreted by the {\tt GMaestro.mak} file and used to build a master
+list of packages called {\tt Fmdirs}.  The build system will attempt
+to build all of the files listed in the various {\tt GPackage.mak}
+files found in the {\tt Fmdirs} directories.  Furthermore, {\tt
+  Fmdirs} will be will be added to the {\tt make} {\tt VPATH}, which
+is the list of directories to search for source files.  The problem
+directory will always be put first in the {\tt VPATH}, so any source
+files placed there override those with the same name found elsewhere
+in the source tree.  Some packages (for instance, the {\tt helmeos}
+EOS) require Fortran include files.  The {\tt Fmincludes} variable
+lists all those directories that contain include files that are
+inserted into the Fortran source at compile time via the {\tt include}
+statement.  Presently, the only instance of this is with the Helmholtz
+general equation of state found in {\tt extern/EOS/helmeos/}.  This is
+automatically handled by the {\tt GMaestro.mak} instructions.
 

### 2010-06-08 docs/architecture/architecture.tex
sync up with Makefile

@@
 
-The final lines in the {\tt GNUmakefile} include the rules to actually
+The final line in the {\tt GNUmakefile} include the rules to actually
 build the executable.
@@
   include $(FPARALLEL)/MAESTRO/GMaestro.mak
-  include $(FPARALLEL)/mk/GMakerules.mak
 \end{verbatim}

### 2010-06-16 docs/FAQ/faq.tex
typo
start a question on multigrid non-convergence

@@
 \end{verbatim}
+
+\item {\em  MAESTRO crashes because the multigrid algorithm fails to
+    converge---how do I get around this?}
+
+  Setting general convergence criteria for multigrid is as much
+  art as science. 
+  First, it is important to determine if the multigrd solver is
+  close to convergence and just dancing around near the desired
+  tolerance, but never reaching it, or if it is no where near 
+  convergence.  For the latter, it may be that the multigrid
+  solver was fed bad data and the problem arose in one of the earlier
+  steps.  To get more detail information from the multigrid solver,
+  set {\tt mg\_verbose} to a positive integer from 1-4 (the higher
+  the number the more information you receive.
+
+  If the multigrid solver is failing during one of the initial
+  ``divu'' iterations, it may be because the velocity is initially
+  zero, so there is no velocity magnitude to use as a reference for
+  convergence, and that ($S - \bar{S}$) is very small (or zero).  In
+  this case, it is usually a good idea to perturb the initial state
+  slightly, so the righthand side is non-zero.
 
@@
   $U^{n+1}$, using a time-centered $\beta_0$, a time-centered $\rho_0$, but
-  an ``$n+1$''-centered $S$.  When then is the resulting $\phi$ (which then
+  an ``$n+1$''-centered $S$.  Why then is the resulting $\phi$ (which then
   defines $\pi$) is at ``$n+1/2$''?}

### 2010-06-18 docs/mg/mg.tex
add documentation on the various eps in MG

\section{Multigrid Solver Tolerances}

Beginning at the start of execution, there are several places where either cell-centered
multigrid or node-centered multigrid solves are performed.

\begin{itemize}

\item {\em initial projection} ({\tt initial\_proj} called from {\tt varden})

  The initial projection creates a first approximation to the velocity
  field by forcing the initial velocity field set by {\tt initveldata}
  to satisfy the elliptic constraint equation.  Since the initial
  velocity may be zero, there is no guarantee that a well-defined
  timestep can be computed at this point, so the source term, $S$,
  used here only involves thermal diffusion and any external heating
  term, $\Hext$---no reactions are included (see paper~III, \S 3.3).

  The initial projection can be disabled with the {\tt do\_initial\_projection}
  runtime parameter.

  The tolerance, $\epsilon$ is set in {\tt initial\_proj} and is:
   \begin{center}
   \begin{tabular}{ll}
   Cartesian:   & $\epsilon = 10^{-12}$ \\
   spherical:   & $\epsilon = 10^{-10}$ \\
   \end{tabular}
   \end{center}

\item {\em ``divu'' iterations} ({\tt divu\_iter} called from {\tt varden})

  The ``divu'' iterations projects the velocity field from the initial
  projection to satisfy the full constraint (including reactions).
  This is an iterative process since the reactions depend on the
  timestep and the timestep depends on the velocity field (see
  paper~III, \S 3.3).  The number of iterations to take is set through
  the {\tt init\_divu\_iter} runtime parameter.

  The tolerance, $\epsilon_\mathrm{divu}$ depends on the iteration, $i$.
  We start with a loose tolerance and progressively get tighter.  The
  tolerances (set in {\tt divu\_iter}) are:
   \begin{center}
   \begin{tabular}{ll}
   Cartesian:   & $\epsilon = \left  \{ \begin{array}{ll} 
                   \min  (10^{-6\phantom{0}}, 10^{-8\phantom{0}} \cdot 10^{(\mathrm{nlevs - 1})} ) & 
                           \quad \mathrm{for}~ i \le \mathrm{init\_divu\_iter} - 2 \\
                   \min  (10^{-8\phantom{0}}, 10^{-10} \cdot 10^{(\mathrm{nlevs - 1})} ) & 
                           \quad \mathrm{for}~ i = \mathrm{init\_divu\_iter} - 1  \\
                   \min  (10^{-10}, 10^{-12} \cdot 10^{(\mathrm{nlevs - 1})} ) & 
                           \quad \mathrm{for}~ i = \mathrm{init\_divu\_iter}   \\
                                 \end{array}
                  \right .$ \\[10mm]
   spherical:   & $\epsilon = \left  \{ \begin{array}{ll} 
                    10^{-6\phantom{0}}  &
                           \quad \mathrm{for}~ i \le \mathrm{init\_divu\_iter} - 2 \, \\
                    10^{-8\phantom{0}}  &
                           \quad \mathrm{for}~ i = \mathrm{init\_divu\_iter} - 1 \, \\
                    10^{-10}  &
                           \quad \mathrm{for}~ i = \mathrm{init\_divu\_iter} \, )\\
                  \end{array}
                  \right .$ \\
   \end{tabular}
   \end{center}

\end{itemize}

In the main algorithm, mulitgrid solves come in during the two MAC projections,
two (optional) thermal diffusion solves, and the final velocity projection.

\begin{itemize}

\item {\em MAC projection}

  The MAC projection forces the edge-centered, half-time advective
  velocities to obey the elliptic constraint.  This is done both in
  the predictor and corrector portions of the main algorithm.

  There are 3 tolerances here: a relative tolerance of $\epsilon =
  10^{-10}$, an absolute tolerance of $\epsilon_\mathrm{abs} =
  \epsilon \cdot ||U^\mathrm{ADV}||_\infty / \Delta x$, and a bottom
  solver tolerance of $\epsilon_\mathrm{bottom} = 10^{-3}$ (these are
  all hard-coded in {\tt mac\_multigrid}).

\item {\em thermal diffusion}

  This uses the same {\tt mac\_multigrid} routine as the MAC
  projection, so it uses the same tolerances.  The only difference is
  that the absolute tolerance is based on the norm of $h$ now, instead
  of $U^\mathrm{ADV}$.

\item {\em velocity projection}

  The final velocity projection uses a tolerance of $\epsilon = \min(
  10^{-10}, 10^{-12} \cdot 10^{(\mathrm{nlevs} - 1)} )$.  This tolerance
  is set in {\tt advance} and passed into {\tt hgproject}.

\end{itemize}

### 2010-07-21 docs/architecture/architecture.tex
Get rid of edge_nodal_flag
update to show Source/

@@
 
- The main MAESTRO algorithm files.  Here you will find the main driver,
- the advection routines, etc.
+ The main MAESTRO algorithm directory.  The individual problem directories
+ lie under here, and all of the main MAESTRO source lives in the {\tt Source/}
+ sub-directory.
 
@@
  \begin{itemize}
+
+ \item {\tt Source}
+
+   The main MAESTRO source.  Here you will find the driver routine, the
+   advection routines, etc.  All MAESTRO problems will compile this
+   source.
 
@@
 
- \item problem directories: {\tt test}, {\tt test2}, {\tt
+ \item problem directories: {\tt test2}, {\tt
    test\_spherical}, $\ldots$
@@
   do comp=1,dm
-     call multifab_build(umac(n,comp), mla%la(n),1,1, &
-                         nodal=edge_nodal_flag(comp,:))
+     call multifab_build_edge(umac(n,comp), mla%la(n),1,1,comp)
   enddo

### 2010-08-04 docs/volume_discrepancy/volume_discrepancy.tex
add some more details on the form of the w0 equation that we solve in make_w0, when we have the volume discrepancy term

@@
 \end{equation}
+
+In solving the $w_0$ evolution equation
+(Eq.~[\ref{eq:w0constraint_vd}]), we expand the divergence, giving
+\begin{equation}
+\nablab \cdotb (w_0 \er)  = \overline{S} - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} - 
+w_0 \er  \cdotb \frac{1}{\beta_0} \nablab \beta_0  - \frac{f}{\gammabar p_0} \frac{p_0 - \overline{p_\mathrm{EOS}}}{\Delta t} \enskip .
+\end{equation}
+Recalling that 
+\begin{equation*}
+\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial r} = 
+ \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}
+\end{equation*}
+(see Paper~I), and that $\psi \equiv D_0 p_0 / Dt \equiv \partial p_0 / \partial t + 
+w_0 \partial p_0 / \partial r$, we have
+\begin{equation}
+\nablab \cdotb (w_0 \er)  = \overline{S} - \frac{1}{\gammabar p_0} \psi -
+ \frac{f}{\gammabar p_0} \frac{p_0 - \overline{p_\mathrm{EOS}}}{\Delta t} \enskip .
+\end{equation}
+This is the form that is solved in {\tt make\_w0}.

### 2010-08-05 docs/planar_invsq_base/basestate.tex
start of the notes on planar 1/r^2 gravity, based on Brendan's derivation notes

\section{\label{app:gravity} Modifications for a $1/r^2$ Plane-Parallel Basestate}

In the plane parallel assumption, we model a layer of thickness
$\Delta R$ a distance $R_\mathrm{base}$ from the center of a star,
under the assumption the $\Delta R \ll R_\mathrm{base}$.  In this
assumption, we can neglect the curvature of the atmosphere.  Here, we 
extend that basic assumption to allow for the gravitational acceleration
to still fall off as $1/r^2$ as we move outward in the envelope.  We assume
that the mass of the envelope is insignificant, and that only the mass of 
the underlying star contributes to the gravitational acceleration.

\subsection{Constraint Equation}

Begin assuming the following equation (derived elsewhere):
\begin{equation}
\nabla \cdot  ({w_0 \er}) = \bar{S} - \frac{1}{\gammabar p_0} \psi
\end{equation}
where
\[ \psi = \frac{D_0 p_0}{D t} = \frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial t}. \]
In Cartesian geometry, the divergence on the left hand side is simply $\partial
w_0 / \partial r$.
Let us now define $\bar{w_0}$ and $\delta w_0$ such that
\[ w_0 = \bar{w_0} + \delta w_0 \quad\quad \mbox{and} \quad\quad
\frac{\partial \ow}{\partial t} = \bar{S}, \]
leaving the equation
\begin{equation}
\frac{\partial \dw}{\partial t} = - \frac{1}{\gammabar p_0} \left ({ \frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r}} \right ).
\end{equation}

If we multiply by $\gammabar p_0$, differentiate by r, then switch the order of $\partial
t$ and $\partial r$ where they appear in the same term, we get
\begin{equation}
\frac{\partial}{\partial r} \left[ \gammabar p_0 \frac{\partial \delta w_0}{\partial r} \right] =
   - \frac{\partial}{\partial t} \frac{\partial p_0}{\partial r} -  \frac{\partial}{\partial r} \left (w_0 \frac{\partial p_0}{\partial r} \right) .
\end{equation}

We then substitute in the equation of hydrostatic equilibrium:
\[ \frac{\partial p_0}{\partial r} = -\rho_0 g \quad\quad \mbox{where} \quad\quad
   g = \frac{G m}{r^2}. \]
We will assume that the mass of the atmosphere is negligible relative to the
mass of the core, which is outside of the simulation domain.  This simplifies
the equation by allowing us to assume that $m$, the enclosed mass, is constant.
So we now have
\begin{equation}
\frac{\partial}{\partial r} \left[ \gammabar p_0 \frac{\partial}{\partial r}{({\delta w_0})} \right]
   = \frac{\partial}{\partial t}{({\rho_0 g}}) + \frac{\partial}{\partial r}{({w_0 \rho_0 g})}
   = \rho_0 \left ({\frac{\partial}{\partial t}{g} + w_0 \frac{\partial}{\partial r}{g}} \right )
         + g \left ({\frac{\partial}{\partial t}{\rho_0} + \frac{\partial}{\partial r}{({w_0 \rho_0})}} \right ).
\end{equation}

We now recall Equation~29 from Paper III:
\[ \frac{\partial}{\partial t}{\rho_0} = - \nabla \cdot ({w_0 \rho_0 \er})
                  - \nabla \cdot ({\etarho \er}), \]
which is, in Cartesian geometry,
\begin{equation}
\frac{\partial \rho_0}{\partial t} = - \frac{\partial}{\partial r}{({w_0 \rho_0})}
                  - \frac{\partial \etarho}{\partial r}.
\end{equation}
Substituting this expression yields
\begin{equation}
\frac{\partial}{\partial r}{} \left[ \gammabar p_0 \frac{\partial}{\partial r}{({\delta w_0})} \right]
   = \rho_0 \frac{D_0 g}{D t} + g \left ({- \frac{\partial}{\partial r}{({w_0 \rho_0})} - \frac{\partial \etarho}{\partial r} + \frac{\partial}{\partial r}{({w_0 \rho_0})}} \right)
   = \rho_0 \frac{D_0 g}{D t} - g \frac{\partial \etarho}{\partial r}.
\end{equation}

We then differentiate the gravitational acceleration:
\begin{eqnarray}
\frac{D_0 g}{D t}
   & = & \frac{D_0}{Dt} \left ({\frac{G m}{r^2}} \right ) \nonumber \\
   & = & G m \left ({\frac{\partial}{\partial t}{({r^{-2}})} + w_0 \frac{\partial}{\partial r}{({r^{-2}})}} \right ) \nonumber \\
   & = & - \frac{2 G m w_0}{r^3} \nonumber \\
   & = & - \frac{2 w_0 g}{r}.
\end{eqnarray}
Substituting in this expression gives our final result:
\begin{equation}
\frac{\partial}{\partial r}{} \left[ \gammabar p_0 \frac{\partial}{\partial r}{({\dw})} \right]
   = - \frac{2 w_0 \rho_0 g}{r} - g \frac{\partial \etarho}{\partial r}
   = \frac{2 w_0}{r} \frac{\partial p_0}{\partial r} - g \frac{\partial \etarho}{\partial r}.
\end{equation}

\subsection{Discretization}

Collecting all of the $\dw$ terms on the left side, our constraint equation
appears as:
\begin{equation}
\frac{\partial}{\partial r} \left [ \gammabar p_0 \frac{\partial \dw}{\partial r} \right ]
 + \frac{2 \rho_0 g \dw}{r} = -\frac{2 \rho_0 g \ow}{r} - g \frac{\partial \etarho}{\partial r}
\end{equation}
%
On a uniform mesh (constant $\dr$, we would discretize this as:
\begin{eqnarray}
\frac{1}{\dr} \left \{ \left [ \gammabar p_0 \frac{\partial \dw}{\partial r} \right ]_j
                     - \left [ \gammabar p_0 \frac{\partial \dw}{\partial r} \right ]_{j-1}
              \right \}
            &+& \left [ \frac{2 \rho_0 g}{r} \dw \right ]_{j-\half} \nonumber \\
            = &-& \left [ \frac{2 \rho_0 g}{r} \ow \right ]_{j-\half}
              - \frac{g_{j-\half}}{\dr} \left [ {\etarho}_j - {\etarho}_{j-1} \right ]
\end{eqnarray}
Expanding the $\partial \dw / \partial r$ terms, we have:
\begin{eqnarray}
\frac{1}{(\dr)^2} \left \{ \left [ (\gammabar p_0)_j
                               \left ({\dw}_{j+\half} - {\dw}_{j-\half} \right ) \right ]
                      \right . &-& \left . \left [ (\gammabar p_0)_{j-1}
                               \left ({\dw}_{j-\half} - {\dw}_{j-\thalf} \right ) \right ]
              \right \}
            + \left [ \frac{2 \rho_0 g}{r} \dw \right ]_{j-\half} \nonumber \\
            = &-& \left [ \frac{2 \rho_0 g}{r} \ow \right ]_{j-\half}
              - \frac{g_{j-\half}}{\dr} \left [ {\etarho}_j - {\etarho}_{j-1} \right ]
\end{eqnarray}
As with the spherical case (multilevel paper, appendix B), we write this in the form:
\begin{equation}
A_j (\dw)_{j-\thalf} + B_j (\dw)_{j-\myhalf} + C_j (\dw)_{j+\myhalf} = F_j,
\end{equation}
then:
\begin{eqnarray}
A_j &=& \frac{1}{\Delta r^2} \left( {\gammabar p_0}\right)_{j-1}, \\
B_j &=& -\frac{1}{\Delta r^2} \left[ \left( {\gammabar p_0}\right)_{j}  + \left( {\gammabar p_0}\right)_{j-1} \right] +  \frac{2}{r_{j-\myhalf}} \left (\rho_0 g \right )_{j-\half}  , \\
C_j &=& \frac{1}{\Delta r^2} \left( {\gammabar p_0}\right)_{j}  , \\
F_j &=&  -\frac{2}{r_{j-\myhalf}} (\rho_0 g)_{j-\half}   (\ow)_{j-\half} - \frac{g_{j-\half}}{\Delta r} \left[ \left( \etarho \right)_{j} - \left( \etarho \right)_{j-1} \right] 
\end{eqnarray}

\subsection{Boundary Conditions}

Together with the assumption that the mass of the envelope does not
contribute to the gravitational acceleration, we assume that as we move
a fluid element in the atmosphere, it does not drive a velocity at the very 
base of the layer.  Therefore, we take $w_0(r_\mathrm{base}) = 0$.

### 2010-08-05 docs/planar_invsq_base/basestate.tex
tweaks
start non-uniform section

@@
 
-Begin assuming the following equation (derived elsewhere):
+We begin with the $w_0$ constraint equation (derived elsewhere), including the volume-discrepancy term:
 \begin{equation}
-\nabla \cdot  ({w_0 \er}) = \bar{S} - \frac{1}{\gammabar p_0} \psi
+\nabla \cdot  ({w_0 \er}) = \bar{S} - \frac{1}{\gammabar p_0} \psi - \frac{f}{\gammabar p_0} \frac{p_0 - \overline{{p_0}_\mathrm{EOS}}}{\dt}
 \end{equation}
@@
 w_0 / \partial r$.
-Let us now define $\bar{w_0}$ and $\delta w_0$ such that
-\[ w_0 = \bar{w_0} + \delta w_0 \quad\quad \mbox{and} \quad\quad
-\frac{\partial \ow}{\partial t} = \bar{S}, \]
+Let us now define $\ow$ and $\dw$ such that
+\[ w_0 = \ow + \dw \]
+We take $\ow$ to satisfy
+\begin{equation}
+\frac{\partial \ow}{\partial t} = \bar{S} - \frac{f}{\gammabar p_0} \frac{p_0 - \overline{{p_0}_\mathrm{EOS}}}{\dt} ,
+\end{equation}
 leaving the equation
@@
 \end{equation}
-
+%
 If we multiply by $\gammabar p_0$, differentiate by r, then switch the order of $\partial
@@
 \end{equation}
-
+%
 We then substitute in the equation of hydrostatic equilibrium:
@@
    = - \frac{2 w_0 \rho_0 g}{r} - g \frac{\partial \etarho}{\partial r}
-   = \frac{2 w_0}{r} \frac{\partial p_0}{\partial r} - g \frac{\partial \etarho}{\partial r}.
 \end{equation}
 
-\subsection{Discretization}
+\subsection{Uniform $\dr$ Discretization}
 
@@
 \begin{eqnarray}
-A_j &=& \frac{1}{\Delta r^2} \left( {\gammabar p_0}\right)_{j-1}, \\
-B_j &=& -\frac{1}{\Delta r^2} \left[ \left( {\gammabar p_0}\right)_{j}  + \left( {\gammabar p_0}\right)_{j-1} \right] +  \frac{2}{r_{j-\myhalf}} \left (\rho_0 g \right )_{j-\half}  , \\
-C_j &=& \frac{1}{\Delta r^2} \left( {\gammabar p_0}\right)_{j}  , \\
-F_j &=&  -\frac{2}{r_{j-\myhalf}} (\rho_0 g)_{j-\half}   (\ow)_{j-\half} - \frac{g_{j-\half}}{\Delta r} \left[ \left( \etarho \right)_{j} - \left( \etarho \right)_{j-1} \right] 
+A_j &=& \frac{1}{\dr^2} \left( {\gammabar p_0}\right)_{j-1}, \\
+B_j &=& -\frac{1}{\dr^2} \left[ \left( {\gammabar p_0}\right)_{j}  + \left( {\gammabar p_0}\right)_{j-1} \right] +  \frac{2}{r_{j-\myhalf}} \left (\rho_0 g \right )_{j-\half}  , \\
+C_j &=& \frac{1}{\dr^2} \left( {\gammabar p_0}\right)_{j}  , \\
+F_j &=&  -\frac{2}{r_{j-\myhalf}} (\rho_0 g)_{j-\half}   (\ow)_{j-\half} - \frac{g_{j-\half}}{\dr} \left[ \left( \etarho \right)_{j} - \left( \etarho \right)_{j-1} \right] 
 \end{eqnarray}
+
+\subsection{Non-Uniform $\dr$ Discretization}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=4in]{\planeinvsqfigpath/grid2}
+\end{figure}
+
+Consider the above non-uniform grid spacing,
+where $\dr_c = 2 \dr_f$.  Here, the discretization of the Laplacian-like term is more complex.
+We want to compute
+\begin{equation}
+\frac{\partial}{\partial r} \left [ \gammabar p_0 \frac{\partial \dw}{\partial r} \right ]_{j-\half}
+\end{equation}
+This is to be centered at $j-\half$, which we accomplish by averaging the two fine grids and then
+differencing:
+\begin{equation}
+\frac{\partial}{\partial r} \left [ \gammabar p_0 \frac{\partial \dw}{\partial r} \right ]_{j-\half} =
+\frac{1}{\dr_c} \left \{ \frac{1}{2} \left [
+      \left ( \gammabar p_0 \frac{\partial \dw}{\partial r} \right )_{j+1} +
+      \left ( \gammabar p_0 \frac{\partial \dw}{\partial r} \right )_{j} \right ]
+   - \left ( \gammabar p_0 \frac{\partial \dw}{\partial r} \right )_{j-1} 
+\right \}
+\end{equation}
 

### 2010-08-05 docs/planar_invsq_base/basestate.tex
non-uniform is not tri-diagonal

@@
 \end{equation}
+Expanding the $\partial \dw / \partial r$ terms results in a equation depending on $\dw$ at
+4 different edge locations---this no longer fits into the tri-diagonal format used in the
+uniform grid case.  In detail, it becomes:
+\begin{eqnarray}
+\frac{\partial}{\partial r} \left [ \gammabar p_0 \frac{\partial \dw}{\partial r} \right ]_{j-\half} &=& 
+\frac{1}{\dr_c} \left \{  \frac{1}{2} \left [
+      \left ( \gammabar p_0 \right )_{j+1} \frac{(\dw)_{j+\thalf} - (\dw)_{j+\half}}{\dr_f} +
+      \left ( \gammabar p_0 \right )_{j}   \frac{(\dw)_{j+\half} - (\dw)_{j-\half}}{\dr_f}  \right ] \right .  \nonumber \\
+&& \qquad \left .  - \left ( \gammabar p_0 \right )_{j-1} \frac{(\dw)_{j-\half} - (\dw)_{j-\thalf}}{\dr_c}  
+\right \}
+\end{eqnarray}
+which has terms proportional to $(\dw)_{j-\thalf}$, $(\dw)_{j-\half}$, $(\dw)_{j+\half}$, and $(\dw)_{j+\thalf}$
 

### 2010-08-09 docs/runtime_parameters/runtime_parameters.tex
update to latest _parameters

--- 2010-04-06 docs/runtime_parameters/runtime_parameters.tex
@@
 
+\verb=  amr_buf_width  = &   the number of buffer zones surrounding a cell tagged for refinement. note that this needs to be >= regrid\_int  &  -1 \\
 \verb=  anelastic_cutoff  = &   The density below which we modify the constraint to look like the anelastic constraint, instead of the low Mach constraint.
@@
 &  3.d6 \\
+\verb=  barrier_timers  = &   Call a parallel\_barrier before each of the simple timers we have coded in advance.f90.  Might cause a very slight overall performance hit.  &  .false. \\
 \verb=  base_cutoff_density  = &   The density below which we keep the initial model constant. Refer to Section \ref{Sec:Base Cutoff Density}  &  3.d6 \\
@@
 \verb=  co_latitude  = &   latitude, in radians, for problems with rotation where the domain is only a subset of a full star.  &  ZERO \\
+\verb=  conductivity_constant  = &   the constant conductivity (only applies if the conductivity/constant module is used)  &  1.0d0 \\
 \verb=  diag_buf_size  = &   number of timesteps to buffer diagnostic output information before writing (note: not implemented for all problems)  &  10 \\
@@
 \verb=  do_initial_projection  = &   Do the initial projection.  &  .true. \\
+\verb=  do_planar_invsq_grav  = &   are we doing 1/r$^2$ gravity for plane-parallel  &  .false. \\
 \verb=  do_smallscale  = &   force $\rho_0 = (\rho h)_0 = 0$, {\tt evolve\_base\_state = F} and {\tt beta\_type} = 3  &  .false. \\
@@
 \verb=  n_cellz  = &   Number of cells for the base level in the z-direction  &  -1 \\
+\verb=  octant  = &   set octant = T if you just want to model an octant of a sphere (note: only takes effect for spherical geometry)  &  .false. \\
 \verb=  perturb_model  = &   Turn on a perturbation in the initial data.  Problem specific.  &  .false. \\
+\verb=  planar_invsq_mass  = &   the point mass for planar 1/r$^2$ gravity  &  0.d0 \\
+\verb=  plot_Hext  = &   plot external heating (Hext) in plotfile  &  .false. \\
 \verb=  plot_base  = &   plot w0\_x, w0\_y, w0\_z, divw0, rho0, rhoh0, h0, and p0 in plotfile  &  .false. \\
@@
 \verb=  plot_deltat  = &   rather than use a plot interva, plot a file after the solution has advanced past plot\_deltat in time  &  0.d0 \\
+\verb=  plot_eta  = &   plot $\etarho$ in plotfile  &  .false. \\
 \verb=  plot_int  = &   plot interval  &  0 \\
+\verb=  plot_omegadot  = &   plot omegadot in plotfile  &  .true. \\
 \verb=  plot_spec  = &   plot species and omegadot in plotfile  &  .true. \\
@@
 \verb=  use_divu_firstdt  = &   Use the divu constraint when computing the first time step.  &  .false. \\
-\verb=  use_eos_coulomb  = &    &  .true. \\
+\verb=  use_eos_coulomb  = &   use the Coulomb corrections between ions (if supported by the EOS)  &  .true. \\
 \verb=  use_etarho  = &   turn on the etarho term as described in flow chart  &  .true. \\
+\verb=  use_hypre  = &   use the hypre library  &  .false. \\
 \verb=  use_soundspeed_firstdt  = &   Use the soundspeed constraint when computing the first time step.  &  .false. \\

### 2010-08-11 docs/thermo_notes/thermo_notes.tex
add a simple derivation of the adiabatic excess using quantities returnable from our EOS

@@
 solution.
+
+\section{Adiabatic Excess}
+\label{Sec:Adiabatic Excess}
+The adiabatic excess, $\Delta\nabla$, is a quantity used to determine
+if a system is stable ($\Delta\nabla < 0$) or unstable ($\Delta\nabla
+> 0$) to convection under the Schwarzschild criterion (i.e. neglecting
+compositional gradients).  Cox and Giuli (see chapter 9) define three
+different ``adiabatic exponents'' that we will use:
+\begin{eqnarray*}
+  \Gamma_1 &\equiv&   \left(\frac{d\ln p}{d\ln\rho}\right)_\text{ad} \\
+  \frac{\Gamma_2}{\Gamma_2-1} &\equiv& 
+  \left(\frac{d\ln p}{d\ln T}\right)_\text{ad} \\
+  \Gamma_3 - 1 &\equiv& \left(\frac{d\ln T}{d\ln\rho}\right)_\text{ad},
+\end{eqnarray*} 
+where the subscript ``ad'' means along an adiabat.  We can combine the
+exponents to get the following relation
+\begin{equation}\label{eq:Gamma relations}
+  \Gamma_1 = \left(\frac{\Gamma_2}{\Gamma_2-1}\right)\left(\Gamma_3-1\right).
+\end{equation}
+
+The adiabatic excess is defined as
+\begin{equation}\label{eq:adiabatic excess}
+  \Delta\nabla = \nabla_\text{actual} - \nabla_\text{ad}
+\end{equation}
+where
+\begin{equation}\label{eq:thermal gradient}
+  \nabla \equiv \frac{d\ln T}{d\ln P}
+\end{equation}
+is the thermal gradient.  It is important to note that these thermal
+gradients are only along the radial direction.  The ``actual''
+gradient can be found from finite differencing the data whereas the
+adiabatic term, $\nabla_\text{ad} = \left(\Gamma_2-1\right) /
+\Gamma_2$, will need to be calculated at each point using
+thermodynamic relations.  Our EOS only returns $\Gamma_1$ so we need
+find another relation to use with \eqref{eq:Gamma relations} to solve
+for the adiabatic excess.
+
+The Schwarzschild criterion does not care about changes in composition
+and we therefore write $p = p(\rho,T)$ and
+\begin{equation}\label{eq:dp}
+  d\ln p = \chi_\rho d\ln\rho + \chi_T d\ln T
+\end{equation}
+where
+\[
+\chi_\rho = \left(\frac{d\ln p}{d\ln\rho}\right)_T,\qquad
+\chi_T = \left(\frac{d\ln p}{d\ln T}\right)_\rho.
+\]
+Dividing \eqref{eq:dp} by $d\ln\rho$ and taking this along an adiabat
+we have
+\begin{equation}\label{eq:dp2}
+  \left(\frac{d\ln p}{d\ln\rho}\right)_\text{ad} = \chi_\rho + \chi_T
+  \left(\frac{d\ln T}{d\ln\rho}\right)_\text{ad}.
+\end{equation}
+Using the $\Gamma$'s, we have
+\begin{equation}\label{eq:Gamma1 relation with Gamma2}
+  \Gamma_1 = \chi_\rho + \chi_T\left(\Gamma_3-1\right).
+\end{equation}
+Combining \eqref{eq:Gamma relations} and \eqref{eq:Gamma1 relation
+  with Gamma2} to eliminate $\Gamma_3$, we have:
+\begin{equation}\label{eq:nabla_ad}
+  \nabla_\text{ad} = \frac{\Gamma_1 - \chi_\rho}{\chi_T\Gamma_1}
+\end{equation}
+which uses only terms which are easily returned from an EOS call.

### 2010-08-17 docs/mg/mg.tex
reference mg_eps module
update the description of the tolerances based on the current version of mg_eps.f90

@@
 
-Beginning at the start of execution, there are several places where either cell-centered
-multigrid or node-centered multigrid solves are performed.
+Beginning at the start of execution, there are several places where
+either cell-centered multigrid or node-centered multigrid solves are
+performed.  The outline below lists the solves one encounters, in order,
+from the start of execution.  The values of the tolerances lists here
+are defined in the {\tt mg\_eps} module.  To set problem-specific values
+of these tolerances, place a local copy of {\tt mg\_eps.f90} in your
+problem directory.
+
+In the initialization, multigrid comes in during the initial projection
+and the ``divu'' iterations.
 
@@
 
-  The tolerance, $\epsilon$ is set in {\tt initial\_proj} and is:
+  The tolerances, {\tt eps\_init\_proj\_cart} and {\tt eps\_init\_proj\_sph}
+  (for Cartesian and spherical respectively) are set in {\tt mg\_eps.f90}
+  and have the default values of:
    \begin{center}
-   \begin{tabular}{ll}
-   Cartesian:   & $\epsilon = 10^{-12}$ \\
-   spherical:   & $\epsilon = 10^{-10}$ \\
+   \begin{tabular}{lll}
+   Cartesian:   &  {\tt eps\_init\_proj\_cart} &= $10^{-12}$ \\
+   spherical:   &  {\tt eps\_init\_proj\_sph}  &= $10^{-10}$
    \end{tabular}
@@
 
-  The tolerance, $\epsilon_\mathrm{divu}$ depends on the iteration, $i$.
+  The overall tolerance, $\epsilon_\mathrm{divu}$ depends on the iteration, $i$.
   We start with a loose tolerance and progressively get tighter.  The
-  tolerances (set in {\tt divu\_iter}) are:
+  tolerances (set in {\tt divu\_iter}) are, for Cartesian:
+  {\small
+   \begin{center}
+   \begin{tabular}{lll}
+   $\epsilon_\mathrm{divu} = \left  \{ \begin{array}{lll} 
+                   \min\, \{& \!\!\!\mathtt{eps\_divu\_cart} \cdot \mathtt{divu\_iter\_factor}^2 \cdot \mathtt{divu\_level\_factor}^{(\mathtt{nlevs}-1)}, \\
+                            & \!\!\!\mathtt{eps\_divu\_cart} \cdot \mathtt{divu\_iter\_factor}^2 \cdot \mathtt{divu\_level\_factor}^2 \, \} & 
+                           \quad \mathrm{for}~ i \le \mathtt{init\_divu\_iter} - 2 \\[2mm]
+                   \min\, \{& \!\!\!\mathtt{eps\_divu\_cart} \cdot \mathtt{divu\_iter\_factor} \cdot \mathtt{divu\_level\_factor}^{(\mathtt{nlevs}-1)}, \\
+                            & \!\!\!\mathtt{eps\_divu\_cart} \cdot \mathtt{divu\_iter\_factor} \cdot \mathtt{divu\_level\_factor}^2 \, \} & 
+                           \quad \mathrm{for}~ i = \mathtt{init\_divu\_iter} - 1  \\[2mm]
+                   \min\, \{& \!\!\!\mathtt{eps\_divu\_cart} \cdot \mathtt{divu\_level\_factor}^{(\mathtt{nlevs}-1)}, \\
+                            & \!\!\!\mathtt{eps\_divu\_cart} \cdot \mathtt{divu\_level\_factor}^2 \, \} & 
+                           \quad \mathrm{for}~ i = \mathtt{init\_divu\_iter}   \\
+                                 \end{array}
+                  \right .$ \\[10mm]
+   \end{tabular}
+   \end{center}
+   }
+   and for spherical:
+   {\small 
    \begin{center}
    \begin{tabular}{ll}
-   Cartesian:   & $\epsilon = \left  \{ \begin{array}{ll} 
-                   \min  (10^{-6\phantom{0}}, 10^{-8\phantom{0}} \cdot 10^{(\mathrm{nlevs - 1})} ) & 
-                           \quad \mathrm{for}~ i \le \mathrm{init\_divu\_iter} - 2 \\
-                   \min  (10^{-8\phantom{0}}, 10^{-10} \cdot 10^{(\mathrm{nlevs - 1})} ) & 
-                           \quad \mathrm{for}~ i = \mathrm{init\_divu\_iter} - 1  \\
-                   \min  (10^{-10}, 10^{-12} \cdot 10^{(\mathrm{nlevs - 1})} ) & 
-                           \quad \mathrm{for}~ i = \mathrm{init\_divu\_iter}   \\
-                                 \end{array}
-                  \right .$ \\[10mm]
-   spherical:   & $\epsilon = \left  \{ \begin{array}{ll} 
-                    10^{-6\phantom{0}}  &
-                           \quad \mathrm{for}~ i \le \mathrm{init\_divu\_iter} - 2 \, \\
-                    10^{-8\phantom{0}}  &
-                           \quad \mathrm{for}~ i = \mathrm{init\_divu\_iter} - 1 \, \\
-                    10^{-10}  &
-                           \quad \mathrm{for}~ i = \mathrm{init\_divu\_iter} \, )\\
+   $\epsilon_\mathrm{divu} = \left  \{ \begin{array}{ll} 
+                     \mathtt{eps\_divu\_sph} \cdot \mathtt{divu\_iter\_factor}^2 &
+                           \quad \mathrm{for}~ i \le \mathtt{init\_divu\_iter} - 2 \, \\[2mm]
+                    \mathtt{eps\_divu\_sph} \cdot \mathtt{divu\_iter\_factor}  & 
+                           \quad \mathrm{for}~ i = \mathtt{init\_divu\_iter} - 1 \, \\[2mm]
+                    \mathtt{eps\_divu\_sph}  &
+                           \quad \mathrm{for}~ i = \mathtt{init\_divu\_iter} \, )\\
                   \end{array}
                   \right .$ \\
+   \end{tabular}
+   \end{center}
+   }
+   The various parameters are set in {\tt mg\_eps.f90} and have the default values of:
+   \begin{center}
+   \begin{tabular}{ll}
+   {\tt eps\_divu\_cart}     &= $10^{-12}$ \\
+   {\tt eps\_divu\_sph}      &= $10^{-10}$ \\
+   {\tt divu\_iter\_factor}  &= 100 \\
+   {\tt divu\_level\_factor} &= 10 
    \end{tabular}
@@
 
-  There are 3 tolerances here: a relative tolerance of $\epsilon =
-  10^{-10}$, an absolute tolerance of $\epsilon_\mathrm{abs} =
-  \epsilon \cdot ||U^\mathrm{ADV}||_\infty / \Delta x$, and a bottom
-  solver tolerance of $\epsilon_\mathrm{bottom} = 10^{-3}$ (these are
-  all hard-coded in {\tt mac\_multigrid}).
+  There are 2 tolerances here: a relative tolerance of $\epsilon =
+  \mathtt{eps\_mac} \cdot \mathtt{mac\_level\_factor}^{(\mathtt{nlevs}-1)}$, and a bottom
+  solver tolerance of $\epsilon_\mathrm{bottom} = \mathtt{eps\_mac\_bottom}$.  These
+  parameters are set in {\tt mg\_eps.f90} and have the default values:
+   \begin{center}
+   \begin{tabular}{ll}
+   {\tt eps\_mac}           &= $10^{-10}$ \\
+   {\tt mac\_level\_factor} &= 10 \\
+   {\tt eps\_mac\_bottom}   &= $10^{-3}$
+   \end{tabular}
+   \end{center}
+
+  %an absolute tolerance of $\epsilon_\mathrm{abs} =
+  %\epsilon \cdot ||U^\mathrm{ADV}||_\infty / \Delta x$
 
@@
 
-  The final velocity projection uses a tolerance of $\epsilon = \min(
-  10^{-10}, 10^{-12} \cdot 10^{(\mathrm{nlevs} - 1)} )$.  This tolerance
-  is set in {\tt advance} and passed into {\tt hgproject}.
+  The final velocity projection uses a tolerance of $\epsilon = \min \{
+  \mathtt{eps\_hg\_max}, \mathtt{eps\_hg} \cdot \mathtt{hg\_level\_factor}^{(\mathtt{nlevs} - 1)} \}$.  This tolerance
+  is set in {\tt hgproject} using the parameter values specified in {\tt mg\_eps.f90}.  The default 
+  parameter values are:
+   \begin{center}
+   \begin{tabular}{ll}
+   {\tt eps\_hg}     &= $10^{-12}$ \\
+   {\tt eps\_hg\_max}      &= $10^{-10}$ \\
+   {\tt hg\_level\_factor}  &= 10 \\
+   \end{tabular}
+   \end{center}
 

### 2010-08-17 docs/mg/mg.tex
update the MAC section

@@
 
-  There are 2 tolerances here: a relative tolerance of $\epsilon =
-  \mathtt{eps\_mac} \cdot \mathtt{mac\_level\_factor}^{(\mathtt{nlevs}-1)}$, and a bottom
-  solver tolerance of $\epsilon_\mathrm{bottom} = \mathtt{eps\_mac\_bottom}$.  These
-  parameters are set in {\tt mg\_eps.f90} and have the default values:
+  There are two tolerances here.  The norm of the residual is required
+  to be reduced by a relative tolerance of $\epsilon =
+  \min \{ \mathtt{eps\_mac\_max}, \mathtt{eps\_mac} \cdot
+  \mathtt{mac\_level\_factor}^{(\mathtt{nlevs}-1)} \}$. A separate
+  tolerance is used for the bottom
+  solver, $\epsilon_\mathrm{bottom} =
+  \mathtt{eps\_mac\_bottom}$.  These parameters are set in {\tt
+    mg\_eps.f90} and have the default values:
    \begin{center}
@@
    {\tt eps\_mac}           &= $10^{-10}$ \\
+   {\tt eps\_mac\_max}           &= $10^{-8}$ \\
    {\tt mac\_level\_factor} &= 10 \\

### 2010-08-17 docs/mg/mg.tex
updated for eps_hg_bottom

@@
   \mathtt{eps\_hg\_max}, \mathtt{eps\_hg} \cdot \mathtt{hg\_level\_factor}^{(\mathtt{nlevs} - 1)} \}$.  This tolerance
-  is set in {\tt hgproject} using the parameter values specified in {\tt mg\_eps.f90}.  The default 
-  parameter values are:
+  is set in {\tt hgproject} using the parameter values specified in {\tt mg\_eps.f90}.  A separate
+  tolerance is used for the bottom
+  solver, $\epsilon_\mathrm{bottom} = \mathtt{eps\_hg\_bottom}$.
+
+  The default parameter values are:
    \begin{center}
@@
    {\tt hg\_level\_factor}  &= 10 \\
+   {\tt eps\_hg\_bottom}   &= $10^{-4}$
    \end{tabular}

### 2010-09-03 docs/initial_models/initial_models.tex
found a typo
write out enforce_HSE for non-constant g

@@
 to the file containing the model data.
-
-During runtime, MAESTRO will outout information about the Check that ``dr of MAESTRO base state'' and ``dr of input file data''
 
@@
 \subsection{Coarse-Fine HSE Discretization}\label{Sec:Coarse-Fine HSE Discretization}
-Integrating the HSE discretization upward, when we encounter a change
-from coarse to fine, here is an equation to integrate up to the
-coarse-fine interface ($l$ is the ``finer'' level, $l-1$ is the
-``coarser'' level)
+When integrating the HSE discretization upward, we must use a different
+differencing procedure at coarse-fine interfaces.  Figure~\ref{fig:ctof} shows
+the transition from coarse (level $l-1$) to fine (level $l$), with the zone
+center indices noted.  
+\begin{figure}[t]
+\centering
+\includegraphics[width=2.5in,angle=90]{\figpath/ctof}
+\caption{\label{fig:ctof} A coarse-fine interface in the 1-d base state}
+\end{figure}
+
+To find the zone-centered pressure in the first fine zone, $p_r^l$, from
+the zone-centered pressure in the coarse zone just below the coarse-fine interface,
+$p_{\sfrac{r}{2}-1}^{l-1}$, we integrate in 2 steps.  We allow for a spatially
+changing gravitational acceleration, for complete generality.
+
+First we integrate up to the
+coarse-fine interface from the coarse-cell center as:
 \begin{equation}
-\frac{p_{r-\myhalf}^l - p_{\sfrac{r}{2}-1}^{l-1}}{\Delta r^{l-1}/2} = \frac{\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}}{2}g ~~~ \rightarrow ~~~ p_{r-\myhalf}^l = p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^{l-1} g}{4}\left(\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}\right).
+\frac{p_{r-\myhalf}^l - p_{\sfrac{r}{2}-1}^{l-1}}{\Delta r^{l-1}/2} = 
+  \frac{\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}}{2}  \,
+  \frac{g_{r-\myhalf}^l + g_{\sfrac{r}{2}-1}^{l-1}}{2} 
 \end{equation}
-Here is an equation to integrate from the coarse-fine interface up to
-the cell center
+We can rewrite this as an expression for the pressure at the coarse-fine interface:
 \begin{equation}
-\frac{p_r^l - p_{r-\myhalf}^l}{\Delta r^l/2} = \frac{\rho_r^l + \rho_{r-\myhalf}^l}{2}g ~~~ \rightarrow ~~~ p_r^l = p_{r-\myhalf}^l + \frac{\Delta r^l g}{4}\left(\rho_r^l + \rho_{r-\myhalf}^l\right).
+ p_{r-\myhalf}^l = p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^{l-1}}{8}
+  \left(\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}\right)
+  \left(g_{r-\myhalf}^l + g_{\sfrac{r}{2}-1}^{l-1}\right).
+  \label{eq:ctoi}
 \end{equation}
-Combining equations gives
+
+Next we integrate up from the coarse-fine interface to the fine-cell center:
 \begin{equation}
-p_r^l = p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^{l-1} g}{4}\left(\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}\right) + \frac{\Delta r^l g}{4}\left(\rho_r^l + \rho_{r-\myhalf}^l\right).
+\frac{p_r^l - p_{r-\myhalf}^l}{\Delta r^l/2} = 
+  \frac{\rho_r^l + \rho_{r-\myhalf}^l}{2} \,
+  \frac{g_r^l + g_{r-\myhalf}^l}{2}
 \end{equation}
+We can rewrite this as an expression for the pressure at the fine-cell center:
+\begin{equation}
+p_r^l = p_{r-\myhalf}^l + \frac{\Delta r^l g}{8}
+  \left(\rho_r^l + \rho_{r-\myhalf}^l\right)
+  \left(g_r^l + g_{r-\myhalf}^l\right).
+  \label{eq:itof}
+\end{equation}
+Combining equations \ref{eq:ctoi} and \ref{eq:itof} gives
+\begin{eqnarray}
+p_r^l = p_{\sfrac{r}{2}-1}^{l-1} &+& 
+     \frac{\Delta r^{l-1}}{8} \left(\rho_{r-\myhalf}^l + \rho_{\sfrac{r}{2}-1}^{l-1}\right)
+                                \left(   g_{r-\myhalf}^l +    g_{\sfrac{r}{2}-1}^{l-1}\right) \nonumber \\
+ &+& \frac{\Delta r^l}{8} \left(\rho_r^l + \rho_{r-\myhalf}^l\right)
+                            \left(   g_r^l +    g_{r-\myhalf}^l\right).
+\end{eqnarray}
 We can simplify using
@@
 \end{equation}
+and by interpolating the cell-centered densities to the coarse-fine interface as:
 \begin{equation}
@@
 \end{equation}
-Simplifying
+Because we carry both the cell- and edge-centered gravitational accelerations, we
+do not need to interpolate $g$ to the interface.
+Simplifying, we have
 \begin{eqnarray}
-p_r^l &=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{\Delta r^l g}{2}\left(\frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_{\sfrac{r}{2}-1}^{l-1}\right) + \frac{\Delta r^l g}{4}\left(\rho_r^l + \frac{2}{3}\rho_r^l + \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}\right) \nonumber \\
-&=& p_{\sfrac{r}{2}-1}^{l-1} + \frac{3\Delta r^l g}{4}\left(\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_r^l\right).\label{Coarse-Fine Stencil}
+p_r^l = p_{\sfrac{r}{2}-1}^{l-1} &+& 
+   \frac{\Delta r^l}{4}\left(\frac{2}{3}\rho_r^l + 
+                             \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1} + 
+                             \rho_{\sfrac{r}{2}-1}^{l-1}\right) 
+                       \left(   g_{r-\myhalf}^l +    g_{\sfrac{r}{2}-1}^{l-1}\right) \nonumber \\
+  &+& \frac{\Delta r^l g}{8}\left(\rho_r^l + 
+                                  \frac{2}{3}\rho_r^l + 
+                                  \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}\right) 
+                       \left(   g_r^l +    g_{r-\myhalf}^l\right) \enskip .
 \end{eqnarray}
+Finally, we note for constant $g$, this simplifies to:
+\begin{equation}
+p_r^l = p_{\sfrac{r}{2}-1}^{l-1} + 
+  \frac{3\Delta r^l g}{4}\left(\rho_{\sfrac{r}{2}-1}^{l-1} + \rho_r^l\right).\label{Coarse-Fine Stencil}
+\end{equation}
 When we encounter a change from fine to coarse, analogously

### 2010-09-03 docs/initial_models/initial_models.tex
fix
add fine-coarse interpolation

@@
    \frac{\Delta r^l}{4}\left(\frac{2}{3}\rho_r^l + 
-                             \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1} + 
-                             \rho_{\sfrac{r}{2}-1}^{l-1}\right) 
+                             \frac{4}{3}\rho_{\sfrac{r}{2}-1}^{l-1} \right)
                        \left(   g_{r-\myhalf}^l +    g_{\sfrac{r}{2}-1}^{l-1}\right) \nonumber \\
-  &+& \frac{\Delta r^l g}{8}\left(\rho_r^l + 
-                                  \frac{2}{3}\rho_r^l + 
+  &+& \frac{\Delta r^l g}{8}\left(\frac{5}{3}\rho_r^l + 
                                   \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}\right) 
@@
 \end{equation}
-When we encounter a change from fine to coarse, analogously
+
+When integrating across a fine-coarse interface (see Figure~\ref{fig:ftoc}), the proceduce is similar.
+\begin{figure}[t]
+\centering
+\includegraphics[width=2.5in,angle=90]{\figpath/ftoc}
+\caption{\label{fig:ftoc} A fine-coarse interface in the 1-d base state}
+\end{figure}
+The expression for general gravity becomes:
+\begin{eqnarray}
+p_{(r+1)/2}^{l-1} = p_{r}^l &+&
+   \frac{\Delta r^l}{4}\left(\frac{2}{3}\rho_r^l + 
+                             \frac{4}{3}\rho_{(r+1)/2}^{l-1} \right)
+                       \left(   g_{(r+1)/2 -\myhalf}^{l-1} +    g_{(r+1)/2}^{l-1}\right) \nonumber \\
+  &+& \frac{\Delta r^l g}{8}\left(\frac{5}{3}\rho_r^l + 
+                                  \frac{1}{3}\rho_{(r+1)/2}^{l-1}\right) 
+                       \left(   g_r^l +    g_{(r+1)/2 -\myhalf}^{l-1} \right) \enskip .
+\end{eqnarray}
+and for spatially-constant gravity, it simplifies to:
 \begin{equation}

### 2010-09-03 docs/initial_models/initial_models.tex
remove stray 'g' factors that Andy pointed out
some updates

@@
 
-MAESTRO needs to be given model data with the same resolution as the base state.
-(Note that for spherical problems we set 
-$\Delta r = \Delta x/\mathtt{drdxfac}$.)  
-We need to generate this from the raw data,
- $\rho^{\raw}, T^{\raw}, p^{\raw}$, and
-$X^{\raw}$.  Here is the raw data file for each test problem:
+We have found that for the best performance, the MAESTRO initialization procedure should be given model data with the same resolution as the base state resolution, $\Delta r$.
+For planar problems, $\Delta r = \Delta x$.
+For multilevel planar problems, we use $\Delta r$ equal to $\Delta x$ at the finest resolution.
+For spherical problems we set $\Delta r = \Delta x/\mathtt{drdxfac}$.
+We generate our initial model either analytically or from raw data,
+ $\rho^{\raw}, T^{\raw}, p^{\raw}$, and $X^{\raw}$.  
+Here is the raw data file for each test problem:
 \begin{eqnarray}
@@
 \begin{equation}
-\frac{p_r - p_{r-1}}{\Delta r} = \frac{\rho_r + \rho_{r-1}}{2}g,\label{HSE Discretization}
+\frac{p_r - p_{r-1}}{\Delta r} = \frac{\rho_r + \rho_{r-1}}{2}g_{r-\myhalf},\label{HSE Discretization}
 \end{equation}
@@
 
-\subsection{Multilevel Plane-Parallel}
-We do the same thing as the single-level case except that we generate
-model data with the same resolution as the finest base state
-resolution.
-
 \section{Creating the Initial Data from the Model Data}
@@
   init\_base\_state} (which is actually a terrible name since we
-are setting the initial data, not the base state), we set the initial
+are computing 1D initial data, which is not quite the same thing 
+as the base satte), we set the initiail
 data $\rho^{\init}, T^{\init}, p^{\init}$ and $X^{\init}$ equal to the
-model data.  Then, we set $h^{\init} =
-h(\rho^{\init},T^{\init},X^{\init})$ through the EOS.  Note that in the code,
-$p^{\init}$ is also overwritten, but the value of $p^{\init}$ does not
-change since we called $p^{\model} =
-p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the
-initial model generation routines in \S~\ref{Sec:Creating the Model Data from Raw
-  Data}.
+model data.  Then, we set $p^{\init},h^{\init} =
+p,h(\rho^{\init},T^{\init},X^{\init})$ through the EOS.  Note that 
+we overwrite the existing value of $p^{\init}$ but this should not change
+the value since we called 
+$p^{\model} = p^{\model}(\rho^{\model},T^{\model},X^{\model})$ at the end of the
+initial model generation routines in 
+\S~\ref{Sec:Creating the Model Data from Raw Data}.
 
-\subsection{Multilevel Plane-Parallel}
-We do the same thing as the single-level case except that we generate
+For multilevel planar problems, we generate
 initial data at the non-finest levels by using linear interpolation
@@
 \begin{equation}
-p_r^l = p_{r-\myhalf}^l + \frac{\Delta r^l g}{8}
+p_r^l = p_{r-\myhalf}^l + \frac{\Delta r^l}{8}
   \left(\rho_r^l + \rho_{r-\myhalf}^l\right)
@@
                        \left(   g_{r-\myhalf}^l +    g_{\sfrac{r}{2}-1}^{l-1}\right) \nonumber \\
-  &+& \frac{\Delta r^l g}{8}\left(\frac{5}{3}\rho_r^l + 
+  &+& \frac{\Delta r^l}{8}\left(\frac{5}{3}\rho_r^l + 
                                   \frac{1}{3}\rho_{\sfrac{r}{2}-1}^{l-1}\right) 
@@
                        \left(   g_{(r+1)/2 -\myhalf}^{l-1} +    g_{(r+1)/2}^{l-1}\right) \nonumber \\
-  &+& \frac{\Delta r^l g}{8}\left(\frac{5}{3}\rho_r^l + 
+  &+& \frac{\Delta r^l}{8}\left(\frac{5}{3}\rho_r^l + 
                                   \frac{1}{3}\rho_{(r+1)/2}^{l-1}\right) 

### 2010-09-03 docs/initial_models/initial_models.tex
finished corrections of initialization procedure.  text in this section can be moved elsewhere now.

@@
 are computing 1D initial data, which is not quite the same thing 
-as the base satte), we set the initiail
+as the base satte), we set the initial
 data $\rho^{\init}, T^{\init}, p^{\init}$ and $X^{\init}$ equal to the
@@
 
+We deal with the edge of the star by tracking the first coarse cell 
+in which the density falls below {\tt base\_cutoff\_density}.  We note 
+the radius of this cell center, and the values of $\rho$, $\rho h$, $X_k$, 
+$p$, and $T$ in this cells.  Then, at every level, if current cell-center 
+is above this radius, we set the state equal to this stored state.  This 
+ensures a consistent treatment of the edge of the star at all levels.
+
 \section{Creating the Base State and Full State}\label{Sec:Creating the Base State and Full State}
@@
   multifab to obtain the full state $\rho, h, X$, and $T$.
-\item If {\tt perturb\_model} = F, skip the remaining steps and
-  instead set $p_0 = p^{\init}$ and $\rho_0 = \rho^{\init}$.
-\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
+\item If {\tt perturb\_model} = T, a user-defined perturbation is
+  added.  This routine should make sure that the EOS is called so that
+  there is some sense of thermodynamic consistency.
 \item Set $\rho_0 = \overline\rho$.
-\item Compute $p_0$ using equation (\ref{HSE Discretization}), using
-  the $p^{\init}$ at the top of the domain as the one required boundary 
-  condition.  Note that we still integrate upwards from the bottom, but then
-  offset the final pressure so the pressure at the top of the domain
-  remains unchanged.
-\item Compute the full state $T,h = T,h(\rho,p_0,X)$.
+\item Compute $p_0$ using {\tt enforce\_HSE}.
+\item Compute $T,h = T,h(\rho,p_0,X_k)$.
 \item Set $(\rho h)_0 = \overline{(\rho h)}$.
+\item Compute $\overline{T}$.  Note that we only use $\overline{T}$ as
+  a diagnostic and as a seed for EOS calls.
 \end{enumerate}
@@
 
-\subsection{Multilevel Plane-Parallel}
-\begin{enumerate}
-\item Fill $\rho^{\init}, h^{\init}, X^{\init}$, and $T^{\init}$ onto
-  a multifab to obtain the full state $\rho, h, X$, and $T$.
-\item If {\tt perturb\_model} = F, skip the next two steps and instead
-  set $\rho_0 = \rho^{\init}$.
-\item Perturb $T$, then compute $\rho,h = \rho,h(p^{\init},T,X)$.
-\item Set $\rho_0 = \overline\rho$.
-\item Compute $p_0$ using equation (\ref{HSE Discretization}), using
-  the pressure at the top of the domain as the initial condition.
-  Note that we still integrate upwards from the bottom, but then
-  offset the final pressure so the pressure at the top of the domain
-  remains unchanged.  Use the coarse-fine interface discretizations
-  given in \S~\ref{Sec:Coarse-Fine HSE Discretization}.
-\item Compute the full state $T,h = T,h(\rho,p_0,X)$.
-\item Set $(\rho h)_0 = \overline{(\rho h)}$.
-\end{enumerate}
-Now $\rho_0 = \overline\rho$, the base state is in HSE, and the full
-state is thermodynamically consistent with $p_0$.
-
-\subsection{Coarse-Fine HSE Discretization}\label{Sec:Coarse-Fine HSE Discretization}
+\subsection{Coarse-Fine {\tt enforce\_HSE} Discretization}\label{Sec:Coarse-Fine HSE Discretization}
 When integrating the HSE discretization upward, we must use a different

### 2010-09-14 docs/lo_density/lo_density.tex
add a discussion of buoyancy_cutoff_density

@@
 
+The {\tt anelastic\_cutoff} is the density below which we modify the constraint.
+
 \begin{itemize}
@@
 \section{base\_cutoff\_density}\label{Sec:Base Cutoff Density}
+
+The {\tt base\_cutoff\_density} is the lowest density that we model.
 
@@
 
+The burning cutoff determines where we call the reaction network to
+get the nuclear energy generation rate and composition changes.  For
+densities below the burning cutoff, we do not call the network.
+
 \begin{itemize}
@@
 \end{itemize}
+
+\section{buoyancy\_cutoff\_factor}
+
+The {\tt buoyancy\_cutoff\_factor} is used to zero out the forcing terms
+to the velocity equation at low densities.
+
+\begin{itemize}
+
+\item In {\tt init\_base\_state} we print out the value of the
+   the density at which the buoyancy cutoff would take effect,
+   {\tt buoyancy\_cutoff\_factor * base\_cutoff\_density}.
+
+\item In {\tt mk\_vel\_force}, we zero out {\tt rhopert}, the
+   perturbational density used in computing the buoyancy force,
+   if $\rho < \mathtt{buoyancy\_cutoff\_factor * base\_cutoff\_density}$.
+
+\item In {\tt mk\_vel\_force}, for spherical problems, we 
+   zero out {\tt centrifugal\_term}, the centrifugal force for
+   rotating stars, if $\rho < \mathtt{buoyancy\_cutoff\_factor *
+   base\_cutoff\_density}$.
+
+\end{itemize}

### 2010-10-05 docs/beta0/beta0.tex
start of a document to describe our understanding of beta_0

The goal of $\beta_0$ is to capture the expansion of a displaced fluid
element due to the stratification of the atmosphere.  MAESTRO computes
$\beta_0$ as:
\begin{equation}
\beta_0(r,t) = \rho_0 \exp\left (  \int_0^r  \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial r^\prime} dr^\prime \right )
\end{equation}

\section{Constant Composition}
Consider an isentropically stratified atmosphere, with a constant
composition as a function of $r$.  If you displace a parcel of fluid,
it will expand adabatically and continue to rise until its density
matches the background density.  
Even if $\gammabar$ is not constant in $r$, following from the
definition of $\beta_0$,
\begin{equation}
\frac{1}{\beta_0} \frac{d\beta_0}{dr} = \frac{1}{\gammabar p_0} \frac{dp_0}{dr}
\end{equation}
and the definition of $\Gamma_1$,
\begin{equation}
\Gamma_1 = \left . \frac{d \log p}{d \log \rho} \right |_s
\end{equation}
So, at constant entropy, from the definition of $\Gamma_1$, it must hold
that
\begin{equation}
\frac{1}{\rho} \frac{d \rho}{dr} = \frac{1}{\Gamma_1 p} \frac{d p}{dz} \enskip .
\end{equation}
Comparing to the defintion of $\beta_0$ then
\begin{equation}
\frac{1}{\beta_0} \frac{\beta_0}{dr} =\frac{1}{\gammabar p_0}\frac{dp_0}{dr} = \frac{1}{\rho_0} \frac{d\rho_0}{dr}  \enskip .
\end{equation}
Therefore, $\beta_0 = \rho_0$.

This means that if we have a constant composition and an
isentropically stratified atmosphere, as we displace a fluid element,
it will always remain neutrally buoyant.

\section{Composition Gradient}

### 2010-10-05 docs/beta0/beta0.tex
more

@@
 Consider an isentropically stratified atmosphere, with a constant
-composition as a function of $r$.  If you displace a parcel of fluid,
-it will expand adabatically and continue to rise until its density
-matches the background density.  
-Even if $\gammabar$ is not constant in $r$, following from the
-definition of $\beta_0$,
+composition as a function of $r$.  If you displace a parcel of fluid
+upwards, it will expand adabatically and continue to rise until its
+density matches the background density.  Even if $\gammabar$ is not
+constant in $r$, following from the definition of $\beta_0$,
 \begin{equation}
@@
 \section{Composition Gradient}
+
+If there is a change in composition with $r$, the situation is more
+complicated.  Consider again an isentropically stratified atmosphere,
+now with a composition gradient.  If you displace a parcel of fluid
+upwards, it will rise.  If there are no processes that change the
+composition (e.g.\ reactions), then the composition in the fluid
+element will remain fixed.  As it rises, it will the ambient medium
+will have a different composition that it has.  In this case, what is
+the path to equilibrium?

### 2010-10-16 docs/architecture/architecture.tex
%ng -> nghost()
add some info on diag and compilation options

@@
 {\tt main.Linux.Intel.exe}, where the specific parts of 
-the name depend on the compilers and OS used.  Each problem
-will have one or more inputs files (for example, {\tt test2/inputs\_2d})
-which override the default values for one or more runtime parameters.
-The code is run simply as:
+the name depend on the compilers and OS used.  MAESTRO can be built
+with support for parallelism (using MPI or OpenMPI), or compiled
+as a serial code.  The top portion
+of the {\tt GNUmakefile} can be editted to choose the build options.
+Setting {\tt MPI := t} will build the code to use MPI for parallelization.
+Setting {\tt OMP := t} will use OpenMP.  If desired, these are used together,
+with MPI distributing the grids across the processors and within a 
+shared-memory node, OpenMP is used to allow many cores to operate
+on the same grid.  A final build option determines whether to build
+the code with support for debugging.  Setting {\tt NDEBUG := t} disables
+debugging features.
+
+Each problem will have one or more inputs files (for example, {\tt
+  test2/inputs\_2d}) which override the default values for one or more
+runtime parameters.  The code is run simply as:
 \begin{verbatim}
@@
 visit can be found at {\tt https://wci.llnl.gov/codes/visit/home.html}\,.
+
+\subsubsection{Diagnostic Files}
+
+By default, MAESTRO outputs global diagnostics each timestep into a file
+called {\tt maestro\_diag.out}.  This includes the maximum Mach number,
+peak temperature, and peak nuclear energy generation rate.  Individual problems
+can provide their own {\tt diag.f90} file to produce custom diagnostic
+output.
 
@@
 \begin{verbatim}
-    ng_sp = s(1)%ng
+    ng_sp = nghost(s(1))
 \end{verbatim}

### 2010-10-19 docs/FAQ/faq.tex
add some recent stuff

@@
   slightly, so the righthand side is non-zero.
+
+  The tolerances used for the various multigrid solves in the code
+  can be overridden on a problem-by-problem basis by putting a
+  copy of {\tt MAESTRO/Source/mg\_eps.f90} into the problem directory
+  and resetting the tolerances.  The role of each of these tolerances
+  is described in {\tt MAESTRO/docs/mg/}.
 
@@
 
+\item {\em Why is $\gammabar$ computed as the average of the full state
+       $\Gamma_1$ instead of computed from the base state density and 
+       pressure via the equation of state?}
+
+ The primary reason is that there is no base state composition.  The
+ base state density is simply the average of the full state density,
+ and the base state pressure is the pressure required for hydrostatic
+ equilibrium.  There is no thermodynamic relationship enforced between
+ these base state quantities.
+
+\item {\em Can I run a full star in 2-d axisymmetric geometry?}
+
+ No.  This is a design decision.  There is no support for axisymmetric
+ coordinates in MAESTRO.  Spherical problems must be run in 3-d.
+
 \end{enumerate}
+
+\section{Analysis}
+
+\begin{enumerate}
+
+\item {\em I want to open a plotfile, derive a new quantity from
+ the data stored there, and write out a new plotfile with this derived 
+ data.  How do I do this?}
+
+ One implementation of this can be found in {\tt
+   fParallel/data\_processing/tutorial/fwrite2d.f90}.  This reads in
+ the plotfile data using the {\tt plotfile\_module} that the {\tt
+   data\_processing} routines rely on, but then builds a multifab
+   and writes the data out to a plotfile using the BoxLib write
+   routines.
+
+\end{enumerate}

### 2010-10-25 docs/thermo_notes/thermo_notes.tex
add some notes on convective stability criteria

@@
 
+\subsection{Convective stability criterion}
+
+Here we look at the criterion for convective stability in the case of 
+non-uniform chemical composition.  This section follows Cox \& Giuli 
+\cite{cg-ed2} closely (see chapter 13).
+
+Consider a fluid parcel that gets displaced upwards (against gravity) from
+a radial location $r$ to $r + \Delta r$.  
+The parcel is stable to convection if the displaced parcel's density is 
+greater than 
+the surrounding fluid and gravity pushes the parcel back towards where it came
+from.  Then the criterion for stability should be
+\begin{eqnarray}
+ \rho_{parcel}(r+\Delta r) - \rho_{background}(r + \Delta r) &>& 0 \\
+ \bigg[\rho_{parcel}(r) + \bigg(\frac{d\rho}{dr}\bigg)_{parcel}\Delta r\bigg] - 
+ \bigg[\rho_{background}(r) + \bigg(\frac{d\rho}{dr}\bigg)_{background}\Delta r\bigg] &>& 0 
+\end{eqnarray}
+Since the parcel originates at r, $\rho_{parcel}(r) = \rho_{background}(r)$ and
+so the stability criterion is
+% \bigg [ \bigg(\frac{d\rho}{dr}\bigg)_{parcel} - \bigg(\frac{d\rho}{dr}\bigg)_{background} \bigg ] \Delta r &>& 0 \\ 
+\begin{equation}
+ \bigg(\frac{d\rho}{dr}\bigg)_{parcel} > \bigg(\frac{d\rho}{dr}\bigg)_{background}
+\label{eqn:basicStability}
+\end{equation}
+
+Since the total pressure, $P$, always increases inward in a star in hydrostatic
+equilibrium, we can use $P$ instead of $r$ as the independent radial variable.  
+Then condition for stability can be written as
+\[
+ \bigg( \frac{d \ln \rho}{d \ln P}\bigg )_{parcel} < \bigg(\frac{d \ln \rho}{d \ln P}\bigg)_{background}
+\]
+
+Using the equation of state $P = P( \rho, T, \bar{\mu})$, where 
+$\bar{\mu}$  is the average mass per molecule, we can write 
+\begin{equation}
+  d \ln P = \frac{\partial \ln P}{\partial \ln \rho} \bigg |_{T, \bar{\mu}}d \ln \rho + \frac{\partial \ln P}{\partial \ln T} \bigg |_{\rho, \bar{\mu}} d \ln T + \frac{\partial \ln P}{\partial \ln \bar{\mu}}\bigg |_{\rho, T} d \ln \bar{\mu}\
+\label{eqn:lnEOS}
+\end{equation}
+For convenience we introduce 
+%\begin{eqnarray}
+\[
+  \chi_{\rho} = \frac{\partial \ln P}{\partial \ln \rho}\bigg |_{T,\bar{\mu}} \qquad
+  \chi_T = \frac{\partial \ln P}{\partial \ln T} \bigg |_{\rho,\bar{\mu}} \qquad
+  \chi_{\bar{\mu}} = \frac{\partial \ln P}{\partial \ln \bar{\mu}} \bigg |_{\rho, T}
+\]
+%\end{eqnarray}
+Then we can rearrange \ref{eqn:lnEOS} to get
+\begin{equation}
+  \frac{d \ln \rho}{\partial \ln P} = \frac{1}{\chi_\rho} - 
+  \frac{\chi_T}{\chi_\rho} \frac{d \ln T}{d \ln P}- \frac{\chi_{\bar{\mu}}}{\chi_\rho}
+  \frac{d \ln \bar{\mu}}{d \ln P}
+\end{equation}
+%
+Then the general stability criterion is
+\begin{equation}
+  \bigg ( \frac{1}{\chi_\rho} - 
+  \frac{\chi_T}{\chi_\rho} \frac{d \ln T}{d \ln P}- \frac{\chi_{\bar{\mu}}}{\chi_\rho}
+  \frac{d \ln \bar{\mu}}{d \ln P} \bigg )_{parcel} < 
+  \bigg ( \frac{1}{\chi_\rho} - 
+  \frac{\chi_T}{\chi_\rho} \frac{d \ln T}{d \ln P}- \frac{\chi_{\bar{\mu}}}{\chi_\rho}
+  \frac{d \ln \bar{\mu}}{d \ln P} \bigg )_{background}
+\label{eqn:genStability}
+\end{equation}
+%
+Here's where various assumptions/simplifications get used.  
+\begin{enumerate}
+\item
+If no assumptions are made, you can't get any further than equation 
+(\ref{eqn:genStability}).  Even in view of an infinitesimally small initial 
+perturbation, you can't, in general, assume the $\chi$'s in parcel are the same 
+as the $\chi$'s in the background.  
+This applies in the case where nuclear reactions and/or ionization change the 
+composition of the parcel.  This case tends not to be of much interest for
+two reasons.  Either composition effects get incorporated implicitly through 
+assuming chemical equilibrium.  Or both of these terms can be neglected in the 
+rising parcel.  This would be justified if the timescale for reactions 
+is long compared 
+to the convective timescale, and either the same is ture for ionization or the 
+fluid is fully ionized. 
+\item
+If we assume that $\bar{\mu}$ remains constant in the parcel, then 
+$\frac{d \ln \bar{\mu}}{d \ln P}$ drops out for the parcel.  In this case,
+we can assume, in view of the arbitrarily small initial perturbation of 
+the parcel, that $\chi_\rho$ and $\chi_T$ to have the same values in the 
+parcel as in the background.  Then the stability criterion becomes
+\begin{equation}
+  \bigg (  \frac{d \ln T}{d \ln P} \bigg )_{parcel} > 
+  \bigg (  \frac{d \ln T}{d \ln P} + \frac{\chi_{\bar{\mu}}}{\chi_T}
+  \frac{d \ln \bar{\mu}}{d \ln P} \bigg )_{background}
+\label{eqn:Ledoux}
+\end{equation}
+The Ledoux stability criterion is obtained by assuming that the parcel moves
+adiabatically.
+\item
+If we assume   
+that the background is in chemical equilibrium and the parcel acheives 
+instantaneous chemical equilibrium, then $\bar{\mu} = \bar{\mu}(\rho,T)$ for
+the background and the parcel.  (Note that we aren't requiring constant 
+composition in the parcel here.)
+The effect of variable composition are then absorbed into $\chi_\rho$ and 
+$\chi_T$.  Again, we can take $\chi_\rho$ and $\chi_T$ to have the same values 
+in the parcel as in the background. The criterion then is 
+\begin{equation}
+\bigg ( \frac{d \ln T}{d \ln P} \bigg )_{parcel} > 
+  \bigg ( \frac{d \ln T}{d \ln P} \bigg )_{background}
+\label{eqn:Schwarz}
+\end{equation}
+We obtain the Schwarzchild criterion for 
+stability if we also assume the parcel moves adiabatically.
+
+The Scharwzchild criterion can be recast in terms of entropy if 
+the EOS is taken as $P(\rho, S)$ instead of $P(\rho, T)$.  Then, in place 
+of equation (\ref{eqn:lnEOS}) we have 
+\begin{equation}
+ d \rho = \frac{\partial \rho}{\partial P} \bigg |_{S} d P + \frac{\partial \rho}{\partial S} \bigg |_{P} dS
+\end{equation}
+We can subsitute this into equation (\ref{eqn:basicStability}) for stability, 
+and assuming the parcel moves adiabatically, we get 
+\begin{equation}
+  \bigg ( \frac{\partial \rho}{\partial S} \bigg |_{P} \frac{dS}{dr} 
+  \bigg )_{background}< 0
+\end{equation}
+One of Maxwell's relations is 
+\begin{equation}
+\frac{\partial \rho^{-1}}{\partial S} \bigg |_{P} = \frac{\partial T}{\partial P} \bigg |_{S}
+\end{equation}
+All thermodynamically stable substances have temperatures that increase upon 
+adiabatic compression, i.e. $\frac{\partial T}{\partial P} \big |_{S} > 0$.
+So Maxwell's relation implies that 
+$\frac{\partial \rho}{\partial S} \big |_{P} < 0$.
+The stability criterion then becomes
+\begin{equation}
+ \bigg ( \frac{d S}{d r} \bigg )_{background} > 0 
+\label{eqn:stabilityEntr}
+\end{equation}
+%
+\end{enumerate}
+
+Determining which stability criterion we want to enforce in creating the 
+initial model is complicated by the phenomenon of semiconvection, which  
+occurs when the Ledoux criterion is statisfied but the Schwarzchild is not, 
+i.e.
+\begin{equation}
+  \bigg (  \frac{d \ln T}{d \ln P} \bigg )_{parcel} < 
+  \bigg (  \frac{d \ln T}{d \ln P} \bigg )_{background} <
+  \bigg (  \frac{d \ln T}{d \ln P} \bigg )_{parcel} - 
+  \bigg ( \frac{\chi_{\bar{\mu}}}{\chi_T} 
+  \frac{d \ln \bar{\mu}}{d \ln P} \bigg )_{background}
+\end{equation}
+(Note that $\chi_{\bar{\mu}}$ is negative, as pressure is inversely proportional 
+to mass per particle, and $\frac{d \ln \bar{\mu}}{d \ln P}$ is positive, since
+nuclear reactions sythesize more massive particles in the center of the star.) 
+In this case, when a rising parcel eventually reaches neutral buoyancy, it will
+have a temperature excess in comparision to it's surroundings.  
+%(Assuming ideal gas, plus black body radiation pressure if wanted.)  
+% I think this qualifcation might be necessary, but want to think about
+% degeneracy pressure too.
+If the parcel can retain
+it's identitiy against diffusive mixing with the background long enough for 
+significant heat exhange to occur, then the parcel's temperature will drop, it 
+will contract increasing it's density, and the parcel will move inwards.
+The time scale of semiconvection is much longer than the timescale of 
+traditional convection.
+
+When we set up an initial model, we want to minimize any initial tendancy 
+towards convective motions, as we want these to be driven by the heating due
+to nuclear reactions,
+not the initial configuration we supply.  Thus I think we want to guard against
+semiconvection as well as ``traditional'' convection by using the stability 
+criterion
+\begin{equation}
+\bigg ( \frac{d \ln T}{d \ln P} \bigg )_{parcel} =
+\frac{d \ln T}{d \ln P} \bigg |_{S,\bar{\mu}} > 
+  \bigg ( \frac{d \ln T}{d \ln P} \bigg )_{background}
+\end{equation}
+Although this looks like the Schwarschild criterion (and, because I'm not 
+entirely sure on vocabulary, it might even be called the Schwarzchild 
+criterion), this does not simplify to equation (\ref{eqn:stabilityEntr})
+because we need to keep the explicit $\bar{\mu}$ dependence in the EOS.
+
+The question of whether we're in chemical equilibrium or not might be a moot
+point since our EOS (or any other part of the code) doesn't enforce chemical 
+equilibrium.  Thus, even  
+in the case of chemical equilbrium, we can't in general 
+drop the explicit $\bar{\mu}$ 
+dependence from our equations.  If we wanted to do that, then we would need
+$\bar{\mu}(\rho,T)$ to be substituted for $\bar{\mu}$ inside the EOS.
+
 \section{Adiabatic Excess}

### 2011-01-18 docs/visualization/visualization.tex
a brief introduction to visualizing MAESTRO data with yt

\section{Visualizing with \amrvis}
\section{Visualizing with \visit}
\section{Visualizing with \yt}
\yt is a Python package for analyzing and visualizing simulation data.
Originally \yt was written for the {\tt Enzo} code, with focus on
cosmological simulation data.  It has since been refactored to remove
the specific {\tt Enzo}isms, and now can work on various data formats,
such as the BoxLib data from {\tt MAESTRO} and {\tt CASTRO}, HDF5 data
from {\tt FLASH}, and data from the SPH code {\tt Gadget}.  For more
information, see the \yt homepage at {\tt http://yt.enzotools.org/}.

\subsection{Installing \yt}
The easiest way to obtain \yt is through the use of an installation script:
\begin{lstlisting}[language=bash,mathescape=false]
> wget http://hg.enzotools.org/yt/raw-file/stable/doc/install_script.sh
> bash install_script.sh
\end{lstlisting}
By default, this \yt install script will download and install, in its
own isolated environment, all the secondary utilities needed to get \yt
up and running.  Note that this currently includes installing {\it
  hdf5 1.6.9, zlib 1.2.3, bzip2 1.0.5, libpng 1.2.43, freetype 2.4.4,
  Python 2.6.3, numpy 1.5.1, matplotlib 1.0.0, mercurial 1.7.1,
  ipython 0.10, h5py 1.2.0, Cython 0.14, Forthon 0.8.4} as well as a
\yt {\it mercurial} bundle of changes.  You can turn off the automatic
installation of any of these particular packages by setting the
appropriate {\tt INST\_}* variable to zero in the install script;
you may have to then change some of *{\tt \_DIR} variables to point to
your own particular installation of that package.  I usually just let
\yt install its own stuff, which ensures things are working properly.

After the install script has finished, and assuming you let \yt install its own packages, you'll need to {\it prepend} some environment variables with \yt locations so that your system finds those first and stops looking.  For example, the \yt specific information in my {\tt .bashrc} file looks like this
\begin{lstlisting}[language=bash,mathescape=false]
# yt stuff
export YT_DEST=~/install/yt/current/yt-x86_64
export PATH=${YT_DEST}/bin/:$PATH
export PYTHONPATH=${YT_DEST}/lib/python2.6/site-packages/:$PYTHONPATH
export LD_LIBRARY_PATH=${YT_DEST}/lib/:$LD_LIBRARY_PATH
\end{lstlisting}
where I have {\tt ~/install/yt/current} as a soft link to my \yt
installation directory.

\subsection{Working with \yt}
The \yt installation provides both an interactive, {\it iPython}-like,
interface or the ability to import \yt modules for use in a batch
script.  The interactive interface should be in your {\tt \$PATH} if
you've followed the instructions in the previous section; to start it,
simply type {\tt iyt} in a terminal.
\begin{lstlisting}[language=Python]
> iyt

Welcome to yt!

In [1]: 
\end{lstlisting}
Codes like {\tt Enzo} use what are called {\it parameter files} to
describe the general information---number of levels, domain
dimensions, time, etc.---of a a dataset.  \yt likes to grab this
information before working on any specific data; this is accomplished
via the convenience method \textcolor{blue}{\tt load}:
\begin{lstlisting}[language=Python]
In [1]: pf = load(``plt00166'')
\end{lstlisting}
This method returns an instance of the {\tt StaticOutput} class.  One
of the easiest ways of handling plots is via a {\tt PlotCollection}
object
\begin{lstlisting}[language=Python]
In [2]: center = (pf.domain_right_edge + pf.domain_left_edge) / 2.0
In [3]: pc = PlotCollection(pf,center)
\end{lstlisting}
By default, the {\tt PlotCollection} constructor places the center of
the plot to be at the location of peak density.  Here we have
calculated the center of the domain by accessing the lower and upper
domain boundaries via the {\tt numpy} arrays {\tt
  pf.domain\_left\_edge} and {\tt pf.domain\_right\_edge},
respectively.  Note that up until this point, \yt has not actually
loaded the AMR dataset.

Now we would like to take some slices of tfromp in the dataset and
generate some 2-d plots.  To do this, we will use the {\tt
  PlotCollection}'s \textcolor{blue}{\tt add\_slice} method:
\begin{lstlisting}[language=Python]
In [4]: pc.add_slice(``tfromp'',0)
In [5]: pc.add_slice(``tfromp'',1)
In [6]: pc.add_slice(``tfromp'',2)
\end{lstlisting}
The first call to \textcolor{blue}{\tt add\_slice} builds an {\tt
  AMRHierarchy} object associated with {\tt pf}.  The {\tt
  AMRHierarchy} object contains information about the actual dataset,
such as its layout in the simulation domain or on disk.  Building the
hierarchy is expensive, but once it is built the data it contains can
be accessed via attributes and dictionary lookup.  In other words, the
subsequent \textcolor{blue}{\tt add\_slice} operations are relatively
cheap.  The first parameter to the \textcolor{blue}{\tt add\_slice}
method is obviously the variable we want; the second (optional)
parameter specifies an coordinate axis orthogonal to the slice being
made---0 for x, 1 for y, 2 for z.  Now we want to save the plots from
the {\tt PlotCollection}; this is done with the \textcolor{blue}{\tt
  save} method, which takes an optional basename for the generated files:
\begin{lstlisting}[language=Python]
In [7]: pc.save(``my_cool_images'')
\end{lstlisting}
This generates the following image files:
\begin{lstlisting}[language=Python]
Out[7]: 
['my_cool_images_Slice_x_tfromp.png',
 'my_cool_images_Slice_y_tfromp.png',
 'my_cool_images_Slice_z_tfromp.png']
\end{lstlisting}
Figure \ref{fig:yt_slice} shows an example of one of the slice images.
Note that this was a quick and dirty generation of the image---there
is a lot of space around the figure, which can be removed with various
options to the \yt methods.  Also, the user can specify derived
variables, log-scaling, annotations, etc. For more information see the
documentation at {\tt http://yt.enzotools.org/doc/}.

\begin{figure}[!h]\label{fig:yt_slice}
\centering
\includegraphics[height=0.4\textheight]{my_cool_images_Slice_x_tfromp_trim}
\caption{Example slice through 3-d dataset with \yt.}
\end{figure}

When writing a script to use in batch mode, one has to manually import
the import modules needed to work with \yt.  As such, all scripts must
import from the {\tt yt.mods} module, which is essentially a
convenience module that sets up the appropriate \yt namespace.  
%% As an aside, if you want to import a module as another variable in
%% the local namespace---e.g. {\tt import this.module as
%% local\_variable}---then this should be done via the various API's
%% for that particular piece of code.  For example,
For completeness, below is a script containing our example above.
\newpage
\begin{lstlisting}[language=Python]
# load our namespace
from yt.mods import *

# the plotfile I'm interested in
fn = ``plt00166''

# load it into a StaticOutput object
pf = load(fn)

# find the center of the domain
center = (pf.domain_right_edge + pf.domain_left_edge) / 2.0

# associate a PlotCollection with the pf
pc = PlotCollection(pf,center)

# add some slices of tfromp
pc.add_slice(``tfromp'',0)
pc.add_slice(``tfromp'',1)
pc.add_slice(``tfromp'',2)

# save our plots to a files with basename
# ``my_cool_images''
pc.save(``my_cool_images'')
\end{lstlisting}

\subsection{Volume Rendering}

### 2011-01-20 docs/runtime_parameters/runtime_parameters.tex
update

@@
 \verb=  bcz_lo  = &   $-z$ boundary condition  &  SLIP\_WALL \\
+\verb=  bds_type  = &   0 = use ppm instead for multi-d integrator \newline 1 = bilinear  &  0 \\
 \verb=  beta_type  = &   what type of coefficient to use inside the velocity divergence constraint.
@@
 &  1 \\
+\verb=  blocking_factor  = &   grids must be an integer multiple of this number, if possible  &  8 \\
 \verb=  buoyancy_cutoff_factor  = &   The multiplicative factor (over base\_cutoff\_density) below which we do zero out the buoyancy term in the momentum equation.  &  5.0 \\
@@
 \verb=  chk_int  = &   Number of timesteps between writing a checkpoint file  &  0 \\
-\verb=  cluster_min_width  = &   The minimum size on a side for a grid created using make\_new\_grids.  &  16 \\
 \verb=  co_latitude  = &   latitude, in radians, for problems with rotation where the domain is only a subset of a full star.  &  ZERO \\
@@
 \verb=  drdxfac  = &   ratio of radial base state zones to Cartesian full state zones for spherical geometry  &  1 \\
+\verb=  drive_initial_convection  = &   freeze the temperature used in the reaction network to the initial value.
+This is useful for developing an initial convective field to carry away the energy, while preventing the reactions from going nonlinear.
+&  .false.
+\\
+\verb=  dump_grid_file  = &    &  .false. \\
 \verb=  enthalpy_pred_type  = &   predict\_rhohprime = 1; \newline predict\_h = 2; \newline predict\_T\_then\_rhohprime = 3; \newline predict\_T\_then\_h = 4; \newline predict\_hprime = 5; \newline predict\_Tprime\_then\_h = 6.
@@
 \verb=  evolve_base_state  = &   turn on (.true.) or off (.false.) basestate evolution  &  .true. \\
+\verb=  fix_base_state  = &   if true, don't call average to reset the base state at all, even during initialization  &  .false. \\
 \verb=  fixed_dt  = &   Fix the time step.  If -1.0, then use the standard time step.  &  -1.0d0 \\
@@
 \\
+\verb=  mach_max_abort  = &   maximum mach number before the code aborts  &  -1.d0 \\
+\verb=  max_dt  = &   This is the maximum dt that is allowed  &  1.0d33 \\
 \verb=  max_dt_growth  = &   The maximum scale factor that the time step is allowed to grow by per time step.  &  1.1d0 \\
 \verb=  max_grid_size  = &   The largest grid size that will be created using make\_new\_grids.  &  64 \\
-\verb=  max_grid_size_base  = &   The largest grid size that will be created using make\_new\_grids for the coarsest level.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
+\verb=  max_grid_size_1  = &   The largest grid size that will be created using make\_new\_grids for the coarsest level.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
+\verb=  max_grid_size_2  = &   The largest grid size that will be created using make\_new\_grids for level 2 (the first refined level).  Defaults to {\tt max\_grid\_size}.  &  -1 \\
+\verb=  max_grid_size_3  = &   The largest grid size that will be created using make\_new\_grids for level 3 (the second refined level) and beyond.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
 \verb=  max_levs  = &   Total number of levels.  1 = single level.  &  1 \\
@@
 \verb=  mg_verbose  = &   Verbsoity of the multigrid solver, but not the bottom solver.  &  0 \\
-\verb=  min_eff  = &   parameter for cluster algorithm for making new grids in adaptive problems  &  0.7d0 \\
+\verb=  min_eff  = &   parameter for cluster algorithm for making new grids in adaptive problems  &  0.9d0 \\
+\verb=  minwidth  = &   The minimum size on a side for a grid created using make\_new\_grids.  &  8 \\
 \verb=  model_file  = &   input model file  &  "model.hse" \\
@@
 \verb=  n_cellz  = &   Number of cells for the base level in the z-direction  &  -1 \\
+\verb=  nuclear_dt_fac  = &   If $T_{max}^n > T_{max}^{n-1}$ set the new dt =   min[dt, dt*{\tt nuclear\_dt\_fac}*( $T_{max}^{n-1}$ / $(T_{max}^n-T_{max}^{n-1})$ ) ] not checkpoint-compatible yet since it wouldn't be backwards compatible  &  -1.0d0 \\
 \verb=  octant  = &   set octant = T if you just want to model an octant of a sphere (note: only takes effect for spherical geometry)  &  .false. \\
@@
 \verb=  plot_Hext  = &   plot external heating (Hext) in plotfile  &  .false. \\
+\verb=  plot_Hnuc  = &   plot nuclear energy generation rate (Hnuc) in plotfile  &  .true. \\
+\verb=  plot_ad_excess  = &   plot the adiabatic excess  &  .false. \\
 \verb=  plot_base  = &   plot w0\_x, w0\_y, w0\_z, divw0, rho0, rhoh0, h0, and p0 in plotfile  &  .false. \\
 \verb=  plot_base_name  = &   prefix to use in plotfile file names  &  "plt" \\
-\verb=  plot_deltat  = &   rather than use a plot interva, plot a file after the solution has advanced past plot\_deltat in time  &  0.d0 \\
+\verb=  plot_cs  = &   plot soundspeed  &  .false. \\
+\verb=  plot_deltat  = &   rather than use a plot interval, plot a file after the solution has advanced past plot\_deltat in time  &  -1.d0 \\
 \verb=  plot_eta  = &   plot $\etarho$ in plotfile  &  .false. \\
+\verb=  plot_gpi  = &   plot pi and grad(pi)  &  .true. \\
+\verb=  plot_h_with_use_tfromp  = &   Turn on storing of enthalpy-based quantities in the plotfile when we are running with {\tt use\_tfromp}  &  .true. \\
 \verb=  plot_int  = &   plot interval  &  0 \\
@@
 \verb=  plot_spec  = &   plot species and omegadot in plotfile  &  .true. \\
+\verb=  plot_sponge_fdamp  = &   plot fdamp rather than sponge assumes sponge has the form 1/(1+dt*{\tt sponge\_kappa}*fdamp)  &  .false. \\
 \verb=  plot_trac  = &   plot tracers in plotfile  &  .false. \\
-\verb=  ppm_type  = &   0 = no ppm \newline 1 = 1985 ppm \newline 2 = 2009 ppm  &  1 \\
+\verb=  ppm_type  = &   0 = no ppm \newline 1 = 1985 ppm \newline 2 = Hybrid Sekora/Colella McCorquodale Colella 2009/2010 ppm  &  1 \\
 \verb=  prob_hi_x  = &   physical coordinates of hi-x corner of problem domain  &  1.d0 \\
@@
 \verb=  restart_into_finer  = &   restart and add a level of refinement  &  .false. \\
+\verb=  restart_with_vel_field  = &   restart the simulation using a result from a {\tt drive\_initial\_convection} = T run note that this uses the restart variable to specify which file to restart from.
+After reading in the velocity information from the restart file, the time and timestep number are zeroed.
+&  .false.
+\\
 \verb=  rotation_radius  = &   radius used for computing centrifugal term in rotation problems  &  1.0d6 \\

### 2011-01-20 docs/runtime_parameters/rp.py
slight tweaking

@@ -5,6 +5,10 @@ import sys
 header=r"""
 \section{Runtime Parameters}
 
+Table~\ref{table:runtime} lists the runtime parameters available to
+all MAESTRO problems.  Problem-specific runtime parameters are not
+shown here.
+
 %%%%%%%%%%%%%%%%
 % symbol table
 %%%%%%%%%%%%%%%%

### 2011-01-28 docs/architecture/architecture.tex
add some notes on ghostcells

@@
 
+Ghostcells are filled through a variety of different routines, depending
+on the objective.
+
+\begin{itemize}
+
+\item {\tt multifab\_fill\_boundary} fills ghost cells for two
+  adjacent grids at the same level, which als includes periodic domain
+  boundary ghost cells.
+
+\item {\tt multifab\_physbc} fills ghostcells at the physical boundaries.
+
+\item {\tt multifab\_fill\_ghost\_cells} is used for multilevel
+  problems, and fills ghostcells in the finer grid (level {\tt n}) by
+  interpolating from data in the coarser grid (level {\tt n-1}).
+
+  This call is usually preceded by a call to {\tt
+    ml\_cc\_restriction\_c} which sets the level {\tt n-1} data to be
+  the average of the level {\tt n} data covering it.
+
+\end{itemize}
+
 \section{Boundary Conditions}

### 2011-02-03 docs/architecture/architecture.tex
updates to ghost cell fill stuff
minor fixes

@@
 the name depend on the compilers and OS used.  MAESTRO can be built
-with support for parallelism (using MPI or OpenMPI), or compiled
+with support for parallelism (using MPI or MPI+OpenMP), or compiled
 as a serial code.  The top portion
@@
 Setting {\tt MPI := t} will build the code to use MPI for parallelization.
-Setting {\tt OMP := t} will use OpenMP.  If desired, these are used together,
+Setting {\tt OMP := t} will use OpenMP.  OpenMP is used together with MPI, 
 with MPI distributing the grids across the processors and within a 
-shared-memory node, OpenMP is used to allow many cores to operate
+shared-memory node, OpenMP allows many cores to operate
 on the same grid.  A final build option determines whether to build
@@
 \begin{verbatim}
-  main.Linux.Intel.exe inputs_2d
-\end{verbatim}
+  ./main.Linux.Intel.exe inputs_2d
+\end{verbatim}
+for a serial executable, and, depending on your MPI environment, something like:
+\begin{verbatim}
+  mpiexec -n 16 ./main.Linux.Intel.mpi.exe inputs_2d
+\end{verbatim}
+for the parallel code.
+
 As the code runs, it will output both plotfiles and checkpoints.
@@
   interpolating from data in the coarser grid (level {\tt n-1}).
-
-  This call is usually preceded by a call to {\tt
-    ml\_cc\_restriction\_c} which sets the level {\tt n-1} data to be
+  This function, by default, will also call {\tt multifab\_fill\_boundary}
+  and {\tt multifab\_physbc} for both levels {\tt n} and {\tt n-1} (you 
+  can override this behavior for speed optimization purposes).
+  This call is usually preceded by a call to 
+  {\tt ml\_cc\_restriction\_c} which sets the level {\tt n-1} data to be
   the average of the level {\tt n} data covering it.

### 2011-02-03 docs/architecture/architecture.tex
updates to the directory structure discussion

@@
  \begin{itemize}
+ \item {\tt conductivity/}
+
+ Various routines for computing the thermal conductivity used in the
+ thermal diffusion part of the algorithm.
+
  \item {\tt constants/}
@@
  gamma\_law\_general/}.
+
+ \item {\tt model\_parser/}
+
+ A simple Fortran module for reading in 1-d initial model files.  This
+ is used by the initialization routines to get the initial model data.
 
@@
    Each problem in MAESTRO gets it own subdirectory.  The {\tt
-   GNUmakefile} includes the instructions on how to build the
-   executable, including what modules in {\tt extern} are used.
+     GNUmakefile} in the problem directory includes the instructions
+   on how to build the executable, including what modules in {\tt
+     extern} are used.
 
@@
    MAESTRO routines (e.g.\ initial conditions via {\tt initdata.f90}.
-   See \S~\ref{sec:makefile} for details on the build system.
+   See \S~\ref{sec:makefile} for details on the build system).
 

### 2011-02-03 docs/architecture/architecture.tex
lots of reorganization/clean-ups

@@
 
+\section{MAESTRO Basics}
+
+MAESTRO models problems that are in tight hydrostatic equilibrium.
+The fluid state is decomposed into a 1-d radial base state that
+describes the hydrostatic structure of the star or atmosphere, and a
+2- or 3-d Cartesian full state, that captures the departures from
+hydrostatic equilibrium.  Two basic geometries are allowed.  A {\em plane-parallel}
+geometry assumes that the domain is thin compared to the radius of curvature
+of the star, and therefore the 1-d base state is perfectly aligned with
+the Cartesian state.  A {\em spherical} geometry is for modeling an entire star.  Here,
+the 1-d base state is not aligned with the Cartesian state.  Figure~\ref{fig:base_state}
+shows this geometry.
+
+\begin{figure}[tb]
+\centering
+\includegraphics[height=2.0in]{base_grid} \hspace{0.5in}
+\includegraphics[height=2.0in]{base_spherical}
+\caption{\label{Fig:base_state} (Left) For multi-level problems
+  in planar geometry, we force a direct alignment between the radial
+  array cell centers and the Cartesian grid cell centers by allowing
+  the radial base state spacing to change with space and time.
+  (Right) For multi-level problems in spherical geometry, since there
+  is no direct alignment between the radial array cell centers and
+  the Cartesian grid cell centers, we choose to fix the radial base
+  state spacing across levels. Figure taken from~\cite{multilevel}}
+\end{figure}
+
+MAESTRO can use adaptive mesh refinement to focus resolution on
+complex regions of flow.  For Cartesian/plane-parallel geometries, all
+cells at the same height must be at the same level of refinement.
+This restriction is to allow for the base state to directly align with
+the Cartesian state everywhere.  For spherical geometries, there is no
+such restriction (again, see Figure~\ref{Fig:base_state}).
+
 \section{Getting Started}
 
-In this section we describe some of the standard problems that come
-with MAESTRO, how to run the code, some basic runtime parameters, and
-how to look at the output.
+In this section we give an overview of MAESTRO, including some of the
+standard problems, how to run the code, some basic runtime parameters,
+and how to look at the output.
 
@@
 
-As the code runs, it will output both plotfiles and checkpoints.
-By default, the plotfiles will be named {\tt plt}{\em nnnnn}, where
-the number {\em nnnnn} is the timestep number when the file was
-outputted.  Similarly, the checkpoints are named {\tt chk}{\em
-  nnnnn}.  BoxLib plotfiles and checkpoints are actually directories,
-with the data stored in subdirectories grouped by refinement level.
-Details of the simulation (number of processors used, output date,
-output directory, runtime parameter values, ...) are stored in
-the {\tt job\_info} file in each plotfile directory.
+As the code runs, it will output both plotfiles and checkpoints as
+well as one of more text diagnostic files (like {\tt
+  maestro\_diag.out}) with integral or extrema information (like
+maximum Mach number) from each timestep.  By default, the plotfiles
+will be named {\tt plt}{\em nnnnn}, where the number {\em nnnnn} is
+the timestep number when the file was outputted.  Similarly, the
+checkpoints are named {\tt chk}{\em nnnnn}.  BoxLib plotfiles and
+checkpoints are actually directories, with the data stored in
+subdirectories grouped by refinement level.  Details of the simulation
+(build information, number of processors used, output date, output
+directory, runtime parameter values, ...) are stored in the {\tt
+  job\_info} file in each plotfile directory.
 
@@
 \item {\tt max\_levs } is the maximum number of grid levels in the AMR
-  hierarchy to us.  {\tt max\_levs = 1} indicates running with only a
+  hierarchy to use.  {\tt max\_levs = 1} indicates running with only a
   single level spanning the whole domain.
@@
   coordinate direction, as measured in terms of number of cells.
+
+  For multilevel problems, the parameter {\tt max\_grid\_size\_1}
+  controls the maximum extent on level 1 (the base grid), {\tt
+    max\_grid\_size\_2} controls the maximum extent on level 2, and
+  {\tt max\_grid\_size\_3} controls the maximum extent on levels 3 and
+  higher.
 
@@
 
-\subsection{Refinement Criteria}
-
 \subsection{Working with the Output}
@@
 A basic {\tt GNUmakefile} begins with:
-\begin{verbatim}
+\begin{lstlisting}[language=bash,mathescape=false]
   NDEBUG := t
@@
   OMP    :=
-\end{verbatim}
-Here, {\tt NDEBUG} is true is we are building an optimized executable.
+\end{lstlisting}
+Here, {\tt NDEBUG} is true if we are building an optimized executable.
 Otherwise, the debug version is built---this typically uses less
@@
 {\tt MPI} and {\tt OMP} are set to true if we want to use either MPI
-or OpenMP for parallelization.
-
+or OpenMP for parallelization.  If {\tt MPI := t}, you will need to
+have the MPI libraries installed, and their location may need to be 
+specified in {\tt MAESTRO/mk/GMakeMPI.mak}.
+
+The next line sets the compiler to be used for compilation:
+\begin{lstlisting}[language=bash,mathescape=false]
+  COMP := Intel
+\end{lstlisting}
 The MAESTRO build system knows what options to use for various
 compiler families.  The {\tt COMP} flag specifies which compiler to
-us.  Allowed values include {\tt Intel}, {\tt gfortran}, {\tt PGI},
+use.  Allowed values include {\tt Intel}, {\tt gfortran}, {\tt PGI},
 and {\tt PathScale}.  The specific details of these choices are
-defined in the {\tt MAESTRO/mk/} directory.
-\begin{verbatim}
-  COMP := Intel
-  MKVERBOSE := t
-\end{verbatim}
+defined in the {\tt MAESTRO/mk/comps/} directory.
+
 {\tt MKVERBOSE} set to true will echo the build commands to the
 terminal as the are executed.
+\begin{lstlisting}[language=bash,mathescape=false]
+  MKVERBOSE := t
+\end{lstlisting}
 
@@
 flags).
-\begin{verbatim}
+\begin{lstlisting}[language=bash,mathescape=false]
   FPARALLEL := ../..
@@
   include $(FPARALLEL)/mk/GMakedefs.mak
-\end{verbatim}
+\end{lstlisting}
+%$
 
@@
 to define the EOS, reaction network, and conductivities to
-use, as well as any extra, problem specific files.  
-\begin{verbatim}
+use, as well as any extra, problem-specific files.  
+\begin{lstlisting}[language=bash,mathescape=false]
 EOS_DIR := extern/EOS/helmeos   
-CONDUCTIVITY_DIR := extern/EOS/conducteos 
-NETWORK_DIR := extern/networks/ignition_chamulak
+CONDUCTIVITY_DIR := extern/conductivity/constant
+NETWORK_DIR := extern/networks/ignition
 
 EXTRA_DIR := extern/random
-\end{verbatim}
+\end{lstlisting}
 Generally, one does not need to include the problem directory itself
@@
 files placed there override those with the same name found elsewhere
-in the source tree.  Some packages (for instance, the {\tt helmeos}
+in the source tree.
+
+Some packages (for instance, the {\tt helmeos}
 EOS) require Fortran include files.  The {\tt Fmincludes} variable
@@
   MAESTRO/\_parameters}).  Sometimes a problem defines their own
-runtime parameters (see \S~\ref{sec:def_runtime_param}), in which
+runtime parameters (for instance {\tt \_parameters.myproblem}), in which
 case, its parameter file should also appear in {\tt
-  PROBIN\_PARAMETERS}.
-\begin{verbatim}
+  PROBIN\_PARAMETERS} (see \S~\ref{sec:def_runtime_param}).
+\begin{lstlisting}[language=bash,mathescape=false]
   PROBIN_PARAMETERS := ../_parameters
-\end{verbatim}
-
-The final line in the {\tt GNUmakefile} include the rules to actually
+\end{lstlisting}
+
+The final line in the {\tt GNUmakefile} includes the rules to actually
 build the executable.
-\begin{verbatim}
+\begin{lstlisting}[language=bash,mathescape=false]
   include $(FPARALLEL)/MAESTRO/GMaestro.mak
-\end{verbatim}
+\end{lstlisting}
+%$
+
+\subsubsection{Handling Problem-Specific Source Files}
 
@@
 no file with the same name elsewhere in the source tree) need to be
-listed in a file {\tt GPackage.mak} in the problem directory.
+listed in a file {\tt GPackage.mak} in the problem directory, and
+the problem-directory needs to be explicitly listed in the {\tt EXTRA\_DIR}
+list in the {\tt GNUmakefile}.
 
@@
 
-MAESTRO initial models are read in by the {\tt init\_base\_state}
-routine.  This expects the initial model to contain a header giving
+MAESTRO initial models are read in by the {\tt extern/model\_parser}
+routines.  This expects the initial model to contain a header giving
 the number of variables and their names, followed by rows of data
@@
 structure is not appropriate for the fancy bottom solver.
+
+\section{Multilevel and Refinement Criteria}
+
+\section{Particles}

### 2011-02-03 docs/architecture/architecture.tex
more making things pretty

@@
 A basic {\tt GNUmakefile} begins with:
-\begin{lstlisting}[language=bash,mathescape=false]
+\begin{lstlisting}[language={[gnu]make},mathescape=false]
   NDEBUG := t
@@
 The next line sets the compiler to be used for compilation:
-\begin{lstlisting}[language=bash,mathescape=false]
+\begin{lstlisting}[language={[gnu]make},mathescape=false]
   COMP := Intel
@@
 terminal as the are executed.
-\begin{lstlisting}[language=bash,mathescape=false]
+\begin{lstlisting}[language={[gnu]make},mathescape=false]
   MKVERBOSE := t
@@
 flags).
-\begin{lstlisting}[language=bash,mathescape=false]
+\begin{lstlisting}[language={[gnu]make},mathescape=false]
   FPARALLEL := ../..
@@
 use, as well as any extra, problem-specific files.  
-\begin{lstlisting}[language=bash,mathescape=false]
+\begin{lstlisting}[language={[gnu]make},mathescape=false]
 EOS_DIR := extern/EOS/helmeos   
@@
   PROBIN\_PARAMETERS} (see \S~\ref{sec:def_runtime_param}).
-\begin{lstlisting}[language=bash,mathescape=false]
+\begin{lstlisting}[language={[gnu]make},mathescape=false]
   PROBIN_PARAMETERS := ../_parameters
@@
 build the executable.
-\begin{lstlisting}[language=bash,mathescape=false]
+\begin{lstlisting}[language={[gnu]make},mathescape=false]
   include $(FPARALLEL)/MAESTRO/GMaestro.mak
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
   subroutine example(s,dx,dt)
@@
     use variables, only: rho_comp
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
     type(multifab) , intent(inout) :: s(:)
     real(kind=dp_t), intent(in   ) :: dx(:,:),dt
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
     ! Local variables
@@
     integer :: lo(dm),hi(dm)
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
     ng_sp = nghost(s(1))
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
     do n=1,nlevs
        do i = 1, s(n)%nboxes
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
           if ( multifab_remote(s(n), i) ) cycle
@@
           hi =  upb(get_box(s(n), i))
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
           select case (dm)
@@
   end subroutine example
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
   subroutine example_2d(density,ng_sp,lo,hi,dx,dt)
@@
   end subroutine example_2d
-\end{verbatim}
+\end{lstlisting}
 
@@
 declared as 
-\begin{verbatim}
-  density(lo(1)-ng_sp:,lo(2)-ng_sp:,lo(3)-ng_sp)
-\end{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
+  density(lo(1)-ng_sp:,lo(2)-ng_sp:,lo(3)-ng_sp:)
+\end{lstlisting}
 and an additional loop over the `z' coordinate (from {\tt lo(3)} to

### 2011-02-03 docs/architecture/architecture.tex
add more on base state arrays

@@
 \includegraphics[height=2.0in]{base_spherical}
-\caption{\label{Fig:base_state} (Left) For multi-level problems
+\caption{\label{fig:base_state} (Left) For multi-level problems
   in planar geometry, we force a direct alignment between the radial
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 nodal in its respective direction:
-\begin{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
   type(multifab) :: umac(dm)
@@
   enddo
-\end{verbatim}
+\end{lstlisting}
 
@@
 $h_0$, the base state enthalpy.
+
+The base state arrays are 2-dimensional, with the first dimension
+giving the level in the AMR hierarchy and the second the radial index
+into the base state.  For spherical geometries, the base state only
+exists at a single level, so the first index will always be 1.  The
+radial index is 0-based, to be consistent with the indexing for the
+Cartesian state data.  For example, the base state density would be
+dimensioned: {\tt rho0(nlevs,0:nr\_fine-1)}.  Here, {\tt nlevs} is the
+number of levels of refinement and {\tt nr\_fine} is the number of
+cells in the radial direction at the finest level of refinement.
+
+For multilevel, plane-parallel geometry, all grids at the same height
+will have the same resolution so that the full state data is always
+aligned with the base state (see Figure~\ref{fig:base_state}).  Base
+state data on coarse grids that are covered by fine grids is not
+guaranteed to be valid.
+
+For spherical problems, the base state resolution, $\Delta r$, is
+generally picked to be finer than the Cartesian grid resolution,
+$\Delta x$, i.e.\ $\Delta r < \Delta x$.  The ratio is controlled
+by the parameter {\tt drdxfac}.
 
@@
 in the allowed range of {\tt 1:nspec}, you do the following query:
-\begin{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
   ic12 = network_species_index("carbon-12")
-\end{verbatim}
+\end{lstlisting}
 If the resulting index is {\tt -1}, then the requested species was not

### 2011-02-07 docs/architecture/architecture.tex
get the whole thing building again
more updates throughout

@@
+\section{MAESTRO Basics}
+
+MAESTRO models problems that are in tight hydrostatic equilibrium.
+The fluid state is decomposed into a 1-d radial base state that
+describes the hydrostatic structure of the star or atmosphere, and a
+2- or 3-d Cartesian full state, that captures the departures from
+hydrostatic equilibrium.  Two basic geometries are allowed.  A {\em
+  plane-parallel} geometry assumes that the domain is thin compared to
+the radius of curvature of the star, and therefore the 1-d base state
+is perfectly aligned with the Cartesian state.  A {\em spherical}
+geometry is for modeling an entire star.  Here, the 1-d base state is
+not aligned with the Cartesian state.  Figure~\ref{fig:base_state}
+shows these geometries.
+
+\begin{figure}[tb]
+\centering
+\includegraphics[height=2.0in]{\archfigpath/base_grid} \hspace{0.5in}
+\includegraphics[height=2.0in]{\archfigpath/base_spherical}
+\caption{\label{fig:base_state} MAESTRO geometries, showing both the
+  1-d base state and the full Cartesian state.  (Left) For multi-level
+  problems in planar geometry, we force a direct alignment between the
+  radial array cell centers and the Cartesian grid cell centers by
+  allowing the radial base state spacing to change with space and
+  time.  (Right) For multi-level problems in spherical geometry, since
+  there is no direct alignment between the radial array cell centers
+  and the Cartesian grid cell centers, we choose to fix the radial
+  base state spacing across levels. Figure taken
+  from~\cite{multilevel}.}
+\end{figure}
+
+MAESTRO can use adaptive mesh refinement to focus resolution on
+complex regions of flow.  For Cartesian/plane-parallel geometries, all
+cells at the same height must be at the same level of refinement.
+This restriction is to allow for the base state to directly align with
+the Cartesian state everywhere.  For spherical geometries, there is no
+such restriction (again, see Figure~\ref{fig:base_state}).
+
 \section{MAESTRO Directory Structure}
 
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 All the files needed to build MAESTRO are contained in the {\tt
-fParallel} directory structure.  Briefly, the sub-directories
-are:
+  fParallel} directory structure.  Problems piece together various
+MAESTRO directories, choosing a reaction network, equation of state,
+and conductivity routine to build an executable.  Briefly, the
+sub-directories are:
 \begin{itemize}
@@
 
-  The multigrid solver.
+  The multigrid solver, with support for both cell-centered and
+  node-centered data.
 
@@
 \end{itemize}
-
-\section{MAESTRO Basics}
-
-MAESTRO models problems that are in tight hydrostatic equilibrium.
-The fluid state is decomposed into a 1-d radial base state that
-describes the hydrostatic structure of the star or atmosphere, and a
-2- or 3-d Cartesian full state, that captures the departures from
-hydrostatic equilibrium.  Two basic geometries are allowed.  A {\em plane-parallel}
-geometry assumes that the domain is thin compared to the radius of curvature
-of the star, and therefore the 1-d base state is perfectly aligned with
-the Cartesian state.  A {\em spherical} geometry is for modeling an entire star.  Here,
-the 1-d base state is not aligned with the Cartesian state.  Figure~\ref{fig:base_state}
-shows this geometry.
-
-\begin{figure}[tb]
-\centering
-\includegraphics[height=2.0in]{base_grid} \hspace{0.5in}
-\includegraphics[height=2.0in]{base_spherical}
-\caption{\label{fig:base_state} (Left) For multi-level problems
-  in planar geometry, we force a direct alignment between the radial
-  array cell centers and the Cartesian grid cell centers by allowing
-  the radial base state spacing to change with space and time.
-  (Right) For multi-level problems in spherical geometry, since there
-  is no direct alignment between the radial array cell centers and
-  the Cartesian grid cell centers, we choose to fix the radial base
-  state spacing across levels. Figure taken from~\cite{multilevel}}
-\end{figure}
-
-MAESTRO can use adaptive mesh refinement to focus resolution on
-complex regions of flow.  For Cartesian/plane-parallel geometries, all
-cells at the same height must be at the same level of refinement.
-This restriction is to allow for the base state to directly align with
-the Cartesian state everywhere.  For spherical geometries, there is no
-such restriction (again, see Figure~\ref{Fig:base_state}).
 
@@
 To build the MAESTRO executable, simply type `{\tt make}' in the
-problem directory.  The executable will have a name like
-{\tt main.Linux.Intel.exe}, where the specific parts of 
-the name depend on the compilers and OS used.  MAESTRO can be built
-with support for parallelism (using MPI or MPI+OpenMP), or compiled
-as a serial code.  The top portion
-of the {\tt GNUmakefile} can be editted to choose the build options.
-Setting {\tt MPI := t} will build the code to use MPI for parallelization.
-Setting {\tt OMP := t} will use OpenMP.  OpenMP is used together with MPI, 
-with MPI distributing the grids across the processors and within a 
-shared-memory node, OpenMP allows many cores to operate
-on the same grid.  A final build option determines whether to build
-the code with support for debugging.  Setting {\tt NDEBUG := t} disables
-debugging features.
+problem directory.  The executable will have a name like {\tt
+  main.Linux.Intel.exe}, where the specific parts of the name depend
+on the compilers and OS used.  MAESTRO can be built with support for
+parallelism (using MPI or MPI+OpenMP), or compiled as a serial code.
+The top portion of the {\tt GNUmakefile} can be editted to choose the
+build options.  Setting {\tt MPI := t} will build the code to use MPI
+for parallelization.  Setting {\tt OMP := t} will use OpenMP.  OpenMP
+is used together with MPI, with MPI distributing the grids across the
+processors and within a shared-memory node, OpenMP allows many cores
+to operate on the same grid.  A final build option determines whether
+to build the code with support for debugging.  Setting {\tt NDEBUG :=
+  t} disables debugging features.
+
+If the {\tt helmeos} equation of state is chosen for the problem, then
+the data table for the equation of state, {\tt
+  extern/EOS/helmeos/helm\_table.dat} needs to be copied to the
+directory in which the code is run.
 
@@
   described in \cite{xrb}.
+
+\item {\tt test\_particles}
+
+  This test exercises the particle advection routine.  A simple
+  circular velocity field, with the magnitude increasing with radius
+  from the center is initialized.  A number of particles are then
+  initialized at various radii from the center and they are advected
+  for one period.  The particle paths should be perfect circles, and
+  the final particle position should overlap with the initial
+  position.
 

### 2011-03-09 docs/runtime_parameters/runtime_parameters.tex
slight tweaking
add the run time parameter use_linear_grav_in_beta.  this adds the option to use piecewise linear gravity in the integration of beta0

@@
 \section{Runtime Parameters}
+
+Table~\ref{table:runtime} lists the runtime parameters available to
+all MAESTRO problems.  Problem-specific runtime parameters are not
+shown here.
 
@@
 &  1 \\
+\verb=  use_linear_grav_in_beta  = &   when integrating $\beta_0$, use a piecewise linear (.true.) representation of gravity or piecewise constant (.false.)  &  .false. \\
 \verb=  blocking_factor  = &   grids must be an integer multiple of this number, if possible  &  8 \\

### 2011-03-09 Source/make_div_coeff.f90

@@ -21,7 +21,7 @@ contains
     use geometry, only: nr_fine, dr, anelastic_cutoff_coord, r_start_coord, r_end_coord, &
          nr, numdisjointchunks, nlevs_radial
     use restrict_base_module
-    use probin_module, only: beta_type
+    use probin_module, only: beta_type, use_linear_grav_in_beta
 
     real(kind=dp_t), intent(  out) :: div_coeff(:,0:)
     real(kind=dp_t), intent(in   ) :: rho0(:,0:), p0(:,0:), gamma1bar(:,0:)
@@ -31,10 +31,14 @@ contains
     integer :: r, n, i, refrat, j
     real(kind=dp_t) :: integral
     real(kind=dp_t) :: beta0_edge(nlevs_radial,0:nr_fine)
-    real(kind=dp_t) :: lambda, mu, nu
-    real(kind=dp_t) :: denom, coeff1, coeff2
+    real(kind=dp_t) :: lambda, mu, nu, kappa
+    real(kind=dp_t) :: denom, coeff1, coeff2, coeff3
     real(kind=dp_t) :: del,dpls,dmin,slim,sflag
     real(kind=dp_t) :: offset
+
+    ! FIXME debugging
+    real(kind=dp_t) :: dbdr, rho_cc
+
 
     div_coeff = 0.d0
 
@@ -121,16 +125,45 @@ contains
 
                 else 
 
-                   denom = nu*gamma1bar(n,r) - mu*p0(n,r)
-                   coeff1 = lambda*gamma1bar(n,r)/mu - rho0(n,r)
-                   coeff2 = lambda*p0(n,r)/nu - rho0(n,r)
+                   if ( use_linear_grav_in_beta ) then
 
-                   integral = (abs(grav_center(n,r))/denom)* &
-                        (coeff1*log( (gamma1bar(n,r) + HALF*mu*dr(n))/ &
-                        (gamma1bar(n,r) - HALF*mu*dr(n))) - &
-                        coeff2*log( (p0(n,r) + HALF*nu*dr(n))/ &
-                        (p0(n,r) - HALF*nu*dr(n))) )
+                      del   = HALF* (grav_center(n,r+1) - grav_center(n,r-1))/dr(n)
+                      dpls  = TWO * (grav_center(n,r+1) - grav_center(n,r  ))/dr(n)
+                      dmin  = TWO * (grav_center(n,r  ) - grav_center(n,r-1))/dr(n)
[…]

### 2011-03-16 docs/beta0/beta0.tex
at Mike's behest, addding my notes on the derivation of \alpha (from paper I) including chemical potential terms and how this affects \beta_0
Note that b/c we are doing things in an operator split fashion, the
system is forced to be in chemical equilibrium during the hydro parts
of the algorithm and therefore the chemical potential terms probably
don't affect the solution of beta

@@
 $\beta_0$ as:
-\begin{equation}
+\begin{equation}\label{eq:beta_0}
 \beta_0(r,t) = \rho_0 \exp\left (  \int_0^r  \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial r^\prime} dr^\prime \right )
@@
 the path to equilibrium?
+
+\section{On the Effect of Chemical Potential}
+\label{Sec:On the Affect of Chemical Potential}
+
+In MAESTRO, we do things in an operator split fashion --- the hydro is
+de-coupled from the burning.  This means that during the hydro parts
+of the algorithm (where $\beta_0$ is used), the system is fixed in
+chemical equilibrium. For completeness, however, here we describe the
+effects of the species' chemical potentials, which were neglected in
+the original derivation of $\beta_0$.  Note that similar terms appear
+in the calculation of things such as specific heats,
+which \textit{may} be important in the burning step --- there appears to
+be very little about this in the literature, but everyone seems to
+assume it makes little difference.
+
+\subsection{Derivation of $\alpha$}
+\label{Sec:Derivation of alpha}
+In paper I, $\alpha$ is defined as
+\begin{equation}\label{eq:alpha}
+\alpha\equiv -\left(
+\frac{(1-\rho h_p)p_T-\rho c_p}{\rho^2c_pp_\rho}\right)
+\end{equation}
+where 
+\[
+h_p \equiv \left(\frac{\partial h}{\partial p}\right)_{T,X}, \quad
+c_p \equiv \left(\frac{\partial h}{\partial T}\right)_{p,X}, \quad
+p_T \equiv \left(\frac{\partial p}{\partial T}\right)_{\rho,X}, \quad
+p_\rho \equiv \left(\frac{\partial p}{\partial \rho}\right)_{T,X}
+\]
+where the subscript $X$ means holding all $X_i$ constant.  In the
+absence of reactions, the $X$ subscript can be dropped from all
+derivatives and with the use of the equation of state $p=p(\rho,T)$,
+$\alpha$ can be written as $\alpha=\alpha(\rho,T)$.  Such a system
+without reactions and in thermal equilibrium could be either a pure
+system of one species, or a system of many species in chemical (and
+therefore \emph{thermodynamic}) equilibrium.  Cox and Giuli (hereafter
+CG) call the former type of system a ``simple system'' and therefore
+the latter a ``non-simple system'' in chemical equilibrium.  The
+analysis in paper I that reduced \eqref{eq:alpha} to
+\begin{equation}\label{eq:alpha_simp_no_rxn} 
+\alpha = \frac{1}{\Gamma_1p_0} 
+\end{equation} 
+used CG's discussion of the various adiabatic $\Gamma$'s.  However,
+their discussion only pertains to ``simple systems'' or ``non-simple
+systems'' in chemical equilibrium.  In general, nuclear reactions will
+be important and therefore this analysis needs to be reformed.
+
+Even in the presence of reactions, \eqref{eq:alpha} can be rewritten
+as was done in paper I:
+\begin{equation}\label{eq:alpha2}
+\alpha = -\frac{1}{p\chi_\rho c_p}\left[\left(\frac{1}{\rho\chi_\rho}
+- \frac{\rho e_\rho}{p\chi_\rho}\right)\frac{p\chi_T}{T} - c_p\right],
+\end{equation}
+where
+\begin{align*}
+\chi_{\rho} &\equiv \left(\frac{\partial\ln p}{\partial\ln\rho}
+\right)_{T,X} \\
+\chi_{T} &\equiv \left(\frac{\partial\ln p}{\partial\ln T}
+\right)_{\rho,X}.
+\end{align*}
+Following the results of paper I, we want to find a relation
+between $p\chi_\rho$ and $\Gamma_1$.
+
+For an equation of state $p=p(\rho,T,X)$ we have
+\[
+d\ln p = \left(\frac{\partial\ln p}{\partial\ln\rho}\right)_{T,X}d\ln\rho +
+\left(\frac{\partial\ln p}{\partial\ln T}\right)_{\rho,X}d\ln T +
+\sum_i\left(\frac{\partial\ln p}{\partial\ln X_i}\right)_{\rho,T,(X_j,j
+\neq i)} d\ln X_i.
+\]
+We define another logarithmic derivative
+\begin{align*}
+\chi_{X_{i}} &\equiv \left(\frac{\partial\ln p}{\partial\ln X_i}
+\right)_{\rho,T,(X_j,j\neq i)}
+\end{align*}
+and therefore
+\[
+d\ln p = \chi_\rho \ d\ln\rho + \chi_T \ d\ln T + \sum_i \chi_{X_i}\ 
+d\ln X_i.
+\]
+From here we get the general statement 
+\[
+\frac{\partial\ln p}{\partial \ln \rho} = \chi_\rho + 
+\chi_T\frac{\partial \ln T}{\partial\ln \rho} +
+\sum_i\chi_{X_i}\frac{\partial\ln X_i}{\partial\ln \rho}
+\]
+which must hold for an adiabatic process as well, and therefore we have
+\begin{equation}\label{eq:gamma1}
+  \Gamma_1 = \chi_\rho + \chi_T\left(\Gamma_3-1\right) 
+  + \sum_i\chi_{X_i}\Gamma_{4,i}
+\end{equation}
+where we use CG's definition of $\Gamma_1$ and $\Gamma_3$ and introduce a 
+fourth gamma function:
+\[
+\Gamma_1 \equiv \left(
+\frac{\partial \ln p}{\partial \ln \rho}\right)_{\text{AD}},\quad
+\Gamma_3-1\equiv \left(
+\frac{\partial \ln T}{\partial \ln \rho}\right)_{\text{AD}},\quad
+\Gamma_{4,i}\equiv \left(
+\frac{\partial\ln X_i}{\partial\ln\rho}\right)_{\text{AD}},
+\]
+where the subscript AD means along an adiabat.  We now derive an expression
+for $\Gamma_3$.
+
+The first law of thermodynamics can be written as
+\[
+dQ = dE + pdV - \sum_i\mu_idN_i
+\]
+where $\mu_i=\left(
+\frac{\partial E}{\partial N_i}\right)_{\text{AD},\rho,(N_j,j\neq i)}$ is
+the chemical potential; or per unit mass we have
+\begin{align*}
+  dq &= de - \frac{p}{\rho^2}d\rho - \sum_i\mu_id
+  \left(\frac{n_i}{\rho}\right)\\
+  &= de - \frac{p}{\rho^2}d\rho - \sum_i
+  \left(
+  \frac{\partial e}{\partial X_i}\right)_{\rho,\text{AD},(X_j,j\neq i)}dX_i
+\end{align*}
+where we have used $X_i \equiv \rho_i/\rho = A_in_i/\rho N_\text{A}$ 
+and the chemical potential has been replaced with 
+$\mu_i = \frac{A_i}{N_\text{A}}\left(\frac{\partial e}{\partial X_i}
+  \right)_{\rho,\text{AD},(X_j,j\neq i)}$.
+Using this and expressing the specific internal energy as $e=e(\rho,T,X)$ 
+we then have
+\[
+dq = c_vdT +
+\left[\left(\frac{\partial e}{\partial \rho}\right)_{T,X} -\frac{p}{\rho^2}
+  \right]d\rho + 
+\sum_i\left[
+  \left(\frac{\partial e}{\partial X_i}\right)_{\rho,T,(X_j,j\neq i)} -
+  \left(\frac{\partial e}{\partial X_i}
+  \right)_{\rho,\text{AD},(X_j,j\neq i)}\right]dX_i
+\]
+and
+\begin{align}\label{eq:gamma3_first}
+\left(\frac{d\ln T}{d\ln\rho}\right)_\text{AD} \equiv \Gamma_3-1
+&= \frac{1}{c_vT}\left[
+\frac{p}{\rho} - \left(\frac{\partial e}{\partial\ln\rho}\right)_{T,X} + 
+\right.{}\nonumber\\
+&\qquad\qquad  \left.\sum_i \left[
+    \left(
+    \frac{\partial e}{\partial X_i}\right)_{\rho,\text{AD},(X_j,j\neq i)} 
+    - 
+    \left(\frac{\partial e}{\partial X_i}\right)_{\rho,T,(X_j,j\neq i)}
+    \right]X_i\Gamma_{4,i}\right]
+\end{align}
+
+Now we need to evaluate $\left(\partial e/\partial \ln\rho\right)_{T,X}$.
+Again using the first law and the fact that $ds=dq/T$ is an exact 
+differential (i.e. mixed derivatives are equal) we have
+\begin{align}\label{eq:dedlnrho}
+  \left(
+  \frac{\partial}{\partial\rho}\left[\frac{c_v}{T}\right]\right)_{T,X} &=
+  \left(\frac{\partial}{\partial T}\left[\frac{1}{T}
+    \left(\frac{\partial e}{\partial\rho}\right)_{T,X} - \frac{p}{T\rho^2}
+    \right]\right)_{\rho,X}{}\nonumber\\
+  \frac{1}{T}\left(\frac{\partial}{\partial\rho}\left(
+  \frac{\partial e}{\partial T}\right)_{\rho,X}\right)_{T,X} &=
+  -\frac{1}{T^2}\left(\frac{\partial e}{\partial\rho}\right)_{T,X} +
+  \frac{1}{T}\left(\frac{\partial}{\partial T}\left(
+  \frac{\partial e}{\partial\rho}\right)_{T,X}\right)_{\rho,X}
+  +\frac{p}{T^2\rho^2} - 
+  \frac{1}{T\rho^2}\left(\frac{\partial p}{\partial T}\right)_{\rho,X}
+  {}\nonumber\\
+  \therefore\quad \left(\frac{\partial e}{\partial\ln \rho}\right)_{T,X} &=
+  \frac{p}{\rho}\left(1-\chi_T\right),
+\end{align}
+exactly the same result if we were to exclude species information.
+Simlarly, we can find an expression for the derivative of energy with 
+respect to composition
+\begin{align*}
+  \left(\frac{\partial}{\partial X_i}\left[
+    \frac{c_v}{T}\right]\right)_{\rho,T,(X_j,j\neq i)} &=
+  \left(\frac{\partial}{\partial T}\left[
+    \frac{1}{T}\left(\frac{\partial e}{\partial X_i}
+      \right)_{\rho,T,(X_j,j\neq i)} - \frac{1}{T}\left(
+      \frac{\partial e}{\partial X_i}\right)_{\rho,\text{AD},(X_j,j\neq i)}
+    \right]\right)_{\rho,X}\\
+  \frac{1}{T}\left(\frac{\partial }{\partial X_i}\left(
+  \frac{\partial e}{\partial T}\right)_{\rho,X}\right)_{\rho,T,(X_j,j\neq i)}
+  &= \frac{1}{T^2}\left[\left(\frac{\partial e}{\partial X_i}
+    \right)_{\rho,\text{AD},(X_j,j\neq i)} - 
+    \left(\frac{\partial e}{\partial X_i}\right)_{\rho,T,(X_j,j\neq i)}
+    \right] + \\
+  &\ \ \ \ \ \frac{1}{T}\left[
+    \left(\frac{\partial}{\partial T}\left(
+    \frac{\partial e}{\partial X_i}\right)_{\rho,T,(X_j,j\neq i)}
+    \right)_{\rho,X} - 
+    \left(\frac{\partial }{\partial T}\left(
+    \frac{\partial e}{\partial X_i}\right)_{\rho,\text{AD},(X_j,j\neq i)}
+    \right)_{\rho,X}\right]\\
+  \therefore\quad 
+  \left(\frac{\partial e}{\partial X_i}\right)_{\rho,T,(X_j,j\neq i)} &=
+  \left(\frac{\partial e}{\partial X_i}
+  \right)_{\rho,\text{AD},(X_j,j\neq i)} - \left(
+  \frac{\partial}{\partial\ln T}\left(
+  \frac{\partial e}{\partial X_i}\right)_{\rho,\text{AD},(X_j,j\neq i)}
+  \right)_{\rho,X}.
+\end{align*}
+Plugging these back into \eqref{eq:gamma3_first} we have
+\begin{equation}\label{eq:gamma3_second}
+  \Gamma_3-1 = \frac{1}{c_vT}\left[\frac{p}{\rho}\chi_T +\sum_i
+    \left(\frac{\partial}{\partial\ln T}\left(
+    \frac{\partial e}{\partial X_i}\right)_{\rho,\text{AD},(X_j,j\neq i)}
+    \right)_{\rho,X}X_i
+    \Gamma_{4,i}\right],
+\end{equation}
+or
+\begin{equation}\label{eq:cv}
+  c_v = \frac{1}{T(\Gamma_3-1)}\left[\frac{p}{\rho}\chi_T +\sum_i
+    \left(\frac{\partial}{\partial\ln T}\left(
+    \frac{\partial e}{\partial X_i}\right)_{\rho,\text{AD},(X_j,j\neq i)}
+    \right)_{\rho,X}X_i
+    \Gamma_{4,i}\right].
+\end{equation}
+
+We can obtain an expression for the specific heat at constant pressure 
+from the enthalpy
+\begin{align*}
+  c_p \equiv \left(\frac{\partial h}{\partial T}\right)_{p,X} &=
+  \left(\frac{\partial e}{\partial T}\right)_{p,X} - \frac{p}{\rho^2}
+  \left(\frac{\partial \rho}{\partial T}\right)_{p,X}\\
+  &= \left(\frac{\partial e}{\partial T}\right)_{p,X} + \frac{p}{\rho^2}
+  \left(\frac{\partial p}{\partial T}\right)_{\rho,X}
+  \left(\frac{\partial \rho}{\partial p}\right)_{T,X}\\
+  &=\left(\frac{\partial e}{\partial T}\right)_{p,X} + \frac{p}{\rho T}
+  \frac{\chi_t}{\chi_\rho}.
+\end{align*}
+The first term on the rhs can be obtained from writing $e=e(p,T,X)$ and
+$p=p(\rho,T,X)$:
+\begin{align*}
+  de &= \left(\frac{\partial e}{\partial p}\right)_{T,X}dp
+  + \left(\frac{\partial e}{\partial T}\right)_{p,X}dT +
+  \sum_i \left(\frac{\partial e}{\partial X_i}\right)_{p,T,(X_j,j\neq i)}
+  dX_i\\
+  dp &= \left(\frac{\partial p}{\partial \rho}\right)_{T,X}d\rho +
+  \left(\frac{\partial p}{\partial T}\right)_{\rho,X}dT + \sum_i 
+  \left(\frac{\partial p}{\partial X_i}\right)_{\rho,T,(X_j,j\neq i)}dX_i\\
+  \therefore \ \left(\frac{\partial e}{\partial T}\right)_{\rho,X} &= 
+  \left(\frac{\partial e}{\partial p}\right)_{T,X}
+  \left(\frac{\partial p}{\partial T}\right)_{\rho,X} + 
+  \left(\frac{\partial e}{\partial T}\right)_{p,X}\\
+  \Rightarrow \ \left(\frac{\partial e}{\partial T}\right)_{p,X} &= c_v -
+  \left(\frac{\partial e}{\partial \rho}\right)_{T,X}
+  \left(\frac{\partial \rho}{\partial p}\right)_{T,X}
+  \left(\frac{\partial p}{\partial T}\right)_{\rho,X}\\
+  &= c_v - \frac{p\chi_T}{\rho T\chi_\rho}\left(1-\chi_T\right)
+\end{align*}
+and 
+\[
+c_p = \frac{p}{\rho T}\frac{\chi_T^2}{\chi_\rho} + c_v
+\]
+Dividing this by \eqref{eq:cv} and using the relation between the 
+$\Gamma$'s, \eqref{eq:gamma1}, we then have
+\begin{align}\label{eq:pchirho}
+  \gamma \equiv \frac{c_p}{c_v} &= 1 + \frac{p(\Gamma_3-1)}{\rho }
+  \frac{\chi_T^2}{\chi_\rho}\left[\frac{p}{\rho}\chi_T +\sum_i
+    \left(\frac{\partial}{\partial\ln T}\left(
+    \frac{\partial e}{\partial X_i}\right)_{\rho,\text{AD},(X_j,j\neq i)}
+    \right)_{\rho,X}X_i
+    \Gamma_{4,i}\right]^{-1}{}\nonumber\\
+  &= 1 + \frac{p\chi_T\left(\Gamma_1 - \chi_\rho -
+    \sum_i \chi_{X_i}\Gamma_{4,i}\right)}{p\chi_\rho\chi_T + \rho
+    \chi_\rho\sum_i \left(
+    \frac{\partial}{\partial \ln T}\left(
+    \frac{\partial e}{\partial X_i}\right)_{\rho,\text{AD},(X_j,j\neq i)}
+    \right)_{\rho,X}X_i\Gamma_{4,i}}{}\nonumber\\
+  &= \frac{p\chi_T\Gamma_1 + \sum_i \left[\rho\chi_\rho\left(
+    \frac{\partial}{\partial \ln T}\left(\frac{\partial e}{\partial X_i}
+    \right)_{\rho,\text{AD},(X_j,j\neq i)}\right)_{\rho,X}X_i - p\chi_T
+    \chi_{X_i}\right]\Gamma_{4,i}}{p\chi_\rho\chi_T + \rho
+    \chi_\rho\sum_i \left(
+    \frac{\partial}{\partial \ln T}\left(
+    \frac{\partial e}{\partial X_i}\right)_{\rho,\text{AD},(X_j,j\neq i)}
+    \right)_{\rho,X}X_i\Gamma_{4,i}}{}\nonumber\\
+  \Rightarrow p\chi_\rho &= \frac{1}{\chi_T\gamma}\left[p\chi_T\Gamma_1 + 
+    \sum_i \left[\rho\chi_\rho\left(1-\gamma\right)\left(
+      \frac{\partial}{\partial \ln T}\left(\frac{\partial e}{\partial X_i}
+      \right)_{\rho,\text{AD},(X_j,j\neq i)}\right)_{\rho,X}X_i - p\chi_T
+      \chi_{X_i}\right]\Gamma_{4,i}\right].
+\end{align}
+
+Plugging \eqref{eq:pchirho} into \eqref{eq:alpha2} and rewriting the 
+partial derivative of $e$ with the help of \eqref{eq:dedlnrho} we have
+\begin{align*}
+\alpha &= -\frac{1}{p\chi_\rho c_p}\left[\left(\frac{1}{\rho\chi_\rho}
+  - \frac{\rho e_\rho}{p\chi_\rho}\right)\frac{p\chi_T}{T} - c_p\right] \\
+&=\frac{\gamma}{c_p}\frac{c_p\chi_T + \left(\rho
+  \left(\frac{\partial e}{\partial\ln\rho}\right)_{T,X}-p\right)
+  \frac{\chi_T^2}{T\rho\chi_\rho}}
+{p\chi_T\Gamma_1 + 
+    \sum_i \left[\rho\chi_\rho\left(1-\gamma\right)\left(
+      \frac{\partial}{\partial \ln T}\left(\frac{\partial e}{\partial X_i}
+      \right)_{\rho,\text{AD},(X_j,j\neq i)}\right)_{\rho,X}X_i - p\chi_T
+      \chi_{X_i}\right]\Gamma_{4,i}}\\
+&=\frac{\gamma}{\Gamma_1 p c_p}\left[\frac{c_p - \frac{p\chi_T^2}
+    {T\rho\chi_\rho}}
+  {1 + \sum_i \left[\frac{\rho\chi_\rho}{p\chi_T}
+      \left(1-\gamma\right)\left(
+      \frac{\partial}{\partial \ln T}\left(\frac{\partial e}{\partial X_i}
+      \right)_{\rho,\text{AD},(X_j,j\neq i)}\right)_{\rho,X}X_i - 
+      \chi_{X_i}\right]\frac{\Gamma_{4,i}}{\Gamma_1}}\right]\\
+&=\left(\frac{1}{\Gamma_1p}\right)
+\left[1 + \sum_i \left[\frac{\rho\chi_\rho}{p\chi_T}
+    \left(1-\gamma\right)\left(
+    \frac{\partial}{\partial \ln T}\left(\frac{\partial e}{\partial X_i}
+    \right)_{\rho,\text{AD},(X_j,j\neq i)}\right)_{\rho,X}X_i - 
+    \chi_{X_i}\right]\frac{\Gamma_{4,i}}{\Gamma_1}\right]^{-1}\\
+\end{align*}
+\[\boxed{
+  \alpha = \frac{1}{\Gamma_1p}\left[1 + \sum_i\left[\frac{\rho^2p_\rho}
+      {pp_T}(1-\gamma)\frac{N_\text{A}}{A_i}
+      \left(\frac{\partial\mu_i}{\partial T}\right)_{\rho,X}X_i - 
+      \chi_{X_i}\right]\frac{\Gamma_{4,i}}{\Gamma_1}\right]^{-1}
+}
+\]
+
+\subsection{Recalling Derivation of $\beta_0$}
+\label{Recalling Derivation of beta0}
+Recall from paper I that $\beta_0$ was derived from the equation 
+\begin{equation}
+\nabla\cdot\mathbf{U} + \alpha\mathbf{U}\cdot\nabla p_0 = \tilde{S}
+\end{equation}
+in such a fashion that we ended up with an equation of the form
+\begin{equation}\label{eq:beta constraint}
+\nabla\cdot\left(\beta_0(r)\mathbf{U}\right) = \beta_0\tilde{S}.
+\end{equation}
+The derivation in Appendix B of paper I for a $\beta_0$ that
+satisfies \eqref{eq:beta constraint} automatically assumed $\alpha
+= \left(\Gamma_{1_0}p_0\right)^{-1}$.  This would have to be modified
+with the above derivation of $\alpha$ to be correct in a non-operator
+split fashion.

### 2011-03-22 docs/lo_density/lo_density.tex
some updates

@@
-These are working notes for the low density parameters in {\tt MAESTRO}.
+These are working notes for the low density parameters in {\tt MAESTRO}.  In low density
+regions, we modify the behavior of the algorithm.  Here is a summary of some parameters,
+and a brief description of what they do.
+\begin{itemize}
+\item {\tt base\_cutoff\_density}, $\rho_{\rm base}$, (real):\\ \\
+Essentially controls the lowest density allowed in the simulation and modifies the behavior
+of several modules.
+\item {\tt base\_cutoff\_density\_coord(:)} (integer array):\\ \\
+For each level in the radial base state array, this is the coordinate of the first cell
+where $\rho_0 < \rho_{\rm base}$.  Slightly more complicated for multilevel problems.
+\item {\tt anelastic\_cutoff}, $\rho_{\rm anelastic}$, (real):\\ \\
+If $\rho_0 < \rho_{\rm anelastic}$, we modify the computation of $\beta_0$ in the
+divergence constraint.
+\item {\tt anelastic\_cutoff\_coord(:)} (integer array):\\ \\
+Anelastic cutoff analogy of {\tt base\_cutoff\_density\_coord(:)}.
+\item {\tt burning\_cutoff\_density}, $\rho_{\rm burning}$,  (real):\\ \\
+If $\rho < \rho_{\rm burning}$, don't call the burner in this cell.
+\item {\tt burning\_cutoff\_density\_coord(:)} (integer array):\\ \\
+Burning cutoff analogy of {\tt base\_cutoff\_density\_coord(:)}.
+\item {\tt buoyancy\_cutoff\_factor} (real):\\ \\
+When computing velocity forcing, set the buoyance term ($\rho-\rho_0$) to 0 if 
+$\rho < $ {\tt buoyancy\_cutoff\_factor} * {\tt base\_cutoff\_density}.
+\item {\tt do\_eos\_h\_above\_cutoff} (logical):\\ \\
+If {\rm true}, at the end of the advection step, for each cell where 
+$\rho < \rho_{\rm base}$, recompute $h = h(\rho,p_0,X)$.
+\end{itemize}
 
@@
 
-At several points in the algorithm, we compute {\tt anelastic\_cutoff\_coord}, 
-{\tt base\_cutoff\_density\_coord}, and {\tt burning\_cutoff\_density\_coord} by 
+At several points in the algorithm, we compute {\tt anelastic\_cutoff\_coord}(:), 
+{\tt base\_cutoff\_density\_coord}(:), and {\tt burning\_cutoff\_density\_coord}(:) by 
 setting the corresponding coordinate equal to $r$ as soon as $\rho_0(r)$ is less than 
-or equal to that particular cutoff value (see Figure \ref{Fig:Cutoff}.
-We compute the cutoff coordinate at the
-coarsest level, and then set the coordinate at the next finer level to be twice that
-of the coarser level.  We set the coordinates at the following points in the code:
+or equal to that particular cutoff value (see Figure \ref{Fig:Cutoff} for the
+single-level case).  Multilevel is more complicated.
+
+We set the coordinates at the following points in the code:
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\item At the beginning of {\tt advance\_timestep}.
+\item After reading the base state from a checkpoint file when restarting.
 
-\item After each call to {\tt advect\_base\_dens}.
+\item After regridding.
 
-\item After we ``{\tt correct base}'' (by setting $\rho_0 = \overline{\rho}$ in
-  plane-parallel and $\rho_0 = \rho_0 + {\bf{Avg}}(\rho - \rho_0)$ in spherical).
+\item After advancing $\rho_0$ with {\tt advect\_base\_dens}.
 
-\item At the beginning of the second-half of the algorithm (Step 6), we reset
+\item After advancing $\rho$ and setting $\rho_0 = \overline{\rho}$.
+
+\item At the beginning of the second-half of the algorithm ({\bf Step 6}), we reset
   the coordinates to the base-time values using $\rho_0^n$.
@@
 
-\item In {\tt sponge}, {\tt anelastic\_cutoff} is used in a problem
-  dependent way.
+\item Some versions of {\tt sponge}, use {\tt anelastic\_cutoff} in a problem dependent way.
 

### 2011-03-22 docs/lo_density/lo_density.tex
first attempt at describing new definition of cutoff densities for multilevel planar

@@
 \section{Computing the Cutoff Values}
+We compute {\tt anelastic\_cutoff\_coord}(:), {\tt base\_cutoff\_density\_coord}(:), 
+and {\tt burning\_cutoff\_density\_coord}(:) in analogous fashion.
 
-At several points in the algorithm, we compute {\tt anelastic\_cutoff\_coord}(:), 
-{\tt base\_cutoff\_density\_coord}(:), and {\tt burning\_cutoff\_density\_coord}(:) by 
-setting the corresponding coordinate equal to $r$ as soon as $\rho_0(r)$ is less than 
-or equal to that particular cutoff value (see Figure \ref{Fig:Cutoff} for the
-single-level case).  Multilevel is more complicated.
-
-We set the coordinates at the following points in the code:
+\subsection{Single-Level Planar or any Spherical}
+Here the base state exists as a single one-dimensional array with constant grid
+spacing $\Delta r$.  Basically, we set the corresponding coordinate equal to $r$ as soon 
+as $\rho_0(r)$ is less than  or equal to that particular cutoff value.
+See Figure \ref{Fig:Cutoff} for a graphical representation.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \includegraphics[width=4.0in]{\lodensfigpath/cutoff}\hspace{0.2in}
-\begin{minipage}[b]{5.0in}
-\caption[Relation between cutoff density and cutoff coordinates]
-{Image of how the cutoff density and cutoff coordinates
-are related.  This is drawn assuming {\tt drdxfac}=5.\vspace{2em}}
-\end{minipage}
+\caption{Image of how the cutoff density and cutoff coordinates
+are related for single-level planar and all spherical problems.}
 \label{Fig:Cutoff}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Note that for single-level planar or any spherical problem, saying $r\ge$ {\tt anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)\le$ {\tt anelastic\_cutoff}.
+Also, saying $r<$ {\tt anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)>$ {\tt anelastic\_cutoff}.
+Ditto for {\tt base\_cutoff\_density} and {\tt base\_cutoff\_density\_coord}.
+
+\subsection{Multilevel Planar}
+In this case, the base state exists as several one-dimensional arrays, each with
+different grid spacing.  Refer to Figure \ref{Fig:Cutoff_Multi} in the following examples.
+The guiding principle is to check whether $\rho_0$ falls below $\rho_{\rm cutoff}$ on the finest
+grid first.  If not, check the next coarser level.  Continue until you reach the base grid,
+in which case $\rho_0$ must fall below $\rho_{\rm cutoff}$ somewhere.  Some examples are in order:
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[hpb]
+\centering
+\includegraphics[width=2.0in]{\lodensfigpath/cutoff_multi}\hspace{0.2in}
+\caption{Multilevel cutoff density example.}
+\label{Fig:Cutoff_Multi}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{itemize}
+
+\item {\bf Example 1:} $\rho_{0,104} > \rho_{\rm cutoff}$ and $\rho_{0,105} < \rho_{\rm cutoff}$.\\ \\
+{\tt cutoff\_density\_coord}(1) = 105\\
+{\tt cutoff\_density\_coord}(2) = 210\\
+{\tt cutoff\_density\_coord}(3) = 420\\ \\
+This is the simplest case in which the cutoff transition happens on the coarsest level.
+In this case, the cutoff coordinates at the finer levels are simply propagated from the
+coarsest level, even though they do not correspond to a valid region.
+
+\item {\bf Example 2:} $\rho_{0,403} > \rho_{\rm cutoff}$ and $\rho_{0,404} < \rho_{\rm cutoff}$.\\ \\
+{\tt cutoff\_density\_coord}(1) = 101\\
+{\tt cutoff\_density\_coord}(2) = 202\\
+{\tt cutoff\_density\_coord}(3) = 404\\ \\
+In this case, the cutoff transition happens where the finest grid is present.  Happily, the
+transition occurs at a location where there is a common grid boundary between all three levels.
+Therefore, we simply propagate the cutoff density coordinate from the finest level downward.
+
+\item {\bf Example 3:} $\rho_{0,404} > \rho_{\rm cutoff}$ and $\rho_{0,405} < \rho_{\rm cutoff}$.\\ \\
+{\tt cutoff\_density\_coord}(1) = 102\\
+{\tt cutoff\_density\_coord}(2) = 203\\
+{\tt cutoff\_density\_coord}(3) = 405\\ \\
+In this case, the cutoff transition happens where the finest grid is present.  However, the
+transition occurs at a location where there NOT is a common grid boundary between all three 
+levels.  We choose to define the cutoff transition at the coarser levels as being at the
+corresponding boundary that is at a larger radius than the location on the finest grid.
+
+\end{itemize}
+
+\section{When are the Cutoff Coordinates Updated?}
+
+At several points in the algorithm, we compute {\tt anelastic\_cutoff\_coord}(:), 
+{\tt base\_cutoff\_density\_coord}(:), and {\tt burning\_cutoff\_density\_coord}(:):
 
@@
 
-\section{Density vs. Coordinate}
+\section{Usage of Cutoff Densities}
 
-Note that saying $r\ge$ {\tt anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)\le$ {\tt anelastic\_cutoff}.
-Also, saying $r<$ {\tt anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)>$ {\tt anelastic\_cutoff}.
-Ditto for {\tt base\_cutoff\_density} and {\tt base\_cutoff\_density\_coord}.
-
-\section{anelastic\_cutoff}\label{Sec:Anelastic Cutoff}
+\subsection{anelastic\_cutoff}\label{Sec:Anelastic Cutoff}
 
@@
 
-\section{base\_cutoff\_density}\label{Sec:Base Cutoff Density}
+\subsection{base\_cutoff\_density}\label{Sec:Base Cutoff Density}
 
@@
 
-\section{burning\_cutoff}
+\subsection{burning\_cutoff}
 
@@
 
-\section{buoyancy\_cutoff\_factor}
+\subsection{buoyancy\_cutoff\_factor}
 

### 2011-03-25 docs/introduction/introduction.tex
update some paper references

@@
 
-\noindent We have two papers in preparation that describe applications of the
-method:
+\noindent We have two papers that describe applications of the
+method to Type Ia supernovae and X-ray bursts:
 \begin{itemize}
+\item {\em Multidimensional Modeling of Type I X-ray Bursts,}
+  C. M. Malone, A. Nonaka, A. S. Almgren, J. B. Bell, \& M. Zingale 2011,
+  ApJ, 728, 118 (henceforth ``the XRB paper'')
 \item {\em The Convective Phase Preceding Type Ia Supernovae,}
-  M. Zingale, A. S. Almgren, J. B. Bell, C. M. Malone, A. Nonaka, \& S. E. Woosley 2010,
-  in preparation (henceforth paper V)
-\item {\em Multidimensional Modeling of Type I X-ray Bursts,}
-  C. M. Malone, A. Nonaka, A. S. Almgren, J. B. Bell, \& M. Zingale 2010,
-  in preparation (henceforth ``the XRB paper'')
+  M. Zingale, A. S. Almgren, J. B. Bell, C. M. Malone, A. Nonaka, \& S. E. Woosley 2011,
+  submitted to ApJ (henceforth paper V)
 \end{itemize}

### 2011-04-02 docs/enthalpy/enthalpy.tex
start of the enthalpy notes

\section{Enthalpy Evolution}

In the low Mach number formulation, the continuity and momentum
equations together with the constraint on the velocity field
completely determine the solution.
\begin{eqnarray}
\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + 
\rho\omegadot_k,\label{eq:species}\\
\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - 
  \frac{1}{\rho}\nabla\pi - 
  \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}
\end{eqnarray}

The addition of an energy equation over-specifies the problem, but its
solution can help us in providing a diagnostic capability for the
solution.  Additionally, work from combustion suggests that low Mach
number methods perform better when considering reactions if we use
an enthalpy equation to determine the temperature.

The velocity constraint equation was constructed by diffrentiating 
the pressure along particle paths, using the equation of state.

The full enthalpy equation, with no approximations, appears as:
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + 
  \frac{Dp}{Dt} + \rho\Hnuc \label{eq:enthalpy}
\end{equation}
Here, $h = e + p/\rho$ is the specific enthalpy, with $e$ the specific
internal energy.  In the low Mach number formulation, we replace $p$
with $p_0$ in the $Dp/Dt$ term, however, the definition of enthalpy
implicitly contains a pressure.  When calling the equation of state,
we take $h$ and $\rho$ as inputs.  The equation of state is expressed
in terms of $T$ and $\rho$, so it iterates until it finds the $h$ that
we desire.  This $h$ will be of the form $h = e + p_\mathrm{EOS}/rho$,
where $p_\mathrm{EOS}$ is the pressure returned from the EOS.  Note that
$p_\mathrm{EOS}$ may not be equal to $p_0$---this causes us to drive
off of the constraint. \MarginPar{not completely sure what to say here}

The mismatch between the pressure implicit in the definition of $h$
and $p_0$ can be seen by substituting $h = e + p/\rho$ into the
enthalpy equation, where we replace $p$ with $p_0$ in the $Dp/Dt$ term:
\begin{eqnarray}
\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + 
  \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} + \frac{\partial p}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) -\nabla\cdot(p\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) - p\nabla\cdot\Ub + \rho\Hnuc + 
  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} \nonumber 
\end{eqnarray}
We note that the last equation appears to be an internal energy
evolution equation, but with an additional forcing term: $D(p_0 - p)/Dt$.

\section{Outstanding Questions}

\begin{itemize}
\item Why do we want to start with enthalpy instead of internal energy?

\item We always say that to substituting $p$ by $p_0$ in the enthalpy equation
   is what filters the acoustics from the system---did we ever prove this?

\item In the EOS, we take $h$ and construct $e$ by subtracting off
  $p$---depending on the inputs to the EOS, we treat the '$p$' part of $h$
  differently.

   if we come into the EOS with $\rho$, $T$, then we get $p$ and $e$ from
   the EOS and compute $h$.

   if we come into the EOS with $\rho$, $h$, then we find a $T$ such that
   $e + p(\rho,T)/\rho = h$, where $p(\rho,T)$ is the pressure from the EOS.
   It is not necessarily equal to $p_0$.

   if we come into the EOS with $\rho$, $p_0$, then find a $T$ that matches
   $p_0$, and then $h = e + p_0/\rho$.  In this case, there is no mismatch
   in the enthalpy equation.

\item do we want to always insist that the '$p$' hidden in the definition of
   $h$ is $p_0$?

\end{itemize}

### 2011-04-03 docs/enthalpy/enthalpy.tex
some small fixes and a comment on why h

@@
 in terms of $T$ and $\rho$, so it iterates until it finds the $h$ that
-we desire.  This $h$ will be of the form $h = e + p_\mathrm{EOS}/rho$,
+we desire.  This $h$ will be of the form $h = e + p_\mathrm{EOS}/\rho$,
 where $p_\mathrm{EOS}$ is the pressure returned from the EOS.  Note that
@@
 
-\begin{itemize}
+\begin{enumerate}
 \item Why do we want to start with enthalpy instead of internal energy?
+
+   We believe that the original desire stems from our experience with smallscale 
+   combustion.  There, stratification is not important and $Dp_0/Dt = 0$, so
+   the enthalpy equation becomes a conservation equation for $(\rho h)$.
 
@@
 \item In the EOS, we take $h$ and construct $e$ by subtracting off
-  $p$---depending on the inputs to the EOS, we treat the '$p$' part of $h$
+  $p$---depending on the inputs to the EOS, we treat the `$p$' part of $h$
   differently.
@@
 
-\item do we want to always insist that the '$p$' hidden in the definition of
+\item do we want to always insist that the `$p$' hidden in the definition of
    $h$ is $p_0$?
 
-\end{itemize}
+\end{enumerate}

### 2011-04-03 docs/enthalpy/enthalpy.tex
Added more stuff about derivation of constraint.

@@
-\section{Enthalpy Evolution}
+\section{Evolution Equations}
 
-In the low Mach number formulation, the continuity and momentum
-equations together with the constraint on the velocity field
-completely determine the solution.
+The compressible and low Mach number formulations of the governing
+equations both share the unapproximated continuity and momentum equations
+shown here (where $p = p_0 + \pi$).
 \begin{eqnarray}
@@
 
-The addition of an energy equation over-specifies the problem, but its
-solution can help us in providing a diagnostic capability for the
-solution.  Additionally, work from combustion suggests that low Mach
-number methods perform better when considering reactions if we use
-an enthalpy equation to determine the temperature.
+In the compressible formulation we complete the system with an energy
+equation as well as an equation of state; in the low Mach number
+formulation we can derive a constraint on the velocity by 
+setting $p_{EOS} = p(\rho,h,X_k) = p_0$ and
+differentiating the equation of state along particle paths.
+In this case adding the energy equation would over-constrain the system, 
+but its solution can help us in providing a diagnostic capability for the
+solution.  We can write the energy equation in terms of internal energy, $e$,
+or enthalpy, $h$; these two equations are analytically equivalent.
+\begin{eqnarray}
+\frac{\partial(\rho h)}{\partial t} + \nabla\cdot(\rho h\Ub) 
+ &=& \frac{Dp}{Dt} + \rho\Hnuc \nonumber \\
+%
+\frac{\partial(\rho e)}{\partial t} + \nabla\cdot(\rho e \Ub)
+ &=& - p\nabla\cdot \Ub + \rho\Hnuc \nonumber 
+\end{eqnarray}
+In low Mach number combustion, $Dp/Dt = 0,$ which makes the enthalpy equation
+preferable to the internal energy equation because we don't need to
+evaluate $p \nabla \cdot \Ub.$
 
-The velocity constraint equation was constructed by diffrentiating 
+\section{Derivation of Constraint Using Enthalpy Equation}
+
+The velocity constraint equation was constructed by differentiating 
 the pressure along particle paths, using the equation of state.  
+Specifically, we wrote $p_{EOS} = p(\rho,T,X_k) = p_0$
+and expressed $DT/Dt$ in terms of $Dh/Dt:$
+\begin{equation}
+\frac{DT}{Dt} = \frac{1}{\rho c_p} \left( (1 - \rho h_p) \frac{D p}{D t}
+- \sum_k \rho \xi_k \omegadot_k + \rho H \right) \enskip , \label{eq:dTdt}
+\end{equation}
+where $c_p = \left.\partial h/\partial T\right|_{p,X_k}$ is the
+specific heat at constant pressure,
+$\xi_k = \left.\partial h/\partial X_k \right|_{p,T}$,
+and $h_p = \left.\partial h/\partial p\right|_{T,X_k}$
+for convenience.
+
+Then we derived
+\begin{eqnarray*}
+\nabla \cdot \ubold &=& \frac{1}{\rho p_\rho} \left(
+- \frac{D p}{D t} + \frac{p_T}{\rho c_p}
+  \left( (1 - \rho h_p) \frac{D p}{D t} - \rho \sum_k \xi_k \omegadot_k + \rho H \right)
++ \sum_k p_{X_k} \omegadot_k \right)  \\
+                 &=& \frac{1}{\rho p_\rho}
+  \left( \frac{p_T}{\rho c_p}(1  - \rho h_p) - 1 \right) \frac{D p}{D t}
+ + \frac{1}{\rho p_\rho} \left(
+  \frac{p_T}{\rho c_p} (\rho H - \rho \sum_k \xi_k   \omegadot_k)
+                               + \sum_k p_{X_k} \omegadot_k \right) \enskip .
+\end{eqnarray*}
+When we derived this expression we explicitly retained the dependence of $h$ on $p$,
+as shown by the presence of the $h_p$ term.
+
+Then, replacing $p$ by $p_0(r)$, $Dp/Dt$ becomes $\ubold \cdot
+\nabla p_0$, and, recalling the definition of $H$, the divergence constraint can
+be written
+\begin{equation}
+\nabla \cdot \ubold + \alpha \ubold \cdot \nabla p_0 =
+\frac{1}{\rho p_\rho} \left(
+   \frac{p_T}{\rho c_p} \left(\nabla \cdot (\kappa \nabla T)
+  - \sum_k\rho (q_k + \xi_k) \omegadot_k \right)
+ + \sum_k p_{X_k} \omegadot_k \right)  \equiv \tilde{S} \enskip \label{eq:full_divu_constraint} ,
+\end{equation}
+where we define
+\begin{equation}
+\alpha(\rho,T) \equiv - \left( \frac{(1 - \rho h_p )p_T - \rho c_p}{\rho^2
+  c_p p_\rho} \right) \enskip . \label{eq:alphadef}
+\end{equation}
+
+\section{Derivation of Constraint Using Energy Equation}
+
+We could also derive the constraint by again differentiating $p_{EOS} = p(\rho,T,X_k) = p_0$
+but this time writing $DT/Dt$ in terms of $De/Dt:$
+
+\begin{equation}
+\frac{DT}{Dt} = \frac{1}{\rho c_v} \left( -e_p \frac{D p}{D t}
+- \sum_k e_k \frac{D X_k}{D t} - p \nabla \cdot U 
++ \rho H \right) \enskip , \label{eq:dTdt}
+\end{equation}
+where $c_v = \left.\partial e/\partial T\right|_{p,X_k}$ is the
+specific heat at constant volume,
+and $e_p = \left.\partial e/\partial p\right|_{T,X_k}$
+for convenience.
+
+Then...
+
+\section{Enthalpy vs Energy Equation}
 
@@
 where $p_\mathrm{EOS}$ is the pressure returned from the EOS.  Note that
-$p_\mathrm{EOS}$ may not be equal to $p_0$---this causes us to drive
-off of the constraint. \MarginPar{not completely sure what to say here}
+$p_\mathrm{EOS}$ may not be equal to $p_0$---this may be what
+causes us to drive off of the constraint.
 
@@
 
-\item We always say that to substituting $p$ by $p_0$ in the enthalpy equation
-   is what filters the acoustics from the system---did we ever prove this?
+\item We substitute $p$ for $p_0$ in two places: 1) in the derivation of the
+constraint, 2) in the enthalpy evolution equation.   Are these substitutions
+consistent with each other?  If they are not consistent, which of these 
+substitutions is responsible for filtering the acoustics from the system?
 

### 2011-04-05 docs/enthalpy/enthalpy.tex
So close ... don't want to lose this.

@@
 
-\section{Derivation of Constraint Using Enthalpy Equation}
+\section{Derivation of Velocity Constraint}
 
-The velocity constraint equation was constructed by differentiating 
-the pressure along particle paths, using the equation of state.  
-Specifically, we wrote $p_{EOS} = p(\rho,T,X_k) = p_0$
-and expressed $DT/Dt$ in terms of $Dh/Dt:$
+Differentiating the equation of state, written in the form, $p =
+p(\rho,T,X_k),$ along particle paths, we can write
+\begin{equation}
+\frac{D p}{Dt}  = p_\rho \frac{D \rho}{Dt} + p_T \frac{D T}{Dt} + \sum_k p_{X_k} \frac{D X_k}{Dt}
+\end{equation}
+Then, by rearranging the terms, we get
+\begin{equation}
+\frac{D \rho}{Dt}  = -\rho \nabla \cdot \Ub = 
+    \frac{1}{p_\rho}
+    \left( \frac{D p}{Dt} - p_T \frac{D T}{Dt}
+                          - \sum_k p_{X_k} {\omegadot}_k \right) \enskip ,
+\end{equation}
+with $p_\rho = \left.\partial p/\partial \rho\right|_{X_k,T}$,
+$p_{X_k} = \left.\partial p/\partial X_k \right|_{T,\rho,(X_j,j\ne k)}$,
+and $p_T = \left.\partial p/\partial T\right|_{\rho,X_k}$.
+
+\subsection{Using the Enthalpy Equation}
+
+Now writing $h = h(p,T,X_k)$ and expanding $Dh/Dt$ and using the enthalpy evolution equation:
+\begin{equation}
+\rho \frac{D h}{Dt}  = \rho \left( h_p \frac{D p}{Dt} + c_p \frac{D T}{Dt} + \sum_k h_{X_k} \frac{D X_k}{Dt} \right)
+                     = \frac{D p}{D t} + \rho \Hnuc
+\end{equation}
+we can express $DT/Dt$ in terms of 
 \begin{equation}
 \frac{DT}{Dt} = \frac{1}{\rho c_p} \left( (1 - \rho h_p) \frac{D p}{D t}
-- \sum_k \rho \xi_k \omegadot_k + \rho H \right) \enskip , \label{eq:dTdt}
+- \sum_k \rho \xi_k \omegadot_k + \rho \Hnuc \right) \enskip , \label{eq:dTdt}
 \end{equation}
@@
 $\xi_k = \left.\partial h/\partial X_k \right|_{p,T}$,
-and $h_p = \left.\partial h/\partial p\right|_{T,X_k}$
-for convenience.
+and $h_p = \left.\partial h/\partial p\right|_{T,X_k}.$
 
-Then we derived
+We could then write,
 \begin{eqnarray*}
@@
 - \frac{D p}{D t} + \frac{p_T}{\rho c_p}
-  \left( (1 - \rho h_p) \frac{D p}{D t} - \rho \sum_k \xi_k \omegadot_k + \rho H \right)
+  \left( (1 - \rho h_p) \frac{D p}{D t} - \rho \sum_k \xi_k \omegadot_k + \rho \Hnuc \right)
 + \sum_k p_{X_k} \omegadot_k \right)  \\
@@
  + \frac{1}{\rho p_\rho} \left(
-  \frac{p_T}{\rho c_p} (\rho H - \rho \sum_k \xi_k   \omegadot_k)
+  \frac{p_T}{\rho c_p} (\rho \Hnuc - \rho \sum_k \xi_k   \omegadot_k)
                                + \sum_k p_{X_k} \omegadot_k \right) \enskip .
@@
 Then, replacing $p$ by $p_0(r)$, $Dp/Dt$ becomes $\ubold \cdot
-\nabla p_0$, and, recalling the definition of $H$, the divergence constraint can
-be written
+\nabla p_0$, and the divergence constraint can be written
 \begin{equation}
@@
 \frac{1}{\rho p_\rho} \left(
-   \frac{p_T}{\rho c_p} \left(\nabla \cdot (\kappa \nabla T)
-  - \sum_k\rho (q_k + \xi_k) \omegadot_k \right)
+   \frac{p_T}{\rho c_p} \left(
+  - \sum_k\rho  \xi_k \omegadot_k + \rho \Hnuc \right)
  + \sum_k p_{X_k} \omegadot_k \right)  \equiv \tilde{S} \enskip \label{eq:full_divu_constraint} ,
@@
 
-\section{Derivation of Constraint Using Energy Equation}
+\subsection{Using the Energy Equation}
 
-We could also derive the constraint by again differentiating $p_{EOS} = p(\rho,T,X_k) = p_0$
-but this time writing $DT/Dt$ in terms of $De/Dt:$
-
+Now writing $e = e(p,T,X_k)$ and expanding $De/Dt$ and using the energy evolution equation:
 \begin{equation}
-\frac{DT}{Dt} = \frac{1}{\rho c_v} \left( -e_p \frac{D p}{D t}
-- \sum_k e_k \frac{D X_k}{D t} - p \nabla \cdot U 
-+ \rho H \right) \enskip , \label{eq:dTdt}
+\rho \frac{D e}{Dt}  = \rho \left( e_p \frac{D p}{Dt} + c_v \frac{D T}{Dt} + \sum_k e_{X_k} \frac{D X_k}{Dt} \right)
+                     = -p \nabla \cdot \Ub + \rho \Hnuc
 \end{equation}
+we can express $DT/Dt$ in terms of 
+\begin{eqnarray*}
+\frac{DT}{Dt} &=& \frac{1}{\rho c_v} \left(
+                  -p \nabla \cdot \Ub + \rho \Hnuc
+                 - \rho e_p \frac{D p}{Dt}
+                 - \rho \sum_k e_{X_k} \omegadot_k   \right)
+\end{eqnarray*}
 where $c_v = \left.\partial e/\partial T\right|_{p,X_k}$ is the
 specific heat at constant volume,
-and $e_p = \left.\partial e/\partial p\right|_{T,X_k}$
-for convenience.
+and $e_p = \left.\partial e/\partial p\right|_{T,X_k}.$
+Then
+\begin{equation}
+-\rho \nabla \cdot \Ub = 
+    \frac{1}{p_\rho} \left( \frac{D p}{Dt} 
+    - \frac{p_T}{\rho c_v} \left(
+                 -p \nabla \cdot \Ub + \rho \Hnuc
+                 - \rho e_p \frac{D p}{Dt}
+                 - \rho \sum_k e_{X_k} \omegadot_k   \right)
+    - \sum_k p_{X_k} {\omegadot}_k \right) \enskip ,
+\end{equation}
+which leads to 
+\begin{equation}
+\left(-\rho -  \frac{p p_T}{\rho c_v p_\rho} \right) \nabla \cdot \Ub = 
+    \frac{1}{p_\rho} \left( (1 + \frac{e_p p_T}{c_v}) \frac{D p}{Dt} 
+                           - \frac{p_T}{\rho c_v} \left(
+                                        \rho \Hnuc
+                                        - \rho \sum_k e_{X_k} \omegadot_k   \right)
+    - \sum_k p_{X_k} {\omegadot}_k \right) \enskip ,
+\end{equation}
 
-Then...
+\section{Comparison of Constraints}
+
+If we set $\omegadot_k = \Hnuc = 0$ for simplicity, then the constraint as derived using $h$ can be written 
+\begin{equation}
+\nabla \cdot \ubold
+- \left( \frac{(1 - \rho h_p )p_T - \rho c_p}{\rho^2
+  c_p p_\rho} \right) \frac{D p_0}{D t} = 0
+\end{equation}
+and the constraint derived using $e$ can be written
+\begin{equation}
+\left(\rho p_\rho + \frac{p p_T}{\rho c_v} \right) \nabla \cdot \Ub +
+      (1 + \frac{e_p p_T}{c_v}) \frac{D p}{Dt} = 0 \enskip .
+\end{equation}
 

### 2011-04-05 docs/enthalpy/enthalpy.tex
Finished section of derviation energy equation .

@@
 \begin{eqnarray*}
-\nabla \cdot \ubold &=& \frac{1}{\rho p_\rho} \left(
+\nabla \cdot \Ub &=& \frac{1}{\rho p_\rho} \left(
 - \frac{D p}{D t} + \frac{p_T}{\rho c_p}
@@
 
-Then, replacing $p$ by $p_0(r)$, $Dp/Dt$ becomes $\ubold \cdot
+Then, replacing $p$ by $p_0(r)$, $Dp/Dt$ becomes $\Ub \cdot
 \nabla p_0$, and the divergence constraint can be written
 \begin{equation}
-\nabla \cdot \ubold + \alpha \ubold \cdot \nabla p_0 =
+\nabla \cdot \Ub + \alpha \Ub \cdot \nabla p_0 =
 \frac{1}{\rho p_\rho} \left(
@@
 \begin{equation}
-\rho \frac{D e}{Dt}  = \rho \left( e_p \frac{D p}{Dt} + c_v \frac{D T}{Dt} + \sum_k e_{X_k} \frac{D X_k}{Dt} \right)
+\rho \frac{D e}{Dt}  = \rho \left( e_p \frac{D p}{Dt} + e_T \frac{D T}{Dt} + \sum_k e_{X_k} \frac{D X_k}{Dt} \right)
                      = -p \nabla \cdot \Ub + \rho \Hnuc
@@
 \begin{eqnarray*}
-\frac{DT}{Dt} &=& \frac{1}{\rho c_v} \left(
+\frac{DT}{Dt} &=& \frac{1}{\rho e_T} \left(
                   -p \nabla \cdot \Ub + \rho \Hnuc
@@
 \end{eqnarray*}
-where $c_v = \left.\partial e/\partial T\right|_{p,X_k}$ is the
+where $e_T = \left.\partial e/\partial T\right|_{p,X_k}$ is the
 specific heat at constant volume,
@@
     \frac{1}{p_\rho} \left( \frac{D p}{Dt} 
-    - \frac{p_T}{\rho c_v} \left(
+    - \frac{p_T}{\rho e_T} \left(
                  -p \nabla \cdot \Ub + \rho \Hnuc
@@
 \begin{equation}
-\left(-\rho -  \frac{p p_T}{\rho c_v p_\rho} \right) \nabla \cdot \Ub = 
-    \frac{1}{p_\rho} \left( (1 + \frac{e_p p_T}{c_v}) \frac{D p}{Dt} 
-                           - \frac{p_T}{\rho c_v} \left(
+\left(-\rho -  \frac{p p_T}{\rho e_T p_\rho} \right) \nabla \cdot \Ub = 
+    \frac{1}{p_\rho} \left( (1 + \frac{e_p p_T}{e_T}) \frac{D p}{Dt} 
+                           - \frac{p_T}{\rho e_T} \left(
                                         \rho \Hnuc
@@
 \begin{equation}
-\nabla \cdot \ubold
+\nabla \cdot \Ub
 - \left( \frac{(1 - \rho h_p )p_T - \rho c_p}{\rho^2
@@
 \begin{equation}
-\left(\rho p_\rho + \frac{p p_T}{\rho c_v} \right) \nabla \cdot \Ub +
-      (1 + \frac{e_p p_T}{c_v}) \frac{D p}{Dt} = 0 \enskip .
+\nabla \cdot \Ub 
++ \left( \frac{\rho e_T + \rho e_p p_T}{\rho^2 e_T p_\rho + p p_T} \right) \frac{D p_0}{D t} = 0
 \end{equation}

### 2011-04-05 docs/enthalpy/enthalpy.tex
Fix sign error in (12)
More comments by Ann.

@@
 \nabla \cdot \Ub
-- \left( \frac{(1 - \rho h_p )p_T - \rho c_p}{\rho^2
++ \left( \frac{(1 - \rho h_p )p_T - \rho c_p}{\rho^2
   c_p p_\rho} \right) \frac{D p_0}{D t} = 0
@@
 + \left( \frac{\rho e_T + \rho e_p p_T}{\rho^2 e_T p_\rho + p p_T} \right) \frac{D p_0}{D t} = 0
+\end{equation}
+
+We note that if we evaluate both constraints for $p = \rho R T,$ with $h = c_p T,$ $e = c_v T,$
+$c_p = c_v + R$ and $\gamma = c_p / c_v,$ then both constraints reduce to 
+\begin{equation}
+\nabla \cdot \Ub + \frac{1}{\gamma p} \frac{D p_0}{D t} = 0
 \end{equation}
@@
 \end{eqnarray}
-We note that the last equation appears to be an internal energy
-evolution equation, but with an additional forcing term: $D(p_0 - p)/Dt$.
+However, if we solve the evolution equation for $e$ we would substitute 
+$p_0$ for $p$ in this equation as well.  Thus, we can pose the situation as the following.  
+If we solve the evolution equation for $h$ then we effectively are solving
+\begin{equation}
+\frac{\partial(\rho e)}{\partial t} +
+  \nabla\cdot(\rho e\Ub) = p \; \nabla\cdot\Ub + \rho\Hnuc + 
+  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} \nonumber 
+\end{equation}
+but if we solve the evolution equation for $e$ we are effectively solving
+\begin{equation}
+\frac{\partial(\rho e)}{\partial t} +
+  \nabla\cdot(\rho e\Ub) = p_0 \nabla\cdot\Ub + \rho\Hnuc \nonumber 
+\end{equation}
+
+The second equation subtracted from the first gives:
+\begin{equation}
+  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} + (p - p_0)  \nabla \cdot \Ub
+\end{equation}
+
+Suppose we solve the current enthalpy equation, but when we call the EOS, we 
+subtract $p_0$ from $\rho h$ and then call the EOS with $e$ instead of $h$.  This is equivalent
+to:
+\begin{eqnarray}
+\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) +
+  \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
+%
+\frac{\partial(\rho e)}{\partial t} + \frac{\partial p_0}{\partial t} &=&
+ -\nabla\cdot(\rho e\Ub) -\nabla\cdot(p_0 \Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
+%
+\frac{\partial(\rho e)}{\partial t} &=&
+ -\nabla\cdot(\rho e\Ub) - p_0 \nabla\cdot\Ub + \rho\Hnuc  \nonumber
+\end{eqnarray}
+which is identical to solving the energy equation.
 
@@
 
-\item We substitute $p$ for $p_0$ in two places: 1) in the derivation of the
-constraint, 2) in the enthalpy evolution equation.   Are these substitutions
-consistent with each other?  If they are not consistent, which of these 
-substitutions is responsible for filtering the acoustics from the system?
+\item Should we call the EOS with $h$ as is, or call the EOS with $e = h - p_0 / \rho$?
 
-\item In the EOS, we take $h$ and construct $e$ by subtracting off
-  $p$---depending on the inputs to the EOS, we treat the `$p$' part of $h$
-  differently.
+\item When we stay on the constraint, i.e. $p_{EOS} = p_0$, then the equations for $e$ and
+for $h$ are identical.   However, once we are off the constraint, do the terms in the
+current evolution equation for $h$ serve to drive us back to the constraint?  Recall
+our current "volume discrepancy factor" acts as a source term which modifies the
+divergence constraint, which effectively modifies both $\rho$ and $T$ (or $e$ or $h$).
+The term in the enthalpy equation only modifies $\rho.$  Is this relevant and/or useful??
 
-   if we come into the EOS with $\rho$, $T$, then we get $p$ and $e$ from
-   the EOS and compute $h$.
+%\item We substitute $p$ for $p_0$ in two places: 1) in the derivation of the
+%constraint, 2) in the enthalpy evolution equation.   Are these substitutions
+%consistent with each other?  If they are not consistent, which of these 
+%substitutions is responsible for filtering the acoustics from the system?
 
-   if we come into the EOS with $\rho$, $h$, then we find a $T$ such that
-   $e + p(\rho,T)/\rho = h$, where $p(\rho,T)$ is the pressure from the EOS.
-   It is not necessarily equal to $p_0$.
+\item Suppose we corrected the $h$ (or $e$ equation) by using the full $p$ instead
+of $p_0$?  Would this be more or less consistent (one could imagine doing this as a 
+correction after solving for $\pi$ earlier in the timestep).
 
-   if we come into the EOS with $\rho$, $p_0$, then find a $T$ that matches
-   $p_0$, and then $h = e + p_0/\rho$.  In this case, there is no mismatch
-   in the enthalpy equation.
-
-\item do we want to always insist that the `$p$' hidden in the definition of
-   $h$ is $p_0$?
+%\item In the EOS, we take $h$ and construct $e$ by subtracting off
+%  $p$---depending on the inputs to the EOS, we treat the `$p$' part of $h$
+%  differently.
+%
+%   if we come into the EOS with $\rho$, $T$, then we get $p$ and $e$ from
+%   the EOS and compute $h$.
+%
+%   if we come into the EOS with $\rho$, $h$, then we find a $T$ such that
+%   $e + p(\rho,T)/\rho = h$, where $p(\rho,T)$ is the pressure from the EOS.
+%   It is not necessarily equal to $p_0$.
+%
+%   if we come into the EOS with $\rho$, $p_0$, then find a $T$ that matches
+%   $p_0$, and then $h = e + p_0/\rho$.  In this case, there is no mismatch
+%   in the enthalpy equation.
 

### 2011-04-05 docs/enthalpy/enthalpy.tex
Add a little more commentary.
More about comparison of current extra terms to volume discrepancy factor.

@@
 \end{eqnarray*}
-where $e_T = \left.\partial e/\partial T\right|_{p,X_k}$ is the
-specific heat at constant volume,
+where $e_T = \left.\partial e/\partial T\right|_{p,X_k}$ 
 and $e_p = \left.\partial e/\partial p\right|_{T,X_k}.$
@@
 \end{equation}
+Note that we can replace $p$ by $p_0$ in the coefficient on the l.h.s. as well as on the r.h.s.
 
@@
 \begin{equation}
-  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} + (p - p_0)  \nabla \cdot \Ub
+  -\frac{D (p - p_0)}{Dt} + (p - p_0)  \nabla \cdot \Ub
 \end{equation}
@@
 The term in the enthalpy equation only modifies $\rho.$  Is this relevant and/or useful??
+Recall that the current "volume discrepancy factor" takes the form of adding 
+to the r.h.s. of the constraint:
+\begin{equation}
+\nabla \cdot(\beta_0\Ub) = \beta_0\left(S-\frac{1}{\overline{\Gamma_1} p_0}
+       \frac{\partial p_0}{\partial t} - \frac{f}{\overline{\Gamma_1} p_0}
+       \frac{p_0-p_\text{EOS}}{\dt}\right) 
+\end{equation}
 

### 2011-04-06 docs/enthalpy/enthalpy.tex
checkin some notes about constant gamma EOS

@@
 The second equation subtracted from the first gives:
-\begin{equation}
-  -\frac{D (p - p_0)}{Dt} + (p - p_0)  \nabla \cdot \Ub
+\begin{equation} \label{eq:difference between h and e equations}
+  \frac{D (p_0 - p)}{Dt} - (p_0 - p)  \nabla \cdot \Ub
 \end{equation}
@@
 which is identical to solving the energy equation.
+
+\subsection{Constant $\gamma$ Gas}
+Going back to the constant $\gamma$, ideal gas EOS, we can rewrite the
+enthalpy equation as a pressure evolution equation
+\begin{eqnarray}
+  \frac{\partial\rho h}{\partial t} + \nabla\cdot\left(\rho h\Ub\right) &=& \frac{Dp_0}{Dt} + \rho H {} \nonumber\\
+  \frac{\gamma}{\gamma-1}\frac{\partial p}{\partial t} + \frac{\gamma}{\gamma-1}\nabla\cdot\left(p\Ub\right) &=& \frac{Dp_0}{Dt} + \rho H {} \nonumber\\
+  \frac{Dp}{Dt} &=& -p\nabla\cdot\Ub + \frac{\gamma}{\gamma-1}\left(\frac{Dp_0}{Dt}+\rho H\right) \label{eq:H:pressure evolution}.
+\end{eqnarray}
+Similarly, we can derive a pressure evolution equation from the energy equation
+\begin{equation}
+  \frac{Dp}{Dt} = -p\nabla\cdot\Ub - \left(\gamma-1\right)p_0\nabla\cdot\Ub + \rho H \label{eq:e:pressure evolution}
+\end{equation}
+Now, if we further make the assumption that $p_0$ is constant,
+$Dp_0/Dt = 0$, then the divergence constraint for such a gas reduces
+to
+\begin{equation}
+\nabla\cdot\Ub = \frac{\gamma-1}{\gamma p_0}H \label{eq:div constraint for constant gamma}.
+\end{equation}
+Plugging this back into either of \eqref{eq:H:pressure evolution} or \eqref{eq:e:pressure evolution} gives
+\begin{equation}
+  \frac{Dp}{Dt} = \frac{\gamma-1}{\gamma}\left(1-\frac{p}{p_0}\right)H. \label{eq:pressure evolution constant gamma}
+\end{equation}
+If $p_0$ is assumed constant and using \eqref{eq:div constraint for
+  constant gamma}, the difference between the enthalpy equation and
+the energy equation, \eqref{eq:difference between h and e equations},
+can be rewritten as
+\[
+-\frac{Dp}{Dt} - \frac{\gamma}{\gamma-1}\left(1-\frac{p}{p_0}\right)H = 0,
+\]
+where the equality holds from \eqref{eq:pressure evolution constant gamma}.
 

### 2011-04-07 docs/enthalpy/enthalpy.tex
was missing a factor of rho
fixed sign error and added some commentary

@@
 \frac{\partial(\rho e)}{\partial t} +
-  \nabla\cdot(\rho e\Ub) = p \; \nabla\cdot\Ub + \rho\Hnuc + 
+  \nabla\cdot(\rho e\Ub) = -p \; \nabla\cdot\Ub + \rho\Hnuc + 
   \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} \nonumber 
@@
 \begin{equation} \label{eq:difference between h and e equations}
-  \frac{D (p_0 - p)}{Dt} - (p_0 - p)  \nabla \cdot \Ub
+  \frac{D (p_0 - p)}{Dt} - (p_0 - p)  \nabla \cdot \Ub = 0,
 \end{equation}
+but this equation is only true, in general, if $p=p_0$.
 
@@
 \end{eqnarray}
-which is identical to solving the energy equation.
+which is identical to solving the energy equation with $p\to p_0$.
 
@@
 \begin{equation}
-\nabla\cdot\Ub = \frac{\gamma-1}{\gamma p_0}H \label{eq:div constraint for constant gamma}.
+\nabla\cdot\Ub = \frac{\gamma-1}{\gamma p_0}\rho H \label{eq:div constraint for constant gamma}.
 \end{equation}
@@
 \begin{equation}
-  \frac{Dp}{Dt} = \frac{\gamma-1}{\gamma}\left(1-\frac{p}{p_0}\right)H. \label{eq:pressure evolution constant gamma}
+  \frac{Dp}{Dt} = \frac{\gamma-1}{\gamma}\left(1-\frac{p}{p_0}\right)\rho H. \label{eq:pressure evolution constant gamma}
 \end{equation}
@@
 \[
--\frac{Dp}{Dt} - \frac{\gamma}{\gamma-1}\left(1-\frac{p}{p_0}\right)H = 0,
+-\frac{Dp}{Dt} - \frac{\gamma}{\gamma-1}\left(1-\frac{p}{p_0}\right)\rho H = 0,
 \]
-where the equality holds from \eqref{eq:pressure evolution constant gamma}.
+where the equality holds from \eqref{eq:pressure evolution constant
+  gamma}. In other words, for the constant $\gamma$ gas we have
+$p=p_0$ as expected.
 

### 2011-04-23 docs/analysis/analysis.tex
add the analysis chapter to the main document
update and add particle stuff

@@
 \section{The {\tt data\_processing} Routines}
+\label{sec:analysis}
 
@@
 
+\item {\tt fvarnames.f90}
+
+  Simply print out the list of variables stored in a plotfile.
+
 \subsection{Data Processing Example}
@@
 
+\subsection{Particle routines}
+
+The {\tt parseparticles.py} routine in the {\tt python/} subdirectory
+can read in MAESTRO particle files containing particle histories
+(usually named {\tt timestamp\_??}.  The driver {\tt test.py} shows
+shows how to use this module to plot particle histories.
+
 \end{itemize}

### 2011-04-23 docs/architecture/architecture.tex
fix directory structure -- docs/ was shown in the wrong place
some updates

@@
  of boxes.  The BoxLib modules define the basic datatypes used
- in MAESTRO.
+ in MAESTRO.  BoxLib also provides the routines that handle the
+ parallelization and I/O.
 
@@
  multidimensional dataset) that operate on BoxLib datasets.
-
-\item {\tt doc/}
-
- Documentation describing the basic algorithm.
 
@@
 
+ \item {\tt docs/}
+
+   Documentation describing the basic algorithm (including this
+   document).
+
+ \item {\tt initial\_models}
+
+   Several sub-directories containing routines to generate
+   1-d initial conditions in hydrostatic equilibrium for 
+   mapping onto the MAESTRO grid.  See \S~\ref{sec:initial_models}
+   for details.
+
  \item {\tt Source}
@@
    source.
-
- \item {\tt initial\_models}
-
-   Several sub-directories containing routines to generate
-   1D initial conditions in hydrostatic equilibrium for 
-   mapping onto the MAESTRO grid.  See \S~\ref{sec:initial_models}
-   for details.
 
@@
   The generic Makefiles that store the compilation flags for
-  various platforms.
+  various platforms.  Platform/compiler-specific options are stored
+  in the {\tt comps/} subdirectory.
 
@@
 runtime parameter can be set either in the inputs file, as {\tt parameter
-  = value} or on the command line by adding {\tt --parameter value} after
+  = value}, or on the command line by adding {\tt --parameter value} after
 listing the inputs file on the command line.  A full list of runtime
@@
       value of this parameter should be the file number to restart from.
+      For example, to restart from the checkpoint file {\tt chk00010},
+      you would set {\tt restart = 10}.
 
@@
 where $\Delta x$ on the base grid is computed as $({\tt prob\_hi\_x}
-- {\tt prob\_lo\_x})/{\tt n\_cellx}$.
+- {\tt prob\_lo\_x})/{\tt n\_cellx}$.  For multilevel problems, the effective
+number of zones on the finest grid in the $x$ direction will be
+${\tt n\_cellx} \cdot 2^{({\tt max\_levels} -1)}$.
 
@@
 in {\tt fParallel/data\_processing}.  The {\tt GNUmakefile} there
-needs to be edited to indicate which of the tools to build.  These
-routines are described in \S~\ref{sec:analysis}.
+needs to be edited to indicate which of the tools to build.  For
+example, to extract the density along a line from the center of a
+plotfile, {\tt plt00200}, in the $y$-direction:
+
+\begin{verbatim}
+fextract.Linux.gfortran.exe -d 2 -v "density" -p plt00200
+\end{verbatim}
+
+These routines are described in \S~\ref{sec:analysis}.
 
@@
   position.
+
+  Particle data is stored separately from the fluid data.  Instead
+  of being part of the plotfiles, the particle data is outputted
+  each timestep into files named {\tt timestamp\_*}, where 
+  the number indicates which processor did the writing.  These
+  particle files can be processed and the particle data plotted
+  using the python routines in {\tt data\_processing/python/}.
 
@@
 
-A MAESTRO application is built from several pacakges (the
+A MAESTRO application is built from several packages (the
 multigrid solver, an EOS, a reaction network, etc.).  The core

### 2011-04-23 docs/architecture/architecture.tex
rework the description of a box

@@
 \end{equation}
+Here, $g_{i+1/2}$ is the edge-centered gravitational acceleration.
 
@@
 finer zones (typically $2\times$).  Only a portion of the domain may
-be covered by the higher levels of refinement.  \MarginPar{Nesting?}
+be covered by the higher levels of refinement.
+
+\MarginPar{Nesting?}
 For parallel computations, the boxes are spread across processors, in
@@
 A \boxtype\ is simply a rectangular domain in space.  Note that boxes
-do not hold the state data themselves.  The datatype for a
-\boxtype\ is
-\begin{verbatim}
-  type box
-     integer :: dim  = 0
-     integer :: lo(MAX_SPACEDIM) =  Huge(1)
-     integer :: hi(MAX_SPACEDIM) = -Huge(1)
-  end type box
-\end{verbatim}
-
-\noindent Here:
-\begin{itemize}
-\item {\tt dim} is the dimensionality of the data
-\item {\tt lo} is the index of the lower bound of the box
-\item {\tt hi} is the index of the upper bound of the box
-\end{itemize}
+do not hold the state data themselves.  A \boxtype\ has a {\tt lo} 
+and {\tt hi} index in each coordinate direction that gives the
+location of the lower-left and upper-right corner with respect to
+a global index space.
 
@@
 Figure~\ref{fig:boxes} shows three boxes in the same level of
-refinement.  The position of the boxes is with respect to a global
+refinement.  The position of the boxes is with respect to the global
 index space at that level.  For example, box 1 in the figure has {\tt
-  lo} = (3,7) and {\tt hi} = (9,12).
+  lo} = (3,7) and {\tt hi} = (9,12).  Note that the global indexing
+is 0-based.
+
+The global index space covers the entire domain at a given resolution.
+For a simulation setup with {\tt n\_cellx = 32} and {\tt n\_celly =
+  32}, the coarsest level (level 1) has $32 \times 32$ zones, and the
+global index space will run from $0, \ldots, 31$ in each coordinate
+direction.  Level 2 will have a global index space running from $0,
+\ldots, 63$ in each coordinate direction (corresponding to $64 \times
+64$ zones if fully refined), and level 3 will have a global index
+space running from $0, \ldots, 127$ in each coordinate direction
+(corresponding to $128\times 128$ zones if fully refined).
 

### 2011-04-24 docs/architecture/architecture.tex
more boxlib descriptions

@@
 
-\MarginPar{Nesting?}
+\MarginPar{Nesting? ghost cells?}
 For parallel computations, the boxes are spread across processors, in
@@
 
-For a \boxtype\ declared as:
+A \boxtype\ declared as:
 \begin{verbatim}
@@
 \end{verbatim}
-
+%
+The upper and lower bounds of the box (in terms of the global
+index space) are found via:
 \begin{itemize}
@@
 \end{itemize}
-
-\subsection{\fab}
-
-A \fab\ is a ``Fortran Array Box''.  It contains the state data in a
-multidimensional array and several \boxtype-types to describe where in
-index-space it lives.  The datatype of a \fab\ is
-\begin{verbatim}
-  type fab
-     integer   :: dim = 0
-     type(box) :: bx
-     type(box) :: pbx
-     type(box) :: ibx
-     integer   :: nc = 1
-     real(kind = dp_t), pointer, dimension(:,:,:,:) :: p => Null()
-  end type fab
-\end{verbatim}
-
-\noindent Here:
-\begin{itemize}
-\item {\tt dim} is the dimensionality of the data
-\item {\tt bx} is a \boxtype\ describing the index space for which this \fab\ is defined
-\item {\tt ibx} is a \boxtype\ that describes the index range of the valid data
-  for the \fab.  This differes from {\tt bx} when the box is nodal in one or more 
-  directions.
-\item {\tt pbx} is a \boxtype\ that extends {\tt ibx} to include
-  any ghostcells, or equivalently, the physical box for the \fab\
-\item {\tt nc} is the number of components (how many variables at  
-  each grid location)
-\item {\tt p} is a pointer to the data, in this case double precision
-  ({\tt kind=dp\_t}).  Other \fab\ types exists for other Fortran
-  data types.
-\end{itemize}
-
-Note that all state data is stored in a four-dimensional array,
-{\tt (nx,ny,nz,nc)} in size, regardless of the dimensionality of the
-problem.  For 2D problems, {\tt nz=1}.
-
-In MAESTRO, we don't usually deal with \fab s alone, but rather
-through \multifab s, described next.
-
-\subsection{\multifab}
-
-A \multifab\ is a collection of \fab s at the same level of
-refinement.  The data type of a \multifab\ is:
-\begin{verbatim}
-  type multifab
-     integer :: dim = 0
-     integer :: nboxes = 0
-     integer :: nc = 1
-     integer :: ng = 0
-     logical, pointer :: nodal(:) => Null()
-     type(layout) :: la
-     type(fab), pointer :: fbs(:) => Null()
-  end type multifab
-\end{verbatim}
 
@@
 
+\subsection{\fab}
+
+A \fab\ is a ``Fortran Array Box''.  It contains the state data in a
+multidimensional array and several \boxtype-types to describe where in
+the global index-space it lives.  
+Note that all state data is stored in a four-dimensional array,
+{\tt (nx,ny,nz,nc)} in size, regardless of the dimensionality of the
+problem.  Here {\tt nc} is the number of components, for instance
+representing different fluid variables.  For 2D problems, {\tt nz=1}.
+
+A \fab\ would represent the data for a single box in the domain.
+In MAESTRO, we don't usually deal with \fab s alone, but rather
+we deal with \multifab s, described next.
+
+\subsection{\multifab}
+
+A \multifab\ is a collection of \fab s at the same level of
+refinement.  This is the primary data structure that MAESTRO
+routine operate on.  A multilevel simulation stores the 
+data in an array of multifabs, where the array index refers
+to the refinement level.
+
+\subsubsection{Working with \multifab s}
+
+To build a \multifab, we need to provide a \layout, the number of
+components to store in the multifab, and the number of ghostcells.  In
+MAESTRO, the hierarchry of grids will be described by a single
+\mllayout.  A \multifab\ can be declared and built at any time in a
+simulation using the \mllayout, thereby allocating space at every
+grid location in the simulation.  The sequence to build a \multifab\
+appears as
+\begin{verbatim}
+  type(multifab) :: mfab(nlevs)
+  ...
+  do n = 1, nlevs
+     call multifab_build(mfab(n), mla%la(n), nc, ng)
+  enddo
+\end{verbatim}
+Here, {\tt nc} is the number of components and {\tt ng} is the number
+of ghostcells.  The \multifab\ is built one level at a time, using the
+\layout\ for that level taken from the \mllayout, {\tt mla}.
+
+A common operation on a \multifab\ is to initialize it to $0$
+everywhere.  This can be done (level-by-level) as
+\begin{verbatim}
+call setval(mfab(n), ZERO, all=.true.)
+\end{verbatim}
+where {\tt ZERO} is the constant 0.0 from {\tt bl\_constants\_module}.
+
+Subroutines to add, multiply, or divide two \multifab s exist, as do
+subroutines to copy from one \multifab\ to another---see {\tt boxlib/multifab.f90}
+for the full list of routines that work with \multifab s.
+
+When you are done working with a \multifab, its memory can be freed by
+calling {\tt multifab\_destroy} on the \multifab.
+
 \subsection{\bctower}
+
+A \bctower holds the information about what boundary conditions are
+in effect for each variable in a MAESTRO simulation.  \MarginPar{better description?}
 

### 2011-04-24 docs/architecture/architecture.tex
more work filling in details

@@
 appears as
-\begin{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
   type(multifab) :: mfab(nlevs)
@@
   enddo
-\end{verbatim}
+\end{lstlisting}
 Here, {\tt nc} is the number of components and {\tt ng} is the number
@@
 everywhere.  This can be done (level-by-level) as
-\begin{verbatim}
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
 call setval(mfab(n), ZERO, all=.true.)
-\end{verbatim}
+\end{lstlisting}
 where {\tt ZERO} is the constant 0.0 from {\tt bl\_constants\_module}.
 
+The procedure for accessing the data in each grid managed by the \multifab\
+is shown in \S~\ref{sec:example}.  
 Subroutines to add, multiply, or divide two \multifab s exist, as do
@@
 
-\subsection{Other Quantities}
-
 \section{MAESTRO Helper Modules}
@@
 \subsection{\tt average}
+
+The {\tt average} module provides a routine {\tt average} that takes
+a multilevel \multifab\ array and averages the full Cartesian data
+onto the 1-d base state.
 
@@
 
-eos module variables
-
-eos input types
+The {\tt eos\_module} declares for convenience the variables that need
+appear in the {\tt eos} call argument list.  MAESTRO routines use
+these module variables in the EOS call to avoid having to declare 
+each quantity in each routine that calls the EOS.
+
+The first argument to the {\tt eos} call is an integer key that
+specifies which thermodynamic variables (in addition to the mass
+fractions) are used as input.  Options include:
+
+   \begin{center}
+   \begin{tabular}{lc}
+   \hline
+   key            & input quantities \\
+   \hline
+   eos\_input\_rt       & $\rho$, $T$ \\
+   eos\_input\_rh       & $\rho$, $h$ \\
+   eos\_input\_tp       & $T$, $p$ \\
+   eos\_input\_rp       & $\rho$, $p$ \\
+   eos\_input\_re       & $\rho$, $e$ \\
+   eos\_input\_ps       & $p$, $s$ \\
+   \hline
+   \end{tabular}
+   \end{center}
 
 \subsection{{\tt fill\_3d\_module}}
+
+The {\tt fill\_3d\_module} provides routines that map from the 1-d
+base state to the full Cartesian 2- or 3-d state.  Variations in the
+routines allow for cell-centered or edge-centered data on either the
+base state or full Cartesian state.
 
@@
 
-\section{Example: Accessing State Data}
+\section{\label{sec:example} Example: Accessing State Data}
 
@@
 {\tt hi(3)}).
-
-\section{Working with Multifabs}
-
-need to describe the routines for adding two multifabs, subtracting,
-etc. and norm, etc.
 

### 2011-05-27 docs/architecture/architecture.tex
tweaks
add some notes on particles

@@
 
-MAESTRO's gridding is handled by the BoxLib library.  The underlying
-idea in BoxLib is to allow for adaptive mesh refinement (AMR)---different
-regions of the domain can have different spatial resolutions.  The
-domain is broken into boxes.  At the coarsest (base) level of
-refinement, the entire computational domain is covered, either by one
-box or broken across many boxes.  Higher levels of refinement have
-finer zones (typically $2\times$).  Only a portion of the domain may
-be covered by the higher levels of refinement.
-
-\MarginPar{Nesting? ghost cells?}
-For parallel computations, the boxes are spread across processors, in
-a fashion designed to put roughly equal amounts of work on each
-processor (load balancing).
-
-On a grid, the data can be stored at cell-centers, on a face/edge, or
-on the corners.  In BoxLib, data that is on an edge is termed `nodal'
-in that direction (see Figure~\ref{fig:dataloc}).  Data that is on the
-corners is nodal in all spatial directions.  In MAESTRO, the state
-data (density, enthalpy, velocity, $\ldots$) is generally
-cell-centered.  Fluxes are nodal in the direction they represent.
-A few quantities are nodal in all directions (e.g.\ $\phi$ used in
-the final velocity projection).
-
 \begin{figure}[t]
@@
 \end{figure}
+
+MAESTRO's gridding is handled by the BoxLib library.  The underlying
+idea in BoxLib is to allow for adaptive mesh refinement (AMR)---different
+regions of the domain can have different spatial resolutions.  The
+domain is broken into boxes.  At the coarsest (base) level of
+refinement, the entire computational domain is covered, either by one
+box or broken across many boxes.  Higher levels of refinement have
+finer zones (typically $2\times$).  Only a portion of the domain may
+be covered by the higher levels of refinement.
+
+\MarginPar{Nesting? ghost cells?}
+For parallel computations, the boxes are spread across processors, in
+a fashion designed to put roughly equal amounts of work on each
+processor (load balancing).
+
+On a grid, the data can be stored at cell-centers, on a face/edge, or
+on the corners.  In BoxLib, data that is on an edge is termed `nodal'
+in that direction (see Figure~\ref{fig:dataloc}).  Data that is on the
+corners is nodal in all spatial directions.  In MAESTRO, the state
+data (density, enthalpy, velocity, $\ldots$) is generally
+cell-centered.  Fluxes are nodal in the direction they represent.
+A few quantities are nodal in all directions (e.g.\ $\phi$ used in
+the final velocity projection).
 
@@
 
-A \bctower holds the information about what boundary conditions are
+A \bctower\ holds the information about what boundary conditions are
 in effect for each variable in a MAESTRO simulation.  \MarginPar{better description?}
@@
 \section{Particles}
+
+MAESTRO has support for Lagrangian particles that are passively
+advected with the velocity field.  These are useful for diagnostics
+and post-processing.  To use particles, particles must be seeded into
+the domain by writing a problem-specific {\tt init\_particles.f90}
+routine.  This routine is called at the start of the simulation.  The
+{\tt init\_particles} routines add particles at specific locations by
+calling the {\tt particle\_module}'s {\tt add} routine when a given
+criteria is met by the fluid state.
+
+When you run the code, particles are enabled by setting {\tt
+  use\_particles = T}.  At the end of each timestep the locations of
+all the particles are written out into a series of files called {\tt
+  timestamp\_NN}, where {\tt NN} is the CPU number on which the
+particle {\em currently} resides.  Particles are always kept on the
+processor containing the state data corresponding to their present
+location.  Several bits of associated data (density, temperature, and
+mass fractions) are stored along with the particle ID and position.
+
+Some simple python scripts allow for the plotting of the particle
+positions.  For example, running {\tt test2} with particles enabled
+will seed particles in the initial hotspot.  To plot the results,
+first set your {\tt PYTHONPATH} environment variable to point to the
+{\tt fParallel/data\_processing/python/} directory, for example:
+\begin{verbatim}
+export PYTHONPATH="/home/username/development/MAESTRO/fParallel/data_processing/python"
+\end{verbatim}
+This will allow python to see the {\tt parseparticles.py} routine.
+For the {\tt test2} problem, the {\tt plotparticles.py} routine shows
+how to plot the particle histories and make an animation of the
+particles colored by the ash mass fraction.  This script is run as:
+\begin{verbatim}
+./plotparticles.py timestamp_*
+\end{verbatim}
+
+Note, these python routines require the NumPy and matplotlib packages.
+On a Fedora Linux system, the necessary routines can be installed via:
+\begin{verbatim}
+yum install python-matplotlib lyx-fonts stix-fonts
+\end{verbatim}

### 2011-06-02 docs/SDC/sdc.tex
add practical SDC implementation notes describing how the burner must change

\section{Implementing an SDC-enabled reaction network (no diffusion)}

Maestro integrates the following system of equations:
\begin{eqnarray}
\frac{\partial(\rho X_k)}{\partial t} &=& 
    -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
\frac{\partial\Ub}{\partial t} &=& 
    -\Ub\cdot\nabla\Ub  - \frac{1}{\rho}\nabla\pi 
    - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
\frac{\partial(\rho h)}{\partial t} &=& 
    -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} 
    + \rho\Hnuc ,\label{eq:enthalpy}
\end{eqnarray}
together with a constraint equation.

In the Strang-split version of Maestro, the reaction network solves just
the reaction portion of the species evolution equations along with a
temperature evolution equation.  This is done using a standard ODE solver
(like {\tt VODE}) on the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T) \\
\frac{dT}{dt}   &=& 
    \frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right ).
\end{eqnarray}
Furthermore, we simplify the problem ``freezing'' the
thermodynamics---i.e., $c_p$ is evaluating at the start of the
integration and held constant thereafter.

In the SDC version, the VODE integration at the end of an SDC
iteration is responsible for updating all the thermodynamic quantities
\{$\rho$, $(\rho h)$, and $(\rho X_k$)\}, including both the advective
fluxes (incorporated via a piecewise constant advective flux source
term) and the reactions.  This provides a much stronger coupling between
the physical processes.  In particular, our system now looks like:
\begin{eqnarray}
\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + A_{\rho X_k} \\
\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + A_{\rho h}
\end{eqnarray}
Here, $A_{\rho X_k}$ and $A_{\rho h}$ are piecewise-constant (in time)
approximations to the change in ${\rho X_k}$ and ${\rho h}$ (respectively)
due to the advection.  These are constructed by calling {\tt density\_advance}
and {\tt enthalpy\_advance} in Maestro and passed into the network solver
during the reaction step.

The RHS routine that the ODE solver operates on will first construct
the density as:
\begin{equation}
\rho = \sum_k (\rho X_k)
\end{equation}
It will then derive the temperature from the equation of state.  If we
are running with {\tt use\_tfromp = T}, then we do
\begin{equation}
T = T(\rho, p_0, X_k)
\end{equation}
otherwise, we do
\begin{equation}
T = T(\rho, h, X_k)
\end{equation}
Note that in constrast to the Strang-split version, here we call the EOS
every time we enter the RHS routine.

The major changes from the non-SDC-enabled burners is the addition of
the advective terms to the system of ODEs, the fact that we integrate
$(\rho X_k)$ instead of just $X_k$, and the need to derive the
temperature from the input state for each RHS evaluation by {\tt
  VODE}.

### 2011-06-04 docs/SDC/sdc.tex
write out the form of A and note that we don't update U
fix typo
add a note about present implementation

@@
 and {\tt enthalpy\_advance} in Maestro and passed into the network solver
-during the reaction step.
+during the reaction step.  These terms approximate the following:
+\begin{eqnarray}
+A_{\rho X_k} &=&  \left [- \nabla \cdot (\rho X_k \Ub) \right ]^{n+1/2} \\
+A_{\rho h}   &=&  \left [- \nabla \cdot (\rho h U) + \frac{Dp_0}{Dt} \right ]^{n+1/2}
+\end{eqnarray}
 
@@
   VODE}.
+
+Note also that the SDC integration by {\tt VODE} does not operate on 
+the velocities at all.  That update is handled in the same fashion 
+as the Strang-split version of the code.
+
+The {\tt ignition\_simple\_SDC} burner shows how to setup the system
+for {\tt use\_tfromp = T} or {\tt F}.  Presently, this implementation
+does not support {\tt evolve\_base\_state = T} (in particular, we 
+need to evolve $p_0$ in the RHS routine).

### 2011-06-14 docs/compilers/compilers.tex
start some documentation on compilers

\section{Compilers}

Here, we list notes on problems with specific compilers and Maestro.

\subsection{Intel}

Overall, these tend to be flaky.

\subsection{gfortran}

Probably the best compilers for Maestro.  Versions 4.5 and 4.6 seem to work
really well and the debugging features are very reliable.

\subsection{PGI}

(on Kraken) The bounds checking routines do not like the way some of
the arrays are allocated and flag them as errors.  For instance, in
{\tt layout.f90}, we allocate a box as:
\begin{verbatim}
    type(box) :: bxs(3**get_dim(b))
\end{verbatim}
but the bounds checking routine seems to fail on this.

(on hopper) There are problems with code compiled with {\tt -fast}
instead of just {\tt -O}.  gcc/gfortran seem to work better and 
are only slighly slower.

\subsection{PathScale}

These are the prefered compilers on jaguar.

### 2011-06-15 docs/SDC/sdc.tex
expand notes to show what else needs to be computed in the network

@@
 
+\subsection{Strang-split version}
+
+In the Strang-split version of MAESTRO, the reaction and diffusion
+processes operate independent of one-another.  The species and
+enthalpy equations are updated through $\Delta t$ using the following
+sequence:
+
+\begin{itemize}
+
+\item React for $\Delta t/2$
+
 In the Strang-split version of Maestro, the reaction network solves just
@@
 \end{eqnarray}
-Furthermore, we simplify the problem ``freezing'' the
-thermodynamics---i.e., $c_p$ is evaluating at the start of the
-integration and held constant thereafter.
+Here, $T$ is evolved solely to evaluate the reaction rates,
+$\omegadot_k(\rho,X_k,T)$.  Furthermore, we simplify the problem
+``freezing'' the thermodynamics---i.e., $c_p$ is evaluating at the
+start of the integration and held constant thereafter.
 
+We update the enthalpy at the end of the integration as:
+\begin{equation}
+h^\mathrm{new} = h^\mathrm{old} + \frac{\Delta t}{2} \Hnuc
+\end{equation}
+
+\item Advect for $\Delta t$
+
+Here we solve
+\begin{eqnarray}
+\frac{\partial(\rho X_k)}{\partial t} &=& 
+    -\nabla\cdot(\rho X_k\Ub) \\
+\frac{\partial(\rho h)}{\partial t} &=& 
+    -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} 
+\end{eqnarray}
+
+Note that no explicit reaction terms appear here.  Since the advection
+takes place using the state updated from the reaction step, the effect
+of the reactions is implicitly contained in the advective update.
+
+\item React for $\Delta t/2$
+
+Finally, we react again, starting with the state left by the advection
+step.
+
+\end{itemize}
+
+\subsection{SDC version}
 In the SDC version, the VODE integration at the end of an SDC
@@
 \end{eqnarray}
+The construction of the interface states used in the advective terms
+uses an iteratively-lagged approximation to the reaction source terms
+in the interface prediction.  This explicitly couples the reaction
+process to the advection process.
 
@@
 
+At the end of the integration, we define 
+\begin{eqnarray}
+I_{\rho X_k} &=& \frac{(\rho X_k)^\mathrm{new} - (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k} \\
+I_{\rho h}  &=& \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h}
+\end{eqnarray}
+These terms are used as the source terms for the advection step in the
+next SDC iteration.
+
+\subsection{Summary of Changes}
 The major changes from the non-SDC-enabled burners is the addition of

### 2011-06-15 docs/pert_notes/pert.tex
start reworking this to allow for other species_pred_types

@@
 \frac{\partial\rho}{\partial t} &=& -\nabla\cdot(\rho\Ub) \nonumber \\
-&=& -\Ub\cdot\nabla\rho - \rho\nabla\cdot\Ub. \label{rho equation}
+&=& -\Ub\cdot\nabla\rho - \rho\nabla\cdot\Ub \, . \label{rho equation}
 \end{eqnarray}
+The species are evolved according to
+\begin{eqnarray}
+\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho\Ub X_k) + \rho \omegadot_k \nonumber \\
+&=& -\Ub\cdot\nabla\rho - \rho\nabla\cdot\Ub + \rho \omegadot_k \, . \label{species equation}
+\end{eqnarray}
+In practice, only the species evolution equation is evolved, and the
+total density is set as
+\begin{equation}
+\rho = \sum_k (\rho X_k)
+\end{equation}
+To advance $(\rho X_k)$ to the next timestep, we need to predict a
+time-centered, interface state.  Algebraically, there are multiple
+paths to get this interface state---we can predict $(\rho X)$ to the
+edges as a single quantity, or predict $\rho$ and $X$ separately
+(either in full or perturbation form).  In the notes below, we use the
+subscript `{\tt edge}' to indicate what quantity was predicted to the
+edges.  In MAESTRO, the different methods of computing $(\rho X)$ on
+edges are controlled by the {\tt species\_pred\_type} parameter.  The
+resulting edge state is shown in the following table:
+
+\begin{table}[h]
+\centering
+\begin{tabular}{c|c}
+\hline
+\hline
+{\tt species\_pred\_type} &   {$(\rho X_k)$ edge state} \\
+\hline \\[-3mm]
+1 / {\tt pred\_rhoprime\_and\_X}  &  $\left(\rho_0^{n+\myhalf,{\rm avg}}
++ \rho'_\mathrm{edge} \right)(X_k)_\mathrm{edge}$ \\[2mm]
+2 / {\tt pred\_rhoX}  &  $(\rho X_k)_\mathrm{edge}$ \\
+\hline
+\end{tabular}
+\end{table}
+We note the labels {\tt pred\_rhoprime\_and\_X} and {\tt pred\_rhoX} are
+provided by the {\tt pred\_parameters} module.
+
+\subsection{Advective force for interface prediction}
+In predicting the interface states, we work with an advection-like
+equation, and call anything other than the advective term a ``force''.
+For example, for a scalar $\phi$, obeying a conservation law with a source term, $f$:
+\begin{equation}
+\frac{\partial \phi}{\partial t} = -\nabla \cdot (\phi \Ub) + f
+\end{equation}
+We write this in advective form as:
+\begin{equation}
+\frac{\partial \phi}{\partial t} = -\Ub \cdot \nabla \phi \, \underbrace{- \phi \nabla \cdot \Ub + f}_{\text{``advective force''}}
+\end{equation}
+Here we identify the non-advection portion of the divergence of the
+flux together with the original force, $f$, as the ``force'' for the
+advective equation.  In predicting the interface states, this advective
+force will be used in computing the interface states.
+
+\subsection{{\tt species\_pred\_type} = {\tt pred\_rhoprime\_and\_X}}
+
+Here we wish to construct $(\rho_0^{n+\myhalf,{\rm avg}}        
++ \rho'_\mathrm{edge})(X_k)_\mathrm{edge}$.
+
 We predict both $\rho'$ and $\rho_0$ to edges separately and later use them to 
@@
 -w_0\frac{\partial\rho_0}{\partial r} 
-\underbrace{-\rho_0\frac{\partial w_0}{\partial r}}_{"\rho_0 ~ \text{force}"}.
+\underbrace{-\rho_0\frac{\partial w_0}{\partial r}}_{``\rho_0 ~ \text{force"}}.
 \label{rho0 equation}
@@
 
-\subsection{Predicting $\rho'$ at edges}
+\subsubsection{Predicting $\rho'$ at edges}
 We define $\rho' = \rho^{(1)} - \rho_0^n$.  Then we predict $\rho'$ to
-edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} and the
+edges using {\tt make\_edge\_scal} in {\tt density\_advance} and the
 underbraced term in (\ref{rhoprime equation}) as the forcing.
 
-\subsection{Predicting $\rho_0$ at edges}\label{Predicting rho0 at edges}
+\subsubsection{Predicting $\rho_0$ at edges}\label{Predicting rho0 at edges}
 There are two ways to predict $\rho_0$ at edges.
@@
 
-\subsection{Computing $\rho$ at edges}\label{Computing rho at edges}
+We note that in essence these choices reflect a hyperbolic (1)
+vs.\ elliptic (2) approach.  In MAESTRO, if we setup a problem with
+$\rho = \rho_0$ initially, and enforce a constraint $\nabla \cdot
+(\rho_0 \Ub) = 0$ (i.e.\ the anelastic constraint), then analytically,
+we should never generate a $\rho'$.  To realize this behavior
+numerically, we use $\rho_0^{n+\myhalf,{\rm avg}}$ in the prediction
+of $(\rho X_k)$ on the edges to be consistent with the use of the
+average of $\rho$ to the interfaces in the projection step at the end
+of the algorithm.
+
+\subsubsection{Computing $\rho$ at edges}\label{Computing rho at edges}
 For the non-radial edges, we directly add $\rho_0^{n+\myhalf,{\rm avg}}$
@@
 
+\subsubsection{Predicting $X_k$ at edges}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{{\tt species\_pred\_type} = {\tt pred\_rhoX}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Advancing $\rho X_k$}\label{Advancing rhoX_k}
@@
 We define $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$.  Then we predict 
-$(\rho h)'$ to edges using {\tt make\_edge\_scal} in {\tt scalar\_advance} 
+$(\rho h)'$ to edges using {\tt make\_edge\_scal} in {\tt enthalpy\_advance} 
 and the underbraced term in (\ref{rhohprime equation}) as the forcing.

### 2011-06-15 docs/pert_notes/pert.tex
finish species_pred_type = 1

@@
 \frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho\Ub X_k) + \rho \omegadot_k \nonumber \\
-&=& -\Ub\cdot\nabla\rho - \rho\nabla\cdot\Ub + \rho \omegadot_k \, . \label{species equation}
+&=& -\Ub\cdot\nabla(\rho X_k) - \rho X_k \nabla\cdot\Ub + \rho \omegadot_k \, . \label{species equation}
 \end{eqnarray}
@@
 \end{equation}
-where the underbraced "$\rho'$ force'' is computed in {\tt modify\_scal\_force}.
+where the underbraced ``$\rho'$ force'' is computed in {\tt modify\_scal\_force}.
+
+We also need $X_k$ at the edges.  Here, we subtract $X_k \times$
+Eq.~\ref{rho equation} from Eq.~\ref{species equation} to obtain
+\begin{equation}
+\frac{\partial X_k}{\partial t} = -\Ub \cdot \nabla X_k + \omegadot_k
+\end{equation}
+When using Strang-splitting, we ignore the $\omegadot_k$ source terms, and
+then the species equation is a pure advection equation with no force.
 
@@
 \subsubsection{Predicting $X_k$ at edges}
+Predicting $X_k$ is straightforward.  We convert the cell-centered
+$(\rho X_k)$ to $X_k$ by dividing by $\rho$ in each zone and then we
+just call {\tt make\_edge\_scal} in {\tt density\_advance} on $X_k$.
+The force seen by {\tt make\_edge\_scal} is 0.
 

### 2011-06-15 docs/pert_notes/pert.tex
add species_pred_type = 2 version

@@
 \begin{equation}
-\frac{\partial \phi}{\partial t} = -\Ub \cdot \nabla \phi \, \underbrace{- \phi \nabla \cdot \Ub + f}_{\text{``advective force''}}
+\frac{\partial \phi}{\partial t} = -\Ub \cdot \nabla \phi \, \underbrace{- \phi \nabla \cdot \Ub}_{\text{``advective force''}} + f
 \end{equation}
 Here we identify the non-advection portion of the divergence of the
-flux together with the original force, $f$, as the ``force'' for the
+flux as the ``advective force'' for the
 advective equation.  In predicting the interface states, this advective
-force will be used in computing the interface states.
+force will be used in computing the interface states.  Depending on the
+nature of $f$, it may either appear in the prediction as well or be absent
+if the process it describes is incorporated by Strang-splitting.
 
@@
 \subsection{{\tt species\_pred\_type} = {\tt pred\_rhoX}}
+
+Here we wish to construct $(\rho X_k)_\mathrm{edge}$ by predicting
+$(\rho X_k)$ to the edges as a single quantity.  We start with
+Eq.~\ref{species equation}:
+\begin{equation}
+\frac{\partial(\rho X_k)}{\partial t} =
+  -\Ub\cdot\nabla(\rho X_k) - \underbrace{\rho X_k \nabla\cdot\Ub}_{``\rho X_k ~\text{force''}} + \rho \omegadot_k \, . \label{species equation labeled}
+\end{equation}
+The edge state is created by calling {\tt make\_edge\_scal} in {\tt
+  density\_advance} with the underbraced term in (\ref{species
+  equation labeled}) as the forcing.  This advective force is computed
+by {\tt modify\_scal\_force} with {\tt fullform = .true.}.  We do not
+consider the $\omegadot_k$ term in the forcing when Strang-splitting.
 

### 2011-06-15 docs/pert_notes/pert.tex
finish species_pred_type stuff (for what's in the code now)

@@
 
+We note that even though it is not needed here, we still predict
+$\rho'$ to the edges since certain enthalpy formulations need it.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 that were already accounted for in {\tt react\_state}, and the
-associated discretization is
+associated discretization is \\
+
+\noindent {\tt species\_pred\_type = pred\_rhoprime\_and\_X}:
 \begin{equation}
 \frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
--\nabla\cdot\left\{\left[\left(\rho_0^{n+\myhalf,{\rm avg}} 
-+ \rho'\right)X_k\right](\Ubt+w_0\eb_r)\right\}.
+-\nabla\cdot\left\{\left[\left({\rho_0}^{n+\myhalf,{\rm avg}}
++ \rho'_\mathrm{edge} \right)(X_k)_\mathrm{edge} \right](\Ubt+w_0\eb_r)\right\}.
+\end{equation} \\
+
+\noindent {\tt species\_pred\_type = pred\_rhoX}:
+\begin{equation}
+\frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
+-\nabla\cdot\left\{\left[\left(\rho X_k \right)_\mathrm{edge} \right](\Ubt+w_0\eb_r)\right\}.
 \end{equation}
@@
 \section{Energy Evolution}
+
+Note: only the default {\tt enthalpy\_pred\_type = 1} is guaranteed to work
+with {\tt species\_pred\_type > 1}. \\
+
 We use {\tt enthalpy\_pred\_type} = 1, in which we predict $(\rho h)'$ at edges.

### 2011-06-15 docs/pert_notes/pert.tex
add species_pred_type = 3 notes

@@
 \centering
-\begin{tabular}{c|c}
+\begin{tabular}{l|c}
 \hline
@@
 + \rho'_\mathrm{edge} \right)(X_k)_\mathrm{edge}$ \\[2mm]
-2 / {\tt pred\_rhoX}  &  $(\rho X_k)_\mathrm{edge}$ \\
+2 / {\tt pred\_rhoX}  &  $(\rho X_k)_\mathrm{edge}$ \\[2mm]
+3 / {\tt pred\_rho\_and\_X}  &  $\rho_\mathrm{edge} (X_k)_\mathrm{edge}$ \\
 \hline
@@
 \end{table}
-We note the labels {\tt pred\_rhoprime\_and\_X} and {\tt pred\_rhoX} are
-provided by the {\tt pred\_parameters} module.
+We note the labels {\tt pred\_rhoprime\_and\_X}, {\tt pred\_rhoX}, and
+{\tt pred\_rho\_and\_X} are provided by the {\tt pred\_parameters}
+module.
 
@@
 \subsubsection{Predicting $X_k$ at edges}
+\label{sec:pert:predict_X}
 Predicting $X_k$ is straightforward.  We convert the cell-centered
@@
 \frac{\partial(\rho X_k)}{\partial t} =
-  -\Ub\cdot\nabla(\rho X_k) - \underbrace{\rho X_k \nabla\cdot\Ub}_{``\rho X_k ~\text{force''}} + \rho \omegadot_k \, . \label{species equation labeled}
+  -\Ub\cdot\nabla(\rho X_k) \, \underbrace{-\rho X_k \nabla\cdot\Ub}_{``\rho X_k ~\text{force''}} + \rho \omegadot_k \, . \label{species equation labeled}
 \end{equation}
@@
 $\rho'$ to the edges since certain enthalpy formulations need it.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{{\tt species\_pred\_type} = {\tt pred\_rho\_and\_X}}
+
+Here we wish to construct $\rho_\mathrm{edge} (X_k)_\mathrm{edge}$
+by predicting $\rho$ and $X_k$ to the edges separately.
+
+Predicting $X_k$ to the edges proceeds exactly as described in
+\S~\ref{sec:pert:predict_X}.
+
+Predicting the full $\rho$ begins with Eq.~\ref{rho equation}:
+\begin{equation}
+\frac{\partial\rho}{\partial t} 
+= -\Ub\cdot\nabla\rho \, \underbrace{- \rho\nabla\cdot\Ub}_{``\rho~\text{force''}} \, . \label{rho equation labeled}
+\end{equation}
+Using this, $\rho$ is predicted to the edges using {\tt
+  make\_edge\_scal} in {\tt density\_advance}, with the underbraced
+force computed in {\tt modify\_scal\_force} with {\tt fullform =
+  .true.}.
 
@@
 -\nabla\cdot\left\{\left[\left(\rho X_k \right)_\mathrm{edge} \right](\Ubt+w_0\eb_r)\right\}.
+\end{equation}
+
+\noindent {\tt species\_pred\_type = pred\_rho\_and\_X}:
+\begin{equation}
+\frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
+-\nabla\cdot\left\{\left[\rho_\mathrm{edge} (X_k)_\mathrm{edge} \right](\Ubt+w_0\eb_r)\right\}.
 \end{equation}

### 2011-06-15 docs/runtime_parameters/runtime_parameters.tex
add change_max_grdi_size_1 to parameters table
update doc

--- 2011-01-20 docs/runtime_parameters/runtime_parameters.tex
@@
 \verb=  cg_verbose  = &   Verbosity of bottom solver  &  0 \\
+\verb=  change_max_grid_size_1  = &   Change the max grid size on the base level on restart  &  .false. \\
 \verb=  check_base_name  = &   prefix to use in checkpoint file names  &  "chk" \\
@@
 \verb=  do_eos_h_above_cutoff  = &   After the advective enthalpy update, recompute the enthalpy if we are below the base cutoff density.  &  .true. \\
+\verb=  do_heating  = &   use analytic heating  &  .false. \\
 \verb=  do_initial_projection  = &   Do the initial projection.  &  .true. \\
@@
 \\
-\verb=  dump_grid_file  = &    &  .false. \\
+\verb=  dump_grid_file  = &   dump out a file, named grids.out, containing the number of grids a each level.  A new line is added each time regrid is called.  &  .false. \\
 \verb=  enthalpy_pred_type  = &   predict\_rhohprime = 1; \newline predict\_h = 2; \newline predict\_T\_then\_rhohprime = 3; \newline predict\_T\_then\_h = 4; \newline predict\_hprime = 5; \newline predict\_Tprime\_then\_h = 6.
@@
 &  1 \\
+\verb=  sdc_iters  = &   how many SDC iteractions to take (requires the code be compiled with {\tt SDC := t}  &  4 \\
 \verb=  single_prec_plotfiles  = &    &  .false. \\
@@
 \verb=  small_temp  = &    &  5.d6 \\
+\verb=  species_pred_type  = &   Which quantities do we predict to the edges for computing the ($\rho X$) edge states?
+\newline {\tt species\_pred\_type} = 1 means predict $\rho^\prime$ and $X$ separately.
+\newline {\tt species\_pred\_type} = 2 means predict the full ($\rho X$) itself.
+\newline {\tt species\_pred\_type} = 3 means predict $\rho$ and $X$ separately.
+&  1 \\
 \verb=  spherical_in  = &   set to 1 if you are doing a spherical problem  &  0 \\
@@
 \verb=  use_hypre  = &   use the hypre library  &  .false. \\
+\verb=  use_linear_grav_in_beta  = &   how to represent gravity in the beta0 integration: .true. = piecewise linear .false. = piecewise constant  &  .false. \\
+\verb=  use_particles  = &   call the particle initialization, advection, etc. routines.  &  .false. \\
 \verb=  use_soundspeed_firstdt  = &   Use the soundspeed constraint when computing the first time step.  &  .false. \\

### 2011-06-15 docs/runtime_parameters/rp.py
update + make the table landscape to make things fit better

@@ -13,12 +13,14 @@ shown here.
 % symbol table
 %%%%%%%%%%%%%%%%
 
+\begin{landscape}
+
 {\small
 
 \renewcommand{\arraystretch}{1.5}
 %
 \begin{center}
-\begin{longtable}{|l|p{3.25in}|l|}
+\begin{longtable}{|l|p{5.25in}|l|}
 \caption[runtime parameters]{runtime parameters.} \label{table:runtime} \\
 %
 \hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
@@ -47,6 +49,8 @@ footer=r"""
 \end{center}
 
 } % ends \small
+
+\end{landscape}
 
 %
 

### 2011-06-16 docs/pert_notes/pert.tex
start reorganizing the enthalpy notes -- now there is a master table that lists all the details for enthalpy prediction depending on the species predition -- only types 1 and 2 so far...

@@
 \hline \\[-3mm]
-1 / {\tt pred\_rhoprime\_and\_X}  &  $\left(\rho_0^{n+\myhalf,{\rm avg}}
+1 / {\tt predict\_rhoprime\_and\_X}  &  $\left(\rho_0^{n+\myhalf,{\rm avg}}
 + \rho'_\mathrm{edge} \right)(X_k)_\mathrm{edge}$ \\[2mm]
-2 / {\tt pred\_rhoX}  &  $(\rho X_k)_\mathrm{edge}$ \\[2mm]
-3 / {\tt pred\_rho\_and\_X}  &  $\rho_\mathrm{edge} (X_k)_\mathrm{edge}$ \\
+2 / {\tt predict\_rhoX}  &  $(\rho X_k)_\mathrm{edge}$ \\[2mm]
+3 / {\tt predict\_rho\_and\_X}  &  $\rho_\mathrm{edge} (X_k)_\mathrm{edge}$ \\
 \hline
@@
 \end{table}
-We note the labels {\tt pred\_rhoprime\_and\_X}, {\tt pred\_rhoX}, and
-{\tt pred\_rho\_and\_X} are provided by the {\tt pred\_parameters}
+We note the labels {\tt predict\_rhoprime\_and\_X}, {\tt predict\_rhoX}, and
+{\tt predict\_rho\_and\_X} are provided by the {\tt pred\_parameters}
 module.
@@
 
-\subsection{{\tt species\_pred\_type} = {\tt pred\_rhoprime\_and\_X}}
+\subsection{{\tt species\_pred\_type} = {\tt predict\_rhoprime\_and\_X}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\tt species\_pred\_type} = {\tt pred\_rhoX}}
+\subsection{{\tt species\_pred\_type} = {\tt predict\_rhoX}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\tt species\_pred\_type} = {\tt pred\_rho\_and\_X}}
+\subsection{{\tt species\_pred\_type} = {\tt predict\_rho\_and\_X}}
 
@@
 
-\noindent {\tt species\_pred\_type = pred\_rhoprime\_and\_X}:
+\noindent {\tt species\_pred\_type = predict\_rhoprime\_and\_X}:
 \begin{equation}
@@
 
-\noindent {\tt species\_pred\_type = pred\_rhoX}:
+\noindent {\tt species\_pred\_type = predict\_rhoX}:
 \begin{equation}
@@
 
-\noindent {\tt species\_pred\_type = pred\_rho\_and\_X}:
+\noindent {\tt species\_pred\_type = predict\_rho\_and\_X}:
 \begin{equation}
@@
 
-We use {\tt enthalpy\_pred\_type} = 1, in which we predict $(\rho h)'$ at edges.
+Maestro solves an enthalpy equation.  
 The full enthalpy equation is
@@
 %\end{equation}
-Due to operator splitting, the call to {\tt react\_state} has already
-been made.  Hence, the goal is to compute an edge state enthalpy
-starting from $(\rho h)^{(1)}$ using an enthalpy equation that does
-not include the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms, where
-were already accounted for in {\tt react\_state}, so our equation becomes
+Due to Strang-splitting of the reactions,the call to {\tt
+  react\_state} has already been made.  Hence, the goal is to compute
+an edge state enthalpy starting from $(\rho h)^{(1)}$ using an
+enthalpy equation that does not include the $\rho H_{\rm nuc}$ and
+$\rho H_{\rm ext}$ terms, where were already accounted for in {\tt
+  react\_state}, so our equation becomes
 \begin{equation}
@@
 \end{eqnarray}
-Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and rearranging terms gives 
-the perturbational enthalpy equation
+Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and
+rearranging terms gives the perturbational enthalpy equation
 \begin{eqnarray}
@@
 \end{eqnarray}
+
+Alternately, we can consider an temperature evolution equation, derived
+from enthalpy, as:
+\begin{equation}
+\frac{\partial T}{\partial t} = -\Ub\cdot\nabla T
++ \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi
++ (\Ubt \cdot \er )\frac{\partial p_0}{\partial r}\right] 
++ \nabla \cdot \kth \nabla T
+- \sum_k\rho\xi_k\omegadot_k 
++ \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.
+\end{equation}
+Again, we neglect the reaction terms, since that will be handled during
+the reaction step, so we can write this as:
+\begin{equation}
+\frac{\partial T}{\partial t} = -\Ub\cdot\nabla T
++ \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi
++ (\Ubt \cdot \er )\frac{\partial p_0}{\partial r}\right] 
++ \nabla \cdot \kth \nabla T \right \} \, .
+\end{equation}
+
+A final alternative is to consider an evolution equation for $h$
+alone.  This can be derived by expanding the derivative of $(\rho h)$
+in Eq.~\ref{rhoh equation} and subtracting off $h \times$ the
+continuity equation (Eq.~\ref{rho equation}):
+\begin{equation}
+\frac{\partial h}{\partial t} = -\Ub \cdot \nabla h + \frac{1}{\rho}
+\left \{ \psi + (\Ubt \cdot \er )\frac{\partial p_0}{\partial r}
++ \nabla \cdot \kth \nabla T \right \} \, .
+\end{equation}
+
+To update the enthalpy, we need to compute an interface state for
+$(\rho h)$.  As with the species evolution, there are multiple
+quantities we can predict to the interfaces to form this state,
+controlled by {\tt enthalpy\_pred\_type}.  A complexity of the
+enthalpy evolution is that the formation of this edge state will
+depend on {\tt species\_pred\_type}.  Table~\ref{table:pred:hoverview}
+gives a summary
+of the {\tt enthalpy\_pred\_type} behavior.
+
+\begin{landscape}
+\begin{table}[t]
+\renewcommand{\arraystretch}{1.5}
+{\small
+\centering
+\begin{tabular}{l|l|c|c|c|c}
+\hline
+\hline
+{\tt species\_pred\_type} &   
+{\tt enthalpy\_pred\_type} &
+{quantity predicted} &
+{intermediate} &
+{species quantity} &
+{final $(\rho h)$} \\[-5pt]
+ &
+ &
+{in {\tt make\_edge\_scal}} &
+{``enthalpy'' edge state} &
+{available in {\tt mkflux}} &
+{edge state} \\
+\hline 
+1 / {\tt predict\_rhoprime\_and\_X}  & 1 / {\tt predict\_rhohprime} &
+  $(\rho h)'$ & $(\rho h)'$ & $X$, $\rho'$ & 
+  $\left [ (\rho h)_0 + (\rho h)' \right ]$ \\
+1 / {\tt predict\_rhoprime\_and\_X}  & 2 / {\tt predict\_h} &
+  $h$ & $h$ & $X$, $\rho'$ & 
+  $\left ( \rho_0 + \rho' \right ) h$ \\
+1 / {\tt predict\_rhoprime\_and\_X}  & 3 / {\tt predict\_T\_then\_rhohprime} &
+  $T$ & $(\rho h)'$ & $X$, $\rho'$ & 
+  $\left [ (\rho h)_0 + (\rho h)' \right ]$ \\
+1 / {\tt predict\_rhoprime\_and\_X}  & 4 / {\tt predict\_T\_then\_h} &
+  $T$ & $h$ & $X$, $\rho'$ & 
+  $\left ( \rho_0 + \rho' \right ) h$ \\
+\hline
+\hline
+2 / {\tt predict\_rhoX}  & 1 / {\tt predict\_rhohprime} &
+  $(\rho h)'$ & $(\rho h)'$ & $(\rho X)$, $\rho'$ & 
+  $\left [ (\rho h)_0 + (\rho h)' \right ]$ \\
+2 / {\tt predict\_rhoX}  & 2 / {\tt predict\_h} &
+  $h$ & $h$ & $(\rho X)$, $\rho'$ & 
+  $\left ( \rho_0 + \rho' \right ) h$ \\
+2 / {\tt predict\_rhoX}  & 3 / {\tt predict\_T\_then\_rhohprime} &
+  $T$ & $(\rho h)'$ & $(\rho X)$, $\rho'$ & 
+  $\left [ (\rho h)_0 + (\rho h)' \right ]$ \\
+2 / {\tt predict\_rhoX}  & 4 / {\tt predict\_T\_then\_h} &
+  $T$ & $h$ & $(\rho X)$, $\rho'$ & 
+  $\left ( \rho_0 + \rho' \right ) h$ \\
+\hline
+\hline
+\end{tabular}
+} % end \small
+\end{table}
+\end{landscape}
+
+\subsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
+
 where the first two terms in $(\rho h)'$ force are computed in 
@@
 
-\subsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
 We define $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$.  Then we predict 

### 2011-06-16 docs/pert_notes/pert.tex
more reorganization

@@
 &=& -w_0\frac{\partial(\rho h)_0}{\partial r} 
-- \underbrace{(\rho h)_0\frac{\partial w_0}{\partial r}+ \psi}_{"(\rho h)_0 ~ \text{force}"}
+- \underbrace{(\rho h)_0\frac{\partial w_0}{\partial r}+ \psi}_{``(\rho h)_0 ~ \text{force}"}
 \enskip .\label{rhoh0 equation}
@@
 - \nabla\cdot[(\rho h)_0\Ubt] + (\Ubt \cdot \er)\frac{\partial p_0}{\partial r}
-+ \nabla\cdot\kth\nabla T}_{"(\rho h)' ~ \text{force}"}, \label{rhohprime equation}
++ \nabla\cdot\kth\nabla T}_{``(\rho h)' ~ \text{force}"}, \label{rhohprime equation}
 \end{eqnarray}
@@
 \frac{\partial T}{\partial t} = -\Ub\cdot\nabla T
+\underbrace{
 + \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi
 + (\Ubt \cdot \er )\frac{\partial p_0}{\partial r}\right] 
-+ \nabla \cdot \kth \nabla T \right \} \, .
++ \nabla \cdot \kth \nabla T \right \} }_{``T~\text{force''}} \, .
+\label{T equation labeled}
 \end{equation}
@@
 \begin{equation}
-\frac{\partial h}{\partial t} = -\Ub \cdot \nabla h + \frac{1}{\rho}
+\frac{\partial h}{\partial t} = -\Ub \cdot \nabla h 
+\underbrace{+ \frac{1}{\rho}
 \left \{ \psi + (\Ubt \cdot \er )\frac{\partial p_0}{\partial r}
-+ \nabla \cdot \kth \nabla T \right \} \, .
++ \nabla \cdot \kth \nabla T \right \} }_{``h~\text{force''}} \, .
+\label{h equation labeled}
 \end{equation}
@@
 
+The general procedure for making the $(\rho h)$ edge state is as follows:
+\begin{enumerate}
+\item predict $(\rho h)'$, $h$, or $T$ to the edges (depending on {\tt
+  enthalpy\_pred\_type} ) using {\tt make\_edge\_scal} and the forces
+  identified in the appropriate evolution equation
+  (Eqs.~\ref{rhohprime equation}, \ref{T equation labeled}, or \ref{h
+    equation labeled} respectively).
+
+  The appropriate forces are summaried in table~\ref{table:pred:hforce}.
+
+\item if we predicted $T$, convert this predicted
+  edge state to an' intermediate ``enthalpy'' state (the quotes
+  indicate that it may be perturbational or full enthalpy) by calling
+  the EOS.
+
+\item construct the final enthalpy edge state in {\tt mkflux}.  The
+  precise construction depends on what species and enthalpy quantities
+  are input to {\tt mkflux}.
+
+\end{enumerate}
+
+\begin{table*}[h]
+\centering
+\caption{Forcing term into {\tt make\_edge\_scal} \newline}
+\label{table:pred:hforce}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{l|c}
+\hline
+\hline
+{\tt enthalpy\_pred\_type} &   {advective force} \\
+\hline
+1 / {\tt predict\_rhohprime} $((\rho h)^\prime)$ &  $(\Ubt \cdot \er) \frac{\partial p_0}{\partial r} - 
+ (\rho h)^\prime \; \nabla \cdot (\Ubt+w_0 \er) - 
+ \nabla \cdot (\Ubt (\rho h)_0) + \nabla \cdot \kth \nabla T$ \\
+2 / {\tt predict\_h}  $(h)$ & $\frac{1}{\rho} \left [\psi + (\Ubt \cdot \er)
+  \frac{\partial p_0}{\partial r} + \nabla \cdot \kth \nabla T \right ]$ \\
+3 / {\tt predict\_T\_then\_rhohprime} $(T)$ & $\frac{1}{\rho c_p} \left \{ (1 - \rho h_p) 
+   \left [\psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} \right ] + \nabla \cdot \kth \nabla T \right \}$ \\
+4 / {\tt predict\_T\_then\_h}  $(T)$ & $\frac{1}{\rho c_p} \left\{ (1 - \rho h_p) \left [\psi + (\Ubt \cdot \er)
+\frac{\partial p_0}{\partial r}\right ] +  \nabla \cdot \kth \nabla T \right\}$ \\
+\hline
+\end{tabular}
+\end{table*}
+
 \begin{landscape}
 \begin{table}[t]
+\caption{Summary of enthalpy edge state construction}
+\label{table:pred:hoverview}
 \renewcommand{\arraystretch}{1.5}
@@
 We use an analogous procedure described in Section \ref{Computing rho
-at edges} for computing $\rho$ at edges to compute $\rho h$ at edges.
-However, instead of computing $(\rho h)_0$ to edges directly, we compute $\rho$
-and $h$ separately, and multiply to get $(\rho h)_0$.
+  at edges} for computing $\rho$ at edges to compute $\rho h$ at
+edges.  For spherical, however, instead of computing $(\rho h)_0$ to
+edges directly, we compute $\rho$ and $h$ separately, and multiply to
+get $(\rho h)_0$.
 
@@
 \end{equation}
-
-\subsection{Summary of Forcing}
-
-\begin{table*}[h]
-\begin{center}
-\caption{Forcing term into {\tt make\_edge\_scal} \newline}
-\begin{tabular}{c|c}
-\hline
-\hline
-{H PRED TYPE} &   {Force} \\
-\hline \\[-3mm]
-1  $((\rho h)^\prime)$ &  $(\Ubt \cdot \er) \frac{\partial p_0}{\partial r} - 
- (\rho h)^\prime \; \nabla \cdot (\Ubt+w_0 \er) - 
- \nabla \cdot (\Ubt (\rho h)_0) + \nabla \cdot \kth \nabla T$ \\[2 mm]
-2  $(h)$ & $\frac{1}{\rho} \left [\psi + (\Ubt \cdot \er)
-  \frac{\partial p_0}{\partial r} + \nabla \cdot \kth \nabla T \right ]$ \\[2 mm]
-3  $(T)$ & $\frac{1}{\rho c_p} \left \{ (1 - \rho h_p) 
-   \left [\psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} \right ] \nabla \cdot \kth \nabla T \right \}$ \\[2 mm]
-4  $(T)$ & $\frac{1}{\rho c_p} \left\{ (1 - \rho h_p) \left [\psi + (\Ubt \cdot \er)
-\frac{\partial p_0}{\partial r}\right ] \nabla \cdot \kth \nabla T \right\}$ \\[2 mm]
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-
-\begin{table*}[h]
-\begin{center}
-\caption{Quantity that goes into and out of {\tt make\_edge\_scal} \newline}
-\begin{tabular}{c|c}
-\hline
-\hline
-{h pred type} & {h/T } \\
-\hline \\[-3mm]
-1 & $(\rho h)^\prime$ \\
-2 & $h$  \\
-3 & $T$  \\
-4 & $T$  \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
 
@@
 \begin{center}
-\caption{Quantity that goes into {\tt mkflux} on edges \newline}
-\begin{tabular}{c|c|c}
-\hline
-\hline
-{h pred type} & {Species } & {h/T } \\
-\hline
-1 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
-2 & $X$ and $\rho^\prime$ & $h$ \\
-3 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
-4 & $X$ and $\rho^\prime$ & $h$ \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-
-\begin{table*}
-\begin{center}
 \caption{Quantity that is created in {\tt mkflux} and passed into {\tt update\_scal}\newline\
@@
 \end{table*}
+
+%% \begin{table*}[h]
+%% \begin{center}
+%% \caption{Quantity that goes into and out of {\tt make\_edge\_scal} \newline}
+%% \begin{tabular}{c|c}
+%% \hline
+%% \hline
+%% {h pred type} & {h/T } \\
+%% \hline \\[-3mm]
+%% 1 & $(\rho h)^\prime$ \\
+%% 2 & $h$  \\
+%% 3 & $T$  \\
+%% 4 & $T$  \\
+%% \hline
+%% \end{tabular}
+%% \end{center}
+%% \end{table*}
+
+%% \begin{table*}
+%% \begin{center}
+%% \caption{Quantity that goes into {\tt mkflux} on edges \newline}
+%% \begin{tabular}{c|c|c}
+%% \hline
+%% \hline
+%% {h pred type} & {Species } & {h/T } \\
+%% \hline
+%% 1 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
+%% 2 & $X$ and $\rho^\prime$ & $h$ \\
+%% 3 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
+%% 4 & $X$ and $\rho^\prime$ & $h$ \\
+%% \hline
+%% \end{tabular}
+%% \end{center}
+%% \end{table*}

### 2011-06-16 docs/pert_notes/pert.tex
more work on the different enthalpy types...

@@
 \section{Density Evolution}
+
+\subsection{Basic equations}
+
 The full density evolution equation is
@@
 \begin{equation}
-\frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
+\frac{\partial\rho X_k}{\partial t} = 
 -\nabla\cdot\left\{\left[\left({\rho_0}^{n+\myhalf,{\rm avg}}
@@
 \begin{equation}
-\frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
+\frac{\partial\rho X_k}{\partial t} = 
 -\nabla\cdot\left\{\left[\left(\rho X_k \right)_\mathrm{edge} \right](\Ubt+w_0\eb_r)\right\}.
@@
 \begin{equation}
-\frac{\partial\rho X_k}{\partial t} = -\nabla\cdot(\rho X_k\Ub) = 
+\frac{\partial\rho X_k}{\partial t} = 
 -\nabla\cdot\left\{\left[\rho_\mathrm{edge} (X_k)_\mathrm{edge} \right](\Ubt+w_0\eb_r)\right\}.
@@
 
-Note: only the default {\tt enthalpy\_pred\_type = 1} is guaranteed to work
-with {\tt species\_pred\_type > 1}. \\
+\subsection{Basic equations}
 
@@
 \end{eqnarray}
+
+\subsubsection{Perturbational enthalpy formulation}
+
 Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and
@@
 \end{eqnarray}
+
+\subsubsection{Temperature formulation}
 
@@
 
+\subsubsection{Pure enthalpy formulation}
+
 A final alternative is to consider an evolution equation for $h$
@@
 
+\subsubsection{Prediction requirements}
+
 To update the enthalpy, we need to compute an interface state for
@@
 enthalpy evolution is that the formation of this edge state will
-depend on {\tt species\_pred\_type}.  Table~\ref{table:pred:hoverview}
-gives a summary
-of the {\tt enthalpy\_pred\_type} behavior.
+depend on {\tt species\_pred\_type}.
 
@@
 
+\noindent Note: the various {\tt enthalpy\_pred\_type}s are only
+guaranteed to work with {\tt species\_pred\_type = 1}.
+
+Table~\ref{table:pred:hoverview}
+gives a summary
+of the {\tt enthalpy\_pred\_type} behavior.
+
 \begin{table*}[h]
@@
 \hline
-1 / {\tt predict\_rhohprime} $((\rho h)^\prime)$ &  $(\Ubt \cdot \er) \frac{\partial p_0}{\partial r} - 
- (\rho h)^\prime \; \nabla \cdot (\Ubt+w_0 \er) - 
- \nabla \cdot (\Ubt (\rho h)_0) + \nabla \cdot \kth \nabla T$ \\
+1 / {\tt predict\_rhohprime} $((\rho h)^\prime)$ &  
+ $-(\rho h)^\prime \; \nabla \cdot (\Ubt+w_0 \er) - 
+ \nabla \cdot (\Ubt (\rho h)_0) + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} + \nabla \cdot \kth \nabla T$ \\
 2 / {\tt predict\_h}  $(h)$ & $\frac{1}{\rho} \left [\psi + (\Ubt \cdot \er)
@@
 
-\subsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{{\tt enthalpy\_pred\_type} = {\tt predict\_rhohprime}}
 
-where the first two terms in $(\rho h)'$ force are computed in 
+Here we wish to construct $\left [ (\rho h)_0 + (\rho
+  h)'_\mathrm{edge} \right ]$ by predicting $(\rho h)'$ to the edges.
+
+\subsubsection{Predicting $(\rho h)'$ at edges}\label{Predicting rhohprime at edges}
+
+We define $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$.  Then we predict 
+$(\rho h)'$ to edges using {\tt make\_edge\_scal} in {\tt enthalpy\_advance} 
+and the underbraced term in (\ref{rhohprime equation}) as the forcing (see
+also table~\ref{table:pred:hforce} for the forcing term).
+The first two terms in $(\rho h)'$ force are computed in 
 {\tt modify\_scal\_force}, and the last two terms are accounted for in
@@
 
-We define $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$.  Then we predict 
-$(\rho h)'$ to edges using {\tt make\_edge\_scal} in {\tt enthalpy\_advance} 
-and the underbraced term in (\ref{rhohprime equation}) as the forcing.
-
-\subsection{Predicting $(\rho h)_0$ at edges}
+\subsubsection{Predicting $(\rho h)_0$ at edges}
 We use an analogous procedure described in Section \ref{Predicting
@@
 
-\subsection{Computing $\rho h$ at edges}
+For spherical, however, instead of computing $(\rho h)_0$ to edges
+directly, we compute $\rho_0$ and $h_0$ separately at the edges, and
+multiply to get $(\rho h)_0$.
+
+\subsubsection{Computing $\rho h$ at edges}
 We use an analogous procedure described in Section \ref{Computing rho
   at edges} for computing $\rho$ at edges to compute $\rho h$ at
-edges.  For spherical, however, instead of computing $(\rho h)_0$ to
-edges directly, we compute $\rho$ and $h$ separately, and multiply to
-get $(\rho h)_0$.
+edges.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{{\tt enthalpy\_pred\_type} = {\tt predict\_h}}
+
+Here we wish to construct $\left (\rho_0 + \rho'_\mathrm{edge} \right
+) h_\mathrm{edge}$ by predicting $h$ to the edges.
+
+\subsubsection{Predicting $h$ at edges}
+
+We define $h = (rho h)^{(1)}/rho^{(1)}$.  Then we predict $h$ to edges
+using {\tt make\_edge\_scal} in {\tt enthalpy\_advance} and the
+underbraced term in Eq.~\ref{h equation labeled} as the forcing (see
+also table~\ref{table:pred:hforce}).  This force is computed by
+{\tt mkrhohforce} and then divided by $\rho$.  Note: {\tt mkrhoforce}
+knows about the different {\tt enthalpy\_pred\_type}s and computes
+the correct force for this type.
+
+\subsubsection{Computing $\rho h$ at edges}
+
+We use the same procedure described in Section \ref{Computing rho at
+  edges} to computing $\rho$ at edges from $\rho_0$ and
+$\rho'_\mathrm{edge}$ and then multiply by $h_\mathrm{edge}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{{\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_rhohprime}}
+
+Here we wish to construct $\left [ (\rho h)_0 + (\rho
+  h)'_\mathrm{edge} \right ]$ by predicting $T$ to the edges and then
+converting this to $(\rho h)'_\mathrm{edge}$ via the EOS.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{{\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_rhohprime}}
+
+Here we wish to construct $\left (\rho_0 + \rho'_\mathrm{edge} \right
+) h_\mathrm{edge}$ by predicting $T$ to the edges and then converting
+this to $h_\mathrm{edge}$ via the EOS.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Advancing $\rho h$}
@@
 \begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) = 
+\frac{\partial(\rho h)}{\partial t} = 
 -\nabla\cdot\left\{\left[(\rho h)_0^{n+\myhalf,{\rm avg}} 
@@
 
-\begin{table*}
+\begin{table*}[h]
 \begin{center}

### 2011-06-16 docs/pert_notes/pert.tex
almost there...

@@
 
-We define $h = (rho h)^{(1)}/rho^{(1)}$.  Then we predict $h$ to edges
+We define $h = (\rho h)^{(1)}/\rho^{(1)}$.  Then we predict $h$ to edges
 using {\tt make\_edge\_scal} in {\tt enthalpy\_advance} and the
@@
 We use the same procedure described in Section \ref{Computing rho at
-  edges} to computing $\rho$ at edges from $\rho_0$ and
+  edges} to computing $\rho_\mathrm{edge}$ from $\rho_0$ and
 $\rho'_\mathrm{edge}$ and then multiply by $h_\mathrm{edge}$.
@@
 
+\subsubsection{Predicting $T$ at edges}
+
+We predict $T$ to edges using {\tt make\_edge\_scal} in {\tt
+  enthalpy\_advance} and the underbraced term in Eq.~\ref{T equation
+  labeled} as the forcing (see also table~\ref{table:pred:hforce}).
+This force is computed by {\tt mktempforce}.
+
+\subsubsection{Converting $T_\mathrm{edge}$ to $(\rho h)'_\mathrm{edge}$}
+
+We the use EOS in {\tt makeHfromRhoT\_edge} (called from {\tt
+  enthalpy\_advance}) to convert from $T_\mathrm{edge}$ to $(\rho
+h)'_\mathrm{edge}$.  For the EOS call, we need $X_\mathrm{edge}$ and
+$\rho_\mathrm{edge}$.  This construction depends on {\tt
+  species\_pred\_type}, since the species edge states may differ
+between the various prediction types (see the ``species quantity''
+column in table~\ref{table:pred:hoverview}).  The EOS inputs are
+constructed as:
+
+\begin{table*}[h]
+\centering
+\caption{EOS states in {\tt makeHfromRhoT\_edge}\newline}
+\label{table:pred:hforce}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{l|c|c}
+\hline
+\hline
+{\tt species\_pred\_type} & {$\rho$ edge state} &   {$X$ edge state}  \\
+\hline
+{\tt predict\_rhoprime\_and\_X} & 
+  $\rho_0^{n+\myhalf,\rm{avg}} + \rho'_\mathrm{edge}$ & 
+  $X_\mathrm{edge}$ \\
+{\tt predict\_rhoX} & 
+  $\rho_\mathrm{edge}$ & 
+  $(\rho X)_\mathrm{edge}/\rho_\mathrm{edge}$ \\
+\hline
+\end{tabular}
+\end{table*}
+
+The output of  {\tt makeHfromRhoT\_edge} is $(\rho h)'_\mathrm{edge}$.
+
+\subsubsection{Computing $\rho h$ at edges}
+
+The computation of the final $(\rho h)$ edge state is done identically
+as the {\tt predict\_rhohprime} version.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_rhohprime}}
+\subsection{{\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_h}}
 
@@
 this to $h_\mathrm{edge}$ via the EOS.
+
+\subsubsection{Predicting $T$ at edges}
+
+The prediction of $T$ to the edges is done identically as the
+{\tt predict\_T\_then\_rhohprime} version.
+
+\subsubsection{Converting $T_\mathrm{edge}$ to $(\rho h)'_\mathrm{edge}$}
+
+This is identical to the {\tt predict\_T\_then\_rhohprime} version,
+except that on output, we compute $h_\mathrm{edge}$.
+
+\subsubsection{Computing $\rho h$ at edges}
+
+The computation of the final $(\rho h)$ edge state is done identically
+as the {\tt predict\_h} version.
 
@@
 
-\begin{table*}[h]
-\begin{center}
-\caption{When predicting temp ... (conversion done by {\tt makeRhoHfromT}) \newline}
-\begin{tabular}{c|c|c|c}
-\hline
-\hline
-{type} & {Inputs } & {EOS inputs} & {output} \\
-\hline
-3 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ &  $(\rho h)^\prime = \rho h - (\rho h)_0$ \\
-   &                   & $   X = $ given & \\
-   &                   &                 & \\
- 4 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ & $h$ \\
-   &                   & $   X = $ given & \\
-   &                   &                 & \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
+%% \begin{table*}[h]
+%% \begin{center}
+%% \caption{When predicting temp ... (conversion done by {\tt makeRhoHfromT}) \newline}
+%% \begin{tabular}{c|c|c|c}
+%% \hline
+%% \hline
+%% {type} & {Inputs } & {EOS inputs} & {output} \\
+%% \hline
+%% 3 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ &  $(\rho h)^\prime = \rho h - (\rho h)_0$ \\
+%%    &                   & $   X = $ given & \\
+%%    &                   &                 & \\
+%%  4 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ & $h$ \\
+%%    &                   & $   X = $ given & \\
+%%    &                   &                 & \\
+%% \hline
+%% \end{tabular}
+%% \end{center}
+%% \end{table*}
 
-\begin{table*}[h]
-\begin{center}
-\caption{Quantity that is created in {\tt mkflux} and passed into {\tt update\_scal}\newline\
-}
-\begin{tabular}{c|c|c|c|c}
-       &            &              &\multicolumn{2}{c}{Forcing Term} \\
-{type} & {Species } & {$(\rho h)$} & {$(\rho X)$} & {$(\rho h)$} \\
-\hline 
-1 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
-2 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
-3 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r}$ \\[2mm]
-4 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
+%% \begin{table*}[h]
+%% \begin{center}
+%% \caption{Quantity that is created in {\tt mkflux} and passed into {\tt update\_scal}\newline\
+%% }
+%% \begin{tabular}{c|c|c|c|c}
+%%        &            &              &\multicolumn{2}{c}{Forcing Term} \\
+%% {type} & {Species } & {$(\rho h)$} & {$(\rho X)$} & {$(\rho h)$} \\
+%% \hline 
+%% 1 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
+%% 2 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
+%% 3 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r}$ \\[2mm]
+%% 4 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
+%% \hline
+%% \end{tabular}
+%% \end{center}
+%% \end{table*}
 

### 2011-06-17 docs/pert_notes/pert.tex
fix multiply defined labels
fix section heading
fix EOS edge states
minor updates

@@
+\section{General form of an advection equation}
+In predicting the interface states, we work with an advection-like
+equation, and call anything other than the advective term a ``force''.
+For example, for a scalar $\phi$, obeying a conservation law with a source term, $f$:
+\begin{equation}
+\frac{\partial \phi}{\partial t} = -\nabla \cdot (\phi \Ub) + f
+\end{equation}
+We write this in advective form as:
+\begin{equation}
+\frac{\partial \phi}{\partial t} = -\Ub \cdot \nabla \phi \, \underbrace{- \phi \nabla \cdot \Ub}_{\text{``advective force''}} + f
+\end{equation}
+Here we identify the non-advection portion of the divergence of the
+flux as the ``advective force'' for the advective equation.  This
+advective force will be used as a source when predicting the edge
+states.  Depending on the nature of $f$, it may either appear in the
+prediction as well or be absent if the process it describes is
+incorporated by Strang-splitting.
+
 %-----------------------------------------------------------------------------
@@
 (either in full or perturbation form).  In the notes below, we use the
-subscript `{\tt edge}' to indicate what quantity was predicted to the
+subscript `{\tt edge}' to indicate what quantity was {\em predicted} to the
 edges.  In MAESTRO, the different methods of computing $(\rho X)$ on
 edges are controlled by the {\tt species\_pred\_type} parameter.  The
-resulting edge state is shown in the following table:
+quantities predicted to edges and the 
+resulting edge state are shown in the table~\ref{table:pred:species}
 
@@
 \centering
-\begin{tabular}{l|c}
+\caption{Summary of species edge state construction}
+\label{table:pred:species}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{l|c|c}
 \hline
 \hline
-{\tt species\_pred\_type} &   {$(\rho X_k)$ edge state} \\
-\hline \\[-3mm]
-1 / {\tt predict\_rhoprime\_and\_X}  &  $\left(\rho_0^{n+\myhalf,{\rm avg}}
-+ \rho'_\mathrm{edge} \right)(X_k)_\mathrm{edge}$ \\[2mm]
-2 / {\tt predict\_rhoX}  &  $(\rho X_k)_\mathrm{edge}$ \\[2mm]
-3 / {\tt predict\_rho\_and\_X}  &  $\rho_\mathrm{edge} (X_k)_\mathrm{edge}$ \\
+{\tt species\_pred\_type} &   {quantities predicted} & {$(\rho X_k)$ edge state} \\[-5pt]
+ & {in {\tt make\_edge\_scal}} & \\
+\hline 
+1 / {\tt predict\_rhoprime\_and\_X}  &  
+  $\rho'_\mathrm{edge}$, $(X_k)_\mathrm{edge}$ &
+  $\left(\rho_0^{n+\myhalf,{\rm avg}}  
+  + \rho'_\mathrm{edge} \right)(X_k)_\mathrm{edge}$ \\
+2 / {\tt predict\_rhoX}  &  
+  $\rho'_\mathrm{edge}$, $(\rho X_k)_\mathrm{edge}$ &
+  $(\rho X_k)_\mathrm{edge}$ \\
+3 / {\tt predict\_rho\_and\_X}  &  
+  $\rho_\mathrm{edge}$, $(X_k)_\mathrm{edge}$ &
+  $\rho_\mathrm{edge} (X_k)_\mathrm{edge}$ \\
 \hline
@@
 module.
-
-\subsection{Advective force for interface prediction}
-In predicting the interface states, we work with an advection-like
-equation, and call anything other than the advective term a ``force''.
-For example, for a scalar $\phi$, obeying a conservation law with a source term, $f$:
-\begin{equation}
-\frac{\partial \phi}{\partial t} = -\nabla \cdot (\phi \Ub) + f
-\end{equation}
-We write this in advective form as:
-\begin{equation}
-\frac{\partial \phi}{\partial t} = -\Ub \cdot \nabla \phi \, \underbrace{- \phi \nabla \cdot \Ub}_{\text{``advective force''}} + f
-\end{equation}
-Here we identify the non-advection portion of the divergence of the
-flux as the ``advective force'' for the
-advective equation.  In predicting the interface states, this advective
-force will be used in computing the interface states.  Depending on the
-nature of $f$, it may either appear in the prediction as well or be absent
-if the process it describes is incorporated by Strang-splitting.
 
@@
 \frac{\partial\rho'}{\partial t} = -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub 
-- \nabla\cdot(\rho_0\Ubt)}_{\rho' ~ \text{force}},
+- \nabla\cdot(\rho_0\Ubt)}_{\rho' ~ \text{force}} \, .
 \label{rhoprime equation}
 \end{equation}
-where the underbraced ``$\rho'$ force'' is computed in {\tt modify\_scal\_force}.
 
@@
 edges using {\tt make\_edge\_scal} in {\tt density\_advance} and the
-underbraced term in (\ref{rhoprime equation}) as the forcing.
+underbraced term in Eq.~\ref{rhoprime equation} as the forcing.  This
+force is computed in {\tt modify\_scal\_force}.
 
@@
 \item if we predicted $T$, convert this predicted
-  edge state to an' intermediate ``enthalpy'' state (the quotes
+  edge state to an intermediate ``enthalpy'' state (the quotes
   indicate that it may be perturbational or full enthalpy) by calling
@@
 {\tt enthalpy\_pred\_type} &
-{quantity predicted} &
+{cell-centered} &
 {intermediate} &
@@
  &
+{quantity predicted} &
+{``enthalpy''} &
+{available in {\tt mkflux}} &
+{edge state} \\[-5pt]
+ &
+ &
 {in {\tt make\_edge\_scal}} &
-{``enthalpy'' edge state} &
-{available in {\tt mkflux}} &
-{edge state} \\
+{edge state} &
+ &
+ \\
 \hline 
 1 / {\tt predict\_rhoprime\_and\_X}  & 1 / {\tt predict\_rhohprime} &
-  $(\rho h)'$ & $(\rho h)'$ & $X$, $\rho'$ & 
-  $\left [ (\rho h)_0 + (\rho h)' \right ]$ \\
+  $(\rho h)'$ & $(\rho h)'_\mathrm{edge}$ & 
+  $X_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
+  $\left [ (\rho h)_0^{n+\myhalf,{\rm avg}} + (\rho h)'_\mathrm{edge} \right ]$ \\
 1 / {\tt predict\_rhoprime\_and\_X}  & 2 / {\tt predict\_h} &
-  $h$ & $h$ & $X$, $\rho'$ & 
-  $\left ( \rho_0 + \rho' \right ) h$ \\
+  $h$ & $h_\mathrm{edge}$ & 
+  $X_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
+  $\left ( \rho_0^{n+\myhalf,{\rm avg}} + \rho'_\mathrm{edge} \right ) h_\mathrm{edge}$ \\
 1 / {\tt predict\_rhoprime\_and\_X}  & 3 / {\tt predict\_T\_then\_rhohprime} &
-  $T$ & $(\rho h)'$ & $X$, $\rho'$ & 
-  $\left [ (\rho h)_0 + (\rho h)' \right ]$ \\
+  $T$ & $(\rho h)'_\mathrm{edge}$ & 
+  $X_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
+  $\left [ (\rho h)_0^{n+\myhalf,{\rm avg}} + (\rho h)'_\mathrm{edge} \right ]$ \\
 1 / {\tt predict\_rhoprime\_and\_X}  & 4 / {\tt predict\_T\_then\_h} &
-  $T$ & $h$ & $X$, $\rho'$ & 
-  $\left ( \rho_0 + \rho' \right ) h$ \\
+  $T$ & $h_\mathrm{edge}$ & 
+  $X_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
+  $\left ( \rho_0^{n+\myhalf,{\rm avg}} + \rho'_\mathrm{edge} \right ) h_\mathrm{edge}$ \\
 \hline
@@
 2 / {\tt predict\_rhoX}  & 1 / {\tt predict\_rhohprime} &
-  $(\rho h)'$ & $(\rho h)'$ & $(\rho X)$, $\rho'$ & 
-  $\left [ (\rho h)_0 + (\rho h)' \right ]$ \\
+  $(\rho h)'$ & $(\rho h)'_\mathrm{edge}$ & 
+  $(\rho X)_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
+  $\left [ (\rho h)_0^{n+\myhalf,{\rm avg}} + (\rho h)'_\mathrm{edge} \right ]$ \\
 2 / {\tt predict\_rhoX}  & 2 / {\tt predict\_h} &
-  $h$ & $h$ & $(\rho X)$, $\rho'$ & 
-  $\left ( \rho_0 + \rho' \right ) h$ \\
+  $h$ & $h_\mathrm{edge}$ & 
+  $(\rho X)_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
+  $\left ( \rho_0^{n+\myhalf,{\rm avg}} + \rho'_\mathrm{edge} \right ) h_\mathrm{edge}$ \\
 2 / {\tt predict\_rhoX}  & 3 / {\tt predict\_T\_then\_rhohprime} &
-  $T$ & $(\rho h)'$ & $(\rho X)$, $\rho'$ & 
-  $\left [ (\rho h)_0 + (\rho h)' \right ]$ \\
+  $T$ & $(\rho h)'_\mathrm{edge}$ & 
+  $(\rho X)_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
+  $\left [ (\rho h)_0^{n+\myhalf,{\rm avg}} + (\rho h)'_\mathrm{edge} \right ]$ \\
 2 / {\tt predict\_rhoX}  & 4 / {\tt predict\_T\_then\_h} &
-  $T$ & $h$ & $(\rho X)$, $\rho'$ & 
-  $\left ( \rho_0 + \rho' \right ) h$ \\
+  $T$ & $h_\mathrm{edge}$ & 
+  $(\rho X)_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
+  $\left ( \rho_0^{n+\myhalf,{\rm avg}} + \rho'_\mathrm{edge} \right ) h_\mathrm{edge}$ \\
 \hline
@@
 \caption{EOS states in {\tt makeHfromRhoT\_edge}\newline}
-\label{table:pred:hforce}
+\label{table:pred:EOSinputs}
 \renewcommand{\arraystretch}{1.5}
@@
 \hline
-{\tt species\_pred\_type} & {$\rho$ edge state} &   {$X$ edge state}  \\
+{\tt species\_pred\_type} & {$\rho$ edge state} &   {$X_k$ edge state}  \\
 \hline
@@
   $\rho_0^{n+\myhalf,\rm{avg}} + \rho'_\mathrm{edge}$ & 
-  $X_\mathrm{edge}$ \\
+  $(X_k)_\mathrm{edge}$ \\
 {\tt predict\_rhoX} & 
-  $\rho_\mathrm{edge}$ & 
-  $(\rho X)_\mathrm{edge}/\rho_\mathrm{edge}$ \\
+  $\sum_k (\rho X_k)_\mathrm{edge}$ & 
+  $(\rho X_k)_\mathrm{edge}/\sum_k (\rho X_k)_\mathrm{edge}$ \\
 \hline
@@
 
-\subsubsection{Converting $T_\mathrm{edge}$ to $(\rho h)'_\mathrm{edge}$}
+\subsubsection{Converting $T_\mathrm{edge}$ to $h_\mathrm{edge}$}
 

### 2011-06-17 docs/pert_notes/pert.tex
finish the update for now.
species_pred_type = 3 is not listed here.

@@
 
-Here we wish to construct $\left [ (\rho h)_0 + (\rho
+Here we wish to construct $\left [ (\rho h)_0^{n+\myhalf,{\rm avg}} + (\rho
   h)'_\mathrm{edge} \right ]$ by predicting $(\rho h)'$ to the edges.
@@
 
-For spherical, however, instead of computing $(\rho h)_0$ to edges
+For spherical, however, instead of computing $(\rho h)_0$ on edges
 directly, we compute $\rho_0$ and $h_0$ separately at the edges, and
@@
 We use the same procedure described in Section \ref{Computing rho at
-  edges} to computing $\rho_\mathrm{edge}$ from $\rho_0$ and
+  edges} for computing $\rho_\mathrm{edge}$ from $\rho_0$ and
 $\rho'_\mathrm{edge}$ and then multiply by $h_\mathrm{edge}$.
@@
 
-We the use EOS in {\tt makeHfromRhoT\_edge} (called from {\tt
+We call the EOS in {\tt makeHfromRhoT\_edge} (called from {\tt
   enthalpy\_advance}) to convert from $T_\mathrm{edge}$ to $(\rho
@@
 
-The output of  {\tt makeHfromRhoT\_edge} is $(\rho h)'_\mathrm{edge}$.
+After calling the EOS, the output of {\tt makeHfromRhoT\_edge} is
+$(\rho h)'_\mathrm{edge}$.
 
@@
 Section \ref{Advancing rhoX_k}.  The forcing term does not include reaction
-source terms accounted for in {\bf React State}.
+source terms accounted for in {\bf React State}, and is the same
+for all {\tt enthalpy\_pred\_type}s.
 \begin{equation}
 \frac{\partial(\rho h)}{\partial t} = 
--\nabla\cdot\left\{\left[(\rho h)_0^{n+\myhalf,{\rm avg}} 
-+ (\rho h)'\right]\left(\Ubt + w_0\eb_r\right)\right\} + (\Ubt \cdot \er)\frac{\partial p_0}{\partial r} + \psi  \enskip .
+-\nabla\cdot\left\{\left \langle (\rho h) \right \rangle_\mathrm{edge}
+ \left(\Ubt + w_0\eb_r\right)\right\} + (\Ubt \cdot \er)\frac{\partial p_0}{\partial r} + \psi  \enskip .
 \end{equation}
+where $\left \langle (\rho h) \right \rangle_\mathrm{edge}$ is the
+edge state for $(\rho h)$ computed as listed in the final column of
+table~\ref{table:pred:hoverview} for the given {\tt enthalpy\_pred\_type}
+and {\tt species\_pred\_type}.
 

### 2011-06-21 docs/architecture/architecture.tex
expand documentation on example code

@@
 MAESTRO models problems that are in tight hydrostatic equilibrium.
-The fluid state is decomposed into a 1-d radial base state that
+The fluid state is decomposed into a 1D radial base state that
 describes the hydrostatic structure of the star or atmosphere, and a
-2- or 3-d Cartesian full state, that captures the departures from
+2- or 3D Cartesian full state, that captures the departures from
 hydrostatic equilibrium.  Two basic geometries are allowed.  A {\em
   plane-parallel} geometry assumes that the domain is thin compared to
-the radius of curvature of the star, and therefore the 1-d base state
+the radius of curvature of the star, and therefore the 1D base state
 is perfectly aligned with the Cartesian state.  A {\em spherical}
-geometry is for modeling an entire star.  Here, the 1-d base state is
+geometry is for modeling an entire star.  Here, the 1D base state is
 not aligned with the Cartesian state.  Figure~\ref{fig:base_state}
@@
 \includegraphics[height=2.0in]{\archfigpath/base_spherical}
-\caption{\label{fig:base_state} MAESTRO geometries, showing both the
-  1-d base state and the full Cartesian state.  (Left) For multi-level
+\caption[MAESTRO geometries]{\label{fig:base_state} MAESTRO geometries, showing both the
+  1D base state and the full Cartesian state.  (Left) For multi-level
   problems in planar geometry, we force a direct alignment between the
@@
 
- Important subdirectories include:
+ Important sub-directories include:
  \begin{itemize}
@@
 
- A simple Fortran module for reading in 1-d initial model files.  This
+ A simple Fortran module for reading in 1D initial model files.  This
  is used by the initialization routines to get the initial model data.
@@
    Several sub-directories containing routines to generate
-   1-d initial conditions in hydrostatic equilibrium for 
+   1D initial conditions in hydrostatic equilibrium for 
    mapping onto the MAESTRO grid.  See \S~\ref{sec:initial_models}
@@
 
-   Each problem in MAESTRO gets it own subdirectory.  The {\tt
+   Each problem in MAESTRO gets it own sub-directory.  The {\tt
      GNUmakefile} in the problem directory includes the instructions
@@
   various platforms.  Platform/compiler-specific options are stored
-  in the {\tt comps/} subdirectory.
+  in the {\tt comps/} sub-directory.
 
@@
 dwarf) onto the Cartesian domain and uses an external heat source to
-cause the star to expand.  This is the 3-d analogue of the {\tt
+cause the star to expand.  This is the 3D analogue of the {\tt
   test\_basestate} unit test (see \S~\ref{sec:unit_tests}).  This
@@
 parallelism (using MPI or MPI+OpenMP), or compiled as a serial code.
-The top portion of the {\tt GNUmakefile} can be editted to choose the
+The top portion of the {\tt GNUmakefile} can be edited to choose the
 build options.  Setting {\tt MPI := t} will build the code to use MPI
@@
 checkpoints are actually directories, with the data stored in
-subdirectories grouped by refinement level.  Details of the simulation
+sub-directories grouped by refinement level.  Details of the simulation
 (build information, number of processors used, output date, output
@@
    \hline
-   periodic             & -1 \\
-   inlet (user-defined) & 11 \\
-   outlet               & 12 \\
-   symmetry             & 13 \\
-   slip wall            & 14 \\
-   no-slip wall         & 15 \\
+   periodic             & $-1$ \\
+   inlet (user-defined) & $11$ \\
+   outlet               & $12$ \\
+   symmetry             & $13$ \\
+   slip wall            & $14$ \\
+   no-slip wall         & $15$ \\
    \hline
@@
 {\tt Amrvis} is an easy-to-use visualization tool developed at LBL for
-2- and 3-d datasets which can plot slices through 3-d datasets as well
-as volume-renderings.  It can also very easily extract 1-d lines
+2- and 3D datasets which can plot slices through 3D datasets as well
+as volume-renderings.  It can also very easily extract 1D lines
 through the dataset along any coordinate direction.  It is distributed
@@
 
-VisIt is a powerful, DOE-supported visualization tool for 2- and 3-d
+VisIt is a powerful, DOE-supported visualization tool for 2- and 3D
 datasets.  It can do contouring, volume rendering, streamlines, ...\
@@
 
-In addition to the problems descibed above which use the full
+In addition to the problems described above which use the full
 capabilities of MAESTRO, there are a number of unit tests that
@@
 
-  This test initializes a 1-d radial base state quantity with a
-  Gaussian distribution, maps it into the 3-d domain (assuming a
+  This test initializes a 1D radial base state quantity with a
+  Gaussian distribution, maps it into the 3D domain (assuming a
   spherical geometry) using the routines provided by the {\tt
     fill\_3d\_module} module, and then calls {\tt average} to put it
-  back onto a 1-d radial array.  This way we test the accuracy of our
-  procedure to map between the 1-d radial and 3-d Cartesian states.
+  back onto a 1D radial array.  This way we test the accuracy of our
+  procedure to map between the 1D radial and 3D Cartesian states.
   The output from this test was described in detail in
@@
 
-Different MAESTRO problems are defined in subdirectories under
+Different MAESTRO problems are defined in sub-directories under
 {\tt MAESTRO/}.  To add a problem, start by creating a new
@@
 by searching the directories listed in the {\tt GNUmakefile}.
-Customised versions of any source files placed in the problem-directory
+Customized versions of any source files placed in the problem-directory
 override those with the same name found elsewhere.  Any unique
@@
 MAESTRO models subsonic flows that are in hydrostatic equilibrium.
-The solution in MAESTRO is broken up into a 1-d base state and the 2-
-or 3-d full state.  The job of the 1-d base state in the algorithm is
+The solution in MAESTRO is broken up into a 1D base state and the 2-
+or 3D full state.  The job of the 1D base state in the algorithm is
 to represent the hydrostatic structure.  The full, Cartesian state
@@
 for MAESTRO.  There are two types of routines here, the first type
-take an existing 1-d initial model produced somewhere else (e.g.\ a
-1-d stellar evolution code), and and map it onto a uniform grid, at
+take an existing 1D initial model produced somewhere else (e.g.\ a
+1D stellar evolution code), and and map it onto a uniform grid, at
 the desired resolution, using the equation of state in MAESTRO, and
@@
 exists so that we don't have to recompute intersections every time we
-do some operation with two multifabs that have that layout, for
+do some operation with two \multifab s that have that layout, for
 example.
@@
 routine operate on.  A multilevel simulation stores the 
-data in an array of multifabs, where the array index refers
+data in an array of \multifab s, where the array index refers
 to the refinement level.
@@
 To build a \multifab, we need to provide a \layout, the number of
-components to store in the multifab, and the number of ghostcells.  In
-MAESTRO, the hierarchry of grids will be described by a single
+components to store in the \multifab\, and the number of ghostcells.  In
+MAESTRO, the hierarchy of grids will be described by a single
 \mllayout.  A \multifab\ can be declared and built at any time in a
@@
 
-The fluid state (density, enthalpy, species) are stored together in a
-cell-centered multi-component \multifab, typically named {\tt sold},
-{\tt s1}, {\tt s2}, or {\tt snew} (depending on which time-level it
-represents).  The enthalpy is stored as $(\rho h)$, and the species
-are stored as partial-densities $(\rho X_k)$.
-
-Individual state variables should be indexed using the integer keys provided 
-by the {\tt variables} module (see \S \ref{sec:variables_module}).  For example,
-the integer {\tt rho\_comp} will always refer to the density component of the state.
+The fluid state (density, enthalpy, species, temperature, and tracer)
+are stored together in a cell-centered multi-component \multifab,
+typically named {\tt sold}, {\tt s1}, {\tt s2}, or {\tt snew}
+(depending on which time-level it represents).  The enthalpy is stored
+as $(\rho h)$, and the species are stored as partial-densities $(\rho
+X_k)$.  The tracer component is not used at present time, but can
+describe an arbitrary advected quantity.
+
+Individual state variables should be indexed using the integer keys
+provided by the {\tt variables} module (see \S
+\ref{sec:variables_module}).  For example, the integer {\tt rho\_comp}
+will always refer to the density component of the state.
+
+Note: the pressure is not carried as part of the `{\tt s}' \multifab s.
 
@@
 a multilevel \multifab\ array and averages the full Cartesian data
-onto the 1-d base state.
+onto the 1D base state.
 
@@
    \hline
-   eos\_input\_rt       & $\rho$, $T$ \\
-   eos\_input\_rh       & $\rho$, $h$ \\
-   eos\_input\_tp       & $T$, $p$ \\
-   eos\_input\_rp       & $\rho$, $p$ \\
-   eos\_input\_re       & $\rho$, $e$ \\
-   eos\_input\_ps       & $p$, $s$ \\
+   {\tt eos\_input\_rt}       & $\rho$, $T$ \\
+   {\tt eos\_input\_rh}       & $\rho$, $h$ \\
+   {\tt eos\_input\_tp}       & $T$, $p$ \\
+   {\tt eos\_input\_rp}       & $\rho$, $p$ \\
+   {\tt eos\_input\_re}       & $\rho$, $e$ \\
+   {\tt eos\_input\_ps}       & $p$, $s$ \\
    \hline
@@
 
-The {\tt fill\_3d\_module} provides routines that map from the 1-d
-base state to the full Cartesian 2- or 3-d state.  Variations in the
+The {\tt fill\_3d\_module} provides routines that map from the 1D
+base state to the full Cartesian 2- or 3D state.  Variations in the
 routines allow for cell-centered or edge-centered data on either the
@@
 
-The species indices are contiguous in the state array, spanning {\tt spec\_comp:spec\_comp-1+nspec}.
-To find a particular species, a query can be made through the {\tt
-  network} module, such as:
+The species indices are contiguous in the state array, spanning {\tt
+  spec\_comp:spec\_comp-1+nspec}.  To find a particular species, a
+query can be made through the {\tt network} module, such as:
 \begin{verbatim}
@@
 
+Other keys in the {\tt variables} modules are reserved for boundary
+conditions ({\tt foextrap\_comp} and {\tt hoextap\_comp}), the
+projection of the pressure ({\tt press\_comp}), or constructing
+the plotfile.
+
 \section{BoxLib Helper Modules}
@@
 
-\section{\label{sec:example} Example: Accessing State Data}
-
-In MAESTRO, the state data is stored in a multifab array (the array
-index refers to the AMR level).  A typical way to extract the state data
-array looks like:
-
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  subroutine example(s,dx,dt)
+\section{\label{sec:example} Example: Accessing State and MAC Data}
+
+In MAESTRO, the state data is stored in a cell-centered multifab array
+(the array index refers to the AMR level) and the MAC velocities are
+stored in a 2D nodal multifab array (with indices referring to the AMR
+level and the velocity component).  Here we demonstrate a typical way
+to extract the state and MAC velocity data.
+
+All MAESTRO routines are contained in a module, to allow for compile-time
+argument checking.
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
+module example_module
+
+contains
+\end{lstlisting}
+
+The main interface to our routine is called {\tt example}---this will
+take the \multifab s containing the data and then pass them to the
+work routines, {\tt example\_2d} or {\tt example\_3d}, depending on
+the dimensionality.  
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
+  subroutine example(mla,s,umac,dx,dt)
 
@@
     use multifab_module
-    use geometry, only: dm, nlevs
+    use ml_layout_module
     use variables, only: rho_comp
@@
 declaring double precision data, and {\tt multifab\_module} defines
-the \multifab\ data type.  The {\tt geometry} module is a MAESTRO
-module that provides the dimensionality ({\tt dm}) and the number of 
-levels ({\tt nlevs}).  The {\tt variables} module is a MAESTRO
+the \multifab\ data type.  The {\tt ml\_layout\_module} defines the
+datatype for a \mllayout---many routines will take an \mllayout
+to allow us to fill ghostcells.  The {\tt variables} module is a MAESTRO
 module that provides integer keys for indexing the state arrays.  In
@@
 
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
+Next we declare the subroutine arguments:
+
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
+    type(ml_layout), intent(in   ) :: mla
     type(multifab) , intent(inout) :: s(:)
+    type(multifab) , intent(inout) :: umac(:,:)
     real(kind=dp_t), intent(in   ) :: dx(:,:),dt
 \end{lstlisting}
-
-\noindent Next we declare the subroutine arguments.  Here, {\tt s(:)}
-is our \multifab\ array that holds the state data.  The array index
-in {\tt s} refers to the AMR level.
-
+Here, {\tt s(:)} is our \multifab\ array that holds the state data.
+with the array index in {\tt s} refers to the AMR level.  The MAC 
+velocities are held in the multifab {\tt umac}, with the array
+indices referring to the AMR level and the component.
+
+Local variable declarations come next:
 \begin{lstlisting}[language={[95]fortran},mathescape=false]
@@
     real(kind=dp_t), pointer :: sp(:,:,:,:)
-    integer :: i,n,ng_sp
-    integer :: lo(dm),hi(dm)
+    real(kind=dp_t), pointer :: ump(:,:,:,:), vmp(:,:,:,:), wmp(:,:,:,:)
+    integer :: i,n,dm,nlevs,ng_sp,ng_um
+    integer :: lo(mla%dim),hi(mla%dim)
 \end{lstlisting}
@@
 {\tt sp}, that will point to a single \fab\ from the
-\multifab\ {\tt s}.
-
+\multifab\ {\tt s}, and a pointer for each component of the MAC
+velocity, {\tt ump}, {\tt vmp}, and {\tt wmp} (for a 2D run,
+we won't use {\tt wmp}).  We note that regardless of the dimensionality,
+these pointers are 4-dimensional: 3 spatial + 1 component.
+
+Next we get the dimensionality and number of levels
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
+    dm = mla%dim
+    nlevs = mla%nlevel
+\end{lstlisting}
+
+Each multifab can have their own number of ghostcells, so we get
+these next:
 \begin{lstlisting}[language={[95]fortran},mathescape=false]
     ng_sp = nghost(s(1))
-\end{lstlisting}
-
-\noindent Here we get the number of ghostcells for this particular
-\multifab.  This will be needed to access the data stored in the
-\fab s.  Note that all levels in a \multifab\ will have the same
-number of ghostcells, so we can use {\tt s(1)} here.
-
+    ng_um = nghost(umac(1,1))
+\end{lstlisting}
+By convention, all levels in a given multifab have the same number of
+ghostcells, so we use level 1 in the {\tt nghost()} call.  We also use
+the same number of ghostcells for each component of the velocity, so
+we only need to consider the first component in the {\tt nghost()}
+call.  The ghostcells will be needed to access the data stored in the
+\fab s.
+
+To access the data, we loop over all the levels, and all the boxes in
+the given level.
 \begin{lstlisting}[language={[95]fortran},mathescape=false]
     do n=1,nlevs
-       do i = 1, s(n)%nboxes
-\end{lstlisting}
-
-\noindent To access the data, we loop over all the levels, and
-all the boxes in the given level.  {\tt s(n)\%nboxes} is simply
-the number of boxes in level {\tt n}.
-
+       do i = 1, nboxes(s(n))
+\end{lstlisting}
+{\tt nboxes(s(n))} is simply the number of boxes in level {\tt n}.
+All data in all \multifab s will have the same box layout, so we only
+need to query the number of boxes from one \multifab.
+
+For a given \boxtype, we get the data and the bounds of the \boxtype.
 \begin{lstlisting}[language={[95]fortran},mathescape=false]
@@
           sp  => dataptr(s(n), i)
+          ump => dataptr(umac(n,1),i)
+          vmp => dataptr(umac(n,2),i)
           lo =  lwb(get_box(s(n), i))
@@
 The actual data array is accessed through the {\tt dataptr} function,
-which takes a \multifab\ (in this case, {\tt s(n)}) and the index of
-the \boxtype\ ({\tt i}) we want.  {\tt sp} is always four-dimensional:
-3 spatial dimensions and 1 for the components.
+which takes a \multifab\ (e.g.\ {\tt s(n)}) and the index of the
+\boxtype\ ({\tt i}) we want.  We see that the $x$ MAC velocity for the
+current box is stored in {\tt ump} and the $y$ MAC velocity is stored
+in {\tt vmp}.  We don't get the $z$ velocity data here, since that
+would not be available for a 2D run---we defer that until we test on
+the dimensionality below.
 
 Finally, the index bounds of the box (just the data, not the ghostcells) are 
-stored in the {\tt dm}-dimensional arrays {\tt lo} and {\tt hi}
-
+stored in the {\tt dm}-dimensional arrays {\tt lo} and {\tt hi}.  These indices
+refer to the current box, and hold for both the state, {\tt sp}, and the MAC
+velocity, {\tt ump} and {\tt vmp}.  However, since the MAC velocity is nodal
+in the component direction, the loops over the valid data will differ
+slight (as we see below).
+
+With the data extracted, we call a subroutine to operate on it.  We use
+different subroutines for the different dimensionalities (and many times
+have a separate routine for spherical geometries).
 \begin{lstlisting}[language={[95]fortran},mathescape=false]
@@
           case (2)
-             call example_2d(sp(:,:,1,rho_comp),ng_sp,lo,hi,dx(n,:),dt)
+             call example_2d(sp(:,:,1,rho_comp),ng_sp, &
+                             ump(:,:,1,1),vmp(:,:,1,1),ng_um, &
+                             lo,hi,dx(n,:),dt)
           case (3)
-             call example_3d(sp(:,:,:,rho_comp),ng_sp,lo,hi,dx(n,:),dt)
+             wmp => dataptr(umac(n,3),i)
+             call example_3d(sp(:,:,:,rho_comp),ng_sp, &
+                             ump(:,:,:,1),vmp(:,:,:,1),wmp(:,:,:,1),ng_um, &
+                             lo,hi,dx(n,:),dt)
           end select
@@
 \end{lstlisting}
-
-\noindent Finally, in this example, we call either the function
+\noindent We call either the function
 {\tt example\_2d} for two-dimensional data or {\tt example\_3d}
@@
 is a 2D problem, and the density component of the state is selected
-(using the integer key {\tt rho\_comp}).
+(using the integer key {\tt rho\_comp}).  The 3D version accesses
+the data as {\tt sp(:,:,:,rho\_comp)}---only the component regarding
+the variable is needed here.  Notice that we also pass through
+the number of ghostcells for each of the quantities.
 
@@
 example\_3d}, which actually operate on the data.  The form of 
-these functions is:
-
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  subroutine example_2d(density,ng_sp,lo,hi,dx,dt)
+the 2D function is:
+
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
+  subroutine example_2d(density,ng_sp, &
+                        umac,vmac,ng_um, &
+                        lo,hi,dx,dt)
 
@@
 
-    integer        , intent(in) :: lo(:),hi(:), ng_sp
+    integer        , intent(in) :: lo(:),hi(:), ng_sp, ng_um
     real(kind=dp_t), intent(in) :: density(lo(1)-ng_sp:,lo(2)-ng_sp:)
+    real(kind=dp_t), intent(in) ::    umac(lo(1)-ng_um:,lo(2)-ng_um:)
+    real(kind=dp_t), intent(in) ::    vmac(lo(1)-ng_um:,lo(2)-ng_um:)
+
     real(kind=dp_t), intent(in) :: dx(:),dt
@@
     real(kind=dp_t) :: x, y
+    real(kind=dp_t) :: dens, u, v
 
@@
 
-          ! operate on density(i,j)
+          dens = density(i,j)
+
+          ! compute cell-centered velocity
+          u = HALF*(umac(i,j) + umac(i+1,j))
+          v = HALF*(vmac(i,j) + vmac(i+1,j))
+
+          ! operate on the data
           ! ...
@@
   end subroutine example_2d
-\end{lstlisting}
-
-\noindent In this function, the bounds of the {\tt density} array
-take into account the {\tt ng\_sp} ghostcells.  The {\tt j} and {\tt i}
-loops loop over all the valid zones.  Coordinate information is 
-computed from {\tt dx} and {\tt prob\_lo} which is the physical
-lower bound of the domain.  {\tt bl\_constants\_module} declares
-useful double-precision constants, like {\tt HALF} (0.5).
+
+end module example_module
+\end{lstlisting}
+
+\noindent In this function, the bounds of the {\tt density} array take
+into account the {\tt ng\_sp} ghostcells and the index space of the
+current box.  Likewise, the MAC velocities refer to the {\tt ng\_um}
+ghostcells.  The {\tt j} and {\tt i} loops loop over all the valid
+zones.  Coordinate information is computed from {\tt dx} and {\tt
+  prob\_lo} which is the physical lower bound of the domain.  {\tt
+  bl\_constants\_module} declares useful double-precision constants,
+like {\tt HALF} (0.5).  Here, we see how to access the density for
+the current zone and compute the cell-centered velocities from the
+MAC velocities.  By convection, for a nodal array, the indices refer 
+to the {\em lower} interface in the nodal direction, so for {\tt umac},
+{\tt umac(i,j)} and {\tt umac(i+1,j)} are the $x$ MAC velocities
+on the lower and upper edge of the zone in the $x$-direction.
 
@@
 {\tt hi(3)}).
+
+In this example, we looped over the valid zones.  If we wished to loop
+over the interfaces bounding the valid zones, in the $x$-direction,
+we would loop as
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
+  do j = lo(2), hi(2)
+     do i = lo(1), hi(1)+1
+        ! access umac(i,j)
+     enddo
+  enddo
+\end{lstlisting}
 

### 2011-06-21 docs/preface/preface.tex
font play
minor changes

@@
 the DOE Office of Mathematics, Information, and Computational Sciences
-under the U.S. Department of Energy under contract
-No.\ DE-AC02-05CH11231 to LBL and by a DOE/Office of Nuclear Physics
-Outstanding Junior Investigator award, grant No.\ DE-FG02-06ER41448,
-to Stony Brook.
+and the SciDAC Program of the DOE Office of High Energy Physics under
+the U.S. Department of Energy under contract No.\ DE-AC02-05CH11231 to
+LBNL and by a DOE/Office of Nuclear Physics grant No.\
+DE-FG02-06ER41448, to Stony Brook. \\
 
-The current version of this User's Guide can be found in the MAESTRO
-source tree under {\tt fParallel/MAESTRO/docs/}.
+\noindent The current version of the MAESTRO User's Guide can be found in 
+the MAESTRO source tree under {\tt fParallel/MAESTRO/docs/}.

### 2011-06-21 docs/introduction/introduction.tex
minor updates + add papers

@@
 hydrostatic equilibrium.  The idea for MAESTRO grew out of our success
-in applying low Mach number combustion methods to small-scale
+in applying low Mach number combustion methods developed for
+terrestrial flames~\cite{DayBell00} to small-scale
 astrophysical flows.  Our original small-scale astrophysical
@@
 
-\noindent We have two papers that describe applications of the
+\noindent We have several papers that describe applications of the
 method to Type Ia supernovae and X-ray bursts:
@@
   submitted to ApJ (henceforth paper V)
+\item {\em High-Resolution Simulations of Convection Preceding Ignition in
+  Type Ia Supernovae Using Adaptive Mesh Refinement,}
+  A. Nonaka, M. Zingale, A. J. Aspden, A. S. Almgren, J. B. Bell, \& S. E. Woosley
 \end{itemize}
@@
 and $p$ is the dynamical pressure.  The time-evolution equation for
-the velocity (eq.~[\ref{eq:incompressible_u}]) can be solved using
+the velocity (Eq.~\ref{eq:incompressible_u}) can be solved using
 techniques similar to those developed for compressible hydrodynamics,
@@
 take this updated velocity and force it to obey the constraint
-equation.  The basic idea follows from the idea that any vector
+equation.  The basic idea follows from the fact that any vector
 field can be expressed as the sum of a divergence-free quantity and
@@
 $U^\star$, and $\phi$ is a scalar.  Taking the divergence of
-eq~(\ref{eq:decomposition}), we have
+Eq.~(\ref{eq:decomposition}), we have
 \begin{equation}
@@
 With appropriate boundary conditions, this Poisson equation can be
-solved for $\phi$, and the final, divergence free velocity can 
+solved for $\phi$, and the final, divergence-free velocity can 
 be computed as

### 2011-06-21 docs/runtime_parameters/rp.py
add shading

@@ -126,7 +126,14 @@ def make_tex_table():
     print header
 
     # sort the parameters and dump them in latex-fashion
+    odd = 1
     for param in sorted(paramsList):
+        if (odd == 1):
+            print "\\rowcolor{tableShade}"
+            odd = 0
+        else:
+            odd = 1
+
         print "\\verb= ", \
             paramsDict[param].var, \
             " = & ", \

### 2011-06-21 docs/pert_notes/pert.tex
predict_rhoX -> predict_rhoprime_and_rhoX

@@
   + \rho'_\mathrm{edge} \right)(X_k)_\mathrm{edge}$ \\
-2 / {\tt predict\_rhoX}  &  
+2 / {\tt predict\_rhoprime\_and\_rhoX}  &  
   $\rho'_\mathrm{edge}$, $(\rho X_k)_\mathrm{edge}$ &
@@
 \end{table}
-We note the labels {\tt predict\_rhoprime\_and\_X}, {\tt predict\_rhoX}, and
+We note the labels {\tt predict\_rhoprime\_and\_X}, {\tt predict\_rhoprime\_and\_rhoX}, and
 {\tt predict\_rho\_and\_X} are provided by the {\tt pred\_parameters}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\tt species\_pred\_type} = {\tt predict\_rhoX}}
+\subsection{{\tt species\_pred\_type} = {\tt predict\_rhoprime\_and\_rhoX}}
 
@@
 
-\noindent {\tt species\_pred\_type = predict\_rhoX}:
+\noindent {\tt species\_pred\_type = predict\_rhoprime\_and\_rhoX}:
 \begin{equation}
@@
 {``enthalpy''} &
-{available in {\tt mkflux}} &
+{available in} &
 {edge state} \\[-5pt]
@@
 {edge state} &
- &
+{\tt mkflux} &
  \\
@@
 \hline
-2 / {\tt predict\_rhoX}  & 1 / {\tt predict\_rhohprime} &
+2 / {\tt predict\_rhoprime\_and\_rhoX}  & 1 / {\tt predict\_rhohprime} &
   $(\rho h)'$ & $(\rho h)'_\mathrm{edge}$ & 
@@
   $\left [ (\rho h)_0^{n+\myhalf,{\rm avg}} + (\rho h)'_\mathrm{edge} \right ]$ \\
-2 / {\tt predict\_rhoX}  & 2 / {\tt predict\_h} &
+2 / {\tt predict\_rhoprime\_and\_rhoX}  & 2 / {\tt predict\_h} &
   $h$ & $h_\mathrm{edge}$ & 
@@
   $\left ( \rho_0^{n+\myhalf,{\rm avg}} + \rho'_\mathrm{edge} \right ) h_\mathrm{edge}$ \\
-2 / {\tt predict\_rhoX}  & 3 / {\tt predict\_T\_then\_rhohprime} &
+2 / {\tt predict\_rhoprime\_and\_rhoX}  & 3 / {\tt predict\_T\_then\_rhohprime} &
   $T$ & $(\rho h)'_\mathrm{edge}$ & 
@@
   $\left [ (\rho h)_0^{n+\myhalf,{\rm avg}} + (\rho h)'_\mathrm{edge} \right ]$ \\
-2 / {\tt predict\_rhoX}  & 4 / {\tt predict\_T\_then\_h} &
+2 / {\tt predict\_rhoprime\_and\_rhoX}  & 4 / {\tt predict\_T\_then\_h} &
   $T$ & $h_\mathrm{edge}$ & 
@@
   $(X_k)_\mathrm{edge}$ \\
-{\tt predict\_rhoX} & 
+{\tt predict\_rhoprime\_and\_rhoX} & 
   $\sum_k (\rho X_k)_\mathrm{edge}$ & 

### 2011-06-21 docs/pert_notes/pert.tex
add species_pred_type = 3 to the enthalpy notes

@@
 
-\subsection{{\tt species\_pred\_type} = {\tt predict\_rhoprime\_and\_X}}
+\subsection{Method 1: {\tt species\_pred\_type} = {\tt predict\_rhoprime\_and\_X}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\tt species\_pred\_type} = {\tt predict\_rhoprime\_and\_rhoX}}
+\subsection{Method 2: {\tt species\_pred\_type} = {\tt predict\_rhoprime\_and\_rhoX}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\tt species\_pred\_type} = {\tt predict\_rho\_and\_X}}
+\subsection{Method 3: {\tt species\_pred\_type} = {\tt predict\_rho\_and\_X}}
 
@@
 
-\noindent Note: the various {\tt enthalpy\_pred\_type}s are only
-guaranteed to work with {\tt species\_pred\_type = 1}.
+\noindent Note: the various {\tt enthalpy\_pred\_type}s have only been
+extensively tested with {\tt species\_pred\_type = 1}.
 
@@
 \hline
+\hline
+3 / {\tt predict\_rho\_and\_X}  & 1 / {\tt predict\_rhohprime} &
+  $(\rho h)'$ & $(\rho h)'_\mathrm{edge}$ & 
+  $X_\mathrm{edge}$, $\rho_\mathrm{edge}$ & 
+  $\left [ (\rho h)_0^{n+\myhalf,{\rm avg}} + (\rho h)'_\mathrm{edge} \right ]$ \\
+3 / {\tt predict\_rho\_and\_X}  & 2 / {\tt predict\_h} &
+  $h$ & $h_\mathrm{edge}$ & 
+  $X_\mathrm{edge}$, $\rho_\mathrm{edge}$ & 
+  $\rho_\mathrm{edge} h_\mathrm{edge}$ \\
+3 / {\tt predict\_rho\_and\_X}  & 3 / {\tt predict\_T\_then\_rhohprime} &
+  $T$ & $(\rho h)'_\mathrm{edge}$ & 
+  $X_\mathrm{edge}$, $\rho_\mathrm{edge}$ & 
+  $\left [ (\rho h)_0^{n+\myhalf,{\rm avg}} + (\rho h)'_\mathrm{edge} \right ]$ \\
+3 / {\tt predict\_rho\_and\_X}  & 4 / {\tt predict\_T\_then\_h} &
+  $T$ & $h_\mathrm{edge}$ & 
+  $X_\mathrm{edge}$, $\rho_\mathrm{edge}$ & 
+  $\rho_\mathrm{edge} h_\mathrm{edge}$ \\
+\hline
 \end{tabular}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\tt enthalpy\_pred\_type} = {\tt predict\_rhohprime}}
+\subsection{Method 1: {\tt enthalpy\_pred\_type} = {\tt predict\_rhohprime}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\tt enthalpy\_pred\_type} = {\tt predict\_h}}
+\subsection{Method 2: {\tt enthalpy\_pred\_type} = {\tt predict\_h}}
 
-Here we wish to construct $\left (\rho_0 + \rho'_\mathrm{edge} \right
-) h_\mathrm{edge}$ by predicting $h$ to the edges.
+Here, the construction of the interface state depends on what species 
+quantities are present.  In all cases, the enthalpy state is found
+by predicting $h$ to the edges.
+
+For {\tt species\_pred\_types}: {\tt predict\_rhoprime\_and\_X} or
+{\tt predict\_rhoprime\_and\_rhoX}, we wish to construct $(\rho_0 +
+\rho'_\mathrm{edge} ) h_\mathrm{edge}$.
+
+For {\tt species\_pred\_types}: {\tt predict\_rho\_and\_X}, we wish to
+construct $\rho_\mathrm{edge} h_\mathrm{edge}$.
 
@@
 
+{\tt species\_pred\_types}: {\tt predict\_rhoprime\_and\_X} or
+{\tt predict\_rhoprime\_and\_rhoX}:\\[1mm]
+%
 We use the same procedure described in Section \ref{Computing rho at
@@
 
+\ \\
+{\tt species\_pred\_types}: {\tt predict\_rho\_and\_X}:\\[1mm]
+%
+We already have $\rho_\mathrm{edge}$ and simply multiply by
+$h_\mathrm{edge}$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_rhohprime}}
+\subsection{Method 3: {\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_rhohprime}}
 
@@
   $(\rho X_k)_\mathrm{edge}/\sum_k (\rho X_k)_\mathrm{edge}$ \\
+{\tt predict\_rho\_and\_X} & 
+  $\rho_\mathrm{edge}$ & 
+  $(X_k)_\mathrm{edge}$ \\
 \hline
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_h}}
+\subsection{Method 4: {\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_h}}
 
-Here we wish to construct $\left (\rho_0 + \rho'_\mathrm{edge} \right
-) h_\mathrm{edge}$ by predicting $T$ to the edges and then converting
-this to $h_\mathrm{edge}$ via the EOS.
+Here, the construction of the interface state depends on what species
+quantities are present.  In all cases, the enthalpy state is found by
+predicting $T$ to the edges and then converting this to
+$h_\mathrm{edge}$ via the EOS.
+
+For {\tt species\_pred\_types}: {\tt predict\_rhoprime\_and\_X} or
+{\tt predict\_rhoprime\_and\_rhoX}, we wish to construct $(\rho_0 +                                             \rho'_\mathrm{edge} ) h_\mathrm{edge}$.
+
+For {\tt species\_pred\_types}: {\tt predict\_rho\_and\_X}, we wish to
+construct $\rho_\mathrm{edge} h_\mathrm{edge}$.
 

### 2011-06-21 docs/preface/preface.tex
add caveat

@@
-MAESTRO is developed at the Center for Computational Sciences and
+This documentation is a work in progress.  MAESTRO is rapidly
+evolving, so portions of the text are likely to be incomplete or
+out-of-date. \\
+
+\noindent MAESTRO is developed at the Center for Computational Sciences and
 Engineering at Lawrence Berkeley National Laboratory and at Stony

### 2011-06-23 docs/analysis/analysis.tex
fix list
move some particle stuff from architecture.  Some updates.

--- 2010-03-04 docs/analysis/analysis.tex
@@
 number of Fortran-based analysis routines for BoxLib datasets.  Many
-of these can be used with both MAESTRO and the compressible
-astrophysics code, Castro.
+of these can be used with both \maestro\ and the compressible
+astrophysics code, \castro.
 
@@
   Laterally average each of the variables in a plotfile (works for
-  both 2-d and 3-d).  This is written with MAESTRO
+  both 2-d and 3-d).  This is written with \maestro\
   plane-parallel geometry plotfiles in mind, and the averaging is done
@@
 
+\item {\tt fvarnames.f90}
+
+  Simply print out the list of variables stored in a plotfile.
+
+\end{itemize}
+
 \subsection{Data Processing Example}
 
-A well-commented example processing routine is provided in {\tt
-  data\_processing/tutorial} as {\tt fspeciesmass2d.f90}.  This
-routine simply computes the total mass of a particular species on the
-entire domain for a 2-d dataset.  It is written to understand a
-multilevel (AMR) dataset, and only considers the finest-available data
-at any physical location in the computational domain.
+The routine {\tt fspeciesmass2d.f90} in {\tt
+  data\_processing/tutorial} serves as a well-commented example of how
+to work with \maestro\ plotfile data.  This routine simply computes
+the total mass of a particular species on the entire domain for a 2-d
+dataset.  It is written to understand a multilevel (AMR) dataset, and
+only considers the finest-available data at any physical location in
+the computational domain.
 
@@
 
-\end{itemize}
+\subsection{Particle routines}
+
+\label{analysis:sec:particles}
+
+The {\tt parseparticles.py} routine in the {\tt python/} subdirectory
+can read in \maestro\ particle files containing particle histories
+(usually named {\tt timestamp\_??}).  See the discussion in \S~\ref{arch:sec:particles} for details on initialization.  The driver {\tt
+  test\_parseparticles.py} shows shows how to use this module to plot
+particle histories.
+
+As a concrete example, running {\tt test2} with particles enabled
+will seed particles in the initial hotspot.  To plot the results,
+first set your {\tt PYTHONPATH} environment variable to point to the
+{\tt fParallel/data\_processing/python/} directory, for example:
+\begin{verbatim}
+export PYTHONPATH="/home/username/development/MAESTRO/fParallel/data_processing/python"
+\end{verbatim}
+This will allow python to see the {\tt parseparticles.py} routine.
+For the {\tt test2} problem, the {\tt plotparticles.py} routine shows
+how to plot the particle histories and make an animation of the
+particles colored by the ash mass fraction.  This script is run as:
+\begin{verbatim}
+./plotparticles.py timestamp_*
+\end{verbatim}
+
+Note, these python routines require the NumPy and matplotlib packages.
+On a Fedora Linux system, the necessary routines can be installed via:
+\begin{verbatim}
+yum install python-matplotlib lyx-fonts stix-fonts
+\end{verbatim}

### 2011-06-23 docs/architecture/architecture.tex
move some particle stuff from architecture.  Define \maestro and \castro
use \maestro

@@
-\section{MAESTRO Basics}
-
-MAESTRO models problems that are in tight hydrostatic equilibrium.
+\section{\maestro\ Basics}
+
+\maestro\ models problems that are in tight hydrostatic equilibrium.
 The fluid state is decomposed into a 1D radial base state that
@@
 \includegraphics[height=2.0in]{\archfigpath/base_spherical}
-\caption[MAESTRO geometries]{\label{fig:base_state} MAESTRO geometries, showing both the
+\caption[\maestro\ geometries]{\label{fig:base_state} \maestro\ geometries, showing both the
   1D base state and the full Cartesian state.  (Left) For multi-level
@@
 
-MAESTRO can use adaptive mesh refinement to focus resolution on
+\maestro\ can use adaptive mesh refinement to focus resolution on
 complex regions of flow.  For Cartesian/plane-parallel geometries, all
@@
 
-\section{MAESTRO Directory Structure}
+\section{\maestro\ Directory Structure}
 
@@
 \includegraphics[width=5.0in]{\archfigpath/maestro_directory2}
-\caption[MAESTRO directory structure]
-{The basic MAESTRO directory structure}
+\caption[\maestro\ directory structure]
+{The basic \maestro\ directory structure}
 \end{figure}
 
-All the files needed to build MAESTRO are contained in the {\tt
+All the files needed to build \maestro\ are contained in the {\tt
   fParallel} directory structure.  Problems piece together various
-MAESTRO directories, choosing a reaction network, equation of state,
+\maestro\ directories, choosing a reaction network, equation of state,
 and conductivity routine to build an executable.  Briefly, the
@@
  of boxes.  The BoxLib modules define the basic datatypes used
- in MAESTRO.  BoxLib also provides the routines that handle the
+ in \maestro.  BoxLib also provides the routines that handle the
  parallelization and I/O.
@@
  A collection of various EOSes for use with the code.  The two
- important ones for MAESTRO are {\tt helmeos/} and {\tt
+ important ones for \maestro\ are {\tt helmeos/} and {\tt
  gamma\_law\_general/}.
@@
 
- Various reaction networks for MAESTRO problems.  In addition to
+ Various reaction networks for \maestro\ problems.  In addition to
  providing a routine to evolve the nuclear species due to reactions,
@@
 
- The main MAESTRO algorithm directory.  The individual problem directories
- lie under here, and all of the main MAESTRO source lives in the {\tt Source/}
+ The main \maestro\ algorithm directory.  The individual problem directories
+ lie under here, and all of the main \maestro\ source lives in the {\tt Source/}
  sub-directory.
@@
    1D initial conditions in hydrostatic equilibrium for 
-   mapping onto the MAESTRO grid.  See \S~\ref{sec:initial_models}
+   mapping onto the \maestro\ grid.  See \S~\ref{sec:initial_models}
    for details.
@@
 
-   The main MAESTRO source.  Here you will find the driver routine, the
-   advection routines, etc.  All MAESTRO problems will compile this
+   The main \maestro\ source.  Here you will find the driver routine, the
+   advection routines, etc.  All \maestro\ problems will compile this
    source.
@@
 
-   Each problem in MAESTRO gets it own sub-directory.  The {\tt
+   Each problem in \maestro\ gets it own sub-directory.  The {\tt
      GNUmakefile} in the problem directory includes the instructions
@@
    This allows problems to have custom versions of the main
-   MAESTRO routines (e.g.\ initial conditions via {\tt initdata.f90}.
+   \maestro\ routines (e.g.\ initial conditions via {\tt initdata.f90}.
    See \S~\ref{sec:makefile} for details on the build system).
@@
   Some simple scripts that are useful for building, running,
-  maintaining MAESTRO.
+  maintaining \maestro.
 
@@
 
-In this section we give an overview of MAESTRO, including some of the
+In this section we give an overview of \maestro, including some of the
 standard problems, how to run the code, some basic runtime parameters,
@@
 
-Different problems in MAESTRO are contained in individual
+Different problems in \maestro\ are contained in individual
 sub-directories under {\tt MAESTRO/}.  The {\tt GNUmakefile}
@@
 transfers heat to the fuel, driving a burning front.  This problem
-does not use the traditional MAESTRO elliptic constraint, but
+does not use the traditional \maestro\ elliptic constraint, but
 rather enforces $\nabla \cdot U = S$ (through the runtime parameter
@@
 
-To build the MAESTRO executable, simply type `{\tt make}' in the
+To build the \maestro\ executable, simply type `{\tt make}' in the
 problem directory.  The executable will have a name like {\tt
   main.Linux.Intel.exe}, where the specific parts of the name depend
-on the compilers and OS used.  MAESTRO can be built with support for
+on the compilers and OS used.  \maestro\ can be built with support for
 parallelism (using MPI or MPI+OpenMP), or compiled as a serial code.
@@
 
-\item {\tt restart} tells MAESTRO to restart from a checkpoint.  The
+\item {\tt restart} tells \maestro\ to restart from a checkpoint.  The
       value of this parameter should be the file number to restart from.
@@
 through the dataset along any coordinate direction.  It is distributed
-separately from the MAESTRO distribution.
+separately from the \maestro\ distribution.
 
@@
 
-By default, MAESTRO outputs global diagnostics each timestep into a file
+By default, \maestro\ outputs global diagnostics each timestep into a file
 called {\tt maestro\_diag.out}.  This includes the maximum Mach number,
@@
 In addition to the problems described above which use the full
-capabilities of MAESTRO, there are a number of unit tests that
-exercise only specific components of the MAESTRO solvers.  The
+capabilities of \maestro, there are a number of unit tests that
+exercise only specific components of the \maestro\ solvers.  The
 tests have their own drivers (a custom {\tt varden.f90}) that
 initialize only the data needed for the specific test and call
-specific MAESTRO routines directly.
+specific \maestro\ routines directly.
 
@@
   This test initializes a Gaussian temperature profile and calls
-  the thermal diffusion routines in MAESTRO to evolve the state 
+  the thermal diffusion routines in \maestro\ to evolve the state 
   considering only diffusion.  The driver estimates a timestep
@@
 
-Different MAESTRO problems are defined in sub-directories under
+Different \maestro\ problems are defined in sub-directories under
 {\tt MAESTRO/}.  To add a problem, start by creating a new
@@
 \end{lstlisting}
-The MAESTRO build system knows what options to use for various
+The \maestro\ build system knows what options to use for various
 compiler families.  The {\tt COMP} flag specifies which compiler to
@@
 
-A MAESTRO application is built from several packages (the
+A \maestro\ application is built from several packages (the
 multigrid solver, an EOS, a reaction network, etc.).  The core
-MAESTRO packages are always included, so a problem only needs
+\maestro\ packages are always included, so a problem only needs
 to define the EOS, reaction network, and conductivities to
@@
 
-The runtime parameters for the core MAESTRO algorithm are listed
+The runtime parameters for the core \maestro\ algorithm are listed
 in {\tt MAESTRO/\_parameters}.  That file is parsed at compile-time
@@
 
-MAESTRO models subsonic flows that are in hydrostatic equilibrium.
-The solution in MAESTRO is broken up into a 1D base state and the 2-
+\maestro\ models subsonic flows that are in hydrostatic equilibrium.
+The solution in \maestro\ is broken up into a 1D base state and the 2-
 or 3D full state.  The job of the 1D base state in the algorithm is
@@
 The routines in {\tt MAESTRO/initial\_model/} prepare an initial model
-for MAESTRO.  There are two types of routines here, the first type
+for \maestro.  There are two types of routines here, the first type
 take an existing 1D initial model produced somewhere else (e.g.\ a
 1D stellar evolution code), and and map it onto a uniform grid, at
-the desired resolution, using the equation of state in MAESTRO, and
-using MAESTRO's discretization of hydrostatic equilibrium.  The second
+the desired resolution, using the equation of state in \maestro, and
+using \maestro's discretization of hydrostatic equilibrium.  The second
 type generate the initial model internally, by integrating the
@@
 
-MAESTRO initial models are read in by the {\tt extern/model\_parser}
+\maestro\ initial models are read in by the {\tt extern/model\_parser}
 routines.  This expects the initial model to contain a header giving
@@
 model should contain the same species data (as mass fractions) as
-defined in the {\tt network} module used by the MAESTRO problem.
+defined in the {\tt network} module used by the \maestro\ problem.
 
@@
 
-MAESTRO's gridding is handled by the BoxLib library.  The underlying
+\maestro's gridding is handled by the BoxLib library.  The underlying
 idea in BoxLib is to allow for adaptive mesh refinement (AMR)---different
@@
 in that direction (see Figure~\ref{fig:dataloc}).  Data that is on the
-corners is nodal in all spatial directions.  In MAESTRO, the state
+corners is nodal in all spatial directions.  In \maestro, the state
 data (density, enthalpy, velocity, $\ldots$) is generally
@@
 A \fab\ would represent the data for a single box in the domain.
-In MAESTRO, we don't usually deal with \fab s alone, but rather
+In \maestro, we don't usually deal with \fab s alone, but rather
 we deal with \multifab s, described next.
@@
 A \multifab\ is a collection of \fab s at the same level of
-refinement.  This is the primary data structure that MAESTRO
+refinement.  This is the primary data structure that \maestro\
 routine operate on.  A multilevel simulation stores the 
@@
 components to store in the \multifab\, and the number of ghostcells.  In
-MAESTRO, the hierarchy of grids will be described by a single
+\maestro\, the hierarchy of grids will be described by a single
 \mllayout.  A \multifab\ can be declared and built at any time in a
@@
 A \bctower\ holds the information about what boundary conditions are
-in effect for each variable in a MAESTRO simulation.  \MarginPar{better description?}
-
-\section{MAESTRO Data Organization}
-
-The state of the star in MAESTRO is described by both a
+in effect for each variable in a \maestro\ simulation.  \MarginPar{better description?}
+
+\section{\maestro\ Data Organization}
+
+The state of the star in \maestro\ is described by both a
 multidimensional state and the 1D base state.  The full
@@
 is simply stored in Fortran arrays.  Here we describe the
-major MAESTRO data-structures.
+major \maestro\ data-structures.
 
@@
 
-\section{MAESTRO Helper Modules}
-
-A number of MAESTRO modules appear frequently throughout the source.
+\section{\maestro\ Helper Modules}
+
+A number of \maestro\ modules appear frequently throughout the source.
 Below, we describe some of the more common functionality of the most
@@
 
-Presently there are 2 equations of state that work with MAESTRO:
+Presently there are 2 equations of state that work with \maestro:
 \begin{itemize}
@@
 The {\tt eos\_module} declares for convenience the variables that need
-appear in the {\tt eos} call argument list.  MAESTRO routines use
+appear in the {\tt eos} call argument list.  \maestro\ routines use
 these module variables in the EOS call to avoid having to declare 
@@
 code ({\tt nspec}), their ordering, and gives their basic properties
-(like atomic number, $Z$, and atomic mass, $A$).  All MAESTRO problems
+(like atomic number, $Z$, and atomic mass, $A$).  All \maestro\ problems
 require a {\tt network} module, even if there are no reactions
@@
 
-In MAESTRO, the state data is stored in a cell-centered multifab array
+In \maestro, the state data is stored in a cell-centered multifab array
 (the array index refers to the AMR level) and the MAC velocities are
@@
 
-All MAESTRO routines are contained in a module, to allow for compile-time
+All \maestro\ routines are contained in a module, to allow for compile-time
 argument checking.
@@
 datatype for a \mllayout---many routines will take an \mllayout
-to allow us to fill ghostcells.  The {\tt variables} module is a MAESTRO
+to allow us to fill ghostcells.  The {\tt variables} module is a \maestro\
 module that provides integer keys for indexing the state arrays.  In
@@
 
-MAESTRO uses the multigrid solver to enforce the divergence constraint
+\maestro\ uses the multigrid solver to enforce the divergence constraint
 both on the half-time edge-centered advective velocities (the ``MAC
@@
 
-MAESTRO has support for Lagrangian particles that are passively
+\label{arch:sec:particles}
+
+\maestro\ has support for Lagrangian particles that are passively
 advected with the velocity field.  These are useful for diagnostics
@@
 Some simple python scripts allow for the plotting of the particle
-positions.  For example, running {\tt test2} with particles enabled
-will seed particles in the initial hotspot.  To plot the results,
-first set your {\tt PYTHONPATH} environment variable to point to the
-{\tt fParallel/data\_processing/python/} directory, for example:
-\begin{verbatim}
-export PYTHONPATH="/home/username/development/MAESTRO/fParallel/data_processing/python"
-\end{verbatim}
-This will allow python to see the {\tt parseparticles.py} routine.
-For the {\tt test2} problem, the {\tt plotparticles.py} routine shows
-how to plot the particle histories and make an animation of the
-particles colored by the ash mass fraction.  This script is run as:
-\begin{verbatim}
-./plotparticles.py timestamp_*
-\end{verbatim}
-
-Note, these python routines require the NumPy and matplotlib packages.
-On a Fedora Linux system, the necessary routines can be installed via:
-\begin{verbatim}
-yum install python-matplotlib lyx-fonts stix-fonts
-\end{verbatim}
+positions.  See \S~\ref{analysis:sec:particles} for details.

### 2011-06-23 docs/initial_models/initial_models.tex
use maestrosymbols

@@
 Here we briefly describe the various routines for generating initial models
-for the main MAESTRO problems and how the initial model is used to initialize
+for the main \maestro\ problems and how the initial model is used to initialize
 both the base state and the full Cartesian state.
@@
 
-We have found that for the best performance, the MAESTRO initialization procedure should be given model data with the same resolution as the base state resolution, $\Delta r$.
-For planar problems, $\Delta r = \Delta x$.
-For multilevel planar problems, we use $\Delta r$ equal to $\Delta x$ at the finest resolution.
-For spherical problems we set $\Delta r = \Delta x/\mathtt{drdxfac}$.
-We generate our initial model either analytically or from raw data,
- $\rho^{\raw}, T^{\raw}, p^{\raw}$, and $X^{\raw}$.  
-Here is the raw data file for each test problem:
+We have found that for the best performance, the \maestro\
+initialization procedure should be given model data with the same
+resolution as the base state resolution, $\Delta r$.  For planar
+problems, $\Delta r = \Delta x$.  For multilevel planar problems, we
+use $\Delta r$ equal to $\Delta x$ at the finest resolution.  For
+spherical problems we set $\Delta r = \Delta x/\mathtt{drdxfac}$.  We
+generate our initial model either analytically or from raw data,
+$\rho^{\raw}, T^{\raw}, p^{\raw}$, and $X^{\raw}$.  Here is the raw
+data file for each test problem:
 \begin{eqnarray}
@@
 The model data is not generated at run-time - it must be generated in
-advance of running any MAESTRO examples.  The inputs file should point
+advance of running any \maestro\ examples.  The inputs file should point
 to the file containing the model data.

### 2011-06-23 docs/introduction/introduction.tex
use maestrosymbols

@@
-\section{History of MAESTRO}
+\section{History of \maestro}
 
-MAESTRO describes the evolution of low Mach number flows that are in
-hydrostatic equilibrium.  The idea for MAESTRO grew out of our success
+\maestro\ describes the evolution of low Mach number flows that are in
+hydrostatic equilibrium.  The idea for \maestro\ grew out of our success
 in applying low Mach number combustion methods developed for
@@
 
-\noindent MAESTRO was developed initially for modeling the convecting
+\noindent \maestro\ was developed initially for modeling the convecting
 phase in a white dwarf preceding the ignition of a Type Ia supernovae.
@@
 large-scale stratification in the star.  The method closest in spirit
-to MAESTRO is the pseudo-incompressible method of
+to \maestro\ is the pseudo-incompressible method of
 Durran~\cite{durran}, developed for terrestrial atmospheric flows
 (assuming an ideal gas).  Part of the complexity of the equations in
-MAESTRO stems from the need to descibe a general equation of state.
+\maestro\ stems from the need to descibe a general equation of state.
 Additionally, since reactions can significantly alter the hydrostatic
@@
 operates on the cell-centered velocities at the new time, again
-enforcing the divergence constraint.  The MAESTRO algorithm performs
+enforcing the divergence constraint.  The \maestro\ algorithm performs
 both of these projections.
 
-The MAESTRO algorithm builds upon these ideas, using a different
+The \maestro\ algorithm builds upon these ideas, using a different
 velocity constraint equation that captures the compressibility
@@
 
-Throughout the papers describing MAESTRO, we've largely kept our
+Throughout the papers describing \maestro, we've largely kept our
 notation consistent.  Table~\ref{table:sym} defines the

### 2011-06-23 docs/lo_density/lo_density.tex
caption fix
use maestrosymbols

@@
-These are working notes for the low density parameters in {\tt MAESTRO}.  In low density
-regions, we modify the behavior of the algorithm.  Here is a summary of some parameters,
-and a brief description of what they do.
+These are working notes for the low density parameters in \maestro.
+In low density regions, we modify the behavior of the algorithm.  Here
+is a summary of some parameters, and a brief description of what they
+do.
 \begin{itemize}
@@
 \includegraphics[width=4.0in]{\lodensfigpath/cutoff}\hspace{0.2in}
-\caption{Image of how the cutoff density and cutoff coordinates
+\caption[Cutoff density and coordinates]{Image of how the cutoff density and cutoff coordinates
 are related for single-level planar and all spherical problems.}

### 2011-06-23 docs/SDC/sdc.tex
typo
use maestrosymbols.tex

@@
 
-Maestro integrates the following system of equations:
+\maestro\ integrates the following system of equations:
 \begin{eqnarray}
@@
 
-In the Strang-split version of MAESTRO, the reaction and diffusion
+In the Strang-split version of \maestro, the reaction and advection
 processes operate independent of one-another.  The species and
@@
 
-In the Strang-split version of Maestro, the reaction network solves just
+In the Strang-split version of \maestro, the reaction network solves just
 the reaction portion of the species evolution equations along with a
@@
 due to the advection.  These are constructed by calling {\tt density\_advance}
-and {\tt enthalpy\_advance} in Maestro and passed into the network solver
+and {\tt enthalpy\_advance} in \maestro\ and passed into the network solver
 during the reaction step.  These terms approximate the following:

### 2011-06-23 docs/managing_jobs/managingjobs.tex
use maestrosymbols.tex

@@
 {\tt process.jaguar} in {\tt wdconvect/job\_scripts/}
-are designed to allow you to run MAESTRO with minimal interaction,
+are designed to allow you to run \maestro\ with minimal interaction,
 while being assured that the data is archived to HPSS on the NCCS

### 2011-06-23 docs/pert_notes/pert.tex
use maestrosymbols.tex

@@
 subscript `{\tt edge}' to indicate what quantity was {\em predicted} to the
-edges.  In MAESTRO, the different methods of computing $(\rho X)$ on
+edges.  In \maestro, the different methods of computing $(\rho X)$ on
 edges are controlled by the {\tt species\_pred\_type} parameter.  The
@@
 We note that in essence these choices reflect a hyperbolic (1)
-vs.\ elliptic (2) approach.  In MAESTRO, if we setup a problem with
+vs.\ elliptic (2) approach.  In \maestro, if we setup a problem with
 $\rho = \rho_0$ initially, and enforce a constraint $\nabla \cdot
@@
 
-Maestro solves an enthalpy equation.  
+\maestro\ solves an enthalpy equation.  
 The full enthalpy equation is

### 2011-06-23 docs/preface/preface.tex
use maestrosymbols.tex

--- 2011-06-21 docs/preface/preface.tex
@@
-MAESTRO is developed at the Center for Computational Sciences and
+This documentation is a work in progress.  \maestro\ is rapidly
+evolving, so portions of the text are likely to be incomplete or
+out-of-date. \\
+
+\noindent \maestro\ is developed at the Center for Computational Sciences and
 Engineering at Lawrence Berkeley National Laboratory and at Stony
@@
 
-\noindent The current version of the MAESTRO User's Guide can be found in 
-the MAESTRO source tree under {\tt fParallel/MAESTRO/docs/}.
+\noindent The current version of the \maestro\ User's Guide can be found in 
+the \maestro\ source tree under {\tt fParallel/MAESTRO/docs/}.

### 2011-06-23 docs/visualization/visualization.tex
add the visualization chapter
use maestrosymbols.tex

@@
 the specific {\tt Enzo}isms, and now can work on various data formats,
-such as the BoxLib data from {\tt MAESTRO} and {\tt CASTRO}, HDF5 data
+such as the BoxLib data from \maestro\ and \castro, HDF5 data
 from {\tt FLASH}, and data from the SPH code {\tt Gadget}.  For more
@@
 \centering
-\includegraphics[height=0.4\textheight]{my_cool_images_Slice_x_tfromp_trim}
+\includegraphics[height=0.4\textheight]{\visfigpath/my_cool_images_Slice_x_tfromp_trim}
 \caption{Example slice through 3-d dataset with \yt.}

### 2011-06-24 docs/architecture/architecture.tex
list of tables

@@
    domain boundaries in the $x$, $y$, and $z$ coordinate directions.
-   The different types are set via integer flags noted below:
-
+   The different types are set via integer flags listed in table~\ref{arch:table:bcflags}.
+
+   \begin{table}[h]
+   \caption{\label{arch:table:bcflags} Boundary condition flags}  
    \begin{center}
@@
    \end{center}
+   \end{table}
 
@@
 specifies which thermodynamic variables (in addition to the mass
-fractions) are used as input.  Options include:
-
+fractions) are used as input.  EOS input options are listed 
+in table~\ref{arch:table:eosinput}.
+
+   \begin{table}[h]
+   \caption{\label{arch:table:eosinput} EOS input flags}
    \begin{center}
@@
 multifabs and other arrays dealing with the scalar quantities.  The
-most commonly used keys are are:
-
+most commonly used keys are are list in table~\ref{arch:table:variables}.
+
+\begin{table}[h]
+\caption{\label{arch:table:variables} Common {\tt variables} module keys}
 \begin{center}
@@
 \end{center}
+\end{table}
 

### 2011-06-27 docs/pert_notes/pert.tex
list of tables
add details about advective form vs. conservative form in make_edge_scal

@@
-\section{General form of an advection equation}
-In predicting the interface states, we work with an advection-like
-equation, and call anything other than the advective term a ``force''.
-For example, for a scalar $\phi$, obeying a conservation law with a source term, $f$:
+\section{Predicting interface states}
+
+The \maestro\ hyperbolic equations come in two forms: advective and conservative.
+The procedure for predicting interface states differs slightly depending on which
+form we are dealing with.
+
+\subsection{Advective form}
+
+Most of the time, we are dealing with equations in advective form.
+Here, a scalar, $\phi$, obeys:
 \begin{equation}
-\frac{\partial \phi}{\partial t} = -\nabla \cdot (\phi \Ub) + f
+\frac{\partial \phi}{\partial t} = -\Ub \cdot \nabla \phi + f
 \end{equation}
-We write this in advective form as:
+where $f$ is the force.  This is the form that the perturbation
+equations take, as well as the equation for $X_k$ itself.
+
+A piecewise linear prediction of $\phi$ to the interface 
+would appear as:
+\begin{eqnarray}
+\phi_{i+1/2,j}^{n+1/2} &=& \phi_{i,j}^n 
+    + \left . \frac{\Delta x}{2} \frac{\partial \phi}{\partial x} \right |_{i,j}
+    + \left . \frac{\Delta t}{2} \frac{\partial \phi}{\partial t} \right |_{i,j} \\
+ &=& \phi_{i,j}^n 
+    + \left . \frac{\Delta x}{2} \frac{\partial \phi}{\partial x} \right |_{i,j}
+    +  \frac{\Delta t}{2} \left ( -u \frac{\partial \phi}{\partial x} 
+                                         -v \frac{\partial \phi}{\partial y} + f \right ) \\
+ &=& \phi_{i,j}^n + \frac{\Delta x}{2} \left ( 1 - \frac{\Delta t}{\Delta x} u \right ) 
+           \frac{\partial \phi}{\partial x} 
+    \underbrace{- \frac{\Delta t}{2} v \frac{\partial \phi}{\partial y}}_{\mathrm{``transverse~term''}} + \frac{\Delta t}{2} f
+\end{eqnarray}
+(see the Godunov notes section for more details).  Here, the
+``transverse term'' is accounted for in {\tt make\_edge\_scal}.  Any
+additional forces should be added to $f$.  For the perturbation form
+of equations, we add additional advection-like terms to $f$ by calling
+{\tt modify\_scal\_force}.  This will be noted below.
+
+\subsection{Conservative form}
+
+A conservative equation takes the form:
 \begin{equation}
-\frac{\partial \phi}{\partial t} = -\Ub \cdot \nabla \phi \, \underbrace{- \phi \nabla \cdot \Ub}_{\text{``advective force''}} + f
+\frac{\partial \phi}{\partial t} = -\nabla \cdot ( \phi \Ub) + f
 \end{equation}
-Here we identify the non-advection portion of the divergence of the
-flux as the ``advective force'' for the advective equation.  This
-advective force will be used as a source when predicting the edge
-states.  Depending on the nature of $f$, it may either appear in the
-prediction as well or be absent if the process it describes is
-incorporated by Strang-splitting.
+%
+Now a piecewise linear prediction of $\phi$ to the interface is 
+\begin{eqnarray}
+\phi_{i+1/2,j}^{n+1/2} &=& \phi_{i,j}^n 
+    + \left . \frac{\Delta x}{2} \frac{\partial \phi}{\partial x} \right |_{i,j}
+    + \left . \frac{\Delta t}{2} \frac{\partial \phi}{\partial t} \right |_{i,j} \\
+ &=& \phi_{i,j}^n 
+    + \left . \frac{\Delta x}{2} \frac{\partial \phi}{\partial x} \right |_{i,j}
+    +  \frac{\Delta t}{2} \left ( -\frac{\partial (\phi u)}{\partial x} 
+                                  -\frac{\partial (\phi v)}{\partial y} + f \right ) \\
+ &=& \phi_{i,j}^n + \frac{\Delta x}{2} \left ( 1 - \frac{\Delta t}{\Delta x} u \right ) 
+           \frac{\partial \phi}{\partial x} 
+    \underbrace{- \frac{\Delta t}{2} \phi \frac{\partial u}{\partial x} 
+                - \frac{\Delta t}{2} \frac{\partial (\phi v)}{\partial y}}_{\mathrm{``transverse~term''}} + \frac{\Delta t}{2} f
+\end{eqnarray}
+Here the ``transverse term'' includes the non-advective portion of the $x$-flux (for the $x$-prediction).
+Both of the underbraced terms are accounted for in {\tt make\_edge\_scal} automatically when
+we call it with {\tt is\_conservative = .true.}.
 
@@
 \centering
-\caption{Forcing term into {\tt make\_edge\_scal} \newline}
+\caption{Forcing term into {\tt make\_edge\_scal}}
 \label{table:pred:hforce}
@@
 \centering
-\caption{EOS states in {\tt makeHfromRhoT\_edge}\newline}
+\caption{EOS states in {\tt makeHfromRhoT\_edge}}
 \label{table:pred:EOSinputs}

### 2011-07-07 docs/visualization/visualization.tex
add list of plotfiles

@@
+\section{Plotfiles}
+
+\maestro\ outputs plotfile specifically for visualization and analysis.
+Table~\ref{vis:table:plotfile} lists the quantities stored in a plotfile.
+Not all of these may be present, dependent on what options were used in 
+creating the plotfile.
+
+By default, plotfiles store double precision data, but if {\tt
+  single\_prec\_plotfiles = .true.} is set, then the data is converted
+to single precision before outputting---this is done to reduce file
+sizes.
+
+\renewcommand{\arraystretch}{1.3}
+
+\begin{center}
+\begin{longtable}{|l|p{2.25in}|p{2.5in}|}
+\caption{Plotfile quantities}
+\label{vis:table:plotfile} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{plotfile variable name}} & 
+       \multicolumn{1}{|c|}{\textbf{description}} & 
+       \multicolumn{1}{|c|}{\textbf{runtime parameter controlling output}} \\ \hline
+\endfirsthead
+
+\multicolumn{3}{c}%          
+{{\tablename\ \thetable{}---continued}} \\
+\hline {plotfile variable name} & 
+       {description} & 
+       {runtime parameter controlling output} \\ \hline
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline
+\endlastfoot
+
+{\tt x\_vel}   & $\ut$  & -- \\
+\hline
+{\tt y\_vel}   & $\vt$  & -- \\
+\hline
+{\tt z\_vel}   & $\wt$  & 3-d runs online \\
+\hline
+{\tt density}  & $\rho$ & -- \\
+\hline
+{\tt rhoh}   & $(\rho h)$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} amd {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+\hline
+{\tt h}      & $(\rho h)/\rho$   & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} amd {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+\hline
+{\tt X(*)}   & $(\rho X_k)/\rho$ & {\tt plot\_spec} \\
+\hline
+{\tt tracer} & tracers           & {\tt plot\_trac} \\
+\hline
+{\tt w0\_x}  & $w_0 \er \cdot \ex$ & {\tt plot\_base} \\ 
+\hline
+{\tt w0\_y}  & $w_0 \er \cdot \ey$ & {\tt plot\_base} \\ 
+\hline
+{\tt w0\_z}  & $w_0 \er \cdot \ez$ & {\tt plot\_base} \\ 
+\hline
+{\tt divw0}  & $\nabla \cdot w_0$  & {\tt plot\_base} \\ 
+\hline
+{\tt rho0}   & $\rho_0$            & {\tt plot\_base} \\ 
+\hline
+{\tt rhoh0}  & $(\rho h)_0$         & {\tt plot\_base} \\ 
+\hline
+{\tt h0}     & $(\rho h)_0/\rho_0$  & {\tt plot\_base} \\ 
+\hline
+{\tt p0}     & $p_0$                & {\tt plot\_base} \\ 
+\hline
+{\tt radial\_velocity}  & $\Ubt \cdot \er + w_0$ & {\tt spherical == 1} \\
+\hline
+{\tt circum\_velocity}  & $| \Ubt - (\Ubt \cdot \er) \er |$ &  {\tt spherical == 1}  \\
+\hline
+{\tt magvel}              & $| \Ubt + w_0 \er |$  & -- \\
+\hline
+{\tt momentum}            & $\rho | \Ubt + w_0 \er |$  & -- \\
+\hline
+{\tt vort}                & $| \nabla \times \Ubt |$   & -- \\
+\hline
+{\tt S}                   & $S$   & -- \\
+\hline
+{\tt rhopert}             & $\rho - \rho_0$  & -- \\
+\hline
+{\tt rhohpert}          & $(\rho h) - (\rho h)_0$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} amd {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+\hline
+{\tt tfromp}              & $T(\rho, p_0, X_k)$  & -- \\
+\hline
+{\tt tfromh}            & $T(\rho, h, X_k)$    & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} amd {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+\hline
+{\tt deltaT}            & $[T(\rho, h, X_k) - T(\rho, p_0, X_k)]/T(\rho, h, X_k)$  & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} amd {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+\hline
+{\tt deltap}            & $|p(\rho,h,X_k) - p_0|/p_0$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} amd {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+\hline
+{\tt tpert}               & $T(\rho,h,X_k) - \overline{T}$ if {\tt use\_tfromp = F}; $T(\rho,p_0,X_k) - \overline{T}$ otherwise & -- \\
+\hline
+{\tt Machnumber}          & $| \Ubt + w_0 \er | / c(\rho,p_0,X_k)$ & -- \\
+\hline
+{\tt soundspeed}        & $c(\rho,p_0,X_k)$ & {\tt plot\_cs} \\
+\hline
+{\tt deltagamma}          & $\Gamma_1(\rho,p_0,X_k) - \gammabar$ & -- \\
+\hline
+{\tt entropy}             & $s(\rho,p_0,X_k)$ & -- \\
+\hline
+{\tt entropypert}         & $[s(\rho,p_0,X_k) - \overline{s}]/\overline{s}$ & -- \\
+\hline
+{\tt sponge} or {\tt sponge\_fdamp} & $1/(1 + \Delta t \kappa f_\mathrm{damp})$ by default; $f_\mathrm{damp}$ if {\tt plot\_sponge\_fdamp = T} & -- \\
+\hline
+{\tt pi}                 & $\pi$ & -- \\
+\hline
+{\tt gpi\_x}           & $\nabla \pi \cdot \ex$ & {\tt plot\_gpi} \\
+\hline 
+{\tt gpi\_y}           & $\nabla \pi \cdot \ey$ & {\tt plot\_gpi} \\
+\hline
+{\tt gpi\_z}           & $\nabla \pi \cdot \ez$ & {\tt plot\_gpi} \\
+\hline
+{\tt pioverp0}         & $\pi / p_0$ & {\tt plot\_base} \\
+\hline
+{\tt p0pluspi}         & $p_0 + \pi$ & {\tt plot\_base} \\
+\hline
+{\tt omegadot(*)}      & $(\rho X_k)/\rho$ & {\tt plot\_omegadot} \\
+\hline
+{\tt enucdot}          & $(\rho \Hnuc)/\rho$ & {\tt plot\_Hnuc} \\
+\hline
+{\tt Hext}             & $(\rho \Hext)/\rho$ & {\tt plot\_Hext} \\
+\hline
+{\tt eta\_rho}         & $\etarho$ & {\tt plot\_eta} \\
+\hline
+{\tt thermal}          & $\nabla \cdot \kappa \nabla T$ & {\tt use\_thermal\_diffusion} \\
+\hline
+{\tt conductivity}     & $\kappa(\rho, T, X_k)$ & {\tt use\_thermal\_diffusion} \\
+\hline
+{\tt ad\_excess}       & $\nabla - \nabla_\mathrm{ad}$ & {\tt plot\_ad\_excess} \\
+\hline
+\end{longtable}
+\end{center}
+
+\renewcommand{\arraystretch}{1.0}
+
 \section{Visualizing with \amrvis}
+
 \section{Visualizing with \visit}
+
 \section{Visualizing with \yt}

### 2011-07-18 docs/visualization/visualization.tex
typo
typos

@@
 \renewcommand{\arraystretch}{1.3}
-
+{\footnotesize
 \begin{center}
 \begin{longtable}{|l|p{2.25in}|p{2.5in}|}
+
 \caption{Plotfile quantities}
@@
 \hline
-{\tt z\_vel}   & $\wt$  & 3-d runs online \\
+{\tt z\_vel}   & $\wt$  & 3-d runs only \\
 \hline
@@
 \hline
-{\tt rhoh}   & $(\rho h)$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} amd {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+{\tt rhoh}   & $(\rho h)$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
 \hline
-{\tt h}      & $(\rho h)/\rho$   & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} amd {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+{\tt h}      & $(\rho h)/\rho$   & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
 \hline
@@
 \hline
-{\tt rhohpert}          & $(\rho h) - (\rho h)_0$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} amd {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+{\tt rhohpert}          & $(\rho h) - (\rho h)_0$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
 \hline
@@
 \hline
-{\tt tfromh}            & $T(\rho, h, X_k)$    & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} amd {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+{\tt tfromh}            & $T(\rho, h, X_k)$    & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
 \hline
-{\tt deltaT}            & $[T(\rho, h, X_k) - T(\rho, p_0, X_k)]/T(\rho, h, X_k)$  & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} amd {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+{\tt deltaT}            & $[T(\rho, h, X_k) - T(\rho, p_0, X_k)]/T(\rho, h, X_k)$  & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
 \hline
-{\tt deltap}            & $|p(\rho,h,X_k) - p_0|/p_0$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} amd {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+{\tt deltap}            & $|p(\rho,h,X_k) - p_0|/p_0$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
 \hline
@@
 \end{center}
+}
 

### 2011-07-19 docs/analysis/analysis.tex
add docstring example for parseparticles

@@
 can read in \maestro\ particle files containing particle histories
-(usually named {\tt timestamp\_??}).  See the discussion in \S~\ref{arch:sec:particles} for details on initialization.  The driver {\tt
-  test\_parseparticles.py} shows shows how to use this module to plot
-particle histories.
+(usually named {\tt timestamp\_??}).  See the discussion in
+\S~\ref{arch:sec:particles} for details on initializing particles in
+\maestro.  The driver {\tt test\_parseparticles.py} shows shows how to
+use this module to plot particle histories.  Additional documentation
+is available from the module itself.  In the python environment,
+type:
+\begin{lstlisting}[language=Python]
+import parseparticles
+help(parseparticles)
+\end{lstlisting}
+to get information on the classes and functions provided by the {\tt
+  parseparticles} module.
 

### 2011-07-23 docs/getting_started/getting_started.tex
split getting started from architecture

--- 2011-06-24 docs/architecture/architecture.tex
@@
-\section{\maestro\ Basics}
-
-\maestro\ models problems that are in tight hydrostatic equilibrium.
-The fluid state is decomposed into a 1D radial base state that
-describes the hydrostatic structure of the star or atmosphere, and a
-2- or 3D Cartesian full state, that captures the departures from
-hydrostatic equilibrium.  Two basic geometries are allowed.  A {\em
-  plane-parallel} geometry assumes that the domain is thin compared to
-the radius of curvature of the star, and therefore the 1D base state
-is perfectly aligned with the Cartesian state.  A {\em spherical}
-geometry is for modeling an entire star.  Here, the 1D base state is
-not aligned with the Cartesian state.  Figure~\ref{fig:base_state}
-shows these geometries.
-
-\begin{figure}[tb]
-\centering
-\includegraphics[height=2.0in]{\archfigpath/base_grid} \hspace{0.5in}
-\includegraphics[height=2.0in]{\archfigpath/base_spherical}
-\caption[\maestro\ geometries]{\label{fig:base_state} \maestro\ geometries, showing both the
-  1D base state and the full Cartesian state.  (Left) For multi-level
-  problems in planar geometry, we force a direct alignment between the
-  radial array cell centers and the Cartesian grid cell centers by
-  allowing the radial base state spacing to change with space and
-  time.  (Right) For multi-level problems in spherical geometry, since
-  there is no direct alignment between the radial array cell centers
-  and the Cartesian grid cell centers, we choose to fix the radial
-  base state spacing across levels. Figure taken
-  from~\cite{multilevel}.}
-\end{figure}
-
-\maestro\ can use adaptive mesh refinement to focus resolution on
-complex regions of flow.  For Cartesian/plane-parallel geometries, all
-cells at the same height must be at the same level of refinement.
-This restriction is to allow for the base state to directly align with
-the Cartesian state everywhere.  For spherical geometries, there is no
-such restriction (again, see Figure~\ref{fig:base_state}).
-
-\section{\maestro\ Directory Structure}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=5.0in]{\archfigpath/maestro_directory2}
-\caption[\maestro\ directory structure]
-{The basic \maestro\ directory structure}
-\end{figure}
-
-All the files needed to build \maestro\ are contained in the {\tt
-  fParallel} directory structure.  Problems piece together various
-\maestro\ directories, choosing a reaction network, equation of state,
-and conductivity routine to build an executable.  Briefly, the
-sub-directories are:
-\begin{itemize}
-\item {\tt boxlib/}
-
- BoxLib is a library for describing meshes consisting of a union
- of boxes.  The BoxLib modules define the basic datatypes used
- in \maestro.  BoxLib also provides the routines that handle the
- parallelization and I/O.
-
-\item {\tt data\_processing/}
-
- Simple Fortran-based analysis routines (e.g.\ extract a line from a
- multidimensional dataset) that operate on BoxLib datasets.
-
-\item {\tt extern/}
-
- External modules, like ODE integrators, equations of state, reaction
- networks
-
- Important sub-directories include:
- \begin{itemize}
- \item {\tt conductivity/}
-
- Various routines for computing the thermal conductivity used in the
- thermal diffusion part of the algorithm.
-
- \item {\tt constants/}
-
- A simple module to define physical constants
-
- \item {\tt EOS/}
-
- A collection of various EOSes for use with the code.  The two
- important ones for \maestro\ are {\tt helmeos/} and {\tt
- gamma\_law\_general/}.
-
- \item {\tt model\_parser/}
-
- A simple Fortran module for reading in 1D initial model files.  This
- is used by the initialization routines to get the initial model data.
-
- \item {\tt networks/}
-
- Various reaction networks for \maestro\ problems.  In addition to
- providing a routine to evolve the nuclear species due to reactions,
- the networks also define the species that are advected by the code.
-
- \item {\tt VODE/}
-
- An integration package for ODEs.  At the moment, this is used 
- for integrating various reaction networks.
-
- \end{itemize}
-
-\item {\tt MAESTRO/}
-
- The main \maestro\ algorithm directory.  The individual problem directories
- lie under here, and all of the main \maestro\ source lives in the {\tt Source/}
- sub-directory.
-
- Important directories under {\tt MAESTRO/} include:
-
- \begin{itemize}
-
- \item {\tt docs/}
-
-   Documentation describing the basic algorithm (including this
-   document).
-
- \item {\tt initial\_models}
-
-   Several sub-directories containing routines to generate
-   1D initial conditions in hydrostatic equilibrium for 
-   mapping onto the \maestro\ grid.  See \S~\ref{sec:initial_models}
-   for details.
-
- \item {\tt Source}
-
-   The main \maestro\ source.  Here you will find the driver routine, the
-   advection routines, etc.  All \maestro\ problems will compile this
-   source.
-
- \item problem directories: {\tt test2}, {\tt
-   test\_spherical}, $\ldots$
-
-   Each problem in \maestro\ gets it own sub-directory.  The {\tt
-     GNUmakefile} in the problem directory includes the instructions
-   on how to build the executable, including what modules in {\tt
-     extern} are used.
-
-   Any file that you place in a sub-directory here takes 
-   precedence over a file of the same name in {\tt MAESTRO/}.
-   This allows problems to have custom versions of the main
-   \maestro\ routines (e.g.\ initial conditions via {\tt initdata.f90}.
-   See \S~\ref{sec:makefile} for details on the build system).
-
- \end{itemize}
-
-\item {\tt mg/}
-
-  The multigrid solver, with support for both cell-centered and
-  node-centered data.
-
-\item {\tt mk/}
-
-  The generic Makefiles that store the compilation flags for
-  various platforms.  Platform/compiler-specific options are stored
-  in the {\tt comps/} sub-directory.
-
-\item {\tt scripts/}
-
-  Some simple scripts that are useful for building, running,
-  maintaining \maestro.
-
-\end{itemize}
-
-\section{Getting Started}
-
-In this section we give an overview of \maestro, including some of the
+In this chapter we give an overview of \maestro, including some of the
 standard problems, how to run the code, some basic runtime parameters,
@@
 
-\subsection{`Standard' Test Problems}
+\section{`Standard' Test Problems}
 
@@
 
-\subsection{Compiling and Running}
+\section{Compiling and Running}
 
@@
 outline the most common runtime parameters.
+
+\section{Common Runtime Parameters}
 
@@
 
-\subsection{Working with the Output}
+\section{Working with the Output}
 
@@
 
-\subsubsection{\tt Amrvis}
+\subsection{\tt Amrvis}
 
@@
 
-\subsubsection{{\tt data\_processing} scripts}
+\subsection{{\tt data\_processing} scripts}
 
@@
 
-\subsubsection{VisIt}
+\subsection{VisIt}
 
@@
 
-\subsubsection{Diagnostic Files}
+\subsection{Diagnostic Files}
 
@@
 output.
-
-\subsection{Unit Tests}
-
-\label{sec:unit_tests}
-
-In addition to the problems described above which use the full
-capabilities of \maestro, there are a number of unit tests that
-exercise only specific components of the \maestro\ solvers.  The
-tests have their own drivers (a custom {\tt varden.f90}) that
-initialize only the data needed for the specific test and call
-specific \maestro\ routines directly.
-
-\begin{itemize}
-\item {\tt test\_advect} \\[-3mm]
-
-  This test initializes a Gaussian density field (no other scalar
-  quantities are used) and a uniform velocity field in any one of the
-  coordinate directions.  The Gaussian profile is advected through
-  the period domain exactly once and the error in the density profile
-  (L2 norm) is computed.  The driver for this problem does this 
-  for every dimension twice (once with a positive velocity and once
-  with a negative velocity).  After all coordinate directions are 
-  tested, the norms are compared to ensure that the error does
-  not show any directional bias.
-
-\item {\tt test\_average} \\[-3mm]
-
-  This test initializes a 1D radial base state quantity with a
-  Gaussian distribution, maps it into the 3D domain (assuming a
-  spherical geometry) using the routines provided by the {\tt
-    fill\_3d\_module} module, and then calls {\tt average} to put it
-  back onto a 1D radial array.  This way we test the accuracy of our
-  procedure to map between the 1D radial and 3D Cartesian states.
-  The output from this test was described in detail in
-  \cite{multilevel}.
-
-\item {\tt test\_basestate} \\[-3mm]
-
-  This test initializes the base state to contain a hydrostatic
-  model and then evolves the state with heating to watch the 
-  hydrostatic adjustment of the atmosphere.  In particular,
-  the base state velocity, $w_0$, is computed in response to 
-  the heating and this is used to advect the base state density
-  and compute the new pressure, $p_0$.  An early version of 
-  this routine was used for the plane-parallel expansion test
-  in \cite{lowMach2}.  This version of the test was also shown
-  for a spherical, self-gravitating star in \cite{multilevel}.
-
-\item {\tt test\_diffusion}
-
-  This test initializes a Gaussian temperature profile and calls
-  the thermal diffusion routines in \maestro\ to evolve the state 
-  considering only diffusion.  The driver estimates a timestep
-  based on the explicit thermal diffusion timescale and loops
-  over calls to the thermal diffusion solver.  A Gaussian remains
-  Gaussian when diffusing, so an explicit error can be computed
-  by comparing to the analytic solution.  This test is 
-  described in \cite{xrb}.
-
-\item {\tt test\_particles}
-
-  This test exercises the particle advection routine.  A simple
-  circular velocity field, with the magnitude increasing with radius
-  from the center is initialized.  A number of particles are then
-  initialized at various radii from the center and they are advected
-  for one period.  The particle paths should be perfect circles, and
-  the final particle position should overlap with the initial
-  position.
-
-  Particle data is stored separately from the fluid data.  Instead
-  of being part of the plotfiles, the particle data is outputted
-  each timestep into files named {\tt timestamp\_*}, where 
-  the number indicates which processor did the writing.  These
-  particle files can be processed and the particle data plotted
-  using the python routines in {\tt data\_processing/python/}.
-
-\end{itemize}
-
-\section{Adding A New Problem}
-\label{sec:adding_problems}
-
-Different \maestro\ problems are defined in sub-directories under
-{\tt MAESTRO/}.  To add a problem, start by creating a new
-sub-directory---this is where you will compile your problem and
-store all the problem-specific files.
-
-The minimum requirement to
-define a new problem would be a {\tt GNUmakefile} which describes how
-to build the application and an input file which lists the runtime
-parameters.  The problem-specific executable is built in the problem
-directory by typing {\tt make}.  Source files are found automatically
-by searching the directories listed in the {\tt GNUmakefile}.
-Customized versions of any source files placed in the problem-directory
-override those with the same name found elsewhere.  Any unique
-source files (and not simply a custom version of a file found
-elsewhere) needs to be listed in a file call {\tt GPackage.mak}
-in the problem-directory.
-
-\subsection{The {\tt GNUmakefile}}
-
-\label{sec:makefile}
-
-A basic {\tt GNUmakefile} begins with:
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  NDEBUG := t
-  MPI    :=
-  OMP    :=
-\end{lstlisting}
-Here, {\tt NDEBUG} is true if we are building an optimized executable.
-Otherwise, the debug version is built---this typically uses less
-optimization and adds various runtime checks through compiler flags.
-{\tt MPI} and {\tt OMP} are set to true if we want to use either MPI
-or OpenMP for parallelization.  If {\tt MPI := t}, you will need to
-have the MPI libraries installed, and their location may need to be 
-specified in {\tt MAESTRO/mk/GMakeMPI.mak}.
-
-The next line sets the compiler to be used for compilation:
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  COMP := Intel
-\end{lstlisting}
-The \maestro\ build system knows what options to use for various
-compiler families.  The {\tt COMP} flag specifies which compiler to
-use.  Allowed values include {\tt Intel}, {\tt gfortran}, {\tt PGI},
-and {\tt PathScale}.  The specific details of these choices are
-defined in the {\tt MAESTRO/mk/comps/} directory.
-
-{\tt MKVERBOSE} set to true will echo the build commands to the
-terminal as the are executed.
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  MKVERBOSE := t
-\end{lstlisting}
-
-The next two lines define where the top of the source tree is
-include the basic Makefile information (like compiler
-flags).
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  FPARALLEL := ../..
-
-  include $(FPARALLEL)/mk/GMakedefs.mak
-\end{lstlisting}
-%$
-
-A \maestro\ application is built from several packages (the
-multigrid solver, an EOS, a reaction network, etc.).  The core
-\maestro\ packages are always included, so a problem only needs
-to define the EOS, reaction network, and conductivities to
-use, as well as any extra, problem-specific files.  
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-EOS_DIR := extern/EOS/helmeos   
-CONDUCTIVITY_DIR := extern/conductivity/constant
-NETWORK_DIR := extern/networks/ignition
-
-EXTRA_DIR := extern/random
-\end{lstlisting}
-Generally, one does not need to include the problem directory itself
-in {\tt EXTRA\_DIR}, unless there are unique source files found there,
-described in a {\tt GPackage.mak} file.  These variables are
-interpreted by the {\tt GMaestro.mak} file and used to build a master
-list of packages called {\tt Fmdirs}.  The build system will attempt
-to build all of the files listed in the various {\tt GPackage.mak}
-files found in the {\tt Fmdirs} directories.  Furthermore, {\tt
-  Fmdirs} will be will be added to the {\tt make} {\tt VPATH}, which
-is the list of directories to search for source files.  The problem
-directory will always be put first in the {\tt VPATH}, so any source
-files placed there override those with the same name found elsewhere
-in the source tree.
-
-Some packages (for instance, the {\tt helmeos}
-EOS) require Fortran include files.  The {\tt Fmincludes} variable
-lists all those directories that contain include files that are
-inserted into the Fortran source at compile time via the {\tt include}
-statement.  Presently, the only instance of this is with the Helmholtz
-general equation of state found in {\tt extern/EOS/helmeos/}.  This is
-automatically handled by the {\tt GMaestro.mak} instructions.
-
-Runtime parameters listed in the {\tt MAESTRO/\_parameters} file are
-parsed at compile time and the file {\tt probin.f90} is written and
-compiled.  This is a Fortran module that holds the values of the
-runtime parameters and makes them available to any routine.  The
-variable {\tt PROBIN\_PARAMETERS} should list all of the files that
-contain runtime parameters for the current problem.  You must always
-include the main {\tt ../\_parameters} file (i.e.\ {\tt
-  MAESTRO/\_parameters}).  Sometimes a problem defines their own
-runtime parameters (for instance {\tt \_parameters.myproblem}), in which
-case, its parameter file should also appear in {\tt
-  PROBIN\_PARAMETERS} (see \S~\ref{sec:def_runtime_param}).
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  PROBIN_PARAMETERS := ../_parameters
-\end{lstlisting}
-
-The final line in the {\tt GNUmakefile} includes the rules to actually
-build the executable.
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  include $(FPARALLEL)/MAESTRO/GMaestro.mak
-\end{lstlisting}
-%$
-
-\subsubsection{Handling Problem-Specific Source Files}
-
-As mentioned above, any source files placed in the problem directory
-override a files with the same name found elsewhere in the source
-tree.  This allows you to create a problem-specific version of any
-routine.  Source files that are unique to this problem (i.e.\ there is
-no file with the same name elsewhere in the source tree) need to be
-listed in a file {\tt GPackage.mak} in the problem directory, and
-the problem-directory needs to be explicitly listed in the {\tt EXTRA\_DIR}
-list in the {\tt GNUmakefile}.
-
-\subsection{Defining Runtime Parameters}
-
-\label{sec:def_runtime_param}
-
-The runtime parameters for the core \maestro\ algorithm are listed
-in {\tt MAESTRO/\_parameters}.  That file is parsed at compile-time
-by the {\tt MAESTRO/write\_probin.py} script
-(along with any other parameter files listed in the {\tt PROBIN\_PARAMETERS}
-variable in the {GNUmakefile}, see \S \ref{sec:makefile}).  The script
-outputs the {\tt probin.f90} source file.  Each line
-in the {\tt \_parameters} file has the form: 
-\vskip 3mm
-{\em parameter} \hskip 10em  {\em data-type} \hskip 10em  {\em value} 
-\vskip 3mm
-\noindent where {\em parameter} is the name of the runtime parameter,
-{\em data-type} is one of \{{\tt character}, {\tt real}, {\tt
-  integer}, {\tt logical}\}, and the {\em value} specifies the default
-value for the runtime parameter.  Comments are indicated by a `{\tt
-  \#}' character and a used to produce documentation about the
-available runtime parameters.
-
-At runtime, the default values for the parameters can be overridden
-either through the inputs file (by adding a line of the form: {\tt
-  parameter = value}) or through a command-line argument (taking the
-form: {\tt --parameter value}).  The {\tt probin\_module} makes the
-values of the runtime parameters available to the various functions
-in the code (see \S~\ref{sec:probin}).
-
-Problem-specify runtime parameters should be defined in the
-problem-directory in a file called {\tt
-  \_parameters}\{{\em.problem-name}\}.  This file should be added
-to the {\tt PROBIN\_PARAMETERS} variable in the {\tt GNUmakefile}.
-
-\subsection{Preparing the Initial Model}
-
-\label{sec:initial_models}
-
-\maestro\ models subsonic flows that are in hydrostatic equilibrium.
-The solution in \maestro\ is broken up into a 1D base state and the 2-
-or 3D full state.  The job of the 1D base state in the algorithm is
-to represent the hydrostatic structure.  The full, Cartesian state
-carries the departures from hydrostatic equilibrium.  The underlying
-formulation of the low Mach number equations assumes that the base
-state is in hydrostatic equilibrium.  At the start of a simulation,
-the initial model is read in and taken as the base state.  Therefore,
-any initial model needs to already be in hydrostatic equilibrium.
-
-The routines in {\tt MAESTRO/initial\_model/} prepare an initial model
-for \maestro.  There are two types of routines here, the first type
-take an existing 1D initial model produced somewhere else (e.g.\ a
-1D stellar evolution code), and and map it onto a uniform grid, at
-the desired resolution, using the equation of state in \maestro, and
-using \maestro's discretization of hydrostatic equilibrium.  The second
-type generate the initial model internally, by integrating the
-condition of hydrostatic equilibrium together with a simplifying
-assumption on the energy (e.g. isothermal or isentropic).  In
-both cases hydrostatic equilibrium is enforced as:
-\begin{equation}
-\frac{p_{i+1} - p_i}{\Delta r} = \frac{1}{2} (\rho_i + \rho_{i+1})
-g_{i+1/2}
-\end{equation}
-Here, $g_{i+1/2}$ is the edge-centered gravitational acceleration.
-
-The {\tt toy\_atm} example provides a simple approximation for a thin
-(plane-parallel) convectively-unstable accreted layer on the surface
-of a star.  This can be used as the starting point for a more complex
-model.
-
-\maestro\ initial models are read in by the {\tt extern/model\_parser}
-routines.  This expects the initial model to contain a header giving
-the number of variables and their names, followed by rows of data
-giving the coordinate and data values at that coordinate.  The initial
-model should contain the same species data (as mass fractions) as
-defined in the {\tt network} module used by the \maestro\ problem.
-
-Full details on which initial model routine matches each problem and
-how the initial models are used to initialize the full state data can
-be found in \S~\ref{sec:initial_models_main}.
-
-\section{BoxLib Data Structures}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=6.5in]{\archfigpath/data_loc2}
-\caption[Data-centerings on the grid]
-  {\label{fig:dataloc} Some of the different data-centerings:
-  (a) cell-centered, (b) nodal in the $x$-direction, and (c) nodal in
-  both the $x$- and $y$-directions.  Note that for nodal data, the
-  integer index corresponds to the lower boundary in that direction.
-  In each of these centerings, the red point has the same indices:\ (1,2).
-  Not shown is the case where data is nodal in the $y$-direction only.}
-\end{figure}
-
-\maestro's gridding is handled by the BoxLib library.  The underlying
-idea in BoxLib is to allow for adaptive mesh refinement (AMR)---different
-regions of the domain can have different spatial resolutions.  The
-domain is broken into boxes.  At the coarsest (base) level of
-refinement, the entire computational domain is covered, either by one
-box or broken across many boxes.  Higher levels of refinement have
-finer zones (typically $2\times$).  Only a portion of the domain may
-be covered by the higher levels of refinement.
-
-\MarginPar{Nesting? ghost cells?}
-For parallel computations, the boxes are spread across processors, in
-a fashion designed to put roughly equal amounts of work on each
-processor (load balancing).
-
-On a grid, the data can be stored at cell-centers, on a face/edge, or
-on the corners.  In BoxLib, data that is on an edge is termed `nodal'
-in that direction (see Figure~\ref{fig:dataloc}).  Data that is on the
-corners is nodal in all spatial directions.  In \maestro, the state
-data (density, enthalpy, velocity, $\ldots$) is generally
-cell-centered.  Fluxes are nodal in the direction they represent.
-A few quantities are nodal in all directions (e.g.\ $\phi$ used in
-the final velocity projection).
-
-To simplify the description of the underlying AMR grid, BoxLib
-provides a number of Fortran types.  We briefly summarize the major
-data types below.
-
-\subsection{\boxtype}
-
-A \boxtype\ is simply a rectangular domain in space.  Note that boxes
-do not hold the state data themselves.  A \boxtype\ has a {\tt lo} 
-and {\tt hi} index in each coordinate direction that gives the
-location of the lower-left and upper-right corner with respect to
-a global index space.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=4.0in]{\archfigpath/index_grid2}
-\caption[Single-level grid structure]
-{\label{fig:boxes} Three boxes that comprise a single level.  At this
-  resolution, the domain is 20$\times$18 zones.  Note that the
-  indexing in BoxLib starts with $0$.}
-\end{figure}
-
-The computational domain is divided into boxes.  The collection of
-boxes with the same resolution comprise a level.
-Figure~\ref{fig:boxes} shows three boxes in the same level of
-refinement.  The position of the boxes is with respect to the global
-index space at that level.  For example, box 1 in the figure has {\tt
-  lo} = (3,7) and {\tt hi} = (9,12).  Note that the global indexing
-is 0-based.
-
-The global index space covers the entire domain at a given resolution.
-For a simulation setup with {\tt n\_cellx = 32} and {\tt n\_celly =
-  32}, the coarsest level (level 1) has $32 \times 32$ zones, and the
-global index space will run from $0, \ldots, 31$ in each coordinate
-direction.  Level 2 will have a global index space running from $0,
-\ldots, 63$ in each coordinate direction (corresponding to $64 \times
-64$ zones if fully refined), and level 3 will have a global index
-space running from $0, \ldots, 127$ in each coordinate direction
-(corresponding to $128\times 128$ zones if fully refined).
-
-\subsubsection{Common Operations on a \boxtype}
-
-A \boxtype\ declared as:
-\begin{verbatim}
-  type(box) :: mybox
-\end{verbatim}
-%
-The upper and lower bounds of the box (in terms of the global
-index space) are found via:
-\begin{itemize}
-
-\item {\tt lo = lwb(mybox)} returns an array, {\tt lo(dm)}, with
-     the box lower bounds
-
-\item {\tt hi = upb(mybox)} returns an array, {\tt hi(dm)}, with
-     the box upper bounds
-
-\end{itemize}
-
-\subsection{\boxarray\ and \mlboxarray}
-
-A \boxarray\ is an array of boxes.  A \mlboxarray\ is a collection of
-\boxarray s at different levels of refinement.
-
-\subsection{\layout\ and \mllayout}
-
-A \layout\ is basically a \boxarray\ that knows information about other
-boxes, or box ``connectivity.''  It contains additional information
-that is used in filling ghost cells from other fine grids or from
-coarser grids.  This information is stored as long as the layout
-exists so that we don't have to recompute intersections every time we
-do some operation with two \multifab s that have that layout, for
-example.
-
-\subsection{\fab}
-
-A \fab\ is a ``Fortran Array Box''.  It contains the state data in a
-multidimensional array and several \boxtype-types to describe where in
-the global index-space it lives.  
-Note that all state data is stored in a four-dimensional array,
-{\tt (nx,ny,nz,nc)} in size, regardless of the dimensionality of the
-problem.  Here {\tt nc} is the number of components, for instance
-representing different fluid variables.  For 2D problems, {\tt nz=1}.
-
-A \fab\ would represent the data for a single box in the domain.
-In \maestro, we don't usually deal with \fab s alone, but rather
-we deal with \multifab s, described next.
-
-\subsection{\multifab}
-
-A \multifab\ is a collection of \fab s at the same level of
-refinement.  This is the primary data structure that \maestro\
-routine operate on.  A multilevel simulation stores the 
-data in an array of \multifab s, where the array index refers
-to the refinement level.
-
-\subsubsection{Working with \multifab s}
-
-To build a \multifab, we need to provide a \layout, the number of
-components to store in the \multifab\, and the number of ghostcells.  In
-\maestro\, the hierarchy of grids will be described by a single
-\mllayout.  A \multifab\ can be declared and built at any time in a
-simulation using the \mllayout, thereby allocating space at every
-grid location in the simulation.  The sequence to build a \multifab\
-appears as
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  type(multifab) :: mfab(nlevs)
-  ...
-  do n = 1, nlevs
-     call multifab_build(mfab(n), mla%la(n), nc, ng)
-  enddo
-\end{lstlisting}
-Here, {\tt nc} is the number of components and {\tt ng} is the number
-of ghostcells.  The \multifab\ is built one level at a time, using the
-\layout\ for that level taken from the \mllayout, {\tt mla}.
-
-A common operation on a \multifab\ is to initialize it to $0$
-everywhere.  This can be done (level-by-level) as
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-call setval(mfab(n), ZERO, all=.true.)
-\end{lstlisting}
-where {\tt ZERO} is the constant 0.0 from {\tt bl\_constants\_module}.
-
-The procedure for accessing the data in each grid managed by the \multifab\
-is shown in \S~\ref{sec:example}.  
-Subroutines to add, multiply, or divide two \multifab s exist, as do
-subroutines to copy from one \multifab\ to another---see {\tt boxlib/multifab.f90}
-for the full list of routines that work with \multifab s.
-
-When you are done working with a \multifab, its memory can be freed by
-calling {\tt multifab\_destroy} on the \multifab.
-
-\subsection{\bctower}
-
-A \bctower\ holds the information about what boundary conditions are
-in effect for each variable in a \maestro\ simulation.  \MarginPar{better description?}
-
-\section{\maestro\ Data Organization}
-
-The state of the star in \maestro\ is described by both a
-multidimensional state and the 1D base state.  The full
-multidimensional state is stored in \multifab s while the base state
-is simply stored in Fortran arrays.  Here we describe the
-major \maestro\ data-structures.
-
-\subsection{`{\tt s}' \multifab s (fluid state)}
-
-The fluid state (density, enthalpy, species, temperature, and tracer)
-are stored together in a cell-centered multi-component \multifab,
-typically named {\tt sold}, {\tt s1}, {\tt s2}, or {\tt snew}
-(depending on which time-level it represents).  The enthalpy is stored
-as $(\rho h)$, and the species are stored as partial-densities $(\rho
-X_k)$.  The tracer component is not used at present time, but can
-describe an arbitrary advected quantity.
-
-Individual state variables should be indexed using the integer keys
-provided by the {\tt variables} module (see \S
-\ref{sec:variables_module}).  For example, the integer {\tt rho\_comp}
-will always refer to the density component of the state.
-
-Note: the pressure is not carried as part of the `{\tt s}' \multifab s.
-
-\subsection{`{\tt u}' \multifab s (fluid velocity)}
-
-The fluid velocity at time-levels $n$ and $n+1$ is stored in
-a cell-centered multi-component \multifab, typically named
-{\tt uold} or {\tt unew}.  Here the {\tt dm}
-components correspond to each coordinate direction.
-
-\subsection{{\tt umac} (the MAC velocity)}
-
-In creating the advective fluxes, we need the time-centered velocity
-through the faces of the zone---the $x$-velocity on the $x$-edges, the
-$y$-velocity on the $y$-edges, etc.\ (see figure~\ref{fig:mac}).  This
-type of velocity discretization is termed the MAC velocity (after the
-``marker-and-cell'' method for free boundaries in incompressible
-flows \cite{harlowwelch:1965}).
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=2.5in]{\archfigpath/mac2}
-\hspace{0.1in}
-\begin{minipage}[b]{3.8in}
-\caption[The MAC grid]
-{\label{fig:mac} The MAC grid for the velocity.  
-Here the $x$-velocity is on the $x$-edges (shown as the 
-blue points) and the $y$-velocity is on the $y$-edges
-(shown as the red points).
-}\ \\
-\end{minipage}
-\end{figure}
-
-The MAC velocities are allocated at each level of refinement, {\tt n},
-by making a \multifab\ array where each of the {\tt dm} components is
-nodal in its respective direction:
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  type(multifab) :: umac(dm)
-
-  do comp=1,dm
-     call multifab_build_edge(umac(n,comp), mla%la(n),1,1,comp)
-  enddo
-\end{lstlisting}
-
-\subsection{Base State Arrays}
-
-The base state is defined by $\rho_0$, $p_0$, and $w_0$.  There is no
-base state composition.  Other arrays are defined as needed, such as
-$h_0$, the base state enthalpy.
-
-The base state arrays are 2-dimensional, with the first dimension
-giving the level in the AMR hierarchy and the second the radial index
-into the base state.  For spherical geometries, the base state only
-exists at a single level, so the first index will always be 1.  The
-radial index is 0-based, to be consistent with the indexing for the
-Cartesian state data.  For example, the base state density would be
-dimensioned: {\tt rho0(nlevs,0:nr\_fine-1)}.  Here, {\tt nlevs} is the
-number of levels of refinement and {\tt nr\_fine} is the number of
-cells in the radial direction at the finest level of refinement.
-
-For multilevel, plane-parallel geometry, all grids at the same height
-will have the same resolution so that the full state data is always
-aligned with the base state (see Figure~\ref{fig:base_state}).  Base
-state data on coarse grids that are covered by fine grids is not
-guaranteed to be valid.
-
-For spherical problems, the base state resolution, $\Delta r$, is
-generally picked to be finer than the Cartesian grid resolution,
-$\Delta x$, i.e.\ $\Delta r < \Delta x$.  The ratio is controlled
-by the parameter {\tt drdxfac}.
-
-\section{\maestro\ Helper Modules}
-
-A number of \maestro\ modules appear frequently throughout the source.
-Below, we describe some of the more common functionality of the most
-popular modules.
-
-\subsection{\tt average}
-
-The {\tt average} module provides a routine {\tt average} that takes
-a multilevel \multifab\ array and averages the full Cartesian data
-onto the 1D base state.
-
-\subsection{{\tt eos\_module}}
-
-The {\tt eos\_module} provides the interface to the equation of 
-state to connect the state variables thermodynamically.  It 
-gets the information about the fluid species from the {\tt network}
-module (for example, the atomic number, $Z$, and atomic weight, $A$,
-of the nuclei).
-
-Presently there are 2 equations of state that work with \maestro:
-\begin{itemize}
-\item {\tt extern/EOS/helmeos/} represents a general stellar equation 
-      of state, consisting of nuclei (as an ideal gas), radiation,
-      and electrons (with arbitrary degeneracy and degree of relativity).
-      This equation of state is that described in \cite{timmes_eos}.
-
-\item {\tt extern/EOS/gamma\_law\_general} assumes an ideal gas with a mixed 
-     composition and a constant ratio of specific heats, $\gamma$:
-      \begin{equation}
-      p = \rho e (\gamma - 1) = \frac{\rho k_B T}{\mu m_p} 
-      \end{equation}
-     where $k_B$ is Boltzmann's constant and $m_p$ is the mass of the
-     proton.
-     The mean molecular weight, $\mu$, is computed assuming 
-     electrically neutral atoms:
-     \begin{equation}
-     \mu = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}
-     \end{equation}
-     An option in the source code itself exists for treating the
-     species as fully-ionized, but there is no runtime-parameter to
-     make this switch.
-\end{itemize}
-
-The {\tt eos\_module} declares for convenience the variables that need
-appear in the {\tt eos} call argument list.  \maestro\ routines use
-these module variables in the EOS call to avoid having to declare 
-each quantity in each routine that calls the EOS.
-
-The first argument to the {\tt eos} call is an integer key that
-specifies which thermodynamic variables (in addition to the mass
-fractions) are used as input.  EOS input options are listed 
-in table~\ref{arch:table:eosinput}.
-
-   \begin{table}[h]
-   \caption{\label{arch:table:eosinput} EOS input flags}
-   \begin{center}
-   \begin{tabular}{lc}
-   \hline
-   key            & input quantities \\
-   \hline
-   {\tt eos\_input\_rt}       & $\rho$, $T$ \\
-   {\tt eos\_input\_rh}       & $\rho$, $h$ \\
-   {\tt eos\_input\_tp}       & $T$, $p$ \\
-   {\tt eos\_input\_rp}       & $\rho$, $p$ \\
-   {\tt eos\_input\_re}       & $\rho$, $e$ \\
-   {\tt eos\_input\_ps}       & $p$, $s$ \\
-   \hline
-   \end{tabular}
-   \end{center}
-   \end{table}
-
-\subsection{{\tt fill\_3d\_module}}
-
-The {\tt fill\_3d\_module} provides routines that map from the 1D
-base state to the full Cartesian 2- or 3D state.  Variations in the
-routines allow for cell-centered or edge-centered data on either the
-base state or full Cartesian state.
-
-\subsection{{\tt geometry}}
-
-\subsection{{\tt network}}
-
-The {\tt network} module defines the number species advected by the
-code ({\tt nspec}), their ordering, and gives their basic properties
-(like atomic number, $Z$, and atomic mass, $A$).  All \maestro\ problems
-require a {\tt network} module, even if there are no reactions
-modeled.  Many different reaction modules (containing different sets
-of isotopes) exist in {\tt extern/networks}.  The particular network
-used by a problem is defined in the problem's {\tt GNUmakefile}.
-
-To find the location of a particular species (for instance, ``carbon-12'')
-in the allowed range of {\tt 1:nspec}, you do the following query:
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  ic12 = network_species_index("carbon-12")
-\end{lstlisting}
-If the resulting index is {\tt -1}, then the requested species was not
-found.
-
-\subsection{{\tt probin\_module}}
-
-\label{sec:probin}
-
-{\tt probin\_module} provides access to the runtime parameters.
-The runtime parameters appear simply as module variables.  To get the 
-value of a parameter, one simply needs to `{\tt use probin\_module}'.
-The preferred method is to add the `{\tt only}' clause to the
-{\tt use} statement and explicitly list only those parameters that
-are used in the routine.  Defining new runtime parameters is
-described in \S~\ref{sec:def_runtime_param}.
-
-\subsection{{\tt variables}}
-
-\label{sec:variables_module}
-
-The {\tt variables} module provides integer keys to index the state
-multifabs and other arrays dealing with the scalar quantities.  The
-most commonly used keys are are list in table~\ref{arch:table:variables}.
-
-\begin{table}[h]
-\caption{\label{arch:table:variables} Common {\tt variables} module keys}
-\begin{center}
-\begin{tabular}{ll}
-\hline
-{\tt rho\_comp}  & density \\
-{\tt rhoh\_comp} & density $\times$ specific enthalpy, $(\rho h)$ \\
-{\tt spec\_comp} & first species partial density, $(\rho X_1)$ \\
-{\tt temp\_comp} & temperature \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-The species indices are contiguous in the state array, spanning {\tt
-  spec\_comp:spec\_comp-1+nspec}.  To find a particular species, a
-query can be made through the {\tt network} module, such as:
-\begin{verbatim}
-  ic12 = network_species_index("carbon-12")
-\end{verbatim}
-and then the \fab\ can be indexed using {\tt spec\_comp-1+ic12} to
-get ``carbon-12''.
-The {\tt variables} module also provides keys for the plotfile
-variables and boundary condition types.
-
-Other keys in the {\tt variables} modules are reserved for boundary
-conditions ({\tt foextrap\_comp} and {\tt hoextap\_comp}), the
-projection of the pressure ({\tt press\_comp}), or constructing
-the plotfile.
-
-\section{BoxLib Helper Modules}
-
-There are a large number of modules in {\tt boxlib/} that provide
-the core functionality for managing grids.  Here we describe
-the most popular such modules.
-
-\subsection{{\tt bl\_types}}
-
-\subsection{{\tt bl\_constants}}
-
-\subsection{{\tt parallel}}
-
-\section{\label{sec:example} Example: Accessing State and MAC Data}
-
-In \maestro, the state data is stored in a cell-centered multifab array
-(the array index refers to the AMR level) and the MAC velocities are
-stored in a 2D nodal multifab array (with indices referring to the AMR
-level and the velocity component).  Here we demonstrate a typical way
-to extract the state and MAC velocity data.
-
-All \maestro\ routines are contained in a module, to allow for compile-time
-argument checking.
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-module example_module
-
-contains
-\end{lstlisting}
-
-The main interface to our routine is called {\tt example}---this will
-take the \multifab s containing the data and then pass them to the
-work routines, {\tt example\_2d} or {\tt example\_3d}, depending on
-the dimensionality.  
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  subroutine example(mla,s,umac,dx,dt)
-
-    use bl_types
-    use multifab_module
-    use ml_layout_module
-    use variables, only: rho_comp
-\end{lstlisting}
-
-\noindent Here, the {\tt bl\_types} and {\tt multifab\_module} modules
-bring in the basic BoxLib data types. Specifically, here, {\tt
-  bl\_types} defines {\tt dp\_t} which is the {\tt kind} used for
-declaring double precision data, and {\tt multifab\_module} defines
-the \multifab\ data type.  The {\tt ml\_layout\_module} defines the
-datatype for a \mllayout---many routines will take an \mllayout
-to allow us to fill ghostcells.  The {\tt variables} module is a \maestro\
-module that provides integer keys for indexing the state arrays.  In
-this case the integer {\tt rho\_comp} refers to the location in the
-state array corresponding to density.
-
-Next we declare the subroutine arguments:
-
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-    type(ml_layout), intent(in   ) :: mla
-    type(multifab) , intent(inout) :: s(:)
-    type(multifab) , intent(inout) :: umac(:,:)
-    real(kind=dp_t), intent(in   ) :: dx(:,:),dt
-\end{lstlisting}
-Here, {\tt s(:)} is our \multifab\ array that holds the state data.
-with the array index in {\tt s} refers to the AMR level.  The MAC 
-velocities are held in the multifab {\tt umac}, with the array
-indices referring to the AMR level and the component.
-
-Local variable declarations come next:
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-    ! Local variables
-    real(kind=dp_t), pointer :: sp(:,:,:,:)
-    real(kind=dp_t), pointer :: ump(:,:,:,:), vmp(:,:,:,:), wmp(:,:,:,:)
-    integer :: i,n,dm,nlevs,ng_sp,ng_um
-    integer :: lo(mla%dim),hi(mla%dim)
-\end{lstlisting}
-
-\noindent Amongst the local variables we define here are a pointer,
-{\tt sp}, that will point to a single \fab\ from the
-\multifab\ {\tt s}, and a pointer for each component of the MAC
-velocity, {\tt ump}, {\tt vmp}, and {\tt wmp} (for a 2D run,
-we won't use {\tt wmp}).  We note that regardless of the dimensionality,
-these pointers are 4-dimensional: 3 spatial + 1 component.
-
-Next we get the dimensionality and number of levels
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-    dm = mla%dim
-    nlevs = mla%nlevel
-\end{lstlisting}
-
-Each multifab can have their own number of ghostcells, so we get
-these next:
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-    ng_sp = nghost(s(1))
-    ng_um = nghost(umac(1,1))
-\end{lstlisting}
-By convention, all levels in a given multifab have the same number of
-ghostcells, so we use level 1 in the {\tt nghost()} call.  We also use
-the same number of ghostcells for each component of the velocity, so
-we only need to consider the first component in the {\tt nghost()}
-call.  The ghostcells will be needed to access the data stored in the
-\fab s.
-
-To access the data, we loop over all the levels, and all the boxes in
-the given level.
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-    do n=1,nlevs
-       do i = 1, nboxes(s(n))
-\end{lstlisting}
-{\tt nboxes(s(n))} is simply the number of boxes in level {\tt n}.
-All data in all \multifab s will have the same box layout, so we only
-need to query the number of boxes from one \multifab.
-
-For a given \boxtype, we get the data and the bounds of the \boxtype.
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-          if ( multifab_remote(s(n), i) ) cycle
-          sp  => dataptr(s(n), i)
-          ump => dataptr(umac(n,1),i)
-          vmp => dataptr(umac(n,2),i)
-          lo =  lwb(get_box(s(n), i))
-          hi =  upb(get_box(s(n), i))
-\end{lstlisting}
-
-\noindent For parallel runs, there is no guarantee that the particular
-box is on the current processor.  The function {\tt multifab\_remote}
-returns {\tt .true.} if the box is off-processor, in which case, we 
-simply skip to the next box.
-
-The actual data array is accessed through the {\tt dataptr} function,
-which takes a \multifab\ (e.g.\ {\tt s(n)}) and the index of the
-\boxtype\ ({\tt i}) we want.  We see that the $x$ MAC velocity for the
-current box is stored in {\tt ump} and the $y$ MAC velocity is stored
-in {\tt vmp}.  We don't get the $z$ velocity data here, since that
-would not be available for a 2D run---we defer that until we test on
-the dimensionality below.
-
-Finally, the index bounds of the box (just the data, not the ghostcells) are 
-stored in the {\tt dm}-dimensional arrays {\tt lo} and {\tt hi}.  These indices
-refer to the current box, and hold for both the state, {\tt sp}, and the MAC
-velocity, {\tt ump} and {\tt vmp}.  However, since the MAC velocity is nodal
-in the component direction, the loops over the valid data will differ
-slight (as we see below).
-
-With the data extracted, we call a subroutine to operate on it.  We use
-different subroutines for the different dimensionalities (and many times
-have a separate routine for spherical geometries).
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-          select case (dm)
-          case (2)
-             call example_2d(sp(:,:,1,rho_comp),ng_sp, &
-                             ump(:,:,1,1),vmp(:,:,1,1),ng_um, &
-                             lo,hi,dx(n,:),dt)
-          case (3)
-             wmp => dataptr(umac(n,3),i)
-             call example_3d(sp(:,:,:,rho_comp),ng_sp, &
-                             ump(:,:,:,1),vmp(:,:,:,1),wmp(:,:,:,1),ng_um, &
-                             lo,hi,dx(n,:),dt)
-          end select
-       enddo    ! end loop over boxes
-
-    enddo    ! end loop over levels
-
-  end subroutine example
-\end{lstlisting}
-\noindent We call either the function
-{\tt example\_2d} for two-dimensional data or {\tt example\_3d}
-for three-dimensional data.  Note that in the two-dimensional
-case, we index the data as {\tt sp(:,:,1,rho\_comp)}.  Here a
-`{\tt 1}' is used as the `z'-coordinate spatial index, since this
-is a 2D problem, and the density component of the state is selected
-(using the integer key {\tt rho\_comp}).  The 3D version accesses
-the data as {\tt sp(:,:,:,rho\_comp)}---only the component regarding
-the variable is needed here.  Notice that we also pass through
-the number of ghostcells for each of the quantities.
-
-This routine will be supplimented with {\tt example\_2d} and {\tt
-example\_3d}, which actually operate on the data.  The form of 
-the 2D function is:
-
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  subroutine example_2d(density,ng_sp, &
-                        umac,vmac,ng_um, &
-                        lo,hi,dx,dt)
-
-    use bl_constants_module
-    use probin_module, only: prob_lo
-
-    integer        , intent(in) :: lo(:),hi(:), ng_sp, ng_um
-    real(kind=dp_t), intent(in) :: density(lo(1)-ng_sp:,lo(2)-ng_sp:)
-    real(kind=dp_t), intent(in) ::    umac(lo(1)-ng_um:,lo(2)-ng_um:)
-    real(kind=dp_t), intent(in) ::    vmac(lo(1)-ng_um:,lo(2)-ng_um:)
-
-    real(kind=dp_t), intent(in) :: dx(:),dt
-
-    integer         :: i, j
-    real(kind=dp_t) :: x, y
-    real(kind=dp_t) :: dens, u, v
-
-    do j = lo(2), hi(2)
-       y = prob_lo(2) + (dble(j) + HALF)*dx(2)
-
-       do i = lo(1), hi(1)
-          x = prob_lo(1) + (dble(i) + HALF)*dx(1)
-
-          dens = density(i,j)
-
-          ! compute cell-centered velocity
-          u = HALF*(umac(i,j) + umac(i+1,j))
-          v = HALF*(vmac(i,j) + vmac(i+1,j))
-
-          ! operate on the data
-          ! ...
-
-       enddo
-    enddo
-
-  end subroutine example_2d
-
-end module example_module
-\end{lstlisting}
-
-\noindent In this function, the bounds of the {\tt density} array take
-into account the {\tt ng\_sp} ghostcells and the index space of the
-current box.  Likewise, the MAC velocities refer to the {\tt ng\_um}
-ghostcells.  The {\tt j} and {\tt i} loops loop over all the valid
-zones.  Coordinate information is computed from {\tt dx} and {\tt
-  prob\_lo} which is the physical lower bound of the domain.  {\tt
-  bl\_constants\_module} declares useful double-precision constants,
-like {\tt HALF} (0.5).  Here, we see how to access the density for
-the current zone and compute the cell-centered velocities from the
-MAC velocities.  By convection, for a nodal array, the indices refer 
-to the {\em lower} interface in the nodal direction, so for {\tt umac},
-{\tt umac(i,j)} and {\tt umac(i+1,j)} are the $x$ MAC velocities
-on the lower and upper edge of the zone in the $x$-direction.
-
-The three-dimensional case is similar, with the {\tt density} array
-declared as 
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  density(lo(1)-ng_sp:,lo(2)-ng_sp:,lo(3)-ng_sp:)
-\end{lstlisting}
-and an additional loop over the `z' coordinate (from {\tt lo(3)} to
-{\tt hi(3)}).
-
-In this example, we looped over the valid zones.  If we wished to loop
-over the interfaces bounding the valid zones, in the $x$-direction,
-we would loop as
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  do j = lo(2), hi(2)
-     do i = lo(1), hi(1)+1
-        ! access umac(i,j)
-     enddo
-  enddo
-\end{lstlisting}
-
-\section{Filling Ghostcells}
-
-Ghostcells are filled through a variety of different routines, depending
-on the objective.
-
-\begin{itemize}
-
-\item {\tt multifab\_fill\_boundary} fills ghost cells for two
-  adjacent grids at the same level, which als includes periodic domain
-  boundary ghost cells.
-
-\item {\tt multifab\_physbc} fills ghostcells at the physical boundaries.
-
-\item {\tt multifab\_fill\_ghost\_cells} is used for multilevel
-  problems, and fills ghostcells in the finer grid (level {\tt n}) by
-  interpolating from data in the coarser grid (level {\tt n-1}).
-  This function, by default, will also call {\tt multifab\_fill\_boundary}
-  and {\tt multifab\_physbc} for both levels {\tt n} and {\tt n-1} (you 
-  can override this behavior for speed optimization purposes).
-  This call is usually preceded by a call to 
-  {\tt ml\_cc\_restriction\_c} which sets the level {\tt n-1} data to be
-  the average of the level {\tt n} data covering it.
-
-\end{itemize}
-
-\section{Boundary Conditions}
-
-\section{Multigrid}
-
-\maestro\ uses the multigrid solver to enforce the divergence constraint
-both on the half-time edge-centered advective velocities (the ``MAC
-projection'') and on the final cell-centered velocities (the ``HG
-projection'').  For the MAC projection, since the velocity data is
-edge-centered (the MAC grid), the projection is cell-centered.  For
-the HG projection, since the velocity data is cell-centered, the
-projection is node-centered.  \MarginPar{should draw a figure} The multigrid solver performs a number
-of V-cycles until the residual drops by 10-12 orders-of-magnitude.
-There are several options that affect how the multigrid solver
-behaves, which we describe below.
-
-\subsection{Special Bottom Solver}
-We have a special bottom solver type that greatly speeds up large problems
-and speeds up smaller problems to a lesser degree.
-In fact, you should use this bottom solver whenever possible, because I
-have yet to find a case that performs worse than our standard bottom solver.
-What this special solver essentially does is take the data from the coarsest 
-level of the original multigrid V-cycle and copy it onto a new grid structure 
-with the same number of total cells in each direction, but with a fewer number of larger 
-grids.  A new V-cycle begins from this point, 
-so we are essentially coarsening this ``new'' problem.
-Now, the coarsest level of the multigrid V-cycle in the ``new'' problem has 
-fewer cells and fewer grids as compared to the original coarsest level.
-
-To enable this solver, set {\tt hg\_bottom\_solver = 4} (for the nodal
-projections) and/or {\tt mg\_bottom\_solver = 4} (for the cell-centered
-projections) in your inputs file.
-
-To understand how this bottom solver works, the first thing you need to know
-is what the grid structure of the coarsest level of your multigrid V-cycle
-looks like (we need to put a description of this in the section above).
-Now, figure out the size of the box you would need if you
-wanted it to fit all the data on the coarsest level.  Now, figure out what
-the largest integer $n$ is so that you can evenly divide the length of this box
-by $2^n$ in every coordinate direction.  If $n < 2$, the program will abort
-since the grid structure is not suitable for this bottom solver.
-
-Now the code will set up a ``new'' problem, using the data at the coarsest level
-of the original problem as the initial data.  The grid structure for this new
-problem has the same number of cells as the coarsest level of the original problem,
-but the data is copied onto a grid structure where each grid has $2^n$ cells
-on each side.  The new V-cycle continues down to the new coarsest level, in
-which each grid has 2 cells on each side.  If you wish to impose a limit on
-the maximum value that $n$ can have, you can do so by setting 
-{\tt max\_mg\_bottom\_nlevs} equal to that value.
-
-{\bf Example 1:} A 3D problem with $384^3$ cells divided into $32^3$ grids, i.e., 
-there is a $12\times 12\times 12$ block of $32^3$ grids.  The coarsest level of the 
-multigrid V-cycle contains $12\times 12\times 12$ grids that have $2^3$ cells, so the 
-entire problem domain has $24^3$ cells.  We see that $n=3$, and create a new problem 
-domain with a $3\times 3\times 3$ block of $8^3$ grids.  The coarsest level of the 
-multigrid V-cycle for the ``new'' problem will be a $3\times 3\times 3$ block of 
-$2^3$ grids.
-
-{\bf Example 2:} A 2D problem with $96\times 384$ cells divided into $48^2$ grids, i.e., 
-there is a $2\times 8$ block of $48^2$ grids.  The coarsest level of the multigrid 
-V-cycle contains $2\times 8$ grids that have $3^2$ cells, so the entire problem 
-domain has $6\times 24$ cells.  We see that $n=1$, so the program aborts since this grid
-structure is not appropriate for the fancy bottom solver.
-
-\section{Multilevel and Refinement Criteria}
-
-\section{Particles}
-
-\label{arch:sec:particles}
-
-\maestro\ has support for Lagrangian particles that are passively
-advected with the velocity field.  These are useful for diagnostics
-and post-processing.  To use particles, particles must be seeded into
-the domain by writing a problem-specific {\tt init\_particles.f90}
-routine.  This routine is called at the start of the simulation.  The
-{\tt init\_particles} routines add particles at specific locations by
-calling the {\tt particle\_module}'s {\tt add} routine when a given
-criteria is met by the fluid state.
-
-When you run the code, particles are enabled by setting {\tt
-  use\_particles = T}.  At the end of each timestep the locations of
-all the particles are written out into a series of files called {\tt
-  timestamp\_NN}, where {\tt NN} is the CPU number on which the
-particle {\em currently} resides.  Particles are always kept on the
-processor containing the state data corresponding to their present
-location.  Several bits of associated data (density, temperature, and
-mass fractions) are stored along with the particle ID and position.
-
-Some simple python scripts allow for the plotting of the particle
-positions.  See \S~\ref{analysis:sec:particles} for details.

### 2011-07-23 docs/architecture/architecture.tex
architecture was getting too big -- split it into getting started and architecture

@@
 
-\section{Getting Started}
-
-In this section we give an overview of \maestro, including some of the
-standard problems, how to run the code, some basic runtime parameters,
-and how to look at the output.
-
-\subsection{`Standard' Test Problems}
-
-Different problems in \maestro\ are contained in individual
-sub-directories under {\tt MAESTRO/}.  The {\tt GNUmakefile}
-in each problem directory lists the components of {\tt MAESTRO}
-that are used to build the executable.  Full details of the
-{\tt GNUmakefile} can be found in \S~\ref{sec:adding_problems}.
-
-Some of the test problems available are:
-\begin{itemize}
-\item {\tt flame} \\[-3mm]
-
-The {\tt flame} problem models a planar laminar flame.  Initially cool
-fuel and hot ash are put in pressure equilibrium.  Thermal diffusion
-transfers heat to the fuel, driving a burning front.  This problem
-does not use the traditional \maestro\ elliptic constraint, but
-rather enforces $\nabla \cdot U = S$ (through the runtime parameter
-{\tt do\_smallscale = T}) as discussed in \cite{SNe}.
-
-\item {\tt spherical\_heat} \\[-3mm]
-
-{\tt spherical\_heat} maps a spherical star (a Chandrasekhar-mass white
-dwarf) onto the Cartesian domain and uses an external heat source to
-cause the star to expand.  This is the 3D analogue of the {\tt
-  test\_basestate} unit test (see \S~\ref{sec:unit_tests}).  This
-problem was discussed in \cite{multilevel}.
-
-\item {\tt test2} \\[-3mm]
-
-{\tt test2} places 3 hots spots in a plane-parallel atmosphere.
-Burning makes these buoyant bubbles which role up.  This problem was
-used in \cite{lowMach3} to compare with compressible solvers.
-
-\item {\tt test\_convect} \\[-3mm]
-
-{\tt test\_convect} drives convection through a plane-parallel
-atmosphere using an externally-specified heat source.  This problem
-was used to compare with compressible solvers in \cite{lowMach3}
-and to test the multilevel algorithm in \cite{multilevel}.
-
-\item {\tt test\_spherical} \\[-3mm]
-
-{\tt test\_spherical} sets up an isentropically stratified star
-and stirs it up with a random velocity field.  The low Mach number
-constraint is replaced with the anelastic constraint (through
-the {\tt beta\_type} runtime parameter).  Analytically, under
-these conditions, the density of the star should not change.
-This test problem was discussed in \cite{lowMach4}.
-
-\end{itemize}
-
-\subsection{Compiling and Running}
-
-To build the \maestro\ executable, simply type `{\tt make}' in the
-problem directory.  The executable will have a name like {\tt
-  main.Linux.Intel.exe}, where the specific parts of the name depend
-on the compilers and OS used.  \maestro\ can be built with support for
-parallelism (using MPI or MPI+OpenMP), or compiled as a serial code.
-The top portion of the {\tt GNUmakefile} can be edited to choose the
-build options.  Setting {\tt MPI := t} will build the code to use MPI
-for parallelization.  Setting {\tt OMP := t} will use OpenMP.  OpenMP
-is used together with MPI, with MPI distributing the grids across the
-processors and within a shared-memory node, OpenMP allows many cores
-to operate on the same grid.  A final build option determines whether
-to build the code with support for debugging.  Setting {\tt NDEBUG :=
-  t} disables debugging features.
-
-If the {\tt helmeos} equation of state is chosen for the problem, then
-the data table for the equation of state, {\tt
-  extern/EOS/helmeos/helm\_table.dat} needs to be copied to the
-directory in which the code is run.
-
-Each problem will have one or more inputs files (for example, {\tt
-  test2/inputs\_2d}) which override the default values for one or more
-runtime parameters.  The code is run simply as:
-\begin{verbatim}
-  ./main.Linux.Intel.exe inputs_2d
-\end{verbatim}
-for a serial executable, and, depending on your MPI environment, something like:
-\begin{verbatim}
-  mpiexec -n 16 ./main.Linux.Intel.mpi.exe inputs_2d
-\end{verbatim}
-for the parallel code.
-
-As the code runs, it will output both plotfiles and checkpoints as
-well as one of more text diagnostic files (like {\tt
-  maestro\_diag.out}) with integral or extrema information (like
-maximum Mach number) from each timestep.  By default, the plotfiles
-will be named {\tt plt}{\em nnnnn}, where the number {\em nnnnn} is
-the timestep number when the file was outputted.  Similarly, the
-checkpoints are named {\tt chk}{\em nnnnn}.  BoxLib plotfiles and
-checkpoints are actually directories, with the data stored in
-sub-directories grouped by refinement level.  Details of the simulation
-(build information, number of processors used, output date, output
-directory, runtime parameter values, ...) are stored in the {\tt
-  job\_info} file in each plotfile directory.
-
-A large number of runtime parameters affect the code's behavior.  Any
-runtime parameter can be set either in the inputs file, as {\tt parameter
-  = value}, or on the command line by adding {\tt --parameter value} after
-listing the inputs file on the command line.  A full list of runtime
-parameters is detailed in \S~\ref{sec:runtime_parameters}.  Below we
-outline the most common runtime parameters.
-
-\subsection{Controlling Timestepping and Output}
-
-Parameters that set the maximum time for the simulation to run
-include:
-\begin{itemize}
-\item {\tt stop\_time} is the maximum simulation time, in seconds,
-      to evolve the system for.
-
-\item {\tt max\_step} is the maximum number of steps to take.
-\end{itemize}
-
-\noindent Parameters affecting the size of the timestep include:
-\begin{itemize}
-\item {\tt cflfac} is a multiplicative factor ({\tt $\le 1$}) 
-      applied to the advective CFL timestep
-
-\item {\tt init\_shrink} is the factor ({\tt $\le 1$}) by which to reduce 
-      the initial timestep from the estimated first timestep.
-\end{itemize}
-
-\noindent Parameters affecting output and restart include:
-\begin{itemize}
-
-\item {\tt restart} tells \maestro\ to restart from a checkpoint.  The
-      value of this parameter should be the file number to restart from.
-      For example, to restart from the checkpoint file {\tt chk00010},
-      you would set {\tt restart = 10}.
-
-\item {\tt plot\_int} is the number of steps to take between
-  outputting a plotfile
-
-\item {\tt plot\_deltat} is the simulation time to evolve between
-  outputting a plotfile.  Note: to output only based on simulation
-  time, set {\tt plot\_int = -1}.
-
-\item {\tt check\_int} is the number of steps to take between
-  outputting a checkpoint.
-
-\end{itemize}
-
-\subsection{Defining the Grid and Boundary Conditions}
-
-Parameters that determine the spatial extent of the grid, 
-the types of boundaries, and the number of computational cells include:
-\begin{itemize}
-
-\item {\tt max\_levs } is the maximum number of grid levels in the AMR
-  hierarchy to use.  {\tt max\_levs = 1} indicates running with only a
-  single level spanning the whole domain.
-
-\item {\tt n\_cellx }, {\tt n\_celly }, {\tt n\_cellz } the size of
-  base level in terms of number of cells, in the $x$, $y$, and $z$
-  coordinate directions.
-
-\item {\tt max\_grid\_size } the maximum extend of a grid, in any
-  coordinate direction, as measured in terms of number of cells.
-
-  For multilevel problems, the parameter {\tt max\_grid\_size\_1}
-  controls the maximum extent on level 1 (the base grid), {\tt
-    max\_grid\_size\_2} controls the maximum extent on level 2, and
-  {\tt max\_grid\_size\_3} controls the maximum extent on levels 3 and
-  higher.
-
-\item {\tt prob\_lo\_x }, {\tt prob\_lo\_y }, {\tt prob\_lo\_z } is
-  the physical coordinate of the lower extent of the domain boundary
-  in the $x$, $y$, and $z$ coordinate directions.
-
-\item {\tt prob\_hi\_x }, {\tt prob\_hi\_y }, {\tt prob\_hi\_z } is
-  the physical coordinate of the upper extent of the domain boundary
-  in the $x$, $y$, and $z$ coordinate directions.
-
-\item {\tt bcx\_lo }, {\tt bcy\_lo }, {\tt bcz\_lo }, 
-      {\tt bcx\_hi }, {\tt bcy\_hi }, {\tt bcz\_hi } are the boundary
-   condition types at the lower (`{\tt lo}') and upper (`{\tt hi}')
-   domain boundaries in the $x$, $y$, and $z$ coordinate directions.
-   The different types are set via integer flags listed in table~\ref{arch:table:bcflags}.
-
-   \begin{table}[h]
-   \caption{\label{arch:table:bcflags} Boundary condition flags}  
-   \begin{center}
-   \begin{tabular}{ll}
-   \hline
-   BC type    & integer flag \\
-   \hline
-   periodic             & $-1$ \\
-   inlet (user-defined) & $11$ \\
-   outlet               & $12$ \\
-   symmetry             & $13$ \\
-   slip wall            & $14$ \\
-   no-slip wall         & $15$ \\
-   \hline
-   \end{tabular}
-   \end{center}
-   \end{table}
-
-\end{itemize}
-
-Note that grid cells must be square, i.e. $\Delta x = \Delta y = \Delta z$
-where $\Delta x$ on the base grid is computed as $({\tt prob\_hi\_x}
-- {\tt prob\_lo\_x})/{\tt n\_cellx}$.  For multilevel problems, the effective
-number of zones on the finest grid in the $x$ direction will be
-${\tt n\_cellx} \cdot 2^{({\tt max\_levels} -1)}$.
-
-\subsection{Working with the Output}
-
-Visualization and analysis are done with on the plotfiles.  A
-number of in-house and externally developed tools can work 
-with BoxLib-formatted plotfiles.
-
-\subsubsection{\tt Amrvis}
-
-{\tt Amrvis} is an easy-to-use visualization tool developed at LBL for
-2- and 3D datasets which can plot slices through 3D datasets as well
-as volume-renderings.  It can also very easily extract 1D lines
-through the dataset along any coordinate direction.  It is distributed
-separately from the \maestro\ distribution.
-
-\subsubsection{{\tt data\_processing} scripts}
-
-Several useful analysis scripts (written in Fortran 90) can be found
-in {\tt fParallel/data\_processing}.  The {\tt GNUmakefile} there
-needs to be edited to indicate which of the tools to build.  For
-example, to extract the density along a line from the center of a
-plotfile, {\tt plt00200}, in the $y$-direction:
-
-\begin{verbatim}
-fextract.Linux.gfortran.exe -d 2 -v "density" -p plt00200
-\end{verbatim}
-
-These routines are described in \S~\ref{sec:analysis}.
-
-\subsubsection{VisIt}
-
-VisIt is a powerful, DOE-supported visualization tool for 2- and 3D
-datasets.  It can do contouring, volume rendering, streamlines, ...\
-directly from BoxLib plotfiles (to open a plotfile in VisIt, point to
-the {\tt Header} file inside the plotfile directory).  Details on
-visit can be found at {\tt https://wci.llnl.gov/codes/visit/home.html}\,.
-
-\subsubsection{Diagnostic Files}
-
-By default, \maestro\ outputs global diagnostics each timestep into a file
-called {\tt maestro\_diag.out}.  This includes the maximum Mach number,
-peak temperature, and peak nuclear energy generation rate.  Individual problems
-can provide their own {\tt diag.f90} file to produce custom diagnostic
-output.
-
-\subsection{Unit Tests}
+\section{Unit Tests}
 

### 2011-07-24 docs/flowchart/flowchart.tex
fixed some small errors
make figures bigger
some restructuring

@@
-Here we outline the algorithm currently implemented in the code.  The latest
-published reference for MAESTRO is the multilevel paper \cite{multilevel}.  In
-this description, we make frequent reference to paper~I \cite{lowMach}, paper~II \cite{lowMach2},
-paper~III \cite{lowMach3}, and paper~IV \cite{lowMach4}.  We also make reference the BDRZW paper \cite{SNe}
-which describes the small-scale astrophysical combustion algrothm.
-
-\section{Changes Between Paper 3 and Paper 4}
-\begin{enumerate}
-\item We defined the mapping of data between a 1D radial array and the 3D Cartesian
-grid for spherical problems (which we improve upon in the multilevel paper).
-\item We update $T$ after the call to {\bf React State}.
-\item We have created a {\tt burning\_cutoff\_density}, where the burning does
-not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
-\item Use corner coupling in advection.
-\item We have an option, controlled by {\tt use\_tfromp}, to update temperature 
-using $T=T(\rho,X_k,p_0)$ rather than $T=T(\rho,h,X_k)$.  The former completely 
-decouples enthalpy from our system.  For spherical problems, we use 
-{\tt use\_tfromp = TRUE}, for planar problems, we use {\tt use\_tfromp = FALSE}.
-\item For spherical problems, we have changed the discretization of 
-$\Ubt\cdot\nabla p_0$ in the enthalpy update to 
-$\nabla\cdot(\Ubt p_0) - p_0\nabla\cdot\Ubt$.
-\item In paper III we discretized the enthalpy evolution equation in
-terms of $T$.  Since then we have discovered that 
-discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
-leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
-This is more like paper II.
-\item We have turned off the evolution of $h$ above the atmosphere and instead
-compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
-\end{enumerate}
-\section{Changes Between Paper 4 and the Multilevel Paper}
-See the multilevel paper for the latest.
-\section{Changes Between the Multilevel Paper and Paper 5}
-\begin{enumerate}
-\item Added rotation.
-\end{enumerate}
-\section{Changes Between Paper 5 and the XRB Paper}
-\begin{enumerate}
-\item We have added thermal diffusion, controlled by {\tt use\_thermal\_diffusion},
-{\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
-\item We added the volume discrepancy term to the velocity constraint equation,
-controlled by the input parameter, {\tt dpdt\_factor}.
-\item For certain problems, we need to set {\tt do\_eos\_h\_above\_cutoff = FALSE}
-to prevent large, unphysical velocities from appearing near the edge of the star.
-\end{enumerate}
+Here we outline the algorithm currently implemented in the code.  The
+latest published reference for MAESTRO is the multilevel
+paper \cite{multilevel}.  In this description, we make frequent
+reference to paper~I \cite{lowMach}, paper~II \cite{lowMach2},
+paper~III \cite{lowMach3}, and paper~IV \cite{lowMach4}.
 
 %-----------------------------------------------------------------------------
-% Future Considerations
+% Equations Summary
 %-----------------------------------------------------------------------------
 
-\section{Future Considerations}
+\section{Summary of the \maestro\ Equation Set}
 
-\begin{itemize}
-\item Should we use a predictor-corrector for updating the full-state density?
-Specifically, after calling {\bf Correct Base}, should we do a full-state density 
-advance and {\bf Correct Base} using the more accurate estimate of $\rho_0^{n+1}$?
-\item We are still exploring the effects of {\tt use\_tfromp = FALSE} for spherical
-problems.  We would eventually like to run in this mode, but $T=T(\rho,X_k,p_0)$ 
-and $T=T(\rho,h,X_k)$ drift away from each other more than we would like.  Our
-attempts at incorporating a {\tt dpdt\_factor} for spherical problems have not 
-been successful.
-\end{itemize}
+Here we summarize the equations solved by \maestro.  We refer the reader
+to papers~I through IV and the multilevel paper for the derivation
+and motivation of the equation set.
 
@@
 \section{Time Advancement Algorithm}\label{Sec:Time Advancement Algorithm}
-Here is the current description of the algorithm, based from the description
-in the multilevel paper.
-The initialization has also been included with more detail than was given in paper III.
+Here is the current description of the algorithm, based from the
+description in the multilevel paper.  The initialization has also been
+included with more detail than was given in paper III.  The main
+driver for a single step of the algorithm is {\tt advance.f90}---refer
+to that code to see the sequence of functions called to implement each
+ste.
+
+\subsection{Notation}
+
+\subsection{Single Step}
 
@@
   and $X_k^{\nph,\star,\pred}$.
-  We convert the perturbational density full state density using
+  We convert the perturbational density to full state density using
 \begin{equation}
@@
 \centering
-\includegraphics[scale=0.6]{\flowfigpath/flowchart}
+\includegraphics[scale=0.75]{\flowfigpath/flowchart}
 \caption[Graphical flowchart of MAESTRO]
@@
 \centering
-\includegraphics[scale=0.6]{\flowfigpath/flowchart_4_8}
+\includegraphics[scale=0.75]{\flowfigpath/flowchart_4_8}
 \caption[Graphical flowchart of the density and enthalpy update steps]
@@
   The thermodynamic state variables and base state variables are
-  indicated in each step.  Red text indicates that q\ uantity was
+  indicated in each step.  Red text indicates that quantity was
   updated during that step.  Note, for {\bf Step 4}, the updated
@@
   \end{enumerate}
+
+%-----------------------------------------------------------------------------
+% changes
+%-----------------------------------------------------------------------------
+
+\section{Changes from Earlier Implementations}
+\subsection{Changes Between Paper 3 and Paper 4}
+\begin{enumerate}
+\item We defined the mapping of data between a 1D radial array and the 3D Cartesian
+grid for spherical problems (which we improve upon in the multilevel paper).
+\item We update $T$ after the call to {\bf React State}.
+\item We have created a {\tt burning\_cutoff\_density}, where the burning does
+not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
+\item Use corner coupling in advection.
+\item We have an option, controlled by {\tt use\_tfromp}, to update temperature 
+using $T=T(\rho,X_k,p_0)$ rather than $T=T(\rho,h,X_k)$.  The former completely 
+decouples enthalpy from our system.  For spherical problems, we use 
+{\tt use\_tfromp = TRUE}, for planar problems, we use {\tt use\_tfromp = FALSE}.
+\item For spherical problems, we have changed the discretization of 
+$\Ubt\cdot\nabla p_0$ in the enthalpy update to 
+$\nabla\cdot(\Ubt p_0) - p_0\nabla\cdot\Ubt$.
+\item In paper III we discretized the enthalpy evolution equation in
+terms of $T$.  Since then we have discovered that 
+discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
+leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
+This is more like paper II.
+\item We have turned off the evolution of $h$ above the atmosphere and instead
+compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
+\end{enumerate}
+
+\subsection{Changes Between Paper 4 and the Multilevel Paper}
+See the multilevel paper for the latest.
+
+\subsection{Changes Between the Multilevel Paper and Paper 5}
+\begin{enumerate}
+\item Added rotation.
+\end{enumerate}
+
+\subsection{Changes Between Paper 5 and the XRB Paper}
+\begin{enumerate}
+\item We have added thermal diffusion, controlled by {\tt use\_thermal\_diffusion},
+{\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
+\item We added the volume discrepancy term to the velocity constraint equation,
+controlled by the input parameter, {\tt dpdt\_factor}.
+\item For certain problems, we need to set {\tt do\_eos\_h\_above\_cutoff = FALSE}
+to prevent large, unphysical velocities from appearing near the edge of the star.
+\end{enumerate}
+
+%-----------------------------------------------------------------------------
+% Future Considerations
+%-----------------------------------------------------------------------------
+
+\section{Future Considerations}
+
+\begin{itemize}
+\item Should we use a predictor-corrector for updating the full-state density?
+Specifically, after calling {\bf Correct Base}, should we do a full-state density 
+advance and {\bf Correct Base} using the more accurate estimate of $\rho_0^{n+1}$?
+\item We are still exploring the effects of {\tt use\_tfromp = FALSE} for spherical
+problems.  We would eventually like to run in this mode, but $T=T(\rho,X_k,p_0)$ 
+and $T=T(\rho,h,X_k)$ drift away from each other more than we would like.  Our
+attempts at incorporating a {\tt dpdt\_factor} for spherical problems have not 
+been successful.
+\end{itemize}

### 2011-09-02 docs/architecture/architecture.tex
update directory structure -- still need to do the graphic

@@
 such restriction (again, see Figure~\ref{fig:base_state}).
+The \maestro\ grids are managed by the \boxlib\ library, which is
+distributed separately.
 
@@
 
-All the files needed to build \maestro\ are contained in the {\tt
-  fParallel} directory structure.  Problems piece together various
-\maestro\ directories, choosing a reaction network, equation of state,
-and conductivity routine to build an executable.  Briefly, the
-sub-directories are:
+Building \maestro\ requires both the \maestro-specific source
+files (distributed in the {\tt MAESTRO/} directory), and the
+\boxlib\ library (distributed in the {\tt BoxLib/} directory.
+\boxlib\ provides both a C++ and a Fortran framework.  \maestro\
+only uses the Fortran framework.
+
+Problems piece together various \maestro\ directories, choosing a
+reaction network, equation of state, and conductivity routine to build
+an executable.  Briefly, the \maestro\ sub-directories are:
 \begin{itemize}
-\item {\tt boxlib/}
-
- BoxLib is a library for describing meshes consisting of a union
- of boxes.  The BoxLib modules define the basic datatypes used
- in \maestro.  BoxLib also provides the routines that handle the
- parallelization and I/O.
-
 \item {\tt data\_processing/}
@@
   Important sub-directories include:
+
   \begin{itemize}
@@
 
- A simple Fortran module for reading in 1D initial model files.  This
- is used by the initialization routines to get the initial model data.
+    A simple Fortran module for reading in 1D initial model files.
+    This is used by the initialization routines to get the initial
+    model data.
 
@@
     Various reaction networks for \maestro\ problems.  In addition to
- providing a routine to evolve the nuclear species due to reactions,
- the networks also define the species that are advected by the code.
+    providing a routine to evolve the nuclear species due to
+    reactions, the networks also define the species that are advected
+    by the code.
 
@@
 
- An integration package for ODEs.  At the moment, this is used 
- for integrating various reaction networks.
+    An integration package for ODEs.  At the moment, this is used for
+    integrating various reaction networks.
 
@@
 
- The main \maestro\ algorithm directory.  The individual problem directories
- lie under here, and all of the main \maestro\ source lives in the {\tt Source/}
- sub-directory.
+  The main \maestro\ algorithm directory.  The individual problem
+  directories lie under here, and all of the main \maestro\ source
+  lives in the {\tt Source/} sub-directory.
 
@@
 
-   Several sub-directories containing routines to generate
-   1D initial conditions in hydrostatic equilibrium for 
-   mapping onto the \maestro\ grid.  See \S~\ref{sec:initial_models}
-   for details.
+    Several sub-directories containing routines to generate 1D initial
+    conditions in hydrostatic equilibrium for mapping onto the
+    \maestro\ grid.  See \S~\ref{sec:initial_models} for details.
 
@@
 
-   The main \maestro\ source.  Here you will find the driver routine, the
-   advection routines, etc.  All \maestro\ problems will compile this
-   source.
-
- \item problem directories: {\tt test2}, {\tt
-   test\_spherical}, $\ldots$
+    The main \maestro\ source.  Here you will find the driver routine,
+    the advection routines, etc.  All \maestro\ problems will compile
+    this source.
+
+  \item problem directories: {\tt test2}, {\tt test\_spherical},
+    $\ldots$
 
@@
 
-   Any file that you place in a sub-directory here takes 
-   precedence over a file of the same name in {\tt MAESTRO/}.
-   This allows problems to have custom versions of the main
-   \maestro\ routines (e.g.\ initial conditions via {\tt initdata.f90}.
-   See \S~\ref{sec:makefile} for details on the build system).
+    Any file that you place in a sub-directory here takes precedence
+    over a file of the same name in {\tt MAESTRO/}.  This allows
+    problems to have custom versions of the main \maestro\ routines
+    (e.g.\ initial conditions via {\tt initdata.f90}.  See
+    \S~\ref{sec:makefile} for details on the build system).
 
@@
 
-\item {\tt mg/}
-
-  The multigrid solver, with support for both cell-centered and
-  node-centered data.
-
-\item {\tt mk/}
-
-  The generic Makefiles that store the compilation flags for
-  various platforms.  Platform/compiler-specific options are stored
-  in the {\tt comps/} sub-directory.
-
-\item {\tt scripts/}
+\end{itemize}
+
+The \boxlib\ directories that \maestro\ uses are:
+\begin{itemize}
+\item {Src/}
+
+  The main \boxlib\ source directory.  In \maestro, we only use the
+  Fortran source files.  The core directories are:
+
+  \begin{itemize}
+
+  \item {\tt F\_BaseLib/}
+
+    The Fortran \boxlib\ files.  This is a library for describing
+    meshes consisting of a union of boxes.  The \boxlib\ modules
+    define the basic datatypes used in \maestro.  \boxlib\ also
+    provides the routines that handle the parallelization and I/O.
+
+  \item {\tt LinearSolvers/}
+
+    The \boxlib\ linear solvers---these are used to solve elliptic
+    problems in the \maestro\ algorithm.
+
+    \begin{itemize}
+
+    \item {\tt F\_MG}
+
+      The Fortran multigrid solver, with support for both
+      cell-centered and node-centered data.
+
+    \end{itemize}
+
+  \end{itemize}
+
+\item {\tt Tools/}
+
+  Various tools used for building \boxlib\ applications.  Here we use:
+
+  \begin{itemize}
+
+  \item {F\_mk/}
+
+    The generic Makefiles that store the compilation flags for various
+    platforms.  Platform/compiler-specific options are stored in the
+    {\tt comps/} sub-directory.
+
+  \item {\tt F\_scripts/}
 
@@
     maintaining \maestro.
+
+  \end{itemize}
 

### 2011-09-07 docs/architecture/architecture.tex
update directory structure figures
move figure
update GNUmakefile text

@@
 \centering
-\includegraphics[width=5.0in]{\archfigpath/maestro_directory2}
+\includegraphics[scale=0.65]{\archfigpath/maestro_directory2}
 \caption[\maestro\ directory structure]
-{The basic \maestro\ directory structure}
+{The basic \maestro\ directory structure.}
 \end{figure}
@@
 \end{itemize}
+
+\begin{figure}[t]
+\centering
+\includegraphics[scale=0.65]{\archfigpath/boxlib_directory2}
+\caption[\boxlib\ directory structure]
+{The basic \boxlib\ directory structure.  The directories used
+by \maestro\ are indicated in red.}
+\end{figure}
 
@@
 
-The next two lines define where the top of the source tree is
-include the basic Makefile information (like compiler
-flags).
+The next line defines where the top of the source tree is located.
 \begin{lstlisting}[language={[gnu]make},mathescape=false]
   FPARALLEL := ../..
-
-  include $(FPARALLEL)/mk/GMakedefs.mak
-\end{lstlisting}
-%$
+\end{lstlisting}
 

### 2011-09-14 docs/FAQ/faq.tex
add note about BOXLIB_HOME directory

@@
      xlf95} for MAESTRO will produce memory leaks.
+
+\item {\em The code doesn't compile, but complains right away that there
+   is ``{\tt No rule to make target `fabio\_c.c', needed by `t/Linux.gfortran.mpi/c.depends'}''}
+
+   The environment variable {\tt BOXLIB\_HOME} needs to be the full path
+   to the {\tt BoxLib} directory.  You cannot use `{\tt $\sim$}' as a shortcut
+   for your home directory.
 

### 2011-10-14 docs/pert_notes/pert.tex
add note about what happens when use_tfromp = T and we try different predicting T to the edges

@@
 
-\noindent Note: the various {\tt enthalpy\_pred\_type}s have only been
+\noindent Note that the various {\tt enthalpy\_pred\_type}s have only been
 extensively tested with {\tt species\_pred\_type = 1}.
+
+Finally, when \maestro\ is run with {\tt use\_tfromp = T}, the
+temperature is derived from the density, basestate pressure ($p_0$),
+and $X_k$.  When run with reactions or external heating, {\tt
+  react\_state} updates the temperature after the reaction/heating
+term is computed.  In {\tt use\_tfromp = T} mode, the temperature will
+not see the heat release, since the enthalpy does not feed in.  Only
+after the hydro update does the temperature gain the effects of the
+heat release due to the adjustment of the density (which in turn sees
+it through the velocity field and $S$).  As a result, the {\tt
+  enthalpy\_pred\_type}s that predict temperature to the interface
+({\tt predict\_T\_then\_rhoprime} and {\tt predict\_T\_then\_h}) will
+not work.  \maestro\ will abort if the code is run with this
+combination of parameters.
 

### 2011-11-22 docs/architecture/architecture.tex
tweak to boxlib description

@@
 
+\maestro's gridding is handled by the BoxLib library, which
+contains the most fundamental objects used to construct parallel
+block-structured AMR applications---different
+regions of the domain can have different spatial resolutions.
+At each level of refinement, the region covered by that level is divided
+into grids, or boxes.  The entire computational domain is covered by
+the coarsest (base) level of refinement, often called level $\ell=0$, either by one
+grid or divided into many grids.
+Higher levels of refinement have cells that are finer by a ``refinement ratio''
+(typically 2).  The grids are properly nested in the sense that the union 
+of grids at level $\ell+1$ is contained in the union of grids at level $\ell$.
+Furthermore, the containment is strict in the sense that, except at physical 
+boundaries, the level $\ell$ grids are large enough to guarantee that there is
+a border at least $n_{\rm buffer}$ level $\ell$ cells wide surrounding each level
+$\ell +1$ grid (grids at all levels are allowed to extend to the physical
+boundaries so the proper nesting is not strict there).  
+For parallel computations, the boxes are spread across processors, in
+a fashion designed to put roughly equal amounts of work on each
+processor (load balancing).
+\MarginPar{ghost cells?}
+
 \begin{figure}[t]
@@
 \end{figure}
-
-\maestro's gridding is handled by the BoxLib library.  The underlying
-idea in BoxLib is to allow for adaptive mesh refinement (AMR)---different
-regions of the domain can have different spatial resolutions.  The
-domain is broken into boxes.  At the coarsest (base) level of
-refinement, the entire computational domain is covered, either by one
-box or broken across many boxes.  Higher levels of refinement have
-finer zones (typically $2\times$).  Only a portion of the domain may
-be covered by the higher levels of refinement.
-
-\MarginPar{Nesting? ghost cells?}
-For parallel computations, the boxes are spread across processors, in
-a fashion designed to put roughly equal amounts of work on each
-processor (load balancing).
-
 On a grid, the data can be stored at cell-centers, on a face/edge, or

### 2011-12-02 docs/visualization/visualization.tex
add particle count and processor number
add python_plotfile info

@@
 {\tt x\_vel}   & $\ut$  & -- \\
-\hline
+%\hline
 {\tt y\_vel}   & $\vt$  & -- \\
-\hline
+%\hline
 {\tt z\_vel}   & $\wt$  & 3-d runs only \\
-\hline
+%\hline
 {\tt density}  & $\rho$ & -- \\
-\hline
+%\hline
 {\tt rhoh}   & $(\rho h)$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
-\hline
+%\hline
 {\tt h}      & $(\rho h)/\rho$   & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
-\hline
+%\hline
 {\tt X(*)}   & $(\rho X_k)/\rho$ & {\tt plot\_spec} \\
-\hline
+%\hline
 {\tt tracer} & tracers           & {\tt plot\_trac} \\
-\hline
+%\hline
 {\tt w0\_x}  & $w_0 \er \cdot \ex$ & {\tt plot\_base} \\ 
-\hline
+%\hline
 {\tt w0\_y}  & $w_0 \er \cdot \ey$ & {\tt plot\_base} \\ 
-\hline
+%\hline
 {\tt w0\_z}  & $w_0 \er \cdot \ez$ & {\tt plot\_base} \\ 
-\hline
+%\hline
 {\tt divw0}  & $\nabla \cdot w_0$  & {\tt plot\_base} \\ 
-\hline
+%\hline
 {\tt rho0}   & $\rho_0$            & {\tt plot\_base} \\ 
-\hline
+%\hline
 {\tt rhoh0}  & $(\rho h)_0$         & {\tt plot\_base} \\ 
-\hline
+%\hline
 {\tt h0}     & $(\rho h)_0/\rho_0$  & {\tt plot\_base} \\ 
-\hline
+%\hline
 {\tt p0}     & $p_0$                & {\tt plot\_base} \\ 
-\hline
+%\hline
 {\tt radial\_velocity}  & $\Ubt \cdot \er + w_0$ & {\tt spherical == 1} \\
-\hline
+%\hline
 {\tt circum\_velocity}  & $| \Ubt - (\Ubt \cdot \er) \er |$ &  {\tt spherical == 1}  \\
-\hline
+%\hline
 {\tt magvel}              & $| \Ubt + w_0 \er |$  & -- \\
-\hline
+%\hline
 {\tt momentum}            & $\rho | \Ubt + w_0 \er |$  & -- \\
-\hline
+%\hline
 {\tt vort}                & $| \nabla \times \Ubt |$   & -- \\
-\hline
+%\hline
 {\tt S}                   & $S$   & -- \\
-\hline
+%\hline
 {\tt rhopert}             & $\rho - \rho_0$  & -- \\
-\hline
+%\hline
 {\tt rhohpert}          & $(\rho h) - (\rho h)_0$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
-\hline
+%\hline
 {\tt tfromp}              & $T(\rho, p_0, X_k)$  & -- \\
-\hline
+%\hline
 {\tt tfromh}            & $T(\rho, h, X_k)$    & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
-\hline
+%\hline
 {\tt deltaT}            & $[T(\rho, h, X_k) - T(\rho, p_0, X_k)]/T(\rho, h, X_k)$  & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
-\hline
+%\hline
 {\tt deltap}            & $|p(\rho,h,X_k) - p_0|/p_0$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
-\hline
+%\hline
 {\tt tpert}               & $T(\rho,h,X_k) - \overline{T}$ if {\tt use\_tfromp = F}; $T(\rho,p_0,X_k) - \overline{T}$ otherwise & -- \\
-\hline
+%\hline
 {\tt Machnumber}          & $| \Ubt + w_0 \er | / c(\rho,p_0,X_k)$ & -- \\
-\hline
+%\hline
 {\tt soundspeed}        & $c(\rho,p_0,X_k)$ & {\tt plot\_cs} \\
-\hline
+%\hline
 {\tt deltagamma}          & $\Gamma_1(\rho,p_0,X_k) - \gammabar$ & -- \\
-\hline
+%\hline
 {\tt entropy}             & $s(\rho,p_0,X_k)$ & -- \\
-\hline
+%\hline
 {\tt entropypert}         & $[s(\rho,p_0,X_k) - \overline{s}]/\overline{s}$ & -- \\
-\hline
+%\hline
 {\tt sponge} or {\tt sponge\_fdamp} & $1/(1 + \Delta t \kappa f_\mathrm{damp})$ by default; $f_\mathrm{damp}$ if {\tt plot\_sponge\_fdamp = T} & -- \\
-\hline
+%\hline
 {\tt pi}                 & $\pi$ & -- \\
-\hline
+%\hline
 {\tt gpi\_x}           & $\nabla \pi \cdot \ex$ & {\tt plot\_gpi} \\
-\hline 
+%\hline 
 {\tt gpi\_y}           & $\nabla \pi \cdot \ey$ & {\tt plot\_gpi} \\
-\hline
+%\hline
 {\tt gpi\_z}           & $\nabla \pi \cdot \ez$ & {\tt plot\_gpi} \\
-\hline
+%\hline
 {\tt pioverp0}         & $\pi / p_0$ & {\tt plot\_base} \\
-\hline
+%\hline
 {\tt p0pluspi}         & $p_0 + \pi$ & {\tt plot\_base} \\
-\hline
+%\hline
 {\tt omegadot(*)}      & $(\rho X_k)/\rho$ & {\tt plot\_omegadot} \\
-\hline
+%\hline
 {\tt enucdot}          & $(\rho \Hnuc)/\rho$ & {\tt plot\_Hnuc} \\
-\hline
+%\hline
 {\tt Hext}             & $(\rho \Hext)/\rho$ & {\tt plot\_Hext} \\
-\hline
+%\hline
 {\tt eta\_rho}         & $\etarho$ & {\tt plot\_eta} \\
-\hline
+%\hline
 {\tt thermal}          & $\nabla \cdot \kappa \nabla T$ & {\tt use\_thermal\_diffusion} \\
-\hline
+%\hline
 {\tt conductivity}     & $\kappa(\rho, T, X_k)$ & {\tt use\_thermal\_diffusion} \\
-\hline
+%\hline
 {\tt ad\_excess}       & $\nabla - \nabla_\mathrm{ad}$ & {\tt plot\_ad\_excess} \\
-\hline
+%\hline
+{\tt particle\_count}  & number of particles in a cell & {\tt use\_particles} \\
+%\hline
+{\tt processor\_number} & processor number containing the cell's data & {\tt plot\_processors} \\
 \end{longtable}
@@
 \section{Visualizing with \visit}
+
+\section{Visualizing with {\tt data\_processing} scripts}
+
+{\tt data\_processing/python\_plotfile} provide a simple commandline tool for doing 
+visualizations of 2-d BoxLib plotfiles.  The main driver
+is written in python and uses a set of Fortran routines, compiled with
+{\tt f2py} to interface with the plotfile data.  To use the routine,
+you will need to have {\tt matplotlib} and {\tt f2py} installed.  On
+a machine running Fedora linux, you can install these packages via
+\begin{verbatim}
+yum install python-matplotlib f2py
+\end{verbatim}
+
+The library required by the python routines can be built by typing
+'{\tt make}' in that directory.  If successful, you should find
+a library {\tt fsnapshot.so}.  A simple plot can be made via:
+\begin{verbatim}
+plotsinglevar.py --log -o test.png plt00000/ tfromp
+\end{verbatim}
+This will make a plot of ``tfromp'' from the plotfile {\tt plt00000} with log scaling,
+and store the output in {\tt test.png}.
+If you don't do `{\tt -o}', then a default output filename consisting of the
+plotfile name + component will be used.
 

### 2011-12-10 docs/FAQ/faq.tex
add abort_maestro

@@
    allocated arrays are automatically deallocated when a subroutine
-   ends.  Fortran 90 does not do this.  Most MAESTRO routines rely 
+   ends.  Fortran 90 does not do this.  Most \maestro\ routines rely 
    on this Fortran 95 feature.
@@
    On IBM machines, for instance, using {\tt xlf90} instead of {\tt
-     xlf95} for MAESTRO will produce memory leaks.
+     xlf95} for \maestro\ will produce memory leaks.
 
@@
 
-\item {\em  MAESTRO crashes because the multigrid algorithm fails to
+\item {\em  \maestro\ crashes because the multigrid algorithm fails to
     converge---how do I get around this?}
@@
 
+\item{\em How can I force \maestro\ to abort?}
+
+  In the output directory, do `{\tt touch .abort\_maestro}'.  This 
+  will cause the code to write out a final checkpoint file, free up
+  any allocated memory, and gracefully exit.  Be sure to remove the 
+  {\tt .abort\_maestro} file before restarting the code in the 
+  same directory.
+
 \end{enumerate}
@@
 
-\item {\em How can I force the code to output a plotfile at the next
-   step?}
+\item {\em How can I force the code to output a plotfile / checkpoint
+  file at the next step?}
 
@@
    restore the code to its normal plotfile behavior.
+
+   Similarly, creating the file {\tt .dump\_checkpoint} will force the
+   output of a checkpoint file.
 
@@
  No.  This is a design decision.  There is no support for axisymmetric
- coordinates in MAESTRO.  Spherical problems must be run in 3-d.
+ coordinates in \maestro.  Spherical problems must be run in 3-d.
 

### 2011-12-13 docs/compilers/compilers.tex
add BDS / Pathscale note
use maestrosymbols.tex and \mastro
update the compiler folklore

@@
 
-Here, we list notes on problems with specific compilers and Maestro.
+Here, we list notes on problems with specific compilers and \maestro.
 
@@
 
-Probably the best compilers for Maestro.  Versions 4.5 and 4.6 seem to work
+Probably the best compilers for \maestro.  Versions 4.5 and 4.6 seem to work
 really well and the debugging features are very reliable.
@@
 
-(on Kraken) The bounds checking routines do not like the way some of
+The general feeling with these compilers is that they produce slower
+code than the PathScale compilers.
+
+\ \\
+Machine-specific notes:
+\begin{itemize}
+
+\item (on Kraken) The bounds checking routines do not like the way some of
 the arrays are allocated and flag them as errors.  For instance, in
@@
 
-(on hopper) There are problems with code compiled with {\tt -fast}
+\item (on hopper) There are problems with code compiled with {\tt -fast}
 instead of just {\tt -O}.  gcc/gfortran seem to work better and 
 are only slighly slower.
+\end{itemize}
 
@@
 
-These are the prefered compilers on jaguar.
+These are the prefered compilers on jaguar, however, the PathScale
+compilers are no longer actively supported on the Crays and therefore
+should be considered depricated.
+
+\ \\
+Machine specific notes:
+\begin{itemize}
+
+\item  To run with BDS, change {\tt -Ofast} to {\tt -O2} in {\tt Linux\_pathscale.mak}
+
+\item (on jaguar) with OpenMP, the Pathscale compiler warns about
+  variables that are shared, but doesn't always give useful line
+  numbers.  In debug mode, add ``-fullwarn'' to the compilation to get
+  the compiler output to tell you where the problems are.
+
+\end{itemize}
+
+\subsection{Cray}
+
+So far, no luck with these compilers (versions 7.4.3 and 8.0.0.135).

### 2011-12-20 docs/pert_notes/pert.tex
update the description of species_pred_type = predict_rhoprime_and_rhoX to be consistent with the recent change.

@@
            \frac{\partial \phi}{\partial x} 
-    \underbrace{- \frac{\Delta t}{2} v \frac{\partial \phi}{\partial y}}_{\mathrm{``transverse~term''}} + \frac{\Delta t}{2} f
+    \underbrace{- \frac{\Delta t}{2} v \frac{\partial \phi}{\partial y}}_{\text{``transverse~term''}} + \frac{\Delta t}{2} f
 \end{eqnarray}
@@
            \frac{\partial \phi}{\partial x} 
-    \underbrace{- \frac{\Delta t}{2} \phi \frac{\partial u}{\partial x} 
-                - \frac{\Delta t}{2} \frac{\partial (\phi v)}{\partial y}}_{\mathrm{``transverse~term''}} + \frac{\Delta t}{2} f
+    \underbrace{- \frac{\Delta t}{2} \phi \frac{\partial u}{\partial x} }_{\text{``non-advective~term''}}
+                \underbrace{- \frac{\Delta t}{2} \frac{\partial (\phi v)}{\partial y}}_{\text{``transverse~term''}} + \frac{\Delta t}{2} f
 \end{eqnarray}
-Here the ``transverse term'' includes the non-advective portion of the $x$-flux (for the $x$-prediction).
-Both of the underbraced terms are accounted for in {\tt make\_edge\_scal} automatically when
-we call it with {\tt is\_conservative = .true.}.
+Here the ``transverse term'' is now in conservative form, and an additional
+term, the non-advective portion of the
+$x$-flux (for the $x$-prediction) appears.  Both of the underbraced terms are
+accounted for in {\tt make\_edge\_scal} automatically when we call it
+with {\tt is\_conservative = .true.}.
 
@@
 Here we wish to construct $(\rho X_k)_\mathrm{edge}$ by predicting
-$(\rho X_k)$ to the edges as a single quantity.  We start with
+$(\rho X_k)$ to the edges as a single quantity.  We recall
 Eq.~\ref{species equation}:
@@
 \frac{\partial(\rho X_k)}{\partial t} =
-  -\Ub\cdot\nabla(\rho X_k) \, \underbrace{-\rho X_k \nabla\cdot\Ub}_{``\rho X_k ~\text{force''}} + \rho \omegadot_k \, . \label{species equation labeled}
+  -\nabla \cdot (\rho \Ub X_k) + \rho \omegadot_k \, . \nonumber
 \end{equation}
 The edge state is created by calling {\tt make\_edge\_scal} in {\tt
-  density\_advance} with the underbraced term in (\ref{species
-  equation labeled}) as the forcing.  This advective force is computed
-by {\tt modify\_scal\_force} with {\tt fullform = .true.}.  We do not
-consider the $\omegadot_k$ term in the forcing when Strang-splitting.
+  density\_advance} with {\tt is\_conservative = .true.}.
+We do not consider the $\rho \omegadot_k$ term in the forcing when
+Strang-splitting.
 
@@
 -\nabla\cdot\left\{\left[\left(\rho X_k \right)_\mathrm{edge} \right](\Ubt+w_0\eb_r)\right\}.
-\end{equation}
+\end{equation} \\
 

### 2011-12-22 docs/compilers/compilers.tex
add information regarding the IBM compilers

@@
 
-\subsection{Intel}
+%-----------------------------------------------------------------------------
+\subsection{Cray}
 
-Overall, these tend to be flaky.
+So far, no luck with these compilers (versions 7.4.3 and 8.0.0.135).
 
+%-----------------------------------------------------------------------------
 \subsection{gfortran}
@@
 
+%-----------------------------------------------------------------------------
+\subsection{IBM (xl* compilers)}
+
+Recent versions of these compilers seem to work ok.  Some notes:
+\begin{itemize}
+\item You need to use {\tt xlf95}, not {\tt xlf90} to compile, since
+  MAESTRO relies on the Fortran 95 feature that arrays are
+  automatically deallocated when you leave a subroutine.
+
+\item Many (all?) IBM machines have a different endian-ness than
+  Intel-like chips.  \boxlib\ plotfiles store information used to
+  dechipher the endian-ness of the data.  To do this properly on
+  the IBM machines, you need to set the preprocessor flag {\tt BL\_AIX},
+  which is interpretted by {\tt fabio\_c.c}.  This can be done for the
+  IBM compilers by adding {\tt -Wp,-DBL\_AIX} to the {\tt CFLAGS} variable.
+
+\end{itemize}
+
+%-----------------------------------------------------------------------------
+\subsection{Intel}
+
+Overall, these tend to be flaky.
+
+%-----------------------------------------------------------------------------
+\subsection{PathScale}
+
+These are the prefered compilers on jaguar, however, the PathScale
+compilers are no longer actively supported on the Crays and therefore
+should be considered depricated.
+
+\ \\
+Machine specific notes:
+\begin{itemize}
+
+\item  To run with BDS, change {\tt -Ofast} to {\tt -O2} in {\tt Linux\_pathscale.mak}
+
+\item (on jaguar) with OpenMP, the Pathscale compiler warns about
+  variables that are shared, but doesn't always give useful line
+  numbers.  In debug mode, add ``-fullwarn'' to the compilation to get
+  the compiler output to tell you where the problems are.
+
+\end{itemize}
+
+%-----------------------------------------------------------------------------
 \subsection{PGI}
@@
 \end{itemize}
-
-\subsection{PathScale}
-
-These are the prefered compilers on jaguar, however, the PathScale
-compilers are no longer actively supported on the Crays and therefore
-should be considered depricated.
-
-\ \\
-Machine specific notes:
-\begin{itemize}
-
-\item  To run with BDS, change {\tt -Ofast} to {\tt -O2} in {\tt Linux\_pathscale.mak}
-
-\item (on jaguar) with OpenMP, the Pathscale compiler warns about
-  variables that are shared, but doesn't always give useful line
-  numbers.  In debug mode, add ``-fullwarn'' to the compilation to get
-  the compiler output to tell you where the problems are.
-
-\end{itemize}
-
-\subsection{Cray}
-
-So far, no luck with these compilers (versions 7.4.3 and 8.0.0.135).

### 2012-01-07 docs/runtime_parameters/runtime_parameters.tex
update + make the table landscape to make things fit better
add shading
add categories to the runtime parameters for documentation purposes. In the _parameters file, the headings define the category for the parameters that follow.  The rp.py script now determines the headers and writes separate tables for each category.

@@
 
+\begin{landscape}
+
 {\small
@@
 \begin{center}
-\begin{longtable}{|l|p{3.25in}|l|}
-\caption[runtime parameters]{runtime parameters.} \label{table:runtime} \\
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ EOS
+ parameters.} \label{table:runtime} \\
 %
@@
 
-\verb=  amr_buf_width  = &   the number of buffer zones surrounding a cell tagged for refinement. note that this needs to be >= regrid\_int  &  -1 \\
-\verb=  anelastic_cutoff  = &   The density below which we modify the constraint to look like the anelastic constraint, instead of the low Mach constraint.
-This prevents velocities from getting out of hand at the edge of the star.
-Refer to Section \ref{Sec:Anelastic Cutoff}.
-&  3.d6 \\
-\verb=  barrier_timers  = &   Call a parallel\_barrier before each of the simple timers we have coded in advance.f90.  Might cause a very slight overall performance hit.  &  .false. \\
-\verb=  base_cutoff_density  = &   The density below which we keep the initial model constant. Refer to Section \ref{Sec:Base Cutoff Density}  &  3.d6 \\
-\verb=  bcx_hi  = &   $+x$ boundary condition (valid values are listed in {\tt  boxlib/bc.f90})  &  SLIP\_WALL \\
-\verb=  bcx_lo  = &   $-x$ boundary condition  &  SLIP\_WALL \\
-\verb=  bcy_hi  = &   $+y$ boundary condition  &  SLIP\_WALL \\
-\verb=  bcy_lo  = &   $-y$ boundary condition  &  SLIP\_WALL \\
-\verb=  bcz_hi  = &   $+z$ boundary condition  &  SLIP\_WALL \\
-\verb=  bcz_lo  = &   $-z$ boundary condition  &  SLIP\_WALL \\
-\verb=  bds_type  = &   0 = use ppm instead for multi-d integrator \newline 1 = bilinear  &  0 \\
-\verb=  beta_type  = &   what type of coefficient to use inside the velocity divergence constraint.
-\newline {\tt beta\_type} = 1 uses $\beta_0$; \newline {\tt beta\_type} = 2 uses $\rho_0$ (anelastic); \newline {\tt beta\_type} = 3 uses 1 (small-scale combustion).
-&  1 \\
-\verb=  blocking_factor  = &   grids must be an integer multiple of this number, if possible  &  8 \\
-\verb=  buoyancy_cutoff_factor  = &   The multiplicative factor (over base\_cutoff\_density) below which we do zero out the buoyancy term in the momentum equation.  &  5.0 \\
-\verb=  burner_threshold_cutoff  = &   Mass fraction cutoff for burner\_threshold\_species  used in burner threshold  &  1.d-10 \\
-\verb=  burner_threshold_species  = &   Name of the species to be used in burner threshold  &  "" \\
-\verb=  cflfac  = &   CFL factor to use in the computation of the advection timestep constraint  &  0.5d0 \\
-\verb=  cg_verbose  = &   Verbosity of bottom solver  &  0 \\
-\verb=  change_max_grid_size_1  = &   Change the max grid size on the base level on restart  &  .false. \\
-\verb=  check_base_name  = &   prefix to use in checkpoint file names  &  "chk" \\
-\verb=  chk_int  = &   Number of timesteps between writing a checkpoint file  &  0 \\
-\verb=  co_latitude  = &   latitude, in radians, for problems with rotation where the domain is only a subset of a full star.  &  ZERO \\
-\verb=  conductivity_constant  = &   the constant conductivity (only applies if the conductivity/constant module is used)  &  1.0d0 \\
-\verb=  diag_buf_size  = &   number of timesteps to buffer diagnostic output information before writing (note: not implemented for all problems)  &  10 \\
-\verb=  dm_in  = &   dimensionality (valid values are 2 or 3)  &  2 \\
-\verb=  do_alltoallv  = &    &  .false. \\
-\verb=  do_burning  = &   turn on (.true.) or off (.false.) burning  &  .true. \\
-\verb=  do_eos_h_above_cutoff  = &   After the advective enthalpy update, recompute the enthalpy if we are below the base cutoff density.  &  .true. \\
-\verb=  do_heating  = &   use analytic heating  &  .false. \\
+\rowcolor{tableShade}
+\verb=  small_dens  = &    &  1.d-5 \\
+\verb=  small_temp  = &    &  5.d6 \\
+\rowcolor{tableShade}
+\verb=  use_eos_coulomb  = &   use the Coulomb corrections between ions (if supported by the EOS)  &  .true. \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ SDC
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  sdc_iters  = &   how many SDC iteractions to take (requires the code be compiled with {\tt SDC := t}  &  4 \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ algorithm initialization
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
 \verb=  do_initial_projection  = &   Do the initial projection.  &  .true. \\
-\verb=  do_planar_invsq_grav  = &   are we doing 1/r$^2$ gravity for plane-parallel  &  .false. \\
-\verb=  do_smallscale  = &   force $\rho_0 = (\rho h)_0 = 0$, {\tt evolve\_base\_state = F} and {\tt beta\_type} = 3  &  .false. \\
-\verb=  do_sponge  = &   Use sponging.  &  .false. \\
-\verb=  dpdt_factor  = &   factor in front of the volume discrepancy term (0.0 = off)  &  0.d0 \\
-\verb=  drdxfac  = &   ratio of radial base state zones to Cartesian full state zones for spherical geometry  &  1 \\
-\verb=  drive_initial_convection  = &   freeze the temperature used in the reaction network to the initial value.
-This is useful for developing an initial convective field to carry away the energy, while preventing the reactions from going nonlinear.
-&  .false.
-\\
-\verb=  dump_grid_file  = &   dump out a file, named grids.out, containing the number of grids a each level.  A new line is added each time regrid is called.  &  .false. \\
-\verb=  enthalpy_pred_type  = &   predict\_rhohprime = 1; \newline predict\_h = 2; \newline predict\_T\_then\_rhohprime = 3; \newline predict\_T\_then\_h = 4; \newline predict\_hprime = 5; \newline predict\_Tprime\_then\_h = 6.
-&  predict\_rhohprime \\
-\verb=  evolve_base_state  = &   turn on (.true.) or off (.false.) basestate evolution  &  .true. \\
-\verb=  fix_base_state  = &   if true, don't call average to reset the base state at all, even during initialization  &  .false. \\
-\verb=  fixed_dt  = &   Fix the time step.  If -1.0, then use the standard time step.  &  -1.0d0 \\
-\verb=  grav_const  = &   the gravitational acceleration (cm~s$^{-2}$) for plane-parallel geometry  &  -1.5d10 \\
-\verb=  hg_bottom_solver  = &   valid values are $\le$ 0  &  -1 \\
-\verb=  hg_dense_stencil  = &   In hgproject, in 2D, use a 9 point Laplacian (.true.) or 5-point Laplacian (.false.).
-In 3D, use a 27 point Laplacian (.true.) or 7-point Laplacian (.false.).
-&  .true.
-\\
 \verb=  init_divu_iter  = &   Number of initial divu iterations.  &  4 \\
+\rowcolor{tableShade}
 \verb=  init_iter  = &   Number of initial pressure iterations.  &  4 \\
-\verb=  init_shrink  = &   the multiplicative factor ($\le 1$) to reduce the initial timestep as computed by the various timestep estimators  &  1.d0 \\
-\verb=  job_name  = &   job name printed in output  &  "" \\
-\verb=  lUsingNFiles  = &   If .true., use nOutFiles processors to write checkpoint and plotfiles.
-Fortran has the unfortunate feature of each processor only being able to write out 1-2GB each without crashing.
-&  .true.
-\\
-\verb=  mach_max_abort  = &   maximum mach number before the code aborts  &  -1.d0 \\
-\verb=  max_dt  = &   This is the maximum dt that is allowed  &  1.0d33 \\
-\verb=  max_dt_growth  = &   The maximum scale factor that the time step is allowed to grow by per time step.  &  1.1d0 \\
-\verb=  max_grid_size  = &   The largest grid size that will be created using make\_new\_grids.  &  64 \\
-\verb=  max_grid_size_1  = &   The largest grid size that will be created using make\_new\_grids for the coarsest level.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
-\verb=  max_grid_size_2  = &   The largest grid size that will be created using make\_new\_grids for level 2 (the first refined level).  Defaults to {\tt max\_grid\_size}.  &  -1 \\
-\verb=  max_grid_size_3  = &   The largest grid size that will be created using make\_new\_grids for level 3 (the second refined level) and beyond.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
-\verb=  max_levs  = &   Total number of levels.  1 = single level.  &  1 \\
-\verb=  max_mg_bottom_nlevels  = &   if mg\_bottom\_solver == 4, then how many mg levels can the bottom solver mgt object have  &  1000 \\
-\verb=  max_step  = &   Maximum number of steps in the simulation.  &  1 \\
-\verb=  mg_bottom_solver  = &   valid values are $\le$ 0  &  -1 \\
-\verb=  mg_verbose  = &   Verbsoity of the multigrid solver, but not the bottom solver.  &  0 \\
-\verb=  min_eff  = &   parameter for cluster algorithm for making new grids in adaptive problems  &  0.9d0 \\
-\verb=  minwidth  = &   The minimum size on a side for a grid created using make\_new\_grids.  &  8 \\
-\verb=  model_file  = &   input model file  &  "model.hse" \\
-\verb=  nOutFiles  = &   If lUsingNFiles = .true., use this many processors to write checkpoint and plotfiles.
-Fortran has the unfortunate feature of each processor only being able to write out 1-2GB each without crashing.
-&  64 \\
-\verb=  n_cellx  = &   Number of cells for the base level in the x-direction  &  -1 \\
-\verb=  n_celly  = &   Number of cells for the base level in the y-direction  &  -1 \\
-\verb=  n_cellz  = &   Number of cells for the base level in the z-direction  &  -1 \\
-\verb=  nuclear_dt_fac  = &   If $T_{max}^n > T_{max}^{n-1}$ set the new dt =   min[dt, dt*{\tt nuclear\_dt\_fac}*( $T_{max}^{n-1}$ / $(T_{max}^n-T_{max}^{n-1})$ ) ] not checkpoint-compatible yet since it wouldn't be backwards compatible  &  -1.0d0 \\
-\verb=  octant  = &   set octant = T if you just want to model an octant of a sphere (note: only takes effect for spherical geometry)  &  .false. \\
-\verb=  perturb_model  = &   Turn on a perturbation in the initial data.  Problem specific.  &  .false. \\
-\verb=  planar_invsq_mass  = &   the point mass for planar 1/r$^2$ gravity  &  0.d0 \\
-\verb=  plot_Hext  = &   plot external heating (Hext) in plotfile  &  .false. \\
-\verb=  plot_Hnuc  = &   plot nuclear energy generation rate (Hnuc) in plotfile  &  .true. \\
-\verb=  plot_ad_excess  = &   plot the adiabatic excess  &  .false. \\
-\verb=  plot_base  = &   plot w0\_x, w0\_y, w0\_z, divw0, rho0, rhoh0, h0, and p0 in plotfile  &  .false. \\
-\verb=  plot_base_name  = &   prefix to use in plotfile file names  &  "plt" \\
-\verb=  plot_cs  = &   plot soundspeed  &  .false. \\
-\verb=  plot_deltat  = &   rather than use a plot interval, plot a file after the solution has advanced past plot\_deltat in time  &  -1.d0 \\
-\verb=  plot_eta  = &   plot $\etarho$ in plotfile  &  .false. \\
-\verb=  plot_gpi  = &   plot pi and grad(pi)  &  .true. \\
-\verb=  plot_h_with_use_tfromp  = &   Turn on storing of enthalpy-based quantities in the plotfile when we are running with {\tt use\_tfromp}  &  .true. \\
-\verb=  plot_int  = &   plot interval  &  0 \\
-\verb=  plot_omegadot  = &   plot omegadot in plotfile  &  .true. \\
-\verb=  plot_spec  = &   plot species and omegadot in plotfile  &  .true. \\
-\verb=  plot_sponge_fdamp  = &   plot fdamp rather than sponge assumes sponge has the form 1/(1+dt*{\tt sponge\_kappa}*fdamp)  &  .false. \\
-\verb=  plot_trac  = &   plot tracers in plotfile  &  .false. \\
-\verb=  ppm_type  = &   0 = no ppm \newline 1 = 1985 ppm \newline 2 = Hybrid Sekora/Colella McCorquodale Colella 2009/2010 ppm  &  1 \\
-\verb=  prob_hi_x  = &   physical coordinates of hi-x corner of problem domain  &  1.d0 \\
-\verb=  prob_hi_y  = &   physical coordinates of hi-y corner of problem domain  &  1.d0 \\
-\verb=  prob_hi_z  = &   physical coordinates of hi-z corner of problem domain  &  1.d0 \\
-\verb=  prob_lo_x  = &   physical coordinates of lo-x corner of problem domain  &  ZERO \\
-\verb=  prob_lo_y  = &   physical coordinates of lo-y corner of problem domain  &  ZERO \\
-\verb=  prob_lo_z  = &   physical coordinates of lo-z corner of problem domain  &  ZERO \\
-\verb=  ref_ratio  = &   Refinement ratio for multilevel problems  &  2 \\
-\verb=  regrid_int  = &   How often we regrid.  &  -1 \\
 \verb=  restart  = &   which file to restart from.  -1 means do not restart  &  -1 \\
+\rowcolor{tableShade}
 \verb=  restart_into_finer  = &   restart and add a level of refinement  &  .false. \\
-\verb=  restart_with_vel_field  = &   restart the simulation using a result from a {\tt drive\_initial\_convection} = T run note that this uses the restart variable to specify which file to restart from.
-After reading in the velocity information from the restart file, the time and timestep number are zeroed.
-&  .false.
-\\
-\verb=  rotation_radius  = &   radius used for computing centrifugal term in rotation problems  &  1.0d6 \\
-\verb=  rotational_frequency  = &   rotational frequency used for computing centrifugal term in rotation problems.  &  ZERO \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ base state mapping
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
 \verb=  s0_interp_type  = &   The interpolation for filling a cell-centered multifab from a 1D bin-centered array.
@@
 &  1 \\
-\verb=  sdc_iters  = &   how many SDC iteractions to take (requires the code be compiled with {\tt SDC := t}  &  4 \\
-\verb=  single_prec_plotfiles  = &    &  .false. \\
-\verb=  slope_order  = &   order of slopes in Godunov algorithm.  Options are 0, 2, or 4.  &  4 \\
-\verb=  small_dens  = &    &  1.d-5 \\
-\verb=  small_temp  = &    &  5.d6 \\
-\verb=  species_pred_type  = &   Which quantities do we predict to the edges for computing the ($\rho X$) edge states?
-\newline {\tt species\_pred\_type} = 1 means predict $\rho^\prime$ and $X$ separately.
-\newline {\tt species\_pred\_type} = 2 means predict the full ($\rho X$) itself.
-\newline {\tt species\_pred\_type} = 3 means predict $\rho$ and $X$ separately.
-&  1 \\
-\verb=  spherical_in  = &   set to 1 if you are doing a spherical problem  &  0 \\
-\verb=  sponge_center_density  = &   Center of the inner sponge.  &  3.d6 \\
-\verb=  sponge_kappa  = &   Parameter for sponge.  Problem dependent.  &  10.d0 \\
-\verb=  sponge_start_factor  = &   The sponge begins at sponge\_center\_density * sponge\_start\_factor.  &  3.333d0 \\
-\verb=  stop_time  = &   simulation stop time  &  -1.d0 \\
-\verb=  temp_diffusion_formulation  = &   How to compute the explicit thermal diffusion term.  \newline 1 = in terms of $T$; \newline 2 = in terms of $\rho,p_0,X$.  &  2 \\
-\verb=  test_set  = &   Fixed grid file.  &  '' \\
-\verb=  the_copy_cache_max  = &    Number of boxassoc layouts we keep in memory to avoid having to recompute the boxassoc, which is computationally expensive.  &  50 \\
-\verb=  the_knapsack_verbosity  = &   Verbosity of the knapscak processor-to-grid algorithm.  &  .false. \\
-\verb=  the_layout_verbosity  = &    &  0 \\
-\verb=  the_sfc_threshold  = &   When assigning processors for grids, this determines whether we use the sfc algorithm or knapsack algorithm.
-If the total number of grids divided by the number of processors is greater than this number, use sfc.
-&  5 \\
-\verb=  thermal_diffusion_type  = &   In the thermal diffusion solver, 1 = Crank-Nicholson; 2 = Backward Euler.  &  1 \\
-\verb=  use_delta_gamma1_term  = &   turns on second order correction to delta gamma1 term  &  .false. \\
-\verb=  use_divu_firstdt  = &   Use the divu constraint when computing the first time step.  &  .false. \\
-\verb=  use_eos_coulomb  = &   use the Coulomb corrections between ions (if supported by the EOS)  &  .true. \\
-\verb=  use_etarho  = &   turn on the etarho term as described in flow chart  &  .true. \\
-\verb=  use_hypre  = &   use the hypre library  &  .false. \\
-\verb=  use_linear_grav_in_beta  = &   how to represent gravity in the beta0 integration: .true. = piecewise linear .false. = piecewise constant  &  .false. \\
-\verb=  use_particles  = &   call the particle initialization, advection, etc. routines.  &  .false. \\
-\verb=  use_soundspeed_firstdt  = &   Use the soundspeed constraint when computing the first time step.  &  .false. \\
-\verb=  use_tfromp  = &   When updating temperature, use $T=T(\rho,p_0,X) $ rather than $T=T(\rho,h,X)$.  &  .false. \\
-\verb=  use_thermal_diffusion  = &   Use thermal diffusion.  &  .false. \\
-\verb=  verbose  = &   General verbosity  &  0 \\
+\rowcolor{tableShade}
 \verb=  w0_interp_type  = &   The interpolation for filling a cell-centered multifab from a 1D edge-centered array.
@@
 
+{\small
+
+\renewcommand{\arraystretch}{1.5}
 %
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ burning
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  burner_threshold_cutoff  = &   Mass fraction cutoff for burner\_threshold\_species  used in burner threshold  &  1.d-10 \\
+\verb=  burner_threshold_species  = &   Name of the species to be used in burner threshold  &  "" \\
+\rowcolor{tableShade}
+\verb=  do_burning  = &   turn on (.true.) or off (.false.) burning  &  .true. \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ general MAESTRO
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  barrier_timers  = &   Call a parallel\_barrier before each of the simple timers we have coded in advance.f90.  Might cause a very slight overall performance hit.  &  .false. \\
+\verb=  do_alltoallv  = &   use the MPI alltoallv call  &  .false. \\
+\rowcolor{tableShade}
+\verb=  job_name  = &   job name printed in output  &  "" \\
+\verb=  the_knapsack_verbosity  = &   Verbosity of the knapscak processor-to-grid algorithm.  &  .false. \\
+\rowcolor{tableShade}
+\verb=  verbose  = &   General verbosity  &  0 \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ grid
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  amr_buf_width  = &   the number of buffer zones surrounding a cell tagged for refinement. note that this needs to be >= regrid\_int  &  -1 \\
+\verb=  bcx_hi  = &   $+x$ boundary condition (valid values are listed in {\tt  boxlib/bc.f90})  &  SLIP\_WALL \\
+\rowcolor{tableShade}
+\verb=  bcx_lo  = &   $-x$ boundary condition  &  SLIP\_WALL \\
+\verb=  bcy_hi  = &   $+y$ boundary condition  &  SLIP\_WALL \\
+\rowcolor{tableShade}
+\verb=  bcy_lo  = &   $-y$ boundary condition  &  SLIP\_WALL \\
+\verb=  bcz_hi  = &   $+z$ boundary condition  &  SLIP\_WALL \\
+\rowcolor{tableShade}
+\verb=  bcz_lo  = &   $-z$ boundary condition  &  SLIP\_WALL \\
+\verb=  blocking_factor  = &   grids must be an integer multiple of this number, if possible  &  8 \\
+\rowcolor{tableShade}
+\verb=  change_max_grid_size_1  = &   Change the max grid size on the base level on restart  &  .false. \\
+\verb=  dm_in  = &   dimensionality (valid values are 2 or 3)  &  2 \\
+\rowcolor{tableShade}
+\verb=  drdxfac  = &   ratio of radial base state zones to Cartesian full state zones for spherical geometry  &  1 \\
+\verb=  dump_grid_file  = &   dump out a file, named grids.out, containing the number of grids a each level. A new line is added each time regrid is called.  &  .false. \\
+\rowcolor{tableShade}
+\verb=  max_grid_size  = &   The largest grid size that will be created using make\_new\_grids.  &  64 \\
+\verb=  max_grid_size_1  = &   The largest grid size that will be created using make\_new\_grids for the coarsest level.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
+\rowcolor{tableShade}
+\verb=  max_grid_size_2  = &   The largest grid size that will be created using make\_new\_grids for level 2 (the first refined level).  Defaults to {\tt max\_grid\_size}.  &  -1 \\
+\verb=  max_grid_size_3  = &   The largest grid size that will be created using make\_new\_grids for level 3 (the second refined level) and beyond.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
+\rowcolor{tableShade}
+\verb=  max_levs  = &   Total number of levels.  1 = single level.  &  1 \\
+\verb=  min_eff  = &   parameter for cluster algorithm for making new grids in adaptive problems  &  0.9d0 \\
+\rowcolor{tableShade}
+\verb=  minwidth  = &   The minimum size on a side for a grid created using make\_new\_grids.  &  8 \\
+\verb=  n_cellx  = &   Number of cells for the base level in the x-direction  &  -1 \\
+\rowcolor{tableShade}
+\verb=  n_celly  = &   Number of cells for the base level in the y-direction  &  -1 \\
+\verb=  n_cellz  = &   Number of cells for the base level in the z-direction  &  -1 \\
+\rowcolor{tableShade}
+\verb=  octant  = &   set octant = T if you just want to model an octant of a sphere (note: only takes effect for spherical geometry)  &  .false. \\
+\verb=  prob_hi_x  = &   physical coordinates of hi-x corner of problem domain  &  1.d0 \\
+\rowcolor{tableShade}
+\verb=  prob_hi_y  = &   physical coordinates of hi-y corner of problem domain  &  1.d0 \\
+\verb=  prob_hi_z  = &   physical coordinates of hi-z corner of problem domain  &  1.d0 \\
+\rowcolor{tableShade}
+\verb=  prob_lo_x  = &   physical coordinates of lo-x corner of problem domain  &  ZERO \\
+\verb=  prob_lo_y  = &   physical coordinates of lo-y corner of problem domain  &  ZERO \\
+\rowcolor{tableShade}
+\verb=  prob_lo_z  = &   physical coordinates of lo-z corner of problem domain  &  ZERO \\
+\verb=  ref_ratio  = &   Refinement ratio for multilevel problems  &  2 \\
+\rowcolor{tableShade}
+\verb=  regrid_int  = &   How often we regrid.  &  -1 \\
+\verb=  spherical_in  = &   set to 1 if you are doing a spherical problem  &  0 \\
+\rowcolor{tableShade}
+\verb=  test_set  = &   Fixed grid file.  &  '' \\
+\verb=  the_copy_cache_max  = &    Number of boxassoc layouts we keep in memory to avoid having to recompute the boxassoc, which is computationally expensive.  &  25 \\
+\rowcolor{tableShade}
+\verb=  the_layout_verbosity  = &    &  0 \\
+\verb=  the_sfc_threshold  = &   When assigning processors for grids, this determines whether we use the sfc algorithm or knapsack algorithm.
+If the total number of grids divided by the number of processors is greater than this number, use sfc.
+&  5 \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ heating
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  do_heating  = &   use analytic heating  &  .false. \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ hydrodynamics
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  anelastic_cutoff  = &   The density below which we modify the constraint to look like the anelastic constraint, instead of the low Mach constraint.
+This prevents velocities from getting out of hand at the edge of the star.
+Refer to Section \ref{Sec:Anelastic Cutoff}.
+&  3.d6 \\
+\verb=  base_cutoff_density  = &   The density below which we keep the initial model constant. Refer to Section \ref{Sec:Base Cutoff Density}  &  3.d6 \\
+\rowcolor{tableShade}
+\verb=  bds_type  = &   0 = use ppm instead for multi-d integrator \newline 1 = bilinear  &  0 \\
+\verb=  beta_type  = &   what type of coefficient to use inside the velocity divergence constraint.
+\newline {\tt beta\_type} = 1 uses $\beta_0$; \newline {\tt beta\_type} = 2 uses $\rho_0$ (anelastic); \newline {\tt beta\_type} = 3 uses 1 (small-scale combustion).
+&  1 \\
+\rowcolor{tableShade}
+\verb=  buoyancy_cutoff_factor  = &   The multiplicative factor (over base\_cutoff\_density) below which we do zero out the buoyancy term in the momentum equation.  &  5.0 \\
+\verb=  co_latitude  = &   latitude, in radians, for problems with rotation where the domain is only a subset of a full star.  &  ZERO \\
+\rowcolor{tableShade}
+\verb=  do_eos_h_above_cutoff  = &   After the advective enthalpy update, recompute the enthalpy if we are below the base cutoff density.  &  .true. \\
+\verb=  do_planar_invsq_grav  = &   are we doing 1/r$^2$ gravity for plane-parallel  &  .false. \\
+\rowcolor{tableShade}
+\verb=  do_smallscale  = &   force $\rho_0 = (\rho h)_0 = 0$, {\tt evolve\_base\_state = F} and {\tt beta\_type} = 3  &  .false. \\
+\verb=  do_sponge  = &   Use sponging.  &  .false. \\
+\rowcolor{tableShade}
+\verb=  dpdt_factor  = &   factor in front of the volume discrepancy term (0.0 = off)  &  0.d0 \\
+\verb=  drive_initial_convection  = &   freeze the temperature used in the reaction network to the initial value.
+This is useful for developing an initial convective field to carry away the energy, while preventing the reactions from going nonlinear.
+&  .false.
+\\
+\rowcolor{tableShade}
+\verb=  enthalpy_pred_type  = &   predict\_rhohprime = 1; \newline predict\_h = 2; \newline predict\_T\_then\_rhohprime = 3; \newline predict\_T\_then\_h = 4; \newline predict\_hprime = 5; \newline predict\_Tprime\_then\_h = 6.
+&  predict\_rhohprime \\
+\verb=  evolve_base_state  = &   turn on (.true.) or off (.false.) basestate evolution  &  .true. \\
+\rowcolor{tableShade}
+\verb=  fix_base_state  = &   if true, don't call average to reset the base state at all, even during initialization  &  .false. \\
+\verb=  grav_const  = &   the gravitational acceleration (cm~s$^{-2}$) for plane-parallel geometry  &  -1.5d10 \\
+\rowcolor{tableShade}
+\verb=  mach_max_abort  = &   maximum mach number before the code aborts  &  -1.d0 \\
+\verb=  planar_invsq_mass  = &   the point mass for planar 1/r$^2$ gravity  &  0.d0 \\
+\rowcolor{tableShade}
+\verb=  plot_sponge_fdamp  = &   plot fdamp rather than sponge assumes sponge has the form 1/(1+dt*{\tt sponge\_kappa}*fdamp)  &  .false. \\
+\verb=  ppm_type  = &   0 = no ppm \newline 1 = 1985 ppm \newline 2 = Hybrid Sekora/Colella McCorquodale Colella 2009/2010 ppm  &  1 \\
+\rowcolor{tableShade}
+\verb=  restart_with_vel_field  = &   restart the simulation using a result from a {\tt drive\_initial\_convection} = T run note that this uses the restart variable to specify which file to restart from.
+After reading in the velocity information from the restart file, the time and timestep number are zeroed.
+&  .false.
+\\
+\verb=  rotation_radius  = &   radius used for computing centrifugal term in rotation problems  &  1.0d6 \\
+\rowcolor{tableShade}
+\verb=  rotational_frequency  = &   rotational frequency used for computing centrifugal term in rotation problems.  &  ZERO \\
+\verb=  slope_order  = &   order of slopes in Godunov algorithm.  Options are 0, 2, or 4.  &  4 \\
+\rowcolor{tableShade}
+\verb=  species_pred_type  = &   Which quantities do we predict to the edges for computing the ($\rho X$) edge states?
+\newline {\tt species\_pred\_type} = 1 means predict $\rho^\prime$ and $X$ separately.
+\newline {\tt species\_pred\_type} = 2 means predict the full ($\rho X$) itself.
+\newline {\tt species\_pred\_type} = 3 means predict $\rho$ and $X$ separately.
+&  1 \\
+\verb=  sponge_center_density  = &   Center of the inner sponge.  &  3.d6 \\
+\rowcolor{tableShade}
+\verb=  sponge_kappa  = &   Parameter for sponge.  Problem dependent.  &  10.d0 \\
+\verb=  sponge_start_factor  = &   The sponge begins at sponge\_center\_density * sponge\_start\_factor.  &  3.333d0 \\
+\rowcolor{tableShade}
+\verb=  use_delta_gamma1_term  = &   turns on second order correction to delta gamma1 term  &  .false. \\
+\verb=  use_etarho  = &   turn on the etarho term as described in flow chart  &  .true. \\
+\rowcolor{tableShade}
+\verb=  use_linear_grav_in_beta  = &   how to represent gravity in the beta0 integration: .true. = piecewise linear .false. = piecewise constant  &  .false. \\
+\verb=  use_tfromp  = &   When updating temperature, use $T=T(\rho,p_0,X) $ rather than $T=T(\rho,h,X)$.  &  .false. \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ linear solvers
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  cg_verbose  = &   Verbosity of bottom solver  &  0 \\
+\verb=  hg_bottom_solver  = &   valid values are $\le$ 0  &  -1 \\
+\rowcolor{tableShade}
+\verb=  hg_dense_stencil  = &   In hgproject, in 2D, use a 9 point Laplacian (.true.) or 5-point Laplacian (.false.).
+In 3D, use a 27 point Laplacian (.true.) or 7-point Laplacian (.false.).
+&  .true.
+\\
+\verb=  max_mg_bottom_nlevels  = &   if mg\_bottom\_solver == 4, then how many mg levels can the bottom solver mgt object have  &  1000 \\
+\rowcolor{tableShade}
+\verb=  mg_bottom_solver  = &   valid values are $\le$ 0  &  -1 \\
+\verb=  mg_verbose  = &   Verbsoity of the multigrid solver, but not the bottom solver.  &  0 \\
+\rowcolor{tableShade}
+\verb=  use_hypre  = &   use the hypre library  &  .false. \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ output
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  check_base_name  = &   prefix to use in checkpoint file names  &  "chk" \\
+\verb=  chk_int  = &   Number of timesteps between writing a checkpoint file  &  0 \\
+\rowcolor{tableShade}
+\verb=  diag_buf_size  = &   number of timesteps to buffer diagnostic output information before writing (note: not implemented for all problems)  &  10 \\
+\verb=  lUsingNFiles  = &   If .true., use nOutFiles processors to write checkpoint and plotfiles.
+Fortran has the unfortunate feature of each processor only being able to write out 1-2GB each without crashing.
+&  .true.
+\\
+\rowcolor{tableShade}
+\verb=  nOutFiles  = &   If lUsingNFiles = .true., use this many processors to write checkpoint and plotfiles.
+Fortran has the unfortunate feature of each processor only being able to write out 1-2GB each without crashing.
+&  64 \\
+\verb=  plot_Hext  = &   plot external heating (Hext) in plotfile  &  .false. \\
+\rowcolor{tableShade}
+\verb=  plot_Hnuc  = &   plot nuclear energy generation rate (Hnuc) in plotfile  &  .true. \\
+\verb=  plot_ad_excess  = &   plot the adiabatic excess  &  .false. \\
+\rowcolor{tableShade}
+\verb=  plot_base  = &   plot w0\_x, w0\_y, w0\_z, divw0, rho0, rhoh0, h0, and p0 in plotfile  &  .false. \\
+\verb=  plot_base_name  = &   prefix to use in plotfile file names  &  "plt" \\
+\rowcolor{tableShade}
+\verb=  plot_cs  = &   plot soundspeed  &  .false. \\
+\verb=  plot_deltat  = &   rather than use a plot interval, plot a file after the solution has advanced past plot\_deltat in time  &  -1.d0 \\
+\rowcolor{tableShade}
+\verb=  plot_eta  = &   plot $\etarho$ in plotfile  &  .false. \\
+\verb=  plot_gpi  = &   plot pi and grad(pi)  &  .true. \\
+\rowcolor{tableShade}
+\verb=  plot_h_with_use_tfromp  = &   Turn on storing of enthalpy-based quantities in the plotfile when we are running with {\tt use\_tfromp}  &  .true. \\
+\verb=  plot_int  = &   plot interval  &  0 \\
+\rowcolor{tableShade}
+\verb=  plot_omegadot  = &   plot omegadot in plotfile  &  .true. \\
+\verb=  plot_processors  = &   create a field in the plotfile storing the processor number for each zone  &  .false. \\
+\rowcolor{tableShade}
+\verb=  plot_spec  = &   plot species and omegadot in plotfile  &  .true. \\
+\verb=  plot_trac  = &   plot tracers in plotfile  &  .false. \\
+\rowcolor{tableShade}
+\verb=  single_prec_plotfiles  = &   store the state data in single precision  &  .false. \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ particles
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  store_particle_vels  = &   store the velocity of the particle?  &  .false. \\
+\verb=  use_particles  = &   call the particle initialization, advection, etc. routines.  &  .false. \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ problem initialization
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  model_file  = &   input model file  &  "model.hse" \\
+\verb=  perturb_model  = &   Turn on a perturbation in the initial data.  Problem specific.  &  .false. \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ thermal diffusion
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  conductivity_constant  = &   the constant conductivity (only applies if the conductivity/constant module is used)  &  1.0d0 \\
+\verb=  temp_diffusion_formulation  = &   How to compute the explicit thermal diffusion term.  \newline 1 = in terms of $T$; \newline 2 = in terms of $\rho,p_0,X$.  &  2 \\
+\rowcolor{tableShade}
+\verb=  thermal_diffusion_type  = &   In the thermal diffusion solver, 1 = Crank-Nicholson; 2 = Backward Euler.  &  1 \\
+\verb=  use_thermal_diffusion  = &   Use thermal diffusion.  &  .false. \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[runtime parameters]{ timestepping 
+ parameters.} \label{table:runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  cflfac  = &   CFL factor to use in the computation of the advection timestep constraint  &  0.5d0 \\
+\verb=  fixed_dt  = &   Fix the time step.  If -1.0, then use the standard time step.  &  -1.0d0 \\
+\rowcolor{tableShade}
+\verb=  init_shrink  = &   the multiplicative factor ($\le 1$) to reduce the initial timestep as computed by the various timestep estimators  &  1.d0 \\
+\verb=  max_dt  = &   This is the maximum dt that is allowed  &  1.0d33 \\
+\rowcolor{tableShade}
+\verb=  max_dt_growth  = &   The maximum scale factor that the time step is allowed to grow by per time step.  &  1.1d0 \\
+\verb=  max_step  = &   Maximum number of steps in the simulation.  &  1 \\
+\rowcolor{tableShade}
+\verb=  nuclear_dt_fac  = &   If $T_{max}^n > T_{max}^{n-1}$ set the new dt =   min[dt, dt*{\tt nuclear\_dt\_fac}*( $T_{max}^{n-1}$ / $(T_{max}^n-T_{max}^{n-1})$ ) ] for example, {\tt nuclear\_dt\_fac} = 0.01 means don't let the max temp grow more than approximately 1 percent not checkpoint-compatible yet since it wouldn't be backwards compatible  &  -1.0d0 \\
+\verb=  small_dt  = &   the minimum allowed timestep -- we abort if dt drops below this value  &  1.d-10 \\
+\rowcolor{tableShade}
+\verb=  stop_time  = &   simulation stop time  &  -1.d0 \\
+\verb=  use_divu_firstdt  = &   Use the divu constraint when computing the first time step.  &  .false. \\
+\rowcolor{tableShade}
+\verb=  use_soundspeed_firstdt  = &   Use the soundspeed constraint when computing the first time step.  &  .false. \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+\end{landscape}
+
+%

### 2012-01-07 docs/runtime_parameters/rp.py

@@ -2,7 +2,7 @@
 import sys
 
 # tex format stuff
-header=r"""
+Mheader=r"""
 \section{Runtime Parameters}
 
 Table~\ref{table:runtime} lists the runtime parameters available to
@@ -14,14 +14,16 @@ shown here.
 %%%%%%%%%%%%%%%%
 
 \begin{landscape}
+"""
 
+header=r"""
 {\small
 
 \renewcommand{\arraystretch}{1.5}
 %
 \begin{center}
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{runtime parameters.} \label{table:runtime} \\
+\caption[runtime parameters]{@@catname@@} \label{table:runtime} \\
 %
 \hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
        \multicolumn{1}{ c|}{\textbf{description}} & 
@@ -49,7 +51,9 @@ footer=r"""
 \end{center}
 
 } % ends \small
+"""
 
+Mfooter=r"""
 \end{landscape}
 
 %
@@ -66,7 +70,14 @@ class Parameter:
         self.var=""
         self.default=""
         self.description=[]
+        self.category=""
 
+    def value(self):
+        """ the value is what we sort based on """
+        return self.category + "." + self.var
+
+    def __cmp__(self, other):
+        return cmp(self.value(), other.value())
 
 
 def make_tex_table():
@@ -78,22 +89,27 @@ def make_tex_table():
         sys.exit(2)
 
     # local storage for the parameters
-    paramsDict={}
     paramsList=[]
     descr=r""
+    category=""
 
     # read in the file
     # skip all lines before the first empty line
     foundFirstParam = False
 
-    for line in f:
+    line = f.readline()
+    while (line):
+
         if not foundFirstParam:
             if line.isspace():
                 # this is the first empty line and we begin reading the file 
                 # from here on out
                 foundFirstParam = True
+                line = f.readline()
                 continue
+
             # no blank line found yet, keep going
+            line = f.readline()
             continue
 
         # land here once we have found the first parameter
@@ -101,11 +117,33 @@ def make_tex_table():
         
         # skip blank lines
         if line.isspace(): 
+            line = f.readline()
             continue
 
+    
+        # look for category definition
+        elif line.startswith("#------"):
+
+            # the next line should be the category definition
+            line = f.readline()
+            index = line.find(":")
+            category = line[index+1:]
+
+            # following this is another #---------
+            line = f.readline()
[…]

### 2012-01-07 docs/runtime_parameters/runtime_parameters.tex
fix table of contents

@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ EOS
+\caption[ EOS
+ parameters.]{ EOS
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ SDC
+\caption[ SDC
+ parameters.]{ SDC
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ algorithm initialization
+\caption[ algorithm initialization
+ parameters.]{ algorithm initialization
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ base state mapping
+\caption[ base state mapping
+ parameters.]{ base state mapping
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ burning
+\caption[ burning
+ parameters.]{ burning
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ general MAESTRO
+\caption[ general MAESTRO
+ parameters.]{ general MAESTRO
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ grid
+\caption[ grid
+ parameters.]{ grid
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ heating
+\caption[ heating
+ parameters.]{ heating
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ hydrodynamics
+\caption[ hydrodynamics
+ parameters.]{ hydrodynamics
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ linear solvers
+\caption[ linear solvers
+ parameters.]{ linear solvers
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ output
+\caption[ output
+ parameters.]{ output
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ particles
+\caption[ particles
+ parameters.]{ particles
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ problem initialization
+\caption[ problem initialization
+ parameters.]{ problem initialization
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ thermal diffusion
+\caption[ thermal diffusion
+ parameters.]{ thermal diffusion
  parameters.} \label{table:runtime} \\
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ timestepping 
+\caption[ timestepping 
+ parameters.]{ timestepping 
  parameters.} \label{table:runtime} \\

### 2012-01-07 docs/runtime_parameters/rp.py

@@ -23,7 +23,7 @@ header=r"""
 %
 \begin{center}
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{@@catname@@} \label{table:runtime} \\
+\caption[@@catname@@]{@@catname@@} \label{table:runtime} \\
 %
 \hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
        \multicolumn{1}{ c|}{\textbf{description}} & 

### 2012-01-07 docs/runtime_parameters/runtime_parameters.tex
fix multiply-defined labels

--- 2012-01-07 docs/runtime_parameters/runtime_parameters.tex
@@
 
-Table~\ref{table:runtime} lists the runtime parameters available to
+The following tables list the runtime parameters available to
 all MAESTRO problems.  Problem-specific runtime parameters are not
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ EOS
- parameters.} \label{table:runtime} \\
+\caption[ EOS
+ parameters.]{ EOS
+ parameters.} \label{table:  EOS
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ SDC
- parameters.} \label{table:runtime} \\
+\caption[ SDC
+ parameters.]{ SDC
+ parameters.} \label{table:  SDC
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ algorithm initialization
- parameters.} \label{table:runtime} \\
+\caption[ algorithm initialization
+ parameters.]{ algorithm initialization
+ parameters.} \label{table:  algorithm initialization
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ base state mapping
- parameters.} \label{table:runtime} \\
+\caption[ base state mapping
+ parameters.]{ base state mapping
+ parameters.} \label{table:  base state mapping
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ burning
- parameters.} \label{table:runtime} \\
+\caption[ burning
+ parameters.]{ burning
+ parameters.} \label{table:  burning
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ general MAESTRO
- parameters.} \label{table:runtime} \\
+\caption[ general MAESTRO
+ parameters.]{ general MAESTRO
+ parameters.} \label{table:  general MAESTRO
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ grid
- parameters.} \label{table:runtime} \\
+\caption[ grid
+ parameters.]{ grid
+ parameters.} \label{table:  grid
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ heating
- parameters.} \label{table:runtime} \\
+\caption[ heating
+ parameters.]{ heating
+ parameters.} \label{table:  heating
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ hydrodynamics
- parameters.} \label{table:runtime} \\
+\caption[ hydrodynamics
+ parameters.]{ hydrodynamics
+ parameters.} \label{table:  hydrodynamics
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ linear solvers
- parameters.} \label{table:runtime} \\
+\caption[ linear solvers
+ parameters.]{ linear solvers
+ parameters.} \label{table:  linear solvers
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ output
- parameters.} \label{table:runtime} \\
+\caption[ output
+ parameters.]{ output
+ parameters.} \label{table:  output
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ particles
- parameters.} \label{table:runtime} \\
+\caption[ particles
+ parameters.]{ particles
+ parameters.} \label{table:  particles
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ problem initialization
- parameters.} \label{table:runtime} \\
+\caption[ problem initialization
+ parameters.]{ problem initialization
+ parameters.} \label{table:  problem initialization
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ thermal diffusion
- parameters.} \label{table:runtime} \\
+\caption[ thermal diffusion
+ parameters.]{ thermal diffusion
+ parameters.} \label{table:  thermal diffusion
+ parameters. runtime} \\
 %
@@
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[runtime parameters]{ timestepping 
- parameters.} \label{table:runtime} \\
+\caption[ timestepping 
+ parameters.]{ timestepping 
+ parameters.} \label{table:  timestepping 
+ parameters. runtime} \\
 %

### 2012-01-07 docs/runtime_parameters/rp.py

@@ -5,7 +5,7 @@ import sys
 Mheader=r"""
 \section{Runtime Parameters}
 
-Table~\ref{table:runtime} lists the runtime parameters available to
+The following tables list the runtime parameters available to
 all MAESTRO problems.  Problem-specific runtime parameters are not
 shown here.
 
@@ -23,7 +23,7 @@ header=r"""
 %
 \begin{center}
 \begin{longtable}{|l|p{5.25in}|l|}
-\caption[@@catname@@]{@@catname@@} \label{table:runtime} \\
+\caption[@@catname@@]{@@catname@@} \label{table: @@catname@@ runtime} \\
 %
 \hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
        \multicolumn{1}{ c|}{\textbf{description}} & 

### 2012-01-10 docs/SDC/sdc.tex
add intro

@@
 \section{Implementing an SDC-enabled reaction network (no diffusion)}
+
+\subsection{Introduction to SDC}
 
@@
 together with a constraint equation.
+
+By default, \maestro\ uses an operator-split approach to coupling
+the different physical processes together (hydro, reactions, diffusion).
+With operator-splitting, we
+react for a half timestep, then perform the hydrodynamics and thermal
+diffusion on this intermediate state, and finally react for the
+remaining half timestep to advance the solution.  For problems where
+the reactions greatly alter the energy balance or composition, this
+operator splitting approach can lead to inaccuracies in the solution.
+This can be more of an issue for low Mach number methods that can take
+large hydrodynamic time steps.
+
+An alternate approach to coupling is spectral deferred corrections
+(SDC). SDC is an iterative approach to couple the various operations
+together, with each operator seeing an iteratively lagged
+approximation of the other operations as a source term.  The SDC
+algorithm converges to an integral representation of the solution in
+time that couples all of the processes together in a self-consistent
+fashion, see \cite{Non11}.
+
+As a first attempt, we will work on coupling hydro and reactions only
+via SDC, with a time-independent base state.
 

### 2012-01-20 docs/compilers/compilers.tex
update gfortran notes

@@
 
-Probably the best compilers for \maestro.  Versions 4.5 and 4.6 seem to work
-really well and the debugging features are very reliable.
+Probably the best compilers for \maestro.  
+Some notes:
+\begin{itemize}
+\item gfortran 4.5 has no known issues.
+\item gfortran 4.6.1 and 4.6.2 on Fedora (15 and 16) have a bug that 
+shows up in multi-processor runs ({\tt toy\_convect} on 4 processors, for instance).
+The symptom is a crash during initialization with the message that
+``message sizes do not match across processes in the collective'' in a
+{\tt parallel\_bcast} call.  This is in {\tt cluster\_f.f90}, and the issue seems
+to be a multiplication error in computing the size to allocate {\tt bxs}.
+The bug is not present when you compile in debug mode, so it seems to be
+related to optimization.
+\end{itemize}
+
+An added benefit of the gfortran compilers is that their debugging
+features are very reliable.
 

### 2012-01-21 docs/architecture/architecture.tex
add comments on runtime parameter docs

@@
   \#}' character and a used to produce documentation about the
-available runtime parameters.
+available runtime parameters.  For the documentation, runtime parameters are grouped together
+in the {\tt \_parameters} file into categories.  The category headings
+are defined by comments in the {\tt \_parameters} file and any comments
+following that heading are placed into that category.  The documentation
+is produced by the script {\tt MAESTRO/docs/runtime\_parameters/rp.py}.
 

### 2012-01-21 docs/SDC/sdc.tex
lots of updating.  Explicitly write out the form of I for species for the 3 different species_pred_types.

@@
 and {\tt enthalpy\_advance} in \maestro\ and passed into the network solver
-during the reaction step.  These terms approximate the following:
+during the reaction step.
+
+\subsubsection{Advective Update}
+
+In the advective update, our goal is to compute $A_{\rho X_k}$ and
+$A_{\rho h}$.  These terms approximate the following:
 \begin{eqnarray}
@@
 uses an iteratively-lagged approximation to the reaction source terms
-in the interface prediction.  This explicitly couples the reaction
-process to the advection process.
+($I_{\rho X_k}$ and $I_{\rho h}$, see below) in the interface
+prediction.  This explicitly couples the reaction process to the
+advection process.
+
+\subsubsection{Final Update}
 
@@
 
-At the end of the integration, we define 
-\begin{eqnarray}
-I_{\rho X_k} &=& \frac{(\rho X_k)^\mathrm{new} - (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k} \\
-I_{\rho h}  &=& \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h}
-\end{eqnarray}
-These terms are used as the source terms for the advection step in the
-next SDC iteration.
+At the end of the integration, we define $I_{\rho X_k}$ and $I_{\rho
+  h}$.  The actual form of these depends on what we are predicting.
+We note that we only need $I_{\rho X_k}$ and $I_{\rho h}$ for the
+prediction of the interface states, not the final conservative update.
+That is because all we need from the advection solver is the
+approximation to $A_{\rho X_k}$ and $A_{\rho h}$, not the final
+updated state.
+
+\paragraph{Species Source Terms.}
+For the species prediction, the form of $I_{\rho X_k}$ depends on
+{\tt species\_pred\_type} (see \S \ref{sec:pred:density}):  
+\begin{itemize}
+\item {\tt species\_pred\_type} = 1 (predict $\rho'$ and $X_k$)
+%
+\begin{equation}
+I_{\rho X_k} = \frac{1}{\rho^{n+\half}} \left [ 
+      \frac{(\rho X_k)^\mathrm{new} - 
+            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
+\end{equation}
+
+\item {\tt species\_pred\_type} = 2 (predict $\rho'$ and $(\rho X_k)$)
+%
+\begin{equation}
+I_{\rho X_k} = \frac{(\rho X_k)^\mathrm{new} - (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k} 
+\end{equation}
+
+\item {\tt species\_pred\_type} = 3 (predict $\rho$ and $X_k$)
+%
+\begin{equation}
+I_{\rho X_k} = \frac{1}{\rho^{n+\half}} \left [ 
+      \frac{(\rho X_k)^\mathrm{new} - 
+            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
+\end{equation}
+
+\end{itemize}
+
+We note that there is no $I$ term for $\rho$ or $\rho'$ prediction, since
+the density evolution equation does not have a reaction source term.
+
+\paragraph{Enthalpy Source Terms.}
+For the enthalpy prediction, the form of $I_{\rho h}$ depends on
+{\tt enthalpy\_pred\_type} (see \S \ref{sec:pred:enthalpy}).:
+\begin{itemize}
+\item {\tt enthalpy\_pred\_type} = 1 \\
+not right -- should be pert form
+\begin{equation}
+I_{\rho h}  = \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h}
+\end{equation}
+\end{itemize}
+
+This is done in {\tt advance.f90} just after the call to {\tt react\_state},
+stored in the \multifab\ {\tt intra}.
+
+These terms are used as the source terms for the
+advection step in the next SDC iteration.
 

### 2012-01-22 docs/flowchart/flowchart.tex
add thermal diffusion to flowchart

@@
 \centering
-\includegraphics[scale=0.75]{\flowfigpath/flowchart}
+\includegraphics[scale=0.65]{\flowfigpath/flowchart}
 \caption[Graphical flowchart of MAESTRO]
@@
   variables that are the same in {\bf Step 6} as they are in
-  {\bf Step 2}, i.e., they are unchanged by the predictor steps.}
+  {\bf Step 2}, i.e., they are unchanged by the predictor steps.
+  The diffusion steps (4a and 8a) are optional, depending on
+  {\tt use\_thermal\_diffusion}.}
 \end{figure}

### 2012-01-22 docs/SDC/sdc.tex
tweaks
add graphical flowchart

@@
 \begin{equation}
-I_{\rho X_k} = \frac{1}{\rho^{n+\half}} \left [ 
+I_{\rho X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
       \frac{(\rho X_k)^\mathrm{new} - 
@@
 not right -- should be pert form
+if we do something in perturbational form, then don't we need to have
+the base state see the reaction term in its prediction?
 \begin{equation}
@@
 
+\ \\
 This is done in {\tt advance.f90} just after the call to {\tt react\_state},
-stored in the \multifab\ {\tt intra}.
-
+stored in the \multifab\ called {\tt intra}.
 These terms are used as the source terms for the

### 2012-01-22 docs/SDC/sdc.tex
add SDC flowchart

@@
 and {\tt enthalpy\_advance} in \maestro\ and passed into the network solver
-during the reaction step.
+during the reaction step.  A flowchart of the \maestro\ SDC algorithm is 
+shown in figure~\ref{fig:sdc:flowchart}.
+
+\begin{figure}[tb]
+\centering
+\includegraphics[scale=0.65]{\sdcfigpath/flowchart_SDC}
+\caption[Graphical flowchart of \maestro\ SDC]
+  {\label{fig:sdc:flowchart} A flowchart of the \maestro SDC algorithm.  The
+  thermodynamic state variables and local velocity are
+  indicated in each step.  The base state is not shown as it is time-independent.
+  Red text indicates that quantity was
+  updated during that step.  The predictor is 
+  outlined by the dotted box.  The blue text indicates state
+  variables that are the same in {\bf Step 6} as they are in
+  {\bf Step 2}, i.e., they are unchanged by the predictor steps.
+  The SDC loop is shown in the gray dotted box.}
+\end{figure}
 

### 2012-01-23 docs/SDC/sdc.tex
latex fixes
define I's for enthalpy

@@
 \caption[Graphical flowchart of \maestro\ SDC]
-  {\label{fig:sdc:flowchart} A flowchart of the \maestro SDC algorithm.  The
+  {\label{fig:sdc:flowchart} A flowchart of the \maestro\ SDC algorithm.  The
   thermodynamic state variables and local velocity are
@@
 \begin{equation}
-I_{\rho X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
+I_{X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
       \frac{(\rho X_k)^\mathrm{new} - 
@@
 I_{\rho X_k} = \frac{(\rho X_k)^\mathrm{new} - (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k} 
+\label{eq:sdc:Irhoo}
 \end{equation}
@@
 \paragraph{Enthalpy Source Terms.}
-For the enthalpy prediction, the form of $I_{\rho h}$ depends on
-{\tt enthalpy\_pred\_type} (see \S \ref{sec:pred:enthalpy}).:
-\begin{itemize}
-\item {\tt enthalpy\_pred\_type} = 1 \\
-not right -- should be pert form
-if we do something in perturbational form, then don't we need to have
-the base state see the reaction term in its prediction?
+The obvious analog to the above form for $I_{\rho X_k}$ is to define
 \begin{equation}
 I_{\rho h}  = \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h}
+\label{eq:sdc:Irhoh}
 \end{equation}
+This can be thought of as an approximation to $(\rho \Hnuc)$,
+but this does not match any of the quantities we predict for the
+enthalpy update ({\tt enthalpy\_pred\_type}, see \S
+\ref{sec:pred:enthalpy}).  The appropriate constructions
+are:
+
+\begin{itemize}
+\item {\tt enthalpy\_pred\_type} = {\tt predict\_rhoprime}
+
+Not implemented.
+
+Here we need an $I_{\rho h}$ term for the $(\rho h)'$ evolution
+equation (see Eq.~\ref{rhohprime equation}).  Defining the source
+terms for this requires a base state enthalpy that knows about the
+source terms---this is currently not implemented.
+
+In particular, should $(\rho h)_0$ see a $\overline{(\rho \Hnuc)}$ term?
+what about an average thermal diffusion?
+
+\item {\tt enthalpy\_pred\_type} = {\tt predict\_h}
+
+This is the most straightforward prediction type.  The SDC solver
+integrates the equation for $(\rho h)$:
+\begin{equation}
+\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt}  + \rho H_{\rm nuc} 
+\end{equation}
+(shown here without diffusion or external heat sources).  Expanding
+the time derivative and divergence, and using the continuity equation
+we see:
+\begin{equation}
+\frac{\partial h}{\partial t} = -\Ub \cdot \nabla h + \frac{1}{\rho} \frac{Dp_0}{Dt}  + \frac{1}{\rho} (\rho H_{\rm nuc}) \label{eq:sdc:h}
+\end{equation}
+Comparing these equations, we see that 
+\begin{equation}
+I_{h}  = \frac{1}{\rho^{n+\myhalf}} \left [
+    \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h} \right ]
+\end{equation}
+
+\item {\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_rhoprime} or 
+      {\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_h}
+
+Both of these {\tt enthalpy\_pred\_type}s predict temperature.  Expressing
+$h = h(p_0,T,X_k)$ and differentiating along particle paths:
+\begin{eqnarray}
+\frac{Dh}{Dt} &=& \left . \frac{\partial h}{\partial T} \right |_{p,X_k} \frac{DT}{Dt} +
+                  \left . \frac{\partial h}{\partial p} \right |_{T,X_k} \frac{Dp_0}{Dt} +
+           \sum_k \left . \frac{\partial h}{\partial X_k} \right |_{p,T} \frac{DX_k}{Dt} \\
+              &=& c_p \frac{DT}{Dt} + h_p  \frac{Dp_0}{Dt} + \sum_k \xi_k \omegadot_k
+\end{eqnarray}
+where $c_p$, $h_p$, and $\xi_k$ are as defined in the table of symbols
+(Table~\ref{table:sym}), and we substitute $DX_k/Dt = \omegadot_k$ (from the species
+continuity equation, Eq.~\ref{species equation}).  Using Eq.~\ref{eq:sdc:h}, we have
+the familiar temperature evolution equation:
+\begin{equation}
+\rho c_p \frac{DT}{Dt} = \underbrace{(1 - \rho h_p) \frac{Dp_0}{Dt}}_{\begin{smallmatrix}\text{already~accounted~for} \\ \text{in~T~prediction}\end{smallmatrix}} - \sum_k \xi_k \rho \omegadot_k + \rho \Hnuc
+\end{equation}
+where the underbraced term is already present in {\tt mktempforce}.  Recognizing that
+Eq.~\ref{eq:sdc:Irhoh} is the SDC approximation to $(\rho \Hnuc)$ and Eq.~\ref{eq:sdc:Irhoo} is the 
+SDC approximation to $(\rho \omegadot_k)$, we can define
+\begin{equation}
+I_T = \frac{1}{\rho^{n+\myhalf} c_p^{n+\myhalf}} \left \{
+  \left [ \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h} \right ]
+  - \sum_k \xi_k^{n+\myhalf} \left [      \frac{(\rho X_k)^\mathrm{new} - 
+            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ] \right \}
+\end{equation}
+
 \end{itemize}

### 2012-01-24 docs/SDC/sdc.tex
tweaks

@@
-\section{Implementing an SDC-enabled reaction network (no diffusion)}
-
-\subsection{Introduction to SDC}
+\section{SDC Overview}
 
@@
 
-\subsection{Strang-split version}
+\section{Strang-split Overview}
 
@@
 
-\subsection{SDC version}
+\section{SDC Without Thermal Diffusion or Base State Evolution}
 In the SDC version, the VODE integration at the end of an SDC
@@
 \begin{eqnarray}
-\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + A_{\rho X_k} \\
-\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + A_{\rho h}
+\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + A_{\rho X_k} \label{eq:sdc:rhoX} \\
+\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + A_{\rho h} \label{eq:sdc:rhoh}
 \end{eqnarray}
@@
 
-\subsubsection{Advective Update}
+\subsection{Advective Update}
 
@@
 
-\subsubsection{Final Update}
+\subsection{Final Update}
 
@@
 
+Finally we integrate the ODE system (Eqs.~\ref{eq:sdc:rhoX} and \ref{eq:sdc:rhoh}).
 At the end of the integration, we define $I_{\rho X_k}$ and $I_{\rho
@@
 
-\paragraph{Species Source Terms.}
+\subsubsection{Species Source Terms.}
 For the species prediction, the form of $I_{\rho X_k}$ depends on
@@
 
-\paragraph{Enthalpy Source Terms.}
+\subsubsection{Enthalpy Source Terms.}
 The obvious analog to the above form for $I_{\rho X_k}$ is to define
@@
 
-\ \\
+\subsubsection{Implementation}
+
 This is done in {\tt advance.f90} just after the call to {\tt react\_state},
@@
 
-\subsection{Summary of Changes}
+\section{Summary of Changes}
 The major changes from the non-SDC-enabled burners is the addition of

### 2012-01-27 docs/visualization/visualization.tex
add sample output
more plotsinglevar documentation

@@
+%-----------------------------------------------------------------------------
+% plotfiles
+%-----------------------------------------------------------------------------
 \section{Plotfiles}
@@
 
+%-----------------------------------------------------------------------------
+% amrvis
+%-----------------------------------------------------------------------------
 \section{Visualizing with \amrvis}
 
+%-----------------------------------------------------------------------------
+% VisIt
+%-----------------------------------------------------------------------------
 \section{Visualizing with \visit}
 
+%-----------------------------------------------------------------------------
+% data_procesing
+%-----------------------------------------------------------------------------
 \section{Visualizing with {\tt data\_processing} scripts}
 
-{\tt data\_processing/python\_plotfile} provide a simple commandline tool for doing 
-visualizations of 2-d BoxLib plotfiles.  The main driver
-is written in python and uses a set of Fortran routines, compiled with
-{\tt f2py} to interface with the plotfile data.  To use the routine,
-you will need to have {\tt matplotlib} and {\tt f2py} installed.  On
-a machine running Fedora linux, you can install these packages via
+{\tt data\_processing/python\_plotfile} provide a simple commandline
+tool for doing visualizations of 2-d \boxlib\ plotfiles, and slices
+through 3-d \boxlib\ plotfiles.  The main driver is written in python
+and uses a set of Fortran routines, compiled with {\tt f2py} to
+interface with the plotfile data.  To use the routine, you will need
+to have {\tt matplotlib} and {\tt f2py} installed.  On a machine
+running Fedora linux, you can install these packages via
 \begin{verbatim}
@@
 This will make a plot of ``tfromp'' from the plotfile {\tt plt00000} with log scaling,
-and store the output in {\tt test.png}.
+and store the output in {\tt test.png}.  See Figure~\ref{fig:python}.
 If you don't do `{\tt -o}', then a default output filename consisting of the
@@
 
+\begin{figure}[t]
+\centering
+\includegraphics[height=0.4\textheight]{\visfigpath/plt00000_tfromp}
+\caption{\label{fig:python} Plot of {\tt test2} done with the python
+  script {\tt plotsinglevar.py}.}
+\end{figure}
+
+If you list 2 different variables after the plotfile name, then they
+will be plotted side-by-side in a single figure.  For example, 
+\begin{verbatim}
+plotsinglevar.py plt00000/ tfromp enucdot
+\end{verbatim}
+produces the output shown in figure~\ref{fig:python_two}.
+
+\begin{figure}[t]
+\centering
+\includegraphics[height=0.4\textheight]{\visfigpath/plt00000_tfromp_enucdot}
+\caption{\label{fig:python_two} Plot of {\tt test2} done with the
+  python script {\tt plotsinglevar.py} showing 2 variables plotted
+  from a single plotfile.}
+\end{figure}
+
+Additional options include `{\tt -m}' to specify the minimum data
+value, `{\tt -M}' to specify the maximum data value, and `{\tt --eps}'
+to make an EPS plot instead of PNG.  Running the script with no parameters
+will give the full list available options.
+
+Limited 3-d support is available.  When run as with a plotfile name
+and variable, it will plot slices ($x$-$y$, $x$-$z$, and $y$-$z$) 
+through the center of the domain.  The option `{\tt --origin}'
+will put the slices through the origin.
+
+%-----------------------------------------------------------------------------
+% yt
+%-----------------------------------------------------------------------------
 \section{Visualizing with \yt}

### 2012-01-28 docs/SDC/sdc.tex
minor fixes

@@
 $\omegadot_k(\rho,X_k,T)$.  Furthermore, we simplify the problem
-``freezing'' the thermodynamics---i.e., $c_p$ is evaluating at the
+``freezing'' the thermodynamics---i.e., $c_p$ and $\xi_k$ are evaluated at the
 start of the integration and held constant thereafter.
@@
 the advective terms to the system of ODEs, the fact that we integrate
-$(\rho X_k)$ instead of just $X_k$, and the need to derive the
+$(\rho X_k)$ instead of just $X_k$, integrate $(\rho h)$ instead
+of $T$, and the need to derive the
 temperature from the input state for each RHS evaluation by {\tt

### 2012-01-29 docs/getting_started/getting_started.tex
more detailed quick start, including how to get boxlib

@@
 and how to look at the output.
+
+%-----------------------------------------------------------------------------
+% Quick Start
+%-----------------------------------------------------------------------------
+\section{Quick Start}
+
+Here we will run the standard {\tt test2} problem (three reacting
+bubbles in a plane-parallel stratified atmosphere) on a single
+processor.  This test problem was shown in paper~3.
+
+\begin{enumerate}
+
+\item Get a copy of \boxlib.
+
+\maestro\ requires the \boxlib\ library to manage the grids and 
+parallelization.  We also rely on the build system in \boxlib\ to
+build a \maestro\ executable.  \boxlib can be obtained via {\tt
+git}\footnote{{\tt git} is an open-source source code version control
+system.  It is a standard part of any Linux distribution.  You may
+need to use your system's package manager to install {\tt git} if it
+is not already present} as:
+\begin{verbatim}
+git clone https://ccse.lbl.gov/pub/Downloads/BoxLib.git
+\end{verbatim}
+
+\item Setup your shell environment.
+
+\maestro\ needs to know where to find \boxlib\, by specifying the 
+{\tt BOXLIB\_HOME} environment variable.
+
+If you shell is {\tt Bash}, add
+\begin{verbatim}
+export BOXLIB_HOME="/path/to/BoxLib/"
+\end{verbatim}
+to your {\tt .bashrc}.
+
+If your shell is{\tt Csh/Tcsh}, add
+\begin{verbatim}
+setenv BOXLIB_HOME /path/to/BoxLib/
+\end{verbatim}
+to your {\tt .cshrc}.
+
+\item Setup the problem's {\tt GNUmakefile}.
+
+In MAESTRO, each problem gets their own sub-directory under {\tt
+MAESTRO/}.  This sub-directory will contain any problem-specific files
+as well as the {\tt GNUmakefile} that specifies how to build the executable.
+Note: we rely on features of GNU {\tt make}.  Full details of the
+{\tt GNUmakefile} can be found in \S~\ref{sec:adding_problems}.  Here
+we will configure for a serial build.
+
+We only need to worry about the options at the very top of the
+{\tt GNUmakefile} for now.  These should be set as follows:
+\begin{itemize}
+\item {\tt NDEBUG := t}
+
+This option determines whether we compile with support for debugging
+(usually also enabling some runtime checks).  Setting {\tt NDEBUG := t}
+turns off debugging.
+
+\item {\tt MPI := }
+
+This determines whether we are doing a parallel build, using the Message
+Passing Interface (MPI) library.  For now, we leave this option empty,
+disabling MPI.  This will build \maestro\ in serial mode, so no MPI
+library needs to be present on the host system.
+
+\item {\tt OMP := }
+
+This determines whether we are using OpenMP to do parallelism within a
+shared memory node.  OpenMP is used together with MPI, with MPI
+distributing the grids across the processors and within a
+shared-memory node, OpenMP allows many cores to operate on the same
+grid.  For now, we leave this option empty, disabling OpenMP.
+
+\item {\tt SDC := }
+
+This option determines whether we want to use the alternate SDC
+(Spectral Deferred Corrections) source for \maestro\ (see
+Chapter \ref{ch:sdc}).  This is experimental, and this option not present
+in all problems.  We leave this blank so we compile the
+default \maestro\ source.
+
+\item {\tt COMP := gfortran}
+
+This option specifies the Fortran compiler.  We will use {\tt gfortran}, which is
+the preferred compiler for \maestro.  Specifying this compiler will automatically
+pull in the compiler settings as specified in {\tt BOXLIB\_HOME/Tools/F\_mk/}.
+
+\end{itemize}
+
+\item Build the executable.
+
+Type {\tt make}.  The build system will first find the dependencies
+amongst all the source files and then build the executable.  When
+finished, the executable will have a name like {\tt
+main.Linux.gfortran.exe}, where the specific parts of the name depend
+on the compilers and OS used.
+
+\item Copy in auxillary files.
+
+This problem uses the {\tt helmeos} equation of state, so 
+we need to copy in the data table for the equation of state, {\tt
+  extern/EOS/helmeos/helm\_table.dat}.
+
+\begin{verbatim}
+cp ../../extern/EOS/helmeos/helm_table.dat .
+\end{verbatim}
+
+\item Run.
+
+Each problem requires an input file.  The inputs file consists of
+lines of the form {\em parameter = value}, where {\em parameter} is
+one of the many runtime parameters \maestro\ knows, and {\em value}
+overrides the default value for that parameter.  For the {\tt test2}
+problem, we will use the inputs file {\tt inputs\_2d}.  An overview of
+some of the more common runtime parameters is given
+in \S~\ref{sec:gettingstarted:runtime}, and a full list of
+all \maestro\ runtime parameters and their default values is given in
+Chapter~\ref{ch:runtimeparameters}.
+
+\maestro\ is run simply as:
+runtime parameters.  The code is run simply as:
+\begin{verbatim}
+ ./main.Linux.Intel.exe inputs_2d
+\end{verbatim}
+We can also override the default value of any runtime parameter by specifying
+it on the commandline as
+\begin{verbatim}
+ ./main.Linux.Intel.exe inputs_2d --parameter value
+\end{verbatim}
+
+As the code runs, a lot of information will pass through the screen.
+For each timestep, each of the steps 1 through 12 shown in
+the \maestro\ flowchart (Chapter~\ref{ch:flowchart}) will be shown
+along with diagnostic information about the solution.  Upon completion
+some memory usage information is printed.
+
+\item Examine the output.
+
+As the code runs, it will output both plotfiles and checkpoints as
+well as one of more text diagnostic files ({\tt maestro\_diag.out} by
+default) with integral or extrema information (like maximum Mach
+number) from each timestep.
+
+By default, the plotfiles will be named {\tt plt}{\em nnnnn}, where
+the number {\em nnnnn} is the timestep number when the file was
+outputted.  Similarly, the checkpoints are named {\tt chk}{\em
+nnnnn}.  \boxlib\ plotfiles and checkpoints are actually directories,
+with the data stored in sub-directories grouped by refinement level.
+Details of the simulation (build information, number of processors
+used, output date, output directory, runtime parameter values, ...)
+are stored in the {\tt job\_info} file in each plotfile directory.
+
+Visualization of results is described in the next section.
+
+\end{enumerate}
+
+\section{Working with the Output}
+
+Visualization and analysis are done with on the plotfiles.  A
+number of in-house and externally developed tools can work 
+with \boxlib-formatted plotfiles.
+
+\subsection{\tt Amrvis}
+
+{\tt Amrvis} is an easy-to-use visualization tool developed at LBL for
+2- and 3D datasets which can plot slices through 3D datasets as well
+as volume-renderings.  It can also very easily extract 1D lines
+through the dataset along any coordinate direction.  It is distributed
+separately from the \maestro\ distribution.
+
+\subsection{{\tt data\_processing} scripts}
+
+Several useful analysis scripts (written in Fortran 90) can be found
+in {\tt fParallel/data\_processing}.  The {\tt GNUmakefile} there
+needs to be edited to indicate which of the tools to build.  For
+example, to extract the density along a line from the center of a
+plotfile, {\tt plt00200}, in the $y$-direction:
+
+\begin{verbatim}
+fextract.Linux.gfortran.exe -d 2 -v "density" -p plt00200
+\end{verbatim}
+
+These routines are described in \S~\ref{sec:analysis}.
+
+\subsection{VisIt}
+
+VisIt is a powerful, DOE-supported visualization tool for 2- and 3D
+datasets.  It can do contouring, volume rendering, streamlines, ...\
+directly from \boxlib\ plotfiles (to open a plotfile in VisIt, point to
+the {\tt Header} file inside the plotfile directory).  Details on
+visit can be found at {\tt https://wci.llnl.gov/codes/visit/home.html}\,.
+
+\subsection{Diagnostic Files}
+
+By default, \maestro\ outputs global diagnostics each timestep into a file
+called {\tt maestro\_diag.out}.  This includes the maximum Mach number,
+peak temperature, and peak nuclear energy generation rate.  Individual problems
+can provide their own {\tt diag.f90} file to produce custom diagnostic
+output.
 
@@
 in each problem directory lists the components of {\tt MAESTRO}
-that are used to build the executable.  Full details of the
-{\tt GNUmakefile} can be found in \S~\ref{sec:adding_problems}.
+that are used to build the executable.
 
@@
 
-\section{Compiling and Running}
-
-To build the \maestro\ executable, simply type `{\tt make}' in the
-problem directory.  The executable will have a name like {\tt
-  main.Linux.Intel.exe}, where the specific parts of the name depend
-on the compilers and OS used.  \maestro\ can be built with support for
-parallelism (using MPI or MPI+OpenMP), or compiled as a serial code.
-The top portion of the {\tt GNUmakefile} can be edited to choose the
-build options.  Setting {\tt MPI := t} will build the code to use MPI
-for parallelization.  Setting {\tt OMP := t} will use OpenMP.  OpenMP
-is used together with MPI, with MPI distributing the grids across the
-processors and within a shared-memory node, OpenMP allows many cores
-to operate on the same grid.  A final build option determines whether
-to build the code with support for debugging.  Setting {\tt NDEBUG :=
-  t} disables debugging features.
-
-If the {\tt helmeos} equation of state is chosen for the problem, then
-the data table for the equation of state, {\tt
-  extern/EOS/helmeos/helm\_table.dat} needs to be copied to the
-directory in which the code is run.
-
-Each problem will have one or more inputs files (for example, {\tt
-  test2/inputs\_2d}) which override the default values for one or more
-runtime parameters.  The code is run simply as:
-\begin{verbatim}
-  ./main.Linux.Intel.exe inputs_2d
-\end{verbatim}
-for a serial executable, and, depending on your MPI environment, something like:
-\begin{verbatim}
-  mpiexec -n 16 ./main.Linux.Intel.mpi.exe inputs_2d
-\end{verbatim}
-for the parallel code.
-
-As the code runs, it will output both plotfiles and checkpoints as
-well as one of more text diagnostic files (like {\tt
-  maestro\_diag.out}) with integral or extrema information (like
-maximum Mach number) from each timestep.  By default, the plotfiles
-will be named {\tt plt}{\em nnnnn}, where the number {\em nnnnn} is
-the timestep number when the file was outputted.  Similarly, the
-checkpoints are named {\tt chk}{\em nnnnn}.  BoxLib plotfiles and
-checkpoints are actually directories, with the data stored in
-sub-directories grouped by refinement level.  Details of the simulation
-(build information, number of processors used, output date, output
-directory, runtime parameter values, ...) are stored in the {\tt
-  job\_info} file in each plotfile directory.
-
-A large number of runtime parameters affect the code's behavior.  Any
-runtime parameter can be set either in the inputs file, as {\tt parameter
-  = value}, or on the command line by adding {\tt --parameter value} after
-listing the inputs file on the command line.  A full list of runtime
-parameters is detailed in \S~\ref{sec:runtime_parameters}.  Below we
-outline the most common runtime parameters.
-
 \section{Common Runtime Parameters}
+\label{sec:gettingstarted:runtime}
 
@@
 ${\tt n\_cellx} \cdot 2^{({\tt max\_levels} -1)}$.
-
-\section{Working with the Output}
-
-Visualization and analysis are done with on the plotfiles.  A
-number of in-house and externally developed tools can work 
-with BoxLib-formatted plotfiles.
-
-\subsection{\tt Amrvis}
-
-{\tt Amrvis} is an easy-to-use visualization tool developed at LBL for
-2- and 3D datasets which can plot slices through 3D datasets as well
-as volume-renderings.  It can also very easily extract 1D lines
-through the dataset along any coordinate direction.  It is distributed
-separately from the \maestro\ distribution.
-
-\subsection{{\tt data\_processing} scripts}
-
-Several useful analysis scripts (written in Fortran 90) can be found
-in {\tt fParallel/data\_processing}.  The {\tt GNUmakefile} there
-needs to be edited to indicate which of the tools to build.  For
-example, to extract the density along a line from the center of a
-plotfile, {\tt plt00200}, in the $y$-direction:
-
-\begin{verbatim}
-fextract.Linux.gfortran.exe -d 2 -v "density" -p plt00200
-\end{verbatim}
-
-These routines are described in \S~\ref{sec:analysis}.
-
-\subsection{VisIt}
-
-VisIt is a powerful, DOE-supported visualization tool for 2- and 3D
-datasets.  It can do contouring, volume rendering, streamlines, ...\
-directly from BoxLib plotfiles (to open a plotfile in VisIt, point to
-the {\tt Header} file inside the plotfile directory).  Details on
-visit can be found at {\tt https://wci.llnl.gov/codes/visit/home.html}\,.
-
-\subsection{Diagnostic Files}
-
-By default, \maestro\ outputs global diagnostics each timestep into a file
-called {\tt maestro\_diag.out}.  This includes the maximum Mach number,
-peak temperature, and peak nuclear energy generation rate.  Individual problems
-can provide their own {\tt diag.f90} file to produce custom diagnostic
-output.

### 2012-01-30 docs/getting_started/getting_started.tex
more quickstart

@@
 
-\item Get a copy of \boxlib.
+\item {\em Get a copy of \boxlib}.
 
@@
 
-\item Setup your shell environment.
+\item {\em Setup your shell environment}.
 
@@
 
-If your shell is{\tt Csh/Tcsh}, add
+If your shell is {\tt Csh/Tcsh}, add
 \begin{verbatim}
@@
 
-\item Setup the problem's {\tt GNUmakefile}.
+\item {\em Setup the problem's {\tt GNUmakefile}}.
 
@@
 
+Change directory to {\tt MAESTRO/test2/}.    
 We only need to worry about the options at the very top of the
@@
 
-\item Build the executable.
+\item {\em Build the executable}.
 
@@
 
-\item Copy in auxillary files.
+\item {\em Copy in auxillary files}.
 
@@
 
-\item Run.
+\item {\em Run!}
 
@@
 
-\item Examine the output.
+\item {\em Examine the output}.
 
@@
 
-\subsection{\tt Amrvis}
+\subsection{\amrvis}
 
-{\tt Amrvis} is an easy-to-use visualization tool developed at LBL for
+\amrvis\ is an easy-to-use visualization tool developed at LBL for
 2- and 3D datasets which can plot slices through 3D datasets as well
@@
 separately from the \maestro\ distribution.
+
+\amrvis\ can be obtained via git as:
+\begin{verbatim}
+git clone https://ccse.lbl.gov/pub/Downloads/Amrvis.git
+\end{verbatim}
+\amrvis\ is built in the C++ \boxlib\ framework (instead of the Fortran 
+\boxlib\ framework that \maestro\ uses).  The build systems are similar,
+but differ in a few ways.
+
+\amrvis\ uses the {\tt Motif} library for defining the GUI.  On a Linux 
+system, you may need to install the {\tt lesstif} package and any
+related development packages (e.g.\ {\tt lesstif-devel}).  Depending
+on your Linux system, you may also need to install {\tt libXpm} 
+and related development packages (e.g.\ {\tt libXpm-devel}).
+
+%need to modify _f95_lib
+
+library line
 
@@
 
-\subsection{VisIt}
+There is also a python visualization method in {\tt
+fParallel/data\_processing/python\_plotfile}.  This is described
+in \S~\ref{sec:vis:python}.
 
-VisIt is a powerful, DOE-supported visualization tool for 2- and 3D
+\subsection{\visit}
+
+\visit\ is a powerful, DOE-supported visualization tool for 2- and 3D
 datasets.  It can do contouring, volume rendering, streamlines, ...\
-directly from \boxlib\ plotfiles (to open a plotfile in VisIt, point to
+directly from \boxlib\ plotfiles (to open a plotfile in \visit, point to
 the {\tt Header} file inside the plotfile directory).  Details on
-visit can be found at {\tt https://wci.llnl.gov/codes/visit/home.html}\,.
+\visit\ can be found at:\newline
+ {\tt https://wci.llnl.gov/codes/visit/home.html}\,. \newline
+The easiest way to get started with \visit\ is to download a precompiled
+binary from the \visit\ webpage.
+
+Once \visit\ is installed, you can open a \boxlib\ plotfile by pointing
+\visit\ to the {\tt Header} file in the plotfile directory.
 

### 2012-01-31 docs/visualization/visualization.tex
small bits to integrate the new quick start
fix captions

@@
 \section{Visualizing with {\tt data\_processing} scripts}
+\label{sec:vis:python}
 
@@
 \includegraphics[height=0.4\textheight]{\visfigpath/plt00000_tfromp}
-\caption{\label{fig:python} Plot of {\tt test2} done with the python
+\caption[Basic plot of {\tt test2} done with {\tt plotsinglevar.py}]
+{\label{fig:python} Plot of {\tt test2} done with the python
   script {\tt plotsinglevar.py}.}
@@
 \includegraphics[height=0.4\textheight]{\visfigpath/plt00000_tfromp_enucdot}
-\caption{\label{fig:python_two} Plot of {\tt test2} done with the
+\caption[Plot of two variables from {\tt test2} done with {\tt plotsinglevar.py}]
+{\label{fig:python_two} Plot of {\tt test2} done with the
   python script {\tt plotsinglevar.py} showing 2 variables plotted

### 2012-02-01 docs/compilers/compilers.tex
fix sectioning
add Hopper OpenMP note

@@
-\section{Compilers}
-
 Here, we list notes on problems with specific compilers and \maestro.
@@
 %-----------------------------------------------------------------------------
-\subsection{Cray}
+\section{Cray}
 
@@
 %-----------------------------------------------------------------------------
-\subsection{gfortran}
+\section{gfortran}
 
@@
 %-----------------------------------------------------------------------------
-\subsection{IBM (xl* compilers)}
+\section{IBM (xl* compilers)}
 
@@
 %-----------------------------------------------------------------------------
-\subsection{Intel}
+\section{Intel}
 
@@
 
+Machine specific notes:
+
+\begin{itemize}
+\item on Hopper, with OpenMP:
+
+From NERSC support ``Intel OpenMP has an extra thread at run time
+beyond OpenMP threads.  This messes up the default affinity settings
+for aprun, since you have a spare thread waiting until all the other
+ones are completed.  You need to break the default affinity by adding
+either one of these arguments to aprun:''
+
+{\tt -cc none} or {\tt -cc numa\_node}.
+\end{itemize}
+
 %-----------------------------------------------------------------------------
-\subsection{PathScale}
+\section{PathScale}
 
@@
 %-----------------------------------------------------------------------------
-\subsection{PGI}
+\section{PGI}
 

### 2012-02-02 docs/visualization/visualization.tex
add contourcompare.py

@@
 {\tt data\_processing/python\_plotfile} provide a simple commandline
-tool for doing visualizations of 2-d \boxlib\ plotfiles, and slices
-through 3-d \boxlib\ plotfiles.  The main driver is written in python
-and uses a set of Fortran routines, compiled with {\tt f2py} to
-interface with the plotfile data.  To use the routine, you will need
-to have {\tt matplotlib} and {\tt f2py} installed.  On a machine
-running Fedora linux, you can install these packages via
+tool for doing visualizations of \boxlib\ plotfiles.  The main drivers
+are written in python and use a set of Fortran routines, compiled with
+{\tt f2py} to interface with the plotfile data.  To use the routine,
+you will need to have {\tt matplotlib} and {\tt f2py} installed.  On a
+machine running Fedora linux, you can install these packages via
 \begin{verbatim}   
@@
 \end{verbatim}
-
+%
 The library required by the python routines can be built by typing
 '{\tt make}' in that directory.  If successful, you should find
-a library {\tt fsnapshot.so}.  A simple plot can be made via:
+a library {\tt fsnapshot.so}.
+
+\subsection{\tt plotsinglevar.py}
+
+{\tt plotsinglevar.py} does visualizations of 2-d \boxlib\ plotfiles,
+and slices through 3-d \boxlib\ plotfiles.  A simple plot can be made
+via:
 \begin{verbatim}
@@
 will put the slices through the origin.
+
+\subsection{\tt contourcompare.py}
+
+{\tt contourcompare.py} takes 2 plotfiles and a variable as arguments
+and plots contours of the 2 datasets on the same set of axes.  This is 
+form comparisons of different runs.  Running the script with no parameters
+will give the full list available options.
 

### 2012-02-05 docs/SDC/sdc.tex
sketching out MAESTRO flow chart

@@
 \begin{eqnarray}
-\frac{\partial(\rho X_k)}{\partial t} &=& 
-    -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
 \frac{\partial\Ub}{\partial t} &=& 
@@
     - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
+\frac{\partial(\rho X_k)}{\partial t} &=& 
+    -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
 \frac{\partial(\rho h)}{\partial t} &=& 
@@
 need to evolve $p_0$ in the RHS routine).
+
+\section{Algorithm Flowchart - ADR with Fixed Base State}
+
+{\bf Step 1:} ({\it Compute advection velocities})\\
+
+{\bf Step 2:} ({\it Advance thermodynamic variables})\\
+
+{\bf Step 2A:} ({\it Predictor})\\
+
+{\bf Step 2A-i:} Compute time-centered edge states for species, density, and enthalpy.
+Use explicitly evaluated diffusion and time-lagged reaction source terms.
+Use these to compute advective fluxes.\\
+
+{\bf Step 2A-ii:} Advance the species and density with the advective fluxes.  Solve a Crank-Nicolson diffusion equation to advance enthalpy, using base-time transport coefficients everywhere.\\
+
+{\bf Step 2A-iii:} Use VODE to advance $\rho X_k$ and $\rho h$, using piecewise-constant, time-centered advective and diffusive flux divergences.\\
+
+{\bf Step 2A-iv:} ({\it Optional}) Compute $S^{n+1,(0)}$, define $S^{n+\myhalf,(0)}$, and perform a MAC projection to obtain an updated advection velocity.\\
+
+{\bf Step 2B:} ({\it Corrector Loop})\\
+
+{\bf Step 2B-i:} Same as {\bf Step 2A-i}, but use iterative-lagged reaction source terms.
+
+{\bf Step 2B-ii:} Advance the species and density with the advective fluxes.  Solve a backward-Euler type correction equation for diffusion to advance $\rho h$.
+
+{\bf Step 2B-iii:} Use VODE to advance $\rho X_k$ and $\rho h$, using piecewise-constant, time-centered advective and diffusive flux divergences.
+Skip to Step 3 if this is the final MISDC iteration.\\
+
+{\bf Step 2B-iv:} ({\it Optional}) Compute $S^{n+1,(0)}$, define $S^{n+\myhalf,(0)}$, and perform a MAC projection to obtain an updated advection velocity.  Then return to {\bf Step 2B-i}.\\
+
+{\bf Step 3:} ({\it Advance the velocity})\\

### 2012-02-06 docs/architecture/architecture.tex
reference evolving boxlib manual
update to describe new directoy layout

@@
   Simple Fortran-based analysis routines (e.g.\ extract a line from a
-  multidimensional dataset) that operate on BoxLib datasets.
+  multidimensional dataset) that operate on BoxLib datasets.  These are
+  described in \S~\ref{sec:analysis}.
+
+  Several sub-directories with python-based routines are also here.
+  These are described both in \S~\ref{sec:analysis} and \S~\ref{sec:vis:python}.
 
@@
   External modules, like ODE integrators, equations of state, reaction
-  networks
+  networks, etc.  Many of these routines were developed by other groups.
 
@@
 
-    An integration package for ODEs.  At the moment, this is used for
-    integrating various reaction networks.
+    The {\tt VODE} \cite{vode} package for integrating ODEs.  At the
+    moment, this is used for integrating various reaction networks.
 
@@
 
-  The main \maestro\ algorithm directory.  The individual problem
-  directories lie under here, and all of the main \maestro\ source
-  lives in the {\tt Source/} sub-directory.
+  The main \maestro\ algorithm directory.
+
+  Each problem in \maestro\ gets it own sub-directory under {\tt
+    SCIENCE/}, {\tt TEST\_PROBLEMS/}, or {\tt UNIT\_TESTS/}.  The {\tt
+    GNUmakefile} in the problem directory includes the instructions on
+  how to build the executable, including what modules in {\tt extern}
+  are used.  Any file that you place in a sub-directory here takes
+  precedence over a file of the same name in {\tt MAESTRO/}.  This
+  allows problems to have custom versions of the main
+  \maestro\ routines (e.g.\ initial conditions via {\tt initdata.f90}.
+  See \S~\ref{sec:makefile} for details on the build system).
 
@@
 
+  \item {\tt SCIENCE/}
+
+    \maestro\ problem directories for science studies.  These are the
+    setups that have been used for science papers in the past, or are
+    the basis for current science studies.
+
   \item {\tt Source}
@@
 
-  \item problem directories: {\tt test2}, {\tt test\_spherical},
-    $\ldots$
-
-    Each problem in \maestro\ gets it own sub-directory.  The {\tt
-      GNUmakefile} in the problem directory includes the instructions
-    on how to build the executable, including what modules in {\tt
-      extern} are used.
-
-    Any file that you place in a sub-directory here takes precedence
-    over a file of the same name in {\tt MAESTRO/}.  This allows
-    problems to have custom versions of the main \maestro\ routines
-    (e.g.\ initial conditions via {\tt initdata.f90}.  See
-    \S~\ref{sec:makefile} for details on the build system).
+  \item {\tt TEST\_PROBLEMS}
+
+    \maestro\ problem directories for simple test problems that have
+    been used while developing \maestro.  Many of these problems have
+    appeared in the various papers describing the \maestro\ algorithm.
+
+  \item {\tt UNIT\_TESTS}
+
+    Special \maestro\ problem directories that test only a single component
+    of the \maestro\ algorithm.  These often have their own main drivers
+    ({\tt varden.f90}) that setup and initialize some data structures and
+    then call only a few of the \maestro\ routines.  See \S~\ref{sec:unit_tests}
+    for details.
 
@@
 provides a number of Fortran types.  We briefly summarize the major
-data types below.
+data types below.  A more extensive introduction to \boxlib\ is 
+provided by the \boxlib\ User's Guide, distributed with the library.
 

### 2012-02-06 docs/getting_started/getting_started.tex
update for the new directory structure

@@
 
-In MAESTRO, each problem gets their own sub-directory under {\tt
-MAESTRO/}.  This sub-directory will contain any problem-specific files
-as well as the {\tt GNUmakefile} that specifies how to build the executable.
-Note: we rely on features of GNU {\tt make}.  Full details of the
-{\tt GNUmakefile} can be found in \S~\ref{sec:adding_problems}.  Here
-we will configure for a serial build.
+In \maestro, each problem lives under one of three sub-directories of
+{\tt MAESTRO/}: {\tt SCIENCE/}, {\tt TEST\_PROBLEMS/}, or {\tt
+UNIT\_TESTS/}.  This problem sub-directory will contain any
+problem-specific files as well as the {\tt GNUmakefile} that specifies
+how to build the executable.  Note: we rely on features of GNU {\tt
+make}.  Full details of the {\tt GNUmakefile} can be found
+in \S~\ref{sec:adding_problems}.  Here we will configure for a serial
+build.
 
-Change directory to {\tt MAESTRO/test2/}.    
+Change directory to {\tt MAESTRO/TEST\_PROBLEMS/test2/}.    
 We only need to worry about the options at the very top of the
@@
 \begin{verbatim}
-cp ../../extern/EOS/helmeos/helm_table.dat .
+cp ../../../extern/EOS/helmeos/helm_table.dat .
 \end{verbatim}

### 2012-02-06 docs/getting_started/getting_started.tex
more syncing

@@
 
-Different problems in \maestro\ are contained in individual
-sub-directories under {\tt MAESTRO/}.  The {\tt GNUmakefile}
-in each problem directory lists the components of {\tt MAESTRO}
-that are used to build the executable.
+Different problems in \maestro\ are contained in one of three
+sub-directories under {\tt MAESTRO/} ({\tt SCIENCE/}, {\tt
+TEST\_PROBLEMS}, or {\tt UNIT\_TESTS}).  The {\tt GNUmakefile} in each
+problem directory lists the components of {\tt MAESTRO} that are used
+to build the executable.  {\tt TEST\_PROBLEMS/} contains simple
+problems that were used in the development of \maestro.  Many
+of these were featured in the papers describing the \maestro\ algorithm.
 

### 2012-02-07 docs/networks/networks.tex
start of section describing networks

\section{Networks}

\maestro\ models multiple species, described by the mass density of
the fluid, $\rho$, and the mass fraction of the species, $X_k \equiv
\rho_k/\rho$, where $\rho_k$ is the mass density of species $k$.  All
\maestro\ problems, regardless of whether they model reactions, need a
network.  In its most basic form, the network supplies the properties
of the species (atomic mass, atomic number) that are interpreted by
the equation of state to compute.

\subsection{{\tt general\_null}}

This is a 'null' network -- i.e. no burning, just define the
properties of the species for thermodynamics.  The twist is that you
can create an inputs file to define what species you want to carry.
For example, the extern/networks/null/ network defines C12, O16, and
Mg24.  To replicate this in {\tt general\_null}, we have the file
{\tt ignition.net} with contents:

\begin{verbatim}
# name       short name    aion     zion
 carbon-12      C12         12.0     6.0
 oxygen-16      O16         16.0     8.0
 magnesium-24   Mg24        24.0    12.0
\end{verbatim}

To use this set of species in your problem, you would set:

\begin{verbatim}
NETWORK_DIR := extern/networks/general_null
GENERAL_NET_INPUTS := ignition.net
\end{verbatim}

It is assumed that the *.net files live in {\tt extern/networks/general\_null/}

Then at compile time, the network.f90 is created using these species and
compiled.  (For the curious, the rule to build {\tt network.f90} lives in
{\tt extern/networks/general\_null/GPackage.mak})

### 2012-02-07 docs/SDC/sdc.tex
minor
some SDC updates

--- 2012-01-28 docs/SDC/sdc.tex
@@
 need to evolve $p_0$ in the RHS routine).
+
+\section{Algorithm Flowchart - ADR with Fixed Base State}
+We now include thermal diffusion and assume the base state is constant in time but not space:
+\begin{eqnarray}
+\frac{\partial\Ub}{\partial t} &=& 
+    -\Ub\cdot\nabla\Ub  - \frac{1}{\rho}\nabla\pi 
+    - \frac{\rho-\rho_0}{\rho} g\eb_r,\\
+\frac{\partial(\rho X_k)}{\partial t} &=& 
+    -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\\
+\frac{\partial(\rho h)}{\partial t} &=&
+    -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} 
+    + \rho\Hnuc
+    + \underbrace{\nabla\cdot\frac{\kth}{c_p}\nabla h - \sum_k\nabla\cdot\frac{\xi_k k_{\rm th}}{c_p}\nabla X_k - \nabla\cdot\frac{h_p k_{\rm th}}{c_p}\nabla p_0}_{\nabla\cdot k_{\rm th}\nabla T}.
+\end{eqnarray}
+
+{\bf Step 1:} ({\it Advance thermodynamic variables}) In this step we integrate $(\rho, \rho X_k, \rho h)$ over the full time step.  We use a predictor-corrector approach, as explained below.\\
+
+{\bf Step 1A:} ({\it Predictor}) In the predictor, we compute a preliminary estimate of $(\rho, \rho X_k, \rho h)$ at the new time level in order to obtain time-advanced values of $S, \beta_0$, and the other thermodynamic coefficients used in the corrector.
+In summary, we compute advection velocities using a Godunov method and a MAC projection, compute advective fluxes, use a Crank-Nicolson-like discretization to compute diffusive fluxes, and integrate the entire system with VODE.
+The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^n$ are computed from the the state at $t^n$.
+Here are the details.\\
+
+{\bf Step 1A-i:} ({\it Compute advection velocities}) Use $\Ub^n$ and a second-order Godunov method to compute time-centered edge velocities, $\uadvone$, with a lagged pressure gradient and explicit buoyancy forcing terms.
+The $\star$ superscript indicates that this field does not satisfy the divergence constrain.
+Compute $S^{n+\myhalf,{\rm pred}}$ by extrapolating in time,
+\begin{equation}
+S^{n+\myhalf,{\rm pred}} = S^n + \frac{\Delta t^n}{2}\frac{S^n - S^{n-1}}{\Delta t^{n-1}},
+\end{equation}
+and project $\uadvone$ to obtain $\uadvpred$, which satisfies
+\begin{equation}
+\nabla\cdot\left(\beta_0^n\uadvpred\right) = S^{n+\myhalf,\rm{pred}}.
+\end{equation}
+
+{\bf Step 1A-ii:} Use $\uadvpred$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,{\rm pred}}$, with time-lagged reaction processes and explicitly evaluated diffusion and thermodynamic pressure as source terms.
+Advance the density using
+\begin{equation}
+\frac{\rho^{n+1,{\rm pred}} - \rho^n}{\Delta t} = -\sum_m\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{\rm pred}}\uadvpred\right].
+\end{equation}
+
+{\bf Step 1A-iii:}   Solve a Crank-Nicolson diffusion equation to advance enthalpy, using base-time transport coefficients everywhere.\\
+
+{\bf Step 1A-iv:} Use VODE to advance ($\rho X_k, \rho h$), using piecewise-constant, time-centered advective and diffusive flux divergences.\\
+
+{\bf Step 1B:} ({\it Corrector}) The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^{n+1,{\rm pred}}$ are computed from the the state at $t^{n+1,{\rm pred}}$.  Here are the details.\\
+
+{\bf Step 1B-i:} ({\it Compute updated advection velocities}) First, compute $S^{n+\myhalf}$ and $\beta_0^{n+\myhalf}$ using
+\begin{equation}
+S^{n+\myhalf} = \frac{S^n + S^{n+1,{\rm pred}}}{2}, \qquad \beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1,{\rm pred}}}{2}
+\end{equation}
+Then, project $\uadvone$ to obtain $\uadv$, which satisfies
+\begin{equation}
+\nabla\cdot\left(\beta_0^{n+\myhalf}\uadv\right) = S^{n+\myhalf}.
+\end{equation}
+
+{\bf Step 1B-ii:} Advance the species and density with the advective fluxes.  Solve a backward-Euler type correction equation for diffusion to advance $\rho h$.
+
+{\bf Step 1B-iii:} Use VODE to advance $\rho X_k$ and $\rho h$, using piecewise-constant, time-centered advective and diffusive flux divergences.
+Skip to Step 3 if this is the final MISDC iteration.\\
+
+{\bf Step 1B-iv:} ({\it Optional}) Compute $S^{n+1,(0)}$, define $S^{n+\myhalf,(0)}$, and perform a MAC projection to obtain an updated advection velocity.  Then return to {\bf Step 2B-i}.\\
+
+{\bf Step 2:} ({\it Advance the velocity})\\

### 2012-02-07 docs/SDC/sdc.tex
first cut at SDC ADR flow chart

--- 2012-01-28 docs/SDC/sdc.tex
@@
 need to evolve $p_0$ in the RHS routine).
+
+\section{Algorithm Flowchart - ADR with Fixed Base State}
+We now include thermal diffusion and assume the base state is constant in time but not space:
+\begin{eqnarray}
+\frac{\partial\Ub}{\partial t} &=& 
+    -\Ub\cdot\nabla\Ub  - \frac{1}{\rho}\nabla\pi 
+    - \frac{\rho-\rho_0}{\rho} g\eb_r,\\
+\frac{\partial(\rho X_k)}{\partial t} &=& 
+    -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:sdc species}\\
+\frac{\partial(\rho h)}{\partial t} &=&
+    -\nabla\cdot(\rho h\Ub) + \underbrace{\Ub\cdot\nabla p_0}_{Dp_0/Dt}
+    + \rho\Hnuc
+    + \underbrace{\nabla\cdot\frac{\kth}{c_p}\nabla h - \sum_k\nabla\cdot\frac{\xi_k k_{\rm th}}{c_p}\nabla X_k - \nabla\cdot\frac{h_p k_{\rm th}}{c_p}\nabla p_0}_{\nabla\cdot k_{\rm th}\nabla T}.\label{eq:sdc enthalpy}
+\end{eqnarray}
+
+{\bf Step 1:} ({\it Advance thermodynamic variables}) In this step we integrate $(\rho, \rho X_k, \rho h)$ over the full time step.  We use a predictor-corrector approach, as explained below.\\
+
+{\bf Step 1A:} ({\it Predictor}) In the predictor, we compute an estimate of $(\rho, \rho X_k, \rho h)$ at the new time level to obtain time-advanced values of $S$, $\beta_0$, and transport coefficients that will be used in the corrector.
+The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^n$ are computed from the the thermodynamic variables at $t^n$.\\
+
+{\bf Step 1A-i:} ({\it Compute advection velocities}) Use $\Ub^n$ and a second-order Godunov method to compute time-centered edge velocities, $\uadvsdcstar$, with a lagged pressure gradient and explicit buoyancy forcing terms.
+The $\star$ superscript indicates that this field does not satisfy the divergence constrain.
+Compute $S^{n+\myhalf,{\rm pred}}$ by extrapolating in time,
+\begin{equation}
+S^{n+\myhalf,{\rm pred}} = S^n + \frac{\Delta t^n}{2}\frac{S^n - S^{n-1}}{\Delta t^{n-1}},
+\end{equation}
+and project $\uadvsdcstar$ to obtain $\uadvsdcpred$, which satisfies
+\begin{equation}
+\nabla\cdot\left(\beta_0^n\uadvsdcpred\right) = S^{n+\myhalf,\rm{pred}}.
+\end{equation}
+
+{\bf Step 1A-ii:} ({\it Compute advective fluxes}) Use $\uadvsdcpred$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,{\rm pred}}$, with time-lagged reactions, explicitly evaluated diffusion, and a time-centered thermodynamic pressure advective derivative as source terms.\\
+
+{\bf Step 1A-iii:} ({\it Compute diffusive fluxes}) Begin by compute a preliminary time-advanced density using
+\begin{equation}
+\frac{\rho^{n+1,{\rm pred}} - \rho^n}{\Delta t} = -\sum_m\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right].
+\end{equation}
+Then, compute preliminary, time-advanced species using
+\begin{equation}
+\frac{\rho^{n+1,{\rm pred}}X_{k,{\rm AD}}^{n+1,{\rm pred}} - (\rho X_k)^n}{\Delta t} = -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right] + I_{R,\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
+\end{equation}
+Then, solve a Crank-Nicolson-type diffusion equation for $h^{n+1,{\rm pred}}_{\rm AD}$, using transport coefficients evaluated at $t^n$ everywhere,
+\begin{eqnarray}
+\frac{\rho^{n+1,{\rm pred}}h^{n+1,{\rm pred}}_{\rm AD} - (\rho h)^n}{\Delta t} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right] + \uadvsdcpred\cdot\nabla p_0 + I_{R,\rho h}^{\rm lagged}\nonumber\\
+&& + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n + \nabla\cdot\frac{\kth^n}{c_p^n}\nabla h_{\rm AD}^{n+1,{\rm pred}}\right)\nonumber\\
+&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,{\rm pred}}\right)\nonumber\\
+&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right)\label{eq:sdc enthalpy 2}
+\end{eqnarray}
+
+{\bf Step 1A-iv:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}^{\rm pred}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{R,\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{\rm pred}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{R,\rho h}^{\rm lagged}$ term.
+Use VODE to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,{\rm pred}}$ using the piecewise-constant advection and diffusion source terms:
+\begin{eqnarray}
+\frac{\partial(\rho X_k)}{\partial t} &=& Q_{\rho X_k}^{\rm pred} + \rho\dot\omega_k\\
+\frac{\partial(\rho h)}{\partial t} &=& Q_{\rho h}^{\rm pred} + \rho\Hnuc.
+\end{eqnarray}
+
+{\bf Step 1B:} ({\it Corrector}) In the corrector, we use the time-advanced state from the predictor to perform a more accurate integration of the thermodynamic variables.
+The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^{n+1,{\rm pred}}$ are computed from $(\rho,\rho X_k,\rho h)^{n+1,{\rm pred}}$.\\
+
+{\bf Step 1B-i:} ({\it Compute advection velocities}) First, compute $S^{n+\myhalf}$ and $\beta_0^{n+\myhalf}$ using
+\begin{equation}
+S^{n+\myhalf} = \frac{S^n + S^{n+1,{\rm pred}}}{2}, \qquad \beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1,{\rm pred}}}{2}
+\end{equation}
+Then, project $\uadvsdcstar$ to obtain $\uadvsdc$, which satisfies
+\begin{equation}
+\nabla\cdot\left(\beta_0^{n+\myhalf}\uadvsdc\right) = S^{n+\myhalf}.
+\end{equation}
+
+{\bf Step 1B-ii:} ({\it Compute advective fluxes}) Use $\uadvsdc$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf}$, with iteratively-lagged reactions, explicitly evaluated diffusion, and a time-centered thermodynamic pressure advective derivative as source terms.\\
+
+{\bf Step 1B-iii:} ({\it Compute diffusive fluxes}) Begin by compute a preliminary time-advanced density using
+\begin{equation}
+\frac{\rho^{n+1} - \rho^n}{\Delta t} = -\sum_m\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf}\uadvsdc\right].
+\end{equation}
+Then, compute preliminary, time-advanced species using
+\begin{equation}
+\frac{\rho^{n+1}X_{k,{\rm AD}}^{n+1} - (\rho X_k)^n}{\Delta t} = -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf}\uadvsdc\right] + I_{R,\rho X_k}^{\rm pred}.\label{eq:sdc species 3}
+\end{equation}
+Then, solve a backward-Euler-type correction equation for $h^{n+1}_{\rm AD}$,
+\begin{eqnarray}
+\frac{\rho^{n+1}h^{n+1}_{\rm AD} - (\rho h)^n}{\Delta t} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf}\uadvsdc\right] + \uadvsdc\cdot\nabla p_0 + I_{R,\rho h}^{\rm pred}\nonumber\\
+&& + \nabla\cdot\frac{\kth^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla h^{n+1}_{\rm AD} + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla h^{n+1,{\rm pred}}\right)\nonumber\\
+&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,{\rm pred}} k_{\rm th}^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla X_{k,{\rm AD}}^{n+1}\right)\nonumber\\
+&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,{\rm pred}}k_{\rm th}^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla p_0\right)\label{eq:sdc enthalpy 3}
+\end{eqnarray}
+
+{\bf Step 1B-iv:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{R,\rho X_k}^{\rm pred}$ term, and define $Q_{\rho h}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{R,\rho h}^{\rm pred}$ term.
+Use VODE to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1}$ using the piecewise-constant advection and diffusion source terms:
+\begin{eqnarray}
+\frac{\partial(\rho X_k)}{\partial t} &=& Q_{\rho X_k} + \rho\dot\omega_k\\
+\frac{\partial(\rho h)}{\partial t} &=& Q_{\rho h} + \rho\Hnuc.
+\end{eqnarray}
+
+{\bf Step 2:} ({\it Advance velocity and dynamic pressure})\\

### 2012-02-08 docs/pert_notes/pert.tex
add SDC to the user's guide.  Define some section labels.
stub for Ryan's notes
some marginpars
Ryan's notes on toy_convect fun

@@
 
-The \maestro\ hyperbolic equations come in two forms: advective and conservative.
-The procedure for predicting interface states differs slightly depending on which
-form we are dealing with.
+The \maestro\ hyperbolic equations come in two forms: advective and
+conservative.  The procedure for predicting interface states differs
+slightly depending on which form we are dealing with.
 
@@
 \section{Density Evolution}
+
+\label{sec:pred:density}
 
@@
 force computed in {\tt modify\_scal\_force} with {\tt fullform =
-  .true.}.
+  .true.}.  \MarginPar{we need to switch this over to doing the conservative prediction}
 
@@
 \section{Energy Evolution}
+
+\label{sec:pred:enthalpy}
 
@@
 {\tt predict\_rhoprime\_and\_rhoX}, we wish to construct $(\rho_0 +
-\rho'_\mathrm{edge} ) h_\mathrm{edge}$.
+\rho'_\mathrm{edge} ) h_\mathrm{edge}$. \MarginPar{alternately, we could construct $\rho_\mathrm{edge} = \sum_k (\rho X_k)_\mathrm{edge}$ for s\_p\_t = 2}
 
@@
 
-%% \begin{table*}[h]
-%% \begin{center}
-%% \caption{When predicting temp ... (conversion done by {\tt makeRhoHfromT}) \newline}
-%% \begin{tabular}{c|c|c|c}
-%% \hline
-%% \hline
-%% {type} & {Inputs } & {EOS inputs} & {output} \\
-%% \hline
-%% 3 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ &  $(\rho h)^\prime = \rho h - (\rho h)_0$ \\
-%%    &                   & $   X = $ given & \\
-%%    &                   &                 & \\
-%%  4 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ & $h$ \\
-%%    &                   & $   X = $ given & \\
-%%    &                   &                 & \\
-%% \hline
-%% \end{tabular}
-%% \end{center}
-%% \end{table*}
+%-----------------------------------------------------------------------------
+% toy_convect
+%-----------------------------------------------------------------------------
+\section{Experience from {\tt toy\_convect}}
 
-%% \begin{table*}[h]
-%% \begin{center}
-%% \caption{Quantity that is created in {\tt mkflux} and passed into {\tt update\_scal}\newline\
-%% }
-%% \begin{tabular}{c|c|c|c|c}
-%%        &            &              &\multicolumn{2}{c}{Forcing Term} \\
-%% {type} & {Species } & {$(\rho h)$} & {$(\rho X)$} & {$(\rho h)$} \\
-%% \hline 
-%% 1 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
-%% 2 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
-%% 3 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r}$ \\[2mm]
-%% 4 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
-%% \hline
-%% \end{tabular}
-%% \end{center}
-%% \end{table*}
+\label{sec:toyconvect}
 
-%% \begin{table*}[h]
-%% \begin{center}
-%% \caption{Quantity that goes into and out of {\tt make\_edge\_scal} \newline}
-%% \begin{tabular}{c|c}
-%% \hline
-%% \hline
-%% {h pred type} & {h/T } \\
-%% \hline \\[-3mm]
-%% 1 & $(\rho h)^\prime$ \\
-%% 2 & $h$  \\
-%% 3 & $T$  \\
-%% 4 & $T$  \\
-%% \hline
-%% \end{tabular}
-%% \end{center}
-%% \end{table*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Why is {\tt toy\_convect} Interesting?}
 
-%% \begin{table*}
-%% \begin{center}
-%% \caption{Quantity that goes into {\tt mkflux} on edges \newline}
-%% \begin{tabular}{c|c|c}
-%% \hline
-%% \hline
-%% {h pred type} & {Species } & {h/T } \\
-%% \hline
-%% 1 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
-%% 2 & $X$ and $\rho^\prime$ & $h$ \\
-%% 3 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
-%% 4 & $X$ and $\rho^\prime$ & $h$ \\
-%% \hline
-%% \end{tabular}
-%% \end{center}
-%% \end{table*}
+The {\tt toy\_convect} problem consists of a carbon-oxygen white dwarf with 
+an accreted layer of solar composition. There is a steep composition 
+gradient between the white dwarf and the accreted layer. The convection 
+that begins as a result of the accretion is extremely sensitive to the 
+amount of mixing.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Initial Observations}
+
+With {\tt use\_tfromp = T} and {\tt cflfac = 0.7} there is a large difference 
+between {\tt species\_pred\_type = 1} and {\tt species\_pred\_type = 2,3} as 
+seen in Figure \ref{fig:spec1_vs_23}. {\tt species\_pred\_type = 1} shows 
+quick heating (peak T vs. t) and there is ok agreement between {\tt tfromh} 
+and {\tt tfromp}. {\tt species\_pred\_type = 2,3} show cooling (peak T vs. t) 
+and {\tt tfromh} looks completely unphysical (see Figure 
+\ref{fig:tfromh_unphysical}). There are also strange filament type features in 
+the momentum plots shown in Figure \ref{fig:mom_filaments}.
+
+% compare spec 1 with spec 2,3
+%-----------------------------------------
+\begin{figure}[!h]
+\centering
+\includegraphics[height=0.3\textheight]{\pertfigpath/spec1_vs_23}
+\caption{Compare {\tt species\_pred\_type = 1,2,3} with {\tt use\_tfromp = 
+T, enthalpy\_pred\_type = 1, cflfac = 0.7}}
+\label{fig:spec1_vs_23}
+\end{figure}
+
+% tfromh is unphysical
+%-----------------------------------------
+\begin{figure}[!h]
+\begin{minipage}[b]{0.5\linewidth}
+\vspace{0pt}
+\centering
+\includegraphics[height=0.3\textheight]{\pertfigpath/tfromh_unphysical}
+\caption{{\tt tfromh} is unphysical when using {\tt species\_pred\_type = 
+2,3, enthalpy\_pred\_type = 1, cflfac = 0.7, use\_tfromp = T}. Shown above 
+is {\tt species\_pred\_type = 2}}
+\label{fig:tfromh_unphysical}
+\end{minipage}
+\hspace{0.5cm}
+% momentum filaments
+%-----------------------------------------
+\begin{minipage}[b]{0.5\linewidth}
+\vspace{0pt}
+\centering
+\includegraphics[height=0.3\textheight]{\pertfigpath/mom_filaments}
+\caption{There are strange filament type features at the bottom of the 
+domain. {\tt species\_pred\_type = 2, enthalpy\_pred\_type = 1, cflfac = 0.7, 
+use\_tfromp = T}}
+\label{fig:mom_filaments}
+\end{minipage}
+\end{figure}
+
+Using {\tt use\_tfromp = F} and {\tt dpdt\_factor $>$ 0} results in many runs 
+crashing very quickly and gives unphyiscal temperature profiles as seen in 
+Figure \ref{fig:tfrompF_unphys}.
+
+% unphysical temp profiles with use_tfromp = F dpdt > 0
+%-----------------------------------------
+\begin{figure}[!h]
+\begin{minipage}[b]{0.5\linewidth}
+\vspace{0pt}
+\centering
+\includegraphics[height=0.3\textheight]{\pertfigpath/tfrompF_unphys}
+\caption{Unphysical temperature profile with {\tt use\_tfromp = F} and 
+{\tt dpdt\_factor = 0.1}. {\tt dpdt\_factor = 0.2,0.3} lead to the code 
+crashing.}
+\label{fig:tfrompF_unphys}
+\end{minipage}
+\hspace{0.5cm}
+% temp vs time of castro ppm 0,1 and maestro spec = 2
+%----------------------------------------------
+\begin{minipage}[b]{0.5\linewidth}
+\vspace{0pt}
+\centering
+\includegraphics[height=0.3\textheight]{\pertfigpath/tfrompF_cfl_1vs7}
+\caption{Compare {\tt cflfac = 0.1} with {\tt cflfac = 0.7} for 
+{\tt use\_tfromp = F, dpdt\_factor = 0.0, species\_pred\_type = 2, 
+enthalpy\_pred\_type = 4}}
+\label{fig:tfrompF_cfl_1vs7}
+\end{minipage}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Change {\tt cflfac} and {\tt enthalpy\_pred\_type}}
+
+With {\tt species\_pred\_type = 1} and {\tt cflfac = 0.1}, 
+there is much less heating (peak T vs. t) than the {\tt cflfac = 0.7} 
+(default). There is also a lower overall Mach number (see Figure 
+\ref{fig:spec1_cfl_1vs7}) with the {\tt cflfac = 0.1} and excellent agreement 
+between {\tt tfromh} and {\tt tfromp}.
+
+{\tt use\_tfromp = F, dpdt\_factor = 0.0, enthalpy\_pred\_type = 3,4 and 
+species\_pred\_type = 2,3} shows cooling (as seen in {\tt use\_tfromp = T}) 
+with a comparable rate of cooling (see Figure \ref{fig:compare_tfromp}) 
+to the {\tt use\_tfromp = T} case. The 
+largest difference between the two runs is that the {\tt use\_tfromp = F} 
+case shows excellent agreement between {\tt tfromh} and {\tt tfromp} with 
+{\tt cflfac = 0.7}. The filaments in the momentum plot of Figure 
+\ref{fig:mom_filaments} are still present.
+
+% Mach number of cfl 0.1 vs cfl 0.7
+%-----------------------------------------
+\begin{figure}[!h]
+\begin{minipage}[!b]{0.5\linewidth}
+\vspace{0pt}
+\centering
+\includegraphics[height=0.3\textheight]{\pertfigpath/spec1_cfl_1vs7}
+\caption{Comparing the Mach number of {\tt cflfac = 0.1} and 
+{\tt cflfac = 0.7}. {\tt species\_pred\_type = 1, enthalpy\_pred\_type = 1}}
+\label{fig:spec1_cfl_1vs7}
+\end{minipage}
+\hspace{0.5cm}
+% comparable cooling rates between tfromp = T and tfromp = F
+%-----------------------------------------
+\begin{minipage}[!b]{0.5\linewidth}
+\vspace{0pt}
+\centering
+\includegraphics[height=0.3\textheight]{\pertfigpath/compare_tfromp}
+\caption{Illustrate the comparable cooling rates between 
+{\tt use\_tfromp = T} and {\tt use\_tfromp = F with dpdt\_factor = 0.0} 
+using {\tt species\_pred\_type = 2, enthalpy\_pred\_type = 3,1}}
+\label{fig:compare_tfromp}
+\end{minipage}
+\end{figure}
+
+For a given {\tt enthalpy\_pred\_type} and {\tt use\_tfromp = F}, 
+{\tt species\_pred\_type = 2} has a lower Mach number (vs. t) compared to 
+{\tt species\_pred\_type = 3}.
+
+Any {\tt species\_pred\_type} with {\tt use\_tfromp = F, dpdt\_factor = 0.0} 
+and {\tt enthalpy\_pred\_type = 1} shows significant heating, although 
+the onset of the heating is delayed in {\tt species\_pred\_type = 2,3} (see 
+Figure \ref{fig:compare_tF_d0_h1_s123}). Only 
+{\tt species\_pred\_type = 1} gives good agreement between {\tt tfromh} and 
+{\tt tfromp}.
+
+Comparing {\tt cflfac = 0.7} and {\tt cflfac = 0.1} with 
+{\tt use\_tfromp = F, dpdt\_factor = 0.0, species\_pred\_type = 2} and 
+{\tt enthalpy\_pred\_type = 4} shows good agreement overall (see Figure 
+\ref{fig:tfrompF_cfl_1vs7}).
+
+% compare tfromp=F%, dpdt=0, h=1, spec=1,2,3
+%-----------------------------------------
+\begin{figure}[!h]
+\begin{minipage}[b]{0.5\linewidth}
+\vspace{0pt}
+\centering
+\includegraphics[height=0.3\textheight]{\pertfigpath/compare_tF_d0_h1_s123}
+\caption{Compare various {\tt species\_pred\_type} with {\tt use\_tfromp = F, 
+dpdt\_factor = 0.0, enthalpy\_pred\_type = 1}}
+\label{fig:compare_tF_d0_h1_s123}
+\end{minipage}
+\hspace{0.5cm}
+% compare tfromp=F,dpdt=0,spec=2,h=4, cfl=0.1,0.7
+%-----------------------------------------
+\begin{minipage}[b]{0.5\linewidth}
+\vspace{0pt}
+\centering
+\includegraphics[height=0.3\textheight]{\pertfigpath/compare_castro}
+\caption{Compare the {\tt castro.ppm\_type CASTRO} runs with the 
+{\tt species\_pred\_type MAESTRO} runs.}
+\label{fig:compare_castro}
+\end{minipage}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Additional Runs}
+
+\subsubsection{{\tt bds\_type = 1}}
+
+Using {\tt bds\_type = 1, use\_tfromp = F, dpdt\_factor = 0.0, 
+species\_pred\_type = 2, enthalpy\_pred\_type = 4} and 
+{\tt cflfac = 0.7} seems to cool off faster, perhaps due to less mixing. 
+There is also no momentum filaments in the bottom of the domain.
+
+\subsubsection{{\tt evolve\_base\_state = F}}
+
+Using {\tt evolve\_base\_state = F, use\_tfromp = F, dpdt\_factor = 0.0, 
+species\_pred\_type = 2} and {\tt enthalpy\_pred\_type = 4} seems to agree 
+well with the normal {\tt evolve\_base\_state = T} run.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{{\tt toy\_convect} in {\tt CASTRO}}
+
+{\tt toy\_convect} was also run using {\tt CASTRO} with 
+{\tt castro.ppm\_type = 0,1}. These runs show temperatures that cool 
+off rather than increase (see Figure \ref{fig:compare_castro}) which 
+suggests using {\tt species\_pred\_type = 2,3} instead of 
+{\tt species\_pred\_type = 1}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Recommendations}
+
+All of these runs suggest that running under {\tt species\_pred\_type = 
+2 or 3, enthalpy\_pred\_type = 3 or 4} with either {{\tt use\_tfromp = F} and 
+{\tt dpdt\_factor = 0.0}} or {\tt use\_tfromp = T} gives the most 
+consistent results.

### 2012-02-08 docs/SDC/sdc.tex
more sensible step labeling

@@
 
-{\bf Step 1:} ({\it Advance thermodynamic variables}) In this step we integrate $(\rho, \rho X_k, \rho h)$ over the full time step.  We use a predictor-corrector approach, as explained below.\\
+The time-advancement is divided into three major steps.
+The first step is the predictor, where we integrate the thermodynamic variables over the full time step.
+The second step is corrector, where we use the results from the predictor to perform a more accurate temporal integration of the thermodynamic variables.
+The third step is the velocity and dynamic pressure update.
 
-{\bf Step 1A:} ({\it Predictor}) In the predictor, we compute an estimate of $(\rho, \rho X_k, \rho h)$ at the new time level to obtain time-advanced values of $S$, $\beta_0$, and transport coefficients that will be used in the corrector.
-The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^n$ are computed from the the thermodynamic variables at $t^n$.\\
+{\bf Step 1:} ({\it Predictor}) In this step, we integrate $(\rho, \rho X_k, \rho h)$ over the full time step.
+The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^n$ are computed from the the thermodynamic variables at $t^n$.
+This step is divided into several sub-steps:\\
 
-{\bf Step 1A-i:} ({\it Compute advection velocities}) Use $\Ub^n$ and a second-order Godunov method to compute time-centered edge velocities, $\uadvsdcstar$, with a lagged pressure gradient and explicit buoyancy forcing terms.
+{\bf Step 1A:} ({\it Compute advection velocities}) Use $\Ub^n$ and a second-order Godunov method to compute time-centered edge velocities, $\uadvsdcstar$, with a time-lagged pressure gradient and explicit buoyancy forcing terms.
 The $\star$ superscript indicates that this field does not satisfy the divergence constrain.
@@
 
-{\bf Step 1A-ii:} ({\it Compute advective fluxes}) Use $\uadvsdcpred$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,{\rm pred}}$, with time-lagged reactions, explicitly evaluated diffusion, and a time-centered thermodynamic pressure advective derivative as source terms.\\
+{\bf Step 1B:} ({\it Compute advective fluxes}) Use $\uadvsdcpred$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,{\rm pred}}$, with time-lagged reactions, explicitly evaluated diffusion, and a time-centered thermodynamic pressure advective derivative as source terms.\\
 
-{\bf Step 1A-iii:} ({\it Compute diffusive fluxes}) Begin by compute a preliminary time-advanced density using
+{\bf Step 1C:} ({\it Compute diffusive fluxes}) Begin by compute a preliminary time-advanced density using
 \begin{equation}
@@
 
-{\bf Step 1A-iv:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}^{\rm pred}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{R,\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{\rm pred}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{R,\rho h}^{\rm lagged}$ term.
+{\bf Step 1D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}^{\rm pred}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{R,\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{\rm pred}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{R,\rho h}^{\rm lagged}$ term.
 Use VODE to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,{\rm pred}}$ using the piecewise-constant advection and diffusion source terms:
@@
 
-{\bf Step 1B:} ({\it Corrector}) In the corrector, we use the time-advanced state from the predictor to perform a more accurate integration of the thermodynamic variables.
-The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^{n+1,{\rm pred}}$ are computed from $(\rho,\rho X_k,\rho h)^{n+1,{\rm pred}}$.\\
+{\bf Step 2:} ({\it Corrector}) In the corrector, we use the time-advanced state from the predictor to perform a more accurate integration of the thermodynamic variables.
+The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^{n+1,{\rm pred}}$ are computed from $(\rho,\rho X_k,\rho h)^{n+1,{\rm pred}}$.
+This step is divided into several sub-steps:\\\\
 
-{\bf Step 1B-i:} ({\it Compute advection velocities}) First, compute $S^{n+\myhalf}$ and $\beta_0^{n+\myhalf}$ using
+{\bf Step 2A:} ({\it Compute advection velocities}) First, compute $S^{n+\myhalf}$ and $\beta_0^{n+\myhalf}$ using
 \begin{equation}
@@
 
-{\bf Step 1B-ii:} ({\it Compute advective fluxes}) Use $\uadvsdc$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf}$, with iteratively-lagged reactions, explicitly evaluated diffusion, and a time-centered thermodynamic pressure advective derivative as source terms.\\
+{\bf Step 2B:} ({\it Compute advective fluxes}) Use $\uadvsdc$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf}$, with iteratively-lagged reactions, explicitly evaluated diffusion, and a time-centered thermodynamic pressure advective derivative as source terms.\\
 
-{\bf Step 1B-iii:} ({\it Compute diffusive fluxes}) Begin by compute a preliminary time-advanced density using
+{\bf Step 2C:} ({\it Compute diffusive fluxes}) Begin by compute a preliminary time-advanced density using
 \begin{equation}
@@
 
-{\bf Step 1B-iv:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{R,\rho X_k}^{\rm pred}$ term, and define $Q_{\rho h}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{R,\rho h}^{\rm pred}$ term.
+{\bf Step 2D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{R,\rho X_k}^{\rm pred}$ term, and define $Q_{\rho h}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{R,\rho h}^{\rm pred}$ term.
 Use VODE to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1}$ using the piecewise-constant advection and diffusion source terms:
@@
 
-{\bf Step 2:} ({\it Advance velocity and dynamic pressure})\\
+{\bf Step 3:} ({\it Advance velocity and dynamic pressure})\\

### 2012-02-08 docs/architecture/architecture.tex
some fixes from Adam Jacobs

@@
 source files (and not simply a custom version of a file found
-elsewhere) needs to be listed in a file call {\tt GPackage.mak}
+elsewhere) needs to be listed in a file called {\tt GPackage.mak}
 in the problem-directory.
@@
 value for the runtime parameter.  Comments are indicated by a `{\tt
-  \#}' character and a used to produce documentation about the
+  \#}' character and are used to produce documentation about the
 available runtime parameters.  For the documentation, runtime parameters are grouped together
@@
 
-Problem-specify runtime parameters should be defined in the
+Problem-specific runtime parameters should be defined in the
 problem-directory in a file called {\tt
@@
 for \maestro.  There are two types of routines here, the first type
-take an existing 1D initial model produced somewhere else (e.g.\ a
-1D stellar evolution code), and and map it onto a uniform grid, at
+modify an existing 1D initial model produced somewhere else (e.g.\ a
+1D stellar evolution code), and map it onto a uniform grid, at
 the desired resolution, using the equation of state in \maestro, and
@@
 giving the coordinate and data values at that coordinate.  The initial
-model should contain the same species data (as mass fractions) as
+model should contain the same species data (in the form of mass fractions) as
 defined in the {\tt network} module used by the \maestro\ problem.

### 2012-02-08 docs/getting_started/getting_started.tex
some fixes from Adam Jacobs

@@
 
-If you shell is {\tt Bash}, add
+If your shell is {\tt Bash}, add
 \begin{verbatim}
@@
 \maestro\ is run simply as:
-runtime parameters.  The code is run simply as:
 \begin{verbatim}
@@
 As the code runs, it will output both plotfiles and checkpoints as
-well as one of more text diagnostic files ({\tt maestro\_diag.out} by
+well as one or more text diagnostic files ({\tt maestro\_diag.out} by
 default) with integral or extrema information (like maximum Mach
@@
 
-Visualization and analysis are done with on the plotfiles.  A
+Visualization and analysis are done on the plotfiles.  A
 number of in-house and externally developed tools can work 
@@
 \visit\ is a powerful, DOE-supported visualization tool for 2- and 3D
-datasets.  It can do contouring, volume rendering, streamlines, ...\
+datasets.  It can do contouring, volume rendering, streamlines, ...\, ,
 directly from \boxlib\ plotfiles (to open a plotfile in \visit, point to
@@
 {\tt test2} places 3 hots spots in a plane-parallel atmosphere.
-Burning makes these buoyant bubbles which role up.  This problem was
+Burning makes these bubbles buoyant, and then roll up.  This problem was
 used in \cite{lowMach3} to compare with compressible solvers.

### 2012-02-08 docs/introduction/introduction.tex
list of tables
update papers
update referneces
some fixes from Adam Jacobs

@@
   M. Zingale, A. S. Almgren, J. B. Bell, C. M. Malone, A. Nonaka, \& S. E. Woosley 2011,
-  submitted to ApJ (henceforth paper V)
+  740, 8 (henceforth paper V)
 \item {\em High-Resolution Simulations of Convection Preceding Ignition in
   Type Ia Supernovae Using Adaptive Mesh Refinement,}
-  A. Nonaka, M. Zingale, A. J. Aspden, A. S. Almgren, J. B. Bell, \& S. E. Woosley
+  A. Nonaka, M. Zingale, A. J. Aspden, A. S. Almgren, J. B. Bell, \& S. E. Woosley 2012, 
+  745, 73
 \end{itemize}
@@
 \cite{leveque}.  The timestep for these methods is limited by the time
-it takes for maximum characteristic speed to traverse one grid cell.
+it takes for the maximum characteristic speed to traverse one grid cell.
 For very subsonic flows, this means that the timestep is dominated by
@@
 \begin{longtable}{|l|p{4.0in}|l|}
-\caption[definition of symbols.]{definition of symbols.} \label{table:sym} \\
+\caption[Definition of symbols.]{Definition of symbols.} \label{table:sym} \\
 %

### 2012-02-10 docs/SDC/sdc.tex
some updates

@@
 \section{SDC Overview}
+
+Spectral deferred corrections (SDC) is an iterative scheme used to integrate
+the thermodynamic variables, $(\rho X_k,\rho h)$, over a time step.  It has
+been shown to be more accurate and efficient than Strang splitting in a
+terrestrial, non-stratified, low Mach number reacting flow solver \cite{Non11},
+so we would like to develop an SDC version of \maestro.
 
@@
     -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} 
-    + \rho\Hnuc ,\label{eq:enthalpy}
+    + \rho\Hnuc + \nabla\cdot k_{\rm th}\nabla T,\label{eq:enthalpy}
 \end{eqnarray}
-together with a constraint equation.
+together with base state evolution equations and a constraint equation.
+By default, \maestro\ advances the thermodynamic variables by coupling
+the different physical processes together (advection, diffusion, reactions) using
+Strang splitting.  Specifically, we integrate the reaction terms 
+over half a time step ignoring contributions from advection and diffusion,
+then from this intermediate state we integrate the advection and diffusion terms over
+a full time step (while ignoring reactions), and finally
+integrate the reactions over a half time step (ignoring advection and diffusion).
+For problems where
+the reactions and/or diffusion greatly alter the energy balance or composition 
+as compared to advection, this operator splitting approach can lead to large 
+splitting errors and highly inaccurate solutions.
+This issue is can be particularly exasperating 
+for low Mach number methods that can take large advection-based time steps.
 
-By default, \maestro\ uses an operator-split approach to coupling
-the different physical processes together (hydro, reactions, diffusion).
-With operator-splitting, we
-react for a half timestep, then perform the hydrodynamics and thermal
-diffusion on this intermediate state, and finally react for the
-remaining half timestep to advance the solution.  For problems where
-the reactions greatly alter the energy balance or composition, this
-operator splitting approach can lead to inaccuracies in the solution.
-This can be more of an issue for low Mach number methods that can take
-large hydrodynamic time steps.
-
-An alternate approach to coupling is spectral deferred corrections
-(SDC). SDC is an iterative approach to couple the various operations
-together, with each operator seeing an iteratively lagged
-approximation of the other operations as a source term.  The SDC
+An alternate approach to advancing the thermodynamic variables is SDC. 
+SDC is an iterative approach to couple the various processes
+together, with each process seeing an 
+approximation of the other processes as a source term.  The SDC
 algorithm converges to an integral representation of the solution in
@@
 
-As a first attempt, we will work on coupling hydro and reactions only
-via SDC, with a time-independent base state.
+As a first attempt, we will work on coupling advection and reactions only
+via SDC, with a base state that is fixed in time, but not space.
 
-\section{Strang-split Overview}
+\section{Strang-Splitting Without Thermal Diffusion or Base State Evolution}
 
-In the Strang-split version of \maestro, the reaction and advection
+In the Strang splitting version of \maestro, the reaction and advection
 processes operate independent of one-another.  The species and
-enthalpy equations are updated through $\Delta t$ using the following
+enthalpy equations are integrated over $\Delta t$ using the following
 sequence:
@@
 
-\item React for $\Delta t/2$
+\item React for $\Delta t/2$:
 
-In the Strang-split version of \maestro, the reaction network solves just
+In the Strang splitting version of \maestro, the reaction network solves just
 the reaction portion of the species evolution equations along with a
-temperature evolution equation.  This is done using a standard ODE solver
-(like {\tt VODE}) on the system:
+temperature evolution equation.  This is done using a standard stiff ODE solver
+package (like {\tt VODE}) on the system:
 \begin{eqnarray}
-\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T) \\
+\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T), \\
 \frac{dT}{dt}   &=& 
@@
 start of the integration and held constant thereafter.
-
-We update the enthalpy at the end of the integration as:
+The density remains constant during this step, i.e., 
+$\rho^\mathrm{new} = \rho^\mathrm{old}$, and we
+update the enthalpy at the end of the integration as:
 \begin{equation}
-h^\mathrm{new} = h^\mathrm{old} + \frac{\Delta t}{2} \Hnuc
+h^\mathrm{new} = h^\mathrm{old} + \frac{\Delta t}{2} \Hnuc.
 \end{equation}
 
-\item Advect for $\Delta t$
+\item Advect for $\Delta t$:
 
-Here we solve
+Beginning with the results from the previous reaction half time step, we integrate
+that state using the equations
 \begin{eqnarray}
 \frac{\partial(\rho X_k)}{\partial t} &=& 
-    -\nabla\cdot(\rho X_k\Ub) \\
+    -\nabla\cdot(\rho X_k\Ub), \\
 \frac{\partial(\rho h)}{\partial t} &=& 
-    -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} 
+    -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt}.
 \end{eqnarray}
-
-Note that no explicit reaction terms appear here.  Since the advection
+Note that no reaction terms appear here.  Since the advection
 takes place using the state updated from the reaction step, the effect
@@
 
-\item React for $\Delta t/2$
+\item React for $\Delta t/2$:
 
@@
 
+Note that \maestro\ uses a predictor-corrector approach.  After integrating $(\rho X_k,\rho h)$ over
+the time step, we use this time-advanced state to get a better estimate of a time-centered $\beta_0$
+and $S$.  We re-compute the advective velocities using an updated divergence constraint and repeat 
+the thermodynamic variable advance.
+
 \section{SDC Without Thermal Diffusion or Base State Evolution}
-In the SDC version, the VODE integration at the end of an SDC
+In the SDC version, the {\tt VODE} integration at the end of an SDC
 iteration is responsible for updating all the thermodynamic quantities
-\{$\rho$, $(\rho h)$, and $(\rho X_k$)\}, including both the advective
-fluxes (incorporated via a piecewise constant advective flux source
-term) and the reactions.  This provides a much stronger coupling between
+including both the advection (incorporated via a piecewise constant advective 
+flux divergence source term) and the reactions.  This provides a much stronger coupling between
 the physical processes.  In particular, our system now looks like:
 \begin{eqnarray}
-\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + A_{\rho X_k} \label{eq:sdc:rhoX} \\
-\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + A_{\rho h} \label{eq:sdc:rhoh}
+\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + \underbrace{\nabla\cdot(\rho X_k\Ub)}_{A_{\rho X_k}}\label{eq:sdc:rhoX} \\
+\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + \underbrace{\nabla\cdot(\rho h\Ub)}_{A_{\rho h}} \label{eq:sdc:rhoh}
 \end{eqnarray}
@@
 A_{\rho X_k} &=&  \left [- \nabla \cdot (\rho X_k \Ub) \right ]^{n+1/2} \\
-A_{\rho h}   &=&  \left [- \nabla \cdot (\rho h U) + \frac{Dp_0}{Dt} \right ]^{n+1/2}
+A_{\rho h}   &=&  \left [- \nabla \cdot (\rho h \Ub) + \frac{Dp_0}{Dt} \right ]^{n+1/2}
 \end{eqnarray}
 The construction of the interface states used in the advective terms
-uses an iteratively-lagged approximation to the reaction source terms
-($I_{\rho X_k}$ and $I_{\rho h}$, see below) in the interface
+uses either a time-lagged or iteratively-lagged approximation to the reaction
+terms ($I_{\rho X_k}$ and $I_{\rho h}$, see below) as a source term in the interface
 prediction.  This explicitly couples the reaction process to the
@@
 \end{equation}
-Note that in constrast to the Strang-split version, here we call the EOS
-every time we enter the RHS routine.
+Note that in contrast to the Strang splitting version, here we call the EOS
+every time we enter the RHS routine, but here we call the EOS to compute temperature
+rather than thermodynamic coefficients.
 
 Finally we integrate the ODE system (Eqs.~\ref{eq:sdc:rhoX} and \ref{eq:sdc:rhoh}).
-At the end of the integration, we define $I_{\rho X_k}$ and $I_{\rho
-  h}$.  The actual form of these depends on what we are predicting.
-We note that we only need $I_{\rho X_k}$ and $I_{\rho h}$ for the
-prediction of the interface states, not the final conservative update.
-That is because all we need from the advection solver is the
-approximation to $A_{\rho X_k}$ and $A_{\rho h}$, not the final
+At the end of the integration, we define $I_{\rho X_k}$ and $I_{\rho h}$.  The actual 
+form of these depends on what quantities we predict to edges during 
+the construction of the advective fluxes.
+Note that we only need $I_{\rho X_k}$ and $I_{\rho h}$ for the
+prediction of the interface states, and not the {\tt VODE} integration.
+This is because all we need from the advection solver is the
+approximation to $A_{\rho X_k}$ and $A_{\rho h}$ and not the final
 updated state.
@@
 of $T$, and the need to derive the
-temperature from the input state for each RHS evaluation by {\tt
-  VODE}.
+temperature from the input state for each RHS evaluation by {\tt VODE}.
 
@@
 the velocities at all.  That update is handled in the same fashion 
-as the Strang-split version of the code.
+as the Strang splitting version of the code.
 
@@
     + \rho\Hnuc
-    + \underbrace{\nabla\cdot\frac{\kth}{c_p}\nabla h - \sum_k\nabla\cdot\frac{\xi_k k_{\rm th}}{c_p}\nabla X_k - \nabla\cdot\frac{h_p k_{\rm th}}{c_p}\nabla p_0}_{\nabla\cdot k_{\rm th}\nabla T}.\label{eq:sdc enthalpy}
+    + \underbrace{\nabla\cdot\frac{\kth}{c_p}\nabla h - \sum_k\nabla\cdot\frac{\xi_k k_{\rm th}}{c_p}\nabla X_k - \nabla\cdot\frac{h_p k_{\rm th}}{c_p}\nabla p_0}_{\nabla\cdot k_{\rm th}\nabla T}.\nonumber\\
+\label{eq:sdc enthalpy}
 \end{eqnarray}
@@
 {\bf Step 1D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}^{\rm pred}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{R,\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{\rm pred}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{R,\rho h}^{\rm lagged}$ term.
-Use VODE to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,{\rm pred}}$ using the piecewise-constant advection and diffusion source terms:
+Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,{\rm pred}}$ using the piecewise-constant advection and diffusion source terms:
 \begin{eqnarray}
@@
 {\bf Step 2D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{R,\rho X_k}^{\rm pred}$ term, and define $Q_{\rho h}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{R,\rho h}^{\rm pred}$ term.
-Use VODE to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1}$ using the piecewise-constant advection and diffusion source terms:
+Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1}$ using the piecewise-constant advection and diffusion source terms:
 \begin{eqnarray}

### 2012-02-10 docs/SDC/sdc.tex
my guess at I_{X_k}

@@
             (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
+\end{equation}
+
+(Andy's Idea) Define $I_{X_k}$ using
+\begin{equation}
+I_{X_k} = \frac{X_k^\mathrm{new} - X_k^\mathrm{old}}{\Delta t} - A_{X_k},
+\end{equation}
+where we first define a state that has only been updated with advection:
+\begin{equation}
+\frac{(\rho X_k)^{(1)} - (\rho X_k)^\mathrm{old}}{\Delta t} = A_{\rho X_k},
+\end{equation}
+and then define the species mass fractions,
+\begin{equation}
+X_k^{(1)} = (\rho X_k)^{(1)} / \sum_k (\rho X_k)^{(1)}, \quad 
+X_k^\mathrm{old} = (\rho X_k)^\mathrm{old} / \sum_k (\rho X_k)^\mathrm{old}, \quad 
+X_k^\mathrm{new} = (\rho X_k)^\mathrm{new} / \sum_k (\rho X_k)^\mathrm{new},
+\end{equation}
+and finally define $A_{X_k}$ using
+\begin{equation}
+\frac{X^{(1)} - X^\mathrm{old}}{\Delta t}= A_{X_k}.
 \end{equation}

### 2012-02-13 docs/SDC/sdc.tex
new version of reaction term for predicting h

@@
 \subsubsection{Species Source Terms.}
-For the species prediction, the form of $I_{\rho X_k}$ depends on
-{\tt species\_pred\_type} (see \S \ref{sec:pred:density}):  
+For the species prediction, the form of $I$ depends on
+{\tt species\_pred\_type} (see \S \ref{sec:pred:density}).
+We note that there is no $I$ term for $\rho$ or $\rho'$ prediction, since
+the density evolution equation does not have a reaction source term.
 \begin{itemize}
 \item {\tt species\_pred\_type} = 1 (predict $\rho'$ and $X_k$)
-%
 \begin{equation}
@@
       \frac{(\rho X_k)^\mathrm{new} - 
-            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
+            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ].
 \end{equation}
-
 (Andy's Idea) Define $I_{X_k}$ using
@@
 \end{equation}
-
 \item {\tt species\_pred\_type} = 2 (predict $\rho'$ and $(\rho X_k)$)
-%
 \begin{equation}
-I_{\rho X_k} = \frac{(\rho X_k)^\mathrm{new} - (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k} 
+I_{\rho X_k} = \frac{(\rho X_k)^\mathrm{new} - (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}.
 \label{eq:sdc:Irhoo}
 \end{equation}
+\item {\tt species\_pred\_type} = 3 (predict $\rho$ and $X_k$)
 
-\item {\tt species\_pred\_type} = 3 (predict $\rho$ and $X_k$)
-%
-\begin{equation}
-I_{X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
-      \frac{(\rho X_k)^\mathrm{new} - 
-            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
-\end{equation}
-
+See {\tt species\_pred\_type} = 1 for the form of $I_{X_k}$.
 \end{itemize}
-
-We note that there is no $I$ term for $\rho$ or $\rho'$ prediction, since
-the density evolution equation does not have a reaction source term.
 
@@
 \begin{itemize}
-\item {\tt enthalpy\_pred\_type} = {\tt predict\_rhoprime}
-
-Not implemented.
+\item {\tt enthalpy\_pred\_type} = {\tt predict\_rhohprime} (not implemented yet)
 
 Here we need an $I_{\rho h}$ term for the $(\rho h)'$ evolution
-equation (see Eq.~\ref{rhohprime equation}).  Defining the source
-terms for this requires a base state enthalpy that knows about the
-source terms---this is currently not implemented.
+equation (see Eq.~\ref{rhohprime equation}).  In this case we will use
+$I_{(\rho h)'} = I_{\rho h}$
 
+In the future, when we enable base state evolution, the base state enthalpy
+evolution equation may need to know about the $I_{\rho h}$ source term.
 In particular, should $(\rho h)_0$ see a $\overline{(\rho \Hnuc)}$ term?
@@
     \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h} \right ]
+\end{equation}
+(Andy's Idea) Form $I_h$ in the same way we would form $I_{X_k}$ from above:
+\begin{equation}
+I_h = \frac{h^\mathrm{new} - h^\mathrm{old}}{\Delta t} - A_h,
+\end{equation}
+where we first define
+\begin{equation}
+\frac{(\rho h)^{(1)} - (\rho h)^\mathrm{old}}{\Delta t} = A_{\rho h},
+\end{equation}
+and then define $h$,
+\begin{equation}
+h^{(1)} = (\rho h)^{(1)} / \sum_k(\rho X_k)^{(1)}, \quad h^\mathrm{old} = (\rho h)^\mathrm{old} / \sum_k(\rho X_k)^\mathrm{old}, \quad h^\mathrm{new} = (\rho h)^\mathrm{new} / \sum_k(\rho X_k)^\mathrm{new},
+\end{equation}
+and finally define $A_h$ using
+\begin{equation}
+I_h = \frac{h^{(1)} - h^\mathrm{old}}{\Delta t} = A_h.
 \end{equation}

### 2012-02-13 docs/SDC/sdc.tex
more sdc source terms

@@
 \item {\tt species\_pred\_type} = 1 (predict $\rho'$ and $X_k$) 
+or 3 (predict $\rho$ and $X_k$)
+
 \begin{equation}
@@
 \end{equation}
-\item {\tt species\_pred\_type} = 3 (predict $\rho$ and $X_k$)
-
-See {\tt species\_pred\_type} = 1 for the form of $I_{X_k}$.
 \end{itemize}
@@
 
-Here we need an $I_{\rho h}$ term for the $(\rho h)'$ evolution
+(Andy's Idea) Here we need an $I_{\rho h}$ term for the $(\rho h)'$ evolution
 equation (see Eq.~\ref{rhohprime equation}).  In this case we will use
-$I_{(\rho h)'} = I_{\rho h}$
+$I_{(\rho h)'} = I_{\rho h}$.  Since we are not evolving the base state, the PDE
+for $(\rho h)_0$ is simply $\partial(\rho h)_0/\partial t = 0$, and thus the
+evolution equation for $(\rho h)'$ is the same as the evolution equation
+for $\rho h$.
 
@@
 
+(Andy's Idea) The idea is to advance the species and enthalpy with advection
+terms only, and compute the resulting temperature, $T^{(1)}$.  Compare that temperature
+with the final temperature computed by the SDC VODE call.  The difference
+between these values is $I_T$.
+\begin{equation}
+I_T = \frac{T^\mathrm{new} - T^\mathrm{old}}{\Delta t} - A_T,
+\end{equation}
+with $A_T$ given by
+\begin{equation}
+\frac{T^{(1)} - T^\mathrm{old}}{\Delta t} = A_T,
+\end{equation}
+and $T^{(1)}$ computed using the equation of state from $\rho^{(1)}, X_k^{(1)}$, 
+and $h^{(1)}$ (or $p_0$, if {\tt use\_tfromp=T}).
 \end{itemize}

### 2012-02-13 docs/SDC/sdc.tex
update the text flow chart

@@
 
-\subsubsection{Species Source Terms.}
+\subsection{Species Source Terms.}
 For the species prediction, the form of $I$ depends on
@@
 
-\subsubsection{Enthalpy Source Terms.}
+\subsection{Enthalpy Source Terms.}
 The obvious analog to the above form for $I_{\rho X_k}$ is to define
@@
 and $T^{(1)}$ computed using the equation of state from $\rho^{(1)}, X_k^{(1)}$, 
-and $h^{(1)}$ (or $p_0$, if {\tt use\_tfromp=T}).
+and $h^{(1)}$ (or $p_0$, if {\tt use\_tfromp = T}).
 \end{itemize}
 
-\subsubsection{Implementation}
+\subsection{Implementation}
 
@@
 
-\section{Summary of Changes}
+\subsection{Summary of Changes}
 The major changes from the non-SDC-enabled burners is the addition of
@@
 The time-advancement is divided into three major steps.
-The first step is the predictor, where we integrate the thermodynamic variables over the full time step.
+The first step is the predictor, where we integrate the thermodynamic variables, $(\rho,\rho X_k,\rho h)$, over the full time step.
 The second step is corrector, where we use the results from the predictor to perform a more accurate temporal integration of the thermodynamic variables.
-The third step is the velocity and dynamic pressure update.
+The third step is the velocity and dynamic pressure update.\\
 
@@
 
-{\bf Step 1A:} ({\it Compute advection velocities}) Use $\Ub^n$ and a second-order Godunov method to compute time-centered edge velocities, $\uadvsdcstar$, with a time-lagged pressure gradient and explicit buoyancy forcing terms.
-The $\star$ superscript indicates that this field does not satisfy the divergence constrain.
+{\bf Step 1A:} ({\it Compute advection velocities}) Use $\Ub^n$ and a second-order Godunov method to compute time-centered edge velocities, $\uadvsdcstar$, with time-lagged dynamic pressure and explicit buoyancy as forcing terms.
+The $\star$ superscript indicates that this field does not satisfy the divergence constraint.
 Compute $S^{n+\myhalf,{\rm pred}}$ by extrapolating in time,
@@
 
-{\bf Step 1B:} ({\it Compute advective fluxes}) Use $\uadvsdcpred$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,{\rm pred}}$, with time-lagged reactions, explicitly evaluated diffusion, and a time-centered thermodynamic pressure advective derivative as source terms.\\
+{\bf Step 1B:} ({\it Compute advective fluxes}) Use $\uadvsdcpred$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,{\rm pred}}$, with time-lagged reactions, explicit diffusion, and time-centered thermodynamic pressure as source terms.\\
 
@@
 \begin{equation}
-\frac{\rho^{n+1,{\rm pred}}X_{k,{\rm AD}}^{n+1,{\rm pred}} - (\rho X_k)^n}{\Delta t} = -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right] + I_{R,\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
+\frac{\rho^{n+1,{\rm pred}}X_{k,{\rm AD}}^{n+1,{\rm pred}} - (\rho X_k)^n}{\Delta t} = -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right] + I_{\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
 \end{equation}
@@
 \begin{eqnarray}
-\frac{\rho^{n+1,{\rm pred}}h^{n+1,{\rm pred}}_{\rm AD} - (\rho h)^n}{\Delta t} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right] + \uadvsdcpred\cdot\nabla p_0 + I_{R,\rho h}^{\rm lagged}\nonumber\\
+\frac{\rho^{n+1,{\rm pred}}h^{n+1,{\rm pred}}_{\rm AD} - (\rho h)^n}{\Delta t} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right] + \uadvsdcpred\cdot\nabla p_0 + I_{\rho h}^{\rm lagged}\nonumber\\
 && + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n + \nabla\cdot\frac{\kth^n}{c_p^n}\nabla h_{\rm AD}^{n+1,{\rm pred}}\right)\nonumber\\
 && + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,{\rm pred}}\right)\nonumber\\
-&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right)\label{eq:sdc enthalpy 2}
+&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right).\label{eq:sdc enthalpy 2}
 \end{eqnarray}
 
-{\bf Step 1D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}^{\rm pred}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{R,\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{\rm pred}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{R,\rho h}^{\rm lagged}$ term.
+{\bf Step 1D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}^{\rm pred}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{\rm pred}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{\rho h}^{\rm lagged}$ term.
 Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,{\rm pred}}$ using the piecewise-constant advection and diffusion source terms:
@@
 \end{eqnarray}
+At this point we can define $I_{\rho X_k}^{\rm pred}$ and $I_{\rho h}^{\rm pred}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the corrector step.\\
 
@@
 \begin{equation}
-S^{n+\myhalf} = \frac{S^n + S^{n+1,{\rm pred}}}{2}, \qquad \beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1,{\rm pred}}}{2}
+S^{n+\myhalf} = \frac{S^n + S^{n+1,{\rm pred}}}{2}, \qquad \beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1,{\rm pred}}}{2}.
 \end{equation}
@@
 
-{\bf Step 2B:} ({\it Compute advective fluxes}) Use $\uadvsdc$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf}$, with iteratively-lagged reactions, explicitly evaluated diffusion, and a time-centered thermodynamic pressure advective derivative as source terms.\\
+{\bf Step 2B:} ({\it Compute advective fluxes}) Use $\uadvsdc$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf}$, with iteratively-lagged reactions, explicit diffusion, and time-centered thermodynamic pressure as source terms.\\
 
-{\bf Step 2C:} ({\it Compute diffusive fluxes}) Begin by compute a preliminary time-advanced density using
+{\bf Step 2C:} ({\it Compute diffusive fluxes}) Begin by computing the time-advanced density using
 \begin{equation}
@@
 \begin{equation}
-\frac{\rho^{n+1}X_{k,{\rm AD}}^{n+1} - (\rho X_k)^n}{\Delta t} = -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf}\uadvsdc\right] + I_{R,\rho X_k}^{\rm pred}.\label{eq:sdc species 3}
+\frac{\rho^{n+1}X_{k,{\rm AD}}^{n+1} - (\rho X_k)^n}{\Delta t} = -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf}\uadvsdc\right] + I_{\rho X_k}^{\rm pred}.\label{eq:sdc species 3}
 \end{equation}
@@
 \begin{eqnarray}
-\frac{\rho^{n+1}h^{n+1}_{\rm AD} - (\rho h)^n}{\Delta t} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf}\uadvsdc\right] + \uadvsdc\cdot\nabla p_0 + I_{R,\rho h}^{\rm pred}\nonumber\\
+\frac{\rho^{n+1}h^{n+1}_{\rm AD} - (\rho h)^n}{\Delta t} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf}\uadvsdc\right] + \uadvsdc\cdot\nabla p_0 + I_{\rho h}^{\rm pred}\nonumber\\
 && + \nabla\cdot\frac{\kth^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla h^{n+1}_{\rm AD} + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla h^{n+1,{\rm pred}}\right)\nonumber\\
@@
 
-{\bf Step 2D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{R,\rho X_k}^{\rm pred}$ term, and define $Q_{\rho h}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{R,\rho h}^{\rm pred}$ term.
+{\bf Step 2D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{\rho X_k}^{\rm pred}$ term, and define $Q_{\rho h}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{\rho h}^{\rm pred}$ term.
 Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1}$ using the piecewise-constant advection and diffusion source terms:
@@
 \end{eqnarray}
+At this point we can define $I_{\rho X_k}^{\rm lagged}$ and $I_{\rho h}^{\rm lagged}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the predictor in the next time step.\\
 
-{\bf Step 3:} ({\it Advance velocity and dynamic pressure})\\
+{\bf Step 3:} ({\it Advance velocity and dynamic pressure}) Same as the original \maestro\ algorithm.

### 2012-02-15 docs/FAQ/faq.tex
why is maestro "hard"

@@
 
+\item {\em Why is \maestro\ so ``hard'' to use (e.g.\ as compared to a
+  compressible code)?}
+
+There are several complexities to the algorithm that don't have
+straightforward compressible counterparts.  These mainly involve the
+role of the base state and the constraint equation.
+
+Care must be taken to setup an initial model/initial base state that
+respects the thermodynamics in \maestro\ and is in hydrostatic equilibrium.
+Best results are attained when the model is processed with the \maestro\
+EOS and reset into HSE, as is done in the {\tt initial\_model} routines.
+Because \maestro\ builds off of the base state, any flaws in that initial 
+state will influence the subsequent behavior of the algorithm.
+
+The constraint equation brings another complexity not seen in compressible
+codes---information is instantly communicated
+across the grid.  In compressible codes you can track down a problem by
+watching where it starts from and watching it move one cell per dt.  In
+\maestro\ things can go wrong in multiple places without it being obvious
+where the root problem is.
+
 \item {\em In the final projection in the algorithm, we project

### 2012-02-27 docs/unit_tests/unit_tests.tex
unit test chapter

--- 2012-01-21 docs/architecture/architecture.tex
@@
-\section{\maestro\ Basics}
+\label{chapter:unit_tests}
 
-\maestro\ models problems that are in tight hydrostatic equilibrium.
-The fluid state is decomposed into a 1D radial base state that
-describes the hydrostatic structure of the star or atmosphere, and a
-2- or 3D Cartesian full state, that captures the departures from
-hydrostatic equilibrium.  Two basic geometries are allowed.  A {\em
-  plane-parallel} geometry assumes that the domain is thin compared to
-the radius of curvature of the star, and therefore the 1D base state
-is perfectly aligned with the Cartesian state.  A {\em spherical}
-geometry is for modeling an entire star.  Here, the 1D base state is
-not aligned with the Cartesian state.  Figure~\ref{fig:base_state}
-shows these geometries.
-
-\begin{figure}[tb]
-\centering
-\includegraphics[height=2.0in]{\archfigpath/base_grid} \hspace{0.5in}
-\includegraphics[height=2.0in]{\archfigpath/base_spherical}
-\caption[\maestro\ geometries]{\label{fig:base_state} \maestro\ geometries, showing both the
-  1D base state and the full Cartesian state.  (Left) For multi-level
-  problems in planar geometry, we force a direct alignment between the
-  radial array cell centers and the Cartesian grid cell centers by
-  allowing the radial base state spacing to change with space and
-  time.  (Right) For multi-level problems in spherical geometry, since
-  there is no direct alignment between the radial array cell centers
-  and the Cartesian grid cell centers, we choose to fix the radial
-  base state spacing across levels. Figure taken
-  from~\cite{multilevel}.}
-\end{figure}
-
-\maestro\ can use adaptive mesh refinement to focus resolution on
-complex regions of flow.  For Cartesian/plane-parallel geometries, all
-cells at the same height must be at the same level of refinement.
-This restriction is to allow for the base state to directly align with
-the Cartesian state everywhere.  For spherical geometries, there is no
-such restriction (again, see Figure~\ref{fig:base_state}).
-The \maestro\ grids are managed by the \boxlib\ library, which is
-distributed separately.
-
-\section{\maestro\ Directory Structure}
-
-\begin{figure}[t]
-\centering
-\includegraphics[scale=0.65]{\archfigpath/maestro_directory2}
-\caption[\maestro\ directory structure]
-{The basic \maestro\ directory structure.}
-\end{figure}
-
-Building \maestro\ requires both the \maestro-specific source
-files (distributed in the {\tt MAESTRO/} directory), and the
-\boxlib\ library (distributed in the {\tt BoxLib/} directory.
-\boxlib\ provides both a C++ and a Fortran framework.  \maestro\
-only uses the Fortran framework.
-
-Problems piece together various \maestro\ directories, choosing a
-reaction network, equation of state, and conductivity routine to build
-an executable.  Briefly, the \maestro\ sub-directories are:
-\begin{itemize}
-\item {\tt data\_processing/}
-
-  Simple Fortran-based analysis routines (e.g.\ extract a line from a
-  multidimensional dataset) that operate on BoxLib datasets.
-
-\item {\tt extern/}
-
-  External modules, like ODE integrators, equations of state, reaction
-  networks
-
-  Important sub-directories include:
-
-  \begin{itemize}
-  \item {\tt conductivity/}
-
-    Various routines for computing the thermal conductivity used in the
-    thermal diffusion part of the algorithm.
-
-  \item {\tt constants/}
-
-    A simple module to define physical constants
-
-  \item {\tt EOS/}
-
-    A collection of various EOSes for use with the code.  The two
-    important ones for \maestro\ are {\tt helmeos/} and {\tt
-    gamma\_law\_general/}.
-
-  \item {\tt model\_parser/}
-
-    A simple Fortran module for reading in 1D initial model files.
-    This is used by the initialization routines to get the initial
-    model data.
-
-  \item {\tt networks/}
-
-    Various reaction networks for \maestro\ problems.  In addition to
-    providing a routine to evolve the nuclear species due to
-    reactions, the networks also define the species that are advected
-    by the code.
-
-  \item {\tt VODE/}
-
-    An integration package for ODEs.  At the moment, this is used for
-    integrating various reaction networks.
-
-  \end{itemize}
-
-\item {\tt MAESTRO/}
-
-  The main \maestro\ algorithm directory.  The individual problem
-  directories lie under here, and all of the main \maestro\ source
-  lives in the {\tt Source/} sub-directory.
-
-  Important directories under {\tt MAESTRO/} include:
-
-  \begin{itemize}
-
-  \item {\tt docs/}
-
-    Documentation describing the basic algorithm (including this
-    document).
-
-  \item {\tt initial\_models}
-
-    Several sub-directories containing routines to generate 1D initial
-    conditions in hydrostatic equilibrium for mapping onto the
-    \maestro\ grid.  See \S~\ref{sec:initial_models} for details.
-
-  \item {\tt Source}
-
-    The main \maestro\ source.  Here you will find the driver routine,
-    the advection routines, etc.  All \maestro\ problems will compile
-    this source.
-
-  \item problem directories: {\tt test2}, {\tt test\_spherical},
-    $\ldots$
-
-    Each problem in \maestro\ gets it own sub-directory.  The {\tt
-      GNUmakefile} in the problem directory includes the instructions
-    on how to build the executable, including what modules in {\tt
-      extern} are used.
-
-    Any file that you place in a sub-directory here takes precedence
-    over a file of the same name in {\tt MAESTRO/}.  This allows
-    problems to have custom versions of the main \maestro\ routines
-    (e.g.\ initial conditions via {\tt initdata.f90}.  See
-    \S~\ref{sec:makefile} for details on the build system).
-
-  \end{itemize}
-
-\end{itemize}
-
-\begin{figure}[t]
-\centering
-\includegraphics[scale=0.65]{\archfigpath/boxlib_directory2}
-\caption[\boxlib\ directory structure]
-{The basic \boxlib\ directory structure.  The directories used
-by \maestro\ are indicated in red.}
-\end{figure}
-
-The \boxlib\ directories that \maestro\ uses are:
-\begin{itemize}
-\item {Src/}
-
-  The main \boxlib\ source directory.  In \maestro, we only use the
-  Fortran source files.  The core directories are:
-
-  \begin{itemize}
-
-  \item {\tt F\_BaseLib/}
-
-    The Fortran \boxlib\ files.  This is a library for describing
-    meshes consisting of a union of boxes.  The \boxlib\ modules
-    define the basic datatypes used in \maestro.  \boxlib\ also
-    provides the routines that handle the parallelization and I/O.
-
-  \item {\tt LinearSolvers/}
-
-    The \boxlib\ linear solvers---these are used to solve elliptic
-    problems in the \maestro\ algorithm.
-
-    \begin{itemize}
-
-    \item {\tt F\_MG}
-
-      The Fortran multigrid solver, with support for both
-      cell-centered and node-centered data.
-
-    \end{itemize}
-
-  \end{itemize}
-
-\item {\tt Tools/}
-
-  Various tools used for building \boxlib\ applications.  Here we use:
-
-  \begin{itemize}
-
-  \item {F\_mk/}
-
-    The generic Makefiles that store the compilation flags for various
-    platforms.  Platform/compiler-specific options are stored in the
-    {\tt comps/} sub-directory.
-
-  \item {\tt F\_scripts/}
-
-    Some simple scripts that are useful for building, running,
-    maintaining \maestro.
-
-  \end{itemize}
-
-\end{itemize}
-
-\section{Unit Tests}
-
-\label{sec:unit_tests}
-
-In addition to the problems described above which use the full
+In addition to the \maestro science problems, which use the full
 capabilities of \maestro, there are a number of unit tests that
-exercise only specific components of the \maestro\ solvers.  The
+exercise only specific components of the \maestro\ solvers.  These
 tests have their own drivers (a custom {\tt varden.f90}) that
@@
 
-\begin{itemize}
-\item {\tt test\_advect} \\[-3mm]
+\section {\tt test\_advect}
 
@@
 
-\item {\tt test\_average} \\[-3mm]
+\section {\tt test\_average}
 
@@
 
-\item {\tt test\_basestate} \\[-3mm]
+\section {\tt test\_basestate}
 
@@
 
-\item {\tt test\_diffusion}
+\section {\tt test\_diffusion}
 
@@
 
-\item {\tt test\_particles}
+\section {\tt test\_eos}
+
+  This test sets up a 3-d cube with $\rho$ varying on one axis, $T$ on
+  another, and the composition on the third.  The EOS is then called
+  in every zone, doing $(\rho, T) \rightarrow  p, h, s, e$  and stores those
+  quantities.  Then it does each of the different EOS types to recover
+  either $T$ or $\rho$ (depending on the type), and stores the new $T$ (or
+  $\rho$) and the relative error with the original value.  A plotfile is
+  stored holding the results and errors.  This allows us to determine
+  whether the EOS inversion routines are working right.
+
+\section {\tt test\_particles}
 
@@
 
-\end{itemize}
-
-\section{Adding A New Problem}
-\label{sec:adding_problems}
-
-Different \maestro\ problems are defined in sub-directories under
-{\tt MAESTRO/}.  To add a problem, start by creating a new
-sub-directory---this is where you will compile your problem and
-store all the problem-specific files.
-
-The minimum requirement to
-define a new problem would be a {\tt GNUmakefile} which describes how
-to build the application and an input file which lists the runtime
-parameters.  The problem-specific executable is built in the problem
-directory by typing {\tt make}.  Source files are found automatically
-by searching the directories listed in the {\tt GNUmakefile}.
-Customized versions of any source files placed in the problem-directory
-override those with the same name found elsewhere.  Any unique
-source files (and not simply a custom version of a file found
-elsewhere) needs to be listed in a file call {\tt GPackage.mak}
-in the problem-directory.
-
-\subsection{The {\tt GNUmakefile}}
-
-\label{sec:makefile}
-
-A basic {\tt GNUmakefile} begins with:
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  NDEBUG := t
-  MPI    :=
-  OMP    :=
-\end{lstlisting}
-Here, {\tt NDEBUG} is true if we are building an optimized executable.
-Otherwise, the debug version is built---this typically uses less
-optimization and adds various runtime checks through compiler flags.
-{\tt MPI} and {\tt OMP} are set to true if we want to use either MPI
-or OpenMP for parallelization.  If {\tt MPI := t}, you will need to
-have the MPI libraries installed, and their location may need to be 
-specified in {\tt MAESTRO/mk/GMakeMPI.mak}.
-
-The next line sets the compiler to be used for compilation:
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  COMP := Intel
-\end{lstlisting}
-The \maestro\ build system knows what options to use for various
-compiler families.  The {\tt COMP} flag specifies which compiler to
-use.  Allowed values include {\tt Intel}, {\tt gfortran}, {\tt PGI},
-and {\tt PathScale}.  The specific details of these choices are
-defined in the {\tt MAESTRO/mk/comps/} directory.
-
-{\tt MKVERBOSE} set to true will echo the build commands to the
-terminal as the are executed.
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  MKVERBOSE := t
-\end{lstlisting}
-
-The next line defines where the top of the source tree is located.
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  FPARALLEL := ../..
-\end{lstlisting}
-
-A \maestro\ application is built from several packages (the
-multigrid solver, an EOS, a reaction network, etc.).  The core
-\maestro\ packages are always included, so a problem only needs
-to define the EOS, reaction network, and conductivities to
-use, as well as any extra, problem-specific files.  
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-EOS_DIR := extern/EOS/helmeos   
-CONDUCTIVITY_DIR := extern/conductivity/constant
-NETWORK_DIR := extern/networks/ignition
-
-EXTRA_DIR := extern/random
-\end{lstlisting}
-Generally, one does not need to include the problem directory itself
-in {\tt EXTRA\_DIR}, unless there are unique source files found there,
-described in a {\tt GPackage.mak} file.  These variables are
-interpreted by the {\tt GMaestro.mak} file and used to build a master
-list of packages called {\tt Fmdirs}.  The build system will attempt
-to build all of the files listed in the various {\tt GPackage.mak}
-files found in the {\tt Fmdirs} directories.  Furthermore, {\tt
-  Fmdirs} will be will be added to the {\tt make} {\tt VPATH}, which
-is the list of directories to search for source files.  The problem
-directory will always be put first in the {\tt VPATH}, so any source
-files placed there override those with the same name found elsewhere
-in the source tree.
-
-Some packages (for instance, the {\tt helmeos}
-EOS) require Fortran include files.  The {\tt Fmincludes} variable
-lists all those directories that contain include files that are
-inserted into the Fortran source at compile time via the {\tt include}
-statement.  Presently, the only instance of this is with the Helmholtz
-general equation of state found in {\tt extern/EOS/helmeos/}.  This is
-automatically handled by the {\tt GMaestro.mak} instructions.
-
-Runtime parameters listed in the {\tt MAESTRO/\_parameters} file are
-parsed at compile time and the file {\tt probin.f90} is written and
-compiled.  This is a Fortran module that holds the values of the
-runtime parameters and makes them available to any routine.  The
-variable {\tt PROBIN\_PARAMETERS} should list all of the files that
-contain runtime parameters for the current problem.  You must always
-include the main {\tt ../\_parameters} file (i.e.\ {\tt
-  MAESTRO/\_parameters}).  Sometimes a problem defines their own
-runtime parameters (for instance {\tt \_parameters.myproblem}), in which
-case, its parameter file should also appear in {\tt
-  PROBIN\_PARAMETERS} (see \S~\ref{sec:def_runtime_param}).
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  PROBIN_PARAMETERS := ../_parameters
-\end{lstlisting}
-
-The final line in the {\tt GNUmakefile} includes the rules to actually
-build the executable.
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  include $(FPARALLEL)/MAESTRO/GMaestro.mak
-\end{lstlisting}
-%$
-
-\subsubsection{Handling Problem-Specific Source Files}
-
-As mentioned above, any source files placed in the problem directory
-override a files with the same name found elsewhere in the source
-tree.  This allows you to create a problem-specific version of any
-routine.  Source files that are unique to this problem (i.e.\ there is
-no file with the same name elsewhere in the source tree) need to be
-listed in a file {\tt GPackage.mak} in the problem directory, and
-the problem-directory needs to be explicitly listed in the {\tt EXTRA\_DIR}
-list in the {\tt GNUmakefile}.
-
-\subsection{Defining Runtime Parameters}
-
-\label{sec:def_runtime_param}
-
-The runtime parameters for the core \maestro\ algorithm are listed
-in {\tt MAESTRO/\_parameters}.  That file is parsed at compile-time
-by the {\tt MAESTRO/write\_probin.py} script
-(along with any other parameter files listed in the {\tt PROBIN\_PARAMETERS}
-variable in the {GNUmakefile}, see \S \ref{sec:makefile}).  The script
-outputs the {\tt probin.f90} source file.  Each line
-in the {\tt \_parameters} file has the form: 
-\vskip 3mm
-{\em parameter} \hskip 10em  {\em data-type} \hskip 10em  {\em value} 
-\vskip 3mm
-\noindent where {\em parameter} is the name of the runtime parameter,
-{\em data-type} is one of \{{\tt character}, {\tt real}, {\tt
-  integer}, {\tt logical}\}, and the {\em value} specifies the default
-value for the runtime parameter.  Comments are indicated by a `{\tt
-  \#}' character and a used to produce documentation about the
-available runtime parameters.  For the documentation, runtime parameters are grouped together
-in the {\tt \_parameters} file into categories.  The category headings
-are defined by comments in the {\tt \_parameters} file and any comments
-following that heading are placed into that category.  The documentation
-is produced by the script {\tt MAESTRO/docs/runtime\_parameters/rp.py}.
-
-At runtime, the default values for the parameters can be overridden
-either through the inputs file (by adding a line of the form: {\tt
-  parameter = value}) or through a command-line argument (taking the
-form: {\tt --parameter value}).  The {\tt probin\_module} makes the
-values of the runtime parameters available to the various functions
-in the code (see \S~\ref{sec:probin}).
-
-Problem-specify runtime parameters should be defined in the
-problem-directory in a file called {\tt
-  \_parameters}\{{\em.problem-name}\}.  This file should be added
-to the {\tt PROBIN\_PARAMETERS} variable in the {\tt GNUmakefile}.
-
-\subsection{Preparing the Initial Model}
-
-\label{sec:initial_models}
-
-\maestro\ models subsonic flows that are in hydrostatic equilibrium.
-The solution in \maestro\ is broken up into a 1D base state and the 2-
-or 3D full state.  The job of the 1D base state in the algorithm is
-to represent the hydrostatic structure.  The full, Cartesian state
-carries the departures from hydrostatic equilibrium.  The underlying
-formulation of the low Mach number equations assumes that the base
-state is in hydrostatic equilibrium.  At the start of a simulation,
-the initial model is read in and taken as the base state.  Therefore,
-any initial model needs to already be in hydrostatic equilibrium.
-
-The routines in {\tt MAESTRO/initial\_model/} prepare an initial model
-for \maestro.  There are two types of routines here, the first type
-take an existing 1D initial model produced somewhere else (e.g.\ a
-1D stellar evolution code), and and map it onto a uniform grid, at
-the desired resolution, using the equation of state in \maestro, and
-using \maestro's discretization of hydrostatic equilibrium.  The second
-type generate the initial model internally, by integrating the
-condition of hydrostatic equilibrium together with a simplifying
-assumption on the energy (e.g. isothermal or isentropic).  In
-both cases hydrostatic equilibrium is enforced as:
-\begin{equation}
-\frac{p_{i+1} - p_i}{\Delta r} = \frac{1}{2} (\rho_i + \rho_{i+1})
-g_{i+1/2}
-\end{equation}
-Here, $g_{i+1/2}$ is the edge-centered gravitational acceleration.
-
-The {\tt toy\_atm} example provides a simple approximation for a thin
-(plane-parallel) convectively-unstable accreted layer on the surface
-of a star.  This can be used as the starting point for a more complex
-model.
-
-\maestro\ initial models are read in by the {\tt extern/model\_parser}
-routines.  This expects the initial model to contain a header giving
-the number of variables and their names, followed by rows of data
-giving the coordinate and data values at that coordinate.  The initial
-model should contain the same species data (as mass fractions) as
-defined in the {\tt network} module used by the \maestro\ problem.
-
-Full details on which initial model routine matches each problem and
-how the initial models are used to initialize the full state data can
-be found in \S~\ref{sec:initial_models_main}.
-
-\section{BoxLib Data Structures}
-
-\maestro's gridding is handled by the BoxLib library, which
-contains the most fundamental objects used to construct parallel
-block-structured AMR applications---different
-regions of the domain can have different spatial resolutions.
-At each level of refinement, the region covered by that level is divided
-into grids, or boxes.  The entire computational domain is covered by
-the coarsest (base) level of refinement, often called level $\ell=0$, either by one
-grid or divided into many grids.
-Higher levels of refinement have cells that are finer by a ``refinement ratio''
-(typically 2).  The grids are properly nested in the sense that the union 
-of grids at level $\ell+1$ is contained in the union of grids at level $\ell$.
-Furthermore, the containment is strict in the sense that, except at physical 
-boundaries, the level $\ell$ grids are large enough to guarantee that there is
-a border at least $n_{\rm buffer}$ level $\ell$ cells wide surrounding each level
-$\ell +1$ grid (grids at all levels are allowed to extend to the physical
-boundaries so the proper nesting is not strict there).  
-For parallel computations, the boxes are spread across processors, in
-a fashion designed to put roughly equal amounts of work on each
-processor (load balancing).
-\MarginPar{ghost cells?}
+  The output from this test can be visualized with the script {\tt
+  plot.py} in the test directory.  The output shows the particle
+  paths (see figure~\ref{fig:unit:particles}).
 
@@
 \centering
-\includegraphics[width=6.5in]{\archfigpath/data_loc2}
-\caption[Data-centerings on the grid]
-  {\label{fig:dataloc} Some of the different data-centerings:
-  (a) cell-centered, (b) nodal in the $x$-direction, and (c) nodal in
-  both the $x$- and $y$-directions.  Note that for nodal data, the
-  integer index corresponds to the lower boundary in that direction.
-  In each of these centerings, the red point has the same indices:\ (1,2).
-  Not shown is the case where data is nodal in the $y$-direction only.}
-\end{figure}
-On a grid, the data can be stored at cell-centers, on a face/edge, or
-on the corners.  In BoxLib, data that is on an edge is termed `nodal'
-in that direction (see Figure~\ref{fig:dataloc}).  Data that is on the
-corners is nodal in all spatial directions.  In \maestro, the state
-data (density, enthalpy, velocity, $\ldots$) is generally
-cell-centered.  Fluxes are nodal in the direction they represent.
-A few quantities are nodal in all directions (e.g.\ $\phi$ used in
-the final velocity projection).
-
-To simplify the description of the underlying AMR grid, BoxLib
-provides a number of Fortran types.  We briefly summarize the major
-data types below.
-
-\subsection{\boxtype}
-
-A \boxtype\ is simply a rectangular domain in space.  Note that boxes
-do not hold the state data themselves.  A \boxtype\ has a {\tt lo} 
-and {\tt hi} index in each coordinate direction that gives the
-location of the lower-left and upper-right corner with respect to
-a global index space.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=4.0in]{\archfigpath/index_grid2}
-\caption[Single-level grid structure]
-{\label{fig:boxes} Three boxes that comprise a single level.  At this
-  resolution, the domain is 20$\times$18 zones.  Note that the
-  indexing in BoxLib starts with $0$.}
+\includegraphics[width=4in]{\unitfigpath/particle_paths} 
+%
+\caption[Particle paths for the {\tt test\_particles} problem]{\label{fig:unit:particles}
+  Particle paths for the {\tt test\_particles} problem.  The initial
+  position of the particles is marked with an $\times$.}
 \end{figure}
 
-The computational domain is divided into boxes.  The collection of
-boxes with the same resolution comprise a level.
-Figure~\ref{fig:boxes} shows three boxes in the same level of
-refinement.  The position of the boxes is with respect to the global
-index space at that level.  For example, box 1 in the figure has {\tt
-  lo} = (3,7) and {\tt hi} = (9,12).  Note that the global indexing
-is 0-based.
+\section {\tt test\_projection}
 
-The global index space covers the entire domain at a given resolution.
-For a simulation setup with {\tt n\_cellx = 32} and {\tt n\_celly =
-  32}, the coarsest level (level 1) has $32 \times 32$ zones, and the
-global index space will run from $0, \ldots, 31$ in each coordinate
-direction.  Level 2 will have a global index space running from $0,
-\ldots, 63$ in each coordinate direction (corresponding to $64 \times
-64$ zones if fully refined), and level 3 will have a global index
-space running from $0, \ldots, 127$ in each coordinate direction
-(corresponding to $128\times 128$ zones if fully refined).
-
-\subsubsection{Common Operations on a \boxtype}
-
-A \boxtype\ declared as:
-\begin{verbatim}
-  type(box) :: mybox
-\end{verbatim}
-%
-The upper and lower bounds of the box (in terms of the global
-index space) are found via:
-\begin{itemize}
-
-\item {\tt lo = lwb(mybox)} returns an array, {\tt lo(dm)}, with
-     the box lower bounds
-
-\item {\tt hi = upb(mybox)} returns an array, {\tt hi(dm)}, with
-     the box upper bounds
-
-\end{itemize}
-
-\subsection{\boxarray\ and \mlboxarray}
-
-A \boxarray\ is an array of boxes.  A \mlboxarray\ is a collection of
-\boxarray s at different levels of refinement.
-
-\subsection{\layout\ and \mllayout}
-
-A \layout\ is basically a \boxarray\ that knows information about other
-boxes, or box ``connectivity.''  It contains additional information
-that is used in filling ghost cells from other fine grids or from
-coarser grids.  This information is stored as long as the layout
-exists so that we don't have to recompute intersections every time we
-do some operation with two \multifab s that have that layout, for
-example.
-
-\subsection{\fab}
-
-A \fab\ is a ``Fortran Array Box''.  It contains the state data in a
-multidimensional array and several \boxtype-types to describe where in
-the global index-space it lives.  
-Note that all state data is stored in a four-dimensional array,
-{\tt (nx,ny,nz,nc)} in size, regardless of the dimensionality of the
-problem.  Here {\tt nc} is the number of components, for instance
-representing different fluid variables.  For 2D problems, {\tt nz=1}.
-
-A \fab\ would represent the data for a single box in the domain.
-In \maestro, we don't usually deal with \fab s alone, but rather
-we deal with \multifab s, described next.
-
-\subsection{\multifab}
-
-A \multifab\ is a collection of \fab s at the same level of
-refinement.  This is the primary data structure that \maestro\
-routine operate on.  A multilevel simulation stores the 
-data in an array of \multifab s, where the array index refers
-to the refinement level.
-
-\subsubsection{Working with \multifab s}
-
-To build a \multifab, we need to provide a \layout, the number of
-components to store in the \multifab\, and the number of ghostcells.  In
-\maestro\, the hierarchy of grids will be described by a single
-\mllayout.  A \multifab\ can be declared and built at any time in a
-simulation using the \mllayout, thereby allocating space at every
-grid location in the simulation.  The sequence to build a \multifab\
-appears as
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  type(multifab) :: mfab(nlevs)
-  ...
-  do n = 1, nlevs
-     call multifab_build(mfab(n), mla%la(n), nc, ng)
-  enddo
-\end{lstlisting}
-Here, {\tt nc} is the number of components and {\tt ng} is the number
-of ghostcells.  The \multifab\ is built one level at a time, using the
-\layout\ for that level taken from the \mllayout, {\tt mla}.
-
-A common operation on a \multifab\ is to initialize it to $0$
-everywhere.  This can be done (level-by-level) as
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-call setval(mfab(n), ZERO, all=.true.)
-\end{lstlisting}
-where {\tt ZERO} is the constant 0.0 from {\tt bl\_constants\_module}.
-
-The procedure for accessing the data in each grid managed by the \multifab\
-is shown in \S~\ref{sec:example}.  
-Subroutines to add, multiply, or divide two \multifab s exist, as do
-subroutines to copy from one \multifab\ to another---see {\tt boxlib/multifab.f90}
-for the full list of routines that work with \multifab s.
-
-When you are done working with a \multifab, its memory can be freed by
-calling {\tt multifab\_destroy} on the \multifab.
-
-\subsection{\bctower}
-
-A \bctower\ holds the information about what boundary conditions are
-in effect for each variable in a \maestro\ simulation.  \MarginPar{better description?}
-
-\section{\maestro\ Data Organization}
-
-The state of the star in \maestro\ is described by both a
-multidimensional state and the 1D base state.  The full
-multidimensional state is stored in \multifab s while the base state
-is simply stored in Fortran arrays.  Here we describe the
-major \maestro\ data-structures.
-
-\subsection{`{\tt s}' \multifab s (fluid state)}
-
-The fluid state (density, enthalpy, species, temperature, and tracer)
-are stored together in a cell-centered multi-component \multifab,
-typically named {\tt sold}, {\tt s1}, {\tt s2}, or {\tt snew}
-(depending on which time-level it represents).  The enthalpy is stored
-as $(\rho h)$, and the species are stored as partial-densities $(\rho
-X_k)$.  The tracer component is not used at present time, but can
-describe an arbitrary advected quantity.
-
-Individual state variables should be indexed using the integer keys
-provided by the {\tt variables} module (see \S
-\ref{sec:variables_module}).  For example, the integer {\tt rho\_comp}
-will always refer to the density component of the state.
-
-Note: the pressure is not carried as part of the `{\tt s}' \multifab s.
-
-\subsection{`{\tt u}' \multifab s (fluid velocity)}
-
-The fluid velocity at time-levels $n$ and $n+1$ is stored in
-a cell-centered multi-component \multifab, typically named
-{\tt uold} or {\tt unew}.  Here the {\tt dm}
-components correspond to each coordinate direction.
-
-\subsection{{\tt umac} (the MAC velocity)}
-
-In creating the advective fluxes, we need the time-centered velocity
-through the faces of the zone---the $x$-velocity on the $x$-edges, the
-$y$-velocity on the $y$-edges, etc.\ (see figure~\ref{fig:mac}).  This
-type of velocity discretization is termed the MAC velocity (after the
-``marker-and-cell'' method for free boundaries in incompressible
-flows \cite{harlowwelch:1965}).
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=2.5in]{\archfigpath/mac2}
-\hspace{0.1in}
-\begin{minipage}[b]{3.8in}
-\caption[The MAC grid]
-{\label{fig:mac} The MAC grid for the velocity.  
-Here the $x$-velocity is on the $x$-edges (shown as the 
-blue points) and the $y$-velocity is on the $y$-edges
-(shown as the red points).
-}\ \\
-\end{minipage}
-\end{figure}
-
-The MAC velocities are allocated at each level of refinement, {\tt n},
-by making a \multifab\ array where each of the {\tt dm} components is
-nodal in its respective direction:
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  type(multifab) :: umac(dm)
-
-  do comp=1,dm
-     call multifab_build_edge(umac(n,comp), mla%la(n),1,1,comp)
-  enddo
-\end{lstlisting}
-
-\subsection{Base State Arrays}
-
-The base state is defined by $\rho_0$, $p_0$, and $w_0$.  There is no
-base state composition.  Other arrays are defined as needed, such as
-$h_0$, the base state enthalpy.
-
-The base state arrays are 2-dimensional, with the first dimension
-giving the level in the AMR hierarchy and the second the radial index
-into the base state.  For spherical geometries, the base state only
-exists at a single level, so the first index will always be 1.  The
-radial index is 0-based, to be consistent with the indexing for the
-Cartesian state data.  For example, the base state density would be
-dimensioned: {\tt rho0(nlevs,0:nr\_fine-1)}.  Here, {\tt nlevs} is the
-number of levels of refinement and {\tt nr\_fine} is the number of
-cells in the radial direction at the finest level of refinement.
-
-For multilevel, plane-parallel geometry, all grids at the same height
-will have the same resolution so that the full state data is always
-aligned with the base state (see Figure~\ref{fig:base_state}).  Base
-state data on coarse grids that are covered by fine grids is not
-guaranteed to be valid.
-
-For spherical problems, the base state resolution, $\Delta r$, is
-generally picked to be finer than the Cartesian grid resolution,
-$\Delta x$, i.e.\ $\Delta r < \Delta x$.  The ratio is controlled
-by the parameter {\tt drdxfac}.
-
-\section{\maestro\ Helper Modules}
-
-A number of \maestro\ modules appear frequently throughout the source.
-Below, we describe some of the more common functionality of the most
-popular modules.
-
-\subsection{\tt average}
-
-The {\tt average} module provides a routine {\tt average} that takes
-a multilevel \multifab\ array and averages the full Cartesian data
-onto the 1D base state.
-
-\subsection{{\tt eos\_module}}
-
-The {\tt eos\_module} provides the interface to the equation of 
-state to connect the state variables thermodynamically.  It 
-gets the information about the fluid species from the {\tt network}
-module (for example, the atomic number, $Z$, and atomic weight, $A$,
-of the nuclei).
-
-Presently there are 2 equations of state that work with \maestro:
-\begin{itemize}
-\item {\tt extern/EOS/helmeos/} represents a general stellar equation 
-      of state, consisting of nuclei (as an ideal gas), radiation,
-      and electrons (with arbitrary degeneracy and degree of relativity).
-      This equation of state is that described in \cite{timmes_eos}.
-
-\item {\tt extern/EOS/gamma\_law\_general} assumes an ideal gas with a mixed 
-     composition and a constant ratio of specific heats, $\gamma$:
-      \begin{equation}
-      p = \rho e (\gamma - 1) = \frac{\rho k_B T}{\mu m_p} 
-      \end{equation}
-     where $k_B$ is Boltzmann's constant and $m_p$ is the mass of the
-     proton.
-     The mean molecular weight, $\mu$, is computed assuming 
-     electrically neutral atoms:
-     \begin{equation}
-     \mu = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}
-     \end{equation}
-     An option in the source code itself exists for treating the
-     species as fully-ionized, but there is no runtime-parameter to
-     make this switch.
-\end{itemize}
-
-The {\tt eos\_module} declares for convenience the variables that need
-appear in the {\tt eos} call argument list.  \maestro\ routines use
-these module variables in the EOS call to avoid having to declare 
-each quantity in each routine that calls the EOS.
-
-The first argument to the {\tt eos} call is an integer key that
-specifies which thermodynamic variables (in addition to the mass
-fractions) are used as input.  EOS input options are listed 
-in table~\ref{arch:table:eosinput}.
-
-   \begin{table}[h]
-   \caption{\label{arch:table:eosinput} EOS input flags}
-   \begin{center}
-   \begin{tabular}{lc}
-   \hline
-   key            & input quantities \\
-   \hline
-   {\tt eos\_input\_rt}       & $\rho$, $T$ \\
-   {\tt eos\_input\_rh}       & $\rho$, $h$ \\
-   {\tt eos\_input\_tp}       & $T$, $p$ \\
-   {\tt eos\_input\_rp}       & $\rho$, $p$ \\
-   {\tt eos\_input\_re}       & $\rho$, $e$ \\
-   {\tt eos\_input\_ps}       & $p$, $s$ \\
-   \hline
-   \end{tabular}
-   \end{center}
-   \end{table}
-
-\subsection{{\tt fill\_3d\_module}}
-
-The {\tt fill\_3d\_module} provides routines that map from the 1D
-base state to the full Cartesian 2- or 3D state.  Variations in the
-routines allow for cell-centered or edge-centered data on either the
-base state or full Cartesian state.
-
-\subsection{{\tt geometry}}
-
-\subsection{{\tt network}}
-
-The {\tt network} module defines the number species advected by the
-code ({\tt nspec}), their ordering, and gives their basic properties
-(like atomic number, $Z$, and atomic mass, $A$).  All \maestro\ problems
-require a {\tt network} module, even if there are no reactions
-modeled.  Many different reaction modules (containing different sets
-of isotopes) exist in {\tt extern/networks}.  The particular network
-used by a problem is defined in the problem's {\tt GNUmakefile}.
-
-To find the location of a particular species (for instance, ``carbon-12'')
-in the allowed range of {\tt 1:nspec}, you do the following query:
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  ic12 = network_species_index("carbon-12")
-\end{lstlisting}
-If the resulting index is {\tt -1}, then the requested species was not
-found.
-
-\subsection{{\tt probin\_module}}
-
-\label{sec:probin}
-
-{\tt probin\_module} provides access to the runtime parameters.
-The runtime parameters appear simply as module variables.  To get the 
-value of a parameter, one simply needs to `{\tt use probin\_module}'.
-The preferred method is to add the `{\tt only}' clause to the
-{\tt use} statement and explicitly list only those parameters that
-are used in the routine.  Defining new runtime parameters is
-described in \S~\ref{sec:def_runtime_param}.
-
-\subsection{{\tt variables}}
-
-\label{sec:variables_module}
-
-The {\tt variables} module provides integer keys to index the state
-multifabs and other arrays dealing with the scalar quantities.  The
-most commonly used keys are are list in table~\ref{arch:table:variables}.
-
-\begin{table}[h]
-\caption{\label{arch:table:variables} Common {\tt variables} module keys}
-\begin{center}
-\begin{tabular}{ll}
-\hline
-{\tt rho\_comp}  & density \\
-{\tt rhoh\_comp} & density $\times$ specific enthalpy, $(\rho h)$ \\
-{\tt spec\_comp} & first species partial density, $(\rho X_1)$ \\
-{\tt temp\_comp} & temperature \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-The species indices are contiguous in the state array, spanning {\tt
-  spec\_comp:spec\_comp-1+nspec}.  To find a particular species, a
-query can be made through the {\tt network} module, such as:
-\begin{verbatim}
-  ic12 = network_species_index("carbon-12")
-\end{verbatim}
-and then the \fab\ can be indexed using {\tt spec\_comp-1+ic12} to
-get ``carbon-12''.
-The {\tt variables} module also provides keys for the plotfile
-variables and boundary condition types.
-
-Other keys in the {\tt variables} modules are reserved for boundary
-conditions ({\tt foextrap\_comp} and {\tt hoextap\_comp}), the
-projection of the pressure ({\tt press\_comp}), or constructing
-the plotfile.
-
-\section{BoxLib Helper Modules}
-
-There are a large number of modules in {\tt boxlib/} that provide
-the core functionality for managing grids.  Here we describe
-the most popular such modules.
-
-\subsection{{\tt bl\_types}}
-
-\subsection{{\tt bl\_constants}}
-
-\subsection{{\tt parallel}}
-
-\section{\label{sec:example} Example: Accessing State and MAC Data}
-
-In \maestro, the state data is stored in a cell-centered multifab array
-(the array index refers to the AMR level) and the MAC velocities are
-stored in a 2D nodal multifab array (with indices referring to the AMR
-level and the velocity component).  Here we demonstrate a typical way
-to extract the state and MAC velocity data.
-
-All \maestro\ routines are contained in a module, to allow for compile-time
-argument checking.
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-module example_module
-
-contains
-\end{lstlisting}
-
-The main interface to our routine is called {\tt example}---this will
-take the \multifab s containing the data and then pass them to the
-work routines, {\tt example\_2d} or {\tt example\_3d}, depending on
-the dimensionality.  
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  subroutine example(mla,s,umac,dx,dt)
-
-    use bl_types
-    use multifab_module
-    use ml_layout_module
-    use variables, only: rho_comp
-\end{lstlisting}
-
-\noindent Here, the {\tt bl\_types} and {\tt multifab\_module} modules
-bring in the basic BoxLib data types. Specifically, here, {\tt
-  bl\_types} defines {\tt dp\_t} which is the {\tt kind} used for
-declaring double precision data, and {\tt multifab\_module} defines
-the \multifab\ data type.  The {\tt ml\_layout\_module} defines the
-datatype for a \mllayout---many routines will take an \mllayout
-to allow us to fill ghostcells.  The {\tt variables} module is a \maestro\
-module that provides integer keys for indexing the state arrays.  In
-this case the integer {\tt rho\_comp} refers to the location in the
-state array corresponding to density.
-
-Next we declare the subroutine arguments:
-
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-    type(ml_layout), intent(in   ) :: mla
-    type(multifab) , intent(inout) :: s(:)
-    type(multifab) , intent(inout) :: umac(:,:)
-    real(kind=dp_t), intent(in   ) :: dx(:,:),dt
-\end{lstlisting}
-Here, {\tt s(:)} is our \multifab\ array that holds the state data.
-with the array index in {\tt s} refers to the AMR level.  The MAC 
-velocities are held in the multifab {\tt umac}, with the array
-indices referring to the AMR level and the component.
-
-Local variable declarations come next:
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-    ! Local variables
-    real(kind=dp_t), pointer :: sp(:,:,:,:)
-    real(kind=dp_t), pointer :: ump(:,:,:,:), vmp(:,:,:,:), wmp(:,:,:,:)
-    integer :: i,n,dm,nlevs,ng_sp,ng_um
-    integer :: lo(mla%dim),hi(mla%dim)
-\end{lstlisting}
-
-\noindent Amongst the local variables we define here are a pointer,
-{\tt sp}, that will point to a single \fab\ from the
-\multifab\ {\tt s}, and a pointer for each component of the MAC
-velocity, {\tt ump}, {\tt vmp}, and {\tt wmp} (for a 2D run,
-we won't use {\tt wmp}).  We note that regardless of the dimensionality,
-these pointers are 4-dimensional: 3 spatial + 1 component.
-
-Next we get the dimensionality and number of levels
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-    dm = mla%dim
-    nlevs = mla%nlevel
-\end{lstlisting}
-
-Each multifab can have their own number of ghostcells, so we get
-these next:
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-    ng_sp = nghost(s(1))
-    ng_um = nghost(umac(1,1))
-\end{lstlisting}
-By convention, all levels in a given multifab have the same number of
-ghostcells, so we use level 1 in the {\tt nghost()} call.  We also use
-the same number of ghostcells for each component of the velocity, so
-we only need to consider the first component in the {\tt nghost()}
-call.  The ghostcells will be needed to access the data stored in the
-\fab s.
-
-To access the data, we loop over all the levels, and all the boxes in
-the given level.
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-    do n=1,nlevs
-       do i = 1, nboxes(s(n))
-\end{lstlisting}
-{\tt nboxes(s(n))} is simply the number of boxes in level {\tt n}.
-All data in all \multifab s will have the same box layout, so we only
-need to query the number of boxes from one \multifab.
-
-For a given \boxtype, we get the data and the bounds of the \boxtype.
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-          if ( multifab_remote(s(n), i) ) cycle
-          sp  => dataptr(s(n), i)
-          ump => dataptr(umac(n,1),i)
-          vmp => dataptr(umac(n,2),i)
-          lo =  lwb(get_box(s(n), i))
-          hi =  upb(get_box(s(n), i))
-\end{lstlisting}
-
-\noindent For parallel runs, there is no guarantee that the particular
-box is on the current processor.  The function {\tt multifab\_remote}
-returns {\tt .true.} if the box is off-processor, in which case, we 
-simply skip to the next box.
-
-The actual data array is accessed through the {\tt dataptr} function,
-which takes a \multifab\ (e.g.\ {\tt s(n)}) and the index of the
-\boxtype\ ({\tt i}) we want.  We see that the $x$ MAC velocity for the
-current box is stored in {\tt ump} and the $y$ MAC velocity is stored
-in {\tt vmp}.  We don't get the $z$ velocity data here, since that
-would not be available for a 2D run---we defer that until we test on
-the dimensionality below.
-
-Finally, the index bounds of the box (just the data, not the ghostcells) are 
-stored in the {\tt dm}-dimensional arrays {\tt lo} and {\tt hi}.  These indices
-refer to the current box, and hold for both the state, {\tt sp}, and the MAC
-velocity, {\tt ump} and {\tt vmp}.  However, since the MAC velocity is nodal
-in the component direction, the loops over the valid data will differ
-slight (as we see below).
-
-With the data extracted, we call a subroutine to operate on it.  We use
-different subroutines for the different dimensionalities (and many times
-have a separate routine for spherical geometries).
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-          select case (dm)
-          case (2)
-             call example_2d(sp(:,:,1,rho_comp),ng_sp, &
-                             ump(:,:,1,1),vmp(:,:,1,1),ng_um, &
-                             lo,hi,dx(n,:),dt)
-          case (3)
-             wmp => dataptr(umac(n,3),i)
-             call example_3d(sp(:,:,:,rho_comp),ng_sp, &
-                             ump(:,:,:,1),vmp(:,:,:,1),wmp(:,:,:,1),ng_um, &
-                             lo,hi,dx(n,:),dt)
-          end select
-       enddo    ! end loop over boxes
-
-    enddo    ! end loop over levels
-
-  end subroutine example
-\end{lstlisting}
-\noindent We call either the function
-{\tt example\_2d} for two-dimensional data or {\tt example\_3d}
-for three-dimensional data.  Note that in the two-dimensional
-case, we index the data as {\tt sp(:,:,1,rho\_comp)}.  Here a
-`{\tt 1}' is used as the `z'-coordinate spatial index, since this
-is a 2D problem, and the density component of the state is selected
-(using the integer key {\tt rho\_comp}).  The 3D version accesses
-the data as {\tt sp(:,:,:,rho\_comp)}---only the component regarding
-the variable is needed here.  Notice that we also pass through
-the number of ghostcells for each of the quantities.
-
-This routine will be supplimented with {\tt example\_2d} and {\tt
-example\_3d}, which actually operate on the data.  The form of 
-the 2D function is:
-
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  subroutine example_2d(density,ng_sp, &
-                        umac,vmac,ng_um, &
-                        lo,hi,dx,dt)
-
-    use bl_constants_module
-    use probin_module, only: prob_lo
-
-    integer        , intent(in) :: lo(:),hi(:), ng_sp, ng_um
-    real(kind=dp_t), intent(in) :: density(lo(1)-ng_sp:,lo(2)-ng_sp:)
-    real(kind=dp_t), intent(in) ::    umac(lo(1)-ng_um:,lo(2)-ng_um:)
-    real(kind=dp_t), intent(in) ::    vmac(lo(1)-ng_um:,lo(2)-ng_um:)
-
-    real(kind=dp_t), intent(in) :: dx(:),dt
-
-    integer         :: i, j
-    real(kind=dp_t) :: x, y
-    real(kind=dp_t) :: dens, u, v
-
-    do j = lo(2), hi(2)
-       y = prob_lo(2) + (dble(j) + HALF)*dx(2)
-
-       do i = lo(1), hi(1)
-          x = prob_lo(1) + (dble(i) + HALF)*dx(1)
-
-          dens = density(i,j)
-
-          ! compute cell-centered velocity
-          u = HALF*(umac(i,j) + umac(i+1,j))
-          v = HALF*(vmac(i,j) + vmac(i+1,j))
-
-          ! operate on the data
-          ! ...
-
-       enddo
-    enddo
-
-  end subroutine example_2d
-
-end module example_module
-\end{lstlisting}
-
-\noindent In this function, the bounds of the {\tt density} array take
-into account the {\tt ng\_sp} ghostcells and the index space of the
-current box.  Likewise, the MAC velocities refer to the {\tt ng\_um}
-ghostcells.  The {\tt j} and {\tt i} loops loop over all the valid
-zones.  Coordinate information is computed from {\tt dx} and {\tt
-  prob\_lo} which is the physical lower bound of the domain.  {\tt
-  bl\_constants\_module} declares useful double-precision constants,
-like {\tt HALF} (0.5).  Here, we see how to access the density for
-the current zone and compute the cell-centered velocities from the
-MAC velocities.  By convection, for a nodal array, the indices refer 
-to the {\em lower} interface in the nodal direction, so for {\tt umac},
-{\tt umac(i,j)} and {\tt umac(i+1,j)} are the $x$ MAC velocities
-on the lower and upper edge of the zone in the $x$-direction.
-
-The three-dimensional case is similar, with the {\tt density} array
-declared as 
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  density(lo(1)-ng_sp:,lo(2)-ng_sp:,lo(3)-ng_sp:)
-\end{lstlisting}
-and an additional loop over the `z' coordinate (from {\tt lo(3)} to
-{\tt hi(3)}).
-
-In this example, we looped over the valid zones.  If we wished to loop
-over the interfaces bounding the valid zones, in the $x$-direction,
-we would loop as
-\begin{lstlisting}[language={[95]fortran},mathescape=false]
-  do j = lo(2), hi(2)
-     do i = lo(1), hi(1)+1
-        ! access umac(i,j)
-     enddo
-  enddo
-\end{lstlisting}
-
-\section{Filling Ghostcells}
-
-Ghostcells are filled through a variety of different routines, depending
-on the objective.
-
-\begin{itemize}
-
-\item {\tt multifab\_fill\_boundary} fills ghost cells for two
-  adjacent grids at the same level, which als includes periodic domain
-  boundary ghost cells.
-
-\item {\tt multifab\_physbc} fills ghostcells at the physical boundaries.
-
-\item {\tt multifab\_fill\_ghost\_cells} is used for multilevel
-  problems, and fills ghostcells in the finer grid (level {\tt n}) by
-  interpolating from data in the coarser grid (level {\tt n-1}).
-  This function, by default, will also call {\tt multifab\_fill\_boundary}
-  and {\tt multifab\_physbc} for both levels {\tt n} and {\tt n-1} (you 
-  can override this behavior for speed optimization purposes).
-  This call is usually preceded by a call to 
-  {\tt ml\_cc\_restriction\_c} which sets the level {\tt n-1} data to be
-  the average of the level {\tt n} data covering it.
-
-\end{itemize}
-
-\section{Boundary Conditions}
-
-\section{Multigrid}
-
-\maestro\ uses the multigrid solver to enforce the divergence constraint
-both on the half-time edge-centered advective velocities (the ``MAC
-projection'') and on the final cell-centered velocities (the ``HG
-projection'').  For the MAC projection, since the velocity data is
-edge-centered (the MAC grid), the projection is cell-centered.  For
-the HG projection, since the velocity data is cell-centered, the
-projection is node-centered.  \MarginPar{should draw a figure} The multigrid solver performs a number
-of V-cycles until the residual drops by 10-12 orders-of-magnitude.
-There are several options that affect how the multigrid solver
-behaves, which we describe below.
-
-\subsection{Special Bottom Solver}
-We have a special bottom solver type that greatly speeds up large problems
-and speeds up smaller problems to a lesser degree.
-In fact, you should use this bottom solver whenever possible, because I
-have yet to find a case that performs worse than our standard bottom solver.
-What this special solver essentially does is take the data from the coarsest 
-level of the original multigrid V-cycle and copy it onto a new grid structure 
-with the same number of total cells in each direction, but with a fewer number of larger 
-grids.  A new V-cycle begins from this point, 
-so we are essentially coarsening this ``new'' problem.
-Now, the coarsest level of the multigrid V-cycle in the ``new'' problem has 
-fewer cells and fewer grids as compared to the original coarsest level.
-
-To enable this solver, set {\tt hg\_bottom\_solver = 4} (for the nodal
-projections) and/or {\tt mg\_bottom\_solver = 4} (for the cell-centered
-projections) in your inputs file.
-
-To understand how this bottom solver works, the first thing you need to know
-is what the grid structure of the coarsest level of your multigrid V-cycle
-looks like (we need to put a description of this in the section above).
-Now, figure out the size of the box you would need if you
-wanted it to fit all the data on the coarsest level.  Now, figure out what
-the largest integer $n$ is so that you can evenly divide the length of this box
-by $2^n$ in every coordinate direction.  If $n < 2$, the program will abort
-since the grid structure is not suitable for this bottom solver.
-
-Now the code will set up a ``new'' problem, using the data at the coarsest level
-of the original problem as the initial data.  The grid structure for this new
-problem has the same number of cells as the coarsest level of the original problem,
-but the data is copied onto a grid structure where each grid has $2^n$ cells
-on each side.  The new V-cycle continues down to the new coarsest level, in
-which each grid has 2 cells on each side.  If you wish to impose a limit on
-the maximum value that $n$ can have, you can do so by setting 
-{\tt max\_mg\_bottom\_nlevs} equal to that value.
-
-{\bf Example 1:} A 3D problem with $384^3$ cells divided into $32^3$ grids, i.e., 
-there is a $12\times 12\times 12$ block of $32^3$ grids.  The coarsest level of the 
-multigrid V-cycle contains $12\times 12\times 12$ grids that have $2^3$ cells, so the 
-entire problem domain has $24^3$ cells.  We see that $n=3$, and create a new problem 
-domain with a $3\times 3\times 3$ block of $8^3$ grids.  The coarsest level of the 
-multigrid V-cycle for the ``new'' problem will be a $3\times 3\times 3$ block of 
-$2^3$ grids.
-
-{\bf Example 2:} A 2D problem with $96\times 384$ cells divided into $48^2$ grids, i.e., 
-there is a $2\times 8$ block of $48^2$ grids.  The coarsest level of the multigrid 
-V-cycle contains $2\times 8$ grids that have $3^2$ cells, so the entire problem 
-domain has $6\times 24$ cells.  We see that $n=1$, so the program aborts since this grid
-structure is not appropriate for the fancy bottom solver.
-
-\section{Multilevel and Refinement Criteria}
-
-\section{Particles}
-
-\label{arch:sec:particles}
-
-\maestro\ has support for Lagrangian particles that are passively
-advected with the velocity field.  These are useful for diagnostics
-and post-processing.  To use particles, particles must be seeded into
-the domain by writing a problem-specific {\tt init\_particles.f90}
-routine.  This routine is called at the start of the simulation.  The
-{\tt init\_particles} routines add particles at specific locations by
-calling the {\tt particle\_module}'s {\tt add} routine when a given
-criteria is met by the fluid state.
-
-When you run the code, particles are enabled by setting {\tt
-  use\_particles = T}.  At the end of each timestep the locations of
-all the particles are written out into a series of files called {\tt
-  timestamp\_NN}, where {\tt NN} is the CPU number on which the
-particle {\em currently} resides.  Particles are always kept on the
-processor containing the state data corresponding to their present
-location.  Several bits of associated data (density, temperature, and
-mass fractions) are stored along with the particle ID and position.
-
-Some simple python scripts allow for the plotting of the particle
-positions.  See \S~\ref{analysis:sec:particles} for details.
+  This tests the hgprojection routine in 2-d.  A divergence-free
+  velocity field is initialized and then ``polluted'' by adding the
+  gradient of a scalar.  The form of the scalar differs depending on
+  the boundary conditions (wall and periodic are supported currently).
+  Finally, the hgproject routine is called to recover the initial
+  divergence-free field.

### 2012-02-27 docs/architecture/architecture.tex
new constants location
add unit test chapter

@@
 
-  \item {\tt constants/}
-
-    A simple module to define physical constants
-
   \item {\tt EOS/}
@@
 
+  \item {\tt constants/}
+
+    A simple module to define physical constants
+
   \item {\tt docs/}
@@
   \end{itemize}
-
-\end{itemize}
-
-\section{Unit Tests}
-
-\label{sec:unit_tests}
-
-In addition to the problems described above which use the full
-capabilities of \maestro, there are a number of unit tests that
-exercise only specific components of the \maestro\ solvers.  The
-tests have their own drivers (a custom {\tt varden.f90}) that
-initialize only the data needed for the specific test and call
-specific \maestro\ routines directly.
-
-\begin{itemize}
-\item {\tt test\_advect} \\[-3mm]
-
-  This test initializes a Gaussian density field (no other scalar
-  quantities are used) and a uniform velocity field in any one of the
-  coordinate directions.  The Gaussian profile is advected through
-  the period domain exactly once and the error in the density profile
-  (L2 norm) is computed.  The driver for this problem does this 
-  for every dimension twice (once with a positive velocity and once
-  with a negative velocity).  After all coordinate directions are 
-  tested, the norms are compared to ensure that the error does
-  not show any directional bias.
-
-\item {\tt test\_average} \\[-3mm]
-
-  This test initializes a 1D radial base state quantity with a
-  Gaussian distribution, maps it into the 3D domain (assuming a
-  spherical geometry) using the routines provided by the {\tt
-    fill\_3d\_module} module, and then calls {\tt average} to put it
-  back onto a 1D radial array.  This way we test the accuracy of our
-  procedure to map between the 1D radial and 3D Cartesian states.
-  The output from this test was described in detail in
-  \cite{multilevel}.
-
-\item {\tt test\_basestate} \\[-3mm]
-
-  This test initializes the base state to contain a hydrostatic
-  model and then evolves the state with heating to watch the 
-  hydrostatic adjustment of the atmosphere.  In particular,
-  the base state velocity, $w_0$, is computed in response to 
-  the heating and this is used to advect the base state density
-  and compute the new pressure, $p_0$.  An early version of 
-  this routine was used for the plane-parallel expansion test
-  in \cite{lowMach2}.  This version of the test was also shown
-  for a spherical, self-gravitating star in \cite{multilevel}.
-
-\item {\tt test\_diffusion}
-
-  This test initializes a Gaussian temperature profile and calls
-  the thermal diffusion routines in \maestro\ to evolve the state 
-  considering only diffusion.  The driver estimates a timestep
-  based on the explicit thermal diffusion timescale and loops
-  over calls to the thermal diffusion solver.  A Gaussian remains
-  Gaussian when diffusing, so an explicit error can be computed
-  by comparing to the analytic solution.  This test is 
-  described in \cite{xrb}.
-
-\item {\tt test\_particles}
-
-  This test exercises the particle advection routine.  A simple
-  circular velocity field, with the magnitude increasing with radius
-  from the center is initialized.  A number of particles are then
-  initialized at various radii from the center and they are advected
-  for one period.  The particle paths should be perfect circles, and
-  the final particle position should overlap with the initial
-  position.
-
-  Particle data is stored separately from the fluid data.  Instead
-  of being part of the plotfiles, the particle data is outputted
-  each timestep into files named {\tt timestamp\_*}, where 
-  the number indicates which processor did the writing.  These
-  particle files can be processed and the particle data plotted
-  using the python routines in {\tt data\_processing/python/}.
 

### 2012-03-01 docs/SDC/sdc.tex
new notation

@@
 
-{\bf Step 1B:} ({\it Compute advective fluxes}) Use $\uadvsdcpred$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,{\rm pred}}$, with time-lagged reactions, explicit diffusion, and time-centered thermodynamic pressure as source terms.\\
+{\bf Step 1B:} ({\it Compute advective fluxes}) Use $\uadvsdcpred$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,{\rm pred}}$, with time-lagged reactions ($I^{n-1\rightarrow n}$), explicit diffusion, and time-centered thermodynamic pressure as source terms.
+Define the advective flux divergences as
+\begin{eqnarray}
+A_{\rho X_k}^{\rm pred} &=& -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right],\\
+A_{\rho h}^{\rm pred} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right].
+\end{eqnarray}
 
@@
 \begin{equation}
-\frac{\rho^{n+1,{\rm pred}} - \rho^n}{\Delta t} = -\sum_m\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right].
+\frac{\rho^{n+1,{\rm pred}} - \rho^n}{\Delta t} = \sum_mA_{\rho X_k}^{\rm pred}.
 \end{equation}
@@
 \begin{equation}
-\frac{\rho^{n+1,{\rm pred}}X_{k,{\rm AD}}^{n+1,{\rm pred}} - (\rho X_k)^n}{\Delta t} = -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right] + I_{\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
+\frac{\rho^{n+1,{\rm pred}}X_{k,{\rm AD}}^{n+1,{\rm pred}} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{\rm pred} + I_{\rho X_k}^{n-1\rightarrow n}.\label{eq:sdc species 2}
 \end{equation}
@@
 \begin{eqnarray}
-\frac{\rho^{n+1,{\rm pred}}h^{n+1,{\rm pred}}_{\rm AD} - (\rho h)^n}{\Delta t} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right] + \uadvsdcpred\cdot\nabla p_0 + I_{\rho h}^{\rm lagged}\nonumber\\
+\frac{\rho^{n+1,{\rm pred}}h^{n+1,{\rm pred}}_{\rm AD} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{\rm pred} + \uadvsdcpred\cdot\nabla p_0 + I_{\rho h}^{n-1\rightarrow n}\nonumber\\
 && + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n + \nabla\cdot\frac{\kth^n}{c_p^n}\nabla h_{\rm AD}^{n+1,{\rm pred}}\right)\nonumber\\
@@
 
-{\bf Step 1D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}^{\rm pred}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{\rm pred}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{\rho h}^{\rm lagged}$ term.
+{\bf Step 1D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}^{\rm pred}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{\rho X_k}^{n-1\rightarrow n}$ term, and define $Q_{\rho h}^{\rm pred}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{\rho h}^{n-1\rightarrow n}$ term.
 Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,{\rm pred}}$ using the piecewise-constant advection and diffusion source terms:
@@
 \end{eqnarray}
-At this point we can define $I_{\rho X_k}^{\rm pred}$ and $I_{\rho h}^{\rm pred}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the corrector step.\\
+At this point we can define $I_{\rho X_k}^{n\rightarrow n+1,{\rm pred}}$ and $I_{\rho h}^{n\rightarrow n+1,{\rm pred}}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the corrector step.\\
 
 {\bf Step 2:} ({\it Corrector}) In the corrector, we use the time-advanced state from the predictor to perform a more accurate integration of the thermodynamic variables.
-The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^{n+1,{\rm pred}}$ are computed from $(\rho,\rho X_k,\rho h)^{n+1,{\rm pred}}$.
+The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^{n+1,{n\rightarrow n+1,{\rm pred}}}$ are computed from $(\rho,\rho X_k,\rho h)^{n+1,{n\rightarrow n+1,{\rm pred}}}$.
 This step is divided into several sub-steps:\\\\
@@
 \begin{equation}
-S^{n+\myhalf} = \frac{S^n + S^{n+1,{\rm pred}}}{2}, \qquad \beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1,{\rm pred}}}{2}.
+S^{n+\myhalf} = \frac{S^n + S^{n+1,{n\rightarrow n+1,{\rm pred}}}}{2}, \qquad \beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1,{n\rightarrow n+1,{\rm pred}}}}{2}.
 \end{equation}
@@
 
-{\bf Step 2B:} ({\it Compute advective fluxes}) Use $\uadvsdc$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf}$, with iteratively-lagged reactions, explicit diffusion, and time-centered thermodynamic pressure as source terms.\\
+{\bf Step 2B:} ({\it Compute advective fluxes}) Use $\uadvsdc$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf}$, with iteratively-lagged reactions ($I^{n\rightarrow n+1,{\rm pred}}$), explicit diffusion, and time-centered thermodynamic pressure as source terms.
+Define the advective flux divergences as
+\begin{eqnarray}
+A_{\rho X_k} &=& -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf}\uadvsdcpred\right],\\
+A_{\rho h} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf}\uadvsdcpred\right].
+\end{eqnarray}
 
@@
 \begin{equation}
-\frac{\rho^{n+1} - \rho^n}{\Delta t} = -\sum_m\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf}\uadvsdc\right].
+\frac{\rho^{n+1} - \rho^n}{\Delta t} = -\sum_mA_{\rho X_k}.
 \end{equation}
@@
 \begin{equation}
-\frac{\rho^{n+1}X_{k,{\rm AD}}^{n+1} - (\rho X_k)^n}{\Delta t} = -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf}\uadvsdc\right] + I_{\rho X_k}^{\rm pred}.\label{eq:sdc species 3}
+\frac{\rho^{n+1}X_{k,{\rm AD}}^{n+1} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k} + I_{\rho X_k}^{n\rightarrow n+1,{\rm pred}}.\label{eq:sdc species 3}
 \end{equation}
@@
 \begin{eqnarray}
-\frac{\rho^{n+1}h^{n+1}_{\rm AD} - (\rho h)^n}{\Delta t} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf}\uadvsdc\right] + \uadvsdc\cdot\nabla p_0 + I_{\rho h}^{\rm pred}\nonumber\\
+\frac{\rho^{n+1}h^{n+1}_{\rm AD} - (\rho h)^n}{\Delta t} &=& A_{\rho h} + \uadvsdc\cdot\nabla p_0 + I_{\rho h}^{n\rightarrow n+1,{\rm pred}}\nonumber\\
 && + \nabla\cdot\frac{\kth^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla h^{n+1}_{\rm AD} + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla h^{n+1,{\rm pred}}\right)\nonumber\\
@@
 
-{\bf Step 2D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{\rho X_k}^{\rm pred}$ term, and define $Q_{\rho h}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{\rho h}^{\rm pred}$ term.
+{\bf Step 2D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{\rho X_k}^{n\rightarrow n+1,{\rm pred}}$ term, and define $Q_{\rho h}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{\rho h}^{n\rightarrow n+1,{\rm pred}}$ term.
 Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1}$ using the piecewise-constant advection and diffusion source terms:
@@
 \end{eqnarray}
-At this point we can define $I_{\rho X_k}^{\rm lagged}$ and $I_{\rho h}^{\rm lagged}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the predictor in the next time step.\\
+At this point we can define $I_{\rho X_k}^{n\rightarrow n+1}$ and $I_{\rho h}^{n\rightarrow n+1}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the predictor in the next time step.\\
 

### 2012-03-14 docs/FAQ/faq.tex
explain what make TAGS and make tags do in the makefiles and how to use them -- very cool feature pointed out by Adam

@@
+\section{Coding}
+
+\begin{enumerate}
+
+\item {\em How do tags work when editing source?}
+
+   Tags allow the editor to follow function/subroutine names an
+   automatically switch you to the source code corresponding to that
+   function.  Using tags in \maestro\ depends on the editor:
+
+   \begin{itemize}
+
+     \item {\tt vi}:
+
+     In the build directory, type `{\tt make tags}'.  Then in {\tt
+       vi}, move the cursor over a function name and use {\tt
+       \verb|^|-]} to bring up the source corresponding to that
+       function.  Use {\tt \verb|^|-t} to go back.  (Here, {\tt \verb|^|-}
+       refers to the Control key.)
+
+     \item {\tt emacs}:
+
+     In the build directory, type `{\tt make TAGS}'.  Then in {\tt
+       emacs}, move the cursor over a function name and use {\tt M-.}\
+       to bring up the source corresponding to that function.  Use
+       {\tt M-*} to go back.  (Here, {\tt M-} refers to the META key.)
+
+   \end{itemize}
+
+\end{enumerate}
+
 \section{Compiling}

### 2012-03-15 docs/architecture/architecture.tex
updated directory structure fig
some BC description

@@
 
-\begin{figure}[t]
+\begin{figure}[]
 \centering
-\includegraphics[scale=0.65]{\archfigpath/maestro_directory2}
+\includegraphics[scale=0.55]{\archfigpath/maestro_ecosystem2}
 \caption[\maestro\ directory structure]
@@
 
+When \maestro\ is run, the boundary condition parameters are read in
+from the input file and used to build the \bctower.  The
+\bctower\ consists of a \bclevel\ object for each level of resolution
+in the simulation.  The \bclevel\ contains 3 difference descriptions
+of the boundary conditions for each box in the domain at that level of
+refinement: {\tt phys\_bc\_level\_array}, {\tt
+  adv\_bc\_level\_array}, and {\tt ell\_bc\_level\_array}.
+
+Each level has a {\tt phys\_bc\_level\_array(0:ngrids,dim,2)} array,
+where {\tt ngrids} is the number of boxes on that level.  This stores
+the physical desciption of the boundary type (outlet, inlet, slipwall,
+etc.)---this description is independent of the variables that live on
+the grid.  The {\tt phys\_bc\_level\_array(0,:,:)} `box' refers to the
+entire domain.  If an edge of a box is not on a physical boundary,
+then it is set to a default value (typically {\tt INTERIOR}).  
+These boundary condition types are used to interpret the actual 
+method to fill the ghostcells for each variable, as described in 
+{\tt adv\_bc\_level\_array} and {\tt ell\_bc\_level\_array}.
+
+The convention is that the first {\tt dm} variables in the
+\bclevel\ refer to the velocity components, and the subsequent slots
+are for the other standard variables described in the {\tt
+  variables\_module}.  For instance, to reference the boundary
+condition for density, one would index with {\tt dm+rho\_comp}.
+For temporary variables that are created on the fly in the various
+routines in \maestro\ there may not be a variable name in {\tt
+  variables\_module} that describes the temporary variable.  In this
+case, the special variables {\tt foextrap\_comp} and {\tt
+  hoextrap\_comp} (first-order and high-order extrapolation) are used.
+
+The actual filling of the ghostcells according to the descriptions
+contained in the \bctower\ is carried out by the {\tt setbc} routine,
+which is called by {\tt multifab\_physbc}.
+
 \section{Multigrid}

### 2012-03-15 docs/architecture/architecture.tex
more BCs

@@
 Each level has a {\tt phys\_bc\_level\_array(0:ngrids,dim,2)} array,
-where {\tt ngrids} is the number of boxes on that level.  This stores
-the physical desciption of the boundary type (outlet, inlet, slipwall,
-etc.)---this description is independent of the variables that live on
-the grid.  The {\tt phys\_bc\_level\_array(0,:,:)} `box' refers to the
-entire domain.  If an edge of a box is not on a physical boundary,
-then it is set to a default value (typically {\tt INTERIOR}).  
-These boundary condition types are used to interpret the actual 
-method to fill the ghostcells for each variable, as described in 
-{\tt adv\_bc\_level\_array} and {\tt ell\_bc\_level\_array}.
-
-The convention is that the first {\tt dm} variables in the
-\bclevel\ refer to the velocity components, and the subsequent slots
-are for the other standard variables described in the {\tt
-  variables\_module}.  For instance, to reference the boundary
-condition for density, one would index with {\tt dm+rho\_comp}.
-For temporary variables that are created on the fly in the various
-routines in \maestro\ there may not be a variable name in {\tt
-  variables\_module} that describes the temporary variable.  In this
-case, the special variables {\tt foextrap\_comp} and {\tt
-  hoextrap\_comp} (first-order and high-order extrapolation) are used.
+where {\tt ngrids} is the number of boxes on that level and the last
+index refers to the lower (1) or upper (2) edge of the box in the given
+dimension.  This stores the physical desciption of the boundary type
+(outlet, inlet, slipwall, etc.)---this description is independent of
+the variables that live on the grid.  The {\tt
+  phys\_bc\_level\_array(0,:,:)} `box' refers to the entire domain.
+If an edge of a box is not on a physical boundary, then it is set to a
+default value (typically {\tt INTERIOR}).  These boundary condition
+types are used to interpret the actual method to fill the ghostcells
+for each variable, as described in {\tt adv\_bc\_level\_array} and
+{\tt ell\_bc\_level\_array}.
+
+Whereas {\tt phys\_bc\_level\_array} provides a physical description
+of the type of boundary, {\tt adv\_bc\_level\_array} describes the
+{\em action} taken (e.g.\ reflect, extrapolate, etc.) for each
+variable when filling boundaries.  {\tt adv\_bc\_level\_array} is
+describes the boundary conditions that are in play for the advection
+(hyperbolic) equations.  The form of this array is {\tt
+  adv\_bc\_level\_array(0:ngrids,dim,2,nvar)} where the new component,
+{\tt nvar}, allows for each state variable that lives on a grid to
+have different boundary condition actions associated with it.  The
+convention is that the first {\tt dm} variables in the \bclevel\ refer
+to the velocity components, and the subsequent slots are for the other
+standard variables described in the {\tt variables\_module}.  For
+instance, to reference the boundary condition for density, one would
+index with {\tt dm+rho\_comp}.  For temporary variables that are
+created on the fly in the various routines in \maestro\ there may not
+be a variable name in {\tt variables\_module} that describes the
+temporary variable.  In this case, the special variables {\tt
+  foextrap\_comp} and {\tt hoextrap\_comp} (first-order and high-order
+extrapolation) are used.
+
+{\tt ell\_bc\_level\_array} is the analog to {\tt adv\_bc\_level\_array}
+for the elliptic solves in \maestro.  This will come into play in the
+multigrid portions of the code.  The actions that are used for
+{\tt ell\_bc\_level\_array} are either Dirichlet or Neumann boundary
+condtions.
 

### 2012-03-15 docs/architecture/architecture.tex
edits

@@
 {\em action} taken (e.g.\ reflect, extrapolate, etc.) for each
-variable when filling boundaries.  {\tt adv\_bc\_level\_array} is
-describes the boundary conditions that are in play for the advection
+variable when filling boundaries.  {\tt adv\_bc\_level\_array} 
+specifically describes the boundary conditions that are in play for the advection
 (hyperbolic) equations.  The form of this array is {\tt
-  adv\_bc\_level\_array(0:ngrids,dim,2,nvar)} where the new component,
+  adv\_bc\_level\_array(0:ngrids,dim,2,nvar)} where the additional component,
 {\tt nvar}, allows for each state variable that lives on a grid to

### 2012-03-15 docs/architecture/architecture.tex
more BC smoothing

@@
 \bctower\ consists of a \bclevel\ object for each level of resolution
-in the simulation.  The \bclevel\ contains 3 difference descriptions
-of the boundary conditions for each box in the domain at that level of
-refinement: {\tt phys\_bc\_level\_array}, {\tt
-  adv\_bc\_level\_array}, and {\tt ell\_bc\_level\_array}.
+in the simulation.  The \bclevel\ contains 3 different descriptions of
+the boundary conditions for each box in the domain at that level of
+refinement: {\tt phys\_bc\_level\_array}, {\tt adv\_bc\_level\_array},
+and {\tt ell\_bc\_level\_array}.  In all cases, the boundary
+conditions are specified via integer values that are defined in {\tt
+  bc\_module} (part of \boxlib).
 
 Each level has a {\tt phys\_bc\_level\_array(0:ngrids,dim,2)} array,
-where {\tt ngrids} is the number of boxes on that level and the last
-index refers to the lower (1) or upper (2) edge of the box in the given
-dimension.  This stores the physical desciption of the boundary type
-(outlet, inlet, slipwall, etc.)---this description is independent of
-the variables that live on the grid.  The {\tt
-  phys\_bc\_level\_array(0,:,:)} `box' refers to the entire domain.
-If an edge of a box is not on a physical boundary, then it is set to a
-default value (typically {\tt INTERIOR}).  These boundary condition
-types are used to interpret the actual method to fill the ghostcells
-for each variable, as described in {\tt adv\_bc\_level\_array} and
-{\tt ell\_bc\_level\_array}.
+where {\tt ngrids} is the number of boxes on that level, {\tt dim} is
+the coordinate direction, and the last index refers to the lower (1)
+or upper (2) edge of the box in the given coordinate direction.  This
+stores the {\em physical desciption} of the boundary type (outlet, inlet,
+slipwall, etc.)---this description is independent of the variables
+that live on the grid.  The {\tt phys\_bc\_level\_array(0,:,:)} `box'
+refers to the entire domain.  If an edge of a box is not on a physical
+boundary, then it is set to a default value (typically {\tt
+  INTERIOR}).  These boundary condition types are used to interpret
+the actual method to fill the ghostcells for each variable, as
+described in {\tt adv\_bc\_level\_array} and {\tt
+  ell\_bc\_level\_array}.
 
 Whereas {\tt phys\_bc\_level\_array} provides a physical description
-of the type of boundary, {\tt adv\_bc\_level\_array} describes the
-{\em action} taken (e.g.\ reflect, extrapolate, etc.) for each
-variable when filling boundaries.  {\tt adv\_bc\_level\_array} 
-specifically describes the boundary conditions that are in play for the advection
-(hyperbolic) equations.  The form of this array is {\tt
-  adv\_bc\_level\_array(0:ngrids,dim,2,nvar)} where the additional component,
-{\tt nvar}, allows for each state variable that lives on a grid to
-have different boundary condition actions associated with it.  The
-convention is that the first {\tt dm} variables in the \bclevel\ refer
-to the velocity components, and the subsequent slots are for the other
+of the type of boundary, the array {\tt adv\_bc\_level\_array}
+describes the {\em action} taken (e.g.\ reflect, extrapolate, etc.)
+for each variable when filling boundaries.  {\tt
+  adv\_bc\_level\_array} specifically describes the boundary
+conditions that are in play for the advection (hyperbolic) equations.
+The form of this array is {\tt
+  adv\_bc\_level\_array(0:ngrids,dim,2,nvar)} where the additional
+component, {\tt nvar}, allows for each state variable that lives on a
+grid to have different boundary condition actions associated with it.
+The convention is that the first {\tt dm} variables in \bclevel (where {\tt dm} is
+the dimensionality of the simulation) refer to the
+velocity components, and the subsequent slots are for the other
 standard variables described in the {\tt variables\_module}.  For
@@
 
-{\tt ell\_bc\_level\_array} is the analog to {\tt adv\_bc\_level\_array}
-for the elliptic solves in \maestro.  This will come into play in the
-multigrid portions of the code.  The actions that are used for
-{\tt ell\_bc\_level\_array} are either Dirichlet or Neumann boundary
-condtions.
+{\tt ell\_bc\_level\_array} is the analog to {\tt
+  adv\_bc\_level\_array} for the elliptic solves in \maestro.  This
+will come into play in the multigrid portions of the code.  The
+actions that are used for {\tt ell\_bc\_level\_array} are either
+Dirichlet or Neumann boundary condtions.  For the velocity
+projections, we are dealing with a pressure-like quantity, $\phi$, so
+the pressure boundary conditions here reflect the behavior we want for
+the velocity.  You want $d\phi/dn=0$ in places where the incoming
+velocity to the elliptic solve must match the velocity you want after
+the projection (e.g.\ inflow, solid walls).  For outlet boundaries,
+setting $\phi = 0$ allows the velocity to adjust as needed to the
+domain. \MarginPar{reference here would be nice}
 

### 2012-03-15 docs/architecture/architecture.tex
add ref for BCs

@@
 A \bctower\ holds the information about what boundary conditions are
-in effect for each variable in a \maestro\ simulation.  \MarginPar{better description?}
+in effect for each variable in a
+\maestro\ simulation.  These are interpretted by the ghost cell filling 
+routines.  See \S~\ref{sec:arch:bcs} for more detail.
 
@@
 \section{Boundary Conditions}
+
+\label{sec:arch:bcs}
 
@@
 setting $\phi = 0$ allows the velocity to adjust as needed to the
-domain. \MarginPar{reference here would be nice}
+domain (see, for example, \cite{almgrenBellSzymczak:1996}).
 

### 2012-03-16 docs/runtime_parameters/runtime_parameters.tex
update

@@
 \verb=  use_eos_coulomb  = &   use the Coulomb corrections between ions (if supported by the EOS)  &  .true. \\
+\verb=  use_eos_e_instead_of_h  = &   In deriving the temperature from the $h$, first subtract off $p_0/\rho$ to define $e$, and use that as the input to the EOS.  &  .false. \\
 
@@
 &  5 \\
+\rowcolor{tableShade}
+\verb=  xhi_boundary_type  = &   $+x$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcx\_hi}  &  "" \\
+\verb=  xlo_boundary_type  = &   $-x$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcx\_lo}  &  "" \\
+\rowcolor{tableShade}
+\verb=  yhi_boundary_type  = &   $+y$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcy\_hi}  &  "" \\
+\verb=  ylo_boundary_type  = &   $-y$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcy\_lo}  &  "" \\
+\rowcolor{tableShade}
+\verb=  zhi_boundary_type  = &   $+z$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcz\_hi}  &  "" \\
+\verb=  zlo_boundary_type  = &   $-z$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcz\_lo}  &  "" \\
 

### 2012-03-18 docs/test_suite/test_suite.tex
documentation for the test suite

\section{Overview}

An automated regression test suite for \maestro\ (or any BoxLib-based
code) written in Python exists in {\tt BoxLib/Tools/C\_util/regtests/}
as {\tt testnew.py}.  The test suite consists of a set of problem
definitions (the \maestro\ problem + their inputs file, etc.).  When
the suite is run the first time, the plotfiles created at the end of
each problem's executation is stored as a benchmark.  After this
initialization, each subsequent run of the test suite compares the
current output of the code, level-by-level and zone-by-zone to the
stored benchmarks (using the {\tt fcompare.f90} routine in {\tt
  AmrPostprocessing/F\_Src/}).  Any differences are flagged as errors.
A web page report is generated by the test suite and provides a
history of the regression testing.  Single-processor and parallel test
problems, compilation tests, self-tests (problems that determine by
themselves whether they were successfully), and testing restarting
from a checkpoint file are supported.

\begin{figure}[t]
\centering
\includegraphics[width=5.0in]{\suiteFigPath/testsuite}
\caption{\label{fig:test_suite_main} Main test suite results page.  Each 
row indicates a single test suite run, arranged by date, and each column
indicates a different test problem. }
\end{figure}

\section{Test Suite Inputs File}

The inputs file for the test suite separates the problems into blocks.
The header of a problem block has the form {\tt [problem-name]}.
Beneath each problem block, there are a number of options set for each
problem.  A separate heading, {\tt [main]}, is used for the suite-wide
options.

An example of the {\tt main} block from {\tt Maestro-tests.ini} is:
\begin{lstlisting}
[main]
boxLibDir      = /home/zingale/tests/BoxLib/
sourceDir      = /home/zingale/tests/MAESTRO/
testTopDir     = /home/zingale/tests/
webTopDir      = /home/www/Maestro/test-suite/
compareToolDir = /home/zingale/tests/AmrPostprocessing/F_Src/
helmeosDir     = /home/zingale/tests/MAESTRO/fParallel/extern/EOS/helmeos/

sourceTree = fParallel

suiteName = Maestro

FCOMP = gfortran

MPIcommand = mpiexec -n @nprocs@ @command@
\end{lstlisting}

The first group of options define the necessary paths.  Here, {\tt
  boxLibDir} points to the location of the \boxlib\ source, {\tt
  sourceDir} points to the top-level \maestro\ directory, which is
expected to contain the {\tt fParallel} subdirectories.  {\tt
  testTopDir} refers to the directory that the suite should use as its
root directory for running the tests and {\tt webTopDir} is the
directory in which to place the web output.  The {\tt fcompare.f90}
comparison tool source is expected to be found in {\tt
  compareToolDir}.  Finally, {\tt helmeosDir} lists the path to the
{\tt helm\_table.dat} file used by the general stellar equation of
state.

Next, we set {\tt sourceTree} to {\tt fParallel}, indicating that
\maestro\ is built using the Fortran \boxlib\ framework.  This option
tells the test suite what build system to use.

{\tt FCOMP} tell the test suite which Fortran compilers to use.  These
override what is listed in any {\tt GNUmakefile} to ensure that the
compiler stays consistent in the tests.

The {\tt suiteName} option simply tells the test suite what name to
prefix to the output directories.  It does not need to match the
program name.

Finally, {\tt MPIcommand} lists the generic manner in which to run an
MPI program on the target system.  If present, the string {\tt @host@}
in the {\tt MPIcommand} will be substituted by the {\tt MPIhost}
string by the test suite.  Similarly the {\tt @nprocs@} string will be
substituted by the number of processors, which is set on a
problem-by-problem basis.  Finally, the {\tt MPIcommand} should
include the string {\tt @command@}, which is where the
\maestro\ executable and inputs file will be substituted.  For single
processor runs, these options are ignored.

Each problem to be run by the test suite gets its own block.  For
example, a test listing for the {\tt test2} problem might look like:

\begin{lstlisting}
[test2]
buildDir = fParallel/MAESTRO/TEST_PROBLEMS/test2/
inputFile = inputs_2d
aux1File = model.hse.cool.coulomb
needsHelmEOS = 1
dim = 2
doVis = 1
visVar = ``tfromp''
compileTest = 0 
restartTest = 0
useMPI = 0
\end{lstlisting}

Here {\tt test2} contained inside the {\tt []} is the name of the
problem, as the test suite will refer to it.  {\tt buildDir} is the
path beneath {\tt sourceDir} where the {\tt make} command should be
executed.  The inputs file is given by {\tt inputFile} and the
necessary model file is listed by {\tt aux1File}. (Not all problems
will need this.  Additional auxillary files can be specified via {\tt
  aux2File} and {\tt aux3File}).  Note that {\tt inputFile} and {\tt aux?File} are
specified relative to the {\tt buildDir}. The dimensionality is specified by {\tt dim}.

A number of options are available:
\begin{itemize}
\item  If the general stellar equation of state is
used, then {\tt needsHelmEOS} should be set to {\tt 1} to ensure that the
EOS table is linked into the run directory.

\item {\tt link1File}, {\tt link2File}, and {\tt link3File} work like
the {\tt aux?File}s, but instead of copying the specified file into
the run directory, a symbolic link is made.

\item If the test is to be run in parallel,
the {\tt useMPI} should be {\tt 1} and {\tt numprocs} should give the number
of processors.

\item To test the compilation of the problem only (and skip running),
set {\tt compileTest} to {\tt 1}.

\item To test the ability of the code to restart, set {\tt restartTest} to
{\tt 1}.  Also set {\tt restartFileNum} to the number of the checkpoint file to restart
from.  The suite will run the problem as usual and then restart from the specified
checkpoint and run to completion again.  The output from the initial run will
then be compared to the output from the restart.  In a restart test, there
is no stored benchmark.

\item Some problems don't output plotfiles, but instead check internally
whether they are successful.  For these, set {\tt selfTest} to {\tt 1},
and set {\tt stSuccessString} to the string output to look for at the
end of the test to determine if it was successful.

\item To add a simple visualization to the test suite webpage, set
  {\tt doVis} to {\tt 1}, and set {\tt visVar} to the name of the
  plotfile variable to visualize.  An image of that field from the
  last plotfile will be appended to the problem's test webpage.

\item Ordinarily, the test suite uses the last plotfile output to compare to.
To force the comparison to a specific file, set {\tt compareFile} to the 
name of the file to compare to.

\item To override some of the options in the problem's {\tt GNUmakefile}
(e.g.\ use a different reaction network), set {\tt addToCompileString}
to the  string to add to the {\tt make} line that compiles the problem.

\end{itemize}

\section{Initializing the Test Suite}

The first time you run the test suite there are no benchmark files to compare to.
Once you generate an inputs file, as described above, you would simply run the
suite as: \\
$~~~~~${\tt ./testnew.py --make\_benchmarks "initial run" ./Castro\_tests.ini} \\
The string following {\tt --make\_benchmarks} is simply a comment that will
be added to the web report.
This command creates three output directories, using the {\tt suiteName} as the prefix.
\begin{itemize}
\item {\tt suiteName-tests} is where the tests are run.  Each time the test
 suite is run, a subdirectory, based on the date, is created, with a subdirectory
 for each test.  All the files necessary to run the test are copied into the
 test subdirectory.

\item {\tt suiteName-web} is where the web-based reports for the test are generated.
 The master webpage is {\tt suiteName-web/index.html}.

\item {\tt suiteName-benchmarks} is where the test benchmark files are stored.  This
 are used for comparison to the current output.
\end{itemize}

\section{Regular Use}

Once the initial benchmarks are created, you can compare the current
version of the code to the stored results by simply doing: \\
$~~~~~${\tt ./testnew.py ./Maestro\_tests.ini} \\ 
This will do a CVS/git update, generate {\tt ChangeLog} files listing all
of the CVS comments for the code, build the test comparison tools, and
then loop over each test, building and running the executable and
comparing the output to the benchmarks as required.

\begin{figure}[t]
\centering
%\includegraphics[width=5.0in]{CastroVerification/test_suite_day}
\caption{\label{fig:test_suite_date} The test suite output for a
  single day's run.  Each row indicates a separate test, showing
  whether they passed or failed.  Clicking on the test name will give
  more information about that particular test on that day. Note: this
  page is from the {\tt Maestro} code, but a {\tt Castro} test suite
  run will produce similar output.}
\end{figure}

Upon completion of all the runs, a web page for this invocation of the
test suite will be generated (see figure~\ref{fig:test_suite_date}),
as well as pages showing the details for each of the problems run.
Test failures indicate that the current output does not match the stored
benchmarks.

\section{Updating Benchmarks}

A test failure means that the current version of the code gives a
different answer than the stored benchmark.  A test can fail either
because a bug was introduced into the code or a bug was fixed or new
feature introduced.

If a bug was introduced into the code recently, then by examing the
test history you can determine the time period in which the bug was
introduced.  The {\tt ChangeLog}s linked to on each test date's webpage
will list all the changes committed to CVS/git up to that point, which is
useful for tracking down the bug.  Once the bug is fixed, rerunning
the suite should generate a `pass'.

If a bug was fixed or a new feature was introduced, and you are
confident that the latest output is correct, then you can tell the
test suite to update the benchmarks.  If you want to do this for all
the test problems, you would do:\\
$~~~~~${\tt ./testnew.py --make\_benchmarks "X bug fixed" ./Castro\_tests.ini} \\
where the string after ``{\tt --make\_benchmarks}'' is a note that is listed
on the regression suite web page describing the reason for the benchmark
update.  Subsequent runs of the test suite will use the new benchmarks.
If you only want to update the benchmarks of a single test, then you
can use the ``{\tt --single\_test test}'' flag on the commandline, where
{\tt test} is the name of the test to update.

### 2012-03-18 docs/test_suite/test_suite.tex
update

@@
 the suite is run the first time, the plotfiles created at the end of
-each problem's executation is stored as a benchmark.  After this
+each problem's execution is stored as a benchmark.  After this
 initialization, each subsequent run of the test suite compares the
@@
 necessary model file is listed by {\tt aux1File}. (Not all problems
-will need this.  Additional auxillary files can be specified via {\tt
+will need this.  Additional auxiliary files can be specified via {\tt
   aux2File} and {\tt aux3File}).  Note that {\tt inputFile} and {\tt aux?File} are
@@
 
-\begin{figure}[t]
-\centering
-%\includegraphics[width=5.0in]{CastroVerification/test_suite_day}
-\caption{\label{fig:test_suite_date} The test suite output for a
-  single day's run.  Each row indicates a separate test, showing
-  whether they passed or failed.  Clicking on the test name will give
-  more information about that particular test on that day. Note: this
-  page is from the {\tt Maestro} code, but a {\tt Castro} test suite
-  run will produce similar output.}
-\end{figure}
-
 Upon completion of all the runs, a web page for this invocation of the
-test suite will be generated (see figure~\ref{fig:test_suite_date}),
-as well as pages showing the details for each of the problems run.
-Test failures indicate that the current output does not match the stored
-benchmarks.
+test suite will be generated as well as pages showing the details for
+each of the problems run.  Test failures indicate that the current
+output does not match the stored benchmarks.
 
@@
 
-If a bug was introduced into the code recently, then by examing the
+If a bug was introduced into the code recently, then by examining the
 test history you can determine the time period in which the bug was

### 2012-04-08 docs/architecture/architecture.tex
remove setbc

@@
 The actual filling of the ghostcells according to the descriptions
-contained in the \bctower\ is carried out by the {\tt setbc} routine,
-which is called by {\tt multifab\_physbc}.
+contained in the \bctower\ is carried out by the {\tt multifab\_physbc}
+routine.
 

### 2012-04-11 docs/analysis/analysis.tex
point to AmrPostprocessing

@@
-\section{The {\tt data\_processing} Routines}
+\section{The {\tt AmrPostprocessing} Routines}
 \label{sec:analysis}
 
-The {\tt fParallel/data\_processing/} directory contains a large
+The {\tt AmrPostprocessing/F\_Src/} directory contains a large
 number of Fortran-based analysis routines for BoxLib datasets.  Many
@@
 The routine {\tt fspeciesmass2d.f90} in {\tt
-  data\_processing/tutorial} serves as a well-commented example of how
+  F\_src/tutorial} serves as a well-commented example of how
 to work with \maestro\ plotfile data.  This routine simply computes
@@
 first set your {\tt PYTHONPATH} environment variable to point to the
-{\tt fParallel/data\_processing/python/} directory, for example:
+{\tt AmrPostprocessing/F\_Src/python/} directory, for example:
 \begin{verbatim}
-export PYTHONPATH="/home/username/development/MAESTRO/fParallel/data_processing/python"
+export PYTHONPATH="/home/username/development/AmrPostprocessing/F_Src/python"
 \end{verbatim}

### 2012-04-12 docs/unit_tests/unit_tests.tex
spacing

@@
 
-In addition to the \maestro science problems, which use the full
+In addition to the \maestro\ science problems, which use the full
 capabilities of \maestro, there are a number of unit tests that

### 2012-04-12 docs/architecture/architecture.tex
add constants

@@
 \begin{itemize}
-\item {\tt data\_processing/}
-
-  Simple Fortran-based analysis routines (e.g.\ extract a line from a
-  multidimensional dataset) that operate on BoxLib datasets.  These are
-  described in \S~\ref{sec:analysis}.
-
-  Several sub-directories with python-based routines are also here.
-  These are described both in \S~\ref{sec:analysis} and \S~\ref{sec:vis:python}.
-
 \item {\tt extern/}
@@
 \end{itemize}
+
+Finally the {\tt AmrPostprocessing} package provides simple
+Fortran-based analysis routines (e.g.\ extract a line from a
+multidimensional dataset) that operate on BoxLib datasets.  These are
+described in \S~\ref{sec:analysis}.  Several sub-directories with
+python-based routines are also here.  These are described both in
+\S~\ref{sec:analysis} and \S~\ref{sec:vis:python}.
 
@@
 
-\subsection{\tt average}
-
-The {\tt average} module provides a routine {\tt average} that takes
+\subsection{\tt average\_module}
+
+The {\tt average\_module} module provides a routine {\tt average} that takes
 a multilevel \multifab\ array and averages the full Cartesian data
@@
 base state or full Cartesian state.
+
+\subsection{\tt fundamental\_constants\_module}
+
+The {\tt fundamental\_constants\_module} provides a simple list of
+various fundamental constants (e.g.\ Newton's gravitational constant)
+in CGS units.
 

### 2012-04-12 docs/introduction/introduction.tex
expanded intro based on feedback from students

@@
 \maestro\ describes the evolution of low Mach number flows that are in
-hydrostatic equilibrium.  The idea for \maestro\ grew out of our success
-in applying low Mach number combustion methods developed for
-terrestrial flames~\cite{DayBell00} to small-scale
-astrophysical flows.  Our original small-scale astrophysical
+hydrostatic equilibrium.  The idea for \maestro\ grew out of our
+success in applying low Mach number combustion methods developed for
+terrestrial flames~\cite{DayBell00} to small-scale astrophysical
+flames (including Landau-Darrieus instability~\cite{SNld},
+Rayleigh-Taylor unstable flames~\cite{SNrt3d}, and flame-turbulence
+interactions~\cite{SNturb}).  Our original small-scale astrophysical
 combustion algorithm is detailed in
@@
 
+\section{Brief Overview of Low Speed Approximations}
+
+There are many low speed formulations of the equations of hydrodynamics
+in use, each with their own applications.  All of these methods share in
+common a constraint equation on the velocity field that augments the
+equations of motion.
+
+\subsection{Incompressible Hydrodynamics}
+
+The simplest low Mach number approximation is incompressible
+hydrodynamics. This approximation is formally the $M \rightarrow 0$
+limit of the Navier-Stokes equations. In incompressible hydrodynamics,
+the velocity satisfies a constraint equation:
+\begin{equation}
+\nabla \cdot \Ub = 0
+\end{equation}
+which acts to instantaneously equilibrate the flow, thereby filtering
+out soundwaves.  The constraint equation implies that
+\begin{equation}
+D\rho/Dt = 0
+\end{equation}
+(through the continuity equation) which says that the density is
+constant along particle paths. This means that there are no
+compressibility effects modeled in this approximation.
+
+\subsection{Anelastic Hydrodynamics}
+
+In the anelastic approximation small amplitude thermodynamic
+perturbations are carried with respect to a static hydrostatic
+background (described by density $\rho_0$).  The density perturbation
+is ignored in the continuity equation, resulting in a constraint
+equation:
+\begin{equation}
+\nabla \cdot (\rho_0 \Ub) = 0
+\end{equation}
+This properly captures the compressibility effects due to the
+stratification of the background. Because there is no evolution
+equation for the perturbational density, approximations are made to
+the buoyancy term in the momentum equation.
+
+\subsection{Low-Mach Number Combustion}
+
+In the low Mach number combustion model, the pressure is decomposed
+into a dynamic, $\pi$, and thermodynamic component, $p_0$, the ratio
+of which is $O(M^2)$. The total pressure is replaced everywhere by the
+thermodynamic pressure, except in the momentum equation. This
+decouples the pressure and density and filters out the sound
+waves. Large amplitude density and temperature fluctuations are
+allowed. The only requirement is that the total pressure stay close to
+the background pressure, which is assumed constant. This requirement
+can be expressed as:
+\begin{equation}
+p = p_0
+\end{equation}
+and differentiating this along particle paths leads to a constraint on
+the velocity field: 
+\begin{equation}
+\nabla \cdot \Ub = S 
+\end{equation}
+This looks like the constraint for incompressible hydrodynamics, but
+now we have a source term, $S$, representing the local compressibility
+effects due to the energy generation and thermal diffusion.  Since the
+background pressure is taken to be constant, we cannot model flows
+that cover a large fraction of a pressure scale height. However, this
+method is ideal for exploring the physics of flames.
+
+\subsection{Pseudo-Incompressible Methods}
+
+The pseudo-incompressible method incorporates both the local changes
+to compressibility due to reaction/heat release, and the large-scale
+changes due to the background stratification.  This was originally
+derived for an ideal gas equation of state for atmospherical flows.
+Allowing the background pressure, $p_0$ to vary (e.g.\ in hydrostatic
+equilibrium), differentiating pressure along particle paths gives:
+\begin{equation}
+\nabla \cdot (p_0^{1/\gamma} \Ub) = H
+\end{equation}
+where $\gamma$ is the ratio of specific heats and $H$ is the source.
+
+\maestro\ is based on this method, generalizing this constraint to an
+arbitrary equation of state and allowing for the time-variation of the
+base state.
+
 \section{Projection Methods 101}
@@
 \begin{equation}
-U_t + \nabla \cdot F(U) = 0
+\Ub_t + \nabla \cdot F(\Ub) = 0
 \end{equation}
-where U is the vector of conserved quantities, $U = (\rho, \rho u,
+where $\Ub$ is the vector of conserved quantities, $\Ub = (\rho, \rho u,
 \rho E)$, with $\rho$ the density, $u$ the velocity, and $E$ the total
@@
 \begin{equation}
-q_t + A(q) q_x = 0
+{\bf q}_t + A({\bf q}) {\bf q}_x = 0
 \end{equation}
-where $q$ are called the primitive variables, and $A$ is the
+where ${\bf q}$ are called the primitive variables, and $A$ is the
 Jacobian, $A \equiv \partial F / \partial U$.  The eigenvalues of the
@@
 \begin{eqnarray}
-U_t &=& -U \cdot \nabla U + \nabla p \label{eq:incompressible_u} \\
-\nabla \cdot U &=& 0
+\Ub_t &=& -\Ub \cdot \nabla \Ub + \nabla p \label{eq:incompressible_u} \\
+\nabla \cdot \Ub &=& 0
 \end{eqnarray}
-Here, $U$ represents the velocity vector%
+Here, $\Ub$ represents the velocity vector%
 %
@@
 incompressible community.  In papers on compressible hydrodynamics,
-$U$ will usually mean the vector of conserved quantities.  In 
-incompressible / low speed papers, $U$ will mean the velocity vector.}
+$\Ub$ will usually mean the vector of conserved quantities.  In 
+incompressible / low speed papers, $\Ub$ will mean the velocity vector.}
 %
@@
 techniques similar to those developed for compressible hydrodynamics,
-updating the old velocity, $U^n$, to the new time-level, $U^\star$.
+updating the old velocity, $\Ub^n$, to the new time-level, $\Ub^\star$.
 Here the `$^\star$' indicates that the updated velocity does not, in
@@
 \begin{equation}
-U^\star = U^d + \nabla \phi \label{eq:decomposition}
+\Ub^\star = \Ub^d + \nabla \phi \label{eq:decomposition}
 \end{equation}
-where $U^d$ is the divergence free portion of the velocity vector,
-$U^\star$, and $\phi$ is a scalar.  Taking the divergence of
+where $\Ub^d$ is the divergence free portion of the velocity vector,
+$\Ub^\star$, and $\phi$ is a scalar.  Taking the divergence of
 Eq.~(\ref{eq:decomposition}), we have
 \begin{equation}
-\nabla^2 \phi = \nabla \cdot U^\star
+\nabla^2 \phi = \nabla \cdot \Ub^\star
 \end{equation}
-(where we used $\nabla \cdot U^d = 0$).
+(where we used $\nabla \cdot \Ub^d = 0$).
 With appropriate boundary conditions, this Poisson equation can be
@@
 \begin{equation}
-U^{n+1} = U^\star - \nabla \phi
+\Ub^{n+1} = \Ub^\star - \nabla \phi
 \end{equation}
+Because soundwaves are filtered, the timestep constraint now depends only
+on $|\Ub|$.
 
@@
 notation consistent.  Table~\ref{table:sym} defines the
-frequently-used quantities and provides their units.
+frequently-used quantities and provides their units.  Additionally,
+for any quantity $\phi$, we denote the average of $\phi$ over a layer
+at constaint radius (or height for plane-parallel simulations) as
+$\overline{\phi}$.
 
@@
 \end{center}
-
 \renewcommand{\arraystretch}{1.0}

### 2012-04-18 docs/flowchart/flowchart.tex
latex fixes
some notes

@@
 to papers~I through IV and the multilevel paper for the derivation
-and motivation of the equation set.
+and motivation of the equation set.  (Note: this `traditional' algorithm
+uses Strang-splitting for the reactions.  An alternate implementation, using
+spectral deferred corrections is outlined in Chapter~\ref{ch:sdc}.)
 
@@
 to that code to see the sequence of functions called to implement each
-ste.
+step.
 
 \subsection{Notation}
+
+reach\_state
+
+enforce\_hse
 
@@
 \item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\star,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$.  We use the equations 
+  for the conservative update of $(\rho X_k)^{(1)}$.
+
+  There are a variety of choices of quantities to predict to the
+  edges (controlled by {\tt species\_pred\_type}---see Chapter~\ref{ch:pert}).
+  By default, we use the equations 
 \begin{eqnarray}
@@
 \item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\star,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$.  The perturbational
+  for the conservative update of $(\rho h)^{(1)}$.  There are a variety
+   of quantities that we can predict to the interfaces here (controlled
+   by {\tt enthalpy\_pred\_type}---see Chapter~\ref{ch:pert}).  For
+   the default case, we use the perturbational
   enthalpy equation, neglecting reactions,
@@
 \end{equation}
-  is used to predict
+  to predict
   $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$ to time-centered edges, 
@@
 \includegraphics[scale=0.65]{\flowfigpath/flowchart}
-\caption[Graphical flowchart of MAESTRO]
+\caption[Graphical flowchart of \maestro]
   {\label{Fig:flowchart} A flowchart of the algorithm.  The

### 2012-04-18 docs/getting_started/getting_started.tex
updates

@@
 parallelization.  We also rely on the build system in \boxlib\ to
-build a \maestro\ executable.  \boxlib can be obtained via {\tt
+build a \maestro\ executable.  \boxlib\ can be obtained via {\tt
 git}\footnote{{\tt git} is an open-source source code version control
@@
 
-This option specifies the Fortran compiler.  We will use {\tt gfortran}, which is
-the preferred compiler for \maestro.  Specifying this compiler will automatically
-pull in the compiler settings as specified in {\tt BOXLIB\_HOME/Tools/F\_mk/}.
+This option specifies the Fortran compiler.  We will use {\tt
+gfortran}, which is the preferred compiler for \maestro.  Specifying
+this compiler will automatically pull in the compiler settings as
+specified in {\tt BOXLIB\_HOME/Tools/F\_mk/}.  (Alternate compiler
+choices include {\tt Intel}, {\tt PGI}, {\tt PathScale}.)
 
@@
 
-Each problem requires an input file.  The inputs file consists of
+Each problem requires an input file.  The inputs file (a Fortran
+namelist) consists of
 lines of the form {\em parameter = value}, where {\em parameter} is
@@
 used, output date, output directory, runtime parameter values, ...)
-are stored in the {\tt job\_info} file in each plotfile directory.
+are stored in the plaintext {\tt job\_info} file in each plotfile directory.
 
@@
 related development packages (e.g.\ {\tt lesstif-devel}).  Depending
-on your Linux system, you may also need to install {\tt libXpm} 
-and related development packages (e.g.\ {\tt libXpm-devel}).
+on your Linux system, you may also need to install {\tt libXpm} and
+related development packages (e.g.\ {\tt libXpm-devel}).  Also, in
+{\tt BoxLib/Tools/C\_mk/Make.defs}, the Fortran libraries are
+explictly linked in by asking {\tt gfortran} for the location of {\tt
+libgfortran.a}.  You may need to explicitly install the static
+libraries for {\tt gfortran} (i.e.\ {\tt libgfortran-static}), or
+change the library line to query for the shared-object version {\tt
+libgfortran.so}.
 
-%need to modify _f95_lib
-
-library line
+Further details on the C++ \boxlib\ build system used by \amrvis\
+can be found in the \boxlib\ documentation.
 

### 2012-04-18 docs/architecture/architecture.tex
fix the MAC velocity example

@@
 \begin{lstlisting}[language={[95]fortran},mathescape=false]
-  type(multifab) :: umac(dm)
-
+  type(multifab) :: umac(nlevel,dm)
+
+  do n=1,nlevel
      do comp=1,dm
         call multifab_build_edge(umac(n,comp), mla%la(n),1,1,comp)
+     enddo
   enddo

### 2012-04-19 docs/unit_tests/unit_tests.tex
add figures showing the projection test output

@@
 
-  This tests the hgprojection routine in 2-d.  A divergence-free
-  velocity field is initialized and then ``polluted'' by adding the
-  gradient of a scalar.  The form of the scalar differs depending on
-  the boundary conditions (wall and periodic are supported currently).
-  Finally, the hgproject routine is called to recover the initial
-  divergence-free field.
+  This tests the projection routines in 2-d---either the hgprojection
+  ({\tt project\_type = 1}) or the MAC projection ({\tt project\_type =
+  2}).  A divergence-free velocity field is initialized and then
+  ``polluted'' by adding the gradient of a scalar.  The form of the
+  scalar differs depending on the boundary conditions (wall and
+  periodic are supported currently).  Finally, the hgproject routine
+  is called to recover the initial divergence-free field.
+  Figure~\ref{fig:unit:projtest} shows the initial field, polluted
+  field, and result of the projection for the hgproject case.
+
+\begin{figure}[t] 
+\centering
+\includegraphics[height=1.8in]{\unitfigpath/wall_u_init_x-velocity} 
+\includegraphics[height=1.8in]{\unitfigpath/wall_u_plus_grad_phi_x-velocity} 
+\includegraphics[height=1.8in]{\unitfigpath/wall_u_new_x-velocity} 
+%
+\caption[Results of the {\tt test\_projection} unit test]{\label{fig:unit:projtest}
+  Initial divergence free velocity field (x-component; left); Velocity field plus
+  gradient of a scalar (x-component; center); and resulting velocity after projecting
+  out the non-divergence free portion (x-component; right).  This is will slipwall
+  boundary conditions on all sides, a 2-level grid with the left half refined and
+  right half coarse, and the hgprojection tested.}
+\end{figure}

### 2012-04-20 docs/unit_tests/unit_tests.tex
add test_advect figures

@@
   not show any directional bias.
+
+\begin{figure}[t] 
+\centering
+\includegraphics[height=1.9in]{\unitfigpath/dens_2d_orig_density}
+\includegraphics[height=1.9in]{\unitfigpath/dens_2d_ppm1_xp_final_density}
+\includegraphics[height=1.9in]{\unitfigpath/dens_2d_ppm1_xp_final_abserror} 
+%
+\caption[Results of the {\tt test\_advect} unit test]{\label{fig:unit:advtest}
+  Initial Gaussian density profile (left); density profile after
+  advecting to the right for one period (center); absolute error
+  between the final and initial density fields, showing the error in
+  the advection scheme (right).}
+\end{figure}
 
@@
 \centering
-\includegraphics[height=1.8in]{\unitfigpath/wall_u_init_x-velocity} 
-\includegraphics[height=1.8in]{\unitfigpath/wall_u_plus_grad_phi_x-velocity} 
-\includegraphics[height=1.8in]{\unitfigpath/wall_u_new_x-velocity} 
+\includegraphics[height=1.9in]{\unitfigpath/wall_u_init_x-velocity} 
+\includegraphics[height=1.9in]{\unitfigpath/wall_u_plus_grad_phi_x-velocity} 
+\includegraphics[height=1.9in]{\unitfigpath/wall_u_new_x-velocity} 
 %

### 2012-04-21 docs/getting_started/getting_started.tex
detail string version of BCs

@@
 are stored in the plaintext {\tt job\_info} file in each plotfile directory.
+
+{\bf Note: unless otherwise specified all quantities in \maestro\ are
+assumed to be in CGS units.}
 
@@
 
-\item {\tt bcx\_lo }, {\tt bcy\_lo }, {\tt bcz\_lo }, 
-      {\tt bcx\_hi }, {\tt bcy\_hi }, {\tt bcz\_hi } are the boundary
-   condition types at the lower (`{\tt lo}') and upper (`{\tt hi}')
-   domain boundaries in the $x$, $y$, and $z$ coordinate directions.
-   The different types are set via integer flags listed in table~\ref{arch:table:bcflags}.
+\item There are two ways to specify boundary conditions---via integer flags
+      or descriptive string names.  If the string names are present,
+      then they will override the integer quantities in determining
+      the boundary conditions.
+
+   \begin{itemize}
+
+   \item {\tt bcx\_lo }, {\tt bcy\_lo }, {\tt bcz\_lo }, {\tt bcx\_hi
+      }, {\tt bcy\_hi }, {\tt bcz\_hi } are the boundary condition
+      types at the lower (`{\tt lo}') and upper (`{\tt hi}') domain
+      boundaries in the $x$, $y$, and $z$ coordinate directions.  The
+      different types are set via integer flags listed in
+      table~\ref{gs:table:bcflags}.
 
    \begin{table}[h]
-   \caption{\label{arch:table:bcflags} Boundary condition flags}  
+   \caption{\label{gs:table:bcflags} Boundary condition flags}  
    \begin{center}
@@
 
+   \item {\tt xlo\_boundary\_type}, {\tt ylo\_boundary\_type}, {\tt
+   zlo\_boundary\_type}, {\tt xhi\_boundary\_type}, {\tt
+   yhi\_boundary\_type}, {\tt zhi\_boundary\_type} are the boundary
+   condition types at the lower and upper domain boundaries in 
+   the $x$, $y$, and $z$ coordinate directions.   The boundary
+   type is set by providing a string name---valid values are listed
+   in table~\ref{gs:table:bcstrings}
+
+   \begin{table}[h]
+   \caption{\label{gs:table:bcstrings} Boundary condition string names}  
+   \begin{center}
+   \begin{tabular}{ll}
+   \hline
+   BC type    & integer flag \\
+   \hline
+   periodic             & ``periodic'' \\
+   inlet (user-defined) & ``inlet'' \\
+   outlet               & ``outlet'' \\
+   symmetry             & ``symmetry'' \\
+   slip wall            & ``slip wall'' \\
+   no-slip wall         & ``no slip wall'' \\
+   \hline
+   \end{tabular}
+   \end{center}
+   \end{table}
+
+   \end{itemize}
+
+  The string-based parameters are a newer option for specifying
+  boundary conditions, and are preferred due to clarity.  The
+  conversion between the string names and integer flags is done
+  by \boxlib\ in the {\tt bc\_module} at the time of initializing
+  the runtime parameters.
+
 \end{itemize}

### 2012-04-24 docs/unit_tests/unit_tests.tex
add note about BDS in 3d

@@
   quantities are used) and a uniform velocity field in any one of the
-  coordinate directions.  The Gaussian profile is advected through
-  the period domain exactly once and the error in the density profile
-  (L2 norm) is computed.  The driver for this problem does this 
-  for every dimension twice (once with a positive velocity and once
-  with a negative velocity).  After all coordinate directions are 
-  tested, the norms are compared to ensure that the error does
-  not show any directional bias.
+  coordinate directions.  The Gaussian profile is advected through the
+  period domain exactly once and the error in the density profile (L2
+  norm) is computed.  The driver for this problem does this for every
+  dimension twice (once with a positive velocity and once with a
+  negative velocity), and loops over all advection methods ({\tt
+  ppm\_type = 0,1,2} and {\tt bds\_type = 1}).  After all coordinate
+  directions are tested, the norms are compared to ensure that the
+  error does not show any directional bias.
+
+  Note: the BDS advection method does not guarantee that the error be
+  independent of the advection direction---small differences can
+  arise.  What's happening here is that within each cell BDS is trying
+  to define a tri-linear profile for rho subject to the constrants in
+  the BDS paper~\cite{bds3d} (constraints 1 and 2 on p.\ 2044 after eq. 3.4).  We
+  do not solve the L2 minimization problem exactly so we iterate up to
+  3 times using a simple heuristic algorithm designed to work toward
+  the constraint.  The iteration will give different results depending
+  on orientation since we work through the corners in arbitrary order.
 

### 2012-04-26 docs/runtime_parameters/rp.py
add a chapter label.  Update the parameter list

@@ -3,6 +3,8 @@ import sys
 
 # tex format stuff
 Mheader=r"""
+\label{ch:parameters}
+
 \section{Runtime Parameters}
 
 The following tables list the runtime parameters available to

### 2012-04-26 docs/architecture/architecture.tex
another MAC fix
some updates

@@
 
-\section{\maestro\ Directory Structure}
+\section{The \maestro\ `Ecosystem'}
 
@@
 \includegraphics[scale=0.55]{\archfigpath/maestro_ecosystem2}
-\caption[\maestro\ directory structure]
-{The basic \maestro\ directory structure.}
+\caption[\maestro\ `ecosystem'] {\label{fig:arch:eco} The basic
+  \maestro\ `ecosystem'.  Here we see the different packages that
+  contribute to building the {\tt test2} problem in \maestro.  The
+  purple lines show the directories that are pulled in through
+  variable Makefile variables ({\tt BOXLIB\_HOME}, {\tt NETWORK\_DIR},
+  {\tt EOS\_DIR}, and {\tt CONDUCTIVITY\_DIR}).  The blue lines show
+  standard modules that are used in most \maestro\ builds.}
 \end{figure}
@@
 files (distributed in the {\tt MAESTRO/} directory), and the
-\boxlib\ library (distributed in the {\tt BoxLib/} directory.
+\boxlib\ library (distributed separately, consisting of the {\tt BoxLib/} directory).
 \boxlib\ provides both a C++ and a Fortran framework.  \maestro\
-only uses the Fortran framework.
+only uses the Fortran portions of \boxlib.  Figure~\ref{fig:arch:eco}
+shows the relationship between the different packages, highlighting
+what goes into defining a specific \maestro\ problem.
 
@@
 
+  \item {\tt Source_SDC}
+
+    Source for an alternate formulation of \maestro\ that removes the
+    splitting between hydrodynamics, diffusion, and reactions (see
+    Chapter~\ref{ch:sdc}).
+
   \item {\tt TEST\_PROBLEMS}
@@
 \includegraphics[scale=0.65]{\archfigpath/boxlib_directory2}
-\caption[\boxlib\ directory structure]
-{The basic \boxlib\ directory structure.  The directories used
-by \maestro\ are indicated in red.}
+\caption[\boxlib\ directory structure] {\label{fig:arch:boxlib} The
+  basic \boxlib\ directory structure.  The directories used by
+  \maestro\ are indicated in red.}
 \end{figure}
 
-The \boxlib\ directories that \maestro\ uses are:
+The \boxlib\ directory structure is shown in
+Figure~\ref{fig:arch:boxlib}.  The subset of the directories that are
+used by \maestro\ are:
 \begin{itemize}
@@
 
-Different \maestro\ problems are defined in sub-directories under
-{\tt MAESTRO/}.  To add a problem, start by creating a new
-sub-directory---this is where you will compile your problem and
-store all the problem-specific files.
-
-The minimum requirement to
-define a new problem would be a {\tt GNUmakefile} which describes how
-to build the application and an input file which lists the runtime
-parameters.  The problem-specific executable is built in the problem
-directory by typing {\tt make}.  Source files are found automatically
-by searching the directories listed in the {\tt GNUmakefile}.
-Customized versions of any source files placed in the problem-directory
-override those with the same name found elsewhere.  Any unique
-source files (and not simply a custom version of a file found
-elsewhere) needs to be listed in a file called {\tt GPackage.mak}
-in the problem-directory.
+Different \maestro\ problems are defined in sub-directories under {\tt
+  MAESTRO/SCIENCE}, {\tt MAESTRO/TEST\_PROBLEMS}, or {\tt
+  MAESTRO/UNIT\_TESTS}.  To add a problem, start by creating a new
+sub-directory---this is where you will compile your problem and store
+all the problem-specific files.
+
+The minimum requirement to define a new problem would be a {\tt
+  GNUmakefile} which describes how to build the application and an
+input file which lists the runtime parameters.  The problem-specific
+executable is built in the problem directory by typing {\tt make}.
+Source files are found automatically by searching the directories
+listed in the {\tt GNUmakefile}.  Customized versions of any source
+files placed in the problem-directory override those with the same
+name found elsewhere.  Any unique source files (and not simply a
+custom version of a file found elsewhere) needs to be listed in a file
+called {\tt GPackage.mak} in the problem-directory (and this needs to
+be told to the build system---see below).
 
@@
 \begin{lstlisting}[language={[gnu]make},mathescape=false]
-  COMP := Intel
+  COMP := gfortran
 \end{lstlisting}
@@
 use.  Allowed values include {\tt Intel}, {\tt gfortran}, {\tt PGI},
-and {\tt PathScale}.  The specific details of these choices are
-defined in the {\tt MAESTRO/mk/comps/} directory.
+{\tt PathScale}, and {\tt Cray}.  The specific details of these
+choices are defined in the {\tt MAESTRO/mk/comps/} directory.
 
@@
 compiled.  This is a Fortran module that holds the values of the
-runtime parameters and makes them available to any routine.  The
-variable {\tt PROBIN\_PARAMETERS} should list all of the files that
-contain runtime parameters for the current problem.  You must always
-include the main {\tt ../\_parameters} file (i.e.\ {\tt
-  MAESTRO/\_parameters}).  Sometimes a problem defines their own
-runtime parameters (for instance {\tt \_parameters.myproblem}), in which
-case, its parameter file should also appear in {\tt
-  PROBIN\_PARAMETERS} (see \S~\ref{sec:def_runtime_param}).
-\begin{lstlisting}[language={[gnu]make},mathescape=false]
-  PROBIN_PARAMETERS := ../_parameters
-\end{lstlisting}
+runtime parameters and makes them available to any routine.  By
+default, the build system looks for a file called {\tt \_parameters}
+in the problem directory and adds those parameters along with the
+master list of \maestro\ parameters ({\tt MAESTRO/\_parameters}) to
+the {\tt probin\_module}.
 
@@
 
-The runtime parameters for the core \maestro\ algorithm are listed
-in {\tt MAESTRO/\_parameters}.  That file is parsed at compile-time
-by the {\tt MAESTRO/write\_probin.py} script
-(along with any other parameter files listed in the {\tt PROBIN\_PARAMETERS}
-variable in the {GNUmakefile}, see \S \ref{sec:makefile}).  The script
-outputs the {\tt probin.f90} source file.  Each line
-in the {\tt \_parameters} file has the form: 
+The runtime parameters for the core \maestro\ algorithm are listed in
+{\tt MAESTRO/\_parameters}.  That file is parsed at compile-time by
+the {\tt MAESTRO/write\_probin.py} script (along with any
+problem-specific parameters).  The script outputs the {\tt probin.f90}
+source file.  Each line in the {\tt \_parameters} file has the form:
 \vskip 3mm
@@
 following that heading are placed into that category.  The documentation
-is produced by the script {\tt MAESTRO/docs/runtime\_parameters/rp.py}.
+(Chapter~\ref{ch:parameters}) is produced by the script 
+{\tt MAESTRO/docs/runtime\_parameters/rp.py}.
 
@@
 Problem-specific runtime parameters should be defined in the
-problem-directory in a file called {\tt
-  \_parameters}\{{\em.problem-name}\}.  This file should be added
-to the {\tt PROBIN\_PARAMETERS} variable in the {\tt GNUmakefile}.
+problem-directory in a file called {\tt \_parameters}.  This file will
+be automatically found at compile time.
 
@@
           u = HALF*(umac(i,j) + umac(i+1,j))
-          v = HALF*(vmac(i,j) + vmac(i+1,j))
+          v = HALF*(vmac(i,j) + vmac(i,j+1))
 

### 2012-04-26 docs/getting_started/getting_started.tex
some updates

@@
 
-\subsection{{\tt data\_processing} scripts}
+\subsection{{\tt AmrPostprocessing} scripts}
 
 Several useful analysis scripts (written in Fortran 90) can be found
-in {\tt fParallel/data\_processing}.  The {\tt GNUmakefile} there
-needs to be edited to indicate which of the tools to build.  For
-example, to extract the density along a line from the center of a
-plotfile, {\tt plt00200}, in the $y$-direction:
+in {\tt AmrPostprocessing/F\_Src/} (distributed separately
+from \maestro).  The {\tt GNUmakefile} there needs to be edited to
+indicate which of the tools to build.  For example, to extract the
+density along a line from the center of a plotfile, {\tt plt00200}, in
+the $y$-direction:
 
@@
 There is also a python visualization method in {\tt
-fParallel/data\_processing/python\_plotfile}.  This is described
+AmrPostprocessing/F\_src/python\_plotfile}.  This is described
 in \S~\ref{sec:vis:python}.
@@
 datasets.  It can do contouring, volume rendering, streamlines, ...\, ,
-directly from \boxlib\ plotfiles (to open a plotfile in \visit, point to
-the {\tt Header} file inside the plotfile directory).  Details on
+directly from \boxlib\ plotfiles.   Details on
 \visit\ can be found at:\newline
@@
 
-By default, \maestro\ outputs global diagnostics each timestep into a file
-called {\tt maestro\_diag.out}.  This includes the maximum Mach number,
-peak temperature, and peak nuclear energy generation rate.  Individual problems
-can provide their own {\tt diag.f90} file to produce custom diagnostic
-output.
+By default, \maestro\ outputs global diagnostics each timestep into a
+file called {\tt maestro\_diag.out}.  This includes the maximum Mach
+number, peak temperature, and peak nuclear energy generation rate.
+Individual problems can provide their own {\tt diag.f90} file to
+produce custom diagnostic output.  This information can be plotted
+directly with {\sl GNUplot}, for example.
 

### 2012-04-26 docs/compilers/compilers.tex
update Cray
fix typo

@@
 
-So far, no luck with these compilers (versions 7.4.3 and 8.0.0.135).
+Cray 8.0.2 compiles and runs \maestro\ fine on the interlagos/jaguarpf 
+machine at OCLF.
 
@@
 \item gfortran 4.5 has no known issues.
-\item gfortran 4.6.1 and 4.6.2 on Fedora (15 and 16) have a bug that 
+\item gfortran 4.6.1--4.6.3 on Fedora (15 and 16) have a bug that 
 shows up in multi-processor runs ({\tt toy\_convect} on 4 processors, for instance).

### 2012-04-26 docs/getting_started/getting_started.tex
add figure

@@
 
-Visualization and analysis are done on the plotfiles.  A
-number of in-house and externally developed tools can work 
-with \boxlib-formatted plotfiles.
+Visualization and analysis are done on the plotfiles.  A number of
+in-house and externally developed tools can work
+with \boxlib-formatted plotfiles.  An example plot of the {\tt test2}
+problem run above is shown in Figure~\ref{fig:gettingstarted:test2}.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=3in]{plt00133_tfromp}
+\caption{\label{fig:gettingstarted:test2} Visualization of the
+final output of the {\tt test2} problem showing the temperature
+field (as derived from the pressure).  This plot was done with
+the {\tt AmrPostprocessing} tools.}
+\end{figure}
 

### 2012-04-26 docs/getting_started/getting_started.tex
add example output fig

--- 2012-04-26 docs/getting_started/getting_started.tex
@@
 
-Visualization and analysis are done on the plotfiles.  A
-number of in-house and externally developed tools can work 
-with \boxlib-formatted plotfiles.
+Visualization and analysis are done on the plotfiles.  A number of
+in-house and externally developed tools can work
+with \boxlib-formatted plotfiles.  An example plot of the {\tt test2}
+problem run above is shown in Figure~\ref{fig:gettingstarted:test2}.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=3in]{\gsfigpath/plt00133_tfromp}
+\caption[Visualization of {\tt test2} output]{\label{fig:gettingstarted:test2} Visualization of the
+final output of the {\tt test2} problem showing the temperature
+field (as derived from the pressure).  This plot was done with
+the {\tt AmrPostprocessing} tools.}
+\end{figure}
 

### 2012-05-18 docs/lo_density/lo_density.tex
fix notes to talk about when we don't hit cutoff density

@@
 The guiding principle is to check whether $\rho_0$ falls below $\rho_{\rm cutoff}$ on the finest
-grid first.  If not, check the next coarser level.  Continue until you reach the base grid,
-in which case $\rho_0$ must fall below $\rho_{\rm cutoff}$ somewhere.  Some examples are in order:
+grid first.  If not, check the next coarser level.  Continue until you reach the base grid.
+Some examples are in order:
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+Note: if $\rho_0$ does not fall below $\rho_{\rm cutoff}$ at any level, we set the cutoff 
+coordinate at the fine level to be first first cell above the domain and propagate the 
+coordinate to the coarser levels.
+
 \section{When are the Cutoff Coordinates Updated?}
@@
 
-\item In {\tt init\_base\_state}, we give a warning if {\tt anelastic\_cutoff}
-  is smaller than the minimum model density.  Then, if ${\tt anelastic\_cutoff} 
-  < {\tt base\_cutoff\_density}$, we abort the program.
-
 \item In {\tt make\_div\_coeff}, for 
@@
 \item In {\tt probin}, {\tt base\_cutoff\_density} is set to $3\times 10^6$ by default.
-
-\item In {\tt init\_base\_state}, we give a warning of {\tt base\_cutoff\_density}
-  is smaller than the minimum model density.  Then, if ${\tt anelastic\_cutoff} 
-  < {\tt base\_cutoff\_density}$ we abort the program.
 

### 2012-05-21 docs/makefiles/makefiles.tex
start of the makefile documentation

\label{ch:make}

\section{Build Process Overview}

The \maestro\ build system relies on the definitions setup by \boxlib.  There are
4 main files that define the macros used by the build:
\begin{itemize}
\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakedefs.mak}

\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakeMPI.mak}

\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakerules.mak}

\item {\tt MAESTRO/GMaestro.mak}
\end{itemize}

The \boxlib\ build system separates the compiler-specific information
from the machine-specific information---this allows for reuse of the
compiler information.

rely on VPATH  / GPackage.mak

dependency generation

include files

files created at build-time: {\tt probin.f90}, {\tt build\_info.f90}, and {\tt network.f90}

\section{\maestro\ Problem Options}

\subsection{Problem-specific Files}

\subsection{Defining Network, EOS, and Conductivity Routines}

\subsection{Core MAESTRO modules}

\subsection{Unit Tests}

\section{Special Targets}

\subsection{Debugging ({\tt print-*})}

\subsection{{\tt clean} and {\tt realclean}}

\section{Extending the Build System}

\subsection{Adding a Machine}

\subsection{Adding a Compiler}

\subsection{Parallel Builds}

### 2012-05-21 docs/makefiles/makefiles.tex
more doc

@@
 
-The \maestro\ build system relies on the definitions setup by \boxlib.  There are
-4 main files that define the macros used by the build:
+The \maestro\ build system uses features of GNU make, which is typically the 
+default on systems.  The main macros that define the build process are 
+split across several files.  The 4 main files are:
 \begin{itemize}
-\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakedefs.mak}
+\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakedefs.mak}:
+  This setups the basic macros, includes the options for the selected
+  compiler, builds the list of object and source files, and defines
+  the compile and link command lines.
 
 \item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakeMPI.mak}
+  This implements any changes to the compiler names and library
+  locations necessary to build a parallel executable with MPI.
 
 \item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakerules.mak}
+  This creates the various build targets and specifies the rules for
+  building the object files, the list of dependencies, and some other
+  lesser-used targets (tags for editors, documentation, etc.)
 
 \item {\tt MAESTRO/GMaestro.mak}
+  This is a \maestro-specific file that gathers all of the various
+  modules that are used to build a typical \maestro\ application
+  and integrates with the \boxlib\ build system.  Every \maestro\
+  problem's makefile will include this file.
+
 \end{itemize}
@@
 from the machine-specific information---this allows for reuse of the
-compiler information.
+compiler information.  The only machine-specific parts of the build system
+are for the MPI library locations, contained in {\tt GMakeMPI.mak}.
 
-rely on VPATH  / GPackage.mak
+The \maestro\ build system relies on the vpath functionality of {\tt
+make}.  In a makefile, the {\tt vpath} variable holds search path used
+to locate source files.  When a particular file is needed, the
+directories listed in {\tt vpath} are searched until the file is
+found.  The first instance of a file is used.  We exploit this feature
+by always putting the build directory first in {\tt vpath} (this is
+done in {\tt GMakerules.mak}).  This means that if a source file is
+placed in the build directory, that copy will override any other
+version in the source path.
 

### 2012-05-21 docs/makefiles/makefiles.tex
more build-system stuff

@@
 
-The \maestro\ build system uses features of GNU make, which is typically the 
-default on systems.  The main macros that define the build process are 
-split across several files.  The 4 main files are:
+The \maestro\ build system uses features of GNU make, which is
+typically the default on systems.  The \maestro\ executable is built
+in the problem's directory (one of the directories under {\tt
+SCIENCE/}, {TEST\_PROBLEMS}, or {UNIT\_TESTS}).  This directory will
+contain a makefile, {\tt GNUmakefile}, that includes all the necessary
+information to build the executable.
+
+The main macros that define the build process are split across several
+files.  The 4 main files are:
 \begin{itemize}
 \item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakedefs.mak}:
+
   This setups the basic macros, includes the options for the selected
@@
 
-\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakeMPI.mak}
+\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakeMPI.mak}:
+
   This implements any changes to the compiler names and library
@@
 
-\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakerules.mak}
+\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakerules.mak}:
+
   This creates the various build targets and specifies the rules for
@@
 
-\item {\tt MAESTRO/GMaestro.mak}
+\item {\tt MAESTRO/GMaestro.mak}:
+
   This is a \maestro-specific file that gathers all of the various
@@
   and integrates with the \boxlib\ build system.  Every \maestro\
-  problem's makefile will include this file.
+  problem's {\tt GNUmakefile} will include this file.
 
 \end{itemize}
+
+\maestro\ gets the location of the \boxlib\ library through the 
+{\tt BOXLIB\_HOME} variable.  This should be set as an environment
+variable in your shell start-up files (e.g.\ {\tt .bashrc} or {\tt
+.cshrc}).
 
@@
 are for the MPI library locations, contained in {\tt GMakeMPI.mak}.
+The compiler flags for the various compilers are listed in the
+files in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.
 
-The \maestro\ build system relies on the vpath functionality of {\tt
+\subsection{Finding Source Files}
+
+The \boxlib\ build system relies on the vpath functionality of {\tt
 make}.  In a makefile, the {\tt vpath} variable holds search path used
@@
 
+In \maestro, the {\tt vpath} variable is set using the macros defined
+in {\tt GMaestro.mak}.  A user does not need to set this variable
+explicitly.  Additional source locations are added in the manner
+described below (see \S~\ref{sec:make:otherfiles}).
+
+\subsection{Dependencies}
+
 dependency generation
@@
 
-files created at build-time: {\tt probin.f90}, {\tt build\_info.f90}, and {\tt network.f90}
+\subsection{Files Created at Compile-time}
+
+Several files are created at build-time:
+\begin{itemize}
+\item {\tt probin.f90}
+
+\item {\tt build\_info.f90}
+
+\item {\tt network.f90} (only for the {\tt general\_null} network)
+\end{itemize}
 
@@
 \subsection{Problem-specific Files}
+\label{sec:make:otherfiles}
 
-\subsection{Defining Network, EOS, and Conductivity Routines}
+\subsection{Defining EOS, Network, and Conductivity Routines}
+
+Each \maestro\ problem needs to define an equation of state, a
+reaction network, and a routine to compute the conductivities (for
+thermal diffusion).  This is true even if the problem is not doing
+reactions of thermal diffusion.  These definitions are specified
+in the problem's {\tt GNUmakefile}.
+
+\begin{itemize}
+\item {\tt EOS\_DIR}:
+
+  This variable points to the path (relative to {\tt fParallel/}) of
+  the equation of state used for the build.  Choices that work
+  with \maestro\ are:
+  \begin{itemize}
+  \item {\tt extern/EOS/helmeos}
+  \item {\tt extern/EOS/gamma\_law\_general}
+  \end{itemize}
+
+\item {\tt CONDUCTIVITY\_DIR}:
+
+  This variable points to the conductivity routine used for the build
+  (relative to {\tt fParallel/}).  Choices that work with \maestro\
+  are:
+  \begin{itemize}
+  \item {\tt extern/conductivity/constant}
+  \item {\tt extern/conductivity/timmes\_stellar}
+  \end{itemize}
+  If diffusion is not being used for the problem, this should be set
+  to {\tt extern/conductivity/constant}.
+
+\item {\tt NETWORK\_DIR}:
+
+  This variable points to the reaction network used for the build
+  (relative to {\tt fParallel/}).  A lot of options are present in
+  {\tt extern/networks/}.  A network is required even if you are not
+  doing reactions, since the network defines the species that are
+  advected and interpreted by the equation of state.
+
+  A special choice, {\tt extern/networks/general\_null} is a general
+  network that simply defines the properties of one or more species.
+  This requires an inputs file, specified by {\tt
+  GENERAL\_NET\_INPUTS}.  This inputs file is read at compile-time and
+  used to build the {\tt network.f90} file that is compiled into the
+  source.
+
+\end{itemize}
 
@@
 \subsection{Debugging ({\tt print-*})}
+
+To see the contents of any variable in the build system, you can build
+the special target {\tt print-{\em varname}}, where {\tt {\em
+varname}} is the name of the variable.  For example, to see what the
+Fortran compiler flags are, you would do:
+\begin{verbatim}
+make print-FFLAGS
+\end{verbatim}
+This would give (for {\tt gfortran}, for example):
+\begin{verbatim}
+FFLAGS is -Jt/Linux.gfortran/m -I t/Linux.gfortran/m -O2 -fno-range-check
+\end{verbatim}
+This functionality is useful for debugging the makefiles.
 

### 2012-05-21 docs/makefiles/makefiles.tex
add info about t/

@@
 The compiler flags for the various compilers are listed in the
-files in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.
+files in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.  The compiler
+is set via the {\tt COMP} variable in the problem's {\tt GNUmakefile}.
+
+There are several options in addition to the compiler that affect the
+build process: {\tt MPI}, {\tt OMP}, and {\tt NDEBUG}---these turn on/off
+MPI parallelization, OpenMP parallelization, and debugging.  Together,
+these choices along with the compiler name are reflected in the name
+of the executable.
+
+When the `{\tt make}' command is run, the object and module files are
+written into a directory {\tt t/{\em OS}.{\em COMP}.{\em other}/},
+where {\tt \em OS} is the operating system detected by the build
+system, {\tt \em COMP} is the compiler used, and {\tt \em other}
+reflects any other build options (MPI, OpenMP, debugging, etc.) used.
+Separating each build into a separate subdirectory under the problem
+directory allows for multiple builds of \maestro\ to exist
+side-by-side in the problem directory.
 
@@
 
-\subsection{Core MAESTRO modules}
+\subsection{Core \maestro\ modules}
 
@@
 
+Typing `{\tt make clean}' deleted the object and module files for the
+current build (i.e., the current choice of {\tt MPI}, {\tt NDEBUG},
+{\tt COMP}, and {\tt OMP}).  This also removes any of the compile-time
+generated source files.  Any other builds are left unchanged.
+
+Typing `{\tt make realclean}' deletes the object and module files for
+all builds---i.e., the entire {\tt t/} directory is removed.
+
 \section{Extending the Build System}
@@
 \subsection{Parallel Builds}
+
+{\tt BOXLIB\_USE\_MPI\_WRAPPERS}

### 2012-05-21 docs/makefiles/makefiles.tex
still more...

@@
 
+The \maestro\ executable is built from source distributed across a
+number of directories.  In each of these directories containing source
+files, there is a {\tt GPackage.mak} file.  This file has a number of
+lines of the form:
+\begin{verbatim}
+f90sources += file.f90
+\end{verbatim}
+where {\tt file.f90} is a source file that should be built when this
+directory is added to the list of build directories.  For old
+fixed-form Fortran files, the files should be added to the {\tt
+fsources} variable instead of {\tt f90sources}.
+
 The \boxlib\ build system relies on the vpath functionality of {\tt
@@
 
-dependency generation
+There is no need to explicitly define the dependencies between the
+source files for Fortran modules.  The scripts in {\tt
+BOXLIB\_HOME/Tools/F\_scripts/} are run at the start of the build
+process and parse all the source files and make an explicit list of
+the dependency pairs.  This process is defined in {\tt
+GMakerules.mak}.
 
-include files
+A few files use explicit `{\tt include}' statements to include Fortran
+source in other source files.  Any include file locations should be
+added to {\tt Fmincludes} variable in the problem's {\tt GNUmakefile}.
+This does not occur frequently.  For the case of the {\tt helmeos}
+equation of state, this is done automatically in {\tt GMaestro.mak}.
 

### 2012-05-22 docs/unit_tests/unit_tests.tex
add a description of test_react

@@
 \end{figure}
+
+\section{\tt test\_react}
+
+  This simply tests the reaction network by calling the \maestro\ {\tt
+  react\_state} routine directly.  The network is selected in the {\tt
+  GNUmakefile} by setting the {\tt NETWORK\_DIR} variable.  A 3d cube
+  is setup with density varying on one axis, temperature varying on
+  another, and the composition varying on the third.  The density and
+  temperature ranges are set in the inputs file.  The composition is
+  read in via an input file.
+
+  A good use of this test is to test whether a burner is threadsafe.
+  This is accomplished by compiling with OpenMP (setting {\tt OMP=t})
+  and the running with 1 thread and multiple threads (this can be done
+  by setting the environment variable {\tt OMP\_NUM\_THREADS} to the 
+  desired number of threads).  Since each zone is independent of the
+  others, the results should be identical regardless of the number
+  of threads.  This can be confirmed using the {\tt fcompare} tool
+  in {\tt AmrPostprocessing}.

### 2012-05-22 docs/makefiles/makefiles.tex
more...

@@
 
+If a problem has a unique file that is needed as part of the build,
+then that file should be added to a {\tt GPackage.mak} file in the
+problem's directory.  Since, by default, problems don't have a {\tt
+GPackage.mak}, the build system needs to be told to look in the
+problem directory for unique sources.  This is accomplished by adding
+the problem's directory to the {\tt EXTRA\_DIR} variable in the
+problem's {\tt GNUmakefile}.
+
+Note that this is not necessary if you place a custom version of 
+a source file in the problem's directory.  Since that file is already
+listed in the {\tt GPackage.mak} in its original location, the build
+system will know that it needs to be built.  Since the {\tt vpath}
+variable puts the problem's directory at the start of the search
+path, the version of the file in the problem's directory will be 
+found first.
+
 \subsection{Defining EOS, Network, and Conductivity Routines}

### 2012-05-22 docs/makefiles/makefiles.tex
more...

@@
 \begin{itemize}
-\item {\tt probin.f90}
+\item {\tt probin.f90}:
 
-\item {\tt build\_info.f90}
+  This is the module that controls runtime parameters.  This is
+  created by the script
+  {\tt write\_probin.py} in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_scripts/}.  The
+  makefile logic for creating it is in {\tt GMaestro.mak}.  At compile
+  time, the problem, main {\tt MAESTRO/}, and {\tt extern/}
+  directories are searched for {\tt \_parameter} files.  These files
+  are parsed and the {\tt probin.f90} file is output containing the 
+  runtime parameters, their default values, and the logic for reading
+  and interpreting the inputs file.
 
-\item {\tt network.f90} (only for the {\tt general\_null} network)
+\item {\tt build\_info.f90}:
+
+  This is a module that contains basic information about the build
+  environment (compiler used, build machine, build directory, compiler
+  flags, etc.).  This is created by the script {\tt make\_build\_info}
+  in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_scripts/} from {\tt
+  GMaestro.mak} by passing in a number of makefile variables.  This is
+  rewritten everytime the code is compiled.  The primary use of this
+  module is writing out the {\tt job\_info} file in the plotfiles.
+
+\item ({\tt network.f90}):
+
+  This is generated at compile time {\em only} for the {\tt
+  general\_null} network.  The {\tt general\_null} network allows the
+  user to define a list of non-reacting species builds the {\tt
+  network.f90} based on this list.  The makefile logic for building
+  the {\tt network.f90} is in the {\tt GPackage.mak} in {\tt
+  extern/networks/general\_null}.  The script {\tt write\_network.py}
+  in that directory does the actual parsing of the species file and
+  outputs the {\tt network.f90}.
+
 \end{itemize}
@@
 
-\subsection{Parallel Builds}
+\subsection{Parallel (MPI) Builds}
 
-{\tt BOXLIB\_USE\_MPI\_WRAPPERS}
+When building with MPI, the build system needs to know about the location of the
+MPI libraries.  If your local MPI has the {\tt mpif90} and {\tt mpicc} wrappers
+installed and working, then \maestro\ can use these simply by setting the
+environment variable:
+\begin{verbatim}
+BOXLIB_USE_MPI_WRAPPERS=1
+\end{verbatim}
+Otherwise, you will need to edit {\tt GMakeMPI.mak} and add a section
+specific to your machine with the compiler and library location.  It
+is best to simply copy an existing similar portion of the makefile and
+adjust it to your system.  Most national supercomputing facilities are
+already supported, and parallel builds on them should work out of the
+box.

### 2012-05-22 docs/makefiles/makefiles.tex
finish for now

@@
 
+Several modules are included in all \maestro\ builds by default.
+From \boxlib, we alway include:
+\begin{itemize}
+\item {\tt \$\{BOXLIB\_HOME\}/Src/F\_BaseLib}
+\item {\tt \$\{BOXLIB\_HOME\}/Src/LinearSolvers/F\_MG}
+\end{itemize}
+
+\noindent From {\tt extern}, we always include
+\begin{itemize}
+\item {\tt extern/model\_parser}
+\item {\tt extern/LAPACK}
+\item {\tt extern/random}
+\item {\tt extern/BLAS}
+\item {\tt \$\{EOS\_DIR\}}
+\item {\tt \$\{NETWORK\_DIR\}}
+\item {\tt \$\{CONDUCTIVITY\_DIR\}}
+\end{itemize}
+for most networks, we automatically add in {\tt extern/VODE} as well.
+
+\noindent From {\tt MAESTRO/}, we add
+\begin{itemize}
+\item {\tt MAESTRO/constants}
+\item {\tt MAESTRO/Source}
+\end{itemize}
+(although see the unit tests section below regarding {\tt MAESTRO/Source}.
+
+\noindent Finally, any extra directories listed in the {\tt EXTRA\_DIR}
+variable are included.
+
+For each of these included directories, {\tt GMaestro.mak} adds the list of source files
+defined in their {\tt GPackage.mak} to the list of files to be compiled.
+It also adds each of these directories to the {\tt vpath} as a directory
+for the build process to search in for source files.
+
 \subsection{Unit Tests}
+
+Sometimes we only want to use a few of the standard \maestro\
+routines, for example in a unit test where we are testing only a small
+part of the \maestro\ algorithm indepenedently.  In this case, we
+don't want to comple all of the files in {\tt MAESTRO/Source}.  If we
+set {\tt UNIT\_TEST := t} in our problem's {\tt GNUmakefile}, then the
+{\tt GPackage.mak} in {\tt MAESTRO/Source} is not read, so those files
+are not automatically put into the list of files to compile.  Instead,
+the problem should create its own {\tt GPackage.mak} listing only the
+subset of files that are to be compiled.  {\tt MAESTRO/Source} is put
+into the {\tt vpath} search path for sources, so those files will
+still be found as needed.
 
@@
 
-\subsection{Adding a Machine}
+\subsection{Adding a Compiler}
 
-\subsection{Adding a Compiler}
+Properties for different compilers are already defined in 
+{\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.
+Each compiler is given its own file.  The appropriate file is included
+into {\tt GMakedefs.mak} by looking at the {\tt COMP} variable and the
+operating system.  These compiler files define the compiler flags for
+both optimized and debug compiling.  Additionally, the variable {\tt
+FCOMP\_VERSION} should be defined there, based on the output from the
+compiler, to provide the compiler version for output into the {\tt
+job\_info} file at runtime.
 

### 2012-05-24 docs/visualization/visualization.tex
point to AmrPostprocessing
fix compilation
quick addition of notes for visit on Lens.  Need to be LaTeX-ified

@@
 
+\subsection{Remote Visualization on Lens}
+
+in .bashrc, add:
+
+export PATH=''/sw/analysis-x64/visit/bin/'':\$PATH
+
+in visit on bender:
+
+setup a new host
+
+   options -> host profiles
+
+      'new'
+
+   under 'host settings' tab:
+
+      host nickname: lens
+      remote host name: lens.ccs.ornl.gov
+
+      make sure your username is correct
+
+      check 'tunnel data connections through SSH'
+
+   under 'launch profiles' tab:
+
+       'new'
+       call it parallel
+
+       under 'parallel' tab:
+
+         check 'lauch parallel engine'
+         parallel launch method: qsub/mpirun
+         partition: computation
+         default number of processors: 8
+         default number of nodes: 2
+         default bank/account: AST006
+         default time limit: 00:30:00
+
+save the host setting by Options -> save settings
+
+then 'open', from the drop down list at the top, select lens.  You
+will be prompted for your password.  After that, you can navigate to
+the directory on lens with the data.
+
 %-----------------------------------------------------------------------------
@@
 %-----------------------------------------------------------------------------
-\section{Visualizing with {\tt data\_processing} scripts}
+\section{Python visualization scripts}
 \label{sec:vis:python}
 
-{\tt data\_processing/python\_plotfile} provide a simple commandline
+{\tt AmrPostprocessing/F\_Src/python\_plotfile} provide a simple commandline
 tool for doing visualizations of \boxlib\ plotfiles.  The main drivers

### 2012-05-24 docs/managing_jobs/managingjobs.tex
add Visit on Lens instructions

@@
 \end{verbatim}
+(Note: if the hsi command prompts you for your password, you will need to
+talk to the NCCS help desk to ask for password-less access to HPSS).
 
@@
 
-Additionally, if you have the path to the {\tt ftime} executable set in the
-script ({\tt ftime.f90} lives in {\tt fParallel/data\_processing/}), then the
+Additionally, if the {\tt ftime} executable is in your path 
+({\tt ftime.f90} lives in {\tt AmrPostprocessing/F\_src/}), then the
 script will create a file called {\tt ftime.out} that lists the name of the
@@
 Finally, right when the job is submitted, the script will tar up all
-of the diagnostic files created by {\tt diag.f90} and archive them on HPSS.
-The {\tt .tar} file is given a name that contains the date-string to allow
-multiple archives to co-exist.
+of the diagnostic files created by {\tt diag.f90} and {\tt ftime.out}
+and archive them on HPSS.  The {\tt .tar} file is given a name that
+contains the date-string to allow multiple archives to co-exist.
 
-This really allows you to run the job and have all of the data stored
-long term automatically.  This way you don't need to worry about
-filesystem purges.  It seems to work well.
-
-Also, the {\tt jaguar.run} submission script has code in it that will look
+The {\tt jaguar.run} submission script has code in it that will look
 at the most recently generated checkpoint files, make sure that they
 were written out correctly (it looks to see if there is a Header file,
-since that is the last thing written), and automatically set the
-{\tt --restart} flag on the run line to restart from the most recent
-checkpoint file.
-
-This allows you to job-chain a bunch of submission and have them wait
-until the previous job finished and then automatically queue up:
+since that is the last thing written), and automatically set the {\tt
+  --restart} flag on the run line to restart from the most recent
+checkpoint file.  This allows you to job-chain a bunch of submission
+and have them wait until the previous job finished and then
+automatically queue up:
 \begin{verbatim}
@@
 is the job submission script (e.g.\ {\tt atlas.run}).
+
+\section{Remote \visit\ Visualization on Lens}
+
+For large data files, visualization with \visit\ should be done with
+a client running on your local machine interfacing with \visit\ running
+on the remote machine.  For the {\tt lens} machine at NCCS, the proper setup
+is described below.
+
+First, on {\tt lens}, in your {\tt .bashrc}, add:
+\begin{verbatim}
+export PATH="/sw/analysis-x64/visit/bin/":$PATH
+\end{verbatim}
+(you would think that you could just add {\tt module load visit} but this
+does not seem to work with \visit.
+
+On your local machine, launch \visit.
+\begin{itemize}
+\item First we setup a new host
+
+  \begin{itemize}
+  \item From the menu, select {\em options $\rightarrow$ host profiles}
+  \item Create a new host by clicking on the {\em New Host} button.
+  \item Enter the {\em Host nickname} as {\tt lens}
+  \item Enter the {\em Remote host name} as {\tt lens.ccs.ornl.gov}
+  \item Make sure that your {\em username} is correct
+  \item Check {\em Tunnel data connections through SSH}
+  \end{itemize}
+
+\item Setup the {\em Launch Profiles}
+  \begin{itemize}
+  \item Click on the {\em Launch Profiles} tab
+  \item Click on the {\em New Profile} button
+  \item Enter the {\em Profile name} as {\tt parallel}
+  \item Click on the {\em Parallel} tab
+  \item Check {\em Launch parallel engine}
+  \item Select the {\em Parallel launch metho} as {\tt qsub/mpirun}
+  \item Set the {\em Partition / Pool / Queue} to {\tt computation}
+  \item Change the {\em Default number of processors} to 8
+  \item Set the {\em Default number of nodes} to 2
+  \item Set the {\em Default Bank / Account} to {\tt AST006}
+  \item Set the {\em Default Time Limit} to {\tt 00:30:00}
+  \end{itemize}
+
+\item Click on {\em Apply} and {\em Post}
+
+\item Save your changes by selecting {\em Options $\rightarrow$ Save Settings}
+\end{itemize}
+
+To do remote visualization, select {\em File $\rightarrow$ Open}.
+From the drop down list at the top, select {\tt lens}.  You will be
+prompted for your password.  After that, you can navigate to the
+directory on lens with the data.

### 2012-05-30 docs/getting_started/getting_started.tex
no longer need to copy in the helm_table.dat

@@
 
-\item {\em Copy in auxillary files}.
-
-This problem uses the {\tt helmeos} equation of state, so 
-we need to copy in the data table for the equation of state, {\tt
-  extern/EOS/helmeos/helm\_table.dat}.
-
-\begin{verbatim}
-cp ../../../extern/EOS/helmeos/helm_table.dat .
-\end{verbatim}
+Note, at the end of the build process, a link will be made in the current
+directory to the data
+table needed for the equation of state ({\tt extern/EOS/helmeos/helm\_table.dat}).
 

### 2012-05-30 docs/pert_notes/pert.tex
some labels
changed "predict_rhoprime_and_rhoX" to "predict_rhoX".  After we predict rhoX to edges, we set rho = sum(rhoX).  Also, in the enthalpy prediction types 2 and 4, we assemble (rho h) on edges using h_edge * sum(rho X_k)_edge instead of h_edge * (rho_0 + rho'_edge)

@@
+\label{ch:pert}
+
 \section{Predicting interface states}
@@
   + \rho'_\mathrm{edge} \right)(X_k)_\mathrm{edge}$ \\
-2 / {\tt predict\_rhoprime\_and\_rhoX}  &  
-  $\rho'_\mathrm{edge}$, $(\rho X_k)_\mathrm{edge}$ &
+2 / {\tt predict\_rhoX}  &  
+  $\sum(\rho X_k)_\mathrm{edge}$, $(\rho X_k)_\mathrm{edge}$ &
   $(\rho X_k)_\mathrm{edge}$ \\
@@
 \end{table}
-We note the labels {\tt predict\_rhoprime\_and\_X}, {\tt predict\_rhoprime\_and\_rhoX}, and
+We note the labels {\tt predict\_rhoprime\_and\_X}, {\tt predict\_rhoX}, and
 {\tt predict\_rho\_and\_X} are provided by the {\tt pred\_parameters}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Method 2: {\tt species\_pred\_type} = {\tt predict\_rhoprime\_and\_rhoX}}
+\subsection{Method 2: {\tt species\_pred\_type} = {\tt predict\_rhoX}}
 
@@
 
-We note that even though it is not needed here, we still predict
-$\rho'$ to the edges since certain enthalpy formulations need it.
+We note that even though it is not needed here, we still compute
+$\rho_\mathrm{edge}=\sum(\rho X_k)_\mathrm{edge}$ at the edges since certain 
+enthalpy formulations need it.
 
@@
 
-\noindent {\tt species\_pred\_type = predict\_rhoprime\_and\_rhoX}:
+\noindent {\tt species\_pred\_type = predict\_rhoX}:
 \begin{equation}
@@
 \hline
-2 / {\tt predict\_rhoprime\_and\_rhoX}  & 1 / {\tt predict\_rhohprime} &
+2 / {\tt predict\_rhoX}  & 1 / {\tt predict\_rhohprime} &
   $(\rho h)'$ & $(\rho h)'_\mathrm{edge}$ & 
-  $(\rho X)_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
+  $(\rho X)_\mathrm{edge}$, $\sum(\rho X)_\mathrm{edge}$ & 
   $\left [ (\rho h)_0^{n+\myhalf,{\rm avg}} + (\rho h)'_\mathrm{edge} \right ]$ \\
-2 / {\tt predict\_rhoprime\_and\_rhoX}  & 2 / {\tt predict\_h} &
+2 / {\tt predict\_rhoX}  & 2 / {\tt predict\_h} &
   $h$ & $h_\mathrm{edge}$ & 
-  $(\rho X)_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
-  $\left ( \rho_0^{n+\myhalf,{\rm avg}} + \rho'_\mathrm{edge} \right ) h_\mathrm{edge}$ \\
-2 / {\tt predict\_rhoprime\_and\_rhoX}  & 3 / {\tt predict\_T\_then\_rhohprime} &
+  $(\rho X)_\mathrm{edge}$, $\sum(\rho X)_\mathrm{edge}$ & 
+  $\sum(\rho X)_\mathrm{edge} h_\mathrm{edge}$ \\
+2 / {\tt predict\_rhoX}  & 3 / {\tt predict\_T\_then\_rhohprime} &
   $T$ & $(\rho h)'_\mathrm{edge}$ & 
-  $(\rho X)_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
+  $(\rho X)_\mathrm{edge}$, $\sum(\rho X)_\mathrm{edge}$ & 
   $\left [ (\rho h)_0^{n+\myhalf,{\rm avg}} + (\rho h)'_\mathrm{edge} \right ]$ \\
-2 / {\tt predict\_rhoprime\_and\_rhoX}  & 4 / {\tt predict\_T\_then\_h} &
+2 / {\tt predict\_rhoX}  & 4 / {\tt predict\_T\_then\_h} &
   $T$ & $h_\mathrm{edge}$ & 
-  $(\rho X)_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
-  $\left ( \rho_0^{n+\myhalf,{\rm avg}} + \rho'_\mathrm{edge} \right ) h_\mathrm{edge}$ \\
+  $(\rho X)_\mathrm{edge}$, $\sum(\rho X)_\mathrm{edge}$ & 
+  $\sum(\rho X)_\mathrm{edge} h_\mathrm{edge}$ \\
 \hline
@@
 
-For {\tt species\_pred\_types}: {\tt predict\_rhoprime\_and\_X} or
-{\tt predict\_rhoprime\_and\_rhoX}, we wish to construct $(\rho_0 +
-\rho'_\mathrm{edge} ) h_\mathrm{edge}$. \MarginPar{alternately, we could construct $\rho_\mathrm{edge} = \sum_k (\rho X_k)_\mathrm{edge}$ for s\_p\_t = 2}
+For {\tt species\_pred\_types}: {\tt predict\_rhoprime\_and\_X}, we wish to construct 
+$(\rho_0 + \rho'_\mathrm{edge} ) h_\mathrm{edge}$.
 
-For {\tt species\_pred\_types}: {\tt predict\_rho\_and\_X}, we wish to
-construct $\rho_\mathrm{edge} h_\mathrm{edge}$.
+For {\tt species\_pred\_types}: {\tt predict\_rho\_and\_X} or
+{\tt predict\_rhoX}, we wish to construct $\rho_\mathrm{edge} h_\mathrm{edge}$.
 
@@
 
-{\tt species\_pred\_types}: {\tt predict\_rhoprime\_and\_X} or
-{\tt predict\_rhoprime\_and\_rhoX}:\\[1mm]
+{\tt species\_pred\_types}: {\tt predict\_rhoprime\_and\_X}:\\[1mm]
 %
@@
 $\rho'_\mathrm{edge}$ and then multiply by $h_\mathrm{edge}$.
+
+\ \\
+{\tt species\_pred\_types}: {\tt predict\_rhoX}:\\[1mm]
+%
+We already have $\sum(\rho X_k)_\mathrm{edge}$ and simply multiply by
+$h_\mathrm{edge}$.
 
@@
   $(X_k)_\mathrm{edge}$ \\
-{\tt predict\_rhoprime\_and\_rhoX} & 
+{\tt predict\_rhoX} & 
   $\sum_k (\rho X_k)_\mathrm{edge}$ & 
@@
 
-For {\tt species\_pred\_types}: {\tt predict\_rhoprime\_and\_X} or
-{\tt predict\_rhoprime\_and\_rhoX}, we wish to construct $(\rho_0 +                                             \rho'_\mathrm{edge} ) h_\mathrm{edge}$.
+For {\tt species\_pred\_types}: {\tt predict\_rhoprime\_and\_X}, we wish to 
+construct $(\rho_0 + \rho'_\mathrm{edge} ) h_\mathrm{edge}$.
+
+For {\tt species\_pred\_types}: {\tt predict\_rhoX}, we wish to
+construct $\sum(\rho X_k)_\mathrm{edge} h_\mathrm{edge}$.
 

### 2012-05-30 Source/density_advance.f90

@@ -4,7 +4,8 @@ module density_advance_module
   use bl_types, only: dp_t
   use multifab_module, only: multifab, multifab_max_c, multifab_min_c, get_layout, &
                              build, destroy, setval, multifab_build_edge, &
-                             multifab_div_div_c, multifab_mult_mult_c
+                             multifab_div_div_c, multifab_mult_mult_c, multifab_copy_c, &
+                             multifab_plus_plus_c
   use ml_layout_module, only: ml_layout
   use define_bc_module, only: bc_level
 
@@ -111,12 +112,9 @@ contains
     ! ** density source term **
 
     ! Make source term for rho or rho' 
-    if ((species_pred_type == predict_rhoprime_and_X) .or. &
-        (species_pred_type == predict_rhoprime_and_rhoX)) then
+    if (species_pred_type == predict_rhoprime_and_X) then
        ! rho' source term
        !   . this is needed for pred_rhoprime_and_X
-       !   . we still do this for predict_rhoprime_and_rhoX since some 
-       !     methods for the enthalpy advancement need rho' edge states
        call modify_scal_force(scal_force,sold,umac,rho0_old, &
                               rho0_edge_old,w0,dx,rho0_old_cart,rho_comp, &
                               mla,the_bc_level)
@@ -131,9 +129,9 @@ contains
     ! ** species source term **
 
     ! for species_pred_types predict_rhoprime_and_X and
-    ! predict_rho_and_X, there is not force for X.
+    ! predict_rho_and_X, there is no force for X.
 
-    ! for predict_rhoprime_and_rhoX, we are predicting (rho X)
+    ! for predict_rhoX, we are predicting (rho X)
     ! as a conservative equation, and there is no force.
 
     if (spherical .eq. 1) then
@@ -161,12 +159,9 @@ contains
        call convert_rhoX_to_X(sold,.true.,mla,the_bc_level)
     endif
 
-    if ((species_pred_type == predict_rhoprime_and_X) .or. &
-        (species_pred_type == predict_rhoprime_and_rhoX)) then
+    if (species_pred_type == predict_rhoprime_and_X) then
        ! convert rho -> rho' in sold
        !   . this is needed for predict_rhoprime_and_X
-       !   . we still do this for predict_rhoprime_and_rhoX since some 
[…]

### 2012-05-30 Source/mkflux.f90

@@ -10,7 +10,7 @@
 !    We have rho' and X, and need a edge-centered base state to
 !    make the final fluxes
 !
-! predict_rhoprime_and_rhoX:
+! predict_rhoX:
 !    We use the (rho X) edge state directly to compute the fluxes.
 !    No base state input needed.
 !
@@ -241,7 +241,7 @@ contains
              sfluxx(i,comp) = &
                   (umac(i)+w0(i))*(rho0_edge+sedgex(i,rho_comp))*sedgex(i,comp)
 
-          else if (species_pred_type == predict_rhoprime_and_rhoX) then
+          else if (species_pred_type == predict_rhoX) then
              ! edge states are (rho X)
              sfluxx(i,comp) = &
                   (umac(i)+w0(i))*sedgex(i,comp)             
@@ -309,7 +309,7 @@ contains
                 sfluxx(i,j,comp) = umac(i,j)* &
                      (rho0_edge+sedgex(i,j,rho_comp))*sedgex(i,j,comp)
 
-             else if (species_pred_type == predict_rhoprime_and_rhoX) then
+             else if (species_pred_type == predict_rhoX) then
                 ! edge states are (rho X)
                 sfluxx(i,j,comp) = umac(i,j)*sedgex(i,j,comp)
 
@@ -334,7 +334,7 @@ contains
                 sfluxy(i,j,comp) = &
                      (vmac(i,j)+w0(j))*(rho0_edge+sedgey(i,j,rho_comp))*sedgey(i,j,comp)
 
-             else if (species_pred_type == predict_rhoprime_and_rhoX) then
+             else if (species_pred_type == predict_rhoX) then
                 ! edge states are (rho X)
                 sfluxy(i,j,comp) = &
                      (vmac(i,j)+w0(j))*sedgey(i,j,comp)
@@ -410,7 +410,7 @@ contains
                    sfluxx(i,j,k,comp) = &
                         umac(i,j,k)*(rho0_edge+sedgex(i,j,k,rho_comp))*sedgex(i,j,k,comp)
 
-                else if (species_pred_type == predict_rhoprime_and_rhoX) then
+                else if (species_pred_type == predict_rhoX) then
                    ! edge states are (rho X)
                    sfluxx(i,j,k,comp) = &
                         umac(i,j,k)*sedgex(i,j,k,comp)          
@@ -434,7 +434,7 @@ contains
[…]

### 2012-05-30 Source/pred_parameters.f90

@@ -4,7 +4,7 @@ module pred_parameters
 
    ! species prediction
    integer, parameter :: predict_rhoprime_and_X    = 1
-   integer, parameter :: predict_rhoprime_and_rhoX = 2
+   integer, parameter :: predict_rhoX              = 2
    integer, parameter :: predict_rho_and_X         = 3
 
    ! enthalpy prediction

### 2012-05-30 Source/rhoh_vs_t.f90

@@ -215,13 +215,10 @@ contains
           
           eos_state%xn(:) = sx(i,spec_comp:spec_comp+nspec-1)
           
-       else if (species_pred_type .eq. predict_rhoprime_and_rhoX) then
+       else if (species_pred_type .eq. predict_rhoX) then
           
-          ! interface states are rho' and (rho X)
-          eos_state%rho = ZERO
-          do n = 1, nspec
-             eos_state%rho = eos_state%rho + sx(i,spec_comp+n-1)
-          enddo
+          ! interface states are rho and (rho X)
+          eos_state%rho = sx(i,rho_comp)
           
           eos_state%xn(:) = sx(i,spec_comp:spec_comp+nspec-1)/eos_state%rho
           
@@ -304,13 +301,10 @@ contains
              
              eos_state%xn(:) = sx(i,j,spec_comp:spec_comp+nspec-1)
              
-          else if (species_pred_type .eq. predict_rhoprime_and_rhoX) then
+          else if (species_pred_type .eq. predict_rhoX) then
              
-             ! interface states are rho' and (rho X)
-             eos_state%rho = ZERO
-             do n = 1, nspec
-                eos_state%rho = eos_state%rho + sx(i,j,spec_comp+n-1)
-             enddo
+             ! interface states are rho and (rho X)
+             eos_state%rho = sx(i,j,rho_comp)
               
              eos_state%xn(:) = sx(i,j,spec_comp:spec_comp+nspec-1)/eos_state%rho
               
@@ -359,13 +353,10 @@ contains
              
              eos_state%xn(:) = sy(i,j,spec_comp:spec_comp+nspec-1)
 
-          else if (species_pred_type .eq. predict_rhoprime_and_rhoX) then
+          else if (species_pred_type .eq. predict_rhoX) then
               
-             ! interface states are rho' and (rho X)
-             eos_state%rho = ZERO
-             do n = 1, nspec
-                eos_state%rho = eos_state%rho + sy(i,j,spec_comp+n-1)
-             enddo
+             ! interface states are rho and (rho X)
+             eos_state%rho = sy(i,j,rho_comp)
               
              eos_state%xn(:) = sy(i,j,spec_comp:spec_comp+nspec-1)/eos_state%rho
               
@@ -453,13 +444,10 @@ contains
 
                 eos_state%xn(:) = sx(i,j,k,spec_comp:spec_comp+nspec-1)
 
[…]

### 2012-05-30 Source_SDC/density_advance.f90

@@ -4,7 +4,8 @@ module density_advance_module
   use bl_types, only: dp_t
   use multifab_module, only: multifab, multifab_max_c, multifab_min_c, get_layout, &
                              build, destroy, setval, multifab_build_edge, &
-                             multifab_div_div_c, multifab_mult_mult_c, multifab_plus_plus_c
+                             multifab_div_div_c, multifab_mult_mult_c,, multifab_copy_c, &
+                             multifab_plus_plus_c
   use ml_layout_module, only: ml_layout
   use define_bc_module, only: bc_level
 
@@ -112,12 +113,9 @@ contains
     ! ** density source term **
 
     ! Make source term for rho or rho' 
-    if ((species_pred_type == predict_rhoprime_and_X) .or. &
-        (species_pred_type == predict_rhoprime_and_rhoX)) then
+    if (species_pred_type == predict_rhoprime_and_X) then
        ! rho' source term
        !   . this is needed for pred_rhoprime_and_X
-       !   . we still do this for predict_rhoprime_and_rhoX since some 
-       !     methods for the enthalpy advancement need rho' edge states
        call modify_scal_force(scal_force,sold,umac,rho0_old, &
                               rho0_edge_old,w0,dx,rho0_old_cart,rho_comp, &
                               mla,the_bc_level)
@@ -132,9 +130,9 @@ contains
     ! ** species source term **
 
     ! for species_pred_types predict_rhoprime_and_X and
-    ! predict_rho_and_X, there is not force for X.
+    ! predict_rho_and_X, there is no force for X.
 
-    ! for predict_rhoprime_and_rhoX, we are predicting (rho X)
+    ! for predict_rhoX, we are predicting (rho X)
     ! as a conservative equation, and there is no force.
 
     if (spherical .eq. 1) then
@@ -167,12 +165,9 @@ contains
        call convert_rhoX_to_X(sold,.true.,mla,the_bc_level)
     endif
 
-    if ((species_pred_type == predict_rhoprime_and_X) .or. &
-        (species_pred_type == predict_rhoprime_and_rhoX)) then
+    if (species_pred_type == predict_rhoprime_and_X) then
        ! convert rho -> rho' in sold
        !   . this is needed for predict_rhoprime_and_X
-       !   . we still do this for predict_rhoprime_and_rhoX since some 
[…]

### 2012-05-31 docs/SDC/sdc.tex
some labels
changed "predict_rhoprime_and_rhoX" to "predict_rhoX".  After we predict rhoX to edges, we set rho = sum(rhoX).  Also, in the enthalpy prediction types 2 and 4, we assemble (rho h) on edges using h_edge * sum(rho X_k)_edge instead of h_edge * (rho_0 + rho'_edge)
final updates for the day
saving progress - new alg notes

--- 2012-02-13 docs/SDC/sdc.tex
@@
+\label{ch:sdc}
+
 \section{SDC Overview}
@@
 \frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + \underbrace{\nabla\cdot(\rho X_k\Ub)}_{A_{\rho X_k}}\label{eq:sdc:rhoX} \\
-\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + \underbrace{\nabla\cdot(\rho h\Ub)}_{A_{\rho h}} \label{eq:sdc:rhoh}
+\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + \underbrace{\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt}}_{A_{\rho h}} \label{eq:sdc:rhoh}
 \end{eqnarray}
@@
 
-{\bf Step 1:} ({\it Predictor}) In this step, we integrate $(\rho, \rho X_k, \rho h)$ over the full time step.
-The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^n$ are computed from the the thermodynamic variables at $t^n$.
-This step is divided into several sub-steps:\\
-
-{\bf Step 1A:} ({\it Compute advection velocities}) Use $\Ub^n$ and a second-order Godunov method to compute time-centered edge velocities, $\uadvsdcstar$, with time-lagged dynamic pressure and explicit buoyancy as forcing terms.
+{\bf Step 1:} ({\it Compute advection velocities})\\ \\
+Use $\Ub^n$ and a second-order Godunov method to compute time-centered edge velocities, $\uadvsdcstar$, with time-lagged dynamic pressure and explicit buoyancy as forcing terms.
 The $\star$ superscript indicates that this field does not satisfy the divergence constraint.
@@
 
-{\bf Step 1B:} ({\it Compute advective fluxes}) Use $\uadvsdcpred$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,{\rm pred}}$, with time-lagged reactions, explicit diffusion, and time-centered thermodynamic pressure as source terms.\\
+{\bf Step 2:} ({\it Predictor})\\ \\
+ In this step, we integrate $(\rho, \rho X_k, \rho h)$ over the full time step.
+The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^n$ are computed from the the thermodynamic variables at $t^n$.
+This step is divided into several sub-steps:\\
 
-{\bf Step 1C:} ({\it Compute diffusive fluxes}) Begin by compute a preliminary time-advanced density using
+{\bf Step 2A:} ({\it Compute advective flux divergences})\\ \\
+Use $\uadvsdcpred$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,(0)}$, with time-lagged reactions ($I^{\rm lagged} = I^{(j_{\rm max})}$ from the previous time step), explicit diffusion, and time-centered thermodynamic pressure as source terms.
+Define the advective flux divergences as
+\begin{eqnarray}
+A_{\rho X_k}^{(0)} &=& -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{(0)}}\uadvsdcpred\right],\\
+A_{\rho h}^{(0)} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,(0)}\uadvsdcpred\right].
+\end{eqnarray}
+Next, use these fluxes to compute a preliminary time-advanced density using
 \begin{equation}
-\frac{\rho^{n+1,{\rm pred}} - \rho^n}{\Delta t} = -\sum_m\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right].
+\frac{\rho^{n+1,(0)} - \rho^n}{\Delta t} = \sum_mA_{\rho X_k}^{(0)}.
 \end{equation}
@@
 \begin{equation}
-\frac{\rho^{n+1,{\rm pred}}X_{k,{\rm AD}}^{n+1,{\rm pred}} - (\rho X_k)^n}{\Delta t} = -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right] + I_{\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
+\frac{\rho^{n+1,(0)}X_{k,{\rm AD}}^{n+1,(0)} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{(0)} + I_{\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
 \end{equation}
-Then, solve a Crank-Nicolson-type diffusion equation for $h^{n+1,{\rm pred}}_{\rm AD}$, using transport coefficients evaluated at $t^n$ everywhere,
+
+{\bf Step 2B:} ({\it Compute diffusive flux divergence})\\ \\
+Solve a Crank-Nicolson-type diffusion equation for $h^{n+1,(0)}_{\rm AD}$, using transport coefficients evaluated at $t^n$ everywhere,
 \begin{eqnarray}
-\frac{\rho^{n+1,{\rm pred}}h^{n+1,{\rm pred}}_{\rm AD} - (\rho h)^n}{\Delta t} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,{\rm pred}}\uadvsdcpred\right] + \uadvsdcpred\cdot\nabla p_0 + I_{\rho h}^{\rm lagged}\nonumber\\
-&& + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n + \nabla\cdot\frac{\kth^n}{c_p^n}\nabla h_{\rm AD}^{n+1,{\rm pred}}\right)\nonumber\\
-&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,{\rm pred}}\right)\nonumber\\
+\frac{\rho^{n+1,(0)}h^{n+1,(0)}_{\rm AD} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(0)} + \uadvsdcpred\cdot\nabla p_0 + I_{\rho h}^{\rm lagged}\nonumber\\
+&& + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n + \nabla\cdot\frac{\kth^n}{c_p^n}\nabla h_{\rm AD}^{n+1,(0)}\right)\nonumber\\
+&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,(0)}\right)\nonumber\\
 && - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right).\label{eq:sdc enthalpy 2}
@@
 
-{\bf Step 1D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}^{\rm pred}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{\rm pred}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{\rho h}^{\rm lagged}$ term.
-Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,{\rm pred}}$ using the piecewise-constant advection and diffusion source terms:
+{\bf Step 2C:} ({\it Advance thermodynamic variables})\\ \\
+Define $Q_{\rho X_k}^{(0)}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{(0)}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{\rho h}^{\rm lagged}$ term.
+Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,(0)}$ using the piecewise-constant advection and diffusion source terms:
 \begin{eqnarray}
-\frac{\partial(\rho X_k)}{\partial t} &=& Q_{\rho X_k}^{\rm pred} + \rho\dot\omega_k\\
-\frac{\partial(\rho h)}{\partial t} &=& Q_{\rho h}^{\rm pred} + \rho\Hnuc.
+\frac{\partial(\rho X_k)}{\partial t} &=& Q_{\rho X_k}^{(0)} + \rho\dot\omega_k\\
+\frac{\partial(\rho h)}{\partial t} &=& Q_{\rho h}^{(0)} + \rho\Hnuc.
 \end{eqnarray}
-At this point we can define $I_{\rho X_k}^{\rm pred}$ and $I_{\rho h}^{\rm pred}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the corrector step.\\
+At this point we can define $I_{\rho X_k}^{(0)}$ and $I_{\rho h}^{(0)}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the corrector step.\\
 
-{\bf Step 2:} ({\it Corrector}) In the corrector, we use the time-advanced state from the predictor to perform a more accurate integration of the thermodynamic variables.
-The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^{n+1,{\rm pred}}$ are computed from $(\rho,\rho X_k,\rho h)^{n+1,{\rm pred}}$.
-This step is divided into several sub-steps:\\\\
-
-{\bf Step 2A:} ({\it Compute advection velocities}) First, compute $S^{n+\myhalf}$ and $\beta_0^{n+\myhalf}$ using
+{\bf Step 3:} ({\it Update advection velocities})\\ \\
+First, compute $S^{n+\myhalf}$ and $\beta_0^{n+\myhalf}$ using
 \begin{equation}
-S^{n+\myhalf} = \frac{S^n + S^{n+1,{\rm pred}}}{2}, \qquad \beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1,{\rm pred}}}{2}.
+S^{n+\myhalf} = \frac{S^n + S^{n+1,(0)}}{2}, \qquad \beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1,(0)}}{2}.
 \end{equation}
@@
 
-{\bf Step 2B:} ({\it Compute advective fluxes}) Use $\uadvsdc$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf}$, with iteratively-lagged reactions, explicit diffusion, and time-centered thermodynamic pressure as source terms.\\
+{\bf Step 4:} ({\it Corrector Loop})\\ \\
+We loop over this step from $j=1,j_{\rm max}$ times.
+In the corrector, we use the time-advanced state from the predictor to perform a more accurate integration of the thermodynamic variables.
+The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^{n+1,(j-1)}$ are computed from $(\rho,\rho X_k,\rho h)^{n+1,(j-1)}$.
+This step is divided into several sub-steps:\\
 
-{\bf Step 2C:} ({\it Compute diffusive fluxes}) Begin by computing the time-advanced density using
+{\bf Step 4A:} ({\it Compute advective flux divergences})\\ \\
+Use $\uadvsdc$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf}$, with iteratively-lagged reactions ($I^{(j-1)}$), explicit diffusion, and time-centered thermodynamic pressure as source terms.
+Define the advective flux divergences as
+\begin{eqnarray}
+A_{\rho X_k}^{(j)} &=& -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,(j)}\uadvsdc\right],\\
+A_{\rho h}^{(j)} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,(j)}\uadvsdc\right].
+\end{eqnarray}
+Next, use these fluxes to compute the final time-advanced density using
 \begin{equation}
-\frac{\rho^{n+1} - \rho^n}{\Delta t} = -\sum_m\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf}\uadvsdc\right].
+\frac{\rho^{n+1} - \rho^n}{\Delta t} = -\sum_mA_{\rho X_k}^{(j)}.
 \end{equation}
@@
 \begin{equation}
-\frac{\rho^{n+1}X_{k,{\rm AD}}^{n+1} - (\rho X_k)^n}{\Delta t} = -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf}\uadvsdc\right] + I_{\rho X_k}^{\rm pred}.\label{eq:sdc species 3}
+\frac{\rho^{n+1}X_{k,{\rm AD}}^{n+1,(j)} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{(j)} + I_{\rho X_k}^{(j-1)}.\label{eq:sdc species 3}
 \end{equation}
-Then, solve a backward-Euler-type correction equation for $h^{n+1}_{\rm AD}$,
+
+{\bf Step 4B:} ({\it Compute diffusive flux divergence})\\ \\
+Solve a backward-Euler-type correction equation for $h^{n+1,(j)}_{\rm AD}$,
 \begin{eqnarray}
-\frac{\rho^{n+1}h^{n+1}_{\rm AD} - (\rho h)^n}{\Delta t} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf}\uadvsdc\right] + \uadvsdc\cdot\nabla p_0 + I_{\rho h}^{\rm pred}\nonumber\\
-&& + \nabla\cdot\frac{\kth^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla h^{n+1}_{\rm AD} + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla h^{n+1,{\rm pred}}\right)\nonumber\\
-&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,{\rm pred}} k_{\rm th}^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla X_{k,{\rm AD}}^{n+1}\right)\nonumber\\
-&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,{\rm pred}}k_{\rm th}^{n+1,{\rm pred}}}{c_p^{n+1,{\rm pred}}}\nabla p_0\right)\label{eq:sdc enthalpy 3}
+\frac{\rho^{n+1}h^{n+1,(j)}_{\rm AD} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(j)} + \uadvsdc\cdot\nabla p_0 + I_{\rho h}^{(j-1)}\nonumber\\
+&& + \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j)}_{\rm AD} + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\
+&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla X_{k,{\rm AD}}^{n+1,(j)}\right)\nonumber\\
+&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,(j-1)}k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla p_0\right)\label{eq:sdc enthalpy 3}
 \end{eqnarray}
 
-{\bf Step 2D:} ({\it Advance thermodynamic variables}) Define $Q_{\rho X_k}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{\rho X_k}^{\rm pred}$ term, and define $Q_{\rho h}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{\rho h}^{\rm pred}$ term.
+{\bf Step 4C:} ({\it Advance thermodynamic variables})\\ \\
+Define $Q_{\rho X_k}^{(j)}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{\rho X_k}^{(j-1)}$ term, and define $Q_{\rho h}^{(j)}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{\rho h}^{(j-1)}$ term.
 Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1}$ using the piecewise-constant advection and diffusion source terms:
 \begin{eqnarray}
-\frac{\partial(\rho X_k)}{\partial t} &=& Q_{\rho X_k} + \rho\dot\omega_k\\
-\frac{\partial(\rho h)}{\partial t} &=& Q_{\rho h} + \rho\Hnuc.
+\frac{\partial(\rho X_k)}{\partial t} &=& Q_{\rho X_k}^{(j)} + \rho\dot\omega_k\\
+\frac{\partial(\rho h)}{\partial t} &=& Q_{\rho h}^{(j)} + \rho\Hnuc.
 \end{eqnarray}
-At this point we can define $I_{\rho X_k}^{\rm lagged}$ and $I_{\rho h}^{\rm lagged}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the predictor in the next time step.\\
+At this point we can define $I_{\rho X_k}^{(j)}$, $I_{\rho h}^{(j)}$, and any other $I$ terms we need depending on 
+our species and enthalpy edge state prediction types, for use in the predictor in the next time step.\\
 
-{\bf Step 3:} ({\it Advance velocity and dynamic pressure}) Same as the original \maestro\ algorithm.
+{\bf Step 5:} ({\it Advance velocity and dynamic pressure})\\ \\
+Similar to the original \maestro\ algorithm, more to come.

### 2012-06-01 docs/SDC/sdc.tex
sdc updates

@@
 \begin{itemize}
-\item {\tt species\_pred\_type} = 1 (predict $\rho'$ and $X_k$) 
-or 3 (predict $\rho$ and $X_k$)
+\item {\tt species\_pred\_type} = 1 ({\tt predict\_rhoprime\_and\_X})
+or 3 ({\tt predict\_rho\_and\_X})
 
@@
 \end{equation}
-\item {\tt species\_pred\_type} = 2 (predict $\rho'$ and $(\rho X_k)$)
+\item {\tt species\_pred\_type} = 2 ({\tt predict\_rhoX})
 \begin{equation}
@@
 \begin{itemize}
-\item {\tt enthalpy\_pred\_type} = {\tt predict\_rhohprime} (not implemented yet)
+\item {\tt enthalpy\_pred\_type} = 1 ({\tt predict\_rhohprime}, not implemented yet)
 
@@
 
-\item {\tt enthalpy\_pred\_type} = {\tt predict\_h}
+\item {\tt enthalpy\_pred\_type} = 2 ({\tt predict\_h} )
 
@@
 
-\item {\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_rhoprime} or 
-      {\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_h}
+\item {\tt enthalpy\_pred\_type} = 3 ({\tt predict\_T\_then\_rhoprime}) or 
+      {\tt enthalpy\_pred\_type} = 4 ({\tt predict\_T\_then\_h} )
 

### 2012-06-05 docs/architecture/architecture.tex
fix compiling
add note about BCs
update the build documentation

@@
 an executable.  Briefly, the \maestro\ sub-directories are:
+
 \begin{itemize}
-\item {\tt extern/}
-
-  External modules, like ODE integrators, equations of state, reaction
-  networks, etc.  Many of these routines were developed by other groups.
-
-  Important sub-directories include:
-
-  \begin{itemize}
-  \item {\tt conductivity/}
-
-    Various routines for computing the thermal conductivity used in the
-    thermal diffusion part of the algorithm.
-
-  \item {\tt EOS/}
-
-    A collection of various EOSes for use with the code.  The two
-    important ones for \maestro\ are {\tt helmeos/} and {\tt
-    gamma\_law\_general/}.
-
-  \item {\tt model\_parser/}
-
-    A simple Fortran module for reading in 1D initial model files.
-    This is used by the initialization routines to get the initial
-    model data.
-
-  \item {\tt networks/}
-
-    Various reaction networks for \maestro\ problems.  In addition to
-    providing a routine to evolve the nuclear species due to
-    reactions, the networks also define the species that are advected
-    by the code.
-
-  \item {\tt VODE/}
-
-    The {\tt VODE} \cite{vode} package for integrating ODEs.  At the
-    moment, this is used for integrating various reaction networks.
-
-  \end{itemize}
 
@@
 
-  \item {\tt initial\_models}
-
-    Several sub-directories containing routines to generate 1D initial
-    conditions in hydrostatic equilibrium for mapping onto the
-    \maestro\ grid.  See \S~\ref{sec:initial_models} for details.
+  \item {\tt Microphysics/}
+
+    The core microphysics routines used by \maestro.  These are organized
+    into the following sub-directories.
+
+    \begin{itemize}
+    \item {\tt conductivity/}
+
+      Various routines for computing the thermal conductivity used in
+      the thermal diffusion part of the algorithm.
+
+    \item {\tt EOS/}
+
+      A collection of various EOSes for use with the code.  The two
+      important ones for \maestro\ are {\tt helmeos/} and {\tt
+        gamma\_law\_general/}.
+
+    \item {\tt networks/}
+
+      Various reaction networks for \maestro\ problems.  In addition
+      to providing a routine to evolve the nuclear species due to
+      reactions, the networks also define the species that are
+      advected by the code.
+
+    \end{itemize}
 
@@
 
-  \item {\tt Source}
+  \item {\tt Source/}
 
@@
 
-  \item {\tt Source_SDC}
+  \item {\tt Source\_SDC/}
 
@@
 
-  \item {\tt TEST\_PROBLEMS}
+  \item {\tt TEST\_PROBLEMS/}
 
@@
 
-  \item {\tt UNIT\_TESTS}
+  \item {\tt UNIT\_TESTS/}
 
@@
     for details.
+
+  \item {\tt Util/}
+
+    Various helper routines exist in this directory.  Some of these
+    are externally developed.
+
+    \begin{itemize}
+
+    \item {\tt BLAS/}
+
+      Linear algebra routines.
+
+    \item {\tt model\_parser/}
+
+      A simple Fortran module for reading in 1D initial model files.
+      This is used by the initialization routines to get the initial
+      model data.
+
+    \item {\tt random/}
+
+      A random number generator.
+
+    \item {\tt VODE/}
+
+      The {\tt VODE} \cite{vode} package for integrating ODEs.  At the
+      moment, this is used for integrating various reaction networks.
+
+   \end{itemize}
 
@@
 
-The next line defines where the top of the source tree is located.
+The next line defines where the top of the \maestro\ source tree is located.
 \begin{lstlisting}[language={[gnu]make},mathescape=false]
-  FPARALLEL := ../..
+  MAESTRO_TOP_DIR := ../..
 \end{lstlisting}
@@
 \begin{lstlisting}[language={[gnu]make},mathescape=false]
-EOS_DIR := extern/EOS/helmeos   
-CONDUCTIVITY_DIR := extern/conductivity/constant
-NETWORK_DIR := extern/networks/ignition
-
-EXTRA_DIR := extern/random
-\end{lstlisting}
+EOS_DIR := helmeos   
+CONDUCTIVITY_DIR := constant
+NETWORK_DIR := ignition_simple
+
+EXTRA_DIR := Util/random
+\end{lstlisting}
+Note that the microphysics packages are listed simply by the name of 
+the directory containing the specific implementation (e.g. {\tt helmeos}).
+By default, the build system will look in {\tt Microphysics/EOS/} for
+the EOS, {\tt Microphysics/conductivity/} for the conductivity routine,
+and {\tt Microphysics/networks/} for the reaction network.  To
+override this default search path, you can set {\tt EOS\_TOP\_DIR},
+{\tt CONDUCTIVITY\_TOP\_DIR}, and {\tt NETWORK\_TOP\_DIR} respectively.
+
 Generally, one does not need to include the problem directory itself
@@
 \begin{lstlisting}[language={[gnu]make},mathescape=false]
-  include $(FPARALLEL)/MAESTRO/GMaestro.mak
+  include $(MAESTRO_TOP_DIR)/GMaestro.mak
 \end{lstlisting}
@@
 The actual filling of the ghostcells according to the descriptions
-contained in the \bctower\ is carried out by the {\tt multifab\_physbc}
-routine.
+contained in the \bctower\ is carried out by the {\tt
+  multifab\_physbc} routine.  When you have an {\tt EXT\_DIR}
+condition in {\tt multifab\_physbc} (an specified in the inputs file
+as {\tt inlet}), the advection solver (via the slope routine) and
+linear solvers will then assume that the value in the ghost cells is
+equal to the value that actually lies on the wall.
 

### 2012-06-05 docs/makefiles/makefiles.tex
new build system

@@
 
-  This variable points to the path (relative to {\tt fParallel/}) of
+  This variable points to the directory (by default, relative to {\tt Microphysics/EOS/}) of
   the equation of state used for the build.  Choices that work
@@
   \end{itemize}
+  To use an EOS contained in a different location, set the variable {\tt EOS\_TOP\_DIR}
+  to point to the directory above the alternate EOS directory.
 
@@
   This variable points to the conductivity routine used for the build
-  (relative to {\tt fParallel/}).  Choices that work with \maestro\
-  are:
+  (by default, relative to {\tt Microphysics/conductivity/}).  Choices
+  that work with \maestro\ are: 
+  %
   \begin{itemize} 
@@
   \end{itemize} 
+  % 
   If diffusion is not being used for the problem, this should be set
-  to {\tt extern/conductivity/constant}.
+  to {\tt extern/conductivity/constant}.  To use an alternate conductivity
+  routine, set the variable {\tt CONDUCTIVITY\_TOP\_DIR} to point
+  to the directory above the alternate conductivity directory.
 
@@
 
-  This variable points to the reaction network used for the build
-  (relative to {\tt fParallel/}).  A lot of options are present in
-  {\tt extern/networks/}.  A network is required even if you are not
-  doing reactions, since the network defines the species that are
-  advected and interpreted by the equation of state.
+  This variable points to the reaction network used for the build (by
+  default, relative to {\tt Microphysics/networks/}).  Several options
+  are present in {\tt extern/networks/}.  A network is required even
+  if you are not doing reactions, since the network defines the
+  species that are advected and interpreted by the equation of state.
 
@@
   source.
+
+  To use an alternate reaction network, set the variable {\tt
+  NETWORK\_TOP\_DIR} to point to the directory above the alternate
+  network.
 

### 2012-06-05 docs/getting_started/getting_started.tex
update for recent directory changes

@@
 specified in {\tt BOXLIB\_HOME/Tools/F\_mk/}.  (Alternate compiler
-choices include {\tt Intel}, {\tt PGI}, {\tt PathScale}.)
+choices include {\tt Intel}, {\tt PGI}, {\tt PathScale}, {\tt Cray}.)
 
@@
 on your Linux system, you may also need to install {\tt libXpm} and
-related development packages (e.g.\ {\tt libXpm-devel}).  Also, in
-{\tt BoxLib/Tools/C\_mk/Make.defs}, the Fortran libraries are
-explictly linked in by asking {\tt gfortran} for the location of {\tt
-libgfortran.a}.  You may need to explicitly install the static
-libraries for {\tt gfortran} (i.e.\ {\tt libgfortran-static}), or
-change the library line to query for the shared-object version {\tt
-libgfortran.so}.
+related development packages (e.g.\ {\tt libXpm-devel}).
 
@@
 produce custom diagnostic output.  This information can be plotted
-directly with {\sl GNUplot}, for example.
+directly with {\sf GNUplot}, for example.
 
@@
 sub-directories under {\tt MAESTRO/} ({\tt SCIENCE/}, {\tt
-TEST\_PROBLEMS}, or {\tt UNIT\_TESTS}).  The {\tt GNUmakefile} in each
+TEST\_PROBLEMS/}, or {\tt UNIT\_TESTS/}).  The {\tt GNUmakefile} in each
 problem directory lists the components of {\tt MAESTRO} that are used
@@
 cause the star to expand.  This is the 3D analogue of the {\tt
-  test\_basestate} unit test (see \S~\ref{sec:unit_tests}).  This
+  test\_basestate} unit test (see \S~\ref{chapter:unit_tests}).  This
 problem was discussed in \cite{multilevel}.
@@
 ${\tt n\_cellx} \cdot 2^{({\tt max\_levels} -1)}$.
+
+\section{Parallel Jobs}
+
+To run in parallel with MPI, you would set {\tt MPI := t} in your {\tt
+GNUmakefile}.  For a machine with working MPI compiler wrappers ({\tt
+mpif90} and {\tt mpicc}), setting the environment variable {\tt
+BOXLIB\_USE\_MPI\_WRAPPERS=1} will have the code build using these
+wrappers.  This is the easiest way to do a parallel build, and should
+work on most Linux systems.
+
+More generally, the build system needs to know about your MPI
+installation.  For popular national computing facilities, this is
+already setup, and the build system looks at the machine hostname to
+set the proper libraries.  For other machines, you may need to edit
+the {\tt GMake.MPI} file in the \boxlib\ build files.  See
+\S~\ref{ch:make} for more details.
+
+OpenMP can be used to parallelize on shared-memory machines (i.e.\
+within a node).  OpenMP support is accomplished through the compiler.
+Setting {\tt OMP := t} in the {\tt GNUmakefile} will enable the proper
+compiler flags to build with OpenMP.  Note: not all \maestro\ modules 
+have OpenMP support.  Microphysics routines need to be written in a
+threadsafe manner.  This can be tested via the {\tt test\_react} unit
+test (see ~\S~\ref{chapter:unit_tests}).

### 2012-06-05 docs/preface/preface.tex
boxlib
minor tweak
update for recent directory changes

@@
 
+\noindent \maestro\ uses the \boxlib\ library,
+developed at the Center for Computational Sciences and
+Engineering at Lawrence Berkeley National Laboratory.  \boxlib\
+is distributed separately from \maestro\ (see https:/\hspace{-0.25em}/ccse.lbl.gov/BoxLib/ )
+\\
+
+\noindent The {\tt helmeos} equation of state was developed by Frank Timmes (ASU).
+The version in \maestro\ was downloaded from his website and the
+interface changes slightly for threadsafeness.  We thank Frank for
+making it available.
+
 \noindent The current version of the \maestro\ User's Guide can be found in 

### 2012-06-06 docs/compilers/compilers.tex
add note about 4.7

@@
 \item gfortran 4.6.1--4.6.3 on Fedora (15 and 16) have a bug that 
-shows up in multi-processor runs ({\tt toy\_convect} on 4 processors, for instance).
-The symptom is a crash during initialization with the message that
-``message sizes do not match across processes in the collective'' in a
-{\tt parallel\_bcast} call.  This is in {\tt cluster\_f.f90}, and the issue seems
-to be a multiplication error in computing the size to allocate {\tt bxs}.
-The bug is not present when you compile in debug mode, so it seems to be
-related to optimization.
+shows up in multi-processor, multilevel runs ({\tt toy\_convect} on 4
+processors, with {\tt inputs\_2d\_ml.test} for instance).  The symptom
+is a crash during initialization with the message that ``message sizes
+do not match across processes in the collective'' in a {\tt
+parallel\_bcast} call.  This is in {\tt cluster\_f.f90}, and the issue
+seems to be a multiplication error in computing the size to allocate
+{\tt bxs}.  The bug is not present when you compile in debug mode, so
+it seems to be related to optimization.
+\item gfortran 4.7.1 (prerelease) seems to have fixed the above issue.
 \end{itemize}

### 2012-06-20 docs/makefiles/makefiles.tex
add not about BOXLIB_ONLY

@@
 
+\subsection{\boxlib-only Tests}
+
+An even more restrictive setting than {\tt UNIT\_TEST := t} is invoked
+by setting {\tt BOXLIB\_ONLY := t}.  This is like the unit test flag,
+but does not include {\tt MAESTRO/Source} in the {\tt vpath} search
+path for sources.  So this is intended for cases where we don't want
+to use any \maestro\ source files.  Typically, this is used in the
+small unit tests that live under the various microphysics solvers.  If
+a {\tt probin.f90} is built for these tests, it will not include all
+the \maestro-specific parameters, but will include any parameters from
+the various microphysics routines.
+
 \section{Special Targets}

### 2012-06-29 docs/architecture/architecture.tex
add note about new EOS interface

@@
 
-The {\tt eos\_module} declares for convenience the variables that need
-appear in the {\tt eos} call argument list.  \maestro\ routines use
-these module variables in the EOS call to avoid having to declare 
-each quantity in each routine that calls the EOS.
+The thermodynamic quantities are stored in a Fortran type {\tt eos\_t},
+which has fields for all the thermodynamic inputs and outputs.  The 
+type definition is brought in through {\tt eos\_type\_module}.
+\footnote{ Note: an older interface to the EOS exists, but is
+  deprecated.  In this mode, the {\tt eos\_module} declares for
+  convenience the variables that need appear in the {\tt eos} call
+  argument list.  \maestro\ routines use these module variables in the
+  EOS call to avoid having to declare each quantity in each routine
+  that calls the EOS.}
 

### 2012-06-29 docs/makefiles/makefiles.tex
update to reflect recent changes

@@
 
-\noindent From {\tt extern}, we always include
+\noindent From {\tt Util}, we always include
 \begin{itemize}
-\item {\tt extern/model\_parser}
-\item {\tt extern/LAPACK}
-\item {\tt extern/random}
-\item {\tt extern/BLAS}
-\item {\tt \$\{EOS\_DIR\}}
-\item {\tt \$\{NETWORK\_DIR\}}
-\item {\tt \$\{CONDUCTIVITY\_DIR\}}
+\item {\tt Util/model\_parser}
+\item {\tt Util/random}
 \end{itemize}
-for most networks, we automatically add in {\tt extern/VODE} as well.
 
-\noindent From {\tt MAESTRO/}, we add
+The microphysics, as described above is also included.  For the networks,
+we include a file called {\tt NETWORK\_REQUIRES} into {\tt GMaestro.mak}
+that tells us whether to also include {\tt Util/VODE} (if {\tt NEED\_VODE := t})
+or {\tt Util/BLAS} (if {\tt NEED\_BLAS := t}).
+
+From {\tt MAESTRO/}, we add
 \begin{itemize}
@@
 
-For each of these included directories, {\tt GMaestro.mak} adds the list of source files
-defined in their {\tt GPackage.mak} to the list of files to be compiled.
-It also adds each of these directories to the {\tt vpath} as a directory
-for the build process to search in for source files.
+For each of these included directories, {\tt GMaestro.mak} adds the
+list of source files defined in their {\tt GPackage.mak} to the list
+of files to be compiled.  It also adds each of these directories to
+the {\tt vpath} as a directory for the build process to search in for
+source files.
 

### 2012-07-05 docs/managing_jobs/managingjobs.tex
add notes on remote visiting a movie

@@
   \item Enter the {\em Remote host name} as {\tt lens.ccs.ornl.gov}
+  \item Enter the {\em Path to Visit installation} as {\tt /sw/analysis-x64/visit} (not sure if this is needed)
   \item Make sure that your {\em username} is correct
@@
 directory on lens with the data.
+
+To make a movie (output sequence of images):
+\begin{itemize}
+\item save a view in \visit\ you like as a session file (File $\rightarrow$ Save session).  
+\item On lens, create a file called {\tt files.visit} which lists all
+  of the files you want to visualize, one per line, with {\tt /Header}
+  after the filename.  This can be done simply as:
+  \begin{verbatim}
+  ls -1 | grep -v processed | awk '{print $1"/Header"}' > files.visit
+  \end{verbatim}
+  %$
+  (note: the {\tt processed} bit is for when you used the script above to 
+  automatically archive the data).
+
+\item Edit the session file, searching for the name of the plotfile you
+  originally visualized, and change it to read {\tt files.visit}.  Make
+  sure that the path is correct.  This may appear multiple times.
+
+\item Relaunch \visit\ locally and restore the session (File $\rightarrow$ Restore session).
+It will render the first image.
+Then reopen (File $\rightarrow$ ReOpen file).
+After this is done, the buttons that allow you to move through the files should become active (black).
+
+\item Resave the session file
+
+\item To generate the frames, you have 2 options:
+
+  \begin{enumerate}
+  \item File $\rightarrow$ Save movie.  Pick {\em New simple movie},
+    then set the format to {\em PNG} and add this to the output box by
+    clicking the right arrow, then in the very last screen, select:
+    {Later, tell me the command to run}.
+
+   \visit\ will pop up a box showing the command to run.  Trying to
+   get the currently running session of \visit\ to generate the frames
+   seems problamatic.  Note: you will probably want to edit out the
+   {\tt -v x.x.x} argument in the commandline to not have it force
+   to use a specific version.
+
+  \item If the session file successfully does the remote visualization
+   as desired, you can run the movie via the commandline with something like:
+
+   \begin{verbatim}
+   visit -movie -format png -geometry 1080x1080 -output subchandra_cutoff3_ \
+       -start 0 -end 44 -sessionfile subchandra_radvel.session
+   \end{verbatim}
+
+  \end{enumerate}
+
+\end{itemize}

### 2012-07-10 docs/test_suite/test_suite.tex
add the test suite
remove references to fParallel

@@
 \centering
-\includegraphics[width=5.0in]{\suiteFigPath/testsuite}
+\includegraphics[width=5.0in]{\suitefigpath/testsuite}
 \caption{\label{fig:test_suite_main} Main test suite results page.  Each 
@@
 compareToolDir = /home/zingale/tests/AmrPostprocessing/F_Src/
-helmeosDir     = /home/zingale/tests/MAESTRO/fParallel/extern/EOS/helmeos/
+helmeosDir     = /home/zingale/tests/MAESTRO/Microphysics/EOS/helmeos/
 
-sourceTree = fParallel
+sourceTree = F_Src
 
@@
   boxLibDir} points to the location of the \boxlib\ source, {\tt
-  sourceDir} points to the top-level \maestro\ directory, which is
-expected to contain the {\tt fParallel} subdirectories.  {\tt
+  sourceDir} points to the top-level \maestro\ directory.  {\tt
   testTopDir} refers to the directory that the suite should use as its
@@
 
-Next, we set {\tt sourceTree} to {\tt fParallel}, indicating that
+Next, we set {\tt sourceTree} to {\tt F\_Src}, indicating that
 \maestro\ is built using the Fortran \boxlib\ framework.  This option
@@
 [test2]
-buildDir = fParallel/MAESTRO/TEST_PROBLEMS/test2/
+buildDir = TEST_PROBLEMS/test2/
 inputFile = inputs_2d

### 2012-07-17 docs/SDC/sdc.tex
notes about the form of the diffusion equations

@@
 \begin{eqnarray}
-\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) + \underbrace{\nabla\cdot(\rho X_k\Ub)}_{A_{\rho X_k}}\label{eq:sdc:rhoX} \\
-\frac{d(\rho h)}{dt}   &=& \rho \Hnuc + \underbrace{\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt}}_{A_{\rho h}} \label{eq:sdc:rhoh}
+\frac{d(\rho X_k)}{dt} &=& \rho \omegadot_k(\rho,X_k,T) - \underbrace{\nabla\cdot(\rho X_k\Ub)}_{A_{\rho X_k}}\label{eq:sdc:rhoX} \\
+\frac{d(\rho h)}{dt}   &=& \rho \Hnuc - \underbrace{\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt}}_{A_{\rho h}} \label{eq:sdc:rhoh}
 \end{eqnarray}
@@
 \subsection{Enthalpy Source Terms.}
-The obvious analog to the above form for $I_{\rho X_k}$ is to define
-\begin{equation}
-I_{\rho h}  = \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h}
-\label{eq:sdc:Irhoh}
-\end{equation}
-This can be thought of as an approximation to $(\rho \Hnuc)$,
-but this does not match any of the quantities we predict for the
-enthalpy update ({\tt enthalpy\_pred\_type}, see \S
-\ref{sec:pred:enthalpy}).  The appropriate constructions
-are:
+The appropriate constructions are:
 
 \begin{itemize}
+\item {\tt enthalpy\_pred\_type} = 0 ({\tt predict\_rhoh})
+\begin{equation}
+I_{\rho h} = \frac{(\rho h)^{\rm new} - (\rho h)^{\rm old}}{\Delta t} - A_{\rho h}.
+\end{equation}
+
 \item {\tt enthalpy\_pred\_type} = 1 ({\tt predict\_rhohprime}, not implemented yet)
@@
 A_{\rho X_k}^{(0)} &=& -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{(0)}}\uadvsdcpred\right],\\
-A_{\rho h}^{(0)} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,(0)}\uadvsdcpred\right].
+A_{\rho h}^{(0)} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,(0)}\uadvsdcpred\right] + \uadvsdcpred\cdot\nabla p_0.
 \end{eqnarray}
-Next, use these fluxes to compute a preliminary time-advanced density using
+Next, use these fluxes to compute the time-advanced density,
 \begin{equation}
-\frac{\rho^{n+1,(0)} - \rho^n}{\Delta t} = \sum_mA_{\rho X_k}^{(0)}.
+\frac{\rho^{n+1} - \rho^n}{\Delta t} = \sum_mA_{\rho X_k}^{(0)}.
 \end{equation}
@@
 \begin{equation}
-\frac{\rho^{n+1,(0)}X_{k,{\rm AD}}^{n+1,(0)} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{(0)} + I_{\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
+\frac{\rho^{n+1}X_{k,{\rm AD}}^{n+1,(0)} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{(0)} + I_{\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
 \end{equation}
@@
 \begin{eqnarray}
-\frac{\rho^{n+1,(0)}h^{n+1,(0)}_{\rm AD} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(0)} + \uadvsdcpred\cdot\nabla p_0 + I_{\rho h}^{\rm lagged}\nonumber\\
+\frac{\rho^{n+1}h^{n+1,(0)}_{\rm AD} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(0)} + I_{\rho h}^{\rm lagged}\nonumber\\
 && + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n + \nabla\cdot\frac{\kth^n}{c_p^n}\nabla h_{\rm AD}^{n+1,(0)}\right)\nonumber\\
 && + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,(0)}\right)\nonumber\\
-&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right).\label{eq:sdc enthalpy 2}
+&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right),\label{eq:sdc enthalpy 2}
+\end{eqnarray}
+which is equivalent to
+\begin{eqnarray}
+\left(\rho^{n+1} - \frac{\Delta t}{2}\nabla\cdot\frac{k_{\rm th}^n}{c_p^n}\nabla\right)h_{\rm AD}^{n+1,(0)} &=& (\rho h)^n + \Delta t\Bigg[A_{\rho h}^{(0)} + I_{\rho h}^{\rm lagged} + \left(\half\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n\right)\nonumber\\
+&&\hspace{0.85in} + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,(0)}\right)\nonumber\\
+&&\hspace{0.85in} - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right)\Bigg].
 \end{eqnarray}
@@
 \end{eqnarray}
-At this point we can define $I_{\rho X_k}^{(0)}$ and $I_{\rho h}^{(0)}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the corrector step.\\
+At this point we can define $I_{\rho X_k}^{(0)}$ and $I_{\rho h}^{(0)}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the corrector step.
+In our first implementation, we are predicting $\rho X_k$ and $\rho h$, in which case we define:
+\begin{eqnarray}
+I_{\rho X_k}^{(0)} &=& \frac{(\rho X_k)^{n+1,(0)} - (\rho X_k)^n}{\Delta t} - Q_{\rho X_k}^{(0)}\\
+I_{\rho h}^{(0)} &=& \frac{(\rho h)^{n+1,(0)} - (\rho h)^n}{\Delta t} - Q_{\rho h}^{(0)}.
+\end{eqnarray}
 
@@
 A_{\rho X_k}^{(j)} &=& -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,(j)}\uadvsdc\right],\\
-A_{\rho h}^{(j)} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,(j)}\uadvsdc\right].
+A_{\rho h}^{(j)} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,(j)}\uadvsdc\right] + \uadvsdc\cdot\nabla p_0.
 \end{eqnarray}
-Next, use these fluxes to compute the final time-advanced density using
-\begin{equation}
-\frac{\rho^{n+1} - \rho^n}{\Delta t} = -\sum_mA_{\rho X_k}^{(j)}.
-\end{equation}
 Then, compute preliminary, time-advanced species using
@@
 \begin{eqnarray}
-\frac{\rho^{n+1}h^{n+1,(j)}_{\rm AD} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(j)} + \uadvsdc\cdot\nabla p_0 + I_{\rho h}^{(j-1)}\nonumber\\
+\frac{\rho^{n+1}h^{n+1,(j)}_{\rm AD} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(j)} + I_{\rho h}^{(j-1)}\nonumber\\
 && + \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j)}_{\rm AD} + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\
 && + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla X_{k,{\rm AD}}^{n+1,(j)}\right)\nonumber\\
-&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,(j-1)}k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla p_0\right)\label{eq:sdc enthalpy 3}
+&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,(j-1)}k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla p_0\right),\label{eq:sdc enthalpy 3}
+\end{eqnarray}
+which is equivalent to
+\begin{eqnarray}
+\left(\rho^{n+1} - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla\right)h^{n+1,(j)}_{\rm AD} &=& (\rho h)^n + \Delta t\Bigg[A_{\rho h}^{(j)} + I_{\rho h}^{(j-1)} \nonumber\\
+&&+ \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\
+&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla X_{k,{\rm AD}}^{n+1,(j)}\right)\nonumber\\
+&&- \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,(j-1)}k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla p_0\right)\Bigg].
 \end{eqnarray}
@@
 Define $Q_{\rho X_k}^{(j)}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{\rho X_k}^{(j-1)}$ term, and define $Q_{\rho h}^{(j)}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{\rho h}^{(j-1)}$ term.
-Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1}$ using the piecewise-constant advection and diffusion source terms:
+Use {\tt VODE} to integerate (\ref{eq:sdc species}) and (\ref{eq:sdc enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,(j)}$ using the piecewise-constant advection and diffusion source terms:
 \begin{eqnarray}
@@
 At this point we can define $I_{\rho X_k}^{(j)}$, $I_{\rho h}^{(j)}$, and any other $I$ terms we need depending on 
-our species and enthalpy edge state prediction types, for use in the predictor in the next time step.\\
+our species and enthalpy edge state prediction types, for use in the predictor in the next time step.  In our first implementation, we are predicting $\rho X_k$ and $\rho h$, in which case we define:
+\begin{eqnarray}
+I_{\rho X_k}^{(j)} &=& \frac{(\rho X_k)^{n+1,(j)} - (\rho X_k)^n}{\Delta t} - Q_{\rho X_k}^{(j)}\\
+I_{\rho h}^{(j)} &=& \frac{(\rho h)^{n+1,(j)} - (\rho h)^n}{\Delta t} - Q_{\rho h}^{(j)}.
+\end{eqnarray}
 

### 2012-07-17 docs/SDC/sdc.tex
typo

@@
 \begin{eqnarray}
-\left(\rho^{n+1} - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla\right)h^{n+1,(j)}_{\rm AD} &=& (\rho h)^n + \Delta t\Bigg[A_{\rho h}^{(j)} + I_{\rho h}^{(j-1)} \nonumber\\
+\left(\rho^{n+1} - \Delta t\nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla\right)h^{n+1,(j)}_{\rm AD} &=& (\rho h)^n + \Delta t\Bigg[A_{\rho h}^{(j)} + I_{\rho h}^{(j-1)} \nonumber\\
 &&+ \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\

### 2012-07-17 docs/flowchart/flowchart.tex
more descriptions of the initialization

@@
 
-\item {\bf Do} {$\nu = 1,...,N_{\rm iters}^{S}$.}
+\item {\em Initial Projection}: if {\tt do\_initial\_projection = T}, then we 
+   first project the velocity field with $\rho = 1$ and $\beta_0^0$ defined 
+   from the initial data.  No sources are included in $S$.  The elliptic
+   equation we solve is 
+   \begin{equation}
+   \nabla \cdot \beta_0^0 \nabla \phi = - \nabla \cdot (\beta_0^0 \Ub^0)
+   \end{equation}
+   This $\phi$ is then used to correct the velocity field.
+
+\item {\em ``Divu'' iterations}: Next we do {\tt init\_divu\_iter} iterations 
+  to project the velocity field using a constraint that sees the reactions. 
+  The initial timestep estimate is provided by {\tt firstdt}, to allow us to 
+  integrate the reaction networks.  In this projection, again the density is 
+  set to 1, and the elliptic equation we solve is:
+   \begin{equation}
+   \nabla \cdot \beta_0^0 \nabla \phi = \beta_0 (S - \Sbar)- \nabla \cdot (\beta_0^0 \Ub^0)
+   \end{equation}
+
+  {\bf Do} {$\nu = 1,...,N_{\rm iters}^{S}$.}
   \begin{enumerate}
@@
 \renewcommand{\labelenumii}{\roman{enumii}.}
-\addtocounter{enumi}{1}
+\addtocounter{enumi}{2}
 \item {\bf Do} {$\nu = 1,...,N_{\rm iters}^{\pi}$.}

### 2012-07-17 docs/flowchart/flowchart.tex
more documenting of the initial iterations

@@
 \addtocounter{enumi}{2}
-\item {\bf Do} {$\nu = 1,...,N_{\rm iters}^{\pi}$.}
+\item {\it Pressure iterations}: Here we do {\tt init\_iter} iterations to get an
+  approximation for the lagged pressure:
+
+  {\bf Do} {$\nu = 1,...,N_{\rm iters}^{\pi}$.}
   \begin{enumerate}
@@
     (The motivation for this form of the projection in the initial pressure iterations
-    is discussed in (cite almgren:bell:crutchfield).)
+    is discussed in \cite{almgren:bell:crutchfield}.)
       We discard the new velocity resulting from this, but keep the new  

### 2012-07-17 docs/SDC/sdc.tex
more sdc fixes

--- 2012-07-17 docs/SDC/sdc.tex
@@
 && + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n + \nabla\cdot\frac{\kth^n}{c_p^n}\nabla h_{\rm AD}^{n+1,(0)}\right)\nonumber\\
-&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,(0)}\right)\nonumber\\
+&& - \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,(0)}\right)\nonumber\\
 && - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right),\label{eq:sdc enthalpy 2}
@@
 \left(\rho^{n+1} - \frac{\Delta t}{2}\nabla\cdot\frac{k_{\rm th}^n}{c_p^n}\nabla\right)h_{\rm AD}^{n+1,(0)} &=& (\rho h)^n + \Delta t\Bigg[A_{\rho h}^{(0)} + I_{\rho h}^{\rm lagged} + \left(\half\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n\right)\nonumber\\
-&&\hspace{0.85in} + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,(0)}\right)\nonumber\\
+&&\hspace{0.85in} - \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,(0)}\right)\nonumber\\
 &&\hspace{0.85in} - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right)\Bigg].
@@
 && + \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j)}_{\rm AD} + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\
-&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla X_{k,{\rm AD}}^{n+1,(j)}\right)\nonumber\\
+&& - \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla X_{k,{\rm AD}}^{n+1,(j)}\right)\nonumber\\
 && - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,(j-1)}k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla p_0\right),\label{eq:sdc enthalpy 3}
@@
 \begin{eqnarray}
-\left(\rho^{n+1} - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla\right)h^{n+1,(j)}_{\rm AD} &=& (\rho h)^n + \Delta t\Bigg[A_{\rho h}^{(j)} + I_{\rho h}^{(j-1)} \nonumber\\
-&&+ \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\
-&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla X_{k,{\rm AD}}^{n+1,(j)}\right)\nonumber\\
-&&- \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,(j-1)}k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla p_0\right)\Bigg].
+\left(\rho^{n+1} - \Delta t\nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla\right)h^{n+1,(j)}_{\rm AD} &=& (\rho h)^n + \Delta t\Bigg[A_{\rho h}^{(j)} + I_{\rho h}^{(j-1)} \nonumber\\
+&& + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\
+&& - \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla X_{k,{\rm AD}}^{n+1,(j)}\right)\nonumber\\
+&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,(j-1)}k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla p_0\right)\Bigg].
 \end{eqnarray}

### 2012-07-17 Source_SDC/advance.f90

@@ -655,8 +655,12 @@ contains
     end if
 
     do n=1,nlevs
-       call multifab_build(sdc_source(n), mla%la(n), nscal, 1)
-       call multifab_copy_c(sdc_source(n), 1, aofs(n), 1, nscal, 1)
+       call multifab_build(sdc_source(n), mla%la(n), nscal, 0)
+       call multifab_setval(sdc_source(n), 0.d0)
+       call multifab_plus_plus_c(sdc_source(n), rhoh_comp, diff_old(n), 1, 0)
+       call multifab_plus_plus_c(sdc_source(n), rhoh_comp, diff_hat(n), 1, 0)
+       call multifab_mult_mult_s(sdc_source(n), 0.5d0, 0)
+       call multifab_copy_c(sdc_source(n), 1, aofs(n), 1, nscal, 0)
     end do
 
     call react_state(mla,tempbar_init,sold,snew,rho_Hext,p0_new, &
@@ -664,10 +668,11 @@ contains
 
     ! extract IR = [ (snew - sold)/dt - sdc_source ] 
     do n=1,nlevs
-       call multifab_copy_c       (intra(n), 1, snew(n), 1,       nscal, 1)
-       call multifab_sub_sub_c    (intra(n), 1, sold(n), 1,       nscal, 1)
-       call multifab_div_div_s_c  (intra(n), 1, dt,               nscal, 1)
-       call multifab_sub_sub_c    (intra(n), 1, sdc_source(n), 1, nscal, 1)
+       call multifab_setval       (intra(n), 0.d0, all=.true.)
+       call multifab_copy_c       (intra(n), 1, snew(n), 1,       nscal, 0)
+       call multifab_sub_sub_c    (intra(n), 1, sold(n), 1,       nscal, 0)
+       call multifab_div_div_s_c  (intra(n), 1, dt,               nscal, 0)
+       call multifab_sub_sub_c    (intra(n), 1, sdc_source(n), 1, nscal, 0)
     end do
 
     ! massage the rhoh intra term into the proper form, depending on

### 2012-07-17 Source_SDC/thermal_conduct.f90

@@ -116,9 +116,9 @@ contains
     call mac_applyop(mla,Lphi,phi,alpha,beta,dx,the_bc_tower, &
                      dm+rhoh_comp,stencil_order,mla%mba%rr)
 
-    ! subtract Lphi from rhs (since applyop returns negative operator)
+    ! add Lphi to rhs
     do n=1,nlevs
-       call multifab_sub_sub_c(rhs(n),1,Lphi(n),1,0)
+       call multifab_plus_plus_c(rhs(n),1,Lphi(n),1,0)
     enddo
 
     !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
@@ -140,9 +140,9 @@ contains
        call mac_applyop(mla,Lphi,phi,alpha,beta,dx,the_bc_tower, &
                         dm+spec_comp+comp-1,stencil_order,mla%mba%rr)
 
-       ! subtract Lphi from rhs (since applyop returns negative operator)
+       ! add Lphi to rhs
        do n=1,nlevs
-          call multifab_sub_sub_c(rhs(n),1,Lphi(n),1,0)
+          call multifab_plus_plus_c(rhs(n),1,Lphi(n),1,0)
        enddo
 
     end do
@@ -166,9 +166,9 @@ contains
        call mac_applyop(mla,Lphi,phi,alpha,beta,dx,the_bc_tower, &
                         dm+spec_comp+comp-1,stencil_order,mla%mba%rr)
 
-       ! subtract Lphi from rhs (since applyop returns negative operator)
+       ! add Lphi to rhs
        do n=1,nlevs
-          call multifab_sub_sub_c(rhs(n),1,Lphi(n),1,0)
+          call multifab_plus_plus_c(rhs(n),1,Lphi(n),1,0)
        enddo
 
     end do
@@ -188,7 +188,7 @@ contains
     call mac_applyop(mla,Lphi,phi,alpha,beta,dx,the_bc_tower, &
                      foextrap_comp,stencil_order,mla%mba%rr)
 
-    ! add Lphi to rhs (since applyop returns negative operator)
+    ! add Lphi to rhs
     do n=1,nlevs
        call multifab_plus_plus_c(rhs(n),1,Lphi(n),1,0)
     enddo    
@@ -208,7 +208,7 @@ contains
     call mac_applyop(mla,Lphi,phi,alpha,beta,dx,the_bc_tower, &
[…]

### 2012-07-18 docs/FAQ/faq.tex
remove references to fParallel
fixed build error
add some bits about debugging and the initial projections

@@
 
+\item {\em Why do the initial projection and ``divu'' iters sometimes
+  have a harder time converging than the multigrid solves in the main algorithm?}
+
+  The initial projection and ``divu'' solve sets the density to $1$
+  (see \S~\ref{sec:flow:initialization}), so the coefficients in the
+  elliptic solve are $O(\beta_0) \sim O(\rho)$.  But in the main
+  algorithm, the coefficients are $O(\beta_0/\rho) \sim O(1)$.  Since
+  $\rho$ can vary a lot, the variation in the coefficients in the
+  initial projection and ``divu'' solve present a harded linear system
+  to solve.
+
 \item {\em How can I obtain profiling infomation for my run?}
@@
 
-\section{I/O}
+\section{Debugging}
 
@@
 \end{verbatim}
+
+\item {\em How can I print out a multifab's contents from within the code?}
+
+  There is a {\tt print} method in {\tt multifab\_module}.  This can
+  be simply called as
+  \begin{verbatim}
+  call print(a)
+  \end{verbatim}
+  where {\tt a} is a multifab (single-level).
+
+\end{enumerate}
+
+\section{I/O}
+
+\begin{enumerate}
 
@@
  One implementation of this can be found in {\tt
-   fParallel/data\_processing/tutorial/fwrite2d.f90}.  This reads in
+   AmrPostprocessing/F\_Src/tutorial/fwrite2d.f90}.  This reads in
  the plotfile data using the {\tt plotfile\_module} that the {\tt

### 2012-07-18 docs/flowchart/flowchart.tex
more notes about initialization
add a section label

@@
 
+\label{sec:flow:initialization}
+
 We start each calculation with user-specified initial values for
@@
 First, we need to construct approximations to $S^0, w_0^{-\myhalf}, \Delta t^0$, 
-and $\Ub^0$.  Start with initial data $X_k^{\initp}, \rho^{\initp},$ $T^{\initp},$ an 
-initial base state, and an initial guess for the velocity, $\Ub^{\initp}$.
+and $\Ub^0$.  Start with initial data $X_k^{\initp}, \rho^{\initp},$ $T^{\initp},$ 
+an initial base state, and an initial guess for the velocity, $\Ub^{\initp}$.
 Set $w_0^0 = 0$ as an initial approximation.  Use the equation of state to 
 determine $(\rho h)^{\initp}$.  Compute $\beta_0^0$ as a function of 
-the initial data.  Then, project $\Ub^{\initp}$ using $\beta_0^0$ and 
-$S = \rho\Hext$, giving $\Ub^{0,0}$.  The next part of the initialization process 
+the initial data.  The next part of the initialization process 
 proceeds as follows.
@@
 \item {\em Initial Projection}: if {\tt do\_initial\_projection = T}, then we 
-   first project the velocity field with $\rho = 1$ and $\beta_0^0$ defined 
-   from the initial data.  No sources are included in $S$.  The elliptic
-   equation we solve is 
+   first project the velocity field with $\rho = 1$ and $\beta_0^0$.
+   The initial projection does not see reactions
+   or external heating, and thus we set $\dot\omega = \Hnuc = \Hext = 0$ in $S$.
+   The reason for ignoring reactions and heating is that we need some kind of 
+   time scale over which to compute the effect of reactions, but we first 
+   need an estimate of the velocity
+   field in order to derive the time step that will be used as a time scale.
+   The elliptic equation we solve is 
    \begin{equation}
-   \nabla \cdot \beta_0^0 \nabla \phi = - \nabla \cdot (\beta_0^0 \Ub^0)
+   \nabla \cdot \beta_0^0 \nabla \phi = \beta_0^0(S - \Sbar)- \nabla \cdot (\beta_0^0 \Ub^{\initp})
    \end{equation}
-   This $\phi$ is then used to correct the velocity field.
+   This $\phi$ is then used to correct the velocity field to obtain $\Ub^{0,0}$.
+   If {\tt do\_initial\_projection = F}, set $\Ub^{0,0} = \Ub^{\initp}$.
 
 \item {\em ``Divu'' iterations}: Next we do {\tt init\_divu\_iter} iterations 
-  to project the velocity field using a constraint that sees the reactions. 
-  The initial timestep estimate is provided by {\tt firstdt}, to allow us to 
-  integrate the reaction networks.  In this projection, again the density is 
-  set to 1, and the elliptic equation we solve is:
-   \begin{equation}
-   \nabla \cdot \beta_0^0 \nabla \phi = \beta_0 (S - \Sbar)- \nabla \cdot (\beta_0^0 \Ub^0)
-   \end{equation}
+  to project the velocity field using a constraint that sees reactions
+  and external heating. 
+  The initial timestep estimate is provided by {\tt firstdt} and
+  $\Ub^{0,0}$, to allow us to compute the effect of reactions over $\Delta t/2$.
 
@@
         $(S^{0,\nu} - \overline{S^{0,\nu}})$ as the source term.  
-        This yields $\Ub^{0,\nu}.$
+        This yields $\Ub^{0,\nu}.$  In this projection, again the density is 
+        set to 1, and the elliptic equation we solve is:
+
+   \begin{equation}
+   \nabla \cdot \beta_0^0 \nabla \phi = \beta_0 (S - \Sbar)- \nabla \cdot (\beta_0^0 \Ub^{0,\nu-1})
+   \end{equation}
 
@@
   \end{enumerate}
+
+The tolerances for these elliptic solves are described in \S~\ref{sec:mgtol}.
 

### 2012-07-20 docs/SDC/sdc.tex

--- 2012-07-17 docs/SDC/sdc.tex
@@
 \begin{equation}
-\frac{\rho^{n+1}X_{k,{\rm AD}}^{n+1,(0)} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{(0)} + I_{\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
+\frac{\rho^{n+1}\widehat{X}_k^{n+1,(0)} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{(0)} + I_{\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
 \end{equation}
@@
 {\bf Step 2B:} ({\it Compute diffusive flux divergence})\\ \\
-Solve a Crank-Nicolson-type diffusion equation for $h^{n+1,(0)}_{\rm AD}$, using transport coefficients evaluated at $t^n$ everywhere,
+Solve a Crank-Nicolson-type diffusion equation for $\widehat{h}^{n+1,(0)}$, using transport coefficients evaluated at $t^n$ everywhere,
 \begin{eqnarray}
-\frac{\rho^{n+1}h^{n+1,(0)}_{\rm AD} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(0)} + I_{\rho h}^{\rm lagged}\nonumber\\
-&& + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n + \nabla\cdot\frac{\kth^n}{c_p^n}\nabla h_{\rm AD}^{n+1,(0)}\right)\nonumber\\
-&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,(0)}\right)\nonumber\\
+\frac{\rho^{n+1}\widehat{h}^{n+1,(0)} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(0)} + I_{\rho h}^{\rm lagged}\nonumber\\
+&& + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n + \nabla\cdot\frac{\kth^n}{c_p^n}\nabla \widehat{h}^{n+1,(0)}\right)\nonumber\\
+&& - \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla\widehat{X}_k^{n+1,(0)}\right)\nonumber\\
 && - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right),\label{eq:sdc enthalpy 2}
@@
 \begin{eqnarray}
-\left(\rho^{n+1} - \frac{\Delta t}{2}\nabla\cdot\frac{k_{\rm th}^n}{c_p^n}\nabla\right)h_{\rm AD}^{n+1,(0)} &=& (\rho h)^n + \Delta t\Bigg[A_{\rho h}^{(0)} + I_{\rho h}^{\rm lagged} + \left(\half\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n\right)\nonumber\\
-&&\hspace{0.85in} + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_{k,{\rm AD}}^{n+1,(0)}\right)\nonumber\\
+\left(\rho^{n+1} - \frac{\Delta t}{2}\nabla\cdot\frac{k_{\rm th}^n}{c_p^n}\nabla\right)\widehat{h}^{n+1,(0)} &=& (\rho h)^n + \Delta t\Bigg[A_{\rho h}^{(0)} + I_{\rho h}^{\rm lagged} + \left(\half\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n\right)\nonumber\\
+&&\hspace{0.85in} - \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla\widehat{X}_k^{n+1,(0)}\right)\nonumber\\
 &&\hspace{0.85in} - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right)\Bigg].
@@
 \begin{equation}
-\frac{\rho^{n+1}X_{k,{\rm AD}}^{n+1,(j)} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{(j)} + I_{\rho X_k}^{(j-1)}.\label{eq:sdc species 3}
+\frac{\rho^{n+1}\widehat{X}_k^{n+1,(j)} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{(j)} + I_{\rho X_k}^{(j-1)}.\label{eq:sdc species 3}
 \end{equation}
@@
 {\bf Step 4B:} ({\it Compute diffusive flux divergence})\\ \\
-Solve a backward-Euler-type correction equation for $h^{n+1,(j)}_{\rm AD}$,
+Solve a backward-Euler-type correction equation for $\widehat{h}^{n+1,(j)}$,
 \begin{eqnarray}
-\frac{\rho^{n+1}h^{n+1,(j)}_{\rm AD} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(j)} + I_{\rho h}^{(j-1)}\nonumber\\
-&& + \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j)}_{\rm AD} + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\
-&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla X_{k,{\rm AD}}^{n+1,(j)}\right)\nonumber\\
+\frac{\rho^{n+1}\widehat{h}^{n+1,(j)} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(j)} + I_{\rho h}^{(j-1)}\nonumber\\
+&& + \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla\widehat{h}^{n+1,(j)} + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\
+&& - \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla\widehat{X}_k^{n+1,(j)}\right)\nonumber\\
 && - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,(j-1)}k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla p_0\right),\label{eq:sdc enthalpy 3}
@@
 \begin{eqnarray}
-\left(\rho^{n+1} - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla\right)h^{n+1,(j)}_{\rm AD} &=& (\rho h)^n + \Delta t\Bigg[A_{\rho h}^{(j)} + I_{\rho h}^{(j-1)} \nonumber\\
-&&+ \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\
-&& + \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla X_{k,{\rm AD}}^{n+1,(j)}\right)\nonumber\\
-&&- \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,(j-1)}k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla p_0\right)\Bigg].
+\left(\rho^{n+1} - \Delta t\nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla\right)\widehat{h}^{n+1,(j)} &=& (\rho h)^n + \Delta t\Bigg[A_{\rho h}^{(j)} + I_{\rho h}^{(j-1)} \nonumber\\
+&& + \half\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\
+&& - \half\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla\widehat{X}_k^{n+1,(j)}\right)\nonumber\\
+&& - \half\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,(j-1)}k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla p_0\right)\Bigg].
 \end{eqnarray}

### 2012-07-23 docs/mg/mg.tex
intro stub
note about MG
add HG centeringsadd HG centeringsadd HG centeringsadd HG centeringsadd HG centeringsadd HG centeringsadd HG centeringsadd HG centeringsadd HG centerings

@@
+\section{Introduction to \maestro\ Multigrid}
+
+\maestro\ uses multigrid to enforce the velocity constraint through
+projections at the half-time (the MAC projection) and end of the time
+step.  Two multigrid solvers are provided by \boxlib---one for
+cell-centered data and one for node-centered (nodal) data.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=5.5in]{MAC_mg2}
+\caption{\label{fig:mg:MAC} Data centerings for the MAC projection}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=5.5in]{HG_mg2}
+\caption{\label{fig:mg:HG} Data centerings for the HG projection}
+\end{figure}
+
+\section{\mgtower}
+
+The \mgtower\ is a special Fortran derived type that carries all the
+information required for a \boxlib\ multigrid solve.
+
+\section{Bottom Solvers}
+
+\section{Convergence Criteria}
+
 \section{Multigrid Solver Tolerances}
+
+\label{sec:mgtol}
 

### 2012-07-24 docs/mg/mg.tex
define mgfigpath
add notes about the different MG solves

@@
 projections at the half-time (the MAC projection) and end of the time
-step.  Two multigrid solvers are provided by \boxlib---one for
-cell-centered data and one for node-centered (nodal) data.
+step (the HG projection).  Two multigrid solvers are provided by
+\boxlib---one for cell-centered data and one for node-centered (nodal)
+data.  Both of these are used in \maestro.
 
@@
 \centering
-\includegraphics[width=5.5in]{MAC_mg2}
+\includegraphics[width=5.5in]{\mgfigpath/MAC_mg2}
 \caption{\label{fig:mg:MAC} Data centerings for the MAC projection}
@@
 
+The MAC projection operates on the advective velocities predicted at
+the cell-interfaces at the half-time.  The edge-centered velocities
+are shown in Figure~\ref{fig:mg:MAC}.  If we consider purely
+incompressible flow, the projection appears as:
+\begin{equation}
+D G \phi = D U
+\end{equation}
+where $D$ is the divergence operator and $G$ is the gradient operator.
+In this discretization, $\phi$ is cell-centered (see
+Figure~\ref{fig:mg:MAC}).  The remaining quantities are discretized as:
+\begin{itemize}
+\item $DU$ is cell-centered, 
+  \begin{equation}
+  (DU)_{i,j} = \frac{u_{i+1/2,j} - u_{i-1/2,j}}{\Delta x} + 
+               \frac{v_{i,j+1/2} - v_{i,j-1/2}}{\Delta y}
+  \end{equation}
+
+\item $G\phi$ is edge-centered, on the MAC grid, as shown in
+  Figure~\ref{fig:mg:MAC}.
+
+\item $DG\phi$ is cell-centered, also shown in Figure~\ref{fig:mg:MAC},
+  computed from $G\phi$ using the same differencing as $DU$.
+
+\end{itemize}
+
 \begin{figure}[t]
 \centering
-\includegraphics[width=5.5in]{HG_mg2}
+\includegraphics[width=5.5in]{\mgfigpath/HG_mg2}
 \caption{\label{fig:mg:HG} Data centerings for the HG projection}
 \end{figure}
+
+The HG projection projects the cell-centered velocities at the end of
+the timestep.  Here, $\phi$ is node-centered.  Figure~\ref{fig:mg:HG}
+shows the locations of the various quantities involved in the HG
+projection.  Again considering simple incompressible flow, we now
+solve:
+\begin{equation}
+L \phi = D U
+\end{equation}
+where $L$ is a discretization of the Laplacian operator.  In this
+sense, the HG projection is an {\em approximate projection}, that is,
+$L \neq DG$ (in discretized form).  The various operations have the
+following centerings:
+\begin{itemize}
+
+\item $DU$ is node-centered.  This is computed as:
+  \begin{equation}
+  (DU)_{i-1/2,j-1/2} = \frac{\frac{1}{2} (u_{i,j} + u_{i,j-1}) -
+                             \frac{1}{2} (u_{i-1,j} + u_{i-1,j-1})}{\Delta x} +
+                       \frac{\frac{1}{2} (v_{i,j} + v_{i-1,j}) -
+                             \frac{1}{2} (v_{i,j-1} + v_{i-1,j-1})}{\Delta y} 
+  \end{equation}
+
+\item $G\phi$ is cell-centered, as shown in Figure~\ref{fig:mg:HG}.
+
+\item $L\phi$ is node-centered.  This is a direct discretization of 
+the Laplacian operator.  By default, \maestro\ uses a dense stencil
+(9-points in 2-d, 27-points in 3-d).  Alternately, a {\em cross}
+stencil can be used (by setting {\tt hg\_dense\_stencil = F}).  This
+uses 5-points in 2-d, 7-points in 3-d.
+
+\end{itemize}
 

### 2012-07-24 docs/mg/mg.tex
information about the mg_tower

@@
 
+The following parameters are specified when building the \mgtower:
+\begin{itemize}
+\item {\tt smoother}: the type of smoother to use.  Choices are listed
+  in {\tt mg\_tower.f90}.  Common options are {\tt
+    MG\_SMOOTHER\_GS\_RB} for red-black Gauss-Seidel and {\tt
+    MG\_SMOOTHER\_JACOBI} for Jacobi.
+
+\item {\tt nu1}: The number of smoothings at each level on the way down
+  the V-cycle.
+
+\item {\tt nu2}: The number of smoothings at each level on the way up
+  the V-cycle.
+
+\item {\tt nub}: The number of smoothing before and after the bottom solver.
+
+\item {\tt gamma}: \MarginPar{what does this mean?}
+
+\item {\tt cycle\_type}: The type of multigrid to do, V-cycles ({\tt
+  MG\_VCycle}), W-cycles ({\tt MG\_WCycle}), full multigrid ({\tt
+  MG\_FCycle}).
+
+\item {\tt omega}:  \MarginPar{description?}
+
+\item {\tt bottom\_solver}: the type of bottom solver to use.  See the next
+ section.
+
+\item {\tt bottom\_max\_iter}: the maximum number of iterations for the 
+  bottom solver
+
+\item {\tt bottom\_solver\_eps}: the tolerance used by the bottom
+  solver.  In \maestro, this is set via {\tt mg\_eps\_module}.
+
+\item {\tt max\_iter}: the maximum number of multigrid cycles.
+
+\item {\tt max\_bottom\_nlevel}: additional coarsening if you use
+  {\tt bottom\_solver} type 4 (see below)
+
+\item {\tt min\_width}: minimum size of grid at coarsest multigrid level
+
+\item {\tt rel\_solver\_eps}: the relative tolerance of the solver (in
+  \maestro, this is set via {\tt mg\_eps\_module}.
+
+\item {\tt abs\_solver\_eps}: the absolute tolerance of the solver (in
+  \maestro, this is set via {\tt mg\_eps\_module}.
+
+\item {\tt verbose}: the verbosity of the multigrid solver.  In \maestro,
+  this is set via the {\tt mg\_verbose} runtime parameter.  Higher
+  numbers give more verbosity.
+
+\item {\tt cg\_verbose}: the verbosity of the bottom solver.  In \maestro,
+  this is set via the {\tt cg\_verbose} runtime parameter.  Higher
+  numbers give more verbosity.
+
+\end{itemize}
+
 \section{Bottom Solvers}
@@
 \section{Convergence Criteria}
+
+All \maestro\ multigrid solves consist of pure V-cycles.
 
@@
 \end{itemize}
+
+\section{General Remarks}
+
+If \maestro\ has trouble converging in the multigrid solves, try
+setting the verbosity {\tt mg\_verbose} or {\tt cg\_verbose} to
+higher values to get more information about the solve.

### 2012-07-24 docs/mg/mg.tex
add discussion of the bottom solvers

@@
 \section{Bottom Solvers}
+
+There are several bottom solvers available to \maestro.  These are set
+through the {\tt mg\_bottom\_solver} (MAC) and {\tt
+  hg\_bottom\_solver} (HG) runtime parameters.  The allowed values
+  are:
+
+\begin{itemize}
+
+\item {\tt mg\_bottom\_solver} / {hg\_bottom\_solver} = 0: smoothing only.
+
+\item {\tt mg\_bottom\_solver} / {hg\_bottom\_solver} = 1: biconjugate
+  gradient stabilized---this is the default.
+
+\item {\tt mg\_bottom\_solver} / {hg\_bottom\_solver} = 2: conjugate
+  gradient method
+
+\item {\tt mg\_bottom\_solver} / {hg\_bottom\_solver} = 4: a special 
+  bottom solver that extends the range of the multigrid coarsening
+  by aggregrating coarse grids on the original mesh together and
+  further coarsening. \MarginPar{does this option then use bicg 
+  at the new bottom?}
+
+\end{itemize}
+
+\MarginPar{any simple discussion on why we might choose one of the
+  other bottom solvers?}
 

### 2012-07-24 docs/architecture/architecture.tex
move special bottom solver section to mg/

@@
 
-\maestro\ uses the multigrid solver to enforce the divergence constraint
-both on the half-time edge-centered advective velocities (the ``MAC
-projection'') and on the final cell-centered velocities (the ``HG
-projection'').  For the MAC projection, since the velocity data is
-edge-centered (the MAC grid), the projection is cell-centered.  For
-the HG projection, since the velocity data is cell-centered, the
-projection is node-centered.  \MarginPar{should draw a figure} The multigrid solver performs a number
-of V-cycles until the residual drops by 10-12 orders-of-magnitude.
-There are several options that affect how the multigrid solver
-behaves, which we describe below.
-
-\subsection{Special Bottom Solver}
-We have a special bottom solver type that greatly speeds up large problems
-and speeds up smaller problems to a lesser degree.
-In fact, you should use this bottom solver whenever possible, because I
-have yet to find a case that performs worse than our standard bottom solver.
-What this special solver essentially does is take the data from the coarsest 
-level of the original multigrid V-cycle and copy it onto a new grid structure 
-with the same number of total cells in each direction, but with a fewer number of larger 
-grids.  A new V-cycle begins from this point, 
-so we are essentially coarsening this ``new'' problem.
-Now, the coarsest level of the multigrid V-cycle in the ``new'' problem has 
-fewer cells and fewer grids as compared to the original coarsest level.
-
-To enable this solver, set {\tt hg\_bottom\_solver = 4} (for the nodal
-projections) and/or {\tt mg\_bottom\_solver = 4} (for the cell-centered
-projections) in your inputs file.
-
-To understand how this bottom solver works, the first thing you need to know
-is what the grid structure of the coarsest level of your multigrid V-cycle
-looks like (we need to put a description of this in the section above).
-Now, figure out the size of the box you would need if you
-wanted it to fit all the data on the coarsest level.  Now, figure out what
-the largest integer $n$ is so that you can evenly divide the length of this box
-by $2^n$ in every coordinate direction.  If $n < 2$, the program will abort
-since the grid structure is not suitable for this bottom solver.
-
-Now the code will set up a ``new'' problem, using the data at the coarsest level
-of the original problem as the initial data.  The grid structure for this new
-problem has the same number of cells as the coarsest level of the original problem,
-but the data is copied onto a grid structure where each grid has $2^n$ cells
-on each side.  The new V-cycle continues down to the new coarsest level, in
-which each grid has 2 cells on each side.  If you wish to impose a limit on
-the maximum value that $n$ can have, you can do so by setting 
-{\tt max\_mg\_bottom\_nlevs} equal to that value.
-
-{\bf Example 1:} A 3D problem with $384^3$ cells divided into $32^3$ grids, i.e., 
-there is a $12\times 12\times 12$ block of $32^3$ grids.  The coarsest level of the 
-multigrid V-cycle contains $12\times 12\times 12$ grids that have $2^3$ cells, so the 
-entire problem domain has $24^3$ cells.  We see that $n=3$, and create a new problem 
-domain with a $3\times 3\times 3$ block of $8^3$ grids.  The coarsest level of the 
-multigrid V-cycle for the ``new'' problem will be a $3\times 3\times 3$ block of 
-$2^3$ grids.
-
-{\bf Example 2:} A 2D problem with $96\times 384$ cells divided into $48^2$ grids, i.e., 
-there is a $2\times 8$ block of $48^2$ grids.  The coarsest level of the multigrid 
-V-cycle contains $2\times 8$ grids that have $3^2$ cells, so the entire problem 
-domain has $6\times 24$ cells.  We see that $n=1$, so the program aborts since this grid
-structure is not appropriate for the fancy bottom solver.
+\maestro\ uses the multigrid solver to enforce the divergence
+constraint both on the half-time edge-centered advective velocities
+(the ``MAC projection'') and on the final cell-centered velocities
+(the ``HG projection'').  For the MAC projection, since the velocity
+data is edge-centered (the MAC grid), the projection is cell-centered.
+For the HG projection, since the velocity data is cell-centered, the
+projection is node-centered.  \MarginPar{should draw a figure} The
+multigrid solver performs a number of V-cycles until the residual
+drops by 10-12 orders-of-magnitude.  There are several options that
+affect how the multigrid solver behaves, which we describe below.
+More detail on the multigrid solvers is given in Chapter~\ref{ch:mg}.
 

### 2012-07-24 docs/mg/mg.tex
move special bottom solver section to mg/

@@
+\label{ch:mg}
+
 \section{Introduction to \maestro\ Multigrid}
@@
   other bottom solvers?}
+
+\subsection{Special Bottom Solver}
+We have a special bottom solver type that greatly speeds up large problems
+and speeds up smaller problems to a lesser degree.
+In fact, you should use this bottom solver whenever possible, because I
+have yet to find a case that performs worse than our standard bottom solver.
+What this special solver essentially does is take the data from the coarsest 
+level of the original multigrid V-cycle and copy it onto a new grid structure 
+with the same number of total cells in each direction, but with a fewer number of larger 
+grids.  A new V-cycle begins from this point, 
+so we are essentially coarsening this ``new'' problem.
+Now, the coarsest level of the multigrid V-cycle in the ``new'' problem has 
+fewer cells and fewer grids as compared to the original coarsest level.
+
+To enable this solver, set {\tt hg\_bottom\_solver = 4} (for the nodal
+projections) and/or {\tt mg\_bottom\_solver = 4} (for the cell-centered
+projections) in your inputs file.
+
+To understand how this bottom solver works, the first thing you need to know
+is what the grid structure of the coarsest level of your multigrid V-cycle
+looks like (we need to put a description of this in the section above).
+Now, figure out the size of the box you would need if you
+wanted it to fit all the data on the coarsest level.  Now, figure out what
+the largest integer $n$ is so that you can evenly divide the length of this box
+by $2^n$ in every coordinate direction.  If $n < 2$, the program will abort
+since the grid structure is not suitable for this bottom solver.
+
+Now the code will set up a ``new'' problem, using the data at the coarsest level
+of the original problem as the initial data.  The grid structure for this new
+problem has the same number of cells as the coarsest level of the original problem,
+but the data is copied onto a grid structure where each grid has $2^n$ cells
+on each side.  The new V-cycle continues down to the new coarsest level, in
+which each grid has 2 cells on each side.  If you wish to impose a limit on
+the maximum value that $n$ can have, you can do so by setting 
+{\tt max\_mg\_bottom\_nlevs} equal to that value.
+
+{\bf Example 1:} A 3D problem with $384^3$ cells divided into $32^3$ grids, i.e., 
+there is a $12\times 12\times 12$ block of $32^3$ grids.  The coarsest level of the 
+multigrid V-cycle contains $12\times 12\times 12$ grids that have $2^3$ cells, so the 
+entire problem domain has $24^3$ cells.  We see that $n=3$, and create a new problem 
+domain with a $3\times 3\times 3$ block of $8^3$ grids.  The coarsest level of the 
+multigrid V-cycle for the ``new'' problem will be a $3\times 3\times 3$ block of 
+$2^3$ grids.
+
+{\bf Example 2:} A 2D problem with $96\times 384$ cells divided into $48^2$ grids, i.e., 
+there is a $2\times 8$ block of $48^2$ grids.  The coarsest level of the multigrid 
+V-cycle contains $2\times 8$ grids that have $3^2$ cells, so the entire problem 
+domain has $6\times 24$ cells.  We see that $n=1$, so the program aborts since this grid
+structure is not appropriate for the fancy bottom solver.
 

### 2012-08-02 docs/MESA/hopper_compile.tex
MESA instructions

Once {\tt MESA} was installed and working properly on a local machine (e.g. 
Bender), I tried installing it on Hopper, which is a Cray XE6 machine. On 
Bender, which is an Intel(R) Xeon(R) X5650 machine, the compiler was 
{\tt gcc} version 4.7.0, on Hopper I tried compiling {\tt MESA} with the 
Cray Compiler using version 8.0.5. The module that was loaded to enable this 
was {\tt PrgEnv-cray}.

There are several steps needed to compile {\tt MAESTRO} with {\tt MESA} on 
Hopper:
\begin{enumerate}
\item Edit the {\tt makefile\_header} in the {\tt utils} directory
\item Edit the {\tt build\_and\_test} script in the {\tt utils} directory
\item Edit the {\tt mtx} makefile
\item Edit the makefiles in {\tt screen}, {\tt utils}, {\tt num} and 
{\tt interp\_1d}
\item Edit the {\tt install} script in the main {\tt MESA} directory
\item Edit the {\tt BoxLib/Tools/F\_mk/GMakedefs.mak} makefile
\item Edit the {\tt BoxLib/Tools/F\_mk/comps/Linux\_cray.mak} file
\item Make sure the makefile {\tt COMP} variable is set to {\tt Cray} in the 
problem directory
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\tt makefile\_header}}
In following the setup outlined in this document, I changed the compilers, 
used the source version of {\tt LAPACK} and {\tt BLAS}, turned off 
{\tt PGPLOT} and turned off the {\tt SE} formatting library. I also added the 
appropriate compiler flags. The results of what I changed are shown below:
\begin{lstlisting}[language=make,mathescape=false]
  # Cray C-compiler
  CC = cc

  # Cray Fortran-Compiler
  FC = ftn

  # if you need special flags for the compiler, define them here:
  SPECIAL_FC_FLAGS = -h mpi0 -target=linux -emf
  # -h mpi0 disables mpi optimization
  # -target=linux specifies the build machine as linux based
  # -emf: the m flag says create NAME.mod files, the f flag says
  #        convert those to name.mod files

  # must explicity define all compilation flags
  FCbasic = $(SPECIAL_FC_FLAGS)
  FCimpno = -eI
  FCchecks =
  FCwarn = -m 3
  FCfixed = -f fixed -N 132
  FCfixed72 = -f fixed -N 72
  FCfree = -f free
  FCopt = -O 1
  FCdebug = -g -O0
  FCstatic =

  FC_fixed_preprocess = -eZ
  FC_free_preprocess = -eZ

  # Cray compilers have omp ON by default
  FCopenmp =
  # to disable omp:
  #FCopenmp = -h noomp 
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\tt build\_and\_test} script}

The {\tt build\_and\_test} script in the {\tt utils} directory is 
responsible for compiling the main module directory as well as compiling 
and running the test directory. We want to turn off the compiling and 
running of the test directory to ensure that {\tt MESA} compiles successfully. 
A portion of the {\tt build\_and\_test} script looks like this:
\begin{lstlisting}[language=bash,mathescape=false]
  cd make
  make
  check_okay
  cd ../test
  ./mk
  check_okay
  if [ -x test_quietly ]
  then
          ./test_quietly
          check_okay
  fi
  ./ck >& diff.txt
\end{lstlisting}
we wish to change it to this:
\begin{lstlisting}[language=bash,mathescape=false]
  cd make
  make
  check_okay
  #cd ../test
  #./mk
  check_okay
  if [ -x test_quietly ]
  then
  #        ./test_quietly
          check_okay
  fi
  #./ck >& diff.txt
\end{lstlisting}
This disables the compiling and running of the test directory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\tt mtx} makefile}

In the {\tt mtx/make/makefile}, there are a few compiler flags that were hard 
coded in. There is a {\tt -w} flag in five places and a {\tt -fno-common} 
flag in one place. The {\tt -w} flag instances look like this:
\begin{lstlisting}[language=make,mathescape=false]
  %.o: $(MODULE_DIR)/lapack_src/%.f
  	$(COMPILE_XTRA) -w $<

  # must turn off optimization for dlamch or can get infinite loop!!!
  dlamch.o: $(MODULE_DIR)/blas_src/dlamch.f
  	$(COMPILE_XTRA_NO_OPT) -w $<
\end{lstlisting}
The {\tt -w} should be deleted. There are three more instances immediately 
after these lines. The {\tt -fno-common} flag appears here:
\begin{lstlisting}[language=make,mathescape=false]
  KLU_C = $(CC) -O3 -fno-common -fexceptions
  KLU_I = -I$(KLU_DIR)
\end{lstlisting}
All three of the manually added flags ({\tt -O3}, {\tt -fno-common} and 
{\tt -fexceptions}) should be deleted from the {\tt KLU\_C} line.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{makefiles in {\tt screen}, {\tt num}, {\tt utils} and 
{\tt interp\_1d}}

When creating the library for these four modules, there is a {\tt LIB\_DEFS} 
variable in the makefile that holds the {\tt *\_def.o} file, e.g. 
{\tt utils\_def.o} or {\tt num\_def.o}. This files is never loaded into the 
library as seen below for the {\tt utils} directory:
\begin{lstlisting}[language=make,mathescape=false]
  LIB = libutils.a
  LIB_DEFS = utils_def.o
  LIB_OBJS = $(UTILS_ISNAN).o utils_nan.o utils_dict.o utils_lib.o

  $(LIB) : $(LIB_DEFS) $(LIB_OBJS)
  	$(LIB_TOOL) $(LIB) $(LIB_OBJS)
\end{lstlisting}
To include the {\tt LIB\_DEFS} file:
\begin{lstlisting}[language=make,mathescape=false]
  LIB = libutils.a
  LIB_DEFS = utils_def.o
  LIB_OBJS = $(UTILS_ISNAN).o utils_nan.o utils_dict.o utils_lib.o

  $(LIB) : $(LIB_DEFS) $(LIB_OBJS)
  	$(LIB_TOOL) $(LIB) $(LIB_DEFS) $(LIB_OBJS)
\end{lstlisting}
This must be done for each makefile in the following directories: 
{\tt screen/make}, {\tt num/make}, {\tt utils/make} and {\tt interp\_1d/make}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\tt install} script}

Certain modules that are unneccessary for use with {\tt MAESTRO} do not need 
to be installed (and also cause some sort of installation error). In the 
main {\tt MESA} directory, edit the {\tt install} script and comment out 
the following four lines: {\tt do\_one sample}, {\tt do\_one star}, 
{\tt do\_one adipls} and {\tt do\_one astero}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\tt BoxLib Linux\_cray.mak} file}

Two flags used in the compilation of {\tt MESA} with the Cray compiler are 
important; the {\tt -e m} or {\tt -em} flag and the {\tt -e f} or {\tt -ef} 
flag. The {\tt -em} flag tells the compiler to create {\tt .mod} files of the 
form {\tt NAME.mod}. The {\tt -e f} flag, which must be used with the 
{\tt -em} flag, instead outputs {\tt .mod} files of the form {\tt name.mod}. 
This is important for {\tt MESA} because the {\tt export} scripts copy the 
various libraries and {\tt .mod} files to the correct directories. The 
{\tt export} script in the {\tt const} directory for example looks like this:
\begin{lstlisting}[language=bash,mathescape=false]
  cp make/const_lib.mod ../include
  cp make/const_def.mod ../include

  cp make/libconst.a ../lib
  cd ../lib
  ranlib libconst.a
\end{lstlisting}
If the {\tt -ef} flag is not enabled, the {\tt .mod} file will be 
{\tt CONST\_LIB.mod} and will not get copied to the {\tt include} directory 
(the library is unaffected). The default for {\tt MAESTRO} is to not use 
the {\tt -ef} flag, but it does use the {\tt -em} flag. To make {\tt MAESTRO} 
and {\tt MESA} compatible, I added the {\tt -ef} flag to the 
{\tt Linux\_cray.mak} file located in {\tt BoxLib/Tools/F\_mk/comps}. I also 
added a few more debugging flags:
\begin{lstlisting}[language=make,mathescape=false]
  FFLAGS   += -J $(mdir) -I $(mdir) -emf
  F90FLAGS += -J $(mdir) -I $(mdir) -emf

  ifdef NDEBUG
    FFLAGS   += -O 1
    F90FLAGS += -O 1
    CFLAGS   += -O 1
  else
    FFLAGS   += -g -O0 -R bps
    F90FLAGS += -g -O0 -R bps
    CFLAGS   += -g -O0 -h bounds
  endif 
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\tt BoxLib GMakedefs.mak} makefile}

In {\tt MAESTRO}, {\tt .mod} files are located in the {\tt \$(tdir)/m} 
directory, where {\tt tdir = t/\$(suf)} and {\tt suf} was in this case 
{\tt Linux.Cray}. This means the {\tt MAESTRO} routines are expecting to find 
module files with the name {\tt t/Linux.Cray/m/network.mod}. The addition of 
the {\tt -ef} flag affects the entire name of the module so the Cray compiler 
was creating files with the name {\tt t/linux.cray/m/network.mod}. To solve 
this, I modified what {\tt suf} is defined to be when using the Cray compiler. 
This was done in the {\tt BoxLib/Tools/F\_mk/GMakedefs.mak} makefile by using 
a simple if statement:
\begin{lstlisting}[language=make,mathescape=false]
  ifeq ($(COMP), Cray)
    tdir = t/$(shell echo $(suf) | tr A-Z a-z)
  else
    tdir = t/$(suf)
  endif
  odir = $(tdir)/o
  mdir = $(tdir)/m
\end{lstlisting}
The shell command {\tt echo \$(suf) | tr A-Z a-z} has the effect of converting 
{\tt \$(suf)} to lowercase.

### 2012-08-02 docs/MESA/instructions.tex
MESA instructions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Intro
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Modules for Experiments in Stellar Astrophysics ({\tt MESA}) is a collection 
of open source, robust and efficient modules developed for a wide range of 
applications in computational stellar astrophysics. The modules include 
macroscopic and microscopic physics such as nuclear reaction rates, opacity, 
equation of state, element diffusion data and atmosphere boundary conditions 
in addition to numerical routines. This document describes how to link the 
{\tt MESA} modules to an outside program such as {\tt MAESTRO}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Setup MESA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Setting Up {\tt MESA}}

To install {\tt MESA}, download the source code from the {\tt MESA} website, 
{\tt http://mesa.sourceforge.net}. Check the website for the latest version, 
which is 4219 with the writing of this document. The codes and algorithms 
discussed here were created when the current version of {\tt MESA} was 
version 4088. Between versions 4088 and 4219, there are minor changes that 
will cause some of the included routines to crash. To download the source code 
associated with version 4088, use the following command:
\begin{lstlisting}[language=bash,mathescape=false]
  svn co -r 4088 http://mesa.svn.sourceforge.net/svnroot/mesa/trunk mesa
\end{lstlisting}
This will download the source into a directory called {\tt mesa} located 
in the directory where this command was run. This step might take some time 
($\sim$10 min depending on the internet connection) possibly because the 
final {\tt mesa} directory will take up roughly 2.2 GB.

Rich Townsend has developed a Software Development Kit (SDK) to help install 
{\tt MESA} called {\tt MESASDK}. The SDK will install certain software 
packages that {\tt MESA} uses and it will also change the {\tt PATH} 
environment variable such that these programs become the default. A tarball 
of the SDK software and indepth instructions for its use are available on 
Townsend's website:
\begin{lstlisting}[language=html,mathescape=false]
  http://www.astro.wisc.edu/~townsend/static.php?ref=mesasdk
\end{lstlisting}
The instructions in this document do not use the SDK in order to avoid 
automatically changing the {\tt PATH} variable.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Setup without SDK
%%%%%%%%%%%%%%%%%%
\subsection{Setup {\tt MESA} Without {\tt MESASDK}}
\label{sec:woSDK}

%-------------------------------------------------------------------------
\subsubsection{Environment Variables (NOT using SDK)}
A few environment variables need to be set (commands are shown for the 
Bourne shell):
\begin{enumerate}
\item Set location of {\tt mesa}: 
{\tt export MESA\_DIR=<absolute-path-to-mesa-directory>}

\item Set OpenMP number of threads: {\tt export OMP\_NUM\_THREADS=n}, a suggested value is the number of cores in the computer
\end{enumerate}

%-------------------------------------------------------------------------
\subsubsection{Modify the Makefile (NOT using SDK)}
The main makefile that contains compiler information including flags is 
located in the {\tt mesa/utils} directory and is called 
{\tt makefile\_header\_non\_mesasdk}. Open it and change the following:
\begin{enumerate}
\item Compilers: First change the compilers labeled {\tt CC} (C compiler) 
and {\tt FC} (Fortran compiler) to the compilers of your choice ({\tt MESA} 
defaults are {\tt gcc} and {\tt ifort}).

\item {\tt UTILS\_ISNAN}: Leave this as {\tt utils\_isnan\_okay}, if {\tt MESA} 
complains upon compiling the {\tt mesa/utils} module, then change it to 
{\tt utils\_isnan\_nope}.

\item {\tt BLAS} and {\tt LAPACK}: Set the location of {\tt LAPACK} and 
{\tt BLAS} provided with the {\tt MESA} source tree using the following 
definitions:
\begin{lstlisting}[language=make,mathescape=false,upquote=true]
  WHICH_LAPACK = USE_SRCS
  LOAD_LAPACK = -lmesalapack
  WHICH_BLAS = USE_SRCS
  LOAD_BLAS = -lmesablas
  MKL_INCLUDE =
\end{lstlisting}

\item {\tt PGPLOT}: Set the variable {\tt USE\_PGPLOT} to {\tt NO} and set 
the variable {\tt LOAD\_PGPLOT} to be empty on the line corresponding to 
the compiler that was chosen. The default for {\tt gcc} and {\tt ifort} is 
{\tt YES}, but for any other compiler the default is {\tt NO}. {\tt PGPLOT} 
controls the pop up images that are used in {\tt mesa/star} (1D stellar 
evolution code).

\item {\tt USE\_SE}: Set the variable {\tt USE\_SE} to {\tt NO} and set both 
{\tt LOAD\_SE} and {\tt INCLUDE\_SE} to be empty. {\tt SE} is a library of 
formats used for writing/reading data in stellar evolution codes.

\item Compiler Flags: Installing {\tt MESA} without OpenMP will result in a 
failed install. {\tt MESA} is capable of being compiled in serial, but it 
must happen after the initial install. If not using {\tt gcc} or {\tt ifort} 
it will be necessary to enter all of the relevant compiler flags.
\end{enumerate}

All {\tt MESA} modules have a makefile that includes the main makefile named 
{\tt makefile\_header}. The file that was just edited was named 
{\tt makefile\_header\_non\_mesasdk} and so the name must be changed in order 
for the edits to take effect. This can be done as follows:
\begin{lstlisting}[language=bash,mathescape=false]
  cp makefile_header_non_mesasdk makefile_header
\end{lstlisting}

At this point, {\tt MESA} is ready to be installed, see Section 
\ref{sec:install}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Install MESA
%%%%%%%%%%%%%%%%
\subsection{Install {\tt MESA}}
\label{sec:install}
If the environment variables were added to the {\tt .bashrc} (or equivalent) 
file, do not forget to source that file before continuing. Now {\tt MESA} 
is ready to be installed. Change directories to the {\tt mesa} directory. 
Run the {\tt install} script by typing {\tt ./install}. The permissions of the 
{\tt install} script may need to be changed in order for it to be executable.
This will untar many data tables used by the equation of state, network and 
chemistry modules. It will also compile and export all modules.

Following the execution of the {\tt ./install} script, {\tt MESA} has been 
installed and the following has completed:
\begin{itemize}
\item SDK only: Supporting software (compilers, linear algebra packages, 
etc.) has been unpacked and installed
\item Data tables have been untarred
\item Linear Algebra packages have been compiled and linked
\item All modules have been compiled (in parallel)
\item All modules have been exported to the {\tt mesa/lib} and 
{\tt mesa/include} directories
\end{itemize}

If you want to run {\tt MESA} in parallel, nothing further needs to be done. 
If you want to make sure all modules are compiled in serial, see Section 
\ref{sec:compile_wo_omp}, which addresses how to compile the modules.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Compile MESA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Compile Modules Without OpenMP}
\label{sec:compile_wo_omp}

In the {\tt mesa/utils} directory, open the {\tt makefile\_header} file. 
Change the {\tt FCopenmp} flag ({\tt gfortran} default = {\tt -fopenmp}) to 
be empty. This will disable OpenMP for all modules.

Anytime a change is made to a makefile or any source code, the module must 
be recompiled, which is a three step process involving the {\tt clean}, 
{\tt mk} and {\tt export} scripts provided in each module directory. The 
first step is to clean out any old files using the {\tt ./clean} script. 
Step two is to make the module using the {\tt ./mk} script. The last step is 
to export the {\tt .mod} files and the library to the right directories using 
the {\tt ./export} script. Section \ref{sec:compile_scripts} contains a 
simple script that is meant to help with compiling modules. It cleans, makes 
and exports every module and should be run in the top {\tt mesa/} directory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional mesa/star OpenMP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Optional: Turn Off {\tt mesa/star} OpenMP}
There are four other places to turn off OpenMP, but they are all contained 
in the {\tt mesa/star} directory, which is the 1-D stellar evolution code. If 
only using a few {\tt MESA} modules that do not include the {\tt mesa/star} 
stellar evolution code, then these last few places do not need to be 
modified in any way.

In {\tt mesa/star/private/micro.f} there are three places 
to set the {\tt use\_omp} flag to False. They occur in the {\tt do\_eos}, 
{\tt do\_neu} and {\tt do\_net} routines. The fourth place to set the 
{\tt use\_omp} flag to False is contained in the {\tt do\_kap} routine in 
{\tt mesa/star/private/opacities.f}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Compiling MESA with an Outside Program
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Compiling {\tt MAESTRO} to Use {\tt MESA}}

Compiling code that depends on {\tt MESA} modules involves linking the code 
with the {\tt .mod} files and libraries that are generated when {\tt MESA} 
is compiled. After the {\tt export} script has been run in all compiled 
directories, the {\tt .mod} files and other supporting routines are 
copied to the {\tt mesa/include} directory while the libraries are copied to 
the {\tt mesa/lib} directory.

The {\tt MESA} libraries have the name {\tt lib(module-name).a}, e.g. 
{\tt libeos.a} or {\tt libnet.a}. The libraries need to be linked using the 
{\tt -L} flag with the {\tt mesa/lib} directory and a list of which libraries 
to load, e.g. {\tt -L\$(MESA\_DIR)/lib -leos -lnet -lnum}. {\tt MESA} 
{\tt .mod} files are linked to code using the {\tt -I} flag and the 
{\tt mesa/include} directory, e.g. {\tt -I\$(MESA\_DIR)/include}.

The order in which the libraries appear is important. If compilation errors 
related to undefined references to a particular {\tt MESA} module such as 
{\tt eos\_lib} occur, check to make sure that the libraries are in the correct 
order. Of course it is also possible that an undefined reference really does 
exist.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% How to setup/envoke/shutdown MESA modules
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Using {\tt MESA} Modules}

This section describes what needs to take place in order to use 
{\tt MESA} modules such as setting up the module, invoking certain routines 
and shutting down the module. This discussion will only include information 
pertaining to the network and equation of state modules.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% EOS
%%%%%%
\subsection{Equation of State}

%-------------------------------------------------------------------------
\subsubsection{Set Up}

There are several steps to setting up the equation of state:
\begin{enumerate}
\item Constant Initialization: This is done by calling the {\tt const\_init} 
subroutine

\item Data Directory: Define where the {\tt mesa/data} directory is: This can 
be done by using the {\tt MESA\_DIR} environment variable that was set 
at installation

\item Chem Initialization: This is done by calling the {\tt chem\_init} 
subroutine: {\tt call chem\_init(data\_dir, `isotopes.data\_real', ierr)}. 
The second argument can also be {\tt `isotopes.data\_approx'}. The ``real'' 
file contains actual values for 
atomic quantities such as atomic weight. The ``approx'' file contains 
approximate values for these quantities, e.g. the mass of the proton and 
neutron are both equal to 1.0. The {\tt ierr} argument is an integer that 
holds the exit status of the subroutine, non-zero means there was an error

\item EOS Initialization: This is done by calling the {\tt eos\_init} 
subroutine: {\tt call eos\_init(data\_dir, eos\_file\_prefix, use\_cache, 
ierr)}. {\tt eos\_file\_prefix} is a character variable and should be 
declared and set equal to `mesa'. {\tt use\_cache} is a logical variable that 
also needs to be declared and set equal to `true'

\item Handle: Obtain a handle associated with the equation of state by 
calling the {\tt alloc\_eos\_handle} function: 
{\tt handle = alloc\_eos\_handle(ierr)}. {\tt handle} should be declared as an 
integer
\end{enumerate}

%-------------------------------------------------------------------------
\subsubsection{Calling the EOS}

The call to the generic {\tt MESA} equation of state looks like:
\begin{lstlisting}[language=fortran,mathescape=false]
  call eosDT_get(handle_eos, Z, X, abar, zbar, nspec, chem_id, net_iso, Xa, &
                 rho, log10(rho), temp, log10(temp), res, d_dlnd, d_dlnT, ierr)
\end{lstlisting}
Where {\tt Z} is the metallicity, {\tt X} is the mass fraction of hydrogen, 
{\tt abar} is the average atomic weight, {\tt zbar} is the average atomic 
number, {\tt nspec} is the number of species in the network, {\tt Xa} is 
the array of mass fractions. See Section \ref{sec:mapping} for a complete 
description of {\tt chem\_id} and {\tt net\_iso}. The results of the EOS call 
are stored in the arrays {\tt res}, {\tt d\_dlnd} and {\tt d\_dlnT}, which 
are all of size {\tt num\_eos\_basic\_results}. The {\tt d\_dlnd} and 
{\tt d\_dlnT} arrays are the derivatives of the {\tt res} array with respect 
to density and temperature, respectively. The values held in the {\tt res} 
array are shown in Table \ref{tab:res}.

\begin{table}[h]
\begin{center}
\begin{tabular}{ccc}
Output & Definition & Units \\ \hline
$P_{gas}$ & gas pressure & ergs cm$^{-3}$ \\
$E$ & internal energy & ergs g$^-{1}$ \\
$S$ & entropy per gram & ergs g$^{-1}$ K$^{-1}$ \\
$dE/d\rho|_T$ & & ergs cm$^3$ g$^{-2}$ \\
$C_V$ & specific heat at constant V $\equiv 1/\rho$ & ergs g$^{-1}$ K$^{-1}$ \\
$dS/d\rho|_T$ & & ergs cm$^3$ g$^{-2}$ K$^{-1}$ \\
$dS/dT|_\rho$ & & ergs g$^{-1}$ K$^{-2}$ \\
$\chi_\rho$ & $\equiv d{\rm ln}P/d{\rm ln}\rho|_T$ & none \\
$\chi_T$ & $\equiv d{\rm ln}P/d{\rm ln}T|_\rho$ & none \\
$C_P$ & specific heat at constant pressure & ergs g$^{-1}$ K$^{-1}$ \\
$\nabla_{ad}$ & adiabatic T gradient with pressure & none \\
$\Gamma_1$ & $\equiv d{\rm ln}P/d{\rm ln}\rho|_S$ & none \\
$\Gamma_3$ & $\equiv d{\rm ln}T/d{\rm ln}\rho|_S + 1$ & none \\
$\eta$ & ratio of electron chemical potential to $k_BT$ & none \\
$\mu$ & mean molecular weight per gas particle & none \\
$1/\mu_e$ & mean number of free electrons per nucleon & none \\ \hline
\end{tabular}
\caption{{\tt eos} output quantities and units}
\label{tab:res}
\end{center}
\end{table}

It is possible to call the HELM equation of state directly. This is done as 
follows:
\begin{lstlisting}[language=fortran,mathescape=false]
  call eos_get_helm_results(handle_eos, Z, X, abar, zbar, dens, log10(dens), &
                            temp, log10(temp), helm_res, ierr)
\end{lstlisting}
The {\tt helm\_res} array is of size {\tt num\_helm\_results} and holds all 
of the output 
values from the call, including all derivatives. 
There are slight differences between the generic EOS call and the HELM EOS 
call: HELM does not need the mass fractions or the maps between 
{\tt MESA} isotope indices and the mass fraction array indices ({\tt chem\_id} 
and {\tt net\_iso}, see Section \ref{sec:mapping}). The HELM EOS returns on 
the order of 400 different quantities which include the values shown in 
Table \ref{tab:res} as well as quantities such as $dP/d(abar)$ and 
$dE/d(zbar)$.

%-------------------------------------------------------------------------
\subsubsection{Shutdown the EOS}

There are two steps to shutting down the equation of state:
\begin{enumerate}
\item Free the handle: Call the {\tt free\_eos\_handle} subroutine: 
{\tt call free\_eos\_handle(handle\_eos)}.

\item Shutdown: Call the {\tt eos\_shutdown} subroutine: 
{\tt call eos\_shutdown}
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Mapping from MAESTRO to MESA using chem_id and net_iso
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mapping {\tt MAESTRO} to {\tt MESA}}
\label{sec:mapping}

{\tt MAESTRO} and {\tt MESA} refer to isotopes in a very similar way; using 
short names of the isotopes such as he4 and ni56. {\tt MAESTRO} 
tends to use capital letters (He4 and Ni56) while {\tt MESA} uses 
lowercase letters (he4 and ni56). A simple routine to convert from 
uppercase to lowercase will be needed.

In order to use {\tt MESA} modules and obtain the correct result, it is 
important to make sure that {\tt MESA} and {\tt MAESTRO} agree on what 
isotope each entry in the mass fraction array refers to. This is done by 
using the {\tt chem\_id} and {\tt net\_iso} variables. Both variables are 
integer pointer arrays of size {\tt num\_chem\_isos}, which is defined in one 
of the chemistry modules. {\tt chem\_id} maps from an isotope in the current 
net to the {\tt MESA} data table that holds the properties of that isotope. 
{\tt net\_iso} is used to map from the {\tt MESA} data tables to an isotope 
in the current network.

For example, assume there are 3 isotopes in the current net: H1, Be7 and C12. 
The {\tt chem\_id} array should look like this: {\tt chem\_id(:) = -1, 
chem\_id(1) = 2, chem\_id(2) = 18, chem\_id(3) = 38}, and the 
{\tt net\_iso} array should look like this: {\tt net\_iso(:) = -1, 
net\_iso(2) = 1, net\_iso(18) = 2, net\_iso(38) = 3}. Where 2, 18 and 38 
refer to the {\tt MESA} data table location of H1, Be7 and C12, respectively.

Care should be taken when setting up these arrays in order to avoid 
disagreement between the {\tt MESA} mass fraction array and the {\tt MAESTRO} 
mass fraction array.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Network
%%%%%%%%%%
\subsection{Network}

%-------------------------------------------------------------------------
\subsubsection{Set Up}

There are significantly more steps involved when initializing the {\tt MESA} 
network. Here is a brief outline of what needs to happen:
\begin{enumerate}
\item Get the {\tt MESA} and {\tt AstroDev} directory locations
\item Convert the short species names to {\tt MESA} format
\item Generate a file that will tell {\tt MESA} which reactions to add based 
on the isotopes in the current network ({\tt call which\_reactions})
\item Initialize {\tt MESA} {\tt chem}, {\tt weak} and {\tt reaclib} modules
\item Setup the map between {\tt MAESTRO} and {\tt MESA}, see Section 
\ref{sec:mapping} for a more in depth explanation
\item Initialize {\tt MESA} {\tt rates} and {\tt net} modules
\item Allocate a network handle
\item Call the {\tt net\_start\_def}, {\tt read\_net\_file}, {\tt net\_ptr} 
and {\tt net\_finish\_def}
\item Output a log file containing the species in the current net and which 
reactions were chosen by {\tt MESA} based on those species
\item Set {\tt which\_rates\_choice} and allocate {\tt which\_rates} array
\item Set entire array equal to {\tt which\_rates\_choice} and call 
{\tt net\_set\_which\_rates}
\item Initialize which linear algebra solver to use
\end{enumerate}

After calling {\tt net\_set\_which\_rates}, the setup routine should call 
{\tt net\_setup\_tables}. This is based on the example 
{\tt one\_zone\_burn.f} program distributed with the {\tt MESA} source code. 
I have not had a successful burn if the {\tt net\_setup\_tables} routine is 
called at setup in the {\tt setup\_mesa\_net.f90} program. I do have 
successful burns when it is placed within the burner program 
{\tt Do\_One\_Zone\_Burn.f90}. To prevent the setup routine from being called 
everytime a burn is to take place, I added a logical flag that signals if the 
{\tt net\_setup\_tables} routine has alredy been called.

%-------------------------------------------------------------------------
\subsubsection{Calling the Network}

Calling the {\tt MESA} network from within the {\tt MAESTRO} wrapper starts 
with calling a modified {\tt MESA} wrapper:
\begin{lstlisting}[language=fortran,mathescape=false]
  results => Xout

  call Do_One_Zone_Burn(dens, temp, dt, Xin, burn_ergs, Xout, results)
\end{lstlisting}
The inputs are density, temperature, how long to burn and starting mass 
fractions. The outputs are the energy generation rate ({\tt burn\_ergs}) 
in units of ergs g$^{-1}$ s$^{-1}$, the final mass fractions and a pointer 
to the final mass fractions ({\tt results}). The pointer is necessary 
because the {\tt MESA} network works with pointers while {\tt MAESTRO} 
uses standard arrays.

To calculate {\tt rho\_omegadot} ($\rho \dot{\omega}$) and {\tt rho\_Hnuc} 
($\rho H_{{\rm nuc}}$) the following code can be used after a call to 
{\tt Do\_One\_Zone\_Burn}:
\begin{lstlisting}[language=fortran,mathescape=false]
  do i=1, nspec
     dX = Xout(i) - Xin(i)
     rho_omegadot(i) = dens * dX / dt
  enddo

  rho_Hnuc = dens * burn_ergs
\end{lstlisting}
There is a support routine called {\tt burn\_solout} that is passed to the 
{\tt MESA} burner. This routine is called after every step of the burn and 
is responsible for calculating the energy generation rate. {\tt burn\_solout} 
is passed as an argument to the {\tt net\_1\_zone\_burn} subroutine (the 
{\tt MESA} burner) and as a result the actual call to the {\tt burn\_solout} 
routine is buried deep within the {\tt MESA} source code.

%-------------------------------------------------------------------------
\subsubsection{Shutdown the Network}

Shutting down the {\tt MESA} network requires only one step:
\begin{enumerate}
\item Deallocate {\tt chem\_id}, {\tt net\_iso} and {\tt which\_rates}
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Compile MESA on Hopper, a Cray XE6 machine
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Install {\tt MESA} on Hopper}
\label{sec:hopper_install}
\input hopper_compile

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Begin the appendix, where all the code examples are
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input appendix

\end{document}

### 2012-08-02 docs/MESA/instructions.tex
add footnote

@@
 
-Modules for Experiments in Stellar Astrophysics ({\tt MESA}) is a collection 
-of open source, robust and efficient modules developed for a wide range of 
-applications in computational stellar astrophysics. The modules include 
-macroscopic and microscopic physics such as nuclear reaction rates, opacity, 
-equation of state, element diffusion data and atmosphere boundary conditions 
-in addition to numerical routines. This document describes how to link the 
-{\tt MESA} modules to an outside program such as {\tt MAESTRO}.
+Modules for Experiments in Stellar Astrophysics ({\tt MESA}) is a
+collection of open source, robust and efficient modules developed for
+a wide range of applications in computational stellar
+astrophysics. The modules include macroscopic and microscopic physics
+such as nuclear reaction rates, opacity, equation of state, element
+diffusion data and atmosphere boundary conditions in addition to
+numerical routines. This document describes how to link the {\tt MESA}
+modules to an outside program such as {\tt MAESTRO}\footnote{This
+  interface work was done by Ryan Orvedahl---questions (and credit)
+  should be directed to him}.
 

### 2012-08-07 docs/MESA/hopper_compile.tex
updates to MESA docs

@@
-Once {\tt MESA} was installed and working properly on a local machine (e.g. 
+Once {\sf MESA} was installed and working properly on a local machine (e.g. 
 Bender), I tried installing it on Hopper, which is a Cray XE6 machine. On 
 Bender, which is an Intel(R) Xeon(R) X5650 machine, the compiler was 
-{\tt gcc} version 4.7.0, on Hopper I tried compiling {\tt MESA} with the 
+{\tt gcc} version 4.7.0, on Hopper I tried compiling {\sf MESA} with the 
 Cray Compiler using version 8.0.5. The module that was loaded to enable this 
@@
 
-There are several steps needed to compile {\tt MAESTRO} with {\tt MESA} on 
+There are several steps needed to compile {\sf MAESTRO} with {\sf MESA} on 
 Hopper:
@@
 \item Edit the {\tt build\_and\_test} script in the {\tt utils} directory
+\item Edit the {\tt chem\_lib.f} file in the {\tt chem} directory
 \item Edit the {\tt mtx} makefile
@@
 {\tt interp\_1d}
-\item Edit the {\tt install} script in the main {\tt MESA} directory
+\item Edit the {\tt install} script in the main {\sf MESA} directory
 \item Edit the {\tt BoxLib/Tools/F\_mk/GMakedefs.mak} makefile
 \item Edit the {\tt BoxLib/Tools/F\_mk/comps/Linux\_cray.mak} file
-\item Make sure the makefile {\tt COMP} variable is set to {\tt Cray} in the 
-problem directory
+\item Install {\sf MESA}
+\item Compile {\sf MAESTRO}
 \end{enumerate}
@@
 and running the test directory. We want to turn off the compiling and 
-running of the test directory to ensure that {\tt MESA} compiles successfully. 
+running of the test directory to ensure that {\sf MESA} compiles successfully. 
 A portion of the {\tt build\_and\_test} script looks like this:
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{{\tt chem\_lib.f} Source File}
+
+Open the {\tt chem\_lib.f} file located in {\tt chem/public}. In the 
+{\tt generate\_nuclide\_set} subroutine, the implied do loop must be removed 
+and replaced with an actual do loop. For an unknown reason, the implied do 
+loop was causing a Segmentation Fault on Hopper with the Cray compiler. The 
+implied loop:
+\begin{lstlisting}[language=fortran,mathescape=false]
+  set = [(nuclide_set(names(i), i), i=1, size(names))]
+\end{lstlisting}
+should be changed to an explicit do loop:
+\begin{lstlisting}[language=fortran,mathescape=false]
+  do i=1, size(names)
+     set(i) = nuclide_set(names(i), i)
+  enddo
+\end{lstlisting}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{{\tt mtx} makefile}
@@
 {\tt utils\_def.o} or {\tt num\_def.o}. This files is never loaded into the 
-library as seen below for the {\tt utils} directory:
+library as seen below for the {\tt utils} directory. Certain modules have a 
+{\tt *\_def.f} file with both definitions and declarations of data that will 
+be needed at runtime, so they must be included in the library. Other modules 
+have a {\tt *\_def.f} file that contains only definitions and no data 
+declarations and therefor only needs to be included at compile time. The Cray 
+compiler on Hopper for whatever reason complained that there were undefined 
+references to certain {\tt *\_def.o} files. The undefined references only 
+occurred for modules whose {\tt *\_def.f} file did not need to be included in 
+the library. The fix was to manually include the {\tt *\_def.o} file in the 
+library for all of the necessary module directories. The {\tt utils} directory 
+is shown here:
 \begin{lstlisting}[language=make,mathescape=false]
@@
 
-Certain modules that are unneccessary for use with {\tt MAESTRO} do not need 
+Certain modules that are unneccessary for use with {\sf MAESTRO} do not need 
 to be installed (and also cause some sort of installation error). In the 
-main {\tt MESA} directory, edit the {\tt install} script and comment out 
+main {\sf MESA} directory, edit the {\tt install} script and comment out 
 the following four lines: {\tt do\_one sample}, {\tt do\_one star}, 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\tt BoxLib Linux\_cray.mak} file}
+\subsection{{\sf BoxLib Linux\_cray.mak} file}
 
-Two flags used in the compilation of {\tt MESA} with the Cray compiler are 
+Two flags used in the compilation of {\sf MESA} with the Cray compiler are 
 important; the {\tt -e m} or {\tt -em} flag and the {\tt -e f} or {\tt -ef} 
@@
 {\tt -em} flag, instead outputs {\tt .mod} files of the form {\tt name.mod}. 
-This is important for {\tt MESA} because the {\tt export} scripts copy the 
+This is important for {\sf MESA} because the {\tt export} scripts copy the 
 various libraries and {\tt .mod} files to the correct directories. The 
@@
 {\tt CONST\_LIB.mod} and will not get copied to the {\tt include} directory 
-(the library is unaffected). The default for {\tt MAESTRO} is to not use 
-the {\tt -ef} flag, but it does use the {\tt -em} flag. To make {\tt MAESTRO} 
-and {\tt MESA} compatible, I added the {\tt -ef} flag to the 
+(the library is unaffected). The default for {\sf MAESTRO} is to not use 
+the {\tt -ef} flag, but it does use the {\tt -em} flag. To make {\sf MAESTRO} 
+and {\sf MESA} compatible, I added the {\tt -ef} flag to the 
 {\tt Linux\_cray.mak} file located in {\tt BoxLib/Tools/F\_mk/comps}. I also 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\tt BoxLib GMakedefs.mak} makefile}
+\subsection{{\sf BoxLib GMakedefs.mak} makefile}
 
-In {\tt MAESTRO}, {\tt .mod} files are located in the {\tt \$(tdir)/m} 
-directory, where {\tt tdir = t/\$(suf)} and {\tt suf} was in this case 
-{\tt Linux.Cray}. This means the {\tt MAESTRO} routines are expecting to find 
+In {\sf MAESTRO}, {\tt .mod} files are located in the {\tt \$(tdir)/m} 
+directory, where {\tt tdir = t/\$(suf)} and {\tt \$(suf)} was in this case 
+{\tt Linux.Cray}. This means the {\sf MAESTRO} routines are expecting to find 
 module files with the name {\tt t/Linux.Cray/m/network.mod}. The addition of 
@@
 was creating files with the name {\tt t/linux.cray/m/network.mod}. To solve 
-this, I modified what {\tt suf} is defined to be when using the Cray compiler. 
+this, I modified what {\tt tdir} is defined to be when using the Cray compiler. 
 This was done in the {\tt BoxLib/Tools/F\_mk/GMakedefs.mak} makefile by using 
@@
 {\tt \$(suf)} to lowercase.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Install {\sf MESA} and Compile {\sf MAESTRO}}
+
+At this point the {\tt install} script in the top {\sf MESA} directory can be 
+executed to install {\sf MESA}. This will compile the source code, generate 
+the libraries and copy the necessary files to the correct directories. Once, 
+{\sf MESA} is installed, {\sf MAESTRO} can be compiled and linked such that 
+it can use {\sf MESA}'s EOS and/or network. Make sure the makefile variable 
+{\tt COMP} is set to {\tt Cray} so the Cray compiler is used.

### 2012-08-07 docs/MESA/instructions.tex
updates to MESA docs

--- 2012-08-02 docs/MESA/instructions.tex
@@
 
-Modules for Experiments in Stellar Astrophysics ({\tt MESA}) is a collection 
+Modules for Experiments in Stellar Astrophysics ({\sf MESA}) is a collection 
 of open source, robust and efficient modules developed for a wide range of 
@@
 in addition to numerical routines. This document describes how to link the 
-{\tt MESA} modules to an outside program such as {\tt MAESTRO}.
+{\sf MESA} modules to an outside program such as {\sf MAESTRO}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Setting Up {\tt MESA}}
+\section{Setting Up {\sf MESA}}
 
-To install {\tt MESA}, download the source code from the {\tt MESA} website, 
+To install {\sf MESA}, download the source code from the {\sf MESA} website, 
 {\tt http://mesa.sourceforge.net}. Check the website for the latest version, 
 which is 4219 with the writing of this document. The codes and algorithms 
-discussed here were created when the current version of {\tt MESA} was 
+discussed here were created when the current version of {\sf MESA} was 
 version 4088. Between versions 4088 and 4219, there are minor changes that 
@@
 Rich Townsend has developed a Software Development Kit (SDK) to help install 
-{\tt MESA} called {\tt MESASDK}. The SDK will install certain software 
-packages that {\tt MESA} uses and it will also change the {\tt PATH} 
+{\sf MESA} called {\sf MESASDK}. The SDK will install certain software 
+packages that {\sf MESA} uses and it will also change the {\tt PATH} 
 environment variable such that these programs become the default. A tarball 
@@
 %%%%%%%%%%%%%%%%%%
-\subsection{Setup {\tt MESA} Without {\tt MESASDK}}
+\subsection{Setup {\sf MESA} Without {\sf MESASDK}}
 \label{sec:woSDK}
@@
 \item Compilers: First change the compilers labeled {\tt CC} (C compiler) 
-and {\tt FC} (Fortran compiler) to the compilers of your choice ({\tt MESA} 
+and {\tt FC} (Fortran compiler) to the compilers of your choice ({\sf MESA} 
 defaults are {\tt gcc} and {\tt ifort}).
 
-\item {\tt UTILS\_ISNAN}: Leave this as {\tt utils\_isnan\_okay}, if {\tt MESA} 
+\item Remove the second instance of {\tt CC} located directly under the 
+{\tt SPECIAL\_FC\_FLAGS} variable
+
+\item {\tt UTILS\_ISNAN}: Leave this as {\tt utils\_isnan\_okay}, if {\sf MESA} 
 complains upon compiling the {\tt mesa/utils} module, then change it to 
@@
 \item {\tt BLAS} and {\tt LAPACK}: Set the location of {\tt LAPACK} and 
-{\tt BLAS} provided with the {\tt MESA} source tree using the following 
+{\tt BLAS} provided with the {\sf MESA} source tree using the following 
 definitions:
@@
 
-\item Compiler Flags: Installing {\tt MESA} without OpenMP will result in a 
-failed install. {\tt MESA} is capable of being compiled in serial, but it 
+\item Compiler Flags: Installing {\sf MESA} without OpenMP will result in a 
+failed install. {\sf MESA} is capable of being compiled in serial, but it 
 must happen after the initial install. If not using {\tt gcc} or {\tt ifort} 
@@
 
-All {\tt MESA} modules have a makefile that includes the main makefile named 
+All {\sf MESA} modules have a makefile that includes the main makefile named 
 {\tt makefile\_header}. The file that was just edited was named 
@@
 
-At this point, {\tt MESA} is ready to be installed, see Section 
+At this point, {\sf MESA} is ready to be installed, see Section 
 \ref{sec:install}.
@@
 %%%%%%%%%%%%%%%%
-\subsection{Install {\tt MESA}}
+\subsection{Install {\sf MESA}}
 \label{sec:install}
 If the environment variables were added to the {\tt .bashrc} (or equivalent) 
-file, do not forget to source that file before continuing. Now {\tt MESA} 
+file, do not forget to source that file before continuing. Now {\sf MESA} 
 is ready to be installed. Change directories to the {\tt mesa} directory. 
@@
 
-Following the execution of the {\tt ./install} script, {\tt MESA} has been 
+Following the execution of the {\tt ./install} script, {\sf MESA} has been 
 installed and the following has completed:
@@
 
-If you want to run {\tt MESA} in parallel, nothing further needs to be done. 
+If you want to run {\sf MESA} in parallel, nothing further needs to be done. 
 If you want to make sure all modules are compiled in serial, see Section 
@@
 in the {\tt mesa/star} directory, which is the 1-D stellar evolution code. If 
-only using a few {\tt MESA} modules that do not include the {\tt mesa/star} 
+only using a few {\sf MESA} modules that do not include the {\tt mesa/star} 
 stellar evolution code, then these last few places do not need to be 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Compiling {\tt MAESTRO} to Use {\tt MESA}}
+\section{Compiling {\sf MAESTRO} to Use {\sf MESA}}
 
-Compiling code that depends on {\tt MESA} modules involves linking the code 
-with the {\tt .mod} files and libraries that are generated when {\tt MESA} 
+Compiling code that depends on {\sf MESA} modules involves linking the code 
+with the {\tt .mod} files and libraries that are generated when {\sf MESA} 
 is compiled. After the {\tt export} script has been run in all compiled 
@@
 
-The {\tt MESA} libraries have the name {\tt lib(module-name).a}, e.g. 
+The {\sf MESA} libraries have the name {\tt lib(module-name).a}, e.g. 
 {\tt libeos.a} or {\tt libnet.a}. The libraries need to be linked using the 
 {\tt -L} flag with the {\tt mesa/lib} directory and a list of which libraries 
-to load, e.g. {\tt -L\$(MESA\_DIR)/lib -leos -lnet -lnum}. {\tt MESA} 
+to load, e.g. {\tt -L\$(MESA\_DIR)/lib -leos -lnet -lnum}. {\sf MESA} 
 {\tt .mod} files are linked to code using the {\tt -I} flag and the 
@@
 The order in which the libraries appear is important. If compilation errors 
-related to undefined references to a particular {\tt MESA} module such as 
+related to undefined references to a particular {\sf MESA} module such as 
 {\tt eos\_lib} occur, check to make sure that the libraries are in the correct 
@@
 
+The wrapper routines that allow {\sf MAESTRO} to use {\sf MESA} modules are 
+located in the {\tt AstroDev/EOS/MESA} and {\tt AstroDev/networks/MESA} 
+directories. In each directory there is a {\tt GPackage.mak} file that 
+specifies the proper {\tt -I} and {\tt -L} compiler flags along with what 
+source files to include in the build. To use the EOS, open the 
+{\tt GNUmakefile} in the {\sf MAESTRO} problem directory and add the 
+following lines:
+\begin{lstlisting}[language=make,mathescape=false]
+  EOS_TOP_DIR := $(ASTRODEV_DIR)/EOS
+  EOS_DIR := MESA
+\end{lstlisting}
+To use the network, add the following lines:
+\begin{lstlisting}[language=make,mathescape=false]
+  NETWORK_TOP_DIR := $(ASTRODEV_DIR)/networks
+  NETWORK_DIR := MESA
+  GENERAL_NET_INPUTS := $(NETWORK_TOP_DIR)/MESA/Net-Files/net_input.net
+\end{lstlisting}
+The {\tt GENERAL\_NET\_INPUTS} variable holds the path to the
+{\tt net\_input.net} file, which does not need to be kept in the
+{\tt Net-Files} directory. The {\tt net\_input.net} file could look like this:
+\begin{lstlisting}[language=make,mathescape=false]
+  # name              short-name    aion   zion
+  hydrogen-1             H1          1.0    1.0
+  helium-4               He4         4.0    2.0
+  carbon-12              C12         12.0   6.0
+  oxygen-16              O16         16.0   8.0
+  neon-20                Ne20        20.0  10.0
+  sodium-23              Na23        23.0  11.0
+  magnesium-23           Mg23        23.0  12.0
+  magnesium-24           Mg24        24.0  12.0
+  silicon-28             Si28        28.0  14.0
+\end{lstlisting}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Using {\tt MESA} Modules}
+\section{Using {\sf MESA} Modules}
 
 This section describes what needs to take place in order to use 
-{\tt MESA} modules such as setting up the module, invoking certain routines 
+{\sf MESA} modules such as setting up the module, invoking certain routines 
 and shutting down the module. This discussion will only include information 
@@
 
-The call to the generic {\tt MESA} equation of state looks like:
+The call to the generic {\sf MESA} equation of state looks like:
 \begin{lstlisting}[language=fortran,mathescape=false]
@@
 call: HELM does not need the mass fractions or the maps between 
-{\tt MESA} isotope indices and the mass fraction array indices ({\tt chem\_id} 
+{\sf MESA} isotope indices and the mass fraction array indices ({\tt chem\_id} 
 and {\tt net\_iso}, see Section \ref{sec:mapping}). The HELM EOS returns on 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Mapping {\tt MAESTRO} to {\tt MESA}}
+\subsection{Mapping {\sf MAESTRO} to {\sf MESA}}
 \label{sec:mapping}
 
-{\tt MAESTRO} and {\tt MESA} refer to isotopes in a very similar way; using 
-short names of the isotopes such as he4 and ni56. {\tt MAESTRO} 
-tends to use capital letters (He4 and Ni56) while {\tt MESA} uses 
+{\sf MAESTRO} and {\sf MESA} refer to isotopes in a very similar way; using 
+short names of the isotopes such as he4 and ni56. {\sf MAESTRO} 
+tends to use capital letters (He4 and Ni56) while {\sf MESA} uses 
 lowercase letters (he4 and ni56). A simple routine to convert from 
@@
 
-In order to use {\tt MESA} modules and obtain the correct result, it is 
-important to make sure that {\tt MESA} and {\tt MAESTRO} agree on what 
+In order to use {\sf MESA} modules and obtain the correct result, it is 
+important to make sure that {\sf MESA} and {\sf MAESTRO} agree on what 
 isotope each entry in the mass fraction array refers to. This is done by 
@@
 of the chemistry modules. {\tt chem\_id} maps from an isotope in the current 
-net to the {\tt MESA} data table that holds the properties of that isotope. 
-{\tt net\_iso} is used to map from the {\tt MESA} data tables to an isotope 
+net to the {\sf MESA} data table that holds the properties of that isotope. 
+{\tt net\_iso} is used to map from the {\sf MESA} data tables to an isotope 
 in the current network.
@@
 net\_iso(2) = 1, net\_iso(18) = 2, net\_iso(38) = 3}. Where 2, 18 and 38 
-refer to the {\tt MESA} data table location of H1, Be7 and C12, respectively.
+refer to the {\sf MESA} data table location of H1, Be7 and C12, respectively.
 
 Care should be taken when setting up these arrays in order to avoid 
-disagreement between the {\tt MESA} mass fraction array and the {\tt MAESTRO} 
+disagreement between the {\sf MESA} mass fraction array and the {\sf MAESTRO} 
 mass fraction array.
@@
 
-There are significantly more steps involved when initializing the {\tt MESA} 
+There are significantly more steps involved when initializing the {\sf MESA} 
 network. Here is a brief outline of what needs to happen:
 \begin{enumerate}
-\item Get the {\tt MESA} and {\tt AstroDev} directory locations
-\item Convert the short species names to {\tt MESA} format
-\item Generate a file that will tell {\tt MESA} which reactions to add based 
+\item Get the {\sf MESA} and {\sf AstroDev} directory locations
+\item Convert the short species names to {\sf MESA} format
+\item Generate a file that will tell {\sf MESA} which reactions to add based 
 on the isotopes in the current network ({\tt call which\_reactions})
-\item Initialize {\tt MESA} {\tt chem}, {\tt weak} and {\tt reaclib} modules
-\item Setup the map between {\tt MAESTRO} and {\tt MESA}, see Section 
+\item Initialize {\sf MESA} {\tt chem}, {\tt weak} and {\tt reaclib} modules
+\item Setup the map between {\sf MAESTRO} and {\sf MESA}, see Section 
 \ref{sec:mapping} for a more in depth explanation
-\item Initialize {\tt MESA} {\tt rates} and {\tt net} modules
+\item Initialize {\sf MESA} {\tt rates} and {\tt net} modules
 \item Allocate a network handle
@@
 \item Output a log file containing the species in the current net and which 
-reactions were chosen by {\tt MESA} based on those species
+reactions were chosen by {\sf MESA} based on those species
 \item Set {\tt which\_rates\_choice} and allocate {\tt which\_rates} array
@@
 {\tt net\_setup\_tables}. This is based on the example 
-{\tt one\_zone\_burn.f} program distributed with the {\tt MESA} source code. 
+{\tt one\_zone\_burn.f} program distributed with the {\sf MESA} source code. 
 I have not had a successful burn if the {\tt net\_setup\_tables} routine is 
@@
 
-Calling the {\tt MESA} network from within the {\tt MAESTRO} wrapper starts 
-with calling a modified {\tt MESA} wrapper:
+Calling the {\sf MESA} network from within the {\sf MAESTRO} wrapper starts 
+with calling a modified {\sf MESA} wrapper:
 \begin{lstlisting}[language=fortran,mathescape=false]
-  results => Xout
-
-  call Do_One_Zone_Burn(dens, temp, dt, Xin, burn_ergs, Xout, results)
+  call Do_One_Zone_Burn(dens, temp, dt, Xin, burn_ergs, Xout)
 \end{lstlisting}
@@
 fractions. The outputs are the energy generation rate ({\tt burn\_ergs}) 
-in units of ergs g$^{-1}$ s$^{-1}$, the final mass fractions and a pointer 
-to the final mass fractions ({\tt results}). The pointer is necessary 
-because the {\tt MESA} network works with pointers while {\tt MAESTRO} 
-uses standard arrays.
+in units of ergs g$^{-1}$ s$^{-1}$ and the final mass fractions.
 
@@
 There is a support routine called {\tt burn\_solout} that is passed to the 
-{\tt MESA} burner. This routine is called after every step of the burn and 
+{\sf MESA} burner. This routine is called after every step of the burn and 
 is responsible for calculating the energy generation rate. {\tt burn\_solout} 
 is passed as an argument to the {\tt net\_1\_zone\_burn} subroutine (the 
-{\tt MESA} burner) and as a result the actual call to the {\tt burn\_solout} 
-routine is buried deep within the {\tt MESA} source code.
+{\sf MESA} burner) and as a result the actual call to the {\tt burn\_solout} 
+routine is buried deep within the {\sf MESA} source code.
 
@@
 
-Shutting down the {\tt MESA} network requires only one step:
+Shutting down the {\sf MESA} network requires only one step:
 \begin{enumerate}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Install {\tt MESA} on Hopper}
+\section{Install {\sf MESA} on Hopper}
 \label{sec:hopper_install}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Begin the appendix, where all the code examples are
+% Handy compilation script
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input appendix
+\section{Compilation Script}
+ \label{sec:compile_scripts}
+
+Useful script written in {\tt BASH} to help with compiling {\sf MESA} modules. 
+The script cleans the module directory, compiles the source code and exports 
+the libraries and {\tt .mod} files to the appropriate directories for every 
+module. This script can be found in {\tt AstroDev/networks/MESA/scripts}. 
+There is also a script to compile just one module. The compile-all-modules 
+script should be run from the top {\sf MESA} directory while the 
+compile-one-module script can be run from anywhere because it takes the 
+relative path to the module.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Script
+%%%%%%%%
+\subsection{Compile All Modules}
+\input compile_all_mods
 

### 2012-08-21 docs/compilers/compilers.tex
compiler notes

@@
 
-Cray 8.0.2 compiles and runs \maestro\ fine on the interlagos/jaguarpf 
+\begin{itemize}
+\item Cray 8.0.2 compiles and runs \maestro\ fine on the interlagos/jaguarpf 
 machine at OCLF.
+\item Cray 5.12 compiles and runs \maestro\ fine on hopper (8/21/12).
+\end{itemize}
 
@@
 \begin{itemize}
+\item gfortran 4.4.3 dies with an out of bounds error in multigrid for
+2+ level runs with effective 128$^3$ resolution or better.
 \item gfortran 4.5 has no known issues.
@@
 it seems to be related to optimization.
-\item gfortran 4.7.1 (prerelease) seems to have fixed the above issue.
+\item gfortran 4.7.1 (prerelease) seems to have fixed the above issue,
+but now wdconvect with 2-level, effective 768$^3$ dies due to an
+unknown seg fault at the beginning of the run.
 \end{itemize}

### 2012-08-30 docs/unit_tests/unit_tests.tex
add 3-d test_projection figure

@@
 
-  This tests the projection routines in 2-d---either the hgprojection
+  This tests the projection routines in 2- and 3-d---either the hgprojection
   ({\tt project\_type = 1}) or the MAC projection ({\tt project\_type =
@@
 %
-\caption[Results of the {\tt test\_projection} unit test]{\label{fig:unit:projtest}
-  Initial divergence free velocity field (x-component; left); Velocity field plus
-  gradient of a scalar (x-component; center); and resulting velocity after projecting
-  out the non-divergence free portion (x-component; right).  This is will slipwall
-  boundary conditions on all sides, a 2-level grid with the left half refined and
-  right half coarse, and the hgprojection tested.}
+\caption[Results of 2-d the {\tt test\_projection} unit test]{\label{fig:unit:projtest}
+  Initial divergence free velocity field (x-component; left); Velocity
+  field plus gradient of a scalar (x-component; center); and resulting
+  velocity after projecting out the non-divergence free portion
+  (x-component; right).  This is with slipwall boundary conditions on
+  all sides, a 2-level grid with the left half refined and right half
+  coarse, and the hgprojection tested.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[height=8in]{\unitfigpath/test_project_3d}
+%
+\caption[Results of the 3-d {\tt test\_projection} unit test]{\label{fig:unit:projtest3d}
+  Projection test in 3-d showing the x-velocity (left), y-velocity
+  (middle), and z-velocity (right) initially (top row), after the
+  gradient of a scalar is added (center row), and the resulting
+  velocity after the projection.  This is with slipwall boundary conditions
+  on all sides, a 2-level grid with an octant refined, and the hgprojection.}
 \end{figure}

### 2012-09-13 docs/architecture/architecture.tex

@@
     do n=1,nlevs
-       do i = 1, nboxes(s(n))
-\end{lstlisting}
-{\tt nboxes(s(n))} is simply the number of boxes in level {\tt n}.
-All data in all \multifab s will have the same box layout, so we only
-need to query the number of boxes from one \multifab.
+       do i = 1, nfabs(s(n))
+\end{lstlisting}
+{\tt nfabs(s(n))} is simply the number of boxes in level {\tt n} on
+the current processor.  Each processor knows which \fab s in its
+\multifab are local to that processor, and this loop will only loop
+over those.
 
@@
 \begin{lstlisting}[language={[95]fortran},mathescape=false]
-          if ( multifab_remote(s(n), i) ) cycle
           sp  => dataptr(s(n), i)
@@
 \end{lstlisting}
-
-\noindent For parallel runs, there is no guarantee that the particular
-box is on the current processor.  The function {\tt multifab\_remote}
-returns {\tt .true.} if the box is off-processor, in which case, we 
-simply skip to the next box.
 

### 2012-09-13 SCIENCE/flame/1d/macproject.f90
changed the_bc_tower class to use local indices for boundary conditions instead of global indices.  Changed the rest of the source code to refer to the local indices.

@@ -808,7 +808,7 @@ contains
       integer       , intent(in   ) :: press_comp
       integer       , intent(in   ) :: ref_ratio(:,:)
 
-      integer :: i,gid
+      integer :: i
       integer :: ng_um,ng_p,ng_b
 
       type(bc_level)           :: bc
@@ -833,7 +833,6 @@ contains
       do n = 1, nlevs
          bc = the_bc_tower%bc_tower_array(n)
          do i = 1, nfabs(rh(n))
-            gid =  global_index(umac(n,1),i)
             ump => dataptr(umac(n,1), i)
             vmp => dataptr(umac(n,2), i)
             php => dataptr( phi(n), i)
@@ -850,12 +849,12 @@ contains
                                  php(:,:,1,1),ng_p, &
                                  bxp(:,:,1,1),byp(:,:,1,1),ng_b, &
                                  lxp(:,:,1,1),hxp(:,:,1,1),lyp(:,:,1,1),hyp(:,:,1,1), &
-                                 dx(n,:),bc%ell_bc_level_array(gid,:,:,press_comp))
+                                 dx(n,:),bc%ell_bc_level_array(i,:,:,press_comp))
                else 
                   call mkumac_2d_base(ump(:,:,1,1),vmp(:,:,1,1), ng_um, & 
                                       php(:,:,1,1), ng_p, &
                                       bxp(:,:,1,1), byp(:,:,1,1), ng_b, &
-                                      dx(n,:),bc%ell_bc_level_array(gid,:,:,press_comp))
+                                      dx(n,:),bc%ell_bc_level_array(i,:,:,press_comp))
                end if
             case (3)
                wmp => dataptr(umac(n,3), i)
@@ -872,12 +871,12 @@ contains
                                  bxp(:,:,:,1), byp(:,:,:,1), bzp(:,:,:,1), ng_b, &
                                  lxp(:,:,:,1),hxp(:,:,:,1),lyp(:,:,:,1),hyp(:,:,:,1), &
                                  lzp(:,:,:,1),hzp(:,:,:,1),dx(n,:),&
-                                 bc%ell_bc_level_array(gid,:,:,press_comp))
+                                 bc%ell_bc_level_array(i,:,:,press_comp))
                else
                   call mkumac_3d_base(ump(:,:,:,1),vmp(:,:,:,1),wmp(:,:,:,1),ng_um,&
                                       php(:,:,:,1), ng_p, &
[…]

### 2012-09-13 Source/enforce_outflow_on_divu_rhs.f90

@@ -19,7 +19,7 @@ contains
       type(multifab) , intent(inout) :: divu_rhs(:)
       type(bc_tower) , intent(in   ) :: the_bc_tower
 
-      integer        :: i,n,ng_d,dm,nlevs,gid
+      integer        :: i,n,ng_d,dm,nlevs
       integer        :: lo(get_dim(divu_rhs(1))),hi(get_dim(divu_rhs(1)))
       type(bc_level) :: bc
       real(kind=dp_t), pointer :: divp(:,:,:,:) 
@@ -32,20 +32,19 @@ contains
       do n = 1, nlevs
          bc = the_bc_tower%bc_tower_array(n)
          do i = 1, nfabs(divu_rhs(n))
-            gid =   global_index(divu_rhs(n),i)
             divp => dataptr(divu_rhs(n) , i)
             lo =    lwb(get_box(divu_rhs(n),i))
             hi =    upb(get_box(divu_rhs(n),i))
             select case (dm)
             case (1)
                call enforce_outflow_1d(divp(:,1,1,1), lo, hi, ng_d, &
-                                       bc%phys_bc_level_array(gid,:,:))
+                                       bc%phys_bc_level_array(i,:,:))
             case (2)
                call enforce_outflow_2d(divp(:,:,1,1), lo, hi, ng_d, &
-                                       bc%phys_bc_level_array(gid,:,:))
+                                       bc%phys_bc_level_array(i,:,:))
             case (3)
                call enforce_outflow_3d(divp(:,:,:,1), lo, hi, ng_d, &
-                                       bc%phys_bc_level_array(gid,:,:))
+                                       bc%phys_bc_level_array(i,:,:))
             end select
          end do
       end do

### 2012-09-13 Source/hgproject.f90

@@ -297,7 +297,7 @@ contains
       real(kind=dp_t), pointer ::  rp(:,:,:,:)
   
       integer :: lo(unew(1)%dim),hi(unew(1)%dim)
-      integer :: i,n,gid,ng_un,ng_uo,ng_rh,ng_gp
+      integer :: i,n,ng_un,ng_uo,ng_rh,ng_gp
 
       type(bl_prof_timer), save :: bpt
 
@@ -311,7 +311,6 @@ contains
       do n = 1, nlevs
          bc = the_bc_tower%bc_tower_array(n)
          do i = 1, nfabs(unew(n))
-            gid =  global_index(unew(n), i)
             lo  =  lwb(get_box(unew(n) , i))
             hi  =  upb(get_box(unew(n) , i))
             unp => dataptr(unew(n)     , i) 
@@ -322,15 +321,15 @@ contains
                case (1)
                  call create_uvec_1d(unp(:,1,1,1), ng_un, uop(:,1,1,1), ng_uo, &
                                       rp(:,1,1,1), ng_rh, gpp(:,1,1,1), ng_gp, &
-                                     lo,hi,dt,bc%phys_bc_level_array(gid,:,:), proj_type)
+                                     lo,hi,dt,bc%phys_bc_level_array(i,:,:), proj_type)
                case (2)
                  call create_uvec_2d(unp(:,:,1,:), ng_un, uop(:,:,1,:), ng_uo, &
                                       rp(:,:,1,1), ng_rh, gpp(:,:,1,:), ng_gp, &
-                                     lo,hi,dt,bc%phys_bc_level_array(gid,:,:), proj_type)
+                                     lo,hi,dt,bc%phys_bc_level_array(i,:,:), proj_type)
                case (3)
                  call create_uvec_3d(unp(:,:,:,:), ng_un, uop(:,:,:,:), ng_uo, &
                                       rp(:,:,:,1), ng_rh, gpp(:,:,:,:), ng_gp, &
-                                     lo,hi,dt, bc%phys_bc_level_array(gid,:,:), proj_type)
+                                     lo,hi,dt, bc%phys_bc_level_array(i,:,:), proj_type)
             end select
          end do
          call multifab_fill_boundary(unew(n))

### 2012-09-13 Source/macproject.f90

@@ -749,7 +749,7 @@ contains
       real(dp_t)    , intent(in   ) :: dx(:,:)
       type(bc_tower), intent(in   ) :: the_bc_tower
 
-      integer :: i,gid,ng_um,ng_p,ng_b,lo(get_dim(rh(1))),hi(get_dim(rh(1))),dm
+      integer :: i,ng_um,ng_p,ng_b,lo(get_dim(rh(1))),hi(get_dim(rh(1))),dm
 
       type(bc_level)           :: bc
       real(kind=dp_t), pointer :: ump(:,:,:,:) 
@@ -776,7 +776,6 @@ contains
       do n = 1, nlevs
          bc = the_bc_tower%bc_tower_array(n)
          do i = 1, nfabs(rh(n))
-            gid =  global_index(umac(n,1),i)
             ump => dataptr(umac(n,1), i)
             php => dataptr( phi(n), i)
             bxp => dataptr(beta(n,1), i)
@@ -787,7 +786,7 @@ contains
                call mkumac_1d(n,ump(:,1,1,1), ng_um, & 
                               php(:,1,1,1), ng_p, &
                               bxp(:,1,1,1), ng_b, &
-                              lo,hi,dx(n,:),bc%ell_bc_level_array(gid,:,:,press_comp))
+                              lo,hi,dx(n,:),bc%ell_bc_level_array(i,:,:,press_comp))
                if (n > 1) then
                   lxp => dataptr(fine_flx(n)%bmf(1,0), i)
                   hxp => dataptr(fine_flx(n)%bmf(1,1), i)
@@ -800,7 +799,7 @@ contains
                call mkumac_2d(n,ump(:,:,1,1),vmp(:,:,1,1), ng_um, & 
                               php(:,:,1,1), ng_p, &
                               bxp(:,:,1,1), byp(:,:,1,1), ng_b, &
-                              lo,hi,dx(n,:),bc%ell_bc_level_array(gid,:,:,press_comp))
+                              lo,hi,dx(n,:),bc%ell_bc_level_array(i,:,:,press_comp))
                if (n > 1) then
                   lxp => dataptr(fine_flx(n)%bmf(1,0), i)
                   hxp => dataptr(fine_flx(n)%bmf(1,1), i)
@@ -818,7 +817,7 @@ contains
                call mkumac_3d(n,ump(:,:,:,1),vmp(:,:,:,1),wmp(:,:,:,1),ng_um,&
[…]

### 2012-09-13 Source/make_edge_scal.f90

@@ -44,7 +44,7 @@ contains
 
     integer                  :: i,scomp,bccomp,n,dm,nlevs
     integer                  :: lo(mla%dim), hi(mla%dim)
-    integer                  :: ng_s,ng_se,ng_um,ng_f,gid
+    integer                  :: ng_s,ng_se,ng_um,ng_f
     real(kind=dp_t), pointer :: sop(:,:,:,:)
     real(kind=dp_t), pointer :: sepx(:,:,:,:)
     real(kind=dp_t), pointer :: sepy(:,:,:,:)
@@ -68,7 +68,6 @@ contains
 
     do n=1,nlevs
        do i = 1, nfabs(s(n))
-          gid  =  global_index(s(n),i)
           sop  => dataptr(s(n),i)
           sepx => dataptr(sedge(n,1),i)
           ump  => dataptr(umac(n,1),i)
@@ -84,7 +83,7 @@ contains
                                         ump(:,1,1,1), ng_um, &
                                        fp(:,1,1,:), ng_f, &
                                        lo, hi, dx(n,:), dt, is_vel, &
-                                       the_bc_level(n)%adv_bc_level_array(gid,:,:,:), &
+                                       the_bc_level(n)%adv_bc_level_array(i,:,:,:), &
                                        scomp, bccomp, is_conservative)
              end do
 
@@ -98,7 +97,7 @@ contains
                                        ump(:,:,1,1), vmp(:,:,1,1), ng_um, &
                                        fp(:,:,1,:), ng_f, &
                                        lo, hi, dx(n,:), dt, is_vel, &
-                                       the_bc_level(n)%adv_bc_level_array(gid,:,:,:), &
+                                       the_bc_level(n)%adv_bc_level_array(i,:,:,:), &
                                        scomp, bccomp, is_conservative)
              end do
 
@@ -114,7 +113,7 @@ contains
                                        ng_se, ump(:,:,:,1), vmp(:,:,:,1), wmp(:,:,:,1), &
                                        ng_um, fp(:,:,:,:), ng_f,  &
                                        lo, hi, dx(n,:), dt, is_vel, &
-                                       the_bc_level(n)%adv_bc_level_array(gid,:,:,:), &
+                                       the_bc_level(n)%adv_bc_level_array(i,:,:,:), &
                                        scomp, bccomp, is_conservative)
              end do
           end select

### 2012-09-13 Source/make_plot_variables.f90

@@ -2161,7 +2161,7 @@ contains
     real(kind=dp_t), pointer:: up(:,:,:,:)
     real(kind=dp_t), pointer:: vp(:,:,:,:)
     integer :: lo(get_dim(vort)),hi(get_dim(vort))
-    integer :: i,ng_u,ng_v,dm,gid
+    integer :: i,ng_u,ng_v,dm
 
     type(bl_prof_timer), save :: bpt
 
@@ -2173,7 +2173,6 @@ contains
     ng_v = nghost(vort)
 
     do i = 1, nfabs(u)
-       gid =  global_index(u,i)
        up  => dataptr(u, i)
        vp  => dataptr(vort, i)
        lo  =  lwb(get_box(u, i))
@@ -2183,10 +2182,10 @@ contains
           call setval(vort,0.d0,comp,1,all=.true.)
        case (2)
           call makevort_2d(vp(:,:,1,comp),ng_v,up(:,:,1,:),ng_u,lo,hi,dx, &
-                           bc%phys_bc_level_array(gid,:,:))
+                           bc%phys_bc_level_array(i,:,:))
        case (3)
           call makevort_3d(vp(:,:,:,comp),ng_v,up(:,:,:,:),ng_u,lo,hi,dx, &
-                           bc%phys_bc_level_array(gid,:,:))
+                           bc%phys_bc_level_array(i,:,:))
        end select
     end do
 

### 2012-09-13 Source/mkutrans.f90

@@ -41,7 +41,7 @@ contains
     real(kind=dp_t), pointer :: w0yp(:,:,:,:)
     real(kind=dp_t), pointer :: w0zp(:,:,:,:)
     integer                  :: lo(mla%dim),hi(mla%dim),dm,nlevs
-    integer                  :: i,n,n_1d,ng_u,ng_uf,ng_ut,ng_w0,gid
+    integer                  :: i,n,n_1d,ng_u,ng_uf,ng_ut,ng_w0
 
     type(bl_prof_timer), save :: bpt
 
@@ -58,7 +58,6 @@ contains
     do n=1,nlevs
 
        do i=1, nfabs(u(n))
-          gid =  global_index(u(n),i)
           up  => dataptr(u(n),i)
           ufp => dataptr(ufull(n),i)
           utp => dataptr(utrans(n,1),i)
@@ -70,16 +69,16 @@ contains
                               ufp(:,1,1,:), ng_uf, &
                               utp(:,1,1,1), ng_ut, w0(n,:), &
                               lo,hi,dx(n,:),dt,&
-                              the_bc_level(n)%adv_bc_level_array(gid,:,:,:), &
-                              the_bc_level(n)%phys_bc_level_array(gid,:,:))
+                              the_bc_level(n)%adv_bc_level_array(i,:,:,:), &
+                              the_bc_level(n)%phys_bc_level_array(i,:,:))
           case (2)
              vtp => dataptr(utrans(n,2),i)
              call mkutrans_2d(up(:,:,1,:), ng_u, &
                               ufp(:,:,1,:), ng_uf, &
                               utp(:,:,1,1), vtp(:,:,1,1), ng_ut, w0(n,:), &
                               lo,hi,dx(n,:),dt,&
-                              the_bc_level(n)%adv_bc_level_array(gid,:,:,:), &
-                              the_bc_level(n)%phys_bc_level_array(gid,:,:))
+                              the_bc_level(n)%adv_bc_level_array(i,:,:,:), &
+                              the_bc_level(n)%phys_bc_level_array(i,:,:))
           case (3)
              vtp => dataptr(utrans(n,2),i)
              wtp => dataptr(utrans(n,3), i)
@@ -96,8 +95,8 @@ contains
                               utp(:,:,:,1), vtp(:,:,:,1), wtp(:,:,:,1), ng_ut, &
                               w0(n_1d,:), w0xp(:,:,:,1), w0yp(:,:,:,1), w0zp(:,:,:,1),&
[…]

### 2012-09-13 Source/multifab_physbc.f90

@@ -23,7 +23,7 @@ contains
 
     ! Local
     integer                  :: lo(get_dim(s)),hi(get_dim(s)),dm
-    integer                  :: i,ng,scomp,bccomp,gid
+    integer                  :: i,ng,scomp,bccomp
     real(kind=dp_t), pointer :: sp(:,:,:,:)
 
     type(bl_prof_timer), save :: bpt
@@ -37,7 +37,6 @@ contains
     call build(bpt, "multifab_physbc")
     
     do i=1,nfabs(s)
-       gid =  global_index(s,i)
        sp  => dataptr(s,i)
        lo  =  lwb(get_box(s,i))
        hi  =  upb(get_box(s,i))
@@ -46,19 +45,19 @@ contains
           do scomp = start_scomp,start_scomp+num_comp-1
              bccomp = start_bccomp + scomp - start_scomp
              call physbc_1d(sp(:,1,1,scomp), lo, hi, ng, &
-                           the_bc_level%adv_bc_level_array(gid,:,:,bccomp),bccomp)
+                           the_bc_level%adv_bc_level_array(i,:,:,bccomp),bccomp)
           end do
        case (2)
           do scomp = start_scomp,start_scomp+num_comp-1
              bccomp = start_bccomp + scomp - start_scomp
              call physbc_2d(sp(:,:,1,scomp), lo, hi, ng, &
-                           the_bc_level%adv_bc_level_array(gid,:,:,bccomp),bccomp)
+                           the_bc_level%adv_bc_level_array(i,:,:,bccomp),bccomp)
           end do
        case (3)
           do scomp = start_scomp,start_scomp+num_comp-1
              bccomp = start_bccomp + scomp - start_scomp
              call physbc_3d(sp(:,:,:,scomp), lo, hi, ng, &
-                           the_bc_level%adv_bc_level_array(gid,:,:,bccomp),bccomp)
+                           the_bc_level%adv_bc_level_array(i,:,:,bccomp),bccomp)
           end do
        end select
     end do

### 2012-09-13 Source/velpred.f90

@@ -32,7 +32,7 @@ contains
     type(bc_level) , intent(in   ) :: the_bc_level(:)
     type(ml_layout), intent(in   ) :: mla
 
-    integer                  :: i,n,n_1d,gid
+    integer                  :: i,n,n_1d
     integer                  :: ng_u,ng_uf,ng_um,ng_ut,ng_f,ng_w0
     integer                  :: lo(mla%dim), hi(mla%dim),dm,nlevs
     real(kind=dp_t), pointer :: uop(:,:,:,:)
@@ -64,7 +64,6 @@ contains
 
     do n=1,nlevs
        do i = 1, nfabs(u(n))
-          gid  =  global_index(u(n),i)
           uop  => dataptr(u(n),i)
           ufp  => dataptr(ufull(n),i)
           ump  => dataptr(umac(n,1),i)
@@ -78,8 +77,8 @@ contains
                              ufp(:,1,1,:), ng_uf, &
                              ump(:,1,1,1), ng_um, &
                              fp(:,1,1,1), ng_f, w0(n,:), lo, hi, dx(n,:), dt, &
-                             the_bc_level(n)%phys_bc_level_array(gid,:,:), &
-                             the_bc_level(n)%adv_bc_level_array(gid,:,:,:))
+                             the_bc_level(n)%phys_bc_level_array(i,:,:), &
+                             the_bc_level(n)%adv_bc_level_array(i,:,:,:))
           case (2)
              vtp  => dataptr(utrans(n,2),i)
              vmp  => dataptr(  umac(n,2),i)
@@ -88,8 +87,8 @@ contains
                              utp(:,:,1,1), vtp(:,:,1,1), ng_ut, &
                              ump(:,:,1,1), vmp(:,:,1,1), ng_um, &
                              fp(:,:,1,:), ng_f, w0(n,:), lo, hi, dx(n,:), dt, &
-                             the_bc_level(n)%phys_bc_level_array(gid,:,:), &
-                             the_bc_level(n)%adv_bc_level_array(gid,:,:,:))
+                             the_bc_level(n)%phys_bc_level_array(i,:,:), &
+                             the_bc_level(n)%adv_bc_level_array(i,:,:,:))
 
           case (3)
              vmp  => dataptr(  umac(n,2),i)
@@ -111,8 +110,8 @@ contains
                              fp(:,:,:,:), ng_f, &
                              w0(n_1d,:),w0xp(:,:,:,1),w0yp(:,:,:,1),w0zp(:,:,:,1), &
[…]

### 2012-09-13 TEST_PROBLEMS/stand_prob/define_bc_tower.f90

@@ -1,7 +1,7 @@
 module define_bc_module
 
   use box_module, only: box
-  use layout_module, only: layout, layout_get_pd, layout_get_box, layout_dim, layout_nboxes
+  use layout_module, only: layout, layout_get_pd, layout_get_box, layout_dim, layout_nlocal
   use bc_module, only: EXT_DIR, FOEXTRAP, HOEXTRAP, REFLECT_EVEN, REFLECT_ODD, BC_DIR, &
                        BC_NEU, BC_PER, BC_INT, INTERIOR, &
                        INLET, NO_SLIP_WALL, OUTLET, PERIODIC, SLIP_WALL, SYMMETRY
@@ -64,7 +64,7 @@ contains
       bct%bc_tower_array(n)%ell_bc_level_array => NULL()
     end if
 
-    ngrids = layout_nboxes(la)
+    ngrids = layout_nlocal(la)
     dm = layout_dim(la)
 
     allocate(bct%bc_tower_array(n)%phys_bc_level_array(0:ngrids,dm,2))
@@ -120,7 +120,7 @@ contains
 
     phys_bc_level = default_value
 
-    ! phys_bc_level(0:nboxes,dim,2) contains the physical name of the
+    ! phys_bc_level(0:nlocal,dim,2) contains the physical name of the
     ! boundary condition type on each edge of each box that makes up
     ! the current level.  The phys_bc_level(0,:,:) 'box' refers to the
     ! entire domain.  If an edge of a box is not on a physical
@@ -136,8 +136,8 @@ contains
        phys_bc_level(i,d,2) = domain_bc(d,2)
     end do
 
-    do i = 1,layout_nboxes(la_level)
-       bx = layout_get_box(la_level,i)
+    do i = 1,layout_nlocal(la_level)
+       bx = layout_get_box(la_level,global_index(la_level,i))
        do d = 1,layout_dim(la_level)
           if (bx%lo(d) == pd%lo(d)) phys_bc_level(i,d,1) = domain_bc(d,1)
           if (bx%hi(d) == pd%hi(d)) phys_bc_level(i,d,2) = domain_bc(d,2)

### 2012-09-13 UNIT_TESTS/test_eos/define_bc_tower.f90

@@ -73,7 +73,7 @@ module define_bc_module
     end if
 
     ncomp = 1
-    ngrids = layout_nboxes(la)
+    ngrids = layout_nlocal(la)
     dm = layout_dim(la)
 
     allocate(bct%bc_tower_array(n)%phys_bc_level_array(0:ngrids,dm,2))
@@ -118,8 +118,8 @@ module define_bc_module
     end do
 
     ! loop over individual grids
-    do i = 1,layout_nboxes(la_level)    ! loop over grids
-       bx = layout_get_box(la_level,i)  ! grab box associated with the grid
+    do i = 1,layout_nlocal(la_level) ! loop over grids
+       bx = layout_get_box(la_level,global_index(la_level,i)) ! grab box associated with the grid
        do d = 1,layout_dim(la_level)    ! loop over directions
           ! if one side of a grid is a domain boundary, set the 
           ! physical boundary condition

### 2012-09-13 docs/MESA/instructions.tex

@@
 {\sf MESA} modules to an outside program such as {\sf MAESTRO}.
+\footnote{This
+  interface work was done by Ryan Orvedahl---questions (and credit)
+  should be directed to him}.
 
@@
 \section{Setting Up {\sf MESA}}
+\label{sec:setupmesa}
 
@@
 {\tt http://mesa.sourceforge.net}. Check the website for the latest version, 
-which is 4219 with the writing of this document. The codes and algorithms 
+which is 4298 with the writing of this document. The codes and algorithms 
 discussed here were created when the current version of {\sf MESA} was 
-version 4088. Between versions 4088 and 4219, there are minor changes that 
+version 4088. Between versions 4088 and 4298, there are minor changes that 
 will cause some of the included routines to crash. To download the source code 
@@
 
-\item {\tt PGPLOT}: Set the variable {\tt USE\_PGPLOT} to {\tt NO} and set 
+\item {\tt PGPLOT}: Set the variable {\tt USE\_PGSTAR} to {\tt NO} and set 
 the variable {\tt LOAD\_PGPLOT} to be empty on the line corresponding to 
@@
 
-Following the execution of the {\tt ./install} script, {\sf MESA} has been 
-installed and the following has completed:
+Following a successful execution of the {\tt ./install} script, 
+{\sf MESA} has been installed and the following has completed:
 \begin{itemize}
-\item SDK only: Supporting software (compilers, linear algebra packages, 
-etc.) has been unpacked and installed
 \item Data tables have been untarred
@@
 \item All modules have been compiled (in parallel)
-\item All modules have been exported to the {\tt mesa/lib} and 
-{\tt mesa/include} directories
+\item All modules and libraries have been exported to the {\tt mesa/include} 
+and {\tt mesa/lib} directories
 \end{itemize}
@@
 Change the {\tt FCopenmp} flag ({\tt gfortran} default = {\tt -fopenmp}) to 
-be empty. This will disable OpenMP for all modules.
+the appropriate value so as to disable OpenMP, e.g. with {\tt gfortran} 
+{\tt FCopenmp} should be empty.
 
@@
 to load, e.g. {\tt -L\$(MESA\_DIR)/lib -leos -lnet -lnum}. {\sf MESA} 
-{\tt .mod} files are linked to code using the {\tt -I} flag and the 
+{\tt .mod} files are linked to the code using the {\tt -I} flag and the 
 {\tt mesa/include} directory, e.g. {\tt -I\$(MESA\_DIR)/include}.
@@
 {\tt net\_input.net} file, which does not need to be kept in the
-{\tt Net-Files} directory. The {\tt net\_input.net} file could look like this:
+{\tt Net-Files} directory. The {\tt net\_input.net} file should have the 
+following format:
 \begin{lstlisting}[language=make,mathescape=false]
@@
 declared and set equal to `mesa'. {\tt use\_cache} is a logical variable that 
-also needs to be declared and set equal to `true'
+also needs to be declared and set equal to {\tt .true.}
 
@@
 {\sf MESA} isotope indices and the mass fraction array indices ({\tt chem\_id} 
-and {\tt net\_iso}, see Section \ref{sec:mapping}). The HELM EOS returns on 
-the order of 400 different quantities which include the values shown in 
+and {\tt net\_iso}, see Section \ref{sec:mapping}). The HELM EOS returns 409 
+different quantities which include the values shown in 
 Table \ref{tab:res} as well as quantities such as $dP/d(abar)$ and 
@@
 integer pointer arrays of size {\tt num\_chem\_isos}, which is defined in one 
-of the chemistry modules. {\tt chem\_id} maps from an isotope in the current 
-net to the {\sf MESA} data table that holds the properties of that isotope. 
-{\tt net\_iso} is used to map from the {\sf MESA} data tables to an isotope 
-in the current network.
+of the chemistry modules. {\tt chem\_id} maps from the index of an isotope in 
+the current net to the {\sf MESA} data table position that holds the 
+properties of that isotope. {\tt net\_iso} is used to map from the {\sf MESA} 
+data table position to the index of an isotope in the current network.
 
@@
 \item Call the {\tt net\_start\_def}, {\tt read\_net\_file}, {\tt net\_ptr} 
-and {\tt net\_finish\_def}
+and {\tt net\_finish\_def} subroutines
 \item Output a log file containing the species in the current net and which 
@@
 
-After calling {\tt net\_set\_which\_rates}, the setup routine should call 
-{\tt net\_setup\_tables}. This is based on the example 
-{\tt one\_zone\_burn.f} program distributed with the {\sf MESA} source code. 
-I have not had a successful burn if the {\tt net\_setup\_tables} routine is 
-called at setup in the {\tt setup\_mesa\_net.f90} program. I do have 
-successful burns when it is placed within the burner program 
-{\tt Do\_One\_Zone\_Burn.f90}. To prevent the setup routine from being called 
-everytime a burn is to take place, I added a logical flag that signals if the 
-{\tt net\_setup\_tables} routine has alredy been called.
+In the {\tt Do\_One\_Zone\_Burn.f90} routine, there are two more ``set up'' 
+routines: {\tt set\_rattab\_range} and {\tt net\_setup\_tables}. I have not 
+found a good way to call these routines only once with the set up of all the 
+other routines because the {\tt set\_rattab\_range} routine depends on 
+temperature. As a result, both routines are called every time a burn is to 
+take place.
 

### 2012-10-02 docs/FAQ/faq.tex
add Warren's question

@@
 \begin{enumerate}
+
+\item {\em Why is everything in its own module?}
+
+   If a subroutine is in a Fortran module, then at compile time, 
+   there is argument checking.  This ensures that the right number
+   and data types of arguments are present.  It makes the code safer.
 
@@
 
+\item {\em Why did we switch all the equations over to the
+  $\tilde{\Ub}$ form instead of just working with $\Ub$?}
+
+This is basically a numerical discretization issue.  Whenever the base
+state aligns with the grid, you should be able to show that you get
+exactly the same answer each way.
+
+When you do a spherical star on a 3d Cartesian grid, though, the $w_0$
+is defined on the radial mesh and the $\tilde{\Ub}$ on the Cartesian
+mesh, and the $w_0$ part never experiences the Cartesian projection,
+for example.  So there are differences in exactly how the $w_0$ component
+appears (projected on the Cartesian mesh vs.\ interpolated from the
+radial mesh)---we made the decision at the time to separate the
+components for that reason.
+
 \end{enumerate}

### 2012-10-02 docs/runtime_parameters/runtime_parameters.tex
add a chapter label.  Update the parameter list
update parameters, remove redundancy in intro

@@
-\section{Runtime Parameters}
-
-The following tables list the runtime parameters available to
-all MAESTRO problems.  Problem-specific runtime parameters are not
-shown here.
+\label{ch:parameters}
 
@@
 \rowcolor{tableShade}
-\verb=  use_eos_coulomb  = &   use the Coulomb corrections between ions (if supported by the EOS)  &  .true. \\
 \verb=  use_eos_e_instead_of_h  = &   In deriving the temperature from the $h$, first subtract off $p_0/\rho$ to define $e$, and use that as the input to the EOS.  &  .false. \\
@@
 \rowcolor{tableShade}
-\verb=  sdc_iters  = &   how many SDC iteractions to take (requires the code be compiled with {\tt SDC := t}  &  4 \\
+\verb=  sdc_couple_mac_velocity  = &   recompute MAC velocity at the beginning of each SDC iter if .false., only recompute MAC velocity at the beginning of the first iter  &  .false. \\
+\verb=  sdc_hold_mac_velocity  = &   if true, never recompute MAC velocity, no matter what the other sdc flags are  &  .false. \\
+\rowcolor{tableShade}
+\verb=  sdc_iters  = &   how many SDC iteractions to take (requires the code be compiled with {\tt SDC := t}  &  1 \\
 
@@
 \rowcolor{tableShade}
+\verb=  do_2d_planar_octant  = &   Set to 1 if using the 2D simplified (planar) model of an octant.  &  0 \\
 \verb=  drdxfac  = &   ratio of radial base state zones to Cartesian full state zones for spherical geometry  &  1 \\
+\rowcolor{tableShade}
 \verb=  dump_grid_file  = &   dump out a file, named grids.out, containing the number of grids a each level. A new line is added each time regrid is called.  &  .false. \\
+\verb=  max_grid_size  = &   The largest grid size that will be created using make\_new\_grids.  &  64 \\
 \rowcolor{tableShade}
-\verb=  max_grid_size  = &   The largest grid size that will be created using make\_new\_grids.  &  64 \\
 \verb=  max_grid_size_1  = &   The largest grid size that will be created using make\_new\_grids for the coarsest level.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
+\verb=  max_grid_size_2  = &   The largest grid size that will be created using make\_new\_grids for level 2 (the first refined level).  Defaults to {\tt max\_grid\_size}.  &  -1 \\
 \rowcolor{tableShade}
-\verb=  max_grid_size_2  = &   The largest grid size that will be created using make\_new\_grids for level 2 (the first refined level).  Defaults to {\tt max\_grid\_size}.  &  -1 \\
 \verb=  max_grid_size_3  = &   The largest grid size that will be created using make\_new\_grids for level 3 (the second refined level) and beyond.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
+\verb=  max_levs  = &   Total number of levels.  1 = single level.  &  1 \\
 \rowcolor{tableShade}
-\verb=  max_levs  = &   Total number of levels.  1 = single level.  &  1 \\
 \verb=  min_eff  = &   parameter for cluster algorithm for making new grids in adaptive problems  &  0.9d0 \\
+\verb=  minwidth  = &   The minimum size on a side for a grid created using make\_new\_grids.  &  8 \\
 \rowcolor{tableShade}
-\verb=  minwidth  = &   The minimum size on a side for a grid created using make\_new\_grids.  &  8 \\
 \verb=  n_cellx  = &   Number of cells for the base level in the x-direction  &  -1 \\
+\verb=  n_celly  = &   Number of cells for the base level in the y-direction  &  -1 \\
 \rowcolor{tableShade}
-\verb=  n_celly  = &   Number of cells for the base level in the y-direction  &  -1 \\
 \verb=  n_cellz  = &   Number of cells for the base level in the z-direction  &  -1 \\
+\verb=  octant  = &   set octant = T if you just want to model an octant of a sphere (note: only takes effect for spherical geometry)  &  .false. \\
 \rowcolor{tableShade}
-\verb=  octant  = &   set octant = T if you just want to model an octant of a sphere (note: only takes effect for spherical geometry)  &  .false. \\
 \verb=  prob_hi_x  = &   physical coordinates of hi-x corner of problem domain  &  1.d0 \\
+\verb=  prob_hi_y  = &   physical coordinates of hi-y corner of problem domain  &  1.d0 \\
 \rowcolor{tableShade}
-\verb=  prob_hi_y  = &   physical coordinates of hi-y corner of problem domain  &  1.d0 \\
 \verb=  prob_hi_z  = &   physical coordinates of hi-z corner of problem domain  &  1.d0 \\
+\verb=  prob_lo_x  = &   physical coordinates of lo-x corner of problem domain  &  ZERO \\
 \rowcolor{tableShade}
-\verb=  prob_lo_x  = &   physical coordinates of lo-x corner of problem domain  &  ZERO \\
 \verb=  prob_lo_y  = &   physical coordinates of lo-y corner of problem domain  &  ZERO \\
+\verb=  prob_lo_z  = &   physical coordinates of lo-z corner of problem domain  &  ZERO \\
 \rowcolor{tableShade}
-\verb=  prob_lo_z  = &   physical coordinates of lo-z corner of problem domain  &  ZERO \\
 \verb=  ref_ratio  = &   Refinement ratio for multilevel problems  &  2 \\
+\verb=  regrid_int  = &   How often we regrid.  &  -1 \\
 \rowcolor{tableShade}
-\verb=  regrid_int  = &   How often we regrid.  &  -1 \\
-\verb=  spherical_in  = &   set to 1 if you are doing a spherical problem  &  0 \\
+\verb=  spherical_in  = &   Set to 1 if you are doing a spherical problem.  &  0 \\
+\verb=  test_set  = &   Fixed grid file.  &  '' \\
 \rowcolor{tableShade}
-\verb=  test_set  = &   Fixed grid file.  &  '' \\
 \verb=  the_copy_cache_max  = &    Number of boxassoc layouts we keep in memory to avoid having to recompute the boxassoc, which is computationally expensive.  &  25 \\
+\verb=  the_layout_verbosity  = &    &  0 \\
 \rowcolor{tableShade}
-\verb=  the_layout_verbosity  = &    &  0 \\
 \verb=  the_sfc_threshold  = &   When assigning processors for grids, this determines whether we use the sfc algorithm or knapsack algorithm.
@@
 &  5 \\
+\verb=  xhi_boundary_type  = &   $+x$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcx\_hi}  &  "" \\
 \rowcolor{tableShade}
-\verb=  xhi_boundary_type  = &   $+x$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcx\_hi}  &  "" \\
 \verb=  xlo_boundary_type  = &   $-x$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcx\_lo}  &  "" \\
+\verb=  yhi_boundary_type  = &   $+y$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcy\_hi}  &  "" \\
 \rowcolor{tableShade}
-\verb=  yhi_boundary_type  = &   $+y$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcy\_hi}  &  "" \\
 \verb=  ylo_boundary_type  = &   $-y$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcy\_lo}  &  "" \\
+\verb=  zhi_boundary_type  = &   $+z$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcz\_hi}  &  "" \\
 \rowcolor{tableShade}
-\verb=  zhi_boundary_type  = &   $+z$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcz\_hi}  &  "" \\
 \verb=  zlo_boundary_type  = &   $-z$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcz\_lo}  &  "" \\
@@
 \rowcolor{tableShade}
-\verb=  enthalpy_pred_type  = &   predict\_rhohprime = 1; \newline predict\_h = 2; \newline predict\_T\_then\_rhohprime = 3; \newline predict\_T\_then\_h = 4; \newline predict\_hprime = 5; \newline predict\_Tprime\_then\_h = 6.
+\verb=  enthalpy_pred_type  = &   predict\_rhoh = 0; \newline predict\_rhohprime = 1; \newline predict\_h = 2; \newline predict\_T\_then\_rhohprime = 3; \newline predict\_T\_then\_h = 4; \newline predict\_hprime = 5; \newline predict\_Tprime\_then\_h = 6.
 &  predict\_rhohprime \\
@@
 \rowcolor{tableShade}
+\verb=  use_alt_energy_fix  = &   modify the momentum equation to have $(\beta_0/\rho) \nabla (\pi/\beta_0)$ instead of just $(1/\rho) \nabla (\pi)$  &  .false. \\
 \verb=  use_delta_gamma1_term  = &   turns on second order correction to delta gamma1 term  &  .false. \\
+\rowcolor{tableShade}
 \verb=  use_etarho  = &   turn on the etarho term as described in flow chart  &  .true. \\
+\verb=  use_linear_grav_in_beta  = &   how to represent gravity in the $\beta_0$ integration: .true. = piecewise linear .false. = piecewise constant  &  .false. \\
 \rowcolor{tableShade}
-\verb=  use_linear_grav_in_beta  = &   how to represent gravity in the beta0 integration: .true. = piecewise linear .false. = piecewise constant  &  .false. \\
 \verb=  use_tfromp  = &   When updating temperature, use $T=T(\rho,p_0,X) $ rather than $T=T(\rho,h,X)$.  &  .false. \\
@@
 \rowcolor{tableShade}
+\verb=  mg_bottom_nu  = &   number of smoothing iterations to do after the multigrid bottom solver  &  10 \\
 \verb=  mg_bottom_solver  = &   valid values are $\le$ 0  &  -1 \\
+\rowcolor{tableShade}
 \verb=  mg_verbose  = &   Verbsoity of the multigrid solver, but not the bottom solver.  &  0 \\
-\rowcolor{tableShade}
 \verb=  use_hypre  = &   use the hypre library  &  .false. \\
@@
 \verb=  plot_omegadot  = &   plot omegadot in plotfile  &  .true. \\
+\verb=  plot_pidivu  = &   plot pi * div(U) -- this is a measure of conservation of energy  &  .false. \\
+\rowcolor{tableShade}
 \verb=  plot_processors  = &   create a field in the plotfile storing the processor number for each zone  &  .false. \\
+\verb=  plot_spec  = &   plot species and omegadot in plotfile  &  .true. \\
 \rowcolor{tableShade}
-\verb=  plot_spec  = &   plot species and omegadot in plotfile  &  .true. \\
 \verb=  plot_trac  = &   plot tracers in plotfile  &  .false. \\
-\rowcolor{tableShade}
 \verb=  single_prec_plotfiles  = &   store the state data in single precision  &  .false. \\

### 2012-10-02 docs/runtime_parameters/rp.py

@@ -5,11 +5,6 @@ import sys
 Mheader=r"""
 \label{ch:parameters}
 
-\section{Runtime Parameters}
-
-The following tables list the runtime parameters available to
-all MAESTRO problems.  Problem-specific runtime parameters are not
-shown here.
 
 %%%%%%%%%%%%%%%%
 % symbol table

### 2012-10-02 docs/visualization/visualization.tex
remove lens stuff -- it is in managing jobs now
add plot_pidivu

--- 2012-02-02 docs/visualization/visualization.tex
@@
 {\tt processor\_number} & processor number containing the cell's data & {\tt plot\_processors} \\
+%\hline
+{\tt pi\_divu}         & $\pi \nabla \cdot \tilde{\Ub}$ (a measure of energy conservation) & {\tt plot\_pidivu} \\
 \end{longtable}

### 2012-10-02 docs/runtime_parameters/runtime_parameters.tex
update

@@
 \verb=  plot_sponge_fdamp  = &   plot fdamp rather than sponge assumes sponge has the form 1/(1+dt*{\tt sponge\_kappa}*fdamp)  &  .false. \\
-\verb=  ppm_type  = &   0 = no ppm \newline 1 = 1985 ppm \newline 2 = Hybrid Sekora/Colella McCorquodale Colella 2009/2010 ppm  &  1 \\
+\verb=  ppm_type  = &   0 = no ppm (piecewise linear slopes instead) \newline 1 = 1984 ppm \newline 2 = Hybrid Sekora/Colella and McCorquodale/Colella 2009/2010 ppm  &  1 \\
 \rowcolor{tableShade}
@@
 \verb=  rotational_frequency  = &   rotational frequency used for computing centrifugal term in rotation problems.  &  ZERO \\
-\verb=  slope_order  = &   order of slopes in Godunov algorithm.  Options are 0, 2, or 4.  &  4 \\
+\verb=  slope_order  = &   order of slopes in piecewise linear Godunov algorithm.  Options are 0, 2, or 4.  &  4 \\
 \rowcolor{tableShade}

### 2012-10-10 docs/architecture/architecture.tex
new directory structure

@@
   contribute to building the {\tt test2} problem in \maestro.  The
+  red directories are part of most standard \maestro\ build.  The
   purple lines show the directories that are pulled in through
-  variable Makefile variables ({\tt BOXLIB\_HOME}, {\tt NETWORK\_DIR},
-  {\tt EOS\_DIR}, and {\tt CONDUCTIVITY\_DIR}).  The blue lines show
-  standard modules that are used in most \maestro\ builds.}
+  various Makefile variables ({\tt BOXLIB\_HOME}, {\tt NETWORK\_DIR},
+  {\tt EOS\_DIR}, and {\tt CONDUCTIVITY\_DIR}).}
 \end{figure}
@@
 
-  \item {\tt constants/}
-
-    A simple module to define physical constants
-
   \item {\tt docs/}
@@
     document).
+
+  \item {\tt Exec/}
+
+    The various problem-setups.
+
+    \begin{itemize}
+    \item {\tt SCIENCE/}
+
+      \maestro\ problem directories for science studies.  These are
+      the setups that have been used for science papers in the past,
+      or are the basis for current science studies.
+
+    \item {\tt TEST\_PROBLEMS/}
+
+      \maestro\ problem directories for simple test problems that have
+      been used while developing \maestro.  Many of these problems
+      have appeared in the various papers describing the
+      \maestro\ algorithm.
+
+    \item {\tt UNIT\_TESTS/}
+
+      Special \maestro\ problem directories that test only a single
+      component of the \maestro\ algorithm.  These often have their
+      own main drivers ({\tt varden.f90}) that setup and initialize
+      some data structures and then call only a few of the
+      \maestro\ routines.  See \S~\ref{sec:unit_tests} for details.
+
+    \end{itemize}
 
@@
 
-  \item {\tt SCIENCE/}
-
-    \maestro\ problem directories for science studies.  These are the
-    setups that have been used for science papers in the past, or are
-    the basis for current science studies.
-
   \item {\tt Source/}
@@
     this source.
-
-  \item {\tt Source\_SDC/}
-
-    Source for an alternate formulation of \maestro\ that removes the
-    splitting between hydrodynamics, diffusion, and reactions (see
-    Chapter~\ref{ch:sdc}).
-
-  \item {\tt TEST\_PROBLEMS/}
-
-    \maestro\ problem directories for simple test problems that have
-    been used while developing \maestro.  Many of these problems have
-    appeared in the various papers describing the \maestro\ algorithm.
-
-  \item {\tt UNIT\_TESTS/}
-
-    Special \maestro\ problem directories that test only a single component
-    of the \maestro\ algorithm.  These often have their own main drivers
-    ({\tt varden.f90}) that setup and initialize some data structures and
-    then call only a few of the \maestro\ routines.  See \S~\ref{sec:unit_tests}
-    for details.
 

### 2012-10-10 docs/architecture/architecture.tex
more updates

@@
 Different \maestro\ problems are defined in sub-directories under {\tt
-  MAESTRO/SCIENCE}, {\tt MAESTRO/TEST\_PROBLEMS}, or {\tt
-  MAESTRO/UNIT\_TESTS}.  To add a problem, start by creating a new
-sub-directory---this is where you will compile your problem and store
-all the problem-specific files.
+  Exec/} in {\tt SCIENCE}, {\tt TEST\_PROBLEMS}, or {\tt UNIT\_TESTS}.
+To add a problem, start by creating a new sub-directory---this is
+where you will compile your problem and store all the problem-specific
+files.
 
@@
 \begin{lstlisting}[language={[gnu]make},mathescape=false]
-  MAESTRO_TOP_DIR := ../..
+  MAESTRO_TOP_DIR := ../../..
 \end{lstlisting}

### 2012-10-10 Docs/MESA/hopper_compile.tex
renname docs Docs

--- 2012-08-07 docs/MESA/hopper_compile.tex
@@
 \begin{enumerate}
+\item Download the {\tt MESA} source code using the command described in 
+Section \ref{sec:setupmesa}
 \item Edit the {\tt makefile\_header} in the {\tt utils} directory
@@
   FCopt = -O 1
-  FCdebug = -g -O0
+  FCdebug = -g -O0 -R bps
   FCstatic =
@@
 and running the test directory. We want to turn off the compiling and 
-running of the test directory to ensure that {\sf MESA} compiles successfully. 
-A portion of the {\tt build\_and\_test} script looks like this:
+running of the test directory to ensure that {\sf MESA} compiles successfully 
+(it failed to install otherwise). A portion of the {\tt build\_and\_test} 
+script looks like this:
 \begin{lstlisting}[language=bash,mathescape=false]
@@
 loop was causing a Segmentation Fault on Hopper with the Cray compiler. The 
-implied loop:
+implied do loop:
 \begin{lstlisting}[language=fortran,mathescape=false]
@@
 variable in the makefile that holds the {\tt *\_def.o} file, e.g. 
-{\tt utils\_def.o} or {\tt num\_def.o}. This files is never loaded into the 
+{\tt utils\_def.o} or {\tt num\_def.o}. This file is never loaded into the 
 library as seen below for the {\tt utils} directory. Certain modules have a 
@@
 \end{lstlisting}
-To include the {\tt LIB\_DEFS} file:
+To include the {\tt LIB\_DEFS} file, change the above code to the following:
 \begin{lstlisting}[language=make,mathescape=false]
@@
 it can use {\sf MESA}'s EOS and/or network. Make sure the makefile variable 
-{\tt COMP} is set to {\tt Cray} so the Cray compiler is used.
+{\tt COMP} is set to {\tt Cray} in the {\sf MAESTRO} problem directory 
+so the Cray compiler is used.

### 2012-10-10 Docs/pert_notes/pert.tex
renname docs Docs

--- 2012-05-30 docs/pert_notes/pert.tex
@@
 underbraced term in Eq.~\ref{rhoprime equation} as the forcing.  This
-force is computed in {\tt modify\_scal\_force}.
+force is computed in {\tt modify\_scal\_force}.  This prediction is
+done in advective form.
 
@@
 just call {\tt make\_edge\_scal} in {\tt density\_advance} on $X_k$.
-The force seen by {\tt make\_edge\_scal} is 0.
+The force seen by {\tt make\_edge\_scal} is 0.  The prediction is
+done in advective form.
 
@@
 %\end{equation}
-Due to Strang-splitting of the reactions,the call to {\tt
+Due to Strang-splitting of the reactions, the call to {\tt
   react\_state} has already been made.  Hence, the goal is to compute
@@
 \frac{\partial(\rho h)}{\partial t} = -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub 
-+ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T. \label{rhoh equation}
++ \underbrace{\psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} + \nabla\cdot\kth\nabla T}_{``(\rho h) ~ \text{force}"} \label{rhoh equation}
 \end{equation}
@@
 \begin{enumerate}
-\item predict $(\rho h)'$, $h$, or $T$ to the edges (depending on {\tt
+\item predict $(rho h)$, $(\rho h)'$, $h$, or $T$ to the edges (depending on {\tt
   enthalpy\_pred\_type} ) using {\tt make\_edge\_scal} and the forces
@@
 \end{enumerate}
-
-\noindent Note that the various {\tt enthalpy\_pred\_type}s have only been
-extensively tested with {\tt species\_pred\_type = 1}.
 
@@
 \hline
+0 / {\tt predict\_rhoh}  $(\rho h)$ & $ \left [\psi + (\Ubt \cdot \er)
+  \frac{\partial p_0}{\partial r} + \nabla \cdot \kth \nabla T \right ]$ \\
 1 / {\tt predict\_rhohprime} $((\rho h)^\prime)$ &  
@@
 \hline 
+1 / {\tt predict\_rhoprime\_and\_X}  & 0 / {\tt predict\_rhoh} &
+  $(\rho h)$ & $(\rho h)_\mathrm{edge}$ & 
+  $X_\mathrm{edge}$, $\rho'_\mathrm{edge}$ & 
+  $(\rho h)_\mathrm{edge}$ \\
 1 / {\tt predict\_rhoprime\_and\_X}  & 1 / {\tt predict\_rhohprime} &
@@
 \hline
+2 / {\tt predict\_rhoX}  & 0 / {\tt predict\_rhoh} &
+  $(\rho h)$ & $(\rho h)_\mathrm{edge}$ & 
+  $(\rho X)_\mathrm{edge}$, $\sum(\rho X)_\mathrm{edge}$ & 
+  $(\rho h)_\mathrm{edge}$ \\
 2 / {\tt predict\_rhoX}  & 1 / {\tt predict\_rhohprime} &
@@
 \hline
+3 / {\tt predict\_rho\_and\_X}  & 0 / {\tt predict\_rhoh} &
+  $(\rho h)$ & $(\rho h)_\mathrm{edge}$ & 
+  $X_\mathrm{edge}$, $\rho_\mathrm{edge}$ & 
+  $(\rho h)_\mathrm{edge}$ \\
 3 / {\tt predict\_rho\_and\_X}  & 1 / {\tt predict\_rhohprime} &
@@
 \end{landscape}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Method 0: {\tt enthalpy\_pred\_type} = {\tt predict\_rhoh}}
+
+Here we wish to construct $(\rho h)_\mathrm{edge}$ by predicting
+$(\rho h)$ to the edges directly.  We use {\tt make\_edge\_scal} with
+{\tt is\_conservative = .true.} on $(\rho h)$, with the underbraced term
+in Eq.~\ref{rhoh equation} as the force (computed in {\tt mkrhohforce}).
 

### 2012-10-10 Docs/test_suite/test_suite.tex
renname docs Docs

--- 2012-07-10 docs/test_suite/test_suite.tex
@@
 An automated regression test suite for \maestro\ (or any BoxLib-based
-code) written in Python exists in {\tt BoxLib/Tools/C\_util/regtests/}
+code) written in Python exists in {\tt BoxLib/Tools/RegressionTesting/}
 as {\tt testnew.py}.  The test suite consists of a set of problem

### 2012-10-10 Docs/unit_tests/unit_tests.tex
renname docs Docs

--- 2012-08-30 docs/unit_tests/unit_tests.tex
@@
 %
-\caption[Results of 2-d the {\tt test\_projection} unit test]{\label{fig:unit:projtest}
+\caption[Results of the 2-d {\tt test\_projection} unit test]{\label{fig:unit:projtest}
   Initial divergence free velocity field (x-component; left); Velocity

### 2012-10-10 Docs/getting_started/getting_started.tex
updates to directory structure

--- 2012-06-05 docs/getting_started/getting_started.tex
@@
 In \maestro, each problem lives under one of three sub-directories of
-{\tt MAESTRO/}: {\tt SCIENCE/}, {\tt TEST\_PROBLEMS/}, or {\tt
+{\tt MAESTRO/Exec}: {\tt SCIENCE/}, {\tt TEST\_PROBLEMS/}, or {\tt
 UNIT\_TESTS/}.  This problem sub-directory will contain any
@@
 
-Change directory to {\tt MAESTRO/TEST\_PROBLEMS/test2/}.    
+Change directory to {\tt MAESTRO/Exec/TEST\_PROBLEMS/test2/}.    
 We only need to worry about the options at the very top of the
@@
 Different problems in \maestro\ are contained in one of three
-sub-directories under {\tt MAESTRO/} ({\tt SCIENCE/}, {\tt
+sub-directories under {\tt MAESTRO/Exec}: ({\tt SCIENCE/}, {\tt
 TEST\_PROBLEMS/}, or {\tt UNIT\_TESTS/}).  The {\tt GNUmakefile} in each
@@
 \begin{itemize}
-\item {\tt flame} \\[-3mm]
+\item {\tt incomp\_shear\_jet} \\[-3mm]
 
-The {\tt flame} problem models a planar laminar flame.  Initially cool
-fuel and hot ash are put in pressure equilibrium.  Thermal diffusion
-transfers heat to the fuel, driving a burning front.  This problem
-does not use the traditional \maestro\ elliptic constraint, but
-rather enforces $\nabla \cdot U = S$ (through the runtime parameter
-{\tt do\_smallscale = T}) as discussed in \cite{SNe}.
-
-\item {\tt spherical\_heat} \\[-3mm]
-
-{\tt spherical\_heat} maps a spherical star (a Chandrasekhar-mass white
-dwarf) onto the Cartesian domain and uses an external heat source to
-cause the star to expand.  This is the 3D analogue of the {\tt
-  test\_basestate} unit test (see \S~\ref{chapter:unit_tests}).  This
-problem was discussed in \cite{multilevel}.
+A simple pure-incompressible shear layer problem.  This is the example
+problem used in \cite{bellcolellaglaz}.  This is useful to see how to
+use \maestro\ as an incompressible solver.
 
@@
 
-\item {\tt test\_spherical} \\[-3mm]
+\end{itemize}
 
-{\tt test\_spherical} sets up an isentropically stratified star
-and stirs it up with a random velocity field.  The low Mach number
-constraint is replaced with the anelastic constraint (through
-the {\tt beta\_type} runtime parameter).  Analytically, under
-these conditions, the density of the star should not change.
-This test problem was discussed in \cite{lowMach4}.
+\section{Distributed Science Problems}
+
+The following problems were used for science studies.  It is
+anticipated that more will be made available with time.
+
+\begin{itemize}
+
+\item {\tt toy\_convect} \\[-3mm]
+
+A nova-like problem for studying convection.  This problem has seen
+extensive use in understanding which prediction types are the best
+when we have sharp species gradients.  See Mike Z or Ryan for details.
+
+\item {\tt wdconvect} \\[-3mm]
+
+Model convection leading up to ignition in the Chandraseskhar-mass SNe
+Ia progenitor model.  This setup was the basis for the simulations
+presented in \cite{lowMach4,wdconvect,wdturb}.
 

### 2012-10-11 Docs/architecture/architecture.tex
some fixes

--- 2012-10-10 docs/architecture/architecture.tex
@@
 
+      A runtime parameter, {\tt use\_eos_coulomb}, is defined in
+      this EOS to enable/disable Coulomb corrections.
+
 \item {\tt extern/EOS/gamma\_law\_general} assumes an ideal gas with a mixed 
@@
 \footnote{ Note: an older interface to the EOS exists, but is
-  deprecated.  In this mode, the {\tt eos\_module} declares for
-  convenience the variables that need appear in the {\tt eos} call
+  deprecated.  In this mode, the {\tt eos\_old_interface} module declares 
+  the variables that need appear in the old-style {\tt eos} call
   argument list.  \maestro\ routines use these module variables in the
   EOS call to avoid having to declare each quantity in each routine
-  that calls the EOS.}
+  that calls the EOS.  Most code has been updated to use the new interface.}
 

### 2012-10-11 Docs/architecture/architecture.tex
fix compilation

@@
 
-      A runtime parameter, {\tt use\_eos_coulomb}, is defined in
+      A runtime parameter, {\tt use\_eos\_coulomb}, is defined in
       this EOS to enable/disable Coulomb corrections.
@@
 \footnote{ Note: an older interface to the EOS exists, but is
-  deprecated.  In this mode, the {\tt eos\_old_interface} module declares 
+  deprecated.  In this mode, the {\tt eos\_old\_interface} module declares 
   the variables that need appear in the old-style {\tt eos} call

### 2012-10-13 Docs/preface/preface.tex
add mailing list

--- 2012-06-05 docs/preface/preface.tex
@@
 \noindent The current version of the \maestro\ User's Guide can be found in 
-the \maestro\ source tree under {\tt fParallel/MAESTRO/docs/}.
+the \maestro\ source tree under {\tt MAESTRO/docs/}.
+
+\noindent Support is available via the mailing list: {\tt maestro-help@bender.astro.sunysb.edu}.  To subscribe to this list or read the archives, visit
+{\tt http://bender.astro.sunysb.edu/mailman/listinfo/maestro-help}.

### 2012-10-16 Docs/architecture/architecture.tex
update
some refs

@@
 
-  Each problem in \maestro\ gets it own sub-directory under {\tt
-    SCIENCE/}, {\tt TEST\_PROBLEMS/}, or {\tt UNIT\_TESTS/}.  The {\tt
-    GNUmakefile} in the problem directory includes the instructions on
-  how to build the executable, including what modules in {\tt extern}
-  are used.  Any file that you place in a sub-directory here takes
-  precedence over a file of the same name in {\tt MAESTRO/}.  This
-  allows problems to have custom versions of the main
-  \maestro\ routines (e.g.\ initial conditions via {\tt initdata.f90}.
-  See \S~\ref{sec:makefile} for details on the build system).
-
   Important directories under {\tt MAESTRO/} include:
@@
 
-  \item {\tt docs/}
+  \item {\tt Docs/}
 
@@
 
-    The various problem-setups.
+    The various problem-setups.  Each problem in \maestro\ gets it own
+    sub-directory under {\tt SCIENCE/}, {\tt TEST\_PROBLEMS/}, or {\tt
+      UNIT\_TESTS/}.  The {\tt GNUmakefile} in the problem directory
+    includes the instructions on how to build the executable,
+    including what modules in {\tt extern} are used.  Any file that
+    you place in a sub-directory here takes precedence over a file of
+    the same name in {\tt MAESTRO/}.  This allows problems to have
+    custom versions of the main \maestro\ routines (e.g.\ initial
+    conditions via {\tt initdata.f90}.  See \S~\ref{sec:makefile} and
+    Chapter~\ref{ch:make} for details on the build system).
 
@@
       some data structures and then call only a few of the
-      \maestro\ routines.  See \S~\ref{sec:unit_tests} for details.
+      \maestro\ routines.  See Chapter~\ref{chapter:unit_tests} for details.
 

### 2012-10-16 Docs/MESA/hopper_compile.tex
latest MESA docs

@@
+These notes are somewhat dated because they pertain only to {\sf MESA} version 
+4088. The above notes are for version 4411. In my experience, version 4088 is 
+much easier to install than 4411. When using the scripts in 
+{\sf AstroDev/MESA/networks/scripts}, make sure the proper list of modules to 
+compile (lines containing {\tt do\_one ....}) is used. Those lists are 
+version dependant.
+
 Once {\sf MESA} was installed and working properly on a local machine (e.g. 

### 2012-10-16 Docs/MESA/instructions.tex
latest MESA docs

--- 2012-09-13 docs/MESA/instructions.tex
@@
 {\tt http://mesa.sourceforge.net}. Check the website for the latest version, 
-which is 4298 with the writing of this document. The codes and algorithms 
+which is 4442 with the writing of this document. The codes and algorithms 
 discussed here were created when the current version of {\sf MESA} was 
-version 4088. Between versions 4088 and 4298, there are minor changes that 
+version 4411. Between versions 4411 and 4442, there are minor changes that 
 will cause some of the included routines to crash. To download the source code 
-associated with version 4088, use the following command:
+associated with version 4411, use the following command:
 \begin{lstlisting}[language=bash,mathescape=false]
-  svn co -r 4088 http://mesa.svn.sourceforge.net/svnroot/mesa/trunk mesa
+  svn co -r 4411 http://mesa.svn.sourceforge.net/svnroot/mesa/trunk mesa
 \end{lstlisting}
@@
 located in the {\tt mesa/utils} directory and is called 
-{\tt makefile\_header\_non\_mesasdk}. Open it and change the following:
+{\tt makefile\_header\_non\_mesasdk}.
+
+All of the following changes apply to 
+variables that already exist. There is no need to explicitly add new code 
+unless you intend to compile {\sf MESA} with something other than 
+{\tt gfortran} or {\tt ifort}. If you want to compile using something 
+different, then the only code that needs to be explicitly added are the 
+definitions of the compiler flags.
+
+Open the {\tt makefile\_header\_non\_mesasdk} and change the following:
 \begin{enumerate}
@@
 \item {\tt PGPLOT}: Set the variable {\tt USE\_PGSTAR} to {\tt NO} and set 
-the variable {\tt LOAD\_PGPLOT} to be empty on the line corresponding to 
-the compiler that was chosen. The default for {\tt gcc} and {\tt ifort} is 
+the variable {\tt LOAD\_PGPLOT} to be empty in the if-construct corresponding 
+to the compiler that was chosen. The default for {\tt gcc} and {\tt ifort} is 
 {\tt YES}, but for any other compiler the default is {\tt NO}. {\tt PGPLOT} 
@@
 \item {\tt USE\_SE}: Set the variable {\tt USE\_SE} to {\tt NO} and set both 
-{\tt LOAD\_SE} and {\tt INCLUDE\_SE} to be empty. {\tt SE} is a library of 
+{\tt LOAD\_SE} and {\tt INCLUDE\_SE} to be empty. These also need to be set 
+in the correct if-construct according to which compiler was chosen. 
+{\tt SE} is a library of 
 formats used for writing/reading data in stellar evolution codes.
 
-\item Compiler Flags: Installing {\sf MESA} without OpenMP will result in a 
+\item Compiler Flags: Find the if-construct corresponding to the chosen 
+compiler. This is where compiler options are defined. It is not necessary to 
+edit any of these if you are using {\tt gfortran} or {\tt ifort}. These are 
+the variables to edit if you wish to add more (or less) debugging flags 
+for example. If using a different compiler, the variables in the last 
+{\tt else} statement (which are all empty except for 
+{\tt FCbasic = UNKNOWN COMPILER}) are what must be edited to hold the compiler 
+flag definitions.
+
+Installing {\sf MESA} without OpenMP will result in a 
 failed install. {\sf MESA} is capable of being compiled in serial, but it 
-must happen after the initial install. If not using {\tt gcc} or {\tt ifort} 
-it will be necessary to enter all of the relevant compiler flags.
+must happen after the initial install.
+
+\item {\tt LOAD\_MATRIX}: {\tt LOAD\_MATRIX} is first defined near the end of 
+the file around line $\sim$275. The purpose of this variable is to load all 
+the libraries associated with linear algebra including {\tt LAPACK}, {\tt BLAS} 
+and a few others. The lines we wish to change look like this:
+\begin{lstlisting}[language=make,mathescape=false]
+  #LOAD_MATRIX = -lmtx -lconst -lmesaklu $(LOAD_LAPACK) $(LOAD_BLAS)
+  # for ifort on mac, we need the following def for LOAD_MATRIX 
+  # because of loader bugs (as of July 2012)
+  LOAD_MATRIX = -lmtx $(MESA_DIR)/const/make/const_def.o \
+      $(MESA_DIR)/const/make/const_lib.o -lmesaklu $(LOAD_LAPACK) $(LOAD_BLAS)
+  LOAD_MESA_NUMERICS = -linterp_2d -linterp_1d -lnum -lutils -lalert \
+      $(LOAD_MATRIX)
+\end{lstlisting}
+We must uncomment the first {\tt LOAD\_MATRIX} and comment out the second one:
+\begin{lstlisting}[language=make,mathescape=false]
+  LOAD_MATRIX = -lmtx -lconst -lmesaklu $(LOAD_LAPACK) $(LOAD_BLAS)
+  # for ifort on mac, we need the following def for LOAD_MATRIX 
+  # because of loader bugs (as of July 2012)
+  #LOAD_MATRIX = -lmtx $(MESA_DIR)/const/make/const_def.o \
+  #    $(MESA_DIR)/const/make/const_lib.o -lmesaklu $(LOAD_LAPACK) $(LOAD_BLAS)
+  LOAD_MESA_NUMERICS = -linterp_2d -linterp_1d -lnum -lutils -lalert \
+       $(LOAD_MATRIX)
+\end{lstlisting}
+
+\item {\tt install} script: In the {\tt MESA} directory, open the {\tt install} 
+script and comment out two lines. We want to change:
+\begin{lstlisting}[language=bash,mathescape=false]
+  do_one atm
+
+  do_one sample
+
+  do_one star
+
+  do_one adipls
+\end{lstlisting}
+to this:
+\begin{lstlisting}[language=bash,mathescape=false]
+  do_one atm
+
+  #do_one sample
+
+  #do_one star
+
+  #do_one adipls
+\end{lstlisting}
+There are routines in the {\tt star} package and {\tt adipls} that rely on 
+the {\tt PGSTAR} modules. We do not wish to use these modules and the easiest 
+way around this is to not install both {\tt star} and {\tt adipls} because 
+neither one is needed to include {\sf MESA} microphysics modules in 
+{\sf MAESTRO}. {\tt star} is the 1D stellar evolution module and {\tt adipls} 
+module is used for asteroseismology. The {\tt sample} module is simply a 
+sample directory that is not needed for using the main microphysics modules.
+
+\item Edit 1 source file: An undefined reference to 
+{\tt omp\_get\_max\_threads} occurs in the {\tt const\_def.f} file located in 
+{\tt const/public/const\_def.f}.
+
+In the {\tt const\_def.f} file, around line $\sim$164, in the 
+{\tt do\_const\_init} subroutine, there is a line that reads:
+\begin{lstlisting}[language=fortran,mathescape=false]
+  omp_num_threads = omp_get_max_threads()
+\end{lstlisting}
+Our first change is to declare a temporary character variable in the 
+({\tt do\_const\_init} subroutine):
+\begin{lstlisting}[language=fortran,mathescape=false]
+  character(len=16) :: temp_max_threads
+\end{lstlisting}
+Then we comment out the above line ({\tt omp\_num\_threads = ...}) and add the 
+following (all in the {\tt do\_const\_init} subroutine):
+\begin{lstlisting}[language=fortran,mathescape=false]
+  !omp_num_threads = omp_get_max_threads()
+  call get_environment_variable("OMP_NUM_THREADS", temp_max_threads)
+  read(temp_max_threads, *) omp_num_threads
+\end{lstlisting}
+
 \end{enumerate}
@@
 pertaining to the network and equation of state modules.
+
+The routines that carry out these processes (setting up EOS, setting up 
+isotope map, shutdown routines, etc.) have already been written and live in 
+{\sf AstroDev/networks/MESA} and {\sf AstroDev/EOS/MESA}. This section 
+outlines the necessary steps taken by those routines in order to set up the 
+EOS and network.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Compilation Script}
+\section{Compilation Scripts}
  \label{sec:compile_scripts}
 
-Useful script written in {\tt BASH} to help with compiling {\sf MESA} modules. 
-The script cleans the module directory, compiles the source code and exports 
-the libraries and {\tt .mod} files to the appropriate directories for every 
-module. This script can be found in {\tt AstroDev/networks/MESA/scripts}. 
-There is also a script to compile just one module. The compile-all-modules 
-script should be run from the top {\sf MESA} directory while the 
-compile-one-module script can be run from anywhere because it takes the 
-relative path to the module.
+Some useful scripts have been written in {\tt BASH} to help with compiling 
+{\sf MESA} modules. They live in the {\sf AstroDev/networks/MESA/scripts} 
+directory. Currently, there are two scripts; one that compiles all modules 
+(which unfortunately is version dependant), and one that compiles a single 
+module.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Script
-%%%%%%%%
-\subsection{Compile All Modules}
-\input compile_all_mods
+The compile-one-module script can be run from any directory because the 
+calling sequence involves the relative path to the {\sf MESA} module:
+\begin{lstlisting}[language=bash,mathescape=false]
+  compile-one-mod.sh ../../../mesa/net
+\end{lstlisting}
+The above command will compile the {\tt net} module without needing to be in 
+the {\sf MESA} directory.
+
+The compile-all-modules script will clean the module directory, compile the 
+source code and export the libraries and {\tt .mod} files to the appropriate 
+directories for every module. To run this script, you must be in the {\sf MESA} 
+directory. The calling sequence would be something like:
+\begin{lstlisting}[language=bash,mathescape=false]
+  cd $MESA_DIR
+  compile-all-mods.sh
+\end{lstlisting}
 

### 2012-10-26 Docs/makefiles/makefiles.tex

--- 2012-06-29 docs/makefiles/makefiles.tex
@@
   environment (compiler used, build machine, build directory, compiler
-  flags, etc.).  This is created by the script {\tt make\_build\_info}
+  flags, etc.).  This is created by the script {\tt makebuildinfo.py}
   in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_scripts/} from {\tt

### 2012-11-09 Docs/architecture/architecture.tex
a little more clarification on the boundary conditions for phi after chatting with Ann

@@
 the pressure boundary conditions here reflect the behavior we want for
-the velocity.  You want $d\phi/dn=0$ in places where the incoming
-velocity to the elliptic solve must match the velocity you want after
-the projection (e.g.\ inflow, solid walls).  For outlet boundaries,
-setting $\phi = 0$ allows the velocity to adjust as needed to the
-domain (see, for example, \cite{almgrenBellSzymczak:1996}).
+the velocity.  After the projection, it is $\nabla \phi$ that modifies
+the velocity field.  At a wall or for inflow conditions, we already
+have the velocity we want at the boundary, so we want the velocity to
+remain unchanged after the projection.  This requires $d\phi/dn=0$ on
+those boundaries.  For outflow, we impose a condition that we do not
+want the boundaries to introduce any tangental acceleration (or
+shear), this is equivalent to setting $\phi = 0$ (then $\partial
+\phi/d\partial t = 0$, with $t$ meaning `tangental').  This allows the
+velocity to adjust as needed to the domain (see, for example,
+\cite{almgrenBellSzymczak:1996}).
 

### 2012-12-19 Docs/managing_jobs/managingjobs.tex
some hopper notes

--- 2012-07-05 docs/managing_jobs/managingjobs.tex
@@
-\section{Automatic Restarting and Archiving of Data}
+\section{Working at OLCF (ORNL)}
+
+\subsection{Automatic Restarting and Archiving of Data}
 
@@
 
-\section{Remote \visit\ Visualization on Lens}
+\subsection{Remote \visit\ Visualization on Lens}
 
@@
 
-On your local machine, launch \visit.
+On your local machine, launch \visit.  Note: this procedure seems to
+work with \visit~2.4.2, but not \visit~2.5.0 for some reason.
 \begin{itemize}
@@
   \item Check {\em Launch parallel engine}
-  \item Select the {\em Parallel launch metho} as {\tt qsub/mpirun}
+  \item Select the {\em Parallel launch method} as {\tt qsub/mpirun}
   \item Set the {\em Partition / Pool / Queue} to {\tt computation}
@@
 \end{itemize}
+
+\section{Working at NERSC}
+
+\subsection{Using the {\tt AmrPostprocesing} python plotting scripts on hopper}
+
+To build the {\tt fsnapshot.so} library, you need to do:
+\begin{verbatim}
+module load gcc
+\end{verbatim}
+{\tt f2py} is already in the path, so the library should then build without issue.
+%
+
+Then edit your {\tt .bashrc.ext} file to set the {\tt PYTHONPATH} to
+the {\tt python\_plotfile} directory, e.g.:
+\begin{verbatim}
+export PYTHONPATH="/global/homes/z/zingale/AmrPostprocessing/F\_Src/python_plotfile"
+\end{verbatim}
+%
+and set the {\tt PATH} to that directory,
+\begin{verbatim}
+export PATH="/global/homes/z/zingale/AmrPostprocessing/F_Src/python_plotfile:$PATH"
+\end{verbatim}
+
+To run the script, you need to do:
+\begin{verbatim}
+module load matplotlib
+module load python
+\end{verbatim}
+
+\subsection{Remote visualization on hopper}
+
+\visit\ is already configured to work with hopper.  If the host does not appear
+in your local version of visit, copy the {\tt host\_nersc\_hopper.xml} file
+from the {\tt .visit/allhosts/} directory under the system's \visit\ install path
+to your {\tt $\mathtt{\sim}$/.visit/hosts/} directory.

### 2012-12-19 Docs/managing_jobs/managingjobs.tex
add more notes about running

@@
 The submission script {\tt jaguar.run} and shell script
-{\tt process.jaguar} in {\tt wdconvect/job\_scripts/}
+{\tt process.jaguar} in {\tt Util/job\_scripts/jaguar/}
 are designed to allow you to run \maestro\ with minimal interaction,
@@
 
-Analogous scripts exist for running on Atlas, with the same general
-procedure.  The command to chain a job on atlas is:
-\begin{verbatim}
-msub -l depend=<JOB-ID> <MSUB SCRIPT>
-\end{verbatim}
-where, again, {\tt <JOB-ID>} is the id number of the job that must
-complete before the current submission runs and {\tt <MSUB SCRIPT>}
-is the job submission script (e.g.\ {\tt atlas.run}).
+The {\tt chainsub.sh} script can be used to automatically launch a number
+of jobs depending on a single, currently queued (or running) job.
 
@@
 
+\subsection{Automatic Restarting and Archiving of Data}
+
+The same set of submission scripts are available for hopper at NERSC in 
+{\tt Util/job\_scripts/hopper/}.  They are mostly identical, with a few
+minor changes for the local system configuration.  The use is the same
+as described above.
+
 \subsection{Using the {\tt AmrPostprocesing} python plotting scripts on hopper}

### 2012-12-19 Docs/managing_jobs/managingjobs.tex
more hopper stuff

@@
 
+\subsection{General Guidelines}
+
+hopper is configured with 24 cores per node, with the cores grouped
+into 4 NUMA nodes of 6 cores each.  Best performance is seen when
+running with 6 threads and 4 MPI tasks per node, using the following
+run line:
+\begin{verbatim}
+aprun -n 128 -N 4 -d 6 -S 1 -ss ./main.Linux.Cray.mpi.omp.ex inputs_3d
+\end{verbatim}
+Here, {\tt -n} is the number of MPI tasks, and {\tt n * d} is the
+total number of cores (which should match the {\tt mppwidth} set at
+the top of the submission script).
+
 \subsection{Automatic Restarting and Archiving of Data}

### 2013-01-15 Docs/managing_jobs/managingjobs.tex
add note about MPI error on hopper

@@
 total number of cores (which should match the {\tt mppwidth} set at
-the top of the submission script).
+the top of the submission script). \\
+
+\noindent (2013-01-15) It's been seen that sometimes an MPICH2 error is 
+generated of the form:
+\begin{verbatim}
+MPIU_nem_gni_get_hugepages(): Unable to mmap 8388608 bytes for file 
+/var/lib/hugetlbfs/global/pagesize-2097152/hugepagefile.MPICH.1.19467.kvs_14619732, 
+err Cannot allocate memory
+\end{verbatim}
+Setting:
+\begin{verbatim}
+export MPICH_GNI_MALLOC_FALLBACK=enabled
+\end{verbatim}
+in your submission script seems to fix this issue.  This will allow
+MPI to continue if there are insufficient huge pages in one NUMA node.
+(Alternately, it was suggested to run without the {\tt -ss} option to
+{\tt aprun}.
 

### 2013-01-20 Docs/visualization/visualization.tex
more details on python plotting scripts

--- 2012-10-02 docs/visualization/visualization.tex
@@
 
-{\tt AmrPostprocessing/F\_Src/python\_plotfile} provide a simple commandline
-tool for doing visualizations of \boxlib\ plotfiles.  The main drivers
+{\tt AmrPostprocessing/F\_Src/python\_plotfile} provide some simple commandline
+tools for doing visualizations of \boxlib\ plotfiles.  The main drivers
 are written in python and use a set of Fortran routines, compiled with
@@
 a library {\tt fsnapshot.so}.
+
+The path to {\tt fsnapshot.so} should be included in your {\tt PYTHONPATH}
+environment variable.  This can be done by adding:
+\begin{verbatim}
+export PYTHONPATH="${PYTHONPATH}:/home/user/AmrPostprocessing/F_Src/python_plotfile}
+\end{verbatim}
+to your {\tt .bashrc}.
+
+It is recommended that you use {\tt matplotlib} version 1.2.0 or
+higher.  If the fonts look strange in the output files, you can try
+installing the {\tt lyx-fonts} package and deleting your {\tt
+  .matplotlib} directory, and trying again.
 
@@
 
-{\tt contourcompare.py} takes 2 plotfiles and a variable as arguments
-and plots contours of the 2 datasets on the same set of axes.  This is 
+{\tt contourcompare.py} takes two or three plotfiles and a single variable as arguments
+and plots contours of the datasets on the same set of axes.  This is 
 form comparisons of different runs.  Running the script with no parameters
 will give the full list available options.
+
+For example:
+\begin{verbatim}
+contourcompare.py tfromp plt00000 other_plt00000
+\end{verbatim}
+will make a contour plot of the variable {\tt tfromp} from the data in
+{\tt plt00000} and {\tt other\_plt00000} shown on the same axes.
+
+\subsection{\tt runtimevis.py}
+
+The {\tt runtimevis.py} script is designed to be run from a submission
+script to produce plots from plotfiles as they are produced.  This is
+accomplished by hooking it into the {\tt process} scripts described in
+Chapter~\ref{ch:managingjobs}.
+
+The script itself reads in an inputs file, {\tt vis.in}, that
+describes the variables to plot.  From 1 to 6 variables can be
+plotting from a plotfile.  The script does its best to organize them
+in columns and rows to maximize the plot area.  The image is always
+output at 1280$\times$720 pixels, corresponding to 720p HD resolution.
+For each variable, a block of the form:
+\begin{verbatim}
+[varname]
+min = 1
+max = 2
+log = 1
+\end{verbatim}
+is supplied.  If {\tt min} or {\tt max} are omitted, then the data
+limits are computed automatically.  If {\tt log} is omitted, then no
+log is taken of the data before plotting.  The script is then run as:
+\begin{verbatim}
+runtimevis.py plt00000
+\end{verbatim}
 
@@
 For completeness, below is a script containing our example above.
-\newpage
 \begin{lstlisting}[language=Python]

### 2013-01-21 Docs/analysis/analysis.tex
renname docs Docs
update path for python plotting routines

--- 2012-04-11 docs/analysis/analysis.tex
@@
 
-\item {\tt fIDLdump.f90}, {\tt fIDLdump3d.f90}
-
-  Create a binary file containing a uniformly gridded, single variable
-  from a plotfile.  This is can then be read into IDL using the routines
-  in {\tt fParallel/scripts/idlbl/}.
-
 \item {\tt fsnapshot2d.f90}, {\tt fsnapshot3d.f90}
@@
 first set your {\tt PYTHONPATH} environment variable to point to the
-{\tt AmrPostprocessing/F\_Src/python/} directory, for example:
+{\tt AmrPostprocessing/python/} directory, for example:
 \begin{verbatim}
-export PYTHONPATH="/home/username/development/AmrPostprocessing/F_Src/python"
+export PYTHONPATH="/home/username/development/AmrPostprocessing/python"
 \end{verbatim}

### 2013-01-21 Docs/managing_jobs/managingjobs.tex
add a label
update path for python plotting routines

@@
+\label{ch:managingjobs}
+
 \section{Working at OLCF (ORNL)}
@@
 \begin{verbatim}
-export PYTHONPATH="/global/homes/z/zingale/AmrPostprocessing/F\_Src/python_plotfile"
+export PYTHONPATH="/global/homes/z/zingale/AmrPostprocessing/python"
 \end{verbatim}
@@
 \begin{verbatim}
-export PATH="/global/homes/z/zingale/AmrPostprocessing/F_Src/python_plotfile:$PATH"
+export PATH="/global/homes/z/zingale/AmrPostprocessing/python:$PATH"
 \end{verbatim}

### 2013-01-21 Docs/visualization/visualization.tex
update path for python plotting routines

@@
 
-{\tt AmrPostprocessing/F\_Src/python\_plotfile} provide some simple commandline
+{\tt AmrPostprocessing/python} provides some simple commandline
 tools for doing visualizations of \boxlib\ plotfiles.  The main drivers
@@
 \begin{verbatim}
-export PYTHONPATH="${PYTHONPATH}:/home/user/AmrPostprocessing/F_Src/python_plotfile}
+export PYTHONPATH="${PYTHONPATH}:/home/user/AmrPostprocessing/python}
 \end{verbatim}

### 2013-01-29 Docs/managing_jobs/managingjobs.tex

@@
 \label{ch:managingjobs}
+
+\section{General Info}
+
+All plotfile directories have a {\tt job\_info} file which lists as
+host of parameters about the simulation, including:
+\begin{itemize}
+\item A descriptive name for the simulation (the {\tt job\_name} runtime
+  parameter
+\item The number of MPI tasks and OpenMP threads
+\item The total amount of CPU-hours used up to this point
+\item The data and time of the plotfile creation and the directory it was written to.
+\item The build date, machine, and directory
+\item The MAESTRO, BoxLib and other relevant git hashes for the source
+\item The directories used in the build
+\item The compilers used and compilation flags
+\item The number of levels and boxes for the grid
+\item The properties of the species carried
+\item The tolerances for the MG solves
+\item The value of all runtime parameters (even those that were not explicitly set in the inputs file)
+\end{itemize}
+This file makes it easy to understand how to recreate the run that
+produced the plotfile.
 

### 2013-02-11 Docs/FAQ/faq.tex

--- 2012-10-02 docs/FAQ/faq.tex
@@
 
+\item {\em Why does ``checkerboarding'' appear in the velocity field,
+especially in regions where the flow is stagnant?}
+
+Checkerboarding can arise from the projection---it doesn't see that
+mode (because it is an approximate projection) so it is unable to
+remove it.  This allows the pattern to slowly build up.  There are
+filtering techniques that can be used to remove these modes, but
+they are not implemented in \maestro.
+
 \end{enumerate}

### 2013-04-03 Docs/MESA/instructions.tex
update MESA docs

@@
 in addition to numerical routines. This document describes how to link the 
-{\sf MESA} modules to an outside program such as {\sf MAESTRO}.
-\footnote{This
+{\sf MESA} modules to an outside program such as {\sf MAESTRO}\footnote{This
   interface work was done by Ryan Orvedahl---questions (and credit)
@@
 {\tt http://mesa.sourceforge.net}. Check the website for the latest version, 
-which is 4442 with the writing of this document. The codes and algorithms 
+which is 4849 when the last upgrades to this document were added. The codes 
+and algorithms 
 discussed here were created when the current version of {\sf MESA} was 
-version 4411. Between versions 4411 and 4442, there are minor changes that 
-will cause some of the included routines to crash. To download the source code 
-associated with version 4411, use the following command:
+version 4088 and have been modified accordingly to keep up with the quickly 
+channging source code. It might be possible to use older versions of 
+{\sf MESA} with 
+the information in this document, although there are minor changes and 
+upgrades between {\sf MESA} versions that may cause the routines and algorithms 
+discussed here to crash. To download the source code associated with version 
+4849, issue the following command:
 \begin{lstlisting}[language=bash,mathescape=false]
-  svn co -r 4411 http://mesa.svn.sourceforge.net/svnroot/mesa/trunk mesa
+  svn co -r 4849 svn://svn.code.sf.net/p/mesa/code/trunk mesa
 \end{lstlisting}
@@
 in the directory where this command was run. This step might take some time 
-($\sim$10 min depending on the internet connection) possibly because the 
-final {\tt mesa} directory will take up roughly 2.2 GB.
+($\sim$3 min depending on the internet connection) possibly because the 
+final {\tt mesa} directory will take up roughly 2.1 GB.
 
@@
 
-\item Set OpenMP number of threads: {\tt export OMP\_NUM\_THREADS=n}, a suggested value is the number of cores in the computer
+\item Set OpenMP number of threads: {\tt export OMP\_NUM\_THREADS=n}, a 
+suggested value is the number of cores in the computer
 \end{enumerate}
@@
 {\tt BLAS} provided with the {\sf MESA} source tree using the following 
-definitions:
+definitions (these should be the default settings):
 \begin{lstlisting}[language=make,mathescape=false,upquote=true]
@@
 
-\item {\tt LOAD\_MATRIX}: {\tt LOAD\_MATRIX} is first defined near the end of 
-the file around line $\sim$275. The purpose of this variable is to load all 
-the libraries associated with linear algebra including {\tt LAPACK}, {\tt BLAS} 
-and a few others. The lines we wish to change look like this:
-\begin{lstlisting}[language=make,mathescape=false]
-  #LOAD_MATRIX = -lmtx -lconst -lmesaklu $(LOAD_LAPACK) $(LOAD_BLAS)
-  # for ifort on mac, we need the following def for LOAD_MATRIX 
-  # because of loader bugs (as of July 2012)
-  LOAD_MATRIX = -lmtx $(MESA_DIR)/const/make/const_def.o \
-      $(MESA_DIR)/const/make/const_lib.o -lmesaklu $(LOAD_LAPACK) $(LOAD_BLAS)
-  LOAD_MESA_NUMERICS = -linterp_2d -linterp_1d -lnum -lutils -lalert \
-      $(LOAD_MATRIX)
-\end{lstlisting}
-We must uncomment the first {\tt LOAD\_MATRIX} and comment out the second one:
-\begin{lstlisting}[language=make,mathescape=false]
-  LOAD_MATRIX = -lmtx -lconst -lmesaklu $(LOAD_LAPACK) $(LOAD_BLAS)
-  # for ifort on mac, we need the following def for LOAD_MATRIX 
-  # because of loader bugs (as of July 2012)
-  #LOAD_MATRIX = -lmtx $(MESA_DIR)/const/make/const_def.o \
-  #    $(MESA_DIR)/const/make/const_lib.o -lmesaklu $(LOAD_LAPACK) $(LOAD_BLAS)
-  LOAD_MESA_NUMERICS = -linterp_2d -linterp_1d -lnum -lutils -lalert \
-       $(LOAD_MATRIX)
-\end{lstlisting}
+%============================================================================
+% This next \item is unnecessary for version 4849 (it was necessary for 4411)
+%============================================================================
+%
+%\item {\tt LOAD\_MATRIX}: {\tt LOAD\_MATRIX} is first defined near the end of 
+%the file around line $\sim$280. The purpose of this variable is to load all 
+%the libraries associated with linear algebra including {\tt LAPACK}, 
+%{\tt BLAS} and a few others. The lines we wish to change look like this:
+%\begin{lstlisting}[language=make,mathescape=false]
+%  #LOAD_MATRIX = -lmtx -lconst -lmesaklu $(LOAD_LAPACK) $(LOAD_BLAS)
+%  # for ifort on mac, we need the following def for LOAD_MATRIX 
+%  # because of loader bugs (as of July 2012)
+%  LOAD_MATRIX = -lmtx $(MESA_DIR)/const/make/const_def.o \
+%      $(MESA_DIR)/const/make/const_lib.o -lmesaklu $(LOAD_LAPACK) $(LOAD_BLAS)
+%  LOAD_MESA_NUMERICS = -linterp_2d -linterp_1d -lnum -lutils -lalert \
+%      $(LOAD_MATRIX)
+%\end{lstlisting}
+%We must uncomment the first {\tt LOAD\_MATRIX} and comment out the second one:
+%\begin{lstlisting}[language=make,mathescape=false]
+%  LOAD_MATRIX = -lmtx -lconst -lmesaklu $(LOAD_LAPACK) $(LOAD_BLAS)
+%  # for ifort on mac, we need the following def for LOAD_MATRIX 
+%  # because of loader bugs (as of July 2012)
+%  #LOAD_MATRIX = -lmtx $(MESA_DIR)/const/make/const_def.o \
+%  #    $(MESA_DIR)/const/make/const_lib.o -lmesaklu $(LOAD_LAPACK) $(LOAD_BLAS)
+%  LOAD_MESA_NUMERICS = -linterp_2d -linterp_1d -lnum -lutils -lalert \
+%       $(LOAD_MATRIX)
+%\end{lstlisting}
 
-\item {\tt install} script: In the {\tt MESA} directory, open the {\tt install} 
-script and comment out two lines. We want to change:
-\begin{lstlisting}[language=bash,mathescape=false]
-  do_one atm
+%============================================================================
+% This next \item is unnecessary for version 4849 (it was necessary for 4411)
+%============================================================================
+%
+%\item {\tt install} script: In the {\tt MESA} directory, open the 
+%{\tt install} script and comment out two lines. We want to change:
+%\begin{lstlisting}[language=bash,mathescape=false]
+%  do_one atm
+%
+%  do_one sample
+%
+%  do_one star
+%
+%  do_one adipls
+%\end{lstlisting}
+%to this:
+%\begin{lstlisting}[language=bash,mathescape=false]
+%  do_one atm
+%
+%  #do_one sample
+%
+%  #do_one star
+%
+%  #do_one adipls
+%\end{lstlisting}
+%There are routines in the {\tt star} package and {\tt adipls} that rely on 
+%the {\tt PGSTAR} modules. We do not wish to use these modules and the easiest 
+%way around this is to not install both {\tt star} and {\tt adipls} because 
+%neither one is needed to include {\sf MESA} microphysics modules in 
+%{\sf MAESTRO}. {\tt star} is the 1D stellar evolution module and {\tt adipls} 
+%module is used for asteroseismology. The {\tt sample} module is simply a 
+%sample directory that is not needed for using the main microphysics modules.
 
-  do_one sample
-
-  do_one star
-
-  do_one adipls
-\end{lstlisting}
-to this:
-\begin{lstlisting}[language=bash,mathescape=false]
-  do_one atm
-
-  #do_one sample
-
-  #do_one star
-
-  #do_one adipls
-\end{lstlisting}
-There are routines in the {\tt star} package and {\tt adipls} that rely on 
-the {\tt PGSTAR} modules. We do not wish to use these modules and the easiest 
-way around this is to not install both {\tt star} and {\tt adipls} because 
-neither one is needed to include {\sf MESA} microphysics modules in 
-{\sf MAESTRO}. {\tt star} is the 1D stellar evolution module and {\tt adipls} 
-module is used for asteroseismology. The {\tt sample} module is simply a 
-sample directory that is not needed for using the main microphysics modules.
-
+%============================================================================
+% This next \item is necessary when compiling in serial
+%============================================================================
+%
 \item Edit 1 source file: An undefined reference to 
 {\tt omp\_get\_max\_threads} occurs in the {\tt const\_def.f} file located in 
-{\tt const/public/const\_def.f}.
+{\tt const/public/const\_def.f} when compiling in serial.
 
-In the {\tt const\_def.f} file, around line $\sim$164, in the 
+In the {\tt const\_def.f} file, around line $\sim$168, in the 
 {\tt do\_const\_init} subroutine, there is a line that reads:
@@
 to export the {\tt .mod} files and the library to the right directories using 
-the {\tt ./export} script. Section \ref{sec:compile_scripts} contains a 
-simple script that is meant to help with compiling modules. It cleans, makes 
-and exports every module and should be run in the top {\tt mesa/} directory.
+the {\tt ./export} script. Section \ref{sec:compile_scripts} discusses a few 
+simple scripts that are meant to help with compiling modules.
 
@@
 
-In {\tt mesa/star/private/micro.f} there are three places 
-to set the {\tt use\_omp} flag to False. They occur in the {\tt do\_eos}, 
-{\tt do\_neu} and {\tt do\_net} routines. The fourth place to set the 
-{\tt use\_omp} flag to False is contained in the {\tt do\_kap} routine in 
-{\tt mesa/star/private/opacities.f}.
+To find where all to set the {\tt use\_omp} flag to False, execute the 
+following command:
+\begin{lstlisting}[language=bash,mathescape=false]
+  find $(MESA_DIR) -name "*.f" -exec grep -Hn --color use_omp {} \;
+\end{lstlisting}
+This will return fortran files located in the {\tt mesa/star/private} 
+directory.
 

### 2013-05-16 Docs/architecture/architecture.tex
some fixes

@@
 is perfectly aligned with the Cartesian state.  A {\em spherical}
-geometry is for modeling an entire star.  Here, the 1D base state is
-not aligned with the Cartesian state.  Figure~\ref{fig:base_state}
-shows these geometries.
+geometry is for modeling an entire star\footnote{Spherical geometry
+  only exists for 3-d. This is a design decision---convection is 3-d.
+  You can however run as an octant}.  Here, the 1D base state is not
+aligned with the Cartesian state.  Figure~\ref{fig:base_state} shows
+these geometries.
 
@@
 
-The routines in {\tt MAESTRO/initial\_model/} prepare an initial model
+The routines in {\tt AstroDev/initial\_models/}\footnote{The {\tt \astrodev} package is currently distributed separately from \maestro} prepare an initial model
 for \maestro.  There are two types of routines here, the first type
@@
 be found in \S~\ref{sec:initial_models_main}.
+
+\subsection{Customizing the Initialization}
+
+The best way to customize the initialization (e.g.\ add perturbations)
+is to copy from one of the existing problems.  The file {\tt
+  initveldata.f90} controls the velocity field initialization and {\tt
+  initscaldata.f90} controls the initialization of the scalars
+($\rho$, $\rho X_k$, $\rho h$).  The {\tt test2} problem is a good
+starting point for plane-parallel and {\tt wdconvect} is a good
+starting point for full stars.
 

### 2013-05-16 Docs/initial_models/initial_models.tex
some updates

--- 2011-06-23 docs/initial_models/initial_models.tex
@@
 {\tt test\_convect} & \rightarrow & \mathrm{none-it~ is~ generated~ analytically} \nonumber \nonumber \\
-{\tt wdconvect} & \rightarrow & {\tt initial\_models/kepler/model\_6.e8.raw} \nonumber \\
+{\tt wdconvect} & \rightarrow & {\tt model\_6.e8.raw} \nonumber \\
 {\tt spherical\_heat} & \rightarrow & \mathrm{none-it~ is~ generated~ analytically} \nonumber \\
-{\tt xrb} & \rightarrow & {\tt initial\_models/xrb/sorted\_xrb.raw} \nonumber
 \end{eqnarray}
@@
 then uses an iterative procedure to modify the model data so that it
-is thermodynamically consistent with the {\tt MAESTRO} equation of
+is thermodynamically consistent with the \maestro\ equation of
 state (EOS), and also satisfies our chosen hydrostatic equilibrium
@@
 \begin{eqnarray}
-{\tt test2} & \rightarrow & {\tt initial\_models/test2/init\_1d.f90} \nonumber \\
-{\tt test\_convect} & \rightarrow & {\tt initial\_models/test2/init\_1d.f90} \nonumber \\
-{\tt wdconvect} & \rightarrow & {\tt initial\_models/kepler/init\_1d.f90} \nonumber \\
-{\tt spherical\_heat} & \rightarrow & {\tt initial\_models/spherical/init\_1d.f90} \nonumber \\
-{\tt xrb} & \rightarrow & {\tt initial\_models/xrb/init\_1d.f90} \nonumber
+{\tt test2} & \rightarrow & {\tt Util/initial\_models/test2/init\_1d.f90} \nonumber \\
+{\tt test\_convect} & \rightarrow & {\tt Util/initial\_models/test2/init\_1d.f90} \nonumber \\
+{\tt wdconvect} & \rightarrow & none\footnotemark \nonumber \\
+{\tt spherical\_heat} & \rightarrow & {\tt Util/initial\_models/spherical/init\_1d.f90} \nonumber \\
 \end{eqnarray}
-The model data is not generated at run-time - it must be generated in
+\footnotetext{the original model routines are not distributed, but the {\tt Util/initial\_models/spherical/init\_1d.f90} routines can work as a simple substitute.}
+The model data is not generated at run-time---it must be generated in
 advance of running any \maestro\ examples.  The inputs file should point

### 2013-05-16 Docs/architecture/architecture.tex
fix

@@
 
-The routines in {\tt AstroDev/initial\_models/}\footnote{The {\tt \astrodev} package is currently distributed separately from \maestro} prepare an initial model
-for \maestro.  There are two types of routines here, the first type
-modify an existing 1D initial model produced somewhere else (e.g.\ a
-1D stellar evolution code), and map it onto a uniform grid, at
-the desired resolution, using the equation of state in \maestro, and
-using \maestro's discretization of hydrostatic equilibrium.  The second
-type generate the initial model internally, by integrating the
-condition of hydrostatic equilibrium together with a simplifying
-assumption on the energy (e.g. isothermal or isentropic).  In
-both cases hydrostatic equilibrium is enforced as:
+The routines in {\tt Util/initial\_models/} prepare an initial model
+for \maestro.  In general, there are two different proceduces that are
+needed.  The first type modify an existing 1D initial model produced
+somewhere else (e.g.\ a 1D stellar evolution code), and map it onto a
+uniform grid, at the desired resolution, using the equation of state
+in \maestro, and using \maestro's discretization of hydrostatic
+equilibrium.  The second type generate the initial model internally,
+by integrating the condition of hydrostatic equilibrium together with
+a simplifying assumption on the energy (e.g. isothermal or
+isentropic).  In both cases hydrostatic equilibrium is enforced as:
 \begin{equation}

### 2013-05-16 Docs/introduction/introduction.tex
update refs
add Gilet paper

--- 2012-04-12 docs/introduction/introduction.tex
@@
 \item {\em The Convective Phase Preceding Type Ia Supernovae,}
-  M. Zingale, A. S. Almgren, J. B. Bell, C. M. Malone, A. Nonaka, \& S. E. Woosley 2011,
+  M. Zingale, A. S. Almgren, J. B. Bell, C. M. Malone, A. Nonaka, \& S. E. Woosley 2011, ApJ,
   740, 8 (henceforth paper V)
@@
   A. Nonaka, M. Zingale, A. J. Aspden, A. S. Almgren, J. B. Bell, \& S. E. Woosley 2012, 
-  745, 73
+  ApJ, 745, 73
+\item {\em Low Mach Number Modeling of Convection in Helium Shells on
+  Sub-Chandrasehkar White Dwarfs. I. Methodology,} M.~Zingale,
+  A.~Nonaka, A.~S.~Almgren, J.~B.~Bell, C.~M.~Malone, \&
+  R.~J.~Orvedahl 2013, ApJ, 764, 97
+\item {\em Low-Mach Number Modeling of Core Convection in Massive Stars ,}
+  C.~Gilet, A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, S.~E.~Woosley, \& M. Zingale
+  2013, ApJ, submitted.
 \end{itemize}

### 2013-05-17 Docs/introduction/introduction.tex
typo

@@
 \noindent We have several papers that describe applications of the
-method to Type Ia supernovae and X-ray bursts:
+method to Type Ia supernovae, X-ray bursts, and stellar evolution:
 \begin{itemize}
@@
   R.~J.~Orvedahl 2013, ApJ, 764, 97
-\item {\em Low-Mach Number Modeling of Core Convection in Massive Stars ,}
+\item {\em Low-Mach Number Modeling of Core Convection in Massive Stars,}
   C.~Gilet, A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, S.~E.~Woosley, \& M. Zingale

### 2013-05-17 Docs/architecture/architecture.tex
update to show initial_models
more docs

@@
       Linear algebra routines.
+
+    \item {\tt initial\_models/}
+
+      Simple routines for generating toy initial models in hydrostatic equilibrium.
 
@@
 
+By separating the layout from the actual data, we can allocate and
+destroy data that lives on the grid as needed.
+
 \subsection{\fab}
@@
 to the refinement level.
+
+All \fab s in a given \multifab\ have the same number of ghost cells,
+but different \multifab s can have different numbers of ghost cells
+(or no ghost cells).
 
@@
 
+The main purpose of this module is to define the Fortran kind {\tt dp\_t}
+which is used throughout the code to declare double precision variables.
+
 \subsection{{\tt bl\_constants}}
 
+This module provides descriptive names for a number of common double precision
+numbers, e.g.\ {\tt ONE = 1.0\_dp\_t}.  This enhances the readability of
+the code.
+
 \subsection{{\tt parallel}}
+
+All MPI calls are wrapped by functions in the {\tt parallel} module.  For
+serial jobs, the wrappers simply do the requested operation on processor.
+By wrapping the calls, we can easily switch between serial and parallel
+builds.
 

### 2013-05-22 Docs/visualization/visualization.tex
start of some file format documentation

@@
 \renewcommand{\arraystretch}{1.0}
+
+\subsection{the \boxlib\ file format}
+
+\maestro\ stores the plotfile data in a hierarchical directory format,
+with each level's data stored in a separate subdirectory.  Some meta-data
+is stored in the top-level to help interpret the structure.  The basic format is:
+\begin{itemize}
+\item {\tt plt00000/}
+  \begin{itemize}
+    \item {\tt Header}
+    \item {\tt job\_info}
+    \item {\tt Level\_00/}
+      \begin{itemize}
+        \item {\tt Cell\_D\_00000}
+        \item {\ldots}
+        \item {\tt Cell\_H}
+      \end{itemize}
+    \item {\tt model\_cc\_00000}
+    \item {\tt model\_ec\_00000}
+  \end{itemize}
+\end{itemize}
+%
+In the main directory, {\tt Header} contains the information required
+to interpret the data stored on disk.  We describe this below.  The
+{\tt job\_info} file is a plaintext file that contains a lot of
+information about the run (where it was run, when it was run, compiler
+options, runtime options, etc.).  It is not needed to interpret the
+data.  Finally, the {\tt model\_cc\_00000} and {\tt model\_ec\_00000}
+contain the \maestro\ basestate information for the cell-centered and
+edge-centered basestate quantities respectively.  These files are not
+typically used for visualization.
+
+\subsubsection{\tt Header}
+
+The main {\tt Header} is written by {\tt fabio\_ml\_multifab\_write\_d}
 

### 2013-07-03 Docs/mg/mg.tex

--- 2012-07-24 docs/mg/mg.tex
@@
 \end{itemize}
+
+In addition to these parameters, the \mgtower\ carries a number of
+\multifab s that carry the solution and stencil for the multigrid
+solve.
+
+The stencil itself is contained in the {\tt ss} \multifab---for each
+zone, this gives the coefficients of the terms in the Laplacian, with
+the convection that the `0' term is located in the current cell and
+the other terms are the $\pm 1$ off the current cell in each direction.
 

### 2013-07-03 Docs/visualization/visualization.tex

@@
 
-The main {\tt Header} is written by {\tt fabio\_ml\_multifab\_write\_d}
+The main {\tt Header} is written by {\tt fabio\_ml\_multifab\_write\_d} by
+processor 0.  The information contained is the following:
+\begin{quote}
+NavierStokes-V1.1 \\
+{\em number of variables} \\
+{\em variable 1 name} \\
+{\em variable 2 name} \\
+$\vdots$ \\
+{\em last variable name} \\
+{\em number of dimensions} \\
+{\em simulation time} \\
+{\em number of levels} (0-based) \\
+{\em physical domain minimum coordinate} (dm numbers) \\
+{\em physical domain maximum coordinate} (dm numbers) \\
+{\em jump in refinement between level 0 and 1} \\
+$\vdots$ \\
+{\em jump in refinement between level $n-2$ and $n-1$} \\
+~\\
+
+\end{quote}
 

### 2013-08-02 Docs/MESA/instructions.tex
latest MESA docs

@@
 {\tt http://mesa.sourceforge.net}. Check the website for the latest version, 
-which is 4849 when the last upgrades to this document were added. The codes 
+which is 5271 when the last upgrades to this document were added. The codes 
 and algorithms 
@@
 upgrades between {\sf MESA} versions that may cause the routines and algorithms 
-discussed here to crash. To download the source code associated with version 
-4849, issue the following command:
+discussed here to crash. These notes are aimed at working with 
+version 4942, to download the source code associated with this version, 
+issue the following command:
 \begin{lstlisting}[language=bash,mathescape=false]
-  svn co -r 4849 svn://svn.code.sf.net/p/mesa/code/trunk mesa
+  svn co -r 4942 svn://svn.code.sf.net/p/mesa/code/trunk mesa
 \end{lstlisting}
@@
 in the directory where this command was run. This step might take some time 
-($\sim$3 min depending on the internet connection) possibly because the 
-final {\tt mesa} directory will take up roughly 2.1 GB.
+($\sim$1-3 min depending on the internet connection) possibly because the 
+final {\tt mesa} directory will take up roughly 1.8 GB.
 
@@
 \end{lstlisting}
-The instructions in this document do not use the SDK in order to avoid 
-automatically changing the {\tt PATH} variable.
+The instructions in this document {\bf \emph{do not}} use the SDK in order 
+to avoid automatically changing the {\tt PATH} variable.
+
+A quick outline for compiling {\sf MESA} for use with {\sf MAESTRO} is 
+shown below. Detailed instructions are discussed in Section \ref{sec:woSDK}.
+\begin{enumerate}
+\item Set environment variables, {\tt MESA\_DIR} and {\tt OMP\_NUM\_THREADS}
+\item Change name of makefile
+\item Makefile: choose compiler, comment out second instance of 
+{\tt CC = gcc}, turn off {\tt USE\_PGSTAR} and {\tt LOAD\_PGPLOT}
+\item Modify the {\tt const/public/const\_def.f} to fix the 
+{\tt omp\_get\_max\_threads()} line
+\item Run {\tt install} script
+\item Turn off OMP
+\item Recompile using the {\tt compile-all-mods.sh} script
+\end{enumerate}
 
@@
 A few environment variables need to be set (commands are shown for the 
-Bourne shell):
+BASH shell):
 \begin{enumerate}
@@
 located in the {\tt mesa/utils} directory and is called 
-{\tt makefile\_header\_non\_mesasdk}.
+{\tt makefile\_header}. All {\sf MESA} modules have a makefile that 
+includes this main makefile. The file that needs to be edited is named 
+{\tt makefile\_header\_non\_mesasdk} and so the name must be changed in order 
+for the edits to take effect. This can be done as follows:
+\begin{lstlisting}[language=bash,mathescape=false]
+  cp makefile_header_non_mesasdk makefile_header
+\end{lstlisting}
 
-All of the following changes apply to 
-variables that already exist. There is no need to explicitly add new code 
+All of the changes listed below apply to variables that already exist. 
+There is no need to explicitly add new code 
 unless you intend to compile {\sf MESA} with something other than 
@@
 
-Open the {\tt makefile\_header\_non\_mesasdk} and change the following:
+Open the {\tt makefile\_header} file and change the following:
 \begin{enumerate}
@@
 
-\item Remove the second instance of {\tt CC} located directly under the 
-{\tt SPECIAL\_FC\_FLAGS} variable
+\item Remove (or comment out) the second instance of {\tt CC} located 
+directly under the {\tt SPECIAL\_FC\_FLAGS} variable
 
@@
 
-\item {\tt BLAS} and {\tt LAPACK}: Set the location of {\tt LAPACK} and 
-{\tt BLAS} provided with the {\sf MESA} source tree using the following 
-definitions (these should be the default settings):
+\item {\tt BLAS} and {\tt LAPACK}: Make sure the location of {\tt LAPACK} and 
+{\tt BLAS} (provided with the {\sf MESA} source tree) are set using the 
+following definitions (these should be the default settings):
 \begin{lstlisting}[language=make,mathescape=false,upquote=true]
@@
 
-\item {\tt USE\_SE}: Set the variable {\tt USE\_SE} to {\tt NO} and set both 
-{\tt LOAD\_SE} and {\tt INCLUDE\_SE} to be empty. These also need to be set 
-in the correct if-construct according to which compiler was chosen. 
+\item {\tt USE\_SE}: Make sure the variable {\tt USE\_SE} is set to {\tt NO} 
+and both {\tt LOAD\_SE} and {\tt INCLUDE\_SE} are empty. These also need to 
+be set in the correct if-construct according to which compiler was chosen. 
 {\tt SE} is a library of 
@@
 {\tt omp\_get\_max\_threads} occurs in the {\tt const\_def.f} file located in 
-{\tt const/public/const\_def.f} when compiling in serial.
+{\tt const/public/const\_def.f} when running the code in serial.
 
@@
 \end{enumerate}
-
-All {\sf MESA} modules have a makefile that includes the main makefile named 
-{\tt makefile\_header}. The file that was just edited was named 
-{\tt makefile\_header\_non\_mesasdk} and so the name must be changed in order 
-for the edits to take effect. This can be done as follows:
-\begin{lstlisting}[language=bash,mathescape=false]
-  cp makefile_header_non_mesasdk makefile_header
-\end{lstlisting}
 
@@
 Change the {\tt FCopenmp} flag ({\tt gfortran} default = {\tt -fopenmp}) to 
-the appropriate value so as to disable OpenMP, e.g. with {\tt gfortran} 
-{\tt FCopenmp} should be empty.
+the appropriate value so as to disable OpenMP, e.g.\ with {\tt gfortran} 
+{\tt FCopenmp} should be empty. Now all modules must be recompiled. This is 
+most easily accomplished with the scripts described in Section 
+\ref{sec:compile_scripts}.
 
@@
 the {\tt ./export} script. Section \ref{sec:compile_scripts} discusses a few 
-simple scripts that are meant to help with compiling modules.
+simple scripts that are meant to help with compiling the modules.
 
@@
 
-The {\sf MESA} libraries have the name {\tt lib(module-name).a}, e.g. 
+The {\sf MESA} libraries have the name {\tt lib(module-name).a}, e.g.\ 
 {\tt libeos.a} or {\tt libnet.a}. The libraries need to be linked using the 
 {\tt -L} flag with the {\tt mesa/lib} directory and a list of which libraries 
-to load, e.g. {\tt -L\$(MESA\_DIR)/lib -leos -lnet -lnum}. {\sf MESA} 
+to load, e.g.\ {\tt -L\$(MESA\_DIR)/lib -leos -lnet -lnum}. {\sf MESA} 
 {\tt .mod} files are linked to the code using the {\tt -I} flag and the 
-{\tt mesa/include} directory, e.g. {\tt -I\$(MESA\_DIR)/include}.
+{\tt mesa/include} directory, e.g.\ {\tt -I\$(MESA\_DIR)/include}.
 
@@
 The wrapper routines that allow {\sf MAESTRO} to use {\sf MESA} modules are 
-located in the {\tt AstroDev/EOS/MESA} and {\tt AstroDev/networks/MESA} 
+located in the {\tt AstroDev/EOS/MESA}, {\tt AstroDev/networks/MESA} and 
+{\tt AstroDev/networks/MESA\_SDC} 
 directories. In each directory there is a {\tt GPackage.mak} file that 
@@
 \end{lstlisting}
-To use the network, add the following lines:
+To use the Strang-Split version of the network, add the following lines:
 \begin{lstlisting}[language=make,mathescape=false]
@@
   NETWORK_DIR := MESA
-  GENERAL_NET_INPUTS := $(NETWORK_TOP_DIR)/MESA/Net-Files/net_input.net
+  GENERAL_NET_INPUTS := $(NETWORK_TOP_DIR)/$(NETWORK_DIR)/Net-Files/net_input.net
+\end{lstlisting}
+To use the Spectral Deferred Corrections (SDC) version of the network add 
+these lines instead:
+\begin{lstlisting}[language=make,mathescape=false]
+  SDC := t
+
+  NETWORK_TOP_DIR := $(ASTRODEV_DIR)/networks
+  NETWORK_DIR := MESA_SDC
+  GENERAL_NET_INPUTS := $(NETWORK_TOP_DIR)/$(NETWORK_DIR)/Net-Files/net_input.net
 \end{lstlisting}
@@
 \section{Using {\sf MESA} Modules}
+\label{sec:using_modules}
 
@@
 The routines that carry out these processes (setting up EOS, setting up 
-isotope map, shutdown routines, etc.) have already been written and live in 
-{\sf AstroDev/networks/MESA} and {\sf AstroDev/EOS/MESA}. This section 
+isotope map, shutdown routines, etc.) {\bf \emph{have already been written}} 
+and live in {\tt AstroDev/networks/MESA}, {\tt AstroDev/networks/MESA\_SDC} 
+and {\tt AstroDev/EOS/MESA}. This section 
 outlines the necessary steps taken by those routines in order to set up the 
@@
 atomic quantities such as atomic weight. The ``approx'' file contains 
-approximate values for these quantities, e.g. the mass of the proton and 
+approximate values for these quantities, e.g.\ the mass of the proton and 
 neutron are both equal to 1.0. The {\tt ierr} argument is an integer that 
@@
 \subsection{Network}
+There are currently two different algorithms that integrate the nuclear 
+reaction network in {\sf MAESTRO}; Strang Splitting and Spectral Deferred 
+Corrections (SDC). The routines to go from {\sf MAESTRO} 
+to {\sf MESA} using Strang are stored in {\tt AstroDev/networks/MESA} while 
+the routines that implement the SDC method are stored in 
+{\tt AstroDev/networks/MESA\_SDC}.
+
+The setup/shutdown procedures for both methods are very similar. The largest 
+difference is that the Strang method uses {\sf MESA} to obtain the right 
+hand sides of the network equations \emph{and} do the integration, while the 
+SDC method uses {\sf MESA} to obtain the right hand sides and uses 
+{\sf MAESTRO} to integrate the equations using {\tt VODE}.
 
@@
 There are significantly more steps involved when initializing the {\sf MESA} 
-network. Here is a brief outline of what needs to happen:
+network. Here is a brief outline of what happens:
 \begin{enumerate}
@@
 \item Generate a file that will tell {\sf MESA} which reactions to add based 
-on the isotopes in the current network ({\tt call which\_reactions})
-\item Initialize {\sf MESA} {\tt chem}, {\tt weak} and {\tt reaclib} modules
-\item Setup the map between {\sf MAESTRO} and {\sf MESA}, see Section 
-\ref{sec:mapping} for a more in depth explanation
+on the isotopes in the current network
+\item Initialize various {\sf MESA} modules
+\item Setup the map between {\sf MAESTRO} and {\sf MESA} (see Section 
+\ref{sec:mapping} for a more in depth explanation)
 \item Initialize {\sf MESA} {\tt rates} and {\tt net} modules
@@
 {\tt net\_set\_which\_rates}
-\item Initialize which linear algebra solver to use
+\item Call {\tt net\_setup\_tables}
+\item Set which linear algebra solver to use
+\item Set the ODE solver to use for the network integration (Strang-Split 
+version only)
 \end{enumerate}
-
-In the {\tt Do\_One\_Zone\_Burn.f90} routine, there are two more ``set up'' 
-routines: {\tt set\_rattab\_range} and {\tt net\_setup\_tables}. I have not 
-found a good way to call these routines only once with the set up of all the 
-other routines because the {\tt set\_rattab\_range} routine depends on 
-temperature. As a result, both routines are called every time a burn is to 
-take place.
 
@@
 
-Calling the {\sf MESA} network from within the {\sf MAESTRO} wrapper starts 
-with calling a modified {\sf MESA} wrapper:
+In the case of Strang Splitting the call to the {\sf MESA} burner looks like 
+this:
 \begin{lstlisting}[language=fortran,mathescape=false]
-  call Do_One_Zone_Burn(dens, temp, dt, Xin, burn_ergs, Xout)
+  call net_1_zone_burn( &
+        handle, solver_choice, species, num_reactions, 0.0_dp_t, burn_tend, &
+        xin, clip, num_times, times, log10Ts_f1, log10Rhos_f1, etas_f1, &
+        dxdt_source_term, rate_factors, category_factors, std_reaction_Qs, &
+        std_reaction_neuQs, screening_mode, theta_e_for_graboske_et_al, &
+        h, max_step_size, max_steps, rtol, atol, itol, &
+        decsol_choice, caller_id, burn_solout, iout, ending_x, &
+        nfcn, njac, nstep, naccpt, nrejct, time_doing_net, ierr)
 \end{lstlisting}
-The inputs are density, temperature, how long to burn and starting mass 
-fractions. The outputs are the energy generation rate ({\tt burn\_ergs}) 
-in units of ergs g$^{-1}$ s$^{-1}$ and the final mass fractions.
+The inputs are density, temperature, how long to burn, starting mass 
+fractions, various rates and parameters as well as some things to control the 
+integration. The outputs are the energy ({\tt burn\_ergs}) in units of ergs/g, 
+the ending mass fractions and various counters such as number of jacobian 
+evaluations. This routine will calculate the right hand sides, pass them 
+to a stiff ODE integrator and return the final results.
 
-To calculate {\tt rho\_omegadot} ($\rho \dot{\omega}$) and {\tt rho\_Hnuc} 
-($\rho H_{{\rm nuc}}$) the following code can be used after a call to 
-{\tt Do\_One\_Zone\_Burn}:
+In the case of the SDC method, the call to the {\sf MESA} routine that returns 
+the instantaneous right hand sides looks like this:
 \begin{lstlisting}[language=fortran,mathescape=false]
-  do i=1, nspec
-     dX = Xout(i) - Xin(i)
-     rho_omegadot(i) = dens * dX / dt
-  enddo
-
-  rho_Hnuc = dens * burn_ergs
+  call burn_derivs(nspec, t, xmass(1:nspec), ydot(1:nspec), lrpar, rpar, &
+                   lipar, ipar, ierr)
 \end{lstlisting}
-There is a support routine called {\tt burn\_solout} that is passed to the 
-{\sf MESA} burner. This routine is called after every step of the burn and 
-is responsible for calculating the energy generation rate. {\tt burn\_solout} 
-is passed as an argument to the {\tt net\_1\_zone\_burn} subroutine (the 
-{\sf MESA} burner) and as a result the actual call to the {\tt burn\_solout} 
-routine is buried deep within the {\sf MESA} source code.
+The inputs are mass fraction, work arrays and the time. The output is the 
+time derivative of the mass fraction. This routine is only responsible for 
+calculating the right hand sides of the network equations, {\tt VODE} is tasked 
+with the integration in this algorithm.
 
@@
 Some useful scripts have been written in {\tt BASH} to help with compiling 
-{\sf MESA} modules. They live in the {\sf AstroDev/networks/MESA/scripts} 
+{\sf MESA} modules. They live in the {\tt AstroDev/networks/MESA/scripts} 
 directory. Currently, there are two scripts; one that compiles all modules 

### 2013-08-05 Docs/runtime_parameters/runtime_parameters.tex
update + fix path to _parameters

--- 2012-10-02 docs/runtime_parameters/runtime_parameters.tex
@@
 \verb=  plot_sponge_fdamp  = &   plot fdamp rather than sponge assumes sponge has the form 1/(1+dt*{\tt sponge\_kappa}*fdamp)  &  .false. \\
+\verb=  ppm_trace_forces  = &   if 1, then perform parabolic reconstruction on the forces used in the prediction and trace under the parabola to the interfaces the amount that can reach the interface over dt  &  0 \\
+\rowcolor{tableShade}
 \verb=  ppm_type  = &   0 = no ppm (piecewise linear slopes instead) \newline 1 = 1984 ppm \newline 2 = Hybrid Sekora/Colella and McCorquodale/Colella 2009/2010 ppm  &  1 \\
-\rowcolor{tableShade}
 \verb=  restart_with_vel_field  = &   restart the simulation using a result from a {\tt drive\_initial\_convection} = T run note that this uses the restart variable to specify which file to restart from.
@@
 \\
+\rowcolor{tableShade}
 \verb=  rotation_radius  = &   radius used for computing centrifugal term in rotation problems  &  1.0d6 \\
+\verb=  rotational_frequency  = &   rotational frequency used for computing centrifugal term in rotation problems.  &  ZERO \\
 \rowcolor{tableShade}
-\verb=  rotational_frequency  = &   rotational frequency used for computing centrifugal term in rotation problems.  &  ZERO \\
 \verb=  slope_order  = &   order of slopes in piecewise linear Godunov algorithm.  Options are 0, 2, or 4.  &  4 \\
-\rowcolor{tableShade}
 \verb=  species_pred_type  = &   Which quantities do we predict to the edges for computing the ($\rho X$) edge states?
@@
 &  1 \\
+\rowcolor{tableShade}
 \verb=  sponge_center_density  = &   Center of the inner sponge.  &  3.d6 \\
+\verb=  sponge_kappa  = &   Parameter for sponge.  Problem dependent.  &  10.d0 \\
 \rowcolor{tableShade}
-\verb=  sponge_kappa  = &   Parameter for sponge.  Problem dependent.  &  10.d0 \\
 \verb=  sponge_start_factor  = &   The sponge begins at sponge\_center\_density * sponge\_start\_factor.  &  3.333d0 \\
+\verb=  stop_initial_convection  = &   timestep beyond which we set {\tt drive\_initial\_convection} = F  &  -1 \\
 \rowcolor{tableShade}
@@
 \rowcolor{tableShade}
+\verb=  hg_cycle_type  = &   Type of cycle used in the nodal multigrid -- 1 = F-cycle, 2 = W-cycle, 3 = V-cycle  &  3 \\
 \verb=  hg_dense_stencil  = &   In hgproject, in 2D, use a 9 point Laplacian (.true.) or 5-point Laplacian (.false.).
@@
 \\
+\rowcolor{tableShade}
 \verb=  max_mg_bottom_nlevels  = &   if mg\_bottom\_solver == 4, then how many mg levels can the bottom solver mgt object have  &  1000 \\
+\verb=  mg_bottom_nu  = &   number of smoothing iterations to do after the multigrid bottom solver  &  10 \\
 \rowcolor{tableShade}
-\verb=  mg_bottom_nu  = &   number of smoothing iterations to do after the multigrid bottom solver  &  10 \\
 \verb=  mg_bottom_solver  = &   valid values are $\le$ 0  &  -1 \\
+\verb=  mg_cycle_type  = &   Type of cycle used in the MAC multigrid -- 1 = F-cycle, 2 = W-cycle, 3 = V-cycle  &  3 \\
+\rowcolor{tableShade}
+\verb=  mg_nu_1  = &   number of smoothing iterations to do going down the V-cycle  &  2 \\
+\verb=  mg_nu_2  = &   number of smoothing iterations to do going up the V-cycle  &  2 \\
 \rowcolor{tableShade}
@@
 \verb=  conductivity_constant  = &   the constant conductivity (only applies if the conductivity/constant module is used)  &  1.0d0 \\
+\verb=  limit_conductivity  = &   apply the conductivity limiting---if T, then we set the thermal coefficients all to 0 for $\rho <$ {\tt buoyancy\_cutoff\_factor} * {\tt base\_cutoff\_density}  &  .false.
+\\
+\rowcolor{tableShade}
 \verb=  temp_diffusion_formulation  = &   How to compute the explicit thermal diffusion term.  \newline 1 = in terms of $T$; \newline 2 = in terms of $\rho,p_0,X$.  &  2 \\
+\verb=  thermal_diffusion_type  = &   In the thermal diffusion solver, 1 = Crank-Nicholson; 2 = Backward Euler.  &  1 \\
 \rowcolor{tableShade}
-\verb=  thermal_diffusion_type  = &   In the thermal diffusion solver, 1 = Crank-Nicholson; 2 = Backward Euler.  &  1 \\
 \verb=  use_thermal_diffusion  = &   Use thermal diffusion.  &  .false. \\

### 2013-08-05 Docs/runtime_parameters/rp.py

@@ -58,7 +58,7 @@ Mfooter=r"""
 """
 
 fParallel="../../.."
-paramFile=fParallel + "/MAESTRO/_parameters"
+paramFile=fParallel + "/MAESTRO/Source/_parameters"
 
 
 # container class for the parameters

### 2013-09-13 Docs/getting_started/getting_started.tex
update path for python plotting routines
add a note about no "~"

@@
 to your {\tt .cshrc}.
+
+Note: you must specify the full path to the BoxLib directory.  Do not
+use ``$\sim$'' to refer to your home directory---the scripts used by
+the build system will not be able to process this.
 
@@
 There is also a python visualization method in {\tt
-AmrPostprocessing/F\_src/python\_plotfile}.  This is described
+AmrPostprocessing/python}.  This is described
 in \S~\ref{sec:vis:python}.

### 2013-09-17 Docs/visualization/visualization.tex
add some details about using Amrvis

@@
 \section{Visualizing with \amrvis}
+
+\amrvis\ is a tool developed together with \boxlib\ to visualize datasets
+from codes built around the \boxlib\ library.  You can download the
+\amrvis\ source from: \\
+\url{https://ccse.lbl.gov/Downloads/downloadAmrvis.html} \\
+
+\amrvis\ exists in the C++ \boxlib\ framework, so the build system is
+slightly different.  A different executable is needed for 2- vs.\ 3-d
+datasets.  Edit the {\tt GNUmakefile} and set the compilers (probably
+{\tt g++} and {\tt gfortran}) and the dimensionality, and turn off any
+of the volume rendering options.  You will need to have the Motif library
+installed on your system (or a replacement, such as {\tt lesstif}.
+
+Once the code is built, you visualize a dataset as:
+\begin{verbatim}
+amrvis3d.Linux.g++.gfortran.ex pltfile
+\end{verbatim}
+where {\tt pltfile} is the name of the plotfile directory.  Different
+variables can be selected from the drop down menu at the top.  Middle
+and right clicking in 3-d select the slice planes, and shift + middle
+or right will extract 1-d lines through the data.  In 2-d, middle and
+right clicking alone extract 1-d lines.
+
+If \amrvis\ cannot find the {\tt Palette} file, then the plots will be
+in grayscale.  To fix this, copy the {\tt amrvis.defaults} and {\tt
+  Palette} files to your home directory and edit {\tt amrvis.defaults} so that
+the {\tt palette} line points to the {\tt Palette} file, e.g.:
+\begin{verbatim}
+palette               /home/username/Palette
+\end{verbatim}
 

### 2013-10-05 Docs/MESA/instructions.tex
hook MESA chapter in
remove upquote in MESA/instructions.tex for now -- needs later latex

@@
 following definitions (these should be the default settings):
-\begin{lstlisting}[language=make,mathescape=false,upquote=true]
+\begin{lstlisting}[language=make,mathescape=false]%,upquote=true]
   WHICH_LAPACK = USE_SRCS
@@
 \label{sec:hopper_install}
-\input hopper_compile
+\input \mesafigpath/hopper_compile
 
@@
 \end{lstlisting}
-
-\end{document}

### 2013-10-07 Docs/managing_jobs/managingjobs.tex
updates

@@
 \item The tolerances for the MG solves
-\item The value of all runtime parameters (even those that were not explicitly set in the inputs file)
+\item Any warnings from the initialization procedure (note: these are not currently stored on restart).
+\item The value of all runtime parameters (even those that were not explicitly set in the inputs file), along with an indicator showing if the default value
+was overridden.
 \end{itemize}

### 2013-12-13 Docs/introduction/introduction.tex
some text about the new energy formulation

@@
   C.~Gilet, A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, S.~E.~Woosley, \& M. Zingale
-  2013, ApJ, submitted.
+  2013, ApJ, 773, 137
 \end{itemize}
@@
 base state.
+
+\subsection{Alternate Energy Formulation}
+
+Several authors~\cite{KP:2012,VLBWZ:2013} showed that with a slightly
+different momentum equation, the low Mach number system can conserve
+an energy (that is, a quantity that looks like the compressible
+energy, but formed using the low Mach number quantities).  This change
+manifests itself as either a change to the buoyancy term or by
+changing $\nabla \pi$ to $\beta_0 \nabla (\pi/\beta_0)$.  Furthermore,
+\cite{VLBWZ:2013} showed that the new formulation better captures the
+vertical propagation of gravity waves.  We explored this in more
+detail and from an asymptotic approach in \cite{MaestroEnergy}.  As of
+Dec.\ 2013, this new formulation is the default in \maestro.
 

### 2014-01-02 Docs/compilers/compilers.tex
renname docs Docs
update for Cray 8.2.x

--- 2012-08-21 docs/compilers/compilers.tex
@@
 machine at OCLF.
-\item Cray 5.12 compiles and runs \maestro\ fine on hopper (8/21/12).
+\item Cray 8.0.7 compiles and runs \maestro\ fine on hopper (08/21/12).
+\item Cray 8.2.1 and 8.2.2 seem to have issues with OpenMP on hopper (01/02/14).
+Basically, in the initial divu iters, we appear to hang.
+Compiling in debug mode works fine, but is slow.
+This suggests that there is a compiler bug.
 \end{itemize}

### 2014-01-03 Docs/compilers/compilers.tex
multifab_volume note
more compiler notes

@@
 \item Cray 8.0.7 compiles and runs \maestro\ fine on hopper (08/21/12).
-\item Cray 8.2.1 and 8.2.2 seem to have issues with OpenMP on hopper (01/02/14).
-Basically, in the initial divu iters, we appear to hang.
-Compiling in debug mode works fine, but is slow.
-This suggests that there is a compiler bug.
+
+\item Cray 8.1.3, 8.2.1, and 8.2.2 seem to have issues with OpenMP on hopper 
+ (01/02/14).  Basically, in the initial divu iters, we appear to hang.
+ Compiling in debug mode works fine, but is slow.  This suggests that
+ there is a compiler bug.
+
+ A separate issue is that the ``volume'' reported by {\tt
+ multifab\_volume()} seems to be wrong.  For a single level run, it
+ should just be $\mathtt{nx} \times \mathtt{ny} \times \mathtt{nz}$,
+ but the value reported is incorrect.
 \end{itemize}

### 2014-01-10 Docs/architecture/architecture.tex
Added to the documentation an explanation of each box's meaning in a fab.

@@
 multidimensional array and several \boxtype-types to describe where in
-the global index-space it lives.  
-Note that all state data is stored in a four-dimensional array,
-{\tt (nx,ny,nz,nc)} in size, regardless of the dimensionality of the
-problem.  Here {\tt nc} is the number of components, for instance
-representing different fluid variables.  For 2D problems, {\tt nz=1}.
+the global index-space it lives:
+\begin{lstlisting}[language={[95]fortran},mathescape=false]
+  type fab
+     ...
+     type(box) :: bx
+     type(box) :: pbx
+     type(box) :: ibx
+  end type fab
+\end{lstlisting}
+{\tt bx} represents the box in the global index-space over which the 
+\fab\ is defined, {\tt pbx} represents the ``physical'' box in the sense
+that it includes {\tt bx} plus ghost cells, and {\tt ibx} is the same as {\tt bx} unless 
+the \fab\ is nodal.  As can be seen in Figure~\ref{fig:dataloc},
+for the same grid nodal data requires one more array element than 
+cell-centered data.  To address this {\tt ibx} is made by growing
+{\tt bx} by one element along all nodal dimensions.
+
+It's important to note that all state data is stored in a four-dimensional array
+{\em regardless of the problem's dimensionality}.  The array is 
+{\tt (nx,ny,nz,nc)} in size, where {\tt nc} is the number of components, for instance
+representing different fluid variables, and {\tt (nx,ny,nz)} are the number of cells
+in each respective spatial dimension.  For 2D problems, {\tt nz=1}.
 

### 2014-01-15 Docs/visualization/visualization.tex
fix copy-paste error for omegadot
some yt cleaning

@@
 %\hline
-{\tt omegadot(*)}      & $(\rho X_k)/\rho$ & {\tt plot\_omegadot} \\
+{\tt omegadot(*)}      & $\dot{\omega}_k$ & {\tt plot\_omegadot} \\
 %\hline
@@
 from {\tt FLASH}, and data from the SPH code {\tt Gadget}.  For more
-information, see the \yt homepage at {\tt http://yt.enzotools.org/}.
+information, see the \yt homepage at \url{http://yt-project.org/} and
+\cite{yt}.
+
+\noindent {\bf Note: for \maestro\ data, you need to use \yt 3.0 or later.}
 
@@
 By default, this \yt install script will download and install, in its
-own isolated environment, all the secondary utilities needed to get \yt
-up and running.  Note that this currently includes installing {\it
-  hdf5 1.6.9, zlib 1.2.3, bzip2 1.0.5, libpng 1.2.43, freetype 2.4.4,
-  Python 2.6.3, numpy 1.5.1, matplotlib 1.0.0, mercurial 1.7.1,
-  ipython 0.10, h5py 1.2.0, Cython 0.14, Forthon 0.8.4} as well as a
-\yt {\it mercurial} bundle of changes.  You can turn off the automatic
+own isolated environment, all the secondary utilities needed to get
+\yt up and running.  Note that this currently includes installing {\it
+  hdf5, zlib, bzip2, libpng, freetype, python, numpy, matplotlib,
+  mercurial, ipython, h5py, Cython, Forthon} as well as a \yt {\it
+  mercurial} bundle of changes.  You can turn off the automatic
 installation of any of these particular packages by setting the
-appropriate {\tt INST\_}* variable to zero in the install script;
-you may have to then change some of *{\tt \_DIR} variables to point to
-your own particular installation of that package.  I usually just let
-\yt install its own stuff, which ensures things are working properly.
+appropriate {\tt INST\_}* variable to zero in the install script; you
+may have to then change some of *{\tt \_DIR} variables to point to
+your own particular installation of that package.  It is usually best
+to just let \yt install its own stuff, which ensures things are
+working properly.
 
-After the install script has finished, and assuming you let \yt install its own packages, you'll need to {\it prepend} some environment variables with \yt locations so that your system finds those first and stops looking.
-For example, the \yt specific information in my {\tt .bashrc} file looks like this
-\begin{lstlisting}[language=bash,mathescape=false]
-# yt stuff
-export YT_DEST=~/install/yt/current/yt-x86_64
-export PATH=${YT_DEST}/bin/:$PATH
-export PYTHONPATH=${YT_DEST}/lib/python2.6/site-packages/:$PYTHONPATH
-export LD_LIBRARY_PATH=${YT_DEST}/lib/:$LD_LIBRARY_PATH
-\end{lstlisting}
-where I have {\tt ~/install/yt/current} as a soft link to my \yt
-installation directory.
+After the install script has finished, and assuming you let \yt
+install its own packages, you'll need to {\it prepend} some
+environment variables with \yt locations so that your system finds
+those first and stops looking.  \yt provides a simple script to
+do this, which will be announced upon successful completion.
+
+%% \begin{lstlisting}[language=bash,mathescape=false]
+%% # yt stuff
+%% export YT_DEST=~/install/yt/current/yt-x86_64
+%% export PATH=${YT_DEST}/bin/:$PATH
+%% export PYTHONPATH=${YT_DEST}/lib/python2.6/site-packages/:$PYTHONPATH
+%% export LD_LIBRARY_PATH=${YT_DEST}/lib/:$LD_LIBRARY_PATH
+%% \end{lstlisting}
+%% where I have {\tt ~/install/yt/current} as a soft link to my \yt
+%% installation directory.
 

### 2014-01-16 Docs/visualization/visualization.tex
hyperrefs
more yt stuff

@@
 \end{verbatim}
-This will make a plot of ``tfromp'' from the plotfile {\tt plt00000} with log scaling,
+This will make a plot of ``{\tt tfromp}'' from the plotfile {\tt plt00000} with log scaling,
 and store the output in {\tt test.png}.  See Figure~\ref{fig:python}.
@@
 \begin{lstlisting}[language=bash,mathescape=false]
-> wget http://hg.enzotools.org/yt/raw-file/stable/doc/install_script.sh
+> wget http://hg.yt-project.org/yt/raw/yt-3.0/doc/install_script.sh
 > bash install_script.sh
@@
 \begin{lstlisting}[language=Python]
-In [1]: pf = load(``plt00166'')
+In [1]: pf = load("plt00166")
 \end{lstlisting}
-This method returns an instance of the {\tt StaticOutput} class.  One
+Note that some older versions of \yt needed an {\tt inputs} file 
+in the same directory as the plotfile, but as of \yt 3.0, all the
+necessary metadata is obtained from the {\tt job\_info} file inside
+the plotfile directory.
+The {\tt load} method returns an instance of the {\tt StaticOutput} class.  One
 of the easiest ways of handling plots is via a {\tt PlotCollection}
@@
 \begin{lstlisting}[language=Python]
-In [4]: pc.add_slice(``tfromp'',0)
-In [5]: pc.add_slice(``tfromp'',1)
-In [6]: pc.add_slice(``tfromp'',2)
+In [4]: pc.add_slice("tfromp",0)
+In [5]: pc.add_slice("tfromp",1)
+In [6]: pc.add_slice("tfromp",2)
 \end{lstlisting}
@@
 \begin{lstlisting}[language=Python]
-In [7]: pc.save(``my_cool_images'')
+In [7]: pc.save("my_cool_images")
 \end{lstlisting}
@@
 variables, log-scaling, annotations, etc. For more information see the
-documentation at {\tt http://yt.enzotools.org/doc/}.
+documentation at \url{http://yt-project.org/docs/dev/}\, .
 
@@
 # the plotfile I'm interested in
-fn = ``plt00166''
+fn = "plt00166"
 
@@
 # add some slices of tfromp
-pc.add_slice(``tfromp'',0)
-pc.add_slice(``tfromp'',1)
-pc.add_slice(``tfromp'',2)
+pc.add_slice("tfromp",0)
+pc.add_slice("tfromp",1)
+pc.add_slice("tfromp",2)
 
 # save our plots to a files with basename
-# ``my_cool_images''
-pc.save(``my_cool_images'')
+# "my_cool_images"
+pc.save("my_cool_images")
 \end{lstlisting}

### 2014-01-22 Docs/preface/preface.tex

@@
 Engineering at Lawrence Berkeley National Laboratory and at Stony
-Brook University.  Development was supported by the SciDAC Program of
+Brook University.  Development is/was supported by the SciDAC Program of
 the DOE Office of Mathematics, Information, and Computational Sciences
@@
 the U.S. Department of Energy under contract No.\ DE-AC02-05CH11231 to
-LBNL and by a DOE/Office of Nuclear Physics grant No.\
-DE-FG02-06ER41448, to Stony Brook.  \\
+LBNL and by a DOE/Office of Nuclear Physics grant Nos.\
+DE-FG02-06ER41448 and DE-FG02-87ER40317 to Stony Brook.  \\
+
+\noindent \maestro\ developers (past and present) include: Ann Almgren, John
+Bell, Candace Gilet, Adam Jacobs, Mike Lijewski, Chris Malone, Andy
+Nonaka, Ryan Orvedahl, Charles Rendleman, \& Michael Zingale
 
 \noindent \maestro\ uses the \boxlib\ library,
-developed at the Center for Computational Sciences and
-Engineering at Lawrence Berkeley National Laboratory.  \boxlib\
-is distributed separately from \maestro\ (see https:/\hspace{-0.25em}/ccse.lbl.gov/BoxLib/ )
+developed at the Center for Computational Sciences and Engineering at
+Lawrence Berkeley National Laboratory.  \boxlib\ is distributed
+separately from \maestro\ (see
+\url{https://ccse.lbl.gov/BoxLib/})
 \\
@@
 \noindent Support is available via the mailing list: {\tt maestro-help@bender.astro.sunysb.edu}.  To subscribe to this list or read the archives, visit
-{\tt http://bender.astro.sunysb.edu/mailman/listinfo/maestro-help}.
+\url{http://bender.astro.sunysb.edu/mailman/listinfo/maestro-help}.

### 2014-03-07 Exec/TEST_PROBLEMS/incomp_shear_jet/base_state.f90
Added notes on the multifab_volume() error and made minor fixes to the incomp_shear_jet test problem.

@@ -92,9 +92,9 @@ contains
        call bl_error("ERROR: Incompressible shear jet is not valid for spherical geometry")
     endif
 
-    !Here's an example of using the parallel module to put the code on hold if
-    !running could result in accessing shared data or resources currently in
-    !use, in this case the terminal.
+    !Here's an example of using the parallel module.  The module assigns one
+    !processor for I/O and parallel_IOProcessor() returns true if the currently
+    !executing processor is the I/O processor.
     if ( parallel_IOProcessor() .and. n == 1) then
        ! Output block for cutoff densities
        write (*,FMT_SEP)

### 2014-03-10 Docs/visualization/visualization.tex
add 2-d yt example

@@
 \section{Visualizing with \yt}
+
 \yt is a Python package for analyzing and visualizing simulation data.
@@
 
+\subsection{2-d datasets}
+
+To visualize 2-d data, you can use the {\tt SlicePlot} function,
+picking the normal direction to be {\tt "z"}:
+\begin{lstlisting}
+from yt.mods import *
+ pf = load("plt00085")
+s = SlicePlot(pf, "z", "tfromp")
+s.save("tfromp.eps")
+\end{lstlisting}
+This generates the figure shown in Figure~\ref{fig:yt2d}.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[height=0.4\textheight]{\visfigpath/tfromp}
+\caption{\label{fig:yt2d} Example slice through 3-d dataset with \yt.}
+\end{figure}
+
 \subsection{Volume Rendering}

### 2014-03-11 Docs/getting_started/getting_started.tex
update the getting started with some new info

@@
 
+\item {\em Get a copy of \maestro}.
+
+If you don't already have a copy of \maestro, you can obtain one
+via {\tt git}\footnote{{\tt git} is an open-source source code version control
+system.  It is a standard part of any Linux distribution.  You may
+need to use your system's package manager to install {\tt git} if it
+is not already present} as:
+\begin{verbatim}
+git clone http://bender.astro.sunysb.edu/Maestro/staging/MAESTRO.git
+\end{verbatim}
+\maestro\ is under active development, so you will want to keep in 
+sync with the changes by periodically pulling from the repository.  Simply
+type
+\begin{verbatim}
+git pull
+\end{verbatim}
+in the {\tt MAESTRO/} directory.
+
 \item {\em Get a copy of \boxlib}.
@@
 build a \maestro\ executable.  \boxlib\ can be obtained via {\tt
-git}\footnote{{\tt git} is an open-source source code version control
-system.  It is a standard part of any Linux distribution.  You may
-need to use your system's package manager to install {\tt git} if it
-is not already present} as:
+git} as:
 \begin{verbatim}
@@
 \end{verbatim}
+
+You will periodic want to update \boxlib\ by doing
+\begin{verbatim}
+git pull
+\end{verbatim}
+in the {\tt BoxLib/} directory.
 
@@
 \visit\ can be found at:\newline
- {\tt https://wci.llnl.gov/codes/visit/home.html}\,. \newline
+ \url{https://wci.llnl.gov/codes/visit/home.html}\,. \newline
 The easiest way to get started with \visit\ is to download a precompiled
@@
 \visit\ to the {\tt Header} file in the plotfile directory.
+
+\subsection{\yt}
+
+\yt\ (version 3.0 and later) can natively read the \maestro\ plotfiles.  See
+the \yt\ documentation or \S~\ref{sec:vis:yt}.
 

### 2014-03-19 Docs/flowchart/flowchart.tex
add the notation needed for the flowchart from the multilevel paper

--- 2012-07-18 docs/flowchart/flowchart.tex
@@
 
-reach\_state
-
-enforce\_hse
+{\bf React State}$[\rho^{\inp},(\rho                                                                   h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp},
+  p_0^{\inp}] \rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp},                                   
+  T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$
+  evolves the species and enthalpy due to reactions through
+  $\Delta t/2$ according to:
+\begin{equation}
+\frac{dX_k}{dt} = \omegadot_k(\rho,X_k,T) ; \qquad
+\frac{dT}{dt}   = \frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right ).
+\end{equation}
+%This system is solved using the stiff ordinary differential equation                                     
+%integration package, {\tt VODE} \citep{vode}.                                                            
+Full details of the
+solution procedure can be found in Paper III. We then define:
+\begin{eqnarray}
+(\rho\omegadot_k)^{\outp} &=& \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\dt/2}, \\
+(\rho h)^{\outp} &=& (\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc)^{\outp} + \frac{\dt}{2} (\rho\Hext)^{\in\
+p}.
+\end{eqnarray}
+where the enthalpy update includes external heat sources $(\rho\Hext)^{\inp}$.
+As introduced in Paper IV, we update the temperature using $T^{\outp} =                                   
+T(\rho^\outp,h^\outp,X_k^\outp)$ for planar geometry or $T^{\outp} =                                      
+T(\rho^\outp,p_0^\inp,X_k^\outp)$ for spherical geometry.
+Note that the density remains unchanged within {\bf React State}, i.e.,
+$\rho^{\outp} = \rho^{\inp}$.
+
+{\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp] \rightarrow
+  [\rho_0^\outp, \rho_0^{\outp,\nph}]$ is the process by which we
+  update the base state density through $\dt$ in time.  We keep the
+  time-centered edge states, $\rho_0^{\outp,\nph}$,
+  since they are used later in discretization of $\etarho$ for planar problems.
+\begin{description}
+\item[planar:] We discretize equation (\ref{eq:Base State Density}) to
+  compute the new base state density,
+\begin{equation}
+\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} \left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} - \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ].
+\end{equation}
+  We compute the time-centered edge states, $\rho_0^{\outp,\nph}$,
+  by discretizing an expanded form of equation (\ref{eq:Base State Density}):
+\begin{equation}
+\frac{\partial \rho_0}{\partial t} + w_0 \frac{\partial \rho_0}{\partial r} = - \rho_0 \frac{\partial w_0}{\partial r},
+\end{equation}
+  where the right hand side is used as the force term.
+\item[spherical:] The base state density update now includes the area factors in the 
+  divergences:
+\begin{equation}
+\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{1}{r_j^2} \frac{\dt}{\dr} \left [ \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} - \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right].
+\end{equation}
+  In order to compute the time-centered edge states, an additional geometric 
+  term is added to the forcing, due to the spherical discretization of 
+  (\ref{eq:Base State Density}):
+\begin{equation}
+\frac{\partial \rho_0}{\partial t} + w_0 \frac{\partial \rho_0}{\partial r} = - \rho_0 \frac{\partial w_0}{\partial r} - \frac{2 \rho_0 w_0}{r}.
+\end{equation}
+\end{description}
+
+{\bf Enforce HSE}$[p_0^{\inp},\rho_0^{\inp}]
+  \rightarrow [p_0^{\outp}]$ has replaced {\bf Advect Base Pressure}
+  from Paper III as the process by which we update the base state
+  pressure.  Rather than discretizing the evolution equation for
+  $p_0$, we enforce hydrostatic equilibrium directly, which is numerically simpler
+  and analytically equivalent.  We first set 
+  $p_{0,j=0}^{\outp} = p_{0,j=0}^{\inp}$ and then update $p_0^\outp$ using:
+\begin{equation}
+p_{0,j+1}^{\outp} = p_{0,j}^{\outp} + \Delta r g_{j+\myhalf}\frac{\left(\rho_{0,j+1}^{\inp}+\rho_{0,j}^{\inp}\right)}{2},
+\end{equation}
+  where $g=g(\rho_0^\inp)$.  As soon as $\rho_{0,j}^\inp < \rho_{\rm cutoff}$, we set 
+  $p_{0,j+1}^\outp = p_{0,j}^\outp$ for all remaining values of $j$.  
+  Then we compare $p_{0,j_{\rm max}}^\outp$ with $p_{0,j_{\rm max}}^\inp$ and offset
+  every element in $p_0^\outp$ so that $p_{0,j_{\rm max}}^\outp = p_{0,j_{\rm max}}^\inp$.  
+  We are effectively using the location where the $\rho_0^\inp$ drops below 
+  $\rho_{\rm cutoff}$ as the starting point for integration.
+
+{\bf Advect Base Enthalpy}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [(\rho h)_0^\outp]$
+  is the process by which we update the base state enthalpy through $\dt$ in time.  
+\begin{description}
+\item[planar:] We discretize equation (\ref{eq:Base State Enthalpy}), neglecting reaction source terms,
+  to compute the new base state enthalpy,
+\begin{equation}
+(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r} \left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} - \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\} + \dt\psi_j^{\inp}.
+\end{equation}
+  We compute the time-centered edge states, $(\rho h)_0^{\nph}$, by discretizing
+  an expanded form of equation (\ref{eq:Base State Enthalpy}):
+\begin{equation}
+\frac{\partial (\rho h)_0}{\partial t} + w_0 \frac{\partial (\rho h)_0}{\partial r} = -(\rho h)_0 \frac{\partial w_0}{\partial r} + \psi.
+\end{equation}
+\item[spherical:]  The base state enthalpy update now includes the area factors 
+  in the divergences:
+\begin{eqnarray}
+(\rho h)_{0,j}^{\outp} &=& (\rho h)_{0,j}^{\inp} \nonumber \\
+&& - \frac{1}{r_j^2} \frac{\dt}{\dr} \left \{ \left[ r^2 (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} - \left[ r^2 (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\} +\dt\psi^{\inp,\nph}.\nonumber\\
+\end{eqnarray}
+  In order to compute the time-centered edge states, an additional geometric 
+  term is added to the forcing, due to the spherical discretization of 
+  (\ref{eq:Base State Enthalpy}):
+\begin{equation}
+\frac{\partial (\rho h)_0}{\partial t} + w_0 \frac{\partial (\rho h)_0}{\partial r} = -(\rho h)_0 \frac{\partial w_0}{\partial r} - \frac{2 (\rho h)_0 w_0}{r} + \psi.
+\end{equation}
+\end{description}
+
+\subsubsection{Computing $w_0$}\label{Sec:Computing w0}
+Here we describe the process by which we compute $w_0$.  The arguments 
+are different for planar and spherical geometries.
+\begin{description}
+
+\item {\bf Compute} {\boldmath $w_0$} {\bf Planar}
+$[\Sbar^{\inp},\gammabar^{\inp}, p_0^{\inp},\psi^{\inp}]\rightarrow [w_0^{\outp}]$:
+
+In Paper III, we showed that $\psi=\etarho g$ for planar geometries, and 
+derived an alternate expression for equation (\ref{eq:w0 divergence}).
+We compute $w_0$ using equation (35) in Paper III using the following 
+discretization:
+\begin{equation}
+\frac{w_{0,j+\myhalf}^\outp-w_{0,j-\myhalf}^\outp}{\Delta r} = \left(\Sbar^{\inp} - \frac{1}{\gammabar^{\inp} p_0^{\inp}}\psi^{\inp}\right)_j,
+\end{equation}
+with $w_{0,-\myhalf}=0$.
+
+\item {\bf Compute} {\boldmath $w_0$} {\bf Spherical}
+  $[\Sbar^{\inp},\gammabar^{\inp},\rho_0^{\inp},p_0^{\inp},\etarho^{\inp}]
+  \rightarrow[w_0^{\outp}]$:
+
+We begin with equation (\ref{eq:w0 divergence}) written in spherical coordinates:
+\begin{equation}
+\frac{1}{r^2}\frac{\partial}{\partial r} \left (r^2 \beta_0 w_0 \right ) = \beta_0 \left ( \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right ).
+\end{equation}
+We expand the spatial derivative and recall from Paper I that
+\begin{equation}
+\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial r} = \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r},
+\end{equation}
+giving:
+\begin{equation}
+\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) = \Sbar - \frac{1}{\gammabar p_0} \underbrace{\left( \frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r} \right)}_{\psi}.\label{eq:psi def}
+\end{equation}
+We solve this equation for $w_0$ as described in Appendix of the multilevel paper.
+
+\end{description}
 
@@
 
-Call {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n]$\\
-\phantom{ }\hfill $\rightarrow [\rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)}]$.
+Call {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n]
+ \rightarrow [\rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)}]$.
 
@@
 Call {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, T^{(2),\star}, 
-(\rho\Hext)^{(2),\star}, p_0^{n+1,\star}]$\\
-\phantom{ }\hfill$\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
+(\rho\Hext)^{(2),\star}, p_0^{n+1,\star}]$ \\
+ ${}\hfill   \rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
 
@@
 \item For spherical geometry, call\\
-{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\overline{\Gamma_1^{\nph,\star}},\rho_0^{\nph,\star},p_0^{\nph,\star},\etarho^{\nph,\star}]$\\
-\phantom{ }\hfill$\rightarrow [w_0^{\nph}]$.
+{\bf Compute} {\boldmath $w_0$} {\bf Spherical}$[\overline{S^{\nph,\star}},\overline{\Gamma_1^{\nph,\star}},\rho_0^{\nph,\star},p_0^{\nph,\star},\etarho^{\nph,\star}]\rightarrow [w_0^{\nph}]$.
 \end{description}
@@
    \begin{equation}
-   \nabla \cdot \beta_0^0 \nabla \phi = \beta_0^0(S - \Sbar)- \nabla \cdot (\beta_0^0 \Ub^{\initp})
+   \nabla \cdot \beta_0^0 \nabla \phi = \underbrace{\beta_0^0(S - \Sbar)}_\mathrm{0~ except~ for~ diffusion} - \nabla \cdot (\beta_0^0 \Ub^{\initp})
    \end{equation}

### 2014-03-19 Docs/flowchart/flowchart.tex
put the notation parts from the multilevel paper here and also put in the delta_gamma1_term stuff from paper III

@@
 \end{itemize}
+
+\subsection{$\Gamma_1$ Variation Changes}
+
+The constraint we derive from requiring the pressure to be close to the background hydrostatic
+pressure takes the form:
+\begin{equation}
+\nablab \cdotb \Ub + \frac{1}{\Gamma_1 p_0} \frac{Dp_0}{Dt} = S \enskip .
+\end{equation}
+The default \maestro\ algorithm replaces $\Gamma_1$ with $\gammabar$, allowing us
+to write this as a divergence constraint.  In paper~III, we explored the effects of
+localized variations in $\Gamma_1$ by writing $\Gamma_1 = \gammabar + \delta \Gamma_1$.
+This gives us:
+\begin{equation}
+\nablab \cdotb \Ub + \frac{1}{(\gammabar + \delta \Gamma_1) \; p_0} 
+  \Ub \cdotb \nablab p_0 = S - \frac{1}{(\gammabar + \delta \Gamma_1) \; p_0} 
+  \frac{\partial p_0}{\partial t} \enskip .
+\end{equation}
+
+Assuming that $\delta \Gamma_1 \ll \gammabar$, we then have
+\begin{equation}
+\nablab \cdotb \Ub + \frac{1}{\gammabar p_0} \Ub \cdotb \nablab p_0 = S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} +
+\frac{\delta \Gamma_1}{\gammabar^2 p_0} \frac{\partial p_0}{\partial t} +
+\frac{\delta \Gamma_1}{\gammabar^2 p_0} \Ub \cdotb \nablab p_0 \enskip , \label{eq:gammafull}
+\end{equation}
+The base state evolution equation is the average of this over a layer
+\begin{equation}
+\nablab \cdot w_0 \er + \frac{1}{\gammabar p_0} w_0 \er \cdotb \nablab p_0 =
+\Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} + \overline{
+  \left ( \frac{\delta \Gamma_1}{\gammabar^2 p_0} \ut \cdotb \nablab p_0
+  \right ) } \enskip .
+\end{equation}
+Subtracting this from equation (\ref{eq:gammafull}), we have
+\begin{equation}
+\nablab \cdotb \Ubt + \frac{1}{\gammabar p_0} \Ubt \cdotb \nablab p_0 = S -
+\Sbar + \frac{\delta \Gamma_1}{\gammabar^2 p_0} \left (\psi + \Ubt \cdotb
+\nablab p_0 \right ) - \overline{ \left ( \frac{\delta \Gamma_1}{\gammabar^2 p_0} \Ubt
+\cdotb \nablab p_0 \right ) } \enskip .
+\end{equation}
+
+These can be written more compactly as:
+\begin{equation}
+\frac{\partial w_0}{\partial r} = \Sbar -\frac{1}{\gammabar p_0}\psi +
+\overline{ \left ( \frac{\delta \Gamma_1}{\gammabar^2 p_0} \ut \cdotb \nablab
+  p_0 \right ) } \enskip , \label{eq:base_w0_with_dgamma1}
+\end{equation}
+for plane-parallel geometries (analogous to eq. [\ref{eq:divw0}]), and
+\begin{equation}
+\nablab \cdotb (\beta_0 \Ubt) = \beta_0 \left [ S - \Sbar + \frac{\delta
+    \Gamma_1}{\gammabar^2 p_0} \psi + \frac{\delta \Gamma_1}{\gammabar^2 p_0}
+  \Ubt \cdotb \nablab p_0 - \overline{ \left ( \frac{\delta
+      \Gamma_1}{\gammabar^2 p_0} \Ubt \cdotb \nablab p_0 \right ) } ~ \right ]
+ \enskip ,  \label{eq:constraint_with_delta_gamma}
+\end{equation}
+
+This constraint is not in a form that can be projected.  To solve this form, we 
+need to use a lagged $\Ubt$ in the righthand side.
+
 \subsection{Thermal Diffusion Changes}

### 2014-03-21 Docs/architecture/architecture.tex
document multigamma + remove some old references to extern/

@@
     includes the instructions on how to build the executable,
-    including what modules in {\tt extern} are used.  Any file that
+    including what modules in {\tt Microphysics/} are used.  Any file that
     you place in a sub-directory here takes precedence over a file of
@@
 statement.  Presently, the only instance of this is with the Helmholtz
-general equation of state found in {\tt extern/EOS/helmeos/}.  This is
+general equation of state found in {\tt Microphysics/EOS/helmeos/}.  This is
 automatically handled by the {\tt GMaestro.mak} instructions.
@@
 
-\maestro\ initial models are read in by the {\tt extern/model\_parser}
+\maestro\ initial models are read in by the {\tt Util/model\_parser}
 routines.  This expects the initial model to contain a header giving
@@
 
-Presently there are 2 equations of state that work with \maestro:
+Presently there are 3 equations of state that work with \maestro:
 \begin{itemize}
-\item {\tt extern/EOS/helmeos/} represents a general stellar equation 
+\item {\tt helmeos/} represents a general stellar equation 
       of state, consisting of nuclei (as an ideal gas), radiation,
@@
 
-\item {\tt extern/EOS/gamma\_law\_general} assumes an ideal gas with a mixed 
+\item {\tt gamma\_law\_general} assumes an ideal gas with a mixed 
      composition and a constant ratio of specific heats, $\gamma$:
@@
      make this switch.
+
+\item {\tt multigamma} is an ideal gas equation of state where each
+  species can have a different value of $\gamma$.  This mainly affects
+  how the internal energy is constructed as each species, represented
+  with a mass fraction $X_k$ will have its contribution to the total
+  specific internal energy take the form of $e = p/\rho/(\gamma_k -                                               
+  1)$.  The main thermodynamic quantities take the form:
+\begin{align}
+p &= \frac{\rho k T}{m_u} \sum_k \frac{X_k}{A_k} \\
+e &= \frac{k T}{m_u} \sum_k \frac{1}{\gamma_k - 1} \frac{X_k}{A_k} \\
+h &= \frac{k T}{m_u} \sum_k \frac{\gamma_k}{\gamma_k - 1} \frac{X_k}{A_k}
+\end{align}
+We recognize that the usual astrophysical $\bar{A}^{-1} = \sum_k                                                  
+X_k/A_k$, but now we have two other sums that involve different
+$\gamma_k$ weightings.
+
+The specific heats are constructed as usual,
+\begin{align}
+c_v &= \left . \frac{\partial e}{\partial T} \right |_\rho =
+    \frac{k}{m_u} \sum_k \frac{1}{\gamma_k - 1} \frac{X_k}{A_k} \\
+c_p &= \left . \frac{\partial h}{\partial T} \right |_p =
+    \frac{k}{m_u} \sum_k \frac{\gamma_k}{\gamma_k - 1} \frac{X_k}{A_k}
+\end{align}
+and it can be seen that the specific gas constant, $R \equiv c_p - c_v$ is
+independent of the $\gamma_i$, and is simply $R = k/m_u\bar{A}$ giving the
+usual relation that $p = R\rho T$.  Furthermore, we can show
+\begin{equation}
+\Gamma_1 \equiv \left . \frac{\partial \log p}{\partial \log \rho} \right |_s =
+   \left ( \sum_k \frac{\gamma_k}{\gamma_k - 1} \frac{X_k}{A_k} \right ) \bigg /
+   \left ( \sum_k \frac{1}{\gamma_k - 1} \frac{X_k}{A_k} \right ) =
+\frac{c_p}{c_v} \equiv \gamma_\mathrm{effective}
+\end{equation}
+and $p = \rho e (\gamma_\mathrm{effective} - 1)$.
+
+This equation of state takes several runtime parameters that can set the
+$\gamma_i$ for a specific species:
+\begin{itemize}
+\item {\tt eos\_gamma\_default}: the default $\gamma$ to apply for
+  all species
+\item {\tt species\_X\_name} and {\tt species\_X\_gamma}: set the $\gamma_i$
+  for the species whose name is given as {\tt species\_X\_name} to the
+  value provided by {\tt species\_X\_gamma}.  Here, {\tt X} can be one
+  of the letters: {\tt a}, {\tt b}, or {\tt c}, allowing us to specify
+  custom $\gamma_i$ for up to three different species.
+\end{itemize}
+
 \end{itemize}
@@
 modeled.  Many different reaction modules (containing different sets
-of isotopes) exist in {\tt extern/networks}.  The particular network
+of isotopes) exist in {\tt Microphysics/networks}.  The particular network
 used by a problem is defined in the problem's {\tt GNUmakefile}.

### 2014-04-01 Docs/flowchart/flowchart.tex
more about the delta_gamma1_term stuff

@@
 
-{\bf React State}$[\rho^{\inp},(\rho                                                                   h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp},
+\begin{itemize}
+\item {\bf React State}$[\rho^{\inp},(\rho                                                                   h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp},
   p_0^{\inp}] \rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp},                                   
@@
 
-{\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp] \rightarrow
+\item {\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp] \rightarrow
   [\rho_0^\outp, \rho_0^{\outp,\nph}]$ is the process by which we
@@
 
-{\bf Enforce HSE}$[p_0^{\inp},\rho_0^{\inp}]
+\item {\bf Enforce HSE}$[p_0^{\inp},\rho_0^{\inp}]
   \rightarrow [p_0^{\outp}]$ has replaced {\bf Advect Base Pressure}
@@
 
-{\bf Advect Base Enthalpy}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [(\rho h)_0^\outp]$
+\item {\bf Advect Base Enthalpy}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow [(\rho h)_0^\outp]$
   is the process by which we update the base state enthalpy through $\dt$ in time.  
@@
 
-\subsubsection{Computing $w_0$}\label{Sec:Computing w0}
+\item {\bf Computing {\boldmath $w_0$}}\label{Sec:Computing w0}
 Here we describe the process by which we compute $w_0$.  The arguments 
@@
 \end{description}
+
+\end{itemize}
 
@@
 
+%-----------------------------------------------------------------------------
 \subsection{$\Gamma_1$ Variation Changes}
 
-The constraint we derive from requiring the pressure to be close to the background hydrostatic
-pressure takes the form:
+The constraint we derive from requiring the pressure to be close to
+the background hydrostatic pressure takes the form:
 \begin{equation}
@@
 \end{equation}
-The default \maestro\ algorithm replaces $\Gamma_1$ with $\gammabar$, allowing us
-to write this as a divergence constraint.  In paper~III, we explored the effects of
-localized variations in $\Gamma_1$ by writing $\Gamma_1 = \gammabar + \delta \Gamma_1$.
-This gives us:
+The default \maestro\ algorithm replaces $\Gamma_1$ with $\gammabar$,
+allowing us to write this as a divergence constraint.  In paper~III,
+we explored the effects of localized variations in $\Gamma_1$ by
+writing $\Gamma_1 = \gammabar + \delta \Gamma_1$.  This gives us:
 \begin{equation}
@@
 \frac{\partial w_0}{\partial r} = \Sbar -\frac{1}{\gammabar p_0}\psi +
-\overline{ \left ( \frac{\delta \Gamma_1}{\gammabar^2 p_0} \ut \cdotb \nablab
+\overline{ \left ( \frac{\delta \Gamma_1}{\gammabar^2 p_0} \Ubt \cdotb \nablab
   p_0 \right ) } \enskip , \label{eq:base_w0_with_dgamma1}
@@
 
-This constraint is not in a form that can be projected.  To solve this form, we 
-need to use a lagged $\Ubt$ in the righthand side.
-
+This constraint is not in a form that can be projected.  To solve this
+form, we need to use a lagged $\Ubt$ in the righthand side.
+
+This change comes into \maestro\ in a variety of steps, summarized here.
+
+\begin{itemize}
+
+\item In {\bf Step 3}, we are doing the ``predictor'' portion of the 
+\maestro\ algorithm, getting the MAC velocity that satisfies the constraint, 
+so we do not try to incorporate the $\delta \Gamma_1$ effect.  We set
+all the $\delta \Gamma_1$ terms in
+Eq.~\ref{eq:constraint_with_delta_gamma} to zero.
+
+\item In {\bf Step 6}, we are computing the new time-centered source, 
+$S^{\nph,\star}$ and the base state velocity, $w_0^\nph$.  Now we can
+incorporate the $\delta \Gamma_1$ effect.  First we construct:
+\begin{equation}
+\frac{\delta \Gamma_1}{\Gamma_1^2 p_0} \Ubt \cdotb \nablab p_0 \approx
+   \frac{\Gamma_1^{n+1,\star} - \overline{\Gamma_1^{n+1,\star}}}
+         {{\overline{\Gamma_1^{n+1,\star}}}^2} \frac{1}{p_0^n} \Ubt^n \cdotb \nablab p_0^n
+\end{equation}
+Then we call {\bf average} to construct the lateral average of this
+\begin{equation}
+\overline{\frac{\delta \Gamma_1}{\Gamma_1^2 p_0} \Ubt \cdotb \nablab p_0} = 
+   \mathbf{Avg} \left (\frac{\Gamma_1^{n+1,\star} - \overline{\Gamma_1^{n+1,\star}}}
+         {{\overline{\Gamma_1^{n+1,\star}}}^2} \frac{1}{p_0^n} \Ubt^n \cdotb \nablab p_0^n \right )
+\end{equation}
+
+Since the average of this is needed in advancing $w_0$, we modify $\overline{S}$ 
+to include this average:
+\begin{equation}
+\overline{S^{\nph,\star}} \leftarrow \overline{S^{\nph,\star}} + 
+   \overline{\frac{\delta \Gamma_1}{\Gamma_1^2 p_0} \Ubt \cdotb \nablab p_0} 
+\end{equation}
+
+\item In {\bf Step 7}, we now include the $\delta \Gamma_1$ term in the righthand
+side for the constraint by solving:
+\begin{equation}
+\nabla \cdot \left(\beta_0^{\nph,\star} \uadvtwo\right) =
+\beta_0^{\nph,\star}\left( S^{\nph,\star} - \overline{S^{\nph,\star}} +
+   \frac{\Gamma_1^{n+1,\star} - \overline{\Gamma_1^{n+1,\star}}} 
+         {{\overline{\Gamma_1^{n+1,\star}}}^2} \frac{1}{p_0^n} 
+         (\psi^{\nph,\star} + \Ubt^n \cdotb \nablab p_0^n)
+\right)
+\end{equation}
+We note that this includes the average of the correction term as shown
+in Eq.~\ref{eq:constraint_with_delta_gamma} because we modified
+$\bar{S}$ to include this already.
+\end{itemize}
+
+%-----------------------------------------------------------------------------
 \subsection{Thermal Diffusion Changes}

### 2014-04-01 Docs/flowchart/flowchart.tex
finish documenting how we handle the delta_gamma1_term

@@
 $\bar{S}$ to include this already.
+
+\item In {\bf Step 10}, we do a construction much like that done in {\bf Step 6},
+but with the time-centerings of some of the quantities changed.
+First we construct:
+\begin{equation}
+\frac{\delta \Gamma_1}{\Gamma_1^2 p_0} \Ubt \cdotb \nablab p_0 \approx
+   \frac{\Gamma_1^{n+1} - \overline{\Gamma_1^{n+1}}}
+         {{\overline{\Gamma_1^{n+1}}}^2} \frac{1}{p_0^{n+1}} \Ubt^n \cdotb \nablab p_0^{n+1}
+\end{equation}
+Then we call {\bf average} to construct the lateral average of this
+\begin{equation}
+\overline{\frac{\delta \Gamma_1}{\Gamma_1^2 p_0} \Ubt \cdotb \nablab p_0} = 
+   \mathbf{Avg} \left (\frac{\Gamma_1^{n+1} - \overline{\Gamma_1^{n+1}}}
+         {{\overline{\Gamma_1^{n+1}}}^2} \frac{1}{p_0^{n+1}} \Ubt^n \cdotb \nablab p_0^{n+1} \right )
+\end{equation}
+
+Again we modify $\overline{S}$ to include this average:
+\begin{equation}
+\overline{S^{n+1}} \leftarrow \overline{S^{n+1}} + 
+   \overline{\frac{\delta \Gamma_1}{\Gamma_1^2 p_0} \Ubt \cdotb \nablab p_0} 
+\end{equation}
+
+\item In {\bf Step 11}, we modify the source of the constraint to include the
+$\delta \Gamma_1$ information.  In particular, we solve:
+\begin{equation}
+\nabla \cdot \left(\beta_0^{\nph} \Ubt^{n+1} \right) 
+= \beta_0^{\nph} \left(S^{n+1} - \overline{S^{n+1}} +
+   \frac{\Gamma_1^{n+1} - \overline{\Gamma_1^{n+1}}} 
+         {{\overline{\Gamma_1^{n+1}}}^2} \frac{1}{p_0^{n+1}}
+         (\psi^{\nph} + \Ubt^n \cdotb \nablab p_0^{n+1})
+\right)
+\end{equation}
+
 \end{itemize}

### 2014-04-02 Docs/getting_started/getting_started.tex
start of an index just for runtime parameters

@@
 \begin{itemize}
-\item {\tt stop\_time} is the maximum simulation time, in seconds,
+\item \runparam{stop\_time} is the maximum simulation time, in seconds,
       to evolve the system for.
 
-\item {\tt max\_step} is the maximum number of steps to take.
+\item \runparam{max\_step} is the maximum number of steps to take.
 \end{itemize}
@@
 \begin{itemize}
-\item {\tt cflfac} is a multiplicative factor ({\tt $\le 1$}) 
+\item \runparam{cflfac} is a multiplicative factor ({\tt $\le 1$}) 
       applied to the advective CFL timestep
 
-\item {\tt init\_shrink} is the factor ({\tt $\le 1$}) by which to reduce 
+\item \runparam{init\_shrink} is the factor ({\tt $\le 1$}) by which to reduce 
       the initial timestep from the estimated first timestep.
@@
 
-\item {\tt restart} tells \maestro\ to restart from a checkpoint.  The
+\item \runparam{restart} tells \maestro\ to restart from a checkpoint.  The
       value of this parameter should be the file number to restart from.
@@
 
-\item {\tt plot\_int} is the number of steps to take between
+\item \runparam{plot\_int} is the number of steps to take between
   outputting a plotfile
 
-\item {\tt plot\_deltat} is the simulation time to evolve between
+\item \runparam{plot\_deltat} is the simulation time to evolve between
   outputting a plotfile.  Note: to output only based on simulation
@@
 
-\item {\tt check\_int} is the number of steps to take between
+\item \runparam{check\_int} is the number of steps to take between
   outputting a checkpoint.
@@
 
-\item {\tt max\_levs } is the maximum number of grid levels in the AMR
+\item \runparam{max\_levs} is the maximum number of grid levels in the AMR
   hierarchy to use.  {\tt max\_levs = 1} indicates running with only a
@@
 
-\item {\tt n\_cellx }, {\tt n\_celly }, {\tt n\_cellz } the size of
+\item \runparam{n\_cellx}, \runparam{n\_celly}, \runparam{n\_cellz} the size of
   base level in terms of number of cells, in the $x$, $y$, and $z$
@@
 
-\item {\tt max\_grid\_size } the maximum extend of a grid, in any
+\item \runparam{max\_grid\_size} the maximum extend of a grid, in any
   coordinate direction, as measured in terms of number of cells.
 
-  For multilevel problems, the parameter {\tt max\_grid\_size\_1}
-  controls the maximum extent on level 1 (the base grid), {\tt
-    max\_grid\_size\_2} controls the maximum extent on level 2, and
-  {\tt max\_grid\_size\_3} controls the maximum extent on levels 3 and
-  higher.
+  For multilevel problems, the parameter \runparam{max\_grid\_size\_1}
+  controls the maximum extent on level 1 (the base
+  grid), \runparam{max\_grid\_size\_2} controls the maximum extent on
+  level 2, and \runparam{max\_grid\_size\_3} controls the maximum extent on
+  levels 3 and higher.
 
-\item {\tt prob\_lo\_x }, {\tt prob\_lo\_y }, {\tt prob\_lo\_z } is
+\item \runparam{prob\_lo\_x}, \runparam{prob\_lo\_y}, \runparam{prob\_lo\_z} is
   the physical coordinate of the lower extent of the domain boundary
@@
 
-\item {\tt prob\_hi\_x }, {\tt prob\_hi\_y }, {\tt prob\_hi\_z } is
+\item \runparam{prob\_hi\_x}, \runparam{prob\_hi\_y}, \runparam{prob\_hi\_z} is
   the physical coordinate of the upper extent of the domain boundary
@@
 
-   \item {\tt bcx\_lo }, {\tt bcy\_lo }, {\tt bcz\_lo }, {\tt bcx\_hi
-      }, {\tt bcy\_hi }, {\tt bcz\_hi } are the boundary condition
-      types at the lower (`{\tt lo}') and upper (`{\tt hi}') domain
-      boundaries in the $x$, $y$, and $z$ coordinate directions.  The
-      different types are set via integer flags listed in
-      table~\ref{gs:table:bcflags}.
+   \item \runparam{bcx\_lo}, \runparam{bcy\_lo}, \runparam{bcz\_lo
+      }, \runparam{bcx\_hi}, \runparam{bcy\_hi}, \runparam{bcz\_hi} are the
+      boundary condition types at the lower (`{\tt lo}') and upper
+      (`{\tt hi}') domain boundaries in the $x$, $y$, and $z$
+      coordinate directions.  The different types are set via integer
+      flags listed in table~\ref{gs:table:bcflags}.
 
@@
 
-   \item {\tt xlo\_boundary\_type}, {\tt ylo\_boundary\_type}, {\tt
-   zlo\_boundary\_type}, {\tt xhi\_boundary\_type}, {\tt
-   yhi\_boundary\_type}, {\tt zhi\_boundary\_type} are the boundary
-   condition types at the lower and upper domain boundaries in 
-   the $x$, $y$, and $z$ coordinate directions.   The boundary
-   type is set by providing a string name---valid values are listed
-   in table~\ref{gs:table:bcstrings}
+   \item \runparam{xlo\_boundary\_type}, \runparam{ylo\_boundary\_type}, \runparam{zlo\_boundary\_type}, \runparam{xhi\_boundary\_type}, \runparam{yhi\_boundary\_type}, \runparam{zhi\_boundary\_type}
+   are the boundary condition types at the lower and upper domain
+   boundaries in the $x$, $y$, and $z$ coordinate directions.  The
+   boundary type is set by providing a string name---valid values are
+   listed in table~\ref{gs:table:bcstrings}
 

### 2014-04-02 Docs/flowchart/flowchart.tex
start of an index just for runtime parameters
more runtime parameters to the index

@@
   There are a variety of choices of quantities to predict to the
-  edges (controlled by {\tt species\_pred\_type}---see Chapter~\ref{ch:pert}).
+  edges (controlled by \runparam{species\_pred\_type}---see Chapter~\ref{ch:pert}).
   By default, we use the equations 
@@
    of quantities that we can predict to the interfaces here (controlled
-   by {\tt enthalpy\_pred\_type}---see Chapter~\ref{ch:pert}).  For
+   by \runparam{enthalpy\_pred\_type}---see Chapter~\ref{ch:pert}).  For
    the default case, we use the perturbational
@@
 This change comes into \maestro\ in a variety of steps, summarized here.
+To enable this portion of the algorithm, set \runparam{use\_delta\_gamma1\_term} {\tt = T}.
 
@@
   The diffusion steps (4a and 8a) are optional, depending on
-  {\tt use\_thermal\_diffusion}.}
+  \runparam{use\_thermal\_diffusion}.}
 \end{figure}
@@
 
-\item {\em Initial Projection}: if {\tt do\_initial\_projection = T}, then we 
+\item {\em Initial Projection}: if \runparam{do\_initial\_projection} {\tt = T}, then we 
    first project the velocity field with $\rho = 1$ and $\beta_0^0$.
@@
 
-\item {\em ``Divu'' iterations}: Next we do {\tt init\_divu\_iter} iterations 
+\item {\em ``Divu'' iterations}: Next we do \runparam{init\_divu\_iter} iterations 
   to project the velocity field using a constraint that sees reactions
@@
 \addtocounter{enumi}{2}
-\item {\it Pressure iterations}: Here we do {\tt init\_iter} iterations to get an
+\item {\it Pressure iterations}: Here we do \runparam{init\_iter} iterations to get an
   approximation for the lagged pressure:
@@
 \item We update $T$ after the call to {\bf React State}.
-\item We have created a {\tt burning\_cutoff\_density}, where the burning does
-not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
+\item We have created a \runparam{burning\_cutoff\_density}, where the burning does
+not happen below this density.  It is presently set to \runparam{base\_cutoff\_density}.
 \item Use corner coupling in advection.
-\item We have an option, controlled by {\tt use\_tfromp}, to update temperature 
+\item We have an option, controlled by \runparam{use\_tfromp}, to update temperature 
 using $T=T(\rho,X_k,p_0)$ rather than $T=T(\rho,h,X_k)$.  The former completely 
@@
 discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
-leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
+leads to better numerical properties.  We use \runparam{enthalpy\_pred\_type = 1}.
 This is more like paper II.
 \item We have turned off the evolution of $h$ above the atmosphere and instead
-compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
+compute $h$ with the EOS using \runparam{do\_eos\_h\_above\_cutoff} {\tt = T}.
 \end{enumerate}
@@
 \begin{enumerate}
-\item We have added thermal diffusion, controlled by {\tt use\_thermal\_diffusion},
-{\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
+\item We have added thermal diffusion, controlled by \runparam{use\_thermal\_diffusion},
+\runparam{temp\_diffusion\_formulation}, and \runparam{thermal\_diffusion\_type}.
 \item We added the volume discrepancy term to the velocity constraint equation,
-controlled by the input parameter, {\tt dpdt\_factor}.
-\item For certain problems, we need to set {\tt do\_eos\_h\_above\_cutoff = FALSE}
+controlled by the input parameter, \runparam{dpdt\_factor}.
+\item For certain problems, we need to set \runparam{do\_eos\_h\_above\_cutoff} {\tt = F}
 to prevent large, unphysical velocities from appearing near the edge of the star.
@@
 advance and {\bf Correct Base} using the more accurate estimate of $\rho_0^{n+1}$?
-\item We are still exploring the effects of {\tt use\_tfromp = FALSE} for spherical
+\item We are still exploring the effects of \runparam{use\_tfromp}  {\tt = F} for spherical
 problems.  We would eventually like to run in this mode, but $T=T(\rho,X_k,p_0)$ 
 and $T=T(\rho,h,X_k)$ drift away from each other more than we would like.  Our
-attempts at incorporating a {\tt dpdt\_factor} for spherical problems have not 
+attempts at incorporating a \runparam{dpdt\_factor} for spherical problems have not 
 been successful.

### 2014-04-02 Docs/architecture/architecture.tex
more indexing

@@
 The global index space covers the entire domain at a given resolution.
-For a simulation setup with {\tt n\_cellx = 32} and {\tt n\_celly =
+For a simulation setup with \runparam{n\_cellx = 32} and \runparam{n\_celly =
   32}, the coarsest level (level 1) has $32 \times 32$ zones, and the
@@
 $\Delta x$, i.e.\ $\Delta r < \Delta x$.  The ratio is controlled
-by the parameter {\tt drdxfac}.
+by the parameter \runparam{drdxfac}.
 
@@
 
-      A runtime parameter, {\tt use\_eos\_coulomb}, is defined in
+      A runtime parameter, \runparam{use\_eos\_coulomb}, is defined in
       this EOS to enable/disable Coulomb corrections.
@@
 \begin{itemize}
-\item {\tt eos\_gamma\_default}: the default $\gamma$ to apply for
+\item \runparam{eos\_gamma\_default}: the default $\gamma$ to apply for
   all species
-\item {\tt species\_X\_name} and {\tt species\_X\_gamma}: set the $\gamma_i$
+\item \runparam{species\_X\_name} and \runparam{species\_X\_gamma}: set the $\gamma_i$
   for the species whose name is given as {\tt species\_X\_name} to the

### 2014-04-02 Docs/unit_tests/unit_tests.tex
more indexing

@@
   dimension twice (once with a positive velocity and once with a
-  negative velocity), and loops over all advection methods ({\tt
-  ppm\_type = 0,1,2} and {\tt bds\_type = 1}).  After all coordinate
-  directions are tested, the norms are compared to ensure that the
-  error does not show any directional bias.
+  negative velocity), and loops over all advection methods
+  (\runparam{ppm\_type} {\tt = 0,1,2} and \runparam{bds\_type} {\tt = 1}).  After
+  all coordinate directions are tested, the norms are compared to
+  ensure that the error does not show any directional bias.
 

### 2014-04-02 Docs/architecture/architecture.tex
more runtime parameters

@@
 The global index space covers the entire domain at a given resolution.
-For a simulation setup with \runparam{n\_cellx = 32} and \runparam{n\_celly =
+For a simulation setup with \runparam{n\_cellx} {\tt = 32} and \runparam{n\_celly} {\tt =
   32}, the coarsest level (level 1) has $32 \times 32$ zones, and the

### 2014-04-03 Docs/unit_tests/unit_tests.tex
code references

@@
 exercise only specific components of the \maestro\ solvers.  These
-tests have their own drivers (a custom {\tt varden.f90}) that
+tests have their own drivers (a custom \code{varden.f90}) that
 initialize only the data needed for the specific test and call
@@
   Gaussian distribution, maps it into the 3D domain (assuming a
-  spherical geometry) using the routines provided by the {\tt
-    fill\_3d\_module} module, and then calls {\tt average} to put it
-  back onto a 1D radial array.  This way we test the accuracy of our
-  procedure to map between the 1D radial and 3D Cartesian states.
-  The output from this test was described in detail in
-  \cite{multilevel}.
+  spherical geometry) using the routines provided by
+  the \code{fill\_3d\_module} module, and then calls \code{average} to
+  put it back onto a 1D radial array.  This way we test the accuracy
+  of our procedure to map between the 1D radial and 3D Cartesian
+  states.  The output from this test was described in detail
+  in \cite{multilevel}.
 
@@
 
-  This simply tests the reaction network by calling the \maestro\ {\tt
-  react\_state} routine directly.  The network is selected in the {\tt
-  GNUmakefile} by setting the {\tt NETWORK\_DIR} variable.  A 3d cube
-  is setup with density varying on one axis, temperature varying on
-  another, and the composition varying on the third.  The density and
-  temperature ranges are set in the inputs file.  The composition is
-  read in via an input file.
+  This simply tests the reaction network by calling
+  the \maestro\ \code{react\_state} routine directly.  The network is
+  selected in the {\tt GNUmakefile} by setting the {\tt NETWORK\_DIR}
+  variable.  A 3d cube is setup with density varying on one axis,
+  temperature varying on another, and the composition varying on the
+  third.  The density and temperature ranges are set in the inputs
+  file.  The composition is read in via an input file.
 

### 2014-04-24 Docs/compilers/compilers.tex
Added notes on the multifab_volume() error and made minor fixes to the incomp_shear_jet test problem.
Edison note

@@
  should just be $\mathtt{nx} \times \mathtt{ny} \times \mathtt{nz}$,
- but the value reported is incorrect.
+ but the value reported is incorrect.  This issue only manifests when
+ using OpenMP, so it is likely related to the OpenMP loop in {\tt
+ boxarray\_volume()}, which is called by \code{multifab\_volume()}.
+ In Cray 8.2.4 the problem appears to be resolved, with  
+ \code{\tt multifab\_volume()} returning the correct cell count.
+
+\item Cray 8.2.6 on NERSC Edison seems to work well with OpenMP.
+
 \end{itemize}

### 2014-05-05 Docs/FAQ/faq.tex

@@
 
+\item {\em {\tt make} issues an error like:
+       {\small
+       \begin{verbatim}
+       .../BoxLib/Tools/F_mk/GMakeMPI.mak:40: Extraneous text after `else' directive
+       .../BoxLib/Tools/F_mk/GMakeMPI.mak:47: *** only one `else' per conditional.  Stop
+      \end{verbatim}
+      }
+      }
+
+   You need to use GNU make version 3.81 or later.  Earlier versions do
+   not support an {\em else-if} clause.
+
 \end{enumerate}
@@
 \end{verbatim}
+\runparamidx{max\_step}
+\runparamidx{init\_iter}
+\runparamidx{init\_divu\_iter}
+\runparamidx{do\_initial\_projection}
 
@@
   steps.  To get more detail information from the multigrid solver,
-  set {\tt mg\_verbose} to a positive integer from 1-4 (the higher
+  set \runparam{mg\_verbose} to a positive integer from 1-4 (the higher
   the number the more information you receive.

### 2014-05-05 Docs/architecture/architecture.tex

@@
     custom versions of the main \maestro\ routines (e.g.\ initial
-    conditions via {\tt initdata.f90}.  See \S~\ref{sec:makefile} and
+    conditions via \code{initdata.f90}.  See \S~\ref{sec:makefile} and
     Chapter~\ref{ch:make} for details on the build system).
@@
       component of the \maestro\ algorithm.  These often have their
-      own main drivers ({\tt varden.f90}) that setup and initialize
+      own main drivers (\code{varden.f90}) that setup and initialize
       some data structures and then call only a few of the
@@
 Runtime parameters listed in the {\tt MAESTRO/\_parameters} file are
-parsed at compile time and the file {\tt probin.f90} is written and
+parsed at compile time and the file \code{probin.f90} is written and
 compiled.  This is a Fortran module that holds the values of the
@@
 master list of \maestro\ parameters ({\tt MAESTRO/\_parameters}) to
-the {\tt probin\_module}.
+the \code{probin\_module}.
 
@@
 the {\tt MAESTRO/write\_probin.py} script (along with any
-problem-specific parameters).  The script outputs the {\tt probin.f90}
+problem-specific parameters).  The script outputs the \code{probin.f90}
 source file.  Each line in the {\tt \_parameters} file has the form:
@@
   parameter = value}) or through a command-line argument (taking the
-form: {\tt --parameter value}).  The {\tt probin\_module} makes the
+form: {\tt --parameter value}).  The \code{probin\_module} makes the
 values of the runtime parameters available to the various functions
@@
 
-\maestro\ initial models are read in by the {\tt Util/model\_parser}
+\maestro\ initial models are read in by the \code{Util/model\_parser}
 routines.  This expects the initial model to contain a header giving
@@
 model should contain the same species data (in the form of mass fractions) as
-defined in the {\tt network} module used by the \maestro\ problem.
+defined in the \code{network} module used by the \maestro\ problem.
 
@@
 The best way to customize the initialization (e.g.\ add perturbations)
-is to copy from one of the existing problems.  The file {\tt
-  initveldata.f90} controls the velocity field initialization and {\tt
-  initscaldata.f90} controls the initialization of the scalars
+is to copy from one of the existing problems.  The file \code{initveldata.f90} controls the velocity field initialization and \code{initscaldata.f90} controls the initialization of the scalars
 ($\rho$, $\rho X_k$, $\rho h$).  The {\tt test2} problem is a good
@@
 \end{lstlisting}
-where {\tt ZERO} is the constant 0.0 from {\tt bl\_constants\_module}.
+where {\tt ZERO} is the constant 0.0 from \code{bl\_constants\_module}.
 
@@
 Subroutines to add, multiply, or divide two \multifab s exist, as do
-subroutines to copy from one \multifab\ to another---see {\tt boxlib/multifab.f90}
+subroutines to copy from one \multifab\ to another---see \code{BoxLib/Src/F\_BaseLib/multifab.f90}
 for the full list of routines that work with \multifab s.
@@
 When you are done working with a \multifab, its memory can be freed by
-calling {\tt multifab\_destroy} on the \multifab.
+calling \code{multifab\_destroy} on the \multifab.
 
@@
 Individual state variables should be indexed using the integer keys
-provided by the {\tt variables} module (see \S
+provided by the \code{variables} module (see \S
 \ref{sec:variables_module}).  For example, the integer {\tt rho\_comp}
@@
 
-The {\tt average\_module} module provides a routine {\tt average} that takes
+The \code{average\_module} module provides a routine {\tt average} that takes
 a multilevel \multifab\ array and averages the full Cartesian data
@@
 
-The {\tt eos\_module} provides the interface to the equation of 
+The \code{eos\_module} provides the interface to the equation of 
 state to connect the state variables thermodynamically.  It 
-gets the information about the fluid species from the {\tt network}
+gets the information about the fluid species from the \code{network}
 module (for example, the atomic number, $Z$, and atomic weight, $A$,
@@
 which has fields for all the thermodynamic inputs and outputs.  The 
-type definition is brought in through {\tt eos\_type\_module}.
+type definition is brought in through \code{eos\_type\_module}.
 \footnote{ Note: an older interface to the EOS exists, but is
   deprecated.  In this mode, the {\tt eos\_old\_interface} module declares 
-  the variables that need appear in the old-style {\tt eos} call
+  the variables that need appear in the old-style \code{eos} call
   argument list.  \maestro\ routines use these module variables in the
@@
 
-The {\tt fill\_3d\_module} provides routines that map from the 1D
+The \code{fill\_3d\_module} provides routines that map from the 1D
 base state to the full Cartesian 2- or 3D state.  Variations in the
@@
 
-The {\tt fundamental\_constants\_module} provides a simple list of
+The \code{fundamental\_constants\_module} provides a simple list of
 various fundamental constants (e.g.\ Newton's gravitational constant)
@@
 
-The {\tt network} module defines the number species advected by the
+The \code{network} module defines the number species advected by the
 code ({\tt nspec}), their ordering, and gives their basic properties
@@
 
-{\tt probin\_module} provides access to the runtime parameters.
+\code{probin\_module} provides access to the runtime parameters.
 The runtime parameters appear simply as module variables.  To get the 
@@
 
-The {\tt variables} module provides integer keys to index the state
+The \code{variables} module provides integer keys to index the state
 multifabs and other arrays dealing with the scalar quantities.  The
@@
   spec\_comp:spec\_comp-1+nspec}.  To find a particular species, a
-query can be made through the {\tt network} module, such as:
+query can be made through the \code{network} module, such as:
 \begin{verbatim}
@@
 get ``carbon-12''.
-The {\tt variables} module also provides keys for the plotfile
+The \code{variables} module also provides keys for the plotfile
 variables and boundary condition types.
@@
 
-All MPI calls are wrapped by functions in the {\tt parallel} module.  For
+All MPI calls are wrapped by functions in the \code{parallel} module.  For
 serial jobs, the wrappers simply do the requested operation on processor.
@@
 
-\noindent Here, the {\tt bl\_types} and {\tt multifab\_module} modules
-bring in the basic BoxLib data types. Specifically, here, {\tt
-  bl\_types} defines {\tt dp\_t} which is the {\tt kind} used for
-declaring double precision data, and {\tt multifab\_module} defines
-the \multifab\ data type.  The {\tt ml\_layout\_module} defines the
-datatype for a \mllayout---many routines will take an \mllayout
-to allow us to fill ghostcells.  The {\tt variables} module is a \maestro\
-module that provides integer keys for indexing the state arrays.  In
-this case the integer {\tt rho\_comp} refers to the location in the
-state array corresponding to density.
+\noindent Here, the {\tt bl\_types} and \code{multifab\_module}
+modules bring in the basic BoxLib data types. Specifically, here,
+\code{bl\_types} defines {\tt dp\_t} which is the {\tt kind} used for
+declaring double precision data, and \code{multifab\_module} defines
+the \multifab\ data type.  The \code{ml\_layout\_module} defines the
+datatype for a \mllayout---many routines will take an \mllayout to
+allow us to fill ghostcells.  The \code{variables} module is a
+\maestro\ module that provides integer keys for indexing the state
+arrays.  In this case the integer {\tt rho\_comp} refers to the
+location in the state array corresponding to density.
 
@@
 By convention, all levels in a given multifab have the same number of
-ghostcells, so we use level 1 in the {\tt nghost()} call.  We also use
+ghostcells, so we use level 1 in the \code{nghost()} call.  We also use
 the same number of ghostcells for each component of the velocity, so
@@
 \end{lstlisting}
-{\tt nfabs(s(n))} is simply the number of boxes in level {\tt n} on
+\code{nfabs(s(n))} is simply the number of boxes in level {\tt n} on
 the current processor.  Each processor knows which \fab s in its
@@
 zones.  Coordinate information is computed from {\tt dx} and {\tt
-  prob\_lo} which is the physical lower bound of the domain.  {\tt
-  bl\_constants\_module} declares useful double-precision constants,
-like {\tt HALF} (0.5).  Here, we see how to access the density for
-the current zone and compute the cell-centered velocities from the
-MAC velocities.  By convection, for a nodal array, the indices refer 
-to the {\em lower} interface in the nodal direction, so for {\tt umac},
-{\tt umac(i,j)} and {\tt umac(i+1,j)} are the $x$ MAC velocities
-on the lower and upper edge of the zone in the $x$-direction.
+  prob\_lo} which is the physical lower bound of the domain.
+\code{bl\_constants\_module} declares useful double-precision
+constants, like {\tt HALF} (0.5).  Here, we see how to access the
+density for the current zone and compute the cell-centered velocities
+from the MAC velocities.  By convection, for a nodal array, the
+indices refer to the {\em lower} interface in the nodal direction, so
+for {\tt umac}, {\tt umac(i,j)} and {\tt umac(i+1,j)} are the $x$ MAC
+velocities on the lower and upper edge of the zone in the
+$x$-direction.
 
@@
 
-\item {\tt multifab\_fill\_boundary} fills ghost cells for two
+\item \code{multifab\_fill\_boundary} fills ghost cells for two
   adjacent grids at the same level, which als includes periodic domain
@@
 
-\item {\tt multifab\_physbc} fills ghostcells at the physical boundaries.
-
-\item {\tt multifab\_fill\_ghost\_cells} is used for multilevel
+\item \code{multifab\_physbc} fills ghostcells at the physical boundaries.
+
+\item \code{multifab\_fill\_ghost\_cells} is used for multilevel
   problems, and fills ghostcells in the finer grid (level {\tt n}) by
@@
   This call is usually preceded by a call to 
-  {\tt ml\_cc\_restriction\_c} which sets the level {\tt n-1} data to be
+  \code{ml\_cc\_restriction\_c} which sets the level {\tt n-1} data to be
   the average of the level {\tt n} data covering it.
@@
 velocity components, and the subsequent slots are for the other
-standard variables described in the {\tt variables\_module}.  For
+standard variables described in the \code{variables\_module}.  For
 instance, to reference the boundary condition for density, one would
@@
 The actual filling of the ghostcells according to the descriptions
-contained in the \bctower\ is carried out by the {\tt
-  multifab\_physbc} routine.  When you have an {\tt EXT\_DIR}
+contained in the \bctower\ is carried out by the \code{multifab\_physbc} routine.  When you have an {\tt EXT\_DIR}
 condition in {\tt multifab\_physbc} (an specified in the inputs file
@@
 and post-processing.  To use particles, particles must be seeded into
-the domain by writing a problem-specific {\tt init\_particles.f90}
+the domain by writing a problem-specific \code{init\_particles.f90}
 routine.  This routine is called at the start of the simulation.  The
 {\tt init\_particles} routines add particles at specific locations by
-calling the {\tt particle\_module}'s {\tt add} routine when a given
+calling the \code{particle\_module}'s {\tt add} routine when a given
 criteria is met by the fluid state.

### 2014-05-05 Docs/flowchart/flowchart.tex

@@
 included with more detail than was given in paper III.  The main
-driver for a single step of the algorithm is {\tt advance.f90}---refer
+driver for a single step of the algorithm is \code{advance.f90}---refer
 to that code to see the sequence of functions called to implement each
@@
   and external heating. 
-  The initial timestep estimate is provided by {\tt firstdt} and
+  The initial timestep estimate is provided by \code{firstdt} and
   $\Ub^{0,0}$, to allow us to compute the effect of reactions over $\Delta t/2$.

### 2014-05-05 Docs/getting_started/getting_started.tex

@@
 number, peak temperature, and peak nuclear energy generation rate.
-Individual problems can provide their own {\tt diag.f90} file to
+Individual problems can provide their own \code{diag.f90} file to
 produce custom diagnostic output.  This information can be plotted
@@
   conversion between the string names and integer flags is done
-  by \boxlib\ in the {\tt bc\_module} at the time of initializing
+  by \boxlib\ in the \code{bc\_module} at the time of initializing
   the runtime parameters.

### 2014-05-05 Docs/initial_models/initial_models.tex

@@
 We deal with the edge of the star by tracking the first coarse cell 
-in which the density falls below {\tt base\_cutoff\_density}.  We note 
+in which the density falls below \runparam{base\_cutoff\_density}.  We note 
 the radius of this cell center, and the values of $\rho$, $\rho h$, $X_k$, 
@@
   multifab to obtain the full state $\rho, h, X$, and $T$.
-\item If {\tt perturb\_model} = T, a user-defined perturbation is
+\item If \runparam{perturb\_model} {\tt = T}, a user-defined perturbation is
   added.  This routine should make sure that the EOS is called so that

### 2014-05-05 Docs/lo_density/lo_density.tex
add note about limit_conductivity

--- 2012-05-18 docs/lo_density/lo_density.tex
@@
 \begin{itemize}
-\item {\tt base\_cutoff\_density}, $\rho_{\rm base}$, (real):\\ \\
+\item \runparam{base\_cutoff\_density}, $\rho_{\rm base}$, (real):\\ \\
 Essentially controls the lowest density allowed in the simulation and modifies the behavior
@@
 where $\rho_0 < \rho_{\rm base}$.  Slightly more complicated for multilevel problems.
-\item {\tt anelastic\_cutoff}, $\rho_{\rm anelastic}$, (real):\\ \\
+
+\item \runparam{anelastic\_cutoff}, $\rho_{\rm anelastic}$, (real):\\ \\
 If $\rho_0 < \rho_{\rm anelastic}$, we modify the computation of $\beta_0$ in the
 divergence constraint.
+
 \item {\tt anelastic\_cutoff\_coord(:)} (integer array):\\ \\
 Anelastic cutoff analogy of {\tt base\_cutoff\_density\_coord(:)}.
-\item {\tt burning\_cutoff\_density}, $\rho_{\rm burning}$,  (real):\\ \\
+
+\item \runparam{burning\_cutoff\_density}, $\rho_{\rm burning}$,  (real):\\ \\
 If $\rho < \rho_{\rm burning}$, don't call the burner in this cell.
+
 \item {\tt burning\_cutoff\_density\_coord(:)} (integer array):\\ \\
 Burning cutoff analogy of {\tt base\_cutoff\_density\_coord(:)}.
-\item {\tt buoyancy\_cutoff\_factor} (real):\\ \\
+
+\item \runparam{buoyancy\_cutoff\_factor} (real):\\ \\
 When computing velocity forcing, set the buoyance term ($\rho-\rho_0$) to 0 if 
 $\rho < $ {\tt buoyancy\_cutoff\_factor} * {\tt base\_cutoff\_density}.
-\item {\tt do\_eos\_h\_above\_cutoff} (logical):\\ \\
+
+\item \runparam{do\_eos\_h\_above\_cutoff} (logical):\\ \\
 If {\rm true}, at the end of the advection step, for each cell where 
@@
 
-Note that for single-level planar or any spherical problem, saying $r\ge$ {\tt anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)\le$ {\tt anelastic\_cutoff}.
-Also, saying $r<$ {\tt anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)>$ {\tt anelastic\_cutoff}.
-Ditto for {\tt base\_cutoff\_density} and {\tt base\_cutoff\_density\_coord}.
+Note that for single-level planar or any spherical problem, saying
+$r\ge$ {\tt anelastic\_cutoff\_coord} is analogous to saying
+$\rho_0(r)\le$ \runparam{anelastic\_cutoff}.  Also, saying $r<$ {\tt
+anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)>$ {\tt
+anelastic\_cutoff}.  Ditto for \runparam{base\_cutoff\_density} and {\tt
+base\_cutoff\_density\_coord}.
 
@@
 
-The {\tt anelastic\_cutoff} is the density below which we modify the constraint.
+The \runparam{anelastic\_cutoff} is the density below which we modify the constraint.
 
@@
 \item in {\tt make\_S}, we set {\tt delta\_gamma1\_term} and {\tt delta\_gamma1} 
-  to zero for $r \ge {\tt anelastic\_cutoff\_coord}$.
+  to zero for $r \ge {\tt anelastic\_cutoff\_coord}$.  This is only relevant
+  if you are running with \runparam{use\_delta\_gamma1\_term} {\tt = T}.
 
@@
 
-The {\tt base\_cutoff\_density} is the lowest density that we model.
+The \runparam{base\_cutoff\_density} is the lowest density that we model.
 
@@
 \item In {\tt update\_scal}, if $\rho \le {\tt base\_cutoff\_density}$
-   and {\tt do\_eos\_h\_above\_cutoff}, we call the EOS to compute $h$.
+   and \runparam{do\_eos\_h\_above\_cutoff}, we call the EOS to compute $h$.
 
@@
 
-\item In {\tt probin}, {\tt burning\_cutoff\_density} is set to 
-  {\tt base\_cutoff\_density}.  There is no option to set 
+\item In {\tt probin}, \runparam{burning\_cutoff\_density} is set to 
+  \runparam{base\_cutoff\_density}.  There is no option to set 
   {\tt burning\_cutoff\_density} using the inputs file.
@@
 
-The {\tt buoyancy\_cutoff\_factor} is used to zero out the forcing terms
+The \runparam{buoyancy\_cutoff\_factor} is used to zero out the forcing terms
 to the velocity equation at low densities.
@@
 
+\item In {\tt make\_explicit\_thermal}, if \runparam{limit\_conductivity} {\tt = T}, then for 
+$\rho < \mathtt{buoyancy\_cutoff\_factor}$ \\ $* \mathtt{base\_cutoff\_density}$, we
+zero out the thermal coefficients, effectively turning off thermal
+diffusion there.
+
 \end{itemize}

### 2014-05-05 Docs/visualization/visualization.tex
update the getting started with some new info

@@
 
-By default, plotfiles store double precision data, but if {\tt
-  single\_prec\_plotfiles = .true.} is set, then the data is converted
-to single precision before outputting---this is done to reduce file
-sizes.
+By default, plotfiles store double precision data, but if
+\runparam{single\_prec\_plotfiles} {\tt = T} is set, then the data is
+converted to single precision before outputting---this is done to
+reduce file sizes.
 
@@
 %\hline
-{\tt rhoh}   & $(\rho h)$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and {\tt plot\_h\_with\_use\_tfromp = T}) \\           
+{\tt rhoh}   & $(\rho h)$ & {\tt use\_tfromp = F} or ({\tt use\_tfromp = T} and \runparam{plot\_h\_with\_use\_tfromp} {\tt = T}) \\           
 %\hline
@@
 %\hline
-{\tt X(*)}   & $(\rho X_k)/\rho$ & {\tt plot\_spec} \\
+{\tt X(*)}   & $(\rho X_k)/\rho$ & \runparam{plot\_spec} \\
 %\hline
-{\tt tracer} & tracers           & {\tt plot\_trac} \\
+{\tt tracer} & tracers           & \runparam{plot\_trac} \\
 %\hline
-{\tt w0\_x}  & $w_0 \er \cdot \ex$ & {\tt plot\_base} \\ 
+{\tt w0\_x}  & $w_0 \er \cdot \ex$ & \runparam{plot\_base} \\ 
 %\hline
@@
 %\hline
-{\tt radial\_velocity}  & $\Ubt \cdot \er + w_0$ & {\tt spherical == 1} \\
+{\tt radial\_velocity}  & $\Ubt \cdot \er + w_0$ & \runparam{spherical} {\tt == 1} \\
 %\hline
@@
 %\hline
-{\tt soundspeed}        & $c(\rho,p_0,X_k)$ & {\tt plot\_cs} \\
+{\tt soundspeed}        & $c(\rho,p_0,X_k)$ & \runparam{plot\_cs} \\
 %\hline
@@
 %\hline
-{\tt sponge} or {\tt sponge\_fdamp} & $1/(1 + \Delta t \kappa f_\mathrm{damp})$ by default; $f_\mathrm{damp}$ if {\tt plot\_sponge\_fdamp = T} & -- \\
+{\tt sponge} or {\tt sponge\_fdamp} & $1/(1 + \Delta t \kappa f_\mathrm{damp})$ by default; $f_\mathrm{damp}$ if \runparam{plot\_sponge\_fdamp = T} & -- \\
 %\hline
@@
 %\hline
-{\tt gpi\_x}           & $\nabla \pi \cdot \ex$ & {\tt plot\_gpi} \\
+{\tt gpi\_x}           & $\nabla \pi \cdot \ex$ & \runparam{plot\_gpi} \\
 %\hline 
@@
 %\hline
-{\tt omegadot(*)}      & $\dot{\omega}_k$ & {\tt plot\_omegadot} \\
+{\tt omegadot(*)}      & $\dot{\omega}_k$ & \runparam{plot\_omegadot} \\
 %\hline
-{\tt enucdot}          & $(\rho \Hnuc)/\rho$ & {\tt plot\_Hnuc} \\
+{\tt enucdot}          & $(\rho \Hnuc)/\rho$ & \runparam{plot\_Hnuc} \\
 %\hline
-{\tt Hext}             & $(\rho \Hext)/\rho$ & {\tt plot\_Hext} \\
+{\tt Hext}             & $(\rho \Hext)/\rho$ & \runparam{plot\_Hext} \\
 %\hline
-{\tt eta\_rho}         & $\etarho$ & {\tt plot\_eta} \\
+{\tt eta\_rho}         & $\etarho$ & \runparam{plot\_eta} \\
 %\hline
-{\tt thermal}          & $\nabla \cdot \kappa \nabla T$ & {\tt use\_thermal\_diffusion} \\
+{\tt thermal}          & $\nabla \cdot \kappa \nabla T$ & \runparam{use\_thermal\_diffusion} \\
 %\hline
@@
 %\hline
-{\tt ad\_excess}       & $\nabla - \nabla_\mathrm{ad}$ & {\tt plot\_ad\_excess} \\
+{\tt ad\_excess}       & $\nabla - \nabla_\mathrm{ad}$ & \runparam{plot\_ad\_excess} \\
 %\hline
-{\tt particle\_count}  & number of particles in a cell & {\tt use\_particles} \\
+{\tt particle\_count}  & number of particles in a cell & \runparam{use\_particles} \\
 %\hline
-{\tt processor\_number} & processor number containing the cell's data & {\tt plot\_processors} \\
+{\tt processor\_number} & processor number containing the cell's data & \runparam{plot\_processors} \\
 %\hline
-{\tt pi\_divu}         & $\pi \nabla \cdot \tilde{\Ub}$ (a measure of energy conservation) & {\tt plot\_pidivu} \\
+{\tt pi\_divu}         & $\pi \nabla \cdot \tilde{\Ub}$ (a measure of energy conservation) & \runparam{plot\_pidivu} \\
 \end{longtable}
@@
 %-----------------------------------------------------------------------------
-\section{Visualizing with \yt}
+\section{\label{sec:vis:yt} Visualizing with \yt}
 

### 2014-05-21 Docs/visualization/visualization.tex

@@
 {\tt AmrPostprocessing/python} provides some simple commandline
-tools for doing visualizations of \boxlib\ plotfiles.  The main drivers
+tools for doing visualizations of \boxlib\ plotfiles (note: a subset
+of these are distributed directly with \boxlib\ in {\tt BoxLib/Tools/Py\_util/}).  The main drivers
 are written in python and use a set of Fortran routines, compiled with
@@
 \noindent {\bf Note: for \maestro\ data, you need to use \yt 3.0 or later.}
+
+Some sample scripts that use \yt\ with \maestro\ data are contained in 
+{\tt MAESTRO/Util/yt/}.
 

### 2014-06-06 Docs/makefiles/makefiles.tex
add versoin # of GNU make
Added some details to the documentation of how Maestro's build system handles dependencies.

@@
 
-The \maestro\ build system uses features of GNU make, which is
-typically the default on systems.  The \maestro\ executable is built
-in the problem's directory (one of the directories under {\tt
-SCIENCE/}, {TEST\_PROBLEMS}, or {UNIT\_TESTS}).  This directory will
-contain a makefile, {\tt GNUmakefile}, that includes all the necessary
-information to build the executable.
+The \maestro\ build system uses features of GNU make (version 3.81 or
+later), which is typically the default on systems.  The \maestro\
+executable is built in the problem's directory (one of the directories
+under {\tt SCIENCE/}, {\tt TEST\_PROBLEMS}, or {\tt UNIT\_TESTS}).  This
+directory will contain a makefile, {\tt GNUmakefile}, that includes
+all the necessary information to build the executable.
 
@@
 process and parse all the source files and make an explicit list of
-the dependency pairs.  This process is defined in {\tt
-GMakerules.mak}.
+the dependency pairs.  The execution of these scripts is triggered
+by including makefiles of the form {\tt *.depends}.  On a fresh build these
+will not exist.  When GNU make cannot find an included makefile it will
+first attempt to build it using any relevant targets before issuing an 
+error.  The targets for the {\tt *.depends} files contain the recipe for
+executing the dependency scripts.  Once these makefiles are built by the
+scripts GNU make will then read the dependencies for the current build
+from them.  This process is defined in {\tt GMakerules.mak}.
 

### 2014-06-19 Docs/compilers/compilers.tex
compiler updates

@@
 
- A separate issue is that the ``volume'' reported by {\tt
+ \item A separate issue is that the ``volume'' reported by {\tt
  multifab\_volume()} seems to be wrong.  For a single level run, it
@@
 
-\item Cray 8.2.6 on NERSC Edison seems to work well with OpenMP.
+\item Cray 8.2.5 works well on titan/OLCF with MPI/OpenMP
+\item Cray 8.2.6 works well on edison/NERSC with MPI/OpenMP.
 
@@
 unknown seg fault at the beginning of the run.
+
+\item There are no known issues with gfortran 4.8.x
+
 \end{itemize}

### 2014-06-29 Docs/getting_started/getting_started.tex
add rt doc

@@
 
+\item {\tt rt} \\ [-3mm]
+
+A Rayleigh-Taylor instability problem.  There are two methods that the
+code is run here, in the standard (using {\tt inputs\_2d}), the base state
+has the stratified atmosphere and we introduce a velocity perturbation
+to start the instability.  The alternate method, {\tt inputs\_2d\_SNe}, uses
+the \runparam{do\_smallscale} runtime parameter to eliminate the base state
+and instead use the incompressible constraint to evolve the system.
+
 \item {\tt test2} \\[-3mm]

### 2014-06-30 Docs/mg/mg.tex
some info on the various multifabs

--- 2012-07-24 docs/mg/mg.tex
@@
 
+\end{itemize}
+
+In addition to these parameters, the \mgtower\ carries a number of
+\multifab s that carry the solution and stencil for the multigrid
+solve.
+
+\begin{itemize}
+
+\item {\tt ss}: The stencil itself---for each zone, this gives the 
+  coefficients of the terms in the Laplacian, with the convection that
+  the `0' term is located in the current cell and the other terms are
+  the $\pm 1$ off the current cell in each direction.
+
+\item {\tt cc}: scratch space (?)
+
+\item {\tt ff}: The source (righthand side) of the elliptic equation
+  we are solving.
+
+\item {\tt dd}: The residual/defect, $ff - L\phi$
+
+\item {\tt uu}: The solution variable ($\phi$ when on the finest level)
+
+\item {\tt mm}: A mask used at coarse/fine interfaces
 \end{itemize}

### 2014-07-18 Docs/managing_jobs/managingjobs.tex
merge the compiler stuff into here

@@
 \item The build date, machine, and directory
-\item The MAESTRO, BoxLib and other relevant git hashes for the source
+\item The \maestro, \boxlib, and other relevant git hashes for the source
 \item The directories used in the build
@@
 
+\section{Linux boxes}
+
+\subsection{gfortran}
+
+gfortran is probably the best compiler for \maestro.  Here are some
+version-specific notes:
+
+\begin{itemize}
+\item {\em gfortran 4.5.x}: gfortran 4.5 has no known issues.
+
+\item {\em gfortran 4.6.x}: gfortran 4.6.1--4.6.3 on Fedora 
+(15 and 16) have a bug that shows up in multi-processor, multilevel
+runs ({\tt toy\_convect} on 4 processors, with {\tt
+inputs\_2d\_ml.test} for instance).  The symptom is a crash during
+initialization with the message that ``message sizes do not match
+across processes in the collective'' in a {\tt parallel\_bcast} call.
+This is in {\tt cluster\_f.f90}, and the issue seems to be a
+multiplication error in computing the size to allocate {\tt bxs}.  The
+bug is not present when you compile in debug mode, so it seems to be
+related to optimization.
+
+\item {\em gfortran 4.7.x}: gfortran 4.7.1 (prerelease) seems to have 
+fixed the above issue, but now wdconvect with 2-level, effective
+768$^3$ dies due to an unknown seg fault at the beginning of the run.
+
+\item {\em gfortran 4.8.x}: There are no known issues with gfortran 4.8.x
+
+\end{itemize}
+
 \section{Working at OLCF (ORNL)}
+
+\subsection{Titan Compilers}
+
+The preferred compilers on Titan are the Cray compilers.  
+Cray 8.2.5 works well on titan/OLCF with MPI/OpenMP.
 
@@
 
+\subsection{edison compilers}
+
+The default compilers on edison are the Intel compilers, but
+it seems that the Cray compilers are the best.
+
+\begin{itemize}
+\item Cray 8.2.6 works well on edison/NERSC with MPI/OpenMP.
+
+\item Cray 8.3.0 generates an internal compiler error on edison/NERSC.
+      To get back to the older Cray compilers, you need to do:
+\begin{verbatim}                                                                                  
+module swap cce/8.3.0 cce/8.2.6                                                                   
+module swap cray-libsci cray-libsci/12.2.0                                                        
+module swap cray-mpich cray-mpich/6.3.1                                                           
+\end{verbatim}
+\end{itemize}
+
 \subsection{General Guidelines}

### 2014-07-21 Docs/managing_jobs/managingjobs.tex
update the edison run info

@@
 
-gfortran is probably the best compiler for \maestro.  Here are some
+gfortran is probably the best-supported compiler for \maestro.  Here are some
 version-specific notes:
@@
 
-The submission script {\tt jaguar.run} and shell script
-{\tt process.jaguar} in {\tt Util/job\_scripts/jaguar/}
+The submission script {\tt titan.run} and shell script
+{\tt process.titan} in {\tt Util/job\_scripts/titan/}
 are designed to allow you to run \maestro\ with minimal interaction,
-while being assured that the data is archived to HPSS on the NCCS
+while being assured that the data is archived to HPSS on the OLCF
 machines.
@@
 (Note: if the hsi command prompts you for your password, you will need to
-talk to the NCCS help desk to ask for password-less access to HPSS).
+talk to the OLCF help desk to ask for password-less access to HPSS).
 
-The script {\tt process.jaguar} is called from {\tt jaguar.run} and
+The script {\tt process.titan} is called from {\tt titan.run} and
 will run in the background and continually wait until checkpoint or
@@
 
-The {\tt jaguar.run} submission script has code in it that will look
+The {\tt titan.run} submission script has code in it that will look
 at the most recently generated checkpoint files, make sure that they
@@
 before the new submission runs and {\tt QSUB SCRIPT} is the submission
-script (e.g.\ {\tt jaguar.run}).
+script (e.g.\ {\tt titan.run}).
 This way you can queue up a bunch of runs and literally leave things
@@
 
-When {\tt process.jaguar} is running, it creates a lockfile (called
+When {\tt process.titan} is running, it creates a lockfile (called
 {\tt process.pid}) that ensures that only one instance of the script
@@
 \subsection{Remote \visit\ Visualization on Lens}
+
+{\em Note: this information may be out-of-date.}
 
@@
 
-hopper is configured with 24 cores per node, with the cores grouped
-into 4 NUMA nodes of 6 cores each.  Best performance is seen when
-running with 6 threads and 4 MPI tasks per node, using the following
-run line:
+edison is configured with 24 cores per node split between two Intel IvyBridge
+12-core processors.  Each processor connects to 1/2 of the node's memory
+and is called a NUMA node, so there are 2 NUMA nodes per edison node.
+Best performance is seen when
+running with 6 threads and 4 MPI tasks per node, as example:
 \begin{verbatim}
-aprun -n 128 -N 4 -d 6 -S 1 -ss ./main.Linux.Cray.mpi.omp.ex inputs_3d
+aprun -n 192 -N 4 -S 2 -d 6 -ss ./main.Linux.Cray.mpi.omp.exe inputs_3d
 \end{verbatim}
-Here, {\tt -n} is the number of MPI tasks, and {\tt n * d} is the
-total number of cores (which should match the {\tt mppwidth} set at
-the top of the submission script). \\
+Here,
+\begin{itemize}
+\item {\tt -S} is the number of MPI tasks per NUMA node (you want to split them
+  across the 2 NUMA nodes per Edison node)
+\item {\tt -n} is the total number of MPI tasks for the entire application
+\item {\tt -N} is the number of MPI tasks PER Edison node
+\item {\tt -d} is the number of OpenMP threads per MPI task (must match 
+{\tt OMP\_NUM\_THREADS})
+\end{itemize}
+The product of {\tt -N} and {\tt -d} should be 24---the total number of cores
+per edison node.
+The script {\tt
+Util/job\_scripts/edison/edison.run} provides an example.
 
+\subsubsection{Older information---still relevant?}
 \noindent (2013-01-15) It's been seen that sometimes an MPICH2 error is 
@@
 
-The same set of submission scripts are available for hopper at NERSC in 
-{\tt Util/job\_scripts/hopper/}.  They are mostly identical, with a few
-minor changes for the local system configuration.  The use is the same
-as described above.
+The same set of submission scripts described for titan are available
+for edison at NERSC in {\tt Util/job\_scripts/edison/}.  They are
+mostly identical, with a few minor changes for the local system
+configuration.  The use is the same as described above.
+
+\subsection{Batch visualization using \yt}
+
+The scripts {\tt parallel-yt} and {\tt process-edison.run} in {\tt
+Util/job\_scripts/edison} show how to invoke \yt\ to loop over a 
+series of plotfiles and do visualization.  A number of tasks are
+run at once on the node, each operating on a separate file.  Note:
+it is important that {\tt aprun} be used to launch the \yt\ script
+to ensure that it is run on the compute node.
 

### 2014-07-21 Docs/mg/mg.tex
flowchart for the cell-centered MG

@@
 
-\item {\tt dd}: The residual/defect, $ff - L\phi$
+\item {\tt dd}: The residual/defect, $f - L\phi$
 
@@
 structure is not appropriate for the fancy bottom solver.
+
+\section{Flowchart}
+
+\MyArrow[1em]{start1}{end1}{colorii,line width=2pt}
+\MyArrow[1em]{start2}{end2}{colorii,line width=2pt}
+
+\subsection{Cell-Centered MG}
+
+\begin{itemize}
+
+\item {\tt stencil\_fill\_cc\_all\_mglevels} / {\tt stencil\_fill\_cc}:
+
+\item {\tt ml\_cc}:
+
+\item {\tt mg\_tower\_v\_cycle} (recursive):
+\begin{itemize}
+
+  \item {\em recursively descend V-cycle}
+  \begin{itemize} 
+  \item \tikzmark{end1} {\tt cc\_mg\_tower\_smoother}:
+
+  \item {\tt compute\_defect}:
+
+  \item\tikzmark{start1} {\tt ml\_restriction}:
+  \end{itemize}
+
+  \item {\tt mg\_tower\_bottom\_solve}:
+
+  \item {\em ascend V-cycle}
+
+  \begin{itemize}
+  \item\tikzmark{end2} {\tt mg\_tower\_prolongation}: Take the solution at level $n-1$ and use it to 
+     correct the solution at level $n$ by representing the data on the finer grid.  This uses
+     linear reconstruction for jumps by $2\times$ and piecewise-constant otherwise.
+
+  \item\tikzmark{start2} {\tt cc\_mg\_tower\_smoother}:
+  \end{itemize}
+
+\end{itemize}
+\item multiple {\tt compute\_defect} calls:
+
+\end{itemize}
 

### 2014-07-22 Docs/mg/mg.tex
more MG doc
lots of MG documentation work

@@
 
-\section{Introduction to \maestro\ Multigrid}
+\section{\boxlib\ Multigrid Philosophy}
 
-\maestro\ uses multigrid to enforce the velocity constraint through
-projections at the half-time (the MAC projection) and end of the time
-step (the HG projection).  Two multigrid solvers are provided by
-\boxlib---one for cell-centered data and one for node-centered (nodal)
-data.  Both of these are used in \maestro.
+Here we describe some of the general ideas behind the \boxlib\ multigrid.
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=5.5in]{\mgfigpath/MAC_mg2}
-\caption{\label{fig:mg:MAC} Data centerings for the MAC projection}
-\end{figure}
+We solve MG on an AMR hierarchy, which means in places we will encounter
+C-F interfaces.  The \boxlib\ MG will modify the stencil for the Laplacian
+at C-F interfaces to ensure the correct solution to the Poisson equation
+across these interfaces.  There is no correction step needed in this
+case (this differs from \cite{ricker:2008}).
 
-The MAC projection operates on the advective velocities predicted at
-the cell-interfaces at the half-time.  The edge-centered velocities
-are shown in Figure~\ref{fig:mg:MAC}.  If we consider purely
-incompressible flow, the projection appears as:
-\begin{equation}
-D G \phi = D U
-\end{equation}
-where $D$ is the divergence operator and $G$ is the gradient operator.
-In this discretization, $\phi$ is cell-centered (see
-Figure~\ref{fig:mg:MAC}).  The remaining quantities are discretized as:
-\begin{itemize}
-\item $DU$ is cell-centered, 
-  \begin{equation}
-  (DU)_{i,j} = \frac{u_{i+1/2,j} - u_{i-1/2,j}}{\Delta x} + 
-               \frac{v_{i,j+1/2} - v_{i,j-1/2}}{\Delta y}
-  \end{equation}
+The MG solver always works with jumps of $2\times$ between levels.  In 
+some codes (but not \maestro) we can have jumps of $4\times$ between
+levels.  \boxlib\ uses a {\tt mini\_cycle} in these cases, effectively
+inserting a multigrid level between the two AMR levels and doing a mini
+V-cycle.
 
-\item $G\phi$ is edge-centered, on the MAC grid, as shown in
-  Figure~\ref{fig:mg:MAC}.
+Support for $\Delta x \ne \Delta y \ne \Delta z$
 
-\item $DG\phi$ is cell-centered, also shown in Figure~\ref{fig:mg:MAC},
-  computed from $G\phi$ using the same differencing as $DU$.
+There are two MG solvers: cell-centered and nodal.
 
-\end{itemize}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=5.5in]{\mgfigpath/HG_mg2}
-\caption{\label{fig:mg:HG} Data centerings for the HG projection}
-\end{figure}
-
-The HG projection projects the cell-centered velocities at the end of
-the timestep.  Here, $\phi$ is node-centered.  Figure~\ref{fig:mg:HG}
-shows the locations of the various quantities involved in the HG
-projection.  Again considering simple incompressible flow, we now
-solve:
-\begin{equation}
-L \phi = D U
-\end{equation}
-where $L$ is a discretization of the Laplacian operator.  In this
-sense, the HG projection is an {\em approximate projection}, that is,
-$L \neq DG$ (in discretized form).  The various operations have the
-following centerings:
-\begin{itemize}
-
-\item $DU$ is node-centered.  This is computed as:
-  \begin{equation}
-  (DU)_{i-1/2,j-1/2} = \frac{\frac{1}{2} (u_{i,j} + u_{i,j-1}) -
-                             \frac{1}{2} (u_{i-1,j} + u_{i-1,j-1})}{\Delta x} +
-                       \frac{\frac{1}{2} (v_{i,j} + v_{i-1,j}) -
-                             \frac{1}{2} (v_{i,j-1} + v_{i-1,j-1})}{\Delta y} 
-  \end{equation}
-
-\item $G\phi$ is cell-centered, as shown in Figure~\ref{fig:mg:HG}.
-
-\item $L\phi$ is node-centered.  This is a direct discretization of 
-the Laplacian operator.  By default, \maestro\ uses a dense stencil
-(9-points in 2-d, 27-points in 3-d).  Alternately, a {\em cross}
-stencil can be used (by setting {\tt hg\_dense\_stencil = F}).  This
-uses 5-points in 2-d, 7-points in 3-d.
-
-\end{itemize}
-
-\section{\mgtower}
+\subsection{\mgtower}
 
@@
 
-\item {\tt mm}: A mask used at coarse/fine interfaces
+\item {\tt mm}: For cell-centered, {\tt mm} takes a direction and
+   tells you whether we are at a Dirichlet or Neumann boundary, or if 
+   we are skewed in that direction.
+
+   For nodal, {\tt mm} simply tells us whether a point is Dirichlet
+   or Neumann.  There is no skew in nodal.
+
 \end{itemize}
 
-\section{Bottom Solvers}
+\subsection{Stencils}
 
-There are several bottom solvers available to \maestro.  These are set
-through the {\tt mg\_bottom\_solver} (MAC) and {\tt
-  hg\_bottom\_solver} (HG) runtime parameters.  The allowed values
-  are:
+There are several different stencil types that we can use for 
+the discretization.  For cell-centered, these are:
+\begin{itemize}
+\item {\tt CC\_CROSS\_STENCIL}: this is the standard cross-stencil---5 points 
+in 2-d and 7 points in 3-d.  For cell-centered MG, this is the default, and
+is usually the best choice.
 
+\item {\tt HO\_CROSS\_STENCIL}: this is a cross-stencil that uses 9 points
+in 2-d and 11-points in 3-d.  For instance, it will use $(i,j)$;
+$(i\pm1,j)$; $(i\pm2,j)$; $(i,j\pm1)$; $(i,j\pm2)$.  This is
+higher-order accurate that {\tt CC\_CROSS\_STENCIL}.
+
+\item {\tt HO\_DENSE\_STENCIL}: this is a dense-stencil---it uses all the
+points (including corners) in $(i\pm1,j\pm1)$, resulting in a 9-point stencil
+in 2-d and 27-point stencil in 3-d.
+\end{itemize}
+
+For the nodal solver, the choices are:
+\begin{itemize}
+\item {\tt ND\_CROSS\_STENCIL}: this is the standard cross-stencil.
+
+\item {\tt ND\_DENSE\_STENCIL}: this is a dense stencil, using
+  all the points in $(i\pm1,j\pm1)$.  The 
+  derivation of this stencil is based on finite-element ideas, defining
+  basis functions on the nodes.  This is developed in 2-d in 
+  \cite{almgrenBellSzymczak:1996}.
+
+\item {\tt ND\_VATER\_STENCIL}: this is an alternate dense stencil derived
+  using a similar finite-element idea as above, but a different control
+  volume.
+\end{itemize}
+For the cell-centered solve, the coefficients for the stencil are computed
+once, at the beginning of the solve.  For the nodal solver, the coefficients
+are hard-coded into the smoothers.
+
+\subsection{Smoothers}
+
+The following smoothers are available (but not necessarily for both the
+cell-centered and nodal solvers):
+\begin{itemize}
+\item {MG\_SMOOTHER\_GS\_RB}: a red-black Gauss-Seidel smoother
+
+\item {MG\_SMOOTHER\_JACOBI}: a Jacobi smoother (not implemented for 
+the dense nodal stencil)
+
+\item {MG\_SMOOTHER\_MINION\_CROSS}
+
+\item {MG\_SMOOTHER\_MINION\_FULL}
+
+\item {MG\_SMOOTHER\_EFF\_RB}
+\end{itemize}
+
+\subsection{Cycling}
+
+The default cycling is a V-cycle, but W-cycles and full multigrid are
+supported as well.
+
+\subsection{Bottom Solvers}
+
+The multigrid cycling coarsens the grids as part of the solve.  When 
+the coarsest grid is reached, the individual boxes that comprise that
+level are coarsened as much as then can, down to $2^3$ zones.  Depending
+on the distribution of sizes of the grids, it may not be possible for
+everything to reach this minimum size.  At this point, the bottom
+solver is invoked.  Most of these will solve the linear system
+on this collection of grids directly.  There is one special bottom
+solver that will define a new box encompassing all of the coarsened
+grids and then put the data on fewer boxes and processors and further
+coarsen the problem, again until we get as close to $2^3$ as possible.
+At that point, one of the other bottom solvers will be called upon
+to solve the problem.
+
+There are several bottom solvers available in \boxlib.  For \maestro.
+These are set through the {\tt mg\_bottom\_solver} (MAC/cell-centered)
+and {\tt hg\_bottom\_solver} (nodal) runtime parameters.
+The allowed values are:
 \begin{itemize}
@@
 
+You should use the special bottom solver (4) whenever possible, even
+if it means changing your gridding strategy (as discussed below) to 
+make it more efficient.
+
 \MarginPar{any simple discussion on why we might choose one of the
@@
 
-\subsection{Special Bottom Solver}
-We have a special bottom solver type that greatly speeds up large problems
-and speeds up smaller problems to a lesser degree.
-In fact, you should use this bottom solver whenever possible, because I
-have yet to find a case that performs worse than our standard bottom solver.
-What this special solver essentially does is take the data from the coarsest 
-level of the original multigrid V-cycle and copy it onto a new grid structure 
-with the same number of total cells in each direction, but with a fewer number of larger 
-grids.  A new V-cycle begins from this point, 
-so we are essentially coarsening this ``new'' problem.
-Now, the coarsest level of the multigrid V-cycle in the ``new'' problem has 
-fewer cells and fewer grids as compared to the original coarsest level.
+\subsubsection{Special Bottom Solver}
+
+The special solver takes the data from the coarsest level of the
+original multigrid V-cycle and copies it onto a new grid structure with
+the same number of total cells in each direction, but with a fewer
+number of larger grids.  A new V-cycle begins from this point, so we
+are essentially coarsening this ``new'' problem.  Now, the coarsest
+level of the multigrid V-cycle in the ``new'' problem has fewer cells
+and fewer grids as compared to the original coarsest level.
 
 To enable this solver, set {\tt hg\_bottom\_solver = 4} (for the nodal
-projections) and/or {\tt mg\_bottom\_solver = 4} (for the cell-centered
-projections) in your inputs file.
+projections) and/or {\tt mg\_bottom\_solver = 4} (for the
+cell-centered projections) in your inputs file.
 
-To understand how this bottom solver works, the first thing you need to know
-is what the grid structure of the coarsest level of your multigrid V-cycle
-looks like (we need to put a description of this in the section above).
-Now, figure out the size of the box you would need if you
-wanted it to fit all the data on the coarsest level.  Now, figure out what
-the largest integer $n$ is so that you can evenly divide the length of this box
-by $2^n$ in every coordinate direction.  If $n < 2$, the program will abort
-since the grid structure is not suitable for this bottom solver.
+To understand how this bottom solver works, the first thing you need
+to know is what the grid structure of the coarsest level of your
+multigrid V-cycle looks like.  Next, figure out the size of the box you
+would need if you wanted it to fit all the data on the coarsest level.
+Finally, figure out what the largest integer $n$ is so that you can evenly
+divide the length of this box by $2^n$ in every coordinate direction.
+If $n < 2$, the program will abort since the grid structure is not
+suitable for this bottom solver.
 
-Now the code will set up a ``new'' problem, using the data at the coarsest level
-of the original problem as the initial data.  The grid structure for this new
-problem has the same number of cells as the coarsest level of the original problem,
-but the data is copied onto a grid structure where each grid has $2^n$ cells
-on each side.  The new V-cycle continues down to the new coarsest level, in
-which each grid has 2 cells on each side.  If you wish to impose a limit on
-the maximum value that $n$ can have, you can do so by setting 
-{\tt max\_mg\_bottom\_nlevs} equal to that value.
+The code will set up a ``new'' problem, using the data at the
+coarsest level of the original problem as the initial data.  The grid
+structure for this new problem has the same number of cells as the
+coarsest level of the original problem, but the data is copied onto a
+grid structure where each grid has $2^n$ cells on each side.  The new
+V-cycle continues down to the new coarsest level, in which each grid
+has 2 cells on each side.  If you wish to impose a limit on the
+maximum value that $n$ can have, you can do so by setting {\tt
+max\_mg\_bottom\_nlevs} equal to that value.
 
-{\bf Example 1:} A 3D problem with $384^3$ cells divided into $32^3$ grids, i.e., 
-there is a $12\times 12\times 12$ block of $32^3$ grids.  The coarsest level of the 
-multigrid V-cycle contains $12\times 12\times 12$ grids that have $2^3$ cells, so the 
-entire problem domain has $24^3$ cells.  We see that $n=3$, and create a new problem 
-domain with a $3\times 3\times 3$ block of $8^3$ grids.  The coarsest level of the 
-multigrid V-cycle for the ``new'' problem will be a $3\times 3\times 3$ block of 
-$2^3$ grids.
+Some grid examples help make this clear:
+\begin{itemize}
+\item {\bf Example 1:} A 3D problem with $384^3$ cells divided into $32^3$
+grids, i.e., there is a $12\times 12\times 12$ block of $32^3$ grids.
+The coarsest level of the multigrid V-cycle contains $12\times
+12\times 12$ grids that have $2^3$ cells, so the entire problem domain
+has $24^3$ cells.  We see that $n=3$, and create a new problem domain
+with a $3\times 3\times 3$ block of $8^3$ grids.  The coarsest level
+of the multigrid V-cycle for the ``new'' problem will be a $3\times
+3\times 3$ block of $2^3$ grids.
 
-{\bf Example 2:} A 2D problem with $96\times 384$ cells divided into $48^2$ grids, i.e., 
-there is a $2\times 8$ block of $48^2$ grids.  The coarsest level of the multigrid 
-V-cycle contains $2\times 8$ grids that have $3^2$ cells, so the entire problem 
-domain has $6\times 24$ cells.  We see that $n=1$, so the program aborts since this grid
+\item {\bf Example 2:} A 2D problem with $96\times 384$ cells divided into
+$48^2$ grids, i.e., there is a $2\times 8$ block of $48^2$ grids.  The
+coarsest level of the multigrid V-cycle contains $2\times 8$ grids
+that have $3^2$ cells, so the entire problem domain has $6\times 24$
+cells.  We see that $n=0$, so the program aborts since this grid
 structure is not appropriate for the fancy bottom solver.
+\end{itemize}
+
+\section{\maestro's Multigrid Use}
+
+\maestro\ uses multigrid to enforce the velocity constraint through
+projections at the half-time (the MAC projection) and end of the time
+step (the HG projection).  Two multigrid solvers are provided by
+\boxlib---one for cell-centered data and one for node-centered (nodal)
+data.  Both of these are used in \maestro.
+
+The MAC projection operates on the advective velocities predicted at
+the cell-interfaces at the half-time.  The edge-centered velocities
+are shown in Figure~\ref{fig:mg:MAC}.  If we consider purely
+incompressible flow, the projection appears as:
+\begin{equation}
+D G \phi = D U
+\end{equation}
+where $D$ is the divergence operator and $G$ is the gradient operator.
+In this discretization, $\phi$ is cell-centered (see
+Figure~\ref{fig:mg:MAC}).  The remaining quantities are discretized as:
+\begin{itemize}
+\item $DU$ is cell-centered, 
+  \begin{equation}
+  (DU)_{i,j} = \frac{u_{i+1/2,j} - u_{i-1/2,j}}{\Delta x} + 
+               \frac{v_{i,j+1/2} - v_{i,j-1/2}}{\Delta y}
+  \end{equation}
+
+\item $G\phi$ is edge-centered, on the MAC grid, as shown in
+  Figure~\ref{fig:mg:MAC}.
+
+\item $DG\phi$ is cell-centered, also shown in Figure~\ref{fig:mg:MAC},
+  computed from $G\phi$ using the same differencing as $DU$.
+
+\end{itemize}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=5.5in]{\mgfigpath/MAC_mg2}
+\caption{\label{fig:mg:MAC} Data centerings for the MAC projection}
+\end{figure}
+
+The HG projection projects the cell-centered velocities at the end of
+the timestep.  Here, $\phi$ is node-centered.  Figure~\ref{fig:mg:HG}
+shows the locations of the various quantities involved in the HG
+projection.  Again considering simple incompressible flow, we now
+solve:
+\begin{equation}
+L \phi = D U
+\end{equation}
+where $L$ is a discretization of the Laplacian operator.  In this
+sense, the HG projection is an {\em approximate projection}, that is,
+$L \neq DG$ (in discretized form).  The various operations have the
+following centerings:
+\begin{itemize}
+
+\item $DU$ is node-centered.  This is computed as:
+  \begin{equation}
+  (DU)_{i-1/2,j-1/2} = \frac{\frac{1}{2} (u_{i,j} + u_{i,j-1}) -
+                             \frac{1}{2} (u_{i-1,j} + u_{i-1,j-1})}{\Delta x} +
+                       \frac{\frac{1}{2} (v_{i,j} + v_{i-1,j}) -
+                             \frac{1}{2} (v_{i,j-1} + v_{i-1,j-1})}{\Delta y} 
+  \end{equation}
+
+\item $G\phi$ is cell-centered, as shown in Figure~\ref{fig:mg:HG}.
+
+\item $L\phi$ is node-centered.  This is a direct discretization of 
+the Laplacian operator.  By default, \maestro\ uses a dense stencil
+(9-points in 2-d, 27-points in 3-d).  Alternately, a {\em cross}
+stencil can be used (by setting {\tt hg\_dense\_stencil = F}).  This
+uses 5-points in 2-d, 7-points in 3-d.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=5.5in]{\mgfigpath/HG_mg2}
+\caption{\label{fig:mg:HG} Data centerings for the HG projection}
+\end{figure}
+
+\end{itemize}
 
@@
 
-\MyArrow[1em]{start1}{end1}{colorii,line width=2pt}
-\MyArrow[1em]{start2}{end2}{colorii,line width=2pt}
+\maestro\ multigrid solves always involve the full AMR hierarchy.
 
 \subsection{Cell-Centered MG}
+
+The flowchart below shows the structure of a cell-centered multigrid
+solve using pure V-cycles.
 
@@
   \item {\em recursively descend V-cycle}
+\MyArrow[1em]{start1}{end1}{colorii,line width=2pt}
   \begin{itemize} 
@@
   \item {\em ascend V-cycle}
-
+\MyArrow[1em]{start2}{end2}{colorii,line width=2pt}
   \begin{itemize}
@@
 
+\subsection{Nodal MG}
+
 \section{Convergence Criteria}
@@
 
-\section{Multigrid Solver Tolerances}
+\subsection{Multigrid Solver Tolerances}
 

### 2014-07-22 Docs/mg/mg.tex
more MG docs

@@
 \begin{itemize}
-\item {MG\_SMOOTHER\_GS\_RB}: a red-black Gauss-Seidel smoother
+\item {\tt MG\_SMOOTHER\_GS\_RB}: a red-black Gauss-Seidel smoother
 
-\item {MG\_SMOOTHER\_JACOBI}: a Jacobi smoother (not implemented for 
+\item {\tt MG\_SMOOTHER\_JACOBI}: a Jacobi smoother (not implemented for 
 the dense nodal stencil)
 
-\item {MG\_SMOOTHER\_MINION\_CROSS}
+\item {\tt MG\_SMOOTHER\_MINION\_CROSS}
 
-\item {MG\_SMOOTHER\_MINION\_FULL}
+\item {\tt MG\_SMOOTHER\_MINION\_FULL}
 
-\item {MG\_SMOOTHER\_EFF\_RB}
+\item {\tt MG\_SMOOTHER\_EFF\_RB}
 \end{itemize}
@@
 
-\item {\tt mg\_bottom\_solver} / {hg\_bottom\_solver} = 0: smoothing only.
+\item {\tt mg\_bottom\_solver} / {\tt hg\_bottom\_solver = 0}: smoothing only.
 
-\item {\tt mg\_bottom\_solver} / {hg\_bottom\_solver} = 1: biconjugate
+\item {\tt mg\_bottom\_solver} / {\tt hg\_bottom\_solver = 1}: biconjugate
   gradient stabilized---this is the default.
 
-\item {\tt mg\_bottom\_solver} / {hg\_bottom\_solver} = 2: conjugate
+\item {\tt mg\_bottom\_solver} / {\tt hg\_bottom\_solver = 2}: conjugate
   gradient method
 
-\item {\tt mg\_bottom\_solver} / {hg\_bottom\_solver} = 4: a special 
+\item {\tt mg\_bottom\_solver} / {\tt hg\_bottom\_solver = 4}: a special 
   bottom solver that extends the range of the multigrid coarsening
   by aggregrating coarse grids on the original mesh together and
-  further coarsening. \MarginPar{does this option then use bicg 
-  at the new bottom?}
+  further coarsening.
 
@@
 \end{itemize}
+
+\section{Flowchart}
+
+\maestro\ multigrid solves always involve the full AMR hierarchy.
+
+\subsection{Cell-Centered MG}
+
+The flowchart below shows the structure of a cell-centered multigrid
+solve using pure V-cycles.
+
+\begin{itemize}
+
+\item {\tt stencil\_fill\_cc\_all\_mglevels} / {\tt stencil\_fill\_cc}:
+  Compute all of the stencil coefficients
+  for the Laplacian operator at all cells.  At the C-F interfaces, the
+  stencil coefficients are modified to know this.
+
+\item {\tt ml\_cc}: The main driver for the cell-centered multigrid.
+   Among other things, this computes the norm that will be used
+   for convergence testing.
+
+\item {\tt mg\_tower\_v\_cycle} (recursive):
+\begin{itemize}
+
+  \item {\em recursively descend V-cycle}
+\MyArrow[1em]{start1}{end1}{colorii,line width=2pt}
+  \begin{itemize} 
+  \item \tikzmark{end1} {\tt cc\_mg\_tower\_smoother}: Smooth the 
+    problem at the current MG level using the desired smoother.
+
+  \item {\tt compute\_defect}: Construct $f - L\phi$.
+
+  \item\tikzmark{start1} {\tt mg\_tower\_restriction}:  Restrict
+    the defect to the coarser level by conservative averaging.
+  \end{itemize}
+
+  \item {\tt mg\_tower\_bottom\_solve}:  Solve the coarsened problem
+    using the chosen bottom solver.
+
+  \item {\em ascend V-cycle}
+\MyArrow[1em]{start2}{end2}{colorii,line width=2pt}
+  \begin{itemize}
+  \item\tikzmark{end2} {\tt mg\_tower\_prolongation}: Take the solution at level $n-1$ and use it to 
+     correct the solution at level $n$ by representing the data on the finer grid.  This uses
+     linear reconstruction for jumps by $2\times$ and piecewise-constant otherwise.
+
+  \item\tikzmark{start2} {\tt cc\_mg\_tower\_smoother}:
+  \end{itemize}
+
+\end{itemize}
+\item {\tt compute\_defect}: This is called multiple times, checking for
+   convergence at each level.
+
+\end{itemize}
+
+\subsection{Nodal MG}
 
@@
 \end{itemize}
-
-\section{Flowchart}
-
-\maestro\ multigrid solves always involve the full AMR hierarchy.
-
-\subsection{Cell-Centered MG}
-
-The flowchart below shows the structure of a cell-centered multigrid
-solve using pure V-cycles.
-
-\begin{itemize}
-
-\item {\tt stencil\_fill\_cc\_all\_mglevels} / {\tt stencil\_fill\_cc}:
-
-\item {\tt ml\_cc}:
-
-\item {\tt mg\_tower\_v\_cycle} (recursive):
-\begin{itemize}
-
-  \item {\em recursively descend V-cycle}
-\MyArrow[1em]{start1}{end1}{colorii,line width=2pt}
-  \begin{itemize} 
-  \item \tikzmark{end1} {\tt cc\_mg\_tower\_smoother}:
-
-  \item {\tt compute\_defect}:
-
-  \item\tikzmark{start1} {\tt ml\_restriction}:
-  \end{itemize}
-
-  \item {\tt mg\_tower\_bottom\_solve}:
-
-  \item {\em ascend V-cycle}
-\MyArrow[1em]{start2}{end2}{colorii,line width=2pt}
-  \begin{itemize}
-  \item\tikzmark{end2} {\tt mg\_tower\_prolongation}: Take the solution at level $n-1$ and use it to 
-     correct the solution at level $n$ by representing the data on the finer grid.  This uses
-     linear reconstruction for jumps by $2\times$ and piecewise-constant otherwise.
-
-  \item\tikzmark{start2} {\tt cc\_mg\_tower\_smoother}:
-  \end{itemize}
-
-\end{itemize}
-\item multiple {\tt compute\_defect} calls:
-
-\end{itemize}
-
-\subsection{Nodal MG}
 

### 2014-07-25 Docs/FAQ/faq.tex

@@
 
+\item {\em How can I debug a parallel (MPI) job with gdb?}
+
+If you only need to use a few processors, the following command will work:
+\begin{verbatim}
+mpiexec -n 4 xterm -e gdb ./main.Linux.gfortran.mpi.exe 
+\end{verbatim}
+where the executable needs to be created with the ``{\tt -g}'' flag to
+the compiler.  This will pop up multiple {\tt xterm}s with {\tt gdb} running
+in each.  You need to then issue:
+\begin{verbatim}
+run inputs
+\end{verbatim}
+where {\tt inputs} is the desired inputs file {\em in each} {\tt xterm}.
+
 \end{enumerate}

### 2014-07-25 Docs/mg/mg.tex

@@
 
-Here we describe some of the general ideas behind the \boxlib\ multigrid.
+Here we describe some of the general ideas behind the \boxlib\
+multigrid (MG).
 
@@
 
+The MG solvers are located in {\tt BoxLib/Src/LinearSolvers/F\_MG/}.
+There are two MG solvers, for cell-centered and nodal data.
+Generally, the routines specific to the cell-centered solver will have
+{\tt cc} in their name, and those specific to the nodal solver will have
+{\tt nd} in their name.
+
 Support for $\Delta x \ne \Delta y \ne \Delta z$
-
-There are two MG solvers: cell-centered and nodal.
 

### 2014-07-25 Docs/flowchart/flowchart.tex
updated to include delta chi update for volume discrepancy for both full and base state velocity

@@
 \item In {\bf Step 2B}, to compute $w_0$, we need to account for the volume discrepancy
-term by computing $p_{\rm EOS} = \overline{p(\rho,h,X_k)^n}$.
+term by first defining $p_{\rm EOS}^n = \overline{p(\rho,h,X_k)^n}$, and then using:
+\begin{equation}
+\frac{\partial w_0^{\nph,\star}}{\partial r} = \overline{S^{\nph,\star\star}} - \frac{1}{\overline{\Gamma_1^n}p_0^n}\psi^{\nmh} - \underbrace{\frac{f}{\overline{\Gamma_1^n}p_0^n}\left(\frac{p_0^n-\overline{p_{\rm EOS}^n}}{\Delta t}\right)}_{\delta\chi_{w_0}}.
+\end{equation}
 \item In {\bf Step 3}, the MAC projection should account for the volume discrepancy term:
@@
 \nabla \cdot \left(\beta_0^n \uadvone\right) = 
-\beta_0^n \left\{ \left(S^{\nph,\star\star} - \overline{S^{\nph,\star\star}}\right)
-+ \frac{f}{\gammabar^n p_0^n}
-\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)^n}}{\Delta t^n}\right]\right\}.
+\beta_0^n \left[ \left(S^{\nph,\star\star} - \overline{S^{\nph,\star\star}}\right)
++ \underbrace{\frac{f}{\gammabar^n p_0^n}
+\left(\frac{p_{\rm EOS}^n - \overline{p_{\rm EOS}^n}}{\Delta t^n}\right)}_{\delta\chi}\right].
 \end{equation}
 \item In {\bf Step 6B}, to compute $w_0$, we need to account for the volume discrepancy
-term by computing $p_{\rm EOS} = \left[\overline{p(\rho,h,X_k)^n} + \overline{p(\rho,h,X_k)^{n+1,\star}}\right]/2$.
+term by first defining $p_{\rm EOS}^{n+1,\star} = p(\rho,h,X_k)^{n+1,\star}$,
+$\overline{\Gamma_1^{\nph,\star}} = (\overline{\Gamma_1^{n}}+\overline{\Gamma_1^{n+1,\star}})/2$,
+and $p^{\nph,\star} = (p^{n}+p^{n+1,\star})/2$, and then using:
+\begin{equation}
+\frac{\partial w_0^{\nph}}{\partial r} = \overline{S^{\nph,\star}} - \frac{1}{\overline{\Gamma_1^{\nph,\star}}p_0^{\nph,\star}}\psi^{\nph,\star} - \frac{f}{\overline{\Gamma_1^{n+1,\star}}p_0^{n+1,\star}}\left(\frac{p_0^{n+1,\star}-\overline{p_{\rm EOS}^{n+1,\star}}}{\Delta t}\right) - \delta\chi_{w_0}
+\end{equation}
 \item In {\bf Step 7}, the MAC projection should account for the volume discrepancy term:
 \begin{equation}
-\nabla \cdot \left(\beta_0^{\nph,\star} \uadvtwo\right) = \beta_0^{\nph,\star}\left\{\left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right) + \frac{f}{\overline{\Gamma_1^{\nph,\star}} p_0^{\nph,\star}} \left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)^{\nph,\star}}}{\Delta t^n}\right]\right\},
+\nabla \cdot \left(\beta_0^{\nph,\star} \uadvtwo\right) = \beta_0^{\nph,\star}\left[\left(S^{\nph,\star} - \overline{S^{\nph,\star}}\right) + \frac{f}{\overline{\Gamma_1^{n+1,\star}} p_0^{n+1,\star}} \left(\frac{p_{\rm EOS}^{n+1,\star} - \overline{p_{\rm EOS}^{n+1,\star}}}{\Delta t^n}\right) + \delta\chi\right],
 \end{equation}

### 2014-08-01 Docs/mg/mg.tex
add a definition of bndry_reg from Ann

@@
 
-\subsection{\mgtower}
+\subsection{Data Structures}
+
+\subsubsection{\mgtower}
 
@@
 \end{itemize}
+
+\subsection{\bndryreg}
+
+A \bndryreg\ is a set of {\tt dim*2} multifabs defined on the coarse cells
+immediately outside of each fine grid.
+
+Ideally we would only include the cells not covered by a different
+fine grid, but in the current implementation, for {\tt N} fine grids we have
+{\tt N*dim*2} multifabs (the factor of 2 is {\tt lo}/{\tt hi}).
+
+We use the \bndryreg to store the correction to the residual to be
+used in the coarse grid relaxation.  The way we use it is that we
+compute a pure coarse grid residual in every coarse cell, then we
+correct it by subtracting the contribution of the coarse flux under
+the coarse-fine edge and adding the contribution of the fine fluxes at
+that same interface.  The \bndryreg\ holds the cell-centered
+contribution from the difference of these edge fluxes and is added to
+the coarse residual multifab.
 

### 2014-08-07 Docs/mg/mg.tex
fix sectioning
nodal flow chart

@@
 
-\subsection{\bndryreg}
+\subsubsection{\bndryreg}
 
@@
 \subsection{Nodal MG}
+
+The flowchart below shows the structure of a cell-centered multigrid
+solve using pure V-cycles.
+
+\begin{itemize}
+
+\item {\tt stencil\_fill\_cc\_all\_mglevels} / {\tt stencil\_fill\_cc}:
+  For the nodal solver, this applies the weights to the 
+  coefficients. \MarginPar{nodal does things differently than cc---need more here}
+
+\item {\tt ml\_nd}: The main driver for the nodal multigrid.
+
+\item {\tt mg\_tower\_v\_cycle} (recursive):
+\begin{itemize}
+
+  \item {\em recursively descend V-cycle}
+\MyArrow[1em]{start3}{end3}{colorii,line width=2pt}
+  \begin{itemize} 
+  \item \tikzmark{end3} {\tt nodal\_smoother\_?d}: Smooth the 
+    problem at the current MG level using the desired smoother.
+
+  \item {\tt compute\_defect}: Construct $f - L\phi$.
+
+  \item\tikzmark{start3} {\tt mg\_tower\_restriction}:  Restrict
+    the defect to the coarser level by simply taking the fine value that
+    lies at the same place as the coarse data.
+  \end{itemize}
+
+  \item {\tt mg\_tower\_bottom\_solve}:  Solve the coarsened problem
+    using the chosen bottom solver.
+
+  \item {\em ascend V-cycle}
+\MyArrow[1em]{start4}{end4}{colorii,line width=2pt}
+  \begin{itemize}
+  \item\tikzmark{end4} {\tt mg\_tower\_prolongation}: For nodal data, the fine grid
+     will have some points at exactly the same place as the coarse data---these are
+     simply copied to the fine grid.  The remain data is interpolated.
+
+  \item\tikzmark{start4} {\tt modal\_smoother\_?d}:
+  \end{itemize}
+
+\end{itemize}
+\item {\tt compute\_defect}: This is called multiple times, checking for
+   convergence at each level.
+
+\end{itemize}
 

### 2014-08-07 Docs/mg/mg.tex
bndry_reg corrections from the boxlib docs

@@
 
-A \bndryreg\ is a set of {\tt dim*2} multifabs defined on the coarse cells
-immediately outside of each fine grid.
+There are two types of \bndryreg\ objects, each of which is a set of
+{\tt dim} multifabs.  The first, which is built with the {\tt
+bndry\_reg\_build} call, is defined on all cells immediately outside
+of each grid.  Each multifab within the \bndryreg\ contains all the
+fabs on both the low and high faces for a given direction.  In the
+context of multigrid, we fill the \bndryreg\ with coarse data before
+interpolating the data to the correct locations to be used in the
+multigrid stencil.  The second type of \bndryreg\ object is defined on
+the coarse cells immediately outside each fine grid, and is defined
+with the {\tt bndry\_reg\_rr\_build} call, is defined on all cells
+immediately outside.  For this latter type, the option is available
+to include only cells not covered by a different fine grid, but this
+is left as an option because it requires additional calculations of
+box intersections.
 
-Ideally we would only include the cells not covered by a different
-fine grid, but in the current implementation, for {\tt N} fine grids we have
-{\tt N*dim*2} multifabs (the factor of 2 is {\tt lo}/{\tt hi}).
+In multigrid, we use the latter type of \bndryreg\ to calculate the
+residual at coarse cells adjacent to fine grids that is used as the
+right hand side in the relaxation step at the coarse level on the way
+down the V-cycle (or other).  The first type of \bndryreg is used to
+hold boundary conditions for the fine grids that are interpolated from
+the coarse grid solution; this is filled on the way back up the
+V-cycle.
 
-We use the \bndryreg to store the correction to the residual to be
-used in the coarse grid relaxation.  The way we use it is that we
-compute a pure coarse grid residual in every coarse cell, then we
-correct it by subtracting the contribution of the coarse flux under
-the coarse-fine edge and adding the contribution of the fine fluxes at
-that same interface.  The \bndryreg\ holds the cell-centered
-contribution from the difference of these edge fluxes and is added to
-the coarse residual multifab.
+To compute the residual in a coarse cell adjacent to a fine grid, we
+first compute the pure coarse residual, then subtract the contribution
+from the coarse face underlying the coarse-fine interface, and add the
+contribution of the fine faces at the coarse-fine interface.
+The \bndryreg\ holds the cell-centered contribution from the
+difference of these edge fluxes and is added to the coarse residual
+multifab.
 

### 2014-08-08 Docs/flowchart/flowchart.tex
start updating the flowchart -- put in the form of the equations we are currently solving.  Need to go through it to make sure it clearly shows the alternate energy formulation stuff.

@@
 spectral deferred corrections is outlined in Chapter~\ref{ch:sdc}.)
+
+\subsection{Base State}
+
+The stratified atmosphere is characterized by a one-dimensional
+time-dependent base state, defined by a base state density, $\rho_0$,
+and a base state pressure, $p_0$, in hydrostatic equilibrum:
+\begin{equation}
+\nabla p_0 = -\rho_0 |g| \er
+\end{equation}
+
+For the time-dependence, we will define a base state velocity, $w_0$,
+which will adjust the base state from one hydrostatic equilibrum to
+another in response to heating.
+
+For convenience, we define a base state enthalphy, $h_0$, as needed
+by laterally averaging the full enthalpy, $h$.
+
+\subsection{Continuity}
+
+Conservation of mass gives the same continuity equation we have with
+compressible flow:
+\begin{equation}
+\frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) = 0
+\end{equation}
+
+Additionally, we carry species around which can react.  The creation and destruction
+of the species is described by their create rate, $\omegadot_k$, and the species
+are defined by their mass fractions, $X_k \equiv \rho_k / \rho$, giving
+\begin{equation}
+\frac{\partial \rho X_k}{\partial t} + \nabla \cdot (\rho \Ub) = \rho \omegadot_k
+\end{equation}
+and
+\begin{equation}
+\sum_k X_k = 1
+\end{equation}
+
+\subsection{Constraint}
+
+The equation of state is cast into an elliptic constraint on the velocity field 
+by differentiating $p_0(\rho, s, X_k)$ along particle paths, giving:
+\begin{equation}
+\nabla \cdot (\beta_0 \Ub) = \beta_0 \left ( S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )
+\end{equation}
+where
+ $\beta_0$ is a density-like variable that carries background stratification, defined as
+\begin{equation}
+\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
+\end{equation}
+and 
+\begin{equation}
+S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext
+\end{equation}
+where $p_{X_k} \equiv \left. \partial p / \partial X_k
+\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h / 
+\partial X_k \right |_{p,T,X_{j,j\ne k}}, 
+p_\rho \equiv \left. \partial p/\partial \rho \right |_{T, X_k}$, and
+$\sigma \equiv p_T/(\rho c_p p_\rho)$, with
+$p_T \equiv \left. \partial p / \partial T \right|_{\rho, X_k}$ and
+$c_p \equiv \left.  \partial h / \partial T
+\right|_{p,X_k}$ is the specific heat at constant pressure.
+
+\subsection{Momentum}
+
+The compressible momentum equation (written in terms of velocity is):
+\begin{equation}
+\rho \frac{\partial \Ub}{\partial t} + \rho \Ub \cdot \nabla \Ub + \nabla p = -\rho |g| \er
+\end{equation}
+Subtracting off the base state, and defining the perturbational
+pressure (sometimes called the dynamic pressure) as $\pi = p -                                    
+p_0$, and perturbational density as $\rho' = \rho - \rho_0$, we have:
+\begin{equation}
+\rho \frac{\partial \Ub}{\partial t} + \rho \Ub \cdot \nabla \Ub + \nabla \pi = -\rho' |g| \er
+\end{equation}
+or
+\begin{equation}
+\frac{\partial \Ub}{\partial t} + \Ub \cdot \nabla \Ub + \frac{1}{\rho} \nabla \pi =
+   -\frac{\rho^\prime}{\rho} |g| \er
+\end{equation}
+This is the form of the momentum equation that we solved in papers I--III and in the multilevel
+paper.
+
+Several authors \cite{KP:2012,VLBWZ:2013} explored the idea of energy
+conservation in a low Mach number system and found that an additional term (which can
+look like a buoyancy) is needed in the low Mach number formulation, yielding:
+\begin{equation}
+\frac{\partial \Ub}{\partial t} + \Ub \cdot \nabla \Ub +
+   \frac{\beta_0}{\rho} \nabla \left (\frac{p^\prime}{\beta_0} \right ) =
+   -\frac{\rho^\prime}{\rho} |g| \er
+\end{equation}
+
+We decompose the full velocity field into a base
+state velocity, $w_0$, that governs the base state dynamics, and a
+local velocity, $\Ubt$, that governs the local dynamics, i.e.,
+\begin{equation}
+\Ub = w_0(r,t)\eb_r + \Ubt(\xb,t).
+\end{equation}
+with
+$\overline{(\Ubt\cdot\eb_r)} = 0$ and
+$w_0 = \overline{(\Ub\cdot\eb_r)}$.
+The velocity evolution equations are then
+\begin{eqnarray}
+\frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
+  r} - \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r},\label{eq:w0
+  evolution}\\ \frac{\partial\Ubt}{\partial t} &=& -\Ub\cdot\nabla\Ubt
+- \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
+\frac{1}{\rho}\nabla\pi +
+\frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r -
+\frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:utildeupd}
+\end{eqnarray}
+where $\pi_0$ is the base state component of the perturbational pressure.
+By laterally averaging to equation (\ref{eq:U divergence}),
+we obtain a divergence constraint for $w_0$:
+\begin{equation}
+\nabla\cdot(\beta_0 w_0 \eb_r) = \beta_0\left(\Sbar - \frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:w0 divergence}
+\end{equation}
+The divergence constraint for $\Ubt$ can be found by subtracting (\ref{eq:w0 divergence})
+into (\ref{eq:U divergence}), resulting in
+\begin{equation}
+\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0\left(S-\Sbar\right).\label{eq:utilde divergence}
+\end{equation}
+
+\subsection{Energy}
+
+Finally, we add an equation for specific enthalpy evolution to our
+system.  Strictly speaking this is not necessary to close the system,
+but it becomes convenient at times to define the temperature.  
+\begin{equation}
+\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc + \rho\Hext,\label{eq:enthalpy}
+\end{equation}
+
+\subsection{Base State and Perturbation Form Equations}
 

### 2014-08-09 Docs/flowchart/flowchart.tex
more work on what we are actually solving

@@
-Here we outline the algorithm currently implemented in the code.  The
-latest published reference for MAESTRO is the multilevel
-paper \cite{multilevel}.  In this description, we make frequent
-reference to paper~I \cite{lowMach}, paper~II \cite{lowMach2},
-paper~III \cite{lowMach3}, and paper~IV \cite{lowMach4}.
+The equation set and solution procedure used by \maestro\ has evolved
+over time.  In this chapter, we outline the algorithm currently
+implemented in the code.  The latest published reference for \maestro
+is the multilevel paper \cite{multilevel}.  In this description, we
+make frequent reference to paper~I \cite{lowMach},
+paper~II \cite{lowMach2}, paper~III \cite{lowMach3}, and
+paper~IV \cite{lowMach4}.
 
@@
 
+The gravitational acceleration, $g$ is either constant or a point-mass
+with a $1/r^2$ dependence (see \S \ref{sec:planarinvsqgravity}) for plane-parallel geometries, or a monopole
+constructed by integrating the base state density for spherical
+geometries.
+
 For the time-dependence, we will define a base state velocity, $w_0$,
@@
 \begin{equation}
-\frac{\partial \rho X_k}{\partial t} + \nabla \cdot (\rho \Ub) = \rho \omegadot_k
+\frac{\partial \rho X_k}{\partial t} + \nabla \cdot (\rho \Ub X_k) = \rho \omegadot_k
 \end{equation}
@@
 
-The equation of state is cast into an elliptic constraint on the velocity field 
-by differentiating $p_0(\rho, s, X_k)$ along particle paths, giving:
+The equation of state is cast into an elliptic constraint on the
+velocity field by differentiating $p_0(\rho, s, X_k)$ along particle
+paths, giving:
 \begin{equation}
@@
 
+In this constraint, $\gammabar$ is the lateral average of
+$\Gamma_1 \equiv d\log p / d\log \rho |_s$.  Using the lateral average
+here makes it possible to cast the constraint as a
+divergence.  \cite{KP:2012} discuss the general case where we want to
+keep the local variations of $\Gamma_1$ (and we explored this in paper
+III).  We also look at this in \S~\ref{sec:flow:gamma1vary}.
+
 \subsection{Momentum}
@@
 Subtracting off the base state, and defining the perturbational
-pressure (sometimes called the dynamic pressure) as $\pi = p -                                    
-p_0$, and perturbational density as $\rho' = \rho - \rho_0$, we have:
+pressure (sometimes called the dynamic pressure) as $\pi = p - p_0$,
+and perturbational density as $\rho' = \rho - \rho_0$, we have:
 \begin{equation}
@@
 \end{equation}
-This is the form of the momentum equation that we solved in papers I--III and in the multilevel
-paper.
+This is the form of the momentum equation that we solved in papers
+I--IV and in the multilevel paper.
 
 Several authors \cite{KP:2012,VLBWZ:2013} explored the idea of energy
-conservation in a low Mach number system and found that an additional term (which can
-look like a buoyancy) is needed in the low Mach number formulation, yielding:
+conservation in a low Mach number system and found that an additional
+term (which can look like a buoyancy) is needed in the low Mach number
+formulation, yielding:
 \begin{equation}
@@
 \end{equation}
-
-We decompose the full velocity field into a base
-state velocity, $w_0$, that governs the base state dynamics, and a
-local velocity, $\Ubt$, that governs the local dynamics, i.e.,
+This is the form that we enforce in \maestro, and the choice is controlled
+by \runparam{use\_alt\_energy\_fix}.
+
+We decompose the full velocity field into a base state velocity,
+$w_0$, that governs the base state dynamics, and a local velocity,
+$\Ubt$, that governs the local dynamics, i.e., 
 \begin{equation}
@@
 $w_0 = \overline{(\Ub\cdot\eb_r)}$.
-The velocity evolution equations are then
+The velocity evolution equations are then \MarginPar{we need the $\beta_0$ here}
 \begin{eqnarray}
@@
 \end{equation}
-The divergence constraint for $\Ubt$ can be found by subtracting (\ref{eq:w0 divergence})
-into (\ref{eq:U divergence}), resulting in
+The divergence constraint for $\Ubt$ can be found by subtracting
+(\ref{eq:w0 divergence}) into (\ref{eq:U divergence}), resulting in
 \begin{equation}
@@
 %-----------------------------------------------------------------------------
-\subsection{$\Gamma_1$ Variation Changes}
+\subsection{\label{sec:flow:gamma1vary} $\Gamma_1$ Variation Changes}
 

### 2014-08-09 Docs/flowchart/flowchart.tex
more equation writing
more equation summarizing

@@
 \frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) = 0
+\label{eq:flow:continuity}
 \end{equation}
@@
 
+The base state density evolution equation can be defined by laterally averaging the 
+full continuity equation, giving:
+\begin{equation}
+\frac{\partial\rho_0}{\partial t} = -\nabla\cdot(\rho_0 w_0 \eb_r),\label{eq:flow:base_density}
+\end{equation}
+
+Subtracting these two yields the evolution equation for the perturbational
+density, $\rho^\prime \equiv \rho - \rho_0$:
+\begin{equation}
+\frac{\partial\rho'}{\partial t} = -\Ub\cdot\nabla\rho' -
+  \rho'\nabla\cdot\Ub - \nabla\cdot\left(\rho_0\Ubt\right) \label{eq:flow:rhoprime}
+\end{equation}
+
+As first discussed in paper III and then refined in the multilevel paper, we capture
+the changes that can occur due to significant convective
+overturning by imposing the constraint that $\overline{\rho'}=0$
+for all time.  This gives
+\begin{equation}
+\frac{\partial\overline{\rho'}}{\partial t} = -\nabla\cdot(\etarho\eb_r).
+\end{equation}
+where
+\begin{equation}
+\etarho = \overline{\left(\rho'\Ub\cdot\eb_r\right)}
+\end{equation}
+In practice, we correct the drift by simply setting $\rho_0 =
+\overline{\rho}$ after the advective update of $\rho$.  However we still need to
+explicitly compute $\etarho$ since it appears in other equations.
+
 \subsection{Constraint}
@@
 \begin{equation}
-\nabla \cdot (\beta_0 \Ub) = \beta_0 \left ( S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )
-\end{equation}
-where
- $\beta_0$ is a density-like variable that carries background stratification, defined as
+\nabla \cdot (\beta_0 \Ub) = 
+   \beta_0 \left ( S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )
+\label{eq:U divergence}
+\end{equation}
+where $\beta_0$ is a density-like variable that carries background
+ stratification, defined as
 \begin{equation}
@@
 
+In practice, we calculate $w_0$ by integrating
+the one-dimensional divergence constraint:
+\begin{equation}
+\frac{\partial w_0}{\partial r} = \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip 
+\end{equation}
+Then we define
+\begin{equation}
+- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + 
+   w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
+\end{equation}
+once $w_0$ at the old and new times is known, and the advective term is computed explicitly.
+Then we can include this for completeness in the update for $\ut.$
+
 \subsection{Momentum}
@@
 \begin{equation}
-\nabla\cdot(\beta_0 w_0 \eb_r) = \beta_0\left(\Sbar - \frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:w0 divergence}
+\nabla\cdot(\beta_0 w_0 \eb_r) = 
+    \beta_0\left(\Sbar - \frac{1}{\gammabar p_0}
+           \frac{\partial p_0}{\partial t}\right).\label{eq:w0 divergence}
 \end{equation}
@@
 \begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc + \rho\Hext,\label{eq:enthalpy}
+\frac{\partial(\rho h)}{\partial t} = 
+   -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc + \rho\Hext,\label{eq:flow:enthalpy}
+\end{equation}
+
+We will often expand $Dp_0/Dt$ as
+\begin{equation}
+\frac{Dp_0}{Dt} = \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r}
+\end{equation}
+where we defined
+\begin{equation}
+\psi \equiv \frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r}
+\end{equation}
+
+At times, we will define a temperature equation by writing $h = h(T,p,X_k)$
+and differentiating:
+\begin{equation}
+\label{eq:tempenthalpypsi}
+\frac{DT}{Dt} = \frac{1}{\rho c_p} \left\{ \left(1 - \rho h_p\right) \left 
+  [ \psi + (\Ubt \cdotb \er) \frac{\partial p_0}{\partial r} \right ]
+ - \sum_k \rho \xi_k {\omegadot}_k
+ + \rho \Hnuc + \rho \Hext \right \} \enskip  .
 \end{equation}
@@
 \subsection{Base State and Perturbation Form Equations}
+
+The base state evolution equations for density and enthalpy can be
+found by averaging (\ref{eq:flow:continuity}) and (\ref{eq:flow:enthalpy})
+respectively over a layer of constant radius, resulting in
+\begin{eqnarray}
+\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot\left[(\rho h)_0w_0\eb_r\right] +
+  \psi + \overline{\rho \Hnuc} + \overline{\rho \Hext}. \label{eq:flow:enthalpy_base}
+\end{eqnarray}
 

### 2014-08-09 Docs/flowchart/flowchart.tex
more equationing

@@
 
-\subsection{Base State and Perturbation Form Equations}
-
 The base state evolution equations for density and enthalpy can be
-found by averaging (\ref{eq:flow:continuity}) and (\ref{eq:flow:enthalpy})
-respectively over a layer of constant radius, resulting in
+found by averaging Eq~\ref{eq:flow:enthalpy}
+over a layer of constant radius, resulting in
 \begin{eqnarray}
@@
 \end{eqnarray}
+Subtracting it from the full enthalpy equation gives:
+\begin{equation}
+\frac{\partial(\rho h)'}{\partial t} = -\Ub\cdot\nabla(\rho h)' - (\rho h)'\nabla\cdot\Ub - 
+  \nabla\cdot\left[(\rho h)_0\Ubt\right] + \Ubt\cdot\nabla p_0 
+   + ( \rho\Hnuc - \overline{\rho \Hnuc}) + (\rho\Hext - \overline{\rho \Hext}) 
+\end{equation}
 

### 2014-08-11 Docs/flowchart/flowchart.tex
more equationing

@@
 \begin{equation}
-S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext
+S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext + \frac{\sigma}{\rho} \nabla \cdot \kth \nabla T
 \end{equation}
@@
 $c_p \equiv \left.  \partial h / \partial T
-\right|_{p,X_k}$ is the specific heat at constant pressure.
+\right|_{p,X_k}$ is the specific heat at constant pressure.   The last
+term is only present if we are using thermal diffusion (\runparam{use\_thermal\_diffusion} {\tt = T}).  In this term, $\kth$ is the thermal conductivity.
 
@@
 
-In practice, we calculate $w_0$ by integrating
-the one-dimensional divergence constraint:
-\begin{equation}
-\frac{\partial w_0}{\partial r} = \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip 
-\end{equation}
-Then we define
-\begin{equation}
-- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + 
-   w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
-\end{equation}
-once $w_0$ at the old and new times is known, and the advective term is computed explicitly.
-Then we can include this for completeness in the update for $\ut.$
-
 \subsection{Momentum}
@@
 $\overline{(\Ubt\cdot\eb_r)} = 0$ and
-$w_0 = \overline{(\Ub\cdot\eb_r)}$.
-The velocity evolution equations are then \MarginPar{we need the $\beta_0$ here}
+$w_0 = \overline{(\Ub\cdot\eb_r)}$---the motivation for this splitting was given in paper~II.
+The velocity evolution equations are then 
 \begin{eqnarray}
 \frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
-  r} - \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r},\label{eq:w0
-  evolution}\\ \frac{\partial\Ubt}{\partial t} &=& -\Ub\cdot\nabla\Ubt
+  r} - \frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r},\label{eq:w0
+  evolution}\\ 
+%
+\frac{\partial\Ubt}{\partial t} &=& -(\Ubt + w_0\er)\cdot\nabla\Ubt
   - \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
 \frac{1}{\rho}\nabla\pi +
-\frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r -
+\frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r}\eb_r -
 \frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:utildeupd}
@@
 
+\subsection{Base State Expansion}
+
+In practice, we calculate $w_0$ by integrating
+the one-dimensional divergence constraint:
+\begin{equation}
+\frac{\partial w_0}{\partial r} = \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip 
+\end{equation}
+Then we define \MarginPar{this term needs the $\beta_0$'s as shown.  Also, is this really
+the form, or should it be $\overline{(\beta_0/\rho_0) \partial (\pi_0/\beta_0)/\partial r}$?}
+\begin{equation}
+- \frac{\beta_0}{\rho_0} \frac{\partial (\pizero/\beta_0)}{\partial r} = \frac{\partial w_0}{\partial t} + 
+   w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
+\end{equation}
+once $w_0$ at the old and new times is known, and the advective term is computed explicitly.
+Then we can include this for completeness in the update for $\ut.$
+
 \subsection{Energy}
@@
 \end{equation}
+
+In paper~III, we showed that for a plane-parallel atmosphere with constant gravity,
+$\psi = \etarho g$
 

### 2014-08-11 Docs/flowchart/flowchart.tex
more work on getting this up to date

@@
 In practice, we calculate $w_0$ by integrating
-the one-dimensional divergence constraint:
-\begin{equation}
+the one-dimensional divergence constraint.  For a plane-parallel atmosphere, the
+evolution is:
+\begin{equation}
+\label{eq:flow:dw0dr_planar}
 \frac{\partial w_0}{\partial r} = \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip 
 \end{equation}
+
 Then we define \MarginPar{this term needs the $\beta_0$'s as shown.  Also, is this really
@@
 \begin{equation}
-\label{eq:tempenthalpypsi}
+\label{eq:flow:temp}
 \frac{DT}{Dt} = \frac{1}{\rho c_p} \left\{ \left(1 - \rho h_p\right) \left 
@@
 \end{equation}
-%This system is solved using the stiff ordinary differential equation                                     
-%integration package, {\tt VODE} \citep{vode}.                                                            
+Here the temperature equation comes from Eq.~\ref{eq:flow:temp} with $Dp_0/Dt = 0$ for
+the burning part of the evolution.
+
 Full details of the
@@
 (\rho\omegadot_k)^{\outp} &=& \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\dt/2}, \\
-(\rho h)^{\outp} &=& (\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc)^{\outp} + \frac{\dt}{2} (\rho\Hext)^{\in\
-p}.
+(\rho h)^{\outp} &=& (\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc)^{\outp} + \frac{\dt}{2} (\rho\Hext)^{\inp}.
 \end{eqnarray}
@@
 T(\rho^\outp,h^\outp,X_k^\outp)$ for planar geometry or $T^{\outp} =
-T(\rho^\outp,p_0^\inp,X_k^\outp)$ for spherical geometry.
+T(\rho^\outp,p_0^\inp,X_k^\outp)$ for spherical geometry, with this behavior
+controlled by \runparam{use\_tfromp}.
 Note that the density remains unchanged within {\bf React State}, i.e.,
@@
 \begin{description}
-\item[planar:] We discretize equation (\ref{eq:Base State Density}) to
+\item[planar:] We discretize equation (\ref{eq:flow:base_density}) to
   compute the new base state density,
@@
 \end{equation}
-  We compute the time-centered edge states, $\rho_0^{\outp,\nph}$,
-  by discretizing an expanded form of equation (\ref{eq:Base State Density}):
+  We compute the time-centered edge states, ${\rho_0}^{\outp,\nph}_{j\pm\myhalf}$,
+  by discretizing an expanded form of equation (\ref{eq:flow:base_density}):
 \begin{equation}
@@
   term is added to the forcing, due to the spherical discretization of 
-  (\ref{eq:Base State Density}):
+  (\ref{eq:flow:base_density}):
 \begin{equation}
@@
 \begin{description}
-\item[planar:] We discretize equation (\ref{eq:Base State Enthalpy}), neglecting reaction source terms,
-  to compute the new base state enthalpy,
+\item[planar:] We discretize equation (\ref{eq:flow:enthalpy_base}), neglecting reaction 
+source terms, to compute the new base state enthalpy,
 \begin{equation}
@@
   We compute the time-centered edge states, $(\rho h)_0^{\nph}$, by discretizing
-  an expanded form of equation (\ref{eq:Base State Enthalpy}):
+  an expanded form of equation (\ref{eq:flow:enthalpy_base}):
 \begin{equation}
@@
   term is added to the forcing, due to the spherical discretization of 
-  (\ref{eq:Base State Enthalpy}):
+  (\ref{eq:flow:enthalpy_base}):
 \begin{equation}
@@
 
-In Paper III, we showed that $\psi=\etarho g$ for planar geometries, and 
-derived an alternate expression for equation (\ref{eq:w0 divergence}).
-We compute $w_0$ using equation (35) in Paper III using the following 
-discretization:
+In Paper III, we showed that $\psi=\etarho g$ for planar geometries,
+and derived derived Eq.~\ref{eq:flow:dw0dr_planar} as an alternate
+expression for Eq.~\ref{eq:w0 divergence}.  We discretize this as:
 \begin{equation}
@@
 \end{equation}
-We solve this equation for $w_0$ as described in Appendix of the multilevel paper.
+We solve this equation for $w_0$ as described in Appendix~B of the multilevel paper.
 

### 2014-08-12 Docs/flowchart/flowchart.tex
more equation work

@@
 \frac{\partial \rho X_k}{\partial t} + \nabla \cdot (\rho \Ub X_k) = \rho \omegadot_k
+\label{eq:flow:rhoX}
 \end{equation}
@@
 \etarho = \overline{\left(\rho'\Ub\cdot\eb_r\right)}
+\label{eq:flow:etarho}
 \end{equation}
@@
 S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext + \frac{\sigma}{\rho} \nabla \cdot \kth \nabla T
+\label{eq:flow:S}
 \end{equation}
@@
 \frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r}\eb_r -
-\frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:utildeupd}
+\frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:flow:utildeupd}
 \end{eqnarray}
@@
    + ( \rho\Hnuc - \overline{\rho \Hnuc}) + (\rho\Hext - \overline{\rho \Hext}) 
+\label{eq:flow:rhohprime}
 \end{equation}
@@
 
-\begin{description}
+\begin{description}\itemsep 1em
 
@@
   time-centered expansion term in the velocity divergence constraint,
-  as given in paper~III equation (19),
-\begin{equation}
-  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext 
-  + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T \enskip .
-\label{eq:defineS} 
-\end{equation}
-
-  For the first time step ($n=0$), we set
+  as given in Eq.~\ref{eq:flow:S}.  For the first time step ($n=0$),
+  we set
 \begin{equation}
@@
 \begin{equation}
--\frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r}, \label{eq:pizero}
+-\frac{\beta_0}{\rho_0} \frac{\partial (\pi_0/\beta_0)}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r}, \label{eq:pizero}
 \end{equation}
@@
 \begin{equation}
-\left ( \frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} \right )^{n,\star} = -\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} - w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star},
+\left ( \frac{\beta_0}{\rho_0} \frac{\partial (\pi_0/\beta_0)}{\partial r} \right )^{n,\star} = -\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} - w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star},
 \end{equation} 
@@
 
-The local velocity field is described by paper III equation (37),
-\begin{equation}
-\frac{\partial\Ubt}{\partial t} = 
-- \left(\Ubt+w_0\right) \cdot \nabla \Ubt
-- \left(\Ubt \cdot \eb_r\right) \frac{\partial w_0}{\partial r} \eb_r
-- \frac{1}{\rho} \nabla\pi
-+ \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \eb_r
-- \frac{(\rho-\rhozero)}{\rho} \; g \; \eb_r  \label{eq:utildeupd}  \enskip .
-\end{equation}
+The local velocity field is described by Eq.~\ref{eq:flow:utildeupd}.
 From this, we compute time-centered edge 
@@
 \end{eqnarray}
-where $\beta_0$ is computed as described in Appendix C of Paper III.
+where $\beta_0$ is computed as described in Appendix C of Paper III (although
+note that an alternate procedure that uses linear reconstruction of $g$ in
+constructing $\beta_0$ is enabled with \runparam{use\_linear\_grav\_in\_beta}).
 
@@
 \item Update $(\rho X_k)$ using a discretized version of the species 
-continuity equation,
+continuity equation, Eq.~\ref{eq:flow:rhoX}, where we omit the
+reaction terms, which were already accounted for in {\bf React State}:
 \begin{equation}
@@
 \end{equation}
-where we omit the reaction terms, which were already 
-accounted for in {\bf React State}.  The update consists of two steps:
+The update consists of two steps:
 
@@
   Then,\\
-  $(\rho X_k)^{\nph,\star,\pred} = (\rho^{\nph,\star,\pred}X_k^{\nph,\star,\pred})$.
+  $(\rho X_k)^{\nph,\star,\pred} = \rho^{\nph,\star,\pred} \, X_k^{\nph,\star,\pred}$.
 
@@
 
-\item Define a radial edge-centered $\etarho^{\nph,\star}$.
+\item Define a radial edge-centered $\etarho^{\nph,\star}$ (Eq.~\ref{eq:flow:etarho}).
 
@@
 \item For spherical geometry, first construct 
-$\etarho^{{\rm cart},\nph,\star} = [\rho'(\Ub\cdot\eb_r)]^{\nph,\star}$ on Cartesian cell centers using:
+$\etarho^{{\rm cart},\nph,\star} =
+[\rho'(\Ub\cdot\eb_r)]^{\nph,\star}$ on Cartesian cell centers using:
 \begin{eqnarray}
@@
 \item Update the enthalpy using a discretized version of the enthalpy
-evolution equation, again omitting the reaction and heating terms
+evolution equation (Eq.~\ref{eq:flow:enthalpy}), again omitting the reaction and heating terms
 since we already accounted for
@@
 \item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\star,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$.  There are a variety
-   of quantities that we can predict to the interfaces here (controlled
-   by \runparam{enthalpy\_pred\_type}---see Chapter~\ref{ch:pert}).  For
-   the default case, we use the perturbational
-  enthalpy equation, neglecting reactions,
+  for the conservative update of $(\rho h)^{(1)}$.  There are a
+   variety of quantities that we can predict to the interfaces here
+   (controlled by \runparam{enthalpy\_pred\_type}---see
+   Chapter~\ref{ch:pert}).  For the default case, we use the
+   perturbational enthalpy equation, Eq.~\ref{eq:flow:rhohprime}, neglecting reactions,
 \begin{equation}
@@
 \end{equation}
+This behavior is controlled by \runparam{do\_eos\_h\_above\_cutoff}.
 
@@
 \end{enumerate}
+As before, this behavior is controlled by \runparam{use\_tfromp}.
 
@@
 \begin{equation}
-S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + \frac{1}{\rho^{n+1,\star} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star} + \sigma \Hnuc^{(2),\star} + \sigma \Hext^{(2),\star},
+S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + 
+   \frac{1}{\rho^{n+1,\star} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star} + 
+   \sigma \Hnuc^{(2),\star} + \sigma \Hext^{(2),\star},
 \end{equation}
@@
   $\rho^{n+1,\star}, X_k^{n+1,\star},$ and $T^{n+1,\star}$ as inputs to
-  the equation of state.  Then, define
+  the equation of state.  If we are using diffusion then we would include the diffusion
+  term in $S$.  Then, define
 \begin{equation}
@@
 \end{equation}
-  Then, $(\rho X_k)^{\nph,\pred} = (\rho^{\nph,\pred}X_k^{\nph,\pred})$.
+  Then, $(\rho X_k)^{\nph,\pred} = \rho^{\nph,\pred} \, X_k^{\nph,\pred}$.
 
@@
 We update the velocity field $\Ubt^n$ to $\Ubt^{n+1,\dagger}$ by discretizing 
-equation (\ref{eq:utildeupd}) as
+equation (\ref{eq:flow:utildeupd}) as
 \begin{eqnarray}

### 2014-08-12 Docs/flowchart/flowchart.tex
finish a pass through the flowchart chapter

@@
    -\frac{\rho^\prime}{\rho} |g| \er
+\label{eq:flow:newmomentum}
 \end{equation}
@@
   - \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
-\frac{1}{\rho}\nabla\pi +
+\frac{\beta_0}{\rho}\nabla\left(\frac{\pi}{\beta_0} \right) +
 \frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r}\eb_r -
@@
 
-In paper~III, we showed that for a plane-parallel atmosphere with constant gravity,
-$\psi = \etarho g$
+When we are using thermal diffusion, there will be an additional term in
+the enthalpy equation (see \S~\ref{sec:flow:diffusion}).
+
+In paper~III, we showed that for a plane-parallel atmosphere with
+constant gravity, $\psi = \etarho g$
 
@@
 \begin{equation}
-\left ( \frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} \right )^n = 
+\left ( \frac{\beta_0}{\rho_0} \frac{\partial (\pi_0/\beta_0)}{\partial r} \right )^n = 
 -\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
@@
 \end{enumerate}
+Again, the actual inputs depend of \runparam{use\_tfromp}.
 
@@
 $X_k^{n+1}$, and $T^{n+1}$ as inputs to the equation of state.
+If we are doing thermal diffusion (\runparam{use\_thermal\_diffusion}{\tt = T})
+then we also include the diffusive term in $S$.
 Then, compute
@@
 &&- \dt \left(\uadvtwo \cdot \eb_r\right)  \left(\frac{\partial w_0}{\partial r} \right)^\nph \eb_r \nonumber \\
-&& + \dt \left[ - \frac{1}{\rho^\nph} \mathbf{G} \pi^\nmh + \left(\frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\right)^n \eb_r - \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],\nonumber \\
+&& + \dt \left[ - \frac{\beta_0^\nph}{\rho^\nph} \mathbf{G} \left ( \frac{\pi}{\beta_0}\right)^\nmh + \left(\frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r}\right)^n \eb_r - \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],\nonumber \\
 \end{eqnarray}
@@
 \subsection{Volume Discrepancy Changes}
+
+Chapter~\ref{ch:volume} describes the reasoning behind the volume discrepancy
+term---a forcing term added to the constraint equation to bring us back to
+the equation of state.  This addition of this term (enabled with \runparam{dpdt\_factor}{\tt = T}) modifies our
+equation set in the following way:
+
 \begin{itemize}
@@
 
-Assuming that $\delta \Gamma_1 \ll \gammabar$, we then have
+Assuming that $\delta \Gamma_1 \ll \gammabar$, we then have \MarginPar{I think that we include to $\delta \Gamma_1^2$ now}
 \begin{equation}
@@
 \end{equation}
-for plane-parallel geometries (analogous to eq. [\ref{eq:divw0}]), and
+for plane-parallel geometries (analogous to
+Eq.~\ref{eq:flow:dw0dr_planar}), and
 \begin{equation}
@@
 \subsection{Thermal Diffusion Changes}
+
+\label{sec:flow:diffusion}
+
+Thermal diffusion was introduced in the XRB paper~\cite{xrb}.  This
+introduces a new term to $S$ as well as the enthalpy equation.
+Treating the enthalpy equation now requires a parabolic solve.  We describe
+that process here.
+
 Immediately after {\bf Step 4H}, diffuse the enthalpy through 
@@
 discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
-leads to better numerical properties.  We use \runparam{enthalpy\_pred\_type = 1}.
+leads to better numerical properties.  We use \runparam{enthalpy\_pred\_type}{\tt = 1}.
 This is more like paper II.
@@
 
-\subsection{Changes Between the Multilevel Paper and Paper 5}
+\subsection{Changes Between the Multilevel Paper and Paper 5~\cite{wdconvect}}
 \begin{enumerate}
@@
 to prevent large, unphysical velocities from appearing near the edge of the star.
+\end{enumerate}
+
+\subsection{Changes Since the XRB Paper}
+
+\begin{enumerate}
+
+\item We switched to the new form of the momentum equation to
+Eq.~\ref{eq:flow:newmomentum} to conserve the low-Mach number form of
+energy.
+
+\item We changed the form of the volume discrepancy term to get better
+agreement between the two temperatures.
+
 \end{enumerate}

### 2014-08-14 Docs/flowchart/flowchart.tex
fix a typo in the delta Gamma_1 term, and clarify an average

@@
 \Sbar - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} + \overline{
-  \left ( \frac{\delta \Gamma_1}{\gammabar^2 p_0} \ut \cdotb \nablab p_0
+  \left ( \frac{\delta \Gamma_1}{\gammabar^2 p_0} \Ubt \cdotb \nablab p_0
   \right ) } \enskip .
 \end{equation}
+where we see that the $[\delta \Gamma_1/(\Gamma_1^2 p_0)] \partial p_0/\partial t$ terms averages to zero, since the average of $\delta\Gamma_1$ term is zero.
 Subtracting this from equation (\ref{eq:gammafull}), we have

### 2014-08-18 Docs/flowchart/flowchart.tex
some latex tweaks

@@
 Subtracting off the base state, and defining the perturbational
-pressure (sometimes called the dynamic pressure) as $\pi = p - p_0$,
-and perturbational density as $\rho' = \rho - \rho_0$, we have:
+pressure (sometimes called the dynamic pressure) as $\pi \equiv p - p_0$,
+and perturbational density as $\rho' \equiv \rho - \rho_0$, we have:
 \begin{equation}
@@
 The velocity evolution equations are then 
-\begin{eqnarray}
-\frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
+\begin{align}
+\frac{\partial w_0}{\partial t} &= -w_0\frac{\partial w_0}{\partial
   r} - \frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r},\label{eq:w0
@@
 %
-\frac{\partial\Ubt}{\partial t} &=& -(\Ubt + w_0\er)\cdot\nabla\Ubt
+\frac{\partial\Ubt}{\partial t} &= -(\Ubt + w_0\er)\cdot\nabla\Ubt
   - \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
@@
 \frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:flow:utildeupd}
-\end{eqnarray}
+\end{align}
 where $\pi_0$ is the base state component of the perturbational pressure.
@@
 
-Assuming that $\delta \Gamma_1 \ll \gammabar$, we then have \MarginPar{I think that we include to $\delta \Gamma_1^2$ now}
-\begin{equation}
-\nablab \cdotb \Ub + \frac{1}{\gammabar p_0} \Ub \cdotb \nablab p_0 = S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} +
-\frac{\delta \Gamma_1}{\gammabar^2 p_0} \frac{\partial p_0}{\partial t} +
-\frac{\delta \Gamma_1}{\gammabar^2 p_0} \Ub \cdotb \nablab p_0 \enskip , \label{eq:gammafull}
-\end{equation}
-The base state evolution equation is the average of this over a layer
+Assuming that $\delta \Gamma_1 \ll \gammabar$, we then
+have 
+\begin{equation}
+\nablab \cdotb \Ub + \frac{1}{\gammabar p_0} \Ub \cdotb \nablab p_0 
+ \left [ 1 - \frac{\delta\Gamma_1}{\gammabar} + \frac{(\delta\Gamma_1)^2}{\gammabar^2} \right ]
+= S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} 
+ \left [ 1 - \frac{\delta\Gamma_1}{\gammabar} + \frac{(\delta\Gamma_1)^2}{\gammabar^2} \right ] \\
+\end{equation}
+Grouping by order of the correction, we have
+\begin{equation}
+\nablab \cdotb \Ub + \frac{1}{\gammabar p_0} \Ub \cdotb \nablab p_0 
+= S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} +
+\underbrace{\frac{\delta \Gamma_1}{\gammabar^2 p_0} 
+            \left [\frac{\partial p_0}{\partial t} + \Ub \cdotb \nablab p_0\right ]}_{\mbox{first order corrections}} 
+ -
+  \underbrace{\frac{(\delta\Gamma_1)^2}{\gammabar^3 p_0}
+            \left [ \frac{\partial p_0}{\partial t} + \Ub \cdotb \nablab p_0 \right ]}_{\mbox{second order corrections}} \enskip , \label{eq:gammafull}
+\end{equation}
+
+\subsubsection{Keeping to First Order in $\delta\Gamma_1$}
+The base state evolution equation is the average of Eq.~\ref{eq:gammafull} over a layer
 \begin{equation}

### 2014-08-20 Docs/flowchart/flowchart.tex

@@
 Subtracting off the base state, and defining the perturbational
-pressure (sometimes called the dynamic pressure) as $\pi \equiv p - p_0$,
-and perturbational density as $\rho' \equiv \rho - \rho_0$, we have:
+pressure (sometimes called the dynamic pressure) as $\pi = p - p_0$,
+and perturbational density as $\rho' = \rho - \rho_0$, we have:
 \begin{equation}
@@
 The velocity evolution equations are then 
-\begin{align}
-\frac{\partial w_0}{\partial t} &= -w_0\frac{\partial w_0}{\partial
+\begin{eqnarray}
+\frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
   r} - \frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r},\label{eq:w0
@@
 %
-\frac{\partial\Ubt}{\partial t} &= -(\Ubt + w_0\er)\cdot\nabla\Ubt
+\frac{\partial\Ubt}{\partial t} &=& -(\Ubt + w_0\er)\cdot\nabla\Ubt
   - \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
@@
 \frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:flow:utildeupd}
-\end{align}
+\end{eqnarray}
 where $\pi_0$ is the base state component of the perturbational pressure.

### 2014-08-20 Docs/runtime_parameters/runtime_parameters.tex

@@
 \verb=  use_eos_e_instead_of_h  = &   In deriving the temperature from the $h$, first subtract off $p_0/\rho$ to define $e$, and use that as the input to the EOS.  &  .false. \\
+\verb=  use_pprime_in_tfromp  = &    &  .false. \\
+\rowcolor{tableShade}
+\verb=  use_tfromp  = &   When updating temperature, use $T=T(\rho,p_0,X) $ rather than $T=T(\rho,h,X)$.  &  .false. \\
 
@@
 \rowcolor{tableShade}
-\verb=  sdc_couple_mac_velocity  = &   recompute MAC velocity at the beginning of each SDC iter if .false., only recompute MAC velocity at the beginning of the first iter  &  .false. \\
-\verb=  sdc_hold_mac_velocity  = &   if true, never recompute MAC velocity, no matter what the other sdc flags are  &  .false. \\
-\rowcolor{tableShade}
+\verb=  sdc_couple_mac_velocity  = &   recompute MAC velocity at the beginning of each SDC iter  &  .false. \\
 \verb=  sdc_iters  = &   how many SDC iteractions to take (requires the code be compiled with {\tt SDC := t}  &  1 \\
@@
 \verb=  do_burning  = &   turn on (.true.) or off (.false.) burning  &  .true. \\
+\verb=  do_subgrid_burning  = &   break a zone into subzones, call the burner in each subzone and then average the result to the original cell  &  .false. \\
 
@@
 \rowcolor{tableShade}
-\verb=  use_alt_energy_fix  = &   modify the momentum equation to have $(\beta_0/\rho) \nabla (\pi/\beta_0)$ instead of just $(1/\rho) \nabla (\pi)$  &  .false. \\
+\verb=  use_alt_energy_fix  = &   modify the momentum equation to have $(\beta_0/\rho) \nabla (\pi/\beta_0)$ instead of just $(1/\rho) \nabla (\pi)$  &  .true. \\
 \verb=  use_delta_gamma1_term  = &   turns on second order correction to delta gamma1 term  &  .false. \\
@@
 \verb=  use_linear_grav_in_beta  = &   how to represent gravity in the $\beta_0$ integration: .true. = piecewise linear .false. = piecewise constant  &  .false. \\
-\rowcolor{tableShade}
-\verb=  use_tfromp  = &   When updating temperature, use $T=T(\rho,p_0,X) $ rather than $T=T(\rho,h,X)$.  &  .false. \\
 
@@
 \verb=  cg_verbose  = &   Verbosity of bottom solver  &  0 \\
-\verb=  hg_bottom_solver  = &   valid values are $\le$ 0  &  -1 \\
+\verb=  hg_bottom_solver  = &   valid values are $\ge$ 0  &  -1 \\
 \rowcolor{tableShade}
@@
 \rowcolor{tableShade}
-\verb=  mg_bottom_solver  = &   valid values are $\le$ 0  &  -1 \\
+\verb=  mg_bottom_solver  = &   valid values are $\ge$ 0  &  -1 \\
 \verb=  mg_cycle_type  = &   Type of cycle used in the MAC multigrid -- 1 = F-cycle, 2 = W-cycle, 3 = V-cycle  &  3 \\

### 2014-09-29 Docs/networks/networks.tex
renname docs Docs
add a note about rprox in the networks chapter

--- 2012-02-07 docs/networks/networks.tex
@@
-\section{Networks}
+\section{Introduction to \maestro\ Networks}
 
@@
 the equation of state to compute.
+
+\section{Notes of Specific Networks}
 
@@
 {\tt extern/networks/general\_null/GPackage.mak})
+
+\subsection{{\tt rprox}}
+
+This is a network introduced in \cite{xrb2} that has 10 species
+approximates hit CNO burning, triple-$\alpha$, and rp-process 
+breakout up through $^{56}\mathrm{Ni}$.  Updated rates from
+ReacLib~\cite{ReacLib} are used.  The overall ideas in this 
+network are based on Appendix C of \cite{wallacewoosley:1981}.

### 2014-10-08 Docs/architecture/architecture.tex

@@
 \end{itemize}
+
+You generally won't see calls in the MAESTRO source code to these subroutines,
+as there is now a special BoxLib subroutine, {\tt ml\_restrict\_and\_fill},
+that takes an array of multifabs at different level, and in order calls:
+(1) {\tt ml\_cc\_restriction\_c}, (2) {\tt multifab\_fill\_boundary},
+(3) {\tt multifab\_physbc}, and (4) {\tt multifab\_fill\_ghost\_cells}.
+These four subroutines are called in such a way to avoid extra
+ghostcell filling, saving on communication time.  You can specify the
+starting component, starting boundary condition component, 
+the number of components, the number of ghost cells,
+and whether or not you want to use the same boundary condition component
+for all variables.
 

### 2014-10-08 Docs/networks/networks.tex

--- 2012-02-07 docs/networks/networks.tex
@@
 {\tt extern/networks/general\_null/GPackage.mak})
+
+\subsection{{\tt ignition\_chamulak}}
+
+This network was introduced in our paper on convection in white dwarfs
+as a model of Type Ia supernovae~\cite{wdconvect}.  It models 
+carbon burning in a regime appropriate for a simmering white dwarf,
+and captures the effects of a much larger network by setting the ash
+state and energetics to the values suggested in \cite{chamulak:2008}.
+
+\subsection{{\tt ignition\_simple}}
+
+This is the original network used in our white dwarf convection
+studies~\cite{lowMach4}.  It includes a single reaction,
+$^{12}\mathrm{C}({}^{12}\mathrm{C},\gamma){}^{24}\mathrm{Mg}$, using
+the rate from Caughlin and Fowler~\cite{caughlan-fowler:1988}.
+
+\subsection{{\tt rprox}}
+
+This is a network introduced in a paper modeling mixed H/He X-ray
+bursts~\cite{xrb2}.  {\tt rprox} that has 10 species approximates hit
+CNO burning, triple-$\alpha$, and rp-process breakout up through
+$^{56}\mathrm{Ni}$.  Updated rates from ReacLib~\cite{ReacLib} are
+used.  The overall ideas in this network are based on Appendix C of
+\cite{wallacewoosley:1981}.

### 2014-10-15 Docs/runtime_parameters/runtime_parameters.tex
removed do_alltoallv parameter becuase BoxLib already stopped doing that 3 years ago
sync with latest

@@
 \verb=  barrier_timers  = &   Call a parallel\_barrier before each of the simple timers we have coded in advance.f90.  Might cause a very slight overall performance hit.  &  .false. \\
-\verb=  do_alltoallv  = &   use the MPI alltoallv call  &  .false. \\
+\verb=  job_name  = &   job name printed in output  &  "" \\
 \rowcolor{tableShade}
-\verb=  job_name  = &   job name printed in output  &  "" \\
 \verb=  the_knapsack_verbosity  = &   Verbosity of the knapscak processor-to-grid algorithm.  &  .false. \\
-\rowcolor{tableShade}
 \verb=  verbose  = &   General verbosity  &  0 \\
@@
 \verb=  dump_grid_file  = &   dump out a file, named grids.out, containing the number of grids a each level. A new line is added each time regrid is called.  &  .false. \\
+\verb=  ignore_fine_in_layout_mapping  = &   Our traditional way of distributing boxes among MPI ranks has a flaw.
+In some cases, it unnecessarily moves grids among processors in regridding.
+Even if the grids do not changes at all, the boxes may end up on different processors after regridding, and the box distribution becomes more and more fragmented over time (i.e.
+neighboring boxes become far away from each other in MPI ranks).
+Setting {\tt ignore\_fine\_in\_layout\_mapping} to true will avoid the flaw.
+&  .true.
+\\
+\rowcolor{tableShade}
 \verb=  max_grid_size  = &   The largest grid size that will be created using make\_new\_grids.  &  64 \\
+\verb=  max_grid_size_1  = &   The largest grid size that will be created using make\_new\_grids for the coarsest level.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
 \rowcolor{tableShade}
-\verb=  max_grid_size_1  = &   The largest grid size that will be created using make\_new\_grids for the coarsest level.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
 \verb=  max_grid_size_2  = &   The largest grid size that will be created using make\_new\_grids for level 2 (the first refined level).  Defaults to {\tt max\_grid\_size}.  &  -1 \\
+\verb=  max_grid_size_3  = &   The largest grid size that will be created using make\_new\_grids for level 3 (the second refined level) and beyond.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
 \rowcolor{tableShade}
-\verb=  max_grid_size_3  = &   The largest grid size that will be created using make\_new\_grids for level 3 (the second refined level) and beyond.  Defaults to {\tt max\_grid\_size}.  &  -1 \\
 \verb=  max_levs  = &   Total number of levels.  1 = single level.  &  1 \\
+\verb=  min_eff  = &   parameter for cluster algorithm for making new grids in adaptive problems  &  0.9d0 \\
 \rowcolor{tableShade}
-\verb=  min_eff  = &   parameter for cluster algorithm for making new grids in adaptive problems  &  0.9d0 \\
 \verb=  minwidth  = &   The minimum size on a side for a grid created using make\_new\_grids.  &  8 \\
+\verb=  n_cellx  = &   Number of cells for the base level in the x-direction  &  -1 \\
 \rowcolor{tableShade}
-\verb=  n_cellx  = &   Number of cells for the base level in the x-direction  &  -1 \\
 \verb=  n_celly  = &   Number of cells for the base level in the y-direction  &  -1 \\
+\verb=  n_cellz  = &   Number of cells for the base level in the z-direction  &  -1 \\
 \rowcolor{tableShade}
-\verb=  n_cellz  = &   Number of cells for the base level in the z-direction  &  -1 \\
 \verb=  octant  = &   set octant = T if you just want to model an octant of a sphere (note: only takes effect for spherical geometry)  &  .false. \\
+\verb=  prob_hi_x  = &   physical coordinates of hi-x corner of problem domain  &  1.d0 \\
 \rowcolor{tableShade}
-\verb=  prob_hi_x  = &   physical coordinates of hi-x corner of problem domain  &  1.d0 \\
 \verb=  prob_hi_y  = &   physical coordinates of hi-y corner of problem domain  &  1.d0 \\
+\verb=  prob_hi_z  = &   physical coordinates of hi-z corner of problem domain  &  1.d0 \\
 \rowcolor{tableShade}
-\verb=  prob_hi_z  = &   physical coordinates of hi-z corner of problem domain  &  1.d0 \\
 \verb=  prob_lo_x  = &   physical coordinates of lo-x corner of problem domain  &  ZERO \\
+\verb=  prob_lo_y  = &   physical coordinates of lo-y corner of problem domain  &  ZERO \\
 \rowcolor{tableShade}
-\verb=  prob_lo_y  = &   physical coordinates of lo-y corner of problem domain  &  ZERO \\
 \verb=  prob_lo_z  = &   physical coordinates of lo-z corner of problem domain  &  ZERO \\
+\verb=  ref_ratio  = &   Refinement ratio for multilevel problems  &  2 \\
 \rowcolor{tableShade}
-\verb=  ref_ratio  = &   Refinement ratio for multilevel problems  &  2 \\
 \verb=  regrid_int  = &   How often we regrid.  &  -1 \\
+\verb=  spherical_in  = &   Set to 1 if you are doing a spherical problem.  &  0 \\
 \rowcolor{tableShade}
-\verb=  spherical_in  = &   Set to 1 if you are doing a spherical problem.  &  0 \\
 \verb=  test_set  = &   Fixed grid file.  &  '' \\
+\verb=  the_copy_cache_max  = &    Number of boxassoc layouts we keep in memory to avoid having to recompute the boxassoc, which is computationally expensive.  &  50 \\
 \rowcolor{tableShade}
-\verb=  the_copy_cache_max  = &    Number of boxassoc layouts we keep in memory to avoid having to recompute the boxassoc, which is computationally expensive.  &  25 \\
 \verb=  the_layout_verbosity  = &    &  0 \\
-\rowcolor{tableShade}
 \verb=  the_sfc_threshold  = &   When assigning processors for grids, this determines whether we use the sfc algorithm or knapsack algorithm.
@@
 &  5 \\
+\rowcolor{tableShade}
 \verb=  xhi_boundary_type  = &   $+x$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcx\_hi}  &  "" \\
+\verb=  xlo_boundary_type  = &   $-x$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcx\_lo}  &  "" \\
 \rowcolor{tableShade}
-\verb=  xlo_boundary_type  = &   $-x$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcx\_lo}  &  "" \\
 \verb=  yhi_boundary_type  = &   $+y$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcy\_hi}  &  "" \\
+\verb=  ylo_boundary_type  = &   $-y$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcy\_lo}  &  "" \\
 \rowcolor{tableShade}
-\verb=  ylo_boundary_type  = &   $-y$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcy\_lo}  &  "" \\
 \verb=  zhi_boundary_type  = &   $+z$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcz\_hi}  &  "" \\
-\rowcolor{tableShade}
 \verb=  zlo_boundary_type  = &   $-z$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcz\_lo}  &  "" \\

### 2014-10-15 Docs/managing_jobs/managingjobs.tex
add note about fno-tree-pre with gfortran 4.8.x

@@
 
-\item {\em gfortran 4.8.x}: There are no known issues with gfortran 4.8.x
+\item {\em gfortran 4.8.x}: This typically works well, but sometimes we get
+an error allocating memory in {\tt cluster\_f.f90}.  The NERSC team looked
+into this an determined it is a compiler bug (affecting atleast 4.8.2 and 4.8.3):
+\begin{quote}
+The code runs without any problem if it is compiled with {\tt -O2 -ftree-vectorize -fno-range-check} (our default) expcept that {\tt cluster\_f.f90} is compiled with {\tt -O2 -ftree-vectorize -fno-range-check -fno-tree-pre}.
+The ``{\tt fno-tree-pre}'' option turns off ``{\tt ftree-pre}'' that is turned on by ``{\tt O2}''.
+
+GCC manual says,
+
+{\tt -ftree-pre} \\
+Perform partial redundancy elimination (PRE) on trees. This flag is enabled by default at {\tt -O2} and {\tt -O3}.
+\end{quote}
 

### 2014-10-15 Docs/managing_jobs/managingjobs.tex
fix gfortran comments

--- 2014-07-21 Docs/managing_jobs/managingjobs.tex
@@
 
-\item {\em gfortran 4.8.x}: There are no known issues with gfortran 4.8.x
+\item {\em gfortran 4.8.x}: This typically works well, but sometimes we get
+an error allocating memory in {\tt cluster\_f.f90}.  This
+is a compiler bug (affecting atleast 4.8.2 and 4.8.3):
+
+The code runs without any problem if it is compiled with {\tt -O2
+-ftree-vectorize -fno-range-check} (our default) but with {\tt
+cluster\_f.f90} compiled with {\tt -O2 -ftree-vectorize
+-fno-range-check -fno-tree-pre}.  The ``{\tt fno-tree-pre}'' option
+turns off ``{\tt ftree-pre}'' that is turned on by ``{\tt O2}'' \\
+
+GCC manual says,
+
+{\tt -ftree-pre} \\
+Perform partial redundancy elimination (PRE) on trees. This flag is enabled by default at {\tt -O2} and {\tt -O3}.
 

### 2014-10-19 Docs/getting_started/getting_started.tex
add new paper ref, add new test problem description

@@
 
+\item {\tt test\_spherical} \\[-3mm]
+
+This problem sets up an isentropically stratified star and stirs it up
+with a random velocity field.  The low Mach number constraint is
+replaced with the anelastic constraint (through
+the \runparam{beta\_type} runtime parameter).  Analytically, under
+these conditions, the density of the star should not change.  This
+test problem was discussed in Maestro paper IV~\cite{lowMach4}.
+
 \end{itemize}

### 2014-10-19 Docs/managing_jobs/scaling/xrb/xrb-3d-edison-20140425/scaling.py

@@ -73,7 +73,7 @@ def scaling():
     pylab.title("NERSC Edison Scaling for 3-d XRB (384 x 384 x 768 zones)")
 
     pylab.ylim(1.,100.)
-    pylab.savefig("xrb_edison_scaling_by_grid.png")
+    pylab.savefig("xrb_edison_scaling_by_grid.eps")
 
 
 

### 2014-10-23 Docs/managing_jobs/managingjobs.tex
add some Blue Waters info

@@
 Cray 8.2.5 works well on titan/OLCF with MPI/OpenMP.
+
+\subsection{Monitoring Allocations}
+
+The {\tt showusage} and {\tt showusage -f} commands give an
+overview of the usage.
 
@@
 to your {\tt $\mathtt{\sim}$/.visit/hosts/} directory.
+
+\section{Using Blue Waters}
+
+\subsection{Overview}
+
+Each node has 2 compute units, and a compute unit has 16 integer / 8
+floating point cores.  They suggest you run using all the integer
+cores.  Furthermore, each compute unit is separated into 2 NUMA nodes.
+
+A recommended configuration is:
+
+\subsection{BW Compilers}
+
+The Cray compilers are the default on blue waters, and version
+8.3.3 works well with \maestro.
+
+\subsection{Monitoring Allocations}
+
+The {\tt usage} command will list your usage and {\tt usage -P {\em account}} will
+list the usage for all users on an account.

### 2014-10-23 Docs/managing_jobs/managingjobs.tex
Flesh out the notes on Blue Waters

--- 2014-10-15 Docs/managing_jobs/managingjobs.tex
@@
 to your {\tt $\mathtt{\sim}$/.visit/hosts/} directory.
+
+\section{Working at NCSA (Blue Waters)}
+
+\subsection{Overview}
+
+Blue Waters consists of 22,640 Cray XE6 compute nodes and 4,224
+Cray XK7 compute nodes.
+
+Each XE node has two AMD Interlagos model 6276 compute units, each of
+which has 16 integer cores (thus, a single node has a total of 32 integer
+cores).  Two integer cores share a multithreaded, 256-bit wide floating 
+point unit (FPU).  If both integer cores have their own thread, each has access 
+to 128-bit floating point processing, whereas if only one thread is 
+assigned the process can access all 256 bits.  In one major science
+application on Blue Waters it was found that having an OpenMP thread for
+each integer core gave the best performance, but when starting a new
+application it's best to experiment.  One OpenMP thread per FPU may
+be better in some cases.
+
+Each compute unit is divided into two NUMA nodes.  Cores in
+the same NUMA region share a pool of L3 cache.  For the same science
+application as before it was found that the best performance was achieved
+by assigning an MPI task to each NUMA node.  Thus, each physical node
+has four MPI tasks.
+
+The XK nodes consist of one AMD Interlagos model 6276 compute unit
+and an NVIDIA GK110 ``Kepler'' GPU accelerator (Tesla K20X).  The
+GPU is configured with 14 streaming multiprocessor units (SMXs), each
+of which has 192 single-precision or 64 double-precision CUDA cores.  Thus
+there are a total of 2688 SP CUDA cores or 896 DP CUDA cores.
+
+For more details, please see 
+\url{https://bluewaters.ncsa.illinois.edu/user-guide}
+
+\subsection{BW Compilers}
+
+The Cray compilers are the default on blue waters, and version
+8.3.3 works well with \maestro.
+
+\subsection{Monitoring Allocations}
+
+The {\tt usage} command will list the current user's usage and 
+{\tt usage -P {\em project}} will
+list the usage for all users in a project allocation named ``project''.

### 2014-10-25 Docs/getting_started/getting_started.tex
update docs

@@
 
+\item {\tt flame} \\[-3mm]
+
+   A combustion-mode problem where we model a thermonuclear flame in a
+   small domain.  This enforces the low Mach combustion constraint
+   div{U} = S.  Hot ash and cool fuel are put into contact and a flame
+   will ignite and propagate across the grid.  Inflow boundary
+   conditions are used to allow for an inflow velocity to be set to
+   keep the laminar flame stationary.
+
+   In this mode, \maestro\ behaves like the code described
+   in \cite{SNe}, which was used for models of Rayleigh-Taylor
+   unstable flames~\cite{SNld,SNrt,SNrt3d}.
+
 \item {\tt toy\_convect} \\[-3mm]

### 2014-11-12 Docs/introduction/introduction.tex
add the notation needed for the flowchart from the multilevel paper
update ref

@@
   ApJ, 728, 118 (henceforth ``the XRB paper'')
+
 \item {\em The Convective Phase Preceding Type Ia Supernovae,}
@@
   740, 8 (henceforth paper V)
+
 \item {\em High-Resolution Simulations of Convection Preceding Ignition in
@@
   ApJ, 745, 73
+
 \item {\em Low Mach Number Modeling of Convection in Helium Shells on
@@
   R.~J.~Orvedahl 2013, ApJ, 764, 97
+
 \item {\em Low-Mach Number Modeling of Core Convection in Massive Stars,}
@@
   2013, ApJ, 773, 137
+
+\item {\em Multidimensional Modeling of Type I X-ray Bursts. II. Two-Dimensional 
+       Convection in a Mixed H/He Accretor,}
+   C.~M.~Malone, M.~Zingale, A.~Nonaka, A.~S.~Almgren, \& J.~B.~Bell 2014, ApJ, 788, 115
+
+\item {\it Comparisons of Two- and Three-Dimensional Convection in Type I X-ray Bursts} 
+M.~Zingale, C.~M.~Malone, A.~Nonaka, A.~S.~Almgren, \& J.~B.~Bell 
+submitted to ApJL.
+
 \end{itemize}

### 2015-01-05 Docs/architecture/architecture.tex
fix typo
add a note about ghost cells for the base state

@@
 {\tt bx} represents the box in the global index-space over which the
-\fab\ is defined, {\tt pbx} represents the ``physical'' box in the sense
-that it includes {\tt bx} plus ghost cells, and {\tt ibx} is the same as {\tt bx} unless 
-the \fab\ is nodal.  As can be seen in Figure~\ref{fig:dataloc},
-for the same grid nodal data requires one more array element than 
-cell-centered data.  To address this {\tt ibx} is made by growing
-{\tt bx} by one element along all nodal dimensions.
-
-It's important to note that all state data is stored in a four-dimensional array
-{\em regardless of the problem's dimensionality}.  The array is 
-{\tt (nx,ny,nz,nc)} in size, where {\tt nc} is the number of components, for instance
-representing different fluid variables, and {\tt (nx,ny,nz)} are the number of cells
-in each respective spatial dimension.  For 2D problems, {\tt nz=1}.
+\fab\ is defined, {\tt pbx} represents the ``physical'' box in the
+sense that it includes {\tt bx} plus ghost cells, and {\tt ibx} is the
+same as {\tt bx} unless the \fab\ is nodal.  As can be seen in
+Figure~\ref{fig:dataloc}, for the same grid nodal data requires one
+more array element than cell-centered data.  To address this {\tt ibx}
+is made by growing {\tt bx} by one element along all nodal dimensions.
+
+It's important to note that all state data is stored in a
+four-dimensional array {\em regardless of the problem's
+  dimensionality}.  The array is {\tt (nx,ny,nz,nc)} in size, where
+{\tt nc} is the number of components, for instance representing
+different fluid variables, and {\tt (nx,ny,nz)} are the number of
+cells in each respective spatial dimension.  For 2D problems, {\tt
+  nz=1}.
 
@@
 
-The procedure for accessing the data in each grid managed by the \multifab\
-is shown in \S~\ref{sec:example}.  
-Subroutines to add, multiply, or divide two \multifab s exist, as do
-subroutines to copy from one \multifab\ to another---see \code{BoxLib/Src/F\_BaseLib/multifab.f90}
-for the full list of routines that work with \multifab s.
+The procedure for accessing the data in each grid managed by the
+\multifab\ is shown in \S~\ref{sec:example}.  Subroutines to add,
+multiply, or divide two \multifab s exist, as do subroutines to copy
+from one \multifab\ to another---see
+\code{BoxLib/Src/F\_BaseLib/multifab.f90} for the full list of
+routines that work with \multifab s.
 
@@
 by the parameter \runparam{drdxfac}.
+
+Note there are no ghost cells for the base state outside of the
+physical domain.  For plane-parallel, multilevel simulations, there
+are ghostcells at the jumps in refinement---these are filled by the
+\code{fill\_code\_base} routine.  The convention when dealing with the
+base state is that we only access it inside of the valid physical
+domain.  Any multi-dimensional quantity that is derived using the base
+state then has its ghost cells filled by the usually \multifab\ ghost
+cell routines.
 

### 2015-01-26 Docs/getting_started/getting_started.tex

@@
 
-If you don't already have a copy of \maestro, you can obtain one
-via {\tt git}\footnote{{\tt git} is an open-source source code version control
-system.  It is a standard part of any Linux distribution.  You may
-need to use your system's package manager to install {\tt git} if it
-is not already present} as:
-\begin{verbatim}
-git clone http://bender.astro.sunysb.edu/Maestro/staging/MAESTRO.git
-\end{verbatim}
+If you don't already have a copy of \maestro, you can obtain one from
+the project's {\sf github}
+page: \url{https://github.com/BoxLib-Codes/MAESTRO}\, .  There are
+several options: you can fork it directly on {\sf github} (recommended
+if you intend to develop the code) or clone it using {\tt git} from
+the project page.
+
 \maestro\ is under active development, so you will want to keep in 
@@
 
+This problem can also be run adaptively.  The \code{tag\_boxes.f90}
+file in the problem directory tags cells for refinement if the
+perturbational temperature, $T^\prime$, exceeds some threshold.
+
 \item {\tt test\_convect} \\[-3mm]

### 2015-02-11 Docs/architecture/architecture.tex
remove the test suite docs -- this is now in BoxLib User's Guide
update location of the Fortran postprocessing tools

@@
 
-Finally the {\tt AmrPostprocessing} package provides simple
+Finally the {\tt BoxLib/Tools/Postprocessing/F\_Src} package provides simple
 Fortran-based analysis routines (e.g.\ extract a line from a
@@
 shear), this is equivalent to setting $\phi = 0$ (then $\partial
-\phi/d\partial t = 0$, with $t$ meaning `tangental').  This allows the
+\phi/\partial t = 0$, with $t$ meaning `tangental').  This allows the
 velocity to adjust as needed to the domain (see, for example,
@@
 positions.  See \S~\ref{analysis:sec:particles} for details.
+
+\section{Regression Testing}
+
+There is an extensive regression test suite for \boxlib\ that works with
+\maestro.  Full details, and a sample \maestro\ configuration file are
+provided in the \boxlib\ User's Guide and source.

### 2015-02-15 Docs/visualization/visualization.tex
update some paths

@@
 \includegraphics[height=0.4\textheight]{\visfigpath/plt00000_tfromp}
-\caption[Basic plot of {\tt test2} done with {\tt plotsinglevar.py}]
-{\label{fig:python} Plot of {\tt test2} done with the python
+\caption[Basic plot of {\tt reacting\_bubble} done with {\tt plotsinglevar.py}]
+{\label{fig:python} Plot of {\tt reacting\_bubble} done with the python
   script {\tt plotsinglevar.py}.}
@@
 \includegraphics[height=0.4\textheight]{\visfigpath/plt00000_tfromp_enucdot}
-\caption[Plot of two variables from {\tt test2} done with {\tt plotsinglevar.py}]
-{\label{fig:python_two} Plot of {\tt test2} done with the
+\caption[Plot of two variables from {\tt reacting\_bubble} done with {\tt plotsinglevar.py}]
+{\label{fig:python_two} Plot of {\tt reacting\_bubble} done with the
   python script {\tt plotsinglevar.py} showing 2 variables plotted

### 2015-02-15 Docs/architecture/architecture.tex

@@
   \maestro\ `ecosystem'.  Here we see the different packages that
-  contribute to building the {\tt test2} problem in \maestro.  The
+  contribute to building the {\tt reacting\_bubble} problem in \maestro.  The
   red directories are part of most standard \maestro\ build.  The
@@
 is to copy from one of the existing problems.  The file \code{initveldata.f90} controls the velocity field initialization and \code{initscaldata.f90} controls the initialization of the scalars
-($\rho$, $\rho X_k$, $\rho h$).  The {\tt test2} problem is a good
+($\rho$, $\rho X_k$, $\rho h$).  The {\tt reacting\_bubble} problem is a good
 starting point for plane-parallel and {\tt wdconvect} is a good

### 2015-02-17 Util/job_scripts/bw/process.py
Fix some process.py bugs

@@ -464,9 +464,17 @@ class Archiver(object):
       for f in os.listdir('.'):
          #Is this a mae file with the right prefix?
          maeregex = '.*(' + prefix + ')([0-9]{5,6})$'
+         if DEBUG:
+            print 'checking ', f
          if isdir(f):
+            if DEBUG:
+               print 'is dir!'
             if match(maeregex, f) and f not in proc_set:
-               if isfile(join(f,'HEADER')):  #Only add fully written files, i.e. those with HEADER
+               if DEBUG:
+                  print 'regex match!'
+               if isfile(join(f,'Header')):  #Only add fully written files, i.e. those with HEADER
+                  if DEBUG:
+                     print 'found ', f
                   maelist.append(f)
       maelist.sort()
       return maelist[:-new_thresh]  #Will return empty list if len <= 2

### 2015-02-17 Docs/getting_started/getting_started.tex

@@
 
-Here we will run the standard {\tt test2} problem (three reacting
-bubbles in a plane-parallel stratified atmosphere) on a single
-processor.  This test problem was shown in paper~3.
+Here we will run the standard {\tt reacting\_bubble} problem (three
+reacting bubbles in a plane-parallel stratified atmosphere) on a
+single processor\footnote{In earlier versions of \maestro\ this
+problem was called {\tt test2}}.  This test problem was shown in
+paper~3.
 
@@
 
-Change directory to {\tt MAESTRO/Exec/TEST\_PROBLEMS/test2/}.    
+Change directory to {\tt MAESTRO/Exec/TEST\_PROBLEMS/reacting\_bubble/}.    
 We only need to worry about the options at the very top of the
@@
 
-Note, at the end of the build process, a link will be made in the current
-directory to the data
-table needed for the equation of state ({\tt extern/EOS/helmeos/helm\_table.dat}).
+Note, at the end of the build process, a link will be made in the
+current directory to the data table needed for the equation of state
+({\tt Microphysics/EOS/helmeos/helm\_table.dat}).
 
@@
 one of the many runtime parameters \maestro\ knows, and {\em value}
-overrides the default value for that parameter.  For the {\tt test2}
+overrides the default value for that parameter.  For the {\tt reacting\_bubble}
 problem, we will use the inputs file {\tt inputs\_2d}.  An overview of
@@
 in-house and externally developed tools can work
-with \boxlib-formatted plotfiles.  An example plot of the {\tt test2}
+with \boxlib-formatted plotfiles.  An example plot of the {\tt reacting\_bubble}
 problem run above is shown in Figure~\ref{fig:gettingstarted:test2}.
@@
 \includegraphics[width=3in]{\gsfigpath/plt00133_tfromp}
-\caption[Visualization of {\tt test2} output]{\label{fig:gettingstarted:test2} Visualization of the
-final output of the {\tt test2} problem showing the temperature
+\caption[Visualization of {\tt reacting\_bubble} output]{\label{fig:gettingstarted:test2} Visualization of the
+final output of the {\tt reacting\_bubble} problem showing the temperature
 field (as derived from the pressure).  This plot was done with
@@
 
-\item {\tt test2} \\[-3mm]
+\item {\tt reacting\_bubble} \\[-3mm]
 
-{\tt test2} places 3 hots spots in a plane-parallel atmosphere.
+{\tt reacting\_bubble} places 3 hots spots in a plane-parallel atmosphere.
 Burning makes these bubbles buoyant, and then roll up.  This problem was

### 2015-02-17 Docs/makefiles/makefiles.tex

@@
   makefile logic for creating it is in {\tt GMaestro.mak}.  At compile
-  time, the problem, main {\tt MAESTRO/}, and {\tt extern/}
-  directories are searched for {\tt \_parameter} files.  These files
+  time, the problem, main {\tt MAESTRO/}, and any microphysics
+  directories (set from the {\tt EOS\_DIR}, {\tt CONDUCTIVITY\_DIR}, and {\tt NETWORK\_DIR} parameters in the {\tt GNUmakefile}
+  are searched for {\tt \_parameter} files.  These files
   are parsed and the {\tt probin.f90} file is output containing the 
@@
   the {\tt network.f90} is in the {\tt GPackage.mak} in {\tt
-  extern/networks/general\_null}.  The script {\tt write\_network.py}
+  Microphysics/networks/general\_null}.  The script {\tt write\_network.py}
   in that directory does the actual parsing of the species file and
@@
 
-  This variable points to the directory (by default, relative to {\tt Microphysics/EOS/}) of
-  the equation of state used for the build.  Choices that work
-  with \maestro\ are:
+  This variable points to the directory (by default, relative to {\tt
+  Microphysics/EOS/}) of the equation of state used for the build.
+  Choices that work with \maestro\ are:
   \begin{itemize}
-  \item {\tt extern/EOS/helmeos}
-  \item {\tt extern/EOS/gamma\_law\_general}
+  \item {\tt helmeos}
+  \item {\tt gamma\_law\_general}
+  \item {\tt multigamma}
   \end{itemize}
-  To use an EOS contained in a different location, set the variable {\tt EOS\_TOP\_DIR}
-  to point to the directory above the alternate EOS directory.
+  To use an EOS contained in a different location, set the variable {\tt
+  EOS\_TOP\_DIR} to point to the directory above the alternate EOS
+  directory.
 
@@
   \begin{itemize} 
-  \item {\tt extern/conductivity/constant} 
-  \item {\tt extern/conductivity/timmes\_stellar} 
+  \item {\tt constant} 
+  \item {\tt timmes\_stellar} 
   \end{itemize} 
@@
   If diffusion is not being used for the problem, this should be set
-  to {\tt extern/conductivity/constant}.  To use an alternate conductivity
+  to {\tt constant}.  To use an alternate conductivity
   routine, set the variable {\tt CONDUCTIVITY\_TOP\_DIR} to point
@@
   default, relative to {\tt Microphysics/networks/}).  Several options
-  are present in {\tt extern/networks/}.  A network is required even
+  are present in {\tt Microphysics/networks/}.  A network is required even
   if you are not doing reactions, since the network defines the
@@
 
-  A special choice, {\tt extern/networks/general\_null} is a general
+  A special choice, {\tt Microphysics/networks/general\_null} is a general
   network that simply defines the properties of one or more species.

### 2015-02-18 Docs/analysis/analysis.tex
update location of the Fortran postprocessing tools

@@
-\section{The {\tt AmrPostprocessing} Routines}
+\section{The {\tt Postprocessing} Routines}
 \label{sec:analysis}
 
-The {\tt AmrPostprocessing/F\_Src/} directory contains a large
+The {\tt BoxLib/Tools/Postprocessing/F\_Src/} directory contains a large
 number of Fortran-based analysis routines for BoxLib datasets.  Many
@@
 
-As a concrete example, running {\tt test2} with particles enabled
+As a concrete example, running {\tt reacting\_bubble} with particles enabled
 will seed particles in the initial hotspot.  To plot the results,
@@
 This will allow python to see the {\tt parseparticles.py} routine.
-For the {\tt test2} problem, the {\tt plotparticles.py} routine shows
+For the {\tt reacting\_bubble} problem, the {\tt plotparticles.py} routine shows
 how to plot the particle histories and make an animation of the

### 2015-02-18 Docs/managing_jobs/managingjobs.tex

@@
 Additionally, if the {\tt ftime} executable is in your path 
-({\tt ftime.f90} lives in {\tt AmrPostprocessing/F\_src/}), then the
+({\tt ftime.f90} lives in {\tt BoxLib/Tools/Postprocessing/F\_src/}), then the
 script will create a file called {\tt ftime.out} that lists the name of the

### 2015-03-15 Docs/managing_jobs/managingjobs.tex
add a note about Edison problems with Cray 8.3.3 to 8.3.7 compilers...
fix typo
add some info on using yt in batch on titan

@@
 
+\subsection{Batch Submission of \yt\ Visualization Scripts}
+
+You can run \yt\ python scripts in the titan batch queues
+to do your visualization.  You need to install \yt\ and all its dependencies
+manually somewhere on a LUSTRE filesystem---this ensures that the compute
+nodes can see it.  A good choice is the project space, since that has
+a longer purge window.
+
+Once \yt\ is install by the {\tt install\_script.sh} script, it will tell
+you what script to source to define the necessary paths.
+
+The scripts {\tt process-titan.run} and {\tt parallel-yt-new} in
+{\tt MAESTRO/Util/job\_scripts/titan/} will manage the yt jobs
+by calling a python script for each file that matches a pattern.
+
 \subsection{Remote \visit\ Visualization on Lens}
@@
       To get back to the older Cray compilers, you need to do:
+
 \begin{verbatim}
@@
 \end{verbatim}
+
+\item Cray 8.3.3--8.3.7 compile, but issue a runtime error:
+\begin{verbatim}
+ BOXLIB ERROR: BOXARRAY_VERIFY_DIM: ba%dim not equal to some boxes dim:  3
+\end{verbatim}
+Again, the solution is to downgrade to the 8.2.6 version.
+
 \end{itemize}

### 2015-03-15 Docs/managing_jobs/managingjobs.tex
more about yt

@@
 \end{verbatim}
-(Note: if the hsi command prompts you for your password, you will need to
-talk to the OLCF help desk to ask for password-less access to HPSS).
+(Note: if the hsi command prompts you for your password, you will need
+to talk to the OLCF help desk to ask for password-less access to
+HPSS).
 
-The script {\tt process.titan} is called from {\tt titan.run} and
-will run in the background and continually wait until checkpoint or
+The script {\tt process.titan} is called from {\tt titan.run} and will
+run in the background and continually wait until checkpoint or
 plotfiles are created (actually, it always leaves the most recent one
@@
 
-Then the script will use {\tt htar} to archive the plotfiles and checkpoints
-to HPSS.  If the {\tt htar} command was successful, then the plotfiles are
-copied into a {\tt plotfile/} subdirectory.  This is actually important,
-since you don't want to try archiving the data a second time and
-overwriting the stored copy, especially if a purge took place.
+Then the script will use {\tt htar} to archive the plotfiles and
+checkpoints to HPSS.  If the {\tt htar} command was successful, then
+the plotfiles are copied into a {\tt plotfile/} subdirectory.  This is
+actually important, since you don't want to try archiving the data a
+second time and overwriting the stored copy, especially if a purge
+took place.
 
-Additionally, if the {\tt ftime} executable is in your path 
-({\tt ftime.f90} lives in {\tt BoxLib/Tools/Postprocessing/F\_src/}), then the
-script will create a file called {\tt ftime.out} that lists the name of the
-plotfile and the corresponding simulation time.
+Additionally, if the {\tt ftime} executable is in your path ({\tt
+ftime.f90} lives in {\tt BoxLib/Tools/Postprocessing/F\_src/}), then
+the script will create a file called {\tt ftime.out} that lists the
+name of the plotfile and the corresponding simulation time.
 
@@
 
-The {\tt titan.run} submission script has code in it that will look
-at the most recently generated checkpoint files, make sure that they
-were written out correctly (it looks to see if there is a Header file,
+The {\tt titan.run} submission script has code in it that will look at
+the most recently generated checkpoint files, make sure that they were
+written out correctly (it looks to see if there is a Header file,
 since that is the last thing written), and automatically set the {\tt
@@
 before the new submission runs and {\tt QSUB SCRIPT} is the submission
-script (e.g.\ {\tt titan.run}).
-This way you can queue up a bunch of runs and literally leave things
-alone and it will restart from the right place automatically and store
-the data as it is generated.
+script (e.g.\ {\tt titan.run}).  This way you can queue up a bunch of
+runs and literally leave things alone and it will restart from the
+right place automatically and store the data as it is generated.
 
@@
 
-The {\tt chainsub.sh} script can be used to automatically launch a number
-of jobs depending on a single, currently queued (or running) job.
+The {\tt chainsub.sh} script can be used to automatically launch a
+number of jobs depending on a single, currently queued (or running)
+job.
 
@@
 
-You can run \yt\ python scripts in the titan batch queues
-to do your visualization.  You need to install \yt\ and all its dependencies
-manually somewhere on a LUSTRE filesystem---this ensures that the compute
-nodes can see it.  A good choice is the project space, since that has
-a longer purge window.
+You can run \yt\ python scripts in the titan batch queues to do your
+visualization.  You need to install \yt\ and all its dependencies
+manually somewhere on a LUSTRE filesystem---this ensures that the
+compute nodes can see it.  A good choice is the project space, since
+that has a longer purge window.
 
-Once \yt\ is install by the {\tt install\_script.sh} script, it will tell
-you what script to source to define the necessary paths.
+Once \yt\ is install by the {\tt install\_script.sh} script, it will
+tell you what script to source to define the necessary paths.
 
-The scripts {\tt process-titan.run} and {\tt parallel-yt-new} in
-{\tt MAESTRO/Util/job\_scripts/titan/} will manage the yt jobs
-by calling a python script for each file that matches a pattern.
+The scripts {\tt vis-titan.run} and {\tt parallel-yt-new} in {\tt
+MAESTRO/Util/job\_scripts/titan/} will manage the yt jobs by calling a
+python script for each file that matches a pattern.  Make sure you set
+the proper path for the \yt\ {\tt activate} script in {\tt
+parallel-yt-new}---this ensures that the paths to the various libraries
+are set on the compute nodes.
+
+Note that the actual visualization command itself is launched by
+{\tt parallel-yt-new} using the {\tt aprun} command.
 
@@
 
-{\em Note: this information may be out-of-date.}
+{\em Note: this information may be out-of-date.  It is recommended that
+ \yt\ be used instead.}
 

### 2015-03-16 Docs/managing_jobs/managingjobs.tex
update rhea vis

@@
 \subsection{Batch Submission of \yt\ Visualization Scripts}
+
+\subsubsection{Rhea---preferred method}
+
+The best way to do visualization is to use rhea, the OLCF vis machine.
+You need to build \yt\ via the {\tt install\_script.sh} script {\em on
+rhea}.  It also must be on a Lustre filesystem, so it is seen by
+the compute node.  It is best to build it in your {\tt \$PROJWORK} directory,
+since that has a longer time between purges.  Once the installation is
+complete, it will tell you what script to source to define the
+necessary paths.
+
+The scripts in {\tt MAESTRO/Util/job\_scripts/rhea/} will handle the
+visualization.  On rhea, the job script gives you access to the
+compute node, and then you can run serial jobs or a parallel job with
+{\tt mpirun}.  The {\tt process-rhea.run} script will request the
+resources and run the {\tt parallel-yt-rhea} script.  {\tt
+parallel-yt-rhea} will launch the visualization process (defined
+via the variables at the top of the script) on all the plotfiles
+found to match the prefix defined in the script.  Several serial
+jobs are run together, with the script using lock files to keep track
+of how many processors are in use.  When processors free, the next
+file in the list is processed, and so on, until there are no more
+plotfiles left to process.  If a {\tt .png} file matching the
+plotfile prefix is found, then that plotfile is skipped.
+
+\subsubsection{Titan}
 

### 2015-03-16 Docs/managing_jobs/managingjobs.tex
more about why we prefer rhea over titan for yt-ing

@@
 You need to build \yt\ via the {\tt install\_script.sh} script {\em on
-rhea}.  It also must be on a Lustre filesystem, so it is seen by
+rhea}.  It also must be on a {\em Lustre filesystem}, so it is seen by
 the compute node.  It is best to build it in your {\tt \$PROJWORK} directory,
@@
 
+Note: the line in {\tt parallel-yt-rhea} that {\tt source}s the \yt\
+{\tt activate} script may need to be modified to point to the
+correct \yt\ installation path.
+
 \subsubsection{Titan}
 
-You can run \yt\ python scripts in the titan batch queues to do your
+You can also run \yt\ python scripts in the titan batch queues to do your
 visualization.  You need to install \yt\ and all its dependencies
-manually somewhere on a LUSTRE filesystem---this ensures that the
+manually somewhere on a {\em Lustre filesystem}---this ensures that the
 compute nodes can see it.  A good choice is the project space, since
@@
 The scripts {\tt vis-titan.run} and {\tt parallel-yt-new} in {\tt
-MAESTRO/Util/job\_scripts/titan/} will manage the yt jobs by calling a
+MAESTRO/Util/job\_scripts/titan/} will manage the \yt jobs by calling a
 python script for each file that matches a pattern.  Make sure you set
@@
 Note that the actual visualization command itself is launched by
-{\tt parallel-yt-new} using the {\tt aprun} command.
+{\tt parallel-yt-new} using the {\tt aprun} command.  This is the
+only way to get a job to run on the compute nodes on titan.  But
+{\tt aprun} can only launch a single job at a time, so this means
+we cannot easily do (trivally) parallel visualization on a node.  For
+this reason, running on rhea is preferred.
 

### 2015-04-23 Util/VBDF/bdf.f90
clean up bdf and omp reference

@@ -66,30 +66,32 @@ module bdf
      real(dp_t), pointer :: atol(:)         ! absolute tolerances
 
      ! state
-     real(dp_t) :: t                        ! current time
-     real(dp_t) :: t1                       ! final time
-     real(dp_t) :: dt                       ! current time step
-     real(dp_t) :: dt_nwt                   ! dt used when building newton iteration matrix
-     integer  :: k                        ! current order
+     real(dp_t), pointer :: t(:)          ! current time
+     real(dp_t) :: t1                     ! final time
+     real(dp_t), pointer :: dt(:)         ! current time step
+     real(dp_t), pointer :: dt_nwt(:)     ! dt used when building newton iteration matrix
+     integer, pointer :: k(:)             ! current order
      integer  :: n                        ! current step
-     integer  :: j_age                    ! age of jacobian
-     integer  :: p_age                    ! age of newton iteration matrix
-     integer  :: k_age                    ! number of steps taken at current order
-     real(dp_t) :: tq(-1:2)                 ! error coefficients (test quality)
-     real(dp_t) :: tq2save
+     !TODO: As of now, j_age and p_age are de-facto scalars.  
+     !   Should decide if vectorized age boosts performance
+     integer, pointer :: j_age(:)         ! age of jacobian
+     integer, pointer :: p_age(:)         ! age of newton iteration matrix
+     integer,   pointer  :: k_age(:)      ! number of steps taken at current order
+     real(dp_t), pointer :: tq(:,:)    ! error coefficients (test quality)
+     real(dp_t), pointer :: tq2save(:)
      logical  :: refactor
 
      real(dp_t), pointer :: J(:,:,:)        ! jacobian matrix
      real(dp_t), pointer :: P(:,:,:)        ! newton iteration matrix
      real(dp_t), pointer :: z(:,:,:)        ! nordsieck histroy array, indexed as (dof, p, n)
      real(dp_t), pointer :: z0(:,:,:)       ! nordsieck predictor array
-     real(dp_t), pointer :: h(:)            ! time steps, h = [ h_n, h_{n-1}, ..., h_{n-k} ]
[…]

### 2015-04-23 Util/VBDF/dev/omp/test_react/react_state.f90

@@ -807,7 +807,6 @@ contains
     logical, save      :: firstCall = .true.
 
     real (kind = dp_t) :: sumX
-    real(kind=dp_t) :: r1, r2
     real (kind = dp_t), parameter :: x_err = 1.d-10
 
     if (firstCall) then
@@ -816,7 +815,6 @@ contains
     endif
 
     ldt = dt
-    r1 = parallel_wtime()
 
     !$OMP PARALLEL DO PRIVATE(i,j,k,cell_valid,rho,x_in,T_in,x_test,x_out,rhowdot,rhoH,sumX,n) FIRSTPRIVATE(ldt) &
     !$OMP SCHEDULE(DYNAMIC,1)
@@ -896,40 +894,7 @@ contains
        enddo
     enddo
     !$OMP END PARALLEL DO
-    r2 = parallel_wtime() - r1
     
-    !TODO: Delete these
-    print *, 'Burner loop runtime:       ', r2
-    print *, '  total wrap runtime:      ', wrap_total
-    print *, '    total build runtime:   ', bdf_build_total
-    print *, '    total advance runtime: ', bdf_advance_total
-    print *, '      total update runtime: ', bdf_update_total
-    print *, '      total predict runtime: ', bdf_predict_total
-    print *, '      total solve runtime: ', bdf_solve_total
-    print *, '      total check runtime: ', bdf_check_total
-    print *, '      total correct runtime: ', bdf_correct_total
-    print *, '      total adjust runtime: ', bdf_adjust_total
-    wrap_total = 0.0_dp_t
-    bdf_build_total = 0.0_dp_t
-    bdf_advance_total = 0.0_dp_t
-    bdf_update_total =  0.0_dp_t
-    bdf_predict_total = 0.0_dp_t
-    bdf_solve_total = 0.0_dp_t 
-    bdf_check_total = 0.0_dp_t 
-    bdf_correct_total = 0.0_dp_t
-    bdf_adjust_total = 0.0_dp_t 
-
-    print *, 'steps:      ', nst
-    print *, 'F evals:    ', nfe
-    print *, 'J evals:    ', nje
-    print *, 'LU decomps: ', nlu
-    print *, 'Solver its: ', nit
-    nst = 0
-    nfe = 0
-    nje = 0
-    nlu = 0
-    nit = 0
-
   end subroutine burner_loop_3d
 
[…]

### 2015-04-23 Docs/FAQ/faq.tex

@@
 
-\item {\em Does it matter if we use a Fortran 90 compiler instead of a 
-   Fortran 95 compiler?}
+\item {\em What version of the Fortran standard is needed?}
 
-   On some machines it does---Fortran 95 specifies that any local
-   allocated arrays are automatically deallocated when a subroutine
-   ends.  Fortran 90 does not do this.  Most \maestro\ routines rely 
-   on this Fortran 95 feature.
+  Some features of Fortran 2003 are used, in particular, the {\tt
+    ISO_C_BINDING} feature of Fortran 2003 is needed to define a long
+  integer type for some MPI operations in Fortran.  This is supported
+  by most Fortran compilers even if they don't support the entire
+  Fortran 2003 standard.
 
-   On IBM machines, for instance, using {\tt xlf90} instead of {\tt
-     xlf95} for \maestro\ will produce memory leaks.
+  We also rely on the Fortran 95 standard that specifies that any
+  local allocated arrays are automatically deallocated when a
+  subroutine ends.  Fortran 90 does not do this.  Most
+  \maestro\ routines rely on this Fortran 95 feature.
 
@@
  One implementation of this can be found in {\tt
-   AmrPostprocessing/F\_Src/tutorial/fwrite2d.f90}.  This reads in
+   BoxLib/Tools/Postprocessing/F\_Src/tutorial/fwrite2d.f90}.  This reads in
  the plotfile data using the {\tt plotfile\_module} that the {\tt

### 2015-04-29 Docs/FAQ/faq.tex
fix _

@@
   Some features of Fortran 2003 are used, in particular, the {\tt
-    ISO_C_BINDING} feature of Fortran 2003 is needed to define a long
+    ISO\_C\_BINDING} feature of Fortran 2003 is needed to define a long
   integer type for some MPI operations in Fortran.  This is supported

### 2015-07-15 Docs/preface/preface.tex

@@
 \noindent The current version of the \maestro\ User's Guide can be found in 
-the \maestro\ source tree under {\tt MAESTRO/docs/}.
+the \maestro\ source tree under {\tt MAESTRO/Docs/}.
 
-\noindent Support is available via the mailing list: {\tt maestro-help@bender.astro.sunysb.edu}.  To subscribe to this list or read the archives, visit
-\url{http://bender.astro.sunysb.edu/mailman/listinfo/maestro-help}.
+\noindent Support is available via the mailing list: {\tt maestro-help@googlegroups.com}.  To subscribe to this list or read the archives, visit
+\url{https://groups.google.com/forum/#!forum/maestro-help}.

### 2015-07-28 Docs/introduction/introduction.tex
update paper list + add notes on what portions of the algorithm are described in any appendices

@@
   paper I)
+
 \item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
@@
   2006, ApJ, 649, 927 (henceforth paper II)
+
 \item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
@@
   2008, ApJ, 684, 449 (henceforth paper III)
+
 \item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection,}
@@
 \noindent We have several papers that describe applications of the
-method to Type Ia supernovae, X-ray bursts, and stellar evolution:
+method to Type Ia supernovae, X-ray bursts, and stellar evolution.  Some of these
+papers have appendices that describe enhancements to the code---these are noted
+below.
 \begin{itemize}
@@
   ApJ, 728, 118 (henceforth ``the XRB paper'')
+
+  This introduces the thermal diffusion portion of the \maestro\ algorithm.
 
@@
 
-\item {\it Comparisons of Two- and Three-Dimensional Convection in Type I X-ray Bursts} 
-M.~Zingale, C.~M.~Malone, A.~Nonaka, A.~S.~Almgren, \& J.~B.~Bell 
-submitted to ApJL.
+\item {\it Comparisons of Two- and Three-Dimensional Convection in
+  Type I X-ray Bursts} M.~Zingale, C.~M.~Malone, A.~Nonaka,
+  A.~S.~Almgren, \& J.~B.~Bell 2015, ApJ, 807, 60.
+
+  This has an appendix that describes the Godunov state construction in more
+  detail than previous papers.
+
+\item {\it Low Mach Number Modeling of Convection in Helium Shells on
+  Sub-Chandrasekhar White Dwarfs II: Bulk Properties of Simple Models,
+} A.~M.~Jacobs, M.~Zingale, A.~Nonaka, A.~S.~Almgren, \& J.~B.~Bell
+  2015, submitted to ApJ
+
+  This has an appendix that shows some test problems for the new energy 
+  formulation in \maestro.
 

### 2015-07-30 Docs/runtime_parameters/runtime_parameters.tex
updated parameters

--- 2014-08-20 Docs/runtime_parameters/runtime_parameters.tex
@@
 \verb=  do_subgrid_burning  = &   break a zone into subzones, call the burner in each subzone and then average the result to the original cell  &  .false. \\
+\rowcolor{tableShade}
+\verb=  reaction_sum_tol  = &   mass fraction sum tolerance (if they don't sum to 1 within this tolerance, we abort)  &  1.d-10 \\
 
@@
 \rowcolor{tableShade}
-\verb=  the_copy_cache_max  = &    Number of boxassoc layouts we keep in memory to avoid having to recompute the boxassoc, which is computationally expensive.  &  25 \\
+\verb=  the_copy_cache_max  = &    Number of boxassoc layouts we keep in memory to avoid having to recompute the boxassoc, which is computationally expensive.  &  128 \\
 \verb=  the_layout_verbosity  = &    &  0 \\
 \rowcolor{tableShade}
+\verb=  the_ml_layout_strategy  = &   This prarmeter has no impact on single-level runs!
+0: This uses either knapsack or sfc on each level.
+Before boxes are    distributed, MPI ranks are sorted according to the memory usage by    all levels.
+This is our traditional approach.
+1: This uses sfc on each level.
+MPI ranks are sorted according to the    memory usage by lower levels.
+&  0 \\
 \verb=  the_sfc_threshold  = &   When assigning processors for grids, this determines whether we use the sfc algorithm or knapsack algorithm.
@@
 &  5 \\
+\rowcolor{tableShade}
+\verb=  use_tpert_in_tagging  = &   pass $T'$ into the tagging routines as the auxillary multifab instead of the default $\rho H_\mathrm{nuc}$.  &  .false. \\
 \verb=  xhi_boundary_type  = &   $+x$ boundary condition name---if this is set, it overrides the integer value set through {\tt bcx\_hi}  &  "" \\
@@
 \rowcolor{tableShade}
+\verb=  mini_plot_base_name  = &   basename for the mini plotfiles  &  "miniplt" \\
+\verb=  mini_plot_deltat  = &   rather than use a mini plot interval, output a mini plotfile every mini\_plot\_deltat in time  &  -1.d0 \\
+\rowcolor{tableShade}
+\verb=  mini_plot_int  = &   mini plot interval  &  -1 \\
+\verb=  mini_plot_var1  = &   names of specific variables to store in the mini plotfile  &  "" \\
+\rowcolor{tableShade}
+\verb=  mini_plot_var2  = &    &  "" \\
+\verb=  mini_plot_var3  = &    &  "" \\
+\rowcolor{tableShade}
+\verb=  mini_plot_var4  = &    &  "" \\
+\verb=  mini_plot_var5  = &    &  "" \\
+\rowcolor{tableShade}
+\verb=  mini_plot_var6  = &    &  "" \\
+\verb=  mini_plot_var7  = &    &  "" \\
+\rowcolor{tableShade}
+\verb=  mini_plot_var8  = &    &  "" \\
+\verb=  mini_plot_var9  = &    &  "" \\
+\rowcolor{tableShade}
 \verb=  nOutFiles  = &   If lUsingNFiles = .true., use this many processors to write checkpoint and plotfiles.
@@
 \verb=  perturb_model  = &   Turn on a perturbation in the initial data.  Problem specific.  &  .false. \\
+\rowcolor{tableShade}
+\verb=  print_init_hse_diag  = &   print out HSE diagnostics as a function of r for the initial model  &  .false. \\
 

### 2015-07-30 Docs/visualization/visualization.tex
add documentation on the mini plotfiles

@@
 \renewcommand{\arraystretch}{1.0}
+
+\subsection{Mini vs.\ regular plotfiles}
+
+\maestro\ can manage two independent sets of plotfiles.  This allows you to
+output the default plotfiles, which contain a lot of variables, sparsely,
+and output mini-plotfiles much more frequently.  A mini-plotfile is controlled
+by the analogous runtime parameters as the main plotfiles:
+
+\begin{itemize}
+\item {\tt mini\_plot\_int} is the interval in steps between successive plotfiles
+\item {\tt mini\_plot\_deltat} is the interval in time between successive plotfiles
+\item {\tt mini\_plot\_base\_name} is the base name that prefixes the plotfiles.  The
+  default is {\tt miniplt}
+\end{itemize}
+
+To set the fields that are stored in the mini plotfile, a second set
+of runtime parameters is used: {\tt mini\_plot\_var1}, {\tt
+  mini\_plot\_var2}, $\ldots$, {\tt mini\_plot\_var9}.  These can be set to
+any of the following:
+\begin{itemize}
+  \item {\tt "density"}
+  \item {\tt "species"}: this gets all of the mass fractions
+  \item the name of an individual species in the network (like {\tt "helium-4"})
+  \item {\tt "radvel"}: this gets both the radial and circumferential velocity
+  \item {\tt "velocity"}: all three componets
+  \item {\tt "temperature"}: this is either $T(\rho,p_0)$ or $T(\rho,h)$, depending
+    on the value of {\tt use\_tfromp}
+  \item {\tt "enuc"}: the nuclear energy generation rate
+  \item {\tt "mach"}: the Mach number
+\end{itemize}
 

### 2015-07-31 Docs/getting_started/getting_started.tex
more on mini plotfiles

@@
 \begin{itemize}
+\item {\tt double\_bubble} \\[-3mm]
+
+A rising bubble problem where the bubble(s) can have a different gamma
+than the surrounding atmosphere.  This uses the {\tt multigamma} EOS.
+
 \item {\tt incomp\_shear\_jet} \\[-3mm]
@@
 use \maestro\ as an incompressible solver.
-
-\item {\tt rt} \\ [-3mm]
-
-A Rayleigh-Taylor instability problem.  There are two methods that the
-code is run here, in the standard (using {\tt inputs\_2d}), the base state
-has the stratified atmosphere and we introduce a velocity perturbation
-to start the instability.  The alternate method, {\tt inputs\_2d\_SNe}, uses
-the \runparam{do\_smallscale} runtime parameter to eliminate the base state
-and instead use the incompressible constraint to evolve the system.
 
@@
 perturbational temperature, $T^\prime$, exceeds some threshold.
+
+\item {\tt rt} \\ [-3mm]
+
+A Rayleigh-Taylor instability problem.  There are two methods that the
+code is run here, in the standard (using {\tt inputs\_2d}), the base state
+has the stratified atmosphere and we introduce a velocity perturbation
+to start the instability.  The alternate method, {\tt inputs\_2d\_SNe}, uses
+the \runparam{do\_smallscale} runtime parameter to eliminate the base state
+and instead use the incompressible constraint to evolve the system.
 
@@
    unstable flames~\cite{SNld,SNrt,SNrt3d}.
+
+\item {\tt flame\_1d} \\[-3mm]
+
+   A 1-d version of the {\tt flame} problem above.  This uses a special
+   elliptic solver in \boxlib\ that only works for a single grid, so
+   no parallel runs are allowed for this problem.
 
@@
 
+\item \runparam{plot\_base\_name} is the basename to use for the
+  plotfile filename.  The step number will be appended to
+  this name.
+
 \end{itemize}
+
+Note that in addition to the normal plotfiles, there are {\em mini} plotfiles
+that store a small subset of the fields, intended to be output more frequently.
+These are described in \S~\ref{vis:sec:miniplotfile}.
 

### 2015-07-31 Docs/visualization/visualization.tex
more on mini plotfiles

@@
 \subsection{Mini vs.\ regular plotfiles}
+\label{vis:sec:miniplotfile}
 
@@
 \begin{itemize}
-\item {\tt mini\_plot\_int} is the interval in steps between successive plotfiles
-\item {\tt mini\_plot\_deltat} is the interval in time between successive plotfiles
-\item {\tt mini\_plot\_base\_name} is the base name that prefixes the plotfiles.  The
+\item \runparam{mini\_plot\_int} is the interval in steps between successive plotfiles
+\item \runparam{mini\_plot\_deltat} is the interval in time between successive plotfiles
+\item \runparam{mini\_plot\_base\_name} is the base name that prefixes the plotfiles.  The
   default is {\tt miniplt}
@@
 To set the fields that are stored in the mini plotfile, a second set
-of runtime parameters is used: {\tt mini\_plot\_var1}, {\tt
+of runtime parameters is used: \runparam{mini\_plot\_var1}, {\tt
   mini\_plot\_var2}, $\ldots$, {\tt mini\_plot\_var9}.  These can be set to

### 2015-09-11 Docs/managing_jobs/managingjobs.tex

@@
 The default compilers on edison are the Intel compilers, but
-it seems that the Cray compilers are the best.
+PGI and Cray also work well
 
 \begin{itemize}
-\item Cray 8.2.6 works well on edison/NERSC with MPI/OpenMP.
+\item Cray 8.4.0 works well on edison/NERSC with MPI/OpenMP
 
-\item Cray 8.3.0 generates an internal compiler error on edison/NERSC.
-      To get back to the older Cray compilers, you need to do:
-
-\begin{verbatim}
-module swap cce/8.3.0 cce/8.2.6
-module swap cray-libsci cray-libsci/12.2.0
-module swap cray-mpich cray-mpich/6.3.1
-\end{verbatim}
-
-\item Cray 8.3.3--8.3.7 compile, but issue a runtime error:
-\begin{verbatim}
- BOXLIB ERROR: BOXARRAY_VERIFY_DIM: ba%dim not equal to some boxes dim:  3
-\end{verbatim}
-Again, the solution is to downgrade to the 8.2.6 version.
+\item Intel 15.0.1 works well on edison/NERSC with MPI/OpenMP
 
@@
 
-edison is configured with 24 cores per node split between two Intel IvyBridge
-12-core processors.  Each processor connects to 1/2 of the node's memory
-and is called a NUMA node, so there are 2 NUMA nodes per edison node.
-Best performance is seen when
-running with 6 threads and 4 MPI tasks per node, as example:
+edison is configured with 24 cores per node split between two Intel
+IvyBridge 12-core processors.  Each processor connects to 1/2 of the
+node's memory and is called a NUMA node, so there are 2 NUMA nodes per
+edison node.  Best performance is seen when running with 6 or 12 threads.
 \begin{verbatim}
-aprun -n 192 -N 4 -S 2 -d 6 -ss ./main.Linux.Cray.mpi.omp.exe inputs_3d
+aprun -n 192 -N 4 -S 2 -d 6 ./main.Linux.Cray.mpi.omp.exe inputs_3d
 \end{verbatim}
@@
 
-\subsubsection{Older information---still relevant?}
-\noindent (2013-01-15) It's been seen that sometimes an MPICH2 error is 
-generated of the form:
-\begin{verbatim}
-MPIU_nem_gni_get_hugepages(): Unable to mmap 8388608 bytes for file 
-/var/lib/hugetlbfs/global/pagesize-2097152/hugepagefile.MPICH.1.19467.kvs_14619732, 
-err Cannot allocate memory
-\end{verbatim}
-Setting:
-\begin{verbatim}
-export MPICH_GNI_MALLOC_FALLBACK=enabled
-\end{verbatim}
-in your submission script seems to fix this issue.  This will allow
-MPI to continue if there are insufficient huge pages in one NUMA node.
-(Alternately, it was suggested to run without the {\tt -ss} option to
-{\tt aprun}.
+Note that with the Intel compilers, you should never use the {\tt -ss}
+option to {\tt aprun}, and should use {\tt -cc depth} for $< 12$
+threads and {\tt -cc numa\_node} for $12$ threads.
 

### 2015-10-12 Docs/managing_jobs/managingjobs.tex
update compiler info

@@
 -fno-range-check -fno-tree-pre}.  The ``{\tt fno-tree-pre}'' option
-turns off ``{\tt ftree-pre}'' that is turned on by ``{\tt O2}'' \\
+turns off ``{\tt ftree-pre}'' that is turned on by ``{\tt O2}''
 
@@
 
-{\tt -ftree-pre} \\
+\begin{quote}
+{\tt -ftree-pre} \newline
 Perform partial redundancy elimination (PRE) on trees. This flag is enabled by default at {\tt -O2} and {\tt -O3}.
+\end{quote}
+
+\item {\em gfortran 5.1.1}: These compilers have no known issues.
 

### 2016-01-15 Docs/preface/preface.tex

@@
 
-\noindent \maestro\ developers (past and present) include: Ann Almgren, John
-Bell, Candace Gilet, Adam Jacobs, Mike Lijewski, Chris Malone, Andy
-Nonaka, Ryan Orvedahl, Charles Rendleman, \& Michael Zingale
+\noindent \maestro\ developers (past and present) include: 
+\begin{quote}
+Ann Almgren \\
+Vince Beckner \\
+John Bell \\
+Max Duarte \\
+Candace Gilet \\
+Adam Jacobs \\
+Daniel Lecoanet \\
+Mike Lijewski \\
+Chris Malone \\
+Andy Nonaka \\
+Ryan Orvedahl \\
+Charles Rendleman \\
+Weiqun Zhang \\
+Michael Zingale
+\end{quote}
 

### 2016-02-09 Docs/FAQ/faq.tex
add docs about the new backtracing stuff + a new general indexing command

@@
 
+\item {\em How can I get more information about floating point exceptions?}
+
+\boxlib\ can intercept floating point exceptions and provide a helpful
+backtrace that shows you where they were generated.  Compile with {\tt
+  TEST=t} \otherindex{GNUmakefile}{TEST=t} in your {\tt GNUmakefile} (this is also
+activated in debug-mode, with {\tt NDEBUG= }).  The floating point
+exceptions include NaNs, divide by zero, and overflow.  When one of
+these is hit, the code will abort, and the backtrace information will
+be output to a file {\tt Backtrace.N}, where {\tt N} is the processor
+number.  BoxLib will also initialize \multifab s with signaliing NaNs
+to help uncover any floating point issues.
+
+Handling of seg faults and interruption by control-C are always
+enabled, so you can get a backtrace from the code by interrupting its
+execution on the commandline (this can be useful for investigating
+deadlocks).
+
+Behind the scenes, \boxlib\ implements this capability via the
+Linux/Unix {\tt feenableexcept} function (this is in {\tt
+  backtrace\_c.cpp} in \boxlib).
+
+Note that due to issues with header files, this backtrace capability
+is not supported with the PGI compilers.
+
 \end{enumerate}

### 2016-02-13 Util/scripts/get_maestro_date.sh
update to be github and branch aware

@@ -4,9 +4,8 @@
 # date/time.  Then output the necessary environment variables to build
 # with this source.
 
+branch=development
 
-username=mzingale
-server=gamera.lbl.gov:/usr/local/gitroot
 
 #-----------------------------------------------------------------------------
 
@@ -16,52 +15,28 @@ pwd=`pwd`
 
 # MAESTRO
 echo "cloning MAESTRO"
-git clone ${username}@${server}/MAESTRO
+git clone ssh://git@github.com/BoxLib-Codes/MAESTRO
 
 echo " "
-echo "resetting to before ${date}"
+echo "resetting to before ${date} on branch ${branch}"
 cd MAESTRO
-hash=`git rev-list -n 1 --before="$date" master`
-git reset --hard ${hash}
+git checkout ${branch}
+hash=`git rev-list -n 1 --before="$date" ${branch}`
+git checkout ${hash}
 
 cd ..
 
 # BoxLib
 echo " "
 echo "cloning BoxLib"
-git clone ${username}@${server}/BoxLib
+git clone ssh://git@github.com/BoxLib-Codes/BoxLib
 
 echo " "
-echo "resetting to before ${date}"
+echo "resetting to before ${date} on branch ${branch}"
 cd BoxLib
-hash=`git rev-list -n 1 --before="$date" master`
-git reset --hard ${hash}
-
-cd ..
-
-# AstroDev
-echo " "
-echo "cloning AstroDev"
-git clone ${username}@${server}/AstroDev
-
-echo " "
-echo "resetting to before ${date}"
-cd AstroDev
-hash=`git rev-list -n 1 --before="$date" master`
-git reset --hard ${hash}
-
-cd ..
-
-# MAESTRO_Exec
-echo " "
-echo "cloning MAESTRO_Exec"
-git clone ${username}@${server}/MAESTRO_Exec
-
-echo " "
-echo "resetting to before ${date}"
-cd MAESTRO_Exec
-hash=`git rev-list -n 1 --before="$date" master`
-git reset --hard ${hash}
+git checkout ${branch}
+hash=`git rev-list -n 1 --before="$date" ${branch}`
+git checkout ${hash}
 
 cd ..
 
@@ -75,7 +50,6 @@ fi
 cat >> exports.sh << EOF 
 export MAESTRO_HOME="${pwd}/MAESTRO"
 export BOXLIB_HOME="${pwd}/BoxLib"
-export ASTRODEV_DIR="${pwd}/AstroDev"
 EOF
 
 echo " "

### 2016-02-24 Docs/managing_jobs/managingjobs.tex
add some notes about compilers
add note about python_base

@@
 
+\item {\em gfortran 5.3.x}: These compilers have no known issues.
+
 \end{itemize}
+
+\subsection{PGI compilers}
+
+The \boxlib\ floating point exception trapping is disabled with PGI compilers, due to 
+problems with PGI using the system header files.
 
@@
 
+Note: in order to compile, you will need to ensure that both the {\tt
+python} and {\tt python\_base} modules are loaded (via the {\tt
+module} command).
+
 \subsection{General Guidelines}

### 2016-02-24 Docs/managing_jobs/managingjobs.tex
some updates for SLURM edison

@@
 \begin{itemize}
-\item Cray 8.4.0 works well on edison/NERSC with MPI/OpenMP
+\item Intel 15.0.1 works well on edison/NERSC with MPI/OpenMP
 
-\item Intel 15.0.1 works well on edison/NERSC with MPI/OpenMP
+\item Cray 8.4.3 triggers a backtrace error in {\tt make\_pi\_cc.f90}---this is 
+  likely an optimization issue, and may be 2-d specific.
+
+  Previously,  Cray 8.4.0 worked well on edison/NERSC with MPI/OpenMP.
 
@@
 edison node.  Best performance is seen when running with 6 or 12 threads.
+
+Note: edison switched to SLURM as the batch system.  Your job is submitted
+using the {\tt sbatch} command.  Options to {\tt sbatch} are specified at the
+top of your submission script with {\tt \#SBATCH} as the prefix.  These options
+can be found on the {\tt sbatch manpage}.  For instance, 
 \begin{verbatim}
-aprun -n 192 -N 4 -S 2 -d 6 ./main.Linux.Cray.mpi.omp.exe inputs_3d
+#SBATCH -N 2
+#SBATCH -J myjob
+#SBATCH -A repo-name
+#SBATCH -p regular
+#SBATCH -t 12:00:00
 \end{verbatim}
-Here,
-\begin{itemize}
-\item {\tt -S} is the number of MPI tasks per NUMA node (you want to split them
-  across the 2 NUMA nodes per Edison node)
-\item {\tt -n} is the total number of MPI tasks for the entire application
-\item {\tt -N} is the number of MPI tasks PER Edison node
-\item {\tt -d} is the number of OpenMP threads per MPI task (must match 
-{\tt OMP\_NUM\_THREADS})
-\end{itemize}
-The product of {\tt -N} and {\tt -d} should be 24---the total number of cores
-per edison node.
-The script {\tt
-Util/job\_scripts/edison/edison.run} provides an example.
+will request 2 nodes ({\tt -N}), under the account {\tt repo-name} ({\tt -J}),
+in the regular queue, and for a 12-hour window {\tt -t}.
 
-Note that with the Intel compilers, you should never use the {\tt -ss}
+If you are using OpenMP, then your script should set {\tt OMP\_NUM\_THREADS}, e.g.,
+\begin{verbatim}
+export OMP_NUM_THREADS=12
+\end{verbatim}
+
+By default, SLURM will change directory into the submission directory.  The 
+job is launched from your script using {\tt srun}, e.g.:
+\begin{verbatim}
+srun -n 48 ./main.Linux.Cray.mpi.exe inputs_3d
+\end{verbatim}
+to run 48 MPI tasks (across the 2 nodes), or 
+\begin{verbatim}
+export OMP_NUM_THREADS=6
+srun -n 8 -c 6 ./main.Linux.Cray.mpi.omp.exe inputs_3d
+\end{verbatim}
+to use 8 MPI tasks each with 6 threads.
+
+The scripts in {\tt Util/job\_scripts/edison/} provides some examples.
+
+{\em Not sure if this still applies: Note that with the Intel compilers, you should never use the {\tt -ss}
 option to {\tt aprun}, and should use {\tt -cc depth} for $< 12$
-threads and {\tt -cc numa\_node} for $12$ threads.
+threads and {\tt -cc numa\_node} for $12$ threads.}
 

### 2016-02-24 Docs/managing_jobs/managingjobs.tex
more compiler info
more slurm
add note about PGI 16

@@
 
-The \boxlib\ floating point exception trapping is disabled with PGI compilers, due to 
-problems with PGI using the system header files.
+The \boxlib\ floating point exception trapping is disabled with PGI
+compilers earlier than version 16, due to problems with PGI using the
+system header files.  From version 16 onward, things should work.
 
@@
 
+\item Intel 16.x currently has issues with the linear solvers in \boxlib\
+  and should be avoided.
+
 \item Cray 8.4.3 triggers a backtrace error in {\tt make\_pi\_cc.f90}---this is 
@@
 
-\subsection{General Guidelines}
+\subsection{Running Jobs}
 
@@
 threads and {\tt -cc numa\_node} for $12$ threads.}
+
+Jobs are submitted with {\tt sbatch}.  A job can be canceled using {\tt scancel}, and
+the status can be checked using {\tt squeue -u {\em username}}.
 

### 2016-03-02 Docs/managing_jobs/managingjobs.tex
no longer need special tweaks for intel
more NERSC notes

@@
 
-{\em Not sure if this still applies: Note that with the Intel compilers, you should never use the {\tt -ss}
-option to {\tt aprun}, and should use {\tt -cc depth} for $< 12$
-threads and {\tt -cc numa\_node} for $12$ threads.}
+To chain jobs, such that one queues up after the previous job finished,
+use the {\tt chainslurm.sh} script in that same directory.
 
-Jobs are submitted with {\tt sbatch}.  A job can be canceled using {\tt scancel}, and
-the status can be checked using {\tt squeue -u {\em username}}.
+Jobs are submitted with {\tt sbatch}.  A job can be canceled using
+{\tt scancel}, and the status can be checked using {\tt squeue -u {\em
+username}}.
 
@@
 The same set of submission scripts described for titan are available
-for edison at NERSC in {\tt Util/job\_scripts/edison/}.  They are
-mostly identical, with a few minor changes for the local system
-configuration.  The use is the same as described above.
+for edison at NERSC in {\tt Util/job\_scripts/edison/}.  In particular,
+the job submission script will set the restart command line parameters
+to restart from the most recent checkpoint file in the output directory.
+
+Note: NERSC does not allow for the {\tt process} script that archives
+to HPSS to run in the main job submission script.  Instead, a separate
+job needs to be run in the ``{\tt xfer}'' queue.  The script {\tt edison.xfer.slurm} 
+in {\tt Util/job\_scripts/edison/} shows how this works.
+
+Jobs in the {\tt xfer} queue start up quickly.  The best approach is
+to start one as you start your main job (or make it dependent on the
+main job).  The sample {\tt process.xrb} script will wait for output
+and then archive it as it is produced, using the techniques described
+for titan above.
+
+To check the status of a job in the {\tt xfer} queue, use:
+\begin{verbatim}
+squeue -u username -M all
+\end{verbatim}
 

### 2016-06-05 Docs/FAQ/faq.tex
add a note about SYSTEM_BLAS

@@
 
-   Tags allow the editor to follow function/subroutine names an
+   Tags allow the editor to follow function/subroutine names and
    automatically switch you to the source code corresponding to that
@@
 backtrace that shows you where they were generated.  Compile with {\tt
-  TEST=t} \otherindex{GNUmakefile}{TEST=t} in your {\tt GNUmakefile} (this is also
+  TEST=t} \otherindex{{\tt GNUmakefile}}{TEST=t} in your {\tt GNUmakefile} (this is also
 activated in debug-mode, with {\tt NDEBUG= }).  The floating point

### 2016-06-17 Docs/FAQ/faq.tex
add note about GPROF=t

@@
 
-  An alternate way to get single-processor timings, when using 
-  the GCC compilers is to add {\tt -pg} to the compilation flags
-  for both {\tt gfortran} and {\tt gcc}.  Upon completion, a
-  file names {\tt gmon.out} will be produced.  This can be 
-  processed by {\tt gprof} running
+  An alternate way to get single-processor timings, when using the GCC
+  compilers is to add {\tt -pg} to the compilation flags for both {\tt
+    gfortran} and {\tt gcc} (this can be accomplished by setting {\tt
+    GPROF := t} in your {\tt GNUmakefile}.  Upon completion, a file
+  names {\tt gmon.out} will be produced.  This can be processed by
+  {\tt gprof} running
 

### 2016-06-27 Docs/makefiles/makefiles.tex

@@
 
-The microphysics, as described above is also included.  For the networks,
-we include a file called {\tt NETWORK\_REQUIRES} into {\tt GMaestro.mak}
-that tells us whether to also include {\tt Util/VODE} (if {\tt NEED\_VODE := t})
-or {\tt Util/BLAS} (if {\tt NEED\_BLAS := t}).
+The microphysics, as described above is also included.  For the
+networks, we include a file called {\tt NETWORK\_REQUIRES} into {\tt
+  GMaestro.mak} that tells us whether to also include {\tt Util/VODE}
+(if {\tt NEED\_VODE := t}).  It is assumed in this case that we need
+BLAS and LINPACK, so these are compiled in from {\tt Util/BLAS}  and
+{\tt Util/LINPACK}.
+
+You can instead link in a system-wide optimized BLAS library by setting
+{\tt SYSTEM\_BLAS := t}\otherindex{{\tt GNUmakefile}}{BLAS} in the {\tt GNUmakefile}.  This adds {\tt -lblas}
+to the link line, and assumes that the library is in your path.  Note
+that for some systems, you should have the static BLAS libraries 
+available.
 

### 2016-06-28 Docs/makefiles/makefiles.tex
a bit more docs on TEST and FSANITIZER

@@
 
+\section{Special Debugging Executables}
+
+\boxlib\ has several options that produce executables that can help
+track down memory issues, uninitialized variables, NaNs, etc.
+
+\begin{itemize}
+
+\item {\tt NDEBUG}
+
+  \otherindex{{\tt GNUmakefile}}{NDEBUG} To generate an executable
+  with debugging information included in the executable (e.g., to be
+  interpreted by the debugger, {\tt gdb}), compile with {\tt NDEBUG
+    :=\ }.  This will usually add {\tt -g} to the compile line and
+  also lower the optimization.  For {\tt gfortran} it will add several
+  options to catch uninitialize variables, bounds errors, etc.
+
+\item {\tt TEST}
+
+  Setting {\tt TEST := t}\otherindex{{\tt GNUmakefile}}{TEST=t} will
+  enable routines in \boxlib\ that intercept floating point
+  exceptions, including NaNs, divide by zero, and overflow.  When one
+  of these is hit, the code will abort, and the backtrace information
+  will be output to a file {\tt Backtrace.N}, where {\tt N} is the
+  processor number.  BoxLib will also initialize \multifab s with
+  signaliing NaNs to help uncover any floating point issues.
+
+  This is also useful to diagnose deadlocks in parallel regions.
+  If the code is hanging, doing ``control-C'' will be intercepted
+  and the code will generate a backtrace which will identify
+  where in the code there was a deadlock.
+
+  Behind the scenes, \boxlib\ implements this capability via the
+  Linux/Unix {\tt feenableexcept} function (this is in {\tt
+    backtrace\_c.cpp} in \boxlib).
+
+\item {\tt FSANITIZER}
+
+  For {\tt gfortran}, {\tt gcc}, {\tt g++}, setting {\tt FSANITIZER :=
+    t}\otherindex{{\tt GNUmakefile}}{FSANITIZER} will enable the
+  address sanitizer support built into GCC.  This is enabled through
+  integration with \url{https://github.com/google/sanitizers} in GCC.
+
+\end{itemize}
+
 \section{Extending the Build System}
@@
 
-Properties for different compilers are already defined in 
-{\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.
-Each compiler is given its own file.  The appropriate file is included
-into {\tt GMakedefs.mak} by looking at the {\tt COMP} variable and the
-operating system.  These compiler files define the compiler flags for
-both optimized and debug compiling.  Additionally, the variable {\tt
-FCOMP\_VERSION} should be defined there, based on the output from the
-compiler, to provide the compiler version for output into the {\tt
-job\_info} file at runtime.
+Properties for different compilers are already defined in {\tt
+  \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.  Each compiler is given its
+own file.  The appropriate file is included into {\tt GMakedefs.mak}
+by looking at the {\tt COMP} variable and the operating system.  These
+compiler files define the compiler flags for both optimized and debug
+compiling.  Additionally, the variable {\tt FCOMP\_VERSION} should be
+defined there, based on the output from the compiler, to provide the
+compiler version for output into the {\tt job\_info} file at runtime.
 

### 2016-06-28 Docs/getting_started/getting_started.tex
update location of boxlib
update getting started with info of Microphysics

@@
 
-If you don't already have a copy of \maestro, you can obtain one from
-the project's {\sf github}
-page: \url{https://github.com/BoxLib-Codes/MAESTRO}\, .  There are
-several options: you can fork it directly on {\sf github} (recommended
-if you intend to develop the code) or clone it using {\tt git} from
-the project page.
+  If you don't already have a copy of \maestro, you can obtain one
+  from the project's {\sf github} page:
+  \url{https://github.com/BoxLib-Codes/MAESTRO}\, .  There are several
+  options: you can fork it directly on {\sf github} (recommended if
+  you intend to develop the code) or clone it using {\tt git} from the
+  project page.
 
   \maestro\ is under active development, so you will want to keep in
-sync with the changes by periodically pulling from the repository.  Simply
-type
+  sync with the changes by periodically pulling from the repository.
+  Simply type
   \begin{verbatim}
@@
 
+\item {\em Get a copy of \microphysics}.
+
+  \maestro\ and its compressible counterpart \castro\ share a
+  common-set of microphysics solvers (nuclear reaction networks and
+  equations of state).  These are kept in a separate repo.
+  \microphysics\ is also available on github and can be obtained 
+  via:
+  \begin{verbatim}
+git clone https://github.com/BoxLib-Codes/Microphysics.git
+  \end{verbatim}
+
+  You will periodic want to update \microphysics\ by doing
+  \begin{verbatim}
+git pull
+  \end{verbatim}
+  in the {\tt Microphysics/} directory
+.
 \item {\em Get a copy of \boxlib}.
@@
   parallelization.  We also rely on the build system in \boxlib\ to
-build a \maestro\ executable.  \boxlib\ can be obtained via {\tt
-git} as:
+  build a \maestro\ executable.  \boxlib\ is also available on github
+  and can be obtained via:
   \begin{verbatim}
-git clone https://ccse.lbl.gov/pub/Downloads/BoxLib.git
+git clone https://github.com/BoxLib-Codes/BoxLib.git
   \end{verbatim}
@@
 
-\maestro\ needs to know where to find \boxlib\, by specifying the 
-{\tt BOXLIB\_HOME} environment variable.
+  \maestro\ needs to know where to find \boxlib, by specifying the
+  {\tt BOXLIB\_HOME} environment variable, and where
+  to find \microphysics, bt specifying the {\tt
+  MICROPHYSICS\_HOME} environment variable.
 
@@
 export BOXLIB_HOME="/path/to/BoxLib/"
+export MICROPHYSICS_HOME="/path/to/Microphysics/"
   \end{verbatim}
@@
 setenv BOXLIB_HOME /path/to/BoxLib/
+setenv MICROPHYSICS_HOME /path/to/Microphysics/
   \end{verbatim}
@@
 
-Note: you must specify the full path to the BoxLib directory.  Do not
-use ``$\sim$'' to refer to your home directory---the scripts used by
-the build system will not be able to process this.
+  Note: you must specify the full path to the {\tt BoxLib/} and {\tt
+    Microphysics/} directory.  Do not use ``$\sim$'' to refer to your
+  home directory---the scripts used by the build system will not be
+  able to process this.
 
@@
 
-In \maestro, each problem lives under one of three sub-directories of
-{\tt MAESTRO/Exec}: {\tt SCIENCE/}, {\tt TEST\_PROBLEMS/}, or {\tt
-UNIT\_TESTS/}.  This problem sub-directory will contain any
-problem-specific files as well as the {\tt GNUmakefile} that specifies
-how to build the executable.  Note: we rely on features of GNU {\tt
-make}.  Full details of the {\tt GNUmakefile} can be found
-in \S~\ref{sec:adding_problems}.  Here we will configure for a serial
-build.
+  In \maestro, each problem lives under one of three sub-directories
+  of {\tt MAESTRO/Exec}: {\tt SCIENCE/}, {\tt TEST\_PROBLEMS/}, or
+  {\tt UNIT\_TESTS/}.  This problem sub-directory will contain any
+  problem-specific files as well as the {\tt GNUmakefile} that
+  specifies how to build the executable.  Note: we rely on features of
+  GNU {\tt make}.  Full details of the {\tt GNUmakefile} can be found
+  in \S~\ref{sec:adding_problems}.  Here we will configure for a
+  serial build.
 
-Change directory to {\tt MAESTRO/Exec/TEST\_PROBLEMS/reacting\_bubble/}.    
-We only need to worry about the options at the very top of the
-{\tt GNUmakefile} for now.  These should be set as follows:
+  Change directory to {\tt
+    MAESTRO/Exec/TEST\_PROBLEMS/reacting\_bubble/}.  We only need to
+  worry about the options at the very top of the {\tt GNUmakefile} for
+  now.  These should be set as follows:
   \begin{itemize}
@@
 
-This option determines whether we compile with support for debugging
-(usually also enabling some runtime checks).  Setting {\tt NDEBUG := t}
-turns off debugging.
+    This option determines whether we compile with support for
+    debugging (usually also enabling some runtime checks).  Setting
+    {\tt NDEBUG := t} turns off debugging.
 
@@
 
-This determines whether we are doing a parallel build, using the Message
-Passing Interface (MPI) library.  For now, we leave this option empty,
-disabling MPI.  This will build \maestro\ in serial mode, so no MPI
-library needs to be present on the host system.
+    This determines whether we are doing a parallel build, using the
+    Message Passing Interface (MPI) library.  For now, we leave this
+    option empty, disabling MPI.  This will build \maestro\ in serial
+    mode, so no MPI library needs to be present on the host system.
 
@@
 
-This determines whether we are using OpenMP to do parallelism within a
-shared memory node.  OpenMP is used together with MPI, with MPI
-distributing the grids across the processors and within a
-shared-memory node, OpenMP allows many cores to operate on the same
-grid.  For now, we leave this option empty, disabling OpenMP.
+    This determines whether we are using OpenMP to do parallelism
+    within a shared memory node.  OpenMP is used together with MPI,
+    with MPI distributing the grids across the processors and within a
+    shared-memory node, OpenMP allows many cores to operate on the
+    same grid.  For now, we leave this option empty, disabling OpenMP.
 
@@
     This option determines whether we want to use the alternate SDC
-(Spectral Deferred Corrections) source for \maestro\ (see
-Chapter \ref{ch:sdc}).  This is experimental, and this option not present
-in all problems.  We leave this blank so we compile the
-default \maestro\ source.
+    (Spectral Deferred Corrections) source for \maestro\ (see Chapter
+    \ref{ch:sdc}).  This is experimental, and this option not present
+    in all problems.  We leave this blank so we compile the default
+    \maestro\ source.
 
@@
     This option specifies the Fortran compiler.  We will use {\tt
-gfortran}, which is the preferred compiler for \maestro.  Specifying
-this compiler will automatically pull in the compiler settings as
-specified in {\tt BOXLIB\_HOME/Tools/F\_mk/}.  (Alternate compiler
-choices include {\tt Intel}, {\tt PGI}, {\tt PathScale}, {\tt Cray}.)
+      gfortran}, which is the preferred compiler for \maestro.
+    Specifying this compiler will automatically pull in the compiler
+    settings as specified in {\tt BOXLIB\_HOME/Tools/F\_mk/}.
+    (Alternate compiler choices include {\tt Intel}, {\tt PGI}, {\tt
+      PathScale}, {\tt Cray}.)
 
@@
   finished, the executable will have a name like {\tt
-main.Linux.gfortran.exe}, where the specific parts of the name depend
-on the compilers and OS used.
+    main.Linux.gfortran.exe}, where the specific parts of the name
+  depend on the compilers and OS used.
 
@@
   Each problem requires an input file.  The inputs file (a Fortran
-namelist) consists of
-lines of the form {\em parameter = value}, where {\em parameter} is
-one of the many runtime parameters \maestro\ knows, and {\em value}
-overrides the default value for that parameter.  For the {\tt reacting\_bubble}
-problem, we will use the inputs file {\tt inputs\_2d}.  An overview of
-some of the more common runtime parameters is given
-in \S~\ref{sec:gettingstarted:runtime}, and a full list of
-all \maestro\ runtime parameters and their default values is given in
+  namelist) consists of lines of the form {\em parameter = value},
+  where {\em parameter} is one of the many runtime parameters
+  \maestro\ knows, and {\em value} overrides the default value for
+  that parameter.  For the {\tt reacting\_bubble} problem, we will use
+  the inputs file {\tt inputs\_2d}.  An overview of some of the more
+  common runtime parameters is given in
+  \S~\ref{sec:gettingstarted:runtime}, and a full list of all
+  \maestro\ runtime parameters and their default values is given in
   Chapter~\ref{ch:runtimeparameters}.
@@
   \end{verbatim}
-We can also override the default value of any runtime parameter by specifying
-it on the commandline as
+  We can also override the default value of any runtime parameter by
+  specifying it on the commandline as
   \begin{verbatim}
@@
   As the code runs, a lot of information will pass through the screen.
-For each timestep, each of the steps 1 through 12 shown in
-the \maestro\ flowchart (Chapter~\ref{ch:flowchart}) will be shown
-along with diagnostic information about the solution.  Upon completion
+  For each timestep, each of the steps 1 through 12 shown in the
+  \maestro\ flowchart (Chapter~\ref{ch:flowchart}) will be shown along
+  with diagnostic information about the solution.  Upon completion
   some memory usage information is printed.
@@
   As the code runs, it will output both plotfiles and checkpoints as
-well as one or more text diagnostic files ({\tt maestro\_diag.out} by
-default) with integral or extrema information (like maximum Mach
+  well as one or more text diagnostic files ({\tt maestro\_diag.out}
+  by default) with integral or extrema information (like maximum Mach
   number) from each timestep.
@@
   outputted.  Similarly, the checkpoints are named {\tt chk}{\em
-nnnnn}.  \boxlib\ plotfiles and checkpoints are actually directories,
-with the data stored in sub-directories grouped by refinement level.
-Details of the simulation (build information, number of processors
-used, output date, output directory, runtime parameter values, ...)
-are stored in the plaintext {\tt job\_info} file in each plotfile directory.
+    nnnnn}.  \boxlib\ plotfiles and checkpoints are actually
+  directories, with the data stored in sub-directories grouped by
+  refinement level.  Details of the simulation (build information,
+  number of processors used, output date, output directory, runtime
+  parameter values, ...)  are stored in the plaintext {\tt job\_info}
+  file in each plotfile directory.
 
-{\bf Note: unless otherwise specified all quantities in \maestro\ are
-assumed to be in CGS units.}
+  {\bf Note: unless otherwise specified all quantities in
+    \maestro\ are assumed to be in CGS units.}
 

### 2016-07-06 Docs/FAQ/faq.tex

@@
 \boxlib\ can intercept floating point exceptions and provide a helpful
-backtrace that shows you where they were generated.  Compile with {\tt
-  TEST=t} \otherindex{{\tt GNUmakefile}}{TEST=t} in your {\tt GNUmakefile} (this is also
-activated in debug-mode, with {\tt NDEBUG= }).  The floating point
-exceptions include NaNs, divide by zero, and overflow.  When one of
-these is hit, the code will abort, and the backtrace information will
-be output to a file {\tt Backtrace.N}, where {\tt N} is the processor
-number.  BoxLib will also initialize \multifab s with signaliing NaNs
-to help uncover any floating point issues.
-
-Handling of seg faults and interruption by control-C are always
-enabled, so you can get a backtrace from the code by interrupting its
-execution on the commandline (this can be useful for investigating
-deadlocks).
-
-Behind the scenes, \boxlib\ implements this capability via the
-Linux/Unix {\tt feenableexcept} function (this is in {\tt
-  backtrace\_c.cpp} in \boxlib).
-
-Note that due to issues with header files, this backtrace capability
-is not supported with the PGI compilers.
+backtrace that shows you where they were generated.  See \S
+\ref{ch:makefiles:special}.
 

### 2016-07-06 Docs/getting_started/getting_started.tex

@@
 
+\section{Development Model}
+
+When you close \maestro\ from github, you will be on the {\tt master}
+branch of the repo.  New changes to \maestro\ are first introduced
+into the {\tt development} branch in the \maestro\ git repository.
+Nightly regression tests are run on {\tt development} to ensure that
+our answers don't change.  Around the first work day of each month, we
+merge from {\tt development} $\rightarrow$ {\tt master} (assuming
+tests pass) and tag the state of the code with a date-based tag {\tt
+  YY-MM}.  We do this on all the other repos in the \boxlib-ecosystem,
+including {\tt BoxLib/}, {\tt Microphysics/}, and {\tt Castro/}.
+
+If you want to contribute to \maestro's development, issue a pull-request
+through github onto the {\tt development} branch.
+
 \section{Parallel Jobs}

### 2016-07-06 Docs/makefiles/makefiles.tex
add runtime parameters for boxlib backtracing
now boxlib_fpe_invalid, boxlib_fpe_zero, and boxlib_fpe_overflow
turn on/off FPE handling through the boxlib backtrace module

@@
 
-\section{Special Debugging Executables}
+\section{Special Debugging Modes}
+
+\label{ch:makefiles:special}
 
@@
 
-  Setting {\tt TEST := t}\otherindex{{\tt GNUmakefile}}{TEST=t} will
-  enable routines in \boxlib\ that intercept floating point
-  exceptions, including NaNs, divide by zero, and overflow.  When one
-  of these is hit, the code will abort, and the backtrace information
-  will be output to a file {\tt Backtrace.N}, where {\tt N} is the
+  Setting {\tt TEST := t}\otherindex{{\tt GNUmakefile}}{TEST} will
+  enable routines in \boxlib\ initialize multifabs and arrays 
+  allowed via {\tt bl\_allocate} to signalliing NaNs.  This behavior
+  is the same as {\tt NDEBUG :=}, but {\tt TEST := t} uses the 
+  same compiler optimizations as a normal build.
+
+  This can be useful with compiler flags that trap floating point
+  exceptions (FPEs), but checks on floating point exceptions can also
+  be enabled through runtime parameters passed to \boxlib's
+  backtrace functionlity:
+  \begin{itemize}
+  \item \runparam{boxlib\_fpe\_invalid}: enabling FPE trapping for
+    invalid operations (e.g. {\tt 0 * inf}, {\tt sqrt(-1)})
+
+  \item \runparam{boxlib\_fpe\_zero}: enable FPE trapping for
+    divide-by-zero
+
+  \item \runparam{boxlib\_fpe\_overflow}: enable FPE trapping for
+    overflow
+  \end{itemize}
+
+\item backtracing
+
+  When exception trapping is enabled (either via \boxlib\ or the
+  compiler), the code will abort, and the backtrace information will
+  be output to a file {\tt Backtrace.N}, where {\tt N} is the
   processor number.  BoxLib will also initialize \multifab s with
@@
   integration with \url{https://github.com/google/sanitizers} in GCC.
+
+  Note: you will need to have the libraries {\tt libasan} and {\tt
+    libubsan} installed on your machine to use this functionality.
 

### 2016-07-06 Docs/runtime_parameters/runtime_parameters.tex

@@
 \label{ch:parameters}
-
-%%%%%%%%%%%%%%%%
-% symbol table
-%%%%%%%%%%%%%%%%
 
@@
 \rowcolor{tableShade}
-\verb=  conductivity_constant  = &   the constant conductivity (only applies if the conductivity/constant module is used)  &  1.0d0 \\
 \verb=  limit_conductivity  = &   apply the conductivity limiting---if T, then we set the thermal coefficients all to 0 for $\rho <$ {\tt buoyancy\_cutoff\_factor} * {\tt base\_cutoff\_density}  &  .false.
 \\
+\verb=  temp_diffusion_formulation  = &   How to compute the explicit thermal diffusion term.  \newline 1 = in terms of $T$; \newline 2 = in terms of $\rho,p_0,X$.  &  2 \\
 \rowcolor{tableShade}
-\verb=  temp_diffusion_formulation  = &   How to compute the explicit thermal diffusion term.  \newline 1 = in terms of $T$; \newline 2 = in terms of $\rho,p_0,X$.  &  2 \\
 \verb=  thermal_diffusion_type  = &   In the thermal diffusion solver, 1 = Crank-Nicholson; 2 = Backward Euler.  &  1 \\
-\rowcolor{tableShade}
 \verb=  use_thermal_diffusion  = &   Use thermal diffusion.  &  .false. \\
@@
 
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[gamma law general  EOS parameters.]{gamma law general  EOS parameters.} \label{table: gamma law general  EOS parameters. runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  eos_assume_neutral  = &   do we assume that all the species are neutral atoms? otherwise, we assume complete ionization  &  .true. \\
+\verb=  eos_gamma  = &   the constant ratio of specific heats ($\gamma$)  &  5.d0/3.d0 \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[generic network parameters parameters.]{generic network parameters parameters.} \label{table: generic network parameters parameters. runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  small_x  = &   cutoff for species mass fractions  &  0.0 \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[constant conductivity parameters.]{constant conductivity parameters.} \label{table: constant conductivity parameters. runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} & 
+       \multicolumn{1}{ c|}{\textbf{description}} & 
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline 
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline 
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb=  conductivity_constant  = &   the constant conductivity (only applies if the conductivity/constant module is used)  &  1.0d0 \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
 \end{landscape}

### 2016-07-10 Docs/managing_jobs/managingjobs.tex
some notes on yt at NERSC

@@
 The preferred compilers on Titan are the Cray compilers.  
-Cray 8.2.5 works well on titan/OLCF with MPI/OpenMP.
+Cray 8.4.0 works well on titan/OLCF with MPI/OpenMP.
 
@@
 To chain jobs, such that one queues up after the previous job finished,
-use the {\tt chainslurm.sh} script in that same directory.
+use the {\tt chainslurm.sh} script in that same directory. You can view
+the job dependency using:
+\begin{verbatim}
+squeue -l -j job-id
+\end{verbatim}
+where {\tt job-id} is the number of the job.
 
@@
 \subsection{Batch visualization using \yt}
+
+\yt\ can be built using the {\tt install\_script.sh}.  It has been
+tested using the build of {\tt yt} from source and dependencies via conda,
+by setting:
+\begin{verbatim}
+INST_CONDA=1
+INST_YT_SOURCE=1
+\end{verbatim}
+in the {\tt install\_script.sh}.  Once these are set, run:
+\begin{verbatim}
+source install_script.sh
+\end{verbatim}
+Note: installation was done in the home directory.
+
+This way of building \yt\ installs it's own python and support
+libraries in a directory, {\tt yt-conda}.  You need to make sure that
+your start-up files (typically {\tt .bashrc.ext} at NERSC) don't {\tt
+  module load} python or any python libraries, as this will interfere
+with the conda installation.  The install script will direct you to
+add the install location to your path.
+
+At the moment, it seems that this does not actually build the yt source,
+so a second step is needed, in {\tt yt-conda/bin/src/yt-hg}, do
+\begin{verbatim}
+python setup.py develop --user
+\end{verbatim}
 

### 2016-07-12 Docs/managing_jobs/managingjobs.tex
update edison yt directions

@@
 
-The scripts {\tt parallel-yt} and {\tt process-edison.run} in {\tt
+The scripts {\tt parallel-yt} and {\tt process-edison.slurm} in {\tt
   Util/job\_scripts/edison} show how to invoke \yt\ to loop over a
-series of plotfiles and do visualization.  A number of tasks are
-run at once on the node, each operating on a separate file.  Note:
-it is important that {\tt aprun} be used to launch the \yt\ script
-to ensure that it is run on the compute node.
+series of plotfiles and do visualization.  A number of tasks are run
+at once on the node, each operating on a separate file.  The {\tt
+  parallel-yt} script then calls {\tt vol.py} to do the volume
+rendering with \yt.  Note: it is important that {\tt srun} be used to
+launch the \yt\ script to ensure that it is run on the compute node.
+
+A simple {\tt test-yt.slurm} script shows how to just call the
+\yt\ python script directly, using one node and 24 threads, again
+using {\tt srun} to execute on the compute node.
 

### 2016-07-13 Docs/architecture/architecture.tex
update EOS stuff to refer to Microphysics

@@
 
-Presently there are 3 equations of state that work with \maestro:
+Presently there is a single EOS that comes with \maestro, {tt gamma\_law\_general},
+but many more are available through the external \microphysics\ repo\footnote{\microphysics\ is
+available at \url{https://github.com/BoxLib-Codes/Microphysics}.  \maestro\ will
+find it via the {\tt MICROPHYSICS\_HOME} environment variable}.  The \microphysics\
+EOSs share the same interface and can be compiled into \maestro\ directly.  
+Here are the more popular EOSs:
 \begin{itemize}
-\item {\tt helmeos/} represents a general stellar equation 
+\item {\tt helmeos} represents a general stellar equation 
       of state, consisting of nuclei (as an ideal gas), radiation,

### 2016-07-13 Docs/managing_jobs/managingjobs.tex
add a note about gfortra 6.1.1 compiler bug

@@
 \begin{itemize}
-\item {\em gfortran 4.5.x}: gfortran 4.5 has no known issues.
-
-\item {\em gfortran 4.6.x}: gfortran 4.6.1--4.6.3 on Fedora 
-(15 and 16) have a bug that shows up in multi-processor, multilevel
-runs ({\tt toy\_convect} on 4 processors, with {\tt
-inputs\_2d\_ml.test} for instance).  The symptom is a crash during
-initialization with the message that ``message sizes do not match
-across processes in the collective'' in a {\tt parallel\_bcast} call.
-This is in {\tt cluster\_f.f90}, and the issue seems to be a
-multiplication error in computing the size to allocate {\tt bxs}.  The
-bug is not present when you compile in debug mode, so it seems to be
-related to optimization.
-
-\item {\em gfortran 4.7.x}: gfortran 4.7.1 (prerelease) seems to have 
-fixed the above issue, but now wdconvect with 2-level, effective
-768$^3$ dies due to an unknown seg fault at the beginning of the run.
 
@@
 
+gfortran 4.8.5 appears to work without issues
+
 \item {\em gfortran 5.1.1}: These compilers have no known issues.
@@
 \item {\em gfortran 5.3.x}: These compilers have no known issues.
+
+\item {\em gfortran 6.1.1}: These compilers have a bug in producing
+  optimized code for \boxlib's {\tt init\_layout\_tilearray}.  At the
+  moment, this is worked around in \boxlib\ by changing the routine
+  from a function to a subroutine.
 

### 2016-07-13 Docs/managing_jobs/managingjobs.tex
more compiler note updates

@@
 
+  gfortran 6.1.1 is used for nightly regression testing.
+
 \end{itemize}
@@
 system header files.  From version 16 onward, things should work.
+
+There are no known issues with PGI 16.5 compilers---these are used
+for nightly regression testing.
 
@@
 
-\item Cray 8.4.3 triggers a backtrace error in {\tt make\_pi\_cc.f90}---this is 
-  likely an optimization issue, and may be 2-d specific.
-
-  Previously,  Cray 8.4.0 worked well on edison/NERSC with MPI/OpenMP.
+\item Cray 8.4.x has worked in the past, but it has not been used
+  at NERSC in a while.
 
@@
 python} and {\tt python\_base} modules are loaded (via the {\tt
-module} command).
+module} command).  \MarginPar{this may have changed with the migration of NERSC to anaconda?}
 

### 2016-07-17 Docs/architecture/architecture.tex
helmeos -> helmholtz

@@
 
-  \item {\tt Microphysics/}
-
-    The core microphysics routines used by \maestro.  These are organized
+  \item {\tt Microphysics/}\footnote{Note: many more \compatible\ routines are available in the separate \microphysics\ git repo}
+
+    The basic microphysics routines used by \maestro.  These are organized
     into the following sub-directories.
@@
 
-      A collection of various EOSes for use with the code.  The two
-      important ones for \maestro\ are {\tt helmeos/} and {\tt
-        gamma\_law\_general/}.
+      The {\tt gamma\_law\_general/}.
 
@@
 
-      Various reaction networks for \maestro\ problems.  In addition
-      to providing a routine to evolve the nuclear species due to
-      reactions, the networks also define the species that are
-      advected by the code.
+      The basic {\tt general\_null} network that defines arbitrary
+      non-reacting species.
 
@@
 \begin{lstlisting}[language={[gnu]make},mathescape=false]
-EOS_DIR := helmeos   
+EOS_DIR := helmholtz   
 CONDUCTIVITY_DIR := constant
@@
 Note that the microphysics packages are listed simply by the name of 
-the directory containing the specific implementation (e.g. {\tt helmeos}).
+the directory containing the specific implementation (e.g. {\tt helmholtz}).
 By default, the build system will look in {\tt Microphysics/EOS/} for
@@
 
-Some packages (for instance, the {\tt helmeos}
+Some packages (for instance, the {\tt helmholtz}
 EOS) require Fortran include files.  The {\tt Fmincludes} variable
@@
 statement.  Presently, the only instance of this is with the Helmholtz
-general equation of state found in {\tt Microphysics/EOS/helmeos/}.  This is
+general equation of state found in {\tt Microphysics/EOS/helmholtz/}.  This is
 automatically handled by the {\tt GMaestro.mak} instructions.
@@
 \begin{itemize}
-\item {\tt helmeos} represents a general stellar equation 
+\item {\tt helmholtz} represents a general stellar equation 
       of state, consisting of nuclei (as an ideal gas), radiation,

### 2016-07-17 Docs/getting_started/getting_started.tex
helmeos -> helmholtz

@@
   current directory to the data table needed for the equation of state
-  ({\tt Microphysics/EOS/helmeos/helm\_table.dat}).
+  ({\tt Microphysics/EOS/helmholtz/helm\_table.dat}).
 

### 2016-07-25 Docs/managing_jobs/managingjobs.tex
update yt @ NERSC instructions

@@
 
-\item Intel 16.x currently has issues with the linear solvers in \boxlib\
-  and should be avoided.
+\item Intel 16.0.0 and 16.0.1 have issues compiling the linear solvers
+  in \boxlib\ and should be avoided.
+
+  Intel 16.0.2 works fine.
 
@@
 This way of building \yt\ installs it's own python and support
-libraries in a directory, {\tt yt-conda}.  You need to make sure that
-your start-up files (typically {\tt .bashrc.ext} at NERSC) don't {\tt
-  module load} python or any python libraries, as this will interfere
-with the conda installation.  The install script will direct you to
-add the install location to your path.
-
-At the moment, it seems that this does not actually build the yt source,
-so a second step is needed, in {\tt yt-conda/bin/src/yt-hg}, do
-\begin{verbatim}
-python setup.py develop --user
-\end{verbatim}
+libraries in a directory, {\tt yt-conda}.  {\bf Important: } You need
+to make sure that your start-up files (typically {\tt .bashrc.ext} at
+NERSC) don't {\tt module load} python or any python libraries, as this
+will interfere with the conda installation.  The install script will
+direct you to add the install location to your path.
 
@@
 using {\tt srun} to execute on the compute node.
+
+If you want to keep up with the development version of \yt, then you
+can update the source in {\tt yt-conda/bin/src/yt-hg}, using:
+\begin{verbatim}
+hg pull
+hg update yt
+\end{verbatim}
+and then rebuild it via:
+\begin{verbatim}
+python setup.py develop --user
+\end{verbatim}
 

### 2016-07-28 Docs/managing_jobs/managingjobs.tex

@@
 
+{\color{red} this section needs to be updated.  See the titan section}
+
 The best way to do visualization is to use rhea, the OLCF vis machine.
@@
 
-You can also run \yt\ python scripts in the titan batch queues to do your
+You can run \yt\ python scripts in the titan batch queues to do your
 visualization.  You need to install \yt\ and all its dependencies
@@
 compute nodes can see it.  A good choice is the project space, since
-that has a longer purge window.
+that has a longer purge window.  The following procedure will setup
+the development version of \yt\ (from source)
+\begin{itemize}
+\item create a directory in your {\tt \$PROJWORK} directory named {\tt yt/}
 
-Once \yt\ is install by the {\tt install\_script.sh} script, it will
-tell you what script to source to define the necessary paths.
+\item in {\tt yt/}, down load the \yt\ install script:
+\begin{verbatim}
+wget https://bitbucket.org/yt_analysis/yt/raw/yt/doc/install_script.sh
+\end{verbatim}
 
-The scripts {\tt vis-titan.run} and {\tt parallel-yt-new} in {\tt
-MAESTRO/Util/job\_scripts/titan/} will manage the \yt jobs by calling a
-python script for each file that matches a pattern.  Make sure you set
-the proper path for the \yt\ {\tt activate} script in {\tt
-parallel-yt-new}---this ensures that the paths to the various libraries
-are set on the compute nodes.
+\item edit the script to use Conda to get the necessary dependencies
+and to build \yt\ from source.  This is accomplished by setting the 
+following variables near the top of the script:
+\begin{verbatim}
+INST_CONDA=1
+INST_YT_SOURCE=1
+\end{verbatim}
 
+\item run the script:
+\begin{verbatim}
+source install_script.sh
+\end{verbatim}
+\end{itemize}
+
+When the script is done, you will have a new python installation in a sub-directory
+called {\tt yt-conda/} and the script will tell you how to modify your path
+in your {\tt .bashrc}
+
+{\bfseries Important: make sure that you are not loading any other
+  python environments in your {\tt .bashrc}, e.g., via modules.}
+
+To test thing out, start up the python shell, and try doing {\tt
+  import yt}.  If there are no errors, then you are good.
+
+The python code {\tt vol.py} and submission script {\tt
+  yt-vis-titan.run} in {\tt MAESTRO/Util/job\_scripts/titan/vis/} show
+how to do a simple visualization in the batch queue using yt.  Note
+that {\tt vol.py} is executable, and that we run it via {\tt aprun} to
+ensure that it is launched on the compute node.
+
+The scripts {\tt vis-titan.run} and {\tt parallel-yt-new} in that
+same directorywill manage the \yt jobs by calling a
+python script for each file that matches a pattern.  
 Note that the actual visualization command itself is launched by
-{\tt parallel-yt-new} using the {\tt aprun} command.  This is the
-only way to get a job to run on the compute nodes on titan.  But
+{\tt parallel-yt-new}, again using the {\tt aprun} command.  But
 {\tt aprun} can only launch a single job at a time, so this means

### 2016-08-01 Docs/FAQ/faq.tex

@@
   compilers is to add {\tt -pg} to the compilation flags for both {\tt
-    gfortran} and {\tt gcc} (this can be accomplished by setting {\tt
-    GPROF := t} in your {\tt GNUmakefile}.  Upon completion, a file
+    gfortran} and {\tt gcc}.  This can be accomplished by setting:
+  \begin{verbatim}
+    GPROF := t
+\end{verbatim}
+  in your {\tt GNUmakefile}.  Upon completion, a file
   names {\tt gmon.out} will be produced.  This can be processed by
@@
 
-  {\tt~~~ gprof} {\em exec-name}
-
-  where {\em exec-name} is the name of the executable.
+  \begin{verbatim}
+    gprof exec-name
+\end{verbatim}
+  where {\em exec-name} is the name of the executable.  More detailed
+  line-by-line timings can be obtained by using the {\tt -l} argument
+  to {\tt gprof}.
 

### 2016-08-01 Docs/makefiles/makefiles.tex

@@
 added to {\tt Fmincludes} variable in the problem's {\tt GNUmakefile}.
-This does not occur frequently.  For the case of the {\tt helmeos}
+This does not occur frequently.  For the case of the {\tt helmholtz}
 equation of state, this is done automatically in {\tt GMaestro.mak}.
@@
   \begin{itemize}
-  \item {\tt helmeos}
+  \item {\tt helmholtz}
   \item {\tt gamma\_law\_general}
@@
 
-\subsection{Debugging ({\tt print-*})}
+\subsection{Debugging}
+
+\subsubsection{({\tt print-*})}
 
@@
 This functionality is useful for debugging the makefiles.
+
+\subsubsection{{\tt file\_locations}}
+
+Source files are found by searching through the {\tt make} {\tt
+  vpath}.  The first instance of the file found in the {\tt vpath}
+is used in the build.  To see which files are used and their locations,
+do:
+\begin{verbatim}
+make file_locations
+\end{verbatim}
+
+This will also show any files that aren't found.  Some are expected
+(e.g., {\tt build\_info.f90} and {\tt probin.f90} are created at
+compile time), but other files that are not found could indicate
+an incomplete {\tt vpath}.
 

### 2016-08-01 Docs/runtime_parameters/runtime_parameters.tex

@@
 \verb= reaction_sum_tol = &  mass fraction sum tolerance (if they don't sum to 1 within this tolerance, we abort) & 1.d-10 \\
+
+\end{longtable}
+\end{center}
+
+} % ends \small
+
+{\small
+
+\renewcommand{\arraystretch}{1.5}
+%
+\begin{center}
+\begin{longtable}{|l|p{5.25in}|l|}
+\caption[ debugging
+ parameters.]{ debugging
+ parameters.} \label{table:  debugging
+ parameters. runtime} \\
+%
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} &
+       \multicolumn{1}{ c|}{\textbf{description}} &
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\tablename\ \thetable{}---continued}} \\
+\hline \multicolumn{1}{|c|}{\textbf{parameter}} &
+       \multicolumn{1}{ c|}{\textbf{description}} &
+       \multicolumn{1}{ c|}{\textbf{default value}} \\ \hline
+\endhead
+
+\multicolumn{3}{|r|}{{\em continued on next page}} \\ \hline
+\endfoot
+
+\hline
+\endlastfoot
+
+\rowcolor{tableShade}
+\verb= boxlib_fpe_invalid = &  enable floating point exception trapping for invalid & .false. \\
+\verb= boxlib_fpe_overflow = &  enable floating point exception trapping for overflow & .false. \\
+\rowcolor{tableShade}
+\verb= boxlib_fpe_zero = &  enable floating point exception trapping for divide by zero & .false. \\
 
@@
 \verb= hg_cycle_type = &  Type of cycle used in the nodal multigrid -- 1 = F-cycle, 2 = W-cycle, 3 = V-cycle & 3 \\
-\verb=  hg_dense_stencil  = &   In hgproject, in 2D, use a 9 point Laplacian (.true.) or 5-point Laplacian (.false.).
-In 3D, use a 27 point Laplacian (.true.) or 7-point Laplacian (.false.).
-&  .true.
-\\
+\verb= hg_dense_stencil = &  In hgproject, in 2D, use a 9 point Laplacian (.true.) or 5-point Laplacian (.false.).  In 3D, use a 27 point Laplacian (.true.) or 7-point Laplacian (.false.). & .true. \\
 \rowcolor{tableShade}

### 2016-08-08 Docs/managing_jobs/managingjobs.tex
document gpu profiling at OLCF

@@
 job.
+
+\subsection{Profiling and Debugging on GPUs}
+To get an idea of how code performs on Titan's GPUs, there are a few tools
+available.  We'll overview a few here.
+
+\subsubsection{{\tt Score-P} with {\tt CUBE} and {\tt vampir}}
+
+{\tt Score-P} is a profiling and tracing tool that can be used to instrument
+code to generate data for other tools to analyze, such as {\tt CUBE} and 
+{\tt vampir}.  These tools have been developed to analyze performance of HPC
+codes that run on several nodes, not specifically for analyzing GPU usage.
+Still, they do support some GPU analysis.  In the next section, we'll discuss
+NVIDA's tools specifically for analyzing GPU usage.
+At the time of writing, {\tt Score-P} usage is fairly well documented on OLCF's
+website here: \url{https://www.olcf.ornl.gov/kb_articles/software-scorep/}.
+We'll review the essentials here, but please see the link for more details.
+
+To instrument a code with {\tt Score-P} you must re-compile.  First, your
+desired modules will need to be loaded.  Please note that \emph{order is
+important} --- you need to load modules needed for compilation before loading 
+{\tt Score-P}.  The {\tt Score-P} module is designed to detect the loaded
+environment and will configure itself based on that.  These tools have been
+tested with the PGI 16.3.0 compilers, and we will use them in our examples.
+One possible set of module loads is
+
+\begin{lstlisting}[language=bash,mathescape=false]
+$ module load PrgEnv-pgi
+$ module swap pgi/15.7.0 pgi/16.3.0
+$ module load cudatoolkit
+$ module load scorep/3.0-rc1
+\end{lstlisting}
+
+In the above we've loaded version 3.0, release candidate 1, which added some
+support for analyzing OpenACC code.  The next step is to compile.  You
+essentially preface your normal compile (and link) line with the {\tt Score-P}
+executable and options.  As an example using the Fortran wrapper required on
+Titan, we have
+
+\begin{lstlisting}[language=bash,mathescape=false]
+$ scorep --cuda --openacc -v ftn program.f90
+\end{lstlisting}
+
+One way to achieve this in \maestro\ is to modify the appropriate make file.  For
+PGI, this would be {\tt \$BOXLIB\_HOME/Tools/F\_mk/comps/Linux\_pgi.mak}.  If
+this proves useful, it may be worth it to build {\tt Score-P} into the build
+infrastructure.
+
+Once compiled, we are ready to generate profiling and tracing data.  Among those
+that develop these tools, note that they draw a distinction between profiling
+and tracing.  Profiling generates a timing (or perhaps other metric) summary of
+the entire program's execution while tracing produces a timeline of the
+execution. {\tt Score-P}'s analysis is configured with environment variables.
+Some of the key configuration variables used in testing include
+
+\begin{lstlisting}[language=bash,mathescape=false]
+export SCOREP_ENABLE_PROFILING=yes
+export SCOREP_ENABLE_TRACING=yes
+export SCOREP_EXPERIMENT_DIRECTORY=profile-trace
+export SCOREP_CUDA_ENABLE=yes,kernel_counter,flushatexit
+export SCOREP_CUDA_BUFFER=200M
+export SCOREP_TOTAL_MEMORY=1G
+export SCOREP_OPENACC_ENABLE=yes
+\end{lstlisting}
+
+For a full listing and definition of configuration variables, execute
+\begin{lstlisting}[language=bash,mathescape=false]
+$ scorep-info config-vars --full 
+\end{lstlisting}
+
+Except for very simple codes, you will never want to enable both tracing and
+profiling.  The overhead is too high, and the code will likely crash or be
+excessively slow.  Typically, it's best to profile first and then trace.  The
+profiling data can be used to help configure tracing (as we'll see shortly).
+
+Once the configuration is set, simply run the code as you normally would.
+Experience suggests you will need to load the same modules that were loaded for
+compilation when executing.  If analysis is being done through a batch script,
+note that you cannot do a simple {\tt module load ...} in the script.  First you
+need to do {\tt source /opt/modules/default/init/bash} in the script, and then
+module loads will work as usual.
+
+After executing, analysis data will be stored in the specified
+{\tt SCOREP\_EXPERIMENT\_DIRECTORY}.  With profiling, you
+will see a file like profile.cubex.  This can be opened with cube 
+(module load cube).
+
+As mentioned, the profiling data can be used to get
+recommended settings for tracing.  Running
+\begin{lstlisting}[language=bash,mathescape=false]
+$ scorep-score -r profile.cubex
+\end{lstlisting}
+will yield output showing estimated sizes for e.g.~{\tt SCOREP\_TOTAL\_MEMORY}.  It
+can also suggest functions to that are called a lot to be filtered out of the
+trace.  If you don't care about them and they're slowing scorep down (or making
+an outrageously large output file), this is useful to do.
+
+Running with tracing enabled will generate a traces.otf2 file that can be
+inspected with vampir (module load vampir)
+
+\subsubsection{{\tt nvprof} and {\tt nvvp}}
+
+NVIDIA provides tools for specifically analyzing how your code utilizing their
+GPUs.  {\tt Score-P} is a fully-featured profiling with some CUDA and OpenACC
+support.  It can be useful for providing context for GPU execution and it allows
+you to, for example, see line numbers for OpenACC directives that are executed.
+{\tt nvprof} will only analyze GPU execution, but in exchange you get much more
+detail than is available with {\tt Score-P}.  {\tt nvvp} is NVIDIA's visual
+profiler.  It can be used to read data generated by {\tt nvprof}.  Most useful
+is the guided analysis it will perform, which analyzes your code's GPU
+performance for bottlenecks and suggests ways to improve performance.  Both are
+provided when you load the {\tt cudatoolkit} module.
+
+With {\tt nvprof}, no instrumentation is necessary.  Instead, you compile
+normally and then run {\tt nvprof} on the executable.  As before, be sure when
+executing to load the modules used for compilation.  Executing {\tt nvprof} on
+Titan's compute nodes requires some unexpected options having to do with how
+{\tt aprun} and {\tt nvprof} interact.
+
+To get a basic overview printed to the terminal on Titan's compute node, execute
+\begin{lstlisting}[language=bash,mathescape=false]
+$ aprun -b nvprof --profile-child-processes ./gpuprogram.exe arg1 arg2... 
+\end{lstlisting}
+
+To generate tracing data for {\tt nvvp}, execute
+\begin{lstlisting}[language=bash,mathescape=false]
+$ aprun -b nvprof --profile-child-processes -o nvprof.timeline.out%p 
+  ./gpuprogram.exe arg1 arg2... 
+\end{lstlisting}
+{\tt nvvp} can then be used to read {\tt nvprof.timeline.out\%p}, where the 
+{\tt \%p} will be replaced with the process ID.
+
+To generate profile-like metric data for {\tt nvvp}, execute
+\begin{lstlisting}[language=bash,mathescape=false]
+$ aprun -b nvprof --profile-child-processes --analysis-metrics 
+  -o nvprof.metrics.out%p ./gpuprogram.exe arg1 arg2... 
+\end{lstlisting}
+This is the output needed for {\tt nvvp}'s guided analysis.
+
+\subsubsection{Target Metrics}
+The output from profilers may be difficult to makes sense of.  The purpose of
+this section is to note different metrics and reasonable targets for them.
+Note that these may be specific to the k20x hardware in Titan.
+\begin{itemize}
+   \item Threads per block: 256-512.  Note that if your code requires many
+      registers per thread, then this will limit the number of threads per block.
+   \item Occupancy: 60\% is a reasonable target.  We have had success with codes
+      even achieving only 23\% occupancy.  
+\end{itemize}
 

### 2016-08-09 Docs/managing_jobs/managingjobs.tex
Polish GPU profiling documentation.

@@
 \begin{lstlisting}[language=bash,mathescape=false]
-$ scorep --cuda --openacc -v ftn program.f90
+$ scorep --cuda --openacc -v ftn gpuprogram.f90
 \end{lstlisting}
@@
 {\tt SCOREP\_EXPERIMENT\_DIRECTORY}.  With profiling, you
-will see a file like profile.cubex.  This can be opened with cube 
-(module load cube).
+will see a file like {\tt profile.cubex}.  This can be opened with cube 
+({\tt module load cube}).
 
-As mentioned, the profiling data can be used to get
-recommended settings for tracing.  Running
+As mentioned, the profiling data can be used to get recommended settings for
+tracing.  Running
 \begin{lstlisting}[language=bash,mathescape=false]
@@
 \end{lstlisting}
-will yield output showing estimated sizes for e.g.~{\tt SCOREP\_TOTAL\_MEMORY}.  It
-can also suggest functions to that are called a lot to be filtered out of the
-trace.  If you don't care about them and they're slowing scorep down (or making
-an outrageously large output file), this is useful to do.
+will yield output showing estimated sizes for e.g.~{\tt SCOREP\_TOTAL\_MEMORY}.
+It also list functions that are called many times.  If you don't care about
+them and they're slowing {\tt Score-P} down (or making an outrageously large output
+file), you can configure {\tt Score-P} to ignore them in its analysis.  To filter a set
+of functions, you need to provide a filter file, for example
+\begin{lstlisting}[language=bash,mathescape=false]
+$ export SCOREP_FILTERING_FILE=scorep.filter
+\end{lstlisting}
+where
+\begin{lstlisting}[language=bash,mathescape=false]
+$ cat scorep.filter
+SCOREP_REGION_NAMES_BEGIN
+ EXCLUDE
+   matmul_sub
+   matvec_sub
+SCOREP_REGION_NAMES_END
+\end{lstlisting}
+This would tell {\tt Score-P} not to trace the routines {\tt matmul\_sub} and 
+{\tt matvec\_sub}.  See the OLCF KnowledgeBase article and/or {\tt Score-P}'s
+help for more, but this doesn't seem to be the best-documented aspect of the
+program.
 
@@
 
-NVIDIA provides tools for specifically analyzing how your code utilizing their
-GPUs.  {\tt Score-P} is a fully-featured profiling with some CUDA and OpenACC
+NVIDIA provides tools for specifically analyzing how your code utilizes their
+GPUs.  {\tt Score-P} is a fully-featured profiler with some CUDA and OpenACC
 support.  It can be useful for providing context for GPU execution and it allows
@@
 normally and then run {\tt nvprof} on the executable.  As before, be sure when
-executing to load the modules used for compilation.  Executing {\tt nvprof} on
+executing to load the modules used at compile-time.  Executing {\tt nvprof} on
 Titan's compute nodes requires some unexpected options having to do with how
@@
 {\tt nvvp} can then be used to read {\tt nvprof.timeline.out\%p}, where the 
-{\tt \%p} will be replaced with the process ID.
+{\tt \%p} will be replaced with the process ID.  You \emph{must} include \%p in
+the output file's name or the code will crash, even if you're not running a
+multi-process code.
 

### 2016-08-12 Docs/managing_jobs/managingjobs.tex

@@
 \end{itemize}
+
+One very useful tool for determining target metrics and what is limiting your
+performance is a spreadsheet developed by NVIDIA to calculate occupancy.  Every
+installation of the CUDA Toolkit should have this occupancy calculator in a
+tools subdirectory.  At time of writing, the calculator is also available at
+this link:
+\url{http://developer.download.nvidia.com/compute/cuda/CUDA_Occupancy_calculator.xls}.
+The document is actually  more than a simple calculator.  It contains quite a
+bit of interesting insight into optimizing a GPU code.  More on occupancy can be
+found here:
+\url{http://docs.nvidia.com/cuda/cuda-c-best-practices-guide/index.html#occupancy}.
 

### 2017-05-23 Docs/managing_jobs/managingjobs.tex

@@
 
-  gfortran 6.1.1 is used for nightly regression testing.
+\item {\em gfortran 6.2}: These compilers work without any known issues.
+
+  gfortran 6.2.1 is used for nightly regression testing.
 
@@
 The scripts {\tt vis-titan.run} and {\tt parallel-yt-new} in that
-same directorywill manage the \yt jobs by calling a
+same directorywill manage the \yt\ jobs by calling a
 python script for each file that matches a pattern.  

### 2017-05-25 Docs/getting_started/getting_started.tex
update to reflect amrex

@@
   from the project's {\sf github} page:
-  \url{https://github.com/BoxLib-Codes/MAESTRO}\, .  There are several
+  \url{https://github.com/AMReX-Astro/MAESTRO}\, .  There are several
   options: you can fork it directly on {\sf github} (recommended if
@@
   \begin{verbatim}
-git clone https://github.com/BoxLib-Codes/Microphysics.git
+git clone https://github.com/starkiller-astro/Microphysics.git
   \end{verbatim}
@@
 .
-\item {\em Get a copy of \boxlib}.
+\item {\em Get a copy of \amrex}.
 
-  \maestro\ requires the \boxlib\ library to manage the grids and
-  parallelization.  We also rely on the build system in \boxlib\ to
-  build a \maestro\ executable.  \boxlib\ is also available on github
+  \maestro\ requires the \amrex\ library to manage the grids and
+  parallelization.  We also rely on the build system in \amrex\ to
+  build a \maestro\ executable.  \amrex\ is also available on github
   and can be obtained via:
   \begin{verbatim}
-git clone https://github.com/BoxLib-Codes/BoxLib.git
+git clone https://github.com/AMReX-Codes/amrex.git
   \end{verbatim}
 
-  You will periodic want to update \boxlib\ by doing
+  You will periodic want to update \amrex\ by doing
   \begin{verbatim}
@@
   \end{verbatim}
-  in the {\tt BoxLib/} directory.
+  in the {\tt amrex/} directory.
 
@@
 
-  \maestro\ needs to know where to find \boxlib, by specifying the
-  {\tt BOXLIB\_HOME} environment variable, and where
+  \maestro\ needs to know where to find \amrex, by specifying the
+  {\tt AMREX\_HOME} environment variable, and where
   to find \microphysics, bt specifying the {\tt
@@
   \begin{verbatim}
-export BOXLIB_HOME="/path/to/BoxLib/"
+export AMREX_HOME="/path/to/amrex/"
 export MICROPHYSICS_HOME="/path/to/Microphysics/"
@@
   \begin{verbatim}
-setenv BOXLIB_HOME /path/to/BoxLib/
+setenv AMREX_HOME /path/to/amrex/
 setenv MICROPHYSICS_HOME /path/to/Microphysics/
@@
 
-  Note: you must specify the full path to the {\tt BoxLib/} and {\tt
+  Note: you must specify the full path to the {\tt AMReX/} and {\tt
     Microphysics/} directory.  Do not use ``$\sim$'' to refer to your
@@
     Specifying this compiler will automatically pull in the compiler
-    settings as specified in {\tt BOXLIB\_HOME/Tools/F\_mk/}.
+    settings as specified in {\tt AMREX\_HOME/Tools/F\_mk/}.
     (Alternate compiler choices include {\tt Intel}, {\tt PGI}, {\tt
@@
   outputted.  Similarly, the checkpoints are named {\tt chk}{\em
-    nnnnn}.  \boxlib\ plotfiles and checkpoints are actually
+    nnnnn}.  \amrex\ plotfiles and checkpoints are actually
   directories, with the data stored in sub-directories grouped by
@@
 Visualization and analysis are done on the plotfiles.  A number of
-in-house and externally developed tools can work
-with \boxlib-formatted plotfiles.  An example plot of the {\tt reacting\_bubble}
-problem run above is shown in Figure~\ref{fig:gettingstarted:test2}.
+in-house and externally developed tools can work with \amrex-formatted
+plotfiles\footnote{The plotfiles are in the same format as those made
+  by the \boxlib\ library upon which \maestro\ was previously based.}.
+An example plot of the {\tt reacting\_bubble} problem run above is
+shown in Figure~\ref{fig:gettingstarted:test2}.
 
@@
 \end{verbatim}
-\amrvis\ is built in the C++ \boxlib\ framework (instead of the Fortran 
-\boxlib\ framework that \maestro\ uses).  The build systems are similar,
+\amrvis\ is built in the C++ \amrex\ framework (instead of the Fortran 
+\amrex\ framework that \maestro\ uses).  The build systems are similar,
 but differ in a few ways.
@@
 
-Further details on the C++ \boxlib\ build system used by \amrvis\
-can be found in the \boxlib\ documentation.
+Further details on the C++ \amrex\ build system used by \amrvis\
+can be found in the \amrex\ documentation.
 
@@
 datasets.  It can do contouring, volume rendering, streamlines, ...\, ,
-directly from \boxlib\ plotfiles.   Details on
+directly from \amrex\ plotfiles.   Details on
 \visit\ can be found at:\newline
@@
 
-Once \visit\ is installed, you can open a \boxlib\ plotfile by pointing
+Once \visit\ is installed, you can open a \amrex\ plotfile by pointing
 \visit\ to the {\tt Header} file in the plotfile directory.
@@
    A 1-d version of the {\tt flame} problem above.  This uses a special
-   elliptic solver in \boxlib\ that only works for a single grid, so
+   elliptic solver in \amrex\ that only works for a single grid, so
    no parallel runs are allowed for this problem.
@@
   conversion between the string names and integer flags is done
-  by \boxlib\ in the \code{bc\_module} at the time of initializing
+  by \amrex\ in the \code{bc\_module} at the time of initializing
   the runtime parameters.
@@
 
-When you close \maestro\ from github, you will be on the {\tt master}
+When you clone \maestro\ from github, you will be on the {\tt master}
 branch of the repo.  New changes to \maestro\ are first introduced
@@
 tests pass) and tag the state of the code with a date-based tag {\tt
-  YY-MM}.  We do this on all the other repos in the \boxlib-ecosystem,
-including {\tt BoxLib/}, {\tt Microphysics/}, and {\tt Castro/}.
+  YY-MM}.  We do this on all the other repos in the \amrex-ecosystem,
+including {\tt amrex/}, {\tt Microphysics/}, and {\tt Castro/}.
 
@@
 To run in parallel with MPI, you would set {\tt MPI := t} in your {\tt
-GNUmakefile}.  For a machine with working MPI compiler wrappers ({\tt
-mpif90} and {\tt mpicc}), setting the environment variable {\tt
-BOXLIB\_USE\_MPI\_WRAPPERS=1} will have the code build using these
-wrappers.  This is the easiest way to do a parallel build, and should
-work on most Linux systems.
+  GNUmakefile}.  For a machine with working MPI compiler wrappers
+({\tt mpif90} and {\tt mpicc}), the build system should find these and
+compile with MPI support automatically.  This is the easiest way to do
+a parallel build, and should work on most Linux systems.
 
@@
 set the proper libraries.  For other machines, you may need to edit
-the {\tt GMake.MPI} file in the \boxlib\ build files.  See
+the {\tt GMake.MPI} file in the \amrex\ build files.  See
 \S~\ref{ch:make} for more details.

### 2017-07-20 Docs/makefiles/makefiles.tex

@@
 \begin{itemize}
-\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakedefs.mak}:
+\item {\tt \$\{AMREX\_HOME\}/Tools/F\_mk/GMakedefs.mak}:
 
@@
 
-\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakeMPI.mak}:
+\item {\tt \$\{AMREX\_HOME\}/Tools/F\_mk/GMakeMPI.mak}:
 
@@
 
-\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakerules.mak}:
+\item {\tt \$\{AMREX\_HOME\}/Tools/F\_mk/GMakerules.mak}:
 
@@
   modules that are used to build a typical \maestro\ application
-  and integrates with the \boxlib\ build system.  Every \maestro\
+  and integrates with the \amrex\ build system.  Every \maestro\
   problem's {\tt GNUmakefile} will include this file.
@@
 
-\maestro\ gets the location of the \boxlib\ library through the 
-{\tt BOXLIB\_HOME} variable.  This should be set as an environment
+\maestro\ gets the location of the \amrex\ library through the 
+{\tt AMREX\_HOME} variable.  This should be set as an environment
 variable in your shell start-up files (e.g.\ {\tt .bashrc} or {\tt
@@
 
-The \boxlib\ build system separates the compiler-specific information
+The \amrex\ build system separates the compiler-specific information
 from the machine-specific information---this allows for reuse of the
@@
 The compiler flags for the various compilers are listed in the
-files in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.  The compiler
+files in {\tt \$\{AMREX\_HOME\}/Tools/F\_mk/comps/}.  The compiler
 is set via the {\tt COMP} variable in the problem's {\tt GNUmakefile}.
@@
 
-The \boxlib\ build system relies on the vpath functionality of {\tt
+The \amrex\ build system relies on the vpath functionality of {\tt
 make}.  In a makefile, the {\tt vpath} variable holds search path used
@@
 source files for Fortran modules.  The scripts in {\tt
-BOXLIB\_HOME/Tools/F\_scripts/} are run at the start of the build
+AMREX\_HOME/Tools/F\_scripts/} are run at the start of the build
 process and parse all the source files and make an explicit list of
@@
   created by the script
-  {\tt write\_probin.py} in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_scripts/}.  The
+  {\tt write\_probin.py} in {\tt \$\{AMREX\_HOME\}/Tools/F\_scripts/}.  The
   makefile logic for creating it is in {\tt GMaestro.mak}.  At compile
@@
   flags, etc.).  This is created by the script {\tt makebuildinfo.py}
-  in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_scripts/} from {\tt
+  in {\tt \$\{AMREX\_HOME\}/Tools/F\_scripts/} from {\tt
   GMaestro.mak} by passing in a number of makefile variables.  This is
@@
 Several modules are included in all \maestro\ builds by default.
-From \boxlib, we alway include:
+From \amrex, we alway include:
 \begin{itemize}
-\item {\tt \$\{BOXLIB\_HOME\}/Src/F\_BaseLib}
-\item {\tt \$\{BOXLIB\_HOME\}/Src/LinearSolvers/F\_MG}
+\item {\tt \$\{AMREX\_HOME\}/Src/F\_BaseLib}
+\item {\tt \$\{AMREX\_HOME\}/Src/LinearSolvers/F\_MG}
 \end{itemize}
@@
 
-\subsection{\boxlib-only Tests}
+\subsection{\amrex-only Tests}
 
 An even more restrictive setting than {\tt UNIT\_TEST := t} is invoked
-by setting {\tt BOXLIB\_ONLY := t}.  This is like the unit test flag,
+by setting {\tt AMREX\_ONLY := t}.  This is like the unit test flag,
 but does not include {\tt MAESTRO/Source} in the {\tt vpath} search
@@
 
-\boxlib\ has several options that produce executables that can help
+\amrex\ has several options that produce executables that can help
 track down memory issues, uninitialized variables, NaNs, etc.
@@
   Setting {\tt TEST := t}\otherindex{{\tt GNUmakefile}}{TEST} will
-  enable routines in \boxlib\ initialize multifabs and arrays 
+  enable routines in \amrex\ initialize multifabs and arrays 
   allowed via {\tt bl\_allocate} to signalliing NaNs.  This behavior
@@
   exceptions (FPEs), but checks on floating point exceptions can also
-  be enabled through runtime parameters passed to \boxlib's
+  be enabled through runtime parameters passed to \amrex's
   backtrace functionlity:
@@
 
-  When exception trapping is enabled (either via \boxlib\ or the
+  When exception trapping is enabled (either via \amrex\ or the
   compiler), the code will abort, and the backtrace information will
   be output to a file {\tt Backtrace.N}, where {\tt N} is the
-  processor number.  BoxLib will also initialize \multifab s with
+  processor number.  \amrex\ will also initialize \multifab s with
   signaliing NaNs to help uncover any floating point issues.
@@
 
-  Behind the scenes, \boxlib\ implements this capability via the
+  Behind the scenes, \amrex\ implements this capability via the
   Linux/Unix {\tt feenableexcept} function (this is in {\tt
-    backtrace\_c.cpp} in \boxlib).
+    backtrace\_c.cpp} in \amrex).
 
@@
 Properties for different compilers are already defined in {\tt
-  \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.  Each compiler is given its
+  \$\{AMREX\_HOME\}/Tools/F\_mk/comps/}.  Each compiler is given its
 own file.  The appropriate file is included into {\tt GMakedefs.mak}
@@
 
-When building with MPI, the build system needs to know about the location of the
-MPI libraries.  If your local MPI has the {\tt mpif90} and {\tt mpicc} wrappers
-installed and working, then \maestro\ can use these simply by setting the
-environment variable:
-\begin{verbatim}
-BOXLIB_USE_MPI_WRAPPERS=1
-\end{verbatim}
-Otherwise, you will need to edit {\tt GMakeMPI.mak} and add a section
-specific to your machine with the compiler and library location.  It
-is best to simply copy an existing similar portion of the makefile and
-adjust it to your system.  Most national supercomputing facilities are
-already supported, and parallel builds on them should work out of the
-box.
+When building with MPI, the build system needs to know about the
+location of the MPI libraries.  If your local MPI has the {\tt mpif90}
+and {\tt mpicc} wrappers installed and working, then \maestro\ will
+attempt to use these.  Otherwise, you will need to edit {\tt
+  GMakeMPI.mak} and add a section specific to your machine with the
+compiler and library location.  It is best to simply copy an existing
+similar portion of the makefile and adjust it to your system.  Most
+national supercomputing facilities are already supported, and parallel
+builds on them should work out of the box.

### 2017-09-08 Docs/mg/mg.tex

@@
 
-\section{\boxlib\ Multigrid Philosophy}
+\section{\amrex\ Multigrid Philosophy}
 
-Here we describe some of the general ideas behind the \boxlib\
+Here we describe some of the general ideas behind the \amrex\
 multigrid (MG).
@@
 We solve MG on an AMR hierarchy, which means in places we will encounter
-C-F interfaces.  The \boxlib\ MG will modify the stencil for the Laplacian
+C-F interfaces.  The \amrex\ MG will modify the stencil for the Laplacian
 at C-F interfaces to ensure the correct solution to the Poisson equation
@@
 some codes (but not \maestro) we can have jumps of $4\times$ between
-levels.  \boxlib\ uses a {\tt mini\_cycle} in these cases, effectively
+levels.  \amrex\ uses a {\tt mini\_cycle} in these cases, effectively
 inserting a multigrid level between the two AMR levels and doing a mini
@@
 
-The MG solvers are located in {\tt BoxLib/Src/LinearSolvers/F\_MG/}.
+The MG solvers are located in {\tt amrex/Src/LinearSolvers/F\_MG/}.
 There are two MG solvers, for cell-centered and nodal data.
@@
 The \mgtower\ is a special Fortran derived type that carries all the
-information required for a \boxlib\ multigrid solve.
+information required for a \amrex\ multigrid solve.
 
@@
 
-There are several bottom solvers available in \boxlib.  For \maestro.
+There are several bottom solvers available in \amrex.  For \maestro.
 These are set through the {\tt mg\_bottom\_solver} (MAC/cell-centered)
@@
 step (the HG projection).  Two multigrid solvers are provided by
-\boxlib---one for cell-centered data and one for node-centered (nodal)
+\amrex---one for cell-centered data and one for node-centered (nodal)
 data.  Both of these are used in \maestro.

### 2018-02-27 Exec/TEST_PROBLEMS/reacting_bubble/init_perturb.f90
boxlib -> amrex in documentation.  update ecosystem flow charts as well

@@ -102,8 +102,8 @@ contains
 
     t0 = s0_init(temp_comp)
 
-    x0 = 3.6d7
-    y0 = 3.6d7
+    x0 = 1.35d7
+    y0 = 1.35d7
     z0 = 8.5d7
 
     r0 = sqrt( (x-x0)**2 + (y-y0)**2 + (z-z0)**2 ) / 2.5d6

### 2018-02-28 Docs/analysis/analysis.tex

@@
 
-The {\tt BoxLib/Tools/Postprocessing/F\_Src/} directory contains a large
-number of Fortran-based analysis routines for BoxLib datasets.  Many
+The {\tt amrex/Tools/Postprocessing/F\_Src/} directory contains a large
+number of Fortran-based analysis routines for AMReX datasets.  Many
 of these can be used with both \maestro\ and the compressible
@@
 
-The following routines are generally applicable for any BoxLib-based
+The following routines are generally applicable for any AMReX-based
 plotfile.  Typing the executable names without any arguments will
@@
   variable.  This assumes that the grids are identical. \\[-3mm]
+
+  With the optional {\tt --zone\_info var} argument, where {\tt var}
+  is the name of a variable, it will also report the full state
+  for the zone where {\tt var} has the largest error.
 
@@
 {\tt fspeciesmass2d.f90} should provide a good starting point for
-writing a new analysis routine for BoxLib data.
+writing a new analysis routine for AMReX data.
 

### 2018-03-12 Docs/FAQ/faq.tex

@@
 
-   The environment variable {\tt BOXLIB\_HOME} needs to be the full path
-   to the {\tt BoxLib} directory.  You cannot use `{\tt $\sim$}' as a shortcut
+   The environment variable {\tt AMREX\_HOME} needs to be the full path
+   to the {\tt amrex/} directory.  You cannot use `{\tt $\sim$}' as a shortcut
    for your home directory.
@@
        \begin{verbatim}
-       .../BoxLib/Tools/F_mk/GMakeMPI.mak:40: Extraneous text after `else' directive
-       .../BoxLib/Tools/F_mk/GMakeMPI.mak:47: *** only one `else' per conditional.  Stop
+       .../amrex/Tools/F_mk/GMakeMPI.mak:40: Extraneous text after `else' directive
+       .../amrex/Tools/F_mk/GMakeMPI.mak:47: *** only one `else' per conditional.  Stop
       \end{verbatim}
@@
 
-\boxlib\ can intercept floating point exceptions and provide a helpful
+\amrex\ can intercept floating point exceptions and provide a helpful
 backtrace that shows you where they were generated.  See \S
@@
  One implementation of this can be found in {\tt
-   BoxLib/Tools/Postprocessing/F\_Src/tutorial/fwrite2d.f90}.  This reads in
+   amrex/Tools/Postprocessing/F\_Src/tutorial/fwrite2d.f90}.  This reads in
  the plotfile data using the {\tt plotfile\_module} that the {\tt
    data\_processing} routines rely on, but then builds a multifab
-   and writes the data out to a plotfile using the BoxLib write
+   and writes the data out to a plotfile using the \amrex\ write
    routines.

### 2018-03-21 Docs/runtime_parameters/runtime_parameters.tex
Clarify description of drdxfac parameter.

@@
 \verb= amr_buf_width = &  the number of buffer zones surrounding a cell tagged for refinement. note that this needs to be >= regrid\_int & -1 \\
-\verb= bcx_hi = &  $+x$ boundary condition (valid values are listed in {\tt  boxlib/bc.f90}) & SLIP\_WALL \\
+\verb= bcx_hi = &  $+x$ boundary condition (valid values are listed in {\tt  amrex/Src/F\_BaseLib/bc.f90}) & SLIP\_WALL \\
 \rowcolor{tableShade}
@@
 \verb= do_2d_planar_octant = &  Set to 1 if using the 2D simplified (planar) model of an octant. & 0 \\
-\verb= drdxfac = &  ratio of radial base state zones to Cartesian full state zones for spherical geometry & 1 \\
+\verb= drdxfac = &  Number of radial base state zones per Cartesian full state zone along the first dimension at the finest level for spherical geometry & 1 \\
 \rowcolor{tableShade}
@@
 \verb= bds_type = &  0 = use ppm instead for multi-d integrator \newline 1 = bilinear & 0 \\
-\verb= beta_type = &  what type of coefficient to use inside the velocity divergence constraint.
+\verb= beta0_type = &  what type of coefficient to use inside the velocity divergence constraint.
 \newline {\tt beta\_type} = 1 uses $\beta_0$; \newline {\tt beta\_type} = 2 uses $\rho_0$ (anelastic); \newline {\tt beta\_type} = 3 uses 1 (small-scale combustion).
@@
 \verb= use_etarho = &  turn on the etarho term as described in flow chart & .true. \\
-\verb= use_linear_grav_in_beta = &  how to represent gravity in the $\beta_0$ integration: .true. = piecewise linear .false. = piecewise constant & .false. \\
+\verb= use_linear_grav_in_beta0 = &  how to represent gravity in the $\beta_0$ integration: .true. = piecewise linear .false. = piecewise constant & .false. \\
 

### 2018-05-01 Docs/runtime_parameters/runtime_parameters.tex

--- 2016-08-01 Docs/runtime_parameters/runtime_parameters.tex
@@
 \rowcolor{tableShade}
+\verb= burning_cutoff_density = &  Burning cutoff density (do not burn if density is below this threshold) Initialized to zero, runtime_init() in probin.f90 will check to see if burning_cutoff_density < 0 and if so will set burning_cutoff_density = base_cutoff_density & -1.0d0 \\
 \verb= do_burning = &  turn on (.true.) or off (.false.) burning & .true. \\
+\rowcolor{tableShade}
 \verb= do_subgrid_burning = &  break a zone into subzones, call the burner in each subzone and then average the result to the original cell & .false. \\
-\rowcolor{tableShade}
 \verb= reaction_sum_tol = &  mass fraction sum tolerance (if they don't sum to 1 within this tolerance, we abort) & 1.d-10 \\
@@
 \verb= bds_type = &  0 = use ppm instead for multi-d integrator \newline 1 = bilinear & 0 \\
-\verb= beta_type = &  what type of coefficient to use inside the velocity divergence constraint.
-\newline {\tt beta\_type} = 1 uses $\beta_0$; \newline {\tt beta\_type} = 2 uses $\rho_0$ (anelastic); \newline {\tt beta\_type} = 3 uses 1 (small-scale combustion).
+\verb= beta0_type = &  what type of coefficient to use inside the velocity divergence constraint.
+\newline {\tt beta0\_type} = 1 uses $\beta_0$; \newline {\tt beta0\_type} = 2 uses $\rho_0$ (anelastic); \newline {\tt beta0\_type} = 3 uses 1 (small-scale combustion).
 & 1 \\
@@
 \rowcolor{tableShade}
-\verb= do_smallscale = &  force $\rho_0 = (\rho h)_0 = 0$, {\tt evolve\_base\_state = F} and {\tt beta\_type} = 3 & .false. \\
+\verb= do_smallscale = &  force $\rho_0 = (\rho h)_0 = 0$, {\tt evolve\_base\_state = F} and {\tt beta0\_type} = 3 & .false. \\
 \verb= do_sponge = &  Use sponging. & .false. \\

### 2018-05-04 Docs/runtime_parameters/runtime_parameters.tex
fix latex issue

@@
 \rowcolor{tableShade}
-\verb= burning_cutoff_density = &  Burning cutoff density (do not burn if density is below this threshold) Initialized to zero, runtime_init() in probin.f90 will check to see if burning_cutoff_density < 0 and if so will set burning_cutoff_density = base_cutoff_density & -1.0d0 \\
+\verb= burning_cutoff_density = &  Burning cutoff density (do not burn if density is below this threshold) Initialized to zero, runtime\_init() in probin.f90 will check to see if burning_cutoff_density < 0 and if so will set burning_cutoff_density = base_cutoff_density & -1.0d0 \\
 \verb= do_burning = &  turn on (.true.) or off (.false.) burning & .true. \\

### 2018-05-04 Docs/runtime_parameters/runtime_parameters.tex
another latex fix

--- 2018-05-01 Docs/runtime_parameters/runtime_parameters.tex
@@
 \rowcolor{tableShade}
-\verb= burning_cutoff_density = &  Burning cutoff density (do not burn if density is below this threshold) Initialized to zero, runtime_init() in probin.f90 will check to see if burning_cutoff_density < 0 and if so will set burning_cutoff_density = base_cutoff_density & -1.0d0 \\
+\verb= burning_cutoff_density = &  Burning cutoff density (do not burn if density is below this threshold) Initialized to zero, runtime\_init() in probin.f90 will check to see if burning\_cutoff\_density < 0 and if so will set burning\_cutoff\_density = base\_cutoff\_density & -1.0d0 \\
 \verb= do_burning = &  turn on (.true.) or off (.false.) burning & .true. \\

### 2018-07-02 Docs/managing_jobs/managingjobs.tex
amrex -> FBoxLib

@@
 \item The build date, machine, and directory
-\item The \maestro, \boxlib, and other relevant git hashes for the source
+\item The \maestro, \fboxlib, and other relevant git hashes for the source
 \item The directories used in the build
@@
 
-\item {\em gfortran 6.1.1}: These compilers have a bug in producing
-  optimized code for \boxlib's {\tt init\_layout\_tilearray}.  At the
-  moment, this is worked around in \boxlib\ by changing the routine
-  from a function to a subroutine.
-
 \item {\em gfortran 6.2}: These compilers work without any known issues.
@@
 
-The \boxlib\ floating point exception trapping is disabled with PGI
+The \fboxlib\ floating point exception trapping is disabled with PGI
 compilers earlier than version 16, due to problems with PGI using the
@@
 Additionally, if the {\tt ftime} executable is in your path ({\tt
-ftime.f90} lives in {\tt BoxLib/Tools/Postprocessing/F\_src/}), then
+ftime.f90} lives in {\tt amrex/Tools/Postprocessing/F\_src/}), then
 the script will create a file called {\tt ftime.out} that lists the
@@
 One way to achieve this in \maestro\ is to modify the appropriate make file.  For
-PGI, this would be {\tt \$BOXLIB\_HOME/Tools/F\_mk/comps/Linux\_pgi.mak}.  If
+PGI, this would be {\tt \$FBOXLIB\_HOME/Tools/F\_mk/comps/Linux\_pgi.mak}.  If
 this proves useful, it may be worth it to build {\tt Score-P} into the build
@@
 
-\item Intel 16.0.0 and 16.0.1 have issues compiling the linear solvers
-  in \boxlib\ and should be avoided.
-
-  Intel 16.0.2 works fine.
+\item Intel 16.0.2 works fine.
 

### 2018-07-02 Docs/preface/preface.tex
amrex -> FBoxLib

@@
 
-\noindent \maestro\ uses the \boxlib\ library,
-developed at the Center for Computational Sciences and Engineering at
-Lawrence Berkeley National Laboratory.  \boxlib\ is distributed
-separately from \maestro\ (see
-\url{https://ccse.lbl.gov/BoxLib/})
-\\
-
-\noindent The {\tt helmeos} equation of state was developed by Frank Timmes (ASU).
-The version in \maestro\ was downloaded from his website and the
-interface changes slightly for threadsafeness.  We thank Frank for
-making it available.
+\noindent \maestro\ uses the \fboxlib\ library (and formerly, the
+\boxlib\ library), developed at the Center for Computational Sciences
+and Engineering at Lawrence Berkeley National Laboratory.  \fboxlib\ is
+distributed separately from \maestro\ (see
+\url{https://github.com/AMReX-Codes/}) \\
 

### 2018-07-03 Docs/getting_started/getting_started.tex

@@
 .
-\item {\em Get a copy of \amrex}.
+\item {\em Get a copy of \fboxlib}.
 
-  \maestro\ requires the \amrex\ library to manage the grids and
-  parallelization.  We also rely on the build system in \amrex\ to
-  build a \maestro\ executable.  \amrex\ is also available on github
+  \maestro\ requires the \fboxlib\ library to manage the grids and
+  parallelization.  We also rely on the build system in \fboxlib\ to
+  build a \maestro\ executable.  \fboxlib\ is also available on github
   and can be obtained via:
   \begin{verbatim}
-git clone https://github.com/AMReX-Codes/amrex.git
+git clone https://github.com/AMReX-Codes/FBoxLib.git
   \end{verbatim}
 
-  You will periodic want to update \amrex\ by doing
+  You will periodic want to update \fboxlib\ by doing
   \begin{verbatim}
@@
   \end{verbatim}
-  in the {\tt amrex/} directory.
+  in the {\tt fboxlib/} directory.
 
@@
 
-  \maestro\ needs to know where to find \amrex, by specifying the
-  {\tt AMREX\_HOME} environment variable, and where
+  \maestro\ needs to know where to find \fboxlib, by specifying the
+  {\tt FBOXLIB\_HOME} environment variable, and where
   to find \microphysics, bt specifying the {\tt
@@
   \begin{verbatim}
-export AMREX_HOME="/path/to/amrex/"
+export FBOXLIB_HOME="/path/to/FBoxLib/"
 export MICROPHYSICS_HOME="/path/to/Microphysics/"
@@
   \begin{verbatim}
-setenv AMREX_HOME /path/to/amrex/
+setenv FBOXLIB_HOME /path/to/FBoxLib/
 setenv MICROPHYSICS_HOME /path/to/Microphysics/
@@
 
-  Note: you must specify the full path to the {\tt AMReX/} and {\tt
+  Note: you must specify the full path to the {\tt FBoxLib/} and {\tt
     Microphysics/} directory.  Do not use ``$\sim$'' to refer to your
@@
     Specifying this compiler will automatically pull in the compiler
-    settings as specified in {\tt AMREX\_HOME/Tools/F\_mk/}.
+    settings as specified in {\tt FBoxLib\_HOME/Tools/F\_mk/}.
     (Alternate compiler choices include {\tt Intel}, {\tt PGI}, {\tt
@@
   outputted.  Similarly, the checkpoints are named {\tt chk}{\em
-    nnnnn}.  \amrex\ plotfiles and checkpoints are actually
+    nnnnn}.  \fboxlib\ plotfiles and checkpoints are actually
   directories, with the data stored in sub-directories grouped by
@@
 Visualization and analysis are done on the plotfiles.  A number of
-in-house and externally developed tools can work with \amrex-formatted
+in-house and externally developed tools can work with \fboxlib-formatted
 plotfiles\footnote{The plotfiles are in the same format as those made
@@
 
-\amrvis\ can be obtained via git as:
+\amrvis\ can be obtained via git from github as:
 \begin{verbatim}
-git clone https://ccse.lbl.gov/pub/Downloads/Amrvis.git
+git clone https://github.com/AMReX-Codes/Amrvis.git
 \end{verbatim}
@@
    A 1-d version of the {\tt flame} problem above.  This uses a special
-   elliptic solver in \amrex\ that only works for a single grid, so
+   elliptic solver in \fboxlib\ that only works for a single grid, so
    no parallel runs are allowed for this problem.
@@
   conversion between the string names and integer flags is done
-  by \amrex\ in the \code{bc\_module} at the time of initializing
+  by \fboxlib\ in the \code{bc\_module} at the time of initializing
   the runtime parameters.
@@
   YY-MM}.  We do this on all the other repos in the \amrex-ecosystem,
-including {\tt amrex/}, {\tt Microphysics/}, and {\tt Castro/}.
+including {\tt FBoxLib/}, {\tt Microphysics/}, and {\tt Castro/}.
 
@@
 set the proper libraries.  For other machines, you may need to edit
-the {\tt GMake.MPI} file in the \amrex\ build files.  See
+the {\tt GMake.MPI} file in the \fboxlib\ build files.  See
 \S~\ref{ch:make} for more details.

### 2018-11-07 Docs/FAQ/faq.tex

@@
 
-   The environment variable {\tt AMREX\_HOME} needs to be the full path
-   to the {\tt amrex/} directory.  You cannot use `{\tt $\sim$}' as a shortcut
+   The environment variable {\tt FBOXLIB\_HOME} needs to be the full path
+   to the {\tt FBoxLib/} directory.  You cannot use `{\tt $\sim$}' as a shortcut
    for your home directory.
@@
 
-\amrex\ can intercept floating point exceptions and provide a helpful
+\fboxlib\ can intercept floating point exceptions and provide a helpful
 backtrace that shows you where they were generated.  See \S

### 2018-11-07 Docs/MESA/hopper_compile.tex

@@
 These notes are somewhat dated because they pertain only to {\sf MESA} version 
-4088. The above notes are for version 4411. In my experience, version 4088 is 
-much easier to install than 4411. When using the scripts in 
+4088. In my experience, version 4088 is 
+very easy to install and follows the same steps as outlined above. When using 
+the scripts in 
 {\sf AstroDev/MESA/networks/scripts}, make sure the proper list of modules to 
@@
 
-Once {\sf MESA} was installed and working properly on a local machine (e.g. 
+Once {\sf MESA} was installed and working properly on a local machine (e.g.\ 
 Bender), I tried installing it on Hopper, which is a Cray XE6 machine. On 
@@
 \begin{enumerate}
-\item Download the {\tt MESA} source code using the command described in 
-Section \ref{sec:setupmesa}
+\item Download the {\tt MESA} source code (Section \ref{sec:setupmesa}) and 
+be sure to use the correct version: {\tt svn co -r 4088 ...}
 \item Edit the {\tt makefile\_header} in the {\tt utils} directory
@@
 \item Edit the {\tt install} script in the main {\sf MESA} directory
-\item Edit the {\tt BoxLib/Tools/F\_mk/GMakedefs.mak} makefile
-\item Edit the {\tt BoxLib/Tools/F\_mk/comps/Linux\_cray.mak} file
+\item Edit the {\tt amrex/Tools/F\_mk/GMakedefs.mak} makefile
+\item Edit the {\tt amrex/Tools/F\_mk/comps/Linux\_cray.mak} file
 \item Install {\sf MESA}
@@
 When creating the library for these four modules, there is a {\tt LIB\_DEFS} 
-variable in the makefile that holds the {\tt *\_def.o} file, e.g. 
+variable in the makefile that holds the {\tt *\_def.o} file, e.g.\ 
 {\tt utils\_def.o} or {\tt num\_def.o}. This file is never loaded into the 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\sf BoxLib Linux\_cray.mak} file}
+\subsection{{\sf AMReX Linux\_cray.mak} file}
 
@@
 and {\sf MESA} compatible, I added the {\tt -ef} flag to the 
-{\tt Linux\_cray.mak} file located in {\tt BoxLib/Tools/F\_mk/comps}. I also 
+{\tt Linux\_cray.mak} file located in {\tt amrex/Tools/F\_mk/comps}. I also 
 added a few more debugging flags:
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\sf BoxLib GMakedefs.mak} makefile}
+\subsection{{\sf AMReX GMakedefs.mak} makefile}
 
@@
 this, I modified what {\tt tdir} is defined to be when using the Cray compiler. 
-This was done in the {\tt BoxLib/Tools/F\_mk/GMakedefs.mak} makefile by using 
+This was done in the {\tt amrex/Tools/F\_mk/GMakedefs.mak} makefile by using 
 a simple if statement:

### 2018-11-07 Docs/makefiles/makefiles.tex

@@
 \begin{itemize}
-\item {\tt \$\{AMREX\_HOME\}/Tools/F\_mk/GMakedefs.mak}:
+\item {\tt \$\{FBOXLIB\_HOME\}/Tools/F\_mk/GMakedefs.mak}:
 
@@
 
-\item {\tt \$\{AMREX\_HOME\}/Tools/F\_mk/GMakeMPI.mak}:
+\item {\tt \$\{FBOXLIB\_HOME\}/Tools/F\_mk/GMakeMPI.mak}:
 
@@
 
-\item {\tt \$\{AMREX\_HOME\}/Tools/F\_mk/GMakerules.mak}:
+\item {\tt \$\{FBOXLIB\_HOME\}/Tools/F\_mk/GMakerules.mak}:
 
@@
   modules that are used to build a typical \maestro\ application
-  and integrates with the \amrex\ build system.  Every \maestro\
+  and integrates with the \fboxlib\ build system.  Every \maestro\
   problem's {\tt GNUmakefile} will include this file.
@@
 
-\maestro\ gets the location of the \amrex\ library through the 
-{\tt AMREX\_HOME} variable.  This should be set as an environment
+\maestro\ gets the location of the \fboxlib\ library through the 
+{\tt FBOXLIB\_HOME} variable.  This should be set as an environment
 variable in your shell start-up files (e.g.\ {\tt .bashrc} or {\tt
@@
 
-The \amrex\ build system separates the compiler-specific information
+The \fboxlib\ build system separates the compiler-specific information
 from the machine-specific information---this allows for reuse of the
@@
 The compiler flags for the various compilers are listed in the
-files in {\tt \$\{AMREX\_HOME\}/Tools/F\_mk/comps/}.  The compiler
+files in {\tt \$\{FBOXLIB\_HOME\}/Tools/F\_mk/comps/}.  The compiler
 is set via the {\tt COMP} variable in the problem's {\tt GNUmakefile}.
@@
 
-The \amrex\ build system relies on the vpath functionality of {\tt
+The \fboxlib\ build system relies on the vpath functionality of {\tt
 make}.  In a makefile, the {\tt vpath} variable holds search path used
@@
 source files for Fortran modules.  The scripts in {\tt
-AMREX\_HOME/Tools/F\_scripts/} are run at the start of the build
+FBOXLIB\_HOME/Tools/F\_scripts/} are run at the start of the build
 process and parse all the source files and make an explicit list of
@@
   created by the script
-  {\tt write\_probin.py} in {\tt \$\{AMREX\_HOME\}/Tools/F\_scripts/}.  The
+  {\tt write\_probin.py} in {\tt \$\{FBOXLIB\_HOME\}/Tools/F\_scripts/}.  The
   makefile logic for creating it is in {\tt GMaestro.mak}.  At compile
@@
   flags, etc.).  This is created by the script {\tt makebuildinfo.py}
-  in {\tt \$\{AMREX\_HOME\}/Tools/F\_scripts/} from {\tt
+  in {\tt \$\{FBOXLIB\_HOME\}/Tools/F\_scripts/} from {\tt
   GMaestro.mak} by passing in a number of makefile variables.  This is
@@
 Several modules are included in all \maestro\ builds by default.
-From \amrex, we alway include:
+From \fboxlib, we alway include:
 \begin{itemize}
-\item {\tt \$\{AMREX\_HOME\}/Src/F\_BaseLib}
-\item {\tt \$\{AMREX\_HOME\}/Src/LinearSolvers/F\_MG}
+\item {\tt \$\{FBOXLIB\_HOME\}/Src/F\_BaseLib}
+\item {\tt \$\{FBOXLIB\_HOME\}/Src/LinearSolvers/F\_MG}
 \end{itemize}
@@
 You can instead link in a system-wide optimized BLAS library by setting
-{\tt SYSTEM\_BLAS := t}\otherindex{{\tt GNUmakefile}}{BLAS} in the {\tt GNUmakefile}.  This adds {\tt -lblas}
+{\tt SYSTEM\_BLAS := t}\otherindex{{\tt GNUmakefile}}{BLAS} in the {\tt GNUmakefile}.  This will
+use the library specified in \otherindex{{\tt GNUmakefile}}{BLAS\_LIBRARY} {\tt BLAS\_LIBRARY}
+(defaulting to {\tt -lopenblas})
 to the link line, and assumes that the library is in your path.  Note
@@
 \end{itemize}
-(although see the unit tests section below regarding {\tt MAESTRO/Source}.
+(although see the unit tests section below regarding {\tt MAESTRO/Source}.)
 
@@
 
-\subsection{\amrex-only Tests}
+\subsection{\fboxlib-only Tests}
 
@@
 
-\amrex\ has several options that produce executables that can help
+\fboxlib\ has several options that produce executables that can help
 track down memory issues, uninitialized variables, NaNs, etc.
@@
   Setting {\tt TEST := t}\otherindex{{\tt GNUmakefile}}{TEST} will
-  enable routines in \amrex\ initialize multifabs and arrays 
+  enable routines in \fboxlib\ initialize multifabs and arrays 
   allowed via {\tt bl\_allocate} to signalliing NaNs.  This behavior
@@
   exceptions (FPEs), but checks on floating point exceptions can also
-  be enabled through runtime parameters passed to \amrex's
+  be enabled through runtime parameters passed to \fboxlib's
   backtrace functionlity:
@@
 
-  When exception trapping is enabled (either via \amrex\ or the
+  When exception trapping is enabled (either via \fboxlib\ or the
   compiler), the code will abort, and the backtrace information will
   be output to a file {\tt Backtrace.N}, where {\tt N} is the
-  processor number.  \amrex\ will also initialize \multifab s with
+  processor number.  \fboxlib\ will also initialize \multifab s with
   signaliing NaNs to help uncover any floating point issues.
@@
 
-  Behind the scenes, \amrex\ implements this capability via the
+  Behind the scenes, \fboxlib\ implements this capability via the
   Linux/Unix {\tt feenableexcept} function (this is in {\tt
-    backtrace\_c.cpp} in \amrex).
+    backtrace\_c.cpp} in \fboxlib).
 
@@
 Properties for different compilers are already defined in {\tt
-  \$\{AMREX\_HOME\}/Tools/F\_mk/comps/}.  Each compiler is given its
+  \$\{FBOXLIB\_HOME\}/Tools/F\_mk/comps/}.  Each compiler is given its
 own file.  The appropriate file is included into {\tt GMakedefs.mak}

### 2018-11-07 Docs/visualization/visualization.tex

@@
 
-\subsection{the \boxlib\ file format}
+\subsection{the \amrex\ file format}
 
@@
 
-\amrvis\ is a tool developed together with \boxlib\ to visualize datasets
-from codes built around the \boxlib\ library.  You can download the
-\amrvis\ source from: \\
-\url{https://ccse.lbl.gov/Downloads/downloadAmrvis.html} \\
+\amrvis\ is a tool developed together with \amrex\ to visualize
+datasets from codes built around the \amrex\ library (including
+\fboxlib\ used here).  You can download the \amrvis\ source from:
+\\ \url{https://ccse.lbl.gov/Downloads/downloadAmrvis.html} \\
 
-\amrvis\ exists in the C++ \boxlib\ framework, so the build system is
+\amrvis\ exists in the C++ \amrex\ framework, so the build system is
 slightly different.  A different executable is needed for 2- vs.\ 3-d
@@
 {\tt AmrPostprocessing/python} provides some simple commandline
-tools for doing visualizations of \boxlib\ plotfiles (note: a subset
-of these are distributed directly with \boxlib\ in {\tt BoxLib/Tools/Py\_util/}).  The main drivers
+tools for doing visualizations of \amrex\ plotfiles (note: a subset
+of these are distributed directly with \amrex\ in {\tt amrex/Tools/Py\_util/}).  The main drivers
 are written in python and use a set of Fortran routines, compiled with
@@
 
-{\tt plotsinglevar.py} does visualizations of 2-d \boxlib\ plotfiles,
-and slices through 3-d \boxlib\ plotfiles.  A simple plot can be made
+{\tt plotsinglevar.py} does visualizations of 2-d \amrex\ plotfiles,
+and slices through 3-d \amrex\ plotfiles.  A simple plot can be made
 via:
@@
 
-\yt is a Python package for analyzing and visualizing simulation data.
-Originally \yt was written for the {\tt Enzo} code, with focus on
-cosmological simulation data.  It has since been refactored to remove
-the specific {\tt Enzo}isms, and now can work on various data formats,
-such as the BoxLib data from \maestro\ and \castro, HDF5 data
-from {\tt FLASH}, and data from the SPH code {\tt Gadget}.  For more
+\yt is a Python package for analyzing and visualizing simulation data,
+and understand that \amrex\ data from \maestro\ and \castro\ (along
+with many other simulation codes). For more
 information, see the \yt homepage at \url{http://yt-project.org/} and

### 2018-11-07 Docs/architecture/architecture.tex

@@
 such restriction (again, see Figure~\ref{fig:base_state}).
-The \maestro\ grids are managed by the \boxlib\ library, which is
+The \maestro\ grids are managed by the \fboxlib\ library\footnote{This is a pure Fortran library that 
+has a lot of similarities to the \cpp\ \amrex\ library}, which is
 distributed separately.
@@
   purple lines show the directories that are pulled in through
-  various Makefile variables ({\tt BOXLIB\_HOME}, {\tt NETWORK\_DIR},
+  various Makefile variables ({\tt FBOXLIB\_HOME}, {\tt NETWORK\_DIR},
   {\tt EOS\_DIR}, and {\tt CONDUCTIVITY\_DIR}).}
@@
 files (distributed in the {\tt MAESTRO/} directory), and the
-\boxlib\ library (distributed separately, consisting of the {\tt BoxLib/} directory).
-\boxlib\ provides both a C++ and a Fortran framework.  \maestro\
-only uses the Fortran portions of \boxlib.  Figure~\ref{fig:arch:eco}
+\fboxlib\ library (distributed separately, consisting of the {\tt FBoxLib/} directory).
+Figure~\ref{fig:arch:eco}
 shows the relationship between the different packages, highlighting
@@
 
-  \item {\tt Microphysics/}\footnote{Note: many more \compatible\ routines are available in the separate \microphysics\ git repo}
+  \item {\tt Microphysics/}\footnote{Note: many more compatible routines are available in the separate \microphysics\ git repo}
 
@@
 \centering
-\includegraphics[scale=0.65]{\archfigpath/boxlib_directory2}
-\caption[\boxlib\ directory structure] {\label{fig:arch:boxlib} The
-  basic \boxlib\ directory structure.  The directories used by
+\includegraphics[scale=0.6]{\archfigpath/amrex_directory2}
+\caption[\amrex\ directory structure] {\label{fig:arch:amrex} The
+  basic \amrex\ directory structure.  The directories used by
   \maestro\ are indicated in red.}
@@
 
-The \boxlib\ directory structure is shown in
-Figure~\ref{fig:arch:boxlib}.  The subset of the directories that are
+The \amrex\ directory structure is shown in
+Figure~\ref{fig:arch:amrex}.  The subset of the directories that are
 used by \maestro\ are:
@@
 
-  The main \boxlib\ source directory.  In \maestro, we only use the
+  The main \amrex\ source directory.  In \maestro, we only use the
   Fortran source files.  The core directories are:
@@
 
-    The Fortran \boxlib\ files.  This is a library for describing
-    meshes consisting of a union of boxes.  The \boxlib\ modules
-    define the basic datatypes used in \maestro.  \boxlib\ also
+    The Fortran \amrex\ files.  This is a library for describing
+    meshes consisting of a union of boxes.  The \amrex\ modules
+    define the basic datatypes used in \maestro.  \amrex\ also
     provides the routines that handle the parallelization and I/O.
@@
 
-    The \boxlib\ linear solvers---these are used to solve elliptic
+    The \amrex\ linear solvers---these are used to solve elliptic
     problems in the \maestro\ algorithm.
@@
 
-  Various tools used for building \boxlib\ applications.  Here we use:
+  Various tools used for building \amrex\ applications.  Here we use:
 
@@
 
-Finally the {\tt BoxLib/Tools/Postprocessing/F\_Src} package provides simple
+Finally the {\tt FBoxLib/Tools/Postprocessing/F\_Src} package provides simple
 Fortran-based analysis routines (e.g.\ extract a line from a
-multidimensional dataset) that operate on BoxLib datasets.  These are
+multidimensional dataset) that operate on \amrex\ datasets.  These are
 described in \S~\ref{sec:analysis}.  Several sub-directories with
@@
 
-\section{BoxLib Data Structures}
-
-\maestro's gridding is handled by the BoxLib library, which
+\section{\fboxlib\ Data Structures}
+
+\maestro's gridding is handled by the \fboxlib\ library, which
 contains the most fundamental objects used to construct parallel
@@
 On a grid, the data can be stored at cell-centers, on a face/edge, or
-on the corners.  In BoxLib, data that is on an edge is termed `nodal'
+on the corners.  In \fboxlib, data that is on an edge is termed `nodal'
 in that direction (see Figure~\ref{fig:dataloc}).  Data that is on the
@@
 
-To simplify the description of the underlying AMR grid, BoxLib
+To simplify the description of the underlying AMR grid, \fboxlib\
 provides a number of Fortran types.  We briefly summarize the major
-data types below.  A more extensive introduction to \boxlib\ is 
-provided by the \boxlib\ User's Guide, distributed with the library.
+data types below.  A more extensive introduction to \fboxlib\ is 
+provided by the \fboxlib\ User's Guide, distributed with the library.
 
@@
   resolution, the domain is 20$\times$18 zones.  Note that the
-  indexing in BoxLib starts with $0$.}
+  indexing in \fboxlib\ starts with $0$.}
 \end{figure}
@@
 from one \multifab\ to another---see
-\code{BoxLib/Src/F\_BaseLib/multifab.f90} for the full list of
+\code{FBoxLib/Src/F\_BaseLib/multifab.f90} for the full list of
 routines that work with \multifab s.
@@
 but many more are available through the external \microphysics\ repo\footnote{\microphysics\ is
-available at \url{https://github.com/BoxLib-Codes/Microphysics}.  \maestro\ will
+available at \url{https://github.com/starkiller-astro/Microphysics}.  \maestro\ will
 find it via the {\tt MICROPHYSICS\_HOME} environment variable}.  The \microphysics\
@@
 
-\section{BoxLib Helper Modules}
-
-There are a large number of modules in {\tt boxlib/} that provide
+\section{\fboxlib\ Helper Modules}
+
+There are a large number of modules in {\tt FBoxLib/} that provide
 the core functionality for managing grids.  Here we describe
@@
 \noindent Here, the {\tt bl\_types} and \code{multifab\_module}
-modules bring in the basic BoxLib data types. Specifically, here,
+modules bring in the basic \fboxlib\ data types. Specifically, here,
 \code{bl\_types} defines {\tt dp\_t} which is the {\tt kind} used for
@@
 You generally won't see calls in the MAESTRO source code to these subroutines,
-as there is now a special BoxLib subroutine, {\tt ml\_restrict\_and\_fill},
+as there is now a special \fboxlib\ subroutine, {\tt ml\_restrict\_and\_fill},
 that takes an array of multifabs at different level, and in order calls:
@@
 conditions are specified via integer values that are defined in {\tt
-  bc\_module} (part of \boxlib).
+  bc\_module} (part of \fboxlib).
 
@@
 
-There is an extensive regression test suite for \boxlib\ that works with
+There is an extensive regression test suite for \amrex\ that works with
 \maestro.  Full details, and a sample \maestro\ configuration file are
-provided in the \boxlib\ User's Guide and source.
+provided in the \amrex\ User's Guide and source.

### 2019-08-20 Docs/mg/mg.tex

@@
 
-\section{\amrex\ Multigrid Philosophy}
+\section{\fboxlib\ Multigrid Philosophy}
 
-Here we describe some of the general ideas behind the \amrex\
+Here we describe some of the general ideas behind the \fboxlib\
 multigrid (MG).
@@
 We solve MG on an AMR hierarchy, which means in places we will encounter
-C-F interfaces.  The \amrex\ MG will modify the stencil for the Laplacian
+C-F interfaces.  The \fboxlib\ MG will modify the stencil for the Laplacian
 at C-F interfaces to ensure the correct solution to the Poisson equation
@@
 some codes (but not \maestro) we can have jumps of $4\times$ between
-levels.  \amrex\ uses a {\tt mini\_cycle} in these cases, effectively
+levels.  \fboxlib\ uses a {\tt mini\_cycle} in these cases, effectively
 inserting a multigrid level between the two AMR levels and doing a mini
@@
 
-The MG solvers are located in {\tt amrex/Src/LinearSolvers/F\_MG/}.
+The MG solvers are located in {\tt FBoxLib/Src/LinearSolvers/F\_MG/}.
 There are two MG solvers, for cell-centered and nodal data.
@@
 The \mgtower\ is a special Fortran derived type that carries all the
-information required for a \amrex\ multigrid solve.
+information required for a \fboxlib\ multigrid solve.
 
@@
 \item {\tt ss}: The stencil itself---for each zone, this gives the 
-  coefficients of the terms in the Laplacian, with the convection that
+  coefficients of the terms in the Laplacian, with the convention that
   the `0' term is located in the current cell and the other terms are
@@
 
-There are several bottom solvers available in \amrex.  For \maestro.
+There are several bottom solvers available in \fboxlib.  For \maestro.
 These are set through the {\tt mg\_bottom\_solver} (MAC/cell-centered)
@@
 step (the HG projection).  Two multigrid solvers are provided by
-\amrex---one for cell-centered data and one for node-centered (nodal)
+\fboxlib---one for cell-centered data and one for node-centered (nodal)
 data.  Both of these are used in \maestro.
